### 2015-09-07 paper/paper.tex
added paper folder

%\documentclass[preprint]{sigplanconf}

\documentclass[preprint]{llncs}

\usepackage{color}

\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
%% \usepackage{mathpartir}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}

\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}

% "define" code highlights for Java and Scala
\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

%%\usepackage{natbib}
%%\bibpunct();A{},
%%\let\cite=\citep

%include lhs2TeX.fmt
%include lhs2TeX.sty
%include forall.fmt

\pagestyle{plain}

%{\theorembodyfont{\sffamily} \newtheorem{theorem}{Theorem}}
%{\theorembodyfont{\sffamily} \newtheorem{lemma}{Lemma}}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\william[1]{\authornote{william}{green}{#1}}
%%\newcommand\tom[1]{\authornote{tom}{red}{#1}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\yanlin[1]{\authornote{yanlin}{green}{#1}}
%%\newcommand\bruno[1]{}

\newcommand{\hl}[1]{\textcolor{red}{#1}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

%\renewcommand{\mathindent}{0pt}

\begin{comment}
\author{Bruno C. d. S. Oliveira and Yanlin Wang}
\institute{The University of Hong Kong\\
\email{\{bruno,ylwang\}@cs.hku.hk}\\
\authorrunning{Bruno Oliveira and Yanlin Wang}} % abbreviated author list (for running head)
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Classless Java}
%%\subtitle{Solving The Expression Problem with Covariant Return Types}
%% Poor Man's Family Polymorphism

\maketitle

\begin{abstract}

\end{abstract}

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

%===============================================================================

\section{Introduction}\label{sec:intro}

\begin{itemize}

\item Using annotattions to implement a rich notion of traits with a 
mechanism to instantiate objects (the of method). Goal 1: is to reduce 
the amount of code that is required to program with interfaces and default 
methods. Goal 2: To provide a convenient means to do multiple inheritance 
in Java.

\item Specify the system more formally.

\item Show that we can model all trait operations

\item Implementation using Lombok

\item Case studies: The expression problem, Trivially
Case Studies from traits paper.

\end{itemize}

\section{Overview}\label{sec:ep}
Yanlin

* Explain what the Mixin annotations do using examples.

* Motivate the use of multiple inheritance in Java.

* Maybe use Marco's example.

\section{Comparing to traits and mixins}
Haoyuan

   - vs traits: we support of methods to create new objects (a replacement to constructors);
   Moreover we have the with and clone methods (we miss more applications for those). Show 
   how to model the operations on traits; discuss operations that we cannot model 
   (example: renaming).

   - vs mixins: we use the trait model of explicitly resolving conflicts. This is arguably 
   better for reasoning.

\section{Formal Semantics}\label{sec:typesafety}
Yanlin and Haoyuan

We need to show 2 things:

1) The dynamic semantics: what's the code that gets generated by a mixin annotation;

2) The type system: what programs to reject; properties: generation of type-safe/checkable code.

\bruno{The implementation is still missing the type system (rejecting some programs)!}

\section{Implementation}
Haoyuan

discuss implementation in lombok; and limitations.

\bruno{The implementation does not support separate compilation yet. Can we fix this?}

\section{Case studies}
Haoyuan and Yanlin

\subsection{A Trivial Solution to the Expression Problem}

\subsection{Other case studies} 
\bruno{The case studies still need to be implemented!}

Collections example from traits paper?

Other case studies using multiple inheritance?

\section{Related Work}\label{sec:related}

- traits (original, variations, scala)
- mixins (original, scala)
- multiple inheritance 
- expression problem
- ...

%===============================================================================
\section{Conclusion}\label{sec:conclusion}

%===============================================================================

\bibliographystyle{splncs}
\bibliography{paper}

\appendix

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "."
%%% End:

### 2015-09-07 paper/computePositions.rb

@@ -1,0 +1,62 @@
+#require 'tempfile'
+require 'fileutils'
+
+# Example 1 - Read File and close
+def scan(fileName, beginpos, endpos, fileloc)
+  file = File.new(fileName, "r")
+  counter = 1
+  while (line = file.gets)
+    if line =~ /BEGIN_(\w+)/
+      throw "Symbol '#{$1}' already defined" if fileloc[$1]
+      fileloc[$1] = fileName
+      beginpos[$1] = counter+1
+      puts "BEGIN #{$1}: #{beginpos[$1]}"
+    end
+    if line =~ /END_(\w+)/
+      endpos[$1] = counter-1
+      puts "END #{$1}: #{endpos[$1]}"
+    end
+    counter = counter + 1
+  end
+  file.close
+end
+
+def process(file, out, beginpos, endpos, fileloc)
+  while (line = file.gets)
+    if line =~ /APPLY:(\w+)=(\w+)/
+      var = $1
+      label = $2
+      line = line.sub( /#{var}=([0-9-]*)/, "#{var}=#{beginpos[label]}-#{endpos[label]}" )
+      line = line.sub( /\{[^{]*\}/, "{#{fileloc[label]}}" )
+    end
+    out.write(line)
+  end
+  file.close
+  out.close
+end
+
+fileloc = {}
+beginpos = {}
+endpos = {}
+
+Dir['../code/ScalaCode/EP/*/*.scala'].each do |file| 
+  scan(file, beginpos, endpos, fileloc)
+end
+Dir['../code/*/*/*/*/*.java'].each do |file| 
+  scan(file, beginpos, endpos, fileloc)
+end
+Dir['../code/*/*/*/*.java'].each do |file| 
+  scan(file, beginpos, endpos, fileloc)
+end
+
+
+
+Dir['*.tex'].each do |file| 
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+

### 2015-09-09 paper/paper.tex
added text: return type refinement
point example

@@
 \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
-\newcommand\william[1]{\authornote{william}{green}{#1}}
-%%\newcommand\tom[1]{\authornote{tom}{red}{#1}}
 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
 \newcommand\yanlin[1]{\authornote{yanlin}{green}{#1}}
-%%\newcommand\bruno[1]{}
-
 \newcommand{\hl}[1]{\textcolor{red}{#1}}
@@
 
-%\renewcommand{\mathindent}{0pt}
+\newcommand\mixin{\textbf{@Mixin}}
 
@@
 
-* Maybe use Marco's example.
+* Maybe use Marco's example (Point example).
+
+\lstinputlisting[linerange=33-42]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
+
+\lstinputlisting[linerange=48-74]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
 
@@
 Haoyuan
+
+   - vs both: we do automatic return type refinement, which has useful applications 
+   (example: Expression Problem)
 

### 2015-09-09 UseMixinLombok/src/test/TestLombok.java

@@ -29,13 +29,48 @@ public class TestLombok {
 
 }
 
+//BEGIN_POINT
 @Mixin
 interface Point {
     int X();
-	int Y();
+    int Y();
     void X(int x);
     void Y(int y);
     Point withX(int x);
     Point withY(int y);
     Point clone();
 }
+//END_POINT
+
+
+/****** generated
+//BEGIN_POINT_OF
+    static Point of(int X, int Y) {
+        return new Point() {
+            int _X = X;
+            public int X() {
+                return _X;
+            }
+            int _Y = Y;
+            public int Y() {
+                return _Y;
+            }
+            public Point withX(int X) {
+                return of(X, Y());
+            }
+            public void X(int X) {
+                _X = X;
+            }
+            public void Y(int Y) {
+                _Y = Y;
+            }
+            public Point withY(int Y) {
+                return of(X(), Y);
+            }
+            public Point clone() {
+                return of(X(), Y());
+            }
+        }; 
+    }  
+//END_POINT_OF
+*/

### 2015-09-09 paper/computePositions.rb

@@ -39,13 +39,7 @@ fileloc = {}
 beginpos = {}
 endpos = {}
 
-Dir['../code/ScalaCode/EP/*/*.scala'].each do |file| 
-  scan(file, beginpos, endpos, fileloc)
-end
-Dir['../code/*/*/*/*/*.java'].each do |file| 
-  scan(file, beginpos, endpos, fileloc)
-end
-Dir['../code/*/*/*/*.java'].each do |file| 
+Dir['../UseMixinLombok/src/test/*.java'].each do |file| 
   scan(file, beginpos, endpos, fileloc)
 end
 

### 2015-09-09 paper/paper.tex
overview: Point example

@@
 
+We provide a Java annotation \mixin to provide default implementations for
+various methods and a mechanism to instantiate objects. \mixin annotation helps
+programmers to write less cumbersome code and instantiate interfaces in Java.
+
+For example, interface \texttt{Point} annotated with \mixin:
+
 \lstinputlisting[linerange=33-42]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
+
+\texttt{Point} has two (conceptually) member fields \texttt{X} and
+\texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
+methods. Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
+\emph{setter} methods. Method \texttt{Point withX(int X)} updates field
+\texttt{X} and returns \textbf{this}.
+
+A typical and trivial implementation that programmers usually do is:
+
+\lstinputlisting[linerange=79-109]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
 
@@
 Collections example from traits paper? 
+(\yanlin{couldn't found source code})
 

### 2015-09-09 UseMixinLombok/src/test/TestLombok.java

@@ -74,3 +74,37 @@ interface Point {
     }  
 //END_POINT_OF
 */
+
+//BEGIN_POINTIMPL
+class PointImpl implements Point {
+    private int _X;
+    private int _Y;
+    public PointImpl(int X, int Y) {
+        this._X = X;
+        this._Y = Y;
+    }
+    public int X() {
+        return _X;
+    }
+    public int Y() {
+        return _Y;
+    }
+    public Point withX(int X) {
+        X(X);
+        return this;
+    }
+    public void X(int X) {
+        _X = X;
+    }
+    public void Y(int Y) {
+        _Y = Y;
+    }
+    public Point withY(int Y) {
+        Y(Y);
+        return this;
+    }
+    public Point clone() {
+        return new PointImpl(_X, _Y);
+    }
+}
+//END_POINTIMPL

### 2015-09-12 paper/paper.tex

@@
 
-\lstinputlisting[linerange=33-42]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
+\lstinputlisting[linerange=33-45]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
 
@@
 
-\lstinputlisting[linerange=79-109]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
+\lstinputlisting[linerange=82-112]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
 
-\lstinputlisting[linerange=48-74]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
+\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
+quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
+may not be reused in a single inheritance language. However, with our approach,
+the \mixin annotation will generate a static method \texttt{of} inside
+\texttt{Point}. \texttt{of} makes use of Java anonymous classes and achieves the
+same implementation as \texttt{PointImpl}.
+
+\lstinputlisting[linerange=50-77]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
+
+Inside the anonymous class in the annotated interface, the following code are
+generated:
+\begin{itemize}
+\item For methods inside the interface with the form \texttt{Tx x()}:
+  \begin{itemize}
+   \item \texttt{x} is the getter method, with return type
+     \texttt{Tx}. Conceptually, it is a member field with name \texttt{x} and
+     type \texttt{Tx}.
+   \item generate member field \texttt{\_x} of type \texttt{Tx}, initialized
+     with \texttt{x}.
+   \item generate implemented getter method:
+       \begin{lstlisting}
+           public Tx x() { return _x; }
+       \end{lstlisting}
+   \end{itemize}
+
+\item For methods inside the interface with the form \texttt{void x(Tx x)}:
+  \begin{itemize}
+    \item check if exist method \texttt{Tx x()}. If not, generate error.
+    \item generate implemented setter method:
+       \begin{lstlisting}
+           pubic void x(Tx x) { this.x = x; }
+       \end{lstlisting}
+    \end{itemize}
+
+\item For methods with the form \texttt{T withX(Tx \_)}:
+  \begin{itemize}
+  \item if there is no \texttt{x} field, or type \texttt{Tx} does not match,
+    then generate error.
+  \item implement `withX` using the `of` method.
+  \end{itemize}
+
+\item For methods with the form of \texttt{T clone()}: Use \texttt{of} method as
+  the constructor, to create a new object with the same field values as the
+  current one.
+
+\item For methods with the form of \texttt{T x(Tx \_)}:
+  \begin{itemize}
+    \item check if exist method \texttt{T x()}, if not, generate error.
+    \item inside the inner class, generate
+      \begin{lstlisting}
+          pubic T x(Tx x) { this.x = x; return this;}
+      \end{lstlisting}
+  \end{itemize}
+\end{itemize}
 

### 2015-09-13 paper/paper.tex
Split tex into files. Text for comparison,

--- 2015-09-09 paper/paper.tex
@@
 
-\section{Introduction}\label{sec:intro}
+\input{sections/Introduction.tex}
 
-\begin{itemize}
+\input{sections/Overview.tex}
 
-\item Using annotattions to implement a rich notion of traits with a 
-mechanism to instantiate objects (the of method). Goal 1: is to reduce 
-the amount of code that is required to program with interfaces and default 
-methods. Goal 2: To provide a convenient means to do multiple inheritance 
-in Java.
+\input{sections/Compare.tex}
 
-\item Specify the system more formally.
+\input{sections/Formalization.tex}
 
-\item Show that we can model all trait operations
+\input{sections/Implementation.tex}
 
-\item Implementation using Lombok
+\input{sections/CaseStudy.tex}
 
-\item Case studies: The expression problem, Trivially
-Case Studies from traits paper.
-
-\end{itemize}
-
-\section{Overview}\label{sec:ep}
-Yanlin
-
-* Explain what the Mixin annotations do using examples.
-
-* Motivate the use of multiple inheritance in Java.
-
-* Maybe use Marco's example (Point example).
-
-\lstinputlisting[linerange=33-42]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
-
-\lstinputlisting[linerange=48-74]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
-
-\section{Comparing to traits and mixins}
-Haoyuan
-
-   - vs both: we do automatic return type refinement, which has useful applications 
-   (example: Expression Problem)
-
-   - vs traits: we support of methods to create new objects (a replacement to constructors);
-   Moreover we have the with and clone methods (we miss more applications for those). Show 
-   how to model the operations on traits; discuss operations that we cannot model 
-   (example: renaming).
-
-   - vs mixins: we use the trait model of explicitly resolving conflicts. This is arguably 
-   better for reasoning.
-
-\section{Formal Semantics}\label{sec:typesafety}
-Yanlin and Haoyuan
-
-We need to show 2 things:
-
-1) The dynamic semantics: what's the code that gets generated by a mixin annotation;
-
-2) The type system: what programs to reject; properties: generation of type-safe/checkable code.
-
-\bruno{The implementation is still missing the type system (rejecting some programs)!}
-
-\section{Implementation}
-Haoyuan
-
-discuss implementation in lombok; and limitations.
-
-\bruno{The implementation does not support separate compilation yet. Can we fix this?}
-
-\section{Case studies}
-Haoyuan and Yanlin
-
-\subsection{A Trivial Solution to the Expression Problem}
-
-\subsection{Other case studies} 
-\bruno{The case studies still need to be implemented!}
-
-Collections example from traits paper?
-
-Other case studies using multiple inheritance?
-
-\section{Related Work}\label{sec:related}
-
-- traits (original, variations, scala)
-- mixins (original, scala)
-- multiple inheritance 
-- expression problem
-- ...
+\input{sections/RelatedWork.tex}
 
 %===============================================================================
-\section{Conclusion}\label{sec:conclusion}
+\input{sections/Conclusion.tex}
 

### 2015-09-13 paper/sections/CaseStudy.tex
Split tex into files. Text for comparison,

\section{Case studies}
Haoyuan and Yanlin

\subsection{A Trivial Solution to the Expression Problem}

\subsection{Other case studies}
\bruno{The case studies still need to be implemented!}

Collections example from traits paper?
(\yanlin{couldn't found source code})

Other case studies using multiple inheritance?

### 2015-09-13 paper/sections/Compare.tex
Split tex into files. Text for comparison,

\section{Comparing to traits and mixins}

\begin{comment}
Haoyuan

   - vs both: we do automatic return type refinement, which has useful applications
   (example: Expression Problem)

   - vs traits: we support of methods to create new objects (a replacement to constructors);
   Moreover we have the with and clone methods (we miss more applications for those). Show
   how to model the operations on traits; discuss operations that we cannot model
   (example: renaming).

   - vs mixins: we use the trait model of explicitly resolving conflicts. This is arguably
   better for reasoning.
\end{comment}

Our approach is based on code generation by Java annotations. The model we generate encourages composability and reusability in object-oriented programming, and is considered to be an alternative to traits or mixins,  meanwhile achieving better performance in some situations. Hence it is necessary for us to make a comparison between this approach and traits (or mixins) we commonly used before.

Our approach is quite different from mixins, in the sense that we use the trait model of explicitly resolving conflicts. Just as [Scharli2003] demonstrated in the paper, mixin inheritance is a good approach of achieving code reuse, nevertheless, the mixin model is not so expressive to resolve conflicts from many mixins. In the trait model, aliases and exclusions are provided for explicit conflict resolution. Such operations can actually be modelled in the mechanism of our approach.

Here we present how the original operations on traits are supported by our model.
\begin{itemize}
\item \textbf{Symmetric sum}: the symmetric composition of two disjoint traits is achieved by simply implementing two interfaces in Java correspondingly, without overriding any method. The composition relies on multiple inheritance on interfaces, which is supported by Java.
\item \textbf{Override}: the overriding operation (also known as asymmetric sum) is modelled by implementing many interfaces, while overriding some methods inside. The code below gives an example of explicitly specifying which super interface to refer to, regarding two methods with the same name.
    \begin{lstlisting}
    interface A { default int m() {return 1;} }
    interface B { default int m() {return 2;} }
    interface C extends A, B { default int m() {return B.super.m();} } 
    \end{lstlisting}
    Here the method \texttt{m()} in interface \texttt{C} simply inherits from \texttt{B.m()}.
\item \textbf{Alias}: an alias operation adds a new name to an old method when creating the new trait. In Java, we just create a new method with reference to the existing method in its super interface. See the example below, where the new method \texttt{k()} is an alias of the existing method \texttt{m()}.
    \begin{lstlisting}
    interface A { default int m() {return 1;} }
    interface B extends A { default int k() {return A.super.m();} }
    \end{lstlisting}
\item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces. See the example below.
    \begin{lstlisting}
    interface A { default int m() {return 1;} }
    interface B extends A { int m(); }
    \end{lstlisting}
\end{itemize}

Besides, we support \texttt{of} methods in our model, as a replacement to the constructors in original traits. Furthermore, we also support \texttt{with} and \texttt{clone} methods as auxiliary constructors, making the creation of instances more flexible and convenient. Conversely, there are certain operations we cannot model, such as method renaming (as in [Reppy2006]), which breaks structural subtyping.

A further feature leads to return type refinement in our model. Generally speaking, when we use inheritance to create a new trait, with the return type of an existing method being refined, the new \texttt{of} method keeps this consistency. This feature is very useful in many applications; we will see how it makes a difference in our Expression Problem example, in Section [Case Study].

### 2015-09-13 paper/sections/Formalization.tex
Split tex into files. Text for comparison,

\section{Formal Semantics}\label{sec:typesafety}
Yanlin and Haoyuan

We need to show 2 things:

1) The dynamic semantics: what's the code that gets generated by a mixin annotation;

2) The type system: what programs to reject; properties: generation of type-safe/checkable code.

\bruno{The implementation is still missing the type system (rejecting some programs)!}

### 2015-09-13 paper/sections/Implementation.tex
Split tex into files. Text for comparison,

\section{Implementation}
Haoyuan

discuss implementation in lombok; and limitations.

\bruno{The implementation does not support separate compilation yet. Can we fix this?}

### 2015-09-13 paper/sections/Introduction.tex
Split tex into files. Text for comparison,

\section{Introduction}\label{sec:intro}

\begin{itemize}

\item Using annotattions to implement a rich notion of traits with a
mechanism to instantiate objects (the of method). Goal 1: is to reduce
the amount of code that is required to program with interfaces and default
methods. Goal 2: To provide a convenient means to do multiple inheritance
in Java.

\item Specify the system more formally.

\item Show that we can model all trait operations

\item Implementation using Lombok

\item Case studies: The expression problem, Trivially
Case Studies from traits paper.

\end{itemize}

### 2015-09-13 paper/sections/Overview.tex
Split tex into files. Text for comparison,

\section{Overview}\label{sec:ep}
Yanlin

* Explain what the Mixin annotations do using examples.

* Motivate the use of multiple inheritance in Java.

* Maybe use Marco's example (Point example).

We provide a Java annotation \mixin to provide default implementations for
various methods and a mechanism to instantiate objects. \mixin annotation helps
programmers to write less cumbersome code and instantiate interfaces in Java.

For example, interface \texttt{Point} annotated with \mixin:

\lstinputlisting[linerange=33-45]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT

\texttt{Point} has two (conceptually) member fields \texttt{X} and
\texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
methods. Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
\emph{setter} methods. Method \texttt{Point withX(int X)} updates field
\texttt{X} and returns \textbf{this}.

A typical and trivial implementation that programmers usually do is:

\lstinputlisting[linerange=82-112]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL

\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
may not be reused in a single inheritance language. However, with our approach,
the \mixin annotation will generate a static method \texttt{of} inside
\texttt{Point}. \texttt{of} makes use of Java anonymous classes and achieves the
same implementation as \texttt{PointImpl}.

\lstinputlisting[linerange=50-77]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF

Inside the anonymous class in the annotated interface, the following code are
generated:
\begin{itemize}
\item For methods inside the interface with the form \texttt{Tx x()}:
  \begin{itemize}
   \item \texttt{x} is the getter method, with return type
     \texttt{Tx}. Conceptually, it is a member field with name \texttt{x} and
     type \texttt{Tx}.
   \item generate member field \texttt{\_x} of type \texttt{Tx}, initialized
     with \texttt{x}.
   \item generate implemented getter method:
       \begin{lstlisting}
           public Tx x() { return _x; }
       \end{lstlisting}
   \end{itemize}

\item For methods inside the interface with the form \texttt{void x(Tx x)}:
  \begin{itemize}
    \item check if exist method \texttt{Tx x()}. If not, generate error.
    \item generate implemented setter method:
       \begin{lstlisting}
           pubic void x(Tx x) { this.x = x; }
       \end{lstlisting}
    \end{itemize}

\item For methods with the form \texttt{T withX(Tx \_)}:
  \begin{itemize}
  \item if there is no \texttt{x} field, or type \texttt{Tx} does not match,
    then generate error.
  \item implement `withX` using the `of` method.
  \end{itemize}

\item For methods with the form of \texttt{T clone()}: Use \texttt{of} method as
  the constructor, to create a new object with the same field values as the
  current one.

\item For methods with the form of \texttt{T x(Tx \_)}:
  \begin{itemize}
    \item check if exist method \texttt{T x()}, if not, generate error.
    \item inside the inner class, generate
      \begin{lstlisting}
          pubic T x(Tx x) { this.x = x; return this;}
      \end{lstlisting}
  \end{itemize}
\end{itemize}

### 2015-09-13 paper/sections/RelatedWork.tex
Split tex into files. Text for comparison,

\section{Related Work}\label{sec:related}

- traits (original, variations, scala)
- mixins (original, scala)
- multiple inheritance
- expression problem
- ...

### 2015-09-15 paper/paper.tex
Attempt to grammar inserted

@@
 \documentclass[preprint]{llncs}
+\input{def}
 
@@
   morekeywords={public,int,interface,implements,default,
-    abstract,case,catch,class,def,%
+    abstract,case,catch,class,def,static,%
     do,else,extends,false,final,finally,%
@@
 \newcommand{\hl}[1]{\textcolor{red}{#1}}
+\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
 
@@
 \newcommand\mixin{\textbf{@Mixin}}
+\usepackage{xspace}
 

### 2015-09-16 paper/sections/Formalization.tex
Attempt to grammar inserted
syntax fix
Expression typing, with comments

@@
 \section{Formal Semantics}\label{sec:typesafety}
+\marco{Tentative Grammar ClassLessJava}
+
+\begin{figure}[h]
+\begin{grammar}
+\production{
+\e
+}{
+  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
+  }{expressions}\\
+\production{
+\obj
+}{
+\QM{new}\C\oR\cR\oC\T_1\f_1\QM;\ldots\T_k\f_k\QM;
+\mh_1\oC\QM{return}\e_1\QM{;}\!\cC
+\ldots
+\mh_n\oC\QM{return}\e_n\QM{;}\!\cC
+\cC
+  }{object creation}\\
+\production{
+\metaVar{I}
+}{
+ \ann\ \QM{interface} \C_0\QM{extends}\C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
+  }{interface declaration}\\
+\production{
+\method
+}{
+ \QM{static}\mh\oC\QM{return}\e\QM{;}\!\cC
+\mid 
+\QM{default}\mh\oC\QM{return}\e\QM{;}\cC
+\mid
+\mh\QM{;}
+  }{method declaration}\\
+\production{
+\mh
+}{
+ \T_0 \m  \oR\T_1\x_1\ldots\T_n\x_n\cR
+  }{method header}\\
+\production{
+\ann
+}{
+  \mixinAnn|\emptyset
+  }{annotations}\\
+\end{grammar}
+\caption{Grammar of ClassLess Java}
+\label{Grammar}
+\end{figure}
+
+In Figure~\ref{Grammar} we show the syntax of ClassLess Java.
+
+To be compatible with java, the concrete syntax for an interface declaration with empty supertype list $\C_1\ldots\C_k$ would also omit the \Q@extends@ keyword.
+
 Yanlin and Haoyuan
@@
 
-\bruno{The implementation is still missing the type system (rejecting some programs)!}
+\bruno{The implementation is still missing the type system (rejecting some
+  programs)!}
+
+Expression Typing:
+
+\[
+\inferrule{}{\Gamma \vdash x \in \Gamma(x)} \quad \textsc{(T-Var)}
+\]
+
+\[
+\inferrule{\Gamma \vdash e_0 \in C_0 \\ 
+  \textsf{mtype}(m,C_0) = \overline{D} \to E \\ 
+  \Gamma \vdash \overline{e} \in \overline{C} \\
+  \overline{C} <: \overline{D} }
+{ \Gamma \vdash e_0.m(\overline{e}) \in E }
+\quad \textsc{(T-Invk)}
+\]
+
+\[
+\inferrule{\textsf{mtype}(m,C) = \overline{D} \to E \\
+\Gamma \vdash \overline{e} \in \overline{C} \\
+\overline{C} <: \overline{D} \\
+\textsf{mmodifier}(m,C) = \textbf{static} }
+{\Gamma \vdash C.m(\overline{e}) \in E}
+\quad \textsc{(T-StaticInvk)}
+\]
+
+\[
+\inferrule{\textsf{mtype}(m,C) = \overline{D} \to E \\
+\Gamma \vdash \overline{e} \in \overline{C} \\
+\overline{C} <: \overline{D} \\
+\textsf{checkSuper}(C.super.m(\overline{e}), A, C) \\\\
+\text{\yanlin{check whether A <: C, where A is the enclosing interface of the
+    expression}} }
+{\Gamma \vdash C.\textbf{super}.m(\overline{e}) \in E}
+\quad \textsc{(T-SuperInvk)}
+\]

### 2015-09-17 paper/sections/Overview.tex
Automatic return type refinement for with, clone methods.

@@
 
-\lstinputlisting[linerange=33-45]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
+\lstinputlisting[linerange=33-42]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
 
@@
 
-\lstinputlisting[linerange=82-112]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
+\lstinputlisting[linerange=81-111]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
 
@@
 
-\lstinputlisting[linerange=50-77]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
+\lstinputlisting[linerange=47-76]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
 

### 2015-09-17 MixinLombok/src/core/lombok/eclipse/handlers/HandleMixin.java

@@ -31,6 +31,43 @@ public class HandleMixin extends EclipseAnnotationHandler<Mixin> {
 	HashMap<String, ArrayList<TypeReference>> fieldType; 
 	HashMap<String, TypeReference> hasDefault;
 	
+	/* Auxiliary: create a method. */
+	private MethodDeclaration newMethod() {
+		MethodDeclaration method = new MethodDeclaration(meDecl.compilationResult);
+		method.annotations = null;
+		method.modifiers = ClassFileConstants.AccDefault;
+		method.typeParameters = null;
+		method.returnType = null;
+		method.selector = "".toCharArray();
+		method.arguments = null;
+		method.binding = null;
+		method.thrownExceptions = null;
+		method.bits |= ECLIPSE_DO_NOT_TOUCH_FLAG;
+		method.statements = null;
+		method.bodyStart = method.declarationSourceStart = method.sourceStart = _ast.sourceStart;
+		method.bodyEnd = method.declarationSourceEnd = method.sourceEnd = _ast.sourceEnd;
+		return method;
+	}
+	
+	/* Auxiliary: throw an error. */
+	private void throwError(String s) { thisAnno.addError(s); }
+	
+	/* Auxiliary: print out information for debugging. */
+	@SuppressWarnings("unused") private void print(String s) {
+		MethodDeclaration print = newMethod();
+		print.modifiers = ClassFileConstants.AccStatic;
+		print.returnType = new SingleTypeReference(TypeBinding.VOID.simpleName, p);
+		print.selector = "print".toCharArray();
+		MessageSend printInfo = new MessageSend();
+		printInfo.arguments = new Expression[] {
+			new StringLiteral(s.toCharArray(), _ast.sourceStart, _ast.sourceEnd, 0)
+		};
+		printInfo.receiver = createNameReference("System.out", _src);
+		printInfo.selector = "println".toCharArray();
+		print.statements = new Statement[] { printInfo };
+		injectMethod(me, print);
+	}
+	
 	@Override public void handle(AnnotationValues<Mixin> annotation, Annotation ast, EclipseNode annotationNode) {
 		pS = ast.sourceStart;
 		pE = ast.sourceEnd;
@@ -41,11 +78,12 @@ public class HandleMixin extends EclipseAnnotationHandler<Mixin> {
 		_src = ast;
 		thisAnno = annotationNode;
[…]

### 2015-09-17 UseMixinLombok/src/test/TestAnimal.java

@@ -46,18 +46,11 @@ public class TestAnimal {
 interface Point2D {
 	int X();
 	int Y();
-	Point2D withX(int val);
-	Point2D withY(int val);
 }
 
 @Mixin
 interface Point3D extends Point2D {
 	int Z();
-	Point3D withZ(int val);
-
-	/* not automatically generated for now */
-	Point3D withX(int val);
-	Point3D withY(int val);
 }
 
 interface Animal {

### 2015-09-17 UseMixinLombok/src/test/TestLombok.java

@@ -36,9 +36,6 @@ interface Point {
     int Y();
     void X(int x);
     void Y(int y);
-    Point withX(int x);
-    Point withY(int y);
-    Point clone();
     default int distance() {
         return (int) Math.sqrt(X() * X() + Y() * Y());
     }
@@ -48,6 +45,13 @@ interface Point {
 /****** generated
 //BEGIN_POINT_OF
     // inside interface Point
+    default Point withX(int X) {
+        return of(X, Y());
+    }
+    default Point withY(int Y) {
+        return of(X(), Y);
+    }
+    Point clone();
     static Point of(int X, int Y) {
         return new Point() {
             int _X = X;
@@ -58,18 +62,13 @@ interface Point {
             public int Y() {
                 return _Y;
             }
-            public Point withX(int X) {
-                return of(X, Y());
-            }
             public void X(int X) {
                 _X = X;
             }
             public void Y(int Y) {
                 _Y = Y;
             }
-            public Point withY(int Y) {
-                return of(X(), Y);
-            }
+            @Override
             public Point clone() {
                 return of(X(), Y());
             }

### 2015-09-20 paper/sections/Overview.tex
shorter example in overview

--- 2015-09-13 paper/sections/Overview.tex
@@
 
-\lstinputlisting[linerange=33-45]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT
+\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT
 
@@
 
-\lstinputlisting[linerange=50-77]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF
+\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
 

### 2015-09-20 UseMixinLombok/src/test/TestLombok.java

@@ -29,7 +29,6 @@ public class TestLombok {
 
 }
 
-//BEGIN_POINT
 @Mixin
 interface Point {
     int X();
@@ -40,10 +39,8 @@ interface Point {
         return (int) Math.sqrt(X() * X() + Y() * Y());
     }
 }
-//END_POINT
 
 /****** generated
-//BEGIN_POINT_OF
     // inside interface Point
     // Yanlin: it generated withX, withY and clone methods even if users are not 
     // requiring them. I think this may not be a proper design decision, because 
@@ -78,8 +75,7 @@ interface Point {
                 return of(X(), Y());
             }
         }; 
-    }  
-//END_POINT_OF
+    }
 */
 
 //BEGIN_POINTIMPL

### 2015-09-20 paper/computePositions.rb

@@ -39,13 +39,22 @@ fileloc = {}
 beginpos = {}
 endpos = {}
 
-Dir['../UseMixinLombok/src/test/*.java'].each do |file| 
+Dir['../UseMixinLombok/src/*/*.java'].each do |file| 
   scan(file, beginpos, endpos, fileloc)
 end
 
 
 
 Dir['*.tex'].each do |file| 
+  #temp = Tempfile.new('compute_positions')
+  tempname = "footempfile.txt"
+  temp = File.new(tempname, "w")
+  process(File.new(file, "r"), temp, beginpos, endpos, fileloc)
+  FileUtils.cp(file, "#{file}-old")
+  FileUtils.cp(tempname, file)
+end
+
+Dir['*/*.tex'].each do |file| 
   #temp = Tempfile.new('compute_positions')
   tempname = "footempfile.txt"
   temp = File.new(tempname, "w")

### 2015-09-20 paper/sections/Overview.tex
revise overview for the shorter Point example.

@@
 various methods and a mechanism to instantiate objects. \mixin annotation helps
-programmers to write less cumbersome code and instantiate interfaces in Java.
+programmers to write less cumbersome code and instantiate interfaces in
+Java. More over, with \mixin annotation, programmers can mimic multiple
+inheritance in a restricted form easily.
 
@@
 \texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
-methods. Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
-\emph{setter} methods. Method \texttt{Point withX(int X)} updates field
-\texttt{X} and returns \textbf{this}.
+methods. % Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
+% \emph{setter} methods. Method \texttt{Point withX(int X)} updates field
+% \texttt{X} and returns \textbf{this}.
 
@@
 
-\lstinputlisting[linerange=82-112]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL
+\lstinputlisting[linerange=33-46]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINTIMPL
 
@@
 
-Inside the anonymous class in the annotated interface, the following code are
+Apart from generating code for \emph{getter} methods, the annotation processor
+can also generate \emph{setter} methods, \emph{with} methods, \texttt{clone}
+methods, etc. We summarize all the forms of method \mixin  supports below. Inside
+the anonymous class in the annotated interface, the following code are
 generated:
+
 \begin{itemize}

### 2015-09-20 UseMixinLombok/src/overview/TestPoint.java

@@ -28,6 +28,25 @@ interface Point {
 //END_POINT_OF
 */
 
+/******* code without @Mixin
+//BEGIN_POINTIMPL
+class PointImpl implements Point {
+    private int _x;
+    private int _y;
+    public PointImpl(int x, int y) {
+        this._x = x;
+        this._y = y;
+    }
+    public int x() {
+        return _x;
+    }
+    public int y() {
+        return _y;
+    }
+}
+//END_POINTIMPL
+ * */
+
 public class TestPoint {
     public static void main(String[] args) {
         Point p1 = Point.of(2, 3);

### 2015-09-20 UseMixinLombok/src/test/TestAnimal.java

@@ -42,29 +42,38 @@ public class TestAnimal {
 	
 }
 
+//BEGIN_POINT2D
 @Mixin 
 interface Point2D {
 	int X();
 	int Y();
 }
+//END_POINT2D
 
+//BEGIN_POINT3D
 @Mixin
 interface Point3D extends Point2D {
 	int Z();
 }
+//END_POINT3D
 
+//BEGIN_ANIMAL
 interface Animal {
 	Point2D point();
 	void point(Point2D val);
 }
+//END_ANIMAL
 
+//BEGIN_HORSE
 @Mixin
 interface Horse extends Animal {
 	default void run() {
 		point(point().withX(point().X() + 20));
 	}
 }
+//END_HORSE
 
+//BEGIN_BIRD
 @Mixin
 interface Bird extends Animal {
 	Point3D point();
@@ -78,6 +87,9 @@ interface Bird extends Animal {
 		point(point().withX(point().X() + 40));
 	}
 }
+//END_BIRD
 
+//BEGIN_PEGASUS
 @Mixin
 interface Pegasus extends Horse, Bird {}
+//END_PEGASUS

### 2015-09-20 paper/paper.tex
Expression typing, with comments
revised overview with ColoredPoint example

@@
 \usepackage[scaled=0.85]{beramono}
+\usepackage{mathpartir}
 
@@
     super,this,throw,trait,true,try,%
-    type,val,var,while,with,yield},
+    type,var,while,with,yield},
   otherkeywords={=>,<-,<\%,<:,>:,\#,@},
@@
 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
-\newcommand\yanlin[1]{\authornote{yanlin}{green}{#1}}
+\newcommand\yanlin[1]{\authornote{yanlin}{pink}{#1}}
 \newcommand{\hl}[1]{\textcolor{red}{#1}}

### 2015-09-20 paper/sections/Overview.tex
revised overview with ColoredPoint example

@@
 \end{itemize}
+
+Another advantage: multiple inheritance.
+
+\lstinputlisting[linerange=46-50]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+
+\lstinputlisting[linerange=54-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+
+\lstinputlisting[linerange=61-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+
+\lstinputlisting[linerange=68-73]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+
+\lstinputlisting[linerange=77-89]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+
+\lstinputlisting[linerange=93-94]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+
+Interface \texttt{Point3D} extends \texttt{Point/Point2D} with a new abstract
+method \texttt{int z()} (treated as a member field). Interface \texttt{Horse}
+shows the usage and advantage of \emph{with} methods by method
+\texttt{run()}. Note that the return type of various methods get refined
+automatically by our annotation processor. Besides \emph{with}, other methods
+(including \emph{clone}, \emph{of}) also do type-refinements automatically.
+
+The \emph{``multiple inheritance''} case appears at interface
+\texttt{Pegasus}. Pegasuses can not only run but also fly! Interface
+\texttt{Pegasus} obtains \texttt{fly} and \texttt{run} functionality through
+interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation, actually
+these is no code at all programmars have to write, which otherwise is hard
+to do in Java-like languages that do not support multiple inheritance.

### 2015-09-20 paper/sections/Compare.tex
texts.

@@
 Such operations can actually be modelled in the mechanism of our approach.
+On the other hand, the total ordering property [Scharli2003] restricts the composition of units of behavior.
+The linearization of mixin inheritance cannot provide a perfect total ordering for various requirements, however, in Java, we have multiple inheritance of interfaces, where methods are inherited at the same time from super interfaces.
 
@@
 The composition relies on multiple inheritance on interfaces, which is supported by Java.
+Below is a simple example.
+    \begin{lstlisting}
+    interface A { int x(); }
+    interface B { int y(); }
+    interface C extends A, B {}
+    \end{lstlisting}
 \item \textbf{Override}: the overriding operation (also known as asymmetric sum) is modelled by implementing many interfaces, while overriding some methods inside.
@@
 
-Besides, we support \texttt{of} methods in our model, as a replacement to the constructors in original traits.
-Furthermore, we also support \texttt{with} and \texttt{clone} methods as auxiliary constructors, making the creation of instances more flexible and convenient.
-Conversely, there are certain operations we cannot model, such as method renaming (as in [Reppy2006]), which breaks structural subtyping.
+Besides, we support more features than the original trait model:
+\begin{itemize}
+\item We provide \texttt{of} methods for the annotated interfaces.
+During annotation processing time, the ``fields'' inside an interface are observed and a static method \texttt{of} is automatically injected to the interface with its arguments correspondingly.
+Such a method is a replacement to the constructors in original traits, making instantiation more convenient to use.
+\item We provide \texttt{with-} and \texttt{clone} methods as auxiliary constructors.
+A \texttt{with-} method is generated for each field, just like record update, returning the new object with that field updated.
+A \texttt{clone} method is generated for the interface, returning a copy of the current object.
+Furthermore, we do automatic return type refinement for these two kind of methods.
+This feature is comparatively useful in big examples, making operations and behaviors more flexible, which we will demonstrate later.
+\item We provide two options for generating setters.
+There are two kind of setters which are commonly used, namely \textit{void setters} and \textit{fluent setters}.
+The only difference is that a fluent setter returns the object itself after setting, thus supporting a pipeline of such operations.
+The generation depends on which type of setter is declared in the interface by users.
+\end{itemize}
 
-A further feature leads to return type refinement in our model.
-Generally speaking, when we use inheritance to create a new trait, with the return type of an existing method being refined, the new \texttt{of} method keeps this consistency.
-This feature is very useful in many applications; we will see how it makes a difference in our Expression Problem example, in Section [Case Study].
+These are the additional features supported by our model, conversely, there are certain operations we cannot model, such as method renaming (as in [Reppy2006]), which breaks structural subtyping.

### 2015-09-21 paper/sections/Overview.tex
revised overview

@@
 Interface \texttt{Point3D} extends \texttt{Point/Point2D} with a new abstract
-method \texttt{int z()} (treated as a member field). Interface \texttt{Horse}
-shows the usage and advantage of \emph{with} methods by method
-\texttt{run()}. Note that the return type of various methods get refined
-automatically by our annotation processor. Besides \emph{with}, other methods
-(including \emph{clone}, \emph{of}) also do type-refinements automatically.
+method \texttt{int z()} (treated as a getter for member field
+\texttt{z}). Interface \texttt{Horse} shows the usage and advantage of
+\emph{with} methods by method \texttt{run()}: method \texttt{withX} returns a
+new point object with field \texttt{x} updated by the argument to
+\texttt{withX}. Without these \texttt{with} methods, operations like
+\texttt{run()} would be much harder to define. Note that the return type of
+various methods get refined automatically by our annotation processor. Besides
+\emph{with}, other methods (including \emph{clone}, \emph{of}) also do
+type-refinements automatically.
 
 The \emph{``multiple inheritance''} case appears at interface
-\texttt{Pegasus}. Pegasuses can not only run but also fly! Interface
-\texttt{Pegasus} obtains \texttt{fly} and \texttt{run} functionality through
-interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation, actually
-these is no code at all programmars have to write, which otherwise is hard
-to do in Java-like languages that do not support multiple inheritance.
+\texttt{Pegasus}. Pegasuses can not only \emph{run} but also \emph{fly}!
+Interface \texttt{Pegasus} obtains \texttt{fly} and \texttt{run} functionality
+through interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation,
+actually these is no code at all programmers have to write. The idea of using
+default methods inside interfaces was proposed in ~\cite{}. It enables us to do
+multiple inheritance, which otherwise is hard to do in Java-like languages that
+do not support multiple inheritance, easily.

### 2015-09-23 paper/run.py
added Context and Subtyping rules

@@ -3,3 +3,4 @@ import subprocess
 
 subprocess.call("ruby computePositions.rb", shell=True)
 subprocess.call("xelatex paper.tex", shell=True)
+subprocess.call("open paper.pdf", shell=True)

### 2015-09-23 paper/sections/Formalization.tex
add spaces in Fig1 Grammar
added Context and Subtyping rules
auxiliary function mtype(m,T)
updated auxiliary mtype

@@
 }{
-\QM{new}\C\oR\cR\oC\T_1\f_1\QM;\ldots\T_k\f_k\QM;
-\mh_1\oC\QM{return}\e_1\QM{;}\!\cC
+\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM;\ldots\T_k\ \f_k\QM;\ 
+\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
 \ldots
-\mh_n\oC\QM{return}\e_n\QM{;}\!\cC
+\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
 \cC
@@
 }{
- \ann\ \QM{interface} \C_0\QM{extends}\C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
+ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
   }{interface declaration}\\
@@
 }{
- \QM{static}\mh\oC\QM{return}\e\QM{;}\!\cC
+ \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
 \mid 
-\QM{default}\mh\oC\QM{return}\e\QM{;}\cC
+\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
 \mid
@@
 }{
- \T_0 \m  \oR\T_1\x_1\ldots\T_n\x_n\cR
+ \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
   }{method header}\\
@@
 \]
+
+Context:
+
+\[ \Gamma = \epsilon \mid (x, C) : \Gamma  \]
+
+Subtyping:
+
+\[ \inferrule{}{T <: T} \]
+
+\[ \inferrule{S <: T \\ T <: U}{S <: U}\]
+
+\[ \inferrule{\emph{ann} \spc \textbf{interface} \spc C_0 \spc \textbf{extends} \spc C_1,...,C_k \{...\}}
+{C_0 <: C_1 \\ ... \\ C_0 <: C_k} \]
+
+Interface Table (IT):
+
+A mapping from types \texttt{T} to interface declarations \texttt{L}. From the
+interface table, we can read off the subtype relation between interfaces. The
+subtype relation is given by the \textbf{extends} clauses in \texttt{IT}.
+
+Auxiliary Definitions:
+
+- \textsf{mtype(m, C)} : the signature of method m in C.
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\ 
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\ 
+  m \notin M}
+{ \textsf{mtype(m,T)} = \emptyset } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\ 
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\ 
+  m \notin M}
+{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
+
+\[ \textsf{mmodifier(m, C)} = ? \]
+
+\[ \textsf{mtype(m, C)} = ? \]
+
+\[ \textsf{checksuper(C.super.m(e), A, C)} = ? \]
+
+\yanlin{todo: add auxiliary functions}
+\yanlin{todo: add subtyping rules}

### 2015-09-23 paper/sections/Overview.tex
updated auxiliary mtype

@@
 
-\lstinputlisting[linerange=46-50]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
-\lstinputlisting[linerange=54-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=61-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=68-73]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=77-89]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=80-92]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=93-94]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-09-24 paper/sections/Formalization.tex
More formalization.

@@
 \yanlin{todo: add subtyping rules}
+
+\subsection{Auxiliary function: \textsf{mbody}}
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\
+  m \notin M}
+{ \textsf{mbody(m,T)} } \]
+
+\subsection{Auxiliary function: \textsf{shadow}}
+
+The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+
+(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence 
+\begin{equation*}
+\begin{array}{ll}
+\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
+\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
+\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
+\end{array}
+\end{equation*}
+
+(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
+\begin{equation*}
+\begin{array}{ll}
+\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
+\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
+\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
+\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
+\end{array}
+\end{equation*}
+
+(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error.
+The other direction $body_2.\textsf{returnType}<:body_1.\textsf{returnType}$ follows the same rule.
+It also gives an error if there is no subtyping relationship between two return types.
+\begin{equation*}
+\begin{array}{ll}
+\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
+& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
+\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
+& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
+\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
+\end{array}
+\end{equation*}
+
+\subsection{Auxiliary function: \textsf{replace}}
+
+The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
+
+\begin{equation*}
+\begin{array}{ll}
+\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
+\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
+\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
+\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
+\end{array}
+\end{equation*}

### 2015-09-25 paper/sections/Formalization.tex
more formalization

@@
-\section{Formal Semantics}\label{sec:typesafety}
+\section{Formal Semantics}\label{sec:formal}
+\subsection{Grammar}
+
 \marco{Tentative Grammar ClassLessJava}
@@
 
-Expression Typing:
+\subsection{Typing Rules}
+\subsubsection{Expression Typing}
 
@@
 
-Context:
+\subsubsection{Context}
 
@@
 
-Subtyping:
+\subsubsection{Subtyping}
 
@@
 
-Interface Table (IT):
-
-A mapping from types \texttt{T} to interface declarations \texttt{L}. From the
+\subsubsection{Interface Table}
+The Interface Table (IT) mapping from types \texttt{T} to interface declarations \texttt{L}. From the
 interface table, we can read off the subtype relation between interfaces. The
@@
 
-Auxiliary Definitions:
+\subsection{Auxiliary Definitions}
+\yanlin{todo: add more auxiliary functions}
 
+\yanlin{todo: add interface context}
+
+\subsubsection{Auxiliary function: \textsf{mtype}}
 - \textsf{mtype(m, C)} : the signature of method m in C.
@@
 
-\[ \textsf{mmodifier(m, C)} = ? \]
-
-\[ \textsf{mtype(m, C)} = ? \]
-
-\[ \textsf{checksuper(C.super.m(e), A, C)} = ? \]
-
-\yanlin{todo: add auxiliary functions}
-\yanlin{todo: add subtyping rules}
-
-\subsection{Auxiliary function: \textsf{mbody}}
+\subsubsection{Auxiliary function: \textsf{mbody}}
 
@@
 
-\subsection{Auxiliary function: \textsf{shadow}}
+\subsubsection{Auxiliary function: \textsf{shadow}}
 
@@
 
-\subsection{Auxiliary function: \textsf{replace}}
+\subsubsection{Auxiliary function: \textsf{replace}}
 

### 2015-09-25 paper/sections/Overview.tex
revised overview, with subsection structure

@@
 \section{Overview}\label{sec:ep}
-Yanlin
 
-* Explain what the Mixin annotations do using examples.
+\subsection{A Running Example: \texttt{Point}}
+Suppose we want to create a point component that models the a point in space,
+that has x-coordinate and y-coordinate. For example, if we create the
+\texttt{Point} interface in Java, it would look like this:
 
-* Motivate the use of multiple inheritance in Java.
+\begin{lstlisting}
+interface Point {
+    int x();
+    int y(); 
+}
+\end{lstlisting}
 
-* Maybe use Marco's example (Point example).
-
-We provide a Java annotation \mixin to provide default implementations for
-various methods and a mechanism to instantiate objects. \mixin annotation helps
-programmers to write less cumbersome code and instantiate interfaces in
-Java. More over, with \mixin annotation, programmers can mimic multiple
-inheritance in a restricted form easily.
-
-For example, interface \texttt{Point} annotated with \mixin:
-
-\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT
-
-\texttt{Point} has two (conceptually) member fields \texttt{X} and
-\texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
-methods. % Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
+\texttt{Point} has two (conceptually) member fields \texttt{x} and \texttt{y},
+representing the two coordinates of a point. Methods \texttt{int x()} and
+\texttt{int y()} serve as \emph{getter} methods. 
+% Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
 % \emph{setter} methods. Method \texttt{Point withX(int X)} updates field
@@
 
-A typical and trivial implementation that programmers usually do is:
+\subsection{Naive Implementation}
+In Java, to implement an interface like \texttt{Point}, a typical and trivial
+approach that programmers usually do is creating a class extending the interface
+and providing implementation for all methods inside. For example, this is the
+implementation for interface \texttt{Point}:
 
@@
 quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
-may not be reused in a single inheritance language. However, with our approach,
-the \mixin annotation will generate a static method \texttt{of} inside
-\texttt{Point}. \texttt{of} makes use of Java anonymous classes and achieves the
-same implementation as \texttt{PointImpl}.
+may not be reused in a single inheritance language.
+
+\subsection{Our Approach}
+Instead of writing a whole another class to provide the implementation for
+\texttt{Point}, we annotate on interface \texttt{Point} directly with \mixin:
+
+\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT
+
+The \mixin annotation will generate a static method \texttt{of} inside
+\texttt{Point}. The method \texttt{of} mimic the functionality of constructors,
+it takes arguments same as constructors and return objects similar to
+constructors. It makes use of Java anonymous classes and achieves the same
+implementation as \texttt{PointImpl}.
+
+With our approach, we provide a Java annotation \mixin to provide default
+implementations for various methods and a mechanism to instantiate
+objects. \mixin annotation helps programmers to write less cumbersome code and
+instantiate interfaces in Java.
 
@@
 
+\subsection{What  \mixin Generates}
 Apart from generating code for \emph{getter} methods, the annotation processor
@@
        \begin{lstlisting}
-           pubic void x(Tx x) { this.x = x; }
+           public void x(Tx x) { this.x = x; }
        \end{lstlisting}
@@
       \begin{lstlisting}
-          pubic T x(Tx x) { this.x = x; return this;}
+          public T x(Tx x) { this.x = x; return this;}
       \end{lstlisting}
@@
 
-Another advantage: multiple inheritance.
+\subsection{More On \mixin}
+Besides the benefit of freeing programmers from writing boilerplate code, our
+\mixin annotation can also allow programs to mimic multiple inheritance in a
+restricted form easily.
 

### 2015-09-25 paper/sections/Formalization.tex
More formalization.

@@
 
+\begin{comment}
 \subsubsection{Auxiliary function: \textsf{mtype}}
@@
 { \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
+\end{comment}
 
@@
 
+$\textsf{mbody}(m,C)$ denotes the actual body of method $m$ that interface $C$ owns. It can either be defined originally in $C$ or in its supertypes, and then passed to $C$ via inheritance.
+
+The body of a method $m$ contains the relevant information with respect to that method, like the type of $m$ as well as the modifier.
+Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
+
+\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
+
 \[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  m \notin M}
+{ \textsf{mbody}(m,C) = \emptyset } \]
+
+\[ \inferrule{
+  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  m \in M}
+{ \textsf{mbody}(m,C) = m } \]
+
+\[ \inferrule{
+  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
   \overline{M} \} \\
-  m \notin M}
-{ \textsf{mbody(m,T)} } \]
+  m \notin M
+  }
+{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]
+
+\[ \inferrule{
+  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\
+  m \in M
+  }
+{ \textsf{mbody}(m,C_0) = \textsf{replace}(m, \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C})) } \]
+
+\subsubsection{Auxiliary function: \textsf{subset}}
+
+\[ \textsf{subset}(\overline{C})=\{\ x\in\overline{C}\ |\ \nexists y\in\overline{C},\ y <: x\ \} \]
 

### 2015-09-25 paper/sections/Formalization.tex
Formalism update

@@
 
-\[ \inferrule{
+\[ \inferrule[(ReviewedRule)]{
   IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \overline{M} \} \\
-  m \in M
+  \methods \} \\
+  \methods(\m)=\method
+\\
+\Aux{tops}(\overline{C})=\C_1\ldots\C_n  
   }
-{ \textsf{mbody}(m,C_0) = \textsf{replace}(m, \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C})) } \]
+{ \mBody(\m,\C_0) = \textsf{replace}(\method, \shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
+ } \]
+\subsubsection{Auxiliary function: \Aux{tops}}
 
-\subsubsection{Auxiliary function: \textsf{subset}}
-
-\[ \textsf{subset}(\overline{C})=\{\ x\in\overline{C}\ |\ \nexists y\in\overline{C},\ y <: x\ \} \]
+\[ \textsf{tops}(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \} \]
 
@@
 
-(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
 \begin{equation*}
 \begin{array}{ll}
-\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
-\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
-\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
+\shadow()=\emptyset\\
+\shadow(\methods)=\method &\mif\ \method \in \methods, \forall \method' \in \methods\setminus\method :  \method' \subtype  \method,\\
+\mbox{ otherwise}\\
+\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBodies}(\methods))=\mh\\
+%\conflictError \shadow \_\\
+\_\subtype\emptyset\\
+%\mh\mbox{\Q@\{return \_;\}@}\shadow \mh & \mif\ \mh\shadow\mh'\\
+\_\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\
+
+\T \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
+
+%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
+%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
+%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
 \end{array}
@@
 
-(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
-\begin{equation*}
-\begin{array}{ll}
-\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
-\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
-\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
-\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
-\end{array}
-\end{equation*}
-
-(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error.
+%(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
+%\begin{equation*}
+%\begin{array}{ll}
+%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
+%\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
+%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
+%\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
+%\end{array}
+%\end{equation*}
+%
+%(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error.
 The other direction $body_2.\textsf{returnType}<:body_1.\textsf{returnType}$ follows the same rule.
 It also gives an error if there is no subtyping relationship between two return types.
-\begin{equation*}
-\begin{array}{ll}
-\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
-\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-\end{array}
-\end{equation*}
+%\begin{equation*}
+%\begin{array}{ll}
+%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
+%& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
+%\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
+%& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
+%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
+%\end{array}
+%\end{equation*}
 

### 2015-09-27 paper/sections/Formalization.tex
little note on updating multiple fields
Method typing
Wops, I was wrong last time,
static behave differetly in interfaces w.r.t. abstract classes.
added \Gamma(this) is method typing
Some work for formalization

@@
 
+\footnote{Future work: updating multiple fields in one method call,
+  \texttt{with(T v)}}
+
 \subsection{Typing Rules}
@@
 \quad \textsc{(T-SuperInvk)}
+\]
+
+\subsubsection{Method Typing} 
+\[ 
+\inferrule 
+{ }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-Meth)}
+\]
+
+\[ 
+\inferrule
+{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0}
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} \\\\ \Gamma \vdash \textbf{this}:I }
+\quad \textsc{(T-MethBody)}
+\]
+
+\[ 
+\inferrule
+{IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-MethExt)}
+\]
+
+\[ 
+\inferrule
+{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\ 
+IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} \\\\ \Gamma \vdash \textbf{this}:I }
+\quad \textsc{(T-MethBodyExt)}
 \]
@@
 
-\subsubsection{Auxiliary function: \textsf{mbody}}
+\subsubsection{Functional notation}
+We allows functional notation for set of methods $\methods$, using the method name $\m$ as a key.
+That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
+For convenience, we define $\methods(\m)=\none$ otherwise.
 
-$\textsf{mbody}(m,C)$ denotes the actual body of method $m$ that interface $C$ owns. It can either be defined originally in $C$ or in its supertypes, and then passed to $C$ via inheritance.
+\subsubsection{Auxiliary function:\mBody}
 
-The body of a method $m$ contains the relevant information with respect to that method, like the type of $m$ as well as the modifier.
-Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
+$\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
 
-\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
-
-\[ \inferrule{
-  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  m \notin M}
-{ \textsf{mbody}(m,C) = \emptyset } \]
-
-\[ \inferrule{
-  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  m \in M}
-{ \textsf{mbody}(m,C) = m } \]
-
-\[ \inferrule{
-  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \overline{M} \} \\
-  m \notin M
-  }
-{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]
+The body of a method $\m$ contains all the relevant information with respect to that method, like the type of $\m$ as well as the modifier.
+We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.
+%
+%\[ \inferrule{
+%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+%  m \notin M}
+%{ \textsf{mbody}(m,C) = \emptyset } \]
+%
+%\[ \inferrule{
+%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+%  m \in M}
+%{ \textsf{mbody}(m,C) = m } \]
+%
+%\[ \inferrule{
+%  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+%  \overline{M} \} \\
+%  m \notin M
+%  }
+%{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]
 
@@
   \methods(\m)=\method
-\\
-\Aux{tops}(\overline{C})=\C_1\ldots\C_n  
+%\\
+%\Aux{tops}(\overline{C})=\C_1\ldots\C_n  
   }
-{ \mBody(\m,\C_0) = \textsf{replace}(\method, \shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
+{ \mBody(\m,\C_0) = \textsf{replace}(\method, %\shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
+\shadow(m,\Aux{tops}(\Cs)) 
 } \]
-\subsubsection{Auxiliary function: \Aux{tops}}
+\subsubsection{Auxiliary function: $\tops(\Cs), \shadow(\m,\Cs),\override(\method,\method')$}
 
-\[ \textsf{tops}(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \} \]
-
-\subsubsection{Auxiliary function: \textsf{shadow}}
+\tops{} leave only the ``needed'' superinterfaces, that is, the ones that are not transitivelly reachable by following another superinterface chain.
 
 The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+%exmples to motivate our design
+%interface A{static String m(){return "A";}}
+%interface C extends A{
+%	default String dm(){
+%		  this.m();//wrong in java
+%		  A.m();
+%		  C.m();//wrong in java
+%		}
+%}
+%
+%
+%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
 
-%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
 \begin{equation*}
 \begin{array}{ll}
-\shadow()=\emptyset\\
+\tops(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \} 
+\\
+\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
+& \mBody(\m,\C_i)\in\methods
+\\ & \mif\ \mBody(\m,\C)=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}
+\\ &\mor\ \mBody(\m,\C)=\mh\QM;
+\\
+
+\shadow()=\none\\
 \shadow(\methods)=\method &\mif\ \method \in \methods, \forall \method' \in \methods\setminus\method :  \method' \subtype  \method,\\
 \mbox{ otherwise}\\
-\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBodies}(\methods))=\mh\\
-%\conflictError \shadow \_\\
-\_\subtype\emptyset\\
+\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBody}(\methods))=\mh\\
+\Aux{removeBody}(\QM{default}\ \mh\mbox{\Q@\{return \_;\}@})
+=\mh\QM;\\
+\Aux{removeBody}(\mh\QM)
+=\mh\QM;
+\\
 %\mh\mbox{\Q@\{return \_;\}@}\shadow \mh & \mif\ \mh\shadow\mh'\\
-\_\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\
+\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\
 
-\T \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
-
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
+\override(\method,\none)=
+\override(\none,\method)=\method\\
+\override(\method,\method')
+=
+\override(\method,\mh\QM;) & \method'=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\\
+\\
+\override(\method,\mh')
+=\method &\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}\\
+& \mh\subtype\mh'\\
 %\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
@@
 \end{equation*}
+
+\subsubsection{Derived notations}
+
+ Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
+
+\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]

### 2015-09-27 paper/sections/Formalization.tex
Shadow and the reviewd rule are ready to discuss together

@@
   IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \methods \} \\
-  \methods(\m)=\method
+  \methods \} 
+%\\
+ % \methods(\m)=\method
 %\\
@@
   }
-{ \mBody(\m,\C_0) = \textsf{replace}(\method, %\shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
-\shadow(m,\Aux{tops}(\Cs)) 
+{ \mBody(\m,\C_0) = \override(\methods(\m), %\shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
+\shadow(\m,\Aux{tops}(\Cs)))
 } \]
@@
 
-The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+\shadow{} choose the most specific version of a method, that is the unique version available, or a conflicted version from a set of possibilities.
+We do not model overloading, so it is an error if multiple versions are available with different parameter types.
+
+%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
 %exmples to motivate our design
@@
 \shadow()=\none\\
-\shadow(\methods)=\method &\mif\ \method \in \methods, \forall \method' \in \methods\setminus\method :  \method' \subtype  \method,\\
-\mbox{ otherwise}\\
-\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBody}(\methods))=\mh\\
-\Aux{removeBody}(\QM{default}\ \mh\mbox{\Q@\{return \_;\}@})
-=\mh\QM;\\
-\Aux{removeBody}(\mh\QM)
-=\mh\QM;
+\shadow(\method)=\method\\
+\shadow(\methods)=\conflicted\mh\QM;&\mif\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
+\Aux{mostSpecific}(\methods)=\method &
+\mif\ \method \in \methods, \forall \method' \in \methods :  \method' \subtype  \method,\\
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
+
+\method \subtype
+\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}=
+\method\subtype\mh\QM;
 \\
-%\mh\mbox{\Q@\{return \_;\}@}\shadow \mh & \mif\ \mh\shadow\mh'\\
-\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\
+\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype\method=
+\mh\QM;\subtype\method
+\\
+\override(\none,\none)=\none\\
+\override(\method,\none)=\method\\
 
-\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
-\override(\method,\none)=
-\override(\none,\method)=\method\\
+\override(\none,\method)=\method &\mif\ \method\neq \conflicted\ \mh\QM;\\
 \override(\method,\method')
 =
-\override(\method,\mh\QM;) & \method'=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\\
-\\
+\override(\method,\mh\QM;) & \method'\in\{\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
 \override(\method,\mh')
@@
 
-\subsubsection{Auxiliary function: \textsf{replace}}
-
-The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
-
-\begin{equation*}
-\begin{array}{ll}
-\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
-\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
-\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-\end{array}
-\end{equation*}
+%\subsubsection{Auxiliary function: \textsf{replace}}
+%
+%The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
+%
+%\begin{equation*}
+%\begin{array}{ll}
+%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
+%\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
+%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
+%\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
+%\end{array}
+%\end{equation*}
 

### 2015-09-29 UseMixinAnnoProcessor/src/testHardCases/Test.java
Ok, now shadow do the right thing.
It was easy, if there is a single method, that is the result,
if there are many methods, they must be all abstract, in that case take
the most specific, otherwise is a conflic and take the most specific as
a conflict

@@ -1,0 +1,23 @@
+package testHardCases;
+
+public class Test {
+  public static void main(String[] arg){
+	  A.m();
+	  C.m();
+  }
+}
+
+interface A1{ default C1 m(){return this.m();}}
+interface C1{}
+interface C2 extends C1{}
+interface A2 extends A1{C1 m();}
+interface A0 {C2 m();}
+interface A3 extends A2,A0,A1{}
+/*interface A{static String m(){return "A";}}
+interface C extends A{
+	default String dm(){
+		  this.m();
+		  A.m();
+		  C.m();
+		}
+}*/

### 2015-10-01 paper/sections/Formalization.tex
mostSpecific
Added rule (Obj)
Ok, now shadow do the right thing.
It was easy, if there is a single method, that is the result,
if there are many methods, they must be all abstract, in that case take
the most specific, otherwise is a conflic and take the most specific as
a conflict
updated Context

@@
 \]
-
+\marco{rule tInvk must check is not static?}
 \[
@@
 
+\[
+\inferrule[(Obj)]{
+\forall i\in 1\ldots n\ \Gamma,\Gamma^{\mh_i}\vdash\e_i\in \_\subtype\C^{\mh_i}\\
+\forall\mh\in \mhs \mbox{ such that }
+\mBody(\m^\mh,\C)=\method;\mh\subtype\mh^\method\\
+\forall\m \mbox{ such that }
+\mBody(\m,\C)\in\{\mh\QM;,\conflicted\mh\QM;\} \exists \mh_i \in\mhs\mbox{ such that }\mh_i\subtype\mh\\
+\Gamma=\Gamma',\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C\\
+\mhs=\mh_1\ldots\mh_n
+}{
+\Gamma' \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM;\ldots\T_k\ \f_k\QM;\
+\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
+\ldots
+\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
+\cC
+\in\C
+}
+\]
+
 \subsubsection{Method Typing}
+
+\marco{I can not understand those rules, when is t-meth and t-meth body called? when is that you state that all methods in all interfaces must be ok?}
 \[ 
@@
 \subsubsection{Context}
+\[ \Gamma :: = \overline{x}:\overline{C} \]
 
-\[ \Gamma = \epsilon \mid (x, C) : \Gamma  \]
+The environment $\Gamma$ is a mapping from variables to types, written
+$\overline{x}:\overline{C}$. Typing statement $\Gamma \vdash t:C$ reads ``in the
+environment $\Gamma$, term $t$ has type $C$.'' We abbreviate typing statements on
+sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
+shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
 
@@
 \subsubsection{Auxiliary function: $\tops(\Cs), \shadow(\m,\Cs),\override(\method,\method')$}
+${}_{}$\\*
+\marco{we need to discuss the names tops, shadow and override}
 
@@
 & \mBody(\m,\C_i)\in\methods
-\\ & \mif\ \mBody(\m,\C)=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}
-\\ &\mor\ \mBody(\m,\C)=\mh\QM;
+\\ & \mif\ \mBody(\m,\C)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
+%\\ &\mor\ \mBody(\m,\C)=\mh\QM;
 \\
@@
 \shadow(\method)=\method\\
+\shadow(\methods)=\Aux{mostSpecific}(\methods) &\mif\   \methods=\mh_1\QM;\ldots\mh_n\QM;\\
 \shadow(\methods)=\conflicted\mh\QM;&\mif\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
 \Aux{mostSpecific}(\methods)=\method &
-\mif\ \method \in \methods, \forall \method' \in \methods :  \method' \subtype  \method,\\
+\mif\ \method \in \methods, \forall \method' \in \methods :  \method \subtype
+                                       \method', \\
 \T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
@@
 \end{equation*}
+
+%\text{\yanlin{shouldn't mostSpecific be: $\forall \method' \in \methods : \method \subtype
+%  \method'$ ?}}
 
@@
 \[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
+
+%\marco{we also need to define a function that gives all the methods of an interface, something line}
+%\[
+%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
+%\]

### 2015-10-01 paper/run.py

@@ -1,6 +1,0 @@
-#!/usr/bin/env python
-import subprocess
-
-subprocess.call("ruby computePositions.rb", shell=True)
-subprocess.call("xelatex paper.tex", shell=True)
-subprocess.call("open paper.pdf", shell=True)

### 2015-10-01 paper/run.sh

@@ -1,0 +1,28 @@
+#!/bin/zsh
+
+if [ $1 ]
+then
+    case "$1" in
+    bib )       
+        latex paper.tex
+        bibtex paper     
+    ;;
+    * )
+        echo "Usage (In terminal with zsh):"
+        echo "\t\t./run.sh"
+        echo "render and open a new pdf"
+        echo "\t\t./run.sh bib"
+        echo "render and open a new pdf with bib changes"
+    esac   
+fi
+
+ruby computePositions.rb
+latex paper.tex
+open *.pdf
+
+rm footempfile.txt
+rm *.bbl
+rm *.blg
+rm *.log
+rm *.aux
+rm *-old

### 2015-10-02 UseMixinAnnoProcessor/src/methodshadowing/test1/Test1.java
example code on method shadowing mechanism in Java

@@ -1,0 +1,27 @@
+package methodshadowing.test1;
+
+interface D0 {
+    int m();
+}
+
+interface D1 extends D0 {
+    default int m() {
+        return 1;
+    }
+}
+
+interface D2 extends D0, D1 {
+    default int n() {
+        return m();
+    }
+}
+
+public class Test1 {
+    public static void main(String[] args) {
+        D2 d2 = new D2() {
+
+        };
+
+        System.out.println(d2.m());
+    }
+}

### 2015-10-02 paper/sections/Formalization.tex
example code on method shadowing mechanism in Java
comments on shadow function
type checking of interfaces

@@
 
-\subsubsection{Method Typing}
+\subsubsection{Typechecking of Methods}
 
 \marco{I can not understand those rules, when is t-meth and t-meth body called? when is that you state that all methods in all interfaces must be ok?}
+\begin{comment}
 \[ 
@@
 \]
+\end{comment}
 
@@
 \]
+
+\subsubsection{Typechecking of Interfaces}
+\[ \inferrule
+{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
+ \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq ERROR \\\\
+ \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
+ \textsf{mbody}(m,I) \neq ERROR }
+{I \text{ OK}}
+\quad \textsc{(T-Intf)}
+ \]
+
+Interface $I$ type checks well, if:
+\begin{itemize}
+\item All its super-interfaces $\overline{J}$ OK.
+\item All methods inside interface $I$ are OK.
+\item All methods that $I$ is inheriting from super-interfaces are OK.      
+\end{itemize}
 
@@
 \subsection{Auxiliary Definitions}
-\yanlin{todo: add more auxiliary functions}
-
-\yanlin{todo: add interface context}
 
@@
 %(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+
+\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
+  two conflicting default methods with same signature?? }
+
+\yanlin{for \textbf{override}, what's the result if arguments for override
+  doesn't match any given rule. For example, override(a static method, mh'). It
+  would return UNDEFINED or generate ERROR??}
 

### 2015-10-03 paper/sections/Formalization.tex
Comments.

--- 2015-10-01 paper/sections/Formalization.tex
@@
 
-\subsubsection{Method Typing}
+\subsubsection{Typechecking of Methods}
 
 \marco{I can not understand those rules, when is t-meth and t-meth body called? when is that you state that all methods in all interfaces must be ok?}
+\begin{comment}
 \[
@@
 %(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+
+\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
+  two conflicting default methods with same signature?? }
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=5in]{screenshot.png}
+\caption{Screenshot.}\label{screenshot_png}
+\end{figure}
+
+\haoyuan{I tried to understand the current algorithm, and did more experiments in eclipse.
+Now I borrow some ideas from the current version, and give a new version of the algorithm in text. See below.
+
+(1) I guess the function \textsf{tops} is not necessary. The first step is still
+\[\textsf{mbody}(m,C_i)\in\overline{meth}\textrm{ (excluding \textbf{static} methods)}\]
+
+(2) Assume the context is ``interface $C_0$ extends $\overline{C}$ \{$meth'$;...\}''. First handle
+\[\textsf{override}(meth',\overline{meth}) \eqno{(*)}\]
+
+(3) If $meth'\ne\none$, $(*)$ returns $meth'$ if
+\[\forall meth\in\overline{meth},meth'\subtype meth\]
+even if there are conflicts in $\overline{meth}$.
+
+(4) If $meth'=\none$, we need to figure out
+\[\textsf{mostSpecific}(\overline{meth})\]
+and it should be the one that ``overrides'' all the others in $\overline{meth}$.
+It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces.
+But for abstract methods, only return types are taken into consideration.
+}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=5in]{table.png}
+\caption{Table.}\label{table_png}
+\end{figure}
+
+\haoyuan{
+I found this from the Internet. See Fig.~\ref{screenshot_png}.
+
+An experiment. See Fig.~\ref{table_png}.
+}
 

### 2015-10-04 paper/sections/Formalization.tex
Translation.

@@
 
+\subsection{Translation Rules}
+
+\[ \inferrule{
+  \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{
+  \methods \} \text{ OK }}
+{ \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods \} \longmapsto \\
+  \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods; \textsf{ofMethod}(C_0); \textsf{otherMethod}(C_0) \}
+} \quad \textsc{(Translation)} \]
+
+\haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}
+
 \subsection{Auxiliary Definitions}
@@
 I found this from the Internet. See Fig.~\ref{screenshot_png}.
-
 An experiment. See Fig.~\ref{table_png}.
@@
 
+\subsubsection{Auxiliary function:???}
+
+\begin{equation*}
+\begin{array}{ll}
+\textsf{ofMethod}(C_0) = ... & ... \\
+\textsf{otherMethod}(C_0) = \methods_1; \methods_2; \methods_3 \hspace{0.1in} & \textsf{resolve}(C_0) = \overline{field} \\
+& \method_1 = \textsf{setter}(field) \\
+& \method_2 = \textsf{with}(field) \\
+& \method_3 = \textsf{clone}(field) \\
+\textsf{resolve}(C_0) = \overline{field} & \mif\ \textsf{methods}(C_0) = \mhs \\
+& \forall field^{\mh} \in \overline{field}, \mh \in \mhs, \textsf{isField}(\mBody(\mh, C_0)) = \QM{true} \\
+& \forall \mh \in \mhs, \mh \textsf{ isField or isSetter or isWith or isClone or isDefault} \\
+\textsf{methods}(C_0) = \mhs\ \bigcup\ \overline{\textsf{methods}(C)} & \mif\ \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
+& \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \} \\
+\textsf{isField}(\method) = \QM{true} & \mif\ \method = T\ m();\ T \ne \QM{void} \\
+\textsf{setter}(field) = \none & \mif\ field^\mh = T\ m(),\ field^{setter} = \QM{void} \\
+\textsf{setter}(field) = T_0\ \QM{set}m(T\ x); & \mif\ field^\mh = T\ m(),\ field^{setter} = T_0 \\
+\textsf{with}(field) = \none & \mif\ field^\mh = T\ m(),\ field^{with} = \none \\
+\textsf{with}(field) = T_0\ \QM{with}m(T\ x); & \mif\ field^\mh = T\ m(),\ field^{with} = T_0 \\
+\textsf{clone}(field) = T_0\ \QM{clone}(); & \mif\ field^\mh = T\ m(),\ field^{clone} = T_0
+\end{array}
+\end{equation*}
+
+\haoyuan{Details? ThisReference? Assignment?}
+
 \subsubsection{Derived notations}

### 2015-10-05 paper/sections/Formalization.tex
Recover.
updated rules

--- 2015-10-03 paper/sections/Formalization.tex
@@
 \Gamma \vdash \overline{e} \in \overline{C} \\
-\overline{C} <: \overline{D} \\
-\textsf{checkSuper}(C.super.m(\overline{e}), A, C) \\\\
-\text{\yanlin{check whether A <: C, where A is the enclosing interface of the
-    expression}} }
+\overline{C} <: \overline{D} \\\\
+\textbf{this}:I \\  I <: C }
 {\Gamma \vdash C.\textbf{super}.m(\overline{e}) \in E}
@@
 \inferrule
-{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\
+{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\
 IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
@@
 {T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
-    I} \\\\ \Gamma \vdash \textbf{this}:I }
+    I} }
 \quad \textsc{(T-MethBodyExt)}
 \]
+
+\subsubsection{Typechecking of Interfaces}
+\[ \inferrule
+{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
+ \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq ERROR \\\\
+ \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
+ \textsf{mbody}(m,I) \neq ERROR }
+{I \text{ OK}}
+\quad \textsc{(T-Intf)}
+ \]
+
+Interface $I$ type checks well, if:
+\begin{itemize}
+\item All its super-interfaces $\overline{J}$ OK.
+\item All methods inside interface $I$ are OK.
+\item All methods that $I$ is inheriting from super-interfaces are OK.      
+\end{itemize}
 

### 2015-10-05 paper/sections/CaseStudy.tex
texts.

@@
 
+The \emph{Expression Problem} (EP)~\cite{wadler98expression-problem} is coined
+by Wadler about modular extensibility issues in software evolution and has been
+a hot topic in programming languages since. Today we know of various solutions
+to the EP that either rely on \emph{new programming language
+  features}~\cite{chambers95multimethods,clifton00multiJava,madsen89virtual,nystrom06j,bruce98astatically,McDirmid01jiazzi,garrigue98polymorphic,zenger01extensible,loh06open,wehr11javaGI},
+or can be used as \emph{design patterns}~\cite{gof} in existing
+languages~\cite{torgersen04theexpression,oliveira06extensible,wouter08datatypes,oliveira09modular,oliveira12extensibility}. In
+this paper we show that the EP can be solved with our \mixin annotation in a
+trivial way. We will use a canonical example of implementing arithmetic
+expressions gradually to illustrate the usage.
+
+\paragraph{\textbf{Initial System}}
+The initial system expresses arithmetic expressions with only literals and
+addition structure:
+
+\lstinputlisting[linerange=23-40]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+
+\texttt{Exp} is the common super-interface with an evaluation operation
+\texttt{eval()} inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend
+interface \texttt{Exp} with default implementations for \texttt{eval}
+operation. The number field (\texttt{x}) of a literal is represented as a getter
+method \texttt{x()} and expression fields (\texttt{e1,e2}) of an addition as
+getter methods \texttt{e1()} and \texttt{e2()}.
+
+\paragraph{\textbf{Adding a New Variant}} 
+It is easy to add new data variants to the code in the initial system in a
+modular way. For example, the following code how to add the subtraction variant.
+
+\lstinputlisting[linerange=44-51]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+
+\paragraph{\textbf{Adding a New Operation}}
+Adding new operations to the previous system is still straightforward (although
+requires longer code). The following code shows an example of adding a new
+operation \texttt{print}.
+
+\lstinputlisting[linerange=55-73]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+
+The basic idea is to define interfaces for extending old interfaces.  The
+interface \texttt{ExpP} extending interface \texttt{Exp} is defined with the
+extra function \texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are
+defined with default implementations of operation \texttt{print()}, extending
+base interfaces \texttt{Lit} and \texttt{Add}, respectively. Importantly, note
+that in \texttt{AddP}, the types of ``\emph{fields}'' \texttt{e1,e2} are refined via
+the \emph{return types refinement} of getter methods \texttt{e1()} and
+\texttt{e2()}.
+
 \subsection{Other case studies}
-\bruno{The case studies still need to be implemented!}
-
-Collections example from traits paper?
-(\yanlin{couldn't found source code})
 

### 2015-10-05 paper/sections/Formalization.tex

--- 2015-10-04 paper/sections/Formalization.tex
@@
 {T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
-    I} \\\\ \Gamma \vdash \textbf{this}:I }
+    I} }
 \quad \textsc{(T-MethBodyExt)}
 \]
+
+\subsubsection{Typechecking of Interfaces}
+\[ \inferrule
+{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
+ \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq ERROR \\\\
+ \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
+ \textsf{mbody}(m,I) \neq ERROR }
+{I \text{ OK}}
+\quad \textsc{(T-Intf)}
+ \]
+
+Interface $I$ type checks well, if:
+\begin{itemize}
+\item All its super-interfaces $\overline{J}$ OK.
+\item All methods inside interface $I$ are OK.
+\item All methods that $I$ is inheriting from super-interfaces are OK.
+\end{itemize}
 

### 2015-10-06 paper/sections/Formalization.tex
revised formalization

--- 2015-10-05 paper/sections/Formalization.tex
@@
 \]
-\marco{rule tInvk must check is not static?}
+%\marco{rule tInvk must check is not static?}
 \[
@@
 \[
-\inferrule[(Obj)]{
+\inferrule{
 \forall i\in 1\ldots n\ \Gamma,\Gamma^{\mh_i}\vdash\e_i\in \_\subtype\C^{\mh_i}\\
@@
 }
+\quad \textsc{(T-Obj)}
 \]
 
-\subsubsection{Typechecking of Methods}
+\subsubsection{Method Typing}
 
-\marco{I can not understand those rules, when is t-meth and t-meth body called? when is that you state that all methods in all interfaces must be ok?}
 \begin{comment}
@@
 \inferrule
-{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\
-IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
+{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\\\
+IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\\\
 \forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
@@
 
-\subsubsection{Typechecking of Interfaces}
+\subsubsection{Interface Typing}
 \[ \inferrule
@@
  \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
- \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq ERROR \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq \error \\\\
  \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
- \textsf{mbody}(m,I) \neq ERROR }
+ \textsf{mbody}(m,I) \neq \error }
 {I \text{ OK}}
@@
 subtype relation is given by the \textbf{extends} clauses in \texttt{IT}.
+
+\subsection{Translation Rules}
+
+\[ \inferrule{
+  \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{
+  \methods \} \text{ OK }}
+{ \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods \} \longmapsto \\
+  \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods; \textsf{ofMethod}(C_0); \textsf{otherMethod}(C_0) \}
+} \quad \textsc{(Translation)} \]
+
+\haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}
 
@@
 & \mh\subtype\mh'\\
+\override(\_,\_)=\error & Otherwise
 %\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
@@
 
+\subsubsection{Auxiliary function:???}
+
+\begin{equation*}
+\begin{array}{ll}
+\textsf{ofMethod}(C_0) = ... & ... \\
+\textsf{otherMethod}(C_0) = \methods_1; \methods_2 \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \overline{field} \\
+& \method_1 = \textsf{with}(field, C_0) \\
+& \method_2 = \textsf{clone}(field, C_0) \\
+\textsf{resolve}(C_0) = \overline{field} & \mif\ \textsf{methods}(C_0) = \mhs \\
+& \forall \mh \in \mhs, \mh \textsf{ isField or isWith or isClone or isDefault} \\
+& \forall field \in \overline{field},\ \mh^{field} = T\ m\oR\cR,\ \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM; \\
+\textsf{methods}(C_0) = \mhs\ \bigcup\ \overline{\textsf{methods}(C)} & \mif\ \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
+& \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \} \\
+\textsf{with}(field, C_0) = T_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \mh^{field} = T\ m\oR\cR,\ \mh' = T'\ \QM{with}m\oR T\ x\cR \\
+& \mBody(\QM{with}m, C_0) \in \{ \mh'\QM;, \QM{default}\ \mh' \mbox{\Q@\{return \_;\}@}\} \\
+& IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \},\ T_0 <: T' \\
+\textsf{clone}(field, C_0) = T_0\ \QM{clone}\oR\cR\QM; & \mif\ IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \}
+\end{array}
+\end{equation*}
+
+\haoyuan{``\textsf{isField or isWith or isClone or isDefault}''?}
+
 \subsubsection{Derived notations}

### 2015-10-09 paper/sections/Formalization.tex
Formalization for interface typing complete. Now is all in a single rule

@@
 \[
-\inferrule{
+\inferrule[(T-Obj)]{
 \forall i\in 1\ldots n\ \Gamma,\Gamma^{\mh_i}\vdash\e_i\in \_\subtype\C^{\mh_i}\\
@@
 }
-\quad \textsc{(T-Obj)}
 \]
 
+\subsubsection{Interface and method Typing}
+
+\[ \inferrule[(T-Intf)]{
+IT(\C_0) = \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n \oC\methods\cC\\\\
+ \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
+\ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
+\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)
+ }{
+I \text{ OK}
+}
+ \]
+
+We use $\dom(\C)$ to denote the set of methods that are defined for type $\C$, that is: $\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
+
+\begin{comment}
 \subsubsection{Method Typing}
 
-\begin{comment}
 \[
@@
 \]
-\end{comment}
 
@@
 
-\subsubsection{Interface Typing}
 \[ \inferrule
@@
 \end{itemize}
-
+\end{comment}
 \subsubsection{Context}
@@
 shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
-
+\marco{Something is wrong, do you really want to repeat $\Gamma$? do we ever really need the bars here?}
 \subsubsection{Subtyping}
@@
 That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
-For convenience, we define $\methods(\m)=\none$ otherwise.
-
+For convenience, we define $\methods(\m)=\none$ otherwise;
+moreover $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
 \subsubsection{Auxiliary function:\mBody}

### 2015-10-09 UseMixinLombok/src/marcoTest/MarcoTest.java

@@ -25,3 +25,7 @@ public class MarcoTest {
         System.out.println(p.toS());
     }
 }
+
+ interface A {void m();}
+interface B {void m();}
+interface C extends A, B {}

### 2015-10-15 paper/sections/Overview.tex
revising overview

@@
 \section{Overview}\label{sec:ep}
+The single inheritance support for classes in Java prohibits doing multiple
+inheritance in Java. Instead, the support of multiple inheritance support for
+interfaces makes us to think the possibility of using interfaces only, throwing
+away classes. But immediately we will face the following three problems of using
+interfaces in Java.
+
+\subsection{Problems with Java Interfaces}\label{sec:intf_prob}
+\paragraph{Instantiation Code}. If we use interfaces, then we cannot
+avoid instantiation code. 
+\begin{lstlisting}
+interface P {
+    default m() { ... }
+}
+\end{lstlisting}
+For example, even if interface \texttt{P} is complete (does not contain any
+un-implemented method), to instantiate it programmers still need to create a
+class \texttt{PImpl} that has a constructor for instantiation:
+\begin{lstlisting}
+class PImpl extends P {}
+\end{lstlisting}
+
+\paragraph{Prohibition of Field Members Inside Interfaces}
+Only methods (abstract, static or default) are allowed to appear in the
+definition of interfaces. The use of interfaces prohibits defining field members
+inside. This limitation of interfaces hinders the adoption of using interfaces
+only. Because the need of using field members almost \emph{always} exists. Hence
+it forces programmers to use classes, but this gives up the ability of multiple
+inheritance in Java. So the trade-off is: if you use interfaces, then no field
+members; if you use classes, then no multiple inheritance.
+
+\paragraph{Manual Type Refinement}
+We will introduce operations \texttt{with} and \texttt{clone}, which requires
+return type refinement in extensions. A \texttt{clone} method
+returns a copy of the current object. For example, interfaces \texttt{I} and its
+extension \texttt{I2}:
+\begin{lstlisting}
+interface I { I clone(); }
+interface I2 extends I { 
+  // inherit I clone() from super-interface I
+  // which should be refined to I2 clone();
+  I2 clone();
+}
+\end{lstlisting}
+
+Interface \texttt{I} contains a \texttt{clone()} method which returns a copy of
+current object (of type \texttt{I}). Interface \texttt{I2} is an extension of
+interface \texttt{I}, so inheriting the \texttt{clone()} method from
+\texttt{I}. Since the current interface is \texttt{I2}, now the return type of
+\texttt{clone()} \texttt{should} be refined to \texttt{I2}. However, interfaces
+in Java does not provide this kind of automatic refinement. In order to achieve
+the same effect, programmers have to manually override it by a new
+\texttt{clone()} method with return type \texttt{I2}.
 
@@
 do not support multiple inheritance, easily.
+
+In summary, our approach solves the three problems of interfaces
+(Section~\ref{sec:intf_prob}):
+\begin{itemize}
+\item Instantiation Code: Our approach solves the instantiation problem by
+  providing \texttt{of} methods, mimicking the behavior of constructors.
+\item No Fields Inside Interfaces: In our approach, abstract methods without
+  arguments stands for fields. The return type of the methods are the types of
+  member fields. Getters and setters are also generated automatically.
+\item Manual Type Refinement: The annotation processor is designed to do this
+  type refinement automatically, for \texttt{with-} methods, \texttt{clone}
+  methods, etc.
+\end{itemize}

### 2015-10-15 paper/sections/Formalization.tex

@@
 }{
-\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM;\ldots\T_k\ \f_k\QM;\
-\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
+\QM{new}\ \C\oR\cR\oC\T_1\ \f_1 = e_1\QM;\ldots\T_k\ \f_k = e_k\QM;\
+\mh_1\oC\QM{return}\ \e_{k+1}\QM{;}\!\cC
 \ldots
-\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
+\mh_n\oC\QM{return}\ \e_{k+n}\QM{;}\!\cC
 \cC
@@
 { \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods \} \longmapsto \\
-  \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods; \textsf{ofMethod}(C_0); \textsf{otherMethod}(C_0) \}
+  \emptyset \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods; \textsf{ofMethod}(C_0); \textsf{otherMethod}(C_0) \}
 } \quad \textsc{(Translation)} \]
@@
 \haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}
+
+\textbf{THEOREM. } If $I \longmapsto I'$ and $I$ OK, then $I'$ OK.
 
@@
 
+\begin{comment}
 \begin{figure}[tbp]
@@
 \end{figure}
-
-\haoyuan{
-I found this from the Internet. See Fig.~\ref{screenshot_png}.
-An experiment. See Fig.~\ref{table_png}.
-}
+\end{comment}
 
@@
 \begin{array}{ll}
-\textsf{ofMethod}(C_0) = ... & ... \\
+\textsf{ofMethod}(C_0) = \QM{static of} \oR T_1\ m_1\QM,\ldots T_k\ m_k\cR\ \QM{\{} &  \\
+\ \ \ \ \QM{return new}\ C_0 \oR\cR\ \QM{\{}  & \mif\ \textsf{resolve}(C_0) = \overline{field} \\
+\ \ \ \ \ \ \ \ T_1\ \_m_1 = m_1\QM;\ldots T_k\ \_m_k = m_k\QM; & \forall field_i \in \overline{field},\ \mh^{field_i} = T_i\ m_i\oR\cR \\
+\ \ \ \ \ \ \ \ T_i\ m_i\oR\cR\ \QM{\{return}\ \_m_i\QM{;\}}\ \ldots & T_i\ m_i\oR\cR \textsf{ included for all } i \\
+\ \ \ \ \ \ \ \ T_0\ \QM{with}m_i\oR T_i\ x_i\cR\ \QM{\{} & T_0\ \QM{with}m_i\oR T_i\ x_i\cR \textsf{ included if with}(field_i, C_0) \textsf{ defined} \\
+\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots x_i\QM,\ldots m_k\oR\cR\cR\QM{;} & \\
+\ \ \ \ \ \ \ \ \QM{\}}\ \ldots & \\
+\ \ \ \ \ \ \ \ T_0\ \QM{clone}\oR\cR\ \QM{\{} & \\
+\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots m_k\oR\cR \cR\QM; & \\
+\ \ \ \ \ \ \ \ \QM{\}} & \\
+\ \ \ \ \QM{\};} & \\
+\QM{\}} & \\
 \textsf{otherMethod}(C_0) = \methods_1; \methods_2 \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \overline{field} \\
@@
 & \method_2 = \textsf{clone}(field, C_0) \\
-\textsf{resolve}(C_0) = \overline{field} & \mif\ \textsf{methods}(C_0) = \mhs \\
-& \forall \mh \in \mhs, \mh \textsf{ isField or isWith or isClone or isDefault} \\
-& \forall field \in \overline{field},\ \mh^{field} = T\ m\oR\cR,\ \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM; \\
-\textsf{methods}(C_0) = \mhs\ \bigcup\ \overline{\textsf{methods}(C)} & \mif\ \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
-& \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \} \\
+\textsf{resolve}(C_0) = \overline{field} & \mif\ \textsf{methods}(C_0) = \methods \\
+& \forall \method \in \methods, \method \textsf{ isField or isWith or isClone or isDefault} \\
+& \forall field \in \overline{field},\ \mh^{field} = T\ m\oR\cR \\
+& \ \ \ \ \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM;\ m\ne\QM{clone} \\
+\textsf{methods}(C_0) = \overline{meth'}\ \bigcup\ \overline{\textsf{methods}(C)} & \mif\ \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
+& \method'\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \} \\
 \textsf{with}(field, C_0) = T_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \mh^{field} = T\ m\oR\cR,\ \mh' = T'\ \QM{with}m\oR T\ x\cR \\
@@
 & IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \},\ T_0 <: T' \\
-\textsf{clone}(field, C_0) = T_0\ \QM{clone}\oR\cR\QM; & \mif\ IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \}
+\textsf{clone}(field, C_0) = T_0\ \QM{clone}\oR\cR\QM; & \mif\ IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
+\method \textsf{ isField} & \mif\ \method = T\ m\oR\cR\QM;\ m\ne\QM{clone} \\
+\method \textsf{ isWith} & \mif \\
+\method \textsf{ isClone} & \mif\ \method = T\ \QM{clone}\oR\cR\QM; \\
+\method \textsf{ isDefault} & \mif\ \method = \QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \\
 \end{array}
 \end{equation*}
-
-\haoyuan{``\textsf{isField or isWith or isClone or isDefault}''?}
 

### 2015-10-19 paper/sections/Formalization.tex
Translation.

--- 2015-10-09 paper/sections/Formalization.tex
@@
 
-\subsection{Translation Rules}
+\subsection{Translation Function}
 
-\[ \inferrule{
-  \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{
-  \methods \} \text{ OK }}
-{ \Gamma \vdash \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods \} \longmapsto \\
-  \mixinAnn \text{ interface } C_0 \text{ extends } \overline{C} \{ \methods; \textsf{ofMethod}(C_0); \textsf{otherMethod}(C_0) \}
-} \quad \textsc{(Translation)} \]
+\[ [\![\mixinAnn\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] = \]
+\[ \emptyset\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods;\ \ofMethod(C_0);\ \otherMethod(C_0) \cC \]
 
 \haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}
+
+\textbf{THEOREM. } If $I \longmapsto I'$ and $I$ OK, then $I'$ OK.
 
@@
 
-\subsubsection{Auxiliary function:???}
+\subsubsection{Auxiliary function: $\ofMethod(C_0)$, $\otherMethod(C_0)$}
 
@@
 \begin{array}{ll}
-\textsf{ofMethod}(C_0) = ... & ... \\
-\textsf{otherMethod}(C_0) = \methods_1; \methods_2 \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \overline{field} \\
-& \method_1 = \textsf{with}(field, C_0) \\
-& \method_2 = \textsf{clone}(field, C_0) \\
-\textsf{resolve}(C_0) = \overline{field} & \mif\ \textsf{methods}(C_0) = \mhs \\
-& \forall \mh \in \mhs, \mh \textsf{ isField or isWith or isClone or isDefault} \\
-& \forall field \in \overline{field},\ \mh^{field} = T\ m\oR\cR,\ \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM; \\
-\textsf{methods}(C_0) = \mhs\ \bigcup\ \overline{\textsf{methods}(C)} & \mif\ \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \} \\
-& \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \} \\
-\textsf{with}(field, C_0) = T_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \mh^{field} = T\ m\oR\cR,\ \mh' = T'\ \QM{with}m\oR T\ x\cR \\
-& \mBody(\QM{with}m, C_0) \in \{ \mh'\QM;, \QM{default}\ \mh' \mbox{\Q@\{return \_;\}@}\} \\
-& IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \},\ T_0 <: T' \\
-\textsf{clone}(field, C_0) = T_0\ \QM{clone}\oR\cR\QM; & \mif\ IT(T_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{ \methods \}
+\ofMethod(C_0) = \QM{static}\ C_0\ \QM{of} \oR T_1\ m_1\QM,\ldots T_k\ m_k\cR\ \QM{\{} &  \\
+\ \ \ \ \QM{return new}\ C_0 \oR\cR\ \QM{\{}  & \mif\ \resolve(C_0) = \methods \\
+\ \ \ \ \ \ \ \ T_1\ \_m_1 = m_1\QM;\ldots T_k\ \_m_k = m_k\QM; & \forall \method_i \in \methods, \method_i = T_i\ m_i\oR\cR\QM; \\
+\ \ \ \ \ \ \ \ T_i\ m_i\oR\cR\ \QM{\{return}\ \_m_i\QM{;\}}\ \ldots & T_i\ m_i\oR\cR \text{ included for all } i \\
+\ \ \ \ \ \ \ \ C_0\ \QM{with}m_i\oR T_i\ x_i\cR\ \QM{\{} & C_0\ \QM{with}m_i\oR T_i\ x_i\cR \text{ included if }\with(\method_i, C_0) \text{ defined} \\
+\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots x_i\QM,\ldots m_k\oR\cR\cR\QM{;} & \\
+\ \ \ \ \ \ \ \ \QM{\}}\ \ldots & \\
+\ \ \ \ \ \ \ \ C_0\ \QM{clone}\oR\cR\ \QM{\{} & C_0\ \QM{clone}\oR\cR \text{ included if }\clone(C_0) \text{ defined} \\
+\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots m_k\oR\cR \cR\QM; & \\
+\ \ \ \ \ \ \ \ \QM{\}} & \\
+\ \ \ \ \QM{\};} & \\
+\QM{\}} & \\
+\otherMethod(C_0) = \overline{\with(\method, C_0)}; \clone(C_0) \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \methods \\
+\resolve(C_0) = \{\mBody(m, C_0)\ |\ \mBody(m, C_0)\ \isField\} & \forall m\in\dom(C_0), \text{one case is satisfied:} \\
+& (1)\ \mBody(m, C_0)\ \isField \\
+& (2)\ \mBody(m, C_0)\ \isWith \\
+& (3)\ \mBody(m, C_0)\ \isClone \\
+& (4)\ \mBody(m, C_0)\ \isDefault \\
+\mBody(m, C_0)\ \isField & \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM;\ m\ne\QM{clone} \\
+\mBody(m, C_0)\ \isWith & \mif\ \mh^{\mBody(m, C_0)} = T'\ \QM{with}m' \oR T\ x\cR, C_0 <: T' \\
+& \mand\ \mBody(m', C_0) = T\ m'\oR\cR\QM;\ m'\ne\QM{clone} \\
+\mBody(m, C_0)\ \isClone & \mif\ \mh^{\mBody(m, C_0)} = T\ \QM{clone}\oR\cR, C_0 <: T \\
+\mBody(m, C_0)\ \isDefault & \mif\ \mBody(m, C_0) = \QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \\
+\with(\method, C_0) = C_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \method = T\ m\oR\cR\QM;\ \mBody(\QM{with}m, C_0)\ \isWith \\
+\clone(C_0) = C_0\ \QM{clone}\oR\cR\QM; & \mif\ \mBody(\clone, C_0)\ \isClone \\
 \end{array}
 \end{equation*}
-
-\haoyuan{``\textsf{isField or isWith or isClone or isDefault}''?}
 

### 2015-10-22 paper/sections/Formalization.tex
Formal part reviewd.
Give it a read and see if it make sense to you.

Still some comment missing, but most(half?) of the text about explaining
the formalism is there now.

Still, not touching the translation.

@@
 \section{Formal Semantics}\label{sec:formal}
-\subsection{Grammar}
-
-\marco{Tentative Grammar ClassLessJava}
-
-\begin{figure}[h]
-\begin{grammar}
-\production{
-\e
-}{
-  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
-  }{expressions}\\
-\production{
-\obj
-}{
-\QM{new}\ \C\oR\cR\oC\T_1\ \f_1 = e_1\QM;\ldots\T_k\ \f_k = e_k\QM;\
-\mh_1\oC\QM{return}\ \e_{k+1}\QM{;}\!\cC
-\ldots
-\mh_n\oC\QM{return}\ \e_{k+n}\QM{;}\!\cC
-\cC
-  }{object creation}\\
-\production{
-\metaVar{I}
-}{
- \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
-  }{interface declaration}\\
-\production{
-\method
-}{
- \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
-\mid
-\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
-\mid
-\mh\QM{;}
-  }{method declaration}\\
-\production{
-\mh
-}{
- \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
-  }{method header}\\
-\production{
-\ann
-}{
-  \mixinAnn|\emptyset
-  }{annotations}\\
-\end{grammar}
-\caption{Grammar of ClassLess Java}
-\label{Grammar}
-\end{figure}
-
-In Figure~\ref{Grammar} we show the syntax of ClassLess Java.
-
-To be compatible with java, the concrete syntax for an interface declaration with empty supertype list $\C_1\ldots\C_k$ would also omit the \Q@extends@ keyword.
-
+\marco{Old notes: new text is later}
 Yanlin and Haoyuan
@@
 
-\subsection{Typing Rules}
-\subsubsection{Expression Typing}
+\begin{figure}[h]
+\begin{grammar}
+\production{
+\e
+}{
+  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
+  }{expressions}\\
+\production{
+\obj
+}{
+\QM{new}\ \C\oR\cR\oC\fields\ 
+\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
+\ldots
+\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
+\cC
+  }{object creation}\\
+\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
+\production{
+\metaVar{I}
+}{
+ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
+  }{interface declaration}\\
+\production{
+\method
+}{
+ \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
+\mid
+\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
+\mid
+\mh\QM{;}\quad
+  }{method declaration}\\
+\production{
+\mh
+}{
+ \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
+  }{method header}\\
+\production{
+\ann
+}{
+  \mixinAnn|\emptyset
+  }{annotations}\\
+\production{\Gamma}{
+\x_1{:}\C_1\ldots\x_n{:}\C_n
+}{environment}
+\end{grammar}
+\caption{Grammar of ClassLess Java}
+\label{Grammar}
+\end{figure}
 
-\[
-\inferrule{}{\Gamma \vdash x \in \Gamma(x)} \quad \textsc{(T-Var)}
-\]
+${}_{}$\\
+\begin{figure}[h]
+$
+\!\!\!\!\!\!\!\!\!\!
+\begin{array}{l}
+\inferrule[(T-Invk)]{
+\forall i\in 0..n  \Gamma \vdash \e_i \in \_<:\C_i \\\\
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} 
+ }{
+ \Gamma \vdash e_0.m(\overline{e}) \in \C }
+\quad
+\inferrule[(T-StaticInvk)]{
+\forall i\in 1..n  \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\textsf{mtypeS}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+%\textsf{mmodifier}(\m,\C_0) = \textbf{static} 
+}{
+\Gamma \vdash \C_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
+\quad
+\inferrule[(T-SuperInvk)]{
+\Gamma(\this) <: \C \\\\
+\forall i\in 1..n  \Gamma \vdash \e_i\in \_<:\C_i \\\\
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
+}{\Gamma \vdash \C_0\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
 
-\[
-\inferrule{\Gamma \vdash e_0 \in C_0 \\
-  \textsf{mtype}(m,C_0) = \overline{D} \to E \\
-  \Gamma \vdash \overline{e} \in \overline{C} \\
-  \overline{C} <: \overline{D} }
-{ \Gamma \vdash e_0.m(\overline{e}) \in E }
-\quad \textsc{(T-Invk)}
-\]
-
-\[
-\inferrule{\textsf{mtype}(m,C) = \overline{D} \to E \\
-\Gamma \vdash \overline{e} \in \overline{C} \\
-\overline{C} <: \overline{D} \\
-\textsf{mmodifier}(m,C) = \textbf{static} }
-{\Gamma \vdash C.m(\overline{e}) \in E}
-\quad \textsc{(T-StaticInvk)}
-\]
-%\marco{rule tInvk must check is not static?}
-\[
-\inferrule{\textsf{mtype}(m,C) = \overline{D} \to E \\
-\Gamma \vdash \overline{e} \in \overline{C} \\
-\overline{C} <: \overline{D} \\\\
-\textbf{this}:I \\  I <: C }
-{\Gamma \vdash C.\textbf{super}.m(\overline{e}) \in E}
-\quad \textsc{(T-SuperInvk)}
-\]
-
-\[
+\\[5ex]
 \inferrule[(T-Obj)]{
-\forall i\in 1\ldots n\ \Gamma,\Gamma^{\mh_i}\vdash\e_i\in \_\subtype\C^{\mh_i}\\
-\forall\mh\in \mhs \mbox{ such that }
-\mBody(\m^\mh,\C)=\method;\mh\subtype\mh^\method\\
-\forall\m \mbox{ such that }
-\mBody(\m,\C)\in\{\mh\QM;,\conflicted\mh\QM;\} \exists \mh_i \in\mhs\mbox{ such that }\mh_i\subtype\mh\\
-\Gamma=\Gamma',\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C\\
-\mhs=\mh_1\ldots\mh_n
+\forall i\in 1..n\ 
+%\Gamma_i
+\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
+\vdash\e_i\in \_\subtype\C^{\mh_i}\\\\
+\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
+\forall i\in 1..k\ \Gamma(\x)=\T_i\\\\
+\forall\m\mbox{ such that }
+\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
+%\forall i\in 1\ldots n\ \Gamma_i=\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
 }{
-\Gamma' \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM;\ldots\T_k\ \f_k\QM;\
+\Gamma \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM=\x_1\QM;\ldots\T_k\ \f_k\QM=\x_k\QM;\
 \mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
@@
 }
-\]
-
-\subsubsection{Interface and method Typing}
-
-\[ \inferrule[(T-Intf)]{
+\quad
+\inferrule[(T-Var)]{
+\Gamma(\x)=\C
+}{
+\Gamma \vdash x \in\C}
+\\[5ex]
+ \inferrule[(T-Intf)]{
 IT(\C_0) = \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n \oC\methods\cC\\\\
@@
 }
- \]
+\end{array}$
+\caption{Typing}
+\label{ET}
+\end{figure}
+\subsection{Grammar and typing rules}${}_{}$\\
+
+In Figure~\ref{Grammar} we show the syntax of ClassLess Java.%
+\footnote{To be compatible with java, the concrete syntax for an interface declaration with empty supertype list $\C_1\ldots\C_n$ would also omit the \Q@extends@ keyword.}
+
+We formalize a minimal version of Java 8, focusing on Interfaces, default methods and object creation literals.
+As you can see, we have no syntax for classes.
+Expressions are the conventional variable and method call$\x$ and$\e\QM.\m\QM(\es)$, then we have conventional static method call
+$\C\QM.\m\QM(\es)$; for simplicity we do not consider the degenerate case of calling a static method over the this receiver as it considered a bad/confusing programming practice.
+Next we have the more interesting super call $\C\QM{.super.}\m\QM(\es)$, whose semantic is to call the (non static) method $\m$ over the $\this$ receiver, but statically dispatching to the version of the method defined in the interface $\C$.
+Finally, we consider the object initialization expression from an interface $\C$, where (for simplicity) all the fields are initialized with a variable present in scope.
+Note how our language is exactly a subset of Java 8.
+We allows the only annotation ... \marco{later}
+We consider an globally present Interface Table (\metaVar{IT}) mapping from interface names $\C$ to interface declarations \metaVar{I}.
+
+The environment $\Gamma$ is a mapping from variables to types.
+As tradition, we allows functional notation for $\Gamma$.
+However, to help us defining auxiliary functions, 
+we also allows functional notation set of methods $\methods$, using the method name $\m$ as a key.
+That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
+For convenience, we define $\methods(\m)=\none$ otherwise;
+moreover $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+
+ Typing statement $\Gamma \vdash \e\in\C$ reads ``in the
+environment $\Gamma$, expression $\e$ has type $\C$.'' 
+As a shortcut, we write $\Gamma \vdash \e\in\C<:\C'$
+instead of $\Gamma \vdash \e\in\C$ and $\C<:\C'$.
+ From the interface table, we can read off the subtype relation between interfaces.
+The subtype relation is given by the \textbf{extends} clauses in the interfaces.
+We omit the definition of the usual, traditional subtyping relation between interfaces.%
+\footnote{
+Notice how there are no classes, thus there is no subclassing.
+We believe that this approach may scratch an old itching point in the long struggle of subtyping versus subclassing:
+According to some authors, from a software engineering perspective, interfaces are just a kind of classes.
+Others consider more opportune to  consider interfaces are pure types.
+In this vision our language would have no subclassing.
+We do not know how to conciliate those two viewpoints and ClassLess Java design.
+We do not have Classes purely in the Java sense.
+}
+
+In Figure~\ref{ET} we show the typing rules.
+We use the following auxiliary notation:
+$\Gamma^\mh$ trivially extract the environment from a method header, collect the types and the names of the method parameters.
+$\m^\mh$ and $\C^\mh$ extracts the method name and the return type from a method header.
+$\mBody(\m,\C)$ return the full method declaration as seen by $\C$, that is the method $\m$ can be declared in $\C$ or inherited from another interface.
+$\mBody(\m,\C)$ will be formally defined later.
+$\textsf{mtype}(\m,\C)$ and 
+$\textsf{mtypeS}(\m,\C)$ return the type signature from a method (using $\mBody(\m,\C)$ internally).
+$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
+$\textsf{mtypeS}(\m,\C)$ only on static ones.
 
 We use $\dom(\C)$ to denote the set of methods that are defined for type $\C$, that is: $\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
+
+We discuss first the most interesting rules, that is
+(t-Obj) and  (t-Intf) \marco{to complete}
+\marco{if we have space we can discuss the other rules too, but is not mandatory}
+
+\subsection{Auxiliary Definitions}
+
+\begin{comment}
+\subsubsection{Auxiliary function: \textsf{mtype}}
+- \textsf{mtype(m, C)} : the signature of method m in C.
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \emptyset } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
+\end{comment}
+
+\subsubsection{Auxiliary function:\mBody}
+Defining \mBody is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
+$\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
+
+The body of a method $\m$ contains all the relevant information with respect to that method, like the type of $\m$ as well as the modifier.
+We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.\\*
+
+$\mBody(\m,\C_0) = \override(\methods(\m), 
+\shadow(\m,\Aux{tops}(\Cs)))
+$\\*
+Where 
+$\metaVar{IT}(C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC\methods\cC$
+
+As you can see, we are delegating the work to three others auxiliary functions: $\tops(\Cs), \shadow(\m,\Cs) and \override(\method,\method')$
+${}_{}$\\*
+\marco{we need to discuss the names tops, shadow and override}
+
+\tops{} leave only the ``needed'' superinterfaces, that is, the ones that are not transitivelly reachable by following another superinterface chain. Formally:\\*
+$\tops(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \}$
+
+\shadow{} choose the most specific version of a method, that is the unique version available, or a conflicted version from a set of possibilities.
+We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:\\*
+\!\!\!\!$\begin{array}{ll}
+\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
+& \mBody(\m,\C_i)\in\methods
+\\ & \mif\ \mBody(\m,\C)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
+%\\ &\mor\ \mBody(\m,\C)=\mh\QM;
+\\
+
+\shadow()=\none\\
+\shadow(\method)=\method\\
+\shadow(\methods)=\Aux{mostSpecific}(\methods) &\mif\   \methods=\mh_1\QM;\ldots\mh_n\QM;\\
+\shadow(\methods)=\conflicted\mh\QM;&\mif\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
+\Aux{mostSpecific}(\methods)=\method &
+\mif\ \method \in \methods, \forall \method' \in \methods :  \method \subtype
+                                       \method', \\
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
+
+\method \subtype
+\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}=
+\method\subtype\mh\QM;
+\\
+\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype\method=
+\mh\QM;\subtype\method
+\end{array}$
+Discuss about override\\*
+\!\!\!\!$\begin{array}{ll}
+\override(\none,\none)=\none\\
+\override(\method,\none)=\method\\
+
+\override(\none,\method)=\method &\mif\ \method\neq \conflicted\ \mh\QM;\\
+\override(\method,\method')
+=
+\override(\method,\mh\QM;) & \method'\in\{\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
+\override(\method,\mh')
+=\method &\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}\\
+& \mh\subtype\mh'
+%\override(\_,\_)=\error & Otherwise
+%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
+%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
+%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
+\end{array}$
+
+%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+%exmples to motivate our design
+%interface A{static String m(){return "A";}}
+%interface C extends A{
+%	default String dm(){
+%		  this.m();//wrong in java
+%		  A.m();
+%		  C.m();//wrong in java
+%		}
+%}
+%
+%
+%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+
+\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
+  two conflicting default methods with same signature?? }
+
+\begin{comment}
+%Now it is correct, but may be we do not need it?
+We abbreviate typing statements on
+sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
+shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
+\end{comment}
+
+\marco{The notation mtype get only non static methods, the notation mtypeS get only static ones}
+
+%\subsubsection{Interface and method Typing}
 
@@
 \end{itemize}
-\end{comment}
-\subsubsection{Context}
-\[ \Gamma :: = \overline{x}:\overline{C} \]
-
-The environment $\Gamma$ is a mapping from variables to types, written
-$\overline{x}:\overline{C}$. Typing statement $\Gamma \vdash t:C$ reads ``in the
-environment $\Gamma$, term $t$ has type $C$.'' We abbreviate typing statements on
-sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
-shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
-\marco{Something is wrong, do you really want to repeat $\Gamma$? do we ever really need the bars here?}
 \subsubsection{Subtyping}
@@
 {C_0 <: C_1 \\ ... \\ C_0 <: C_k} \]
-
-\subsubsection{Interface Table}
-The Interface Table (IT) mapping from types \texttt{T} to interface declarations \texttt{L}. From the
-interface table, we can read off the subtype relation between interfaces. The
-subtype relation is given by the \textbf{extends} clauses in \texttt{IT}.
+\end{comment}
+%\subsubsection{Interface Table}
 
@@
 \textbf{THEOREM. } If $I \longmapsto I'$ and $I$ OK, then $I'$ OK.
-
-\subsection{Auxiliary Definitions}
-
-\begin{comment}
-\subsubsection{Auxiliary function: \textsf{mtype}}
-- \textsf{mtype(m, C)} : the signature of method m in C.
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
-{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  m \notin M}
-{ \textsf{mtype(m,T)} = \emptyset } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\
-  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
-{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \overline{M} \} \\
-  m \notin M}
-{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
-\end{comment}
-
-\subsubsection{Functional notation}
-We allows functional notation for set of methods $\methods$, using the method name $\m$ as a key.
-That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
-For convenience, we define $\methods(\m)=\none$ otherwise;
-moreover $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
-\subsubsection{Auxiliary function:\mBody}
-
-$\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
-
-The body of a method $\m$ contains all the relevant information with respect to that method, like the type of $\m$ as well as the modifier.
-We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.
-%
-%\[ \inferrule{
-%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-%  m \notin M}
-%{ \textsf{mbody}(m,C) = \emptyset } \]
-%
-%\[ \inferrule{
-%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-%  m \in M}
-%{ \textsf{mbody}(m,C) = m } \]
-%
-%\[ \inferrule{
-%  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-%  \overline{M} \} \\
-%  m \notin M
-%  }
-%{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]
-
-\[ \inferrule[(ReviewedRule)]{
-  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \methods \}
-%\\
- % \methods(\m)=\method
-%\\
-%\Aux{tops}(\overline{C})=\C_1\ldots\C_n
-  }
-{ \mBody(\m,\C_0) = \override(\methods(\m), %\shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
-\shadow(\m,\Aux{tops}(\Cs)))
-} \]
-\subsubsection{Auxiliary function: $\tops(\Cs), \shadow(\m,\Cs),\override(\method,\method')$}
-${}_{}$\\*
-\marco{we need to discuss the names tops, shadow and override}
-
-\tops{} leave only the ``needed'' superinterfaces, that is, the ones that are not transitivelly reachable by following another superinterface chain.
-
-\shadow{} choose the most specific version of a method, that is the unique version available, or a conflicted version from a set of possibilities.
-We do not model overloading, so it is an error if multiple versions are available with different parameter types.
-
-%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
-%exmples to motivate our design
-%interface A{static String m(){return "A";}}
-%interface C extends A{
-%	default String dm(){
-%		  this.m();//wrong in java
-%		  A.m();
-%		  C.m();//wrong in java
-%		}
-%}
-%
-%
-%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
-
-\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
-  two conflicting default methods with same signature?? }
 
@@
 \end{comment}
-
-\begin{equation*}
-\begin{array}{ll}
-\tops(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \}
-\\
-\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
-& \mBody(\m,\C_i)\in\methods
-\\ & \mif\ \mBody(\m,\C)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
-%\\ &\mor\ \mBody(\m,\C)=\mh\QM;
-\\
-
-\shadow()=\none\\
-\shadow(\method)=\method\\
-\shadow(\methods)=\Aux{mostSpecific}(\methods) &\mif\   \methods=\mh_1\QM;\ldots\mh_n\QM;\\
-\shadow(\methods)=\conflicted\mh\QM;&\mif\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
-\Aux{mostSpecific}(\methods)=\method &
-\mif\ \method \in \methods, \forall \method' \in \methods :  \method \subtype
-                                       \method', \\
-\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
-
-\method \subtype
-\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}=
-\method\subtype\mh\QM;
-\\
-\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype\method=
-\mh\QM;\subtype\method
-\\
-\override(\none,\none)=\none\\
-\override(\method,\none)=\method\\
-
-\override(\none,\method)=\method &\mif\ \method\neq \conflicted\ \mh\QM;\\
-\override(\method,\method')
-=
-\override(\method,\mh\QM;) & \method'\in\{\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
-\override(\method,\mh')
-=\method &\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}\\
-& \mh\subtype\mh'\\
-\override(\_,\_)=\error & Otherwise
-%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
-%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
-%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
-\end{array}
-\end{equation*}
 

### 2015-10-26 paper/sections/Formalization.tex
Little formalization and formatting improvements, also better explanations

@@
 We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:\\*
-\!\!\!\!$\begin{array}{ll}
+\!\!\!\!$\begin{array}{l}
 \shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
-& \mBody(\m,\C_i)\in\methods
-\\ & \mif\ \mBody(\m,\C)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
+\whereNote
+ \mwhere\ \mBody(\m,\C_i)\in\methods
+\ \mif\ \mBody(\m,\C_i)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
 %\\ &\mor\ \mBody(\m,\C)=\mh\QM;
@@
 \shadow(\method)=\method\\
-\shadow(\methods)=\Aux{mostSpecific}(\methods) &\mif\   \methods=\mh_1\QM;\ldots\mh_n\QM;\\
-\shadow(\methods)=\conflicted\mh\QM;&\mif\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
-\Aux{mostSpecific}(\methods)=\method &
-\mif\ \method \in \methods, \forall \method' \in \methods :  \method \subtype
+\shadow(\methods)=\Aux{mostSpecific}(\methods) 
+\whereNote
+\mwhere\   \methods\mbox{ is of the form }\mh_1\QM;\ldots\mh_n\QM;\\
+\shadow(\methods)=\conflicted\mh\QM;
+\whereNote
+\mwhere\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
+\Aux{mostSpecific}(\methods)=\method
+\whereNote
+\mwhere\ \method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype
                                        \method', \\
-\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR
+\whereNote
+ \mwhere\ \T\subtype \T'\\
 
@@
 \end{array}$
-Discuss about override\\*
+${}_{}$\\*
+Where $\Aux{mostSpecific}$ return the most specific method using return type specialization as introduced in Java ??
+We just check the subtype between method headers, so we discard information abut method implementation.\\*
+${}_{}$\\*
+The override function models how the implementation in an interface can override implementation in the superinterface; even in case of a conflict.
+Note how we use the special value $\none$, and how (forth case) overriding can solve a conflict.
+\\*
 \!\!\!\!$\begin{array}{ll}
@@
 
-\override(\none,\method)=\method &\mif\ \method\neq \conflicted\ \mh\QM;\\
+\override(\none,\method)=\method 
+\whereNote
+\mwhere\ \method\mbox{ not of the form } \conflicted\ \mh\QM;\\
 \override(\method,\method')
 =
-\override(\method,\mh\QM;) & \method'\in\{\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
-\override(\method,\mh')
-=\method &\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}\\
-& \mh\subtype\mh'
+\method 
+\whereNote
+\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
+%\override(\method,\mh')
+%=\method 
+%\whereNote
+%\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}, \mh\subtype\mh'
 %\override(\_,\_)=\error & Otherwise

### 2015-10-29 paper/sections/Overview.tex
revised overview

--- 2015-09-25 paper/sections/Overview.tex
@@
 
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=4in]{PegasusDetail.png}
+\caption{Pegasus Example.}\label{fig:pegasus}
+\end{figure}
+
+\subsection{A Running Example: Animals}
+Suppose we want to model an animal system as shown in Fig~\ref{fig:pegasus}. In
+order to model \texttt{Pegasus}, \emph{multiple inheritance} is clearly needed
+in this example. Therefore, in Java we use interfaces instead of classes. An
+\texttt{Animal} has a \texttt{point} field, indicating the current position of
+the animal. \texttt{Horse} and \texttt{Bird} are subtypes of \texttt{Animal},
+with methods \texttt{run()} and \texttt{fly()}, respectively. Note that
+\texttt{Bird} can fly, therefore refines the type of \texttt{point} from
+\texttt{Point2D} to \texttt{Point3D}. Pegasus (one of the best known creatures
+in Greek mythology) can not only \emph{run} but also \emph{fly}! This is the
+place where \emph{``multiple inheritance''} is needed, because \texttt{Pegasus}
+need to obtain \texttt{fly} and \texttt{run} functionality from both
+\texttt{Horse} and \texttt{Bird}.
+
+\subsubsection{Model field members and constructors}
+Firstly, to model \texttt{Point2D} that has x-coordinate and y-coordinate by an
+interface, we immediately run into the problem of expressing the fields
+\texttt{x} and \texttt{y}. Since in Java there is no way to define member fields
+inside interfaces, we propose to simulate fields by abstract methods inside
+interfaces. Method \texttt{withX, withY} creates a new instance of
+\texttt{Point2D} with updated field \texttt{x,y}, respectively.
+
+% Firstly, to model \texttt{Animal} by an interface, we immediately run into the
+% problem of expressing the field \texttt{point}. Since in Java there is no way to
+% define member fields inside interfaces, we propose to simulate fields by
+% abstract methods inside interfaces:
+
+\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+
+\paragraph{Instantiation}
+In Java, to implement an interface like \texttt{Point2D}, a typical and trivial
+approach that programmers usually do is creating a class extending the interface
+and providing implementation for all methods inside. For example, this is the
+implementation for interface \texttt{Point2D}:
+
+\begin{lstlisting}
+class Point2DImpl implements Point2D {
+    private int _x;
+    private int _y;
+    public Point2DImpl(int x, int y) {
+        this._x = x;
+        this._y = y;
+    }
+    public int x() {
+        return _x;
+    }
+    public int y() {
+        return _y;
+    }
+    public Point2D withX(int x) {
+        x(x);
+        return this;
+    }
+    public void x(int x) {
+        _x = x;
+    }
+    public void y(int y) {
+        _y = y;
+    }
+    public Point2D withY(int y) {
+        y(y);
+        return this;
+    }
+}
+\end{lstlisting}
+
+\texttt{Point2DImpl} implements \texttt{Point2D} and provides a constructor with
+quite mechanical code. What's worse, the implementation in \texttt{Point2DImpl}
+may not be reused in a single inheritance language.
+
+\paragraph{Our Approach}
+Instead of writing a whole another class to provide the implementation for
+\texttt{Point2D}, we annotate on interface \texttt{Point2D} directly with
+\mixin. The \mixin annotation will generate a static method \texttt{of} inside
+the interface. The \texttt{of} method mimics the functionality of constructors,
+it takes arguments same as constructors and return objects similar to
+constructors. It makes use of Java anonymous classes and achieves the same
+implementation as \texttt{Point2DImpl}.
+
+With our approach, we provide a Java annotation \mixin to provide default
+implementations for various methods (e.g. getters, setters, with-methods, etc)
+and a mechanism to instantiate objects. \mixin annotation helps programmers to
+write less cumbersome code and instantiate interfaces in Java.
+
+\begin{lstlisting}
+  // inside interface Point2D
+  static Point of(int x, int y) {
+      return new Point() {
+          int _x = x;
+          int _y = y;
+          public int x() {
+            return _x;
+          }
+          public int y() {
+            return _y;
+          }
+          public Point2D withX(int x) {
+            x(x);
+            return this;
+          }
+          public void x(int x) {
+            _x = x;
+          }
+          public void y(int y) {
+            _y = y;
+          }
+          public Point2D withY(int y) {
+            y(y);
+            return this;
+          }
+    }
+  }
+\end{lstlisting}
+
+%\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
+
+\subsubsection{Model concrete methods}
+Now we proceed to define \texttt{Animal} with \texttt{point} ``member
+field''. Again, we model this member field with getter and setter methods:
+\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+
+For interface \texttt{Horse}, concrete implementation of \texttt{run()} method
+need to be defined. Before Java 8, concrete methods (except for static methods)
+are not allowed to appear in interfaces as in classes. But Java 8 introduces
+\emph{default methods}, which allow for implementation insides interfaces. We
+are making use of default methods to implement \texttt{run()}.  Interface also
+\texttt{Horse} shows the usage and advantage of \emph{with} methods by method
+\texttt{run()}: method \texttt{withX} returns a new point object with field
+\texttt{x} updated by the argument to \texttt{withX}. Without these
+\texttt{with} methods, operations like \texttt{run()} would be much harder to
+define.
+
+\subsubsection{Type Refinement}
+\texttt{Bird} needs a 3D position, thus we define \texttt{Point3D} and
+\texttt{Bird} as follows:
+\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=80-92]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+
+Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
+\texttt{int z()} (treated as a getter for member field \texttt{z}). Note that
+the return type of various methods (e.g. with- methods, getters) get refined
+either by covariant method overriding or automatically by our annotation
+processor. Besides \emph{with}, other methods (including \emph{clone},
+\emph{of}) also do type-refinements automatically.
+
+\subsubsection{Multiple Inheritance}
+The \emph{``multiple inheritance''} case appears at interface
+\texttt{Pegasus}. Pegasuses can not only \emph{run} but also \emph{fly}!
+Interface \texttt{Pegasus} obtains \texttt{fly} and \texttt{run} functionality
+through interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation,
+actually these is no code at all programmers have to write. The idea of using
+default methods inside interfaces was proposed in ~\cite{}. It enables us to do
+multiple inheritance, which otherwise is hard to do in Java-like languages that
+do not support multiple inheritance, easily.
+
+\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+
+\begin{comment}
 \subsection{A Running Example: \texttt{Point}}
@@
 \lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
+\end{comment}
 
@@
 
+\begin{comment}
 \subsection{More On \mixin}
@@
 do not support multiple inheritance, easily.
+\end{comment}

### 2015-11-02 paper/sections/Overview.tex
Overview improved.
We need to take a decision about the
 with(T val) methods.

--- 2015-09-25 paper/sections/Overview.tex
@@
+\marco{an improvement of this may become part of intro?}
+
+Before Java 8, concrete methods and static methods where not allowed to appear in interfaces.
+Java 8 allows static interface methods and introduces
+\emph{default methods}, which allow for implementation insides interfaces.
+This had an important positive consequence that was probably overlooked by the Java design team: the concept of class (in java) is now redundant and unneeded.
+We define a subset of Java, called ClassLess Java, where programs and (reusable) libraries can be easily defined and used.
+To avoid for some syntactic boilerplate caused by Java not being originally designed to work without classes, we introduce a new annotation:\mixin  provide default
+implementations for various methods (e.g. getters, setters, with-methods)
+and a mechanism to instantiate objects. \mixin annotation helps programmers to
+write less cumbersome code while coding in ClassLess Java;
+indeed we think the obtained gain is so high that ClassLess Java with \mixin annotation can be less cumbersome than full Java.
+
 \section{Overview}\label{sec:ep}
 
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=4in]{PegasusDetail.png}
+\caption{Pegasus Example.}\label{fig:pegasus}
+\end{figure}
+
+\subsection{A Running Example: Animals}
+\marco{fig1 is good but is very big, not sure if it worth its space. If we keep it, we need to do it with a vectorial graphic, or better inside latex (for example with tikz)}
+To propose a standard example, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
+Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
+Birds can \Q@fly@, thus their location need to be a three dimensional \Q@Point3D@.
+Finally, we model \Q@Pegasus@ as a kind of \Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s.\footnote{
+Some research argues in favour of using subtyping for modelling taxonomies, other research argue against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+
+%Suppose we want to model an animal system as shown in Fig~\ref{fig:pegasus}. In
+%order to model \texttt{Pegasus}, \emph{multiple inheritance} is clearly needed
+%in this example. Therefore, in Java we use interfaces instead of classes. An
+%\texttt{Animal} has a \texttt{point} field, indicating the current position of
+%the animal. \texttt{Horse} and \texttt{Bird} are subtypes of \texttt{Animal},
+%with methods \texttt{run()} and \texttt{fly()}, respectively. Note that
+%\texttt{Bird} can fly, therefore refines the type of \texttt{point} from
+%\texttt{Point2D} to \texttt{Point3D}. Pegasus (one of the best known creatures
+%in Greek mythology) can not only \emph{run} but also \emph{fly}! This is the
+%place where \emph{``multiple inheritance''} is needed, because \texttt{Pegasus}
+%need to obtain \texttt{fly} and \texttt{run} functionality from both
+%\texttt{Horse} and \texttt{Bird}.
+%---------------
+% We can put more in dept discussion while discussing the classes in details.
+%--------------
+\subsubsection{\Q@Point2D@: simple immutable data with fields}
+Since in ClassLess Java there are no classes, our \Q@Point2D@ must be an interface, as for example:
+\begin{lstlisting}
+interface Point2D{ int x(); int y();}
+\end{lstlisting}
+That is, since Java disallows fields inside interfaces,
+ we simulate state by abstract methods.
+However, how can we create an instance of \Q@Point2D@ without using classes? 
+we can just provide an implementation for \Q@x()@ and \Q@y()@ as in
+\begin{lstlisting}
+new Point2D(){ public int x(){return 4;} public int y(){return 2;}}
+\end{lstlisting}
+However we are not required to use this cumbersome syntax\footnote{Available in Java from version ...} for every object allocation.
+As programmes do, for ease or reuse, we can encapsulate boring long repetitive code in a method. A static factory method in this case is appropriate:
+\begin{lstlisting}
+interface Point2D{ int x(); int y();
+  static Point2D of(int x, int y){return new Point2D(){
+    public int x(){return x;} public int y(){return y;}};}
+  }
+\end{lstlisting}
+This obvious constructor code can be automatically generated by our \mixin annotation.
+
+By annotating the interface \Q@Point2D@, the annotation will generate the shown static method \texttt{of}, mimicking the functionality of a simple minded constructor:
+By looking the the methods that need implementation it first detects what are the fields, then 
+generate an \Q@of@ method with one argument for each of them.
+
+That is, we can just write
+\begin{lstlisting}
+@Mixin interface Point2D{ int x(); int y();}
+\end{lstlisting}
+
+More precisely, a field or factory parameter is generated for every no-args method requiring implementation whose name does not have special meaning\footnote{Formal definition of \mixin behaviour is provided later.}.
+An example of code using \Q@Point2D@ is 
+\Q@Point2D.of(42,myPoint.y())@
+where we return a new point, using \Q@42@ as x-coordinate, and 
+taking all the other informations (only \Q@y@ in this case) from another point.
+This pattern is very common when programming with immutable data-structures; it is so common that we decided to support it in our code generation as \Q@with-@ methods, that is:
+\begin{lstlisting}
+@Mixin interface Point2D{ int x(); int y();
+  Point2D withX(int val); 
+  Point2D withY(int val);}
+\end{lstlisting}
+is equivalent to 
+\begin{lstlisting}
+@Mixin interface Point2D{ int x(); int y();
+  default Point2D withX(int val){ return Point2D.of(val,this.y());}
+  default Point2D withY(int val){ return Point2D.of(this.x(),val);}  }
+\end{lstlisting}
+Notice how we can provide implementation for methods in an interface, 
+and our annotation generate code only if the method \emph{needs} implementation. In this way is trivial for the programmer to personalize the behaviour if they have any special need.
+
+%Firstly, to model \texttt{Point2D} that has x-coordinate and y-coordinate by an
+%interface, we immediately run into the problem of expressing the fields
+%\texttt{x} and \texttt{y}.
+%interfaces. Method \texttt{withX, withY} creates a new instance of
+%\texttt{Point2D} with updated field \texttt{x,y}, respectively.
+
+% Firstly, to model \texttt{Animal} by an interface, we immediately run into the
+% problem of expressing the field \texttt{point}. Since in Java there is no way to
+% define member fields inside interfaces, we propose to simulate fields by
+% abstract methods inside interfaces:
+
+%\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+
+%\paragraph{Instantiation}
+%In Java, to implement an interface like \texttt{Point2D}, a typical and trivial
+%approach that programmers usually do is creating a class extending the interface
+%and providing implementation for all methods inside. For example, this is the
+%implementation for interface \texttt{Point2D}:
+
+%\begin{lstlisting}
+%class Point2DImpl implements Point2D {
+%    private int _x;
+%    private int _y;
+%    public Point2DImpl(int x, int y) {
+%        this._x = x;
+%        this._y = y;
+%    }
+%    public int x() {
+%        return _x;
+%    }
+%    public int y() {
+%        return _y;
+%    }
+% %Your implementation of with is wrong
+%    public Point2D withX(int x) {
+%        x(x);
+%        return this;
+%    }
+%    public void x(int x) {
+%        _x = x;
+%    }
+%    public void y(int y) {
+%        _y = y;
+%    }
+%    public Point2D withY(int y) {
+%        y(y);
+%        return this;
+%    }
+%}
+%\end{lstlisting}
+%
+%\texttt{Point2DImpl} implements \texttt{Point2D} and provides a constructor with
+%quite mechanical code. What's worse, the implementation in \texttt{Point2DImpl}
+%may not be reused in a single inheritance language.
+
+%
+%\begin{lstlisting}
+%  // inside interface Point2D
+%  static Point of(int x, int y) {
+%      return new Point() {
+%          int _x = x;
+%          int _y = y;
+%          public int x() {
+%            return _x;
+%          }
+%          public int y() {
+%            return _y;
+%          }
+%          public Point2D withX(int x) {
+%            x(x);
+%            return this;
+%          }
+%          public void x(int x) {
+%            _x = x;
+%          }
+%          public void y(int y) {
+%            _y = y;
+%          }
+%          public Point2D withY(int y) {
+%            y(y);
+%            return this;
+%          }
+%    }
+%  }
+%\end{lstlisting}
+%
+%\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
+
+\subsubsection{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
+\Q@Points2D@ are mathematical entities, thus we chosen to use immutable data structure to model them; however animals are real world entities, and when an animal moves, it is the same animal that have now a different location.
+We model this with mutable state.
+%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
+%field''.
+%%Not again, before we used only getters!
+% Again, we model this member field with getter and setter methods:
+\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\marco{TODO: change point into ``location''}
+Here we declare abstract getter and setter for the mutable ``field'' \Q@location@.
+As you see, we do not apply the \mixin annotation. This is morally equivalent to an abstract class in full Java; that is we do not provide a convenient way to instantiate it.
+
+\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+
+For \texttt{Horse}, concrete implementation of \texttt{run()} method
+need to be defined in a default method, where 
+we also show the convenience of \emph{with} methods.
+% by method
+%\texttt{run()}: method \texttt{withX} returns a new point object with field
+%\texttt{x} updated by the argument to \texttt{withX}. Without these
+%\texttt{with} methods, operations like \texttt{run()} would be much harder to define.
+
+\subsubsection{\Q@Bird@:  field type refinement}
+\Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
+as follows:
+\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+
+Just to give you a feeling on how much boring code \mixin is generating,  we show the correspondent code without then \mixin annotation.\footnote{Can you spot our intentional typo?
+This kind of code is so boring that our brain refuse to process it.
+Programmers should not be required to write this kind of code.}
+\begin{lstlisting}
+interface Point3D extends Point2D{
+  Point3D withX(int val);
+  Point3D withY(int val);
+  Point3D withZ(int val);
+  public static Point3D of(int x, int y, int z){return new Point3D(){
+    public int x(){return x;} 
+    public int y(){return y;}
+    public int z(){return z;}
+    public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
+    public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
+    public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
+    }
+  }
+}
+\end{lstlisting}
+Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
+\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected.
+If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
+Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
+
+\Q@Bird@s are \Q@Animal@s, 
+but while \Q@Animal@s, only need 2D locations,
+\Q@Bird@s need  3D locations.
+In Java if we define an animal class with a field we have a set of unsatisfactory options in front of us:
+\begin{itemize}
+\item Define a \Q@Point3D@ field in \Q@Animal@:
+this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
+\item Define a \Q@Point2D@ field in \Q@Animal@ and...
+\item Define an extra \Q@int z@ field in \Q@Bird@.
+This solution is very at-hoc, requires to basically duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside of \mbox{\Q@Bird@}.
+Again, there are many reasons this would be bad, the most dramatic is that it would not scale to a scenario when the programmer of \Q@Bird@ and the programmer of \Q@Point3D@ are different.
+\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects in the field and cast
+the value out of the \Q@Point2D@ field to \Q@Point3D@ when implementing the overridden getter.
+This solution scales to the multiple programmers approach, but requires ugly casts and can be implemented in a wrong way leading to bugs.
+\end{itemize}
+
+Many here would think that a language extension is needed.
+Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}:
+
+\lstinputlisting[linerange=80-92]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+
+%Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
+%\texttt{int z()} (treated as a getter for member field \texttt{z}). Note that
+%the return type of various methods (e.g. with- methods, getters) get refined
+%either by covariant method overriding or automatically by our annotation
+%processor. Besides \emph{with}, other methods (including \emph{clone},
+%\emph{of}) also do type-refinements automatically.
+
+Note how
+\begin{itemize}
+\item by covariant method overriding we refine the type of our field
+\item by \emph{overloading} we define a new setter with the more precise type
+\item by default method we provide an implementation for the setter called with the old signature.
+\end{itemize}
+The cast is trivially safe since it is guarded by an \Q@instanceof@ test.
+The chain of \Q@with-@ methods is not fully satisfactory: it requires \Q@Bird@ to know of all the fields of \Q@Point2D@. We are planning to extend our \mixin annotation so that 
+we can write \Q@point(point().with(val))@, where the \Q@with(T val)@ method would behave like calling all the \Q@with-@ where a compliant getter exists in \Q@T@.
+\marco{We need to discuss this, we may just implement it? --in this case the text would just become:}
+In Point3D:
+\begin{lstlisting}
+@Mixin interface Point3D extends Point2D{
+  int z(); Point3D withZ(int val);
+  Point3D with(Point val);
+  }
+\end{lstlisting}
+expandend in 
+\begin{lstlisting}
+interface Point3D extends Point2D{
+  ....
+  Point3D with(Point2D val){
+    if(val instanceof Point3D){return (Point3D)val;}
+    return Point3D.of(val.x(),val.y(),this.z());
+    }
+  ...
+  }
+\end{lstlisting}
+and here some discussion about the new with method,
+and then the Bird becomes 
+\begin{lstlisting}
+@Mixin interface Bird extends Animal{
+  Point3D point();
+  void point(Point3D val);
+  default void point(Point2D val){ return point(point().with(val));}
+  default void fly(){...}
+  }
+\end{lstlisting}
+\marco{what do we think?}
+\subsubsection{\Q@Pegasus@ and multiple Inheritance}
+The \emph{``multiple inheritance''} case appears at interface
+\texttt{Pegasus}.
+Pegasus (one of the best known creatures
+in Greek mythology) can not only \emph{run} but also \emph{fly}! 
+It would be handy if they could obtain \texttt{fly} and \texttt{run} functionality from both \texttt{Horse} and \texttt{Bird}.
+
+ Using interfaces and \mixin annotation,
+this can happen transparently.
+
+\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+
+Note how  even the non trivial pattern for field type refinement is transparently composed, and our pegasus have a \Q@Point3D@ \Q@location@.
+Note that this works because \Q@Horse@ do not perform any field type refinement , otherwise we may have to choose/create a common subtype in order for pegasus to exists.
+
+\begin{comment}
 \subsection{A Running Example: \texttt{Point}}

### 2015-11-03 paper/sections/RelatedWork.tex
related work

@@
 \section{Related Work}\label{sec:related}
+
+Multiple Inheritance in Java
+
+default methods and the paper that we found. 
+traits programming in java 8.
+- proposals for extending Java with traits/multiple inheritance/...
+
+Multiple Inheritance
 
@@
 - multiple inheritance
+- ...
+
+ThisType/MyType/Extensibility
+
+We are using covariant-return types to simulate some uses 
+of ThisType/MyType. Only works on positive positions, whereas 
+MyType is more general, as it works on any positions. Nevertheless 
+our approach is still useful (for example for fluent interfaces).
+
 - expression problem
-- ...
+
+Type-Directed Translations/Syntactic Sugar
+
+- SugarJ
+
+We can model certain types of language extensions with annotations 
+only, but those extensions do not introduce new syntax: they 
+merely do automatic code generation.
+
+- Meta-programming techniques?
+
+Annotations/Code Generation/Lombok
+
+What is the difference between our work and existing annotations? 
+What is it that we can do that, those annotations cannot?
+
+Fluent Interfaces
+
+Talk about existing libraries that use fluent interfaces (do any of 
+those use any code generation to assist in the implementation?).

### 2015-11-03 paper/sections/RelatedWork.tex
related work (in progress)

@@
 \section{Related Work}\label{sec:related}
+In this section we discuss related work and comparison to Classless Java.
 
-Multiple Inheritance in Java
+\subsection{Multiple Inheritance in Java}
+Since Java 8 default methods are introduced, which enables concrete method
+implementation to be defined (via the \textbf{default} keyword) inside
+interface. Since in Java supports extension on multiple interfaces (instead of
+classes), the introduction of default methods opens the gate of doing multiple
+inheritance in Java by using interfaces. Details on achieving multiple
+inheritance by mimicking trait programing through interfaces are summarized by
+Bono.et. al.~\cite{bono14}. There is also formalization proposed by
+Goetz and Field~\cite{goetz12fdefenders} to formalize defender (default) methods
+in Java. However this formalization is limited since they only model exactly one
+method inside classes/interfaces, where in our formalization multiple methods
+may be defined.
 
-default methods and the paper that we found. 
-traits programming in java 8.
-- proposals for extending Java with traits/multiple inheritance/...
+There are proposals for extending Java with traits. For example, FeatherTrait
+Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the calculus of
+Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
+trait-based inheritance in Java. \yanlin{Drawbacks to be added...}
 
-Multiple Inheritance
+\paragraph{Multiple Inheritance}
+Traits~\cite{} and mixins~\cite{} provides novel models of new programming
+architecture in object oriented programming paradim. As shown in
+section~\ref{sec:comparison}, although they solved some multiple inheritance
+problems, they are extremely restricting expressiveness.
 
@@
 - mixins (original, scala)
+- c++ solution : virtual inheritance
 - multiple inheritance
@@
 
-ThisType/MyType/Extensibility
+\paragraph{ThisType/MyType/Extensibility}
 
@@
 
-Type-Directed Translations/Syntactic Sugar
+\paragraph{Type-Directed Translations/Syntactic Sugar}
 
@@
 
-- Meta-programming techniques?
+\paragraph{Meta-programming techniques}
+- MetaFJig*
 
@@
 
-Fluent Interfaces
+\paragraph{Fluent Interfaces}
 

### 2015-11-04 paper/sections/RelatedWork.tex
related work on multiple inheritance

@@
 
+\yanlin{need to discuss: whether the comments for FTJ is fair.}
 There are proposals for extending Java with traits. For example, FeatherTrait
@@
 Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
-trait-based inheritance in Java. \yanlin{Drawbacks to be added...}
+trait-based inheritance in Java. However, language extensions (including FTJ)
+have a natural drawback: the programmer has to learn new syntax. In contrast,
+our approach is completely compatible with the current Java language, so that
+programmers don't need to pay any learning cost to adapt to this new classless
+programming pattern. Another drawback which is particular for FTJ is that FTJ
+doesn't have type for traits, hence the correctness check of trait is done when
+type-checking classes. This choice makes the design of FTJ simpler but lost
+typechecking efficiency (one trait will be potentially checked multiple times if
+it is used in multiple classes).
 
-\paragraph{Multiple Inheritance}
-Traits~\cite{} and mixins~\cite{} provides novel models of new programming
-architecture in object oriented programming paradim. As shown in
+\subsection{Multiple Inheritance in General}
+Multiple inheritance is a great feature to have in a programming language, but
+it is difficult to model and implement. These difficulties include the famous
+diamond (fork-join) problem, conflicting methods, yo-yo problem, etc. In order
+to model multiple inheritance, lots of models have been proposed in the past few
+years, including mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
+Both traits and mixins provide novel models of new programming architecture in
+object oriented programming paradigm.
+
+1) Mixins model named components that can be applied to various classes as
+reusable functionality units. However, the well-known limitation of mixins is
+that they suffer from linearization constraints, which hinders the
+expressiveness of mixins and the ability of resolving conflicts.  As shown in
 section~\ref{sec:comparison}, although they solved some multiple inheritance
@@
 
-- traits (original, variations, scala)
-- mixins (original, scala)
-- c++ solution : virtual inheritance
-- multiple inheritance
-- ...
+2) Traits are a novel model to differentiate two functionalities out of classes:
+units of reusable components and object factories. In trait models, traits are
+responsible for units of reusable components, which contain pure methods as
+reusable functionalities; whereas classes act as object factories, which
+requires various functionalities from multiple traits. However drawbacks of
+traits are obvious: traits do not allow state; traits can't be instantiated
+directly.
+
+3) C++ and Scala also try to provide solutions to multiple inheritance, but both
+suffer from object initialization problems. Virtual inheritance in C++ provides
+another solution to multiple inheritance (especially the diamond problem by
+keeping only one copy of the base class)~\cite{ellis1990annotated}, however
+suffers from object initialization problem as pointed out by Malayeri et
+al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
+superclasses, which would potentially cause serious semantic error. Scala
+solution (traits) avoids this problem by disallowing constructor parameters,
+causing no ambiguity in cases such as diamond problem. The drawback of this
+elegant design is limited expressiveness.
 
 \paragraph{ThisType/MyType/Extensibility}
+\emph{This} is a polymorphic type variable.  Typical Type Systems
+forClass-Based Object-Oriented PLs. Class names as types.  Inheritance as
+subtyping Resulting in difficulty in reusing classes with recursive interfaces
+by inheritance Standard (non)solution: downcasts Self types (often called MyType
+[Bruce et al.])  OCaml
+
+Does our approach support binary methods?
 

### 2015-11-04 paper/sections/RelatedWork.tex
related work on MyType

@@
 \paragraph{ThisType/MyType/Extensibility}
-\emph{This} is a polymorphic type variable.  Typical Type Systems
-forClass-Based Object-Oriented PLs. Class names as types.  Inheritance as
-subtyping Resulting in difficulty in reusing classes with recursive interfaces
-by inheritance Standard (non)solution: downcasts Self types (often called MyType
-[Bruce et al.])  OCaml
+Typical object-oriented programming languages (e.g. Java) use class names as
+types and inheritance as subtyping relations. This gives difficulty when
+extending classes/interfaces with recursive types. Problems such as binary
+methods, return type refinement, etc. emerge and hinders software extensibility
+greatly.
 
-Does our approach support binary methods?
+\emph{MyType}~\cite{bruce1994paradigmatic} was proposed by Bruce and Foster,
+which represents the type of \textbf{this} and changes its meaning along with
+inheritance. The addition of MyType to a language will allow easy definition of
+binary methods, methods with recursive types (i.e., the same type of the
+receiver appears in the argument or return positions of methods), etc. MyType
+greatly enhances the expressiveness and extensibility of object-oriented
+programming languages.
 
-We are using covariant-return types to simulate some uses 
-of ThisType/MyType. Only works on positive positions, whereas 
-MyType is more general, as it works on any positions. Nevertheless 
-our approach is still useful (for example for fluent interfaces).
-
-- expression problem
+In our approach, we are using covariant-return types to simulate some uses of
+MyType. But our approach only works on method positive positions, whereas MyType
+is more general, as it works on any positions. Nevertheless our approach is
+still useful for modeling fluent interfaces and solving expression
+problems,etc.
 
@@
 
-- SugarJ
+- SugarJ~\cite{erdweg11sugarj}
 

### 2015-11-05 paper/sections/Formalization.tex
Typos on sect4
Typo2
comments on the formalization

@@
   programs)!}
+
+\bruno{Some notes for discussion regarding formalization:
+\begin{itemize}
+\item Syntax and Rules: 
+We can make it more compact and consistent by use overline 
+notation more, instead of \ldots.
+\item Order of premisses: It may be worthwhile to swap the order of 
+some premisses, so that it follows the order of the explanations.
+\item T-Invk: I think we need to type-check $e_0$ separately. Otherwise 
+$C_0$ shows up magically.
+\item T-Obj: small fixes on 3rd line. Suggestion, create two 
+auxiliary functions: $sigvalid(C,\overline{mh})$ and $alldefined(C,\overline{mh})$
+\item T-INTF: $C_0$ is $I$. Type of this was $C$, modified it to $C_0$.
+In the last line, should it be $dom(\overline{meth})$ or just $\overline{meth}$?
+
+\item tops and shadow: tops seems redundant; 
+we use tops to collect classes, so just that in the next step we iterate 
+over those classes to collect methods that satisfy some property. 
+I think we can merge those 2 steps.
+
+\item I have trouble following the definition of shadow; I think there 
+are errors there.
+
+\end{itemize}
+}
 
@@
 \inferrule[(T-Invk)]{
-\forall i\in 0..n  \Gamma \vdash \e_i \in \_<:\C_i \\\\
+\forall i\in 0..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
@@
  }{
- \Gamma \vdash e_0.m(\overline{e}) \in \C }
+ \Gamma \vdash \e_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C }
 \quad
 \inferrule[(T-StaticInvk)]{
-\forall i\in 1..n  \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\forall i\in 1..n\  \Gamma \vdash \e_i\in \_<:\C_i \\\\
 \textsf{mtypeS}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
@@
 \inferrule[(T-SuperInvk)]{
-\Gamma(\this) <: \C \\\\
-\forall i\in 1..n  \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\Gamma(\this) <: \C_0 \\\\
+\forall i\in 1..n\ \Gamma \vdash \e_i\in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
@@
 \forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
-\forall i\in 1..k\ \Gamma(\x)=\T_i\\\\
+\forall i\in 1..k\ \Gamma(\x_i)\subtype\T_i\\\\
 \forall\m\mbox{ such that }
@@
  \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
-\ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
+\ \Gamma^{\mh},\,\QM{this}{:}\C_0\vdash\e\in \_\subtype\C^{\mh} \\\\
 \dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)
@@
 \whereNote
-\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \}\\
+\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
+\method\subtype\method'
+\\
 %\override(\method,\mh')

### 2015-11-05 paper/sections/Introduction.tex
Notes for the introduction

@@
 \section{Introduction}\label{sec:intro}
+
+\bruno{Notes for the introduction.}
+
+Default methods were introduced in Java; they allow a simple form 
+of multiple-inheritance similar to traits. However, trying to 
+use default methods to simulate multiple inheritance quickly becomes 
+cumbersome
+
+\begin{itemize}
+
+\item interfaces have no constructors; classes are still required to
+create objects, leading to boilerplate code for initialization.
+
+\item interfaces do not have instance fields; using setter and getter 
+methods is a way out of this; but this requires tedious boilerplate 
+classes that latter implement those methods.
+
+\item Useful methods, general purpose methods, such as $clone$, or 
+\emph{fluent} setters~\cite{} can only be defined 
+with access to the internal state, which is not available in 
+an interface. Moreover such methods require special care in the 
+presence of subtypes.
+
+\end{itemize}
+
+This paper proposes \emph{concrete interfaces}: a simple extension 
+to Java interfaces where all interface methods have an implementation. 
+Concrete interfaces have a single constructor that is derived from the 
+interface; they support various common utility methods (such as getters 
+and setters, or clone methods) without boilerplate; and they automatically 
+they with utility methods in the presence of subtyping.
+
+To formalize concrete interfaces, we propose FeatherweightConcrete (FC): 
+a FeatherweightJava-style calculus, which captures the essence 
+of interfaces with default methods. The semantics of concrete interfaces 
+is given as a type-directed translation from FC to FC itself. 
+In the resulting FC code all concrete interfaces are translated into 
+regular FC (and Java) interfaces with default methods. The translation 
+is proved to be type-preserving, ensuring that the tranlation does 
+not introduce new errors. One interesting aspect of the theorem is that 
+if we ignore the concrete annotations, the program still type-checks 
+under the regular typing rules of FC.
+
+We implement concrete interfaces, purely using Java annotations. As such 
+no new syntax is required, and existing Java tools (such as IDE's) work 
+with our implementation.
 

### 2015-11-05 paper/sections/Formalization.tex
Translation refactoring.
Refactor more.
Translation function fixed

@@
 \production{
-\metaVar{I}
+\II
 }{
@@
 
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=5in]{table.png}
+\caption{Table.}\label{table_png}
+\end{figure}
+
 %The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
@@
 %(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
-
-\yanlin{for \textbf{shadow} conflicted mh;, one missing case is when there are
-  two conflicting default methods with same signature?? }
 
@@
 
-\[ [\![\mixinAnn\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] = \]
-\[ \emptyset\ \QM{interface}\ C_0\ \QM{extends}\ \Cs \oC \methods;\ \ofMethod(C_0);\ \otherMethod(C_0) \cC \]
-
-\haoyuan{Not $\{ \methods; ... \}$, actually. Should avoid duplication.}
-
-\textbf{THEOREM. } If $I \longmapsto I'$ and $I$ OK, then $I'$ OK.
+$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =$\\*
+${}_{}$\quad$\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC 
+\methods\ 
+\ofMethod(\C_0) \ 
+\otherMethod(\C_0)\ \cC
+$\\*\tab
+where  $\valid(\C_0)$
+%\item $\overline{T\ m\oR\cR\QM;} = \fieldsFunc(C_0)$
+%\item $p_i = \isWith(\mh^{\mBody(\QM{with#}m_i, C_0)}, C_0)$
+%\item $q = \isClone(\mh^{\mBody(\clone, C_0)}, C_0)$
+%\end{itemize}
 
@@
 }
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=5in]{table.png}
-\caption{Table.}\label{table_png}
-\end{figure}
 \end{comment}
@@
 
-\subsubsection{Auxiliary function: $\ofMethod(C_0)$, $\otherMethod(C_0)$}
+\subsubsection{Auxiliary function: $\valid, \ofMethod$, $\otherMethod$}${}_{}$\\*
 
-\begin{equation*}
-\begin{array}{ll}
-\ofMethod(C_0) = \QM{static}\ C_0\ \QM{of} \oR T_1\ m_1\QM,\ldots T_k\ m_k\cR\ \QM{\{} &  \\
-\ \ \ \ \QM{return new}\ C_0 \oR\cR\ \QM{\{}  & \mif\ \resolve(C_0) = \methods \\
-\ \ \ \ \ \ \ \ T_1\ \_m_1 = m_1\QM;\ldots T_k\ \_m_k = m_k\QM; & \forall \method_i \in \methods, \method_i = T_i\ m_i\oR\cR\QM; \\
-\ \ \ \ \ \ \ \ T_i\ m_i\oR\cR\ \QM{\{return}\ \_m_i\QM{;\}}\ \ldots & T_i\ m_i\oR\cR \text{ included for all } i \\
-\ \ \ \ \ \ \ \ C_0\ \QM{with}m_i\oR T_i\ x_i\cR\ \QM{\{} & C_0\ \QM{with}m_i\oR T_i\ x_i\cR \text{ included if }\with(\method_i, C_0) \text{ defined} \\
-\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots x_i\QM,\ldots m_k\oR\cR\cR\QM{;} & \\
-\ \ \ \ \ \ \ \ \QM{\}}\ \ldots & \\
-\ \ \ \ \ \ \ \ C_0\ \QM{clone}\oR\cR\ \QM{\{} & C_0\ \QM{clone}\oR\cR \text{ included if }\clone(C_0) \text{ defined} \\
-\ \ \ \ \ \ \ \ \ \ \ \ \QM{return this.of}\oR m_1\oR\cR\QM,\ldots m_k\oR\cR \cR\QM; & \\
-\ \ \ \ \ \ \ \ \QM{\}} & \\
-\ \ \ \ \QM{\};} & \\
-\QM{\}} & \\
-\otherMethod(C_0) = \overline{\with(\method, C_0)}; \clone(C_0) \hspace{0.1in} & \mif\ \textsf{resolve}(C_0) = \methods \\
-\resolve(C_0) = \{\mBody(m, C_0)\ |\ \mBody(m, C_0)\ \isField\} & \forall m\in\dom(C_0), \text{one case is satisfied:} \\
-& (1)\ \mBody(m, C_0)\ \isField \\
-& (2)\ \mBody(m, C_0)\ \isWith \\
-& (3)\ \mBody(m, C_0)\ \isClone \\
-& (4)\ \mBody(m, C_0)\ \isDefault \\
-\mBody(m, C_0)\ \isField & \mif\ \mBody(m, C_0) = T\ m\oR\cR\QM;\ m\ne\QM{clone} \\
-\mBody(m, C_0)\ \isWith & \mif\ \mh^{\mBody(m, C_0)} = T'\ \QM{with}m' \oR T\ x\cR, C_0 <: T' \\
-& \mand\ \mBody(m', C_0) = T\ m'\oR\cR\QM;\ m'\ne\QM{clone} \\
-\mBody(m, C_0)\ \isClone & \mif\ \mh^{\mBody(m, C_0)} = T\ \QM{clone}\oR\cR, C_0 <: T \\
-\mBody(m, C_0)\ \isDefault & \mif\ \mBody(m, C_0) = \QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \\
-\with(\method, C_0) = C_0\ \QM{with}m\oR T\ x\cR\QM; & \mif\ \method = T\ m\oR\cR\QM;\ \mBody(\QM{with}m, C_0)\ \isWith \\
-\clone(C_0) = C_0\ \QM{clone}\oR\cR\QM; & \mif\ \mBody(\clone, C_0)\ \isClone \\
-\end{array}
-\end{equation*}
+\noindent$\valid(C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
+$\isField(\method)$,
+$\isWith(\method, \C_0)$
+or $\isClone(\method, \C_0)$.
+%or $\isImplemented(\method)$.
+
+We write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
+$\m$ must be lowercase and is turned in uppercase upon merging.
+For example \QM{with#foo}=\QM{withFoo}.
+Special names $\specialName(\m)$ are \QM{clone},\QM{with} and all the identifiers of form $\QM{with#}\m$.
+
+We formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
+To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that all the methods do not start with underscore, and we prefix method names with underscore to obtain valid field/parameter names.\\*
+\noindent$\begin{array}{l}
+\ofMethod(\C_0) = \
+ \QM{static}\ C_0\ \QM{of} \oR \C_1\ \m_1\QM,\ldots \C_n\ \m_n\cR\
+\QM{\{}
+\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
+\tab\tab\tab\tab\tab\tab\tab C_1\ \QM_\m_1 = \m_1\QM;\ldots \C_n\ \QM_\m_n = \m_n\QM; \\
+\tab\tab\tab\tab\tab\tab\tab
+\C_1\ \m_1\oR\cR\ \QM{\{return _}\m_1\QM{;\}}\ \ldots 
+\C_n\ \m_n\oR\cR\ \QM{\{return _}\m_n\QM{;\}}\\
+\tab\tab\tab\tab\tab\tab\tab\withMethod(\m_1,\C_0,\es_1)\ldots\withMethod(\m_n,\C_0,\es_n)\\
+\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
+\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
+\end{array}$
+\\*
+with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
+$\es_i=\QM{this.}\m_1\QM{(),}\ldots \QM{this.}\m_{1-i}\QM{(),}\QM{_val,}\QM{this.}\m_{i+1}\QM{(),}\ldots \QM{this.}\m_n\oR\cR$\\*
+and $\es=\QM{this.}\m_1\QM{(),}\ldots\ldots \QM{this.}\m_n\oR\cR$\\*
+
+\noindent{}Other auxiliary functions:\\*
+\noindent$\begin{array}{ll}
+\withMethod:&\withMethod(\C,\m,\C_0,\es)=\method=
+\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
+&\mbox{iff }\method<:\mBody(\QM{with#}\m,\C_0)\\
+&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
+\cloneMethod:&\cloneMethod(\C_0,\es)=\method=
+\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+&\mbox{iff }\method<:\mBody(\QM{clone},\C_0)\\
+&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
+
+\otherMethod:& \C_0\ \QM{with#}\m\oR \C_i\ \x_i\cR\QM;\in
+\otherMethod(\C_0) 
+\\&
+ \mbox{iff }
+\C\ \m\QM{();}\in \fieldsFunc(\C_0)\mbox{ and } \QM{with#}\m\notin\dom(\C_0)\\
+&\C_0\ \QM{clone}\oR\cR\QM;\in
+\otherMethod(\C_0)  \mbox{ iff } \isClone( \mBody(\m, \C_0))\\
+
+%other methos was wrong: we need the with to be NOT there to generate it.
+%no use isWith, any method with that name must be absent
+
+\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
+\method\in \dom(\C_0)\ \mand\ \isField(\method)
+\\
+\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
+\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
+\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
+& \mand\ \mnot\ \specialName(\m)\\
+\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
+%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
+%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+\end{array}$
+
+\subsubsection{Results}
+\textbf{THEOREM. } 
+For a given $\II_0\ldots\II_n$ interface table such that
+$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n, \II$ OK.
 

### 2015-11-05 paper/sections/Formalization.tex
Improved formalism and auxiliary functions

@@
 }{
- \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_k \oC\method_1\ldots\method_n\cC
+ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\ \methods\ \cC
   }{interface declaration}\\
@@
 \mid
-\mh\QM{;}\quad
+\mh\QM{;}
   }{method declaration}\\
@@
 \inferrule[(T-Invk)]{
-\forall i\in 0..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
+ \Gamma \vdash \e_0 \in \C_0 \\\\
+\forall i\in 1..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
@@
 \ \Gamma^{\mh},\,\QM{this}{:}\C_0\vdash\e\in \_\subtype\C^{\mh} \\\\
+ \forall \QM{static}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
+\ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
 \dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)
  }{
-I \text{ OK}
+\C_0 \text{ OK}
 }
@@
 
-\subsubsection{Auxiliary function:\mBody}
-Defining \mBody is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
+Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
 $\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
@@
 
-\tops{} leave only the ``needed'' superinterfaces, that is, the ones that are not transitivelly reachable by following another superinterface chain. Formally:\\*
-$\tops(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \}$
+\tops{} recover from the interface table only the ``needed'' methods, that is, 
+the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:\\*
+$\method\in\tops(\m,\Cs) \mbox{ iff }\C\in\Cs\ ,\method=\mBody(\m,\C), \method\mbox{ not a static method}$\\*
+${}_{}$\tab and $\nexists \C'\in\Cs\setminus\C$ such that $\C' \subtype \C$.
 
@@
 \!\!\!\!$\begin{array}{l}
-\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
-\whereNote
- \mwhere\ \mBody(\m,\C_i)\in\methods
-\ \mif\ \mBody(\m,\C_i)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
+%\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
+%\whereNote
+% \mwhere\ \mBody(\m,\C_i)\in\methods
+%\ \mif\ \mBody(\m,\C_i)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
 %\\ &\mor\ \mBody(\m,\C)=\mh\QM;
-\\
+%\\
 
@@
 \shadow(\method)=\method\\
-\shadow(\methods)=\Aux{mostSpecific}(\methods)
-\whereNote
-\mwhere\   \methods\mbox{ is of the form }\mh_1\QM;\ldots\mh_n\QM;\\
+\shadow(\overline{\mh\QM;})=\Aux{mostSpecific}(\overline{\mh\QM;})\\
 \shadow(\methods)=\conflicted\mh\QM;
 \whereNote
-\mwhere\ \Aux{mostSpecific}(\methods))\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
+\mwhere\ 
+\methods\mbox{ not of the form }\overline{\mh\QM;}\mbox{ and }
+\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
 \Aux{mostSpecific}(\methods)=\method

### 2015-11-06 paper/paper.tex
revised overview, with subsection structure
type checking of interfaces
Comments.
Going forward: formalization

@@
     super,this,throw,trait,true,try,%
-    type,var,while,with,yield},
+    type,var,while,yield},
   otherkeywords={=>,<-,<\%,<:,>:,\#,@},
@@
 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
-\newcommand\yanlin[1]{\authornote{yanlin}{pink}{#1}}
+\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
 \newcommand{\hl}[1]{\textcolor{red}{#1}}
 \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
+\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
 
@@
 
-\newcommand\mixin{\textbf{@Mixin}}
+\newcommand\mixin{\textbf{@Mixin} }
 \usepackage{xspace}

### 2015-11-06 paper/sections/Formalization.tex
Going forward: formalization

@@
 \ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
-\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)
+\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
  }{
@@
 Note how our language is exactly a subset of Java 8.
-We allows the only annotation ... \marco{later}
+We allows the only annotation \mixin since is the only one interesting in this article.
 We consider an globally present Interface Table (\metaVar{IT}) mapping from interface names $\C$ to interface declarations \metaVar{I}.
@@
 We discuss first the most interesting rules, that is
-(t-Obj) and  (t-Intf) \marco{to complete}
+\rn{(t-Obj)} and  \rn{(t-Intf)} \marco{Bruno, you can explain t-obj, if you can make it understandable, I go for t-inft}
+
+Rule \rn{(t-Obj)}....
+
+Rule \rn{(t-intf)} check that an interface $\C_0$ is correctly typed.
+First we check that the body of all the default and static methods are well typed.
+Then we check that
+$\dom(\C_0)$ is the same of $\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$.
+This is not a trivial check, since $\dom(\C)$ is defined using $\mBody$, that is undefined in many cases: notably if a method $\method\in\methods$ is not compatible with some method in
+$\dom(\C_1)\ldots\dom(\C_n)$ or if any method in both $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$)  is conflicting.
+
 \marco{if we have space we can discuss the other rules too, but is not mandatory}
@@
 
-\subsection{Translation Function}
+\subsection{What  \mixin Generates}
+We now show what the \mixin annotation generates. We present a formal definition for
+most of the generated methods; however in our formalism we do not consider setters, so we do not formally define the generation of the setters. 
+We do not include casts or \Q@instanceof@, so we do not include the \Q@with@ method.
 
-$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =$\\*
-${}_{}$\quad$\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC 
-\methods\ 
-\ofMethod(\C_0) \ 
-\otherMethod(\C_0)\ \cC
-$\\*\tab
-where  $\valid(\C_0)$
-%\item $\overline{T\ m\oR\cR\QM;} = \fieldsFunc(C_0)$
-%\item $p_i = \isWith(\mh^{\mBody(\QM{with#}m_i, C_0)}, C_0)$
-%\item $q = \isClone(\mh^{\mBody(\clone, C_0)}, C_0)$
-%\end{itemize}
+\subsubsection{Translation Function}${}_{}$\\*
+
+\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =
+\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC 
+\methods\ \methods' \cC
+$\\*${}_{}$\tab
+where  $\valid(\C_0)$,\Q@of@$\notin\dom(\methods)$ and $\methods'=\ofMethod(\C_0) \ \otherMethod(\C_0,\methods)$.\\*
+
+To translate an annotated interface, we add the \Q@of@ method, and then we add some other methods.
+However, first of all we check if the interface is valid for annotation:
+
+\noindent$\valid(\C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
+$\isField(\method)$,
+$\isWith(\method, \C_0)$
+or $\isClone(\method, \C_0)$.
+
+That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.
+To complete our formal definition we would need to add setters and \Q@with@.
+Moreover, we check that the method \Q@of@ is not already defined by the user.
+In our simplified formalization we consider this to be just an error.
+In our prototype we keep \Q@overloading@ into account, and so we check that an of method with the same signature of the one we would like to generate is not already present.\marco{Do we check it?}
+
+In the following we will write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
+$\m$ must be lower-case and is turned in upper-case upon merging.
+For example \QM{with#foo}=\QM{withFoo}.
+Special names $\specialName(\m)$ are \QM{clone}, \QM{with} and all the identifiers of form $\QM{with#}\m$.
+
+\subsubsection{$\ofMethod$}${}_{}$\\*
+We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
+To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that all the methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
+\noindent$\begin{array}{l}
+\ofMethod(\C_0) = \
+ \QM{static}\ C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
+\QM{\{}
+\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
+\tab\tab\tab\tab\tab\tab\tab C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
+\tab\tab\tab\tab\tab\tab\tab
+\C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots 
+\C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
+\tab\tab\tab\tab\tab\tab\tab\withMethod(\m_1,\C_0,\es_1)\ldots\withMethod(\m_n,\C_0,\es_n)\\
+\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
+\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
+\end{array}$
+\\*
+with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
+$\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
+and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
+
+The function $\fieldsFunc(\C_0)$ (formally defined later) denotes all the fields in the current interface.
+For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
+  \begin{itemize}
+   \item $\m_i$ is the field name, and have type $\C_i$.
+   \item $\m_i$\QM{()} is the getter, that just return the current field value.
+   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
+    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
+   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
+   \item To complete our generation, we need to generate setters, fluent setters and the with method.
+   \item \marco{should we just formalize setters?}
+   \end{itemize}
+
+\subsubsection{Other auxiliary functions}${}_{}$\\*
+
+\noindent$\begin{array}{ll}
+\withMethod:&\withMethod(\C,\m,\C_0,\es)=
+\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
+&\mbox{iff }
+\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;\\
+&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
+\cloneMethod:&\cloneMethod(\C_0,\es)=
+\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+&\mbox{iff }
+\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
+&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
+\end{array}$
+As you can see above, \Q@with-@ and \Q@clone@ methods are generated if needed.
+We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$.
+Note that we do not need to check if its header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
+
+\noindent$\begin{array}{ll}
+\otherMethod:& \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods) 
+\\&
+ \mbox{iff }
+\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isWith(\mBody(\QM{with#}\m, \C_0)) 
+\\&\mbox{ and } \QM{with#}\m\notin\dom(\methods)\\
+&\C_0\ \QM{clone}\oR\cR\QM;\in
+\otherMethod(\C_0)
+\\&  \mbox{ iff } 
+\isClone(\mBody(\QM{clone}, \C_0)) 
+\mbox{ and } \QM{clone}\notin\dom(\methods)\\
+\end{array}$
+
+Other methods that we need to generate in the interface are \Q@with-@ and \Q@clone@.
+A complete formalization would also generate the \Q@with@.
+This is needed only if we need to refine the return type.
+To discover if this is the case, we check if such \Q@with-@ or \Q@clone@ is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.
+
+\noindent$\begin{array}{ll}
+\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
+\method\in \dom(\C_0)\ \mand\ \isField(\method)
+\\
+\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
+\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
+\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
+& \mand\ \mnot\ \specialName(\m)\\
+\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
+%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
+%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+\end{array}$
+
+Our prototype also generate setters; informally
+\begin{itemize}
+\item For methods inside the interface with the form \Q@void @$\m$\QM($\C\ \x$\QM{);}:
+  \begin{itemize}
+    \item Check if exist method $\C\ \m$\Q@();@. If not, generate error (that is, is not $\valid(\C_0)$).
+    \item Generate implemented setter method inside \Q@of@:\\*
+           \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@
+    \end{itemize}
+\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
+  \begin{itemize}
+    \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
+    \item Generate implemented setter method inside \Q@of@:\\*
+           \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
+    \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
+  \end{itemize}
+\end{itemize}
+
+Note how the second kind of setter is what is called a fluent setter~\cite{the lombock thread, something about fluent stuff}.
+This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}.
+
+\subsubsection{Results}
+\textbf{THEOREM. } 
+For a given $\II_0\ldots\II_n$ interface table such that
+$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
+
+To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer
+\marco{should we repeat the 3 points in the email?}
+
+%\subsubsection{Derived notations}
+% Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
+%\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
+
+%\marco{we also need to define a function that gives all the methods of an interface, something line}
+%\[
+%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
+%\]
 
@@
 %\end{equation*}
-
-\subsubsection{Auxiliary function: $\valid, \ofMethod$, $\otherMethod$}${}_{}$\\*
-
-\noindent$\valid(C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
-$\isField(\method)$,
-$\isWith(\method, \C_0)$
-or $\isClone(\method, \C_0)$.
-%or $\isImplemented(\method)$.
-
-We write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
-$\m$ must be lowercase and is turned in uppercase upon merging.
-For example \QM{with#foo}=\QM{withFoo}.
-Special names $\specialName(\m)$ are \QM{clone},\QM{with} and all the identifiers of form $\QM{with#}\m$.
-
-We formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
-To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that all the methods do not start with underscore, and we prefix method names with underscore to obtain valid field/parameter names.\\*
-\noindent$\begin{array}{l}
-\ofMethod(\C_0) = \
- \QM{static}\ C_0\ \QM{of} \oR \C_1\ \m_1\QM,\ldots \C_n\ \m_n\cR\
-\QM{\{}
-\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
-\tab\tab\tab\tab\tab\tab\tab C_1\ \QM_\m_1 = \m_1\QM;\ldots \C_n\ \QM_\m_n = \m_n\QM; \\
-\tab\tab\tab\tab\tab\tab\tab
-\C_1\ \m_1\oR\cR\ \QM{\{return _}\m_1\QM{;\}}\ \ldots 
-\C_n\ \m_n\oR\cR\ \QM{\{return _}\m_n\QM{;\}}\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\m_1,\C_0,\es_1)\ldots\withMethod(\m_n,\C_0,\es_n)\\
-\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
-\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
-\end{array}$
-\\*
-with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
-$\es_i=\QM{this.}\m_1\QM{(),}\ldots \QM{this.}\m_{1-i}\QM{(),}\QM{_val,}\QM{this.}\m_{i+1}\QM{(),}\ldots \QM{this.}\m_n\oR\cR$\\*
-and $\es=\QM{this.}\m_1\QM{(),}\ldots\ldots \QM{this.}\m_n\oR\cR$\\*
-
-\noindent{}Other auxiliary functions:\\*
-\noindent$\begin{array}{ll}
-\withMethod:&\withMethod(\C,\m,\C_0,\es)=\method=
-\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
-&\mbox{iff }\method<:\mBody(\QM{with#}\m,\C_0)\\
-&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\cloneMethod:&\cloneMethod(\C_0,\es)=\method=
-\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
-&\mbox{iff }\method<:\mBody(\QM{clone},\C_0)\\
-&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
-
-\otherMethod:& \C_0\ \QM{with#}\m\oR \C_i\ \x_i\cR\QM;\in
-\otherMethod(\C_0) 
-\\&
- \mbox{iff }
-\C\ \m\QM{();}\in \fieldsFunc(\C_0)\mbox{ and } \QM{with#}\m\notin\dom(\C_0)\\
-&\C_0\ \QM{clone}\oR\cR\QM;\in
-\otherMethod(\C_0)  \mbox{ iff } \isClone( \mBody(\m, \C_0))\\
-
-%other methos was wrong: we need the with to be NOT there to generate it.
-%no use isWith, any method with that name must be absent
-
-\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
-\method\in \dom(\C_0)\ \mand\ \isField(\method)
-\\
-\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
-\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
-\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
-& \mand\ \mnot\ \specialName(\m)\\
-\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
-%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
-%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-\end{array}$
-
-\subsubsection{Results}
-\textbf{THEOREM. } 
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n, \II$ OK.
-
-\subsubsection{Derived notations}
-
- Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
-
-\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
-
-%\marco{we also need to define a function that gives all the methods of an interface, something line}
-%\[
-%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
-%\]

### 2015-11-06 paper/sections/Overview.tex
Going forward: formalization

@@
   Point3D withZ(int val);
+  Point3D with(Point2D val);
   public static Point3D of(int x, int y, int z){return new Point3D(){
@@
     public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
+    public Point3D with(Point2D val){
+      if(val instanceof Point3D){return (Point3D)val;}
+      return Point3D.of(val.x(),val.y(),this.z());
       }
+    };
   }
@@
 Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
+\Q@with-@ methods (functionally) update a field/property at a time.
+This can be inefficient, and sometime hard to maintain.
+Often we want to update many fields at the same time, for example using another object as source.
+Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
+it takes a certain type and in the current object update all field in that type that match fields in the current type.
+The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
+The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test.
+The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.
 
@@
 
-\lstinputlisting[linerange=80-92]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=80-88]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 \end{itemize}
-The cast is trivially safe since it is guarded by an \Q@instanceof@ test.
-The chain of \Q@with-@ methods is not fully satisfactory: it requires \Q@Bird@ to know of all the fields of \Q@Point2D@. We are planning to extend our \mixin annotation so that 
-we can write \Q@point(point().with(val))@, where the \Q@with(T val)@ method would behave like calling all the \Q@with-@ where a compliant getter exists in \Q@T@.
-\marco{We need to discuss this, we may just implement it? --in this case the text would just become:}
-In Point3D:
-\begin{lstlisting}
-@Mixin interface Point3D extends Point2D{
-  int z(); Point3D withZ(int val);
-  Point3D with(Point val);
-  }
-\end{lstlisting}
-expandend in 
-\begin{lstlisting}
-interface Point3D extends Point2D{
-  ....
-  Point3D with(Point2D val){
-    if(val instanceof Point3D){return (Point3D)val;}
-    return Point3D.of(val.x(),val.y(),this.z());
-    }
-  ...
-  }
-\end{lstlisting}
-and here some discussion about the new with method,
-and then the Bird becomes 
-\begin{lstlisting}
-@Mixin interface Bird extends Animal{
-  Point3D point();
-  void point(Point3D val);
-  default void point(Point2D val){ return point(point().with(val));}
-  default void fly(){...}
-  }
-\end{lstlisting}
-\marco{what do we think?}
+
+From the type perspective, the key is the covariant method overriding; however from the semantic perspective the key is the implementation for the setter with the old signature:
+by using the \Q@with@ method we may use the information for \Q@z@ already stored in the object to forge an appropriate \Q@Point3D@ to store.
+Note how all the informations about what fields sits in \Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the \Q@with@ method, and is transparent for the implementer of \Q@Bird@.
+
 \subsubsection{\Q@Pegasus@ and multiple Inheritance}
@@
 
-\subsection{What  \mixin Generates}
-Apart from generating code for \emph{getter} methods, the annotation processor
-can also generate \emph{setter} methods, \emph{with} methods, \texttt{clone}
-methods, etc. We summarize all the forms of method \mixin  supports below. Inside
-the anonymous class in the annotated interface, the following code are
-generated:
-
-\begin{itemize}
-\item For methods inside the interface with the form \texttt{Tx x()}:
-  \begin{itemize}
-   \item \texttt{x} is the getter method, with return type
-     \texttt{Tx}. Conceptually, it is a member field with name \texttt{x} and
-     type \texttt{Tx}.
-   \item generate member field \texttt{\_x} of type \texttt{Tx}, initialized
-     with \texttt{x}.
-   \item generate implemented getter method:
-       \begin{lstlisting}
-           public Tx x() { return _x; }
-       \end{lstlisting}
-   \end{itemize}
-
-\item For methods inside the interface with the form \texttt{void x(Tx x)}:
-  \begin{itemize}
-    \item check if exist method \texttt{Tx x()}. If not, generate error.
-    \item generate implemented setter method:
-       \begin{lstlisting}
-           public void x(Tx x) { this.x = x; }
-       \end{lstlisting}
-    \end{itemize}
-
-\item For methods with the form \texttt{T withX(Tx \_)}:
-  \begin{itemize}
-  \item if there is no \texttt{x} field, or type \texttt{Tx} does not match,
-    then generate error.
-  \item implement `withX` using the `of` method.
-  \end{itemize}
-
-\item For methods with the form of \texttt{T clone()}: Use \texttt{of} method as
-  the constructor, to create a new object with the same field values as the
-  current one.
-
-\item For methods with the form of \texttt{T x(Tx \_)}:
-  \begin{itemize}
-    \item check if exist method \texttt{T x()}, if not, generate error.
-    \item inside the inner class, generate
-      \begin{lstlisting}
-          public T x(Tx x) { this.x = x; return this;}
-      \end{lstlisting}
-  \end{itemize}
-\end{itemize}
-
 \begin{comment}

### 2015-11-06 UseMixinLombok/src/test/TestAnimal.java

@@ -55,7 +55,7 @@ interface Point2D {
 //BEGIN_POINT3D
 @Mixin
 interface Point3D extends Point2D {
-	int z();
+	int z(); Point3D with(Point val);
 	Point3D withZ(int z);
 }
 //END_POINT3D
@@ -81,16 +81,16 @@ interface Horse extends Animal {
 interface Bird extends Animal {
 	Point3D point();
 	void point(Point3D val);
-	default void point(Point2D val) {
-		if(val instanceof Point3D) { point((Point3D) val); }
-		Point3D newPoint = point().withX(val.x()).withY(val.y());
-		point(newPoint);
-	}
+	default void point(Point2D val) { point(point().with(val));}
 	default void fly() {
 		point(point().withX(point().x() + 40));
 	}
 }
 //END_BIRD
+
+
+
+
 
 //BEGIN_PEGASUS
 @Mixin

### 2015-11-09 paper/sections/Implementation.tex

@@
 \bruno{The implementation does not support separate compilation yet. Can we fix this?}
+
+\haoyuan{http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html}
+
+\haoyuan{https://blog.frankel.ch/lombok-reduces-your-boilerplate-code}
+
+\haoyuan{https://projectlombok.org/}
+
+Our implementation is based on an extension to the Lombok framework[ref].
+Lombok is a Java library aiming at reducing boilerplate code by using code generation via annotations, whereas such code generation is performed at compile time to modify bytecode.
+There are a number of annotations provided by the original Lombok, including \textbf{@Getter}, \textbf{@Setter}, \textbf{@ToString} for generating getters, setters and \QM{toString} methods respectively, and so on.
+And during compile time, these methods are only generated in class files, with neither the source code modified nor new Java files generated.
+
+Furthermore, Lombok provides a number of interfaces for users to create custom transformations, as extensions to the original framework.
+By realizing these transformations, with new annotations supported, users can generate new field and method members, and inject them to current class types automatically at compile time.
+More specifically, a transformation is based on a handler, which acts on the AST from parsing the annotated node and returns a modified AST for analysis and generation afterwards.
+Such a handler can either be a Javac handler or an Eclipse handler.
+At this stage our implementation only realizes the Eclipse handler and our experiments are all conducted in Eclipse.
+
+In Eclipse, with an interface annotated by \mixin, the automatic annotation processing is performed and the information of the interface from compilation is captured in the ``Outline'' window, including all the methods inside the interface as well as the generated ones.
+The custom transformation is easy and convenient to use, however, a certain limitation is that our implementation doesn't support separate compilation yet at this stage, for in the implementation it is hard to capture a type declaration from its reference.

### 2015-11-09 paper/sections/Formalization.tex

@@
 \section{Formal Semantics}\label{sec:formal}
+
+\subsection{Method Overriding Bug in Java}
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=5in]{table.png}
+\caption{Table.}\label{fig:javabug}
+\end{figure}
+
+Figure~\ref{fig:javabug} shows the method overriding mechanism in Java. In the
+example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+named as \texttt{m} which are either abstract or default methods. If we define a
+new interface \texttt{M} that extends two of these interfaces, then the method
+overriding result is shown in the table. For example, row2 col2 means \texttt{M
+  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+following this way of interpretation. The interesting case in row4 col2 where
+\texttt{M extends A1,B3}, the result should be \texttt{ERROR}, however in Java
+the experimental result is \texttt{B3.m}. Because if we think consistently, this
+case is esencially a method conflicting situation with abstract methods and
+default methods, which is no different than the first case.
+
+\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+In our approach, we overcome this bug by detecting method correctness when
+typechecking all interfaces.
+
 \marco{Old notes: new text is later}
@@
 
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=5in]{table.png}
-\caption{Table.}\label{table_png}
-\end{figure}
-
 %The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.

### 2015-11-09 paper/sections/RelatedWork.tex

@@
 
-\paragraph{ThisType/MyType/Extensibility}
+\subsection{ThisType/MyType/Extensibility}
 Typical object-oriented programming languages (e.g. Java) use class names as
@@
 
-\paragraph{Type-Directed Translations/Syntactic Sugar}
+\subsection{Type-Directed Translations/Syntactic Sugar}
+Language extensions are often implemented as syntactic sugar of the base
+language. For example, Scala compiler supports XML syntax in normal Scala code
+directly (after Scala ?, users need to import \texttt{scala-xml} library
+manually). However, this approach is hard in terms of implementation, because it
+requires extending the compiler. Also, this approach does not support combining
+multiple extensions into one.
 
-- SugarJ~\cite{erdweg11sugarj}
+SugarJ~\cite{erdweg11sugarj} is a Java-based extensible programming language
+that allows programmers to extend the base language with custom language
+features by definitions in meta-DSLs (SDF, Stratego, etc). 
+\yanlin{Is new syntax really a ``drawback''? I think for some system, like
+  SugarJ, one of purpose IS to introduce these new syntax.}  Drawbacks: new
+syntax. To create the extension, programmers have to work with multiple
+languages (SDF, stratego, etc) while our approach works totally in Java
+environment.
 
@@
 - MetaFJig*
+\yanlin{Maybe Marco can help with the text for MetaFJig?}
 

### 2015-11-09 paper/sections/Formalization.tex
Setters and clone

@@
 }{
-  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\obj
+  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\x\QM=\e\QM;\e'\mid\obj
   }{expressions}\\
@@
 }{
-\Gamma \vdash x \in\C}
+\Gamma \vdash \x \in\C}
 \\[5ex]
+
+\quad
+\inferrule[(T-update)]{
+\Gamma \vdash \e\in\_<:\Gamma(\x)\\\\
+\Gamma \vdash \e'\in\C
+}{
+\Gamma \vdash \x\QM=\e\QM;\e'\in\C }
+\quad
  \inferrule[(T-Intf)]{
@@
 Next we have the more interesting super call $\C\QM{.super.}\m\QM(\es)$, whose semantic is to call the (non static) method $\m$ over the $\this$ receiver, but statically dispatching to the version of the method defined in the interface $\C$.
+We model field updating as $\x\QM=\e\QM;\e'$, where we expect $\x$ to be a field name.
+After updating the field $\x$ using the value of $\e$, the expression $\e'$ is executed.
+To blend the statement based nature of Java and the expression based nature of our language, we consider a method body of the form \Q@return@ $\x\QM=\e\QM;\e'$
+to represent  $\x\QM=\e\QM;\QM{return}\ \e'$ in Java.
 Finally, we consider the object initialization expression from an interface $\C$, where (for simplicity) all the fields are initialized with a variable present in scope.
@@
 We now show what the \mixin annotation generates. We present a formal definition for
-most of the generated methods; however in our formalism we do not consider setters, so we do not formally define the generation of the setters. 
-We do not include casts or \Q@instanceof@, so we do not include the \Q@with@ method.
+most of the generated methods; however in our formalism we do not consider
+% setters, so we do not formally define the generation of the setters. 
+%We do not include 
+casts or \Q@instanceof@, so we do not include the \Q@with@ method.
+For the same reason we do not include \Q@void@ returning setters, since they are just a minor variation over the more interesting fluent setters, and they would require special handling just for the conventional \Q@void@ type.
 
@@
 $\isWith(\method, \C_0)$
-or $\isClone(\method, \C_0)$.
+or
+$\isSetter(\method,\C_0)$ 
+%$\isClone(\method, \C_0)$.
 
 That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.
-To complete our formal definition we would need to add setters and \Q@with@.
+
+%To complete our formal definition we would need to add setters and \Q@with@.
 Moreover, we check that the method \Q@of@ is not already defined by the user.
@@
 For example \QM{with#foo}=\QM{withFoo}.
-Special names $\specialName(\m)$ are \QM{clone}, \QM{with} and all the identifiers of form $\QM{with#}\m$.
+Special names $\specialName(\m)$ are  \QM{with} and all the identifiers of form $\QM{with#}\m$.
 
@@
 We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
-To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that all the methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
+To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
 \noindent$\begin{array}{l}
@@
 \C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\m_1,\C_0,\es_1)\ldots\withMethod(\m_n,\C_0,\es_n)\\
-\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
+\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
+\tab\tab\tab\tab\tab\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
+%\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
+\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_0)\\
 \tab\tab\tab\tab\tab\tab\QM{\};\}} \\
@@
 with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
-$\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
-and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
+and $\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
+%and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
 
@@
     the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
-   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
-   \item To complete our generation, we need to generate setters, fluent setters and the with method.
-   \item \marco{should we just formalize setters?}
+\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
+%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
+   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
+   %\item \marco{should we just formalize setters?}
    \end{itemize}
@@
 &\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\cloneMethod:&\cloneMethod(\C_0,\es)=
-\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+\setterMethod:&\setterMethod(\C,\m,\C_0)=
+\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+ \m\QM{= _val;return this;\}} \\
 &\mbox{iff }
-\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
-&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
+\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;\\
+&\setterMethod(\C,\m,\C_0)=\emptyset\mbox{ otherwise}\\
+%\cloneMethod:&\cloneMethod(\C_0,\es)=
+%\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+%&\mbox{iff }
+%\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
+%&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
 \end{array}$
-As you can see above, \Q@with-@ and \Q@clone@ methods are generated if needed.
+
+As you can see above, \Q@with-@ and setter methods are generated if needed.
 We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$.
@@
 \\&\mbox{ and } \QM{with#}\m\notin\dom(\methods)\\
-&\C_0\ \QM{clone}\oR\cR\QM;\in
-\otherMethod(\C_0)
-\\&  \mbox{ iff } 
-\isClone(\mBody(\QM{clone}, \C_0)) 
-\mbox{ and } \QM{clone}\notin\dom(\methods)\\
+& \C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods) 
+\\&
+ \mbox{iff }
+\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isSetter(\mBody(\QM_\m, \C_0)) 
+\\&\mbox{ and } \QM_\m\notin\dom(\methods)\\
+%&\C_0\ \QM{clone}\oR\cR\QM;\in
+%\otherMethod(\C_0)
+%\\&  \mbox{ iff } 
+%\isClone(\mBody(\QM{clone}, \C_0)) 
+%\mbox{ and } \QM{clone}\notin\dom(\methods)\\
 \end{array}$
 
-Other methods that we need to generate in the interface are \Q@with-@ and \Q@clone@.
-A complete formalization would also generate the \Q@with@.
+Other methods that we need to generate in the interface are \Q@with-@ and setters. %\Q@clone@.
+%A complete formalization would also generate the \Q@with@.
 This is needed only if we need to refine the return type.
-To discover if this is the case, we check if such \Q@with-@ or \Q@clone@ is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.
+To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
+ is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.
 
@@
 & \mand\ \mnot\ \specialName(\m)\\
-\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
+\isSetter:&\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)
+\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
+& \mand\ \mnot\ \specialName(\m)\\
+
+%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
 %\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
@@
 
-Our prototype also generate setters; informally
+We have not formally modelled non fluent setters and the \Q@with@ method; informally
 \begin{itemize}
@@
            \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@
+    Note how there is no need to refine the return type for non fluent setters, thus we do not need to generate the method header in the interface body itself.
     \end{itemize}
-\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
+\item For methods with the form $\C'\ $\QM{with(}$\C\ \x$\QM{);}:
   \begin{itemize}
-    \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
-    \item Generate implemented setter method inside \Q@of@:\\*
-           \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
-    \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
-  \end{itemize}
+    \item As for before, check that $\C'$ is a supertype of the current interface type $\C_0$.
+    \item Generate implemented \Q@with@ method inside \Q@of@:\\*
+           \Q@public @$\C_0\ $\Q@with(@$\C$\Q@ _val) { @\\*
+           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;@\\*
+           \Q@return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
+where with $\m_1\ldots\m_n$  fields of $\C_0$,
+$\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method; otherwise
+$\e_i=\m_i$.
+    \item If needed, as for \Q@with-@ and setters, generate the method header with refined return type in the interface.
  \end{itemize}
 
-Note how the second kind of setter is what is called a fluent setter~\cite{the lombock thread, something about fluent stuff}.
-This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}.
+%\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
+ % \begin{itemize}
+  %  \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
+   % \item Generate implemented setter method inside \Q@of@:\\*
+    %       \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
+   % \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
+ % \end{itemize}
+\end{itemize}
+
+\marco{ insert somewhere description of  fluent setter~\cite{the lombock thread, something about fluent stuff}.
+This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}}.
 

### 2015-11-10 paper/sections/Implementation.tex
revised Implementation section

@@
 \section{Implementation}
+\begin{comment}
 Haoyuan
@@
 \haoyuan{https://projectlombok.org/}
+\end{comment}
 
-Our implementation is based on an extension to the Lombok framework[ref].
-Lombok is a Java library aiming at reducing boilerplate code by using code generation via annotations, whereas such code generation is performed at compile time to modify bytecode.
-There are a number of annotations provided by the original Lombok, including \textbf{@Getter}, \textbf{@Setter}, \textbf{@ToString} for generating getters, setters and \QM{toString} methods respectively, and so on.
-And during compile time, these methods are only generated in class files, with neither the source code modified nor new Java files generated.
+One of the downside of Java is the code verbosity, e.g., getter and setter
+methods for fields, unneeded type annotation, etc. Lombok project~\cite{} is a
+Java tool that aims at removing (or reducing) these boilerplate code as far as
+possible, via annotations. There are a number of annotations provided by the
+original Lombok, including \textbf{@Getter}, \textbf{@Setter},
+\textbf{@ToString} for generating getters, setters and \QM{toString} methods,
+respectively.
 
-Furthermore, Lombok provides a number of interfaces for users to create custom transformations, as extensions to the original framework.
-By realizing these transformations, with new annotations supported, users can generate new field and method members, and inject them to current class types automatically at compile time.
-More specifically, a transformation is based on a handler, which acts on the AST from parsing the annotated node and returns a modified AST for analysis and generation afterwards.
-Such a handler can either be a Javac handler or an Eclipse handler.
-At this stage our implementation only realizes the Eclipse handler and our experiments are all conducted in Eclipse.
+Furthermore, Lombok provides a number of interfaces for users to create custom
+transformations, as extensions to the original framework. By realizing these
+transformations, with new annotations supported, users can generate new field
+and method members, and inject them to current class types automatically at
+compile time. More specifically, a transformation is based on a handler, which
+acts on the AST from parsing the annotated node and returns a modified AST for
+analysis and generation afterwards. Such a handler can either be a Javac handler
+or an Eclipse handler.
 
-In Eclipse, with an interface annotated by \mixin, the automatic annotation processing is performed and the information of the interface from compilation is captured in the ``Outline'' window, including all the methods inside the interface as well as the generated ones.
-The custom transformation is easy and convenient to use, however, a certain limitation is that our implementation doesn't support separate compilation yet at this stage, for in the implementation it is hard to capture a type declaration from its reference.
+Our implementation is based on extension to Lombok. In Eclipse, with an
+interface annotated by \mixin, the automatic annotation processing is performed
+and the information of the interface from compilation is captured in the
+``Outline'' window, including all the methods inside the interface as well as
+the generated ones. The custom transformation is easy and convenient to use.
+
+The biggest reasons to use Lombok rather than using conventional Java annotation
+processor are:
+\begin{itemize}
+\item Lombok is byte-code based instead of source-code based, which makes client
+  code concise and easy to maintain. Such code generation is performed at
+  compile time to modify bytecode. And during compile time, these methods are
+  only generated in class files, with neither the source code modified nor new
+  Java files generated.
+\item Moreover, and probably more importantly, Lombok is capable of generating
+  code \emph{inside} a class/interface.This is the ability that conventional
+  Java annotation processor can not provide.
+\end{itemize}
+
+The implementation using Lombok has certain limitation: 
+\begin{itemize}
+\item Our implementation does not support separate compilation yet at this
+  stage, for in the implementation it is hard to capture a type declaration from
+  its reference.
+\item At this stage our implementation only realizes the Eclipse handler and our
+  experiments are all conducted in Eclipse.
+\end{itemize}

### 2015-11-10 paper/sections/CaseStudy.tex
Setters and clone
UnitTest.

@@
 
-\lstinputlisting[linerange=23-40]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=25-42]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 
@@
 
-\lstinputlisting[linerange=44-51]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=46-53]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=55-73]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=57-75]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 
@@
 \subsection{Other case studies}
-
-Other case studies using multiple inheritance?
+\marco{should we put here the DB with fluent setters?}

### 2015-11-10 UseMixinLombok/src/UnitTest/AutoTest.java

@@ -1,0 +1,28 @@
+package UnitTest;
+
+import static org.junit.Assert.assertEquals;
+import org.junit.Test;
+import test.TestExpression;
+import test.TestLombok;
+
+public class AutoTest {
+	@Test
+	public void testExpression() {
+		String s = "3" + "\n";
+		s += "7" + "\n";
+		s += "-1" + "\n";
+		s += "3 = 3" + "\n";
+		s += "(3 + 4) = 7" + "\n";
+		assertEquals(s, TestExpression.runTest());
+	}
+	@Test
+	public void testLombok() {
+		String s = "p1 = Point.of(2, 3);" + "\n";
+		s += "p2 = p1.clone();" + "\n";
+		s += "print p1: 2 3" + "\n";
+		s += "p1.x(4); p1.y(5);" + "\n";
+		s += "print p1: 4 5" + "\n";
+		s += "print p2: 2 3" + "\n";
+		assertEquals(s, TestLombok.runTest());
+	}
+}

### 2015-11-10 UseMixinLombok/src/jooq/Database.java

@@ -1,0 +1,32 @@
+package jooq;
+
+import lombok.Mixin;
+
+@Mixin
+interface Database {
+	String select();
+	Database select(String select);
+	String from();
+	Database from(String from);
+	String where();
+	Database where(String where);
+	default Database runTest() {
+		return Database.of("", "", "")
+					   .select("a, b")
+					   .from("Table")
+					   .where("c > 10");
+	}
+}
+
+@Mixin
+interface ExtendedDatabase extends Database {
+	String orderBy();
+	ExtendedDatabase orderBy(String orderBy);
+	default ExtendedDatabase runTest2() {
+		return ExtendedDatabase.of("", "", "", "")
+							   .select("a, b")
+							   .from("Table")
+							   .where("c > 10")
+							   .orderBy("b");
+	}
+}

### 2015-11-10 UseMixinLombok/src/test/TestExpression.java

@@ -3,19 +3,21 @@ package test;
 import lombok.Mixin;
 
 public class TestExpression {
-	public static void main(String[] args) {
+	public static String runTest() {
+		String res = "";
 		Lit e1 = Lit.of(3);
-		System.out.println(e1.eval());
+		res += e1.eval() + "\n"; // 3
 		Lit e2 = Lit.of(4);
 		Add e3 = Add.of(e1, e2);
-		System.out.println(e3.eval());
+		res += e3.eval() + "\n"; // 7
 		Sub e4 = Sub.of(e1, e2);
-		System.out.println(e4.eval());
+		res += e4.eval() + "\n"; // -1
 		LitP e5 = LitP.of(3);
 		LitP e6 = LitP.of(4);
 		AddP e7 = AddP.of(e5, e6);
-		System.out.println(e5.print() + " = " + e5.eval());
-		System.out.println(e7.print() + " = " + e7.eval());
+		res += e5.print() + " = " + e5.eval() + "\n"; // 3 = 3
+		res += e7.print() + " = " + e7.eval() + "\n"; // (3 + 4) = 7
+		return res;
 	}
 }
 

### 2015-11-10 paper/sections/Formalization.tex
revised method overriding mechanism description.

@@
 \section{Formal Semantics}\label{sec:formal}
+
+\subsection{Method Overriding Mechanism}
+We summarize the mechanism for method overriding below (all methods considered
+here are methods with the same name and argument types):
+
+1. Two abstract methods with subtyping relation do not cause conflict, and the
+most specific one wins. Subtyping relation for two methods means they have the
+same arguments' types and one return type is a subtype of the other one. For
+example:
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+Method \texttt{Integer m()} in \texttt{B} wins over method \texttt{Object m()}
+in \texttt{A} because \texttt{Integer <: Object}.
+
+2. A default method conflict with another default method or abstract method. For
+example:
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+
+3. The method in the current interface wins over any methods defined in its
+super-interfaces, provided that the method conform to the subtype of all methods
+in its super-interfaces, i.e., the method is the most specific one. For example:
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { int m(); } //accepted
+\end{lstlisting}
 
@@
 
-Figure~\ref{fig:javabug} shows the method overriding mechanism in Java. In the
+Figure~\ref{fig:javabug} shows the method overriding bug in Java. In the
 example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all

### 2015-11-10 UseMixinLombok/src/methodshadowing/test1/Test1.java

@@ -1,16 +1,30 @@
 package methodshadowing.test1;
 
-interface D0 {
+
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { int m(); } //accepted
+
+interface X0 {
     int m();
 }
 
-interface D1 extends D0 {
+interface X1 extends X0 {
     default int m() {
         return 1;
     }
 }
 
-interface D2 extends D0, D1 {
+interface X2 extends X0, X1 {
     default int n() {
         return m();
     }
@@ -18,7 +32,7 @@ interface D2 extends D0, D1 {
 
 public class Test1 {
     public static void main(String[] args) {
-        D2 d2 = new D2() {
+        X2 d2 = new X2() {
 
         };
 

### 2015-11-10 UseMixinLombok/src/methodshadowing/test5/Test5.java

@@ -6,8 +6,8 @@ interface A1 { int m(); }
 interface B1 { default int m() { return 0; }}
 interface B2 extends B1 { int m(); }
 interface B3 extends B2 { default int m() {return 1;}}
-interface M1 extends A1, B3 {} //correct
-interface M2 extends A1, B1 {} //error
+interface M1 extends A1, B3 {} //accepted, but should be rejected
+interface M2 extends A1, B1 {} //rejected
 //END_JAVABUG
 
 public class Test5 {

### 2015-11-10 paper/sections/Compare.tex
related work (in progress)
Some work on comparision and related work. Section Future work now exists

@@
-\section{Comparing to traits and mixins}
-
+%\section{Comparing to traits and mixins}\label{sec:comparison}
+\subsection{Multiple Inheritance in ClassLess Java}
 \begin{comment}
@@
 
-Our approach is based on code generation by Java annotations.
-The model we generate encourages composability and reusability in object-oriented programming, and is considered to be an alternative to traits or mixins,  meanwhile achieving better performance in some situations.
-Hence it is necessary for us to make a comparison between this approach and traits (or mixins) we commonly used before.
+Our approach is based on forbidding the use of certain language constructs (classes), in order to rely on the modular composition offered by the rest of the language (interfaces).
+Since Java was designed for classes, a direct class-less programming style is verbose and feel unnatural. However, annotation driven code generation is enough to overcome this difficulty, and
+the resulting programming style encourages modularity, composability and reusability, by keeping a strong object oriented feel.
+We consider our approach as an alternative to traits or mixins.
 
-Our approach is quite different from mixins, in the sense that we use the trait model of explicitly resolving conflicts.
-Just as [Scharli2003] demonstrated in the paper, mixin inheritance is a good approach of achieving code reuse, nevertheless, the mixin model is not so expressive to resolve conflicts from many mixins.
-In the trait model, aliases and exclusions are provided for explicit conflict resolution.
-Such operations can actually be modelled in the mechanism of our approach.
-On the other hand, the total ordering property [Scharli2003] restricts the composition of units of behavior.
-The linearization of mixin inheritance cannot provide a perfect total ordering for various requirements, however, in Java, we have multiple inheritance of interfaces, where methods are inherited at the same time from super interfaces.
+We now compare our approach to the original operations on traits, showing how are supported by our model.\footnote{This comparison shares  similarities with the encoding shown in \cite{bono14};
+we differ in the 
+proposed representation of
+ \textbf{exclusion}.
+The author of ~\cite{bono14} agree that our revised version is cleaner, typesafe and more direct.
+}
 
-Here we present how the original operations on traits are supported by our model.
 \begin{itemize}
@@
     \end{lstlisting}
-\item \textbf{Override}: the overriding operation (also known as asymmetric sum) is modelled by implementing many interfaces, while overriding some methods inside.
-The code below gives an example of explicitly specifying which super interface to refer to, regarding two methods with the same name.
+\item \textbf{Override}: the overriding operation (also known as asymmetric/preferential sum) is modelled by implementing many interfaces, while overriding some methods inside.
+The code below gives an example of explicitly specifying which super interface take precedence, regarding conflict on a specific method.
     \begin{lstlisting}
@@
 Such a method is a replacement to the constructors in original traits, making instantiation more convenient to use.
-\item We provide \texttt{with-} and \texttt{clone} methods as auxiliary constructors.
-A \texttt{with-} method is generated for each field, just like record update, returning the new object with that field updated.
-A \texttt{clone} method is generated for the interface, returning a copy of the current object.
-Furthermore, we do automatic return type refinement for these two kind of methods.
-This feature is comparatively useful in big examples, making operations and behaviors more flexible, which we will demonstrate later.
+That is, in our approach there are only interfaces, our model requires a single concept, while the trait model requires traits \emph{and} classes.
+
+\item We provide \texttt{with-} methods as auxiliary constructors. A \texttt{with-} method is generated for each field, just like record update, returning the new object with that field updated
+%. A \texttt{clone} method is generated for the interface, returning a copy of the current object.
+Furthermore, we do automatic return type refinement for these kind of methods.
+This feature is comparatively useful in big examples, making operations and behaviours more flexible.%, which we will demonstrate later.
 \item We provide two options for generating setters.
@@
 These are the additional features supported by our model, conversely, there are certain operations we cannot model, such as method renaming (as in [Reppy2006]), which breaks structural subtyping.
+
+There are other limitations of our current approach, but they may be addressed in future work. See \marco{reference to future work when there}

### 2015-11-10 paper/sections/Overview.tex
Some work on comparision and related work. Section Future work now exists

@@
 
-Just to give you a feeling on how much boring code \mixin is generating,  we show the correspondent code without then \mixin annotation.\footnote{Can you spot our intentional typo?
-This kind of code is so boring that our brain refuse to process it.
-Programmers should not be required to write this kind of code.}
-\begin{lstlisting}
+Just to give you a feeling on how much boring code \mixin is generating,  we show the correspondent code without then \mixin annotation\footnote{
+As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
+swap the arguments of one of the many calls of \Q@Point3D.of@.
+}.\begin{lstlisting}
 interface Point3D extends Point2D{

### 2015-11-10 paper/sections/RelatedWork.tex
Some work on comparision and related work. Section Future work now exists

@@
 
-\subsection{Multiple Inheritance in Java}
-Since Java 8 default methods are introduced, which enables concrete method
-implementation to be defined (via the \textbf{default} keyword) inside
-interface. Since in Java supports extension on multiple interfaces (instead of
-classes), the introduction of default methods opens the gate of doing multiple
-inheritance in Java by using interfaces. Details on achieving multiple
-inheritance by mimicking trait programing through interfaces are summarized by
-Bono.et. al.~\cite{bono14}. There is also formalization proposed by
+\subsection{Formalization of Java8}
+\marco{this go later, when we compare our formalization}
+There is also formalization proposed by
 Goetz and Field~\cite{goetz12fdefenders} to formalize defender (default) methods
@@
 
-\yanlin{need to discuss: whether the comments for FTJ is fair.}
-There are proposals for extending Java with traits. For example, FeatherTrait
-Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the calculus of
-Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
-trait-based inheritance in Java. However, language extensions (including FTJ)
-have a natural drawback: the programmer has to learn new syntax. In contrast,
-our approach is completely compatible with the current Java language, so that
-programmers don't need to pay any learning cost to adapt to this new classless
-programming pattern. Another drawback which is particular for FTJ is that FTJ
-doesn't have type for traits, hence the correctness check of trait is done when
-type-checking classes. This choice makes the design of FTJ simpler but lost
-typechecking efficiency (one trait will be potentially checked multiple times if
-it is used in multiple classes).
+%\yanlin{need to discuss: whether the comments for FTJ is fair.}
 
-\subsection{Multiple Inheritance in General}
-Multiple inheritance is a great feature to have in a programming language, but
-it is difficult to model and implement. These difficulties include the famous
-diamond (fork-join) problem, conflicting methods, yo-yo problem, etc. In order
-to model multiple inheritance, lots of models have been proposed in the past few
+%No, was not fair, I know that work, it does modular type checking, and 
+%more work after that they extend with more safe/modular typing things.
+%also, Java8 IS a language extension on Java
+%However, language extensions (including FTJ)
+%have a natural drawback: the programmer has to learn new syntax. In contrast,
+%our approach is completely compatible with the current Java language, so that
+%programmers don't need to pay any learning cost to adapt to this new classless
+%programming pattern. Another drawback which is particular for FTJ is that FTJ
+%doesn't have type for traits, hence the correctness check of trait is done when
+%type-checking classes. This choice makes the design of FTJ simpler but lost
+%typechecking efficiency (one trait will be potentially checked multiple times if
+%it is used in multiple classes).
+
+\subsection{Multiple Inheritance in Object oriented languages}
+Many authors have argued in favour or against multiple inheritance.
+It provides expressive power, but it is difficult to model, and implement and can create programs that are hard to reason about.
+These difficulties include the famous
+diamond (fork-join) problem, conflicting methods, and the yo-yo problem.
+\marco{add citation on the text over}
+
+To conciliate the need for expressive power and the need for simplicity, 
+lots of models have been proposed in the past few
 years, including mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
-Both traits and mixins provide novel models of new programming architecture in
-object oriented programming paradigm.
+They provide novel programming architecture models in the
+OO paradigm. 
+\begin{enumerate}
+\item  Mixins allows to name components that can be applied to various classes as
+reusable functionality units. However,  they suffer from linearisation: the order of mixin application is relevant in often subtitle and undesired ways. constraints. This hinders their
+usability and their ability of resolving conflicts:
+ the linearisation (total ordering) of mixin inheritance cannot provide a satisfactory resolution in all cases and restricts the flexibility of mixin composition.
 
-1) Mixins model named components that can be applied to various classes as
-reusable functionality units. However, the well-known limitation of mixins is
-that they suffer from linearization constraints, which hinders the
-expressiveness of mixins and the ability of resolving conflicts.  As shown in
-section~\ref{sec:comparison}, although they solved some multiple inheritance
-problems, they are extremely restricting expressiveness.
+ To fight those limitations, an algebra of mixin operators is introduced~\cite{CMS elena zucca davide ancona}, but this raised the complexity of the approach, especially when constructors and fields are considered~\cite{Ecoop FJigsaw marco}.
 
-2) Traits are a novel model to differentiate two functionalities out of classes:
-units of reusable components and object factories. In trait models, traits are
-responsible for units of reusable components, which contain pure methods as
-reusable functionalities; whereas classes act as object factories, which
-requires various functionalities from multiple traits. However drawbacks of
-traits are obvious: traits do not allow state; traits can't be instantiated
-directly.
+\item 
+Simplifying the mixin algebras approach, traits draw a strong line between units of reuse (traits) and object factories (classes)
+In this model, traits~\cite{Scharli2003} units of reusable code,  containing only methods as
+reusable functionalities. Thus, no state/state initialization is considered.
 
-3) C++ and Scala also try to provide solutions to multiple inheritance, but both
+Classes act as object factories, requiring functionalities from multiple traits.
+Traits offers a trait algebra  with operations like sum, alias and exclusion, provided for explicit conflict resolution.
+
+Concluding, (pure) traits do not allow state and they do not offers any reuse instrument to ensure that state is coherently initialized when finally defined in classes.
+Traits can't be instantiated. This model requires two concepts (traits and classes) to coexist and cooperate.
+
+Some authors see this as good language design fostering good software development by helping programmers to think about the structure of their programs.
+However, other authors see the need of two concepts and the absence of state as drawbacks of this model.
+
+\item C++ and Scala also try to provide solutions to multiple inheritance, but both
 suffer from object initialization problems. Virtual inheritance in C++ provides
@@
 superclasses, which would potentially cause serious semantic error. Scala
-solution (traits) avoids this problem by disallowing constructor parameters,
-causing no ambiguity in cases such as diamond problem. The drawback of this
-elegant design is limited expressiveness.
+solution (very similar to linearised mixins, but misleadingly called traits in the language) avoids this problem by disallowing constructor parameters,
+causing no ambiguity in cases such as diamond problem. 
+This approach has limited expressiveness, and suffers from all the problems of linearised mixin composition.
+Python also offers multiple inheritance via linearised mixins. Indeed python any class is implicitly a mixin, and mixin composition informally expressed as\\*
+\Q@ class A use B,C {...new methods...}@\\*
+can be expressed in python as \\*
+\Q@ class Aux: ...new methods...@\\*
+\Q@ class A(B,C,Aux):  pass@ 
+\end{enumerate}
+
+\subsection{Multiple Inheritance in Java}
+Since Java 8 default methods are introduced, which enables concrete method
+implementation to be defined (via the \textbf{default} keyword) inside
+interfaces. Since Java supports implementation of multiple interfaces (instead of extension of a single
+class), the introduction of default methods opens the gate for various flavours of multiple
+inheritance in Java, using interfaces. 
+Former work by Bono.et. al.~\cite{bono14}.
+provides details on mimicking traits~\cite{sharly thesis} through interfaces.
+
+There are proposals for extending Java (before Java8) with traits. For example, FeatherTrait
+Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the calculus of
+Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
+trait-based inheritance in Java.
+Except for few, mostly syntactic details, their work can be emulated/rephrased with Java8 interface.
+
+\input{sections/Compare.tex}
 
@@
 those use any code generation to assist in the implementation?).
+
+\section{Future work}
+
+\subsection{Private state/state initialization}
+The biggest limitation of our approach is the absence in Java8 of support for private/protected methods in interfaces.That is, in Java8 all members of interfaces are public, including static methods.
+Since we use abstract methods to encode the state, our state is always all public; however is impossible for the user to know if a certain method maps directly to a field or if it have a default implementation.
+If the use wants a constructor that does not directly maps to the fields, (as for secondary constructors in Scala) he can simply define its own \Q@of@ method and delegate on the generated one, as in
+\begin{lstlisting}
+@Mixin interface Point{
+  int x(); int y();
+  static Point of(int val){return Point.of(val,val);}  
+  }
+\end{lstlisting}
+However, the generated \Q@of@ method would also be present and public.
+If a future version of Java was to support \emph{static private methods in interfaces} we could extend our code generation to handle also encapsulation.
+Currently, is possible to use a public nested class with private static methods inside, but this is ugly and cumbersome.
+We are considering if our annotation processor can take code with \Q!@Private!
+annotation and turn it into static private methods of a nested class.
+In this extension,  also the of method could be made private following the same pattern.
+
+\subsection{Class invariants in ClassLess Java}
+Since the objects are created by automatically generate methods,
+another limitation of our current approach is that there is no place where the user can dynamically check for class invariants.
+In Java often we see code like
+\begin{lstlisting}
+class Point{ int x; int y;
+  Point(int x; int y){this.x=x;this.y=y; assert this.checkInvariant();}
+  private boolean checkInvariant(){... x>0,y>0...}
+\end{lstlisting}
+
+We are considering an extension of our annotation where 
+default methods with the special name \Q@checkInvariant()@ will be called inside of the \Q@of@ methods.
+if multiple interfaces are implemented, and more then one offers
+\Q@checkInvariant()@,  a composed implementation could be automatically generated, composing by \Q@&&@ the various competing implementations.
+
+\subsection{Clone, toString, equals and hashCode}
+Methods originally defined in \Q@Object@, as \Q@clone@ and \Q@toString@ can be supported by our approach, but they need special care.
+If an interface annotated with \mixin ask an implementation for \Q@clone@, \Q@toString@ \Q@equls@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters.
+Note how this would solve most of the Java ugliness related to \Q@clone@ methods.}
+
+However, if the user wish to provide its own implementation, since the method is also implemented in \Q@Object@ we would have a conflict, that we have to explicitly resolve inside of \Q@of@, by implementing the method and delegating it to the user implementation, thus
+
+\begin{lstlisting}
+@Mixin interface Point{ int x(); int y();
+  default Point clone(){ return Point.of(0,0);}//user defined clone
+  }
+\end{lstlisting} 
+Would expand into
+
+\begin{lstlisting}
+interface Point{ int x(); int y();
+  default Point clone(){ return Point.of(0,0);}//user defined clone
+  public static Point of(int _x,int_y){
+    return new Point(){...
+      public Point clone(){ return Point.super.clone();}
+      };  }  }
+\end{lstlisting}

### 2015-11-10 paper/sections/RelatedWork.tex
Notes added

@@
 \subsection{ThisType/MyType/Extensibility}
+\marco{ this section is unfair.
+There are unsolved problems with ThisType in negative positions, namelly it breaks subtyping.
+The most common solution is to allow calling some methods only when the exact type is known.
+This demolish most advantages of interfaces.}
 Typical object-oriented programming languages (e.g. Java) use class names as

### 2015-11-10 paper/sections/RelatedWork.tex
Text reorganization

@@
 In this section we discuss related work and comparison to Classless Java.
-
-\subsection{Formalization of Java8}
-\marco{this go later, when we compare our formalization}
-There is also formalization proposed by
-Goetz and Field~\cite{goetz12fdefenders} to formalize defender (default) methods
-in Java. However this formalization is limited since they only model exactly one
-method inside classes/interfaces, where in our formalization multiple methods
-may be defined.
 
@@
 
+\subsection{Formalization of Java8}
+We provide a simple and well designed formalization for a subset of Java including default/static interface methods and object initialization literals (often called anonymous local inner classes).
+A similar formalization was drafted by\marco{ I use the term draft because I seams to remember it was just a technical report, I'm right?}
+Goetz and Field~\cite{goetz12fdefenders} to formalize defender (default) methods
+in Java. However this formalization is limited to model exactly one
+method inside classes/interfaces.
+
+As a evidence of the attention and care present in our formalization work, while double-checking the behaviour of Java in side cases we have discovered a likely bug in the current \texttt{javac}.\marco{refer to before when we explain the issue.}
+
 \section{Future work}

### 2015-11-11 paper/sections/RelatedWork.tex
added references

@@
 These difficulties include the famous
-diamond (fork-join) problem, conflicting methods, and the yo-yo problem.
-\marco{add citation on the text over}
-
+diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, and the yo-yo problem~\cite{taenzer1989problems}.
+\marco{add citation on the text over.}
+\yanlin{added.}
 To conciliate the need for expressive power and the need for simplicity, 

### 2015-11-11 paper/sections/RelatedWork.tex
related work on Lombok

@@
 \paragraph{Meta-programming techniques}
-- MetaFJig*
+- MetaFJig*~\cite{}
 \yanlin{Maybe Marco can help with the text for MetaFJig?}
@@
 What is it that we can do that, those annotations cannot?
+
+-Lombok~\cite{} project provides a set of predefined annotations, including
+constructor generators similar as ours (\textbf{@NoArgsConstructor},
+\textbf{@RequiredArgsConstructor} and \textbf{@AllArgsConstructor}). They
+generate various kinds of constructors for \emph{classes}, with or without
+constructor arguments. This set of annotations is of great use, especially when
+they are used together with other features provided in Lombok (e.g.,
+\textbf{@Data}). Moreover, the implementation of these annotations in Lombok
+gives us hints on how to implement \mixin. However, none of these annotations
+can model what we are doing with \mixin - generating a constructor for
+\emph{interfaces}. Apart from constructors, \mixin also provides other
+convenient features (including generating fluent setters, type refinement,etc)
+that the base Lombok project could not provide.
 

### 2015-11-11 paper/sections/RelatedWork.tex
related work on fluent interfaces

@@
 
-\paragraph{Meta-programming techniques}
+\subsection{Meta-programming techniques}
 - MetaFJig*~\cite{}
@@
 
-\paragraph{Fluent Interfaces}
-
+\subsection{Fluent Interfaces}
 Talk about existing libraries that use fluent interfaces (do any of
 those use any code generation to assist in the implementation?).
+
+Since the style of fluent interfaces was invented in Smalltalk as method
+cascading, more and more languages came to support fluent interfaces, including
+JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
+interfaces, programmers have to either handwritten everything or create a
+wrapper around the original non-fluent interfaces, using \textbf{this}. In Java,
+several libraries help create fluent APIs easily (by code generation), including
+jOOQ, op4j, fluflu, JaQue, etc. These libraries have given us hints on the
+implementation of fluent setters in ClassLess Java.
 

### 2015-11-11 paper/paper.tex
more related work
Some work on comparision and related work. Section Future work now exists
Abstract first draft

@@
 
+Java 8 introduced \emph{default methods}, allowing interfaces to
+have method implementations. When combined with (multiple) interface
+inheritance, default methods provide a basic form of multiple
+inheritance. However, using this combination to simulate multiple
+inheritance quickly becomes cumbersome, and appears to be quite
+restricted.
+
+This paper shows that, with some minor syntactic sugar, default
+methods and interface inheritance are in fact be very expressive. Our
+proposed syntax sugar, called \emph{concrete interfaces}, enables
+powerful object-oriented idioms, using multiple inheritance, to be
+expressed directly in Java. Concrete interfaces avoid some syntactic
+boilerplate caused by Java being originally designed to work mainly
+with classes. Moreover, concrete interfaces come with their own object
+instantiation mechanism, providing an alternative to class
+constructors. As a result many Java programs can be built without
+using a single class! An implementation of concrete interfaces using 
+Java annotations and a formalization of the static and dynamic
+semantics are presented. Moreover the usefulness of concrete
+interfaces is illustrated through various examples and case studies.
+\bruno{a first draft of the abstract. still missing ``lightweight
+  language extensions''.}
+
 \end{abstract}
@@
 
-\input{sections/Compare.tex}
-
 \input{sections/Formalization.tex}

### 2015-11-12 paper/sections/Introduction.tex
Draft text for the introduction

--- 2015-09-13 paper/sections/Introduction.tex
@@
 \section{Introduction}\label{sec:intro}
+
+Java 8 introduced \emph{default methods}, allowing interfaces to have
+method implementations. The main motivation behind the introduction of
+default methods in Java 8 is \emph{interface evolution}. That is, to
+allow interfaces to be extended over time, while preserving backwards
+compatibility. It soon became clear that default methods could also be
+used to emulate something similar to \emph{traits}~\cite{}. The
+original notion of traits by Scharli et al. prescribes, among other
+things, that: 1) a trait provides a set of methods that implement
+behaviour; and 2) a trait does not specify any state variables, so the
+methods provided by traits do not access state variables directly.  A
+detailed description of how to emulate trait-oriented programming in
+Java 8 can be found in the work by Bono et al.~\cite{}.  The Java 8
+team designing default interfaces, was also fully aware of that
+secondary use of traits, but it was not their objective to model
+traits: ``The key goal of adding default methods to Java was
+"interface evolution", not "poor man's traits"''~\cite{mailing list}. As
+a result they were happy to support the secondary use of default
+methods as long as it did not make the implementation and language
+more complex.
+
+Still, the design has been criticised for being too conservative and 
+not very useful in its current form to model traits or multiple 
+inheritance~\cite{}. \bruno{mention scala traits here?}
+Indeed, our own personal experience of combining default methods 
+and multiple interface inheritance to achieve multiple implementation 
+inheritance is that many workarounds and boilerplate code are needed. 
+In particular, we encountered difficulties because:
 
@@
 
-\item Using annotattions to implement a rich notion of traits with a
-mechanism to instantiate objects (the of method). Goal 1: is to reduce
-the amount of code that is required to program with interfaces and default
-methods. Goal 2: To provide a convenient means to do multiple inheritance
-in Java.
+\item {\em Interfaces have no constructors.} As a result classes are 
+still required to create objects, leading to substantial boilerplate 
+code for initialization.
 
-\item Specify the system more formally.
+\item {\em Interfaces do not have state.} This creates a tension between 
+ using multiple inheritance and having state. Using setter and
+  getter methods is a way out of this tension, but this workaround
+  requires tedious boilerplate classes that latter implements those
+  methods.
 
-\item Show that we can model all trait operations
-
-\item Implementation using Lombok
-
-\item Case studies: The expression problem, Trivially
-Case Studies from traits paper.
+\item {\em Useful, general purpose methods, require special care in
+  the presence of subtyping.} Methods such as $clone$, or
+  \emph{fluent} setters~\cite{}, not only require access to the
+  internal state of an object, but they also require their types to be
+  refined in subtypes.
 
 \end{itemize}
+
+\noindent Clearly, a way around those difficulties would be to try 
+to change the language and just remove these limitations. Scala's 
+own notion of traits, for example, allows state in traits. Ofcourse 
+this would complicate the language and require changes in the compiler, 
+and this would go beyond the goals of Java 8 development team.
+
+This paper takes a different approach. Rather than trying to get
+around the diffulculties by extending the language in quite
+fundamental ways, we show that, with some minor syntactic sugar,
+default methods and interface inheritance are in fact be very
+expressive. Our proposed syntax sugar, called \emph{concrete
+  interfaces}, enables powerful object-oriented idioms, using multiple
+inheritance, to be expressed directly in the current Java
+language. Concrete interfaces avoid some syntactic boilerplate caused
+by Java being originally designed to work mainly with
+classes. Moreover, concrete interfaces come with their own object
+instantiation mechanism, providing an alternative to class
+constructors. Finally concrete interfaces support various common
+utility methods (such as getters and setters, or clone methods)
+without boilerplate; and they automatically they with utility methods
+in the presence of subtyping. As a result many Java programs can be 
+built without using a single class! \bruno{integrate more of Marco's text.}
+
+To formalize concrete interfaces, we propose FeatherweightConcrete (FC): 
+a FeatherweightJava-style calculus, which captures the essence 
+of interfaces with default methods. The semantics of concrete interfaces 
+is given as a type-directed translation from FC to FC itself. 
+In the resulting FC code all concrete interfaces are translated into 
+regular FC (and Java) interfaces with default methods. The translation 
+is proved to be type-preserving, ensuring that the tranlation does 
+not introduce new errors. One interesting aspect of the theorem is that 
+if we ignore the concrete annotations, the program still type-checks 
+under the regular typing rules of FC.\bruno{double-check!}
+
+An implementation of concrete interfaces using Java annotations and a
+formalization of the static and dynamic semantics are presented. We
+implement concrete interfaces, purely using Java annotations. As such
+no new syntax is required, and existing Java tools (such as IDE's)
+work with our implementation.
+
+Moreover the usefulness of concrete
+interfaces is illustrated through various examples and case studies.
+
+The contributions of this paper are:
+
+\begin{itemize}
+
+\item {\bf Concrete interfaces:} A simple language feature that allows 
+various powerful programming idioms in Java 8 to be expressed conveniently. 
+Concrete interfaces are pure sugar: they can be desugared into standard Java 
+code, without loss of modularity or type-safety.
+
+\item {\bf Featherweight Concrete (FC):} A simple formal calculus that models 
+the essential features of Java 8 interfaces with default methods, and 
+can be used to formally define the desugaring of concrete interfaces. 
+We prove ...
+
+\item {\bf Implementation and Case Studies:} We have a prototype implementation 
+  of concrete interfaces, using Java annotations. Moreover the usefulness
+  of concrete interfaces is illustrated through various examples and
+  case studies.
+
+\end{itemize}
+
+\bruno{Useful reference, perhaps, on the discussion of why not override object 
+methods: 
+\url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}

### 2015-11-12 paper/sections/CaseStudy.tex
code adjusted for case study

@@
 
-\lstinputlisting[linerange=25-42]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=25-35]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 
@@
 
-\lstinputlisting[linerange=46-53]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=39-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=57-75]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=48-57]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 

### 2015-11-12 UseMixinLombok/src/game/TestGame.java

@@ -3,6 +3,7 @@ package game;
 import static java.lang.System.out;
 import lombok.Mixin;
 
+/* Defines a base-door with no particular features */
 @Mixin interface TDoor {
     public boolean locked();
     public int doorMaxCoins();
@@ -33,13 +34,11 @@ import lombok.Mixin;
     }
 }
 
-@Mixin
-interface TCounter {
-    /* State references */
+/* Provides a counter that after a limit releases coins */
+@Mixin interface TCounter {
     public int counter();
     public void counter(int c);
     public int limit();
-    /* Coin management */
     public int counterMaxCoins();
     default void incrementCounter() {
         counter(counter() + 1);
@@ -58,9 +57,9 @@ interface TCounter {
     }
 }
 
+/* Provides a chest that contains coins */
 interface TChest {
-    /* Coin management */
-    public int chestMaxCoins(); /* Opens the chest */
+    public int chestMaxCoins();
     default int open() {
         out.print("The chest is now opened!");
         double rnd = Math.random();
@@ -70,8 +69,9 @@ interface TChest {
         return c;
     }
 }
+
+/* Provides an enchantment that can give or take coins */
 interface TEnchantment {
-    /* Coin management */
     public int enchantMaxCoins();
     /** An enchantment can give coins
     (max getEnchantMaxCoins()) or
@@ -79,7 +79,7 @@ interface TEnchantment {
     default int applyEnchantment() {
         out.println("\nThis is an enchantment!");
         out.print("\"If the luck is up, ");
-        out.println("of coins you’ll have a cup,");
+        out.println("of coins you will have a cup,");
         out.print("but if no luck you got, ");
         out.println("you are gonna lose a lot.\"");
         int max = enchantMaxCoins();
@@ -97,8 +97,8 @@ interface TEnchantment {
     }
 }
 
-@Mixin
-interface TEnchantedDoor extends TDoor, TEnchantment {
+/* Puts together a door and an enchantment */
+@Mixin interface TEnchantedDoor extends TDoor, TEnchantment {
     /** When you open an enchanted door,
     you break the enchantment and so
     you apply it. **/
@@ -109,9 +109,9 @@ interface TEnchantedDoor extends TDoor, TEnchantment {
         return coins;
     }
 }
-@Mixin
-interface TChestedDoor
-        extends TDoor, TChest {
+
+/* Puts together a door and a chest */ 
+@Mixin interface TChestedDoor extends TDoor, TChest {
     /** When you open a chested door,
[…]

### 2015-11-12 UseMixinLombok/src/gameori/ChestedDoor.java

@@ -1,26 +1,0 @@
-package gameori;
-public class ChestedDoor implements TChestedDoor {
-    /* Fields for the door */private boolean locked;
-    /* Glue Code for TDoor */public boolean getLocked()
-    {
-        return this.locked;
-    }
-
-    /* Glue code for coin management */public int getDoorMaxCoins()
-    {
-        return 120;
-    }
-    public int getChestMaxCoins()
-    {
-        return 250;
-    }
-    /* Constructor */
-    public ChestedDoor(boolean l)
-    {
-        setLocked(l);
-    }
-    /* Other helpful methods */public void setLocked(boolean l)
-    {
-        this.locked = l;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/DoorsRoom.java

@@ -1,25 +1,0 @@
-package gameori;
-public class DoorsRoom {
-    private TDoor leftDoor;
-    private TDoor rightDoor;
-    private TDoor frontDoor; /* Constructor */
-    public DoorsRoom(TDoor l, TDoor r,
-            TDoor f) {
-        leftDoor = l;
-        rightDoor = r;
-        frontDoor = f;
-}
-    /* Getters */
-    public TDoor getLeftDoor()
-    {
-        return leftDoor;
-    }
-    public TDoor getRightDoor()
-    {
-        return rightDoor;
-    }
-    public TDoor getFrontDoor()
-    {
-        return frontDoor;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/EnchantedDoor.java

@@ -1,24 +1,0 @@
-package gameori;
-public class EnchantedDoor implements TEnchantedDoor {
-    /* Fields for the door */
-    private boolean locked;
-    /* Glue Code for TDoor */
-    public boolean getLocked() {
-        return locked;
-    }
-    /* Glue code for coin management */
-    public int getDoorMaxCoins() {
-        return 120;
-    }
-    public int getEnchantMaxCoins() {
-        return 150;
-    }
-    /* Constructor */
-    public EnchantedDoor(boolean l) {
-        setLocked(l);
-    }
-    /* Other helpful methods */
-    public void setLocked(boolean l) {
-        this.locked = l;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/KnockDoor.java

@@ -1,45 +1,0 @@
-package gameori;
-public class KnockDoor implements TKnockDoor { /* Fields for the door */
-    private boolean locked;
-    /* Fields for the counter */
-    private int counter;
-    private int limit;
-    /* Glue Code for TDoor */public boolean getLocked()
-    {
-        return this.locked;
-    }
-    /* Glue code for TCounter */public int getCounter()
-    {
-        return this.counter;
-    }
-    public void setCounter(int c)
-    {
-        this.counter = c;
-    }
-    public int getLimit()
-    {
-        return this.limit;
-    }
-    /* Glue code for coin management */public int getDoorMaxCoins()
-    {
-        return 120;
-    }
-    public int getCounterMaxCoins()
-    {
-        return 500;
-    } /* Constructor */
-    public KnockDoor(boolean l, int li) {
-        setCounter(0);
-        setLocked(l);
-        setLimit(li);
-    }
-    /* Other helpful methods */
-    private void setLocked(boolean l)
-    {
-        this.locked = l;
-    }
-    private void setLimit(int l)
-    {
-        this.limit = l;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/Player.java

@@ -1,31 +1,0 @@
-package gameori;
-public class Player {
-    private final String nickname;
-    private int bag = 150; //contains the coins /* Constructor */
-    public Player(String n) {
-        nickname = n;
-    }
-    /* Setters and getters */public int getCoins()
-    {
-        return bag;
-    }
-    public String getNickname()
-    {
-        return this.nickname;
-    }
-    /* Helpful methods */
-    public void addInBag(int amount)
-    {
-        this.bag += amount;
-    }
-    public void removeFromBag(int amount)
-    {
-        this.bag -= amount;
-    }
-    public String toString() {
-        String s = "I’m " + getNickname();
-        s += " and i’ve got " + getCoins();
-        s += " coins in my bag.";
-        return s;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TChest.java

@@ -1,15 +1,0 @@
-package gameori;
-import java.lang.Math;
-import static java.lang.System.out;
-/* Provides a chest that contains coins */public interface TChest {
-    /* Coin management */
-    public int getChestMaxCoins(); /* Opens the chest */
-    default int open() {
-        out.print("The chest is now opened!");
-        double rnd = Math.random();
-        int c = (int) (rnd * getChestMaxCoins());
-        out.print("You got " + c);
-        out.println(" coins from the chest.");
-        return c;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TChestedDoor.java

@@ -1,16 +1,0 @@
-package gameori;
-/* Puts together a door and a chest */ public interface TChestedDoor
-extends TDoor, TChest {
-    /** When you open a chested door,
-you also get the prize from the chest.
-This overrides the TDoor’s open(). **/
-    default int open() {
-        int coins = TDoor.super.open(); if(coins > 0) //if the door is open
-            coins += openChest();
-        return coins;
-    }
-    /* Alias for open() from TChest */
-    default int openChest() {
-        return TChest.super.open();
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TCounter.java

@@ -1,27 +1,0 @@
-package gameori;
-import java.lang.Math;
-import static java.lang.System.out;
-/* Provides a counter that after
-a limit releases coins */public interface TCounter {
-    /* State references */
-    public int getCounter();
-    public void setCounter(int c);
-    public int getLimit();
-    /* Coin management */
-    public int getCounterMaxCoins();
-    default void incrementCounter() {
-        setCounter(getCounter() + 1);
-    }
-    default void decrementCounter() {
-        setCounter(getCounter() - 1);
-    }
-    default boolean hasReachedLimit() {
-        return getCounter() >= getLimit();
-    }
-    default int releaseCoins() {
-        double rnd = Math.random();
-        int cns = (int) (rnd * getCounterMaxCoins()) + 1;
-        out.println("You got " + cns + " coins.");
-        return cns;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TDoor.java

@@ -1,40 +1,0 @@
-package gameori;
-import static java.lang.System.out;
-
-/* Defines a base-door with
-no particular features */
-public interface TDoor {
-    /* State references */
-    public boolean getLocked();
-    /* Coin management */
-    public int getDoorMaxCoins();
-    /* Tells if the door is locked or not */
-    default boolean isLocked()
-    {
-        return getLocked();
-    }
-    /* Tries to open the door */
-    default int open() {
-        if (!isLocked()) {
-            out.println("The door has been opened!");
-            double rnd = Math.random();
-            int cns = (int) (rnd * getDoorMaxCoins()) + 1;
-            out.println("You got " + cns + " coins.");
-            return cns;
-        }
-        else {
-            out.println("This door is locked.");
-            return -1;
-        }
-    }
-    /* Performs a knock on the door */
-    default int knock() {
-        out.print("Door says: ");
-        out.print("How you dare, ");
-        out.println("I am the one who knocks!");
-        int c = (Math.random() < 0.8) ? 0 : 1;
-        if (c > 0)
-            out.println("Ow! You got a free coin!");
-        return c;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TEnchantedDoor.java

@@ -1,13 +1,0 @@
-package gameori;
-/* Puts together a door and an enchantment */
-public interface TEnchantedDoor extends TDoor, TEnchantment {
-    /** When you open an enchanted door,
-    you break the enchantment and so
-    you apply it. **/
-    default int open() {
-        int coins = TDoor.super.open();
-        if (coins > 0) //if the door is open
-            coins += applyEnchantment();
-        return coins;
-    }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TEnchantment.java

@@ -1,25 +1,0 @@
-package gameori;
-import java.lang.Math;
-import static java.lang.System.out;
-/* Provides an enchantment that
-can give or take coins */ public interface TEnchantment {
-    /* Coin management */
-    public int getEnchantMaxCoins(); /** An enchantment can give coins
-(max getEnchantMaxCoins()) or
-remove coins (max -getEnchantMaxCoins()) **/ 
-    default int applyEnchantment() {
-        out.println("\nThis is an enchantment!");
-        out.print("\"If the luck is up, ");
-        out.println("of coins you’ll have a cup,");
-        out.print("but if no luck you got, ");
-        out.println("you are gonna lose a lot.\"");
-        int max = getEnchantMaxCoins();
-    double rnd = Math.random();
-    int cns = -max + (int)(rnd*((max*2)+1)); if(cns >= 0) {
-        out.print("Ohoh! You got "+cns);
-        out.println(" coins!"); }
-    else {
-        out.print("You lost "+Math.abs(cns)); out.println(" coins!");
-    }
-    return cns; }
-}

### 2015-11-12 UseMixinLombok/src/gameori/TKnockDoor.java

@@ -1,21 +1,0 @@
-package gameori;
-/* Puts together a door and a counter */
-import static java.lang.System.out;
-public interface TKnockDoor extends TDoor, TCounter {
-    /** Every know makes the counter increment.
-     * If the limit is reached, more coins are released. **/
-    default int knock() {
-        int coins = TDoor.super.knock(); incrementCounter(); if(hasReachedLimit()) {
-        out.print("Ohh! A special drop for you!");
-        coins += releaseCoins();
-        } else {
-            //Let’s give a suggestion to the player 
-            out.print("Don’t challenge me... ");
-            int c = getLimit();
-            String sug = "never knock a door ";
-            sug = sug + "more then " + c + " times.";
-            out.println(sug);
-        }
-        return coins;
-    }
-}

### 2015-11-13 paper/paper.tex
Draft text for the introduction
Introduction

--- 2015-11-06 paper/paper.tex
@@
 
+Java 8 introduced \emph{default methods}, allowing interfaces to
+have method implementations. When combined with (multiple) interface
+inheritance, default methods provide a basic form of multiple
+inheritance. However, using this combination to simulate multiple
+inheritance quickly becomes cumbersome, and appears to be quite
+restricted.
+
+This paper shows that, with some minor syntactic sugar, default
+methods and interface inheritance are in fact be very expressive. Our
+proposed syntax sugar, called \emph{object interfaces}, enables
+powerful object-oriented idioms, using multiple inheritance, to be
+expressed directly in Java. Object interfaces avoid some syntactic
+boilerplate caused by Java being originally designed to work mainly
+with classes. Moreover, object interfaces come with their own object
+instantiation mechanism, providing an alternative to class
+constructors. As a result many Java programs can be built without
+defining a single class! An implementation of object interfaces using 
+Java annotations and a formalization of the static and dynamic
+semantics are presented. Moreover the usefulness of object
+interfaces is illustrated through various examples and case studies.
+\bruno{a first draft of the abstract. still missing ``lightweight
+  language extensions''.}
+
 \end{abstract}

### 2015-11-13 paper/sections/Introduction.tex
Introduction

--- 2015-09-13 paper/sections/Introduction.tex
@@
 \section{Introduction}\label{sec:intro}
+
+Java 8 introduced \emph{default methods}, allowing interfaces to have
+method implementations. The main motivation behind the introduction of
+default methods in Java 8 is \emph{interface evolution}. That is, to
+allow interfaces to be extended over time, while preserving backwards
+compatibility. It soon became clear that default methods could also be
+used to emulate something similar to \emph{traits}~\cite{}. The
+original notion of traits by Scharli et al. prescribes, among other
+things, that: 1) a trait provides a set of methods that implement
+behaviour; and 2) a trait does not specify any state variables, so the
+methods provided by traits do not access state variables
+directly. Java 8 interfaces follow similar principles too. Indeed, a
+detailed description of how to emulate trait-oriented programming in
+Java 8 can be found in the work by Bono et al.~\cite{}.  The Java 8
+team designing default methods, was also fully aware of that secondary
+use of interfaces, but it was not their objective to model traits:
+``The key goal of adding default methods to Java was "interface
+evolution", not "poor man's traits"''~\cite{mailing list}. As a result
+they were happy to support the secondary use of interfaces with
+default methods as long as it did not make the implementation and
+language more complex.
+
+Still, the design has been criticised for being too conservative and 
+not very useful in its current form to model traits or multiple 
+inheritance~\cite{}. 
+Indeed, our own personal experience of combining default methods 
+and multiple interface inheritance in Java to achieve multiple implementation 
+inheritance is that many workarounds and boilerplate code are needed. 
+In particular, we encountered difficulties because:
 
@@
 
-\item Using annotattions to implement a rich notion of traits with a
-mechanism to instantiate objects (the of method). Goal 1: is to reduce
-the amount of code that is required to program with interfaces and default
-methods. Goal 2: To provide a convenient means to do multiple inheritance
-in Java.
+\item {\em Interfaces have no constructors.} As a result classes are 
+still required to create objects, leading to substantial boilerplate 
+code for initialization.
 
-\item Specify the system more formally.
+\item {\em Interfaces do not have state.} This creates a tension between 
+ using multiple inheritance and having state. Using setter and
+  getter methods is a way out of this tension, but this workaround
+  requires tedious boilerplate classes that latter implements those
+  methods.
 
-\item Show that we can model all trait operations
-
-\item Implementation using Lombok
-
-\item Case studies: The expression problem, Trivially
-Case Studies from traits paper.
+\item {\em Useful, general purpose methods, require special care in
+  the presence of subtyping.} Methods such as $clone$, or
+  \emph{fluent} setters~\cite{}, not only require access to the
+  internal state of an object, but they also require their types to be
+  refined in subtypes.
 
 \end{itemize}
+
+\noindent Clearly, a way around those difficulties would be to try to
+change the language and just remove these limitations. Scala's own
+notion of traits, for example, allows state in traits. Ofcourse adding
+state (and other features) to interfaces would complicate the language
+and require changes in the compiler, and this would go beyond the
+goals of Java 8 development team.
+
+This paper takes a different approach. Rather than trying to get
+around the difficulties by changing the language in fundamental ways,
+we show that, with some minor syntactic sugar, default methods and
+interface inheritance are in fact be very expressive. Our proposed
+syntax sugar enables powerful object-oriented idioms, using multiple
+inheritance, to be expressed directly without modifications to the
+current Java language. We call our sugar \emph{object interfaces},
+because such interfaces can be instantiated directly, without the need
+for an explict class definition. Object interfaces come with their own
+object instantiation mechanism, providing an alternative to class
+constructors. Moreover, object interfaces avoid some syntactic
+boilerplate caused by Java being originally designed to work mainly
+with classes. This includes support for various common utility methods
+(such as getters and setters, or clone methods), even in the presence
+of subtyping. As a result, with object interfaces, many Java programs
+can be built without using a single class!  \bruno{integrate more of
+  Marco's text.}
+
+To formalize object interfaces, we propose Featherweight Object Interfaces (FOI): 
+a FeatherweightJava-style calculus, which captures the essence 
+of interfaces with default methods. The semantics of object interfaces 
+is given as a type-directed translation from FOI to itself. 
+In the resulting FOI code all object interfaces are translated into 
+regular FOI (and Java) interfaces with default methods. The translation 
+is proved to be type-preserving, ensuring that the translation does 
+not introduce type-errors. One interesting aspect of the theorem is that, 
+if we ignore the object annotations, the program still type-checks 
+under the regular typing rules of FOI (and Java).\bruno{double-check!}
+
+We also present a prototype implementation of object interfaces using
+Java annotations. One benefit of an annotation-based implementation 
+is that existing Java tools (such as IDE's) work with out-of-the-box
+with our implementation. As a result we could experiment object
+interfaces with several interesting Java programs, and conduct various
+case studies. \bruno{mention the following or not? maybe good if we
+  had something about tuning.} Ofcourse, another implementation option would be to
+extend Java, with a single qualifier keyword {\bf object}, so that
+object interfaces would be an intrinsic part of the language. We leave
+this for future work.
+
+To evaluate the usefulness of object interfaces, we illustrate 3
+applications and case studies. The first application is a simple 
+solution to the Expression Problem~\cite{}, supporting independent 
+extensibility~\cite{}, and without boilerplate code. The second
+application is to show how embedded DSLs using fluent interfaces~\cite{} 
+can be easily defined using object interfaces. Finally, the last
+application is a larger case study for a simple Maze game implemented with 
+multiple inheritance. For the last application we show that there is a
+significant reduction in the numbers of lines of code when compared 
+to an existing implementation~\cite{} using plain Java 8. 
+Noteworthy, is the fact that all applications are implemented 
+without defining a single class.
+
+In summary, the contributions of this paper are:
+
+\begin{itemize}
+
+\item {\bf Object interfaces:} A simple language feature that allows 
+various powerful programming idioms to be expressed conveniently in
+Java. Object interfaces are pure sugar: they can be desugared into standard Java 
+code, without loss of modularity or type-safety.
+
+\item {\bf Featherweight Object (FC):} A simple formal calculus that models 
+the essential features of Java 8 interfaces with default methods, and 
+can be used to formally define the desugaring of object interfaces. 
+We prove ...
+
+\item {\bf Implementation and Case Studies:} We have a prototype implementation 
+  of object interfaces, using Java annotations. Moreover the usefulness
+  of object interfaces is illustrated through various examples and
+  case studies.
+
+\end{itemize}
+
+\bruno{Useful reference, perhaps, on the discussion of why not override object 
+methods: 
+\url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}

### 2015-11-13 paper/sections/Introduction.tex
Some revision on the Introduction

@@
 
-To formalize object interfaces, we propose Featherweight Object Interfaces (FOI): 
+To formalize object interfaces, we propose Featherweight Object
+Interfaces (FOI)\bruno{should we just call it Featherweight Classless
+  Java?}: 
 a FeatherweightJava-style calculus, which captures the essence 
@@
 under the regular typing rules of FOI (and Java).\bruno{double-check!}
+
+FOI's usefulness goes beyond serving as a calculus to formalize object
+interfaces. During the development process of FOI, we encountered an 
+incoherent behaviour in Java's implementation of default methods that 
+could be classified as a bug. \bruno{Should we report bug to the Java
+  developers mailing list?}
 
@@
 
-\item {\bf Object interfaces:} A simple language feature that allows 
+\item {\bf Object Interfaces:} A simple language feature that allows 
 various powerful programming idioms to be expressed conveniently in
@@
 
-\item {\bf Featherweight Object (FC):} A simple formal calculus that models 
+\item {\bf Featherweight Object Interfaces (FOI):} A simple formal calculus that models 
 the essential features of Java 8 interfaces with default methods, and 
@@
 We prove ...
+
+\item {\bf A Bug in the Implementation of Java 8:}\bruno{state this as
+  a contrib?}
 
@@
 \url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}
+\bruno{Maybe Marco's text can be in the conclusion, which I think is
+  ok to be more provocative. I tried to be a bit more contained in the
+Introduction.}

### 2015-11-13 paper/paper.tex
latex table
LIPICS format

@@
 %\documentclass[preprint]{sigplanconf}
+%\documentclass[preprint]{llncs}
+\documentclass[a4paper,UKenglish]{lipics}
 
-\documentclass[preprint]{llncs}
 \input{def}
 
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+\usepackage[table,xcdraw]{xcolor}
 \usepackage{color}
-
 \usepackage{amsmath}
@@
 
-\pagestyle{plain}
+%\pagestyle{plain}
 
@@
 \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
-\newcommand{\hl}[1]{\textcolor{red}{#1}}
+%\newcommand{\hl}[1]{\textcolor{red}{#1}}
 \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
@@
 
-\begin{comment}
-\author{Bruno C. d. S. Oliveira and Yanlin Wang}
-\institute{The University of Hong Kong\\
-\email{\{bruno,ylwang\}@cs.hku.hk}\\
-\authorrunning{Bruno Oliveira and Yanlin Wang}} % abbreviated author list (for running head)
-\end{comment}
+% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Classless Java}
+\titlerunning{Classless Java}
+
+\author[1]{John Q. Open}
+\author[2]{Joan R. Access}
+\affil[1]{Dummy University Computing Laboratory\\
+  Address, Country\\
+  \texttt{open@dummyuni.org}}
+\affil[2]{Department of Informatics, Dummy College\\
+  Address, Country\\
+  \texttt{access@dummycollege.org}}
+\authorrunning{J.\,Q. Open and J.\,R. Access} 
+% mandatory. First: Use abbreviated first/middle names. Second (only in severe
+% cases): Use first author plus 'et. al.'
+
+\Copyright{John Q. Open and Joan R. Access}
+% mandatory, please use full first names. LIPIcs license is "CC-BY";
+% http://creativecommons.org/licenses/by/3.0/
+
+\subjclass{Dummy classification -- please refer to
+  \url{http://www.acm.org/about/class/ccs98-html}}
+% mandatory: Please choose ACM 1998 classifications from
+% http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of
+% Computation".
+\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
+% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\serieslogo{}%please provide filename (without suffix)
+\volumeinfo%(easychair interface)
+  {Billy Editor and Bill Editors}% editors
+  {2}% number of editors: 1, 2, ....
+  {Conference title on which this volume is based on}% event
+  {1}% volume
+  {1}% issue
+  {1}% starting page number
+\EventShortName{}
+\DOI{10.4230/LIPIcs.xxx.yyy.p}% to be completed by the volume editor
+% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \begin{document}
-
-\title{Classless Java}
-%%\subtitle{Solving The Expression Problem with Covariant Return Types}
-%% Poor Man's Family Polymorphism
-
 \maketitle
@@
 
-\bibliographystyle{splncs}
+%\bibliographystyle{splncs}
+\bibliographystyle{plain}
 \bibliography{paper}

### 2015-11-13 paper/sections/RelatedWork.tex
LIPICS format

@@
 
- To fight those limitations, an algebra of mixin operators is introduced~\cite{CMS elena zucca davide ancona}, but this raised the complexity of the approach, especially when constructors and fields are considered~\cite{Ecoop FJigsaw marco}.
+ To fight those limitations, an algebra of mixin operators is introduced~\cite{CMS-elena-zucca-davide-ancona}, but this raised the complexity of the approach, especially when constructors and fields are considered~\cite{Ecoop-FJigsaw-marco}.
 
@@
 Former work by Bono.et. al.~\cite{bono14}.
-provides details on mimicking traits~\cite{sharly thesis} through interfaces.
+provides details on mimicking traits~\cite{sharly-thesis} through interfaces.
 

### 2015-11-14 paper/sections/Introduction.tex
Ok, pdflatex is now working fine. I submit a couple of typo fix to introduction/abstract and some new text.

--- 2015-11-13 paper/sections/Introduction.tex
@@
 change the language and just remove these limitations. Scala's own
-notion of traits, for example, allows state in traits. Ofcourse adding
+notion of traits, for example, allows state in traits. Of course adding
 state (and other features) to interfaces would complicate the language
@@
 
-To formalize object interfaces, we propose Featherweight Object Interfaces (FOI): 
+To formalize object interfaces, we propose Classless Java (CJ)\marco{I propose to call it directly ClassLess Java, the  Featherweight part is in being classless.}: 
 a FeatherweightJava-style calculus, which captures the essence 
 of interfaces with default methods. The semantics of object interfaces 
-is given as a type-directed translation from FOI to itself. 
-In the resulting FOI code all object interfaces are translated into 
-regular FOI (and Java) interfaces with default methods. The translation 
+is given as a type-directed translation from CJ to itself. 
+In the resulting CJ code all object interfaces are translated into 
+regular CJ (and Java) interfaces with default methods. The translation 
 is proved to be type-preserving, ensuring that the translation does 
@@
 if we ignore the object annotations, the program still type-checks 
-under the regular typing rules of FOI (and Java).\bruno{double-check!}
-
+under the regular typing rules of CJ (and Java).\bruno{double-check!}
+\marco{ I do not understand what you mean.
+1) annotations are part of java and CJ, so you do not need to remove them 2) programs well typed without expansion are well typed with expansion, here you seams to state the opposite}
+CJ's usefulness goes beyond serving as a calculus to formalize object
+interfaces. During the development process of CJ, we encountered an 
+incoherent behaviour in Java's implementation of default methods that 
+could be classified as a bug. \bruno{Should we report bug to the Java
+  developers mailing list?}
+\marco{we need to put the exact javac version}
 We also present a prototype implementation of object interfaces using
@@
 interfaces with several interesting Java programs, and conduct various
-case studies. \bruno{mention the following or not? maybe good if we
-  had something about tuning.} Ofcourse, another implementation option would be to
-extend Java, with a single qualifier keyword {\bf object}, so that
-object interfaces would be an intrinsic part of the language. We leave
-this for future work.
+case studies. 
+\marco{ I removed your text here and I add after to try talking about tuning}
+%I think that the qualifier object goes AGAINST the idea of tuning.
+%\bruno{mention the following or not? maybe good if we
+%  had something about tuning.} Of course, another implementation option would be to
+%extend Java, with a single qualifier keyword {\bf object}, so that
+%object interfaces would be an intrinsic part of the language. We leave
+%this for future work.
 
@@
 
+Moreover, we believe that our approach is an example of a more general idea, that we call \emph{Language Tuning},
+a concept that sits in the middle between a lightweight language extension and a glorified library.
+Thanks to Java dynamic class loading and annotation processing,  annotations can trigger type-driven transformations on classes during either loading or compilation.
+There are many examples of such Language Tuning libraries,~\cite[lomboc,anotherImSearching].
+Language tuning can offers many features usually implemented by a real language extension, but they do not modify the syntax; pre-existing tools can work transparently on the tuned language.
+
 In summary, the contributions of this paper are:
@@
 
-\item {\bf Object interfaces:} A simple language feature that allows 
+\item {\bf Object Interfaces:} A simple feature that allows 
 various powerful programming idioms to be expressed conveniently in
-Java. Object interfaces are pure sugar: they can be desugared into standard Java 
+Java. Object interfaces are pure sugar: they are desugared into standard Java 
 code, without loss of modularity or type-safety.
 
-\item {\bf Featherweight Object (FC):} A simple formal calculus that models 
+\item {\bf ClassLess Java (CJ):} A simple formal calculus that models 
 the essential features of Java 8 interfaces with default methods, and 
@@
 We prove ...
+
+\item {\bf A Bug in the Implementation of Java 8:}\bruno{state this as
+  a contrib?}\marco{yes, especially if we "suggest solutions"}
 
@@
 
+\item{\bf Language Tuning} We identify this concept and we describe Project Lomboc and object interfaces as examples of Language Tuning.
 \end{itemize}

### 2015-11-14 UseMixinLombok/src/marcoTest/StaticTest.java

@@ -1,0 +1,19 @@
+package marcoTest;
+
+public interface StaticTest {
+  int foo();
+  class Foo{private static StaticTest of(int foo){
+    return new StaticTest(){public int foo(){return foo;}};
+  }}
+  static StaticTest of(int foo){
+    if(foo<0)foo=-foo;
+    return Foo.of(foo);
+  }
+  
+}
+
+interface StaticSee extends StaticTest{
+  static StaticTest boom(){
+    return StaticTest.Foo.of(-2);
+    }
+  }

### 2015-11-15 paper/sections/Formalization.tex
latex table
fix
Proof in progress.

@@
 \subsection{Method Overriding Bug in Java}
-\begin{figure}[tbp]
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
 \centering
-\includegraphics[width=5in]{table.png}
-\caption{Table.}\label{fig:javabug}
-\end{figure}
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & ERROR                        & ERROR       & ERROR       \\ \hline
+\textbf{B2}        & A1/B2                        & ERROR       & A3/B2       \\ \hline
+\textbf{B3}        & \cellcolor[HTML]{C0C0C0}B3.m & ERROR       & ERROR       \\ \hline
+\end{tabular}
+\caption{M extends $A_i, B_j$}
+\label{table:javabug}
+\end{table}
 
-Figure~\ref{fig:javabug} shows the method overriding bug in Java. In the
+Table~\ref{table:javabug} shows the method overriding bug in Java. In the
 example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
@@
 \inferrule[(T-Invk)]{
- \Gamma \vdash \e_0 \in \C_0 \\\\
+ \Gamma \vdash \e \in \C \\\\
 \forall i\in 1..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
-%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static}
+  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
+%\textsf{mmodifier}(\m,\C) \neq \textbf{static}
  }{
- \Gamma \vdash \e_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C }
+ \Gamma \vdash \e\QM.\m\QM(\overline{\e}\QM) \in \C' }
 \quad
@@
 \forall i\in 1..n\  \Gamma \vdash \e_i\in \_<:\C_i \\\\
-\textsf{mtypeS}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
-%\textsf{mmodifier}(\m,\C_0) = \textbf{static}
+\textsf{mtypeS}(\m,\C) \!=\! \overline{\C} \!\to\! \C'
+%\textsf{mmodifier}(\m,\C) = \textbf{static}
 }{
-\Gamma \vdash \C_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
+\Gamma \vdash \C\QM.\m\QM(\overline{\e}\QM) \in \C'}
 \quad
 \inferrule[(T-SuperInvk)]{
-\Gamma(\this) <: \C_0 \\\\
+\Gamma(\this) <: \C \\\\
 \forall i\in 1..n\ \Gamma \vdash \e_i\in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
 %\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
-}{\Gamma \vdash \C_0\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
+}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\overline{\e}\QM) \in \C'}
 
@@
 %\subsubsection{Interface Table}
-
-\subsection{What  \mixin Generates}
-We now show what the \mixin annotation generates. We present a formal definition for
-most of the generated methods; however in our formalism we do not consider
-% setters, so we do not formally define the generation of the setters. 
-%We do not include 
-casts or \Q@instanceof@, so we do not include the \Q@with@ method.
-For the same reason we do not include \Q@void@ returning setters, since they are just a minor variation over the more interesting fluent setters, and they would require special handling just for the conventional \Q@void@ type.
-
-\subsubsection{Translation Function}${}_{}$\\*
-
-\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =
-\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC 
-\methods\ \methods' \cC
-$\\*${}_{}$\tab
-where  $\valid(\C_0)$,\Q@of@$\notin\dom(\methods)$ and $\methods'=\ofMethod(\C_0) \ \otherMethod(\C_0,\methods)$.\\*
-
-To translate an annotated interface, we add the \Q@of@ method, and then we add some other methods.
-However, first of all we check if the interface is valid for annotation:
-
-\noindent$\valid(\C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
-$\isField(\method)$,
-$\isWith(\method, \C_0)$
-or
-$\isSetter(\method,\C_0)$ 
-%$\isClone(\method, \C_0)$.
-
-That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.
-
-%To complete our formal definition we would need to add setters and \Q@with@.
-Moreover, we check that the method \Q@of@ is not already defined by the user.
-In our simplified formalization we consider this to be just an error.
-In our prototype we keep \Q@overloading@ into account, and so we check that an of method with the same signature of the one we would like to generate is not already present.\marco{Do we check it?}
-
-In the following we will write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
-$\m$ must be lower-case and is turned in upper-case upon merging.
-For example \QM{with#foo}=\QM{withFoo}.
-Special names $\specialName(\m)$ are  \QM{with} and all the identifiers of form $\QM{with#}\m$.
-
-\subsubsection{$\ofMethod$}${}_{}$\\*
-We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
-To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
-\noindent$\begin{array}{l}
-\ofMethod(\C_0) = \
- \QM{static}\ C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
-\QM{\{}
-\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
-\tab\tab\tab\tab\tab\tab\tab C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
-\tab\tab\tab\tab\tab\tab\tab
-\C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots 
-\C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
-\tab\tab\tab\tab\tab\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
-%\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_0)\\
-\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
-\end{array}$
-\\*
-with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
-and $\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
-%and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
-
-The function $\fieldsFunc(\C_0)$ (formally defined later) denotes all the fields in the current interface.
-For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
-  \begin{itemize}
-   \item $\m_i$ is the field name, and have type $\C_i$.
-   \item $\m_i$\QM{()} is the getter, that just return the current field value.
-   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
-    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
-\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
-%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
-   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
-   %\item \marco{should we just formalize setters?}
-   \end{itemize}
-
-\subsubsection{Other auxiliary functions}${}_{}$\\*
-
-\noindent$\begin{array}{ll}
-\withMethod:&\withMethod(\C,\m,\C_0,\es)=
-\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
-&\mbox{iff }
-\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;\\
-&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\setterMethod:&\setterMethod(\C,\m,\C_0)=
-\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
- \m\QM{= _val;return this;\}} \\
-&\mbox{iff }
-\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;\\
-&\setterMethod(\C,\m,\C_0)=\emptyset\mbox{ otherwise}\\
-%\cloneMethod:&\cloneMethod(\C_0,\es)=
-%\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
-%&\mbox{iff }
-%\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
-%&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\end{array}$
-
-As you can see above, \Q@with-@ and setter methods are generated if needed.
-We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$.
-Note that we do not need to check if its header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
-
-\noindent$\begin{array}{ll}
-\otherMethod:& \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods) 
-\\&
- \mbox{iff }
-\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isWith(\mBody(\QM{with#}\m, \C_0)) 
-\\&\mbox{ and } \QM{with#}\m\notin\dom(\methods)\\
-& \C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods) 
-\\&
- \mbox{iff }
-\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isSetter(\mBody(\QM_\m, \C_0)) 
-\\&\mbox{ and } \QM_\m\notin\dom(\methods)\\
-%&\C_0\ \QM{clone}\oR\cR\QM;\in
-%\otherMethod(\C_0)
-%\\&  \mbox{ iff } 
-%\isClone(\mBody(\QM{clone}, \C_0)) 
-%\mbox{ and } \QM{clone}\notin\dom(\methods)\\
-\end{array}$
-
-Other methods that we need to generate in the interface are \Q@with-@ and setters. %\Q@clone@.
-%A complete formalization would also generate the \Q@with@.
-This is needed only if we need to refine the return type.
-To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
- is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.
-
-\noindent$\begin{array}{ll}
-\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
-\method\in \dom(\C_0)\ \mand\ \isField(\method)
-\\
-\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
-\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
-\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
-& \mand\ \mnot\ \specialName(\m)\\
-\isSetter:&\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)
-\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
-& \mand\ \mnot\ \specialName(\m)\\
-
-%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
-%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
-%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-\end{array}$
-
-We have not formally modelled non fluent setters and the \Q@with@ method; informally
-\begin{itemize}
-\item For methods inside the interface with the form \Q@void @$\m$\QM($\C\ \x$\QM{);}:
-  \begin{itemize}
-    \item Check if exist method $\C\ \m$\Q@();@. If not, generate error (that is, is not $\valid(\C_0)$).
-    \item Generate implemented setter method inside \Q@of@:\\*
-           \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@
-    Note how there is no need to refine the return type for non fluent setters, thus we do not need to generate the method header in the interface body itself.
-    \end{itemize}
-\item For methods with the form $\C'\ $\QM{with(}$\C\ \x$\QM{);}:
-  \begin{itemize}
-    \item As for before, check that $\C'$ is a supertype of the current interface type $\C_0$.
-    \item Generate implemented \Q@with@ method inside \Q@of@:\\*
-           \Q@public @$\C_0\ $\Q@with(@$\C$\Q@ _val) { @\\*
-           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;@\\*
-           \Q@return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
-where with $\m_1\ldots\m_n$  fields of $\C_0$,
-$\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method; otherwise
-$\e_i=\m_i$.
-    \item If needed, as for \Q@with-@ and setters, generate the method header with refined return type in the interface.
- \end{itemize}
-
-%\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
- % \begin{itemize}
-  %  \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
-   % \item Generate implemented setter method inside \Q@of@:\\*
-    %       \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
-   % \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
- % \end{itemize}
-\end{itemize}
-
-\marco{ insert somewhere description of  fluent setter~\cite{the lombock thread, something about fluent stuff}.
-This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}}.
-
-\subsubsection{Results}
-\textbf{THEOREM. } 
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
-
-To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer
-\marco{should we repeat the 3 points in the email?}
-
-%\subsubsection{Derived notations}
-% Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
-%\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
-
-%\marco{we also need to define a function that gives all the methods of an interface, something line}
-%\[
-%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
-%\]
-
-\begin{comment}
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=5in]{screenshot.png}
-\caption{Screenshot.}\label{screenshot_png}
-\end{figure}
-
-\haoyuan{I tried to understand the current algorithm, and did more experiments in eclipse.
-Now I borrow some ideas from the current version, and give a new version of the algorithm in text. See below.
-
-(1) I guess the function \textsf{tops} is not necessary. The first step is still
-\[\textsf{mbody}(m,C_i)\in\overline{meth}\textrm{ (excluding \textbf{static} methods)}\]
-
-(2) Assume the context is ``interface $C_0$ extends $\overline{C}$ \{$meth'$;...\}''. First handle
-\[\textsf{override}(meth',\overline{meth}) \eqno{(*)}\]
-
-(3) If $meth'\ne\none$, $(*)$ returns $meth'$ if
-\[\forall meth\in\overline{meth},meth'\subtype meth\]
-even if there are conflicts in $\overline{meth}$.
-
-(4) If $meth'=\none$, we need to figure out
-\[\textsf{mostSpecific}(\overline{meth})\]
-and it should be the one that ``overrides'' all the others in $\overline{meth}$.
-It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces.
-But for abstract methods, only return types are taken into consideration.
-}
-\end{comment}
-
-%\text{\yanlin{shouldn't mostSpecific be: $\forall \method' \in \methods : \method \subtype
-%  \method'$ ?}}
-
-%(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
-%\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
-%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
-%\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}
-%
-%(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error.
-The other direction $body_2.\textsf{returnType}<:body_1.\textsf{returnType}$ follows the same rule.
-It also gives an error if there is no subtyping relationship between two return types.
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-%& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-%\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-%& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
-%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}
-
-%\subsubsection{Auxiliary function: \textsf{replace}}
-%
-%The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
-%
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
-%\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
-%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-%\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}

### 2015-11-15 paper/sections/Translation.tex
Commit again.

\section{What  \mixin Generates}\label{sec:translation}

We now show what the \mixin annotation generates. We present a formal definition for
most of the generated methods; however in our formalism we do not consider
% setters, so we do not formally define the generation of the setters.
%We do not include
casts or \Q@instanceof@, so we do not include the \Q@with@ method.
For the same reason we do not include \Q@void@ returning setters, since they are just a minor variation over the more interesting fluent setters, and they would require special handling just for the conventional \Q@void@ type.

\subsection{Translation Function}${}_{}$\\*

\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =
\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \methods' \cC
$\\*${}_{}$\tab
where  $\valid(\C_0)$,\Q@of@$\notin\dom(\methods)$ and $\methods'=\ofMethod(\C_0) \ \otherMethod(\C_0,\methods)$.\\*

To translate an annotated interface, we add the \Q@of@ method, and then we add some other methods.
However, first of all we check if the interface is valid for annotation:

\noindent$\valid(\C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
$\isField(\method)$,
$\isWith(\method, \C_0)$
or
$\isSetter(\method,\C_0)$
%$\isClone(\method, \C_0)$.

That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.

%To complete our formal definition we would need to add setters and \Q@with@.
Moreover, we check that the method \Q@of@ is not already defined by the user.
In our simplified formalization we consider this to be just an error.
In our prototype we keep \Q@overloading@ into account, and so we check that an of method with the same signature of the one we would like to generate is not already present.\marco{Do we check it?}

In the following we will write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
$\m$ must be lower-case and is turned in upper-case upon merging.
For example \QM{with#foo}=\QM{withFoo}.
Special names $\specialName(\m)$ are  \QM{with} and all the identifiers of form $\QM{with#}\m$.

\subsection{$\ofMethod$}${}_{}$\\*
We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
\noindent$\begin{array}{l}
\ofMethod(\C_0) = \
 \QM{static}\ C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
\QM{\{}
\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
\tab\tab\tab\tab\tab\tab\tab C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
\tab\tab\tab\tab\tab\tab\tab
\C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots
\C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
\tab\tab\tab\tab\tab\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
%\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_0)\\
\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
\end{array}$
\\*
with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
and $\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
%and $\es=\m_1\QM,\ldots\QM, \m_n$\\*

The function $\fieldsFunc(\C_0)$ (formally defined later) denotes all the fields in the current interface.
For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
  \begin{itemize}
   \item $\m_i$ is the field name, and have type $\C_i$.
   \item $\m_i$\QM{()} is the getter, that just return the current field value.
   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
   %\item \marco{should we just formalize setters?}
   \end{itemize}

\subsection{Other auxiliary functions}${}_{}$\\*

\noindent$\begin{array}{ll}
\withMethod:&\withMethod(\C,\m,\C_0,\es)=
\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
&\mbox{iff }
\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;\\
&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
\setterMethod:&\setterMethod(\C,\m,\C_0)=
\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
 \m\QM{= _val;return this;\}} \\
&\mbox{iff }
\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;\\
&\setterMethod(\C,\m,\C_0)=\emptyset\mbox{ otherwise}\\
%\cloneMethod:&\cloneMethod(\C_0,\es)=
%\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
%&\mbox{iff }
%\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
%&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
\end{array}$

\haoyuan{method name of setter?}

As you can see above, \Q@with-@ and setter methods are generated if needed.
We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. Note that we do not need to check if its header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.

\noindent$\begin{array}{ll}
\otherMethod:& \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)
\\&
 \mbox{iff }
\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isWith(\mBody(\QM{with#}\m, \C_0))
\\&\mbox{ and } \QM{with#}\m\notin\dom(\methods)\\
& \C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)
\\&
 \mbox{iff }
\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isSetter(\mBody(\QM_\m, \C_0))
\\&\mbox{ and } \QM_\m\notin\dom(\methods)\\
%&\C_0\ \QM{clone}\oR\cR\QM;\in
%\otherMethod(\C_0)
%\\&  \mbox{ iff }
%\isClone(\mBody(\QM{clone}, \C_0))
%\mbox{ and } \QM{clone}\notin\dom(\methods)\\
\end{array}$

Other methods that we need to generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.

\noindent$\begin{array}{ll}
\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
\method\in \dom(\C_0)\ \mand\ \isField(\method)
\\
\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
& \mand\ \mnot\ \specialName(\m)\\
\isSetter:&\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)
\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
& \mand\ \mnot\ \specialName(\m)\\

%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
\end{array}$

We have not formally modelled non fluent setters and the \Q@with@ method; informally
\begin{itemize}
\item For methods inside the interface with the form \Q@void @$\m$\QM($\C\ \x$\QM{);}:
  \begin{itemize}
    \item Check if exist method $\C\ \m$\Q@();@. If not, generate error (that is, is not $\valid(\C_0)$).
    \item Generate implemented setter method inside \Q@of@:\\*
           \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@
    Note how there is no need to refine the return type for non fluent setters, thus we do not need to generate the method header in the interface body itself.
    \end{itemize}
\item For methods with the form $\C'\ $\QM{with(}$\C\ \x$\QM{);}:
  \begin{itemize}
    \item As for before, check that $\C'$ is a supertype of the current interface type $\C_0$.
    \item Generate implemented \Q@with@ method inside \Q@of@:\\*
           \Q@public @$\C_0\ $\Q@with(@$\C$\Q@ _val) { @\\*
           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;@\\*
           \Q@return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
where with $\m_1\ldots\m_n$  fields of $\C_0$,
$\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method; otherwise
$\e_i=\m_i$.
    \item If needed, as for \Q@with-@ and setters, generate the method header with refined return type in the interface.
 \end{itemize}

%\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
 % \begin{itemize}
  %  \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
   % \item Generate implemented setter method inside \Q@of@:\\*
    %       \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
   % \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
 % \end{itemize}
\end{itemize}

\marco{ insert somewhere description of  fluent setter~\cite{the lombock thread, something about fluent stuff}.
This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}}.

\subsection{Results}
\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.

To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer
\marco{should we repeat the 3 points in the email?}\\

\noindent\textbf{LEMMA. }
If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
\begin{proof}
First assume that $\II_0$ is of form $$\mixinAnn\ \QM{interface}\ C_0\ \QM{extends}\ \overline{C}\ \oC \methods\ \cC$$
By the rule \rn{t-Invk} in Figure~\ref{ET}, we divide it into two parts.

\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\II_0]\!]$, the type of the return value is compatible with the method's return type.

Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(C_0)$.

By the definition of $\ofMethod(C_0)$, the return value is an object $$\QM{return new}\ C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
To prove it is of type $C_0$, we use the typing rule \rn{t-Obj}.
\begin{itemize}
\item For field typing,
\[\Gamma(\QM_ m_i)=C_i\subtype C_i\]
\item For method typing of getters,
\[\Gamma, m_i:C_i, \QM{this}:C_0, \Gamma^{\mh_i}\vdash m_i\in C_i=C^{\mh_i}\]
\item For method typing of nonempty \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
\[\Gamma, \overline{m_i:C_i}, \QM{this}:C_0, \es_i:\overline{C}_i, \Gamma^{\mh_i}\vdash C_0\QM{.of}\oR\es_i\cR\in C_0=C^{\mh_i}\]
\item For method typing of nonempty setter methods, \haoyuan{what about the \QM{with} method?}
\[\Gamma, \QM{this}:C_0, \Gamma^{\mh_i}\vdash\QM{this}\in C_0=C^{\mh_i}\]
\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, C_0)$,
 \begin{itemize}
 \item For getters, \[C_i\ m_i\oR\cR\QM; \in \fieldsFunc(C_0) \Rightarrow C_i\ m_i\oR\cR\QM; \in \dom(C_0)\]
 \item For nonempty \QM{with-} methods,
  \[\begin{array}{l}
  \mBody(\QM{with#}\m_i,C_0) \mbox{ is of form }\mh\QM; \\ \valid(C_0)
  \end{array}\Rightarrow \isWith\Rightarrow C_0\subtype\mBody(m^{\mh_i}, C_0)\]
 \item For nonempty setters,
  \[\begin{array}{l}
  \mBody(\m_i,C_0) \mbox{ is of form }\mh\QM; \\ \valid(C_0)
  \end{array}\Rightarrow \isSetter\Rightarrow C_0\subtype\mBody(m^{\mh_i}, C_0)\]
 \end{itemize}
\item To prove that such a created object indeed implements all the abstract methods in $\dom(C_0)$, we simply refer to $\valid(C_0)$, since it guarantees each abstract method $\method$ to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases, hence it is of type $C_0$ by \rn{t-Obj}.
\end{itemize}

\noindent\textbf{Part II.} Next we check $\dom(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)\cup\dom(\methods')$.

\end{proof}

%\subsubsection{Derived notations}
% Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
%\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]

%\marco{we also need to define a function that gives all the methods of an interface, something line}
%\[
%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
%\]

\begin{comment}
\begin{figure}[tbp]
\centering
\includegraphics[width=5in]{screenshot.png}
\caption{Screenshot.}\label{screenshot_png}
\end{figure}

\haoyuan{I tried to understand the current algorithm, and did more experiments in eclipse.
Now I borrow some ideas from the current version, and give a new version of the algorithm in text. See below.

(1) I guess the function \textsf{tops} is not necessary. The first step is still
\[\textsf{mbody}(m,C_i)\in\overline{meth}\textrm{ (excluding \textbf{static} methods)}\]

(2) Assume the context is ``interface $C_0$ extends $\overline{C}$ \{$meth'$;...\}''. First handle
\[\textsf{override}(meth',\overline{meth}) \eqno{(*)}\]

(3) If $meth'\ne\none$, $(*)$ returns $meth'$ if
\[\forall meth\in\overline{meth},meth'\subtype meth\]
even if there are conflicts in $\overline{meth}$.

(4) If $meth'=\none$, we need to figure out
\[\textsf{mostSpecific}(\overline{meth})\]
and it should be the one that ``overrides'' all the others in $\overline{meth}$. It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces. But for abstract methods, only return types are taken into consideration.
}
\end{comment}

%\text{\yanlin{shouldn't mostSpecific be: $\forall \method' \in \methods : \method \subtype
%  \method'$ ?}}

%(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
%\begin{equation*}
%\begin{array}{ll}
%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
%\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
%\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
%\end{array}
%\end{equation*}
%
%(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error. The other direction $body_2.\textsf{returnType}<:body_1.\textsf{returnType}$ follows the same rule. It also gives an error if there is no subtyping relationship between two return types.
%\begin{equation*}
%\begin{array}{ll}
%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
%& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
%\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
%& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
%\end{array}
%\end{equation*}

%\subsubsection{Auxiliary function: \textsf{replace}}
%
%The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
%
%\begin{equation*}
%\begin{array}{ll}
%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
%\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
%\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
%\end{array}
%\end{equation*}

### 2015-11-15 paper/run.sh
Pegasus vector-graph

@@ -4,7 +4,7 @@ if [ $1 ]
 then
     case "$1" in
     bib )       
-        xelatex paper.tex
+        pdflatex paper.tex
         bibtex paper     
     ;;
     * )
@@ -17,7 +17,7 @@ then
 fi
 
 ruby computePositions.rb
-xelatex paper.tex
+pdflatex paper.tex
 open paper.pdf
 
 rm footempfile.txt

### 2015-11-15 paper/sections/CaseStudy.tex
text for Game case study

@@
 
-\paragraph{\textbf{Initial System}}
+\subsubsection{Initial System}
 The initial system expresses arithmetic expressions with only literals and
@@
 
-\paragraph{\textbf{Adding a New Variant}} 
+\subsubsection{Adding a New Variant}
 It is easy to add new data variants to the code in the initial system in a
@@
 
-\paragraph{\textbf{Adding a New Operation}}
+\subsubsection{Adding a New Operation}
 Adding new operations to the previous system is still straightforward (although
@@
 
+\subsection{A Small Game Example}
+We now present a variant of a classic game example, which is often used to
+evaluate reuse ability related to inheritance and design patterns. The code
+using our approach and the solution by Bono et. al~\cite{bono14} in online, for
+space constraints, we will omit it here. In the game, there is a player with the
+goal of collecting as many coins as possible. She may enter a room with several
+doors to be chosen among. This is a good example because it involves code reuse
+(different kinds of doors inherit a common type, with different features and
+behaviour), multiple inheritance (a special kind of door may require features
+from two other door types). It also shows how to model symmetric sum, override
+and alias as trait-oriented programming. Moreover, the code is provided in
+~\cite{bono14} so that we can have a clear and fair comparison with our
+approach, as shown in Table~\ref{table:codereduction}.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{ccc}
+\hline
+            & SLOC   & \# of classes/interfaces \\ \hline
+Bono et al. & 335    & 14                       \\
+Ours        & 199    & 11                       \\
+\rowcolor[HTML]{C0C0C0} 
+Reduced by  & 40.6\% & 21.4\%                   \\ \hline
+\end{tabular}
+\caption{Code Reduction and Reuse.}
+\label{table:codereduction}
+\end{table}
+
 \subsection{Other case studies}

### 2015-11-15 paper/sections/Formalization.tex
revised method overriding mechanism in Java
Some work on the table for java bug

@@
 \section{Formal Semantics}\label{sec:formal}
-
-\subsection{Method Overriding Mechanism}
-We summarize the mechanism for method overriding below (all methods considered
-here are methods with the same name and argument types):
-
-1. Two abstract methods with subtyping relation do not cause conflict, and the
-most specific one wins. Subtyping relation for two methods means they have the
-same arguments' types and one return type is a subtype of the other one. For
-example:
-\begin{lstlisting}
-interface A1 { Object m(); }
-interface B1 { Integer m();}
-interface C1 extends A1,B1 {} //accepted
-\end{lstlisting}
-Method \texttt{Integer m()} in \texttt{B} wins over method \texttt{Object m()}
-in \texttt{A} because \texttt{Integer <: Object}.
-
-2. A default method conflict with another default method or abstract method. For
-example:
-\begin{lstlisting}
-interface A2 { default int m() {return 1;}}
-interface B2 { int m(); }
-interface C2 { default int m() {return 2;}}
-interface D2 extends A2,B2 {} //rejected due to conflicting methods
-interface E2 extends A2,C2 {} //rejected due to conflicting methods
-\end{lstlisting}
-
-3. The method in the current interface wins over any methods defined in its
-super-interfaces, provided that the method conform to the subtype of all methods
-in its super-interfaces, i.e., the method is the most specific one. For example:
-\begin{lstlisting}
-interface D3 extends A2,B2 { int m(); } //accepted
-interface E3 extends A2,C2 { int m(); } //accepted
-\end{lstlisting}
-
-\subsection{Method Overriding Bug in Java}
-% Please add the following required packages to your document preamble:
-% \usepackage[table,xcdraw]{xcolor}
-% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
-\textbf{B1}        & ERROR                        & ERROR       & ERROR       \\ \hline
-\textbf{B2}        & A1/B2                        & ERROR       & A3/B2       \\ \hline
-\textbf{B3}        & \cellcolor[HTML]{C0C0C0}B3.m & ERROR       & ERROR       \\ \hline
-\end{tabular}
-\caption{M extends $A_i, B_j$}
-\label{table:javabug}
-\end{table}
-
-Table~\ref{table:javabug} shows the method overriding bug in Java. In the
-example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
-named as \texttt{m} which are either abstract or default methods. If we define a
-new interface \texttt{M} that extends two of these interfaces, then the method
-overriding result is shown in the table. For example, row2 col2 means \texttt{M
-  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
-\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
-  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
-following this way of interpretation. The interesting case in row4 col2 where
-\texttt{M extends A1,B3}, the result should be \texttt{ERROR}, however in Java
-the experimental result is \texttt{B3.m}. Because if we think consistently, this
-case is esencially a method conflicting situation with abstract methods and
-default methods, which is no different than the first case.
-
-\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
-
-In our approach, we overcome this bug by detecting method correctness when
-typechecking all interfaces.
 
@@
 
-As you can see, we are delegating the work to three others auxiliary functions: $\tops(\Cs), \shadow(\m,\Cs) and \override(\method,\method')$
+As you can see, we are delegating the work to three others auxiliary functions:\\*
+ $\tops(\Cs), \shadow(\m,\Cs)$ and $\override(\method,\method')$
 ${}_{}$\\*
@@
 \method \subtype
-\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}=
+\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}=
 \method\subtype\mh\QM;
 \\
-\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype\method=
+\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method=
 \mh\QM;\subtype\method
@@
 ${}_{}$\\*
-Where $\Aux{mostSpecific}$ return the most specific method using return type specialization as introduced in Java ??
+Where $\Aux{mostSpecific}$ return the most specific method using return type specialization as introduced in Java\marco{insert the version number}
 We just check the subtype between method headers, so we discard information abut method implementation.\\*
@@
 \whereNote
-\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
+\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
 \method\subtype\method'
@@
 %\subsubsection{Interface Table}
+
+\subsection{Method Overriding Mechanism}
+We summarize the mechanism for method overriding below (all methods considered
+here are methods with the same name and argument types):
+\begin{enumerate}
+\item Two abstract methods with subtyping relation do not cause conflict, and the
+most specific one wins. Subtyping relation for two methods means they have the
+same arguments' types and one return type is a subtype of the other one. For
+example:
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+Method \texttt{Integer m()} in \texttt{B} wins over method \texttt{Object m()}
+in \texttt{A} because \texttt{Integer <: Object}.
+
+\item A default method conflict with another default method or abstract method. For
+example:
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is in contrast with what happens in most trait models, where \Q@D2@ would be accepted, and the implementation in \Q@A2@ would be part of the behaviour of \Q@D2@.
+
+\item The method in the current interface wins over any methods defined in its
+super-interfaces, provided that the method conform to the subtype of all methods
+in its super-interfaces, i.e., the method is the most specific one. For example:
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { int m(); } //accepted
+\end{lstlisting}
+\end{enumerate}
+\subsection{Method Overriding Incoherence in Java}
+% Please add the following required packages to your document preamble:
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+While trying to formally encode the Java specification we have done some tests to clarify corner case behaviour.
+Consider the following correct declarations:
+\begin{lstlisting}
+interface A1{T m(); }
+interface A2 extends A1{default T m(){ ... } }
+interface A3 extends A2{T m(); }
+
+interface B1{default T m(){ ... } }
+interface B2 extends B1{T m(); }
+interface B3 extends B2{default T m(){ ... } }
+\end{lstlisting}
+
+What happens if we define a new interface \Q@M@ implementing one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & error                        & error      &error       \\ \hline
+\textbf{B2}        & accepted                        & error       & accepted       \\ \hline
+\textbf{B3}        & \cellcolor[HTML]{C0C0C0}accepted &error       & error      \\ \hline
+\end{tabular}
+\\*
+ and we have found that javac\footnote{We refer to \texttt{javac} as in JDK version is\marco{add}}
+have a very strange behaviour in the following case
+
+Table~\ref{table:javabug} shows the method overriding bug in Java. In the
+example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+named as \texttt{m} which are either abstract or default methods. If we define a
+new interface \texttt{M} that extends two of these interfaces, then the method
+overriding result is shown in the table. For example, row2 col2 means \texttt{M
+  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+following this way of interpretation. The interesting case in row4 col2 where
+\texttt{M extends A1,B3}, the result should be \texttt{ERROR}, however in Java
+the experimental result is \texttt{B3.m}. Because if we think consistently, this
+case is esencially a method conflicting situation with abstract methods and
+default methods, which is no different than the first case.
+
+\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+In our approach, we overcome this bug by detecting method correctness when
+typechecking all interfaces. Internally, algorithm in the auxiliary function
+$\mBody(\m,\C)$ apply the correct method overriding mechanism and returns the
+correct method signature.

### 2015-11-15 paper/sections/Overview.tex
Pegasus vector-graph
Some work on the table for java bug

@@
 
-\begin{figure}[tbp]
+\begin{figure}
 \centering
-\includegraphics[width=4in]{PegasusDetail.png}
+\includegraphics[height=4cm]{PegasusDetail.pdf}
 \caption{Pegasus Example.}\label{fig:pegasus}
@@
  we simulate state by abstract methods.
+By using the new Java8 \Q@default@ methods, we can easily add implemented methods inside the interface. For example:
+\begin{lstlisting}
+interface Point2D{ int x(); int y();
+  default float distanceFromOrigin(){ return Math.sqrt(x()*x()+y()*y());}
+  }
+\end{lstlisting}
+resemble a class that have fields (getters) and methods.
 However, how can we create an instance of \Q@Point2D@ without using classes? 
@@
 @Mixin interface Point2D{ int x();    int y();
-  Point2D withX(int val); 
-  Point2D withY(int val);}
+  Point2D withX(int val);    Point2D withY(int val);}
 \end{lstlisting}
-is equivalent to 
+is expanded by our annotation into
+
+\begin{lstlisting}
+interface Point2D{ int x(); int y();
+  static Point2D of(int _x, int _y){return new Point2D(){
+    int x=_x; int y=_y;
+    public int x(){return x;}   public int y(){return y;}
+    Point2D withX(int val){ return Point2D.of(val,this.y());}    
+    Point2D withY(int val){ return Point2D.of(this.x(),val);}  
+    };}
+  Point2D withX(int val);    Point2D withY(int val);}
+\end{lstlisting}
+If the programmer would like a different implementation, they can just provide it using \Q@default@; that is:
 \begin{lstlisting}
 @Mixin interface Point2D{ int x(); int y();
-  default Point2D withX(int val){ return Point2D.of(val,this.y());}
-  default Point2D withY(int val){ return Point2D.of(this.x(),val);}  }
+  default Point2D withX(int val){ ... };    default Point2D withY(int val){ ... }}
 \end{lstlisting}
-Notice how we can provide implementation for methods in an interface, 
+is expanded into
+\begin{lstlisting}
+interface Point2D{ int x(); int y();
+  static Point2D of(int _x, int _y){return new Point2D(){
+    int x=_x;    int y=_y;
+    public int x(){return x;}    public int y(){return y;}
+    };}
+  default Point2D withX(int val){ ... }    default Point2D withY(int val){ ... }}
+\end{lstlisting}
+
+That is, we can provide implementation for methods in an interface, 
 and our annotation generate code only if the method \emph{needs} implementation. In this way is trivial for the programmer to personalize the behaviour if they have any special need.
@@
 
-Just to give you a feeling on how much boring code \mixin is generating,  we show the correspondent code without then \mixin annotation\footnote{
+Just to give you a feeling on how much boring code \mixin is generating,  we show the expanded code without then \mixin annotation\footnote{
 As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
@@
 interface Point3D extends Point2D{
-  Point3D withX(int val);
-  Point3D withY(int val);
-  Point3D withZ(int val);
+  Point3D withX(int val);    Point3D withY(int val);    Point3D withZ(int val);
   Point3D with(Point2D val);
-  public static Point3D of(int x, int y, int z){return new Point3D(){
-    public int x(){return x;} 
-    public int y(){return y;}
-    public int z(){return z;}
+  public static Point3D of(int _x, int _y, int _z){
+    int x=_x;    int y=_y;    int z=_z;
+    return new Point3D(){
+      public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
       public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
@@
         return Point3D.of(val.x(),val.y(),this.z());
-      }
-    };
-  }
-}
+        }      };    }  }
 \end{lstlisting}
@@
 this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
-\item Define a \Q@Point2D@ field in \Q@Animal@ and...
+\item Define a \Q@Point2D@ field in \Q@Animal@ and:
 \item Define an extra \Q@int z@ field in \Q@Bird@.

### 2015-11-15 paper/sections/Introduction.tex
added references

@@
 compatibility. It soon became clear that default methods could also be
-used to emulate something similar to \emph{traits}~\cite{}. The
+used to emulate something similar to \emph{traits}~\cite{scharli03traits}. The
 original notion of traits by Scharli et al. prescribes, among other
@@
 detailed description of how to emulate trait-oriented programming in
-Java 8 can be found in the work by Bono et al.~\cite{}.  The Java 8
+Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
 team designing default methods, was also fully aware of that secondary
@@
 ``The key goal of adding default methods to Java was "interface
-evolution", not "poor man's traits"''~\cite{mailing list}. As a result
+evolution", not "poor man's traits"''~\cite{goetz13default}. As a result
 they were happy to support the secondary use of interfaces with
@@
   the presence of subtyping.} Methods such as $clone$, or
-  \emph{fluent} setters~\cite{}, not only require access to the
+  \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
   internal state of an object, but they also require their types to be
@@
 applications and case studies. The first application is a simple 
-solution to the Expression Problem~\cite{}, supporting independent 
-extensibility~\cite{}, and without boilerplate code. The second
+solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
+extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
 application is to show how embedded DSLs using fluent interfaces~\cite{} 
@@
 significant reduction in the numbers of lines of code when compared 
-to an existing implementation~\cite{} using plain Java 8. 
+to an existing implementation~\cite{bono14} using plain Java 8. 
 Noteworthy, is the fact that all applications are implemented 

### 2015-11-15 paper/run.sh

@@ -21,7 +21,7 @@ pdflatex paper.tex
 open paper.pdf
 
 rm footempfile.txt
-rm *.bbl
-rm *.blg
+# rm *.bbl
+# rm *.blg
 rm *.log
 rm *-old

### 2015-11-15 paper/sections/Formalization.tex
Ready to read table text

@@
 What happens if we define a new interface \Q@M@ implementing one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
-
+\noindent
 \begin{tabular}{|l|l|l|l|}
@@
 \textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
-\textbf{B1}        & error                        & error      &error       \\ \hline
-\textbf{B2}        & accepted                        & error       & accepted       \\ \hline
-\textbf{B3}        & \cellcolor[HTML]{C0C0C0}accepted &error       & error      \\ \hline
+\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
+\textbf{B3}        & \cellcolor[HTML]{C0C0C0} accepted? &conflict  error       & conservative error      \\ \hline
 \end{tabular}
 \\*
- and we have found that javac\footnote{We refer to \texttt{javac} as in JDK version is\marco{add}}
-have a very strange behaviour in the following case
+We try to classify the results out of the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both implemented. This is also considered an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both abstract. This is also considered correct in all trait models.
+\item \textbf{conservative error}
+ happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ is implemented in only one side.
+The is different from what we would expect in a trait model, but it may make sense with the conservative idea that a method defined in an interface
+should not silently satisfy a method in another one.
+\end{itemize}
 
-Table~\ref{table:javabug} shows the method overriding bug in Java. In the
-example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
-named as \texttt{m} which are either abstract or default methods. If we define a
-new interface \texttt{M} that extends two of these interfaces, then the method
-overriding result is shown in the table. For example, row2 col2 means \texttt{M
-  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
-\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
-  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
-following this way of interpretation. The interesting case in row4 col2 where
-\texttt{M extends A1,B3}, the result should be \texttt{ERROR}, however in Java
-the experimental result is \texttt{B3.m}. Because if we think consistently, this
-case is esencially a method conflicting situation with abstract methods and
-default methods, which is no different than the first case.
+We are very confused on the result for \textbf{B3},\textbf{A1}, and we can not find any rational for this behaviour.
+\footnote{We refer to the behaviour of \texttt{javac} as in JDK version is\marco{add}.}
 
-\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
+%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+%named as \texttt{m} which are either abstract or default methods. If we define a
+%new interface \texttt{M} that extends two of these interfaces, then the method
+%overriding result is shown in the table. For example, row2 col2 means \texttt{M
+%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+%following this way of interpretation.
 
-In our approach, we overcome this bug by detecting method correctness when
-typechecking all interfaces. Internally, algorithm in the auxiliary function
-$\mBody(\m,\C)$ apply the correct method overriding mechanism and returns the
-correct method signature.
+% The interesting case in row4 col2 where
+To be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
+We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
+We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
+%the experimental result is \texttt{B3.m}. Because if we think consistently, this
+
+%\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+In our approach, we choose to not model this strange behaviour (a bug?).
+Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
+The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java change its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.

### 2015-11-15 paper/sections/Implementation.tex
Minor changes to implementation section. I think it was already pretty good

@@
 ``Outline'' window, including all the methods inside the interface as well as
-the generated ones. The custom transformation is easy and convenient to use.
+the generated ones.
+ The custom transformation is easy and convenient to use.
+ For example this means that the IDE functionality 
+ content assist autocomplete will work for the newly generated methods.
 
 The biggest reasons to use Lombok rather than using conventional Java annotation
-processor are:
+processor is
 \begin{itemize}
-\item Lombok is byte-code based instead of source-code based, which makes client
-  code concise and easy to maintain. Such code generation is performed at
-  compile time to modify bytecode. And during compile time, these methods are
-  only generated in class files, with neither the source code modified nor new
-  Java files generated.
+%this seamd like a commercial spot
+%\item Lombok is byte-code based instead of source-code based, which makes client
+%  code concise and easy to maintain. Such code generation is performed at
+%  compile time to modify bytecode.
+\item Lombock modify the generation process of the class files.
+neither the source code is modified nor new  Java files are generated.
 \item Moreover, and probably more importantly, Lombok is capable of generating
   code \emph{inside} a class/interface.This is the ability that conventional
-  Java annotation processor can not provide.
+  Java annotation processor do not provide.
 \end{itemize}
 
-The implementation using Lombok has certain limitation: 
+Our prototype implementation using Lombok has certain limitation: 
 \begin{itemize}
-\item Our implementation does not support separate compilation yet at this
+\item We do not support separate compilation yet at this
   stage, for in the implementation it is hard to capture a type declaration from
   its reference.
-\item At this stage our implementation only realizes the Eclipse handler and our
-  experiments are all conducted in Eclipse.
+\item At this stage our implementation only realizes the Eclipse handler and our experiments are all conducted in Eclipse.
 \end{itemize}

### 2015-11-15 paper/sections/Compare.tex
Formatting fix in compare

@@
 
-We now compare our approach to the original operations on traits, showing how are supported by our model.\footnote{This comparison shares  similarities with the encoding shown in \cite{bono14};
-we differ in the 
-proposed representation of
+A comparison on how to to emulate the original operations on traits using Java8 can be find in\cite{bono14}.
+We briefly recall the main points of their encoding; however we propose a different representation of
  \textbf{exclusion}.
 The author of ~\cite{bono14} agree that our revised version is cleaner, typesafe and more direct.
-}
 
+\newcommand\shortItem{\vspace{-1ex}}
 \begin{itemize}
+\begin{comment}%I shoreten up a lot after. I think this version is better but longer, and this content is just repeating bono at all.
 \item \textbf{Symmetric sum}: the symmetric composition of two disjoint traits is achieved by simply implementing two interfaces in Java correspondingly, without overriding any method.
@@
     \end{lstlisting}
-\item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces. See the example below.
+\end{comment}
+
+\item \textbf{Symmetric sum} can be obtained by simple multiple inheritance between interfaces.
+\shortItem
+    \begin{lstlisting}
+    interface A { int x(); }    interface B { int y(); }    interface C extends A, B {}
+    \end{lstlisting}
+\shortItem
+\item \textbf{Overriding} a conflict is obtained by specifying which super interface take precedence.
+\shortItem
+    \begin{lstlisting}
+    interface A { default int m() {return 1;} } 
+    interface B { default int m() {return 2;} }
+    interface C extends A, B { default int m() {return B.super.m();} }
+    \end{lstlisting}
+\shortItem
+\item \textbf{Alias} is creating  a new method delegating to the existing super interface.
+\shortItem
+    \begin{lstlisting}
+    interface A { default int m() {return 1;} }
+    interface B extends A { default int k() {return A.super.m();} }
+    \end{lstlisting}
+\shortItem
+
+\item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces.
+\shortItem
     \begin{lstlisting}
@@
     \end{lstlisting}
+\shortItem
 \end{itemize}

### 2015-11-15 paper/sections/RelatedWork.tex
This type improved

@@
 \subsection{ThisType/MyType/Extensibility}
-\marco{ this section is unfair.
+%\marco{ this section is unfair.
 There are unsolved problems with ThisType in negative positions, namelly it breaks subtyping.
@@
 This demolish most advantages of interfaces.}
-Typical object-oriented programming languages (e.g. Java) use class names as
-types and inheritance as subtyping relations. This gives difficulty when
-extending classes/interfaces with recursive types. Problems such as binary
-methods, return type refinement, etc. emerge and hinders software extensibility
-greatly.
 
-\emph{MyType}~\cite{bruce1994paradigmatic} was proposed by Bruce and Foster,
-which represents the type of \textbf{this} and changes its meaning along with
-inheritance. The addition of MyType to a language will allow easy definition of
-binary methods, methods with recursive types (i.e., the same type of the
-receiver appears in the argument or return positions of methods), etc. MyType
-greatly enhances the expressiveness and extensibility of object-oriented
-programming languages.
+In certain situations, our approach allows automatic refinement for return types.
+This is part of a bigger topic in class based languages: expressing and preserving type recursion and (nominal/structural) subtyping at the same time.
 
-In our approach, we are using covariant-return types to simulate some uses of
-MyType. But our approach only works on method positive positions, whereas MyType
-is more general, as it works on any positions. Nevertheless our approach is
-still useful for modeling fluent interfaces and solving expression
-problems,etc.
+One famous attempt in this direction is provided by 
+\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of \textbf{this}, changing its meaning along with
+inheritance.
+However when invoking a method with MyType in a parameter position, the exact type of the receiver must be known.
+This is huge limitation in class based object oriented programming,
+and is exasperated by interface-only programming as we propose: no type is ever exact since classes are not part of the language.
+A recent article~\cite{igarashi Matching MyType to subtyping}
+lights up this topic, since
+they propose two features: exact statements and non-inheritable methods; both related to our work:
+any method generated inside of the \Q@of@ method is indeed non-inheritable, since there is no class name to extends from, and exact statements (a form of wild-card capture on the exact run-time type)
+could capture the ``exact type'' of an object even in a class-less environment.
+
+% The addition of MyType to a language will allow easy definition of
+%binary methods, methods with recursive types (i.e., the same type of the
+%receiver appears in the argument or return positions of methods), etc. MyType
+%greatly enhances the expressiveness and extensibility of object-oriented
+%programming languages.
+
+%In our approach, we are using covariant-return types to simulate some uses of
+%MyType. But our approach only works on method positive positions, whereas MyType
+%is more general, as it works on any positions. Nevertheless our approach is
+%still useful for modeling fluent interfaces and solving expression
+%problems,etc.
 
@@
 
-\subsection{Private state/state initialization}
+\subsection{Private state}
 The biggest limitation of our approach is the absence in Java8 of support for private/protected methods in interfaces.That is, in Java8 all members of interfaces are public, including static methods.
@@
 In this extension,  also the of method could be made private following the same pattern.
+
+\subsection{State initialization}
+
+As discuss before, the user can trivially define its own \Q@of@ method, and initialize a portion of the state with default values.
+However, the initialization code would not be reused/reusable, and subinterfaces would have to repeat such initialization code.
+If a field has no setters, a simple alternative is to just define the ``field'' as a default method as in 
+\begin{lstlisting}
+@Mixin interface Box{ default int val(){return 0;} }
+\end{lstlisting}
+if setters are required, a possible extension of our code expansion could recognize a field if the getter is provided and the setter is required, and could generate the following code:
+\begin{lstlisting}
+interface Box{ 
+  default int val(){return 0;} 
+  void val(int _val);
+  static Box of(){return new Box(){
+    int val=Box.super.val();
+    int val(){return val;}
+    void val(int _val){val=_val;}
+    };}}
+\end{lstlisting}
+We are unsure of the value of this solution: is very tricky, the user define a method that (contrary to our usual expectation) is actually overridden in a way that the behaviour change, but change only after the first setter is called, plus this code would cache the result instead of re-computing it every time.
+This can be very relevant and tricky in a non functional setting.
 

### 2015-11-15 paper/sections/RelatedWork.tex
Some more work on last part of related work

@@
 \subsection{Type-Directed Translations/Syntactic Sugar}
+\marco{I'm tring to merge this and the next one}
 Language extensions are often implemented as syntactic sugar of the base
@@
 
-\subsection{Meta-programming techniques}
+\subsection{Meta-programming competes with Language extensions}
+The most obvious solution to add features to a language, is language extension.
+They are often implemented as syntactic sugar of the base
+language.
+For example, the Scala compiler was extended to directly supports XML syntax.
+However, this approach does not support combining multiple extensions into one.
+We are de facto creating a fork in the language, and rarely the new fork gain enough traction to become the main language release.
+On this topic we mention SugarJ~\cite{erdweg11sugarj}; a Java-based extensible language allowing programmers to extend it with custom features by definitions in meta-DSLs (SDF, Stratego, etc).
+
+On the other side, when the starting language have a flexible enough syntax and a fast and powerful enough reflection, we may just need to play with operator overloading and other language tricks to discover that the language feature we need can be expressed a a simple library in our language.
+For example, consider SQL alchemy in python.
+
+Java-like language tends to sits in the middle of this two extremes:
+libraries can not influence the type system, so many solutions valid in python or other languages could not be applicable, or may be applicable at the cost of loosing safety.
+
+We can model certain types of language extensions with annotations 
+only, but those extensions do not introduce new syntax: they 
+merely do automatic code generation.
+
 - MetaFJig*~\cite{}

### 2015-11-16 paper/sections/Translation.tex
First draft of translation and proof.

@@
 
-\noindent\textbf{LEMMA. }
+\noindent\textbf{LEMMA 1. }
+If $\II_0$ satisfies $\dom(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom'(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)\cup\dom(\methods')$.
+\begin{proof}
+It suffices to prove $\dom'(C_0)=\dom(C_0)\cup\dom(\methods')$.
+\begin{itemize}
+\item If $m\notin\dom(\methods')$, by the definition of $\mBody$, it is obtained from $\override$.
+The first argument of $\override$, namely $\methods\cup\methods'(m)$, is equal to $\methods(m)$, and the second argument is not changed as well, thus the result of $\override$ is not changed during translation.
+Hence $m\in\dom'(C_0)$ iff $m\in\dom(C_0)$.
+\item If $m\in\dom(\methods')$, we are to prove $m\in\dom'(C_0)$.
+So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.
+On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
+Hence $m\in\dom'(C_0)$ as well.
+\end{itemize}
+\end{proof}
+
+\noindent\textbf{LEMMA 2. }
 If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
@@
 
-\noindent\textbf{Part II.} Next we check $\dom(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)\cup\dom(\methods')$.
+\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)\cup\dom(\methods')$$
+And actually it is guaranteed by \textbf{LEMMA 1}.
 
+\end{proof}
+
+\begin{proof}
+\textbf{(THEOREM) LEMMA 2} already proves that $[\![\II_0]\!]$ is OK.
+On the other hand, if some $\II$ is not a subtype of $\II_0$, the generated code in translation has no way to affect the domain of $\II$, which finishes our proof.
 \end{proof}

### 2015-11-16 paper/sections/Translation.tex
proof in appendix.

--- 2015-11-15 paper/sections/Translation.tex
@@
 
-\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC \methods \cC ]\!] =
+\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!] =
 \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
@@
 \ofMethod(\C_0) = \
- \QM{static}\ C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
+ \QM{static}\ \C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
 \QM{\{}
-\QM{return new}\ C_0 \oR\cR\ \QM{\{} \\
-\tab\tab\tab\tab\tab\tab\tab C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
+\QM{return new}\ \C_0 \oR\cR\ \QM{\{} \\
+\tab\tab\tab\tab\tab\tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
 \tab\tab\tab\tab\tab\tab\tab
@@
 %and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
+\haoyuan{Should we include the $\withMethod(\C_0)$ in the formalization?}
 
@@
 
-\haoyuan{method name of setter?}
+\haoyuan{method name of setter? $m$ or $\_m$?}
 
@@
 
-\noindent\textbf{LEMMA. }
+To prove the theorem we introduce two lemmas below. The complete proof is available in Appendix \ref{subsec:proof1}, \ref{subsec:proof2} and \ref{subsec:proof3}.\\
+
+\noindent\textbf{LEMMA 1. }
+If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom'(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.\\
+
+\noindent\textbf{LEMMA 2. }
 If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
-\begin{proof}
-First assume that $\II_0$ is of form $$\mixinAnn\ \QM{interface}\ C_0\ \QM{extends}\ \overline{C}\ \oC \methods\ \cC$$
-By the rule \rn{t-Invk} in Figure~\ref{ET}, we divide it into two parts.
-
-\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\II_0]\!]$, the type of the return value is compatible with the method's return type.
-
-Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(C_0)$.
-
-By the definition of $\ofMethod(C_0)$, the return value is an object $$\QM{return new}\ C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
-To prove it is of type $C_0$, we use the typing rule \rn{t-Obj}.
-\begin{itemize}
-\item For field typing,
-\[\Gamma(\QM_ m_i)=C_i\subtype C_i\]
-\item For method typing of getters,
-\[\Gamma, m_i:C_i, \QM{this}:C_0, \Gamma^{\mh_i}\vdash m_i\in C_i=C^{\mh_i}\]
-\item For method typing of nonempty \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
-\[\Gamma, \overline{m_i:C_i}, \QM{this}:C_0, \es_i:\overline{C}_i, \Gamma^{\mh_i}\vdash C_0\QM{.of}\oR\es_i\cR\in C_0=C^{\mh_i}\]
-\item For method typing of nonempty setter methods, \haoyuan{what about the \QM{with} method?}
-\[\Gamma, \QM{this}:C_0, \Gamma^{\mh_i}\vdash\QM{this}\in C_0=C^{\mh_i}\]
-\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, C_0)$,
- \begin{itemize}
- \item For getters, \[C_i\ m_i\oR\cR\QM; \in \fieldsFunc(C_0) \Rightarrow C_i\ m_i\oR\cR\QM; \in \dom(C_0)\]
- \item For nonempty \QM{with-} methods,
-  \[\begin{array}{l}
-  \mBody(\QM{with#}\m_i,C_0) \mbox{ is of form }\mh\QM; \\ \valid(C_0)
-  \end{array}\Rightarrow \isWith\Rightarrow C_0\subtype\mBody(m^{\mh_i}, C_0)\]
- \item For nonempty setters,
-  \[\begin{array}{l}
-  \mBody(\m_i,C_0) \mbox{ is of form }\mh\QM; \\ \valid(C_0)
-  \end{array}\Rightarrow \isSetter\Rightarrow C_0\subtype\mBody(m^{\mh_i}, C_0)\]
- \end{itemize}
-\item To prove that such a created object indeed implements all the abstract methods in $\dom(C_0)$, we simply refer to $\valid(C_0)$, since it guarantees each abstract method $\method$ to satisfy $\isField$, $\isWith$ or $\isSetter$.
-But that object includes all implementations for those cases, hence it is of type $C_0$ by \rn{t-Obj}.
-\end{itemize}
-
-\noindent\textbf{Part II.} Next we check $\dom(C_0)=\dom(C_1)\cup\ldots\cup\dom(C_n)\cup\dom(\methods)\cup\dom(\methods')$.
-
-\end{proof}
 

### 2015-11-16 paper/sections/RelatedWork.tex
Discussion on MetaFjig

@@
 
-We can model certain types of language extensions with annotations 
-only, but those extensions do not introduce new syntax: they 
-merely do automatic code generation.
+Here (compile/load time) code generation come at the rescue: 
+if for a certain feature (\mixin in our case) it is possible to use the original language syntax to
+\emph{express-describe} any specific instantiation of such feature
+(annotating a class and provide getters), then we can insert in the compilation process a tool that exam and enrich the code before compilation.
+No need to modify the original, source, we can work on temporary enriched source.
+Java is a particular good candidate for this kind of manipulation since it already provide ways to define and integrate such tools in its own compilation process.
 
-- MetaFJig*~\cite{}
-\yanlin{Maybe Marco can help with the text for MetaFJig?}
+Other languages offers even stronger support to safe code manipulation:
+Template Haskell~\cite{}, F\# (type providers)~\cite{} and MetaFjig (Active Libraries)~\cite{}
+all allows to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
+In particular, MetaFjig offers a property called \emph{meta-level-soundness}.
+In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code.
+This is roughly equivalent to what we state and manually proof in Lemma 2 for our particular transformation.
+Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.
 

### 2015-11-17 paper/sections/RelatedWork.tex
added references. revise related work.

--- 2015-11-15 paper/sections/RelatedWork.tex
@@
 \subsection{Multiple Inheritance in Object oriented languages}
-Many authors have argued in favour or against multiple inheritance.
-It provides expressive power, but it is difficult to model, and implement and can create programs that are hard to reason about.
-These difficulties include the famous
-diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, and the yo-yo problem~\cite{taenzer1989problems}.
-\marco{add citation on the text over.}
-\yanlin{added.}
-To conciliate the need for expressive power and the need for simplicity, 
-lots of models have been proposed in the past few
-years, including mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
-They provide novel programming architecture models in the
-OO paradigm. 
+Many authors have argued in favour or against multiple inheritance.  It provides
+expressive power, but it is difficult to model, and implement and can create
+programs that are hard to reason about.  These difficulties include the famous
+diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods,
+and the yo-yo problem~\cite{taenzer1989problems}.  To conciliate the need for
+expressive power and the need for simplicity, lots of models have been proposed
+in the past few years, including mixins~\cite{bracha90mixin} and
+traits~\cite{scharli03traits}.  They provide novel programming architecture
+models in the OO paradigm.
+
 \begin{enumerate}
 \item Mixins allows to name components that can be applied to various classes as
-reusable functionality units. However,  they suffer from linearisation: the order of mixin application is relevant in often subtitle and undesired ways. constraints. This hinders their
-usability and their ability of resolving conflicts:
- the linearisation (total ordering) of mixin inheritance cannot provide a satisfactory resolution in all cases and restricts the flexibility of mixin composition.
+  reusable functionality units. However, they suffer from linearisation: the
+  order of mixin application is relevant in often subtitle and undesired
+  ways. constraints. This hinders their usability and their ability of resolving
+  conflicts: the linearisation (total ordering) of mixin inheritance cannot
+  provide a satisfactory resolution in all cases and restricts the flexibility
+  of mixin composition.
 
- To fight those limitations, an algebra of mixin operators is introduced~\cite{CMS-elena-zucca-davide-ancona}, but this raised the complexity of the approach, especially when constructors and fields are considered~\cite{Ecoop-FJigsaw-marco}.
+  To fight those limitations, an algebra of mixin operators is
+  introduced~\cite{ancona2002calculus}, but this raised the complexity of the
+  approach, especially when constructors and fields are
+  considered~\cite{marco09FJigsaw}.
 
-\item 
-Simplifying the mixin algebras approach, traits draw a strong line between units of reuse (traits) and object factories (classes)
-In this model, traits~\cite{Scharli2003} units of reusable code,  containing only methods as
-reusable functionalities. Thus, no state/state initialization is considered.
+\item Simplifying the mixin algebras approach, traits draw a strong line between
+  units of reuse (traits) and object factories (classes) In this model,
+  traits~\cite{scharli03traits} units of reusable code, containing only methods
+  as reusable functionalities. Thus, no state/state initialization is
+  considered.
 
-Classes act as object factories, requiring functionalities from multiple traits.
-Traits offers a trait algebra  with operations like sum, alias and exclusion, provided for explicit conflict resolution.
+  Classes act as object factories, requiring functionalities from multiple
+  traits.  Traits offers a trait algebra with operations like sum, alias and
+  exclusion, provided for explicit conflict resolution.
 
-Concluding, (pure) traits do not allow state and they do not offers any reuse instrument to ensure that state is coherently initialized when finally defined in classes.
-Traits can't be instantiated. This model requires two concepts (traits and classes) to coexist and cooperate.
+  Concluding, (pure) traits do not allow state and they do not offers any reuse
+  instrument to ensure that state is coherently initialized when finally defined
+  in classes.  Traits can't be instantiated. This model requires two concepts
+  (traits and classes) to coexist and cooperate.
 
-Some authors see this as good language design fostering good software development by helping programmers to think about the structure of their programs.
-However, other authors see the need of two concepts and the absence of state as drawbacks of this model.
+  Some authors see this as good language design fostering good software
+  development by helping programmers to think about the structure of their
+  programs.  However, other authors see the need of two concepts and the absence
+  of state as drawbacks of this model.
 
-\item C++ and Scala also try to provide solutions to multiple inheritance, but both
-suffer from object initialization problems. Virtual inheritance in C++ provides
-another solution to multiple inheritance (especially the diamond problem by
-keeping only one copy of the base class)~\cite{ellis1990annotated}, however
-suffers from object initialization problem as pointed out by Malayeri et
-al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
+\item C++ and Scala also try to provide solutions to multiple inheritance, but
+  both suffer from object initialization problems. Virtual inheritance in C++
+  provides another solution to multiple inheritance (especially the diamond
+  problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
+  however suffers from object initialization problem as pointed out by Malayeri
+  et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
   superclasses, which would potentially cause serious semantic error. Scala
-solution (very similar to linearised mixins, but misleadingly called traits in the language) avoids this problem by disallowing constructor parameters,
-causing no ambiguity in cases such as diamond problem. 
-This approach has limited expressiveness, and suffers from all the problems of linearised mixin composition.
+  solution (very similar to linearised mixins, but misleadingly called traits in
+  the language) avoids this problem by disallowing constructor parameters,
+  causing no ambiguity in cases such as diamond problem.  This approach has
+  limited expressiveness, and suffers from all the problems of linearised mixin
+  composition.
   Python also offers multiple inheritance via linearised mixins. Indeed python any class is implicitly a mixin, and mixin composition informally expressed as\\*
@@
 \subsection{Multiple Inheritance in Java}
-Since Java 8 default methods are introduced, which enables concrete method
-implementation to be defined (via the \textbf{default} keyword) inside
-interfaces. Since Java supports implementation of multiple interfaces (instead of extension of a single
-class), the introduction of default methods opens the gate for various flavours of multiple
-inheritance in Java, using interfaces. 
-Former work by Bono.et. al.~\cite{bono14}.
-provides details on mimicking traits~\cite{sharly-thesis} through interfaces.
+Since Java 8 default methods are introduced, concrete method implementation are
+allowed to be defined (via the \textbf{default} keyword) inside
+interfaces. Since Java supports implementation of multiple interfaces (instead
+of extension of a single class), the introduction of default methods opens the
+gate for various flavours of multiple inheritance in Java, using interfaces.
+Former work by Bono.et. al.~\cite{bono14}. provides details on mimicking traits
+through interfaces.
 
-There are proposals for extending Java (before Java8) with traits. For example, FeatherTrait
-Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the calculus of
-Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
-trait-based inheritance in Java.
-Except for few, mostly syntactic details, their work can be emulated/rephrased with Java8 interface.
+There are proposals for extending Java (before Java8) with traits. For example,
+FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the
+calculus of Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed
+traits, adding trait-based inheritance in Java.  Except for few, mostly
+syntactic details, their work can be emulated/rephrased with Java8 interface.
 
@@
 One famous attempt in this direction is provided by
-\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of \textbf{this}, changing its meaning along with
-inheritance.
-However when invoking a method with MyType in a parameter position, the exact type of the receiver must be known.
-This is huge limitation in class based object oriented programming,
-and is exasperated by interface-only programming as we propose: no type is ever exact since classes are not part of the language.
-A recent article~\cite{igarashi Matching MyType to subtyping}
-lights up this topic, since
-they propose two features: exact statements and non-inheritable methods; both related to our work:
-any method generated inside of the \Q@of@ method is indeed non-inheritable, since there is no class name to extends from, and exact statements (a form of wild-card capture on the exact run-time type)
-could capture the ``exact type'' of an object even in a class-less environment.
+\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
+\textbf{this}, changing its meaning along with inheritance.  However when
+invoking a method with MyType in a parameter position, the exact type of the
+receiver must be known.  This is huge limitation in class based object oriented
+programming, and is exasperated by interface-only programming as we propose: no
+type is ever exact since classes are not part of the language.  A recent
+article~\cite{Saito2013933} lights up this topic, since they propose two
+features: exact statements \yanlin{do we support exact statements?} and
+non-inheritable methods; both related to our work: any method generated inside
+of the \Q@of@ method is indeed non-inheritable, since there is no class name to
+extends from, and exact statements (a form of wild-card capture on the exact
+run-time type) could capture the ``exact type'' of an object even in a
+class-less environment.
 

### 2015-11-17 paper/sections/Appendix.tex

\section{Appendix}\label{sec:appendix}

\subsection{Proof of LEMMA 1}\label{subsec:proof1}
\begin{proof}
First for convenience, assume that $\II_0$ is of form $$\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC$$
It suffices to prove $\dom'(\C_0)=\dom(\C_0)\cup\dom(\methods')$.
\begin{itemize}
\item If $m\notin\dom(\methods')$, by the definition of $\mBody$, it is obtained from $\override$. The first argument of $\override$, namely $\methods\cup\methods'(m)$, is equal to $\methods(m)$, and the second argument is not changed as well, thus the result of $\override$ is not changed during translation. Hence $m\in\dom'(\C_0)$ iff $m\in\dom(\C_0)$.
\item If $m\in\dom(\methods')$, we are to prove $m\in\dom'(\C_0)$. So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$. On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship. Hence $m\in\dom'(\C_0)$ as well.
\end{itemize}
\end{proof}

\subsection{Proof of LEMMA 2}\label{subsec:proof2}
\begin{proof}
By the rule \rn{t-Invk} in Figure~\ref{ET}, we divide it into two parts.

\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\II_0]\!]$, the type of the return value is compatible with the method's return type.

Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.

By the definition of $\ofMethod(\C_0)$, the return value is an object $$\QM{return new}\ \C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
To prove it is of type $\C_0$, we use the typing rule \rn{t-Obj}.
\begin{itemize}
\item For field typing,
\[\Gamma(\QM_ m_i)=\C_i\subtype \C_i\]
\item For method typing of getters,
\[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i=\C^{\mh_i}\]
\item For method typing of nonempty \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
\[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0=C^{\mh_i}\]
\item For method typing of nonempty setter methods, \haoyuan{what about the \QM{with} method?}
\[\Gamma, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash\QM{this}\in \C_0=\C^{\mh_i}\]
\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
 \begin{itemize}
 \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
 \item For nonempty \QM{with-} methods,
  \[\begin{array}{l}
  \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isWith\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \item For nonempty setters,
  \[\begin{array}{l}
  \mBody(\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isSetter\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \end{itemize}
\item To prove that such a created object indeed implements all the abstract methods in $\dom(\C_0)$, we simply refer to $\valid(\C_0)$, since it guarantees each abstract method $\method$ to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
\end{itemize}

\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$$
And actually it is guaranteed by \textbf{LEMMA 1}.

\end{proof}

\subsection{Proof of THEOREM}\label{subsec:proof3}
\begin{proof}
\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$, the generated code in translation has no way to affect the domain of $\II$, which finishes our proof.
\end{proof}

### 2015-11-17 paper/sections/Translation.tex

@@
            \Q@public @$\C_0\ $\Q@with(@$\C$\Q@ _val) { @\\*
-           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;@\\*
-           \Q@return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
+           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;}@\\*
+${}_{}$\Q@  return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
 where with $\m_1\ldots\m_n$  fields of $\C_0$,

### 2015-11-17 paper/sections/Appendix.tex
Typo.

@@
 \begin{proof}
-First for convenience, assume that $\II_0$ is of form $$\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC$$
-It suffices to prove $\dom'(\C_0)=\dom(\C_0)\cup\dom(\methods')$.
+It suffices to prove $\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\methods')$.
 \begin{itemize}
@@
 The first argument of $\override$, namely $\methods\cup\methods'(m)$, is equal to $\methods(m)$, and the second argument is not changed as well, thus the result of $\override$ is not changed during translation.
-Hence $m\in\dom'(\C_0)$ iff $m\in\dom(\C_0)$.
-\item If $m\in\dom(\methods')$, we are to prove $m\in\dom'(\C_0)$.
+Hence $m\in\dom([\![\II_0]\!])$ iff $m\in\dom(\C_0)$.
+\item If $m\in\dom(\methods')$, we are to prove $m\in\dom([\![\II_0]\!])$.
 So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.
 On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
-Hence $m\in\dom'(\C_0)$ as well.
+Hence $m\in\dom([\![\II_0]\!])$ as well.
 \end{itemize}
@@
 \begin{proof}
-By the rule \rn{t-Invk} in Figure~\ref{ET}, we divide it into two parts.
+By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide it into two parts.
 

### 2015-11-17 paper/sections/Appendix.tex
Comments.

@@
 \subsection{Proof of LEMMA 1}\label{subsec:proof1}
+\textbf{LEMMA 1. }
+Assume that
+\begin{align*}
+\II_0\ &=\ \mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
+[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \methods' \cC
+\end{align*}
+If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.
 \begin{proof}
@@
 \subsection{Proof of LEMMA 2}\label{subsec:proof2}
+\textbf{LEMMA 2. }
+If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
 \begin{proof}
@@
 
+\haoyuan{How to more explicitly say the default methods are ok in the new interface table, so that it is sufficient to only prove the generated ofMethod is ok?}
+
 By the definition of $\ofMethod(\C_0)$, the return value is an object $$\QM{return new}\ \C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
 To prove it is of type $\C_0$, we use the typing rule \rn{t-Obj}.
+
+\haoyuan{The order of conditions in \rn{t-Obj} is not clear below.}
+
 \begin{itemize}
@@
 \[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i=\C^{\mh_i}\]
-\item For method typing of nonempty \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
+\item For method typing of \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
 \[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0=C^{\mh_i}\]
-\item For method typing of nonempty setter methods, \haoyuan{what about the \QM{with} method?}
+\item For method typing of setter methods, \haoyuan{what about the \QM{with} method?}
 \[\Gamma, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash\QM{this}\in \C_0=\C^{\mh_i}\]
@@
  \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
- \item For nonempty \QM{with-} methods,
+ \item For \QM{with-} methods,
   \[\begin{array}{l}
@@
   \end{array}\Rightarrow \isWith\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
- \item For nonempty setters,
+ \item For setters,
   \[\begin{array}{l}
@@
 
-\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$$
+\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$$
 And actually it is guaranteed by \textbf{LEMMA 1}.
@@
 \subsection{Proof of THEOREM}\label{subsec:proof3}
+\textbf{THEOREM. }
+For a given $\II_0\ldots\II_n$ interface table such that
+$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
 \begin{proof}

### 2015-11-17 paper/sections/Translation.tex
Typo.
Comments.
Text for theorem.

@@
 
-To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer
-\marco{should we repeat the 3 points in the email?}\\
+\haoyuan{In theorem: $I_0$ is valid and $\QM{of}\notin\dom{I_0}$. To prove the theorem we need an extra lemma: if the translation is well defined, then $[\![\II_0]\!]$ is well formed.}
+
+To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer for safety:
+\begin{itemize}
+\item \textit{Self coherence}: the generated code itself is well-formed; there are no type errors.
+\item \textit{Client coherence}: the client code doesn't fail if it was not failing before code generation.
+\item \textit{Heir coherence}: the heirs are safe. 
+\end{itemize}
+The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed.
+In Java it implies not touching the annotated interface, and surely an alternative approach is to create a new interface and generate what we need.
+But since we are using Lombok for the translation, the transformation is applied to the annotated interface itself.
+But heir coherence is indeed an issue.
+See below how this is affected by our approach:
+
+\begin{lstlisting}
+interface A { int x(); A withX(int x); }
+@Mixin interface B extends A {}
+interface C extends B { A withX(int x); }
+\end{lstlisting}
+
+Here is the client code for our example.
+With interface \Q@B@ annotated with \mixin, by the translation rule, it will generate a method ``\Q@B withX(int x);@'' inside, as an automatic return type refinement.
+But in that case, interface \Q@C@ doesn't type check.
+
+Now we go back and read the statement of our theorem again.
+It illustrates that a program with the translation of \mixin applied still type checks if it did, except for the subtypes of the annotated interface (type safety of them is not guaranteed).
+
+\haoyuan{Text needs improving.}
 
@@
 \noindent\textbf{LEMMA 1. }
-If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom'(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.\\
+Assume that
+\begin{align*}
+\II_0\ &=\ \mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
+[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \methods' \cC
+\end{align*}
+If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.\\
 

### 2015-11-17 paper/sections/Overview.tex

@@
 
-\section{Overview}\label{sec:ep}
+\section{A Running Example: Animals}\label{sec:ep}
 
@@
 
-\subsection{A Running Example: Animals}
-\marco{fig1 is good but is very big, not sure if it worth its space. If we keep it, we need to do it with a vectorial graphic, or better inside latex (for example with tikz)}
+\marco{fig1 is now better. Where we cite it in the text? can we make it so that is more "linear"? point2d <- point3D,   animal <- (horse over bird ) <- pegasus}
 To propose a standard example, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
@@
 %--------------
-\subsubsection{\Q@Point2D@: simple immutable data with fields}
+\subsection{\Q@Point2D@: simple immutable data with fields}
 Since in ClassLess Java there are no classes, our \Q@Point2D@ must be an interface, as for example:
@@
 
-By annotating the interface \Q@Point2D@, the annotation will generate the shown static method \texttt{of}, mimicking the functionality of a simple minded constructor:
+By annotating the interface \Q@Point2D@, the annotation will generate a variation of the shown static method \texttt{of}, mimicking the functionality of a simple minded constructor:
 By looking the the methods that need implementation it first detects what are the fields, then 
@@
 \end{lstlisting}
-is expanded by our annotation into
+is expanded by our annotation into\footnote{
+Note how we actually generate a real field \Q@int x=_x;@.
+This provide a more uniform translation that can work also for mutable datastructures, where setters are required.
+}
 
@@
 
-\subsubsection{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
+\subsection{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
 \Q@Points2D@ are mathematical entities, thus we chosen to use immutable data structure to model them; however animals are real world entities, and when an animal moves, it is the same animal that have now a different location.
@@
 
-\subsubsection{\Q@Bird@:  field type refinement}
+\subsection{\Q@Bird@:  field type refinement}
+
+\subsubsection*{\Q@Point3D@ and properties updater}
 \Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
@@
 
+\subsubsection*{Unsatisfactory class based solutions to field type refinement}
+
 \Q@Bird@s are \Q@Animal@s, 
@@
 
-Many here would think that a language extension is needed.
-Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}:
+Many readers would now think that a language extension is needed.
+Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}.
+\subsubsection*{Our approach to field type refinement}
 
@@
 
-Note how
+In our proposed solution,
 \begin{itemize}
-\item by covariant method overriding we refine the type of our field
+\item by covariant method overriding we refine the field type
 \item by \emph{overloading} we define a new setter with the more precise type
@@
 
-\subsubsection{\Q@Pegasus@ and multiple Inheritance}
+\subsection{\Q@Pegasus@ and multiple Inheritance}
 The \emph{``multiple inheritance''} case appears at interface

### 2015-11-17 paper/sections/RelatedWork.tex

@@
 article~\cite{Saito2013933} lights up this topic, since they propose two
-features: exact statements \yanlin{do we support exact statements?} and
+features: exact statements \yanlin{do we support exact statements?}\marco{no we do not. When is that it may seams we do?} and
 non-inheritable methods; both related to our work: any method generated inside
@@
 
+\begin{comment}
 \subsection{Type-Directed Translations/Syntactic Sugar}
@@
 merely do automatic code generation. 
-
+\end{comment}
 \subsection{Meta-programming competes with Language extensions}
-The most obvious solution to add features to a language, is language extension.
-They are often implemented as syntactic sugar of the base
+The most obvious solution to add features to a language is language extension.
+They are often implemented as syntactic extensions that can be desugared to the base
 language.
@@
 
-On the other side, when the starting language have a flexible enough syntax and a fast and powerful enough reflection, we may just need to play with operator overloading and other language tricks to discover that the language feature we need can be expressed a a simple library in our language.
+On the other side, when the starting language have a flexible enough syntax and a fast and powerful enough reflection, we may just need to play with operator overloading and other language tricks to discover that the language feature we need can be expressed as a simple library in our language.
 For example, consider SQL alchemy in python.
@@
 (annotating a class and provide getters), then we can insert in the compilation process a tool that exam and enrich the code before compilation.
-No need to modify the original, source, we can work on temporary enriched source.
-Java is a particular good candidate for this kind of manipulation since it already provide ways to define and integrate such tools in its own compilation process.
+No need to modify the original source; for example we can work on temporary files.
+Java is a particular good candidate for this kind of manipulation since it already provide ways to define and integrate such tools in its own compilation process: in this way there is no need of temporary files, and there is a well defined way of putting multiple extensions together.
 
@@
 
+\marco{text under still to revise. Sould all the lombock discussion go in the implementation section? -- just to keep stuff together?}
 Annotations/Code Generation/Lombok
@@
 interface Box{ 
-  default int val(){return 0;} 
-  void val(int _val);
-  static Box of(){return new Box(){
+  default int val(){return 0;} //provided
+  void val(int _val);//provided
+  static Box of(){return new Box(){//generated
     int val=Box.super.val();
@@
 \subsection{Class invariants in ClassLess Java}
-Since the objects are created by automatically generate methods,
+Since the objects are created by automatically generated methods,
 another limitation of our current approach is that there is no place where the user can dynamically check for class invariants.
@@
 Methods originally defined in \Q@Object@, as \Q@clone@ and \Q@toString@ can be supported by our approach, but they need special care.
-If an interface annotated with \mixin ask an implementation for \Q@clone@, \Q@toString@ \Q@equls@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters.
+If an interface annotated with \mixin ask an implementation for \Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters.
 Note how this would solve most of the Java ugliness related to \Q@clone@ methods.}

### 2015-11-18 paper/sections/Conclusion.tex
Minor changes in Introduction

--- 2015-11-10 paper/sections/Overview.tex
@@
-\marco{an improvement of this may become part of intro?}
+\section{Conclusion}\label{sec:conclusion}
 
-Before Java 8, concrete methods and static methods where not allowed to appear in interfaces.
-Java 8 allows static interface methods and introduces
-\emph{default methods}, which allow for implementation insides interfaces.
-This had an important positive consequence that was probably overlooked by the Java design team: the concept of class (in java) is now redundant and unneeded.
-We define a subset of Java, called ClassLess Java, where programs and (reusable) libraries can be easily defined and used.
-To avoid for some syntactic boilerplate caused by Java not being originally designed to work without classes, we introduce a new annotation:\mixin  provide default
-implementations for various methods (e.g. getters, setters, with-methods)
-and a mechanism to instantiate objects. \mixin annotation helps programmers to
-write less cumbersome code while coding in ClassLess Java;
-indeed we think the obtained gain is so high that ClassLess Java with \mixin annotation can be less cumbersome than full Java.
-
-\section{Overview}\label{sec:ep}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=4in]{PegasusDetail.png}
-\caption{Pegasus Example.}\label{fig:pegasus}
-\end{figure}
-
-\subsection{A Running Example: Animals}
-\marco{fig1 is good but is very big, not sure if it worth its space. If we keep it, we need to do it with a vectorial graphic, or better inside latex (for example with tikz)}
-To propose a standard example, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
-Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
-Birds can \Q@fly@, thus their location need to be a three dimensional \Q@Point3D@.
-Finally, we model \Q@Pegasus@ as a kind of \Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s.\footnote{
-Some research argues in favour of using subtyping for modelling taxonomies, other research argue against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
-
-%Suppose we want to model an animal system as shown in Fig~\ref{fig:pegasus}. In
-%order to model \texttt{Pegasus}, \emph{multiple inheritance} is clearly needed
-%in this example. Therefore, in Java we use interfaces instead of classes. An
-%\texttt{Animal} has a \texttt{point} field, indicating the current position of
-%the animal. \texttt{Horse} and \texttt{Bird} are subtypes of \texttt{Animal},
-%with methods \texttt{run()} and \texttt{fly()}, respectively. Note that
-%\texttt{Bird} can fly, therefore refines the type of \texttt{point} from
-%\texttt{Point2D} to \texttt{Point3D}. Pegasus (one of the best known creatures
-%in Greek mythology) can not only \emph{run} but also \emph{fly}! This is the
-%place where \emph{``multiple inheritance''} is needed, because \texttt{Pegasus}
-%need to obtain \texttt{fly} and \texttt{run} functionality from both
-%\texttt{Horse} and \texttt{Bird}.
-%---------------
-% We can put more in dept discussion while discussing the classes in details.
-%--------------
-\subsubsection{\Q@Point2D@: simple immutable data with fields}
-Since in ClassLess Java there are no classes, our \Q@Point2D@ must be an interface, as for example:
-\begin{lstlisting}
-interface Point2D{ int x(); int y();}
-\end{lstlisting}
-That is, since Java disallows fields inside interfaces,
- we simulate state by abstract methods.
-However, how can we create an instance of \Q@Point2D@ without using classes? 
-we can just provide an implementation for \Q@x()@ and \Q@y()@ as in
-\begin{lstlisting}
-new Point2D(){ public int x(){return 4;} public int y(){return 2;}}
-\end{lstlisting}
-However we are not required to use this cumbersome syntax\footnote{Available in Java from version ...} for every object allocation.
-As programmes do, for ease or reuse, we can encapsulate boring long repetitive code in a method. A static factory method in this case is appropriate:
-\begin{lstlisting}
-interface Point2D{ int x(); int y();
-  static Point2D of(int x, int y){return new Point2D(){
-    public int x(){return x;} public int y(){return y;}};}
-  }
-\end{lstlisting}
-This obvious constructor code can be automatically generated by our \mixin annotation.
-
-By annotating the interface \Q@Point2D@, the annotation will generate the shown static method \texttt{of}, mimicking the functionality of a simple minded constructor:
-By looking the the methods that need implementation it first detects what are the fields, then 
-generate an \Q@of@ method with one argument for each of them.
-
-That is, we can just write
-\begin{lstlisting}
-@Mixin interface Point2D{ int x(); int y();}
-\end{lstlisting}
-
-More precisely, a field or factory parameter is generated for every no-args method requiring implementation whose name does not have special meaning\footnote{Formal definition of \mixin behaviour is provided later.}.
-An example of code using \Q@Point2D@ is 
-\Q@Point2D.of(42,myPoint.y())@
-where we return a new point, using \Q@42@ as x-coordinate, and 
-taking all the other informations (only \Q@y@ in this case) from another point.
-This pattern is very common when programming with immutable data-structures; it is so common that we decided to support it in our code generation as \Q@with-@ methods, that is:
-\begin{lstlisting}
-@Mixin interface Point2D{ int x(); int y();
-  Point2D withX(int val); 
-  Point2D withY(int val);}
-\end{lstlisting}
-is equivalent to 
-\begin{lstlisting}
-@Mixin interface Point2D{ int x(); int y();
-  default Point2D withX(int val){ return Point2D.of(val,this.y());}
-  default Point2D withY(int val){ return Point2D.of(this.x(),val);}  }
-\end{lstlisting}
-Notice how we can provide implementation for methods in an interface, 
-and our annotation generate code only if the method \emph{needs} implementation. In this way is trivial for the programmer to personalize the behaviour if they have any special need.
-
-%Firstly, to model \texttt{Point2D} that has x-coordinate and y-coordinate by an
-%interface, we immediately run into the problem of expressing the fields
-%\texttt{x} and \texttt{y}.
-%interfaces. Method \texttt{withX, withY} creates a new instance of
-%\texttt{Point2D} with updated field \texttt{x,y}, respectively.
-
-% Firstly, to model \texttt{Animal} by an interface, we immediately run into the
-% problem of expressing the field \texttt{point}. Since in Java there is no way to
-% define member fields inside interfaces, we propose to simulate fields by
-% abstract methods inside interfaces:
-
-%\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
-
-%\paragraph{Instantiation}
-%In Java, to implement an interface like \texttt{Point2D}, a typical and trivial
-%approach that programmers usually do is creating a class extending the interface
-%and providing implementation for all methods inside. For example, this is the
-%implementation for interface \texttt{Point2D}:
-
-%\begin{lstlisting}
-%class Point2DImpl implements Point2D {
-%    private int _x;
-%    private int _y;
-%    public Point2DImpl(int x, int y) {
-%        this._x = x;
-%        this._y = y;
-%    }
-%    public int x() {
-%        return _x;
-%    }
-%    public int y() {
-%        return _y;
-%    }
-% %Your implementation of with is wrong
-%    public Point2D withX(int x) {
-%        x(x);
-%        return this;
-%    }
-%    public void x(int x) {
-%        _x = x;
-%    }
-%    public void y(int y) {
-%        _y = y;
-%    }
-%    public Point2D withY(int y) {
-%        y(y);
-%        return this;
-%    }
-%}
-%\end{lstlisting}
-%
-%\texttt{Point2DImpl} implements \texttt{Point2D} and provides a constructor with
-%quite mechanical code. What's worse, the implementation in \texttt{Point2DImpl}
-%may not be reused in a single inheritance language.
-
-%
-%\begin{lstlisting}
-%  // inside interface Point2D
-%  static Point of(int x, int y) {
-%      return new Point() {
-%          int _x = x;
-%          int _y = y;
-%          public int x() {
-%            return _x;
-%          }
-%          public int y() {
-%            return _y;
-%          }
-%          public Point2D withX(int x) {
-%            x(x);
-%            return this;
-%          }
-%          public void x(int x) {
-%            _x = x;
-%          }
-%          public void y(int y) {
-%            _y = y;
-%          }
-%          public Point2D withY(int y) {
-%            y(y);
-%            return this;
-%          }
-%    }
-%  }
-%\end{lstlisting}
-%
-%\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
-
-\subsubsection{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
-\Q@Points2D@ are mathematical entities, thus we chosen to use immutable data structure to model them; however animals are real world entities, and when an animal moves, it is the same animal that have now a different location.
-We model this with mutable state.
-%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
-%field''.
-%%Not again, before we used only getters!
-% Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
-\marco{TODO: change point into ``location''}
-Here we declare abstract getter and setter for the mutable ``field'' \Q@location@.
-As you see, we do not apply the \mixin annotation. This is morally equivalent to an abstract class in full Java; that is we do not provide a convenient way to instantiate it.
-
-\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
-
-For \texttt{Horse}, concrete implementation of \texttt{run()} method
-need to be defined in a default method, where 
-we also show the convenience of \emph{with} methods.
-% by method
-%\texttt{run()}: method \texttt{withX} returns a new point object with field
-%\texttt{x} updated by the argument to \texttt{withX}. Without these
-%\texttt{with} methods, operations like \texttt{run()} would be much harder to define.
-
-\subsubsection{\Q@Bird@:  field type refinement}
-\Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
-as follows:
-\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
-
-Just to give you a feeling on how much boring code \mixin is generating,  we show the correspondent code without then \mixin annotation\footnote{
-As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
-swap the arguments of one of the many calls of \Q@Point3D.of@.
-}.\begin{lstlisting}
-interface Point3D extends Point2D{
-  Point3D withX(int val);
-  Point3D withY(int val);
-  Point3D withZ(int val);
-  Point3D with(Point2D val);
-  public static Point3D of(int x, int y, int z){return new Point3D(){
-    public int x(){return x;} 
-    public int y(){return y;}
-    public int z(){return z;}
-    public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
-    public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
-    public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
-    public Point3D with(Point2D val){
-      if(val instanceof Point3D){return (Point3D)val;}
-      return Point3D.of(val.x(),val.y(),this.z());
-      }
-    };
-  }
-}
-\end{lstlisting}
-Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
-\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected.
-If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
-Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
-\Q@with-@ methods (functionally) update a field/property at a time.
-This can be inefficient, and sometime hard to maintain.
-Often we want to update many fields at the same time, for example using another object as source.
-Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
-it takes a certain type and in the current object update all field in that type that match fields in the current type.
-The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
-The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test.
-The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.
-
-\Q@Bird@s are \Q@Animal@s, 
-but while \Q@Animal@s, only need 2D locations,
-\Q@Bird@s need  3D locations.
-In Java if we define an animal class with a field we have a set of unsatisfactory options in front of us:
-\begin{itemize}
-\item Define a \Q@Point3D@ field in \Q@Animal@:
-this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
-\item Define a \Q@Point2D@ field in \Q@Animal@ and...
-\item Define an extra \Q@int z@ field in \Q@Bird@.
-This solution is very at-hoc, requires to basically duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside of \mbox{\Q@Bird@}.
-Again, there are many reasons this would be bad, the most dramatic is that it would not scale to a scenario when the programmer of \Q@Bird@ and the programmer of \Q@Point3D@ are different.
-\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects in the field and cast
-the value out of the \Q@Point2D@ field to \Q@Point3D@ when implementing the overridden getter.
-This solution scales to the multiple programmers approach, but requires ugly casts and can be implemented in a wrong way leading to bugs.
-\end{itemize}
-
-Many here would think that a language extension is needed.
-Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}:
-
-\lstinputlisting[linerange=80-88]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
-
-%Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
-%\texttt{int z()} (treated as a getter for member field \texttt{z}). Note that
-%the return type of various methods (e.g. with- methods, getters) get refined
-%either by covariant method overriding or automatically by our annotation
-%processor. Besides \emph{with}, other methods (including \emph{clone},
-%\emph{of}) also do type-refinements automatically.
-
-Note how
-\begin{itemize}
-\item by covariant method overriding we refine the type of our field
-\item by \emph{overloading} we define a new setter with the more precise type
-\item by default method we provide an implementation for the setter called with the old signature.
-\end{itemize}
-
-From the type perspective, the key is the covariant method overriding; however from the semantic perspective the key is the implementation for the setter with the old signature:
-by using the \Q@with@ method we may use the information for \Q@z@ already stored in the object to forge an appropriate \Q@Point3D@ to store.
-Note how all the informations about what fields sits in \Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the \Q@with@ method, and is transparent for the implementer of \Q@Bird@.
-
-\subsubsection{\Q@Pegasus@ and multiple Inheritance}
-The \emph{``multiple inheritance''} case appears at interface
-\texttt{Pegasus}.
-Pegasus (one of the best known creatures
-in Greek mythology) can not only \emph{run} but also \emph{fly}! 
-It would be handy if they could obtain \texttt{fly} and \texttt{run} functionality from both \texttt{Horse} and \texttt{Bird}.
-
- Using interfaces and \mixin annotation,
-this can happen transparently.
-
-\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
-
-Note how  even the non trivial pattern for field type refinement is transparently composed, and our pegasus have a \Q@Point3D@ \Q@location@.
-Note that this works because \Q@Horse@ do not perform any field type refinement , otherwise we may have to choose/create a common subtype in order for pegasus to exists.
-
-\begin{comment}
-\subsection{A Running Example: \texttt{Point}}
-Suppose we want to create a point component that models the a point in space,
-that has x-coordinate and y-coordinate. For example, if we create the
-\texttt{Point} interface in Java, it would look like this:
-
-\begin{lstlisting}
-interface Point {
-    int x();
-    int y(); 
-}
-\end{lstlisting}
-
-\texttt{Point} has two (conceptually) member fields \texttt{x} and \texttt{y},
-representing the two coordinates of a point. Methods \texttt{int x()} and
-\texttt{int y()} serve as \emph{getter} methods. 
-% Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
-% \emph{setter} methods. Method \texttt{Point withX(int X)} updates field
-% \texttt{X} and returns \textbf{this}.
-
-\subsection{Naive Implementation}
-In Java, to implement an interface like \texttt{Point}, a typical and trivial
-approach that programmers usually do is creating a class extending the interface
-and providing implementation for all methods inside. For example, this is the
-implementation for interface \texttt{Point}:
-
-\lstinputlisting[linerange=33-46]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINTIMPL
-
-\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
-quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
-may not be reused in a single inheritance language.
-
-\subsection{Our Approach}
-Instead of writing a whole another class to provide the implementation for
-\texttt{Point}, we annotate on interface \texttt{Point} directly with \mixin:
-
-\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT
-
-The \mixin annotation will generate a static method \texttt{of} inside
-\texttt{Point}. The method \texttt{of} mimic the functionality of constructors,
-it takes arguments same as constructors and return objects similar to
-constructors. It makes use of Java anonymous classes and achieves the same
-implementation as \texttt{PointImpl}.
-
-With our approach, we provide a Java annotation \mixin to provide default
-implementations for various methods and a mechanism to instantiate
-objects. \mixin annotation helps programmers to write less cumbersome code and
-instantiate interfaces in Java.
-
-\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
-\end{comment}
-
-\begin{comment}
-\subsection{More On \mixin}
-Besides the benefit of freeing programmers from writing boilerplate code, our
-\mixin annotation can also allow programs to mimic multiple inheritance in a
-restricted form easily.
-
-\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
-
-\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
-
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
-
-\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
-
-\lstinputlisting[linerange=80-88]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
-
-\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
-
-Interface \texttt{Point3D} extends \texttt{Point/Point2D} with a new abstract
-method \texttt{int z()} (treated as a getter for member field
-\texttt{z}). Interface \texttt{Horse} shows the usage and advantage of
-\emph{with} methods by method \texttt{run()}: method \texttt{withX} returns a
-new point object with field \texttt{x} updated by the argument to
-\texttt{withX}. Without these \texttt{with} methods, operations like
-\texttt{run()} would be much harder to define. Note that the return type of
-various methods get refined automatically by our annotation processor. Besides
-\emph{with}, other methods (including \emph{clone}, \emph{of}) also do
-type-refinements automatically.
-
-The \emph{``multiple inheritance''} case appears at interface
-\texttt{Pegasus}. Pegasuses can not only \emph{run} but also \emph{fly}!
-Interface \texttt{Pegasus} obtains \texttt{fly} and \texttt{run} functionality
-through interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation,
-actually these is no code at all programmers have to write. The idea of using
-default methods inside interfaces was proposed in ~\cite{}. It enables us to do
-multiple inheritance, which otherwise is hard to do in Java-like languages that
-do not support multiple inheritance, easily.
-\end{comment}
+Before Java 8, concrete methods and static methods where not allowed
+to appear in interfaces.  Java 8 allows static interface methods and
+introduces \emph{default methods}, which allow for implementation
+insides interfaces. This had an important positive consequence that
+was probably overlooked by the Java design team: the concept of class
+(in java) is now redundant and unneeded.  We define a subset of Java,
+called ClassLess Java, where programs and (reusable) libraries can be
+easily defined and used.  To avoid for some syntactic boilerplate
+caused by Java not being originally designed to work without classes,
+we introduce a new annotation:\mixin provide default implementations
+for various methods (e.g. getters, setters, with-methods) and a
+mechanism to instantiate objects. \mixin annotation helps programmers
+to write less cumbersome code while coding in ClassLess Java; indeed
+we think the obtained gain is so high that ClassLess Java with \mixin
+annotation can be less cumbersome than full Java.\bruno{May need rewriting}

### 2015-11-18 paper/sections/Introduction.tex
Minor changes in Introduction

@@
 \noindent Clearly, a way around those difficulties would be to try to
-change the language and just remove these limitations. Scala's own
+change Java and just remove these limitations. Scala's own
 notion of traits, for example, allows state in traits. Of course adding
@@
 \end{itemize}
-
-\bruno{Useful reference, perhaps, on the discussion of why not override object 
-methods: 
-\url{http://mail.openjdk.java.net/pipermail/lambda-dev/2013-March/008435.html}}
-\bruno{Maybe Marco's text can be in the conclusion, which I think is
-  ok to be more provocative. I tried to be a bit more contained in the
-Introduction.}

### 2015-11-18 paper/sections/Overview.tex
Minor changes in Introduction

@@
-\marco{an improvement of this may become part of intro?}
-
-Before Java 8, concrete methods and static methods where not allowed to appear in interfaces.
-Java 8 allows static interface methods and introduces
-\emph{default methods}, which allow for implementation insides interfaces.
-This had an important positive consequence that was probably overlooked by the Java design team: the concept of class (in java) is now redundant and unneeded.
-We define a subset of Java, called ClassLess Java, where programs and (reusable) libraries can be easily defined and used.
-To avoid for some syntactic boilerplate caused by Java not being originally designed to work without classes, we introduce a new annotation:\mixin  provide default
-implementations for various methods (e.g. getters, setters, with-methods)
-and a mechanism to instantiate objects. \mixin annotation helps programmers to
-write less cumbersome code while coding in ClassLess Java;
-indeed we think the obtained gain is so high that ClassLess Java with \mixin annotation can be less cumbersome than full Java.
-
 \section{A Running Example: Animals}\label{sec:ep}

### 2015-11-18 paper/sections/Formalization.tex
polishing formalization

@@
 \section{Formal Semantics}\label{sec:formal}
-
-\marco{Old notes: new text is later}
-Yanlin and Haoyuan
-
-We need to show 2 things:
-
-1) The dynamic semantics: what's the code that gets generated by a mixin annotation;
-
-2) The type system: what programs to reject; properties: generation of type-safe/checkable code.
-
-\bruno{The implementation is still missing the type system (rejecting some
-  programs)!}
 
@@
 \begin{itemize}
-\item Syntax and Rules:
-We can make it more compact and consistent by use overline
-notation more, instead of \ldots.
 \item Order of premisses: It may be worthwhile to swap the order of
 some premisses, so that it follows the order of the explanations.
-\item T-Invk: I think we need to type-check $e_0$ separately. Otherwise
-$C_0$ shows up magically.
 \item T-Obj: small fixes on 3rd line. Suggestion, create two
@@
 I think we can merge those 2 steps.
-
-\item I have trouble following the definition of shadow; I think there
-are errors there.
 
@@
 }{
- \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\ \methods\ \cC
+ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC
   }{interface declaration}\\
@@
 \mid
-\QM{default}\ \mh\oC\QM{return}\ \e\QM{;}\cC
+\QM{default}\ \mh\ \oC\QM{return}\ \e\QM{;}\cC
 \mid
@@
 
+\bruno{Use space better in this figure. }
+
 ${}_{}$\\
@@
 \begin{array}{l}
+
+%% T-Invk
 \inferrule[(T-Invk)]{
  \Gamma \vdash \e \in \C \\\\
-\forall i\in 1..n\  \Gamma \vdash \e_i \in \_<:\C_i \\\\
+\forall i\in 1..n\\   \Gamma \vdash \e_i \in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
@@
  \Gamma \vdash \e\QM.\m\QM(\overline{\e}\QM) \in \C' }
-\quad
+\quad\quad
+
+%%T-StaticInvk
 \inferrule[(T-StaticInvk)]{
-\forall i\in 1..n\  \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\forall i\in 1..n\  \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
 \textsf{mtypeS}(\m,\C) \!=\! \overline{\C} \!\to\! \C'
@@
 \Gamma \vdash \C\QM.\m\QM(\overline{\e}\QM) \in \C'}
-\quad
+\quad\quad
+
+%%T-SuperInvk
 \inferrule[(T-SuperInvk)]{
 \Gamma(\this) <: \C \\\\
-\forall i\in 1..n\ \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\forall i\in 1..n\ \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
@@
 
+%%T-Var
 \\[5ex]
+\inferrule[(T-Var)]{
+\Gamma(\x)=\C
+}{
+\Gamma \vdash \x \in\C}
+\quad\quad\quad
+
+%%T-Obj
 \inferrule[(T-Obj)]{
@@
 }
-\quad
-\inferrule[(T-Var)]{
-\Gamma(\x)=\C
-}{
-\Gamma \vdash \x \in\C}
 \\[5ex]
 
+%%T-update
 \quad
@@
 \Gamma \vdash \x\QM=\e\QM;\e'\in\C }
-\quad
+\quad\quad\quad\quad\quad
+
+%%T-Intf
  \inferrule[(T-Intf)]{
-IT(\C_0) = \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n \oC\methods\cC\\\\
+IT(\C) = \ann\ \QM{interface}\ \C\ \QM{extends}\ \overline{I} \
+\oC\methods\ \cC\\\\
  \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
-\ \Gamma^{\mh},\,\QM{this}{:}\C_0\vdash\e\in \_\subtype\C^{\mh} \\\\
+\ \ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
  \forall \QM{static}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
-\ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
-\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
+\ \ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
+\dom(\C)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
  }{
-\C_0 \text{ OK}
+\C \text{ OK}
 }
 \end{array}$
-\caption{Typing}
+\caption{CJ Typing}
 \label{ET}

### 2015-11-18 paper/sections/Formalization.tex
Formalization partly revised

@@
 
-\bruno{Some notes for discussion regarding formalization:
-\begin{itemize}
-\item Order of premisses: It may be worthwhile to swap the order of
-some premisses, so that it follows the order of the explanations.
-\item T-Obj: small fixes on 3rd line. Suggestion, create two
-auxiliary functions: $sigvalid(C,\overline{mh})$ and $alldefined(C,\overline{mh})$
-\item T-INTF: $C_0$ is $I$. Type of this was $C$, modified it to $C_0$.
-In the last line, should it be $dom(\overline{meth})$ or just $\overline{meth}$?
+This section presents a formalization of ClassLess Java: a minimal
+FeatherweightJava-like calculus which models the essence of Java 
+interfaces with default methods. This calculus also supports a minimal
+extension to Java to account for the \mixin annotation for object
+interfaces. The syntax, typing rules and method lookup mechanisms 
+of CJ are modelled. Finally, using our formalization, we discuss
+examples of real Java programs that seem to be incoherent\bruno{revise
+later}.
 
-\item tops and shadow: tops seems redundant;
-we use tops to collect classes, so just that in the next step we iterate
-over those classes to collect methods that satisfy some property.
-I think we can merge those 2 steps.
-
-\end{itemize}
-}
-
-\footnote{Future work: updating multiple fields in one method call,
-  \texttt{with(T v)}}
-
-\begin{figure}[h]
+\begin{figure}[t]
 \begin{grammar}
@@
 
-\bruno{Use space better in this figure. }
+\subsection{Syntax and Typing}\marco{Future work: updating multiple fields in one method call,
+  \texttt{with(T v)}}
 
-${}_{}$\\
-\begin{figure}[h]
+\paragraph{Syntax}
+Figure~\ref{Grammar} shows the syntax of ClassLess Java.  \footnote{To
+  be compatible with java, the concrete syntax for an interface
+  declaration with empty supertype list $\C_1\ldots\C_n$ would also
+  omit the \Q@extends@ keyword.}  The syntax formalizes a minimal
+version of Java 8, focusing on interfaces, default methods and object
+creation literals.  There is no syntax for classes.  Expressions
+consist of conventional constructs such as variables ($\x$), method
+calls ($\e\QM.\m\QM(\es)$) and static method calls
+($\C\QM.\m\QM(\es)$). For simplicity the degenerate case of calling a
+static method over the $\this$ receiver is not considered.  A more
+interesting type of expressions are super calls
+($\C\QM{.super.}\m\QM(\es)$), whose semantic is to call the (non
+static) method $\m$ over the $\this$ receiver, but statically
+dispatching to the version of the method defined in the interface
+$\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
+modelled. In the syntax of field updates $\x$ is expected to be a
+field name. After updating the field $\x$ using the value of $\e$, the
+expression $\e'$ is executed. To blend the statement based nature of
+Java and the expression based nature of our language, we consider a
+method body of the form \Q@return@ $\x\QM=\e\QM;\e'$ to represent
+$\x\QM=\e\QM;\QM{return}\ \e'$ in Java.  Finally, there is an object
+initialization expression from an interface $\C$, where (for
+simplicity) all the fields are initialized with a variable present in
+scope. Note how our language is a subset of Java 8.  The single
+non-Java 8 piece of syntax is the \mixin annotation, which is the only
+one interesting piece of syntax in this article.  Following standard
+practise, we consider a global Interface Table (\metaVar{IT}) mapping
+from interface names $\C$ to interface declarations \metaVar{I}.
+
+The environment $\Gamma$ is a mapping from variables to types.  As
+usual, we allow a functional notation for $\Gamma$ to do variable
+lookup.  Moreover, to help us defining auxiliary functions, a functional
+notation is also allowed for a set of methods $\methods$, using the
+method name $\m$ as a key.  That is, we define $\methods(\m)=\method$
+iff there is a unique $\method\in\methods$ whose name is $\m$.  For
+convenience, we define $\methods(\m)=\none$ otherwise; moreover
+$\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+
+\paragraph{Typing} 
+Typing statement $\Gamma \vdash \e\in\C$ reads ``in the environment
+$\Gamma$, expression $\e$ has type $\C$.''. 
+Before discussing the typing rules we discuss some of the used notation.
+As a shortcut, we write
+$\Gamma \vdash \e\in\C<:\C'$ instead of $\Gamma \vdash \e\in\C$ and
+$\C<:\C'$.  From the interface table, we can read off the subtype
+relation between interfaces. The subtype relation is given by the
+\textbf{extends} clauses in the interfaces. We omit the definition of
+the usual\bruno{where can it be found? add reference here?}, traditional subtyping relation between interfaces.%
+\footnote{ Notice how there are no classes, thus there is no
+  subclassing.  We believe that this approach may scratch an old
+  itching point in the long struggle of subtyping versus subclassing:
+  According to some authors, from a software engineering perspective,
+  interfaces are just a kind of classes. Others consider more
+  opportune to consider interfaces are pure types. In this vision our
+  language would have no subclassing. We do not know how to conciliate
+  those two viewpoints and ClassLess Java design. We do not have
+  Classes purely in the Java sense.}\bruno{is a footnote here the
+  appropriate place to have this discussion?}
+The auxiliary notation $\Gamma^\mh$ trivially
+extracts the environment from a method header, by collecting the all types
+and names of the method parameters.  The
+notation $\m^\mh$ and $\C^\mh$ denotes, respectivelly, extracting the
+method name and the return type from a method header. $\mBody(\m,\C)$,
+defined in Section~\ref{},
+returns the full method declaration as seen by $\C$, that is the
+method $\m$ can be declared in $\C$ or inherited from another
+interface. 
+$\textsf{mtype}(\m,\C)$ and $\textsf{mtypeS}(\m,\C)$ return the type
+signature from a method (using $\mBody(\m,\C)$ internally).
+$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
+$\textsf{mtypeS}(\m,\C)$ only on static ones. We use $\dom(\C)$ to
+denote the set of methods that are defined for type $\C$, that is:
+$\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
+
+\begin{figure}[t]
 $
-\!\!\!\!\!\!\!\!\!\!
 \begin{array}{l}
@@
  \Gamma \vdash \e \in \C \\\\
-\forall i\in 1..n\\   \Gamma \vdash \e_i \in \_<:\C_i \\\\
+\forall i\in 1..n\ \ \Gamma \vdash \e_i \in \_<:\C_i \\\\
   \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
@@
 \Gamma \vdash \x \in\C}
-\quad\quad\quad
+\quad\quad
 
@@
 \inferrule[(T-Obj)]{
-\forall i\in 1..n\
+\forall i\in 1..k\ \ \Gamma(\x_i)\subtype\T_i\\\\
+\forall i\in 1..n\ \ 
 %\Gamma_i
@@
 \vdash\e_i\in \_\subtype\C^{\mh_i}\\\\
-\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
-\forall i\in 1..k\ \Gamma(\x_i)\subtype\T_i\\\\
-\forall\m\mbox{ such that }
-\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
+\sigvalid(\mh_1\ldots\mh_n,I)\quad\quad\quad\quad
+%\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
+\alldefined(\mh_1\ldots\mh_n,I)
+%\forall\m\mbox{ such that }
+%\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
 %\forall i\in 1\ldots n\ \Gamma_i=\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
@@
 \Gamma \vdash \x\QM=\e\QM;\e'\in\C }
-\quad\quad\quad\quad\quad
+\quad\quad\quad
 
@@
 \end{figure}
-\subsection{Grammar and typing rules}${}_{}$\\
 
-In Figure~\ref{Grammar} we show the syntax of ClassLess Java.%
-\footnote{To be compatible with java, the concrete syntax for an interface declaration with empty supertype list $\C_1\ldots\C_n$ would also omit the \Q@extends@ keyword.}
+In Figure~\ref{ET} we show the typing rules.  We discuss first the
+most interesting rules, that is \rn{t-Obj} and \rn{t-Intf}. Rule
+\rn{t-Obj} is the most complex typing rule. Firstly, we need to
+ensure that all field initializations are type correct, by looking up the type of
+each variable assigned to a field in the typing environment and verifying that such type is a
+subtype of the field type. Secondly, we check that all method bodies are
+well-typed. To do this the enviroment used to check the method body
+needs to be extended appropriately: we add all fields and their types;
+add $\this:I$; and add the arguments (and types) of the respective
+method. Finally, we need to check that all method headers are valid
+with respect to the superinterfaces of $I$; and we also need to check
+that all abstract and conflicted methods (that is methods, that need
+to be explicitly overriden) in the superinterfaces have been
+implemented. The definitions $\sigvalid$ and $\alldefined$ take care
+of this: \\
 
-We formalize a minimal version of Java 8, focusing on Interfaces, default methods and object creation literals.
-As you can see, we have no syntax for classes.
-Expressions are the conventional variable and method call$\x$ and$\e\QM.\m\QM(\es)$, then we have conventional static method call
-$\C\QM.\m\QM(\es)$; for simplicity we do not consider the degenerate case of calling a static method over the this receiver as it considered a bad/confusing programming practice.
-Next we have the more interesting super call $\C\QM{.super.}\m\QM(\es)$, whose semantic is to call the (non static) method $\m$ over the $\this$ receiver, but statically dispatching to the version of the method defined in the interface $\C$.
-We model field updating as $\x\QM=\e\QM;\e'$, where we expect $\x$ to be a field name.
-After updating the field $\x$ using the value of $\e$, the expression $\e'$ is executed.
-To blend the statement based nature of Java and the expression based nature of our language, we consider a method body of the form \Q@return@ $\x\QM=\e\QM;\e'$
-to represent  $\x\QM=\e\QM;\QM{return}\ \e'$ in Java.
-Finally, we consider the object initialization expression from an interface $\C$, where (for simplicity) all the fields are initialized with a variable present in scope.
-Note how our language is exactly a subset of Java 8.
-We allows the only annotation \mixin since is the only one interesting in this article.
-We consider an globally present Interface Table (\metaVar{IT}) mapping from interface names $\C$ to interface declarations \metaVar{I}.
+$
+\begin{array}{lll}
+\sigvalid(\mh_1\ldots\mh_n,I) & = &
+\forall i\in 1..n\ \ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\
 
-The environment $\Gamma$ is a mapping from variables to types.
-As tradition, we allows functional notation for $\Gamma$.
-However, to help us defining auxiliary functions,
-we also allows functional notation set of methods $\methods$, using the method name $\m$ as a key.
-That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
-For convenience, we define $\methods(\m)=\none$ otherwise;
-moreover $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+\alldefined(\mh_1\ldots\mh_n,I) & = &
+\forall\m\mbox{ such that } \mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
+\end{array}
+$\bruno{Marco, you probably want to double-check the explanation. Any
+  more explanation needed?}
+\bruno{Also, according to the text above, mbody returns the full
+  method body, so in sigvalid, shouldn't we use mtype instead? The
+  same for alldefined, should it be mtype? Still not very confident
+  about alldefined. }
 
- Typing statement $\Gamma \vdash \e\in\C$ reads ``in the
-environment $\Gamma$, expression $\e$ has type $\C$.''
-As a shortcut, we write $\Gamma \vdash \e\in\C<:\C'$
-instead of $\Gamma \vdash \e\in\C$ and $\C<:\C'$.
- From the interface table, we can read off the subtype relation between interfaces.
-The subtype relation is given by the \textbf{extends} clauses in the interfaces.
-We omit the definition of the usual, traditional subtyping relation between interfaces.%
-\footnote{
-Notice how there are no classes, thus there is no subclassing.
-We believe that this approach may scratch an old itching point in the long struggle of subtyping versus subclassing:
-According to some authors, from a software engineering perspective, interfaces are just a kind of classes.
-Others consider more opportune to  consider interfaces are pure types.
-In this vision our language would have no subclassing.
-We do not know how to conciliate those two viewpoints and ClassLess Java design.
-We do not have Classes purely in the Java sense.
-}
-
-In Figure~\ref{ET} we show the typing rules.
-We use the following auxiliary notation:
-$\Gamma^\mh$ trivially extract the environment from a method header, collect the types and the names of the method parameters.
-$\m^\mh$ and $\C^\mh$ extracts the method name and the return type from a method header.
-$\mBody(\m,\C)$ return the full method declaration as seen by $\C$, that is the method $\m$ can be declared in $\C$ or inherited from another interface.
-$\mBody(\m,\C)$ will be formally defined later.
-$\textsf{mtype}(\m,\C)$ and
-$\textsf{mtypeS}(\m,\C)$ return the type signature from a method (using $\mBody(\m,\C)$ internally).
-$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
-$\textsf{mtypeS}(\m,\C)$ only on static ones.
-
-We use $\dom(\C)$ to denote the set of methods that are defined for type $\C$, that is: $\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
-
-We discuss first the most interesting rules, that is
-\rn{(t-Obj)} and  \rn{(t-Intf)} \marco{Bruno, you can explain t-obj, if you can make it understandable, I go for t-inft}
-
-Rule \rn{(t-Obj)}....
-
-Rule \rn{(t-intf)} check that an interface $\C_0$ is correctly typed.
-First we check that the body of all the default and static methods are well typed.
-Then we check that
-$\dom(\C_0)$ is the same of $\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$.
-This is not a trivial check, since $\dom(\C)$ is defined using $\mBody$, that is undefined in many cases: notably if a method $\method\in\methods$ is not compatible with some method in
-$\dom(\C_1)\ldots\dom(\C_n)$ or if any method in both $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$)  is conflicting.
-
+The rule \rn{(t-intf)} checks that an interface $\C$ is correctly
+typed.  First we check that the body of all the default and static
+methods are well typed.  Then we check that $\dom(\C)$ is the same of
+$\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$.  This is not a
+trivial check, since $\dom(\C)$ is defined using $\mBody$, that is
+undefined in many cases: notably if a method $\method\in\methods$ is
+not compatible with some method in $\dom(\C_1)\ldots\dom(\C_n)$ or if
+any method in both $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$) is
+conflicting.
 \marco{if we have space we can discuss the other rules too, but is not mandatory}

### 2015-11-19 paper/sections/Implementation.tex
section5 Implementation revised

@@
 \section{Implementation}
-\begin{comment}
-Haoyuan
-
-discuss implementation in lombok; and limitations.
-
-\bruno{The implementation does not support separate compilation yet. Can we fix this?}
-
-\haoyuan{http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html}
-
-\haoyuan{https://blog.frankel.ch/lombok-reduces-your-boilerplate-code}
-
-\haoyuan{https://projectlombok.org/}
-\end{comment}
 
@@
 
-The biggest reasons to use Lombok rather than using conventional Java annotation
-processor is
+ The biggest reasons to use Lombok rather than using a conventional Java
+ annotation processor are:
 \begin{itemize}
@@
 %  compile time to modify bytecode.
-\item Lombock modify the generation process of the class files.
-neither the source code is modified nor new  Java files are generated.
+\item Lombok modifies the generation process of the class files. Neither the
+  source code is modified nor new Java files are generated.
 \item Moreover, and probably more importantly, Lombok is capable of generating
@@
 
+\paragraph{Limitation}
 Our prototype implementation using Lombok has certain limitation: 
 \begin{itemize}
-\item We do not support separate compilation yet at this
-  stage, for in the implementation it is hard to capture a type declaration from
-  its reference.
-\item At this stage our implementation only realizes the Eclipse handler and our experiments are all conducted in Eclipse.
+\item The prototype does not support separate compilation yet. Currently all
+  related interfaces have to appear in a single Java file. Therefore, changes to
+  a single interface would require re-compiling the whole file. This compilation
+  limitation is not caused by our algorithm. It is Lombok implementation related
+  issue: in Lombok it is hard to capture a type declaration from its reference,
+  even harder when the type declaration is in other files (we have not found a
+  way to do this currently).
+\item At this stage our implementation only realizes the Eclipse handler and our
+  experiments are all conducted in Eclipse. The implementation for
+  \texttt{javac} is missing.
 \end{itemize}
+
+\paragraph{Comparison with other Lombok annotations}
+Lombok project provides a set of predefined annotations, including constructor
+generators similar as ours (e.g., \textbf{@NoArgsConstructor},
+\textbf{@RequiredArgsConstructor} and \textbf{@AllArgsConstructor}). They
+generate various kinds of constructors for \emph{classes}, with or without
+constructor arguments. This set of annotations is of great use, especially when
+used together with other features provided in Lombok (e.g.,
+\textbf{@Data}). Moreover, the implementation of these annotations in Lombok
+gives us hints on how to implement \mixin. However, none of these annotations
+can model what we are doing with \mixin - generating constructor-methods
+(\textbf{of}) for \emph{interfaces}. Apart from constructors, \mixin also
+provides other convenient features (including generating fluent setters, type
+refinement,etc) which the base Lombok project could not provide.

### 2015-11-19 paper/sections/RelatedWork.tex
section5 Implementation revised

@@
 
-\marco{text under still to revise. Sould all the lombock discussion go in the implementation section? -- just to keep stuff together?}
-Annotations/Code Generation/Lombok
-
-What is the difference between our work and existing annotations?
-What is it that we can do that, those annotations cannot?
-
--Lombok~\cite{} project provides a set of predefined annotations, including
-constructor generators similar as ours (\textbf{@NoArgsConstructor},
-\textbf{@RequiredArgsConstructor} and \textbf{@AllArgsConstructor}). They
-generate various kinds of constructors for \emph{classes}, with or without
-constructor arguments. This set of annotations is of great use, especially when
-they are used together with other features provided in Lombok (e.g.,
-\textbf{@Data}). Moreover, the implementation of these annotations in Lombok
-gives us hints on how to implement \mixin. However, none of these annotations
-can model what we are doing with \mixin - generating a constructor for
-\emph{interfaces}. Apart from constructors, \mixin also provides other
-convenient features (including generating fluent setters, type refinement,etc)
-that the base Lombok project could not provide.
-
 \subsection{Fluent Interfaces}

### 2015-11-20 paper/sections/Overview.tex
Section 2 (work in progress)

@@
 
-%Suppose we want to model an animal system as shown in Fig~\ref{fig:pegasus}. In
-%order to model \texttt{Pegasus}, \emph{multiple inheritance} is clearly needed
-%in this example. Therefore, in Java we use interfaces instead of classes. An
-%\texttt{Animal} has a \texttt{point} field, indicating the current position of
-%the animal. \texttt{Horse} and \texttt{Bird} are subtypes of \texttt{Animal},
-%with methods \texttt{run()} and \texttt{fly()}, respectively. Note that
-%\texttt{Bird} can fly, therefore refines the type of \texttt{point} from
-%\texttt{Point2D} to \texttt{Point3D}. Pegasus (one of the best known creatures
-%in Greek mythology) can not only \emph{run} but also \emph{fly}! This is the
-%place where \emph{``multiple inheritance''} is needed, because \texttt{Pegasus}
-%need to obtain \texttt{fly} and \texttt{run} functionality from both
-%\texttt{Horse} and \texttt{Bird}.
-%---------------
-% We can put more in dept discussion while discussing the classes in details.
-%--------------
-\subsection{\Q@Point2D@: simple immutable data with fields}
-Since in ClassLess Java there are no classes, our \Q@Point2D@ must be an interface, as for example:
+\subsection{Simple Traits with Default Methods}
+
+Suppose that we want to model an animal system as shown in
+Fig~\ref{fig:pegasus}. \texttt{Horse} and \texttt{Bird} are subtypes
+of \texttt{Animal}, with methods \texttt{run()} and \texttt{fly()},
+respectively. Pegasus (one of the best known creatures in Greek
+mythology) can not only \emph{run} but also \emph{fly}! This is the
+place where \emph{``multiple inheritance''} is needed, because
+\texttt{Pegasus} needs to obtain \texttt{fly} and \texttt{run}
+functionality from both \texttt{Horse} and \texttt{Bird}.
+Ignoring points, for the moment, a 
+first attempt to model the animal system in Java 8 is shown next:
+
+\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+
+Note that the implementations of the methods \texttt{run}
+and \texttt{fly} are defined inside interfaces, using default
+methods. Moreover, because interfaces support multiple interface
+inheritance, the interface for \texttt{Pegasus} can inherit behaviour
+from both \texttt{Horse} and \texttt{Bird}. Although Java interfaces
+do not allow instance fields, no form of state is needed so far to
+model the animal system.  Therefore this example provides a simple
+illustratation how a basic form of multiple inheritance, similar to
+what can be done with traits, can be accomplished in Java 8.
+
+\bruno{
+Note that \texttt{Bird} can fly, therefore refines the
+type of \texttt{point} from \texttt{Point2D} to
+\texttt{Point3D}. 
+}
+
+\paragraph{Instantiation}
+To use \texttt{Horse}, \texttt{Bird} and \texttt{Pegasus}, some
+objects must be instantiated first. A first problem with using
+interfaces to model the animal system is simply that interfaces 
+cannot directly be instantiated: a class is needed. Therefore, we have
+to first create some classes, such as:
+
+\lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
+
+\noindent in order to create animal objects. With those classes, 
+a Pegasus animal, can be created using the class constructor:
+
+\begin{lstlisting}
+Pegasus p = new PegasusImpl();
+\end{lstlisting}
+
+\noindent There are a couple of annoyances here. Firstly, the sole
+purpose of the classes is to provide a way to instantiate the
+objects. Although (in this case) it takes only one line code to
+provide each of those classes, this code is essentially boilerplate
+code, which does not add behaviour to the system. Secondly, 
+the namespace gets filled with three additional types. For example, 
+both \texttt{Horse} and \texttt{HorseImpl} are needed: \texttt{Horse}
+is needed because it needs to be an inferface so that \texttt{Pegasus}
+can use multiple inheritance; and \texttt{HorseImpl} is needed to
+provide object instantiation.
+
+Note that, for this very simple animal system, plain Java 8 anonymous
+classes can be used to avoid these problems.  We could have simply
+instantiated Pegasus using:
+
+\begin{lstlisting}
+Pegasus p = new Pegasus() {}; // anonymous class
+\end{lstlisting}
+
+However, as we shall see, once the system gets a little more
+complicated, the code for instantiation quickly becomes more 
+complex and verbose (even with anonymous classes).
+
+\subsection{Object Interfaces and Instantiation}
+
+To model the animal system with object interfaces all that a user
+needs to do is to add a \mixin annotation to the \texttt{Horse}, 
+\texttt{Bird}, and \texttt{Pegasus} interfaces.  For example, 
+for \texttt{Bird} and \texttt{Pegasus} the code would be as follows:
+
+\lstinputlisting[linerange=11-14]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+
+The effect of the annotations is that a static \emph{factory} method called
+\texttt{of} is automatically added to the interfaces. With the
+\texttt{of} method a Pegasus object is instantiated as follows:
+
+\begin{lstlisting}
+Pegasus p = Pegasus.of();
+\end{lstlisting}
+
+\noindent The \texttt{of} method provides an alternative to a
+constructor, which is missing from interfaces. The following code,
+shows the code correponding to the \texttt{Pegasus} interface
+after the \mixin annotation is processed:
+
+\begin{lstlisting}
+interface Pegasus extends Horse, Bird { // generated code not visible to users
+	static Pegasus of() {return new Pegasus() {};}
+}
+\end{lstlisting}
+
+\noindent Note that the generated code is transparent to a user, which
+only sees the original code with the \mixin annotation. Compared to the pure
+Java solution in Section~\ref{}, the solution using object interfaces
+has the advantage of providing a direct mechanism for object
+instantiation, which avoids adding boilerplate classes to the
+namespace.
+
+\subsection{Object Interfaces with State}
+
+The animal system modelled so far, is a simplified version of the
+system presented in Figure~\ref{}. \bruno{emphasize that, so far the
+  need for \mixin is not justified?} Lets now move on to modelling 
+the complete animal system. In order to do this, animals will include 
+a notion of points, representing their location in space.
+
+\paragraph{\Q@Point2D@: simple immutable data with fields}
+Two dimensional points should keep track of their coordinates. 
+The usual approach to model points in Java would be to use a class 
+with fields for the coordinates. However here we will illustrate how 
+points are modelled with interfaces, providing a simple 
+example of how to model state with interfaces. Since Java disallows fields 
+inside interfaces, the state is modelled using abstract (getter) methods:
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-That is, since Java disallows fields inside interfaces,
- we simulate state by abstract methods.
-By using the new Java8 \Q@default@ methods, we can easily add implemented methods inside the interface. For example:
+
+\noindent Unfortunatelly, creating a new point object is quite cumbersome, even 
+when using anonymous classes:
+
 \begin{lstlisting}
-interface Point2D{ int x(); int y();
-  default float distanceFromOrigin(){ return Math.sqrt(x()*x()+y()*y());}
-  }
+Point2D p = new Point2D(){ public int x(){return 4;} public int y(){return 2;}}
 \end{lstlisting}
-resemble a class that have fields (getters) and methods.
-However, how can we create an instance of \Q@Point2D@ without using classes? 
-we can just provide an implementation for \Q@x()@ and \Q@y()@ as in
-\begin{lstlisting}
-new Point2D(){ public int x(){return 4;} public int y(){return 2;}}
-\end{lstlisting}
-However we are not required to use this cumbersome syntax\footnote{Available in Java from version ...} for every object allocation.
-As programmes do, for ease or reuse, we can encapsulate boring long repetitive code in a method. A static factory method in this case is appropriate:
+
+However we are not required to use this cumbersome
+syntax\footnote{Available in Java from version ...} for every object
+allocation.  As programmers do, for ease or reuse, we can encapsulate
+boring long repetitive code in a method. A static factory method in
+this case is appropriate:
 \begin{lstlisting}
@@
 \end{lstlisting}
-This obvious constructor code can be automatically generated by our \mixin annotation.
 
-By annotating the interface \Q@Point2D@, the annotation will generate a variation of the shown static method \texttt{of}, mimicking the functionality of a simple minded constructor:
-By looking the the methods that need implementation it first detects what are the fields, then 
-generate an \Q@of@ method with one argument for each of them.
+\paragraph{\Q@Point2D@ with Object Interfaces}
+This obvious constructor code can be automatically generated by our
+\mixin annotation.  By annotating the interface \Q@Point2D@, the
+annotation will generate a variation of the shown static method
+\texttt{of}, mimicking the functionality of a simple minded
+constructor: By looking at the methods that need implementations it
+first detects what are the fields, then generates an \Q@of@ method with
+one argument for each of them.That is, we can just write
 
-That is, we can just write
 \begin{lstlisting}
@@
 
-More precisely, a field or factory parameter is generated for every no-args method requiring implementation whose name does not have special meaning\footnote{Formal definition of \mixin behaviour is provided later.}.
-An example of code using \Q@Point2D@ is 
-\Q@Point2D.of(42,myPoint.y())@
-where we return a new point, using \Q@42@ as x-coordinate, and 
-taking all the other informations (only \Q@y@ in this case) from another point.
-This pattern is very common when programming with immutable data-structures; it is so common that we decided to support it in our code generation as \Q@with-@ methods, that is:
+More precisely, a field or factory parameter is generated for every
+no-args method requiring implementation whose name does not have
+special meaning\footnote{Formal definition of \mixin behaviour is
+  provided later.}.  An example of code using \Q@Point2D@ is
+\Q@Point2D.of(42,myPoint.y())@ where we return a new point, using
+\Q@42@ as x-coordinate, and taking all the other informations (only
+\Q@y@ in this case) from another point.
+
+\paragraph{\texttt{with} methods in Object Interfaces}
+This pattern is very common
+when programming with immutable data-structures; it is so common that
+we decided to support it in our code generation as \Q@with-@ methods,
+that is:
 \begin{lstlisting}

### 2015-11-20 UseMixinLombok/src/pegasusJava/Main.java

@@ -25,5 +25,4 @@ public interface Main {
 		Bird b = new BirdImpl();
 		Pegasus p = new PegasusImpl();	
 	}
-
 }

### 2015-11-20 paper/sections/Overview.tex
Commit in conflic, hope is easy to solve

--- 2015-11-18 paper/sections/Overview.tex
@@
 \lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
-
+\marco{Fix over: with(Point must be with(Point2D
+and it should be after withZ, and on more lines}
 Just to give you a feeling on how much boring code \mixin is generating,  we show the expanded code without then \mixin annotation\footnote{

### 2015-11-20 paper/sections/CaseStudy.tex
revised method overriding mechanism in Java

@@
-\section{Case studies}
+\section{Case Studies}
 Haoyuan and Yanlin
@@
 
-\lstinputlisting[linerange=25-35]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 
@@
 It is easy to add new data variants to the code in the initial system in a
-modular way. For example, the following code how to add the subtraction variant.
+modular way. For example, the following code shows how to add the subtraction variant.
 
-\lstinputlisting[linerange=39-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=48-53]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=48-57]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=57-66]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 

### 2015-11-21 paper/sections/Overview.tex
More polishing on Section 2

@@
 
-\subsection{Simple Traits with Default Methods}
+\subsection{Simple Multiple Inheritance with Default Methods}
 
@@
 The animal system modelled so far, is a simplified version of the
-system presented in Figure~\ref{}. \bruno{emphasize that, so far the
-  need for \mixin is not justified?} Lets now move on to modelling 
+system presented in Figure~\ref{}, and still does not appear to 
+justify the \mixin annotation. Lets now move on to modelling 
 the complete animal system. In order to do this, animals will include 
-a notion of points, representing their location in space.
+a notion of points, representing their location in space. As we shall
+see, modelling stateful components using plain Java 8 quickly becomes 
+cumbersome. The \mixin annotation comes to rescue here and 
+allows avoiding significant amounts of boilerplate code.
 
@@
 
-However we are not required to use this cumbersome
-syntax\footnote{Available in Java from version ...} for every object
-allocation.  As programmers do, for ease or reuse, we can encapsulate
-boring long repetitive code in a method. A static factory method in
-this case is appropriate:
+However programmers are not required to use this cumbersome syntax for
+every object allocation. As programmers do, for ease or reuse, the
+boring long repetitive code can be encapsulate in a method. A
+generalization of the \texttt{of} static factory
+method is appropriate in this case:
 \begin{lstlisting}
@@
 
+\vspace{-5pt}
 \paragraph{\Q@Point2D@ with Object Interfaces}
@@
 first detects what are the fields, then generates an \Q@of@ method with
-one argument for each of them.That is, we can just write
+one argument for each of them. That is, we can just write
 
@@
 
-More precisely, a field or factory parameter is generated for every
-no-args method requiring implementation whose name does not have
-special meaning\footnote{Formal definition of \mixin behaviour is
-  provided later.}.  An example of code using \Q@Point2D@ is
-\Q@Point2D.of(42,myPoint.y())@ where we return a new point, using
-\Q@42@ as x-coordinate, and taking all the other informations (only
-\Q@y@ in this case) from another point.
+\noindent More precisely, a field or factory parameter is generated for every
+no-args method requiring implementation (except for methods whose name have
+special meaning).  An example of code using \Q@Point2D@ is:
+
+\begin{lstlisting}
+Point2D p = Point2D.of(42,myPoint.y());
+\end{lstlisting}
+
+\noindent where we return a new point, using \Q@42@ as x-coordinate,
+and taking all the other informations (only \Q@y@ in this case) from
+another point.
 
 \paragraph{\texttt{with} methods in Object Interfaces}
-This pattern is very common
+The pattern of creating a new piece of data by reusing most
+information from an old piece of data is very common
 when programming with immutable data-structures; it is so common that
-we decided to support it in our code generation as \Q@with-@ methods,
+we decided to support it in our code generation as \Q@with@ methods,
 that is:
 \begin{lstlisting}
-@Mixin interface Point2D{ int x();    int y();
-  Point2D withX(int val);    Point2D withY(int val);}
+@Mixin interface Point2D { 
+  int x();    int y(); // getters
+  Point2D withX(int val);    Point2D withY(int val); // with methods
+}
 \end{lstlisting}
+
+\noindent Using \texttt{with} methods, the point \texttt{p} could have
+been created instead using:
+
+\begin{lstlisting}
+Point2D p = myPoint.withX(42);
+\end{lstlisting}
+
+\noindent If there is a large number of fields, \texttt{with} methods
+will save programmers from writting large amounts of tedious code that
+simply copies the values of fields. 
+\begin{comment}
 is expanded by our annotation into\footnote{
@@
 \end{lstlisting}
-If the programmer would like a different implementation, they can just provide it using \Q@default@; that is:
+\end{comment}
+Moreover, if the programmer wants a different implementation, they
+just provide it using a \Q@default@ method. For example:
 \begin{lstlisting}
-@Mixin interface Point2D{ int x(); int y();
+@Mixin interface Point2D{ 
+  int x(); int y();
   default Point2D withX(int val){ ... };    default Point2D withY(int val){ ... }}
 \end{lstlisting}
+
+\noindent can be used by programmers to define their own behaviour for 
+the \texttt{with} methods.
+
+\begin{comment}
 is expanded into
@@
 and our annotation generate code only if the method \emph{needs} implementation. In this way is trivial for the programmer to personalize the behaviour if they have any special need.
+\end{comment}
 
@@
 
-\subsection{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
-\Q@Points2D@ are mathematical entities, thus we chosen to use immutable data structure to model them; however animals are real world entities, and when an animal moves, it is the same animal that have now a different location.
-We model this with mutable state.
+\paragraph{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
+Two dimensional points are mathematical entities, thus we chosen to
+use immutable data structure to model them. However animals are real
+world entities, and when an animal moves, it is the same animal that
+now has a different location. We model this with mutable state.
+
 %Now we proceed to define \texttt{Animal} with \texttt{point} ``member
@@
 \lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
-\marco{TODO: change point into ``location''}
-Here we declare abstract getter and setter for the mutable ``field'' \Q@location@.
-As you see, we do not apply the \mixin annotation. This is morally equivalent to an abstract class in full Java; that is we do not provide a convenient way to instantiate it.
 
-\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\noindent Here we declare abstract getter and setter for the mutable ``field'' \Q@location@.
+The \mixin annotation is not used. This is morally
+equivalent to an abstract class in full Java; and there is no
+convenient way to instantiate it.
+For \texttt{Horse}, the \mixin annotation is used and a concrete implementation of \texttt{run()} method
+is defined using a default method, which also further illustrates the convenience of \emph{with} methods.
 
-For \texttt{Horse}, concrete implementation of \texttt{run()} method
-need to be defined in a default method, where 
-we also show the convenience of \emph{with} methods.
+\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+
 % by method
@@
 
-\lstinputlisting[linerange=80-88]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 
-\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=71-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=80-88]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=96-97]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-11-21 UseMixinLombok/src/test/TestAnimal.java

@@ -7,29 +7,29 @@ public class TestAnimal {
 	public static void main(String[] args) {
 		Horse horse = Horse.of(Point2D.of(0, 0));
 		System.out.print("horse initialized at ");
-		print(horse.point());
+		print(horse.location());
 		horse.run();
 		System.out.print("horse.run(); now at ");
-		print(horse.point());
+		print(horse.location());
 		System.out.println();
 		
 		Bird bird = Bird.of(Point3D.of(10, 10, 10));
 		System.out.print("bird initialized at ");
-		print(bird.point());
+		print(bird.location());
 		bird.fly();
 		System.out.print("bird.fly(); now at ");
-		print(bird.point());
+		print(bird.location());
 		System.out.println();
 		
 		Pegasus pegasus = Pegasus.of(Point3D.of(100, 100, 100));
 		System.out.print("pegasus initialized at ");
-		print(pegasus.point());
+		print(pegasus.location());
 		pegasus.run();
 		System.out.print("pegasus.run(); now at ");
-		print(pegasus.point());
+		print(pegasus.location());
 		pegasus.fly();
 		System.out.print("pegasus.fly(); now at ");
-		print(pegasus.point());
+		print(pegasus.location());
 	}
 	
 	static void print(Point2D p) {
@@ -62,28 +62,26 @@ interface Point3D extends Point2D {
 
 //BEGIN_ANIMAL
 interface Animal {
-	Point2D point();
-	void point(Point2D val);
+	Point2D location();
+	void location(Point2D val);
 }
 //END_ANIMAL
 
 //BEGIN_HORSE
 @Mixin
 interface Horse extends Animal {
-	default void run() {
-		point(point().withX(point().x() + 20));
-	}
+	default void run() {location(location().withX(location().x() + 20));}
 }
 //END_HORSE
 
 //BEGIN_BIRD
 @Mixin
 interface Bird extends Animal {
-	Point3D point();
+	Point3D location();
 	void point(Point3D val);
-	default void point(Point2D val) { point(point().with(val));}
+	default void location(Point2D val) { location(location().with(val));}
 	default void fly() {
-		point(point().withX(point().x() + 40));
+		location(location().withX(location().x() + 40));
 	}
 }
 //END_BIRD

### 2015-11-21 paper/sections/Overview.tex
More on Section 2

@@
 
-%\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+%\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+
+\noindent Creating and using a horse animal is quite simple:
+
+\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=USINGHORSE
+
+\noindent Note how the \texttt{of}, \texttt{withX} and
+\texttt{location} methods (all generated automatically) provide a
+basic interface for dealing with animals.
+
+\paragraph{Summary} Dealing with state (whether mutable or immutable)
+in object interfaces relies on a notion of abstract state, where only
+methods to interact with state are available to users. Object
+interfaces provides support for four different types abstract state
+operations, which are summarized in Figure~\ref{fig:abstractstate}.
+The abstract state operations are determined by naming conventions and the
+types of the methods. Fluent setters are a variant of mutable setters,
+and are discussed in the case study in Section~\ref{}.
+
+\begin{figure}
+\begin{tabular}{|l|l|l|}
+\hline
+ & \textbf{Example}                  & \textbf{Description } \\ \hline
+\textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \hline
+{\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
+object; updates field \texttt{x} to \texttt{val}.             \\ \hline
+\textbf{mutable setters}        & \Q@void x(int val);@ & Sets the field
+\texttt{x} to a  new value \texttt{val}.        \\ \hline
+\textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
+\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
+\end{tabular}
+
+\caption{Abstract state operations, for a field \texttt{x}, allowed by the \mixin
+  annotation. }
+
+\label{fig:abstractstate}
+
+\end{figure}
 
@@
 
+\subsection{Object Interfaces and Subtyping}
+
 \subsection{\Q@Bird@:  field type refinement}
@@
 as follows:
-\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 \marco{Fix over: with(Point must be with(Point2D
@@
 
-\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 
-\subsection{\Q@Pegasus@ and multiple Inheritance}
+\subsection{Advanced Multiple Inheritance with Object Interfaces}
 The \emph{``multiple inheritance''} case appears at interface
@@
 
-\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
-\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-11-21 UseMixinLombok/src/test/TestAnimal.java

@@ -5,7 +5,11 @@ import lombok.Mixin;
 public class TestAnimal {
 	
 	public static void main(String[] args) {
-		Horse horse = Horse.of(Point2D.of(0, 0));
+//BEGIN_USINGHORSE
+Point2D p = Point2D.of(0, 0);
+Horse horse = Horse.of(p);
+horse.location(p.withX(42));
+//END_USINGHORSE
 		System.out.print("horse initialized at ");
 		print(horse.location());
 		horse.run();

### 2015-11-21 paper/sections/Formalization.tex
Section 2 (work in progress)
Commit in conflic, hope is easy to solve
Macros to make formalization auxiliary functions more coherents
Formalization more coherent

@@
 }{
- \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC
+ \ann\ \QM{interface}\ \C\ \QM{extends}\ \Cs\ \oC \methods\ \cC
   }{interface declaration}\\
@@
 \forall i\in 1..n\ \ \Gamma \vdash \e_i \in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
+  \textsf{mtype}(\m,\C) \!=\! \C_1\ldots\C_n \!\!\to\! \C'
 %\textsf{mmodifier}(\m,\C) \neq \textbf{static}
  }{
- \Gamma \vdash \e\QM.\m\QM(\overline{\e}\QM) \in \C' }
+ \Gamma \vdash \e\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C' }
 \quad\quad
@@
 \forall i\in 1..n\  \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
-\textsf{mtypeS}(\m,\C) \!=\! \overline{\C} \!\to\! \C'
+\textsf{mtypeS}(\m,\C) \!=\! \C_1\ldots\C_n \!\to\! \C'
 %\textsf{mmodifier}(\m,\C) = \textbf{static}
 }{
-\Gamma \vdash \C\QM.\m\QM(\overline{\e}\QM) \in \C'}
+\Gamma \vdash \C\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C'}
 \quad\quad
@@
 \forall i\in 1..n\ \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C) \!=\! \overline{\C} \!\!\to\! \C'
+  \textsf{mtype}(\m,\C) \!=\! \C_1\ldots\C_n \!\!\to\! \C'
 %\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
-}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\overline{\e}\QM) \in \C'}
+}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C'}
 
@@
  \inferrule[(T-Intf)]{
-IT(\C) = \ann\ \QM{interface}\ \C\ \QM{extends}\ \overline{I} \
+IT(\C) = \ann\ \QM{interface}\ \C\ \QM{extends}\ \C_1\ldots\C_n \
 \oC\methods\ \cC\\\\
- \forall \QM{default}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
+ \forall \QM{default}\ \mh\ \oC\QM{return}\ \e\QM;\cC \in \methods,
 \ \ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
- \forall \QM{static}\ \mh\oC\QM{return}\ \e\QM;\cC \in \methods,
+ \forall \QM{static}\ \mh\ \oC\QM{return}\ \e\QM;\cC \in \methods,
 \ \ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
@@
 add $\this:I$; and add the arguments (and types) of the respective
-method. Finally, we need to check that all method headers are valid
-with respect to the superinterfaces of $I$; and we also need to check
-that all abstract and conflicted methods (that is methods, that need
-to be explicitly overriden) in the superinterfaces have been
-implemented. The definitions $\sigvalid$ and $\alldefined$ take care
-of this: \\
+method.
+Now we need to check if the object is a valid extension for that specific interface.
+This can be logically diveded in two steps:
+First  we check that all method headers are valid
+with respect to the corresponding method already present in $\C$;
 
-$
-\begin{array}{lll}
-\sigvalid(\mh_1\ldots\mh_n,I) & = &
-\forall i\in 1..n\ \ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\
+\noindent$\begin{array}{l}
+\InTextDef{10em}{\sigvalid(\mh_1\ldots\mh_n,I)
+}{
+\forall i\in 1..n\ \ \mh_i\QM;\subtype\mBody(\m^{\mh_i},\C)
+}
+\end{array}$
 
-\alldefined(\mh_1\ldots\mh_n,I) & = &
-\forall\m\mbox{ such that } \mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
-\end{array}
-$\bruno{Marco, you probably want to double-check the explanation. Any
-  more explanation needed?}
-\bruno{Also, according to the text above, mbody returns the full
-  method body, so in sigvalid, shouldn't we use mtype instead? The
-  same for alldefined, should it be mtype? Still not very confident
-  about alldefined. }
+Here we require that for all newly declared methods, there is a
+method with the same name defined in the interface $\C$, and that such method is a subtype of the newly introduced one.
+We formally define here subtyping between method, in a more general form that will be useful later.
+
+\noindent$\begin{array}{l}
+\InTextDef{18em}{
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR
+}{
+\T\subtype \T'
+}\\
+\InTextDef{18em}{\method \subtype
+\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}
+}{
+\method\subtype\mh\QM;
+}\\
+\InTextDef{18em}{\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method
+}{
+\mh\QM;\subtype\method
+}\\
+\end{array}$
+
+Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
+That is, we allows return type specialization as introduced in Java\marco{insert the version number}.
+Default methods are subtypes if their method headers are subtypes.
+
+Finally, all abstract methods in the interface (that is methods that need
+to be explicitly overridden) have been
+implemented. That is, we define a method with the same name.
+
+\noindent$\begin{array}{l}
+\InTextDef{11em}{\alldefined(\mh_1\ldots\mh_n,I)}{
+\forall\m\mbox{ such that } \mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m}
+\end{array}$
 
@@
 We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.\\*
-
-$\mBody(\m,\C_0) = \override(\methods(\m),
+\noindent$\begin{array}{l}
+\InTextDef{15ex}{
+\mBody(\m,\C_0)
+}{
+\override(\methods(\m),
 \shadow(\m,\Aux{tops}(\Cs)))
-$\\*
-Where
-$\metaVar{IT}(C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs \oC\methods\cC$
+}\\
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\cC}
+\end{array}$
 
@@
 the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:\\*
-$\method\in\tops(\m,\Cs) \mbox{ iff }\C\in\Cs\ ,\method=\mBody(\m,\C), \method\mbox{ not a static method}$\\*
-${}_{}$\tab and $\nexists \C'\in\Cs\setminus\C$ such that $\C' \subtype \C$.
+
+\noindent$\begin{array}{l}
+\InTextAssert{25ex}{
+\method\in\tops(\m,\Cs)}\\
+\InTextWith{\C\in\Cs\ ,\method=\mBody(\m,\C), \method\mbox{ not a static method and } \nexists \C'\in\Cs\setminus\C \mbox{ such that } \C' \subtype \C}
+\end{array}$
 
@@
 We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:\\*
-\!\!\!\!$\begin{array}{l}
-%\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
-%\whereNote
-% \mwhere\ \mBody(\m,\C_i)\in\methods
-%\ \mif\ \mBody(\m,\C_i)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}
-%\\ &\mor\ \mBody(\m,\C)=\mh\QM;
-%\\
+\noindent
+$\begin{array}{l}
+\InTextDef{15ex}{
+\shadow()}{\none}\\
+\InTextDef{15ex}{\shadow(\method)}{\method}\\
+\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
+\InTextDef{15ex}{\shadow(\methods)}{\conflicted\mh\QM;}\\
+\InTextWith{\methods\mbox{ not of the form }\overline{\mh\QM;}\mbox{ and }
+\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
+\end{array}$
 
-\shadow()=\none\\
-\shadow(\method)=\method\\
-\shadow(\overline{\mh\QM;})=\Aux{mostSpecific}(\overline{\mh\QM;})\\
-\shadow(\methods)=\conflicted\mh\QM;
-\whereNote
-\mwhere\
-\methods\mbox{ not of the form }\overline{\mh\QM;}\mbox{ and }
-\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}\\
-\Aux{mostSpecific}(\methods)=\method
-\whereNote
-\mwhere\ \method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype
-                                       \method', \\
-\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR
-\whereNote
- \mwhere\ \T\subtype \T'\\
+Where $\Aux{mostSpecific}$ return the most specific method , that is a method whose type is the subtype of all the others.
+Since method subtyping is a partial ordering, this may be not defined, this in turn makes \shadow{}, and the whole \mBody{} not defined for that specific $\m$.
+Rule  \rn{(t-intf)}  rely on this behaviour.
 
-\method \subtype
-\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}=
-\method\subtype\mh\QM;
-\\
-\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method=
-\mh\QM;\subtype\method
+\noindent$\begin{array}{l}
+\InTextDef{18ex}{\Aux{mostSpecific}(\methods)}{\method}\\
+\InTextWith{\method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype \method'}
+
 \end{array}$
-${}_{}$\\*
-Where $\Aux{mostSpecific}$ return the most specific method using return type specialization as introduced in Java\marco{insert the version number}
-We just check the subtype between method headers, so we discard information abut method implementation.\\*
+
 ${}_{}$\\*
@@
 Note how we use the special value $\none$, and how (forth case) overriding can solve a conflict.
-\\*
-\!\!\!\!$\begin{array}{ll}
-\override(\none,\none)=\none\\
-\override(\method,\none)=\method\\
 
-\override(\none,\method)=\method
-\whereNote
-\mwhere\ \method\mbox{ not of the form } \conflicted\ \mh\QM;\\
-\override(\method,\method')
-=
-\method
-\whereNote
-\mwhere\ \method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
-\method\subtype\method'
-\\
-%\override(\method,\mh')
-%=\method
-%\whereNote
-%\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}, \mh\subtype\mh'
-%\override(\_,\_)=\error & Otherwise
-%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
-%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
-%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
+\noindent$\begin{array}{l}
+\InTextDef{18ex}{\override(\none,\none)}{\none}\\
+\InTextDef{18ex}{\override(\method,\none)}{\method}\\
+
+\InTextDef{18ex}{\override(\none,\method)}{\method}\\
+\InTextWith{\method\mbox{ not of the form } \conflicted\ \mh\QM;}\\
+\InTextDef{18ex}{\override(\method,\method')}{\method}\\
+\InTextWith{\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
+\method\subtype\method'}
 \end{array}$
@@
 \end{comment}
-
-\marco{The notation mtype get only non static methods, the notation mtypeS get only static ones}
-
-%\subsubsection{Interface and method Typing}
 
@@
 
-%\lstinputlisting[linerange=5-10]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
 

### 2015-11-22 paper/sections/CaseStudy.tex
EP case study text revised with independent extensibility

@@
 by Wadler about modular extensibility issues in software evolution and has been
-a hot topic in programming languages since. Today we know of various solutions
-to the EP that either rely on \emph{new programming language
-  features}~\cite{chambers95multimethods,clifton00multiJava,madsen89virtual,nystrom06j,bruce98astatically,McDirmid01jiazzi,garrigue98polymorphic,zenger01extensible,loh06open,wehr11javaGI},
-or can be used as \emph{design patterns}~\cite{gof} in existing
-languages~\cite{torgersen04theexpression,oliveira06extensible,wouter08datatypes,oliveira09modular,oliveira12extensibility}. In
-this paper we show that the EP can be solved with our \mixin annotation in a
-trivial way. We will use a canonical example of implementing arithmetic
-expressions gradually to illustrate the usage.
+a hot topic in programming languages since. The requirements of a solution to
+the EP can be summarized as: 
+\begin{itemize}
+\item \emph{Extensibility in both dimensions}: A solution must allow
+  the addition of new data variants and new operations and support
+  extending existing operations.
+\item \emph{Strong static type safety}: A solution must prevent
+  applying an operation to a data variant which it cannot handle using
+  static checks.
+\item \emph{No modification or duplication}: Existing code must not be
+  modified nor duplicated.
+\item \emph{Separate compilation and type-checking}: Safety checks or
+  compilation steps must not be deferred until link or runtime.
+\item \emph{Independent extensibility}~\cite{zenger05independentlyextensible}:
+  It should be possible to combine independently developed extensions so that
+  they can be used jointly.
+\end{itemize}
 
-\subsubsection{Initial System}
+Today we know of various solutions to the EP that either rely on \emph{new
+  programming language features}
+% ~\cite{chambers95multimethods, clifton00multiJava, madsen89virtual, nystrom06j,
+%   bruce98astatically, McDirmid01jiazzi, garrigue98polymorphic,
+%   zenger01extensible, loh06open, wehr11javaGI}, 
+or can be used as \emph{design patterns}%~\cite{gof}
+in existing languages.
+% ~\cite{torgersen04theexpression, oliveira06extensible,
+%  wouter08datatypes, oliveira09modular, oliveira12extensibility}. 
+Now we show that the EP can be solved with our \mixin annotation in a trivial
+way. We will use the canonical example of implementing arithmetic expressions
+gradually to illustrate the usage.
+
+\paragraph{Initial System}
 The initial system expresses arithmetic expressions with only literals and
@@
 
-\subsubsection{Adding a New Variant}
+\paragraph{Adding a New Variant}
 It is easy to add new data variants to the code in the initial system in a
-modular way. For example, the following code shows how to add the subtraction variant.
+modular way. For example, the following code shows how to add the subtraction
+variant.
 
@@
 
-\subsubsection{Adding a New Operation}
+\paragraph{Adding a New Operation}
 Adding new operations to the previous system is still straightforward (although
@@
 base interfaces \texttt{Lit} and \texttt{Add}, respectively. Importantly, note
-that in \texttt{AddP}, the types of ``\emph{fields}'' \texttt{e1,e2} are refined via
-the \emph{return types refinement} of getter methods \texttt{e1()} and
+that in \texttt{AddP}, the types of ``\emph{fields}'' \texttt{e1,e2} are refined
+via the \emph{return types refinement} of getter methods \texttt{e1()} and
 \texttt{e2()}.
+
+\paragraph{Independent Extensibility}
+To show that our approach supports independent extensibility, we first define a
+new operation \texttt{collectLit}(which collects all
+literal components in an expression) on expressions:
+
+\lstinputlisting[linerange=70-86]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_COLLECTLIT
+
+\noindent Now we combine the two extensions (\texttt{print} and
+\texttt{collectLit}) together:
+
+\lstinputlisting[linerange=90-94]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+
+\noindent \texttt{ExpPC} is the new expression interface supporting
+\texttt{print} and \texttt{collectLit} operations; \texttt{LitPC} and
+\texttt{AddPC} are the extended variants. Notice that except for the routine of
+\textbf{extends} clauses, no glue code is required. Return types of
+\texttt{e1,e2} are also automatically refined to \texttt{ExpPC}.
 

### 2015-11-22 UseMixinLombok/src/test/TestExpression.java

@@ -26,7 +26,7 @@ public class TestExpression {
 	    
 	    //independent extensibility
 	    AddPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
-	    System.out.println(e8.print() + " = " + e8.eval() + "\nLiterals" + e8.collectLit().toString());
+	    System.out.println(e8.print() + " = " + e8.eval() + " Literals: " + e8.collectLit().toString());
 	}
 }
 

### 2015-11-22 paper/sections/Formalization.tex

@@
+\section{Interaction of interface methods with interface composition}
+From Java8 interfaces can have three type of methods:
+abstract methods, default methods, static methods.
+\begin{itemize}
+\item Static methods are handled in a very clean way: they are visible only on the interface they are explicitly declared. This means that 
+\begin{lstlisting}
+interface A0 { static int m(){return 1;} }
+interface B0 extends A0 {}
+...
+B0.m()//ill typed
+\end{lstlisting}
+Note how this is different w.r.t. the way static methods are handled in classes.
+In this way static methods have simply no interaction with interface composition (\Q@extends@ or \Q@implements@).
+
+\item Abstract methods composition is accepted when there is one that is the most specific.  For example in 
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+Method \texttt{Integer m()} is visible in \Q@C1@.
+
+\item Default methods conflict with any other default or abstract method. For
+example:
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is in contrast with what happens in most trait models, where \Q@D2@ would be accepted, and the implementation in \Q@A2@ would be part of the behaviour of \Q@D2@.
+
+\item The method in the current interface wins over any methods defined in its
+super-interfaces, provided that the method conform to the subtype of all methods
+in its super-interfaces, i.e., the method is the most specific one.
+This also override rejection due to conflicting methods. For example:
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { int m(); } //accepted
+\end{lstlisting}
+\end{itemize}
+
+While trying to formally encode the Java specification we have done some tests to clarify corner case behaviour.
+Consider the following correct declarations:
+\begin{lstlisting}
+interface A1{T m(); }
+interface A2 extends A1{default T m(){ ... } }
+interface A3 extends A2{T m(); }
+
+interface B1{default T m(){ ... } }
+interface B2 extends B1{T m(); }
+interface B3 extends B2{default T m(){ ... } }
+\end{lstlisting}
+
+What happens if we define a new interface \Q@M@ extending one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
+\noindent
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
+\textbf{B3}        & \cellcolor[HTML]{C0C0C0} conservative error &conflict  error       & conservative error      \\ \hline
+\end{tabular}
+\\*
+We try to classify the results out of the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both implemented. This is also considered an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both abstract. This is also considered correct in all trait models.
+\item \textbf{conservative error}
+ happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ is implemented in only one side.
+This is different from what we would expect in a trait model, but is coherent with the conservative idea that a method defined in an interface
+should not silently satisfy a method in another one.
+\end{itemize}
+
+During our experimentation, we found a bug in ECJ (eclipse version of javac): The case \textbf{B3},\textbf{A1}
+is accepted by ECJ4.5.1 and rejected by javac.
+By email communication with Brian Goetz (lead Java8 designer) we have confirmation that the expected behaviour is rejection, and that this is a bug in ECJ.
+%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
+%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+%named as \texttt{m} which are either abstract or default methods. If we define a
+%new interface \texttt{M} that extends two of these interfaces, then the method
+%overriding result is shown in the table. For example, row2 col2 means \texttt{M
+%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+%following this way of interpretation.
+
+% The interesting case in row4 col2 where
+%Indeed, to be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
+%We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
+%We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
+%the experimental result is \texttt{B3.m}. Because if we think consistently, this
+
+%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+%In our approach, we choose to not model this strange behaviour (a bug?).
+%Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
+%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java change its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
+
 \section{Formal Semantics}\label{sec:formal}
@@
 FeatherweightJava-like calculus which models the essence of Java 
-interfaces with default methods. This calculus also supports a minimal
-extension to Java to account for the \mixin annotation for object
-interfaces. The syntax, typing rules and method lookup mechanisms 
-of CJ are modelled. Finally, using our formalization, we discuss
-examples of real Java programs that seem to be incoherent\bruno{revise
-later}.
+interfaces with default methods.
 
@@
 
-\subsection{Syntax and Typing}\marco{Future work: updating multiple fields in one method call,
-  \texttt{with(T v)}}
+\subsection{Syntax}
 
-\paragraph{Syntax}
-Figure~\ref{Grammar} shows the syntax of ClassLess Java.  \footnote{To
+Figure~\ref{Grammar} shows the syntax of ClassLess Java.%
+\footnote{To
   be compatible with java, the concrete syntax for an interface
@@
 version of Java 8, focusing on interfaces, default methods and object
-creation literals.  There is no syntax for classes.  Expressions
+creation literals.  There is no syntax for classes.
+To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all maps to a single set of identifiers as in Java.  Expressions
 consist of conventional constructs such as variables ($\x$), method
@@
 static) method $\m$ over the $\this$ receiver, but statically
-dispatching to the version of the method defined in the interface
+dispatching to the version of the method as visible in the interface
 $\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
@@
 simplicity) all the fields are initialized with a variable present in
-scope. Note how our language is a subset of Java 8.  The single
-non-Java 8 piece of syntax is the \mixin annotation, which is the only
-one interesting piece of syntax in this article.  Following standard
+scope. Note how our language is a subset of Java 8.
+% The single
+%non-Java 8 piece of syntax is the \mixin annotation, which is the only
+%one interesting piece of syntax in this article.
+%??? how is @Mixin not java8? in the sense that is java5?
+  Following standard
 practise, we consider a global Interface Table (\metaVar{IT}) mapping
-from interface names $\C$ to interface declarations \metaVar{I}.
+from interface names $\C$ to interface declarations $\II$.
 
@@
 
-\paragraph{Typing} 
+\subsection{Typing}
+
 Typing statement $\Gamma \vdash \e\in\C$ reads ``in the environment
@@
 \textbf{extends} clauses in the interfaces. We omit the definition of
-the usual\bruno{where can it be found? add reference here?}, traditional subtyping relation between interfaces.%
-\footnote{ Notice how there are no classes, thus there is no
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.%
+\footnote{\marco{find a better place for} Notice how there are no classes, thus there is no
   subclassing.  We believe that this approach may scratch an old
@@
   those two viewpoints and ClassLess Java design. We do not have
-  Classes purely in the Java sense.}\bruno{is a footnote here the
-  appropriate place to have this discussion?}
+  Classes purely in the Java sense.}
 The auxiliary notation $\Gamma^\mh$ trivially
@@
 \inferrule[(T-Invk)]{
- \Gamma \vdash \e \in \C \\\\
+ \Gamma \vdash \e \in \C_0 \\\\
 \forall i\in 1..n\ \ \Gamma \vdash \e_i \in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C) \!=\! \C_1\ldots\C_n \!\!\to\! \C'
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
 %\textsf{mmodifier}(\m,\C) \neq \textbf{static}
  }{
- \Gamma \vdash \e\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C' }
+ \Gamma \vdash \e\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C }
 \quad\quad
@@
 \forall i\in 1..n\  \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
-\textsf{mtypeS}(\m,\C) \!=\! \C_1\ldots\C_n \!\to\! \C'
+\textsf{mtypeS}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\to\! \C
 %\textsf{mmodifier}(\m,\C) = \textbf{static}
 }{
-\Gamma \vdash \C\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C'}
+\Gamma \vdash \C_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
 \quad\quad
@@
 \inferrule[(T-SuperInvk)]{
-\Gamma(\this) <: \C \\\\
+\Gamma(\this) <: \C_0 \\\\
 \forall i\in 1..n\ \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
-  \textsf{mtype}(\m,\C) \!=\! \C_1\ldots\C_n \!\!\to\! \C'
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
 %\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
-}{\Gamma \vdash \C\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C'}
+}{\Gamma \vdash \C_0\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
 
@@
 %\subsubsection{Interface Table}
-
-\subsection{Method Overriding Mechanism}
-We summarize the mechanism for method overriding below (all methods considered
-here are methods with the same name and argument types):
-\begin{enumerate}
-\item Two abstract methods with subtyping relation do not cause conflict, and the
-most specific one wins. Subtyping relation for two methods means they have the
-same arguments' types and one return type is a subtype of the other one. For
-example:
-\begin{lstlisting}
-interface A1 { Object m(); }
-interface B1 { Integer m();}
-interface C1 extends A1,B1 {} //accepted
-\end{lstlisting}
-Method \texttt{Integer m()} in \texttt{B} wins over method \texttt{Object m()}
-in \texttt{A} because \texttt{Integer <: Object}.
-
-\item A default method conflict with another default method or abstract method. For
-example:
-\begin{lstlisting}
-interface A2 { default int m() {return 1;}}
-interface B2 { int m(); }
-interface C2 { default int m() {return 2;}}
-interface D2 extends A2,B2 {} //rejected due to conflicting methods
-interface E2 extends A2,C2 {} //rejected due to conflicting methods
-\end{lstlisting}
-Note how this is in contrast with what happens in most trait models, where \Q@D2@ would be accepted, and the implementation in \Q@A2@ would be part of the behaviour of \Q@D2@.
-
-\item The method in the current interface wins over any methods defined in its
-super-interfaces, provided that the method conform to the subtype of all methods
-in its super-interfaces, i.e., the method is the most specific one. For example:
-\begin{lstlisting}
-interface D3 extends A2,B2 { int m(); } //accepted
-interface E3 extends A2,C2 { int m(); } //accepted
-\end{lstlisting}
-\end{enumerate}
-\subsection{Method Overriding Incoherence in Java}
-% Please add the following required packages to your document preamble:
-% \usepackage[table,xcdraw]{xcolor}
-% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-While trying to formally encode the Java specification we have done some tests to clarify corner case behaviour.
-Consider the following correct declarations:
-\begin{lstlisting}
-interface A1{T m(); }
-interface A2 extends A1{default T m(){ ... } }
-interface A3 extends A2{T m(); }
-
-interface B1{default T m(){ ... } }
-interface B2 extends B1{T m(); }
-interface B3 extends B2{default T m(){ ... } }
-\end{lstlisting}
-
-What happens if we define a new interface \Q@M@ implementing one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
-\noindent
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
-\textbf{B1}        & conservative error                        & conflict error      & conservative error       \\ \hline
-\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
-\textbf{B3}        & \cellcolor[HTML]{C0C0C0} accepted? &conflict  error       & conservative error      \\ \hline
-\end{tabular}
-\\*
-We try to classify the results out of the table:
-\begin{itemize}
-\item \textbf{conflict error} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both implemented. This is also considered an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both abstract. This is also considered correct in all trait models.
-\item \textbf{conservative error}
- happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ is implemented in only one side.
-The is different from what we would expect in a trait model, but it may make sense with the conservative idea that a method defined in an interface
-should not silently satisfy a method in another one.
-\end{itemize}
-
-We are very confused on the result for \textbf{B3},\textbf{A1}, and we can not find any rational for this behaviour.
-\footnote{We refer to the behaviour of \texttt{javac} as in JDK version is\marco{add}.}
-
-%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
-%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
-%named as \texttt{m} which are either abstract or default methods. If we define a
-%new interface \texttt{M} that extends two of these interfaces, then the method
-%overriding result is shown in the table. For example, row2 col2 means \texttt{M
-%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
-%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
-%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
-%following this way of interpretation.
-
-% The interesting case in row4 col2 where
-To be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
-We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
-We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
-%the experimental result is \texttt{B3.m}. Because if we think consistently, this
-
-%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
-
-In our approach, we choose to not model this strange behaviour (a bug?).
-Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
-The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java change its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.

### 2015-11-22 paper/sections/CaseStudy.tex
case study section summary

@@
 \section{Case Studies}
-Haoyuan and Yanlin
+In this section we conduct three case studies which reveals various advantages
+using \mixin annotation. 1st case study provides a easy and straightforward way
+to solve the Expression Problem. 2nd case study models a simple game and the
+comparison between our approach and others shows that code amount is reduced
+tremendously using \mixin. 3rd case study models an embedded DSL for SQL
+languages with fluent interfaces and shows that \mixin not only removes verbol
+code for defining fluent interfaces but also does implicit refinement along
+extension.
 

### 2015-11-22 paper/sections/Translation.tex
more fixing

@@
 
-\noindent$[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!] =
+\noindent$\begin{array}{l}
+\InTextDef{18em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
+}{
 \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
 \methods\ \methods' \cC
-$\\*${}_{}$\tab
-where  $\valid(\C_0)$,\Q@of@$\notin\dom(\methods)$ and $\methods'=\ofMethod(\C_0) \ \otherMethod(\C_0,\methods)$.\\*
+}\\
+\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0) \cup \otherMethod(\C_0,\methods)}
+\end{array}$
 
@@
 \noindent$\begin{array}{l}
-\ofMethod(\C_0) = \
+\InTextDef{5em}{\ofMethod(\C_0)}{
  \QM{static}\ \C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
 \QM{\{}
-\QM{return new}\ \C_0 \oR\cR\ \QM{\{} \\
-\tab\tab\tab\tab\tab\tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
-\tab\tab\tab\tab\tab\tab\tab
+\QM{return new}\ \C_0 \oR\cR\ \QM{\{} }\\
+\tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
+\tab\tab
 \C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots
 \C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
-\tab\tab\tab\tab\tab\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
-%\tab\tab\tab\tab\tab\tab\tab\cloneMethod(\C_0,\es)\\
-\tab\tab\tab\tab\tab\tab\tab\withMethod(\C_0)\\
-\tab\tab\tab\tab\tab\tab\QM{\};\}} \\
+\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
+\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
+%\tab\tab\cloneMethod(\C_0,\es)\\
+%\tab\tab\withMethod(\C_0)\\
+\tab\QM{\};\}} \\
+\InTextWith{\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}
+\mbox{and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n}
 \end{array}$
-\\*
-with $\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}$,\\*
-and $\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n$\\*
-%and $\es=\m_1\QM,\ldots\QM, \m_n$\\*
-\haoyuan{Should we include the $\withMethod(\C_0)$ in the formalization?}
 
@@
 
-\noindent$\begin{array}{ll}
-\withMethod:&\withMethod(\C,\m,\C_0,\es)=
+\noindent$\begin{array}{l}
+\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{
 \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return} \C_0\QM{.of(}\es\QM{);\}} \\
-&\mbox{iff }
-\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;\\
-&\withMethod(\C,\m,\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\setterMethod:&\setterMethod(\C,\m,\C_0)=
+\QM{return} \C_0\QM{.of(}\es\QM{);\}}} \\
+\InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;}\\
+\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{\emptyset\mbox{ otherwise}}\\
+\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{
 \C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
- \m\QM{= _val;return this;\}} \\
-&\mbox{iff }
-\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;\\
-&\setterMethod(\C,\m,\C_0)=\emptyset\mbox{ otherwise}\\
+ \m\QM{= _val;return this;\}}} \\
+\InTextWith{
+\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;}\\
+\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{\emptyset\mbox{ otherwise}}\\
 %\cloneMethod:&\cloneMethod(\C_0,\es)=
@@
 
-\haoyuan{method name of setter? $m$ or $\_m$?}
-
 As you can see above, \Q@with-@ and setter methods are generated if needed.
@@
 \noindent$\begin{array}{ll}
-\otherMethod:& \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)
-\\&
- \mbox{iff }
-\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isWith(\mBody(\QM{with#}\m, \C_0))
-\\&\mbox{ and } \QM{with#}\m\notin\dom(\methods)\\
-& \C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)
-\\&
- \mbox{iff }
-\C\ \m\QM{();}\in \fieldsFunc(\C_0), \isSetter(\mBody(\QM_\m, \C_0))
-\\&\mbox{ and } \QM_\m\notin\dom(\methods)\\
+\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{
+ \isWith(\mBody(\QM{with#}\m, \C_0))}\\
+\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
+\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0))}\\
+\InTextWith{
+\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+\mbox{ and } \QM_\m\notin\dom(\methods)}\\
 %&\C_0\ \QM{clone}\oR\cR\QM;\in
@@
 \noindent$\begin{array}{ll}
-\fieldsFunc:&\method\in\fieldsFunc(\C_0) \mbox{ iff }
-\method\in \dom(\C_0)\ \mand\ \isField(\method)
-\\
-\isField:&\isField(\C\ \m\oR\cR\QM;)\tab \mif\ \mnot\ \specialName(\m)\\
-\isWith:&\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)
-\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
-& \mand\ \mnot\ \specialName(\m)\\
-\isSetter:&\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)
-\tab \mif\ \C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;\\
-& \mand\ \mnot\ \specialName(\m)\\
+\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
+\isField(\method)\mand\ 
+%}\\\InTextWith{
+\method\in \dom(\C_0)
+}\\
+
+\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
+\mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+ \mand\ \mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+ \mand\ \mnot\ \specialName(\m)}\\
 
@@
 
-\haoyuan{In theorem: $I_0$ is valid and $\QM{of}\notin\dom{I_0}$. To prove the theorem we need an extra lemma: if the translation is well defined, then $[\![\II_0]\!]$ is well formed.}
+\haoyuan{In theorem: $\C_0$ is valid and $\QM{of}\notin\dom(\C_0)$. To prove the theorem we need an extra lemma: if the translation is well defined, then $[\![\II_0]\!]$ is well formed.}
 
-To understand this theorem statement, we need to understand that there are three kind of guarantees that we can offer for safety:
+To understand this theorem statement, we need to understand three kind of guarantees that we can offer for safety:
 \begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-formed; there are no type errors.
-\item \textit{Client coherence}: the client code doesn't fail if it was not failing before code generation.
-\item \textit{Heir coherence}: the heirs are safe. 
+\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not written.
+\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
+\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well typed if they was well typed without the instrumentation.
 \end{itemize}

### 2015-11-22 paper/sections/Formalization.tex

@@
 \textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
-\textbf{B3}        & \cellcolor[HTML]{C0C0C0} conservative error &conflict  error       & conservative error      \\ \hline
+\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
 \end{tabular}
@@
 Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
-$\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
+$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
 
-The body of a method $\m$ contains all the relevant information with respect to that method, like the type of $\m$ as well as the modifier.
 We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.\\*

### 2015-11-22 paper/sections/CaseStudy.tex
section for SQL language case study

@@
 
-\subsection{Other case studies}
-\marco{should we put here the DB with fluent setters?}
+\subsection{Embedded DSLs with Fluent Interfaces}
+Since the style of fluent interfaces was invented in Smalltalk as method
+cascading, more and more languages came to support fluent interfaces, including
+JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
+interfaces, programmers have to either handwritten everything or create a
+wrapper around the original non-fluent interfaces, using \textbf{this}. In Java,
+several libraries help create fluent APIs easily (by code generation), including
+jOOQ, op4j, fluflu, JaQue, etc. These libraries have given us hints on the
+implementation of fluent setters in ClassLess Java.
+
+When creating fluent interfaces, our approach has two main advantages:
+\begin{enumerate}
+\item Instead of forcing programmers to hand write code using \textbf{return
+    this}, our approach with \mixin annotation removes this verbosity and
+  automatically generate fluent setters.
+\item Along the extension direction, the return types of fluent setters are
+  automatically refined.
+\end{enumerate}
+
+\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
+The first query language \texttt{Database} (conceptually) models
+\texttt{select}, \texttt{from} and \texttt{where} clause:
+\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
+
+\noindent Note that we are only illustrating the idea of modeling embedded DSLs
+with the database example, no concrete implementation of database connection nor
+operations are provided in this case study. One benefit that fluent methods give
+us is method chaining convenience. So when making a query using \texttt{select},
+\texttt{from} and \texttt{where}, instead of writing code like:
+\lstinputlisting[linerange=26-29]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=NORMAL_QUERY
+
+\noindent We can write more concise code below:
+\lstinputlisting[linerange=35-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
+
+\noindent Now we extend the first query language with a new feature
+\texttt{orderBy} which orders the result records by a field that users
+specified. The key is to extend interface \texttt{Database} with only the
+feature we need to add and the return type of fluent setters in
+\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
+\lstinputlisting[linerange=17-20]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+
+% \lstinputlisting[linerange=44-48]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2

### 2015-11-22 paper/sections/RelatedWork.tex
section for SQL language case study

@@
 
-\subsection{Fluent Interfaces}
-Talk about existing libraries that use fluent interfaces (do any of
-those use any code generation to assist in the implementation?).
-
-Since the style of fluent interfaces was invented in Smalltalk as method
-cascading, more and more languages came to support fluent interfaces, including
-JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
-interfaces, programmers have to either handwritten everything or create a
-wrapper around the original non-fluent interfaces, using \textbf{this}. In Java,
-several libraries help create fluent APIs easily (by code generation), including
-jOOQ, op4j, fluflu, JaQue, etc. These libraries have given us hints on the
-implementation of fluent setters in ClassLess Java.
-
 \subsection{Formalization of Java8}

### 2015-11-22 paper/sections/Overview.tex

@@
 \begin{figure}
-\centering
-\includegraphics[height=4cm]{PegasusDetail.pdf}
-\caption{Pegasus Example.}\label{fig:pegasus}
+\begin{tabular}{l|l}
+\includegraphics[height=4cm]{PegasusDetail.pdf}\hspace{20pt} &
+\begin{minipage}{8cm}
+\vspace{-90pt}
+\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+\end{minipage}
+\end{tabular}
+\caption{The animal system. Complete animal system on the left, and code for simplified animal system on the right. }\label{fig:pegasus}
 \end{figure}
@@
 Suppose that we want to model an animal system as shown in
-Fig~\ref{fig:pegasus}. \texttt{Horse} and \texttt{Bird} are subtypes
+the left side of Figure~\ref{fig:pegasus}. \texttt{Horse} and \texttt{Bird} are subtypes
 of \texttt{Animal}, with methods \texttt{run()} and \texttt{fly()},
@@
 Ignoring points, for the moment, a 
-first attempt to model the animal system in Java 8 is shown next:
-
-\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
-
+first attempt to model the animal system in Java 8 is given on the right-side 
+of Figure~\ref{fig:pegasus}.
 Note that the implementations of the methods \texttt{run}
@@
 
-\bruno{
-Note that \texttt{Bird} can fly, therefore refines the
-type of \texttt{point} from \texttt{Point2D} to
-\texttt{Point3D}. 
-}
-
 \paragraph{Instantiation}
@@
 
-\lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
+\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
 
@@
 provide object instantiation. 
-
 Note that, for this very simple animal system, plain Java 8 anonymous
@@
 
-However, as we shall see, once the system gets a little more
+\noindent However, as we shall see, once the system gets a little more
 complicated, the code for instantiation quickly becomes more 
@@
 The animal system modelled so far, is a simplified version of the
-system presented in Figure~\ref{}, and still does not appear to 
+system presented in the left-side of Figure~\ref{fig:pegasus}, and still does not appear to 
 justify the \mixin annotation. Lets now move on to modelling 
@@
 points are modelled with interfaces, providing a simple 
-example of how to model state with interfaces. Since Java disallows fields 
+example of how to model immutable state with interfaces. Since Java disallows fields 
 inside interfaces, the state is modelled using abstract (getter) methods:
@@
 
-However programmers are not required to use this cumbersome syntax for
+\noindent However programmers are not required to use this cumbersome syntax for
 every object allocation. As programmers do, for ease or reuse, the
@@
 when programming with immutable data-structures; it is so common that
-we decided to support it in our code generation as \Q@with@ methods,
+it is supported in our code generation as \Q@with@ methods,
 that is:
@@
 
-\subsection{\Q@Bird@:  field type refinement}
+%\subsection{\Q@Bird@:  field type refinement}
 
+\Q@Bird@s are \Q@Animal@s, but while \Q@Animal@s only need 2D
+locations, \Q@Bird@s need 3D locations. Therefore when the \texttt{Bird}
+interface extends the animal interface, the notion of points needs to
+be refined. Such kind of refinement usually poses a challenge
+with typical class-based approaches. Fortunatelly, with object interfaces,
+we are able to provide a simple and effective solution to that problem.
+
+\paragraph{Unsatisfactory class-based solutions to field type refinement}
+In Java if we define an animal class with a field we have a set of unsatisfactory options in front of us:
+\begin{itemize}
+\item Define a \Q@Point3D@ field in \Q@Animal@:
+this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
+\item Define a \Q@Point2D@ field in \Q@Animal@ and define an extra \Q@int z@ field in \Q@Bird@.
+This solution is very ad-hoc, requires to basically duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside of \mbox{\Q@Bird@}.
+Again, there are many reasons this would be bad, the most dramatic is that it would not scale to a scenario when the programmer of \Q@Bird@ and the programmer of \Q@Point3D@ are different.
+\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects in the field and cast
+the value out of the \Q@Point2D@ field to \Q@Point3D@ when implementing the overridden getter.
+This solution scales to the multiple programmers approach, but requires ugly casts and can be implemented in a wrong way leading to bugs.
+\end{itemize}
+
+Many readers would now think that a language extension is needed. Instead, with object interfaces, 
+another approach is possible.
+%Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}.
+
+\paragraph{Field type refinement with object interfaces}
+In our proposed solution,
+\begin{itemize}
+\item by covariant method overriding we refine the field type
+\item by \emph{overloading} we define a new setter with the more precise type
+\item by default method we provide an implementation for the setter called with the old signature.
+\end{itemize}
+
+\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+
+%Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
+%\texttt{int z()} (treated as a getter for member field \texttt{z}). Note that
+%the return type of various methods (e.g. with- methods, getters) get refined
+%either by covariant method overriding or automatically by our annotation
+%processor. Besides \emph{with}, other methods (including \emph{clone},
+%\emph{of}) also do type-refinements automatically.
+
+From the type perspective, the key is the covariant method overriding; however from the semantic perspective the key is the implementation for the setter with the old signature:
+by using the \Q@with@ method we may use the information for \Q@z@ already stored in the object to forge an appropriate \Q@Point3D@ to store.
+Note how all the informations about what fields sits in \Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the \Q@with@ method, and is transparent for the implementer of \Q@Bird@.
+
+\bruno{See how to fit this in.}
 \subsubsection*{\Q@Point3D@ and properties updater}
@@
 The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.
-
-\subsubsection*{Unsatisfactory class based solutions to field type refinement}
-
-\Q@Bird@s are \Q@Animal@s, 
-but while \Q@Animal@s, only need 2D locations,
-\Q@Bird@s need  3D locations.
-In Java if we define an animal class with a field we have a set of unsatisfactory options in front of us:
-\begin{itemize}
-\item Define a \Q@Point3D@ field in \Q@Animal@:
-this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
-\item Define a \Q@Point2D@ field in \Q@Animal@ and:
-\item Define an extra \Q@int z@ field in \Q@Bird@.
-This solution is very at-hoc, requires to basically duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside of \mbox{\Q@Bird@}.
-Again, there are many reasons this would be bad, the most dramatic is that it would not scale to a scenario when the programmer of \Q@Bird@ and the programmer of \Q@Point3D@ are different.
-\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects in the field and cast
-the value out of the \Q@Point2D@ field to \Q@Point3D@ when implementing the overridden getter.
-This solution scales to the multiple programmers approach, but requires ugly casts and can be implemented in a wrong way leading to bugs.
-\end{itemize}
-
-Many readers would now think that a language extension is needed.
-Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}.
-\subsubsection*{Our approach to field type refinement}
-
-\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
-
-%Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
-%\texttt{int z()} (treated as a getter for member field \texttt{z}). Note that
-%the return type of various methods (e.g. with- methods, getters) get refined
-%either by covariant method overriding or automatically by our annotation
-%processor. Besides \emph{with}, other methods (including \emph{clone},
-%\emph{of}) also do type-refinements automatically.
-
-In our proposed solution,
-\begin{itemize}
-\item by covariant method overriding we refine the field type
-\item by \emph{overloading} we define a new setter with the more precise type
-\item by default method we provide an implementation for the setter called with the old signature.
-\end{itemize}
-
-From the type perspective, the key is the covariant method overriding; however from the semantic perspective the key is the implementation for the setter with the old signature:
-by using the \Q@with@ method we may use the information for \Q@z@ already stored in the object to forge an appropriate \Q@Point3D@ to store.
-Note how all the informations about what fields sits in \Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the \Q@with@ method, and is transparent for the implementer of \Q@Bird@.
 

### 2015-11-22 paper/sections/RelatedWork.tex
related work (in progress)

@@
 
-\subsection{Multiple Inheritance in Object oriented languages}
+\subsection{Multiple Inheritance in Object Oriented Languages}
 Many authors have argued in favour or against multiple inheritance.  It provides
-expressive power, but it is difficult to model, and implement and can create
+expressive power, but it is difficult to model and implement, and can create
 programs that are hard to reason about.  These difficulties include the famous
-diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods,
-and the yo-yo problem~\cite{taenzer1989problems}.  To conciliate the need for
+diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
+%and the yo-yo problem~\cite{taenzer1989problems}.  
+To conciliate the need for
 expressive power and the need for simplicity, lots of models have been proposed
@@
   provide a satisfactory resolution in all cases and restricts the flexibility
-  of mixin composition.
-
-  To fight those limitations, an algebra of mixin operators is
-  introduced~\cite{ancona2002calculus}, but this raised the complexity of the
-  approach, especially when constructors and fields are
-  considered~\cite{marco09FJigsaw}.
-
+  of mixin composition.  To fight those limitations, an algebra of mixin
+  operators is introduced~\cite{ancona2002calculus}, but this raised the
+  complexity of the approach, especially when constructors and fields are
+  considered~\cite{marco09FJigsaw}.  Our approach does not have the
+  linearisation problem, in that the semantics of Java \textbf{extends} clause
+  is unordered and symmetric.
 \item Simplifying the mixin algebras approach, traits draw a strong line between
@@
   programs.  However, other authors see the need of two concepts and the absence
-  of state as drawbacks of this model.
+  of state as drawbacks of this model. Our approach takes interfaces as units of
+  reuse, and meanwhile generates factory method for instantiation.
 
@@
   composition.
-  Python also offers multiple inheritance via linearised mixins. Indeed python any class is implicitly a mixin, and mixin composition informally expressed as\\*
+  Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
   \Q@ class A use B,C {...new methods...}@\\*
@@
   \Q@ class A(B,C,Aux): pass@
+  Our approach not only does not involve linearised mixin problem, but also
+  support proper constructor mechanism.
 \end{enumerate}
@@
 allowed to be defined (via the \textbf{default} keyword) inside
-interfaces. Since Java supports implementation of multiple interfaces (instead
-of extension of a single class), the introduction of default methods opens the
+interfaces. 
+% Since Java supports implementation of multiple interfaces (instead
+% of extension of a single class), 
+The introduction of default methods opens the
 gate for various flavours of multiple inheritance in Java, using interfaces.

### 2015-11-22 paper/sections/Formalization.tex
More progress on formalization english

@@
 \item Static methods are handled in a very clean way: they are visible only on the interface they are explicitly declared. This means that 
+the following code is ill-typed.
 \begin{lstlisting}
@@
 \item Abstract methods composition is accepted when there is one that is the most specific.  For example in 
+method \texttt{Integer m()} is visible in \Q@C1@.
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-Method \texttt{Integer m()} is visible in \Q@C1@.
 
 \item Default methods conflict with any other default or abstract method. For
-example:
+example the following code is rejected due to conflicting methods.
 \begin{lstlisting}
@@
 in its super-interfaces, i.e., the method is the most specific one.
-This also override rejection due to conflicting methods. For example:
+This also override rejection due to conflicting methods. For example, the following code is accepted, but would be rejected (see before) if the method \Q@m@ was not redefined.
 \begin{lstlisting}
 interface D3 extends A2,B2 { int m(); } //accepted
-interface E3 extends A2,C2 { int m(); } //accepted
+interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
 \end{lstlisting}
@@
 Figure~\ref{Grammar} shows the syntax of ClassLess Java.%
-\footnote{To
-  be compatible with java, the concrete syntax for an interface
-  declaration with empty supertype list $\C_1\ldots\C_n$ would also
-  omit the \Q@extends@ keyword.}  The syntax formalizes a minimal
+  The syntax formalizes a minimal
 version of Java 8, focusing on interfaces, default methods and object
@@
 scope. Note how our language is a subset of Java 8.
+To  be compatible with java, the concrete syntax for an interface
+  declaration with empty supertype list  would also
+  omit the \Q@extends@ keyword.
 % The single
@@
 $\Gamma \vdash \e\in\C<:\C'$ instead of $\Gamma \vdash \e\in\C$ and
-$\C<:\C'$.  From the interface table, we can read off the subtype
-relation between interfaces. The subtype relation is given by the
-\textbf{extends} clauses in the interfaces. We omit the definition of
-the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.%
+$\C<:\C'$.
+
+We omit the definition of
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation, that can be read off the interface table and given by the
+\textbf{extends} clauses in the interfaces. %
 \footnote{\marco{find a better place for} Notice how there are no classes, thus there is no
@@
 method with the same name defined in the interface $\C$, and that such method is a subtype of the newly introduced one.
-We formally define here subtyping between method, in a more general form that will be useful later.
+We define subtyping between methods in a general form that will be useful also later.
 
@@
 Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
-That is, we allows return type specialization as introduced in Java\marco{insert the version number}.
+That is, we allows return type specialization as introduced in Java5.
 Default methods are subtypes if their method headers are subtypes.
@@
 conflicting.
-\marco{if we have space we can discuss the other rules too, but is not mandatory}
 
@@
 ${}_{}$\\*
-\marco{we need to discuss the names tops, shadow and override}
 
-\tops{} recover from the interface table only the ``needed'' methods, that is,
+\tops{} recovers from the interface table only the ``needed'' methods, that is,
 the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:\\*
@@
 \noindent$\begin{array}{l}
-\InTextAssert{25ex}{
-\method\in\tops(\m,\Cs)}\\
-\InTextWith{\C\in\Cs\ ,\method=\mBody(\m,\C), \method\mbox{ not a static method and } \nexists \C'\in\Cs\setminus\C \mbox{ such that } \C' \subtype \C}
+\InTextAssert{35ex}{
+\mBody(\m,\C)\in\tops(\m,\Cs)}{
+}\\
+\InTextWith{
+  \mBody(\m,\C)\mbox{ not a static method and } 
+  \C\in\Cs\ \mbox{ such that }
+\forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C
+}
 \end{array}$
@@
 Since method subtyping is a partial ordering, this may be not defined, this in turn makes \shadow{}, and the whole \mBody{} not defined for that specific $\m$.
-Rule  \rn{(t-intf)}  rely on this behaviour.
+Rule  \rn{t-intf}  relies on this behaviour.
 
@@
 \noindent$\begin{array}{l}
-\InTextDef{18ex}{\override(\none,\none)}{\none}\\
-\InTextDef{18ex}{\override(\method,\none)}{\method}\\
+\InTextDef{20ex}{\override(\none,\none)}{\none}\\
+\InTextDef{20ex}{\override(\method,\none)}{\method}\\
 
-\InTextDef{18ex}{\override(\none,\method)}{\method}\\
+\InTextDef{20ex}{\override(\none,\method)}{\method}\\
 \InTextWith{\method\mbox{ not of the form } \conflicted\ \mh\QM;}\\
-\InTextDef{18ex}{\override(\method,\method')}{\method}\\
+\InTextDef{20ex}{\override(\method,\method')}{\method}\\
 \InTextWith{\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},

### 2015-11-22 paper/sections/Translation.tex
More progress on formalization english

@@
 or
-$\isSetter(\method,\C_0)$
+$\isSetter(\method,\C_0)$. \ 
 %$\isClone(\method, \C_0)$.
-
 That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.

### 2015-11-23 paper/sections/Overview.tex
Bug seems fixed. lib file lombok.jar updated, run it before testing.

@@
 
-%\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+%\lstinputlisting[linerange=41-47]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 
-\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 as follows:
-\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 \marco{Fix over: with(Point must be with(Point2D
@@
 
-\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+\lstinputlisting[linerange=41-47]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
-\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-11-23 UseMixinLombok/src/test/TestExpression.java

@@ -2,7 +2,7 @@ package test;
 
 import java.util.ArrayList;
 import java.util.List;
-import lombok.Mixin;
+import lombok.Obj;
 
 public class TestExpression {
 	public static String runTest() {
@@ -32,11 +32,11 @@ public class TestExpression {
 
 //BEGIN_EXPRESSION_INIT
 interface Exp { int eval(); }
-@Mixin interface Lit extends Exp {
+@Obj interface Lit extends Exp {
 	int x();
 	default int eval() { return x(); }
 }
-@Mixin interface Add extends Exp {
+@Obj interface Add extends Exp {
 	Exp e1(); Exp e2();
 	default int eval() {
 		return e1().eval() + e2().eval();
@@ -45,7 +45,7 @@ interface Exp { int eval(); }
 //END_EXPRESSION_INIT
 
 //BEGIN_EXPRESSION_SUB
-@Mixin interface Sub extends Exp {
+@Obj interface Sub extends Exp {
 	Exp e1(); Exp e2();
 	default int eval() {
 		return e1().eval() - e2().eval();
@@ -55,10 +55,10 @@ interface Exp { int eval(); }
 
 //BEGIN_EXPRESSION_PRINT
 interface ExpP extends Exp { String print(); }
-@Mixin interface LitP extends Lit, ExpP {
+@Obj interface LitP extends Lit, ExpP {
 	default String print() {return "" + x();}
 }
-@Mixin interface AddP extends Add, ExpP {
+@Obj interface AddP extends Add, ExpP {
     ExpP e1(); ExpP e2();
 	default String print() {
 		return "(" + e1().print() + " + " + e2().print() + ")";
@@ -68,14 +68,14 @@ interface ExpP extends Exp { String print(); }
 
 //BEGIN_EXPRESSION_COLLECTLIT
 interface ExpC extends Exp { List<Integer> collectLit(); }
-@Mixin interface LitC extends Lit, ExpC {
+@Obj interface LitC extends Lit, ExpC {
     default List<Integer> collectLit() {
         List<Integer> list = new ArrayList<Integer>();
         list.add(x());
         return list;
     }
 }
-@Mixin interface AddC extends Add, ExpC {
+@Obj interface AddC extends Add, ExpC {
     ExpC e1(); ExpC e2();
     default List<Integer> collectLit() {
[…]

### 2015-11-23 paper/sections/Appendix.tex
Comments on the proof

@@
 \subsection{Proof of LEMMA 1}\label{subsec:proof1}
+\marco{this seams to be just a sublemma for Lemma2, I think you can mention it in the appendix only.}
 \textbf{LEMMA 1. }
@@
 \item If $m\in\dom(\methods')$, we are to prove $m\in\dom([\![\II_0]\!])$.
-So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.
-On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
+So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.\marco{need to add here that we know of is not in dom(..)} On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method,
+ we can apply the last clause of override, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
 Hence $m\in\dom([\![\II_0]\!])$ as well.
@@
 
+\marco{ we need to state (may be as a new lemma):
+forall expressions e, if under interface table I0,I1..In
+$\Gamma\vdash e\in\C$, then  under interface table [[I0]],I1..In
+$\Gamma\vdash e\in\_<:\C$
+ProofScatch:
+(1) observing otherMethods(), we see that [I] is structurally a subtype of I, thus all the expressions typed in an interface table with I, are well typed in an interface table with [I]. 
+The only interesting case is method call (normal, super and static)
+consider $\Gamma\vdash e_0.m(e_1..e_n)\in\C$, with $e_0$ of type I0,
+if it is well typed in the first table, then in the other table it is still well typed:
+the method still exists and its argument types are the same type.
+The expressions 1..n may be now typed in a subtype, but this is ok since the premise of the rule ask for $\_<:\C$
+}
 \subsection{Proof of LEMMA 2}\label{subsec:proof2}
@@
 \begin{proof}
+
 By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide it into two parts.
@@
 
-Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.
-
-\haoyuan{How to more explicitly say the default methods are ok in the new interface table, so that it is sufficient to only prove the generated ofMethod is ok?}
+Since $\II_0$ OK, \marco{by newLemma} all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}.
+It suffices to prove that it still holds for $\ofMethod(\C_0)$.
 
@@
 \item For method typing of getters,
-\[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i=\C^{\mh_i}\]
+\[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i\]
+We know that $\C_i=\C^{\mh_i}$ since...\marco{finish}.
 \item For method typing of \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
-\[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0=C^{\mh_i}\]
-\item For method typing of setter methods, \haoyuan{what about the \QM{with} method?}
+\marco{the next judgement is not ok, state something like:
+We consider the with method for a arbitrary i field.
+In this case the method body we have to type would be...
+and here you can replace $e_i$ with the correct specialized expression}
+\[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
+We know that $\C_0=C^{\mh_i}$ since ..\marco{ finish}
+\item For method typing of setter methods, \haoyuan{what about the \QM{with} method?}\marco{we do not formalize it}
 \[\Gamma, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash\QM{this}\in \C_0=\C^{\mh_i}\]
+\marco{no, you also need to show that the part x=var is correct, and again push I= out}
+\marco{for under, now that the explanation have been improved, use the names of the new auxiliary functions}
 \item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
@@
  \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
- \item For \QM{with-} methods,
+ \item For \QM{with-} methods,\marco{ use if-then or other english forms instead of the logic arrow}
   \[\begin{array}{l}
@@
 And actually it is guaranteed by \textbf{LEMMA 1}.
+\marco{may be you may just delete lemma1 and put its proof here?}
 
@@
 \begin{proof}
-\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK.
-On the other hand, if some $\II$ is not a subtype of $\II_0$, the generated code in translation has no way to affect the domain of $\II$, which finishes our proof.
+\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$, 
+\marco{by the newLemma, we know that all its methods are still well typed, and } the generated code in translation has no way to affect the domain of $\II$,\marco{ so rule t-interf can still be applied,} which finishes our proof.
 \end{proof}

### 2015-11-23 paper/sections/Translation.tex
Comments on the proof

@@
 For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$
+$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
+if  
+$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
+then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$\\*
 $\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
@@
 \begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not written.
+\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not wrote.
+In our case it means that either \mixin{} produces in controlled way an understandable error, or the class can be successfully annotate and the generated code is well typed.
+We guarantee \textit{Self coherence}.
+
 \item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
+That is, the annotation do not remove any functionality, is just
+adding more behaviour.
+We guarantee \textit{Client coherence}.
+
 \item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well typed if they was well typed without the instrumentation.
-\end{itemize}
-The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed.
-In Java it implies not touching the annotated interface, and surely an alternative approach is to create a new interface and generate what we need.
-But since we are using Lombok for the translation, the transformation is applied to the annotated interface itself.
-But heir coherence is indeed an issue.
-See below how this is affected by our approach:
+This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
+We do not guarantee \textit{Heir coherence}.
+Indeed consider the following example
 
@@
 
-Here is the client code for our example.
-With interface \Q@B@ annotated with \mixin, by the translation rule, it will generate a method ``\Q@B withX(int x);@'' inside, as an automatic return type refinement.
-But in that case, interface \Q@C@ doesn't type check.
+By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
+This would break \Q@C@.
 
-Now we go back and read the statement of our theorem again.
+\end{itemize}
+%The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed.
+In Java it implies not touching the annotated interface, and surely an alternative approach is to create a new interface and generate what we need.
+But since we are using Lombok for the translation, the transformation is applied to the annotated interface itself.
+But heir coherence is indeed an issue.
+See below how this is affected by our approach:
+
+%Now we go back and read the statement of our theorem again.
 It illustrates that a program with the translation of \mixin applied still type checks if it did, except for the subtypes of the annotated interface (type safety of them is not guaranteed).
 
-\haoyuan{Text needs improving.}
+%\haoyuan{Text needs improving.}
 
@@
 
+\marco{ you need to use the Lemma/theorem macros}
 \noindent\textbf{LEMMA 2. }
 If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
+This is what we defined as self coherence before.
 

### 2015-11-24 paper/paper.tex
Ok, pdflatex is now working fine. I submit a couple of typo fix to introduction/abstract and some new text.
Proof in progress.
proof in appendix.
Minor changes in Introduction
Formalization partly revised
changes

@@
 \newcommand\mixin{\textbf{@Mixin} }
+\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
 \usepackage{xspace}
@@
 % Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Classless Java}
+\title{Classless Java: Tuning Java Interfaces}
 \titlerunning{Classless Java}
@@
 
+\begin{comment}
 \Copyright{John Q. Open and Joan R. Access}
@@
 % Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{comment}
 
@@
 
-This paper shows that, with some minor syntactic sugar, default
-methods and interface inheritance are in fact be very expressive. Our
-proposed syntax sugar, called \emph{object interfaces}, enables
-powerful object-oriented idioms, using multiple inheritance, to be
-expressed directly in Java. Object interfaces avoid some syntactic
-boilerplate caused by Java being originally designed to work mainly
-with classes. Moreover, object interfaces come with their own object
-instantiation mechanism, providing an alternative to class
-constructors. As a result many Java programs can be built without
-defining a single class! An implementation of object interfaces using 
-Java annotations and a formalization of the static and dynamic
-semantics are presented. Moreover the usefulness of object
-interfaces is illustrated through various examples and case studies.
-\bruno{a first draft of the abstract. still missing ``lightweight
-  language extensions''.}
+This paper shows that, with a simple language feature, default methods
+and interface inheritance are in fact very expressive. Our proposed
+language feature, called \emph{object interfaces}, enables powerful
+object-oriented idioms, using multiple inheritance, to be expressed
+conveniently in Java. Object interfaces refine conventional Java
+interfaces in three different ways. Firstly, object interfaces have
+their own object instantiation mechanism, providing an alternative to
+class constructors. Secondly, object interfaces support \emph{abstract
+  state operations}, providing a way to use multiple inheritance with
+state in Java. Finally, object interfaces allow type refinements that
+are often tricky to model in conventional class-based
+approaches. Interestingly, object interfaces do not require changes to
+the runtime, and they also do not introduce any new syntax (apart from
+a small annotation): all three features are achieved by reinterpreting
+existing Java syntax, and are translated into regular Java code
+without loss of type-safety. Since no new syntax is introduced, it
+would be incorrect to call object interfaces a language extension or
+syntactic sugar. So we use the term \emph{language tuning} to
+characterize this kind of language feature. An implementation of
+object interfaces using Java annotations and a formalization of the
+static and dynamic semantics are presented. Moreover the usefulness of
+object interfaces is illustrated through various examples.
 
@@
 
+\input{sections/Translation.tex}
+
 \input{sections/Implementation.tex}
@@
 \appendix
+\input{sections/Appendix.tex}
 

### 2015-11-24 paper/sections/Introduction.tex
changes

@@
 
-Still, the design has been criticised for being too conservative and 
-not very useful in its current form to model traits or multiple 
-inheritance~\cite{}. 
+Still, the design is quite conservative and appears to be quite limited
+ in its current form to model advanced forms of multiple inheritance.
 Indeed, our own personal experience of combining default methods 
@@
   getter methods is a way out of this tension, but this workaround
-  requires tedious boilerplate classes that latter implements those
+  requires tedious boilerplate classes that latter implement those
   methods.
@@
 
-\noindent Clearly, a way around those difficulties would be to try to
+\noindent Clearly, a way around those difficulties would be to
 change Java and just remove these limitations. Scala's own
@@
 around the difficulties by changing the language in fundamental ways,
-we show that, with some minor syntactic sugar, default methods and
+we show that, with a simple language feature, default methods and
 interface inheritance are in fact be very expressive. Our proposed
-syntax sugar enables powerful object-oriented idioms, using multiple
-inheritance, to be expressed directly without modifications to the
-current Java language. We call our sugar \emph{object interfaces},
-because such interfaces can be instantiated directly, without the need
-for an explict class definition. Object interfaces come with their own
-object instantiation mechanism, providing an alternative to class
-constructors. Moreover, object interfaces avoid some syntactic
-boilerplate caused by Java being originally designed to work mainly
-with classes. This includes support for various common utility methods
-(such as getters and setters, or clone methods), even in the presence
-of subtyping. As a result, with object interfaces, many Java programs
-can be built without using a single class!  \bruno{integrate more of
-  Marco's text.}
+language feature enables powerful object-oriented idioms, using
+multiple inheritance. We call the language feature \emph{object
+  interfaces}, because such interfaces can be instantiated directly,
+without the need for an explict class definition. Moreover, object
+interfaces support \emph{abstract state operations}, providing a way
+to use multiple inheritance with state in Java. The abstract state
+operations include various common utility methods (such as getters and
+setters, or clone-like methods). In the presence of subtyping, such
+operations often require special care, as their types need to be
+refined. Object interfaces provide support for type-refinement and can
+automatically produce code that deals with type-refinement
+adequately. With object interfaces, many Java programs can be built
+without using a single class!
 
-To formalize object interfaces, we propose Classless Java (CJ)\marco{I propose to call it directly ClassLess Java, the  Featherweight part is in being classless.}: 
-a FeatherweightJava-style calculus, which captures the essence 
+Object interfaces do not require changes to the
+runtime, and they also do not introduce any new syntax (apart from a
+small annotation). All three features of object interfaces are
+achieved by reinterpreting existing Java syntax, and are translated
+into regular Java code without loss of type-safety. Since no new
+syntax is introduced, it would be incorrect to call object interfaces
+a language extension or syntactic sugar. So we use the term
+\emph{language tuning} instead. Language tuning sits in between a
+lightweight language extension and a glorified library. Language
+tuning can offer many features usually implemented by a real language
+extension, but because it does not modify the language syntax
+pre-existing tools can work transparently on the tuned language.  To
+exploit the full-benefits of language tuning, our prototype
+implementation of object interfaces uses Java annotations and AST
+rewriting, allowing existing Java tools (such as IDE's) to work
+out-of-the-box with our implementation. As a result we could
+experiment object interfaces with several interesting Java programs,
+and conduct various case studies.
+
+To formalize object interfaces, we propose Classless Java (CJ): a
+FeatherweightJava-style~\cite{Igarashi01FJ} calculus, which captures the essence
 of interfaces with default methods. The semantics of object interfaces
-is given as a type-directed translation from CJ to itself. 
-In the resulting CJ code all object interfaces are translated into 
-regular CJ (and Java) interfaces with default methods. The translation 
-is proved to be type-preserving, ensuring that the translation does 
-not introduce type-errors. One interesting aspect of the theorem is that, 
-if we ignore the object annotations, the program still type-checks 
-under the regular typing rules of CJ (and Java).\bruno{double-check!}
-\marco{ I do not understand what you mean.
-1) annotations are part of java and CJ, so you do not need to remove them 2) programs well typed without expansion are well typed with expansion, here you seams to state the opposite}
-CJ's usefulness goes beyond serving as a calculus to formalize object
-interfaces. During the development process of CJ, we encountered an 
-incoherent behaviour in Java's implementation of default methods that 
-could be classified as a bug. \bruno{Should we report bug to the Java
-  developers mailing list?}
-\marco{we need to put the exact javac version}
-We also present a prototype implementation of object interfaces using
-Java annotations. One benefit of an annotation-based implementation 
-is that existing Java tools (such as IDE's) work with out-of-the-box
-with our implementation. As a result we could experiment object
-interfaces with several interesting Java programs, and conduct various
-case studies. 
-\marco{ I removed your text here and I add after to try talking about tuning}
-%I think that the qualifier object goes AGAINST the idea of tuning.
-%\bruno{mention the following or not? maybe good if we
-%  had something about tuning.} Of course, another implementation option would be to
-%extend Java, with a single qualifier keyword {\bf object}, so that
-%object interfaces would be an intrinsic part of the language. We leave
-%this for future work.
+is given as a type-directed translation from CJ to itself. In the
+resulting CJ code all object interfaces are translated into regular CJ
+(and Java) interfaces with default methods. The translation is proved
+to be type-preserving, ensuring that the translation does not
+introduce type-errors. CJ's usefulness goes beyond serving as a
+calculus to formalize object interfaces. During the development
+process of CJ, we encountered a bug in the implementation of default
+methods for the Eclipse Compiler for Java (ECJ). For the program revealing the 
+bug, ECJ behaves differently from both our formalization and Oracle's 
+Java compiler.
 
@@
 extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
-application is to show how embedded DSLs using fluent interfaces~\cite{} 
+application is to show how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
 can be easily defined using object interfaces. Finally, the last
@@
 
-Moreover, we believe that our approach is an example of a more general idea, that we call \emph{Language Tuning},
-a concept that sits in the middle between a lightweight language extension and a glorified library.
-Thanks to Java dynamic class loading and annotation processing,  annotations can trigger type-driven transformations on classes during either loading or compilation.
-There are many examples of such Language Tuning libraries,~\cite[lomboc,anotherImSearching].
-Language tuning can offers many features usually implemented by a real language extension, but they do not modify the syntax; pre-existing tools can work transparently on the tuned language.
-
 In summary, the contributions of this paper are:
@@
 
-\item {\bf Object Interfaces:} A simple feature that allows 
-various powerful programming idioms to be expressed conveniently in
-Java. Object interfaces are pure sugar: they are desugared into standard Java 
-code, without loss of modularity or type-safety.
+\item {\bf Object Interfaces:} A simple feature that allows various
+  powerful multiple-inheritanc eprogramming idioms to be expressed
+  conveniently in Java.
 
@@
 the essential features of Java 8 interfaces with default methods, and 
-can be used to formally define the desugaring of object interfaces. 
-We prove ...
+can be used to formally define the translation of object interfaces. 
+We prove a type-preservation theorem, and we present a Java program 
+that reveals a bug in the ECJ implementation of default methods.
 
-\item {\bf A Bug in the Implementation of Java 8:}\bruno{state this as
-  a contrib?}\marco{yes, especially if we "suggest solutions"}
+\item {\bf Implementation and Case Studies:} We have a prototype
+  implementation of object interfaces, using Java
+  annotations. Moreover the usefulness of object interfaces is
+  illustrated through various examples and case studies.
 
-\item {\bf Implementation and Case Studies:} We have a prototype implementation 
-  of object interfaces, using Java annotations. Moreover the usefulness
-  of object interfaces is illustrated through various examples and
-  case studies.
+\item{\bf Language Tuning} We identify this concept and we describe
+object interfaces as examples of Language Tuning. We also discuss 
+how other existing approaches, such as the annotations in project 
+Lombok~\cite{lombok}, can be viewed as language tuning.
 
-\item{\bf Language Tuning} We identify this concept and we describe Project Lomboc and object interfaces as examples of Language Tuning.
 \end{itemize}

### 2015-11-24 paper/sections/Overview.tex
changes

@@
 
-\marco{fig1 is now better. Where we cite it in the text? can we make it so that is more "linear"? point2d <- point3D,   animal <- (horse over bird ) <- pegasus}
 To propose a standard example, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.

### 2015-11-24 paper/sections/Implementation.tex
implementation revised

@@
-\section{Implementation}
+\section{Implementation with Lombok}
 
-One of the downside of Java is the code verbosity, e.g., getter and setter
-methods for fields, unneeded type annotation, etc. Lombok project~\cite{} is a
-Java tool that aims at removing (or reducing) these boilerplate code as far as
-possible, via annotations. There are a number of annotations provided by the
-original Lombok, including \textbf{@Getter}, \textbf{@Setter},
-\textbf{@ToString} for generating getters, setters and \QM{toString} methods,
-respectively.
+Our implementation is based on extension to Lombok. Lombok project~\cite{lombok}
+is a Java tool that aims at removing (or reducing) Java boilerplate code as far
+as possible, via annotations. Code verbosity is one of the downside of Java,
+usually caused by getter and setter methods for fields, unneeded type
+annotation, etc. There are a number of annotations provided by the original
+Lombok, including \textbf{@Getter}, \textbf{@Setter}, \textbf{@ToString} for
+generating getters, setters and \QM{toString} methods, respectively.
 
@@
 
-Our implementation is based on extension to Lombok. In Eclipse, with an
-interface annotated by \mixin, the automatic annotation processing is performed
-and the information of the interface from compilation is captured in the
-``Outline'' window, including all the methods inside the interface as well as
-the generated ones.
- The custom transformation is easy and convenient to use.
- For example this means that the IDE functionality 
- content assist autocomplete will work for the newly generated methods.
+The annotation we created is \mixin. In Eclipse, with an interface annotated by
+\mixin, the automatic annotation processing is performed and the information of
+the interface from compilation is captured in the ``Outline'' window, including
+all the methods inside the interface as well as the generated ones.  The custom
+transformation is easy and convenient to use.  For example this means that the
+IDE functionality content assist autocomplete will work for the newly generated
+methods.
 
@@
   \texttt{javac} is missing.
+\item The current implementation does not take type-parameter polymorphism into
+  consideration, thus does not support generics yet.
 \end{itemize}

### 2015-11-24 paper/sections/CaseStudy.tex
Case Studies section ready to review

@@
 In this section we conduct three case studies which reveals various advantages
-using \mixin annotation. 1st case study provides a easy and straightforward way
-to solve the Expression Problem. 2nd case study models a simple game and the
-comparison between our approach and others shows that code amount is reduced
-tremendously using \mixin. 3rd case study models an embedded DSL for SQL
-languages with fluent interfaces and shows that \mixin not only removes verbol
-code for defining fluent interfaces but also does implicit refinement along
-extension.
+using \mixin annotation. The first case study provides a easy and
+straightforward way to solve the Expression Problem. The second one models a
+simple game and the comparison between our approach and others shows that code
+amount is reduced tremendously using \mixin. The third one models an embedded
+DSL for SQL languages with fluent interfaces and shows that \mixin not only
+removes verbol code for defining fluent interfaces but also does implicit
+refinement along extension.
 
@@
 %  wouter08datatypes, oliveira09modular, oliveira12extensibility}. 
-Now we show that the EP can be solved with our \mixin annotation in a trivial
-way. We will use the canonical example of implementing arithmetic expressions
-gradually to illustrate the usage.
+A new solution~\cite{eptrivially} based on covariant type refinement solves EP
+and works in Java-like languages. It does not introduce new language features,
+does not even require generics support. However one of the limitation is certain
+code verbosity caused by ``Final'' classes for instantiation. Now we show that
+our \mixin annotation helps to remove these mechanical code in a trivial way. We
+will use the canonical example of implementing arithmetic expressions gradually
+as in \cite{eptrivially} to illustrate the usage.
 
@@
 We now present a variant of a classic game example, which is often used to
-evaluate reuse ability related to inheritance and design patterns. The code
+evaluate code reuse ability related to inheritance and design patterns. The code
 using our approach and the solution by Bono et. al~\cite{bono14} in online, for
@@
 behaviour), multiple inheritance (a special kind of door may require features
-from two other door types). It also shows how to model symmetric sum, override
-and alias as trait-oriented programming. Moreover, the code is provided in
-~\cite{bono14} so that we can have a clear and fair comparison with our
-approach, as shown in Table~\ref{table:codereduction}.
+from two other door types) and it also shows how to model operations
+\texttt{symmetric sum}, \texttt{override} and \texttt{alias} as trait-oriented
+programming. Moreover, the code is provided in ~\cite{bono14} so that we can
+have a clear and fair comparison with our approach, as shown in
+Table~\ref{table:codereduction}. We reduced the interfaces/classes used by
+21.4\% (from 14 to 11) due to the replacement of instantiation classes with
+generated \texttt{of} methods. We reduced source lines of code (SLOC) by
+40\%. To ensure a fair comparison, we use the same code style as Bono's.
 

### 2015-11-24 paper/sections/Introduction.tex

@@
 automatically produce code that deals with type-refinement
-adequately. With object interfaces, many Java programs can be built
-without using a single class!
+adequately. %With object interfaces, many Java programs can be built
+%without using a single class!
 
-Object interfaces do not require changes to the
-runtime, and they also do not introduce any new syntax (apart from a
+Object interfaces do not require changes to the Java runtime or compiler, 
+and they also do not introduce any new syntax (apart from a
 small annotation). All three features of object interfaces are
@@
 Noteworthy, is the fact that all applications are implemented 
-without defining a single class.
+without defining a single class!
 
@@
 \item {\bf Object Interfaces:} A simple feature that allows various
-  powerful multiple-inheritanc eprogramming idioms to be expressed
+  powerful multiple-inheritance programming idioms to be expressed
   conveniently in Java.
@@
   implementation of object interfaces, using Java
-  annotations. Moreover the usefulness of object interfaces is
+  annotations and AST rewriting. Moreover the usefulness of object interfaces is
   illustrated through various examples and case studies.
 
-\item{\bf Language Tuning} We identify this concept and we describe
-object interfaces as examples of Language Tuning. We also discuss 
+\item{\bf Language Tuning} We identify the concept of language tuning 
+and we describe object interfaces as an example. We also discuss 
 how other existing approaches, such as the annotations in project 
-Lombok~\cite{lombok}, can be viewed as language tuning.
+Lombok~\cite{lombok}, which can be viewed as language tuning.
 

### 2015-11-24 paper/sections/RelatedWork.tex
related work ready for review (except for S8.4)

@@
 %and the yo-yo problem~\cite{taenzer1989problems}.  
-To conciliate the need for
-expressive power and the need for simplicity, lots of models have been proposed
-in the past few years, including mixins~\cite{bracha90mixin} and
-traits~\cite{scharli03traits}.  They provide novel programming architecture
-models in the OO paradigm.
+To conciliate the need for expressive power and the need for simplicity, lots of
+models have been proposed in the past few years, including C++ virtual
+inheritance, mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
+They provide novel programming architecture models in the OO paradigm. In terms
+of restrictions set on these models, C++ virtual inheritance aims at a more
+general model, mixins added some restrictions on the model, and trait model is
+the most restricted one (excluding states, instantiation, etc).
 
-\begin{enumerate}
-\item Mixins allows to name components that can be applied to various classes as
-  reusable functionality units. However, they suffer from linearisation: the
-  order of mixin application is relevant in often subtitle and undesired
-  ways. constraints. This hinders their usability and their ability of resolving
-  conflicts: the linearisation (total ordering) of mixin inheritance cannot
-  provide a satisfactory resolution in all cases and restricts the flexibility
-  of mixin composition.  To fight those limitations, an algebra of mixin
-  operators is introduced~\cite{ancona2002calculus}, but this raised the
-  complexity of the approach, especially when constructors and fields are
-  considered~\cite{marco09FJigsaw}.  Our approach does not have the
-  linearisation problem, in that the semantics of Java \textbf{extends} clause
-  is unordered and symmetric.
-\item Simplifying the mixin algebras approach, traits draw a strong line between
-  units of reuse (traits) and object factories (classes) In this model,
-  traits~\cite{scharli03traits} units of reusable code, containing only methods
-  as reusable functionalities. Thus, no state/state initialization is
-  considered.
+\paragraph{C++ Approach.}
+C++ tries to provide a general solution to multiple inheritance. Virtual
+inheritance in C++ provides another solution to multiple inheritance (especially
+the diamond problem by keeping only one copy of the base
+class)~\cite{ellis1990annotated}, however suffers from object initialization
+problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
+constructor calls to virtual superclasses, which would potentially cause serious
+semantic error.
 
-  Classes act as object factories, requiring functionalities from multiple
-  traits.  Traits offers a trait algebra with operations like sum, alias and
-  exclusion, provided for explicit conflict resolution.
-
-  Concluding, (pure) traits do not allow state and they do not offers any reuse
-  instrument to ensure that state is coherently initialized when finally defined
-  in classes.  Traits can't be instantiated. This model requires two concepts
-  (traits and classes) to coexist and cooperate.
-
-  Some authors see this as good language design fostering good software
-  development by helping programmers to think about the structure of their
-  programs.  However, other authors see the need of two concepts and the absence
-  of state as drawbacks of this model. Our approach takes interfaces as units of
-  reuse, and meanwhile generates factory method for instantiation.
-
-\item C++ and Scala also try to provide solutions to multiple inheritance, but
-  both suffer from object initialization problems. Virtual inheritance in C++
-  provides another solution to multiple inheritance (especially the diamond
-  problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
-  however suffers from object initialization problem as pointed out by Malayeri
-  et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
-  superclasses, which would potentially cause serious semantic error. Scala
-  solution (very similar to linearised mixins, but misleadingly called traits in
-  the language) avoids this problem by disallowing constructor parameters,
-  causing no ambiguity in cases such as diamond problem.  This approach has
-  limited expressiveness, and suffers from all the problems of linearised mixin
-  composition.
+\paragraph{Mixins.}
+Mixins are a more restricted model than C++ approach. It allows to name
+components that can be applied to various classes as reusable functionality
+units. However, they suffer from linearisation: the order of mixin application
+is relevant in often subtitle and undesired ways. This hinders their usability
+and the ability of resolving conflicts: the linearisation (total ordering) of
+mixin inheritance cannot provide a satisfactory resolution in all cases and
+restricts the flexibility of mixin composition. To fight those limitations, an
+algebra of mixin operators is introduced~\cite{ancona2002calculus}, but this
+raised the complexity of the approach, especially when constructors and fields
+are considered~\cite{marco09FJigsaw}. Scala solution adopts linearised mixins,
+(but misleadingly called traits in the language) avoids object initialization
+problem by disallowing constructor parameters, causing no ambiguity in cases
+such as diamond problem. However this approach has limited expressiveness, and
+suffers from all the problems of linearised mixin composition.
 Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
@@
 \Q@ class A(B,C,Aux): pass@
-  Our approach not only does not involve linearised mixin problem, but also
-  support proper constructor mechanism.
-\end{enumerate}
+
+\noindent Our approach does not involve linearised mixin problem, in that the semantics of
+Java \textbf{extends} clause is unordered and symmetric. Also,our approach
+supports proper constructor mechanism.
+
+\paragraph{Traits.}
+Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
+line between units of reuse (traits) and object factories (classes). In this
+model, traits as units of reusable code, contain only methods as reusable
+functionalities. Thus, no state/state initialization is considered. Classes act
+as object factories, requiring functionalities from multiple traits.  Traits
+offer a trait algebra with operations like sum, alias and exclusion, provided
+for explicit conflict resolution.  Concluding, (pure) traits do not allow state
+and they do not offers any reuse instrument to ensure that state is coherently
+initialized when finally defined in classes.  Traits can not be
+instantiated. This model requires two concepts (traits and classes) to coexist
+and cooperate.  Some authors see this as good language design fostering good
+software development by helping programmers to think about the structure of
+their programs.  However, other authors see the need of two concepts and the
+absence of state as drawbacks of this model. Our approach takes interfaces as
+units of reuse, and meanwhile generates factory method for instantiation.
+
+% \begin{enumerate}
+% \item Mixins allows to name components that can be applied to various classes as
+%   reusable functionality units. However, they suffer from linearisation: the
+%   order of mixin application is relevant in often subtitle and undesired
+%   ways. constraints. This hinders their usability and their ability of resolving
+%   conflicts: the linearisation (total ordering) of mixin inheritance cannot
+%   provide a satisfactory resolution in all cases and restricts the flexibility
+%   of mixin composition.  To fight those limitations, an algebra of mixin
+%   operators is introduced~\cite{ancona2002calculus}, but this raised the
+%   complexity of the approach, especially when constructors and fields are
+%   considered~\cite{marco09FJigsaw}.  Our approach does not have the
+%   linearisation problem, in that the semantics of Java \textbf{extends} clause
+%   is unordered and symmetric.
+% \item Simplifying the mixin algebras approach, traits draw a strong line between
+%   units of reuse (traits) and object factories (classes) In this model,
+%   traits~\cite{scharli03traits} units of reusable code, containing only methods
+%   as reusable functionalities. Thus, no state/state initialization is
+%   considered.
+
+%   Classes act as object factories, requiring functionalities from multiple
+%   traits.  Traits offers a trait algebra with operations like sum, alias and
+%   exclusion, provided for explicit conflict resolution.
+
+%   Concluding, (pure) traits do not allow state and they do not offers any reuse
+%   instrument to ensure that state is coherently initialized when finally defined
+%   in classes.  Traits can't be instantiated. This model requires two concepts
+%   (traits and classes) to coexist and cooperate.
+
+%   Some authors see this as good language design fostering good software
+%   development by helping programmers to think about the structure of their
+%   programs.  However, other authors see the need of two concepts and the absence
+%   of state as drawbacks of this model. Our approach takes interfaces as units of
+%   reuse, and meanwhile generates factory method for instantiation.
+
+% \item C++ and Scala also try to provide solutions to multiple inheritance, but
+%   both suffer from object initialization problems. Virtual inheritance in C++
+%   provides another solution to multiple inheritance (especially the diamond
+%   problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
+%   however suffers from object initialization problem as pointed out by Malayeri
+%   et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
+%   superclasses, which would potentially cause serious semantic error. Scala
+%   solution (very similar to linearised mixins, but misleadingly called traits in
+%   the language) avoids this problem by disallowing constructor parameters,
+%   causing no ambiguity in cases such as diamond problem.  This approach has
+%   limited expressiveness, and suffers from all the problems of linearised mixin
+%   composition.
+%   Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
+%   \Q@ class A use B,C {...new methods...}@\\*
+%   can be expressed in python as \\*
+%   \Q@ class Aux: ...new methods...@\\*
+%   \Q@ class A(B,C,Aux): pass@
+%   Our approach not only does not involve linearised mixin problem, but also
+%   support proper constructor mechanism.
+% \end{enumerate}
 
@@
 
-In certain situations, our approach allows automatic refinement for return types.
-This is part of a bigger topic in class based languages: expressing and preserving type recursion and (nominal/structural) subtyping at the same time.
+In certain situations, our approach allows automatic refinement for return
+types. This is part of a bigger topic in class based languages: expressing and
+preserving type recursion and (nominal/structural) subtyping at the same time.
 
@@
 article~\cite{Saito2013933} lights up this topic, since they propose two
-features: exact statements \yanlin{do we support exact statements?}\marco{no we do not. When is that it may seams we do?} and
-non-inheritable methods; both related to our work: any method generated inside
-of the \Q@of@ method is indeed non-inheritable, since there is no class name to
-extends from, and exact statements (a form of wild-card capture on the exact
-run-time type) could capture the ``exact type'' of an object even in a
-class-less environment.
+features: exact statements \yanlin{do we support exact statements?}\marco{no we
+  do not. When is that it may seams we do?} and non-inheritable methods; both
+related to our work: any method generated inside of the \Q@of@ method is indeed
+non-inheritable, since there is no class name to extends from, and exact
+statements (a form of wild-card capture on the exact run-time type) could
+capture the ``exact type'' of an object even in a class-less environment.
 
@@
 \end{comment}
-\subsection{Meta-programming competes with Language extensions}
-The most obvious solution to add features to a language is language extension.
-They are often implemented as syntactic extensions that can be desugared to the base
-language.
-For example, the Scala compiler was extended to directly supports XML syntax.
-However, this approach does not support combining multiple extensions into one.
-We are de facto creating a fork in the language, and rarely the new fork gain enough traction to become the main language release.
-On this topic we mention SugarJ~\cite{erdweg11sugarj}; a Java-based extensible language allowing programmers to extend it with custom features by definitions in meta-DSLs (SDF, Stratego, etc).
+\subsection{Meta-programming Competes with Language extensions}
+The most obvious solution to features addition to a language is language
+extension.  It is often implemented as syntactic extensions that can be
+desugared to the base language. For example, the Scala compiler was extended to
+directly supports XML syntax. However, this approach does not support combining
+multiple extensions into one. We are de facto creating a fork in the language,
+and rarely the new fork gain enough traction to become the main language
+release.  On this topic we mention SugarJ~\cite{erdweg11sugarj} - a Java-based
+extensible language allowing programmers to extend it with custom features by
+definitions in meta-DSLs (SDF, Stratego, etc).
 
-On the other side, when the starting language have a flexible enough syntax and a fast and powerful enough reflection, we may just need to play with operator overloading and other language tricks to discover that the language feature we need can be expressed as a simple library in our language.
-For example, consider SQL alchemy in python.
+On the other side, when the starting language have a flexible enough syntax and
+a fast and powerful enough reflection mechanism, we may just need to play with
+operator overloading and other language tricks to discover that the language
+feature we need can be expressed as a simple library in our language. For
+example, consider SQLAlchemy in python.
 
-Java-like language tends to sits in the middle of this two extremes:
-libraries can not influence the type system, so many solutions valid in python or other languages could not be applicable, or may be applicable at the cost of loosing safety.
+Java-like languages tend to sit in the middle of two extremes: libraries
+can not influence the type system, so many solutions valid in python or other
+languages could not be applicable, or may be applicable at the cost of loosing
+safety.
 
-Here (compile/load time) code generation come at the rescue: 
-if for a certain feature (\mixin in our case) it is possible to use the original language syntax to
-\emph{express-describe} any specific instantiation of such feature
-(annotating a class and provide getters), then we can insert in the compilation process a tool that exam and enrich the code before compilation.
-No need to modify the original source; for example we can work on temporary files.
-Java is a particular good candidate for this kind of manipulation since it already provide ways to define and integrate such tools in its own compilation process: in this way there is no need of temporary files, and there is a well defined way of putting multiple extensions together.
+Here (compile/load time) code generation comes at the rescue: if for a certain
+feature (\mixin in our case) it is possible to use the original language syntax
+to \emph{express-describe} any specific instantiation of such feature
+(annotating a class and provide getters), then we can insert in the compilation
+process a tool that exam and enrich the code before compilation. No need to
+modify the original source; for example we can work on temporary files
+\yanlin{Here what does temporary files mean? could explain more?}. Java is
+a particular good candidate for this kind of manipulation since it already
+provide ways to define and integrate such tools in its own compilation process:
+in this way there is no need of temporary files, and there is a well defined way
+of putting multiple extensions together.
 
-Other languages offers even stronger support to safe code manipulation:
-Template Haskell~\cite{}, F\# (type providers)~\cite{} and MetaFjig (Active Libraries)~\cite{}
-all allows to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
+Other languages offer even stronger support to safe code manipulation:
+Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
+all allow to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
 In particular, MetaFjig offers a property called \emph{meta-level-soundness}.
@@
 \subsection{Formalization of Java8}
-We provide a simple and well designed formalization for a subset of Java including default/static interface methods and object initialization literals (often called anonymous local inner classes).
-A similar formalization was drafted by\marco{ I use the term draft because I seams to remember it was just a technical report, I'm right?}
-Goetz and Field~\cite{goetz12fdefenders} to formalize defender (default) methods
-in Java. However this formalization is limited to model exactly one
-method inside classes/interfaces.
+We provide a simple and well designed formalization for a subset of Java
+including default/static interface methods and object initialization literals
+(often called anonymous local inner classes).  A similar formalization was
+drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize defender
+(default) methods in Java. However this formalization is limited to model
+exactly one method inside classes/interfaces.
 
-As a evidence of the attention and care present in our formalization work, while double-checking the behaviour of Java in side cases we have discovered a likely bug in the current \texttt{javac}.\marco{refer to before when we explain the issue.}
+% As an evidence of the attention and care present in our formalization work,
+% while double-checking the behaviour of Java in side cases we have discovered a
+% likely bug in the current \texttt{javac}.\marco{refer to before when we explain
+%   the issue.}
+\yanlin{bug claim removed.}
 
-\section{Future work}
+\section{Future work}\label{sec:futurework}
 

### 2015-11-25 paper/sections/CaseStudy.tex
Future Work ready for review.

@@
 \subsection{A Small Game Example}
-We now present a variant of a classic game example, which is often used to
+We now present a variant of a classic game example, which is used to
 evaluate code reuse ability related to inheritance and design patterns. The code

### 2015-11-25 paper/sections/FutureWork.tex
Future Work ready for review.

\section{Future work}\label{sec:futurework}

\subsection{Keywords in Classes} %{Private state}
\yanlin{briefly mentioned ``synchronized'' keyword, didn't put the workaround
  here, it's in the code.}
The biggest limitation of our approach is the inability to model all features
for classes (private, protected, synchronized, etc.). For example, the absence
of the support for private/protected methods in Java8 interfaces forces all
members of interfaces to be public, including static methods. Since we use
abstract methods to encode the state, our state is always all public; however it
is impossible for the user to know if a certain method maps directly to a field
or if it have a default implementation.  If the use wants a constructor that
does not directly maps to the fields, (as for secondary constructors in Scala)
he can simply define its own \Q@of@ method and delegate on the generated one, as
in
\begin{lstlisting}
@Mixin interface Point{
  int x(); int y();
  static Point of(int val){return Point.of(val,val);}  
}
\end{lstlisting}
However, the generated \Q@of@ method would also be present and public.  If a
future version of Java was to support \emph{static private methods in
  interfaces} we could extend our code generation to handle also encapsulation.
Currently, it is possible to use a public nested class with private static
methods inside, but this is ugly and cumbersome. We are considering if our
annotation processor can take code with \Q!@Private! annotation and turn it into
static private methods of a nested class. In this extension, also the \Q@of@
method could be made private following the same pattern.

\subsection{State initialization}
As discuss before, the user can trivially define its own \Q@of@ method, and
initialize a portion of the state with default values.  However, the
initialization code would not be reusable, and subinterfaces would have
to repeat such initialization code.  If a field has no setters, a simple
alternative is to just define the ``field'' as a default method as in
\begin{lstlisting}
@Mixin interface Box{ default int val(){return 0;} }
\end{lstlisting}
If setters are required, a possible extension of our code expansion could
recognize a field if the getter is provided and the setter is required, and
could generate the following code: \yanlin{both of the two generated val()
  method should be public to pass compilation. Did you omit the public keyword
  on purpose to make the code clean? }
\begin{lstlisting}
interface Box{ 
  default int val(){return 0;} //provided
  void val(int _val);//provided
  static Box of(){return new Box(){//generated
    int val=Box.super.val();
    int val(){return val;}
    void val(int _val){val=_val;}
    };}}
\end{lstlisting}
We are unsure of the value of this solution: is very tricky, the user define a
method that (contrary to our usual expectation) is actually overridden in a way
that the behaviour changes, but change only after the first setter is called,
plus this code would cache the result instead of re-computing it every
time. This can be very relevant and tricky in a non functional setting.

\subsection{Class Invariants in ClassLess Java}
Since objects are created by automatically generated methods, another limitation
of our current approach is that there is no place where the user can dynamically
check for class invariants. In Java often we see code like
\begin{lstlisting}
class Point{ int x; int y;
  Point(int x; int y){this.x=x;this.y=y; assert this.checkInvariant();}
  private boolean checkInvariant(){... x>0,y>0...}
}
\end{lstlisting}

We are considering an extension of our annotation where 
default methods with the special name \Q@checkInvariant()@ will be called inside the \Q@of@ methods.
If multiple interfaces are implemented, and more then one offers
\Q@checkInvariant()@,  a composed implementation could be automatically generated, composing by \Q@&&@ the various competing implementations.

\subsection{Clone, toString, equals and hashCode}
Methods originally defined in Java class \Q@Object@, as \Q@clone@ and \Q@toString@, can be supported by our approach, but they need special care. If an interface annotated with \mixin asks an implementation for \Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters. Note how this would solve most of the Java ugliness related to \Q@clone@ methods.}

However, if the user wish to provide his own implementation, since the method is
also implemented in \Q@Object@, a conflict would arise. Then the user has to
explicitly resolve the conflict inside \Q@of@, by implementing the method and
delegating it to the user implementation, thus
\begin{lstlisting}
@Mixin interface Point{ int x(); int y();
  default Point clone(){ return Point.of(0,0);}//user defined clone
}
\end{lstlisting} 
Would expand into

\begin{lstlisting}
interface Point{ int x(); int y();
  default Point clone(){ return Point.of(0,0);}//user defined clone
  public static Point of(int _x,int_y){
    return new Point(){...
      public Point clone(){ return Point.super.clone();}
      };  }  }
\end{lstlisting}

### 2015-11-25 paper/sections/RelatedWork.tex
Future Work ready for review.

@@
 \yanlin{bug claim removed.}
-
-\section{Future work}\label{sec:futurework}
-
-\subsection{Private state}
-The biggest limitation of our approach is the absence in Java8 of support for private/protected methods in interfaces.That is, in Java8 all members of interfaces are public, including static methods.
-Since we use abstract methods to encode the state, our state is always all public; however is impossible for the user to know if a certain method maps directly to a field or if it have a default implementation.
-If the use wants a constructor that does not directly maps to the fields, (as for secondary constructors in Scala) he can simply define its own \Q@of@ method and delegate on the generated one, as in
-\begin{lstlisting}
-@Mixin interface Point{
-  int x(); int y();
-  static Point of(int val){return Point.of(val,val);}  
-  }
-\end{lstlisting}
-However, the generated \Q@of@ method would also be present and public.
-If a future version of Java was to support \emph{static private methods in interfaces} we could extend our code generation to handle also encapsulation.
-Currently, is possible to use a public nested class with private static methods inside, but this is ugly and cumbersome.
-We are considering if our annotation processor can take code with \Q!@Private!
-annotation and turn it into static private methods of a nested class.
-In this extension,  also the of method could be made private following the same pattern.
-
-\subsection{State initialization}
-
-As discuss before, the user can trivially define its own \Q@of@ method, and initialize a portion of the state with default values.
-However, the initialization code would not be reused/reusable, and subinterfaces would have to repeat such initialization code.
-If a field has no setters, a simple alternative is to just define the ``field'' as a default method as in 
-\begin{lstlisting}
-@Mixin interface Box{ default int val(){return 0;} }
-\end{lstlisting}
-if setters are required, a possible extension of our code expansion could recognize a field if the getter is provided and the setter is required, and could generate the following code:
-\begin{lstlisting}
-interface Box{ 
-  default int val(){return 0;} //provided
-  void val(int _val);//provided
-  static Box of(){return new Box(){//generated
-    int val=Box.super.val();
-    int val(){return val;}
-    void val(int _val){val=_val;}
-    };}}
-\end{lstlisting}
-We are unsure of the value of this solution: is very tricky, the user define a method that (contrary to our usual expectation) is actually overridden in a way that the behaviour change, but change only after the first setter is called, plus this code would cache the result instead of re-computing it every time.
-This can be very relevant and tricky in a non functional setting.
-
-\subsection{Class invariants in ClassLess Java}
-Since the objects are created by automatically generated methods,
-another limitation of our current approach is that there is no place where the user can dynamically check for class invariants.
-In Java often we see code like
-\begin{lstlisting}
-class Point{ int x; int y;
-  Point(int x; int y){this.x=x;this.y=y; assert this.checkInvariant();}
-  private boolean checkInvariant(){... x>0,y>0...}
-\end{lstlisting}
-
-We are considering an extension of our annotation where 
-default methods with the special name \Q@checkInvariant()@ will be called inside of the \Q@of@ methods.
-if multiple interfaces are implemented, and more then one offers
-\Q@checkInvariant()@,  a composed implementation could be automatically generated, composing by \Q@&&@ the various competing implementations.
-
-\subsection{Clone, toString, equals and hashCode}
-Methods originally defined in \Q@Object@, as \Q@clone@ and \Q@toString@ can be supported by our approach, but they need special care.
-If an interface annotated with \mixin ask an implementation for \Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters.
-Note how this would solve most of the Java ugliness related to \Q@clone@ methods.}
-
-However, if the user wish to provide its own implementation, since the method is also implemented in \Q@Object@ we would have a conflict, that we have to explicitly resolve inside of \Q@of@, by implementing the method and delegating it to the user implementation, thus
-
-\begin{lstlisting}
-@Mixin interface Point{ int x(); int y();
-  default Point clone(){ return Point.of(0,0);}//user defined clone
-  }
-\end{lstlisting} 
-Would expand into
-
-\begin{lstlisting}
-interface Point{ int x(); int y();
-  default Point clone(){ return Point.of(0,0);}//user defined clone
-  public static Point of(int _x,int_y){
-    return new Point(){...
-      public Point clone(){ return Point.super.clone();}
-      };  }  }
-\end{lstlisting}

### 2015-11-25 UseMixinLombok/src/test/TestSynchronize.java

@@ -41,3 +41,14 @@ interface B {
 public interface TestSynchronize {
 
 }
+
+
+interface Box{ 
+    default int val(){return 0;} //provided
+    void val(int _val);//provided
+    static Box of(){return new Box(){//generated
+        int val=Box.super.val();
+        public int val(){return val;}
+        public void val(int _val){val=_val;}
+    };}
+}

### 2015-11-25 paper/sections/Implementation.tex
Revised Implementation

--- 2015-11-19 paper/sections/Implementation.tex
@@
-\section{Implementation}
+\section{Implementation with Lombok}
 
-One of the downside of Java is the code verbosity, e.g., getter and setter
-methods for fields, unneeded type annotation, etc. Lombok project~\cite{} is a
-Java tool that aims at removing (or reducing) these boilerplate code as far as
-possible, via annotations. There are a number of annotations provided by the
+Our implementation is based on extension to Lombok. The Lombok
+project~\cite{lombok} is a Java tool that aims at removing (or
+reducing) Java boilerplate code as far as possible, via
+annotations. There are a number of annotations provided by the
 original Lombok, including \textbf{@Getter}, \textbf{@Setter},
-\textbf{@ToString} for generating getters, setters and \QM{toString} methods,
-respectively.
+\textbf{@ToString} for generating getters, setters and \QM{toString}
+methods, respectively.  Furthermore, Lombok provides a number of
+interfaces for users to create custom transformations, as extensions
+to the original framework.
+A transformation is based on a handler, which acts on the AST from
+parsing the annotated node and returns a modified AST for analysis and
+generation afterwards. Such a handler can either be a Javac handler or
+an Eclipse handler.
 
-Furthermore, Lombok provides a number of interfaces for users to create custom
-transformations, as extensions to the original framework. By realizing these
-transformations, with new annotations supported, users can generate new field
-and method members, and inject them to current class types automatically at
-compile time. More specifically, a transformation is based on a handler, which
-acts on the AST from parsing the annotated node and returns a modified AST for
-analysis and generation afterwards. Such a handler can either be a Javac handler
-or an Eclipse handler.
-
-Our implementation is based on extension to Lombok. In Eclipse, with an
-interface annotated by \mixin, the automatic annotation processing is performed
-and the information of the interface from compilation is captured in the
-``Outline'' window, including all the methods inside the interface as well as
-the generated ones.
- The custom transformation is easy and convenient to use.
- For example this means that the IDE functionality 
- content assist autocomplete will work for the newly generated methods.
-
- The biggest reasons to use Lombok rather than using a conventional Java
+The annotation we created is \mixin. In Eclipse, with an interface annotated by
+\mixin, the automatic annotation processing is performed transparently and the information of
+the interface from compilation is captured in the ``Outline'' window. This includes
+all the methods inside the interface as well as the generated ones.  The custom
+transformation is easy and convenient to use.  For example this means that the
+IDE functionality for content assist and autocomplete will work for the newly generated
+methods. The biggest reasons to use Lombok rather than using a conventional Java
  annotation processor are:
@@
 %  compile time to modify bytecode.
-\item Lombok modifies the generation process of the class files. Neither the
-  source code is modified nor new Java files are generated.
+\item Lombok modifies the generation process of the class files, by directly modifying the AST. 
+Neither the source code is modified nor new Java files are generated.
 \item Moreover, and probably more importantly, Lombok is capable of generating
-  code \emph{inside} a class/interface.This is the ability that conventional
-  Java annotation processor do not provide.
+  code \emph{inside} a class/interface. This is the ability that conventional
+  Java annotation processors do not provide.
 \end{itemize}
 
-\paragraph{Limitation}
-Our prototype implementation using Lombok has certain limitation: 
+\paragraph{Limitations}
+Our prototype implementation using Lombok has certain limitations: 
 \begin{itemize}
@@
   a single interface would require re-compiling the whole file. This compilation
-  limitation is not caused by our algorithm. It is Lombok implementation related
+  limitation is not caused by our algorithm. It is a Lombok implementation related
   issue: in Lombok it is hard to capture a type declaration from its reference,
   even harder when the type declaration is in other files (we have not found a
-  way to do this currently).
+  way to do this yet).
 \item At this stage our implementation only realizes the Eclipse handler and our
@@
   \texttt{javac} is missing.
+\item The current implementation does not take type-parameters into
+  consideration, thus it does not support generics yet.
 \end{itemize}
@@
 \paragraph{Comparison with other Lombok annotations}
-Lombok project provides a set of predefined annotations, including constructor
+The Lombok project provides a set of predefined annotations, including constructor
 generators similar as ours (e.g., \textbf{@NoArgsConstructor},
@@
 provides other convenient features (including generating fluent setters, type
-refinement,etc) which the base Lombok project could not provide.
+refinement,etc), which the base Lombok project does not provide. 
+Finally, while \mixin is formalized, none of Lombok's annotations have been 
+studied in a formal way.

### 2015-11-25 paper/sections/CaseStudy.tex
Game example

--- 2015-11-22 paper/sections/CaseStudy.tex
@@
 \section{Case Studies}
-In this section we conduct three case studies which reveals various advantages
-using \mixin annotation. 1st case study provides a easy and straightforward way
-to solve the Expression Problem. 2nd case study models a simple game and the
-comparison between our approach and others shows that code amount is reduced
-tremendously using \mixin. 3rd case study models an embedded DSL for SQL
-languages with fluent interfaces and shows that \mixin not only removes verbol
-code for defining fluent interfaces but also does implicit refinement along
-extension.
+In this section we conduct three case studies which reveals various
+advantages using \mixin annotation. The first case study provides a
+simple way to solve the Expression Problem, while supporting multiple,
+independent extensions in Java.  The second case study show how to model an
+embedded DSL for SQL languages with fluent interfaces. Finally, the third 
+case study models a simple game, and compares our implementation with an 
+existing one, showing that the amount of code is reduced significantly using \mixin.
 
-\subsection{A Trivial Solution to the Expression Problem}
+\subsection{A Trivial Solution to the Expression Problem with Object Interfaces}
 
-The \emph{Expression Problem} (EP)~\cite{wadler98expression-problem} is coined
-by Wadler about modular extensibility issues in software evolution and has been
-a hot topic in programming languages since. The requirements of a solution to
-the EP can be summarized as: 
-\begin{itemize}
-\item \emph{Extensibility in both dimensions}: A solution must allow
-  the addition of new data variants and new operations and support
-  extending existing operations.
-\item \emph{Strong static type safety}: A solution must prevent
-  applying an operation to a data variant which it cannot handle using
-  static checks.
-\item \emph{No modification or duplication}: Existing code must not be
-  modified nor duplicated.
-\item \emph{Separate compilation and type-checking}: Safety checks or
-  compilation steps must not be deferred until link or runtime.
-\item \emph{Independent extensibility}~\cite{zenger05independentlyextensible}:
-  It should be possible to combine independently developed extensions so that
-  they can be used jointly.
-\end{itemize}
-
-Today we know of various solutions to the EP that either rely on \emph{new
-  programming language features}
-% ~\cite{chambers95multimethods, clifton00multiJava, madsen89virtual, nystrom06j,
-%   bruce98astatically, McDirmid01jiazzi, garrigue98polymorphic,
-%   zenger01extensible, loh06open, wehr11javaGI}, 
-or can be used as \emph{design patterns}%~\cite{gof}
-in existing languages.
-% ~\cite{torgersen04theexpression, oliveira06extensible,
-%  wouter08datatypes, oliveira09modular, oliveira12extensibility}. 
-Now we show that the EP can be solved with our \mixin annotation in a trivial
-way. We will use the canonical example of implementing arithmetic expressions
-gradually to illustrate the usage.
+The \emph{Expression Problem} (EP)~\cite{wadler98expression-problem}
+is a well-known problem about modular extensibility issues in software
+evolution. Recently, a new solution~\cite{eptrivially} that uses only
+covariant type refinement that has been shown to work in Java. When
+that solution is modelled with interfaces and default methods, it can
+even provide independent extensibility: that is, the ability to
+assemble a system, from multiple, independently developed extensions.
+Unfortunatelly, a plain Java solution has alot of code verbosity in
+Java due to code required for instantiation, which makes the solution 
+cumbersome to use. The \mixin annotation helps to remove such code, and 
+makes the solution quite easy to use.
+\bruno{We should try to do some space saving in here. You could, for example 
+combine the code for the initial system with the code for a new operation in a figure/table 
+with 2 colums. Then just refer to the Table/Figure in the text.}
 
 \paragraph{Initial System}
-The initial system expresses arithmetic expressions with only literals and
-addition structure:
+In the formulation of the EP there is an initial system that models
+arithmetic expressions with only literals and addition, and an initial 
+operation:
 
@@
 
-\texttt{Exp} is the common super-interface with an evaluation operation
+\noindent \texttt{Exp} is the common super-interface with an evaluation operation
 \texttt{eval()} inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend
-interface \texttt{Exp} with default implementations for \texttt{eval}
-operation. The number field (\texttt{x}) of a literal is represented as a getter
-method \texttt{x()} and expression fields (\texttt{e1,e2}) of an addition as
+interface \texttt{Exp} with default implementations for the \texttt{eval}
+operation. The number field \texttt{x} of a literal is represented as a getter
+method \texttt{x()} and expression fields (\texttt{e1} and \texttt{e2}) of an addition as
 getter methods \texttt{e1()} and \texttt{e2()}.
 
-\paragraph{Adding a New Variant}
-It is easy to add new data variants to the code in the initial system in a
-modular way. For example, the following code shows how to add the subtraction
-variant.
+\paragraph{Adding a New Type of Expressions}
+It is easy to add new types of expressions to the code in the initial system in a
+modular way. For example, the following code shows how to add subtraction.
 
@@
 
-\paragraph{Adding a New Operation}
-Adding new operations to the previous system is still straightforward (although
-requires longer code). The following code shows an example of adding a new
-operation \texttt{print}.
+\paragraph{Adding a New Operation} The difficulty of the EP in object-oriented 
+languages arizes from adding new operations. For example, adding a pretty printing 
+operation, would typically involve changing all of the existing code. However, a solution 
+to the EP forbids this and it also forbids using casts. In other words, it should be possible 
+to add the pretty printing operation in a type-safe and modular way.  
+It turns out that, this can be done easily, using our object interfaces approach. 
+The following code shows how to add the new operation \texttt{print}.
 
 \lstinputlisting[linerange=57-66]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
+\bruno{please indent code properly (the getters are not aligned with default))}
 
-The basic idea is to define interfaces for extending old interfaces.  The
-interface \texttt{ExpP} extending interface \texttt{Exp} is defined with the
-extra function \texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are
-defined with default implementations of operation \texttt{print()}, extending
+\noindent  The interface \texttt{ExpP} extending interface \texttt{Exp} is defined with the
+extra method \texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are
+defined with default implementations of \texttt{print()}, extending
 base interfaces \texttt{Lit} and \texttt{Add}, respectively. Importantly, note
-that in \texttt{AddP}, the types of ``\emph{fields}'' \texttt{e1,e2} are refined
-via the \emph{return types refinement} of getter methods \texttt{e1()} and
-\texttt{e2()}.
+that in \texttt{AddP}, the types of the ``\emph{fields}'' (i.e. the getter methods) \texttt{e1} 
+and \texttt{e2} are refined. If the types were not refined then the body of the \texttt{print}
+method in \texttt{AddP} would fail to type-check.
 
 \paragraph{Independent Extensibility}
-To show that our approach supports independent extensibility, we first define a
-new operation \texttt{collectLit}(which collects all
-literal components in an expression) on expressions:
+To show that our approach supports independent extensibility~\cite{odersky}, we first define a
+new operation \texttt{collectLit} (which collects all
+literal components in an expression) on expressions. For space reasons, 
+we omit the definitions of the methods: 
+\bruno{No need to show full code in the paper. You can just put \ldots in the body of the collectLit methods.}
 
@@
 
-\subsection{A Small Game Example}
-We now present a variant of a classic game example, which is often used to
-evaluate reuse ability related to inheritance and design patterns. The code
-using our approach and the solution by Bono et. al~\cite{bono14} in online, for
-space constraints, we will omit it here. In the game, there is a player with the
-goal of collecting as many coins as possible. She may enter a room with several
-doors to be chosen among. This is a good example because it involves code reuse
-(different kinds of doors inherit a common type, with different features and
-behaviour), multiple inheritance (a special kind of door may require features
-from two other door types). It also shows how to model symmetric sum, override
-and alias as trait-oriented programming. Moreover, the code is provided in
-~\cite{bono14} so that we can have a clear and fair comparison with our
-approach, as shown in Table~\ref{table:codereduction}.
+Note that the code for instantiation is automatically generated by \mixin. 
+So, for example creating a simple expression of type \texttt{ExpPC} is 
+as simple as:
 
-\begin{table}[h]
-\centering
-\begin{tabular}{ccc}
-\hline
-            & SLOC   & \# of classes/interfaces \\ \hline
-Bono et al. & 335    & 14                       \\
-Ours        & 199    & 11                       \\
-\rowcolor[HTML]{C0C0C0} 
-Reduced by  & 40.6\% & 21.4\%                   \\ \hline
-\end{tabular}
-\caption{Code Reduction and Reuse.}
-\label{table:codereduction}
-\end{table}
+\bruno{Add code here}
+
+\noindent In contrast, in a pure Java solution, the tedious instantiation code would need 
+to be defined manually.
 
@@
 JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
-interfaces, programmers have to either handwritten everything or create a
-wrapper around the original non-fluent interfaces, using \textbf{this}. In Java,
-several libraries help create fluent APIs easily (by code generation), including
-jOOQ, op4j, fluflu, JaQue, etc. These libraries have given us hints on the
-implementation of fluent setters in ClassLess Java.
+interfaces, programmers have to either handwrite everything or create a
+wrapper around the original non-fluent interfaces using \textbf{this}. In Java,
+several libraries help create fluent APIs easily (by code generation)\bruno{do they use code 
+generation to produce the fluent interfaces?}, including
+jOOQ, op4j, fluflu, JaQue, etc.
 
-When creating fluent interfaces, our approach has two main advantages:
+The \mixin annotation can also be used to create fluent interfaces. 
+When creating fluent interfaces with \mixin, there are two main advantages:
 \begin{enumerate}
@@
 \noindent We use embedded DSLs of two simple SQL query languages to illustrate.
-The first query language \texttt{Database} (conceptually) models
-\texttt{select}, \texttt{from} and \texttt{where} clause:
+The first query language \texttt{Database}  models
+\texttt{select}, \texttt{from} and \texttt{where} clauses:
 \lstinputlisting[linerange=6-13]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
+\bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
+to have to use the of explicitly to initialized all fields.}
 
-\noindent Note that we are only illustrating the idea of modeling embedded DSLs
-with the database example, no concrete implementation of database connection nor
-operations are provided in this case study. One benefit that fluent methods give
-us is method chaining convenience. So when making a query using \texttt{select},
-\texttt{from} and \texttt{where}, instead of writing code like:
-\lstinputlisting[linerange=26-29]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=NORMAL_QUERY
+\noindent The main benefit that fluent methods give
+us is the convinience of method chaining:
 
-\noindent We can write more concise code below:
 \lstinputlisting[linerange=35-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
+\bruno{write query in 1 line}
 
-\noindent Now we extend the first query language with a new feature
+\noindent Note how all the logic for the fluent setters is automatically provided by the \mixin annotation.
+
+\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
 \texttt{orderBy} which orders the result records by a field that users
-specified. The key is to extend interface \texttt{Database} with only the
-feature we need to add and the return type of fluent setters in
+specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
+feature, and the return type of fluent setters in
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
 \lstinputlisting[linerange=17-20]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+This way, when a \texttt{ExtendedDatabase} query created, 
+all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling 
+\texttt{orderBy}.
 
-% \lstinputlisting[linerange=44-48]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\lstinputlisting[linerange=44-48]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\bruno{write query in 1 line}
+
+\subsection{A Small Game Example}
+The last case study is a simple Maze game, which is used\bruno{where?}
+to evaluate code reuse ability related to inheritance and design
+patterns. In the game, there is a player with the goal of collecting
+as many coins as possible. She may enter a room with several doors to
+be chosen among. This is a good example because it involves code reuse
+(different kinds of doors inherit a common type, with different
+features and behaviour), multiple inheritance (a special kind of door
+may require features from two other door types) and it also shows how
+to model operations \texttt{symmetric sum}, \texttt{override} and
+\texttt{alias} as trait-oriented programming. The game has been
+implemented using plain Java 8 and default methods by Bono
+et. al~\cite{bono14}, and the code for that implementation is
+available online. We reimplemented the game using \mixin. Due to space
+constraints, we omit the code here. The following table summarizes
+the number of lines of code and classes/interfaces in each implementation:
+
+\vspace{5pt}
+\hspace{80pt}\begin{tabular}{ccc}
+\hline
+            & SLOC   & \# of classes/interfaces \\ \hline
+Bono et al. & 335    & 14                       \\
+Ours        & 199    & 11                       \\
+\rowcolor[HTML]{C0C0C0} 
+Reduced by  & 40.6\% & 21.4\%                   \\ \hline
+\end{tabular}
+\vspace{5pt}
+
+\noindent The \mixin annotation allowed us to reduce the interfaces/classes used
+in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
+reductions was due to the replacement of instantiation classes with
+generated \texttt{of} methods. The number of source lines of code (SLOC)
+was reduced by 40\%\bruno{explain why? we generate code for getters/setters?}. 
+To ensure a fair comparison, we used the same coding style as Bono's.

### 2015-11-25 paper/sections/CaseStudy.tex
Game example
fix bibtex errors

@@
 
-The \emph{Expression Problem} (EP)~\cite{wadler98expression-problem}
+The \emph{Expression Problem} (EP)~\cite{wadler98expression}
 is a well-known problem about modular extensibility issues in software
@@
 \paragraph{Independent Extensibility}
-To show that our approach supports independent extensibility~\cite{odersky}, we first define a
+To show that our approach supports independent extensibility~\cite{zenger05independentlyextensible}, we first define a
 new operation \texttt{collectLit} (which collects all
@@
 was reduced by 40\%\bruno{explain why? we generate code for getters/setters?}. 
-To ensure a fair comparison, we used the same coding style as Bono's.
+To ensure a fair comparison, we used the same coding style as Bono et al.'s.

### 2015-11-25 paper/sections/RelatedWork.tex
Revised Related work (in progress)

@@
 \subsection{Multiple Inheritance in Object Oriented Languages}
-Many authors have argued in favour or against multiple inheritance.  It provides
+Many authors have argued in favour or against multiple inheritance.  Multiple inheritance provides
 expressive power, but it is difficult to model and implement, and can create
@@
 %and the yo-yo problem~\cite{taenzer1989problems}.  
-To conciliate the need for expressive power and the need for simplicity, lots of
-models have been proposed in the past few years, including C++ virtual
+To conciliate the need for expressive power and the need for simplicity, many
+models have been proposed in over the years, including C++ virtual
 inheritance, mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
@@
 the most restricted one (excluding states, instantiation, etc).
+\bruno{How about Malayeri and Aldrich's paper? shouldn't that be discussed?}
 
@@
 constructor calls to virtual superclasses, which would potentially cause serious
-semantic error.
+semantic errors. \bruno{What happens in our approach for the same case?}
 
 \paragraph{Mixins.}
-Mixins are a more restricted model than C++ approach. It allows to name
+Mixins are a more restricted model than the C++ approach. Mixins allow to name
 components that can be applied to various classes as reusable functionality
 units. However, they suffer from linearisation: the order of mixin application
-is relevant in often subtitle and undesired ways. This hinders their usability
+is relevant in often subtle and undesired ways. This hinders their usability
 and the ability of resolving conflicts: the linearisation (total ordering) of
@@
 raised the complexity of the approach, especially when constructors and fields
-are considered~\cite{marco09FJigsaw}. Scala solution adopts linearised mixins,
-(but misleadingly called traits in the language) avoids object initialization
+are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearised mixins.
+Scala avoids the object initialization
 problem by disallowing constructor parameters, causing no ambiguity in cases
 such as diamond problem. However this approach has limited expressiveness, and
-suffers from all the problems of linearised mixin composition.
+suffers from all the problems of linearised mixin composition. Other languages, such as 
+Python, also use linearized mixins.
+\begin{comment}
 Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
@@
 \Q@ class A(B,C,Aux): pass@ 
+\end{comment}
+\noindent Java interfaces and default methods does not use
+linearisation: the semantics of Java \textbf{extends} clause in
+interfaces is unordered and symmetric.
 
-\noindent Our approach does not involve linearised mixin problem, in that the semantics of
-Java \textbf{extends} clause is unordered and symmetric. Also,our approach
-supports proper constructor mechanism.
+\bruno{rrevise latter?}
+However, in pure Java, there is no mechanism for creating objects in
+interfaces. Also, our approach supports proper constructor mechanism.
 
-\paragraph{Traits.}
-Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
-line between units of reuse (traits) and object factories (classes). In this
-model, traits as units of reusable code, contain only methods as reusable
-functionalities. Thus, no state/state initialization is considered. Classes act
-as object factories, requiring functionalities from multiple traits.  Traits
-offer a trait algebra with operations like sum, alias and exclusion, provided
-for explicit conflict resolution.  Concluding, (pure) traits do not allow state
-and they do not offers any reuse instrument to ensure that state is coherently
-initialized when finally defined in classes.  Traits can not be
-instantiated. This model requires two concepts (traits and classes) to coexist
-and cooperate.  Some authors see this as good language design fostering good
-software development by helping programmers to think about the structure of
-their programs.  However, other authors see the need of two concepts and the
-absence of state as drawbacks of this model. Our approach takes interfaces as
-units of reuse, and meanwhile generates factory method for instantiation.
+\paragraph{Traits and Java's default methods}
+Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
+strong line between units of reuse (traits) and object factories
+(classes). In this model, traits as units of reusable code, contain
+only methods as reusable entities. Thus, no state and state
+initialization are considered. Classes act as object factories,
+requiring functionalities from multiple traits. Java 8 interfaces with
+default methods are closely related to traits: concrete method
+implementation are allowed (via the \textbf{default} keyword) inside
+interfaces. The introduction of default methods opens the gate for
+various flavours of multiple inheritance in Java, using interfaces.
+Traits offer a trait algebra with operations like sum, alias and exclusion, provided
+for explicit conflict resolution.  
+Former work by Bono et al.~\cite{bono14}. provides details on
+mimicking the trait algebra through Java 8 interfaces. 
+We briefly recall the main points of their encoding;
+however we propose a different representation of \textbf{exclusion}.
+The author of ~\cite{bono14} agree that our revised version is
+cleaner, typesafe and more direct.\bruno{do we want do say this?}
+
+\newcommand\shortItem{\vspace{-1ex}}
+\begin{itemize}
+\item \textbf{Symmetric sum} can be obtained by simple multiple inheritance between interfaces.
+\shortItem
+    \begin{lstlisting}
+    interface A { int x(); }    interface B { int y(); }    interface C extends A, B {}
+    \end{lstlisting}
+\shortItem
+\item \textbf{Overriding} a conflict is obtained by specifying which super interface take precedence.
+\shortItem
+    \begin{lstlisting}
+    interface A { default int m() {return 1;} } 
+    interface B { default int m() {return 2;} }
+    interface C extends A, B { default int m() {return B.super.m();} }
+    \end{lstlisting}
+\shortItem
+\item \textbf{Alias} is creating  a new method delegating to the existing super interface.
+\shortItem
+    \begin{lstlisting}
+    interface A { default int m() {return 1;} }
+    interface B extends A { default int k() {return A.super.m();} }
+    \end{lstlisting}
+\shortItem
+
+\item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces.
+\shortItem
+    \begin{lstlisting}
+    interface A { default int m() {return 1;} }
+    interface B extends A { int m(); }
+    \end{lstlisting}
+\shortItem
+\end{itemize}
+
+There are also proposals for extending Java (before Java8) with
+traits. For example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by
+Liquori et al. extends the calculus of Featherweight Java
+(FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
+trait-based inheritance in Java.  Except for few, mostly syntactic
+details, their work can be emulated/rephrased with Java8 interfaces.
+There are also extensions to the original trait model, which have operations 
+that default methods and interfaces cannot model, such as method renaming
+(as in [Reppy2006]), which breaks structural subtyping.
+
+\paragraph{Traits vs Object Interfaces}
+We consider object interfaces to be an alternative to traits or mixins.
+In the trait model two concepts (traits and classes) coexist and
+cooperate. Some authors~\bruno{who?} see this as good language design
+fostering good software development by helping programmers to think
+about the structure of their programs.  However, other authors see the
+need of two concepts and the absence of state as drawbacks of this
+model~\cite{malayeri}.  Object interfaces are units of reuse, and at
+the same time they provide factory methods for instantiation and
+support state.  Our approach promotes the use of interfaces instead of
+classes, in order to rely on the modular composition offered
+by interfaces. Since Java was designed for classes, a direct class-less
+programming style is verbose and feels unnatural. However, annotation
+driven code generation is enough to overcome this difficulty, and the
+resulting programming style encourages modularity, composability and
+reusability, by keeping a strong object oriented feel. In that sense,
+we promote object interfaces as being both units of reusable code and
+object factories. Our practical experience is that, in Java,
+separating the two notions leads to alot of boilerplate code, and is
+quite limiting when multiple inheritance with state is required. The
+use of abstract state operations avoids the key difficulties
+associated with multiple inheritance and state, while still being quite expressive. 
+Moreover the ability to support constructors 
+adds additional expressivity, which is not available in approaches such 
+as Scala's traits/mixins.
+
+%Nevertheless, we
+%believe that in other languages the separation can be more practical
+%than in Java, and we do not necessarally advocate that merging the two
+%notions is a better approach in general.
+
+%There are other limitations of our current approach, but they may be addressed
+%in future work (see Section~\ref{sec:futurework}).
+
+\begin{comment}
+Besides, we support more features than the original trait model:
+\begin{itemize}
+\item We provide \texttt{of} methods for the annotated interfaces.
+During annotation processing time, the ``fields'' inside an interface are observed and a static method \texttt{of} is automatically injected to the interface with its arguments correspondingly.
+Such a method is a replacement to the constructors in original traits, making instantiation more convenient to use.
+That is, in our approach there are only interfaces, our model requires a single concept, while the trait model requires traits \emph{and} classes.
+
+\item We provide \texttt{with-} methods as auxiliary constructors. A \texttt{with-} method is generated for each field, just like record update, returning the new object with that field updated
+%. A \texttt{clone} method is generated for the interface, returning a copy of the current object.
+Furthermore, we do automatic return type refinement for these kind of methods.
+This feature is comparatively useful in big examples, making operations and behaviours more flexible.%, which we will demonstrate later.
+\item We provide two options for generating setters.
+There are two kind of setters which are commonly used, namely \textit{void setters} and \textit{fluent setters}.
+The only difference is that a fluent setter returns the object itself after setting, thus supporting a pipeline of such operations.
+The generation depends on which type of setter is declared in the interface by users.
+\end{itemize}
+\end{comment}
 
@@
 
-\subsection{Multiple Inheritance in Java}
-Since Java 8 default methods are introduced, concrete method implementation are
-allowed to be defined (via the \textbf{default} keyword) inside
-interfaces. 
-% Since Java supports implementation of multiple interfaces (instead
-% of extension of a single class), 
-The introduction of default methods opens the
-gate for various flavours of multiple inheritance in Java, using interfaces.
-Former work by Bono.et. al.~\cite{bono14}. provides details on mimicking traits
-through interfaces.
-
-There are proposals for extending Java (before Java8) with traits. For example,
-FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends the
-calculus of Featherweight Java (FJ)~\cite{Igarashi01FJ} with statically-typed
-traits, adding trait-based inheritance in Java.  Except for few, mostly
-syntactic details, their work can be emulated/rephrased with Java8 interface.
-
-\input{sections/Compare.tex}
+% \input{sections/Compare.tex}
 

### 2015-11-25 paper/sections/CaseStudy.tex
EP case study revised

@@
 makes the solution quite easy to use.
-\bruno{We should try to do some space saving in here. You could, for example 
-combine the code for the initial system with the code for a new operation in a figure/table 
-with 2 colums. Then just refer to the Table/Figure in the text.}
+% \bruno{We should try to do some space saving in here. You could, for example 
+% combine the code for the initial system with the code for a new operation in a figure/table 
+% with 2 colums. Then just refer to the Table/Figure in the text.}
 
@@
 
-\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\begin{lstlisting}[multicols=2]
+interface Exp { int eval(); }
+@Obj interface Lit extends Exp {
+	int x();
+	default int eval() { return x(); }
+}
+@Obj interface Add extends Exp {
+	Exp e1(); Exp e2();
+	default int eval() {
+		return e1().eval() + e2().eval();
+	}
+}
+\end{lstlisting}
+
+% \begin{multicols}{2}
+% \lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+% \end{multicols}
 
@@
 
-\lstinputlisting[linerange=48-53]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=57-66]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
-\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
-\bruno{please indent code properly (the getters are not aligned with default))}
+\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+%\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
+%\bruno{please indent code properly (the getters are not aligned with default))}
 
@@
 we omit the definitions of the methods:
-\bruno{No need to show full code in the paper. You can just put \ldots in the body of the collectLit methods.}
 
-\lstinputlisting[linerange=70-86]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_COLLECTLIT
+\begin{lstlisting}[]
+interface ExpC extends Exp { List<Integer> collectLit(); }
+@Obj interface LitC extends Lit, ExpC {
+    default List<Integer> collectLit() { ... }}
+@Obj interface AddC extends Add, ExpC {
+    ExpC e1(); ExpC e2(); //return type refined!
+    default List<Integer> collectLit() { ... }}
+\end{lstlisting}
 
@@
 
-\lstinputlisting[linerange=90-94]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=85-87]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 
@@
 as simple as:
-
-\bruno{Add code here}
-
+\begin{lstlisting}
+ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
+\end{lstlisting}
 \noindent In contrast, in a pure Java solution, the tedious instantiation code would need 

### 2015-11-25 paper/sections/RelatedWork.tex
related work

@@
 
-In certain situations, our approach allows automatic refinement for return
-types. This is part of a bigger topic in class based languages: expressing and
+In certain situations, object interfaces allow automatic refinement for \emph{return
+types}. This is part of a bigger topic in class based languages: expressing and
 preserving type recursion and (nominal/structural) subtyping at the same time.
@@
 invoking a method with MyType in a parameter position, the exact type of the
-receiver must be known.  This is huge limitation in class based object oriented
-programming, and is exasperated by interface-only programming as we propose: no
-type is ever exact since classes are not part of the language. A recent
-article~\cite{Saito2013933} lights up this topic, since they propose two
-features: exact statements \yanlin{do we support exact statements?}\marco{no we
-  do not. When is that it may seams we do?} and non-inheritable methods; both
+receiver must be known.  This is a big limitation in class based object oriented
+programming, and is exasperated by the interface-based programming we propose: no
+type is ever going to be exact since classes are not explicitly used. A recent
+article~\cite{Saito2013933} lights up this topic, proposing two
+new features: exact statements and non-inheritable methods. Both are
 related to our work: any method generated inside of the \Q@of@ method is indeed
-non-inheritable, since there is no class name to extends from, and exact
+non-inheritable, since there is no class name to extend from, and exact
 statements (a form of wild-card capture on the exact run-time type) could
 capture the ``exact type'' of an object even in a class-less environment.
+%Interestingly, our use of covariant type refinement is enough to emulate
 
@@
 \end{comment}
-\subsection{Meta-programming Competes with Language extensions}
-The most obvious solution to features addition to a language is language
+
+\subsection{Meta-programming Competes with Language Extensions}
+\bruno{needs to be polished better, I think}
+The most obvious solution to adding features to a language is language
 extension.  It is often implemented as syntactic extensions that can be
@@
 
-On the other side, when the starting language have a flexible enough syntax and
+On the other side, when the base language has a flexible enough syntax and
 a fast and powerful enough reflection mechanism, we may just need to play with
@@
 can not influence the type system, so many solutions valid in python or other
-languages could not be applicable, or may be applicable at the cost of loosing
+languages are not be applicable, or may be applicable at the cost of losing
 safety.
@@
 (annotating a class and provide getters), then we can insert in the compilation
-process a tool that exam and enrich the code before compilation. No need to
+process a tool that examins and enrichs the code before compilation. No need to
 modify the original source; for example we can work on temporary files
@@
 \subsection{Formalization of Java8}
-We provide a simple and well designed formalization for a subset of Java
+We provide a simple formalization for a subset of Java
 including default/static interface methods and object initialization literals
@@
 (default) methods in Java. However this formalization is limited to model
-exactly one method inside classes/interfaces.
+exactly one method inside classes/interfaces.\bruno{explain why having multiple 
+methods is significantly more challenging.}
 
@@
 %   the issue.}
-\yanlin{bug claim removed.}

### 2015-11-25 paper/sections/FutureWork.tex

@@
 
-\subsection{Keywords in Classes} %{Private state}
+\subsection{Qualifiers in Methods} %{Private state}
 \yanlin{briefly mentioned ``synchronized'' keyword, didn't put the workaround
   here, it's in the code.}
-The biggest limitation of our approach is the inability to model all features
-for classes (private, protected, synchronized, etc.). For example, the absence
+The biggest limitation of our approach is the inability to model qualifiers
+for class methods (private, protected, synchronized, etc.). For example, the absence
 of the support for private/protected methods in Java8 interfaces forces all
 members of interfaces to be public, including static methods. Since we use
-abstract methods to encode the state, our state is always all public; however it
+abstract methods to encode the state, our state is always all public. Still, because 
+the state can only be accessed by methods, it 
 is impossible for the user to know if a certain method maps directly to a field
-or if it have a default implementation.  If the use wants a constructor that
+or if it has a default implementation.  If the user wants a constructor that
 does not directly maps to the fields, (as for secondary constructors in Scala)
@@
 Currently, it is possible to use a public nested class with private static
-methods inside, but this is ugly and cumbersome. We are considering if our
-annotation processor can take code with \Q!@Private! annotation and turn it into
+methods inside, but this is ugly and cumbersome. One possibility is that the
+annotation processor takes code with a \Q!@Private! annotation, and turns it into
 static private methods of a nested class. In this extension, also the \Q@of@
@@
 \subsection{State initialization}
-As discuss before, the user can trivially define its own \Q@of@ method, and
+As discussed before, the user can trivially define its own \Q@of@ method, and
 initialize a portion of the state with default values.  However, the
@@
 time. This can be very relevant and tricky in a non functional setting.
+\bruno{Yanlin please polish text/break long sentences.}
 
+\begin{comment}
 \subsection{Class Invariants in ClassLess Java}
@@
 \Q@checkInvariant()@,  a composed implementation could be automatically generated, composing by \Q@&&@ the various competing implementations.
+\end{comment}
+\bruno{removed the invariants stuff; we need space, I think.}
 
 \subsection{Clone, toString, equals and hashCode}
-Methods originally defined in Java class \Q@Object@, as \Q@clone@ and \Q@toString@, can be supported by our approach, but they need special care.
-If an interface annotated with \mixin asks an implementation for \Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily generate one from the fields.\footnote{In particular, for clone we can do automatic return type refinement as we do for \Q@with-@ and fluent setters.
-Note how this would solve most of the Java ugliness related to \Q@clone@ methods.}
 
-However, if the user wish to provide his own implementation, since the method is
-also implemented in \Q@Object@, a conflict would arise. Then the user has to
+Methods originally defined in Java class \Q@Object@, as \Q@clone@ and
+\Q@toString@, can be supported by our approach, but they need special
+care. If an interface annotated with \mixin asks an implementation for
+\Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily
+generate one from the fields.\footnote{In particular, for clone we can
+  do automatic return type refinement as we do for \Q@with-@ and
+  fluent setters. Note how this would solve most of the Java ugliness
+  related to \Q@clone@ methods.}
+However, if the user wishes to provide his own implementation, since the method is
+also implemented in \Q@Object@, a conflict arises. The user has to
 explicitly resolve the conflict inside \Q@of@, by implementing the method and

### 2015-11-25 paper/sections/RelatedWork.tex

@@
 \bruno{needs to be polished better, I think}
+\bruno{Mention/Compare with Language Virtualization work in Scala. They also re-interpret 
+syntax.}
 The most obvious solution to adding features to a language is language
@@
 extensible language allowing programmers to extend it with custom features by
-definitions in meta-DSLs (SDF, Stratego, etc).
+definitions in meta-DSLs (SDF, Stratego, etc).\bruno{so, what's the point that 
+we want to make with SugarJ?}
 

### 2015-11-25 paper/sections/CaseStudy.tex
Fluent interface case study revised

@@
 JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
-interfaces, programmers have to either handwrite everything or create a
-wrapper around the original non-fluent interfaces using \textbf{this}. In Java,
-several libraries help create fluent APIs easily (by code generation)\bruno{do they use code 
-generation to produce the fluent interfaces?}, including
-jOOQ, op4j, fluflu, JaQue, etc.
+interfaces, programmers have to either handwrite everything or create a wrapper
+around the original non-fluent interfaces using \textbf{this}. In Java, there
+are several libraries (including jOOQ, op4j, fluflu, JaQue, etc) providing useful
+fluent APIs. However most of them only provide a fixed set of predefined fluent
+interfaces. Fluflu enables the creation of a fluent API and implements control
+over method chaining by using Java annotations. However methods that returns
+\textbf{this} are still hand-written.
 
-The \mixin annotation can also be used to create fluent interfaces. 
-When creating fluent interfaces with \mixin, there are two main advantages:
+The \mixin annotation can also be used to create fluent interfaces.  When
+creating fluent interfaces with \mixin, there are two main advantages:
 \begin{enumerate}
@@
 \texttt{select}, \texttt{from} and \texttt{where} clauses:
-\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
-\bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
-to have to use the of explicitly to initialized all fields.}
+\lstinputlisting[linerange=6-11]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
+% \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
+% to have to use the of explicitly to initialized all fields.}
 
@@
 
-\lstinputlisting[linerange=35-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
-\bruno{write query in 1 line}
+\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
 
@@
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
-\lstinputlisting[linerange=17-20]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+\lstinputlisting[linerange=15-18]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
 This way, when a \texttt{ExtendedDatabase} query created, 
@@
 
-\lstinputlisting[linerange=44-48]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
-\bruno{write query in 1 line}
+\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
 

### 2015-11-25 paper/sections/CaseStudy.tex
conflict

--- 2015-11-25 paper/sections/CaseStudy.tex
@@
 
-\subsection{A Small Game Example}
+\subsection{A Maze Game}
 The last case study is a simple Maze game, which is used\bruno{where?}

### 2015-11-25 paper/sections/CaseStudy.tex
game case study revised. run.sh script should be working on your machines.

@@
 
-\subsection{A Maze Game}
-The last case study is a simple Maze game, which is used\bruno{where?}
+\subsection{A Small Game Example}
+The last case study is a simple Maze game, which is often
+used~\cite{gof,bono14,vicentedesign} 
+\bruno{where?}
+\yanlin{the last one ``design patterns in OCaml''is not a proper paper, not sure
+whether to cite.}
 to evaluate code reuse ability related to inheritance and design
@@
 generated \texttt{of} methods. The number of source lines of code (SLOC)
-was reduced by 40\%\bruno{explain why? we generate code for getters/setters?}. 
+was reduced by 40\% due to both the removal of instantiation overhead and
+generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}. 
 To ensure a fair comparison, we used the same coding style as Bono et al.'s.

### 2015-11-25 paper/run.sh

@@ -1,4 +1,4 @@
-#!/bin/zsh
+#!/bin/bash
 
 if [ $1 ]
 then

### 2015-11-25 paper/sections/CaseStudy.tex
added related work: Malayeri & Aldrich (CZ)

@@
 
-\subsection{A Small Game Example}
-The last case study is a simple Maze game, which is often
-used~\cite{gof,bono14,vicentedesign} 
-\bruno{where?}
-\yanlin{the last one ``design patterns in OCaml''is not a proper paper, not sure
-whether to cite.}
+\subsection{A Maze Game}
+The last case study is a simplified variant of Maze game, which is often
+used~\cite{gof,bono14} 
 to evaluate code reuse ability related to inheritance and design

### 2015-11-25 paper/sections/Formalization.tex
SOme fixes thanks to H!
A couple of typo fixed, ready for tomorrow!

@@
 $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquelly indentfied by their name.
 
@@
 We omit the definition of
-the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation, that can be read off the interface table and given by the
-\textbf{extends} clauses in the interfaces. %
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation. %
 \footnote{\marco{find a better place for} Notice how there are no classes, thus there is no
@@
 and names of the method parameters.  The
-notation $\m^\mh$ and $\C^\mh$ denotes, respectivelly, extracting the
+notation $\m^\mh$ and $\C^\mh$ denotes respectivelly, extracting the
 method name and the return type from a method header. $\mBody(\m,\C)$,
@@
 
-In Figure~\ref{ET} we show the typing rules.  We discuss first the
+In Figure~\ref{ET} we show the typing rules.  We discuss the
 most interesting rules, that is \rn{t-Obj} and \rn{t-Intf}. Rule
@@
 \InTextDef{18em}{
-\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR\QM; \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR\QM;
 }{
@@
 \override(\methods(\m),
-\shadow(\m,\Aux{tops}(\Cs)))
+\shadow(\m,\tops(\m,\Cs\ )))
 }\\
@@
 \InTextAssert{35ex}{
-\mBody(\m,\C)\in\tops(\m,\Cs)}{
+\method\in\tops(\m,\Cs)}{
 }\\
 \InTextWith{
-  \mBody(\m,\C)\mbox{ not a static method and } 
-  \C\in\Cs\ \mbox{ such that }
+  \mBody(\m,\C)=\method,\ \method\mbox{ not a static method, and }}\\ 
+\tab\tab  \C\in\Cs\ \mbox{ such that }
 \forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C
-}
+
 \end{array}$

### 2015-11-25 paper/sections/Translation.tex
Last little fix, now formalization is ok for me.
fix bibtex errors
A couple of typo fixed, ready for tomorrow!

@@
 In our simplified formalization we consider this to be just an error.
-In our prototype we keep \Q@overloading@ into account, and so we check that an of method with the same signature of the one we would like to generate is not already present.\marco{Do we check it?}
+In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}
 
@@
 We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
-To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.\\*
+To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.
+
 \noindent$\begin{array}{l}
@@
 \C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return} \C_0\QM{.of(}\es\QM{);\}}} \\
+\QM{return}\ \C_0\QM{.of(}\es\QM{);\}}} \\
 \InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;}\\
@@
 %}\\\InTextWith{
-\method\in \dom(\C_0)
+\method=\mBody(\_,\C_0)
 }\\
@@
   \begin{itemize}
+   \item $\C$ must be an interface type (no classes or primitive types).
     \item As for before, check that $\C'$ is a supertype of the current interface type $\C_0$.
@@
 
-\marco{ insert somewhere description of  fluent setter~\cite{the lombock thread, something about fluent stuff}.
+\marco{ insert somewhere description of  fluent setter~\cite{the-lombock-thread-something-about-fluent-stuff}.
 This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}}.

### 2015-11-25 paper/sections/Appendix.tex
Remove the feature of checking existence of ofMethod. Is buggy and nontrivial.

@@
 
-\subsection{Proof of LEMMA 1}\label{subsec:proof1}
-\marco{this seams to be just a sublemma for Lemma2, I think you can mention it in the appendix only.}
+\subsection{LEMMA 1 and Proof}\label{subsec:proof1}
 \textbf{LEMMA 1. }
@@
 \item If $m\in\dom(\methods')$, we are to prove $m\in\dom([\![\II_0]\!])$.
-So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.\marco{need to add here that we know of is not in dom(..)} On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method,
+So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.\marco{need to add here that we know of is not in dom(..)}\haoyuan{BTW, the condition of the translation should be $\QM{of}\notin\dom(\methods)$ or $\QM{of}\notin\dom(\C_0)$?} On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method,
  we can apply the last clause of override, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
@@
 }
+
+\subsection{LEMMA 2 and Proof}\label{subsec:proofs}
+\textbf{LEMMA 2. }
+For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
+\begin{proof}
+The proof is based on induction. By the grammar shown in Figure~\ref{Grammar}, there are three cases for an arbitrary expression $e$:
+\begin{itemize}
+\item A variable or a field update. The type preservation for $e$ is ensured by induction.
+\item A method call (normal, static or super).
+Such a method won't be ``removed'' by the translation and is still there, and the types of arguments remain unchanged.
+The only difference is that the return type can be refined, in which case the expression $e$ is typed in a subtype of $I$.?
+\item An object creation.
+\end{itemize}
+\end{proof}
+
 \subsection{Proof of LEMMA 2}\label{subsec:proof2}

### 2015-11-25 paper/sections/Translation.tex
Remove the feature of checking existence of ofMethod. Is buggy and nontrivial.

--- 2015-11-23 paper/sections/Translation.tex
@@
 $\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
-if  
+if $\II_0$ has \mixin,
 $\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
@@
 $\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
-
-\haoyuan{In theorem: $\C_0$ is valid and $\QM{of}\notin\dom(\C_0)$. To prove the theorem we need an extra lemma: if the translation is well defined, then $[\![\II_0]\!]$ is well formed.}
 
@@
 To prove the theorem we introduce two lemmas below. The complete proof is available in Appendix \ref{subsec:proof1}, \ref{subsec:proof2} and \ref{subsec:proof3}.\\
-
-\noindent\textbf{LEMMA 1. }
-Assume that
-\begin{align*}
-\II_0\ &=\ \mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
-[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \methods' \cC
-\end{align*}
-If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.\\
 

### 2015-11-25 paper/sections/RelatedWork.tex

@@
 models have been proposed in over the years, including C++ virtual
-inheritance, mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
+inheritance, mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and
+hybrid model such as CZ~\cite{malayeri2009cz}.
 They provide novel programming architecture models in the OO paradigm. In terms
@@
 \bruno{How about Malayeri and Aldrich's paper? shouldn't that be discussed?}
+\yanlin{Newly added discussion on CZ. please double check.}
 
@@
 interfaces. Also, our approach supports proper constructor mechanism.
+
+\paragraph{CZ.}
+Malayeri and Aldrich proposed a model CZ~\cite{malayeri2009cz} which aims to do
+multiple inheritance without diamond problem. They divide inheritance into two
+separate concepts: inheritance dependency (using \texttt{require}) and
+implementation inheritance (using \texttt{extends}). Using a combination of
+\texttt{requires} and \texttt{extends}, a program with diamond inheritance can be
+transformed to one without diamonds. Also in this approach, fields and multiple
+inheritance can coexist. However as shown in the example in the article, the
+untangling of inheritance also untangles class structure. In CZ, not only the
+number of classes (from 4 to 6 in the stream example), but also the class
+hirarchy complexity increases. Our approach does not complicate the structure,
+and (conceptual) fields coexist with multiple inheritance.
 

### 2015-11-26 paper/sections/RelatedWork.tex
related work before 8.3

@@
 constructor calls to virtual superclasses, which would potentially cause serious
-semantic errors. \bruno{What happens in our approach for the same case?}
+semantic errors. \bruno{What happens in our approach for the same
+  case?}\yanlin{explained below please double-check.} In our
+approach, \mixin annotation processor has the full control of object
+initialization mechanism, and this mechanism is clean and transparent to
+users. Moreover, if not satisfied with the default generated \texttt{of} method,
+customized factory methods are also allowed to be defined.
 
@@
 The author of ~\cite{bono14} agree that our revised version is
-cleaner, typesafe and more direct.\bruno{do we want do say this?}
+cleaner, typesafe and more direct.\bruno{do we want do say this?}\yanlin{need to
+  ask Marco.}
 
@@
 In the trait model two concepts (traits and classes) coexist and
-cooperate. Some authors~\bruno{who?} see this as good language design
+cooperate. Some authors~\bruno{who?}\yanlin{need to ask Marco} see this as good language design
 fostering good software development by helping programmers to think
@@
 need of two concepts and the absence of state as drawbacks of this
-model~\cite{malayeri}.  Object interfaces are units of reuse, and at
+model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
 the same time they provide factory methods for instantiation and
@@
 article~\cite{Saito2013933} lights up this topic, proposing two
-new features: exact statements and non-inheritable methods. Both are
-related to our work: any method generated inside of the \Q@of@ method is indeed
-non-inheritable, since there is no class name to extend from, and exact
+new features: exact statements and nonheritable methods. Both are
+related to our work: 1) any method generated inside of the \Q@of@ method is indeed
+non-inheritable since there is no class name to extend from; 2)exact
 statements (a form of wild-card capture on the exact run-time type) could

### 2015-11-26 paper/sections/Overview.tex
Overview

@@
 types of the methods. Fluent setters are a variant of mutable setters,
-and are discussed in the case study in Section~\ref{}.
+and are discussed in the case study in Section~\ref{sec:dsls}.
 
@@
 \paragraph{Field type refinement with object interfaces}
-In our proposed solution,
+Object interfaces address the challange of type-refinement as follows:
 \begin{itemize}
-\item by covariant method overriding we refine the field type
-\item by \emph{overloading} we define a new setter with the more precise type
-\item by default method we provide an implementation for the setter called with the old signature.
+\item by covariant method overriding we refine the type of the location field to \texttt{Point3D};
+\item by \emph{overloading} we define a new setter for the location field with the more precise type;
+\item using a default method we provide an implementation for the setter called 
+with the old signature.
 \end{itemize}
@@
 
-From the type perspective, the key is the covariant method overriding; however from the semantic perspective the key is the implementation for the setter with the old signature:
-by using the \Q@with@ method we may use the information for \Q@z@ already stored in the object to forge an appropriate \Q@Point3D@ to store.
-Note how all the informations about what fields sits in \Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the \Q@with@ method, and is transparent for the implementer of \Q@Bird@.
+\noindent From the type perspective, the key is the covariant method
+overriding of \texttt{location()}. However from the semantic
+perspective the key is the implementation for the setter with the old
+signature.
 
-\bruno{See how to fit this in.}
-\subsubsection*{\Q@Point3D@ and properties updater}
-\Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
-as follows:
+\paragraph{\Q@Point3D@ and properties updater}
+To implement the old setter in a convenient way, \mixin supports one
+last type of operations: property updater \texttt{with}
+methods. Unlike the \texttt{withX} (where \texttt{X} stands for a
+field name) methods presented so far, property updaters take several
+fields at once, contained in an interface, and copy those fields into
+fields of another interface.
+The \Q@Point3D@ interface is defined as follows:
 \lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
-\marco{Fix over: with(Point must be with(Point2D
-and it should be after withZ, and on more lines}
-Just to give you a feeling on how much boring code \mixin is generating,  we show the expanded code without then \mixin annotation\footnote{
-As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
-swap the arguments of one of the many calls of \Q@Point3D.of@.
-}.\begin{lstlisting}
+\noindent By using the new \Q@with@ method we may use the information for
+\Q@z@ already stored in the object to forge an appropriate \Q@Point3D@
+to store.  Note how all the informations about what fields sits in
+\Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the
+\Q@with@ method, and is transparent for the implementer of \Q@Bird@.
+
+\paragraph{Generated Boilerplate}
+Just to give a feeling on how much boring code \mixin is generating,
+we show the generated code for the \texttt{Point3D} interface 
+Figure~\ref{fig:boilerplate}. The generated code is very repetitive,
+writing such code by hand can easily induce bugs, for example a
+distracted programmer may swap the arguments of one of the many calls of
+\Q@Point3D.of@.
+\begin{figure}
+\begin{lstlisting}
 interface Point3D extends Point2D{
@@
 \end{lstlisting}
+\caption{Generated boilerplate code.}
+\label{fig:boilerplate}
+\end{figure}
 Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
@@
 If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
-Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
+Again, the philosophy is that if the programmer provides something directly, \mixin does not touch it.
 \Q@with-@ methods (functionally) update a field/property at a time.
@@
 \subsection{Advanced Multiple Inheritance with Object Interfaces}
-The \emph{``multiple inheritance''} case appears at interface
-\texttt{Pegasus}.
-Pegasus (one of the best known creatures
-in Greek mythology) can not only \emph{run} but also \emph{fly}! 
-It would be handy if they could obtain \texttt{fly} and \texttt{run} functionality from both \texttt{Horse} and \texttt{Bird}.
-
- Using interfaces and \mixin annotation,
-this can happen transparently.
+Finally, we can define Pegasus as simply as we did in the simplified (and stateless) 
+version in the right-side of Figure~\ref{fig:pegasus}:
 
@@
 
-Note how  even the non trivial pattern for field type refinement is transparently composed, and our pegasus have a \Q@Point3D@ \Q@location@.
-Note that this works because \Q@Horse@ do not perform any field type refinement , otherwise we may have to choose/create a common subtype in order for pegasus to exists.
+\noindent Note how even the non trivial pattern for field type refinement is
+transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
+\Q@location@. This works because \Q@Horse@ do not perform
+any field type refinement, otherwise we may have to choose/create a
+common subtype in order for \texttt{Pegasus} to exists.
 

### 2015-11-26 paper/sections/CaseStudy.tex

@@
 
-\subsection{Embedded DSLs with Fluent Interfaces}
+\subsection{Embedded DSLs with Fluent Interfaces}\label{sec:dsls}
 Since the style of fluent interfaces was invented in Smalltalk as method

### 2015-11-26 paper/sections/RelatedWork.tex
revising related work

@@
 definitions in meta-DSLs (SDF, Stratego, etc).\bruno{so, what's the point that 
-we want to make with SugarJ?}
+we want to make with SugarJ?}\yanlin{we were talking about lightweight language
+sugar, so SugarJ is a comparison to ours. Where using SugarJ requires learning
+various DSLs like SDF, stratego, etc, in comparison, our approach is just using
+Java. }
 
@@
 
-Here (compile/load time) code generation comes at the rescue: if for a certain
+Here (compile/load time)\yanlin{only compile time should do?} code generation comes at the rescue: if for a certain
 feature (\mixin in our case) it is possible to use the original language syntax

### 2015-11-27 paper/sections/Translation.tex

@@
 $\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
+there is no $\C<:\C_0$, there is no \Q@new @$\C_0$\Q@(){..}@,
 then in the interface table
 $[\![\II_0]\!]\II_1\ldots\II_n$\\*
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.
 

### 2015-11-27 paper/sections/RelatedWork.tex
more related work

@@
 The author of ~\cite{bono14} agree that our revised version is
-cleaner, typesafe and more direct.\bruno{do we want do say this?}\yanlin{need to
-  ask Marco.}
+cleaner, typesafe and more direct.
+%\bruno{do we want do say this?}\yanlin{need to ask Marco.}
 
@@
 In the trait model two concepts (traits and classes) coexist and
-cooperate. Some authors~\bruno{who?}\yanlin{need to ask Marco} see this as good language design
+cooperate. Some authors~\cite{bettini2013pure}\yanlin{There are many trait related papers of ferruccio
+  damiani, please check whether citing this one is ok.} 
+see this as good language design
 fostering good software development by helping programmers to think
@@
 statements (a form of wild-card capture on the exact run-time type) could
-capture the ``exact type'' of an object even in a class-less environment.
-%Interestingly, our use of covariant type refinement is enough to emulate
-
+capture the ``exact type'' of an object even in a class-less
+environment. 
+Admittedly, MyType greatly enhances the expressiveness and extensibility of
+object-oriented programming languages. We are using covariant-return types to
+simulate some uses of MyType. But our approach only works on method positive
+positions, whereas MyType is more general, as it works on any
+positions. Nevertheless our approach is still useful for modeling fluent
+interfaces and solving expression problems,etc.
 % The addition of MyType to a language will allow easy definition of
 %binary methods, methods with recursive types (i.e., the same type of the
-%receiver appears in the argument or return positions of methods), etc. MyType
-%greatly enhances the expressiveness and extensibility of object-oriented
-%programming languages.
-
-%In our approach, we are using covariant-return types to simulate some uses of
-%MyType. But our approach only works on method positive positions, whereas MyType
-%is more general, as it works on any positions. Nevertheless our approach is
-%still useful for modeling fluent interfaces and solving expression
-%problems,etc.
+%receiver appears in the argument or return positions of methods), etc.
 
@@
 
-Here (compile/load time)\yanlin{only compile time should do?} code generation comes at the rescue: if for a certain
-feature (\mixin in our case) it is possible to use the original language syntax
-to \emph{express-describe} any specific instantiation of such feature
-(annotating a class and provide getters), then we can insert in the compilation
-process a tool that examins and enrichs the code before compilation. No need to
-modify the original source; for example we can work on temporary files
-\yanlin{Here what does temporary files mean? could explain more?}. Java is
-a particular good candidate for this kind of manipulation since it already
-provide ways to define and integrate such tools in its own compilation process:
-in this way there is no need of temporary files, and there is a well defined way
-of putting multiple extensions together.
+Here compile time code generation comes at the rescue: if for a certain feature
+(\mixin in our case) it is possible to use the original language syntax to
+\emph{express-describe} any specific instantiation of such feature (annotating a
+class and provide getters), then we can insert in the compilation process a tool
+that examins and enrichs the code before compilation. No need to modify the
+original source; for example we can work on temporary files Java is a particular
+good candidate for this kind of manipulation since it already provide ways to
+define and integrate such tools in its own compilation process: in this way
+there is no need of temporary files, and there is a well defined way of putting
+multiple extensions together.
 
@@
 \subsection{Formalization of Java8}
-We provide a simple formalization for a subset of Java
-including default/static interface methods and object initialization literals
-(often called anonymous local inner classes).  A similar formalization was
-drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize defender
-(default) methods in Java. However this formalization is limited to model
-exactly one method inside classes/interfaces.\bruno{explain why having multiple 
-methods is significantly more challenging.}
+We provide a simple formalization for a subset of Java including default/static
+interface methods and object initialization literals (often called anonymous
+local inner classes).  A similar formalization was drafted by Goetz and
+Field~\cite{goetz12fdefenders} to formalize defender (default) methods in
+Java. However this formalization is limited to model exactly one method inside
+classes/interfaces.\bruno{explain why having multiple methods is significantly
+  more challenging.} Our code generation generates several methods that may
+interacts with each other and methods defined by user. Thus implementing and
+formalizing multiple methods inside interfaces adds complexity greatly.
 

### 2015-11-27 paper/sections/FutureWork.tex
Future work

@@
 
-\subsection{Qualifiers in Methods} %{Private state}
-\yanlin{briefly mentioned ``synchronized'' keyword, didn't put the workaround
-  here, it's in the code.}
+\subsection{Qualifiers in Methods} %{Private state
 The biggest limitation of our approach is the inability to model qualifiers
@@
 recognize a field if the getter is provided and the setter is required, and
-could generate the following code: \yanlin{both of the two generated val()
-  method should be public to pass compilation. Did you omit the public keyword
-  on purpose to make the code clean? }
-\begin{lstlisting}
-interface Box{ 
-  default int val(){return 0;} //provided
-  void val(int _val);//provided
-  static Box of(){return new Box(){//generated
-    int val=Box.super.val();
-    int val(){return val;}
-    void val(int _val){val=_val;}
-    };}}
-\end{lstlisting}
-We are unsure of the value of this solution: is very tricky, the user define a
-method that (contrary to our usual expectation) is actually overridden in a way
-that the behaviour changes, but change only after the first setter is called,
-plus this code would cache the result instead of re-computing it every
-time. This can be very relevant and tricky in a non functional setting.
-\bruno{Yanlin please polish text/break long sentences.}
+could generate the following code:
+\lstinputlisting[linerange=4-11]{../UseMixinLombok/src/test/TestStateInitialization.java}% APPLY:linerange=STATE_INIT
+
+We are unsure about the value of this workaround because of its trickiness. In
+order to define a state with initialization, users have to learn this pattern:
+defining a method which will be overridden later with behaviour change, but
+change only after the first setter is called. Moreover, this code would cache
+the result instead of re-computing it every time. This can be very tricky in a
+non functional setting.
+%\bruno{Yanlin please polish text/break long sentences.}
 
@@
 \end{comment}
-\bruno{removed the invariants stuff; we need space, I think.}
+%\bruno{removed the invariants stuff; we need space, I think.}
 
@@
   default Point clone(){ return Point.of(0,0);}//user defined clone
-  public static Point of(int _x,int_y){
+  public static Point of(int _x, int _y){
     return new Point(){...
       public Point clone(){ return Point.super.clone();}
-      };  }  }
+    };}}
 \end{lstlisting}

### 2015-11-27 UseMixinLombok/src/test/TestStateInitialization.java

@@ -1,0 +1,15 @@
+package test;
+
+//BEGIN_STATE_INIT
+interface Box{ 
+    default int val(){return 0;} //provided
+    void val(int _val);//provided
+    static Box of(){return new Box(){//generated
+        int val=Box.super.val();
+        public int val(){return val;}
+        public void val(int _val){val=_val;}
+    };}}
+//END_STATE_INIT
+
+public class TestStateInitialization {
+}

### 2015-11-27 UseMixinLombok/src/test/TestSynchronize.java

@@ -41,14 +41,3 @@ interface B {
 public interface TestSynchronize {
 
 }
-
-
-interface Box{ 
-    default int val(){return 0;} //provided
-    void val(int _val);//provided
-    static Box of(){return new Box(){//generated
-        int val=Box.super.val();
-        public int val(){return val;}
-        public void val(int _val){val=_val;}
-    };}
-}

### 2015-11-27 paper/sections/CaseStudy.tex
Re read before korea completed

@@
 is a well-known problem about modular extensibility issues in software
-evolution. Recently, a new solution~\cite{eptrivially} that uses only
-covariant type refinement that has been shown to work in Java. When
-that solution is modelled with interfaces and default methods, it can
-even provide independent extensibility: that is, the ability to
-assemble a system, from multiple, independently developed extensions.
-Unfortunatelly, a plain Java solution has alot of code verbosity in
-Java due to code required for instantiation, which makes the solution 
-cumbersome to use. The \mixin annotation helps to remove such code, and 
-makes the solution quite easy to use.
+evolution. Recently, a new solution~\cite{eptrivially}, \marco{repharesed} using only
+covariant type refinement, has been shown.
+When such solution is modelled with interfaces and default methods, it can
+even provide independent extensibility: the ability to
+assemble a system from multiple, independently developed, extensions.
+Unfortunatelly, the required instantiation code makes a plain Java solution verbose and cumbersome to use.
+The \mixin annotation is enough to remove the boilerplate code, making the presented approach very appealing.
 % \bruno{We should try to do some space saving in here. You could, for example 
@@
 interface Exp { int eval(); }
+
 @Obj interface Lit extends Exp {
@@
 \paragraph{Adding a New Type of Expressions}
-It is easy to add new types of expressions to the code in the initial system in a
-modular way. For example, the following code shows how to add subtraction.
+In the object oriented paradigm, it is easy to add new types of expressions. For example, the following code shows how to add subtraction.
 
@@
 
+\marco{text under need to be improved}
 \paragraph{Adding a New Operation} The difficulty of the EP in object-oriented 
-languages arizes from adding new operations. For example, adding a pretty printing 
+languages arises from adding new operations. For example, adding a pretty printing 
 operation, would typically involve changing all of the existing code. However, a solution 
@@
 
+\marco{ use the twoc colums here too}
 \lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT

### 2015-11-27 paper/sections/Formalization.tex
Re read before korea completed

@@
 is accepted by ECJ4.5.1 and rejected by javac.
-By email communication with Brian Goetz (lead Java8 designer) we have confirmation that the expected behaviour is rejection, and that this is a bug in ECJ.
+By email communication with Brian Goetz (lead Java8 designer) we have confirmation that the expected behaviour is rejection, and that this is a bug in ECJ.\footnote{This bug is fixed in the developer version ECJ???, not yet released as stable.}
 %Table~\ref{table:javabug} shows the method overriding bug in Java. In the

### 2015-11-27 paper/sections/FutureWork.tex
Re read before korea completed

@@
 \begin{lstlisting}
-@Mixin interface Point{
-  int x(); int y();
-  static Point of(int val){return Point.of(val,val);}  
-}
+@Obj interface Point{  int x(); int y();
+  static Point of(int val){return Point.of(val,val);}  }
 \end{lstlisting}
@@
 methods inside, but this is ugly and cumbersome. One possibility is that the
-annotation processor takes code with a \Q!@Private! annotation, and turns it into
+annotation processor takes methods with a \Q!@Private! annotation, and turns it into
 static private methods of a nested class. In this extension, also the \Q@of@
@@
 \begin{lstlisting}
-@Mixin interface Box{ default int val(){return 0;} }
+@Obj interface Box{ default int val(){return 0;} }
 \end{lstlisting}
@@
 We are unsure about the value of this workaround because of its trickiness. In
-order to define a state with initialization, users have to learn this pattern:
-defining a method which will be overridden later with behaviour change, but
-change only after the first setter is called. Moreover, this code would cache
+order to define a state with initialization, users have to defining a method trusting that it will be overridden later with a behavioural change, but
+such change is visible only after the first time the setter is called. Moreover, this code would cache
 the result instead of re-computing it every time. This can be very tricky in a
@@
 However, if the user wishes to provide his own implementation, since the method is
-also implemented in \Q@Object@, a conflict arises. The user has to
-explicitly resolve the conflict inside \Q@of@, by implementing the method and
+also implemented in \Q@Object@, a conflict arises. The generated code can resolve the conflict inside \Q@of@, by implementing the method and
 delegating it to the user implementation, thus
 \begin{lstlisting}
-@Mixin interface Point{ int x(); int y();
+@Obj interface Point{ int x(); int y();
   default Point clone(){ return Point.of(0,0);}//user defined clone

### 2015-11-27 paper/sections/Overview.tex
Re read before korea completed

@@
 \begin{lstlisting}
-@Mixin interface Point2D{ int x(); int y();}
+@Obj interface Point2D{ int x(); int y();}
 \end{lstlisting}
@@
 \begin{lstlisting}
-@Mixin interface Point2D { 
+@Obj interface Point2D { 
   int x();    int y(); // getters
@@
 \begin{lstlisting}
-@Mixin interface Point2D{ 
+@Obj interface Point2D{ 
   int x(); int y();
@@
 The \Q@Point3D@ interface is defined as follows:
-\lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=51-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 \noindent By using the new \Q@with@ method we may use the information for

### 2015-11-27 paper/sections/RelatedWork.tex
Re read before korea completed

@@
 \emph{express-describe} any specific instantiation of such feature (annotating a
-class and provide getters), then we can insert in the compilation process a tool
+class and providing getters), then we can insert in the compilation process a tool
 that examins and enrichs the code before compilation. No need to modify the
-original source; for example we can work on temporary files Java is a particular
+original source; for example we can work on temporary files. Java is a particular
 good candidate for this kind of manipulation since it already provide ways to
@@
 formalizing multiple methods inside interfaces adds complexity greatly.
-
+\marco{I do not see this as a good explanation, I think we may just remove the `one method limitation discussion'.
+Is their work modelling interactions between abstract and default methods?
+like making a default method abstract again?}
 % As an evidence of the attention and care present in our formalization work,

### 2015-11-27 paper/sections/Translation.tex
Re read before korea completed

@@
 \subsection{Results}
+Informally, in a type correct interface table, if we translate a specific interface, then
+the resulting table is type correct, or there is a subtype of the translated interface in the original table.
 \textbf{THEOREM. }
@@
 interface A { int x(); A withX(int x); }
-@Mixin interface B extends A {}
+@Obj interface B extends A {}
 interface C extends B { A withX(int x); }

### 2015-11-27 UseMixinLombok/src/pegasusJava/Main.java

@@ -4,12 +4,12 @@ package pegasusJava;
 
 //BEGIN_PEGASUS_JAVA
 interface Animal {} // no points yet!
-interface Horse {
+interface Horse extends Animal{
 	default void run() {
 		System.out.println("I can run!");
 	};
 }
-interface Bird {
+interface Bird extends Animal{
 	default void fly() {
 		System.out.println("I can fly!");
 	}

### 2015-11-27 UseMixinLombok/src/pegasusLombok/Main.java

@@ -4,11 +4,11 @@ import lombok.Obj;
 
 
 interface Animal {}
-@Obj interface Horse {
+@Obj interface Horse extends Animal{
 	default void run() {System.out.println("I can run!");};
 }
 //BEGIN_PEGASUS_LOMBOK
-@Obj interface Bird {
+@Obj interface Bird extends Animal{
 	default void fly() {System.out.println("I can fly!");}
 }
 @Obj interface Pegasus extends Horse, Bird {}

### 2015-11-27 UseMixinLombok/src/test/TestAnimal.java

@@ -50,11 +50,11 @@ interface Point2D {
 //BEGIN_POINT3D
 @Obj
 interface Point3D extends Point2D {
-	int z(); Point3D with(Point2D val);
+	int z();
 	Point3D withZ(int z);
+	Point3D with(Point2D val);
 }
 //END_POINT3D
-
 //BEGIN_ANIMAL
 interface Animal {
 	Point2D location();

### 2015-11-29 paper/sections/Translation.tex
Revise translation with @ObjWeak added.

@@
 
-\subsection{Translation Function}${}_{}$\\*
+\hl{Furthermore, we introduce \textbf{@ObjWeak}, which is only an annotation for illustrating the translation, hence is not implemented in Lombok.
+By this, the real translation is divided into two parts: \textbf{@Obj} automatically refines the return types only, and \textbf{@ObjWeak} generates the constructor method $\QM{of}$.
+In the source code, the only annotation \textbf{@Obj} combines both of the work and finishes the translation at a time.}
+
+\subsection{Translation Function}
 
@@
 }{
+\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
+\methods\ \methods' \cC
+}\\
+\InTextWith{\methods'=\otherMethod(\C_0,\methods)}
+\end{array}$
+
+\hl{To translate an interface annotated by \mixin, we add some other methods for type refinement, then add a new annotation \textbf{@ObjWeak} to the interface.}
+
+\noindent$\begin{array}{l}
+\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
+}{
 \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
@@
 }\\
-\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0) \cup \otherMethod(\C_0,\methods)}
+\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0)}
 \end{array}$
 
-To translate an annotated interface, we add the \Q@of@ method, and then we add some other methods.
+\hl{To translate an interface annotated by \textbf{@ObjWeak}, we add the $\QM{of}$ method.}
 However, first of all we check if the interface is valid for annotation:
@@
 In our simplified formalization we consider this to be just an error.
-In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}
+In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}\haoyuan{Not sure if it is $\QM{of}\notin\dom(\methods)$ or $\QM{of}\notin\dom(\C_0)$.
+It is an error if a static method overrides an instance method.}
 
@@
 
-\subsection{$\ofMethod$}${}_{}$\\*
+\subsection{$\otherMethod$}
+The definition of $\otherMethod(\C_0,\methods)$ is as follows:
+
+\haoyuan{I think $\QM{with#}\m\notin\dom(\methods)$ is not enough. The withM method with return type $I_0$ can also be defined already in a super interface. Like:
+\[\begin{array}{l}\QM{
+interface A}\ \oC\QM{int x(); B withX(int x);}\cC\\
+\mixinAnn\ \QM{interface B extends A}\ \oC\cC\end{array}\]}
+
+\noindent$\begin{array}{ll}
+\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{
+ \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
+\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
+\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0)}\\
+\InTextWith{
+\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+\mbox{ and } \QM_\m\notin\dom(\methods)}\\
+\end{array}$
+
+The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
+%A complete formalization would also generate the \Q@with@.
+We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
+To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
+ is required by $\C_0$, but is not already present in the methods \hl{directly declared in} $\C_0$.
+
+\noindent$\begin{array}{ll}
+\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
+\isField(\method)\ \mand\
+%}\\\InTextWith{
+\method=\mBody(\_,\C_0)
+}\\
+
+\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
+\mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)}\\
+
+%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
+%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
+%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+\end{array}$
+
+\subsection{$\ofMethod$}
 We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
@@
 
-The function $\fieldsFunc(\C_0)$ (formally defined later) denotes all the fields in the current interface.
+The function $\fieldsFunc(\C_0)$ (defined before) denotes all the fields in the current interface.
 For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
@@
 
-\subsection{Other auxiliary functions}${}_{}$\\*
+The auxiliary functions are defined below:\\*
 
@@
 
-As you can see above, \Q@with-@ and setter methods are generated if needed.
-We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$.
-Note that we do not need to check if its header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
-
-\noindent$\begin{array}{ll}
-\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{
- \isWith(\mBody(\QM{with#}\m, \C_0))}\\
-\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
-\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
-\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0))}\\
-\InTextWith{
-\C\ \m\QM{();}\in \fieldsFunc(\C_0)
-\mbox{ and } \QM_\m\notin\dom(\methods)}\\
-%&\C_0\ \QM{clone}\oR\cR\QM;\in
-%\otherMethod(\C_0)
-%\\&  \mbox{ iff }
-%\isClone(\mBody(\QM{clone}, \C_0))
-%\mbox{ and } \QM{clone}\notin\dom(\methods)\\
-\end{array}$
-
-Other methods that we need to generate in the interface are \Q@with-@ and setters. %\Q@clone@.
-%A complete formalization would also generate the \Q@with@.
-This is needed only if we need to refine the return type.
-To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
- is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.
-
-\noindent$\begin{array}{ll}
-\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
-\isField(\method)\mand\
-%}\\\InTextWith{
-\method=\mBody(\_,\C_0)
-}\\
-
-\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
-\mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
- \mand\ \mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
- \mand\ \mnot\ \specialName(\m)}\\
-
-%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
-%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
-%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-\end{array}$
+\subsection{Other features}
 

### 2015-11-29 paper/sections/Appendix.tex
Proof.

--- 2015-11-17 paper/sections/Appendix.tex
@@
 
-\subsection{Proof of LEMMA 1}\label{subsec:proof1}
+\textbf{THEOREM. }
+For a given $\II_0\ldots\II_n$ interface table such that
+$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
+if $\II_0$ has \textbf{@ObjWeak},
+$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
+then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$\\*
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.\\
+
+\noindent For convenience, assume that
+\begin{align*}
+\II_0\ &=\ \weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
+[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \method' \cC
+\end{align*}
+
+\subsection{LEMMA 1 and Proof}\label{subsec:lemma1}
+
 \textbf{LEMMA 1. }
-Assume that
-\begin{align*}
-\II_0\ &=\ \mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
-[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \methods' \cC
-\end{align*}
-If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.
+For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} and $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
 \begin{proof}
-It suffices to prove $\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\methods')$.
+The proof is based on induction. By the grammar shown in Figure~\ref{Grammar}, there are three cases for an arbitrary expression $e$:
 \begin{itemize}
-\item If $m\notin\dom(\methods')$, by the definition of $\mBody$, it is obtained from $\override$.
-The first argument of $\override$, namely $\methods\cup\methods'(m)$, is equal to $\methods(m)$, and the second argument is not changed as well, thus the result of $\override$ is not changed during translation.
-Hence $m\in\dom([\![\II_0]\!])$ iff $m\in\dom(\C_0)$.
-\item If $m\in\dom(\methods')$, we are to prove $m\in\dom([\![\II_0]\!])$.
-So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$.
-On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship.
-Hence $m\in\dom([\![\II_0]\!])$ as well.
+\item A variable or a field update. The type preservation for $e$ is ensured by induction.
+\item A method call (normal, static or super).
+Such a method won't be ``removed'' by the translation and is still there, and the types remain unchanged.
+The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any method call.
+\item An object creation.
+Adding the $\QM{of}$ method doesn't introduce unimplemented methods to an interface, moreover, the static method is not inheritable, hence after translation such an object creation still type checks and has the right type by induction.
 \end{itemize}
@@
 
-\subsection{Proof of LEMMA 2}\label{subsec:proof2}
+\subsection{LEMMA 2 and Proof}\label{subsec:lemma2}
+
 \textbf{LEMMA 2. }
-If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
+If $\II_0$ has \textbf{@ObjWeak} and is OK, then $[\![\II_0]\!]$ OK.
 \begin{proof}
-By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide it into two parts.
+
+By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide the proof into two parts.
 
@@
 
-Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.
-
-\haoyuan{How to more explicitly say the default methods are ok in the new interface table, so that it is sufficient to only prove the generated ofMethod is ok?}
+Since $\II_0$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}.
+It suffices to prove that it still holds for $\ofMethod(\C_0)$.
 
@@
 
-\haoyuan{The order of conditions in \rn{t-Obj} is not clear below.}
+\begin{itemize}
+\item All field initializations are type correct.
+By the definition of $\ofMethod(\C_0)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
+\item All method bodies are well-typed.
+    \begin{itemize}
+    \item Typing of the $i$-getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]
+        We know that $\C_i=\C^{\mh_i}$ since the $i$-th getter has its return type the same as the corresponding field $m_i$.
+    \item Typing of the \QM{with-} method of an arbitrary field $m_i$.
+By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C_0\QM{.of(}\es_i\QM{);\}}\]
+        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C_0.\QM{of}$ method.
+Hence \[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
+        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
+    \item Typing of the $i$-setter \QM_$m_i$. If the \QM_$m_i$ method is well-defined, it has the form
+        \[\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
+        By \rn{t-Update}, the assignment ``$\m_i\QM{= _val;}$'' is correct since $\m_i$ and $\QM{_val}$ have the same type $I_i$, and the return type is decided by \QM{this}.
+\[\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash\QM{this}\in \C_0\]
+        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
+        \haoyuan{Can we write $\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i, \C^{\mh_i} = \C_0 \vdash\QM{this}\in \C_0 = \C^{\mh_i}$?}\\
+        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C_0$?}
+    \end{itemize}
+\item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
+    \haoyuan{TODO}
+\item All abstract methods in the domain of $\C_0$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
+    Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$.
+But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
+\end{itemize}
+
+\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$
+Since $\II_0$ OK, we have $\dom(\II_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\method')$$
+This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so $\method'$ doesn't overlap with $\dom(\C_0)$.
+The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$, which finishes our proof.
+\end{proof}
+
+\subsection{Notes}
 
 \begin{itemize}
-\item For field typing,
-\[\Gamma(\QM_ m_i)=\C_i\subtype \C_i\]
-\item For method typing of getters,
-\[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i=\C^{\mh_i}\]
-\item For method typing of \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
-\[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0=C^{\mh_i}\]
-\item For method typing of setter methods, \haoyuan{what about the \QM{with} method?}
-\[\Gamma, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash\QM{this}\in \C_0=\C^{\mh_i}\]
 \item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
@@
  \end{itemize}
-\item To prove that such a created object indeed implements all the abstract methods in $\dom(\C_0)$, we simply refer to $\valid(\C_0)$, since it guarantees each abstract method $\method$ to satisfy $\isField$, $\isWith$ or $\isSetter$.
-But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
 \end{itemize}
-
-\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$$
-And actually it is guaranteed by \textbf{LEMMA 1}.
-
-\end{proof}
 
@@
 \begin{proof}
-\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK.
-On the other hand, if some $\II$ is not a subtype of $\II_0$, the generated code in translation has no way to affect the domain of $\II$, which finishes our proof.
+\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$,
+\marco{by the newLemma, we know that all its methods are still well typed, and } the generated code in translation has no way to affect the domain of $\II$,\marco{ so rule t-interf can still be applied,} which finishes our proof.\haoyuan{And the well-formedness.}
 \end{proof}

### 2015-11-29 paper/sections/Translation.tex
More.
Proof.

@@
 \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
-\methods\ \methods' \cC
+\methods\ \method' \cC
 }\\
-\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0)}
+\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \method'=\ofMethod(\C_0)}
 \end{array}$
@@
 
-\subsection{$\ofMethod$}
+\subsection{$\ofMethod$}\label{subsec:ofmethod}
 We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
@@
 
-The auxiliary functions are defined below:\\*
+The auxiliary functions are defined below. Note that we do not need to check if some header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
 

### 2015-11-30 paper/sections/Appendix.tex
Proof in appendix ready for comments.

@@
+\newpage
+
 \section{Appendix}\label{sec:appendix}
+
+\subsection{THEOREM}\label{subsec:theorem}
 
@@
     \begin{itemize}
-    \item Typing of the $i$-getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]
+    \item Typing of the $i$-th getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]
         We know that $\C_i=\C^{\mh_i}$ since the $i$-th getter has its return type the same as the corresponding field $m_i$.
@@
         We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
-    \item Typing of the $i$-setter \QM_$m_i$. If the \QM_$m_i$ method is well-defined, it has the form
+    \item Typing of the $i$-th setter \QM_$m_i$. If the \QM_$m_i$ method is well-defined, it has the form
         \[\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
@@
 \item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
-    \haoyuan{TODO}
+    \begin{itemize}
+    \item For the $i$-th getter $m_i$,
+        \begin{align*}
+        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C_0)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; = \mBody(m_i,\C_0)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C_0)
+        \end{align*}
+    \item For the $\QM{with#}\m_i$ method,
+        \begin{align*}
+        &\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
+        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
+        \mimply\hspace{.2in}& \isWith(\mBody(\QM{with#}\m_i,\C_0),\C_0)\\
+        \mimply\hspace{.2in}& \C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM{with#}\m_i,\C_0)
+        \end{align*}
+    \item For the $i$-th setter $\QM_\m_i$,
+        \begin{align*}
+        &\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
+        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
+        \mimply\hspace{.2in}& \isSetter(\mBody(\QM_\m_i,\C_0),\C_0)\\
+        \mimply\hspace{.2in}& \C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM_\m_i,\C_0)
+        \end{align*}
+    \end{itemize}
 \item All abstract methods in the domain of $\C_0$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
@@
 This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so $\method'$ doesn't overlap with $\dom(\C_0)$.
-The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$, which finishes our proof.
+The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$.
+Also note that after translation, there are still no methods with conflicted names, since the \QM{of} method was previously not in the domain, hence $[\![\II_0]\!]$ is well-formed, which finishes our proof.
 \end{proof}
 
-\subsection{Notes}
-
-\begin{itemize}
-\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
- \begin{itemize}
- \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
- \item For \QM{with-} methods,\marco{ use if-then or other english forms instead of the logic arrow}
-  \[\begin{array}{l}
-  \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
-  \end{array}\Rightarrow \isWith\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
- \item For setters,
-  \[\begin{array}{l}
-  \mBody(\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
-  \end{array}\Rightarrow \isSetter\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
- \end{itemize}
-\end{itemize}
-
 \subsection{Proof of THEOREM}\label{subsec:proof3}
-\textbf{THEOREM. }
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
 \begin{proof}
-\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$,
-\marco{by the newLemma, we know that all its methods are still well typed, and } the generated code in translation has no way to affect the domain of $\II$,\marco{ so rule t-interf can still be applied,} which finishes our proof.\haoyuan{And the well-formedness.}
+\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, for any $\II\ne\II_0$, by \textbf{LEMMA 1}, we know that all its methods
+are still well-typed, and the generated code in translation of \textbf{@ObjWeak} is only a static method \QM{of}, which has no way to affect the domain
+of $\II$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
 \end{proof}

### 2015-11-30 paper/sections/CaseStudy.tex
code for expression problem updated.

@@
 \paragraph{Initial System}
+\yanlin{code figure updated.}
 In the formulation of the EP there is an initial system that models
 arithmetic expressions with only literals and addition, and an initial 
-operation:
+operation \texttt{eval} for expression evaluation.
 
-\begin{lstlisting}[multicols=2]
+\noindent As shown in Figure~\ref{fig:ep}, \texttt{Exp} is the common
+super-interface with an evaluation operation \texttt{eval()}
+inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend interface
+\texttt{Exp} with default implementations for the \texttt{eval} operation. The
+number field \texttt{x} of a literal is represented as a getter method
+\texttt{x()} and expression fields (\texttt{e1} and \texttt{e2}) of an addition
+as getter methods \texttt{e1()} and \texttt{e2()}.
+
+\begin{figure}
+\begin{tabular}{l|l}
+\begin{minipage}{8cm}
+%\vspace{-90pt}
+\begin{lstlisting}[]
 interface Exp { int eval(); }
-
 @Obj interface Lit extends Exp {
@@
 \end{lstlisting}
-
-% \begin{multicols}{2}
-% \lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
-% \end{multicols}
-
-\noindent \texttt{Exp} is the common super-interface with an evaluation operation
-\texttt{eval()} inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend
-interface \texttt{Exp} with default implementations for the \texttt{eval}
-operation. The number field \texttt{x} of a literal is represented as a getter
-method \texttt{x()} and expression fields (\texttt{e1} and \texttt{e2}) of an addition as
-getter methods \texttt{e1()} and \texttt{e2()}.
+\end{minipage}
+&
+\begin{minipage}{8cm}
+\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\end{minipage}
+\end{tabular}
+\caption{The Expression Problem. Initial system on the left, and code for adding
+  print operation on the right. }\label{fig:ep}
+\end{figure}
 
@@
 It turns out that, this can be done easily, using our object interfaces approach. 
-The following code shows how to add the new operation \texttt{print}.
-
-\marco{ use the twoc colums here too}
-\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
-%\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
-%\bruno{please indent code properly (the getters are not aligned with default))}
+The code in Figure~\ref{fig:ep} shows how to add the new operation \texttt{print}.
 

### 2015-11-30 paper/sections/CaseStudy.tex
compressing text

@@
 
-The \emph{Expression Problem} (EP)~\cite{wadler98expression}
-is a well-known problem about modular extensibility issues in software
-evolution. Recently, a new solution~\cite{eptrivially}, \marco{repharesed} using only
-covariant type refinement, has been shown.
-When such solution is modelled with interfaces and default methods, it can
-even provide independent extensibility: the ability to
-assemble a system from multiple, independently developed, extensions.
-Unfortunatelly, the required instantiation code makes a plain Java solution verbose and cumbersome to use.
-The \mixin annotation is enough to remove the boilerplate code, making the presented approach very appealing.
-% \bruno{We should try to do some space saving in here. You could, for example 
-% combine the code for the initial system with the code for a new operation in a figure/table 
-% with 2 colums. Then just refer to the Table/Figure in the text.}
+The \emph{Expression Problem} (EP)~\cite{wadler98expression} is a well-known
+problem about modular extensibility issues in software evolution. Recently, a
+new solution~\cite{eptrivially}, using only covariant type refinement, has been
+shown.  When such solution is modelled with interfaces and default methods, it
+can even provide independent extensibility: the ability to assemble a system
+from multiple, independently developed, extensions.  Unfortunatelly, the
+required instantiation code makes a plain Java solution verbose and cumbersome
+to use. The \mixin annotation is enough to remove the boilerplate code, making
+the presented approach very appealing.
 
 \paragraph{Initial System}
-\yanlin{code figure updated.}
 In the formulation of the EP there is an initial system that models
@@
 \texttt{select}, \texttt{from} and \texttt{where} clauses:
-\lstinputlisting[linerange=6-11]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
+\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
 % \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
@@
 
-\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
+\lstinputlisting[linerange=31-31]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
 
@@
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
-\lstinputlisting[linerange=15-18]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+\lstinputlisting[linerange=14-16]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
 This way, when a \texttt{ExtendedDatabase} query created, 
@@
 
-\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\lstinputlisting[linerange=37-37]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
 

### 2015-11-30 paper/sections/Introduction.tex
compressing text

@@
 To evaluate the usefulness of object interfaces, we illustrate 3
-applications and case studies. The first application is a simple 
+applications (case studies). The first application is a simple 
 solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
 extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
-application is to show how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
-can be easily defined using object interfaces. Finally, the last
+application is shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
+can be easily defined using object interfaces. The last
 application is a larger case study for a simple Maze game implemented with 
@@
 to an existing implementation~\cite{bono14} using plain Java 8. 
-Noteworthy, is the fact that all applications are implemented 
+Noteworthy, all applications are implemented 
 without defining a single class!
@@
 In summary, the contributions of this paper are:
-
 \begin{itemize}

### 2015-11-30 paper/sections/Overview.tex
code for expression problem updated.
compressing text

@@
 To use \texttt{Horse}, \texttt{Bird} and \texttt{Pegasus}, some
-objects must be instantiated first. A first problem with using
+objects must be created first. A first problem with using
 interfaces to model the animal system is simply that interfaces 
-cannot directly be instantiated: a class is needed. Therefore, we have
-to first create some classes, such as:
+cannot be directly instantiated. Classes are needed for instantiation, such as:
 
@@
 
-\noindent in order to create animal objects. With those classes, 
-a Pegasus animal, can be created using the class constructor:
+\noindent to create animal objects. Now a \texttt{Pegasus} animal can be created
+using the class constructor:
 
@@
 
-\noindent There are a couple of annoyances here. Firstly, the sole
-purpose of the classes is to provide a way to instantiate the
+\noindent There is a lot of annoyance here. Firstly, the sole
+purpose of classes is to provide a way to instantiate
 objects. Although (in this case) it takes only one line code to
@@
 interface Pegasus extends Horse, Bird { // generated code not visible to users
-	static Pegasus of() {return new Pegasus() {};}
-}
+	static Pegasus of() {return new Pegasus() {};}}
 \end{lstlisting}
@@
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-Two dimensional points should keep track of their coordinates. 
-The usual approach to model points in Java would be to use a class 
-with fields for the coordinates. However here we will illustrate how 
-points are modelled with interfaces, providing a simple 
-example of how to model immutable state with interfaces. Since Java disallows fields 
-inside interfaces, the state is modelled using abstract (getter) methods:
+Two dimensional points should keep track of their coordinates.  The usual
+approach to model points in Java would be to use a class with fields for the
+coordinates. However here we will illustrate how points are modelled with
+interfaces. Since Java disallows fields inside interfaces, states are modeled
+using abstract (getter) methods:
 
@@
 \noindent Unfortunatelly, creating a new point object is quite cumbersome, even
-when using anonymous classes:
+with anonymous classes:
 
@@
 
-\noindent However programmers are not required to use this cumbersome syntax for
-every object allocation. As programmers do, for ease or reuse, the
-boring long repetitive code can be encapsulate in a method. A
-generalization of the \texttt{of} static factory
-method is appropriate in this case:
+\noindent However it is not required to use this cumbersome syntax for every
+object allocation. As programmers do, for ease or reuse, the boring
+repetitive code can be encapsulated in a method. A generalization of the
+\texttt{of} static factory method is appropriate in this case:
 \begin{lstlisting}
@@
   static Point2D of(int x, int y){return new Point2D(){
-    public int x(){return x;} public int y(){return y;}};}
-}
+    public int x(){return x;} public int y(){return y;}};}}
 \end{lstlisting}
@@
 This obvious constructor code can be automatically generated by our
-\mixin annotation.  By annotating the interface \Q@Point2D@, the
-annotation will generate a variation of the shown static method
-\texttt{of}, mimicking the functionality of a simple minded
-constructor: By looking at the methods that need implementations it
+\mixin annotation.  By annotating the interface \Q@Point2D@, a variation of the shown static method
+\texttt{of} will be generated, mimicking the functionality of a simple-minded
+constructor: by looking at the methods that need implementation it
 first detects what are the fields, then generates an \Q@of@ method with
-one argument for each of them. That is, we can just write
+one parameter for each of them. That is, we can just write
 
@@
   int x();    int y(); // getters
-  Point2D withX(int val);    Point2D withY(int val); // with methods
-}
+  Point2D withX(int val);    Point2D withY(int val); // with methods }
 \end{lstlisting}
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 \subsection{Object Interfaces and Subtyping}
-
-%\subsection{\Q@Bird@:  field type refinement}
-
 \Q@Bird@s are \Q@Animal@s, but while \Q@Animal@s only need 2D
@@
 be refined. Such kind of refinement usually poses a challenge
-with typical class-based approaches. Fortunatelly, with object interfaces,
+in typical class-based approaches. Fortunatelly, with object interfaces,
 we are able to provide a simple and effective solution to that problem.
@@
 
-\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 
-\paragraph{Generated Boilerplate}
+\paragraph{Generated boilerplate}
 Just to give a feeling on how much boring code \mixin is generating,
@@
                 if(val instanceof Point3D){return (Point3D)val;}
-        return Point3D.of(val.x(),val.y(),this.z());
-        }      };    }  }
+                return Point3D.of(val.x(),val.y(),this.z()); }};}}
 \end{lstlisting}
@@
 
-\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=51-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=51-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=59-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=66-69]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=73-81]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=85-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-11-30 paper/sections/RelatedWork.tex
compressing text

@@
 the most restricted one (excluding states, instantiation, etc).
-\bruno{How about Malayeri and Aldrich's paper? shouldn't that be discussed?}
-\yanlin{Newly added discussion on CZ. please double check.}
 
@@
 constructor calls to virtual superclasses, which would potentially cause serious
-semantic errors. \bruno{What happens in our approach for the same
-  case?}\yanlin{explained below please double-check.} In our
-approach, \mixin annotation processor has the full control of object
-initialization mechanism, and this mechanism is clean and transparent to
-users. Moreover, if not satisfied with the default generated \texttt{of} method,
-customized factory methods are also allowed to be defined.
+semantic errors. In our approach, \mixin annotation processor has the full
+control of object initialization mechanism, and this mechanism is clean and
+transparent to users. Moreover, if not satisfied with the default generated
+\texttt{of} method, customized factory methods are also allowed to be defined.
 
@@
 
-\bruno{rrevise latter?}
-However, in pure Java, there is no mechanism for creating objects in
-interfaces. Also, our approach supports proper constructor mechanism.
+% However, in pure Java, there is no mechanism for creating objects in
+% interfaces. Also, our approach supports proper constructor mechanism.
 
@@
 \paragraph{Traits vs Object Interfaces}
-We consider object interfaces to be an alternative to traits or mixins.
-In the trait model two concepts (traits and classes) coexist and
-cooperate. Some authors~\cite{bettini2013pure}\yanlin{There are many trait related papers of ferruccio
-  damiani, please check whether citing this one is ok.} 
-see this as good language design
-fostering good software development by helping programmers to think
-about the structure of their programs.  However, other authors see the
-need of two concepts and the absence of state as drawbacks of this
-model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
-the same time they provide factory methods for instantiation and
-support state.  Our approach promotes the use of interfaces instead of
-classes, in order to rely on the modular composition offered
-by interfaces. Since Java was designed for classes, a direct class-less
-programming style is verbose and feels unnatural. However, annotation
-driven code generation is enough to overcome this difficulty, and the
-resulting programming style encourages modularity, composability and
-reusability, by keeping a strong object oriented feel. In that sense,
-we promote object interfaces as being both units of reusable code and
-object factories. Our practical experience is that, in Java,
-separating the two notions leads to alot of boilerplate code, and is
-quite limiting when multiple inheritance with state is required. The
-use of abstract state operations avoids the key difficulties
-associated with multiple inheritance and state, while still being quite expressive. 
-Moreover the ability to support constructors 
-adds additional expressivity, which is not available in approaches such 
-as Scala's traits/mixins.
+We consider object interfaces to be an alternative to traits or mixins.  In the
+trait model two concepts (traits and classes) coexist and cooperate. Some
+authors~\cite{bettini2013pure} see this as good language design fostering good
+software development by helping programmers to think about the structure of
+their programs.  However, other authors see the need of two concepts and the
+absence of state as drawbacks of this model~\cite{malayeri2009cz}.  Object
+interfaces are units of reuse, and at the same time they provide factory methods
+for instantiation and support state.  Our approach promotes the use of
+interfaces instead of classes, in order to rely on the modular composition
+offered by interfaces. Since Java was designed for classes, a direct class-less
+programming style is verbose and feels unnatural. However, annotation driven
+code generation is enough to overcome this difficulty, and the resulting
+programming style encourages modularity, composability and reusability, by
+keeping a strong object oriented feel. In that sense, we promote object
+interfaces as being both units of reusable code and object factories. Our
+practical experience is that, in Java, separating the two notions leads to alot
+of boilerplate code, and is quite limiting when multiple inheritance with state
+is required. The use of abstract state operations avoids the key difficulties
+associated with multiple inheritance and state, while still being quite
+expressive.  Moreover the ability to support constructors adds additional
+expressivity, which is not available in approaches such as Scala's
+traits/mixins.
 

### 2015-11-30 UseMixinLombok/src/jooq/DatabaseTest.java

@@ -7,15 +7,13 @@ import lombok.Obj;
 	String select(); Database select(String select);
 	String from(); 	 Database from(String from);
 	String where();  Database where(String where);
-	static Database of() {return of("", "", "");}
-}
+	static Database of() {return of("", "", "");}}
 //END_FLUENT_DATABASE
 
 //BEGIN_FLUENT_DATABASE_EXT
 @Obj interface ExtendedDatabase extends Database {
 	String orderBy(); ExtendedDatabase orderBy(String orderBy);
-	static ExtendedDatabase of() {return of("", "", "","");}
-}
+	static ExtendedDatabase of() {return of("", "", "","");}}
 //END_FLUENT_DATABASE_EXT
 
 public interface DatabaseTest {

### 2015-11-30 UseMixinLombok/src/test/TestAnimal.java

@@ -56,32 +56,22 @@ interface Point3D extends Point2D {
 }
 //END_POINT3D
 //BEGIN_ANIMAL
-interface Animal {
-	Point2D location();
-	void location(Point2D val);
-}
+interface Animal {Point2D location(); void location(Point2D val);}
 //END_ANIMAL
 
 //BEGIN_HORSE
-@Obj
-interface Horse extends Animal {
-	default void run() {location(location().withX(location().x() + 20));}
-}
+@Obj interface Horse extends Animal {
+	default void run() {location(location().withX(location().x() + 20));}}
 //END_HORSE
 
 //BEGIN_BIRD
-@Obj
-interface Bird extends Animal {
-	Point3D location();
+@Obj interface Bird extends Animal {
+	Point3D location();	
 	void location(Point3D val);
 	default void location(Point2D val) { location(location().with(val));}
-	default void fly() {
-		location(location().withX(location().x() + 40));
-	}
-}
+	default void fly() {location(location().withX(location().x() + 40));}}
 //END_BIRD
 
 //BEGIN_PEGASUS
-@Obj
-interface Pegasus extends Horse, Bird {}
+@Obj interface Pegasus extends Horse, Bird {}
 //END_PEGASUS

### 2015-12-02 paper/sections/Overview.tex
revising text

@@
 \begin{figure}
-\begin{tabular}{l|l}
+\begin{tabular}{c|c}
 \includegraphics[height=4cm]{PegasusDetail.pdf}\hspace{20pt} &
-\begin{minipage}{8cm}
+\begin{minipage}{7cm}
 \vspace{-90pt}
 \lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+%basicstyle=\ttfamily\scriptsize
 \end{minipage}
@@
 \paragraph{\Q@Point2D@ with Object Interfaces}
-This obvious constructor code can be automatically generated by our
-\mixin annotation.  By annotating the interface \Q@Point2D@, a variation of the shown static method
-\texttt{of} will be generated, mimicking the functionality of a simple-minded
-constructor: by looking at the methods that need implementation it
-first detects what are the fields, then generates an \Q@of@ method with
-one parameter for each of them. That is, we can just write
+This obvious constructor code can be automatically generated by our \mixin
+annotation.  By annotating the interface \Q@Point2D@, a variation of the shown
+static method \texttt{of} will be generated, mimicking the functionality of a
+simple-minded constructor: by looking at the abstract methods it first detects
+what are the fields, then generates an \Q@of@ method with one parameter for each
+of them. That is, we can just write
 
@@
 \noindent More precisely, a field or factory parameter is generated for every
-no-args method requiring implementation (except for methods whose name have
-special meaning).  An example of code using \Q@Point2D@ is:
-
+abstract method that takes no parameters (except for methods with special
+names). An example of using \Q@Point2D@ is:
 \begin{lstlisting}
@@
 \end{lstlisting}
-
 \noindent where we return a new point, using \Q@42@ as x-coordinate,
@@
 \paragraph{\texttt{with} methods in Object Interfaces}
-The pattern of creating a new piece of data by reusing most
-information from an old piece of data is very common
-when programming with immutable data-structures; it is so common that
-it is supported in our code generation as \Q@with@ methods,
-that is:
+The pattern of creating a new object by reusing most information from an old
+object is very common when programming with immutable data-structures. It is
+supported in our code generation as \Q@with@ methods, for example:
 \begin{lstlisting}
@@
 
-\noindent Using \texttt{with} methods, the point \texttt{p} could have
-been created instead using:
+\noindent Using \texttt{with} methods, the point \texttt{p} can be created instead using:
 
@@
 \noindent If there is a large number of fields, \texttt{with} methods
-will save programmers from writting large amounts of tedious code that
-simply copies the values of fields. 
+will save programmers from writing large amount of tedious code that
+simply copies field values. 
 \begin{comment}
@@
 \end{comment}
-Moreover, if the programmer wants a different implementation, they
-just provide it using a \Q@default@ method. For example:
+Moreover, if the programmer wants a different implementation, they may 
+just provide new \texttt{with} methods using \Q@default@:
 \begin{lstlisting}
@@
 \end{lstlisting}
-
-\noindent can be used by programmers to define their own behaviour for 
-the \texttt{with} methods.
 
@@
 \paragraph{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
-Two dimensional points are mathematical entities, thus we chosen to
-use immutable data structure to model them. However animals are real
-world entities, and when an animal moves, it is the same animal that
-now has a different location. We model this with mutable state.
+2D points are mathematical entities, thus we choose immutable data structure to
+model them. However animals are real world entities, and when an animal moves,
+it is the \emph{same} animal with a different location. We model this with
+mutable state.
 
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=57-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\noindent Here we declare abstract getter and setter for the mutable ``field'' \Q@location@.
-The \mixin annotation is not used. This is morally
-equivalent to an abstract class in full Java; and there is no
-convenient way to instantiate it.
-For \texttt{Horse}, the \mixin annotation is used and a concrete implementation of \texttt{run()} method
-is defined using a default method, which also further illustrates the convenience of \emph{with} methods.
+\noindent Here we declare abstract getter and setter for the mutable ``field''
+\Q@location@.  Without the \mixin annotation, there is no convenient way to
+instantiate \texttt{Animal}.  For \texttt{Horse}, the \mixin annotation is used
+and an implementation of \texttt{run()} method is defined using \Q@default@,
+which further illustrates the convenience of \emph{with} methods:
 
-\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=61-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\noindent Creating and using a horse animal is quite simple:
+\noindent Creating and using \texttt{Horse} is quite simple:
 
@@
 
-\paragraph{Summary} Dealing with state (whether mutable or immutable)
-in object interfaces relies on a notion of abstract state, where only
-methods to interact with state are available to users. Object
-interfaces provides support for four different types abstract state
-operations, which are summarized in Figure~\ref{fig:abstractstate}.
-The abstract state operations are determined by naming conventions and the
-types of the methods. Fluent setters are a variant of mutable setters,
-and are discussed in the case study in Section~\ref{sec:dsls}.
+\paragraph{Summary} Dealing with states (mutable or not) in object interfaces
+relies on a notion of abstract state, where only methods that interact with
+states are available to users. Object interfaces provide support for four
+different types of abstract state operations, which are summarized in
+Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
+naming conventions and the types of the methods. Fluent setters are a variant of
+mutable setters, and are discussed in detail in Section~\ref{sec:dsls}.
 
@@
 in typical class-based approaches. Fortunatelly, with object interfaces,
-we are able to provide a simple and effective solution to that problem.
+we are able to provide a simple and effective solution.
 
 \paragraph{Unsatisfactory class-based solutions to field type refinement}
-In Java if we define an animal class with a field we have a set of unsatisfactory options in front of us:
+In Java if we define an animal class with a field we have a set of
+unsatisfactory options in front of us:
+
 \begin{itemize}
-\item Define a \Q@Point3D@ field in \Q@Animal@:
-this is bad since all animals would require more that is needed, and also it may requires the programmer to predict the future, or it may require to adapt the old code to accommodate for new evolutions.
-\item Define a \Q@Point2D@ field in \Q@Animal@ and define an extra \Q@int z@ field in \Q@Bird@.
-This solution is very ad-hoc, requires to basically duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside of \mbox{\Q@Bird@}.
-Again, there are many reasons this would be bad, the most dramatic is that it would not scale to a scenario when the programmer of \Q@Bird@ and the programmer of \Q@Point3D@ are different.
-\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects in the field and cast
-the value out of the \Q@Point2D@ field to \Q@Point3D@ when implementing the overridden getter.
-This solution scales to the multiple programmers approach, but requires ugly casts and can be implemented in a wrong way leading to bugs.
+
+\item Define a \Q@Point3D@ field in \Q@Animal@: this is bad since all animals
+  would require more than needed.
+  %Also it requires the programmer to predict the future, or 
+  Also it requires adapting the old code to accommodate for new evolutions.
+
+\item Define a \Q@Point2D@ field in \Q@Animal@ and define an extra \Q@int z@
+  field in \Q@Bird@.  This solution is very ad-hoc, requires to basically
+  duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside \Q@Bird@.
+  %Again, there are many reasons this would be bad, 
+  The most dramatic criticism is that it would not scale to a scenario when
+  \Q@Bird@ and \Q@Point3D@ are from different programmers.
+
+\item Redefine getters and setters in \Q@Bird@, always put \Q@Point3D@ objects
+  in the field and cast the value out of the \Q@Point2D@ field to \Q@Point3D@
+  when implementing the overridden getter.  This solution scales to the multiple
+  programmers approach, but requires ugly casts and can be implemented in a
+  wrong way leading to bugs.
 \end{itemize}
@@
 \paragraph{Field type refinement with object interfaces}
-Object interfaces address the challange of type-refinement as follows:
+Object interfaces address the challenge of type-refinement as follows:
 \begin{itemize}
-\item by covariant method overriding we refine the type of the location field to \texttt{Point3D};
-\item by \emph{overloading} we define a new setter for the location field with the more precise type;
-\item using a default method we provide an implementation for the setter called 
-with the old signature.
+\item by \emph{covariant method overriding} the return type of
+  \texttt{location()} is refined to \texttt{Point3D};
+\item by \emph{overloading} a new setter for location is defined with a more
+  precise type;
+\item we provide a \Q@default@ setter with the old signature.
 \end{itemize}
 
-\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=66-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 \paragraph{\Q@Point3D@ and properties updater}
-To implement the old setter in a convenient way, \mixin supports one
-last type of operations: property updater \texttt{with}
-methods. Unlike the \texttt{withX} (where \texttt{X} stands for a
-field name) methods presented so far, property updaters take several
-fields at once, contained in an interface, and copy those fields into
-fields of another interface.
 The \Q@Point3D@ interface is defined as follows:
-\lstinputlisting[linerange=51-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+
+\lstinputlisting[linerange=51-54]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+
 \noindent By using the new \Q@with@ method we may use the information for
@@
 
+\Q@with-@ methods (functionally) update a field at a time.  This can be
+inefficient, and sometime hard to maintain.  Often we want to update multiple
+fields simultaneously, for example using another object as source.  Following
+this idea, the method \Q@with(Point2D)@ is an example of a (functional)
+properties updater: it takes a certain type of object and returns a copy of the
+current object with all fields that match the fields of parameter type
+updated. The idea is that we want the result to be still like \Q@this@, but
+modified as similar as possible to the parameter.
+
+% To implement the old setter in a convenient way, \mixin supports one
+% last type of operations: property updater \texttt{with}
+% methods. Unlike the \texttt{withX} (where \texttt{X} stands for a
+% field name) methods presented so far, property updaters take several
+% fields at once, contained in an interface, and copy those fields into
+% fields of another interface.
+
 \paragraph{Generated boilerplate}
-Just to give a feeling on how much boring code \mixin is generating,
-we show the generated code for the \texttt{Point3D} interface 
-Figure~\ref{fig:boilerplate}. The generated code is very repetitive,
-writing such code by hand can easily induce bugs, for example a
-distracted programmer may swap the arguments of one of the many calls of
-\Q@Point3D.of@.
+
 \begin{figure}
-\begin{lstlisting}
-interface Point3D extends Point2D{
-    Point3D withX(int val); Point3D withY(int val); Point3D withZ(int val);
-    Point3D with(Point2D val);
-    public static Point3D of(int _x, int _y, int _z){
-        int x=_x;    int y=_y;    int z=_z;
-        return new Point3D(){
-            public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
-            public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
-            public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
-            public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
-            public Point3D with(Point2D val){
-                if(val instanceof Point3D){return (Point3D)val;}
-                return Point3D.of(val.x(),val.y(),this.z()); }};}}
-\end{lstlisting}
+\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
 \caption{Generated boilerplate code.}
@@
 \end{figure}
-Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
-\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected.
-If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
-Again, the philosophy is that if the programmer provides something directly, \mixin does not touch it.
-\Q@with-@ methods (functionally) update a field/property at a time.
-This can be inefficient, and sometime hard to maintain.
-Often we want to update many fields at the same time, for example using another object as source.
-Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
-it takes a certain type and in the current object update all field in that type that match fields in the current type.
-The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
-The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test.
-The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.
+
+Just to give a feeling on how much boring code \mixin is generating, we show the
+generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. The
+generated code is very mechanical, writing such code by hand is error-prune. For
+example a distracted programmer may swap the arguments of calls to
+\Q@Point3D.of@.  Note how \Q@with-@ methods are automatically refined in their
+return type, so that code like \Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@
+will be accepted and behave as expected. If the programmer wishes to suppress
+this behaviour and keep the signature as it was, it is sufficient to redefine
+the \Q@with-@ methods in the new interface repeating the old signature.  Again,
+the philosophy is that if the programmer provides something directly, \mixin
+does not touch it.  The cast in \Q@with(Point2D)@ is trivially safe since it is
+guarded by an \Q@instanceof@ test. The idea is that if the parameter is a
+subtype of the current exact type, then we can just return the parameter, as
+something that is just ``more'' than \Q@this@.
 
@@
 
-\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=51-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=51-54]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=59-59]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=57-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=63-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=61-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=68-72]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=66-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=76-76]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-12-02 UseMixinLombok/src/test/TestAnimal.java

@@ -48,12 +48,10 @@ interface Point2D {
 //END_POINT2D
 
 //BEGIN_POINT3D
-@Obj
-interface Point3D extends Point2D {
+@Obj interface Point3D extends Point2D {
 	int z();
 	Point3D withZ(int z);
-	Point3D with(Point2D val);
-}
+	Point3D with(Point2D val); }
 //END_POINT3D
 //BEGIN_ANIMAL
 interface Animal {Point2D location(); void location(Point2D val);}
@@ -75,3 +73,22 @@ interface Animal {Point2D location(); void location(Point2D val);}
 //BEGIN_PEGASUS
 @Obj interface Pegasus extends Horse, Bird {}
 //END_PEGASUS
+
+class ManualCode {
+//BEGIN_GENERATED_POINT3D
+interface Point3D extends Point2D{
+    Point3D withX(int val); Point3D withY(int val); Point3D withZ(int val);
+    Point3D with(Point2D val);
+    public static Point3D of(int _x, int _y, int _z){
+        int x=_x; int y=_y; int z=_z;
+        return new Point3D(){
+            public int x(){return x;} public int y(){return y;} public int z(){return z;}
+            public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
+            public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
+            public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
+            public Point3D with(Point2D val){
+                if(val instanceof Point3D){return (Point3D)val;}
+                return Point3D.of(val.x(),val.y(),this.z()); }};}}
+//END_GENERATED_POINT3D
+
+}

### 2015-12-02 paper/sections/Implementation.tex
Game example
small stuff

@@
-\section{Implementation with Lombok}
+\section{Implementation}
 
-Our implementation is based on extension to Lombok. The Lombok
+Our implementation is based on an extension to Lombok. The Lombok
 project~\cite{lombok} is a Java tool that aims at removing (or
-reducing) Java boilerplate code as far as possible, via
+reducing) Java boilerplate code via
 annotations. There are a number of annotations provided by the

### 2015-12-02 paper/sections/RelatedWork.tex

@@
 
-There are also proposals for extending Java (before Java8) with
-traits. For example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by
-Liquori et al. extends the calculus of Featherweight Java
-(FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
-trait-based inheritance in Java.  Except for few, mostly syntactic
-details, their work can be emulated/rephrased with Java8 interfaces.
-There are also extensions to the original trait model, which have operations 
-that default methods and interfaces cannot model, such as method renaming
-(as in [Reppy2006]), which breaks structural subtyping.
+There are also proposals for extending Java (before Java8) with traits. For
+example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends
+the calculus of Featherweight Java (FJ)~\cite{Igarashi01FJ} with
+statically-typed traits, adding trait-based inheritance in Java.  Except for
+few, mostly syntactic details, their work can be emulated/rephrased with Java8
+interfaces.  There are also extensions to the original trait model, which have
+operations that default methods and interfaces cannot model, such as method
+renaming in \cite{reppy2006foundation}, which breaks structural subtyping.
 
@@
 trait model two concepts (traits and classes) coexist and cooperate. Some
-authors~\cite{bettini2013pure} see this as good language design fostering good
+authors~\cite{BettiniDSS13} see this as good language design fostering good
 software development by helping programmers to think about the structure of

### 2015-12-02 paper/sections/Translation.tex

@@
 
+Two translation functions for the two parts are presented as follows:
+
 \noindent$\begin{array}{l}
@@
 
-\hl{To translate an interface annotated by \mixin, we add some other methods for type refinement, then add a new annotation \textbf{@ObjWeak} to the interface.}
+\noindent\hl{To translate an interface annotated by \mixin, we add some other methods for type refinement, then add a new annotation \textbf{@ObjWeak} to the interface.}
 
@@
 
-\hl{To translate an interface annotated by \textbf{@ObjWeak}, we add the $\QM{of}$ method.}
+\noindent\hl{To translate an interface annotated by \textbf{@ObjWeak}, we add the $\QM{of}$ method.}
 However, first of all we check if the interface is valid for annotation:

### 2015-12-02 paper/sections/Overview.tex
less space around code

@@
 
-\lstinputlisting[linerange=11-14]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+\lstinputlisting[linerange=9-15]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
 
@@
 
-\paragraph{Generated boilerplate}
-
 \begin{figure}
@@
 
+\paragraph{Generated boilerplate}
 Just to give a feeling on how much boring code \mixin is generating, we show the

### 2015-12-02 UseMixinLombok/src/test/TestAnimal.java

@@ -48,8 +48,7 @@ interface Point2D {
 //END_POINT2D
 
 //BEGIN_POINT3D
-@Obj
-interface Point3D extends Point2D {
+@Obj interface Point3D extends Point2D {
 	int z(); 
 	Point3D withZ(int z);
 	Point3D with(Point2D val); }

### 2015-12-02 paper/sections/Formalization.tex
Revise translation with @ObjWeak added.
revising text

--- 2015-11-25 paper/sections/Formalization.tex
@@
-\section{Interaction of interface methods with interface composition}
-From Java8 interfaces can have three type of methods:
-abstract methods, default methods, static methods.
+\section{Interaction of Interface Methods with Interface Composition}
+From Java 8 interfaces can have three type of methods: abstract, default, and
+static methods.
 \begin{itemize}
-\item Static methods are handled in a very clean way: they are visible only on the interface they are explicitly declared. This means that 
-the following code is ill-typed.
+
+\item Static methods are handled in a very clean way: they are visible only in
+  the interface they are explicitly defined. This means the following code
+  is ill-typed.
 \begin{lstlisting}
@@
 interface B0 extends A0 {}
-...
-B0.m()//ill typed
+... B0.m()//ill typed
 \end{lstlisting}
-Note how this is different w.r.t. the way static methods are handled in classes.
-In this way static methods have simply no interaction with interface composition (\Q@extends@ or \Q@implements@).
+This is different from the way static methods are handled in classes. Here
+static methods have simply no interaction with interface
+composition (\Q@extends@ or \Q@implements@).
 
-\item Abstract methods composition is accepted when there is one that is the most specific.  For example in 
-method \texttt{Integer m()} is visible in \Q@C1@.
-
+\item Abstract methods composition is accepted when exist a most specific one.
+  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
 \begin{lstlisting}
@@
 \item Default methods conflict with any other default or abstract method. For
-example the following code is rejected due to conflicting methods.
+  example the following code is rejected due to methods conflict.
 \begin{lstlisting}
@@
 \end{lstlisting}
-Note how this is in contrast with what happens in most trait models, where \Q@D2@ would be accepted, and the implementation in \Q@A2@ would be part of the behaviour of \Q@D2@.
+Note how this is different from what happens in most trait models, where \Q@D2@
+would be accepted, and the implementation in \Q@A2@ would be part of the
+behaviour of \Q@D2@.
 
-\item The method in the current interface wins over any methods defined in its
-super-interfaces, provided that the method conform to the subtype of all methods
-in its super-interfaces, i.e., the method is the most specific one.
-This also override rejection due to conflicting methods. For example, the following code is accepted, but would be rejected (see before) if the method \Q@m@ was not redefined.
+\item A method in the current interface wins over any method in its
+super-interfaces, provided that the method%  conforms to the subtype of all methods
+% in its super-interfaces, i.e., the method
+is the most specific one. This method also overrides conflict due to
+inheritance. For example, the following code is accepted, but would be rejected
+(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
+\texttt{E3}.
 \begin{lstlisting}
@@
 \end{lstlisting}
+
 \end{itemize}
@@
 
-What happens if we define a new interface \Q@M@ extending one \Q@A@${}_i$ and one \Q@B@${}_i$? we have 9 cases, that can fit nicely a table:\\*
+What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
+\Q@B@${}_i$? The 9 cases are shown here:\\*
 \noindent
@@
 \begin{itemize}
-\item \textbf{conflict error} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both implemented. This is also considered an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ are both abstract. This is also considered correct in all trait models.
-\item \textbf{conservative error}
- happens when the method from \Q@A@${}_i$ and one \Q@B@${}_i$ is implemented in only one side.
-This is different from what we would expect in a trait model, but is coherent with the conservative idea that a method defined in an interface
-should not silently satisfy a method in another one.
+\item \textbf{conflict error} happens when the methods from \Q@A@${}_i$ and
+  \Q@B@${}_i$ are both implemented, which is also an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the methods from \Q@A@${}_i$
+  and \Q@B@${}_i$ are both abstract, which is also considered correct in all
+  trait models.
+\item \textbf{conservative error} happens when only one method from \Q@A@${}_i$
+  or \Q@B@${}_i$ is implemented, which is different from what we would expect in
+  a trait model, but is coherent with the conservative idea that a method
+  defined in an interface should not silently satisfy a method in another one.
 \end{itemize}
 
-During our experimentation, we found a bug in ECJ (eclipse version of javac): The case \textbf{B3},\textbf{A1}
-is accepted by ECJ4.5.1 and rejected by javac.
-By email communication with Brian Goetz (lead Java8 designer) we have confirmation that the expected behaviour is rejection, and that this is a bug in ECJ.
+During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
+ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
+(leading Java 8 designer) we have confirmed that the expected behaviour is
+rejection, hence this is a bug in ECJ.\footnote{This bug is fixed in the ECJ
+  developer brunch , but not released as a stable version yet.}
 %Table~\ref{table:javabug} shows the method overriding bug in Java. In the
@@
 }{
-  \mixinAnn|\emptyset
+  \mixinAnn|\weakAnn|\emptyset
   }{annotations}\\

### 2015-12-02 paper/sections/Overview.tex
revising text

--- 2015-12-02 paper/sections/Overview.tex
@@
 \subsection{Advanced Multiple Inheritance with Object Interfaces}
-Finally, we can define Pegasus as simply as we did in the simplified (and stateless) 
-version in the right-side of Figure~\ref{fig:pegasus}:
+Finally, we can define \texttt{Pegasus} as simple as we did in the simplified
+(and stateless) version on the right-side of Figure~\ref{fig:pegasus}:
 
@@
 transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
-\Q@location@. This works because \Q@Horse@ do not perform
-any field type refinement, otherwise we may have to choose/create a
-common subtype in order for \texttt{Pegasus} to exists.
+\Q@location@.%  This works because \Q@Horse@ do not perform
+% any field type refinement, otherwise we may have to choose/create a
+% common subtype in order for \texttt{Pegasus} to exists.
 

### 2015-12-02 UseMixinLombok/src/test/TestExpression.java

@@ -34,14 +34,11 @@ public class TestExpression {
 interface Exp { int eval(); }
 @Obj interface Lit extends Exp {
 	int x();
-	default int eval() { return x(); }
-}
+	default int eval() { return x(); }}
 @Obj interface Add extends Exp {
     Exp e1(); Exp e2();
     default int eval() {
-		return e1().eval() + e2().eval();
-	}
-}
+        return e1().eval() + e2().eval();}}
 //END_EXPRESSION_INIT
 
 //BEGIN_EXPRESSION_SUB
@@ -52,13 +49,11 @@ interface Exp { int eval(); }
 //BEGIN_EXPRESSION_PRINT
 interface ExpP extends Exp { String print(); }
 @Obj interface LitP extends Lit, ExpP {
-    default String print() {return "" + x();}
-}
+    default String print() {return "" + x();}}
 @Obj interface AddP extends Add, ExpP {
     ExpP e1(); ExpP e2();//return type refined!
     default String print() {
-		return "(" + e1().print() + " + " + e2().print() + ")";
-	}}
+        return "("+e1().print()+" + "+e2().print()+")";}}
 //END_EXPRESSION_PRINT
 
 //BEGIN_EXPRESSION_COLLECTLIT

### 2015-12-02 paper/paper.tex
Future Work ready for review.
Revised Implementation

@@
 
+\usepackage{listings}
+\usepackage{multicol}
 \usepackage{microtype}%if unwanted, comment out or use option "draft"
@@
 keywordstyle=\ttfamily\bfseries,
+aboveskip=3pt,
+belowskip=-1pt
 % commentstyle=\color{Gray},
@@
 
-\newcommand\mixin{\textbf{@Mixin} }
+\newcommand\mixin{\textbf{@Obj}\xspace}
 \renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
+\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
 \usepackage{xspace}
@@
 
+\input{sections/FutureWork.tex}
+
 %===============================================================================

### 2015-12-02 paper/sections/Formalization.tex
S3 summary

@@
 \section{Interaction of Interface Methods with Interface Composition}
+Before formalizing our annotation, we summarize methods interaction/shadowing 
+mechanism in Java 8 interfaces and especially. We show several interesting cases
+and summarize compilation result of method composition into 3 categories:
+conflict error, accepted (both abstract) and conservative error. Meanwhile we
+also show different composition handling mechanism among traits, javac and ECJ.
+
 From Java 8 interfaces can have three type of methods: abstract, default, and
@@
 (leading Java 8 designer) we have confirmed that the expected behaviour is
-rejection, hence this is a bug in ECJ.\footnote{This bug is fixed in the ECJ
-  developer brunch , but not released as a stable version yet.}
+rejection, hence this is a bug in ECJ.\footnote{This bug was also reported by
+  others and is fixed in the ECJ developer brunch, but not released as a stable
+  version yet.}
 %Table~\ref{table:javabug} shows the method overriding bug in Java. In the

### 2015-12-02 paper/sections/Overview.tex
small fix

@@
 naming conventions and the types of the methods. Fluent setters are a variant of
-mutable setters, and are discussed in detail in Section~\ref{sec:dsls}.
+(void) setters, and are discussed in detail in Section~\ref{sec:dsls}.
 
@@
 object; updates field \texttt{x} to \texttt{val}.             \\ \hline
-\textbf{mutable setters}        & \Q@void x(int val);@ & Sets the field
+\textbf{setters}        & \Q@void x(int val);@ & Sets the field
 \texttt{x} to a  new value \texttt{val}.        \\ \hline

### 2015-12-02 paper/sections/Introduction.tex
Revisions

@@
 \item {\em Useful, general purpose methods, require special care in
-  the presence of subtyping.} Methods such as $clone$, or
+  the presence of subtyping.} Methods such as
   \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
-  internal state of an object, but they also require their types to be
+  internal state of an object, but they also require their return types to be
   refined in subtypes.
@@
 of interfaces with default methods. The semantics of object interfaces
-is given as a type-directed translation from CJ to itself. In the
+is given as a syntax-directed translation from CJ to itself. In the
 resulting CJ code all object interfaces are translated into regular CJ
@@
 to be type-preserving, ensuring that the translation does not
-introduce type-errors. CJ's usefulness goes beyond serving as a
-calculus to formalize object interfaces. During the development
-process of CJ, we encountered a bug in the implementation of default
-methods for the Eclipse Compiler for Java (ECJ). For the program revealing the 
-bug, ECJ behaves differently from both our formalization and Oracle's 
-Java compiler.
+introduce type-errors. 
+%CJ's usefulness goes beyond serving as a
+%calculus to formalize object interfaces. During the development
+%process of CJ, we encountered a bug in the implementation of default
+%methods for the Eclipse Compiler for Java (ECJ). For the program revealing the 
+%bug, ECJ behaves differently from both our formalization and Oracle's 
+%Java compiler.
 
 To evaluate the usefulness of object interfaces, we illustrate 3
-applications (case studies). The first application is a simple 
+applications. The first application is a simple 
 solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
@@
 can be used to formally define the translation of object interfaces. 
-We prove a type-preservation theorem, and we present a Java program 
-that reveals a bug in the ECJ implementation of default methods.
+We prove that several properties hold for the translation, including a type-preservation.
 

### 2015-12-02 paper/sections/Overview.tex
Revisions

--- 2015-12-02 paper/sections/Overview.tex
@@
 \end{figure}
+\bruno{Yanlin: figure needs to rename point to location}
 
-To propose a standard example, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
+This section illustrates how the \mixin annotation enables expressive
+idioms using multiple inheritance. 
+To propose a standard example for multiple inheritance, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
 Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
 Birds can \Q@fly@, thus their location need to be a three dimensional \Q@Point3D@.
-Finally, we model \Q@Pegasus@ as a kind of \Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s.\footnote{
+Finally, we model \Q@Pegasus@ (one of the best known creatures in Greek
+mythology) as a kind of \Q@Animal@ with the skills
+of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
+the basic system in given in the left side of Figure~\ref{fig:pegasus}.\footnote{
 Some research argues in favour of using subtyping for modelling taxonomies, other research argue against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
 
-\subsection{Simple Multiple Inheritance with Default Methods}
+\subsection{Simple Multiple Inheritance with Default
+  Methods}\label{sec:simple}
 
-Suppose that we want to model an animal system as shown in
-the left side of Figure~\ref{fig:pegasus}. \texttt{Horse} and \texttt{Bird} are subtypes
+Before we model the final animal system, we first start with a
+simplified version without locations. This version serves the purpose of illustrating how 
+Java 8 default methods can already model simple forms of multiple inheritance.
+\texttt{Horse} and \texttt{Bird} are subtypes
 of \texttt{Animal}, with methods \texttt{run()} and \texttt{fly()},
-respectively. Pegasus (one of the best known creatures in Greek
-mythology) can not only \emph{run} but also \emph{fly}! This is the
+respectively. Pegasus can not only \emph{run} but also \emph{fly}! This is the
 place where \emph{``multiple inheritance''} is needed, because
@@
 functionality from both \texttt{Horse} and \texttt{Bird}.
-Ignoring points, for the moment, a 
-first attempt to model the animal system in Java 8 is given on the right-side 
+A first attempt to model the animal system in Java 8 is given on the right-side 
 of Figure~\ref{fig:pegasus}.
@@
 do not allow instance fields, no form of state is needed so far to
-model the animal system.  Therefore this example provides a simple
-illustratation how a basic form of multiple inheritance, similar to
-what can be done with traits, can be accomplished in Java 8.
+model the animal system.
 
@@
 interfaces to model the animal system is simply that interfaces 
-cannot be directly instantiated. Classes are needed for instantiation, such as:
+cannot be directly instantiated. Classes, such as:
 
@@
 
-\noindent to create animal objects. Now a \texttt{Pegasus} animal can be created
+\noindent are needed for instantiation. Now a \texttt{Pegasus} animal can be created
 using the class constructor:
@@
 
-\noindent There is a lot of annoyance here. Firstly, the sole
+\noindent There is a some annoyances here. Firstly, the sole
 purpose of classes is to provide a way to instantiate
@@
 needs to do is to add a \mixin annotation to the \texttt{Horse}, 
-\texttt{Bird}, and \texttt{Pegasus} interfaces.  For example, 
-for \texttt{Bird} and \texttt{Pegasus} the code would be as follows:
+\texttt{Bird}, and \texttt{Pegasus} interfaces:
 
@@
 
-The effect of the annotations is that a static \emph{factory} method called
+\noindent The effect of the annotations is that a static \emph{factory} method called
 \texttt{of} is automatically added to the interfaces. With the
@@
 only sees the original code with the \mixin annotation. Compared to the pure
-Java solution in Section~\ref{}, the solution using object interfaces
+Java solution in Section~\ref{sec:simple}, the solution using object interfaces
 has the advantage of providing a direct mechanism for object
@@
 
-The animal system modelled so far, is a simplified version of the
+The animal system modelled so far is a simplified version of the
 system presented in the left-side of Figure~\ref{fig:pegasus}, and still does not appear to 
 justify the \mixin annotation. Lets now move on to modelling 
-the complete animal system. In order to do this, animals will include 
+the complete animal system. In order to do this animals will include 
 a notion of points, representing their location in space. As we shall
@@
 coordinates. However here we will illustrate how points are modelled with
-interfaces. Since Java disallows fields inside interfaces, states are modeled
+interfaces. Since Java disallows fields inside interfaces, state is modeled
 using abstract (getter) methods:
@@
 
-\noindent Unfortunatelly, creating a new point object is quite cumbersome, even
+\noindent Unfortunatelly, creating a new point object is cumbersome, even
 with anonymous classes:
@@
 
-\noindent However it is not required to use this cumbersome syntax for every
+\noindent However this cumbersome syntax is not required for every
 object allocation. As programmers do, for ease or reuse, the boring
@@
 \paragraph{\Q@Point2D@ with Object Interfaces}
-This obvious constructor code can be automatically generated by our \mixin
+This obvious ``constructor'' code can be automatically generated by our \mixin
 annotation.  By annotating the interface \Q@Point2D@, a variation of the shown
 static method \texttt{of} will be generated, mimicking the functionality of a
-simple-minded constructor: by looking at the abstract methods it first detects
+simple-minded constructor. \mixin first looks at the abstract methods and detects
 what are the fields, then generates an \Q@of@ method with one parameter for each
@@
 
-\paragraph{\texttt{with} methods in Object Interfaces}
+\paragraph{\texttt{wither} methods in Object Interfaces}
 The pattern of creating a new object by reusing most information from an old
-object is very common when programming with immutable data-structures. It is
-supported in our code generation as \Q@with@ methods, for example:
+object is very common when programming with immutable
+data-structures. As such, it is
+supported in our code generation as \Q@with@ methods. For example:
 \begin{lstlisting}
@@
 \end{comment}
-Moreover, if the programmer wants a different implementation, they may 
-just provide new \texttt{with} methods using \Q@default@:
+Moreover, if the programmer wants a different implementation, he may 
+provide an alternative implementation using \Q@default@ methods:
 \begin{lstlisting}
@@
 instantiate \texttt{Animal}.  For \texttt{Horse}, the \mixin annotation is used
-and an implementation of \texttt{run()} method is defined using \Q@default@,
-which further illustrates the convenience of \emph{with} methods:
+and an implementation of \texttt{run()} is defined using a \Q@default@
+method. The implementation of \texttt{run()} further illustrates the convenience of \emph{with} methods:
 
@@
 
-\paragraph{Summary} Dealing with states (mutable or not) in object interfaces
+\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
 relies on a notion of abstract state, where only methods that interact with
-states are available to users. Object interfaces provide support for four
+state are available to users. Object interfaces provide support for four
 different types of abstract state operations, which are summarized in
@@
 locations, \Q@Bird@s need 3D locations. Therefore when the \texttt{Bird}
-interface extends the animal interface, the notion of points needs to
+interface extends the \Q@Animal@ interface, the notion of points needs to
 be refined. Such kind of refinement usually poses a challenge
@@
 \item Define a \Q@Point2D@ field in \Q@Animal@ and define an extra \Q@int z@
-  field in \Q@Bird@.  This solution is very ad-hoc, requires to basically
+  field in \Q@Bird@.  This solution is very ad-hoc, requiring to basically
   duplicate the difference between \Q@Point2D@ and \Q@Point3D@ inside \Q@Bird@.
@@
   precise type;
-\item we provide a \Q@default@ setter with the old signature.
+\item a \Q@default@ setter implementation with the old signature is provided.
 \end{itemize}
+
+Thus, with object interfaces, the code for the \Q@Bird@ interface is:
 
@@
 perspective the key is the implementation for the setter with the old
-signature.
+signature (\Q@location(Point2D)@). The key to the setter
+implementation is a new type of \@with@ method, called a property updater.
 
-\paragraph{\Q@Point3D@ and properties updater}
+\paragraph{\Q@Point3D@ and properties updaters}
 The \Q@Point3D@ interface is defined as follows:
@@
 
-\noindent By using the new \Q@with@ method we may use the information for
-\Q@z@ already stored in the object to forge an appropriate \Q@Point3D@
-to store.  Note how all the informations about what fields sits in
-\Q@Point3D@ and what in \Q@Point2D@ is properly encapsulated in the
-\Q@with@ method, and is transparent for the implementer of \Q@Bird@.
-
-\Q@with-@ methods (functionally) update a field at a time.  This can be
-inefficient, and sometime hard to maintain.  Often we want to update multiple
+\noindent Note that this interface includes a different type of
+\Q@with@ method, which takes a \Q@Point2D@ as an argument.
+The old wither methods (such as \Q@withX@) functionally update a field at a time.  This can be
+inefficient, and sometimes hard to maintain.  Often we want to update multiple
 fields simultaneously, for example using another object as source.  Following
@@
 current object with all fields that match the fields of parameter type
-updated. The idea is that we want the result to be still like \Q@this@, but
-modified as similar as possible to the parameter.
+updated. The idea is that we want the result to still be like \Q@this@, but
+modified to be as similar as possible to the parameter.
+
+With the new \Q@with@ method we may use the information for
+\Q@z@ already stored in the object to forge an appropriate \Q@Point3D@
+to store. Note how all the information about what fields sit in
+\Q@Point3D@ and in \Q@Point2D@ is properly encapsulated in the
+\Q@with@ method, and is transparent to the implementer of \Q@Bird@.
 
@@
 generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. The
-generated code is very mechanical, writing such code by hand is error-prune. For
+generated code is very mechanical, writing such code by hand is error-prone. For
 example a distracted programmer may swap the arguments of calls to
 \Q@Point3D.of@.  Note how \Q@with-@ methods are automatically refined in their
-return type, so that code like \Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@
-will be accepted and behave as expected. If the programmer wishes to suppress
+return type, so that code like:
+
+\begin{lstlisting}
+Point3D p=Point3D.of(1,2,3); p=p.withX(42);
+\end{lstlisting}
+
+\noindent will be accepted and behave as expected. If the programmer wishes to suppress
 this behaviour and keep the signature as it was, it is sufficient to redefine
@@
 \subsection{Advanced Multiple Inheritance with Object Interfaces}
-Finally, we can define \texttt{Pegasus} as simple as we did in the simplified
-(and stateless) version on the right-side of Figure~\ref{fig:pegasus}:
+Finally, we can define Pegasus as simply as we did in the simplified (and stateless) 
+version in the right-side of Figure~\ref{fig:pegasus}:
 
@@
 transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
-\Q@location@.%  This works because \Q@Horse@ do not perform
-% any field type refinement, otherwise we may have to choose/create a
-% common subtype in order for \texttt{Pegasus} to exists.
+\Q@location@. This works because \Q@Horse@ do not perform
+any field type refinement, otherwise we may have to choose/create a
+common subtype in order for \texttt{Pegasus} to exists.
 

### 2015-12-02 UseMixinLombok/src/pegasusLombok/Main.java

@@ -7,7 +7,7 @@ interface Animal {}
 
 //BEGIN_PEGASUS_LOMBOK
 @Obj interface Horse extends Animal{
-	default void run() {System.out.println("I can run!");};
+	default void run() {System.out.println("I can run!");}
 }
 @Obj interface Bird extends Animal{
 	default void fly() {System.out.println("I can fly!");}

### 2015-12-02 paper/sections/CaseStudy.tex

@@
 problem about modular extensibility issues in software evolution. Recently, a
-new solution~\cite{eptrivially}, using only covariant type refinement, has been
-shown.  When such solution is modelled with interfaces and default methods, it
+new solution~\cite{eptrivially} using only covariant type refinement is
+proposed. When this solution is modeled with interfaces and default methods, it
 can even provide independent extensibility: the ability to assemble a system
-from multiple, independently developed, extensions.  Unfortunatelly, the
+from multiple, independently developed extensions. Unfortunatelly, the
 required instantiation code makes a plain Java solution verbose and cumbersome
@@
 \begin{tabular}{l|l}
-\begin{minipage}{8cm}
-%\vspace{-90pt}
-\begin{lstlisting}[]
-interface Exp { int eval(); }
-@Obj interface Lit extends Exp {
-	int x();
-	default int eval() { return x(); }
-}
-@Obj interface Add extends Exp {
-	Exp e1(); Exp e2();
-	default int eval() {
-		return e1().eval() + e2().eval();
-	}
-}
-\end{lstlisting}
-\end{minipage}
+\lstinputlisting[linerange=34-41]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 &
-\begin{minipage}{8cm}
-\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
-\end{minipage}
+\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
@@
 
-\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=45-46]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=85-87]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=80-82]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 

### 2015-12-02 paper/sections/Formalization.tex
Fiexed incoherence in declarations
Section 3

--- 2015-12-02 paper/sections/Formalization.tex
@@
 \section{Interaction of Interface Methods with Interface Composition}
-From Java 8 interfaces can have three type of methods: abstract, default, and
+Before formalizing Classless Java, it is helpful to informally discuss the behaviour
+of methods in Java 8 interfaces.
+
+%We show several interesting cases
+%and summarize compilation result of method composition into 3 categories:
+%conflict error, accepted (both abstract) and conservative error. Meanwhile we
+%also show different composition handling mechanism among traits, javac and ECJ.
+
+\subsection{Methods in Java 8 Interfaces}
+In Java 8 interfaces there are three types of methods: abstract, default, and
 static methods.
-\begin{itemize}
 
-\item Static methods are handled in a very clean way: they are visible only in
-  the interface they are explicitly defined. This means the following code
+\paragraph{Static methods} are handled in a very clean way: they are visible only in
+  the interface in which they are explicitly defined. This means the following code
   is ill-typed.
@@
 
-\item Abstract methods composition is accepted when exist a most specific one.
+\paragraph{Abstract method} composition is accepted when there exist a most specific one.
   For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
@@
 
-\item Default methods conflict with any other default or abstract method. For
-  example the following code is rejected due to methods conflict.
+\paragraph{Default methods} conflict with any other default or abstract method. For
+  example the following code is rejected due to method conflicts.
 \begin{lstlisting}
@@
 
-\item A method in the current interface wins over any method in its
-super-interfaces, provided that the method%  conforms to the subtype of all methods
-% in its super-interfaces, i.e., the method
+\paragraph{Resolving conflicts:}
+A method in the current interface wins over any method in its
+super-interfaces, provided that the method
 is the most specific one. This method also overrides conflict due to
@@
 
-\end{itemize}
-
-While trying to formally encode the Java specification we have done some tests to clarify corner case behaviour.
-Consider the following correct declarations:
+\subsection{Classifying Method Composition}
+When interfaces are composed and methods with the same name (and
+signature) exist, there are 3 possible outcomes.
+We illustrate the possible outcomes using the following (correct) declarations:
 \begin{lstlisting}
@@
 
+\noindent 
 What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
-\Q@B@${}_i$? The 9 cases are shown here:\\*
+\Q@B@${}_j$? Nine representative cases are shown next:\\*
 \noindent
@@
 \\*
-We try to classify the results out of the table:
+We try to classify the results in the table:
 \begin{itemize}
-\item \textbf{conflict error} happens when the methods from \Q@A@${}_i$ and
-  \Q@B@${}_i$ are both implemented, which is also an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the methods from \Q@A@${}_i$
-  and \Q@B@${}_i$ are both abstract, which is also considered correct in all
+\item \textbf{conflict error} happens when the methods from are both implemented, which is also an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the methods from are both abstract, which is also considered correct in all
   trait models.
-\item \textbf{conservative error} happens when only one method from \Q@A@${}_i$
-  or \Q@B@${}_i$ is implemented, which is different from what we would expect in
+\item \textbf{conservative error} happens when only one method from is implemented, which is different from what we would expect in
   a trait model, but is coherent with the conservative idea that a method
@@
 
-During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
 the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
@@
 (leading Java 8 designer) we have confirmed that the expected behaviour is
-rejection, hence this is a bug in ECJ.\footnote{This bug is fixed in the ECJ
-  developer brunch , but not released as a stable version yet.}
+rejection, hence this is a bug in ECJ. This bug was also reported by
+  others and is fixed in the ECJ developer branch, but not released as a stable
+  version yet.
 %Table~\ref{table:javabug} shows the method overriding bug in Java. In the
@@
 \noindent$\begin{array}{l}
-\InTextAssert{35ex}{
+\InTextDef{22ex}{
 \method\in\tops(\m,\Cs)}{
+}{
+\mBody(\m,\C)=\method
 }\\
 \InTextWith{
-  \mBody(\m,\C)=\method,\ \method\mbox{ not a static method, and }}\\
-\tab\tab  \C\in\Cs\ \mbox{ such that }
-\forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C
+ \method\mbox{ not a static method, and }
+ \C\in\Cs\ \mbox{ such that }
+\forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C}
 

### 2015-12-02 paper/sections/Formalization.tex
small change in S3

@@
 
-\subsection{Classifying Method Composition}
+\subsection{Classifying Outcomes of Interface Composition}
 When interfaces are composed and methods with the same name (and

### 2015-12-02 paper/sections/Appendix.tex
Proof in the appendix, fixed and added notes.

@@
 \begin{proof}
-The proof is based on induction. By the grammar shown in Figure~\ref{Grammar}, there are three cases for an arbitrary expression $e$:
+By induction on the typing rules: by the grammar shown in Figure~\ref{Grammar}, there are 6 cases for an arbitrary expression $e$:
 \begin{itemize}
-\item A variable or a field update. The type preservation for $e$ is ensured by induction.
+\item Variables are typed in the same exact way.
+\item Field update. The type preservation is ensured by induction.
 \item A method call (normal, static or super).
-Such a method won't be ``removed'' by the translation and is still there, and the types remain unchanged.
-The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any method call.
+The corresponding method declaration won't be ``removed'' by the translation, also the types remain unchanged.
+The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
 \item An object creation.
@@
 \textbf{LEMMA 2. }
-If $\II_0$ has \textbf{@ObjWeak} and is OK, then $[\![\II_0]\!]$ OK.
+If $\II_0$ has \textbf{@ObjWeak} and $\C_0$ OK, then $[\![\II_0]\!]$ OK.
 \begin{proof}
@@
         $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C_0.\QM{of}$ method.
-Hence \[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
+Hence \[\Gamma, m_1:\C_1\ldots m_n:\C_n, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
         We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
@@
         \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C_0$?}
+      \marco{Not without defining what that would mean. Do we need to? it seams good as it is now...}
     \end{itemize}
@@
         \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C_0)
-        \end{align*}
+        \end{align*}\marco{I do not understand this point, I think ofMethod is not what you want to write, but I can not understand what you want do do...}
     \item For the $\QM{with#}\m_i$ method,
@@
     Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$.
-But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
+But that object includes all implementations for those cases\marco{you need to expand here, proof that we have all the implementations for all cases by using the definition of ofMethod}, hence it is of type $\C_0$ by \rn{t-Obj}.
 \end{itemize}
@@
 
-\subsection{Proof of THEOREM}\label{subsec:proof3}
+\subsection{Proof of THEOREM}
+\marco{the therem statement must be here}
+\label{subsec:proof3}
 \begin{proof}

### 2015-12-02 paper/sections/Translation.tex
Proof in the appendix, fixed and added notes.

--- 2015-11-29 paper/sections/Translation.tex
@@
 
-\hl{Furthermore, we introduce \textbf{@ObjWeak}, which is only an annotation for illustrating the translation, hence is not implemented in Lombok.
-By this, the real translation is divided into two parts: \textbf{@Obj} automatically refines the return types only, and \textbf{@ObjWeak} generates the constructor method $\QM{of}$.
-In the source code, the only annotation \textbf{@Obj} combines both of the work and finishes the translation at a time.}
-
-\subsection{Translation Function}
+We chose to modularly divide our translation in two parts, so that we can discuss about formal properties of those parts later.
+To this aim we introduce the annotation \Q:@ObjWeak:.
+Its role is only in the translation process, hence is not part of the classless java language.
+\Q:@ObjWeak: generates the constructor method $\QM{of}$, while 
+\mixin automatically refines the return types and calls \Q:@ObjWeak:.
 
 \noindent$\begin{array}{l}
-\InTextDef{18em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
+\InTextDef{16em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
 }{
-\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
+[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
 \methods\ \methods' \cC
+]\!]}\\
+\InTextWith{\methods'=\otherMethod(\C_0,\methods)}\\
+
+\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
+}{
+\QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
+\methods\ \ofMethod(\C_0) \cC
 }\\
-\InTextWith{\methods'=\otherMethod(\C_0,\methods)}
+\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\C_0) }
 \end{array}$
 
-\hl{To translate an interface annotated by \mixin, we add some other methods for type refinement, then add a new annotation \textbf{@ObjWeak} to the interface.}
-
-\noindent$\begin{array}{l}
-\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
-}{
-\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
-\methods\ \method' \cC
-}\\
-\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \method'=\ofMethod(\C_0)}
-\end{array}$
-
-\hl{To translate an interface annotated by \textbf{@ObjWeak}, we add the $\QM{of}$ method.}
-However, first of all we check if the interface is valid for annotation:
+As you see, we explicitly check if the interface is valid for annotation:
 
@@
 %$\isClone(\method, \C_0)$.
+
 That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.
-
-%To complete our formal definition we would need to add setters and \Q@with@.
 Moreover, we check that the method \Q@of@ is not already defined by the user.
 In our simplified formalization we consider this to be just an error.
-In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}\haoyuan{Not sure if it is $\QM{of}\notin\dom(\methods)$ or $\QM{of}\notin\dom(\C_0)$.
-It is an error if a static method overrides an instance method.}
+In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.
 
@@
 \mixinAnn\ \QM{interface B extends A}\ \oC\cC\end{array}\]}
+\marco{It is ok, in that case we just regenerate it identical}
 
@@
 \otherMethod(\C_0,\methods)}{
- \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
-\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+ \isWith(\mBody(\QM{with#}\m, \C_0), \C_0),}\\
+\tab\tab{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
 \mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
 \InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0)}\\
-\InTextWith{
+\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
+\tab\tab{
 \C\ \m\QM{();}\in \fieldsFunc(\C_0)
@@
 
-\marco{ insert somewhere description of  fluent setter~\cite{the-lombock-thread-something-about-fluent-stuff}.
-This allows for convenient and chains of setters, as we will show later \marco{insert forward reference when available}}.
+\subsection{Guaratees}
+Now we try to understand what is the meaning of saying that our approach is correct, and more in genera, what it means to say that a Language tuning is correct.
+We identify three types of guarancees: 
+\begin{itemize}
+\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not wrote.
+In our case it means that either \mixin{} produces in controlled way an understandable error, or the class can be successfully annotate and the generated code is well typed.
+We guarantee \textit{Self coherence}.
 
+\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
+That is, the annotation do not remove any functionality, is just
+adding more behaviour.
+We guarantee \textit{Client coherence}.
+
+\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well typed if they was well typed without the instrumentation.
+This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
+\mixin  do not guarantee \textit{Heir coherence}.
+Indeed consider the following example
+
+\begin{lstlisting}
+interface A { int x(); A withX(int x); }
+@Obj interface B extends A {}
+interface C extends B { A withX(int x); }
+\end{lstlisting}
+
+By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
+This would break \Q@C@.
+\end{itemize}
 \subsection{Results}
@@
 
-To understand this theorem statement, we need to understand three kind of guarantees that we can offer for safety:
-\begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not wrote.
-In our case it means that either \mixin{} produces in controlled way an understandable error, or the class can be successfully annotate and the generated code is well typed.
-We guarantee \textit{Self coherence}.
-
-\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
-That is, the annotation do not remove any functionality, is just
-adding more behaviour.
-We guarantee \textit{Client coherence}.
-
-\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well typed if they was well typed without the instrumentation.
-This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
-We do not guarantee \textit{Heir coherence}.
-Indeed consider the following example
-
-\begin{lstlisting}
-interface A { int x(); A withX(int x); }
-@Obj interface B extends A {}
-interface C extends B { A withX(int x); }
-\end{lstlisting}
-
-By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@.
-
-\end{itemize}
 %The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed.

### 2015-12-02 paper/sections/Conclusion.tex
revision

@@
 
-Before Java 8, concrete methods and static methods where not allowed
+Before Java 8, concrete methods and static methods were not allowed
 to appear in interfaces.  Java 8 allows static interface methods and
-introduces \emph{default methods}, which allow for implementation
-insides interfaces. This had an important positive consequence that
-was probably overlooked by the Java design team: the concept of class
-(in java) is now redundant and unneeded.  We define a subset of Java,
-called ClassLess Java, where programs and (reusable) libraries can be
+introduces \emph{default methods}, which enables implementations
+inside interfaces. This had an important positive consequence that
+was probably overlooked: the concept of class
+(in Java) is now (almost) redundant and unneeded.  We define a subset of Java,
+called Classless Java, where programs and (reusable) libraries can be
 easily defined and used.  To avoid for some syntactic boilerplate
 caused by Java not being originally designed to work without classes,
-we introduce a new annotation:\mixin provide default implementations
+we introduce a new annotation, \mixin, which provides default implementations
 for various methods (e.g. getters, setters, with-methods) and a
-mechanism to instantiate objects. \mixin annotation helps programmers
-to write less cumbersome code while coding in ClassLess Java; indeed
-we think the obtained gain is so high that ClassLess Java with \mixin
-annotation can be less cumbersome than full Java.\bruno{May need rewriting}
+mechanism to instantiate objects. The \mixin annotation helps programmers
+to write less cumbersome code while coding in Classless Java. Indeed
+we think the obtained gain is so high that Classless Java with the \mixin
+annotation can be less cumbersome than full Java.
+Interestingly, without classes there is also no subclassing. This scratches an old
+  itching point in the long struggle of subtyping versus subclassing:
+  According to some authors, from a software engineering perspective,
+  interfaces are just a kind of classes. Others consider more
+  opportune to consider interfaces are pure types. We do not know how to conciliate
+  those two viewpoints and ClassLess Java design. We do not have
+  Classes purely in the Java sense.\bruno{Marco: you probably should
+    revise the conclusion}

### 2015-12-02 paper/sections/Formalization.tex
revision

@@
 FeatherweightJava-like calculus which models the essence of Java
-interfaces with default methods.
+interfaces with default methods. This formalization is used in
+Section~\ref{sec:translation} to define the semantics of object interfaces.
 
@@
 
-Figure~\ref{Grammar} shows the syntax of ClassLess Java.%
+Figure~\ref{Grammar} shows the syntax of ClassLess Java.
 The syntax formalizes a minimal
-version of Java 8, focusing on interfaces, default methods and object
+subset of Java 8, focusing on interfaces, default methods and object
 creation literals.  There is no syntax for classes.
-To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all maps to a single set of identifiers as in Java.  Expressions
+To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all map to a single set of identifiers as in Java.  Expressions
 consist of conventional constructs such as variables ($\x$), method
@@
 simplicity) all the fields are initialized with a variable present in
-scope. Note how our language is a subset of Java 8.
-To  be compatible with java, the concrete syntax for an interface
+scope. 
+To  be fully compatible with Java, the concrete syntax for an interface
   declaration with empty supertype list  would also
@@
 $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
-For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquelly indentfied by their name.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquelly indentified by their name.
 
@@
 We omit the definition of
-the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation. %
-\footnote{\marco{find a better place for} Notice how there are no classes, thus there is no
-  subclassing.  We believe that this approach may scratch an old
-  itching point in the long struggle of subtyping versus subclassing:
-  According to some authors, from a software engineering perspective,
-  interfaces are just a kind of classes. Others consider more
-  opportune to consider interfaces are pure types. In this vision our
-  language would have no subclassing. We do not know how to conciliate
-  those two viewpoints and ClassLess Java design. We do not have
-  Classes purely in the Java sense.}
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.
 The auxiliary notation $\Gamma^\mh$ trivially
@@
 method name and the return type from a method header. $\mBody(\m,\C)$,
-defined in Section~\ref{},
+defined in Section~\ref{sec:auxiliary},
 returns the full method declaration as seen by $\C$, that is the
@@
 
-\subsection{Auxiliary Definitions}
+\subsection{Auxiliary Definitions}\label{sec:auxiliary}
 

### 2015-12-02 paper/sections/CaseStudy.tex
revise on related work. TODO: ‘one method limitation discussion’

@@
 \paragraph{Adding a New Type of Expressions}
-In the object oriented paradigm, it is easy to add new types of expressions. For example, the following code shows how to add subtraction.
+In the OO paradigm, it is easy to add new types of expressions. For example, the
+following code shows how to add subtraction.
 
@@
 
-\marco{text under need to be improved}
-\paragraph{Adding a New Operation} The difficulty of the EP in object-oriented 
+%\marco{text under need to be improved}
+\paragraph{Adding a New Operation} The difficulty of the EP in OO  
 languages arises from adding new operations. For example, adding a pretty printing 
-operation, would typically involve changing all of the existing code. However, a solution 
-to the EP forbids this and it also forbids using casts. In other words, it should be possible 
-to add the pretty printing operation in a type-safe and modular way.  
-It turns out that, this can be done easily, using our object interfaces approach. 
-The code in Figure~\ref{fig:ep} shows how to add the new operation \texttt{print}.
-
-\noindent  The interface \texttt{ExpP} extending interface \texttt{Exp} is defined with the
-extra method \texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are
-defined with default implementations of \texttt{print()}, extending
-base interfaces \texttt{Lit} and \texttt{Add}, respectively. Importantly, note
-that in \texttt{AddP}, the types of the ``\emph{fields}'' (i.e. the getter methods) \texttt{e1} 
-and \texttt{e2} are refined. If the types were not refined then the body of the \texttt{print}
-method in \texttt{AddP} would fail to type-check.
+operation would typically change all existing code. However, a solution 
+%to the EP forbids this and it also forbids using casts. In other words, it should be possible 
+should add the pretty printing operation in a type-safe and modular way.  It
+turns out to be easily achieved with the assitance of \mixin.  The code in
+Figure~\ref{fig:ep} (rhs) shows how to add the new operation \texttt{print}.
+Interface \texttt{ExpP} extends \texttt{Exp} with the extra method
+\texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are defined with
+default implementations of \texttt{print()}, extending base interfaces
+\texttt{Lit} and \texttt{Add}, respectively. Importantly, note that in
+\texttt{AddP}, the types of ``\emph{fields}'' (i.e. the getter methods)
+\texttt{e1} and \texttt{e2} are refined. If the types were not refined then
+\texttt{print()} method in \texttt{AddP} would fail to type-check.
 

### 2015-12-02 paper/sections/RelatedWork.tex
fix some spacing issues
revising text
revise on related work. TODO: ‘one method limitation discussion’

@@
 \subsection{Multiple Inheritance in Object Oriented Languages}
-Many authors have argued in favour or against multiple inheritance.  Multiple inheritance provides
-expressive power, but it is difficult to model and implement, and can create
-programs that are hard to reason about. These difficulties include the famous
-diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
-%and the yo-yo problem~\cite{taenzer1989problems}.  
-To conciliate the need for expressive power and the need for simplicity, many
-models have been proposed in over the years, including C++ virtual
-inheritance, mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and
-hybrid model such as CZ~\cite{malayeri2009cz}.
-They provide novel programming architecture models in the OO paradigm. In terms
-of restrictions set on these models, C++ virtual inheritance aims at a more
-general model, mixins added some restrictions on the model, and trait model is
-the most restricted one (excluding states, instantiation, etc).
+Many authors have argued in favour or against multiple inheritance.  Multiple
+inheritance is very expressive, but difficult to model and implement, and can
+cause difficulty (including the famous diamond (fork-join)
+problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning
+programs. To conciliate the need for expressive power and simplicity, many
+models have been proposed, including C++ virtual inheritance,
+mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid model
+such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
+models in the OO paradigm. In terms of restrictions set on these models, C++
+virtual inheritance aims at a relative general model; mixin model add some
+restrictions on the model; and trait model is the most restricted one (excluding
+states, instantiation, etc).
 
 \paragraph{C++ Approach.}
-C++ tries to provide a general solution to multiple inheritance. Virtual
-inheritance in C++ provides another solution to multiple inheritance (especially
-the diamond problem by keeping only one copy of the base
-class)~\cite{ellis1990annotated}, however suffers from object initialization
-problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
-constructor calls to virtual superclasses, which would potentially cause serious
-semantic errors. In our approach, \mixin annotation processor has the full
-control of object initialization mechanism, and this mechanism is clean and
-transparent to users. Moreover, if not satisfied with the default generated
-\texttt{of} method, customized factory methods are also allowed to be defined.
+C++ tries to provide a general solution to multiple inheritance by virtual
+inheritance that deals with diamond problem by keeping only one copy of the base
+class~\cite{ellis1990annotated}, however suffers from object initialization
+problem as shown in~\cite{malayeri2009cz}. It bypasses all constructor calls to
+virtual superclasses, which would potentially cause serious semantic errors. In
+our approach, \mixin annotation processor has the full control over object
+initialization, whose mechanism is transparent to users. Moreover, customized
+factory methods are also allowed, if users are not satisfied with the default
+generated \texttt{of} method.
 
 \paragraph{Mixins.}
-Mixins are a more restricted model than the C++ approach. Mixins allow to name
+Mixins are a more restricted model than C++ approach. Mixins allow to name
 components that can be applied to various classes as reusable functionality
-units. However, they suffer from linearisation: the order of mixin application
-is relevant in often subtle and undesired ways. This hinders their usability
-and the ability of resolving conflicts: the linearisation (total ordering) of
-mixin inheritance cannot provide a satisfactory resolution in all cases and
-restricts the flexibility of mixin composition. To fight those limitations, an
+units. However, % they suffer from linearisation: the order of mixin application
+% is relevant in often subtle and undesired ways. This hinders their usability
+% and the ability of resolving conflicts:
+the linearisation (total ordering) of
+mixin inheritance cannot provide a satisfactory resolution in some cases and
+restricts the flexibility of mixin composition. To fight this limitation, an
 algebra of mixin operators is introduced~\cite{ancona2002calculus}, but this
-raised the complexity of the approach, especially when constructors and fields
+raises the complexity, especially when constructors and fields
 are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearised mixins.
@@
 multiple inheritance without diamond problem. They divide inheritance into two
-separate concepts: inheritance dependency (using \texttt{require}) and
-implementation inheritance (using \texttt{extends}). Using a combination of
-\texttt{requires} and \texttt{extends}, a program with diamond inheritance can be
-transformed to one without diamonds. Also in this approach, fields and multiple
-inheritance can coexist. However as shown in the example in the article, the
-untangling of inheritance also untangles class structure. In CZ, not only the
-number of classes (from 4 to 6 in the stream example), but also the class
-hirarchy complexity increases. Our approach does not complicate the structure,
-and (conceptual) fields coexist with multiple inheritance.
+separate concepts: inheritance dependency and implementation inheritance. Using
+a combination of \texttt{requires} and \texttt{extends}, a program with diamond
+inheritance can be transformed to one without diamonds. Moreover, fields and
+multiple inheritance can coexist. However as shown in the example in the
+article, the untangling of inheritance also untangles class structure. In CZ,
+not only the number of classes (from 4 to 6 in the stream example), but also the
+class hirarchy complexity increases. \mixin does not complicate the
+structure, and fields can also coexist with multiple inheritance.
 
 \paragraph{Traits and Java's default methods}
-Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
-strong line between units of reuse (traits) and object factories
-(classes). In this model, traits as units of reusable code, contain
-only methods as reusable entities. Thus, no state and state
-initialization are considered. Classes act as object factories,
-requiring functionalities from multiple traits. Java 8 interfaces with
-default methods are closely related to traits: concrete method
-implementation are allowed (via the \textbf{default} keyword) inside
-interfaces. The introduction of default methods opens the gate for
-various flavours of multiple inheritance in Java, using interfaces.
-Traits offer a trait algebra with operations like sum, alias and exclusion, provided
-for explicit conflict resolution.  
-Former work by Bono et al.~\cite{bono14}. provides details on
-mimicking the trait algebra through Java 8 interfaces. 
-We briefly recall the main points of their encoding;
-however we propose a different representation of \textbf{exclusion}.
-The author of ~\cite{bono14} agree that our revised version is
-cleaner, typesafe and more direct.
-%\bruno{do we want do say this?}\yanlin{need to ask Marco.}
+Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
+line between units of reuse and object factories. Traits, as units of reusable
+code, contain only methods as reusable functionality. ignoring states and state
+initialization. Classes, as object factories, require functionalities from
+(multiple) traits. Java 8 interfaces with default methods are closely related to
+traits: concrete method implementation are allowed (via the \textbf{default}
+keyword) inside interfaces. The introduction of default methods opens the gate
+for various flavours of multiple inheritance in Java. Traits offer an algebra
+with operations like sum, alias and exclusion, providing explicit conflict
+resolution. Former work~\cite{bono14} provides details on mimicking the trait
+algebra through Java 8 interfaces.  We briefly recall the main points of their
+encoding; however we propose a different representation of \textbf{exclusion}.
+The author of ~\cite{bono14} agrees that our revised version is cleaner,
+typesafe and more direct.
 
@@
 \item \textbf{Symmetric sum} can be obtained by simple multiple inheritance between interfaces.
-\shortItem
     \begin{lstlisting}
@@
 \item \textbf{Overriding} a conflict is obtained by specifying which super interface take precedence.
-\shortItem
     \begin{lstlisting}
@@
 \item \textbf{Alias} is creating  a new method delegating to the existing super interface.
-\shortItem
     \begin{lstlisting}
@@
 \item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces.
-\shortItem
     \begin{lstlisting}
@@
 There are also proposals for extending Java (before Java8) with traits. For
-example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by Liquori et al. extends
-the calculus of Featherweight Java (FJ)~\cite{Igarashi01FJ} with
-statically-typed traits, adding trait-based inheritance in Java.  Except for
-few, mostly syntactic details, their work can be emulated/rephrased with Java8
-interfaces.  There are also extensions to the original trait model, which have
-operations that default methods and interfaces cannot model, such as method
-renaming in \cite{reppy2006foundation}, which breaks structural subtyping.
+example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} extends
+FJ~\cite{Igarashi01FJ} with statically-typed traits, adding trait-based
+inheritance in Java.  Except for few, mostly syntactic details, their work can
+be emulated with Java8 interfaces. There are also extensions to the original
+trait model, which have operations that default methods and interfaces cannot
+model, such as method renaming in \cite{reppy2006foundation}, which breaks
+structural subtyping.
 
@@
 interfaces instead of classes, in order to rely on the modular composition
-offered by interfaces. Since Java was designed for classes, a direct class-less
-programming style is verbose and feels unnatural. However, annotation driven
-code generation is enough to overcome this difficulty, and the resulting
-programming style encourages modularity, composability and reusability, by
-keeping a strong object oriented feel. In that sense, we promote object
-interfaces as being both units of reusable code and object factories. Our
-practical experience is that, in Java, separating the two notions leads to alot
-of boilerplate code, and is quite limiting when multiple inheritance with state
-is required. The use of abstract state operations avoids the key difficulties
-associated with multiple inheritance and state, while still being quite
-expressive.  Moreover the ability to support constructors adds additional
-expressivity, which is not available in approaches such as Scala's
-traits/mixins.
+offered by interfaces. Since Java was designed for classes, a direct classless
+programming style is verbose and unnatural. However, annotation driven code
+generation is enough to overcome this difficulty, and the resulting programming
+style encourages modularity, composability and reusability, by keeping a strong
+OO feel. In that sense, we promote object interfaces as being both units of
+reusable code and object factories. Our practical experience is that, in Java,
+separating the two notions leads to a lot of boilerplate code, and is quite
+limiting when multiple inheritance with states is required.  Abstract state
+operations avoid the key difficulties associated with multiple inheritance and
+states, while still being quite expressive.  Moreover the ability to support
+constructors adds additional expressivity, which is not available in approaches
+such as Scala's traits/mixins.
 
@@
 \subsection{Meta-programming Competes with Language Extensions}
-\bruno{needs to be polished better, I think}
-\bruno{Mention/Compare with Language Virtualization work in Scala. They also re-interpret 
-syntax.}
 The most obvious solution to adding features to a language is language
-extension.  It is often implemented as syntactic extensions that can be
-desugared to the base language. For example, the Scala compiler was extended to
-directly supports XML syntax. However, this approach does not support combining
-multiple extensions into one. We are de facto creating a fork in the language,
-and rarely the new fork gain enough traction to become the main language
-release.  On this topic we mention SugarJ~\cite{erdweg11sugarj} - a Java-based
-extensible language allowing programmers to extend it with custom features by
-definitions in meta-DSLs (SDF, Stratego, etc).\bruno{so, what's the point that 
-we want to make with SugarJ?}\yanlin{we were talking about lightweight language
-sugar, so SugarJ is a comparison to ours. Where using SugarJ requires learning
-various DSLs like SDF, stratego, etc, in comparison, our approach is just using
-Java. }
+extension. It is often implemented as syntactic extensions that can be desugared
+to the base language. For example, the Scala compiler was extended to directly
+supports XML syntax. However, this approach does not support combining multiple
+extensions into one. Scala-Virtualized is an extension to Scala which allows to
+blend shallow and deep embedding of DSLs. It redefines Scala structures to
+method calls, which can be overridden by DSL implementors. Therefore
+Scala-Virtualized also reinterprets syntax. These extensions are admittedly
+useful, however they are essentially creating a fork of the language, and rarely
+the new fork gains enough traction to become the main language release. % On this
+% topic we mention
+SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language allowing
+programmers to extend it with custom features by definitions in meta-DSLs (SDF,
+Stratego, etc). Using SugarJ requires learning various DSLs while our approach
+is just using Java. However its intention is different from ours: one of the
+purpose of SugarJ is to allow easy defition of sugars while our approach aims to
+implement the desired program transformation without just in Java.
 

### 2015-12-02 paper/sections/Formalization.tex
Section 4

@@
 Now we need to check if the object is a valid extension for that specific interface.
-This can be logically diveded in two steps:
+This can be logically divided into two steps.
 First  we check that all method headers are valid
@@
 
-Here we require that for all newly declared methods, there is a
+\noindent Here we require that for all newly declared methods, there is a
 method with the same name defined in the interface $\C$, and that such method is a subtype of the newly introduced one.
@@
 
-Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
-That is, we allows return type specialization as introduced in Java5.
+\noindent Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
+That is, we allows return type specialization as introduced in Java 5.
 Default methods are subtypes if their method headers are subtypes.
-
-Finally, all abstract methods in the interface (that is methods that need
+Secondly, all abstract methods in the interface (that is methods that need
 to be explicitly overridden) have been
@@
 
-The rule \rn{(t-intf)} checks that an interface $\C$ is correctly
+The rule \rn{t-intf} checks that an interface $\C$ is correctly
 typed.  First we check that the body of all the default and static
@@
 Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
-$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
-
-We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.\\*
+$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
+interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
 \noindent$\begin{array}{l}
@@
 \override(\methods(\m),
-\shadow(\m,\tops(\m,\Cs\ )))
+\shadow(\tops(\m,\Cs\ )))
 }\\
@@
 
-As you can see, we are delegating the work to three others auxiliary functions:\\*
- $\tops(\Cs), \shadow(\m,\Cs)$ and $\override(\method,\method')$
-${}_{}$\\*
+\noindent The definition of $mBody$ delegates the work to three others auxiliary functions:
+ $\tops(m,\Cs)$, $\shadow(\methods)$ and $\override(\method,\method')$.
 
-\tops{} recovers from the interface table only the ``needed'' methods, that is,
-the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:\\*
+\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
+the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:
 
@@
 
-\shadow{} choose the most specific version of a method, that is the unique version available, or a conflicted version from a set of possibilities.
-We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:\\*
+\paragraph{\shadow{}} chooses the most specific version of a method,
+that is the unique version available, or a conflicted version from a
+set of possibilities. Internally a special modifier $\conflicted$ is used to denote the case of two methods with conflicting implementation.
+We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
 \noindent
@@
 
-Where $\Aux{mostSpecific}$ return the most specific method , that is a method whose type is the subtype of all the others.
-Since method subtyping is a partial ordering, this may be not defined, this in turn makes \shadow{}, and the whole \mBody{} not defined for that specific $\m$.
-Rule  \rn{t-intf}  relies on this behaviour.
+\noindent The definition $\Aux{mostSpecific}$ returns the most
+specific method, that is a method whose type is the subtype of all the
+others.  Since method subtyping is a partial ordering,
+$\Aux{mostSpecific}$ may not be defined, this in turn makes \shadow{},
+and the whole \mBody{} not defined for that specific $\m$. Rule
+\rn{t-intf} relies on this behaviour.
 
@@
 \InTextWith{\method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype \method'}
-
 \end{array}$
 
-${}_{}$\\*
-The override function models how the implementation in an interface can override implementation in the superinterface; even in case of a conflict.
-Note how we use the special value $\none$, and how (forth case) overriding can solve a conflict.
+\paragraph{override} models how the implementation in an interface can
+override implementation in the superinterface, even in case of a
+conflict.  Note how we use the special value $\none$, and how (forth
+case) overriding can solve a conflict.
 

### 2015-12-02 paper/sections/Implementation.tex
Section 6

@@
 
-Our implementation is based on an extension to Lombok. The Lombok
+Our implementation is based on an extension of Lombok. The Lombok
 project~\cite{lombok} is a Java tool that aims at removing (or
@@
 to the original framework.
-A transformation is based on a handler, which acts on the AST from
-parsing the annotated node and returns a modified AST for analysis and
+A transformation is based on a handler, which acts on the AST for the 
+annotated node and returns a modified AST for analysis and
 generation afterwards. Such a handler can either be a Javac handler or

### 2015-12-02 paper/sections/CaseStudy.tex
Case studies

@@
 \section{Case Studies}
-In this section we conduct three case studies which reveals various
-advantages using \mixin annotation. The first case study provides a
+In this section we conduct three case studies which reveal various
+advantages of using the \mixin annotation. The first case study provides a
 simple way to solve the Expression Problem, while supporting multiple,
@@
 problem about modular extensibility issues in software evolution. Recently, a
-new solution~\cite{eptrivially} using only covariant type refinement is
+new solution~\cite{eptrivially} using only covariant type refinement was
 proposed. When this solution is modeled with interfaces and default methods, it
-can even provide independent extensibility: the ability to assemble a system
+can even provide independent extensibility~\cite{zenger05independentlyextensible}: the ability to assemble a system
 from multiple, independently developed extensions. Unfortunatelly, the
@@
 operation \texttt{eval} for expression evaluation.
-
-\noindent As shown in Figure~\ref{fig:ep}, \texttt{Exp} is the common
+As shown in Figure~\ref{fig:ep}, \texttt{Exp} is the common
 super-interface with an evaluation operation \texttt{eval()}
@@
   print operation on the right. }\label{fig:ep}
-\end{figure}
+\end{figure}\bruno{Yanlin: code is overflowing the margins. Note that
+  there are 2 spare lines that can be used on the right!}
 
@@
 \paragraph{Independent Extensibility}
-To show that our approach supports independent extensibility~\cite{zenger05independentlyextensible}, we first define a
-new operation \texttt{collectLit} (which collects all
-literal components in an expression) on expressions. For space reasons, 
+To show that our approach supports independent extensibility, we first define a
+new operation \texttt{collectLit}, which collects all
+literal components in an expression. For space reasons, 
 we omit the definitions of the methods:
@@
 interface ExpC extends Exp { List<Integer> collectLit(); }
-@Obj interface LitC extends Lit, ExpC {
-    default List<Integer> collectLit() { ... }}
-@Obj interface AddC extends Add, ExpC {
-    ExpC e1(); ExpC e2(); //return type refined!
-    default List<Integer> collectLit() { ... }}
+@Obj interface LitC extends Lit, ExpC {...}
+@Obj interface AddC extends Add, ExpC {ExpC e1(); ExpC e2(); ...}
 \end{lstlisting}
@@
 Note that the code for instantiation is automatically generated by \mixin. 
-So, for example creating a simple expression of type \texttt{ExpPC} is 
+Creating a simple expression of type \texttt{ExpPC} is 
 as simple as:
@@
   automatically generate fluent setters.
-\item Along the extension direction, the return types of fluent setters are
+\item The approach supports extensibility: the return types of fluent setters are
   automatically refined.
@@
 in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
-reductions was due to the replacement of instantiation classes with
+reductions were due to the replacement of instantiation classes with
 generated \texttt{of} methods. The number of source lines of code (SLOC)

### 2015-12-02 paper/sections/Formalization.tex
Revising formalization.

@@
 \override(\methods(\m),
-\shadow(\tops(\m,\Cs\ )))
+\shadow(\tops(\m,\C_0)))
 }\\
 \InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\cC}
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}
 \end{array}$
 
-\noindent The definition of $mBody$ delegates the work to three others auxiliary functions:
- $\tops(m,\Cs)$, $\shadow(\methods)$ and $\override(\method,\method')$.
+\noindent The definition of $\mBody$ delegates the work to three others auxiliary functions:
+ $\tops(m,\C_0)$, $\shadow(\methods)$ and $\override(\method,\method')$.
 
 \paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
-the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:
+the non-static ones that are not transitively reachable by following another, less specific, superinterface chain.
+\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
+super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
+Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
+Formally:
 
 \noindent$\begin{array}{l}
-\InTextDef{22ex}{
-\method\in\tops(\m,\Cs)}{
+\InTextDef{13ex}{
+\tops(\m,\C_0)}{
 }{
-\mBody(\m,\C)=\method
+\{\ \get(\m, \C)\ |\ \C\in\super(\C_0), \get(\m, \C) \ne\emptyset,\mbox{ and }
 }\\
-\InTextWith{
- \method\mbox{ not a static method, and }
- \C\in\Cs\ \mbox{ such that }
-\forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C}
-
+\hspace{2.1in}\forall\C'\in\super(\C_0)\mbox{ and }\get(\m,\C')\ne\emptyset: \C'<:\C\Rightarrow\C'=\C\ \}\\
+\InTextDef{13ex}{
+\super(\C_0)}{
+}{
+\{\ \C'\ |\ \C'\in\Cs,\mbox{ or }\exists\C\in\Cs,\C'\in\super(\C)\ \}
+}\\
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}\\
+\InTextDef{13ex}{
+\get(\m,\C_0)}{
+}{
+\method
+}\\
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
 \end{array}$
@@
 conflict.  Note how we use the special value $\none$, and how (forth
-case) overriding can solve a conflict.
+case) overriding can solve a conflict. \hl{The notation in the last case below
+means $\method'$ is not static, and is supertype of $\method$.}
 

### 2015-12-02 paper/sections/Translation.tex
Revising formalization.

@@
 
-\haoyuan{I think $\QM{with#}\m\notin\dom(\methods)$ is not enough. The withM method with return type $I_0$ can also be defined already in a super interface. Like:
-\[\begin{array}{l}\QM{
-interface A}\ \oC\QM{int x(); B withX(int x);}\cC\\
-\mixinAnn\ \QM{interface B extends A}\ \oC\cC\end{array}\]}
-\marco{It is ok, in that case we just regenerate it identical}
-
 \noindent$\begin{array}{ll}
-\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
+\InTextDef{16em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
 \otherMethod(\C_0,\methods)}{
- \isWith(\mBody(\QM{with#}\m, \C_0), \C_0),}\\
-\tab\tab{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+ \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
+\hspace{2.65in}{
 \mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
-\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
+\InTextDef{16em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
 \otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
-\tab\tab{
-\C\ \m\QM{();}\in \fieldsFunc(\C_0)
+\hspace{2.65in}{
 \mbox{ and } \QM_\m\notin\dom(\methods)}\\
@@
 To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
- is required by $\C_0$, but is not already present in the methods \hl{directly declared in} $\C_0$.
+ is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$, \hl{since overloading is not modelled}.
 

### 2015-12-02 paper/sections/Appendix.tex
Revise formalization and proof.

@@
 \section{Appendix}\label{sec:appendix}
-
-\subsection{THEOREM}\label{subsec:theorem}
-
-\textbf{THEOREM. }
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
-if $\II_0$ has \textbf{@ObjWeak},
-$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
-then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$\\*
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.\\
-
-\noindent For convenience, assume that
-\begin{align*}
-\II_0\ &=\ \weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
-[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \method' \cC
-\end{align*}
 
@@
 \item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
+    For convenience, we use ``$\method$ in $\ofMethod(\C_0)$'' to denote that $\method$ is one of the implemented methods in the return expression of $\ofMethod(\C_0)$, namely \Q@new@ $\C_0$\Q@(){...}@.
     \begin{itemize}
@@
         \begin{align*}
-        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
         \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C_0)\\
@@
         \end{align*}\marco{I do not understand this point, I think ofMethod is not what you want to write, but I can not understand what you want do do...}
+        \haoyuan{Right, sorry for that, I mean if a getter is created by ofMethod, then it is a field of I0, and finally for this part our aim is to prove sigvalid, and the definition of sigvalid asks you to prove $\mh\QM;<:\mBody(...)$, see the last line.}
     \item For the $\QM{with#}\m_i$ method,
         \begin{align*}
-        &\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        &\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
         \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
@@
         \begin{align*}
-        &\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\in\ofMethod(\C_0)\\
+        &\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
         \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
@@
     Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$.
-But that object includes all implementations for those cases\marco{you need to expand here, proof that we have all the implementations for all cases by using the definition of ofMethod}, hence it is of type $\C_0$ by \rn{t-Obj}.
+But that object includes all implementations for those cases.
+A getter $m_i$ is generated if it satisfies $\isField$; a \Q@with-@ method is generated for the case $\isWith$, by the definition of $\withMethod$; a setter for $\isSetter$, similarly, by the definition of $\setterMethod$.
+Hence it is of type $\C_0$ by \rn{t-Obj}.
 \end{itemize}
@@
 
-\subsection{Proof of THEOREM}
-\marco{the therem statement must be here}
-\label{subsec:proof3}
+\subsection{THEOREM and Proof}\label{subsec:theorem}
+\textbf{THEOREM. }
+For a given $\II_0\ldots\II_n$ interface table such that
+$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
+if $\II_0$ has \textbf{@ObjWeak},
+$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
+then in the interface table
+$[\![\II_0]\!]\II_1\ldots\II_n$\\*
+$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.
 \begin{proof}

### 2015-12-02 paper/sections/Translation.tex
Revise formalization and proof.

@@
 \subsection{Guaratees}
-Now we try to understand what is the meaning of saying that our approach is correct, and more in genera, what it means to say that a Language tuning is correct.
-We identify three types of guarancees:
+Now we try to understand what is the meaning of saying that our approach is correct, and more in general, what it means to say that a language tuning is correct.\haoyuan{Actually only self coherence is supported?}
+We identify three types of guarantees:
 \begin{itemize}
@@
 adding more behaviour.
-We guarantee \textit{Client coherence}.
+\mixin  does not guarantee \textit{Client coherence}.
 
@@
 This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
-\mixin  do not guarantee \textit{Heir coherence}.
-Indeed consider the following example
+\mixin  does not guarantee \textit{Heir coherence}.
+Indeed consider the following example:
 
@@
 
-By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@.
+\noindent By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
+This would break \Q@C@. And it's similar why client coherence is not supported as well. In the client code, an expression
+that uses ``\Q@new B(){...}@'' to create an object of \Q@B@ gives an implementation of \Q@A withX(int x);@, but after the return
+type is refined to \Q@B@, there is a type error.
 \end{itemize}
+
+\noindent Consequently, type refinement is a common case in client code, and hence is included by \mixin as a feature for convenient use, whereas it simultaneously brings some risk in throwing type errors, though they are likely to be fixed easily by users.
+On the other hand, a property of \mixin is informally concluded as: for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation.
+Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
+The formal theorem and proof is presented afterwards.
+
 \subsection{Results}
-Informally, in a type correct interface table, if we translate a specific interface, then
-the resulting table is type correct, or there is a subtype of the translated interface in the original table.
-\textbf{THEOREM. }
+
+\haoyuan{We use \textbf{@ObjWeak} or \Q:@ObjWeak:?}
+
+The theorem is presented below. Lemmas and the complete proof are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
+
+\noindent\textbf{THEOREM. }
 For a given $\II_0\ldots\II_n$ interface table such that
 $\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
-if $\II_0$ has \mixin,
+if $\II_0$ has \textbf{@ObjWeak},
 $\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
-there is no $\C<:\C_0$, there is no \Q@new @$\C_0$\Q@(){..}@,
 then in the interface table
@@
 %\haoyuan{Text needs improving.}
-
-To prove the theorem we introduce two lemmas below. The complete proof is available in Appendix \ref{subsec:proof1}, \ref{subsec:proof2} and \ref{subsec:proof3}.\\
-
-\marco{ you need to use the Lemma/theorem macros}
-\noindent\textbf{LEMMA 2. }
-If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
-This is what we defined as self coherence before.
 

### 2015-12-02 paper/sections/Appendix.tex
Small fix in lemmas.

@@
 \textbf{LEMMA 1. }
-For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} and $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
+For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} annotation and $[\![\II_0]\!] = \II'_0$, then under the interface table $\II'_0\II_1\ldots\II_n$, $\Gamma\vdash e\in\C$.
 \begin{proof}
@@
 \textbf{LEMMA 2. }
-If $\II_0$ has \textbf{@ObjWeak} and $\C_0$ OK, then $[\![\II_0]\!]$ OK.
+If $\II_0$ has \textbf{@ObjWeak} annotation and $\C_0$ OK, then $[\![\II_0]\!]$ OK.
 \begin{proof}

### 2015-12-03 paper/sections/Overview.tex
updated figure 1. compress text

@@
 \begin{tabular}{c|c}
-\includegraphics[height=4cm]{PegasusDetail.pdf}\hspace{20pt} &
+\includegraphics[height=3.5cm]{PegasusDetail.pdf}\hspace{20pt} &
 \begin{minipage}{7cm}
 \vspace{-90pt}
-\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+\lstinputlisting[linerange=6-15]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
 %basicstyle=\ttfamily\scriptsize
@@
 \end{tabular}
-\caption{The animal system. Complete animal system on the left, and code for simplified animal system on the right. }\label{fig:pegasus}
+\caption{The animal system (left: complete structure, right: code for simplified animal system).}\label{fig:pegasus}
 \end{figure}
-\bruno{Yanlin: figure needs to rename point to location}
 
@@
 
-\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
+\lstinputlisting[linerange=19-21]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
 
@@
 
-\lstinputlisting[linerange=9-15]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
 
@@
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-Two dimensional points should keep track of their coordinates.  The usual
+2D points should keep track of their coordinates.  The usual
 approach to model points in Java would be to use a class with fields for the
-coordinates. However here we will illustrate how points are modelled with
+coordinates. However here we will illustrate how points are modeled with
 interfaces. Since Java disallows fields inside interfaces, state is modeled
@@
 
-%\lstinputlisting[linerange=41-47]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+%\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=57-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=61-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 
-\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=USINGHORSE
+\lstinputlisting[linerange=10-12]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=USINGHORSE
 
@@
 
-\lstinputlisting[linerange=66-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 
-\lstinputlisting[linerange=51-54]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
@@
 
-\lstinputlisting[linerange=41-47]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
 
-\lstinputlisting[linerange=51-54]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=57-57]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=61-62]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=66-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
 

### 2015-12-03 UseMixinLombok/src/pegasusJava/Main.java

@@ -1,18 +1,16 @@
 package pegasusJava;
 
-//import static java.lang.System.out;
+import static java.lang.System.out;
 
 //BEGIN_PEGASUS_JAVA
 interface Animal {} // no points yet!
 interface Horse extends Animal{
-	default void run() {
-		System.out.println("I can run!");
-	};
+    default void run() {
+        out.println("I can run!");}
 }
 interface Bird extends Animal{
-	default void fly() {
-		System.out.println("I can fly!");
-	}
+    default void fly() {
+        out.println("I can fly!");}
 }
 interface Pegasus extends Horse, Bird {}
 //END_PEGASUS_JAVA
@@ -24,9 +22,9 @@ class PegasusImpl implements Pegasus {}
 //END_PEGASUS_INST
 
 public interface Main {
-	public static void main(String[] args) {
-		Horse h = new HorseImpl();
-		Bird b = new BirdImpl();
-		Pegasus p = new PegasusImpl();	
-	}
+    public static void main(String[] args) {
+        Horse h = new HorseImpl();
+        Bird b = new BirdImpl();
+        Pegasus p = new PegasusImpl();	
+    }
 }

### 2015-12-03 UseMixinLombok/src/pegasusLombok/Main.java

@@ -1,17 +1,13 @@
 package pegasusLombok;
 
 import lombok.Obj;
-
+import static java.lang.System.out;
 
 interface Animal {}
 
 //BEGIN_PEGASUS_LOMBOK
-@Obj interface Horse extends Animal{
-	default void run() {System.out.println("I can run!");}
-}
-@Obj interface Bird extends Animal{
-	default void fly() {System.out.println("I can fly!");}
-}
+@Obj interface Horse extends Animal{ default void run() {out.println("running!");}}
+@Obj interface Bird extends Animal{	default void fly() {out.println("flying!");}}
 @Obj interface Pegasus extends Horse, Bird {}
 //END_PEGASUS_LOMBOK
 

### 2015-12-03 UseMixinLombok/src/test/TestAnimal.java

@@ -1,6 +1,7 @@
 package test;
 
 import lombok.Obj;
+import static java.lang.System.out;
 
 public class TestAnimal {
 	
@@ -64,8 +65,7 @@ interface Animal {Point2D location(); void location(Point2D val);}
 
 //BEGIN_BIRD
 @Obj interface Bird extends Animal {
-	Point3D location();	
-	void location(Point3D val);
+	Point3D location();	void location(Point3D val);
 	default void location(Point2D val) { location(location().with(val));}
 	default void fly() {location(location().withX(location().x() + 40));}}
 //END_BIRD

### 2015-12-03 UseMixinLombok/src/test/TestExpression.java

@@ -5,36 +5,36 @@ import java.util.List;
 import lombok.Obj;
 
 public class TestExpression {
-	public static String runTest() {
-		String res = "";
-		Lit e1 = Lit.of(3);
-		res += e1.eval() + "\n"; // 3
-		Lit e2 = Lit.of(4);
-		Add e3 = Add.of(e1, e2);
-		res += e3.eval() + "\n"; // 7
-		Sub e4 = Sub.of(e1, e2);
-		res += e4.eval() + "\n"; // -1
-		LitP e5 = LitP.of(3);
-		LitP e6 = LitP.of(4);
-		AddP e7 = AddP.of(e5, e6);
-		res += e5.print() + " = " + e5.eval() + "\n"; // 3 = 3
-		res += e7.print() + " = " + e7.eval() + "\n"; // (3 + 4) = 7
-		return res;
-	}
-	public static void main(String[] args) {
-	    System.out.println(runTest());
-	    
-	    //independent extensibility
-	    ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
-	    System.out.println(e8.print() + " = " + e8.eval() + " Literals: " + e8.collectLit().toString());
-	}
+    public static String runTest() {
+        String res = "";
+        Lit e1 = Lit.of(3);
+        res += e1.eval() + "\n"; // 3
+        Lit e2 = Lit.of(4);
+        Add e3 = Add.of(e1, e2);
+        res += e3.eval() + "\n"; // 7
+        Sub e4 = Sub.of(e1, e2);
+        res += e4.eval() + "\n"; // -1
+        LitP e5 = LitP.of(3);
+        LitP e6 = LitP.of(4);
+        AddP e7 = AddP.of(e5, e6);
+        res += e5.print() + " = " + e5.eval() + "\n"; // 3 = 3
+        res += e7.print() + " = " + e7.eval() + "\n"; // (3 + 4) = 7
+        return res;
+    }
+    public static void main(String[] args) {
+        System.out.println(runTest());
+
+        //independent extensibility
+        ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
+        System.out.println(e8.print() + " = " + e8.eval() + " Literals: " + e8.collectLit().toString());
+    }
 }
 
 //BEGIN_EXPRESSION_INIT
 interface Exp { int eval(); }
 @Obj interface Lit extends Exp {
-	int x();
-	default int eval() { return x(); }}
+    int x();
[…]

### 2015-12-03 paper/sections/CaseStudy.tex
more revise and compress

@@
 &
-\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=50-57]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
-\caption{The Expression Problem. Initial system on the left, and code for adding
-  print operation on the right. }\label{fig:ep}
-\end{figure}\bruno{Yanlin: code is overflowing the margins. Note that
-  there are 2 spare lines that can be used on the right!}
+\caption{The Expression Problem (left: initial system, right: code for adding
+  print operation).}\label{fig:ep}
+\end{figure}% \bruno{Yanlin: code is overflowing the margins. Note that
+  % there are 2 spare lines that can be used on the right!}
 
@@
 
-\lstinputlisting[linerange=80-82]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=81-83]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 

### 2015-12-03 paper/sections/Formalization.tex
more revise and compress

@@
 \begin{itemize}
-\item \textbf{conflict error} happens when the methods from are both implemented, which is also an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the methods from are both abstract, which is also considered correct in all
+\item \textbf{conflict error} happens when the methods from both are implemented, which is also an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the methods from both are abstract, which is also considered correct in all
   trait models.
-\item \textbf{conservative error} happens when only one method from is implemented, which is different from what we would expect in
+\item \textbf{conservative error} happens when only one method is implemented
+  (leaving another one abstract), which is different from what we would expect in
   a trait model, but is coherent with the conservative idea that a method

### 2015-12-03 paper/sections/FutureWork.tex
more revise and compress

@@
 for class methods (private, protected, synchronized, etc.). For example, the absence
-of the support for private/protected methods in Java8 interfaces forces all
+of the support for private/protected methods in Java 8 interfaces forces all
 members of interfaces to be public, including static methods. Since we use
-abstract methods to encode the state, our state is always all public. Still, because 
+abstract methods to encode state, our state is always all public. Still, because 
 the state can only be accessed by methods, it 
@@
 does not directly maps to the fields, (as for secondary constructors in Scala)
-he can simply define its own \Q@of@ method and delegate on the generated one, as
-in
+he can simply define its own \Q@of@ method and delegate on the generated one:
 \begin{lstlisting}

### 2015-12-03 paper/sections/FutureWork.tex
revise future work

@@
 Methods originally defined in Java class \Q@Object@, as \Q@clone@ and
-\Q@toString@, can be supported by our approach, but they need special
-care. If an interface annotated with \mixin asks an implementation for
-\Q@clone@, \Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily
-generate one from the fields.\footnote{In particular, for clone we can
-  do automatic return type refinement as we do for \Q@with-@ and
-  fluent setters. Note how this would solve most of the Java ugliness
-  related to \Q@clone@ methods.}
-However, if the user wishes to provide his own implementation, since the method is
-also implemented in \Q@Object@, a conflict arises. The generated code can resolve the conflict inside \Q@of@, by implementing the method and
-delegating it to the user implementation, thus
+\Q@toString@, can be supported by our approach with special care. If an
+interface annotated with \mixin asks an implementation for \Q@clone@,
+\Q@toString@, \Q@equals@ or \Q@hashCode@ we can easily generate one from the
+fields.\footnote{In particular, for clone we can do automatic return type
+  refinement as we do for \Q@with-@ and fluent setters. Note how this would
+  solve most of the Java ugliness related to \Q@clone@ methods.}  However, if
+the user wishes to provide his own implementation, since the method is also
+implemented in \Q@Object@, a conflict arises. The generated code can resolve the
+conflict inside \Q@of@, by implementing the method and delegating it to the user
+implementation, thus
+
 \begin{lstlisting}
 @Obj interface Point{ int x(); int y();
-  default Point clone(){ return Point.of(0,0);}//user defined clone
-}
+  default Point clone(){ return Point.of(0,0);}}//user defined clone
 \end{lstlisting} 
-Would expand into
+would expand into
 
 \begin{lstlisting}
-interface Point{ int x(); int y();
-  default Point clone(){ return Point.of(0,0);}//user defined clone
+interface Point{ ...
   public static Point of(int _x, int _y){
     return new Point(){...
-      public Point clone(){ return Point.super.clone();}
-    };}}
+      public Point clone(){ return Point.super.clone();}};}}
 \end{lstlisting}

### 2015-12-03 paper/sections/FutureWork.tex
now it is within 25 pages. comments on future work.

@@
 
+\yanlin{future work on state initialization removed for now. Because I think the
+  other two future works are fine. But this one is still not clear and
+  controversial. The proposed workaround forces users to make decision: if he
+  wants state with initialization then he also has to define the setter. Or
+  else, it is ambiguous: could be interpreted as a state initializer or just
+  a normal default method.}
+\begin{comment}
 \subsection{State initialization}
@@
 %\bruno{Yanlin please polish text/break long sentences.}
+\end{comment}
 

### 2015-12-03 paper/sections/Appendix.tex
Big progress on formalism/proofs

--- 2015-12-02 paper/sections/Appendix.tex
@@
 \textbf{LEMMA 1. }
-For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} and $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
+For any expression $e$ under an interface table $\II$ IT where $\Gamma\vdash e\in\C^\II$, $\II$ has \textbf{@ObjWeak} annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\C^\II$.
 \begin{proof}
@@
 The corresponding method declaration won't be ``removed'' by the translation, also the types remain unchanged.
-The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
+The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
 \item An object creation.
@@
 \end{proof}
-
+\textbf{LEMMA 1b. }
+For any expression $e$ under an interface table $\II$ IT where there is no heir of $\C^\II$,  $\Gamma\vdash e\in\C^\II$, $\II$ has \mixin annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\_<:\C^\II$.
+\begin{proof}
+The proof follows the same scheme of the Lemma1, but for the case of method call the return type may be refined with a subtype.
+This is still ok since we require $\_<:\C^\II$.
+On the other side, this weaker result still allows the application on the method call typing rules, since in the premises the types of the actual parameter are required to be a subtype of the formal one.
+\end{proof}
 \subsection{LEMMA 2 and Proof}\label{subsec:lemma2}
@@
 \textbf{LEMMA 2. }
-If $\II_0$ has \textbf{@ObjWeak} and $\C_0$ OK, then $[\![\II_0]\!]$ OK.
+If $\II$ has \textbf{@ObjWeak} annotation and $\C^{\II}$ OK in $\II$ IT, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
 \begin{proof}
@@
 
-\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\II_0]\!]$, the type of the return value is compatible with the method's return type.
+\noindent\textbf{Part I.} For each default or static method in the domain of $[\![\C^\II]\!]$, the type of the return value is compatible with the method's return type.
 
-Since $\II_0$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}.
-It suffices to prove that it still holds for $\ofMethod(\C_0)$.
+Since $\II$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II]\!]$, except for the new method \QM{of}.
+It suffices to prove that it still holds for $\ofMethod(\C)$.
 
-By the definition of $\ofMethod(\C_0)$, the return value is an object $$\QM{return new}\ \C_0 \oR\cR\ \QM{\{}...\ \QM{\}}$$
-To prove it is of type $\C_0$, we use the typing rule \rn{t-Obj}.
+By the definition of $\ofMethod(\C)$, the return value is an object $$\QM{return new}\ \C^\II \oR\cR\ \QM{\{}...\ \QM{\}}$$
+To prove it is of type $\C^\II$, we use the typing rule \rn{t-Obj}.
 
@@
 \item All field initializations are type correct.
-By the definition of $\ofMethod(\C_0)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
+By the definition of $\ofMethod(\C^\II)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
 \item All method bodies are well-typed.
     \begin{itemize}
-    \item Typing of the $i$-th getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]
+    \item Typing of the $i$-th getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C^\II \vdash m_i\in \C_i\]
         We know that $\C_i=\C^{\mh_i}$ since the $i$-th getter has its return type the same as the corresponding field $m_i$.
     \item Typing of the \QM{with-} method of an arbitrary field $m_i$.
-By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C_0\QM{.of(}\es_i\QM{);\}}\]
-        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C_0.\QM{of}$ method.
-Hence \[\Gamma, m_1:\C_1\ldots m_n:\C_n, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
-        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
+By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C^\II\QM{.of(}\es_i\QM{);\}}\]
+        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C^\II.\QM{of}$ method.
+Hence \[\Gamma, m_1:\C_1\ldots m_n:\C_n, \QM{this}:\C^\II,\ \QM{_val} : \C_i \vdash \C^\II\QM{.of}\oR\es_i\cR\in \C^\II\]
+        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
     \item Typing of the $i$-th setter \QM_$m_i$. If the \QM_$m_i$ method is well-defined, it has the form
-        \[\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
+        \[\C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
         By \rn{t-Update}, the assignment ``$\m_i\QM{= _val;}$'' is correct since $\m_i$ and $\QM{_val}$ have the same type $I_i$, and the return type is decided by \QM{this}.
-\[\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash\QM{this}\in \C_0\]
-        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
-        \haoyuan{Can we write $\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i, \C^{\mh_i} = \C_0 \vdash\QM{this}\in \C_0 = \C^{\mh_i}$?}\\
-        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C_0$?}
+\[\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i \vdash\QM{this}\in \C^\II\]
+        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
+        \haoyuan{Can we write $\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i, \C^{\mh_i} = \C^\II \vdash\QM{this}\in \C^\II = \C^{\mh_i}$?}\\
+        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C^\II$?}
       \marco{Not without defining what that would mean. Do we need to? it seams good as it is now...}
     \end{itemize}
-\item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
-    For convenience, we use ``$\method$ in $\ofMethod(\C_0)$'' to denote that $\method$ is one of the implemented methods in the return expression of $\ofMethod(\C_0)$, namely \Q@new@ $\C_0$\Q@(){...}@.
+\item All method headers are valid with respect to the domain of $\C^\II$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
+    For convenience, we use ``$\method$ in $\ofMethod(\C^\II)$'' to denote that $\method$ is one of the implemented methods in the return expression of $\ofMethod(\C^\II)$, namely \Q@new@ $\C^\II$\Q@(){...}@.
     \begin{itemize}
@@
         \begin{align*}
-        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
-        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C_0)\\
-        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; = \mBody(m_i,\C_0)\\
-        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C_0)
-        \end{align*}\marco{I do not understand this point, I think ofMethod is not what you want to write, but I can not understand what you want do do...}
-        \haoyuan{Right, sorry for that, I mean if a getter is created by ofMethod, then it is a field of I0, and finally for this part our aim is to prove sigvalid, and the definition of sigvalid asks you to prove $\mh\QM;<:\mBody(...)$, see the last line.}
+        &\C_i\ \m_i\oR\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; \in \fieldsFunc(\C^\II)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; = \mBody(m_i,\C^\II)\\
+        \mimply\hspace{.2in}& \C_i\ \m_i\oR\cR\QM; <: \mBody(m_i,\C^\II)
+        \end{align*}
     \item For the $\QM{with#}\m_i$ method,
         \begin{align*}
-        &\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
-        \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
-        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
-        \mimply\hspace{.2in}& \isWith(\mBody(\QM{with#}\m_i,\C_0),\C_0)\\
-        \mimply\hspace{.2in}& \C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM{with#}\m_i,\C_0)
+        &\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
+        \mimply\hspace{.2in}& \mBody(\QM{with#}\m_i,\C^\II) \mbox{ is of form }\mh\QM;\\
+        \mbox{with}\hspace{.2in}& \valid(\C^\II)\\
+        \mimply\hspace{.2in}& \isWith(\mBody(\QM{with#}\m_i,\C^\II),\C^\II)\\
+        \mimply\hspace{.2in}& \C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM{with#}\m_i,\C^\II)
         \end{align*}
@@
         \begin{align*}
-        &\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C_0)\\
-        \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C_0) \mbox{ is of form }\mh\QM;\\
-        \mbox{with}\hspace{.2in}& \valid(\C_0)\\
-        \mimply\hspace{.2in}& \isSetter(\mBody(\QM_\m_i,\C_0),\C_0)\\
-        \mimply\hspace{.2in}& \C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM_\m_i,\C_0)
+        &\C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{...\}}\mbox{ in }\ofMethod(\C^\II)\\
+        \mimply\hspace{.2in}& \mBody(\QM_\m_i,\C^\II) \mbox{ is of form }\mh\QM;\\
+        \mbox{with}\hspace{.2in}& \valid(\C^\II)\\
+        \mimply\hspace{.2in}& \isSetter(\mBody(\QM_\m_i,\C^\II),\C^\II)\\
+        \mimply\hspace{.2in}& \C^\II\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\QM; <: \mBody(\QM_\m_i,\C^\II)
         \end{align*}
     \end{itemize}
-\item All abstract methods in the domain of $\C_0$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
-    Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$.
+\item All abstract methods in the domain of $\C^\II$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
+    Here we simply refer to $\valid(\C^\II)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$.
 But that object includes all implementations for those cases.
 A getter $m_i$ is generated if it satisfies $\isField$; a \Q@with-@ method is generated for the case $\isWith$, by the definition of $\withMethod$; a setter for $\isSetter$, similarly, by the definition of $\setterMethod$.
-Hence it is of type $\C_0$ by \rn{t-Obj}.
+Hence it is of type $\C^\II$ by \rn{t-Obj}.
 \end{itemize}
 
-\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$
-Since $\II_0$ OK, we have $\dom(\II_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\method')$$
-This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so $\method'$ doesn't overlap with $\dom(\C_0)$.
-The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$.
-Also note that after translation, there are still no methods with conflicted names, since the \QM{of} method was previously not in the domain, hence $[\![\II_0]\!]$ is well-formed, which finishes our proof.
+\noindent\textbf{Part II.} Next we check that in $[\![\II]\!]$, $$\dom([\![\II]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$
+
+Since $\II$ OK, we have $\dom(\II)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II]\!])=\dom(\C^\II)\cup\dom(\method')$$
+This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so $\method'$ doesn't overlap with $\dom(\C^\II)$.
+The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II]\!])$ is only an extension to $\dom(\C)$ with the \QM{of} method, namely $\method'$.
+Also note that after translation, there are still no methods with conflicted names, since the \QM{of} method was previously not in the domain, hence $[\![\II]\!]$ is well-formed, which finishes our proof.
+\end{proof}
+
+\textbf{LEMMA 2b. }
+If $\II$ has \mixin annotation $\C^{\II}$ OK in $\II$ IT
+and there is no heir of $\C^\II$, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
+\begin{proof}
+\noindent\textbf{Part I.} Similarly to what already argued for Lemma2, 
+since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$,IT.
+The translation function delegate its work to \Q!@ObjWeak! in such way that we can refer to 
+Lemma2  to complete this part. Note that all the methods that are added (directly) by \mixin are abstract, and thus there is no body to typecheck.
+
+\noindent\textbf{Part II.}
+Similar to what already argued for Lemma2, but we need to notice that the the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
+Thus by last clause of the definition of $\override(\_)$, $\mBody(\_)$ is defined on the same method names.
 \end{proof}
@@
 \subsection{THEOREM and Proof}\label{subsec:theorem}
-\textbf{THEOREM. }
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
-if $\II_0$ has \textbf{@ObjWeak},
-$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
-then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$\\*
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.
+
+\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
+If a given interface table $\II$ IT is OK\\*
+ where $\II$ has \Q!@ObjWeak!, 
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
+then the interface table $[\![\II]\!]$ IT is OK.
+
 \begin{proof}
-\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, for any $\II\ne\II_0$, by \textbf{LEMMA 1}, we know that all its methods
+\textbf{LEMMA 2} already proves that $[\![\II]\!]$ is OK. On the other hand, for any $\II'\in $ IT$\setminus\II$, by \textbf{LEMMA 1}, we know that all its methods
 are still well-typed, and the generated code in translation of \textbf{@ObjWeak} is only a static method \QM{of}, which has no way to affect the domain
-of $\II$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
+of $\II'$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
 \end{proof}
+
+\noindent\textbf{THEOREM }\Q!@Obj!\\*
+If a given interface table $\II$ IT is OK\\*
+ where $\II$ has \Q!@Obj!, 
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
+then the interface table $[\![\II]\!]$ IT is OK.
+
+\begin{proof}
+Similar to what already argued for \textbf{THEOREM }\Q!@ObjWeak!
+we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finsh by \textbf{THEOREM }\Q!@ObjWeak!.
+\end{proof}

### 2015-12-03 paper/sections/Translation.tex
Big progress on formalism/proofs

@@
 %A complete formalization would also generate the \Q@with@.
-We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
+We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only in order to refine the return type.
 To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
- is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$, \hl{since overloading is not modelled}.
+ is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$. \marco{no, overloading is not the issue here\hl{since overloading is not modelled}}.
 
@@
 adding more behaviour.
-\mixin  does not guarantee \textit{Client coherence}.
+We guarantee \textit{Client coherence}.
 
@@
 
-\noindent By the translation rule, \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@. And it's similar why client coherence is not supported as well. In the client code, an expression
-that uses ``\Q@new B(){...}@'' to create an object of \Q@B@ gives an implementation of \Q@A withX(int x);@, but after the return
-type is refined to \Q@B@, there is a type error.
+\noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
+This would break \Q@C@. 
+Similarly, an expression of the form \Q@new B(){.. A withX(int x){..}}@ would compile before translation, but would be ill typed after the translation.
 \end{itemize}
 
-\noindent Consequently, type refinement is a common case in client code, and hence is included by \mixin as a feature for convenient use, whereas it simultaneously brings some risk in throwing type errors, though they are likely to be fixed easily by users.
-On the other hand, a property of \mixin is informally concluded as: for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation.
-Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
-The formal theorem and proof is presented afterwards.
+\noindent Our automatic type refinement 
+is a usefull and convenient feature, but is not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantic  to be well typed.
 
 \subsection{Results}
+To formally characterize the behaviour of our annotation, and the two level of guarancees that we offer, we provide some notations and two theorems:
+\begin{itemize}
+\item We say that an interface table 
+IT is OK if under such interface table, all interfaces are OK.
+\item Since interface tables are just represented as sequences of interfaces we can write IT = $\II$ IT' to talk about a specific interface in a table.
+\item We denote with $\C^{\II}$ the name of an interface.
+\item An interface table contains a hair of $\C$ if there is any interface that extends it, or any \Q@new@ that instantiate it.
+\end{itemize}
 
-\haoyuan{We use \textbf{@ObjWeak} or \Q:@ObjWeak:?}
+\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
+If a given interface table $\II$ IT is OK\\*
+ where $\II$ has \Q!@ObjWeak!, 
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
+then the interface table $[\![\II]\!]$ IT is OK.
 
-The theorem is presented below. Lemmas and the complete proof are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
+\noindent\textbf{THEOREM }\Q!@Obj!\\*
+If a given interface table $\II$ IT is OK\\*
+ where $\II$ has \Q!@Obj!, 
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
+then the interface table $[\![\II]\!]$ IT is OK.
 
-\noindent\textbf{THEOREM. }
-For a given $\II_0\ldots\II_n$ interface table such that
-$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
-if $\II_0$ has \textbf{@ObjWeak},
-$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
-then in the interface table
-$[\![\II_0]\!]\II_1\ldots\II_n$\\*
-$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.
-
-%The first two levels of safety are essential to our lightweight framework, nevertheless, the third one is not needed.
-In Java it implies not touching the annotated interface, and surely an alternative approach is to create a new interface and generate what we need.
-But since we are using Lombok for the translation, the transformation is applied to the annotated interface itself.
-But heir coherence is indeed an issue.
-See below how this is affected by our approach:
-
-%Now we go back and read the statement of our theorem again.
-It illustrates that a program with the translation of \mixin applied still type checks if it did, except for the subtypes of the annotated interface (type safety of them is not guaranteed).
-
-%\haoyuan{Text needs improving.}
-
-%\subsubsection{Derived notations}
-% Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
-%\[ \textsf{mbody}(m,C) = \textit{modifier } E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \] \[ \Rightarrow \textsf{mtype}(m,C) = \overline{D} \to E,\ \textsf{mmodifier}(m,C) = \textit{modifier}\]
-
-%\marco{we also need to define a function that gives all the methods of an interface, something line}
-%\[
-%\Aux{methodsOf}_\C=\{\m|\mBody(\m,\C)=\method\}
-%\]
+Informally, this means that for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation.
+Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
+The formal theorems proofs are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
 

### 2015-12-06 paper/sections/Formalization.tex
Now needed is good, (lets hope). I think it passes the H example!

@@
 \override(\methods(\m),
-\shadow(\tops(\m,\C_0)))
+\shadow(\tops(\m,\Cs)))
 }\\
 \InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n\ \oC\methods\ \cC \mbox{ and }
+ \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n
+
+}
 \end{array}$
 
-\noindent The definition of $\mBody$ delegates the work to three others auxiliary functions:
+\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to three others auxiliary functions:
  $\tops(m,\C_0)$, $\shadow(\methods)$ and $\override(\method,\method')$.
@@
 \paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
-the non-static ones that are not transitively reachable by following another, less specific, superinterface chain.
-\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
-super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
-Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
+the non-static ones that are not reachable by another, less specific, superinterface.
+Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
+%\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
+%super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
+%Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
 Formally:
@@
 \noindent$\begin{array}{l}
-\InTextDef{13ex}{
-\tops(\m,\C_0)}{
+\InTextDef{27ex}{
+\method\in\tops(\m,\C_0\dots\C_n)}{
  }{
-\{\ \get(\m, \C)\ |\ \C\in\super(\C_0), \get(\m, \C) \ne\emptyset,\mbox{ and }
+\metaVar{IT}(\C_0)(\m)=\method,  \method\mbox{ default or abstract method}
  }\\
-\hspace{2.1in}\forall\C'\in\super(\C_0)\mbox{ and }\get(\m,\C')\ne\emptyset: \C'<:\C\Rightarrow\C'=\C\ \}\\
-\InTextDef{13ex}{
-\super(\C_0)}{
-}{
-\{\ \C'\ |\ \C'\in\Cs,\mbox{ or }\exists\C\in\Cs,\C'\in\super(\C)\ \}
-}\\
-\InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}\\
-\InTextDef{13ex}{
-\get(\m,\C_0)}{
-}{
-\method
-}\\
-\InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
+\InTextWith{
+\not\exists i\in1..n \mbox{ such that } 
+\metaVar{IT}(\C_i)(\m)\mbox{ is a default or abstract method, and }
+\C_i \subtype \C_0}
+
+%\\
+%\InTextDef{13ex}{
+%\tops(\m,\C_0)}{
+%}{
+%\{\ \get(\m, \C)\ |\ \C\in\super(\C_0), \get(\m, \C) \ne\emptyset,\mbox{ and }
+%}\\
+%\hspace{2.1in}\forall\C'\in\super(\C_0)\mbox{ and }\get(\m,\C')\ne\emptyset: \C'<:\C\Rightarrow\C'=\C\ \}\\
+%\InTextDef{13ex}{
+%\super(\C_0)}{
+%}{
+%\{\ \C'\ |\ \C'\in\Cs,\mbox{ or }\exists\C\in\Cs,\C'\in\super(\C)\ \}
+%}\\
+%\InTextWith{\metaVar{IT}(\C_0) =
+%\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}\\
+%\InTextDef{13ex}{
+%\get(\m,\C_0)}{
+%}{
+%\method
+%}\\
+%\InTextWith{\metaVar{IT}(\C_0) =
+%\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
 \end{array}$

### 2015-12-06 paper/sections/Translation.tex
section 5.5 reviewd

@@
 \subsection{Guaratees}
-Now we try to understand what is the meaning of saying that our approach is correct, and more in general, what it means to say that a language tuning is correct.\haoyuan{Actually only self coherence is supported?}
-We identify three types of guarantees:
+To understand to what extents our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
 \begin{itemize}
 \item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not wrote.
-In our case it means that either \mixin{} produces in controlled way an understandable error, or the class can be successfully annotate and the generated code is well typed.
+In our case it means that either \mixin{} produces (in a controlled way) an understandable error, or the class can be successfully annotated and the generated code is well typed.
 We guarantee \textit{Self coherence}.
@@
 \mixin  does not guarantee \textit{Heir coherence}.
-Indeed consider the following example:
+Indeed consider the following:
 
@@
 \noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@. 
-Similarly, an expression of the form \Q@new B(){.. A withX(int x){..}}@ would compile before translation, but would be ill typed after the translation.
+This would break \Q@C@. \\*
+Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@'' would compile before translation, but would be ill typed after the translation.
 \end{itemize}
 
-\noindent Our automatic type refinement 
-is a usefull and convenient feature, but is not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantic  to be well typed.
+\noindent This means that our automatic type refinement 
+is a usefull and convenient feature, but is not transparent to the heirs of the annotated interface.
+They need to be aware of the annotation semantic  and provide the right type while refining methods.
 
@@
 \begin{itemize}
-\item We say that an interface table 
+\item We denote with $\C^{\II}$ the name of an interface.
+\item An interface table 
 IT is OK if under such interface table, all interfaces are OK.
-\item Since interface tables are just represented as sequences of interfaces we can write IT = $\II$ IT' to talk about a specific interface in a table.
-\item We denote with $\C^{\II}$ the name of an interface.
-\item An interface table contains a hair of $\C$ if there is any interface that extends it, or any \Q@new@ that instantiate it.
+\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
+\item IT contains a heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
 \end{itemize}
@@
 Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
-The formal theorems proofs are available in Appendix~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
+The formal theorems proofs are available in Appendix.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
 

### 2015-12-06 paper/paper.tex
Reviewd abstract and intro

@@
 approaches. Interestingly, object interfaces do not require changes to
-the runtime, and they also do not introduce any new syntax (apart from
-a small annotation): all three features are achieved by reinterpreting
+the runtime, and they also do not introduce any new syntax: all three features are achieved by reinterpreting
 existing Java syntax, and are translated into regular Java code

### 2015-12-06 paper/sections/Introduction.tex
Reviewd abstract and intro

@@
 we show that, with a simple language feature, default methods and
-interface inheritance are in fact be very expressive. Our proposed
+interface inheritance are in fact very expressive. Our proposed
 language feature enables powerful object-oriented idioms, using
@@
 to be type-preserving, ensuring that the translation does not
-introduce type-errors. 
+introduce type-errors in client code. 
 %CJ's usefulness goes beyond serving as a
@@
 extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
-application is shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
+application shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
 can be easily defined using object interfaces. The last
@@
 can be used to formally define the translation of object interfaces. 
-We prove that several properties hold for the translation, including a type-preservation.
+We prove several properties of the translation.
 
@@
 how other existing approaches, such as the annotations in project 
-Lombok~\cite{lombok}, which can be viewed as language tuning.
+Lombok~\cite{lombok},  can be viewed as language tuning.
 

### 2015-12-06 paper/paper.tex
Section 2 revied

@@
   morekeywords={public,int,interface,implements,default,
-    abstract,case,catch,class,def,static,%
+    abstract,case,void,catch,class,def,static,%
     do,else,extends,false,final,finally,%
@@
 
-\newcommand\mixin{\textbf{@Obj}\xspace}
+\newcommand\mixin{\mixinAnn\xspace}
 \renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}

### 2015-12-06 paper/sections/Overview.tex
Section 2 revied

@@
 
-This section illustrates how the \mixin annotation enables expressive
+This section illustrates how the \mixinAnn{} annotation enables expressive
 idioms using multiple inheritance. 
@@
 
-Before we model the final animal system, we first start with a
+Before modelling the complete animal system, we  start with a
 simplified version without locations. This version serves the purpose of illustrating how 
@@
 To model the animal system with object interfaces all that a user
-needs to do is to add a \mixin annotation to the \texttt{Horse}, 
+needs to do is to add a \mixinAnn{} annotation to the \texttt{Horse}, 
 \texttt{Bird}, and \texttt{Pegasus} interfaces:
@@
 shows the code correponding to the \texttt{Pegasus} interface
-after the \mixin annotation is processed:
+after the \mixinAnn{} annotation is processed:
 
@@
 system presented in the left-side of Figure~\ref{fig:pegasus}, and still does not appear to 
-justify the \mixin annotation. Lets now move on to modelling 
-the complete animal system. In order to do this animals will include 
-a notion of points, representing their location in space. As we shall
+justify the \mixinAnn{} annotation. Lets now move on to modelling 
+the complete animal system, where \Q@Animal@s will include a \Q@location@, representing their position in space. As we shall
 see, modelling stateful components using plain Java 8 quickly becomes 
-cumbersome. The \mixin annotation comes to rescue here and 
+cumbersome. The \mixinAnn{} annotation comes to rescue here and 
 allows avoiding significant amounts of boilerplate code.
@@
   int x();    int y(); // getters
-  Point2D withX(int val);    Point2D withY(int val); // with methods }
+  Point2D withX(int val);    Point2D withY(int val); /* with methods*/ }
 \end{lstlisting}
@@
 
-\begin{comment}
+%\begin{comment}
+\marco{I re-enabled this code, I think is needed for understandability}
 is expanded into
@@
 
-That is, we can provide implementation for methods in an interface, 
-and our annotation generate code only if the method \emph{needs} implementation. In this way is trivial for the programmer to personalize the behaviour if they have any special need.
-\end{comment}
+\noindent We can provide implementation for methods; 
+we generate code only for methods needing implementation. Thus, programmers can easily tailor the behaviour on their special needs.
+%\end{comment}
 
@@
 
-Many readers would now think that a language extension is needed. Instead, with object interfaces, 
-another approach is possible.
-%Instead, with our language \emph{restriction} the problem can be easily challenged\footnote{Often in programming languages ``freedom is slavery''.}.
+We are tempted to assume that a language extension is needed. 
+%Instead, with object interfaces, another approach is possible
+Instead, with our language \emph{restriction}, another approach is possible;often in programming languages ``freedom is slavery''.
 
@@
 \noindent From the type perspective, the key is the covariant method
-overriding of \texttt{location()}. However from the semantic
+overriding of \texttt{location()}. However, from the semantic
 perspective the key is the implementation for the setter with the old
 signature (\Q@location(Point2D)@). The key to the setter
-implementation is a new type of \@with@ method, called a property updater.
+implementation is a new type of \Q@with@ method, called a property updater.
 
@@
 
-\noindent Note that this interface includes a different type of
-\Q@with@ method, which takes a \Q@Point2D@ as an argument.
-The old wither methods (such as \Q@withX@) functionally update a field at a time.  This can be
+\noindent \Q@Point3D@ includes a
+\Q@with@ method, taking a \Q@Point2D@ as an argument.
+Other wither methods (such as \Q@withX@) functionally update a field at a time.  This can be
 inefficient, and sometimes hard to maintain.  Often we want to update multiple
@@
 current object with all fields that match the fields of parameter type
-updated. The idea is that we want the result to still be like \Q@this@, but
+updated. The idea is that the result should be like \Q@this@, but
 modified to be as similar as possible to the parameter.
@@
 \subsection{Advanced Multiple Inheritance with Object Interfaces}
-Finally, we can define \texttt{Pegasus} as simple as we did in the simplified
-(and stateless) version on the right-side of Figure~\ref{fig:pegasus}:
-
-\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
-
-\noindent Note how even the non trivial pattern for field type refinement is
+Finally, defining \texttt{Pegasus} is as simple as we did in the simplified
+(and stateless) version on the right-side of Figure~\ref{fig:pegasus}.
+ Note how even the non trivial pattern for field type refinement is
 transparently composed, and \texttt{Pegasus} has a \Q@Point3D@
@@
 % common subtype in order for \texttt{Pegasus} to exists.
+
+\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-12-06 paper/sections/Formalization.tex
Section 2 revied
Section 3 reviewd

@@
 \section{Interaction of Interface Methods with Interface Composition}
-Before formalizing Classless Java, it is helpful to informally discuss the behaviour
-of methods in Java 8 interfaces.
+%Before formalizing Classless Java, it is helpful to informally discuss the behaviour
+%of methods in Java 8 interfaces.
 
@@
 \subsection{Classifying Outcomes of Interface Composition}
-When interfaces are composed and methods with the same name (and
-signature) exist, there are 3 possible outcomes.
-We illustrate the possible outcomes using the following (correct) declarations:
+%When interfaces are composed and methods with the same name (and
+%signature) exist, there are 3 possible outcomes.
+%
+We now try to classify possible outcomes for composition of methods with the same name (and
+signature).
+We will use the following (correct) declarations:
 \begin{lstlisting}
@@
 What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
-\Q@B@${}_j$? Nine representative cases are shown next:\\*
+\Q@B@${}_j$?\\*
+% Nine representative cases are shown next:\\*
 \noindent
@@
 \\*
-We try to classify the results in the table:
+%We try to classify the results in the table:
 \begin{itemize}

### 2015-12-06 paper/sections/Formalization.tex
revised S3 and S4

@@
 \section{Interaction of Interface Methods with Interface Composition}
-%Before formalizing Classless Java, it is helpful to informally discuss the behaviour
-%of methods in Java 8 interfaces.
+Before formalizing Classless Java and object interfaces, it is helpful
+to informally discuss the behaviour of methods in Java 8
+interfaces. In particular it is useful to understand how Java 8
+interfaces differ from conventional trait models.
 

### 2015-12-06 paper/sections/Overview.tex
Overview/Intro
revised S3 and S4

@@
 \end{figure}
+\bruno{Yanlin: We need space, but we cannot make the figure {\bf unreadable}; the
+  diagram fonts are too small now. Please increase the size of the
+  picture a bit. You can also use more space in the Java code. Since
+  you are on it please fix Point2D: there's no reason for the space
+  between x and y.}
 
-This section illustrates how the \mixinAnn{} annotation enables expressive
-idioms using multiple inheritance. 
+This section illustrates how object interfaces, expressed using the \mixinAnn{} annotation, enables powerful
+idioms using multiple inheritance in Java. 
 To propose a standard example for multiple inheritance, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
@@
 the basic system in given in the left side of Figure~\ref{fig:pegasus}.\footnote{
-Some research argues in favour of using subtyping for modelling taxonomies, other research argue against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
 
@@
 
-\noindent There is a some annoyances here. Firstly, the sole
+\noindent There are some annoyances here. Firstly, the sole
 purpose of classes is to provide a way to instantiate
@@
 Moreover, if the programmer wants a different implementation, he may 
-provide an alternative implementation using \Q@default@ methods:
+provide an alternative implementation using \Q@default@ methods. For example:
 \begin{lstlisting}
@@
 %\begin{comment}
-\marco{I re-enabled this code, I think is needed for understandability}
-is expanded into
+%\marco{I re-enabled this code, I think is needed for understandability}
+\noindent is expanded into
 \begin{lstlisting}
@@
 
-\noindent We can provide implementation for methods; 
-we generate code only for methods needing implementation. Thus, programmers can easily tailor the behaviour on their special needs.
+\noindent If an implementation is provided for methods,
+only code for methods needing implementation is generated. Thus,
+programmers can easily tailor the behaviour for their special needs.
 %\end{comment}
@@
 naming conventions and the types of the methods. Fluent setters are a variant of
-(void) setters, and are discussed in detail in Section~\ref{sec:dsls}.
+conventional setters, and are discussed in detail in Section~\ref{sec:dsls}.
 
@@
 
-We are tempted to assume that a language extension is needed. 
+We may tempted to assume that a language extension is needed. 
 %Instead, with object interfaces, another approach is possible
-Instead, with our language \emph{restriction}, another approach is possible;often in programming languages ``freedom is slavery''.
+Instead, the \emph{restriction} of (object) interfaces to have no
+fields, enlighten us that another approach is possible; often in programming languages ``freedom is slavery''.
 
@@
 properties updater: it takes a certain type of object and returns a copy of the
-current object with all fields that match the fields of parameter type
-updated. The idea is that the result should be like \Q@this@, but
+current object with all fields that match the fields of updated
+parameter type. The idea is that the result should be like \Q@this@, but
 modified to be as similar as possible to the parameter.
@@
 generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. The
-generated code is very mechanical, writing such code by hand is error-prone. For
+generated code is mechanical. Writing such code by hand is error-prone. For
 example a distracted programmer may swap the arguments of calls to

### 2015-12-06 paper/sections/Formalization.tex
Fixed again needed (sob)
revised S3 and S4

@@
 
-This section presents a formalization of ClassLess Java: a minimal
+This section presents a formalization of Classless Java: a minimal
 FeatherweightJava-like calculus which models the essence of Java
@@
 \end{grammar}
-\caption{Grammar of ClassLess Java}
+\caption{Grammar of Classless Java}
 \label{Grammar}
@@
 
-Figure~\ref{Grammar} shows the syntax of ClassLess Java.
+Figure~\ref{Grammar} shows the syntax of Classless Java.
 The syntax formalizes a minimal
@@
 \noindent Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
-That is, we allows return type specialization as introduced in Java 5.
+That is, we allow return type specialization as introduced in Java 5.
 Default methods are subtypes if their method headers are subtypes.
-Secondly, all abstract methods in the interface (that is methods that need
+Secondly, we check that all abstract methods in the interface (that is methods that need
 to be explicitly overridden) have been
@@
 
-\paragraph{override} models how the implementation in an interface can
-override implementation in the superinterface, even in case of a
+\paragraph{override} models how a method in an interface can
+override an implementation in the superinterface, even in case of a
 conflict.  Note how we use the special value $\none$, and how (forth
-case) overriding can solve a conflict. \hl{The notation in the last case below
-means $\method'$ is not static, and is supertype of $\method$.}
+case) overriding can solve a conflict. The notation in the last case below
+means $\method'$ is not static, and is supertype of $\method$.
 

### 2015-12-06 paper/sections/Formalization.tex
Recovered from conflicts, should be all ok
Doh, big fix in shadow and override

@@
 \override(\methods(\m),
-\shadow(\tops(\m,\Cs)))
+\tops(\m,\Cs))
 }\\
@@
 
-\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to three others auxiliary functions:
- $\tops(m,\C_0)$, $\shadow(\methods)$ and $\override(\method,\method')$.
+\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to two others auxiliary functions:
+ $\tops(m,\C_0)$%, $\shadow(\methods)$
+ and $\override(\method,\method')$.
 
@@
 Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
+In the definition we denote by $\Aux{originalMethod}(\m,\C)=\method$ the non-static method called $\m$ defined directly in $\C$.
 %\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
@@
 \noindent$\begin{array}{l}
-\InTextDef{27ex}{
-\method\in\tops(\m,\C_0\dots\C_n)}{
- }{
-\metaVar{IT}(\C_0)(\m)=\method,  \method\mbox{ default or abstract method}
+\InTextDef{20ex}{
+\Aux{originalMethod}(\m,\C_0)}{
+\method
 }\\
-\InTextWith{
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
+ \\
+
+\InTextDef{44ex}{
+\Aux{originalMethod}(\m,\C_0)\in\tops(\m,\C_0\dots\C_n)}{
+ }{}\\
+\tab\tab
 \not\exists i\in1..n \mbox{ such that } 
-\metaVar{IT}(\C_i)(\m)\mbox{ is a default or abstract method, and }
-\C_i \subtype \C_0}
+\Aux{originalMethod}(\m,\C_i) \mbox{ is defined and }
+\C_i \subtype \C_0
 
@@
 
-\paragraph{\shadow{}} chooses the most specific version of a method,
-that is the unique version available, or a conflicted version from a
-set of possibilities. Internally a special modifier $\conflicted$ is used to denote the case of two methods with conflicting implementation.
-We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
-\noindent
-$\begin{array}{l}
-\InTextDef{15ex}{
-\shadow()}{\none}\\
-\InTextDef{15ex}{\shadow(\method)}{\method}\\
-\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
-\InTextDef{15ex}{\shadow(\methods)}{\conflicted\mh\QM;}\\
-\InTextWith{\methods\mbox{ not of the form }\overline{\mh\QM;}\mbox{ and }
-\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
+%\paragraph{\shadow{}} chooses the most specific version of a method,
+%that is the unique version available, or a conflicted version from a
+%set of possibilities. In this case all the conflicting defintions are returned.
+%We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
+%
+%\noindent
+%$\begin{array}{l}
+%\InTextDef{15ex}{
+%\shadow()}{\none}\\
+%\InTextDef{15ex}{\shadow(\method)}{\method}\\
+%\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
+%\InTextDef{15ex}{\shadow(\methods)}{\methods}\\
+%\InTextWith{\mbox{ the former case not applicable} }
+%%\methods\mbox{ not of the form }\overline{\mh\QM;}
+%%\mbox{ and }
+%%\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
+%\end{array}$
+%
+\paragraph{override} models how a method in an interface can
+override an implementation in the superinterface, even in case of a
+conflict.  Note how we use the special value $\none$, and how (forth
+case) overriding can solve a conflict.
+% The notation in the last case below
+%means $\method'$ is not static, and is supertype of $\method$.
+
+\noindent$\begin{array}{l}
+\InTextDef{20ex}{\override(\none,\emptyset)}{\none}\\
+\InTextDef{20ex}{\override(\method,\emptyset)}{\method}\\
+\InTextDef{20ex}{\override(\none,\method)}{\method}\\
+\InTextDef{20ex}{\override(\none,\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
+\InTextDef{20ex}{\override(\method,\methods)}{\method}\\
+\InTextWith{
+\forall \method' \in \methods \ \ 
+%\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
+\method\subtype\method'
+}
 \end{array}$
@@
 others.  Since method subtyping is a partial ordering,
-$\Aux{mostSpecific}$ may not be defined, this in turn makes \shadow{},
-and the whole \mBody{} not defined for that specific $\m$. Rule
-\rn{t-intf} relies on this behaviour.
+$\Aux{mostSpecific}$ may not be defined, this in turn forces us to rely on the last clause of $\override$; otherwise
+the whole \mBody{} would be not defined for that specific $\m$.
+Rule \rn{t-intf} relies on this behaviour.
 
@@
 \InTextWith{\method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype \method'}
-\end{array}$
-
-\paragraph{override} models how a method in an interface can
-override an implementation in the superinterface, even in case of a
-conflict.  Note how we use the special value $\none$, and how (forth
-case) overriding can solve a conflict. The notation in the last case below
-means $\method'$ is not static, and is supertype of $\method$.
-
-\noindent$\begin{array}{l}
-\InTextDef{20ex}{\override(\none,\none)}{\none}\\
-\InTextDef{20ex}{\override(\method,\none)}{\method}\\
-
-\InTextDef{20ex}{\override(\none,\method)}{\method}\\
-\InTextWith{\method\mbox{ not of the form } \conflicted\ \mh\QM;}\\
-\InTextDef{20ex}{\override(\method,\method')}{\method}\\
-\InTextWith{\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
-\method\subtype\method'}
 \end{array}$

### 2015-12-06 paper/sections/RelatedWork.tex
Review 7-start of 8

@@
 cause difficulty (including the famous diamond (fork-join)
-problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning
+problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about 
 programs. To conciliate the need for expressive power and simplicity, many
@@
 line between units of reuse and object factories. Traits, as units of reusable
-code, contain only methods as reusable functionality. ignoring states and state
+code, contain only methods as reusable functionality, ignoring states and state
 initialization. Classes, as object factories, require functionalities from
-(multiple) traits. Java 8 interfaces with default methods are closely related to
+(multiple) traits. Java 8 interfaces are closely related to
 traits: concrete method implementation are allowed (via the \textbf{default}
@@
 for various flavours of multiple inheritance in Java. Traits offer an algebra
-with operations like sum, alias and exclusion, providing explicit conflict
+of composition operations like sum, alias and exclusion, providing explicit conflict
 resolution. Former work~\cite{bono14} provides details on mimicking the trait

### 2015-12-06 paper/sections/RelatedWork.tex
Reviewd up to 8. Last part of 8 have notes we need to discuss

@@
 purpose of SugarJ is to allow easy defition of sugars while our approach aims to
-implement the desired program transformation without just in Java.
+implement the desired program transformation in pure Java.
 
@@
 Java-like languages tend to sit in the middle of two extremes: libraries
-can not influence the type system, so many solutions valid in python or other
+can not influence the type system, so many solutions valid in python or other dynamic
 languages are not be applicable, or may be applicable at the cost of losing
@@
 In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code.
-This is roughly equivalent to what we state and manually proof in Lemma 2 for our particular transformation.
+This is roughly equivalent to
+\textit{Self coherence}, that we manually proof in the Appendix.
 Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.

### 2015-12-06 paper/sections/Conclusion.tex
First round over.
Do we cite cofoja ?

@@
   opportune to consider interfaces are pure types. We do not know how to conciliate
-  those two viewpoints and ClassLess Java design. We do not have
-  Classes purely in the Java sense.\bruno{Marco: you probably should
-    revise the conclusion}
+  those two viewpoints and ClassLess Java design.
+  We do not have classes purely in the Java sense.
+
+ Our language restriction encourage to code in a more flexible way by forcing to either keep a higher
+ abstraction level (interfaces are a more abstract concept than classes), or to rely on
+ concrete object initialization (the \Q@new I(){...}@ construct).
+
+ More importantly, we identify the concept of \emph{language tuning}.
+ We identify libraries that are already performing language tuning (Lombock and Cofoja), and 
+ we forecast many different kinds of language tuning will emerge on suitable platforms like Java or the C\# CLR.
+ We identify varius kinds of safety guarancees that can be offered by language tuning, but the door is open for more flavours of safety guarantees to emerge.

### 2015-12-06 paper/sections/Implementation.tex
review 6 (boh, is ok? non convincing but I do not know what to improve) and exp problem
More discussion on language tuning in section 6, plus CoFoJa

@@
 Our implementation is based on an extension of Lombok. The Lombok
-project~\cite{lombok} is a Java tool that aims at removing (or
+project~\cite{lombok} is a Java library that aims at removing (or
 reducing) Java boilerplate code via
@@
 studied in a formal way.
+
+\paragraph{Lombok do Language tuning}
+We consider Lombock to be the most developed example of language tuning; while they do not introduce a specific term for what they are doing,
+their slogan \emph{``Spice up your java''} seams to be in line with the philosophy of language tuning.
+Some other example of language tuning in Lombok include the \Q@val@ type, similar to \Q@auto@ in C\# or C++04.
+Another library doing language tuning is CoFoJa~\cite{}, where annotations are used to insert pre-post conditions in generated bytecode.

### 2015-12-06 paper/sections/Introduction.tex
Overview/Intro
revised S1 S2 (typos)

@@
   interfaces}, because such interfaces can be instantiated directly,
-without the need for an explict class definition. Moreover, object
+without the need for an explicit class definition. Moreover, object
 interfaces support \emph{abstract state operations}, providing a way
@@
 
-\item{\bf Language Tuning} We identify the concept of language tuning 
+\item{\bf Language Tuning:} We identify the concept of language tuning 
 and we describe object interfaces as an example. We also discuss 

### 2015-12-06 paper/sections/Overview.tex
refined figure
revised S1 S2 (typos)

@@
 \begin{tabular}{c|c}
-\includegraphics[height=3.5cm]{PegasusDetail.pdf}\hspace{20pt} &
+\includegraphics[height=3.8cm]{PegasusDetail.pdf}\hspace{20pt} &
 \begin{minipage}{7cm}
@@
 idioms using multiple inheritance in Java. 
-To propose a standard example for multiple inheritance, we show \Q@Animal@s with a two dimensional \Q@Point2D@ representing their \Q@location@.
+To propose a standard example for multiple inheritance, we show \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
 Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
-Birds can \Q@fly@, thus their location need to be a three dimensional \Q@Point3D@.
-Finally, we model \Q@Pegasus@ (one of the best known creatures in Greek
+Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s.
+Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
 mythology) as a kind of \Q@Animal@ with the skills
@@
 both \texttt{Horse} and \texttt{HorseImpl} are needed: \texttt{Horse}
-is needed because it needs to be an inferface so that \texttt{Pegasus}
+is needed because it needs to be an interface so that \texttt{Pegasus}
 can use multiple inheritance; and \texttt{HorseImpl} is needed to
@@
 constructor, which is missing from interfaces. The following code,
-shows the code correponding to the \texttt{Pegasus} interface
+shows the code corresponding to the \texttt{Pegasus} interface
 after the \mixinAnn{} annotation is processed:
@@
 The animal system modelled so far is a simplified version of the
-system presented in the left-side of Figure~\ref{fig:pegasus}, and still does not appear to 
-justify the \mixinAnn{} annotation. Lets now move on to modelling 
-the complete animal system, where \Q@Animal@s will include a \Q@location@, representing their position in space. As we shall
+system presented in the left-side of Figure~\ref{fig:pegasus}, which still does not appear to 
+justify the \mixinAnn{} annotation.  Now moving on to modelling 
+the complete animal system where \Q@Animal@s will include a \Q@location@
+representing their position in space, we shall
 see, modelling stateful components using plain Java 8 quickly becomes 
@@
 \noindent where we return a new point, using \Q@42@ as x-coordinate,
-and taking all the other informations (only \Q@y@ in this case) from
+and taking all the other information (only \Q@y@ in this case) from
 another point.
@@
 \noindent If there is a large number of fields, \texttt{with} methods
-will save programmers from writing large amount of tedious code that
+will save programmers from writing large amounts of tedious code that
 simply copies field values. 
@@
 Note how we actually generate a real field \Q@int x=_x;@.
-This provide a more uniform translation that can work also for mutable datastructures, where setters are required.
+This provides a more uniform translation that can work also for mutable data structures, where setters are required.
 }
@@
 be refined. Such kind of refinement usually poses a challenge
-in typical class-based approaches. Fortunatelly, with object interfaces,
+in typical class-based approaches. Fortunately, with object interfaces,
 we are able to provide a simple and effective solution.
@@
 
-We may tempted to assume that a language extension is needed. 
+We may be tempted to assume that a language extension is needed. 
 %Instead, with object interfaces, another approach is possible
@@
 (and stateless) version on the right-side of Figure~\ref{fig:pegasus}.
- Note how even the non trivial pattern for field type refinement is
+ Note how even the non-trivial pattern for field type refinement is
 transparently composed, and \texttt{Pegasus} has a \Q@Point3D@

### 2015-12-06 paper/sections/Formalization.tex
revised S3,4,5,6

@@
 static method over the $\this$ receiver is not considered.  A more
-interesting type of expressions are super calls
-($\C\QM{.super.}\m\QM(\es)$), whose semantic is to call the (non
-static) method $\m$ over the $\this$ receiver, but statically
+interesting type of expressions is super calls
+($\C\QM{.super.}\m\QM(\es)$), whose semantic is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
 dispatching to the version of the method as visible in the interface
@@
 $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
-For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquelly indentified by their name.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their name.
 
@@
 subtype of the field type. Secondly, we check that all method bodies are
-well-typed. To do this the enviroment used to check the method body
+well-typed. To do this the environment used to check the method body
 needs to be extended appropriately: we add all fields and their types;

### 2015-12-06 paper/sections/Implementation.tex
revised S3,4,5,6

--- 2015-12-02 paper/sections/Implementation.tex
@@
 studied in a formal way.
+
+\paragraph{Lombok does Language tuning}
+We consider Lombok to be the most developed example of language tuning; while they do not introduce a specific term for what they are doing,
+their slogan \emph{``Spice up your java''} seems to be in line with the philosophy of language tuning.
+Some other examples of language tuning in Lombok include the \Q@val@ type, similar to \Q@auto@ in C\# or C++04.
+Another library doing language tuning is CoFoJa~\cite{}, where annotations are used to insert pre-post conditions in generated bytecode.

### 2015-12-06 paper/sections/Translation.tex
revised S3,4,5,6

@@
 \subsection{$\ofMethod$}\label{subsec:ofmethod}
-We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory.
-To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.
+We now formally define $\ofMethod$, the function that generates the method
+\QM{of}, that behaves like a factory. To avoid boring digressions about
+well-known ways to find unique names, for the sake of this formalization we
+assume that no-args methods start with an underscore, and we prefix method names with underscore to obtain valid  parameter names.
 
@@
    \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
-    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
+    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This corresponds to the expressions $\es_i$.
 \item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
@@
 \begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not wrote.
-In our case it means that either \mixin{} produces (in a controlled way) an understandable error, or the class can be successfully annotated and the generated code is well typed.
+\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not written.
+In our case it means that either \mixin{} produces (in a controlled way) an understandable error, or the class can be successfully annotated and the generated code is well-typed.
 We guarantee \textit{Self coherence}.
 
-\item \textit{Client coherence}: all the client code (as for example method calls) that is well typed without the generation/instrumentation process is well typed also after the generation.
-That is, the annotation do not remove any functionality, is just
+\item \textit{Client coherence}: all the client code (as for example method calls) that is well-typed without the generation/instrumentation process is well-typed also after the generation.
+That is, the annotation does not remove any functionality, is just
 adding more behaviour.
@@
 
-\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well typed if they was well typed without the instrumentation.
+\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well-typed if they were well-typed without the instrumentation.
 This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
@@
 This would break \Q@C@. \\*
-Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@'' would compile before translation, but would be ill typed after the translation.
+Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@'' would compile before translation, but would be ill-typed after the translation.
 \end{itemize}
@@
 \subsection{Results}
-To formally characterize the behaviour of our annotation, and the two level of guarancees that we offer, we provide some notations and two theorems:
+To formally characterize the behaviour of our annotation, and the two level of guarantees that we offer, we provide some notations and two theorems:
 \begin{itemize}
@@
 \item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
-\item IT contains a heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
+\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
 \end{itemize}

### 2015-12-06 paper/sections/CaseStudy.tex
review 6 (boh, is ok? non convincing but I do not know what to improve) and exp problem
Review 7-start of 8
revised S7,8,9,10

@@
 simple way to solve the Expression Problem, while supporting multiple,
-independent extensions in Java.  The second case study show how to model an
+independent extensions in Java.  The second case study shows how to model an
 embedded DSL for SQL languages with fluent interfaces. Finally, the third 
@@
 can even provide independent extensibility~\cite{zenger05independentlyextensible}: the ability to assemble a system
-from multiple, independently developed extensions. Unfortunatelly, the
+from multiple, independently developed extensions. Unfortunately, the
 required instantiation code makes a plain Java solution verbose and cumbersome
@@
 should add the pretty printing operation in a type-safe and modular way.  It
-turns out to be easily achieved with the assitance of \mixin.  The code in
+turns out to be easily achieved with the assistance of \mixin.  The code in
 Figure~\ref{fig:ep} (rhs) shows how to add the new operation \texttt{print}.
@@
 \textbf{extends} clauses, no glue code is required. Return types of
-\texttt{e1,e2} are also automatically refined to \texttt{ExpPC}.
+\texttt{e1,e2} must be refined to \texttt{ExpPC}.
 
@@
 \end{lstlisting}
-\noindent In contrast, in a pure Java solution, the tedious instantiation code would need 
+\noindent Without our approach, tedious instantiation code would need 
 to be defined manually.
@@
 \lstinputlisting[linerange=37-37]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\marco{put on two lines}
 

### 2015-12-06 paper/sections/Conclusion.tex
revised S7,8,9,10

@@
  More importantly, we identify the concept of \emph{language tuning}.
- We identify libraries that are already performing language tuning (Lombock and Cofoja), and 
+ We identify libraries that are already performing language tuning (Lombok and Cofoja), and 
  we forecast many different kinds of language tuning will emerge on suitable platforms like Java or the C\# CLR.
- We identify varius kinds of safety guarancees that can be offered by language tuning, but the door is open for more flavours of safety guarantees to emerge.
+ We identify various kinds of safety guarantees that can be offered by language tuning, but the door is open for more flavours of safety guarantees to emerge.

### 2015-12-06 paper/sections/FutureWork.tex
First round over.
Do we cite cofoja ?
revised S7,8,9,10

@@
 We are unsure about the value of this workaround because of its trickiness. In
-order to define a state with initialization, users have to defining a method trusting that it will be overridden later with a behavioural change, but
+order to define a state with initialization, users have to define a method trusting that it will be overridden later with a behavioural change, but
 such change is visible only after the first time the setter is called. Moreover, this code would cache
 the result instead of re-computing it every time. This can be very tricky in a
-non functional setting.
+non-functional setting.
 %\bruno{Yanlin please polish text/break long sentences.}
@@
 \begin{lstlisting}
-interface Point{ ...
+interface Point{ ...//as before
   public static Point of(int _x, int _y){

### 2015-12-06 paper/sections/RelatedWork.tex
revised S7,8,9,10

@@
 models in the OO paradigm. In terms of restrictions set on these models, C++
-virtual inheritance aims at a relative general model; mixin model add some
+virtual inheritance aims at a relative general model; mixin model adds some
 restrictions on the model; and trait model is the most restricted one (excluding
@@
 \end{comment}
-\noindent Java interfaces and default methods does not use
+\noindent Java interfaces and default methods do not use
 linearisation: the semantics of Java \textbf{extends} clause in
@@
 not only the number of classes (from 4 to 6 in the stream example), but also the
-class hirarchy complexity increases. \mixin does not complicate the
+class hierarchy complexity increases. \mixin does not complicate the
 structure, and fields can also coexist with multiple inheritance.
@@
 offered by interfaces. Since Java was designed for classes, a direct classless
-programming style is verbose and unnatural. However, annotation driven code
+programming style is verbose and unnatural. However, annotation-driven code
 generation is enough to overcome this difficulty, and the resulting programming
@@
 %. A \texttt{clone} method is generated for the interface, returning a copy of the current object.
-Furthermore, we do automatic return type refinement for these kind of methods.
+Furthermore, we do automatic return type refinement for this kind of methods.
 This feature is comparatively useful in big examples, making operations and behaviours more flexible.%, which we will demonstrate later.
 \item We provide two options for generating setters.
-There are two kind of setters which are commonly used, namely \textit{void setters} and \textit{fluent setters}.
+There are two kinds of setters which are commonly used, namely \textit{void setters} and \textit{fluent setters}.
 The only difference is that a fluent setter returns the object itself after setting, thus supporting a pipeline of such operations.
@@
 In certain situations, object interfaces allow automatic refinement for \emph{return
-types}. This is part of a bigger topic in class based languages: expressing and
+types}. This is part of a bigger topic in class-based languages: expressing and
 preserving type recursion and (nominal/structural) subtyping at the same time.
@@
 is just using Java. However its intention is different from ours: one of the
-purpose of SugarJ is to allow easy defition of sugars while our approach aims to
+purposes of SugarJ is to allow easy defition of sugars while our approach aims to
 implement the desired program transformation in pure Java.
@@
 can not influence the type system, so many solutions valid in python or other dynamic
-languages are not be applicable, or may be applicable at the cost of losing
+languages are not applicable, or have the cost of losing
 safety.
@@
 class and providing getters), then we can insert in the compilation process a tool
-that examins and enrichs the code before compilation. No need to modify the
+that examines and enriches the code before compilation. No need to modify the
 original source; for example we can work on temporary files. Java is a particular
@@
 define and integrate such tools in its own compilation process: in this way
-there is no need of temporary files, and there is a well defined way of putting
+there is no need of temporary files, and there is a well-defined way of putting
 multiple extensions together.
@@
 In particular, MetaFjig offers a property called \emph{meta-level-soundness}.
-In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code.
+In short this property ensures by construction that library code (even if wrong or unreasonable) would never generate ill-typed code.
 This is roughly equivalent to
 \textit{Self coherence}, that we manually proof in the Appendix.
-Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.
+Since MetaFjig is not working on annotated classes, there is no direct
+equivalence on the overall theorem of the shown safety.
 

### 2015-12-07 paper/sections/Translation.tex
Revised Section 5

@@
 
-We now show what the \mixin annotation generates. We present a formal definition for
-most of the generated methods; however in our formalism we do not consider
+This section shows what the \mixin annotation generates, and presents a 
+formal definition for most of the generated methods. Since 
+Classless Java does not consider
 % setters, so we do not formally define the generation of the setters.
 %We do not include
-casts or \Q@instanceof@, so we do not include the \Q@with@ method.
-For the same reason we do not include \Q@void@ returning setters, since they are just a minor variation over the more interesting fluent setters, and they would require special handling just for the conventional \Q@void@ type.
+casts or \Q@instanceof@, the \Q@with@ method is not included in the
+formal translation. For the same reason \Q@void@ returning setters are
+not included, since they are just a minor variation over the more
+interesting fluent setters, and they would require special handling
+just for the conventional \Q@void@ type.
+\bruno{Haoyuan: ObjWeak is a poor name. I suggest ObjOf. 
+Please revise (and also revise the proof).}
 
-We chose to modularly divide our translation in two parts, so that we can discuss about formal properties of those parts later.
-To this aim we introduce the annotation \Q:@ObjWeak:.
-Its role is only in the translation process, hence is not part of the classless java language.
-\Q:@ObjWeak: generates the constructor method $\QM{of}$, while
-\mixin automatically refines the return types and calls \Q:@ObjWeak:.
+\subsection{Translation}
+
+For the purposes of the formalization, the translation is divided in
+two parts.  The two parts are important so that we can discuss the
+formal properties later. To this aim we introduce the annotation
+\Q:@ObjWeak:. Its role is only in the translation process, hence is
+not part of the Classless Java language.  \Q:@ObjWeak: generates the
+constructor method $\QM{of}$, while \mixin automatically refines the
+return types and calls \Q:@ObjWeak:.
 
@@
 
-As you see, we explicitly check if the interface is valid for annotation:
+\noindent Note that it is necessary to explicitly check if the interface is valid 
+for annotation:
 
@@
 %$\isClone(\method, \C_0)$.
+\bruno{Isn't this a definition; Shouldn't it have the bullet points before as 
+all other definitions?}
 
-That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.
+\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.\bruno{I don't understand what the previous sentence means!}
 Moreover, we check that the method \Q@of@ is not already defined by the user.
-In our simplified formalization we consider this to be just an error.
-In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.
+In the formalization an existing definition of the \Q@of@ method is an error. However,
+in the prototype (which also needs to account for overloading), the check is more complex
+as it additionally checks that an \Q@of@ method with the same signature of the one being 
+generated is not already present.
 
-In the following we will write $\QM{with#}\m$ to append $\m$ to \QM{with}, following the camelCase rule, so the first letter of
-$\m$ must be lower-case and is turned in upper-case upon merging.
-For example \QM{with#foo}=\QM{withFoo}.
-Special names $\specialName(\m)$ are  \QM{with} and all the identifiers of form $\QM{with#}\m$.
+In the following we will write $\QM{with#}\m$ to append $\m$ to
+\QM{with}, following the camelCase rule. That is, the first letter of $\m$
+must be lower-case and is turned in upper-case upon merging.  For
+example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
+are \QM{with} and all the identifiers of form $\QM{with#}\m$.
 
-\subsection{$\otherMethod$}
+\paragraph{The $\otherMethod$ method:}
 The definition of $\otherMethod(\C_0,\methods)$ is as follows:
@@
 
-The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
+\noindent The methods generated in the interface are \Q@with-@ and setters. %\Q@clone@.
 %A complete formalization would also generate the \Q@with@.
-We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only in order to refine the return type.
-To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
- is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$. \marco{no, overloading is not the issue here\hl{since overloading is not modelled}}.
+The methods are generated when they are unimplemented in $\C_0$, because
+the return types need to be refined.
+To discover whether the methods need to be generated, 
+we check if such \Q@with-@ or setter methods %\Q@clone@
+are required by $\C_0$, but are not already present in the methods directly declared in $\C_0$.
 
 \noindent$\begin{array}{ll}
-\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
-\isField(\method)\ \mand\
-%}\\\InTextWith{
-\method=\mBody(\_,\C_0)
-}\\
-
-\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
-\mnot\ \specialName(\m)}\\
 \InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
@@
 \end{array}$
+\bruno{no explanation of the above auxiliary definitions?}
 
-\subsection{$\ofMethod$}\label{subsec:ofmethod}
-We now formally define $\ofMethod$, the function that generates the method
-\QM{of}, that behaves like a factory. To avoid boring digressions about
+\paragraph{The $\ofMethod$ method:}\label{subsec:ofmethod}
+The $\ofMethod$ function generates the method
+\QM{of}, which behaves like a factory. To avoid boring digressions about
 well-known ways to find unique names, for the sake of this formalization we
@@
 
-The function $\fieldsFunc(\C_0)$ (defined before) denotes all the fields in the current interface.
-For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
+\noindent Note that, the function $\fieldsFunc(\C_0)$ denotes all the fields in the current interface:
+
+\noindent$\begin{array}{ll}
+\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
+\isField(\method)\ \mand\
+%}\\\InTextWith{
+\method=\mBody(\_,\C_0)
+}\\
+
+\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
+\mnot\ \specialName(\m)}\\
+\end{array}$
+
+\noindent For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
   \begin{itemize}
-   \item $\m_i$ is the field name, and have type $\C_i$.
-   \item $\m_i$\QM{()} is the getter, that just return the current field value.
+   \item $\m_i$ is the field name, and has type $\C_i$.
+   \item $\m_i$\QM{()} is the getter, and just returns the current field value.
    \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
@@
 
-We have not formally modelled non fluent setters and the \Q@with@ method; informally
+We do not formally model non fluent setters and the \Q@with@ method. 
+An informal explanation of how those methods are generated is given next:
 \begin{itemize}
@@
   \begin{itemize}
-    \item Check if exist method $\C\ \m$\Q@();@. If not, generate error (that is, is not $\valid(\C_0)$).
+    \item Check if method $\C\ \m$\Q@();@ exists. If not, generate error (that is, is not $\valid(\C_0)$).
     \item Generate implemented setter method inside \Q@of@:\\*
-           \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@
+           \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@\\*
     Note how there is no need to refine the return type for non fluent setters, thus we do not need to generate the method header in the interface body itself.
@@
    \item $\C$ must be an interface type (no classes or primitive types).
-    \item As for before, check that $\C'$ is a supertype of the current interface type $\C_0$.
+    \item As before, check that $\C'$ is a supertype of the current interface type $\C_0$.
     \item Generate implemented \Q@with@ method inside \Q@of@:\\*
@@
 
-\subsection{Guaratees}
-To understand to what extents our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
+\subsection{Guarantees}
+To understand to what extent our approach is correct (and more generally, what it means to say that a language tuning is correct) we identify three types of guarantees:
 \begin{itemize}
@@
 \noindent This means that our automatic type refinement 
-is a usefull and convenient feature, but is not transparent to the heirs of the annotated interface.
-They need to be aware of the annotation semantic  and provide the right type while refining methods.
+is a useful and convenient feature, but is not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantic  and provide the right type while refining methods.
 
@@
 then the interface table $[\![\II]\!]$ IT is OK.
+\bruno{The formatting of the theorems seems rather odd; why are the lines breaking?
+Also naming the theorems after the annotations seems strange! Giving a proper name 
+to the theorems seems a much better idea.}
 
@@
 The formal theorems proofs are available in Appendix.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
+\bruno{For both theorems?}
 

### 2015-12-07 paper/sections/Implementation.tex
revised S6

--- 2015-12-02 paper/sections/Implementation.tex
@@
 studied in a formal way.
+
+\paragraph{Lombok does Language tuning}
+We consider Lombok to be the most developed example of language
+tuning.  While the authors of Lombok do not introduce a specific term
+for what they are doing, their slogan \emph{``Spice up your java''}
+seems to be in line with the philosophy of language tuning. Some
+other examples of language tuning in Lombok include the \Q@val@ type,
+similar to \Q@auto@ in C\# or C++04.  Another library doing language
+tuning is CoFoJa~\cite{}\bruno{reference}, where annotations are used to insert
+pre-post conditions in generated bytecode.

### 2015-12-07 paper/sections/CaseStudy.tex
revised S7

@@
 %to the EP forbids this and it also forbids using casts. In other words, it should be possible 
-should add the pretty printing operation in a type-safe and modular way.  It
+should add the pretty printing operation in a type-safe and modular way. This
 turns out to be easily achieved with the assistance of \mixin.  The code in
-Figure~\ref{fig:ep} (rhs) shows how to add the new operation \texttt{print}.
+Figure~\ref{fig:ep} (on the right) shows how to add the new operation \texttt{print}.
 Interface \texttt{ExpP} extends \texttt{Exp} with the extra method
@@
 \texttt{e1} and \texttt{e2} are refined. If the types were not refined then
-\texttt{print()} method in \texttt{AddP} would fail to type-check.
+the \texttt{print()} method in \texttt{AddP} would fail to type-check.
 
@@
 cascading, more and more languages came to support fluent interfaces, including
-JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
+JavaScript, Java, C++, D, Ruby, Scala, etc. In most languages, to create fluent
 interfaces, programmers have to either handwrite everything or create a wrapper
@@
     this}, our approach with \mixin annotation removes this verbosity and
-  automatically generate fluent setters.
+  automatically generates fluent setters.
 \item The approach supports extensibility: the return types of fluent setters are
@@
 specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
-feature, and the return type of fluent setters in
+features, and the return type of fluent setters in
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
@@
 to model operations \texttt{symmetric sum}, \texttt{override} and
-\texttt{alias} as trait-oriented programming. The game has been
+\texttt{alias} from trait-oriented programming. The game has been
 implemented using plain Java 8 and default methods by Bono

### 2015-12-07 paper/sections/FutureWork.tex
revised bib and lots of small changes

@@
 
-\yanlin{future work on state initialization removed for now. Because I think the
-  other two future works are fine. But this one is still not clear and
-  controversial. The proposed workaround forces users to make decision: if he
-  wants state with initialization then he also has to define the setter. Or
-  else, it is ambiguous: could be interpreted as a state initializer or just
-  a normal default method.}
+% \yanlin{future work on state initialization removed for now. Because I think the
+%   other two future works are fine. But this one is still not clear and
+%   controversial. The proposed workaround forces users to make decision: if he
+%   wants state with initialization then he also has to define the setter. Or
+%   else, it is ambiguous: could be interpreted as a state initializer or just
+%   a normal default method.}
 \begin{comment}

### 2015-12-07 paper/sections/Translation.tex
revised bib and lots of small changes

--- 2015-12-06 paper/sections/Translation.tex
@@
 
-\subsection{Guaratees}
+\subsection{Guarantees}
 To understand to what extents our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:

### 2015-12-07 UseMixinLombok/src/jooq/DatabaseTest.java

@@ -34,7 +34,8 @@ Database query1 = Database.of().select("a, b").from("Table").where("c > 10");
     }
     static ExtendedDatabase runTest2() {
 //BEGIN_FLUENT_QUERY2
-ExtendedDatabase query2 = ExtendedDatabase.of().select("a, b").from("Table").where("c > 10").orderBy("b");
+ExtendedDatabase query2 = ExtendedDatabase.of().select("a, b").from("Table")
+                                               .where("c > 10").orderBy("b");
 //END_FLUENT_QUERY2
         return query2;
     }

### 2015-12-07 paper/sections/RelatedWork.tex
Related Wotk

@@
 models have been proposed, including C++ virtual inheritance,
-mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid model
+mixins~\cite{bracha90mixin}, traits~\cite{scharli03traits}, and hybrid models
 such as CZ~\cite{malayeri2009cz}.  They provide novel programming architecture
 models in the OO paradigm. In terms of restrictions set on these models, C++
-virtual inheritance aims at a relative general model; mixin model adds some
-restrictions on the model; and trait model is the most restricted one (excluding
+virtual inheritance aims at a relative general model; the mixin model adds some
+restrictions; and the trait model is the most restricted one (excluding
 states, instantiation, etc).
@@
 \paragraph{C++ Approach.}
-C++ tries to provide a general solution to multiple inheritance by virtual
-inheritance that deals with diamond problem by keeping only one copy of the base
-class~\cite{ellis1990annotated}, however suffers from object initialization
-problem as shown in~\cite{malayeri2009cz}. It bypasses all constructor calls to
-virtual superclasses, which would potentially cause serious semantic errors. In
-our approach, \mixin annotation processor has the full control over object
-initialization, whose mechanism is transparent to users. Moreover, customized
-factory methods are also allowed, if users are not satisfied with the default
-generated \texttt{of} method.
+C++ tries to provide a general solution to multiple inheritance by
+virtual inheritance, dealing with the diamond problem by keeping only
+one copy of the base class~\cite{ellis1990annotated}. However C++
+suffers from the object initialization problem~\cite{malayeri2009cz}.
+It bypasses all constructor calls to virtual superclasses, which can
+potentially cause serious semantic errors. In our approach, the \mixin
+annotation has full control over object initialization,
+and the mechanism is transparent to users. Moreover, customized factory
+methods are also allowed: if users are not satisfied with the default
+generated \texttt{of} method, they can implement their own.
 
 \paragraph{Mixins.}
-Mixins are a more restricted model than C++ approach. Mixins allow to name
+Mixins are a more restricted model than the C++ approach. Mixins allow to name
 components that can be applied to various classes as reusable functionality
@@
 \paragraph{CZ.}
-Malayeri and Aldrich proposed a model CZ~\cite{malayeri2009cz} which aims to do
-multiple inheritance without diamond problem. They divide inheritance into two
-separate concepts: inheritance dependency and implementation inheritance. Using
-a combination of \texttt{requires} and \texttt{extends}, a program with diamond
-inheritance can be transformed to one without diamonds. Moreover, fields and
-multiple inheritance can coexist. However as shown in the example in the
-article, the untangling of inheritance also untangles class structure. In CZ,
-not only the number of classes (from 4 to 6 in the stream example), but also the
-class hierarchy complexity increases. \mixin does not complicate the
-structure, and fields can also coexist with multiple inheritance.
+Malayeri and Aldrich proposed a model CZ~\cite{malayeri2009cz} which
+aims to do multiple inheritance without the diamond problem. They
+divide inheritance into two separate concepts: inheritance dependency
+and implementation inheritance. Using a combination of
+\texttt{requires} and \texttt{extends}, a program with diamond
+inheritance can be transformed to one without diamonds. Moreover,
+fields and multiple inheritance can coexist. However untangling
+inheritance also untangles the class structure. In CZ, not only the
+number of classes, but also the class hierarchy complexity
+increases. \mixin does not complicate the class structure, and state
+can also coexist with multiple inheritance.
 
@@
 (multiple) traits. Java 8 interfaces are closely related to
-traits: concrete method implementation are allowed (via the \textbf{default}
+traits: concrete method implementations are allowed (via the \textbf{default}
 keyword) inside interfaces. The introduction of default methods opens the gate
@@
 encoding; however we propose a different representation of \textbf{exclusion}.
-The author of ~\cite{bono14} agrees that our revised version is cleaner,
-typesafe and more direct.
+The first author of~\cite{bono14} agreed (via personal communication) that 
+our revised version for exclusion is cleaner, typesafe and more direct.
 
@@
 
-There are also proposals for extending Java (before Java8) with traits. For
+There are also proposals for extending Java with traits. For
 example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} extends
@@
 inheritance in Java.  Except for few, mostly syntactic details, their work can
-be emulated with Java8 interfaces. There are also extensions to the original
+be emulated with Java 8 interfaces. There are also extensions to the original
 trait model, which have operations that default methods and interfaces cannot
-model, such as method renaming in \cite{reppy2006foundation}, which breaks
+model, such as method renaming~\cite{reppy2006foundation}, which breaks
 structural subtyping.
@@
 separating the two notions leads to a lot of boilerplate code, and is quite
-limiting when multiple inheritance with states is required.  Abstract state
+limiting when multiple inheritance with state is required.  Abstract state
 operations avoid the key difficulties associated with multiple inheritance and
-states, while still being quite expressive.  Moreover the ability to support
-constructors adds additional expressivity, which is not available in approaches
+state, while still being quite expressive.  Moreover the ability to support
+constructors adds expressivity, which is not available in approaches
 such as Scala's traits/mixins.
@@
 
-\subsection{ThisType/MyType/Extensibility}
+\subsection{ThisType and MyType}
 %\marco{ this section is unfair.
@@
 related to our work: 1) any method generated inside of the \Q@of@ method is indeed
-non-inheritable since there is no class name to extend from; 2)exact
+non-inheritable since there is no class name to extend from; 2) exact
 statements (a form of wild-card capture on the exact run-time type) could
@@
 Admittedly, MyType greatly enhances the expressiveness and extensibility of
-object-oriented programming languages. We are using covariant-return types to
-simulate some uses of MyType. But our approach only works on method positive
-positions, whereas MyType is more general, as it works on any
-positions. Nevertheless our approach is still useful for modeling fluent
-interfaces and solving expression problems,etc.
+object-oriented programming languages. Object interfaces use covariant-return types to
+simulate some uses of MyType. But our approach only works for refining
+return types, whereas MyType is more general, as it also works for
+parameter types. Nevertheless, as illustrated with our examples and 
+case studies, object interfaces are still very useful in many 
+practical applications.
 % The addition of MyType to a language will allow easy definition of
@@
 
-\subsection{Meta-programming Competes with Language Extensions}
-The most obvious solution to adding features to a language is language
-extension. It is often implemented as syntactic extensions that can be desugared
-to the base language. For example, the Scala compiler was extended to directly
-supports XML syntax. However, this approach does not support combining multiple
-extensions into one. Scala-Virtualized is an extension to Scala which allows to
-blend shallow and deep embedding of DSLs. It redefines Scala structures to
-method calls, which can be overridden by DSL implementors. Therefore
-Scala-Virtualized also reinterprets syntax. These extensions are admittedly
-useful, however they are essentially creating a fork of the language, and rarely
-the new fork gains enough traction to become the main language release. % On this
+\subsection{Meta-Programming Competes with Language Extensions}
+The most obvious solution to adding features to a language is via syntactic
+extensions. Syntactic extensions are often implemented as 
+desugarings to the base language. For example,
+the Scala compiler was extended to directly support XML syntax.
+However, when syntactic extensions are independently created it is hard
+to combine multiple extensions into one.
+SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language that
+aims at making the creation and composition of syntactic sugar
+extensions easy, by allowing programmers to extend Java with custom
+features (typically for DSLs). However SugarJ's goals are different 
+from language tuning: SugarJ aims at creating and composing new syntax; 
+whereas language tuning merely reinterprets existing syntax. It is 
+clear that reinterpreting existing syntax only can be limiting 
+for some applications. However, when language tuning is possible 
+it has the advantage that existing tools for the language work 
+out-of-the-box (since the syntax is still the same); and composition
+of independently developed tunings is straightforward.
+
+Scala-Virtualized~\cite{}\bruno{reference!} is an extension to Scala, 
+which allows blending shallow and deep embedding of DSLs. It redefines 
+some of Scala's language constructs to method calls, 
+which can be overridden by DSL implementors. Thus Scala Virtualized  
+can also reinterpret syntax, and be seen as a form of language tuning. 
+However, although many Scala's language constructs are supported, 
+not all language constructs can be virtualized.
+
+%These extensions are admittedly useful, however they are essentially 
+%creating a fork of the language, and rarely the new fork gains enough 
+%traction to become the main language release. % On this
 % topic we mention
-SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language allowing
-programmers to extend it with custom features by definitions in meta-DSLs (SDF,
-Stratego, etc). Using SugarJ requires learning various DSLs while our approach
-is just using Java. However its intention is different from ours: one of the
-purposes of SugarJ is to allow easy defition of sugars while our approach aims to
-implement the desired program transformation in pure Java.
 
-On the other side, when the base language has a flexible enough syntax and
-a fast and powerful enough reflection mechanism, we may just need to play with
-operator overloading and other language tricks to discover that the language
-feature we need can be expressed as a simple library in our language. For
-example, consider SQLAlchemy in python.
+When the base language has a flexible enough syntax and a fast and
+powerful enough reflection mechanism, we may just need to play with
+operator overloading and other language tricks to discover that the
+language feature we need can be expressed as a simple library in our
+language. An example of this is SQLAlchemy\cite{}\bruno{reference} in
+Python, which ...\bruno{does what? describe what SQLAlchemy is}
+without requiring any syntactic extensions to Python.
+Java-like languages tend to sit in the middle of two extremes:
+libraries can not influence the type system, so many solutions valid
+in Python or other dynamic languages are not applicable, or have the
+cost of losing type-safety.
 
-Java-like languages tend to sit in the middle of two extremes: libraries
-can not influence the type system, so many solutions valid in python or other dynamic
-languages are not applicable, or have the cost of losing
-safety.
-
-Here compile time code generation comes at the rescue: if for a certain feature
-(\mixin in our case) it is possible to use the original language syntax to
-\emph{express-describe} any specific instantiation of such feature (annotating a
+In Java-like languages compile time code generation comes at the rescue: if, for a certain feature
+(\mixin in our case), it is possible to use the original language syntax to
+\emph{express/describe} any specific instantiation of such feature (annotating a
 class and providing getters), then we can insert in the compilation process a tool
 that examines and enriches the code before compilation. No need to modify the
-original source; for example we can work on temporary files. Java is a particular
+original source (for example we can work on temporary files). 
+Java is a particular
 good candidate for this kind of manipulation since it already provide ways to
-define and integrate such tools in its own compilation process: in this way
+define and integrate such tools in its own compilation process via annotation
+processing. In this way
 there is no need of temporary files, and there is a well-defined way of putting
@@
 
-Other languages offer even stronger support to safe code manipulation:
+Other languages offer even stronger support for safe code manipulation:
 Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
-all allow to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
+all allow to execute code at compile time.
+They generate code/classes that is transparently integrated in the program that is being generated/processed/compiled.
 In particular, MetaFjig offers a property called \emph{meta-level-soundness}.
-In short this property ensures by construction that library code (even if wrong or unreasonable) would never generate ill-typed code.
+In short this property ensures by construction that library code (even if wrong or unreasonable) never generates ill-typed code.
 This is roughly equivalent to
-\textit{Self coherence}, that we manually proof in the Appendix.
+\textit{Self coherence}, that we have to manually prove.
 Since MetaFjig is not working on annotated classes, there is no direct
-equivalence on the overall theorem of the shown safety.
+equivalence on the overall theorem of the shown safety.\bruno{Don't understand 
+the end of the sentence!}
 

### 2015-12-07 paper/sections/Conclusion.tex
Revised Future Work and Conclusion

@@
 called Classless Java, where programs and (reusable) libraries can be
-easily defined and used.  To avoid for some syntactic boilerplate
+easily defined and used.  To avoid syntactic boilerplate
 caused by Java not being originally designed to work without classes,
@@
   interfaces are just a kind of classes. Others consider more
-  opportune to consider interfaces are pure types. We do not know how to conciliate
-  those two viewpoints and ClassLess Java design.
-  We do not have classes purely in the Java sense.
-
- Our language restriction encourage to code in a more flexible way by forcing to either keep a higher
- abstraction level (interfaces are a more abstract concept than classes), or to rely on
- concrete object initialization (the \Q@new I(){...}@ construct).
+  opportune to consider interfaces as pure types. We do not know how to conciliate
+  those two viewpoints and Classless Java design.
+  Classless Java does not have classes purely in the Java sense.  
+ Classless Java encourages coding in a more flexible way by either
+ keeping a higher abstraction level (interfaces are a more abstract
+ concept than classes), or relying on concrete object initialization
+ (the \Q@new I(){...}@ construct).
 

### 2015-12-07 paper/sections/FutureWork.tex
Revised Future Work and Conclusion

@@
 for class methods (private, protected, synchronized, etc.). For example, the absence
-of the support for private/protected methods in Java 8 interfaces forces all
+of support for private/protected methods in Java 8 interfaces forces all
 members of interfaces to be public, including static methods. Since we use

### 2015-12-07 paper/sections/Formalization.tex
typos

@@
 \begin{itemize}
-\item \textbf{conflict error} happens when the methods from both are implemented, which is also an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the methods from both are abstract, which is also considered correct in all
+\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
+\item \textbf{both abstract, accepted} happens when the methods from both interfaces are abstract, which is also considered correct in all
   trait models.

### 2015-12-07 paper/sections/Introduction.tex
revised bib and lots of small changes
typos

@@
 (and Java) interfaces with default methods. The translation is proved
-to be type-preserving, ensuring that the translation does not
+to be type-safe, ensuring that the translation does not
 introduce type-errors in client code. 
@@
 can be used to formally define the translation of object interfaces. 
-We prove several properties of the translation.
+We prove several properties of the translation\footnote{Proofs and prototype implementation are available in
+  the supplementary materials.}.
 

### 2015-12-07 paper/sections/RelatedWork.tex
sqlalchemy

@@
 language. An example of this is SQLAlchemy\cite{}\bruno{reference} in
-Python, which ...\bruno{does what? describe what SQLAlchemy is}
+Python, which uses operator overloading to dynamically turn normal python expressions into database queries
 without requiring any syntactic extensions to Python.

### 2015-12-07 paper/sections/Overview.tex
revised bib and lots of small changes

@@
 \end{figure}
-\bruno{Yanlin: We need space, but we cannot make the figure {\bf unreadable}; the
-  diagram fonts are too small now. Please increase the size of the
-  picture a bit. You can also use more space in the Java code. Since
-  you are on it please fix Point2D: there's no reason for the space
-  between x and y.}
 
@@
 representing their position in space, we shall
-see, modelling stateful components using plain Java 8 quickly becomes 
+see that modelling stateful components using plain Java 8 quickly becomes 
 cumbersome. The \mixinAnn{} annotation comes to rescue here and 

### 2015-12-07 paper/sections/Translation.tex

--- 2015-12-07 paper/sections/Translation.tex
@@
 
-\noindent$\valid(\C_0)$  holds if $\forall \m\in\dom(\C_0),$ if $\mh\QM; = \mBody(\m, \C_0),$ one case is satisfied:
-$\isField(\method)$,
-$\isWith(\method, \C_0)$
-or
-$\isSetter(\method,\C_0)$. \
-%$\isClone(\method, \C_0)$.
-\bruno{Isn't this a definition; Shouldn't it have the bullet points before as 
-all other definitions?}
+\noindent$\begin{array}{ll}
+\InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; = \mBody(\m, \C_0),\mbox{ one case is satisfied:}}\\
+\tab\tab
+\isField(\method), \isWith(\method, \C_0) \mbox{ or }
+\isSetter(\method,\C_0)\\
+\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
+\mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
+\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)}\\
 
-\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement.\bruno{I don't understand what the previous sentence means!}
+%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
+%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
+%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
+\end{array}$
+
+\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement: it is eiter a field getter, a with method or a setter method.
+
+\bruno{I don't understand what the previous sentence means!}
 Moreover, we check that the method \Q@of@ is not already defined by the user.
@@
 
-\noindent$\begin{array}{ll}
-\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
-
-%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
-%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
-%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-\end{array}$
-\bruno{no explanation of the above auxiliary definitions?}
-
 \paragraph{The $\ofMethod$ method:}\label{subsec:ofmethod}
@@
 \method=\mBody(\_,\C_0)
-}\\
-
-\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
-\mnot\ \specialName(\m)}\\
+}
 \end{array}$
@@
 \subsection{Guarantees}
-To understand to what extent our approach is correct (and more generally, what it means to say that a language tuning is correct) we identify three types of guarantees:
+To understand to what extents our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
+
 \begin{itemize}

### 2015-12-07 paper/sections/RelatedWork.tex
scala virtualized
sqlalchemy ref

--- 2015-12-07 paper/sections/RelatedWork.tex
@@
 
-Scala-Virtualized~\cite{}\bruno{reference!} is an extension to Scala, 
+Scala-Virtualized~\cite{Rompf2012scalavirtualized} is an extension to Scala, 
 which allows blending shallow and deep embedding of DSLs. It redefines 
@@
 language feature we need can be expressed as a simple library in our
-language. An example of this is SQLAlchemy\cite{}\bruno{reference} in
-Python, which ...\bruno{does what? describe what SQLAlchemy is}
+language. An example of this is SQLAlchemy\cite{bayer2012sqlalchemy} in
+Python, which uses operator overloading to dynamically turn normal python expressions into database queries
 without requiring any syntactic extensions to Python.

### 2015-12-07 paper/sections/Formalization.tex
latest

@@
 
-\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to two others auxiliary functions:
+\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to two other auxiliary functions:
  $\tops(m,\C_0)$%, $\shadow(\methods)$

### 2015-12-07 paper/sections/Translation.tex
formatting

@@
 \noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
-If a given interface table $\II$ IT is OK\\*
+If a given interface table $\II$ IT is OK 
  where $\II$ has \Q!@ObjWeak!, 
@@
 
-\noindent\textbf{THEOREM }\Q!@Obj!\\*
-If a given interface table $\II$ IT is OK\\*
+\noindent\textbf{THEOREM }\Q!@Obj! 
+If a given interface table $\II$ IT is OK 
  where $\II$ has \Q!@Obj!, 

### 2015-12-07 paper/sections/Appendix.tex
ObjWeak renamed to ObjOf. Small fixes. All annotations in code style.

@@
 \textbf{LEMMA 1. }
-For any expression $e$ under an interface table $\II$ IT where $\Gamma\vdash e\in\C^\II$, $\II$ has \textbf{@ObjWeak} annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\C^\II$.
+For any expression $e$ under an interface table $\II$ IT where $\Gamma\vdash e\in\C^\II$, $\II$ has $\weakAnn$ annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\C^\II$.
 \begin{proof}
@@
 The corresponding method declaration won't be ``removed'' by the translation, also the types remain unchanged.
-The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
+The only work $\weakAnn$ does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C^\II)$, so adding $\QM{of}$ method has no way to affect any formerly well typed method call.
 \item An object creation.
@@
 \textbf{LEMMA 2. }
-If $\II$ has \textbf{@ObjWeak} annotation and $\C^{\II}$ OK in $\II$ IT, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
+If $\II$ has $\weakAnn$ annotation and $\C^{\II}$ OK in $\II$ IT, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
 \begin{proof}
@@
 since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$,IT.
-The translation function delegate its work to \Q!@ObjWeak! in such way that we can refer to 
+The translation function delegate its work to $\weakAnn$ in such way that we can refer to
 Lemma2  to complete this part. Note that all the methods that are added (directly) by \mixin are abstract, and thus there is no body to typecheck.
@@
 
-\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
+\noindent\textbf{THEOREM }$\weakAnn$\\*
 If a given interface table $\II$ IT is OK\\*
- where $\II$ has \Q!@ObjWeak!, 
+ where $\II$ has $\weakAnn$,
 $\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
@@
 \textbf{LEMMA 2} already proves that $[\![\II]\!]$ is OK. On the other hand, for any $\II'\in $ IT$\setminus\II$, by \textbf{LEMMA 1}, we know that all its methods
-are still well-typed, and the generated code in translation of \textbf{@ObjWeak} is only a static method \QM{of}, which has no way to affect the domain
+are still well-typed, and the generated code in translation of $\weakAnn$ is only a static method \QM{of}, which has no way to affect the domain
 of $\II'$, so after translation rule \rn{t-Intf} can still be applied, which finishes our proof.
@@
 \begin{proof}
-Similar to what already argued for \textbf{THEOREM }\Q!@ObjWeak!
-we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finsh by \textbf{THEOREM }\Q!@ObjWeak!.
+Similar to what already argued for \textbf{THEOREM }$\weakAnn$
+we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finish by \textbf{THEOREM }$\weakAnn$.
 \end{proof}

### 2015-12-07 paper/sections/Formalization.tex
ObjWeak renamed to ObjOf. Small fixes. All annotations in code style.

@@
 \noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to two other auxiliary functions:
- $\tops(m,\C_0)$%, $\shadow(\methods)$
- and $\override(\method,\method')$.
+ $\tops(m,\Cs)$ and $\override(\method,\methods)$.
 

### 2015-12-07 paper/sections/Implementation.tex
revised bib and lots of small changes
added cofoja reference
ObjWeak renamed to ObjOf. Small fixes. All annotations in code style.

@@
 annotations. There are a number of annotations provided by the
-original Lombok, including \textbf{@Getter}, \textbf{@Setter},
-\textbf{@ToString} for generating getters, setters and \QM{toString}
+original Lombok, including \Q:@Getter:, \Q:@Setter:,
+\Q:@ToString: for generating getters, setters and \QM{toString}
 methods, respectively.  Furthermore, Lombok provides a number of
@@
 The Lombok project provides a set of predefined annotations, including constructor
-generators similar as ours (e.g., \textbf{@NoArgsConstructor},
-\textbf{@RequiredArgsConstructor} and \textbf{@AllArgsConstructor}). They
+generators similar as ours (e.g., \Q:@NoArgsConstructor:,
+\Q:@RequiredArgsConstructor: and \Q:@AllArgsConstructor:). They
 generate various kinds of constructors for \emph{classes}, with or without
@@
 used together with other features provided in Lombok (e.g.,
-\textbf{@Data}). Moreover, the implementation of these annotations in Lombok
+\Q:@Data:). Moreover, the implementation of these annotations in Lombok
 gives us hints on how to implement \mixin. However, none of these annotations
@@
 similar to \Q@auto@ in C\# or C++04.  Another library doing language
-tuning is CoFoJa~\cite{}\bruno{reference}, where annotations are used to insert
+tuning is CoFoJa~\cite{cofoja}, where annotations are used to insert
 pre-post conditions in generated bytecode.

### 2015-12-07 paper/sections/Translation.tex
ObjWeak renamed to ObjOf. Small fixes. All annotations in code style.

@@
 formal properties later. To this aim we introduce the annotation
-\Q:@ObjWeak:. Its role is only in the translation process, hence is
-not part of the Classless Java language.  \Q:@ObjWeak: generates the
+$\weakAnn$. Its role is only in the translation process, hence is
+not part of the Classless Java language.  $\weakAnn$ generates the
 constructor method $\QM{of}$, while \mixin automatically refines the
-return types and calls \Q:@ObjWeak:.
+return types and calls $\weakAnn$.
 
@@
 
-\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
+\InTextDef{16em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
 }{
@@
 
-\noindent\textbf{THEOREM }\Q!@ObjWeak!\\*
+\noindent\textbf{THEOREM }$\weakAnn$\\*
 If a given interface table $\II$ IT is OK
- where $\II$ has \Q!@ObjWeak!, 
+ where $\II$ has $\weakAnn$,
 $\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
@@
 Informally, this means that for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation.
-Conversely, the second step of \mixin, namely what \Q:@ObjWeak: does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
+Conversely, the second step of \mixin, namely what $\weakAnn$ does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
 The formal theorems proofs are available in Appendix.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\

### 2015-12-07 paper/sections/RelatedWork.tex
Draft for the section, formalization of Java8.

@@
 Field~\cite{goetz12fdefenders} to formalize defender (default) methods in
-Java. However this formalization is limited to model exactly one method inside
-classes/interfaces.\bruno{explain why having multiple methods is significantly
-  more challenging.} Our code generation generates several methods that may
-interacts with each other and methods defined by user. Thus implementing and
-formalizing multiple methods inside interfaces adds complexity greatly.
+Java.
+In their formalization, classes and interfaces can have only one method \Q@m()@ without arguments, so as to simplify method overloading and renaming.
+Our work expands and generalizes the situation to support multiple methods with some arguments inside interfaces, in which case the methods that \mixin generates are better modelled.
+Meanwhile, features like multiple inheritance of interfaces and reabstraction of default methods are also preserved in our formalization.
+Furthermore, one advantage of the \Q@of@ method provided by \mixin is that it's generated as a static method, considered as a clean way without affecting its heirs.
+It is indeed supported since static methods are included in our formalization.
+\bruno{explain why having multiple methods is significantly
+  more challenging.}
+%Our code generation generates several methods that may
+%interacts with each other and methods defined by user. Thus implementing and
+%formalizing multiple methods inside interfaces adds complexity greatly.
 \marco{I do not see this as a good explanation, I think we may just remove the `one method limitation discussion'.

### 2015-12-07 paper/sections/CaseStudy.tex
revised bib and lots of small changes
compressed to 25 pages (toggled out author notes) again.

@@
 As shown in Figure~\ref{fig:ep}, \texttt{Exp} is the common
-super-interface with an evaluation operation \texttt{eval()}
+super-interface with operation \texttt{eval()}
 inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend interface
@@
 %to the EP forbids this and it also forbids using casts. In other words, it should be possible 
-should add the pretty printing operation in a type-safe and modular way. This
+should add operations in a type-safe and modular way. This
 turns out to be easily achieved with the assistance of \mixin.  The code in
@@
 
-\lstinputlisting[linerange=37-37]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
-\marco{put on two lines}
+\lstinputlisting[linerange=37-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
 

### 2015-12-07 paper/sections/Conclusion.tex
compressed to 25 pages (toggled out author notes) again.

@@
   itching point in the long struggle of subtyping versus subclassing:
-  According to some authors, from a software engineering perspective,
+  according to some authors, from a software engineering perspective,
   interfaces are just a kind of classes. Others consider more

### 2015-12-07 paper/sections/Implementation.tex
compressed to 25 pages (toggled out author notes) again.

@@
 \item Moreover, and probably more importantly, Lombok is capable of generating
-  code \emph{inside} a class/interface. This is the ability that conventional
-  Java annotation processors do not provide.
+  code \emph{inside} a class/interface, which conventional
+  Java annotation processors do not support.
 \end{itemize}

### 2015-12-07 paper/sections/Overview.tex
compressed to 25 pages (toggled out author notes) again.

@@
 
-\noindent will be accepted and behave as expected. If the programmer wishes to suppress
-this behaviour and keep the signature as it was, it is sufficient to redefine
-the \Q@with-@ methods in the new interface repeating the old signature.  Again,
-the philosophy is that if the programmer provides something directly, \mixin
-does not touch it.  The cast in \Q@with(Point2D)@ is trivially safe since it is
-guarded by an \Q@instanceof@ test. The idea is that if the parameter is a
-subtype of the current exact type, then we can just return the parameter, as
-something that is just ``more'' than \Q@this@.
+\noindent will be accepted. If the programmer wishes to suppress this behaviour
+and keep the signature as it was, it is sufficient to redefine the \Q@with-@
+methods in the new interface repeating the old signature.  Again, the philosophy
+is that if the programmer provides something directly, \mixin does not touch it.
+The cast in \Q@with(Point2D)@ is trivially safe because of the \Q@instanceof@
+test. The idea is that if the parameter is a subtype of the current exact type,
+then we can just return the parameter, as something that is just ``more'' than
+\Q@this@.
 

### 2015-12-07 paper/sections/RelatedWork.tex
compressed to 25 pages (toggled out author notes) again.

@@
 such as diamond problem. However this approach has limited expressiveness, and
-suffers from all the problems of linearised mixin composition. Other languages, such as
-Python, also use linearized mixins.
+suffers from all the problems of linearised mixin composition. % Other languages, such as
+% Python, also use linearized mixins.
 \begin{comment}
@@
 be emulated with Java 8 interfaces. There are also extensions to the original
-trait model, which have operations that default methods and interfaces cannot
-model, such as method renaming~\cite{reppy2006foundation}, which breaks
-structural subtyping.
+trait model, with operations (e.g. renaming~\cite{reppy2006foundation}, which breaks
+structural subtyping) that default methods and interfaces cannot
+model.
 
@@
 Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
-all allow to execute code at compile time.
-They generate code/classes that is transparently integrated in the program that is being generated/processed/compiled.
-In particular, MetaFjig offers a property called \emph{meta-level-soundness}.
-In short this property ensures by construction that library code (even if wrong or unreasonable) never generates ill-typed code.
-This is roughly equivalent to
+all allow to execute code at compile time. They generate code that is
+transparently integrated in the program that is being
+generated/processed/compiled. In particular, MetaFjig offers a property called
+\emph{meta-level-soundness}, ensuring by construction that library code (even if wrong or unreasonable) never generates ill-typed code. This is roughly equivalent to
 \textit{Self coherence}, that we have to manually prove.

### 2015-12-07 paper/sections/RelatedWork.tex
local

@@
 restrictions; and the trait model is the most restricted one (excluding
-states, instantiation, etc).
+state, instantiation, etc).
 
@@
 line between units of reuse and object factories. Traits, as units of reusable
-code, contain only methods as reusable functionality, ignoring states and state
+code, contain only methods as reusable functionality, ignoring state and state
 initialization. Classes, as object factories, require functionalities from
@@
 \subsection{Formalization of Java8}
-We provide a simple formalization for a subset of Java including default/static
-interface methods and object initialization literals (often called anonymous
-local inner classes).  A similar formalization was drafted by Goetz and
-Field~\cite{goetz12fdefenders} to formalize defender (default) methods in
-Java.
-In their formalization, classes and interfaces can have only one method \Q@m()@ without arguments, so as to simplify method overloading and renaming.
-Our work expands and generalizes the situation to support multiple methods with some arguments inside interfaces, in which case the methods that \mixin generates are better modelled.
-Meanwhile, features like multiple inheritance of interfaces and reabstraction of default methods are also preserved in our formalization.
-Furthermore, one advantage of the \Q@of@ method provided by \mixin is that it's generated as a static method, considered as a clean way without affecting its heirs.
-It is indeed supported since static methods are included in our formalization.
-\bruno{explain why having multiple methods is significantly
-  more challenging.}
-%Our code generation generates several methods that may
-%interacts with each other and methods defined by user. Thus implementing and
-%formalizing multiple methods inside interfaces adds complexity greatly.
-\marco{I do not see this as a good explanation, I think we may just remove the `one method limitation discussion'.
-Is their work modelling interactions between abstract and default methods?
-like making a default method abstract again?}
-% As an evidence of the attention and care present in our formalization work,
-% while double-checking the behaviour of Java in side cases we have discovered a
-% likely bug in the current \texttt{javac}.\marco{refer to before when we explain
-%   the issue.}
+We provide a simple formalization for a subset of Java including
+default/static interface methods and object initialization literals
+(often called anonymous local inner classes).  A similar formalization
+was drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize
+defender (default) methods in Java. In their formalization, classes
+and interfaces can have only one method \Q@m()@ without arguments, so
+as to simplify method overloading and renaming. Classless Java is more
+general, as it supports multiple methods with arguments, it supports
+static methods, and features such as multiple inheritance of
+interfaces and reabstraction of default methods are also preserved.

### 2015-12-07 paper/sections/Translation.tex
local

@@
 \subsection{Guarantees}
-To understand to what extents our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
+To understand to what extent our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
 

### 2015-12-07 paper/sections/RelatedWork.tex
small things

@@
 language feature we need can be expressed as a simple library in our
-language. An example of this is SQLAlchemy\cite{bayer2012sqlalchemy} in
+language. An example of this is SQLAlchemy~\cite{bayer2012sqlalchemy} in
 Python, which uses operator overloading to dynamically turn normal python expressions into database queries
@@
 
-\subsection{Formalization of Java8}
+\subsection{Formalization of Java 8}
 We provide a simple formalization for a subset of Java including

### 2015-12-07 paper/sections/Translation.tex
Small fix theorem formatting

@@
 just for the conventional \Q@void@ type.
-\bruno{Haoyuan: ObjWeak is a poor name. I suggest ObjOf.
-Please revise (and also revise the proof).}
 
@@
 
-\bruno{I don't understand what the previous sentence means!}
 Moreover, we check that the method \Q@of@ is not already defined by the user.
@@
 in the prototype (which also needs to account for overloading), the check is more complex
-as it additionally checks that an \Q@of@ method with the same signature of the one being
+as it just checks that an \Q@of@ method with the same signature of the one being
 generated is not already present.
@@
 
-\noindent\textbf{THEOREM }$\weakAnn$\\*
+\noindent\textbf{THEOREM }$\weakAnn$ tuning\\*
 If a given interface table $\II$ IT is OK
  where $\II$ has $\weakAnn$,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,
 then the interface table $[\![\II]\!]$ IT is OK.
 
-\noindent\textbf{THEOREM }\Q!@Obj!
+\noindent\textbf{THEOREM }\mixin tuning\\*
 If a given interface table $\II$ IT is OK
  where $\II$ has \Q!@Obj!,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,
 then the interface table $[\![\II]\!]$ IT is OK.
-\bruno{The formatting of the theorems seems rather odd; why are the lines breaking?
-Also naming the theorems after the annotations seems strange! Giving a proper name
-to the theorems seems a much better idea.}
 

### 2015-12-07 paper/sections/RelatedWork.tex
Fixed sentences

@@
 \textit{Self coherence}, that we have to manually prove.
-Since MetaFjig is not working on annotated classes, there is no direct
-equivalence on the overall theorem of the shown safety.\bruno{Don't understand
-the end of the sentence!}
+Since MetaFjig is not working on annotated classes, there is not a ``semantic with/without annotations''. Our \mixin tuning theorem does not make sense in such context.
 
@@
 static methods, and features such as multiple inheritance of
-interfaces and reabstraction of default methods are also preserved.
+interfaces and reabstraction of default methods are also modelled.

### 2015-12-07 paper/sections/Translation.tex
Connecting text
revised Marco text

@@
 
-Informally, this means that for a client program that typechecks before a translation is applied, if the annotated type has no subtypes, and no objects of that type are created, then type safety is guaranteed after the successful translation.
-Conversely, the second step of \mixin, namely what $\weakAnn$ does in the formalization, is guaranteed to be type-safe for the three kinds of coherence.
-The formal theorems proofs are available in Appendix.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
-\bruno{For both theorems?}
+Informally, the theorems mean that for a client program that
+typechecks before a translation is applied, if the annotated type has
+no subtypes, and no objects of that type are created, then type safety
+is guaranteed after the successful translation. Conversely, the second
+step of \mixin, namely what $\weakAnn$ does in the formalization, is
+guaranteed to be type-safe for the three kinds of coherence.  \mixin
+tuning is more interesting: Since \mixin does not guarantee heir
+coherence, we exclude explicitly the presence of heirs. In this way
+\mixin tuning guarantees only self and client coherence. The formal
+theorem proofs are available in the supplementary materials.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
 

### 2015-12-07 paper/sections/FutureWork.tex
Added new version of supporting privateness

@@
 
+\subsection{Encapsulation}
+The biggest limitation of our approach is the inability to model visibility restrictions. For example, the absence
+of support for private/protected methods in Java 8 interfaces forces all
+members of interfaces to be public, including static methods. Since we use
+abstract methods to encode state, our state is always all public. Still, because 
+the state can only be accessed by methods, it 
+is impossible for the user to know if a certain method maps directly to a field
+or if it has a default implementation.  If the user wants a constructor that
+does not directly maps to the fields, (as for secondary constructors in Scala)
+he can simply define its own \Q@of@ method and delegate on the generated one:
+\begin{lstlisting}
+@Obj interface Point{  int x(); int y();
+  static Point of(int val){return Point.of(val,val);}  }
+\end{lstlisting}
+However, the generated \Q@of@ method would also be present and public.  If a
+future version of Java was to support \emph{static private methods in
+  interfaces} we could extend our code generation to handle also encapsulation.
+
+However, interfaces as a whole can have public or 
+package private (java default) visiblity.
+
+We can add a second annotation 
+\Q!@Exposed!
+that leverage on this edge:
+An
+interface without exposed works as usual, but if any method of a public \mixin interface is annoted with \Q!@Exposed!, we can apply a translation where a new  (package private) interface type is introduced, and the orginal interface become just a facade.
+For example:
+\begin{lstlisting}
+@Obj public interface Person{
+  void name(String val);
+  @Exposed default void rename(String newName){ if(/*valid name*/){ this.name(val);}}
+  @Exposed String name();
+  @Exposed static Person from(String val){ if(/*valid name*/){return Person.of(val);}
+    throw /*invalid name*/}  }
+\end{lstlisting}
+becomes
+\begin{lstlisting}
+public interface Person{
+  void rename(String newName)
+  String name();
+  static Person from(String val){ return Person$.from(val);} }
+
+@Obj interface Person$ extends Person{//will be further expanded by @Obj
+  void name(String val);
+  default void rename(String newName){ if(/*valid name*/){ this.name(val);}}
+  String name();
+  static Person from(String val){ if(/*valid name*/){return Person$.of(val);}
+    throw /*invalid name*/}  }
+\end{lstlisting}
+
+ This is not a perfect solution, since \Q@Person$@ can still be seen inside the \Q@Person@ package and heirs of \Q@Person$@, however it is surprising we achive such of a good result without any language support for privatness in interfaces.
+
+\begin{comment}
 \subsection{Qualifiers in Methods} %{Private state
@@
 method could be made private following the same pattern.
-
+\end{comment}
+\marco{old version still commented if we decided to swap back}
 % \yanlin{future work on state initialization removed for now. Because I think the

### 2015-12-07 paper/sections/Translation.tex
small fix

@@
 no subtypes, and no objects of that type are created, then type safety
-is guaranteed after the successful translation. Conversely, the second
+is guaranteed after the successful translation.
+
+The second
 step of \mixin, namely what $\weakAnn$ does in the formalization, is
-guaranteed to be type-safe for the three kinds of coherence.  \mixin
+guaranteed to be type-safe for the three kinds of coherence by the $\weakAnn$  tuning theorem.  
+\mixin
 tuning is more interesting: Since \mixin does not guarantee heir

### 2015-12-08 paper/sections/FutureWork.tex
S9.1 switched to old version for space reason, need to discuss

@@
 
+\begin{comment}
 \subsection{Encapsulation}
@@
 
-We can add a second annotation 
-\Q!@Exposed!
-that leverage on this edge:
-An
-interface without exposed works as usual, but if any method of a public \mixin interface is annoted with \Q!@Exposed!, we can apply a translation where a new  (package private) interface type is introduced, and the orginal interface become just a facade.
-For example:
+We can add a second annotation \Q!@Exposed!  that leverages on this edge: An
+interface without exposed works as usual, but if any method of a public \mixin
+interface is annoted with \Q!@Exposed!, we can apply a translation where a new
+(package private) interface type is introduced, and the orginal interface become
+just a facade.  For example:
 \begin{lstlisting}
@@
  This is not a perfect solution, since \Q@Person$@ can still be seen inside the \Q@Person@ package and heirs of \Q@Person$@, however it is surprising we achive such of a good result without any language support for privatness in interfaces.
+\end{comment}
 
-\begin{comment}
 \subsection{Qualifiers in Methods} %{Private state
@@
 method could be made private following the same pattern.
-\end{comment}
+
 \marco{old version still commented if we decided to swap back}

### 2015-12-08 paper/sections/CaseStudy.tex
spellcheck one more time

@@
 advantages of using the \mixin annotation. The first case study provides a
-simple way to solve the Expression Problem, while supporting multiple,
+simple way to solve the Expression Problem while supporting multiple,
 independent extensions in Java.  The second case study shows how to model an
@@
 \paragraph{Initial System}
-In the formulation of the EP there is an initial system that models
+In the formulation of the EP, there is an initial system that models
 arithmetic expressions with only literals and addition, and an initial 

### 2015-12-08 paper/sections/FutureWork.tex
spellcheck one more time

@@
 However, interfaces as a whole can have public or 
-package private (java default) visiblity.
+package private (java default) visibility.
 
@@
 interface without exposed works as usual, but if any method of a public \mixin
-interface is annoted with \Q!@Exposed!, we can apply a translation where a new
-(package private) interface type is introduced, and the orginal interface become
+interface is annotated with \Q!@Exposed!, we can apply a translation where a new
+(package private) interface type is introduced, and the original interface become
 just a facade.  For example:
@@
 
- This is not a perfect solution, since \Q@Person$@ can still be seen inside the \Q@Person@ package and heirs of \Q@Person$@, however it is surprising we achive such of a good result without any language support for privatness in interfaces.
+This is not a perfect solution, since
+\Q@Person$@ can still be seen inside the \Q@Person@ package and heirs of
+\Q@Person$@,
+however it is surprising we achieve such of a good result without any language
+support for privacy in interfaces.
 \end{comment}

### 2015-12-08 paper/sections/Translation.tex
spellcheck one more time

@@
 
-We do not formally model non fluent setters and the \Q@with@ method.
+We do not formally model non-fluent setters and the \Q@with@ method.
 An informal explanation of how those methods are generated is given next:
@@
            \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@\\*
-    Note how there is no need to refine the return type for non fluent setters, thus we do not need to generate the method header in the interface body itself.
+    Note how there is no need to refine the return type for non-fluent setters, thus we do not need to generate the method header in the interface body itself.
     \end{itemize}

### 2015-12-08 paper/sections/Overview.tex
typos
bundle. small fixes. move on to theorem macros.

@@
 of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
-the basic system in given in the left side of Figure~\ref{fig:pegasus}.\footnote{
+the basic system is given in the left side of Figure~\ref{fig:pegasus}.\footnote{
 Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
@@
 \noindent There are some annoyances here. Firstly, the sole
-purpose of classes is to provide a way to instantiate
-objects. Although (in this case) it takes only one line code to
+purpose of the classes is to provide a way to instantiate
+objects. Although (in this case) it takes only one line of code to
 provide each of those classes, this code is essentially boilerplate

### 2015-12-08 paper/sections/Introduction.tex
spellcheck one more time
manual check

@@
 default methods in Java 8 is \emph{interface evolution}. That is, to
-allow interfaces to be extended over time, while preserving backwards
+allow interfaces to be extended over time while preserving backwards
 compatibility. It soon became clear that default methods could also be
@@
 
-\item {\em Useful, general purpose methods, require special care in
+\item {\em Useful, general purpose methods require special care in
   the presence of subtyping.} Methods such as
@@
 implementation of object interfaces uses Java annotations and AST
-rewriting, allowing existing Java tools (such as IDE's) to work
+rewriting, allowing existing Java tools (such as IDEs) to work
 out-of-the-box with our implementation. As a result we could
@@
 
-\item {\bf ClassLess Java (CJ):} A simple formal calculus that models 
+\item {\bf Classless Java (CJ):} A simple formal calculus that models 
 the essential features of Java 8 interfaces with default methods, and 
@@
 \item{\bf Language Tuning:} We identify the concept of language tuning 
-and we describe object interfaces as an example. We also discuss 
+and describe object interfaces as an example. We also discuss 
 how other existing approaches, such as the annotations in project 

### 2015-12-08 paper/sections/Overview.tex
manual check

--- 2015-12-07 paper/sections/Overview.tex
@@
 \noindent There are some annoyances here. Firstly, the sole
-purpose of classes is to provide a way to instantiate
+purpose of the classes is to provide a way to instantiate
 objects. Although (in this case) it takes only one line code to
@@
 
-\noindent Unfortunatelly, creating a new point object is cumbersome, even
+\noindent Unfortunately, creating a new point object is cumbersome, even
 with anonymous classes:

### 2015-12-08 paper/sections/RelatedWork.tex
spellcheck one more time
manual check

@@
 code, contain only methods as reusable functionality, ignoring state and state
-initialization. Classes, as object factories, require functionalities from
+initialization. Classes, as object factories, require functionality from
 (multiple) traits. Java 8 interfaces are closely related to
@@
 %believe that in other languages the separation can be more practical
-%than in Java, and we do not necessarally advocate that merging the two
+%than in Java, and we do not necessarily advocate that merging the two
 %notions is a better approach in general.
@@
 %\marco{ this section is unfair.
-There are unsolved problems with ThisType in negative positions, namelly it breaks subtyping.
+There are unsolved problems with ThisType in negative positions, namely it breaks subtyping.
 The most common solution is to allow calling some methods only when the exact type is known.
-This demolish most advantages of interfaces.}
+This demolishes most advantages of interfaces.}
 
@@
 some of Scala's language constructs to method calls,
-which can be overridden by DSL implementors. Thus Scala Virtualized
+which can be overridden by DSL implementors. Thus Scala-Virtualized
 can also reinterpret syntax, and be seen as a form of language tuning.

### 2015-12-08 paper/sections/Translation.tex
manual check

@@
 
-\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement: it is eiter a field getter, a with method or a setter method.
+\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement: it is either a field getter, a with method or a setter method.
 

### 2015-12-08 paper/sections/Appendix.tex

@@
 \[\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i \vdash\QM{this}\in \C^\II\]
-        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
-        \haoyuan{Can we write $\Gamma, \QM{this}:\C^\II,\ \QM{_val} : \C_i, \C^{\mh_i} = \C^\II \vdash\QM{this}\in \C^\II = \C^{\mh_i}$?}\\
-        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C^\II$?}
-      \marco{Not without defining what that would mean. Do we need to? it seams good as it is now...}
+        We know that $\C^\II=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.
     \end{itemize}
@@
 \begin{proof}
-\noindent\textbf{Part I.} Similarly to what already argued for Lemma2,
-since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$,IT.
-The translation function delegate its work to $\weakAnn$ in such way that we can refer to
-Lemma2  to complete this part. Note that all the methods that are added (directly) by \mixin are abstract, and thus there is no body to typecheck.
+\noindent\textbf{Part I.} Similarly to what already argued for \textbf{Lemma 2},
+since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$ IT.
+The translation function delegates its work to $\weakAnn$ in such way that we can refer to
+\textbf{Lemma 2} to complete this part. Note that all the methods added (directly) by \mixin are abstract, and thus there is no body to typecheck.
 
 \noindent\textbf{Part II.}
-Similar to what already argued for Lemma2, but we need to notice that the the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
+Similar to what we already argued for \textbf{Lemma 2}, but we need to notice that the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
 Thus by last clause of the definition of $\override(\_)$, $\mBody(\_)$ is defined on the same method names.
@@
 
-\noindent\textbf{THEOREM }$\weakAnn$\\*
-If a given interface table $\II$ IT is OK\\*
+\noindent\textbf{THEOREM }$\weakAnn$ tuning\\*
+If a given interface table $\II$ IT is OK
  where $\II$ has $\weakAnn$,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,\\*
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,
 then the interface table $[\![\II]\!]$ IT is OK.
@@
 
-\noindent\textbf{THEOREM }\Q!@Obj!\\*
-If a given interface table $\II$ IT is OK\\*
+\noindent\textbf{THEOREM }\mixin tuning\\*
+If a given interface table $\II$ IT is OK
  where $\II$ has \Q!@Obj!,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,\\*
+$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,
 then the interface table $[\![\II]\!]$ IT is OK.
@@
 \begin{proof}
-Similar to what already argued for \textbf{THEOREM }$\weakAnn$
-we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finish by \textbf{THEOREM }$\weakAnn$.
+Similar to what already argued for \textbf{THEOREM }$\weakAnn$ tuning
+we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finish by \textbf{THEOREM }$\weakAnn$ tuning.
 \end{proof}

### 2015-12-08 paper/sections/Appendix.tex
Theorems and lemmas formatting.

@@
 
-\textbf{LEMMA 1. }
+\begin{lem}[a]\label{lemma1a}
 For any expression $e$ under an interface table $\II$ IT where $\Gamma\vdash e\in\C^\II$, $\II$ has $\weakAnn$ annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\C^\II$.
+\end{lem}
 \begin{proof}
@@
 \end{proof}
-\textbf{LEMMA 1b. }
+\begin{lem2}[b]\label{lemma1b}
 For any expression $e$ under an interface table $\II$ IT where there is no heir of $\C^\II$,  $\Gamma\vdash e\in\C^\II$, $\II$ has \mixin annotation and $[\![\II]\!] = \II'$, then under the interface table $\II'$ IT, $\Gamma\vdash e\in\_<:\C^\II$.
+\end{lem2}
 \begin{proof}
-The proof follows the same scheme of the Lemma1, but for the case of method call the return type may be refined with a subtype.
+The proof follows the same scheme of Lemma~\ref{lemma1a} (a), but for the case of method call the return type may be refined with a subtype.
 This is still ok since we require $\_<:\C^\II$.
@@
 
-\textbf{LEMMA 2. }
+\begin{lem}[a]\label{lemma2a}
 If $\II$ has $\weakAnn$ annotation and $\C^{\II}$ OK in $\II$ IT, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
+\end{lem}
 \begin{proof}
@@
 
-Since $\II$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II]\!]$, except for the new method \QM{of}.
+Since $\II$ OK, and by Lemma~\ref{lemma1a} (a), all the existing default and static methods are well typed in $[\![\II]\!]$, except for the new method \QM{of}.
 It suffices to prove that it still holds for $\ofMethod(\C)$.
@@
 
-\textbf{LEMMA 2b. }
+\begin{lem2}[b]\label{lemma2b}
 If $\II$ has \mixin annotation $\C^{\II}$ OK in $\II$ IT
 and there is no heir of $\C^\II$, then $[\![\II]\!]$ OK in $[\![\II]\!]$ IT.
+\end{lem2}
 \begin{proof}
-\noindent\textbf{Part I.} Similarly to what already argued for \textbf{Lemma 2},
-since $\II$ OK, and by \textbf{LEMMA 1b}, all the existing default and static methods are well typed in $[\![\II]\!]$ IT.
+\noindent\textbf{Part I.} Similarly to what already argued for Lemma~\ref{lemma2a} (a),
+since $\II$ OK, and by Lemma~\ref{lemma1b} (b), all the existing default and static methods are well typed in $[\![\II]\!]$ IT.
 The translation function delegates its work to $\weakAnn$ in such way that we can refer to
-\textbf{Lemma 2} to complete this part. Note that all the methods added (directly) by \mixin are abstract, and thus there is no body to typecheck.
+Lemma~\ref{lemma2a} (a) to complete this part. Note that all the methods added (directly) by \mixin are abstract, and thus there is no body to typecheck.
 
 \noindent\textbf{Part II.}
-Similar to what we already argued for \textbf{Lemma 2}, but we need to notice that the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
+Similar to what we already argued for Lemma~\ref{lemma2a} (a), but we need to notice that the newly added methods are valid refinements for already present methods in $\dom(\C^\II)$ before the translation.
 Thus by last clause of the definition of $\override(\_)$, $\mBody(\_)$ is defined on the same method names.
@@
 
-\noindent\textbf{THEOREM }$\weakAnn$ tuning\\*
+\begin{thm2}[@ObjOf tuning]\label{theorem1}
 If a given interface table $\II$ IT is OK
@@
 then the interface table $[\![\II]\!]$ IT is OK.
+\end{thm2}
 
 \begin{proof}
-\textbf{LEMMA 2} already proves that $[\![\II]\!]$ is OK. On the other hand, for any $\II'\in $ IT$\setminus\II$, by \textbf{LEMMA 1}, we know that all its methods
+Lemma~\ref{lemma2a} (a) already proves that $[\![\II]\!]$ is OK. On the other hand, for any $\II'\in $ IT$\setminus\II$, by Lemma~\ref{lemma1a} (a), we know that all its methods
 are still well-typed, and the generated code in translation of $\weakAnn$ is only a static method \QM{of}, which has no way to affect the domain
@@
 
-\noindent\textbf{THEOREM }\mixin tuning\\*
+\begin{thm2}[@Obj tuning]\label{theorem2}
 If a given interface table $\II$ IT is OK
- where $\II$ has \Q!@Obj!,
+ where $\II$ has $\mixinAnn$,
 $\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,
 then the interface table $[\![\II]\!]$ IT is OK.
+\end{thm2}
 
 \begin{proof}
-Similar to what already argued for \textbf{THEOREM }$\weakAnn$ tuning
-we can apply \textbf{LEMMA 2b} and \textbf{LEMMA 1b}. Then we finish by \textbf{THEOREM }$\weakAnn$ tuning.
+Similar to what already argued for Theorem~\ref{theorem1},
+we can apply Lemma~\ref{lemma2b} (b) and Lemma~\ref{lemma1b} (b). Then we finish by Theorem~\ref{theorem1}.
 \end{proof}

### 2015-12-08 paper/sections/Translation.tex
capitalize title
removed some line breaks
Theorems and lemmas formatting.

@@
 
-\subsection{Other features}
+\subsection{Other Features}
 
@@
 To understand to what extent our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
-
 \begin{itemize}
@@
 
-\noindent\textbf{THEOREM }$\weakAnn$ tuning\\*
+\begin{thm}[@ObjOf tuning]
 If a given interface table $\II$ IT is OK
@@
 then the interface table $[\![\II]\!]$ IT is OK.
+\end{thm}
 
-\noindent\textbf{THEOREM }\mixin tuning\\*
+\begin{thm}[@Obj tuning]
 If a given interface table $\II$ IT is OK
- where $\II$ has \Q!@Obj!,
+ where $\II$ has $\mixinAnn$,
 $\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,
 then the interface table $[\![\II]\!]$ IT is OK.
+\end{thm}
 

### 2015-12-08 paper/sections/Formalization.tex
spellcheck one more time
bundle. small fixes. move on to theorem macros.
small fix in S4

@@
 
-\paragraph{Abstract method} composition is accepted when there exist a most specific one.
+\paragraph{Abstract method} composition is accepted when there exists a most specific one.
   For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
@@
 %Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
-%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java change its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
+%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java changes its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
 
@@
 interesting type of expressions is super calls
-($\C\QM{.super.}\m\QM(\es)$), whose semantic is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
+($\C\QM{.super.}\m\QM(\es)$), whose semantics is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
 dispatching to the version of the method as visible in the interface
@@
 usual, we allow a functional notation for $\Gamma$ to do variable
-lookup.  Moreover, to help us defining auxiliary functions, a functional
+lookup.  Moreover, to help us define auxiliary functions, a functional
 notation is also allowed for a set of methods $\methods$, using the
@@
 $\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
-For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their name.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their names.
 

### 2015-12-08 paper/paper-without-appendix.tex
clean up and updated pdfs

--- 2015-12-06 paper/paper.tex
@@
 \usepackage{pstricks,pst-node,pst-tree}
-\usepackage{theorem}
 %% \usepackage{mathpartir}
@@
 \usepackage{mathpartir}
+
+\theoremstyle{plain}
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}{Lemma}
+\newtheorem{thm2}{Theorem}
+\newtheorem{lem2}{Lemma}
 
@@
 %\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
-\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
-\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
-\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
-%\newcommand{\hl}[1]{\textcolor{red}{#1}}
-\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
-\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
+
+%%%%========== authornotes marco begin ================
+%%use the following 5 lines for showing notes
+% \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
+% \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
+% \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
+% \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
+% \newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
+%%use the following 4 lines for hiding notes
+\newcommand\bruno[1]{}
+\newcommand\yanlin[1]{}
+\newcommand\marco[1]{}
+\newcommand\haoyuan[1]{}
+%%%%========== authornotes marco end ====================
 
@@
 
-\author[1]{John Q. Open}
-\author[2]{Joan R. Access}
-\affil[1]{Dummy University Computing Laboratory\\
-  Address, Country\\
-  \texttt{open@dummyuni.org}}
-\affil[2]{Department of Informatics, Dummy College\\
-  Address, Country\\
-  \texttt{access@dummycollege.org}}
-\authorrunning{J.\,Q. Open and J.\,R. Access}
+\author[1]{}
+% \author[2]{Joan R. Access}
+% \affil[1]{Dummy University Computing Laboratory\\
+%   Address, Country\\
+%   \texttt{open@dummyuni.org}}
+% \affil[2]{Department of Informatics, Dummy College\\
+%   Address, Country\\
+%   \texttt{access@dummycollege.org}}
+% \authorrunning{J.\,Q. Open and J.\,R. Access}
 % mandatory. First: Use abbreviated first/middle names. Second (only in severe
@@
 
-\begin{abstract}
-
-Java 8 introduced \emph{default methods}, allowing interfaces to
-have method implementations. When combined with (multiple) interface
-inheritance, default methods provide a basic form of multiple
-inheritance. However, using this combination to simulate multiple
-inheritance quickly becomes cumbersome, and appears to be quite
-restricted.
-
-This paper shows that, with a simple language feature, default methods
-and interface inheritance are in fact very expressive. Our proposed
-language feature, called \emph{object interfaces}, enables powerful
-object-oriented idioms, using multiple inheritance, to be expressed
-conveniently in Java. Object interfaces refine conventional Java
-interfaces in three different ways. Firstly, object interfaces have
-their own object instantiation mechanism, providing an alternative to
-class constructors. Secondly, object interfaces support \emph{abstract
-  state operations}, providing a way to use multiple inheritance with
-state in Java. Finally, object interfaces allow type refinements that
-are often tricky to model in conventional class-based
-approaches. Interestingly, object interfaces do not require changes to
-the runtime, and they also do not introduce any new syntax: all three features are achieved by reinterpreting
-existing Java syntax, and are translated into regular Java code
-without loss of type-safety. Since no new syntax is introduced, it
-would be incorrect to call object interfaces a language extension or
-syntactic sugar. So we use the term \emph{language tuning} to
-characterize this kind of language feature. An implementation of
-object interfaces using Java annotations and a formalization of the
-static and dynamic semantics are presented. Moreover the usefulness of
-object interfaces is illustrated through various examples.
-
-\end{abstract}
-
-%if False
-
-%\category{D.3.2}{Programming Languages}
-%                {Language Classifications}
-%                [Functional Languages]
-%\category{F.3.3}{Logics and Meanings of Programs}
-%                {Studies of Program Constructs}
-%                []
-
-%\terms
-%Languages
-%
-%\keywords
-%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
-%interference
-
-%endif
-
-%===============================================================================
+\input{sections/Abstract.tex}
 
@@
 
-\appendix
-\input{sections/Appendix.tex}
-
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "."
-%%% End:

### 2015-12-08 paper/sections/Abstract.tex
clean up and updated pdfs

--- 2015-12-06 paper/paper.tex
@@
-%\documentclass[preprint]{sigplanconf}
-%\documentclass[preprint]{llncs}
-\documentclass[a4paper,UKenglish]{lipics}
+\begin{abstract}
+  Java 8 introduced \emph{default methods}, allowing interfaces to have method
+  implementations. When combined with (multiple) interface inheritance, default
+  methods provide a basic form of multiple inheritance. However, using this
+  combination to simulate multiple inheritance quickly becomes cumbersome, and
+  appears to be quite restricted.
 
-\input{def}
-
-\usepackage{listings}
-\usepackage{multicol}
-\usepackage{microtype}%if unwanted, comment out or use option "draft"
-\usepackage[table,xcdraw]{xcolor}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{stmaryrd}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{fancyvrb}
-\usepackage{url}
-\usepackage{pstricks,pst-node,pst-tree}
-\usepackage{theorem}
-%% \usepackage{mathpartir}
-\usepackage{bbm}
-\usepackage{pgf}
-\usepackage{multirow}
-
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{wrapfig}
-
-\usepackage[T1]{fontenc}
-\usepackage[scaled=0.85]{beramono}
-\usepackage{mathpartir}
-
-% "define" code highlights for Java and Scala
-\lstdefinelanguage{JavaScala}{
-  morekeywords={public,int,interface,implements,default,
-    abstract,case,catch,class,def,static,%
-    do,else,extends,false,final,finally,%
-    for,if,implicit,import,match,mixin,%
-    new,null,object,override,package,%
-    private,protected,requires,return,sealed,%
-    super,this,throw,trait,true,try,%
-    type,var,while,yield},
-  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
-  sensitive=true,
-  morecomment=[l]{//},
-  morecomment=[n]{/*}{*/},
-  morestring=[b]",
-  morestring=[b]',
-  morestring=[b]"""
-}
-
-\lstset{ %
-language=JavaScala,                % choose the language of the code
-columns=flexible,
-lineskip=-1pt,
-basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
-numbers=none,                   % where to put the line-numbers
-numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
-stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
-numbersep=5pt,                  % how far the line-numbers are from the code
-backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
-showspaces=false,               % show spaces adding particular underscores
-showstringspaces=false,         % underline spaces within strings
-showtabs=false,                 % show tabs within strings adding particular underscores
-morekeywords={var},
-%  frame=single,                   % adds a frame around the code
-tabsize=2,                  % sets default tabsize to 2 spaces
-captionpos=none,                   % sets the caption-position to bottom
-breaklines=true,                % sets automatic line breaking
-breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
-title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
-escapeinside={(*}{*)},          % if you want to add a comment within your code
-keywordstyle=\ttfamily\bfseries,
-aboveskip=3pt,
-belowskip=-1pt
-% commentstyle=\color{Gray},
-% stringstyle=\color{Green}
-}
-
-%%\usepackage{natbib}
-%%\bibpunct();A{},
-%%\let\cite=\citep
-
-%include lhs2TeX.fmt
-%include lhs2TeX.sty
-%include forall.fmt
-
-%\pagestyle{plain}
-
-%{\theorembodyfont{\sffamily} \newtheorem{theorem}{Theorem}}
-%{\theorembodyfont{\sffamily} \newtheorem{lemma}{Lemma}}
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}{Lemma}
-%\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
-\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
-\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
-\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
-%\newcommand{\hl}[1]{\textcolor{red}{#1}}
-\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
-\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
-
-\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
-\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
-\newcommand\tsem[1]{\llbracket #1 \rrbracket}
-\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
-\newcommand\nat[0]{\mathbb{N}}
-\newcommand\unit[0]{\mathbbm{1}}
-
-\newcommand\mixin{\textbf{@Obj}\xspace}
-\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
-\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
-\usepackage{xspace}
-
-% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Classless Java: Tuning Java Interfaces}
-\titlerunning{Classless Java}
-
-\author[1]{John Q. Open}
-\author[2]{Joan R. Access}
-\affil[1]{Dummy University Computing Laboratory\\
-  Address, Country\\
-  \texttt{open@dummyuni.org}}
-\affil[2]{Department of Informatics, Dummy College\\
-  Address, Country\\
-  \texttt{access@dummycollege.org}}
-\authorrunning{J.\,Q. Open and J.\,R. Access}
-% mandatory. First: Use abbreviated first/middle names. Second (only in severe
-% cases): Use first author plus 'et. al.'
-
-\begin{comment}
-\Copyright{John Q. Open and Joan R. Access}
-% mandatory, please use full first names. LIPIcs license is "CC-BY";
-% http://creativecommons.org/licenses/by/3.0/
-
-\subjclass{Dummy classification -- please refer to
-  \url{http://www.acm.org/about/class/ccs98-html}}
-% mandatory: Please choose ACM 1998 classifications from
-% http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of
-% Computation".
-\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
-% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\serieslogo{}%please provide filename (without suffix)
-\volumeinfo%(easychair interface)
-  {Billy Editor and Bill Editors}% editors
-  {2}% number of editors: 1, 2, ....
-  {Conference title on which this volume is based on}% event
-  {1}% volume
-  {1}% issue
-  {1}% starting page number
-\EventShortName{}
-\DOI{10.4230/LIPIcs.xxx.yyy.p}% to be completed by the volume editor
-% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-\maketitle
-
-\begin{abstract}
-
-Java 8 introduced \emph{default methods}, allowing interfaces to
-have method implementations. When combined with (multiple) interface
-inheritance, default methods provide a basic form of multiple
-inheritance. However, using this combination to simulate multiple
-inheritance quickly becomes cumbersome, and appears to be quite
-restricted.
-
-This paper shows that, with a simple language feature, default methods
-and interface inheritance are in fact very expressive. Our proposed
-language feature, called \emph{object interfaces}, enables powerful
-object-oriented idioms, using multiple inheritance, to be expressed
-conveniently in Java. Object interfaces refine conventional Java
-interfaces in three different ways. Firstly, object interfaces have
-their own object instantiation mechanism, providing an alternative to
-class constructors. Secondly, object interfaces support \emph{abstract
-  state operations}, providing a way to use multiple inheritance with
-state in Java. Finally, object interfaces allow type refinements that
-are often tricky to model in conventional class-based
-approaches. Interestingly, object interfaces do not require changes to
-the runtime, and they also do not introduce any new syntax: all three features are achieved by reinterpreting
-existing Java syntax, and are translated into regular Java code
-without loss of type-safety. Since no new syntax is introduced, it
-would be incorrect to call object interfaces a language extension or
-syntactic sugar. So we use the term \emph{language tuning} to
-characterize this kind of language feature. An implementation of
-object interfaces using Java annotations and a formalization of the
-static and dynamic semantics are presented. Moreover the usefulness of
-object interfaces is illustrated through various examples.
-
+  This paper shows that, with a simple language feature, default methods and
+  interface inheritance are in fact very expressive. Our proposed language
+  feature, called \emph{object interfaces}, enables powerful object-oriented
+  idioms, using multiple inheritance, to be expressed conveniently in
+  Java. Object interfaces refine conventional Java interfaces in three different
+  ways. Firstly, object interfaces have their own object instantiation
+  mechanism, providing an alternative to class constructors. Secondly, object
+  interfaces support \emph{abstract state operations}, providing a way to use
+  multiple inheritance with state in Java. Finally, object interfaces allow type
+  refinements that are often tricky to model in conventional class-based
+  approaches. Interestingly, object interfaces do not require changes to the
+  runtime, and they also do not introduce any new syntax: all three features are
+  achieved by reinterpreting existing Java syntax, and are translated into
+  regular Java code without loss of type-safety. Since no new syntax is
+  introduced, it would be incorrect to call object interfaces a language
+  extension or syntactic sugar. So we use the term \emph{language tuning} to
+  characterize this kind of language feature. An implementation of object
+  interfaces using Java annotations and a formalization of the static and
+  dynamic semantics are presented. Moreover the usefulness of object interfaces
+  is illustrated through various examples.
 \end{abstract}
-
-%if False
-
-%\category{D.3.2}{Programming Languages}
-%                {Language Classifications}
-%                [Functional Languages]
-%\category{F.3.3}{Logics and Meanings of Programs}
-%                {Studies of Program Constructs}
-%                []
-
-%\terms
-%Languages
-%
-%\keywords
-%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
-%interference
-
-%endif
-
-%===============================================================================
-
-\input{sections/Introduction.tex}
-
-\input{sections/Overview.tex}
-
-\input{sections/Formalization.tex}
-
-\input{sections/Translation.tex}
-
-\input{sections/Implementation.tex}
-
-\input{sections/CaseStudy.tex}
-
-\input{sections/RelatedWork.tex}
-
-\input{sections/FutureWork.tex}
-
-%===============================================================================
-\input{sections/Conclusion.tex}
-
-%===============================================================================
-
-%\bibliographystyle{splncs}
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\appendix
-\input{sections/Appendix.tex}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "."
-%%% End:

### 2015-12-08 paper/run.sh

@@ -5,7 +5,8 @@ then
     case "$1" in
     bib )       
         pdflatex paper.tex
-        bibtex paper     
+        bibtex paper
+        bibtex paper-without-appendix     
     ;;
     * )
         echo "Usage (In terminal with zsh):"
@@ -19,6 +20,9 @@ fi
 ruby computePositions.rb
 pdflatex paper.tex
 open paper.pdf
+
+pdflatex paper-without-appendix.tex
+open paper-without-appendix.pdf
 
 rm footempfile.txt
 # rm *.bbl

### 2015-12-08 paper/paper.tex
Revised Section 5
revised bib and lots of small changes
toggle author notes
ready to submit
Theorems and lemmas formatting.
fix a small error in Formalization

@@
 %\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
+
+%%%%========== authornotes marco begin ================
+%%use the following 5 lines for showing notes
 \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
@@
 \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
-%\newcommand{\hl}[1]{\textcolor{red}{#1}}
 \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
 \newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
+%%use the following 4 lines for hiding notes
+% \newcommand\bruno[1]{}
+% \newcommand\yanlin[1]{}
+% \newcommand\marco[1]{}
+% \newcommand\haoyuan[1]{}
+%%%%========== authornotes marco end ====================
 

### 2015-12-08 paper/sections/Formalization.tex
fix a small error in Formalization

@@
 $\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
-$\textsf{mtypeS}(\m,\C)$ only on static ones. We use $\dom(\C)$ to
+$\textsf{mtypeS}(\m,\C)$ only for static ones. We use $\dom(\C)$ to
 denote the set of methods that are defined for type $\C$, that is:
@@
 
-\noindent Here we require that for all newly declared methods, there is a
-method with the same name defined in the interface $\C$, and that such method is a subtype of the newly introduced one.
-We define subtyping between methods in a general form that will be useful also later.
+\noindent Here we require that for all newly declared methods, there is a method
+with the same name defined in the interface $\C$, and that such method is a
+supertype of the newly introduced one. We define subtyping between methods in a
+general form that will be useful also later.
 

### 2015-12-08 paper/sections/Formalization.tex
revising S4.2

@@
 First  we check that all method headers are valid
-with respect to the corresponding method already present in $\C$;
+with respect to the corresponding method already present in $\C$:
 
@@
 supertype of the newly introduced one. We define subtyping between methods in a
-general form that will be useful also later.
+general form that will also be useful later.
 
@@
 
-\noindent Two method headers are subtypes if all the parameter types are the same and the return types are subtypes.
-That is, we allow return type specialization as introduced in Java 5.
-Default methods are subtypes if their method headers are subtypes.
-Secondly, we check that all abstract methods in the interface (that is methods that need
-to be explicitly overridden) have been
-implemented. That is, we define a method with the same name.
+\noindent We allow return type specialization as introduced in Java 5. A method
+header with return type $I$ is a subtype of another method header with return
+type $I'$ if all parameter types are the same, and $I <: I'$. A default method
+$\method_1$ is a subtype of another default method $\method_2$ iff
+$\mh^{\method_1}$ is a subtype of $\mh^{\method_2}$. Secondly, we check that all
+abstract methods (which need to be explicitly overridden) in the interface have
+been implemented: %That is, we define a method with the same name.
 
@@
 The rule \rn{t-intf} checks that an interface $\C$ is correctly
-typed.  First we check that the body of all the default and static
-methods are well typed.  Then we check that $\dom(\C)$ is the same of
+typed.  First we check that the body of all default and static
+methods are well typed.  Then we check that $\dom(\C)$ is the same as
 $\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$.  This is not a
-trivial check, since $\dom(\C)$ is defined using $\mBody$, that is
+trivial check, since $\dom(\C)$ is defined using $\mBody$, which would be 
 undefined in many cases: notably if a method $\method\in\methods$ is
 not compatible with some method in $\dom(\C_1)\ldots\dom(\C_n)$ or if
-any method in both $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$) is
-conflicting.
+there are methods in any $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$) conflict.
 

### 2015-12-08 paper/sections/Formalization.tex
S4.3 with comments

@@
 
-Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
+Defining \mBody{} is not trivial, and requires quite a lot of attention to the
+specific model of Java interfaces, and to how it differs w.r.t. Java Class model.
 $\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
@@
 
+\yanlin{I think the explaination for needed might be hard to understand (for
+  reviewer). need revising.}
 \paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
-the non-static ones that are not reachable by another, less specific, superinterface.
+the non-static ones that are not reachable by another, less specific superinterface.
 Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
@@
 override an implementation in the superinterface, even in case of a
-conflict.  Note how we use the special value $\none$, and how (forth
+conflict. Note how the special value $\none$ is used, and how (the 4th
 case) overriding can solve a conflict.
@@
 \InTextWith{
-\forall \method' \in \methods \ \
+\forall \method' \in \methods : 
 %\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
@@
 \end{array}$
-
+\yanlin{added a colon}
 \noindent The definition $\Aux{mostSpecific}$ returns the most
-specific method, that is a method whose type is the subtype of all the
+specific method whose type is the subtype of all the
 others.  Since method subtyping is a partial ordering,
 $\Aux{mostSpecific}$ may not be defined, this in turn forces us to rely on the last clause of $\override$; otherwise
-the whole \mBody{} would be not defined for that specific $\m$.
+the whole \mBody{} would not be defined for that specific $\m$.
 Rule \rn{t-intf} relies on this behaviour.

### 2015-12-08 paper/sections/Translation.tex
revising S5.1

@@
 
-This section shows what the \mixin annotation generates, and presents a
+This section shows what the \mixin annotation generates and presents a
 formal definition for most of the generated methods. Since
@@
 
-For the purposes of the formalization, the translation is divided in
-two parts.  The two parts are important so that we can discuss the
-formal properties later. To this aim we introduce the annotation
+For the purposes of the formalization, the translation is divided into
+two parts for more convenient discussion on formal properties later. To this aim we introduce the annotation
 $\weakAnn$. Its role is only in the translation process, hence is
@@
 \noindent$\begin{array}{ll}
-\InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; = \mBody(\m, \C_0),\mbox{ one case is satisfied:}}\\
+\InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; =
+            \mBody(\m, \C_0),\mbox{ one of the following cases is satisfied:}}\\
 \tab\tab
@@
 
-\noindent That is, we can categorize all the \emph{not implemented} methods in a pattern that we know how to implement: it is either a field getter, a with method or a setter method.
+\noindent That is, we can categorize all abstract methods in a pattern that we
+know how to implement: it is either a field getter, a with method or a setter.
+\yanlin{\emph{not implemented} -> abstract}
 
@@
 
-In the following we will write $\QM{with#}\m$ to append $\m$ to
-\QM{with}, following the camelCase rule. That is, the first letter of $\m$
-must be lower-case and is turned in upper-case upon merging.  For
+In the following sections  we will write $\QM{with#}\m$ to append $\m$ to
+\QM{with}, following the camelCase rule. The first letter of $\m$
+must be lower-case and is changed to upper-case upon appending. For
 example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
-are \QM{with} and all the identifiers of form $\QM{with#}\m$.
+are \QM{with} and all identifiers of the form $\QM{with#}\m$.
 
 \paragraph{The $\otherMethod$ method:}
-The definition of $\otherMethod(\C_0,\methods)$ is as follows:
+$\otherMethod(\C_0,\methods)$ is defined as follows:
 
@@
 the return types need to be refined.
-To discover whether the methods need to be generated,
+To determine whether the methods need to be generated,
 we check if such \Q@with-@ or setter methods %\Q@clone@
-are required by $\C_0$, but are not already present in the methods directly declared in $\C_0$.
+are required by $\C_0$, but not declared directly in $\C_0$.
 
 \paragraph{The $\ofMethod$ method:}\label{subsec:ofmethod}
-The $\ofMethod$ function generates the method
-\QM{of}, which behaves like a factory. To avoid boring digressions about
-well-known ways to find unique names, for the sake of this formalization we
-assume that no-args methods start with an underscore, and we prefix method names with underscore to obtain valid  parameter names.
+The function $\ofMethod$ generates the method
+\QM{of}, as an object factory. To avoid boring digressions into 
+well-known ways to find unique names, %for the sake of this formalization 
+we assume that no-args methods start with an underscore, and we prefix method names with underscore to obtain valid  parameter names.
 

### 2015-12-08 paper/sections/Translation.tex
fix S5.1

@@
 \paragraph{The $\ofMethod$ method:}\label{subsec:ofmethod}
-The function $\ofMethod$ generates the method
-\QM{of}, as an object factory. To avoid boring digressions into 
-well-known ways to find unique names, %for the sake of this formalization 
-we assume that no-args methods start with an underscore, and we prefix method names with underscore to obtain valid  parameter names.
+The function $\ofMethod$ generates the method \QM{of}, as an object factory. To
+avoid boring digressions into well-known ways to find unique
+names, %for the sake of this formalization
+we assume that all methods with no parameters do not start with an underscore,
+and we prefix method names with underscores to obtain valid parameter names for
+\QM{of}.
 
@@
 \tab\QM{\};\}} \\
-\InTextWith{\fieldsFunc(\C_0)=\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();}
-\mbox{and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n}
+\InTextWith{\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();} = \fieldsFunc(\C_0) 
+\mbox{ and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n}
 \end{array}$
@@
    \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
-    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This corresponds to the expressions $\es_i$.
+    the current value for all the fields except for $\m_i$. Such new value is
+    provided as a parameter. This corresponds to the expressions $\es_i$.
 \item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
@@
 \QM{return}\ \C_0\QM{.of(}\es\QM{);\}}} \\
-\InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ is of form }\mh\QM;}\\
+\InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ having the form }\mh\QM;}\\
 \InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{\emptyset\mbox{ otherwise}}\\
@@
 \InTextWith{
-\mBody(\QM_\m,\C_0) \mbox{ is of form }\mh\QM;}\\
+\mBody(\QM_\m,\C_0) \mbox{ having the form }\mh\QM;}\\
 \InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{\emptyset\mbox{ otherwise}}\\

### 2015-12-08 paper/sections/Translation.tex
revised S5.2

@@
   \begin{itemize}
-    \item Check if method $\C\ \m$\Q@();@ exists. If not, generate error (that is, is not $\valid(\C_0)$).
-    \item Generate implemented setter method inside \Q@of@:\\*
+    \item Check if method $\C\ \m$\Q@();@ exists. If not, generate error (that
+      is, $\valid(\C_0)$ is false).
+    \item Generate the implemented setter method inside \Q@of@:\\*
       \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@\\*
-    Note how there is no need to refine the return type for non-fluent setters, thus we do not need to generate the method header in the interface body itself.
+      There is no need to refine the return type for non-fluent setters, thus we
+      do not need to generate the method header in the interface body itself.
     \end{itemize}
@@
 ${}_{}$\Q@  return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
-where with $\m_1\ldots\m_n$  fields of $\C_0$,
-$\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method; otherwise
-$\e_i=\m_i$.
-    \item If needed, as for \Q@with-@ and setters, generate the method header with refined return type in the interface.
+ with $\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method where
+ $\m_1\ldots\m_n$ are fields of $\C_0$; otherwise $\e_i=\m_i$.
+    \item If needed, as for \Q@with-@ and setters, generate the method headers with refined return types in the interface.
  \end{itemize}

### 2015-12-08 paper/sections/FutureWork.tex
subsection -> paragraph

@@
 \section{Future work}\label{sec:futurework}
-
+In this section we discuss potential future work (not a full list).
 \begin{comment}
@@
 
-\subsection{Qualifiers in Methods} %{Private state
+%\subsection{Qualifiers in Methods} %{Private state
+\paragraph{Qualifiers in Methods} %{Private state
 The biggest limitation of our approach is the inability to model qualifiers
@@
 
-\marco{old version still commented if we decided to swap back}
 % \yanlin{future work on state initialization removed for now. Because I think the
@@
 
-\subsection{Clone, toString, equals and hashCode}
-
+%\subsection{Clone, toString, equals and hashCode}
+\paragraph{Clone, toString, equals and hashCode}
 Methods originally defined in Java class \Q@Object@, as \Q@clone@ and

### 2015-12-08 paper/sections/Implementation.tex
spellcheck one more time
capitalize title
subsection -> paragraph

@@
 \begin{itemize}
-%this seamd like a commercial spot
+%this seems like a commercial spot
 %\item Lombok is byte-code based instead of source-code based, which makes client
@@
 provides other convenient features (including generating fluent setters, type
-refinement,etc), which the base Lombok project does not provide.
+refinement, etc), which the base Lombok project does not provide.
 Finally, while \mixin is formalized, none of Lombok's annotations have been
@@
 
-\paragraph{Lombok does Language tuning}
+\paragraph{Lombok does language tuning}
 We consider Lombok to be the most developed example of language

### 2015-12-08 paper/sections/RelatedWork.tex
removed some line breaks
subsection -> paragraph

@@
 
-\subsection{Multiple Inheritance in Object Oriented Languages}
+\paragraph{Multiple Inheritance in Object Oriented Languages}
 Many authors have argued in favour or against multiple inheritance.  Multiple
@@
 
-\paragraph{C++ Approach.}
+%\paragraph{C++ Approach.}
 C++ tries to provide a general solution to multiple inheritance by
@@
 
-\paragraph{Mixins.}
+%\paragraph{Mixins.}
 Mixins are a more restricted model than the C++ approach. Mixins allow to name
@@
 
-\paragraph{CZ.}
+%\paragraph{CZ.}
 Malayeri and Aldrich proposed a model CZ~\cite{malayeri2009cz} which
@@
 
-\paragraph{Traits and Java's default methods}
+%\paragraph{Traits and Java's default methods}
 Simplifying the mixins approach, traits~\cite{scharli03traits} draw a strong
@@
 our revised version for exclusion is cleaner, typesafe and more direct.
-
 \newcommand\shortItem{\vspace{-1ex}}
@@
 
-\paragraph{Traits vs Object Interfaces}
+\emph{Traits vs Object Interfaces.}
 We consider object interfaces to be an alternative to traits or mixins.  In the
@@
 
-\subsection{ThisType and MyType}
+\paragraph{ThisType and MyType}
 %\marco{ this section is unfair.
@@
 This demolishes most advantages of interfaces.}
-
 In certain situations, object interfaces allow automatic refinement for \emph{return
@@
 
-\subsection{Meta-Programming Competes with Language Extensions}
+\paragraph{Meta-Programming Competes with Language Extensions}
 The most obvious solution to adding features to a language is via syntactic
@@
 
-\subsection{Formalization of Java 8}
+\paragraph{Formalization of Java 8}
 We provide a simple formalization for a subset of Java including

### 2015-12-08 paper/sections/Translation.tex
S5.3 revised

@@
 \begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; type errors are not present in code the user have not written.
-In our case it means that either \mixin{} produces (in a controlled way) an understandable error, or the class can be successfully annotated and the generated code is well-typed.
+\item \textit{Self coherence}: the generated code itself is well-typed; that is,
+  type errors are not present in code the user has not written.
+In our case it means that either \mixin{} produces (in a controlled way) an
+understandable error, or the interface can be successfully annotated and the generated code is well-typed.
 We guarantee \textit{Self coherence}.
 
-\item \textit{Client coherence}: all the client code (as for example method calls) that is well-typed without the generation/instrumentation process is well-typed also after the generation.
-That is, the annotation does not remove any functionality, is just
-adding more behaviour.
+\item \textit{Client coherence}: all the client code (for example method calls)
+  that is well-typed before code generation is also well-typed after the generation.
+The annotation just adds more behaviour without removing any functionality.
 We guarantee \textit{Client coherence}.
 
-\item \textit{Heir coherence}: Interfaces (and in general classes) inheriting from instrumented code are well-typed if they were well-typed without the instrumentation.
-This would require to not add any (default or abstract) method to the annotated interfaces, including type refinement.
+\item \textit{Heir coherence}: interfaces (and in general classes) inheriting the instrumented code are well-typed if they were well-typed without the instrumentation.
+This would forbid adding any (default or abstract) method to the annotated interfaces, including type refinement.
 \mixin  does not guarantee \textit{Heir coherence}.
@@
 This would break \Q@C@. \\*
-Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@'' would compile before translation, but would be ill-typed after the translation.
+Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
+would be correct before translation, but would be ill-typed after the translation.
 \end{itemize}
@@
 \noindent This means that our automatic type refinement
-is a useful and convenient feature, but is not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantic  and provide the right type while refining methods.
+is a useful and convenient feature, but not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantics and provide the right type while refining methods.
 

### 2015-12-08 paper/sections/Translation.tex
S5.4 revised

@@
 \subsection{Results}
-To formally characterize the behaviour of our annotation, and the two level of guarantees that we offer, we provide some notations and two theorems:
+To formally characterize the behaviour of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
 \begin{itemize}
@@
 Informally, the theorems mean that for a client program that
-typechecks before a translation is applied, if the annotated type has
+typechecks before the translation is applied, if the annotated type has
 no subtypes, and no objects of that type are created, then type safety
@@
 \mixin
-tuning is more interesting: Since \mixin does not guarantee heir
-coherence, we exclude explicitly the presence of heirs. In this way
+tuning is more interesting: since \mixin does not guarantee heir
+coherence, we explicitly exclude the presence of heirs. In this way
 \mixin tuning guarantees only self and client coherence. The formal

### 2015-12-08 paper/sections/CaseStudy.tex
manual check

@@
 \noindent The main benefit that fluent methods give
-us is the convinience of method chaining:
+us is the convenience of method chaining:
 
@@
 \lstinputlisting[linerange=14-16]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
-This way, when a \texttt{ExtendedDatabase} query created, 
+In this way, when a query created using \texttt{ExtendedDatabase}, 
 all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling 

### 2015-12-08 paper/sections/Formalization.tex
Added a newline for formatting
remove @ObjOf in grammar

@@
 }{
-  \mixinAnn|\weakAnn|\emptyset
+  \mixinAnn|\emptyset
   }{annotations}\\
@@
 \paragraph{override} models how a method in an interface can
-override an implementation in the superinterface, even in case of a
-conflict. Note how the special value $\none$ is used, and how (the 4th
+override implementations in its superinterfaces, even in the case of
+conflicts. Note how the special value $\none$ is used, and how (the 4th
 case) overriding can solve a conflict.
@@
 \end{array}$
+
 \yanlin{added a colon}
@@
 
+\paragraph{Example}
+To illustrate the mechanism of $\mBody$ in a clearer and more intuitive way, we present an example here.
+\begin{lstlisting}
+interface A { Object m(); }
+interface B extends A { default Object m() {return new Object(){};} }
+interface C extends A {}
+interface D extends B, C { String m(); }
+\end{lstlisting}
+We start to compute $\mBody(\QM{m},\QM{D})$ inside out, and step by step:
+\begin{itemize}
+\item First the full set of supertypes of $\QM{D}$ is obtained recursively and results in $\{\QM{A,B,C}\}$.
+\item To compute $\tops(\QM{m},\{\QM{A,B,C}\})$, $\QM{C}$ can be omitted, since $\QM{m}$ is not declared directly in $\QM{C}$, hence $\Aux{originalMethod}(\QM{m},\QM{C})$ is undefined.
+And $\QM{B}$ is a subtype of $\QM{A}$ (thus $\QM{B.m}$ hides $\QM{A.m}$), so $\tops(\QM{m},\{\QM{A,B,C}\})$ equals $\QM{B.m}$.
+\item Now the final step is to compute $\override(\QM{D.m},\QM{B.m})$, and this concludes by the fifth case of $\override$ that $\QM{D.m}$ hides $\QM{B.m}$ successfully ($\QM{String}$ is a subtype of $\QM{Object}$) and becomes the most specific method.
+So finally we get $\mBody(\QM{m},\QM{D}) = \QM{D.m}$.
+\end{itemize}
 %The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.

### 2015-12-08 paper/sections/Introduction.tex
text driving Marco crazy

@@
   getter methods is a way out of this tension, but this workaround
-  requires tedious boilerplate classes that latter implement those
+  requires tedious boilerplate classes that later implement those
   methods.
@@
 Object interfaces do not require changes to the Java runtime or compiler, 
-and they also do not introduce any new syntax (apart from a
-small annotation). All three features of object interfaces are
+and they also do not introduce any new syntax. All three features of object interfaces are
 achieved by reinterpreting existing Java syntax, and are translated

### 2015-12-08 paper/sections/Overview.tex
removed some line breaks

--- 2015-12-08 paper/sections/Overview.tex
@@
 of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
-the basic system is given in the left side of Figure~\ref{fig:pegasus}.\footnote{
-Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+the basic system is given in the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
 
@@
 
-\noindent If an implementation is provided for methods,
-only code for methods needing implementation is generated. Thus,
-programmers can easily tailor the behaviour for their special needs.
+\noindent Only code for methods needing implementation is generated. Thus,
+programmers can easily customize the behaviour for their special needs.
 %\end{comment}
@@
 unsatisfactory options in front of us:
-
 \begin{itemize}
-
 \item Define a \Q@Point3D@ field in \Q@Animal@: this is bad since all animals
@@
 properties updater: it takes a certain type of object and returns a copy of the
-current object with all fields that match the fields of updated
-parameter type. The idea is that the result should be like \Q@this@, but
-modified to be as similar as possible to the parameter.
+current object where all the fields that match fields in the parameter
+object are updated to the corresponding value in the parameter. The idea is that
+the result should be like \Q@this@, but modified to be as similar as possible to the parameter.
 

### 2015-12-08 paper/sections/Translation.tex

@@
 This section shows what the \mixin annotation generates and presents a
-formal definition for most of the generated methods. Since
-Classless Java does not consider
+formal definition for most of the generated methods. Since the formalized part
+of Classless Java does not consider
 % setters, so we do not formally define the generation of the setters.
@@
 \subsection{Translation}
-
 For the purposes of the formalization, the translation is divided into
@@
 
-In the following sections  we will write $\QM{with#}\m$ to append $\m$ to
+We write $\QM{with#}\m$ to append $\m$ to
 \QM{with}, following the camelCase rule. The first letter of $\m$
@@
 
-\paragraph{The $\otherMethod$ method:}
+\paragraph{The $\otherMethod$ function:}
 $\otherMethod(\C_0,\methods)$ is defined as follows:
@@
 
-\paragraph{The $\ofMethod$ method:}\label{subsec:ofmethod}
+\paragraph{The $\ofMethod$ function:}\label{subsec:ofmethod}
 The function $\ofMethod$ generates the method \QM{of}, as an object factory. To
@@
 %}\\\InTextWith{
-\method=\mBody(\_,\C_0)
+\method=\mBody(m^\method,\C_0)
 }
@@
    \item $\m_i$\QM{()} is the getter, and just returns the current field value.
-   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
+   \item if a method \Q@with#@$\m_i()$ is required, then it is implemented by calling the \Q@of@ method using
     the current value for all the fields except for $\m_i$. Such new value is

### 2015-12-08 UseMixinLombok/src/pegasusJava/Main.java
small fixes

@@ -6,11 +6,11 @@ import static java.lang.System.out;
 interface Animal {} // no points yet!
 interface Horse extends Animal{
     default void run() {
-        out.println("I can run!");}
+        out.println("running!");}
 }
 interface Bird extends Animal{
     default void fly() {
-        out.println("I can fly!");}
+        out.println("flying!");}
 }
 interface Pegasus extends Horse, Bird {}
 //END_PEGASUS_JAVA

### 2015-12-08 paper/sections/Formalization.tex
small fixes
revise the newly added example

@@
 \subsection{Methods in Java 8 Interfaces}
-In Java 8 interfaces there are three types of methods: abstract, default, and
-static methods.
+In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
 
@@
 \paragraph{Example}
-To illustrate the mechanism of $\mBody$ in a clearer and more intuitive way, we present an example here.
+To illustrate the mechanism of $\mBody$ in a more intuitive way, we present an example here.
 \begin{lstlisting}
@@
 \item First the full set of supertypes of $\QM{D}$ is obtained recursively and results in $\{\QM{A,B,C}\}$.
-\item To compute $\tops(\QM{m},\{\QM{A,B,C}\})$, $\QM{C}$ can be omitted, since $\QM{m}$ is not declared directly in $\QM{C}$, hence $\Aux{originalMethod}(\QM{m},\QM{C})$ is undefined.
+\item To compute $\tops(\QM{m},\{\QM{A,B,C}\})$, $\QM{C}$ will be omitted, since $\QM{m}$ is not declared directly in $\QM{C}$, hence $\Aux{originalMethod}(\QM{m},\QM{C})$ is undefined.
 And $\QM{B}$ is a subtype of $\QM{A}$ (thus $\QM{B.m}$ hides $\QM{A.m}$), so $\tops(\QM{m},\{\QM{A,B,C}\})$ equals $\QM{B.m}$.
-\item Now the final step is to compute $\override(\QM{D.m},\QM{B.m})$, and this concludes by the fifth case of $\override$ that $\QM{D.m}$ hides $\QM{B.m}$ successfully ($\QM{String}$ is a subtype of $\QM{Object}$) and becomes the most specific method.
+\item The final step is to compute $\override(\QM{D.m},\QM{B.m})$, and this matches the fifth case of $\override$ that $\QM{D.m}$ hides $\QM{B.m}$ successfully ($\QM{String}$ is a subtype of $\QM{Object}$) and becomes the most specific method.
 So finally we get $\mBody(\QM{m},\QM{D}) = \QM{D.m}$.

### 2015-12-08 paper/sections/Formalization.tex
4th -> 5th
Revised text example H

--- 2015-12-08 paper/sections/Formalization.tex
@@
 \paragraph{override} models how a method in an interface can
-override an implementation in the superinterface, even in case of a
-conflict. Note how the special value $\none$ is used, and how (the 4th
+override implementations in its superinterfaces, even in the case of
+conflicts. Note how the special value $\none$ is used, and how (the 5th
 case) overriding can solve a conflict.
@@
 \end{array}$
+
+To illustrate the mechanism of $\mBody$, we present an example. We compute $\mBody(\QM{m},\QM{D})$:
+\begin{lstlisting}
+interface A { Object m(); }
+interface B extends A { default Object m() {return this.m();} }
+interface C extends A {}
+interface D extends B, C { String m(); }
+\end{lstlisting}
+\vspace{-2ex}
+\begin{itemize}
+\item First $\{\QM{A,B,C}\}$, the full set of supertypes of $\QM{D}$ is obtained.
+\item Then we compute 
+$\tops(\QM{m},\{\QM{A,B,C}\})= $
+\Q@default Object m() {...}@,
+ that is $\QM{B.m}$.
+That is, we do not consider either
+$\QM{C.m}$ (since $\QM{m}$ is not declared directly in $\QM{C}$, hence $\Aux{originalMethod}(\QM{m},\QM{C})$ is undefined) or $\QM{B.m}$ 
+(that is a subtype of $\QM{A}$, thus $\QM{B.m}$ hides $\QM{A.m}$).
+\item The final step computes $\override(\QM{D.m},\QM{B.m})=\QM{D.m}$,
+by the last case of $\override$ we get that $\QM{D.m}$ hides $\QM{B.m}$ successfully ($\QM{String}$ is a subtype of $\QM{Object}$). Finally we get $\mBody(\QM{m},\QM{D}) = \QM{D.m}$.
+\end{itemize}
 

### 2015-12-09 paper/sections/Introduction.tex
small fix

@@
 state (and other features) to interfaces would complicate the language
-and require changes in the compiler, and this would go beyond the
+and require changes to the compiler, and this would go beyond the
 goals of Java 8 development team.
@@
 pre-existing tools can work transparently on the tuned language.  To
-exploit the full-benefits of language tuning, our prototype
-implementation of object interfaces uses Java annotations and AST
+exploit the full benefits of language tuning, our prototype
+implementation of object interfaces uses Java annotations to do AST
 rewriting, allowing existing Java tools (such as IDEs) to work

### 2015-12-09 paper/sections/Overview.tex
Spacing1
Done Spacing
small fix

@@
 \begin{figure}
+\saveSpaceFig
 \begin{tabular}{c|c}
@@
 
-This section illustrates how object interfaces, expressed using the \mixinAnn{} annotation, enables powerful
+This section illustrates how object interfaces, expressed using the \mixinAnn{} annotation, enable powerful
 idioms using multiple inheritance in Java.
@@
 of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
-the basic system is given in the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
 
@@
 functionality from both \texttt{Horse} and \texttt{Bird}.
-A first attempt to model the animal system in Java 8 is given on the right-side
+A first attempt to model the animal system in Java 8 is given on the right side
 of Figure~\ref{fig:pegasus}.
@@
 \begin{figure}
+\saveSpaceFig
 \begin{tabular}{|l|l|l|}
@@
 \begin{figure}
+\saveSpaceFig
 \lstinputlisting[linerange=79-91]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D

### 2015-12-09 paper/sections/Overview.tex
S2 small fixes

@@
 classes can be used to avoid these problems.  We could have simply
-instantiated Pegasus using:
+instantiated \texttt{Pegasus} using:
 
@@
 \texttt{of} is automatically added to the interfaces. With the
-\texttt{of} method a Pegasus object is instantiated as follows:
+\texttt{of} method a \texttt{Pegasus} object is instantiated as follows:
 
@@
 \noindent The \texttt{of} method provides an alternative to a
-constructor, which is missing from interfaces. The following code,
+constructor, which is missing from interfaces. The following code 
 shows the code corresponding to the \texttt{Pegasus} interface
@@
 
-\noindent Note that the generated code is transparent to a user, which
+\noindent Note that the generated code is not transparent to a user, who 
 only sees the original code with the \mixin annotation. Compared to the pure
@@
 the complete animal system where \Q@Animal@s will include a \Q@location@
-representing their position in space, we shall
+representing their positions in space, we shall
 see that modelling stateful components using plain Java 8 quickly becomes
 cumbersome. The \mixinAnn{} annotation comes to rescue here and
-allows avoiding significant amounts of boilerplate code.
+ avoids significant amounts of boilerplate code.
 
@@
 \vspace{-5pt}
-\paragraph{\Q@Point2D@ with Object Interfaces}
+\paragraph{\Q@Point2D@ with object interfaces}
 This obvious ``constructor'' code can be automatically generated by our \mixin
@@
 
-\paragraph{\texttt{wither} methods in Object Interfaces}
+\paragraph{\texttt{with-} methods in object interfaces}
 The pattern of creating a new object by reusing most information from an old
@@
 data-structures. As such, it is
-supported in our code generation as \Q@with@ methods. For example:
+supported in our code generation as \Q@with-@ methods. For example:
 \begin{lstlisting}
@@
 
-\noindent Using \texttt{with} methods, the point \texttt{p} can be created instead using:
+\noindent Using \texttt{with-} methods, the point \texttt{p} can also be created
+by:
 
@@
 \paragraph{Summary} Dealing with state (mutable or not) in object interfaces
-relies on a notion of abstract state, where only methods that interact with
+relies on a notion of abstract state, where only \emph{methods} that interact with
 state are available to users. Object interfaces provide support for four
@@
 \paragraph{Unsatisfactory class-based solutions to field type refinement}
-In Java if we define an animal class with a field we have a set of
+In Java if we want to define an animal class with a field we have a set of
 unsatisfactory options in front of us:
@@
 Instead, the \emph{restriction} of (object) interfaces to have no
-fields, enlighten us that another approach is possible; often in programming languages ``freedom is slavery''.
+fields enlightens us that another approach is possible; often in programming languages ``freedom is slavery''.
 
@@
 \begin{itemize}
-\item by \emph{covariant method overriding} the return type of
+\item by \emph{covariant method overriding}, the return type of
   \texttt{location()} is refined to \texttt{Point3D};
-\item by \emph{overloading} a new setter for location is defined with a more
+\item by \emph{overloading}, a new setter for location is defined with a more
   precise type;
@@
 \Q@with@ method, taking a \Q@Point2D@ as an argument.
-Other wither methods (such as \Q@withX@) functionally update a field at a time.  This can be
+Other wither methods (such as \Q@withX@) functionally update a field one at a time.  This can be
 inefficient, and sometimes hard to maintain.  Often we want to update multiple
@@
 to store. Note how all the information about what fields sit in
-\Q@Point3D@ and in \Q@Point2D@ is properly encapsulated in the
+\Q@Point3D@ and \Q@Point2D@ is properly encapsulated in the
 \Q@with@ method, and is transparent to the implementer of \Q@Bird@.
@@
 \paragraph{Generated boilerplate}
-Just to give a feeling on how much boring code \mixin is generating, we show the
-generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. The
-generated code is mechanical. Writing such code by hand is error-prone. For
+Just to give a feeling of how much mechanical code \mixin is generating, we show the
+generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. Writing such code by hand is error-prone. For
 example a distracted programmer may swap the arguments of calls to
@@
 Finally, defining \texttt{Pegasus} is as simple as we did in the simplified
-(and stateless) version on the right-side of Figure~\ref{fig:pegasus}.
+(and stateless) version on the right of Figure~\ref{fig:pegasus}.
  Note how even the non-trivial pattern for field type refinement is

### 2015-12-09 paper/sections/Formalization.tex
Done Spacing
comments removed
small fixes

@@
 \textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
-\textbf{B2}        & both abstract, accepted                        & conservative error       & both abstract, accepted       \\ \hline
+\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
 \textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
@@
 \item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
-\item \textbf{both abstract, accepted} happens when the methods from both interfaces are abstract, which is also considered correct in all
+\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
   trait models.
@@
 \begin{figure}[t]
+\saveSpaceFig
 \begin{grammar}
@@
 \begin{figure}[t]
+\saveSpaceFig
 $
@@
 
-\yanlin{I think the explaination for needed might be hard to understand (for
-  reviewer). need revising.}
 \paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
@@
 
-\yanlin{added a colon}
 \noindent The definition $\Aux{mostSpecific}$ returns the most

### 2015-12-09 paper/sections/Introduction.tex
small fixes

@@
 ``The key goal of adding default methods to Java was "interface
-evolution", not "poor man's traits"''~\cite{goetz13default}. As a result
+evolution", not "poor man's traits"''~\cite{goetz13default}. As a result, 
 they were happy to support the secondary use of interfaces with
@@
 
-\item {\em Interfaces have no constructors.} As a result classes are 
+\item {\em Interfaces have no constructors.} As a result, classes are 
 still required to create objects, leading to substantial boilerplate 
@@
 rewriting, allowing existing Java tools (such as IDEs) to work
-out-of-the-box with our implementation. As a result we could
+out-of-the-box with our implementation. As a result, we could
 experiment object interfaces with several interesting Java programs,
@@
 is given as a syntax-directed translation from CJ to itself. In the
-resulting CJ code all object interfaces are translated into regular CJ
+resulting CJ code, all object interfaces are translated into regular CJ
 (and Java) interfaces with default methods. The translation is proved
@@
   implementation of object interfaces, using Java
-  annotations and AST rewriting. Moreover the usefulness of object interfaces is
+  annotations and AST rewriting. Moreover, the usefulness of object interfaces is
   illustrated through various examples and case studies.

### 2015-12-09 paper/sections/Abstract.tex
small fixes
typos, small fixes

@@
   methods provide a basic form of multiple inheritance. However, using this
-  combination to simulate multiple inheritance quickly becomes cumbersome, and
+  combination to simulate more advanced forms of multiple inheritance quickly becomes cumbersome, and
   appears to be quite restricted.
@@
   interfaces using Java annotations and a formalization of the static and
-  dynamic semantics are presented. Moreover the usefulness of object interfaces
+  dynamic semantics are presented. Moreover, the usefulness of object interfaces
   is illustrated through various examples.

### 2015-12-09 paper/sections/CaseStudy.tex
Done Spacing
typos, small fixes

@@
 \begin{figure}
+\saveSpaceFig
 \begin{tabular}{l|l}
@@
 \subsection{A Maze Game}
-The last case study is a simplified variant of Maze game, which is often
+The last case study is a simplified variant of a Maze game, which is often
 used~\cite{gof,bono14} 

### 2015-12-09 paper/sections/Formalization.tex
typos, small fixes

@@
 each variable assigned to a field in the typing environment and verifying that such type is a
-subtype of the field type. Secondly, we check that all method bodies are
+subtype of the field type. Finally, we check that all method bodies are
 well-typed. To do this the environment used to check the method body
@@
 
-\noindent The definition of $\mBody$ reconstruct the full set of supertypes $\Cs$ and then delegates the work to two other auxiliary functions:
+\noindent The definition of $\mBody$ reconstructs the full set of supertypes $\Cs$ and then delegates the work to two other auxiliary functions:
  $\tops(m,\Cs)$ and $\override(\method,\methods)$.

### 2015-12-09 paper/sections/Overview.tex
small fixes
typos, small fixes

@@
 
-\noindent Note that the generated code is not transparent to a user, who 
+\noindent Note that the generated code is transparent to a user, who 
 only sees the original code with the \mixin annotation. Compared to the pure
@@
 and an implementation of \texttt{run()} is defined using a \Q@default@
-method. The implementation of \texttt{run()} further illustrates the convenience of \emph{with} methods:
+method. The implementation of \texttt{run()} further illustrates the convenience of \texttt{with} methods:
 
@@
 naming conventions and the types of the methods. Fluent setters are a variant of
-conventional setters, and are discussed in detail in Section~\ref{sec:dsls}.
+conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.
 

### 2015-12-09 paper/sections/Conclusion.tex
small things
spell grammar check

@@
 mechanism to instantiate objects. The \mixin annotation helps programmers
-to write less cumbersome code while coding in Classless Java. Indeed
+to write less cumbersome code while coding in Classless Java. Indeed, 
 we think the obtained gain is so high that Classless Java with the \mixin
@@
 
- More importantly, we identify the concept of \emph{language tuning}.
+ More generally, we identify the concept of \emph{language tuning}.
  We identify libraries that are already performing language tuning (Lombok and Cofoja), and 

### 2015-12-09 paper/sections/Translation.tex
comments removed
spell grammar check

@@
 know how to implement: it is either a field getter, a with method or a setter.
-\yanlin{\emph{not implemented} -> abstract}
 
@@
    \item $\m_i$ is the field name, and has type $\C_i$.
-   \item $\m_i$\QM{()} is the getter, and just returns the current field value.
+   \item $\m_i$\QM{()} is the getter and just returns the current field value.
    \item if a method \Q@with#@$\m_i()$ is required, then it is implemented by calling the \Q@of@ method using
@@
 typechecks before the translation is applied, if the annotated type has
-no subtypes, and no objects of that type are created, then type safety
+no subtypes and no objects of that type are created, then type safety
 is guaranteed after the successful translation.

### 2015-12-09 paper/sections/CaseStudy.tex
unify US/UK English

@@
 JavaScript, Java, C++, D, Ruby, Scala, etc. In most languages, to create fluent
-interfaces, programmers have to either handwrite everything or create a wrapper
+interfaces, programmers have to either hand-write everything or create a wrapper
 around the original non-fluent interfaces using \textbf{this}. In Java, there
@@
 (different kinds of doors inherit a common type, with different
-features and behaviour), multiple inheritance (a special kind of door
+features and behavior), multiple inheritance (a special kind of door
 may require features from two other door types) and it also shows how

### 2015-12-09 paper/sections/Conclusion.tex
unify US/UK English

@@
  we forecast many different kinds of language tuning will emerge on suitable platforms like Java or the C\# CLR.
- We identify various kinds of safety guarantees that can be offered by language tuning, but the door is open for more flavours of safety guarantees to emerge.
+ We identify various kinds of safety guarantees that can be offered by language tuning, but the door is open for more flavors of safety guarantees to emerge.

### 2015-12-09 paper/sections/Formalization.tex
unify US/UK English

@@
 Before formalizing Classless Java and object interfaces, it is helpful
-to informally discuss the behaviour of methods in Java 8
+to informally discuss the behavior of methods in Java 8
 interfaces. In particular it is useful to understand how Java 8
@@
 would be accepted, and the implementation in \Q@A2@ would be part of the
-behaviour of \Q@D2@.
+behavior of \Q@D2@.
 
@@
 ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
-(leading Java 8 designer) we have confirmed that the expected behaviour is
+(leading Java 8 designer) we have confirmed that the expected behavior is
 rejection, hence this is a bug in ECJ. This bug was also reported by
@@
 $\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
-modelled. In the syntax of field updates $\x$ is expected to be a
+modeled. In the syntax of field updates $\x$ is expected to be a
 field name. After updating the field $\x$ using the value of $\e$, the
@@
   Following standard
-practise, we consider a global Interface Table (\metaVar{IT}) mapping
+practice, we consider a global Interface Table (\metaVar{IT}) mapping
 from interface names $\C$ to interface declarations $\II$.
@@
 and names of the method parameters.  The
-notation $\m^\mh$ and $\C^\mh$ denotes respectivelly, extracting the
+notation $\m^\mh$ and $\C^\mh$ denotes respectively, extracting the
 method name and the return type from a method header. $\mBody(\m,\C)$,
@@
 the whole \mBody{} would not be defined for that specific $\m$.
-Rule \rn{t-intf} relies on this behaviour.
+Rule \rn{t-intf} relies on this behavior.
 

### 2015-12-09 paper/sections/FutureWork.tex
unify US/UK English

@@
 We are unsure about the value of this workaround because of its trickiness. In
-order to define a state with initialization, users have to define a method trusting that it will be overridden later with a behavioural change, but
+order to define a state with initialization, users have to define a method trusting that it will be overridden later with a behavioral change, but
 such change is visible only after the first time the setter is called. Moreover, this code would cache

### 2015-12-09 paper/sections/Introduction.tex
typos, small fixes
unify US/UK English

@@
 things, that: 1) a trait provides a set of methods that implement
-behaviour; and 2) a trait does not specify any state variables, so the
+behavior; and 2) a trait does not specify any state variables, so the
 methods provided by traits do not access state variables
@@
 
-\noindent Clearly, a way around those difficulties would be to
-change Java and just remove these limitations. Scala's own
-notion of traits, for example, allows state in traits. Of course adding
-state (and other features) to interfaces would complicate the language
-and require changes to the compiler, and this would go beyond the
-goals of Java 8 development team.
+\noindent Clearly, a way around those difficulties would be to change
+Java and just remove these limitations. Scala's own notion of
+traits~\cite{scala-overview}, for example, allows state in traits. Of
+course adding state (and other features) to interfaces would
+complicate the language and require changes to the compiler, and this
+would go beyond the goals of Java 8 development team.
 

### 2015-12-09 paper/sections/Overview.tex
unify US/UK English

@@
 of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
-the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favour of using subtyping for modelling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
 
@@
 
-Before modelling the complete animal system, we  start with a
+Before modeling the complete animal system, we  start with a
 simplified version without locations. This version serves the purpose of illustrating how
@@
 methods. Moreover, because interfaces support multiple interface
-inheritance, the interface for \texttt{Pegasus} can inherit behaviour
+inheritance, the interface for \texttt{Pegasus} can inherit behavior
 from both \texttt{Horse} and \texttt{Bird}. Although Java interfaces
@@
 provide each of those classes, this code is essentially boilerplate
-code, which does not add behaviour to the system. Secondly,
+code, which does not add behavior to the system. Secondly,
 the namespace gets filled with three additional types. For example,
@@
 
-The animal system modelled so far is a simplified version of the
+The animal system modeled so far is a simplified version of the
 system presented in the left-side of Figure~\ref{fig:pegasus}, which still does not appear to
-justify the \mixinAnn{} annotation.  Now moving on to modelling
+justify the \mixinAnn{} annotation.  Now moving on to modeling
 the complete animal system where \Q@Animal@s will include a \Q@location@
 representing their positions in space, we shall
-see that modelling stateful components using plain Java 8 quickly becomes
+see that modeling stateful components using plain Java 8 quickly becomes
 cumbersome. The \mixinAnn{} annotation comes to rescue here and
@@
 \noindent Only code for methods needing implementation is generated. Thus,
-programmers can easily customize the behaviour for their special needs.
+programmers can easily customize the behavior for their special needs.
 %\end{comment}
@@
 
-\noindent will be accepted. If the programmer wishes to suppress this behaviour
+\noindent will be accepted. If the programmer wishes to suppress this behavior
 and keep the signature as it was, it is sufficient to redefine the \Q@with-@

### 2015-12-09 paper/sections/RelatedWork.tex
spell grammar check
unify US/UK English

@@
 \paragraph{Multiple Inheritance in Object Oriented Languages}
-Many authors have argued in favour or against multiple inheritance.  Multiple
+Many authors have argued in favor or against multiple inheritance.  Multiple
 inheritance is very expressive, but difficult to model and implement, and can
@@
 components that can be applied to various classes as reusable functionality
-units. However, % they suffer from linearisation: the order of mixin application
+units. However, % they suffer from linearization: the order of mixin application
 % is relevant in often subtle and undesired ways. This hinders their usability
 % and the ability of resolving conflicts:
-the linearisation (total ordering) of
+the linearization (total ordering) of
 mixin inheritance cannot provide a satisfactory resolution in some cases and
@@
 raises the complexity, especially when constructors and fields
-are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearised mixins.
+are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearized mixins.
 Scala avoids the object initialization
@@
 such as diamond problem. However this approach has limited expressiveness, and
-suffers from all the problems of linearised mixin composition. % Other languages, such as
+suffers from all the problems of linearized mixin composition. % Other languages, such as
 % Python, also use linearized mixins.
 \begin{comment}
-Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
+Python also offers multiple inheritance via linearized mixins. Indeed, in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
 \Q@ class A use B,C {...new methods...}@\\*
@@
 \noindent Java interfaces and default methods do not use
-linearisation: the semantics of Java \textbf{extends} clause in
+linearization: the semantics of Java \textbf{extends} clause in
 interfaces is unordered and symmetric.
@@
 keyword) inside interfaces. The introduction of default methods opens the gate
-for various flavours of multiple inheritance in Java. Traits offer an algebra
+for various flavors of multiple inheritance in Java. Traits offer an algebra
 of composition operations like sum, alias and exclusion, providing explicit conflict
@@
 Furthermore, we do automatic return type refinement for this kind of methods.
-This feature is comparatively useful in big examples, making operations and behaviours more flexible.%, which we will demonstrate later.
+This feature is comparatively useful in big examples, making operations and behaviors more flexible.%, which we will demonstrate later.
 \item We provide two options for generating setters.
@@
 % \item Mixins allows to name components that can be applied to various classes as
-%   reusable functionality units. However, they suffer from linearisation: the
+%   reusable functionality units. However, they suffer from linearization: the
 %   order of mixin application is relevant in often subtitle and undesired
 %   ways. constraints. This hinders their usability and their ability of resolving
-%   conflicts: the linearisation (total ordering) of mixin inheritance cannot
+%   conflicts: the linearization (total ordering) of mixin inheritance cannot
 %   provide a satisfactory resolution in all cases and restricts the flexibility
@@
 %   considered~\cite{marco09FJigsaw}.  Our approach does not have the
-%   linearisation problem, in that the semantics of Java \textbf{extends} clause
+%   linearization problem, in that the semantics of Java \textbf{extends} clause
 %   is unordered and symmetric.
@@
 some of Scala's language constructs to method calls,
-which can be overridden by DSL implementors. Thus Scala-Virtualized
+which can be overridden by DSL implementer. Thus Scala-Virtualized
 can also reinterpret syntax, and be seen as a form of language tuning.
@@
 static methods, and features such as multiple inheritance of
-interfaces and reabstraction of default methods are also modelled.
+interfaces and reabstraction of default methods are also modeled.

### 2015-12-09 paper/sections/Translation.tex
unify US/UK English

@@
     provided as a parameter. This corresponds to the expressions $\es_i$.
-\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
+\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modeling overloading.
 %   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
@@
   that is well-typed before code generation is also well-typed after the generation.
-The annotation just adds more behaviour without removing any functionality.
+The annotation just adds more behavior without removing any functionality.
 We guarantee \textit{Client coherence}.
@@
 \subsection{Results}
-To formally characterize the behaviour of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
+To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
 \begin{itemize}

### 2015-12-09 paper/sections/CaseStudy.tex
code reformatted. updated bundle.zip, appendix.pdf and paper-without-appendix.pdf

@@
 \begin{tabular}{l|l}
-\lstinputlisting[linerange=34-41]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 &
-\lstinputlisting[linerange=50-57]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=53-63]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
@@
 
-\lstinputlisting[linerange=45-46]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=81-83]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=87-89]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 

### 2015-12-09 paper/sections/Overview.tex
code reformatted. updated bundle.zip, appendix.pdf and paper-without-appendix.pdf

@@
 \begin{tabular}{c|c}
-\includegraphics[height=3.8cm]{PegasusDetail.pdf}\hspace{20pt} &
+\includegraphics[height=4cm]{PegasusDetail.pdf}\hspace{20pt} &
 \begin{minipage}{7cm}
 \vspace{-90pt}
-\lstinputlisting[linerange=6-15]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
 %basicstyle=\ttfamily\scriptsize
@@
 
-\lstinputlisting[linerange=19-21]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
+\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
 

### 2015-12-09 UseMixinLombok/src/pegasusJava/Main.java

@@ -6,11 +6,13 @@ import static java.lang.System.out;
 interface Animal {} // no points yet!
 interface Horse extends Animal{
     default void run() {
-        out.println("running!");}
+        out.println("running!");
+    }
 }
 interface Bird extends Animal{
     default void fly() {
-        out.println("flying!");}
+        out.println("flying!");
+    }
 }
 interface Pegasus extends Horse, Bird {}
 //END_PEGASUS_JAVA

### 2015-12-09 UseMixinLombok/src/test/TestExpression.java

@@ -34,11 +34,14 @@ public class TestExpression {
 interface Exp { int eval(); }
 @Obj interface Lit extends Exp {
     int x();
-    default int eval() { return x(); }}
+    default int eval() {return x();}
+}
 @Obj interface Add extends Exp {
     Exp e1(); Exp e2();
     default int eval() {
-        return e1().eval() + e2().eval();}}
+        return e1().eval() + e2().eval();
+    }
+}
 //END_EXPRESSION_INIT
 
 //BEGIN_EXPRESSION_SUB
@@ -49,12 +52,15 @@ interface Exp { int eval(); }
 //BEGIN_EXPRESSION_PRINT
 interface ExpP extends Exp {String print();}
 @Obj interface LitP extends Lit, ExpP {
-    default String print() {return "" + x();}}
+    default String print() {return "" + x();}
+}
 @Obj interface AddP extends Add, ExpP {
-    ExpP e1(); ExpP e2();//return type refined!
+    ExpP e1(); //return type refined! 
+    ExpP e2(); //return type refined!
     default String print() {
         return "(" + e1().print() + " + " 
-                + e2().print() + ")";}}
+                + e2().print() + ")";}
+}
 //END_EXPRESSION_PRINT
 
 //BEGIN_EXPRESSION_COLLECTLIT

### 2015-12-09 paper/sections/Overview.tex
better code format

@@
 interface Pegasus extends Horse, Bird { // generated code not visible to users
-	static Pegasus of() {return new Pegasus() {};}}
+    static Pegasus of() { return new Pegasus() {}; }
+}
 \end{lstlisting}
@@
 interface Point2D{ int x(); int y();
-  static Point2D of(int x, int y){return new Point2D(){
-    public int x(){return x;} public int y(){return y;}};}}
+    static Point2D of(int x, int y) { return new Point2D() {
+        public int x(){return x;} 
+        public int y(){return y;}};}
+}
 \end{lstlisting}
@@
     int x();    int y(); // getters
-  Point2D withX(int val);    Point2D withY(int val); /* with methods*/ }
+    Point2D withX(int val);    Point2D withY(int val); //with methods 
+}
 \end{lstlisting}
@@
     int x(); int y();
-  default Point2D withX(int val){ ... };    default Point2D withY(int val){ ... }}
+    default Point2D withX(int val){ ... }  default Point2D withY(int val){ ... }}
 \end{lstlisting}

### 2015-12-09 paper/sections/FutureWork.tex
revised two sentences
consistent code format, all updated

@@
 \section{Future work}\label{sec:futurework}
-In this section we discuss potential future work (not a full list).
+In this section we discuss potential future work.
 \begin{comment}
@@
 \begin{lstlisting}
-@Obj interface Point{  int x(); int y();
+@Obj interface Point {  int x(); int y();
     static Point of(int val){return Point.of(val,val);} }
@@
 \begin{lstlisting}
-@Obj interface Point{ int x(); int y();
-  default Point clone(){ return Point.of(0,0);}}//user defined clone
+@Obj interface Point { int x(); int y();
+    default Point clone() { return Point.of(0,0); } } //user defined clone
 \end{lstlisting} 
@@
 \begin{lstlisting}
-interface Point{ ...//as before
-  public static Point of(int _x, int _y){
-    return new Point(){...
-      public Point clone(){ return Point.super.clone();}};}}
+interface Point { ...//as before
+    public static Point of(int _x, int _y) {
+        return new Point() {...
+            public Point clone() { return Point.super.clone();}};} }
 \end{lstlisting}

### 2015-12-09 paper/sections/Overview.tex
consistent code format, all updated

@@
 \begin{lstlisting}
-interface Point2D{ int x(); int y();}
+interface Point2D { int x(); int y(); }
 \end{lstlisting}
@@
 \begin{lstlisting}
-Point2D p = new Point2D(){ public int x(){return 4;} public int y(){return 2;}}
+Point2D p = new Point2D() { public int x(){return 4;} public int y(){return 2;} }
 \end{lstlisting}
@@
 \begin{lstlisting}
-interface Point2D{ int x(); int y();
+interface Point2D { int x(); int y();
     static Point2D of(int x, int y) { return new Point2D() {
@@
 \begin{lstlisting}
-@Obj interface Point2D{ int x(); int y();}
+@Obj interface Point2D { int x(); int y(); }
 \end{lstlisting}
@@
 \begin{lstlisting}
-interface Point2D{ int x(); int y();
-  static Point2D of(int _x, int _y){return new Point2D(){
+interface Point2D { int x(); int y();
+    static Point2D of(int _x, int _y){ return new Point2D(){
         int x=_x; int y=_y;
-    public int x(){return x;}   public int y(){return y;}
-    Point2D withX(int val){ return Point2D.of(val,this.y());}
-    Point2D withY(int val){ return Point2D.of(this.x(),val);}
-    };}
-  Point2D withX(int val);    Point2D withY(int val);}
+        public int x(){return x;}   public int y(){ return y; }
+        Point2D withX(int val){ return Point2D.of(val,this.y()); }
+        Point2D withY(int val){ return Point2D.of(this.x(),val); }
+    }; }
+  Point2D withX(int val);    Point2D withY(int val); }
 \end{lstlisting}
@@
 \begin{lstlisting}
-@Obj interface Point2D{
+@Obj interface Point2D {
     int x(); int y();
-    default Point2D withX(int val){ ... }  default Point2D withY(int val){ ... }}
+    default Point2D withX(int val){ ... } default Point2D withY(int val){ ... } }
 \end{lstlisting}
@@
 \begin{lstlisting}
-interface Point2D{ int x(); int y();
-  static Point2D of(int _x, int _y){return new Point2D(){
+interface Point2D { 
+    int x(); int y();
+    default Point2D withX(int val){ ... } default Point2D withY(int val){ ... }
+    static Point2D of(int _x, int _y){ return new Point2D(){
         int x=_x;    int y=_y;
-    public int x(){return x;}    public int y(){return y;}
-    };}
-  default Point2D withX(int val){ ... }    default Point2D withY(int val){ ... }}
+        public int x(){return x;}    public int y(){return y;} }; } }
 \end{lstlisting}
@@
 \begin{lstlisting}
-Point3D p=Point3D.of(1,2,3); p=p.withX(42);
+Point3D p = Point3D.of(1,2,3); p = p.withX(42);
 \end{lstlisting}

### 2015-12-09 UseMixinLombok/src/jooq/DatabaseTest.java

@@ -4,16 +4,16 @@ import lombok.Obj;
 
 //BEGIN_FLUENT_DATABASE
 @Obj interface Database {
-	String select(); Database select(String select);
-	String from(); 	 Database from(String from);
-	String where();  Database where(String where);
-	static Database of() {return of("", "", "");}}
+    String select(); Database select(String select);
+    String from();   Database from(String from);
+    String where();  Database where(String where);
+    static Database of() {return of("", "", "");} }
 //END_FLUENT_DATABASE
 
 //BEGIN_FLUENT_DATABASE_EXT
 @Obj interface ExtendedDatabase extends Database {
-	String orderBy(); ExtendedDatabase orderBy(String orderBy);
-	static ExtendedDatabase of() {return of("", "", "","");}}
+    String orderBy(); ExtendedDatabase orderBy(String orderBy);
+    static ExtendedDatabase of() {return of("", "", "","");} }
 //END_FLUENT_DATABASE_EXT
 
 public interface DatabaseTest {

### 2015-12-09 UseMixinLombok/src/pegasusJava/Main.java

@@ -4,12 +4,12 @@ import static java.lang.System.out;
 
 //BEGIN_PEGASUS_JAVA
 interface Animal {} // no points yet!
-interface Horse extends Animal{
+interface Horse extends Animal {
     default void run() {
         out.println("running!");
     }
 }
-interface Bird extends Animal{
+interface Bird extends Animal {
     default void fly() {
         out.println("flying!");
     }

### 2015-12-09 UseMixinLombok/src/test/TestAnimal.java

@@ -4,70 +4,70 @@ import lombok.Obj;
 import static java.lang.System.out;
 
 public class TestAnimal {
-	
-	public static String runTest() {
+    
+    public static String runTest() {
 //BEGIN_USINGHORSE
 Point2D p = Point2D.of(0, 0);
 Horse horse = Horse.of(p);
 horse.location(p.withX(42));
 //END_USINGHORSE
-		String res = "";
-		res += "horse initialized at " + print(horse.location()) + "\n";
-		horse.run();
-		res += "horse.run(); now at " + print(horse.location()) + "\n\n";
-		Bird bird = Bird.of(Point3D.of(10, 10, 10));
-		res += "bird initialized at " + print(bird.location()) + "\n";
-		bird.fly();
-		res += "bird.fly(); now at " + print(bird.location()) + "\n\n";
-		Pegasus pegasus = Pegasus.of(Point3D.of(100, 100, 100));
-		res += "pegasus initialized at " + print(pegasus.location()) + "\n";
-		pegasus.run();
-		res += "pegasus.run(); now at " + print(pegasus.location()) + "\n";
-		pegasus.fly();
-		res += "pegasus.fly(); now at " + print(pegasus.location()) + "\n";
-		return res;
-	}
-	
-	static String print(Point2D p) {
-		return p.x() + ", " + p.y();
-	}
-	
-	static String print(Point3D p) {
-		return p.x() + ", " + p.y() + ", " + p.z();
-	}
-	
+        String res = "";
+        res += "horse initialized at " + print(horse.location()) + "\n";
+        horse.run();
+        res += "horse.run(); now at " + print(horse.location()) + "\n\n";
+        Bird bird = Bird.of(Point3D.of(10, 10, 10));
+        res += "bird initialized at " + print(bird.location()) + "\n";
+        bird.fly();
+        res += "bird.fly(); now at " + print(bird.location()) + "\n\n";
+        Pegasus pegasus = Pegasus.of(Point3D.of(100, 100, 100));
+        res += "pegasus initialized at " + print(pegasus.location()) + "\n";
+        pegasus.run();
+        res += "pegasus.run(); now at " + print(pegasus.location()) + "\n";
+        pegasus.fly();
+        res += "pegasus.fly(); now at " + print(pegasus.location()) + "\n";
+        return res;
+    }
+    
+    static String print(Point2D p) {
+        return p.x() + ", " + p.y();
+    }
+    
+    static String print(Point3D p) {
[…]

### 2015-12-09 UseMixinLombok/src/test/TestExpression.java

@@ -45,8 +45,8 @@ interface Exp { int eval(); }
 //END_EXPRESSION_INIT
 
 //BEGIN_EXPRESSION_SUB
-@Obj interface Sub extends Exp {Exp e1(); Exp e2();
-    default int eval() {return e1().eval() - e2().eval();}}
+@Obj interface Sub extends Exp { Exp e1(); Exp e2();
+    default int eval() {return e1().eval() - e2().eval();} }
 //END_EXPRESSION_SUB
 
 //BEGIN_EXPRESSION_PRINT
@@ -86,5 +86,5 @@ interface ExpC extends Exp { List<Integer> collectLit(); }
 //BEGIN_INDEPENDENT_EXTENSIBILITY
 interface ExpPC extends ExpP, ExpC {}
 @Obj interface LitPC extends ExpPC, LitP, LitC {}
-@Obj interface AddPC extends ExpPC, AddP, AddC {ExpPC e1(); ExpPC e2();}
+@Obj interface AddPC extends ExpPC, AddP, AddC { ExpPC e1(); ExpPC e2(); }
 //END_INDEPENDENT_EXTENSIBILITY

### 2015-12-09 UseMixinLombok/src/test/TestStateInitialization.java

@@ -1,14 +1,14 @@
 package test;
 
 //BEGIN_STATE_INIT
-interface Box{ 
-    default int val(){return 0;} //provided
-    void val(int _val);//provided
-    static Box of(){return new Box(){//generated
-        int val=Box.super.val();
-        public int val(){return val;}
-        public void val(int _val){val=_val;}
-    };}}
+interface Box { 
+    default int val() { return 0; } //provided
+    void val(int _val); //provided
+    static Box of() { return new Box() { //generated
+        int val = Box.super.val();
+        public int val() { return val; }
+        public void val(int _val) { val = _val; }
+    };} }
 //END_STATE_INIT
 
 public class TestStateInitialization {

### 2015-12-09 paper/sections/Overview.tex
clean up packages

@@
 \vspace{-90pt}
-\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_JAVA
+\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
 %basicstyle=\ttfamily\scriptsize
@@
 
-\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasusJava/Main.java}% APPLY:linerange=PEGASUS_INST
+\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
 
@@
 
-\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/pegasusLombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
 
@@
 
-%\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+%\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=POINT2D
 
@@
 %
-%\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
+%\lstinputlisting[linerange=-]{} % APPLY:linerange=POINT_OF
 
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 
-\lstinputlisting[linerange=10-12]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=USINGHORSE
+\lstinputlisting[linerange=10-12]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=USINGHORSE
 
@@
 
-\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 
@@
 
-\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=POINT3D
 
@@
 \saveSpaceFig
-\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
+\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
 \caption{Generated boilerplate code.}
@@
 
-\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=33-46]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINTIMPL
+\lstinputlisting[linerange=-]{} % APPLY:linerange=POINTIMPL
 
@@
 
-\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT
+\lstinputlisting[linerange=-]{} % APPLY:linerange=POINT
 
@@
 
-\lstinputlisting[linerange=15-27]{../UseMixinLombok/src/overview/TestPoint.java} % APPLY:linerange=POINT_OF
+\lstinputlisting[linerange=-]{} % APPLY:linerange=POINT_OF
 \end{comment}
@@
 
-\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D
+\lstinputlisting[linerange=42-48]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=POINT2D
 
-\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
+\lstinputlisting[linerange=52-55]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=POINT3D
 
-\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2015-12-09 UseMixinLombok/src/casestudy/ep/TestExpression.java

@@ -1,4 +1,4 @@
-package test;
+package casestudy.ep;
 
 import java.util.ArrayList;
 import java.util.List;

### 2015-12-09 UseMixinLombok/src/casestudy/mazegame/java8/TestGame.java

@@ -1,4 +1,4 @@
-package gameori;
+package casestudy.mazegame.java8;
 import static java.lang.System.out;
 
 /* Defines a base-door with no particular features */

### 2015-12-09 UseMixinLombok/src/casestudy/mazegame/lombok/TestGame.java

@@ -1,4 +1,4 @@
-package game;
+package casestudy.mazegame.lombok;
 
 import static java.lang.System.out;
 import lombok.Obj;

### 2015-12-09 UseMixinLombok/src/casestudy/sql/DatabaseTest.java

@@ -1,4 +1,4 @@
-package jooq;
+package casestudy.sql;
 
 import lombok.Obj;
 

### 2015-12-09 UseMixinLombok/src/futurework/TestClone.java

@@ -1,0 +1,34 @@
+package futurework;
+
+import lombok.Obj;
+
+
+//now mixin generate clone always. This prevent the user 
+//to define its own
+/*public @Mixin interface CloneTest{
+  int foo();
+  default CloneTest clone(){ return of(2);}
+  public static void main(String[] arg){
+    CloneTest t=CloneTest.of(3);
+    t=t.clone();
+    System.out.println(t.foo());
+  } 
+}*/
+
+public interface TestClone{
+    int foo();
+    public static TestClone of(int val){
+        return new TestClone(){
+            public int foo(){return val;}
+            //public CloneTest clone(){//This is what we should generate. it look sort of sad and hard to explain...
+            //  return CloneTest.super.clone();}
+            //we may want to do similar thing for toString.... may be we just write it in future work?
+        };
+    }
+    default TestClone clone(){ return of(2);}
+    public static void main(String[] arg){
+        TestClone t=TestClone.of(3);
+        t=t.clone();
+        System.out.println(t.foo());
+    } 
+}

### 2015-12-09 UseMixinLombok/src/futurework/TestStateInitialization.java

@@ -1,4 +1,4 @@
-package test;
+package futurework;
 
 //BEGIN_STATE_INIT
 interface Box { 

### 2015-12-09 UseMixinLombok/src/futurework/TestSynchronize.java

@@ -1,4 +1,4 @@
-package test;
+package futurework;
 
 /*
  \yanlin{ what would happen if Classless Java is used in parallel programming

### 2015-12-09 UseMixinLombok/src/marcoTest/CloneTest.java

@@ -1,34 +1,0 @@
-package marcoTest;
-
-import lombok.Obj;
-
-
-//now mixin generate clone always. This prevent the user 
-//to define its own
-/*public @Mixin interface CloneTest{
-  int foo();
-  default CloneTest clone(){ return of(2);}
-  public static void main(String[] arg){
-    CloneTest t=CloneTest.of(3);
-    t=t.clone();
-    System.out.println(t.foo());
-  } 
-}*/
-
-public interface CloneTest{
-  int foo();
-  public static CloneTest of(int val){
-    return new CloneTest(){
-      public int foo(){return val;}
-      //public CloneTest clone(){//This is what we should generate. it look sort of sad and hard to explain...
-      //  return CloneTest.super.clone();}
-      //we may want to do similar thing for toString.... may be we just write it in future work?
-      };
-  }
-  default CloneTest clone(){ return of(2);}
-  public static void main(String[] arg){
-    CloneTest t=CloneTest.of(3);
-    t=t.clone();
-    System.out.println(t.foo());
-  } 
-}

### 2015-12-09 UseMixinLombok/src/marcoTest/MarcoTest.java

@@ -1,31 +1,0 @@
-package marcoTest;
-
-import lombok.Obj;
-
-@Obj
-interface Point {
-    int x();
-    int y();
-    Point withX(int x);
-    Point withY(int y); //need discussion: whether this line is needed.
-    default String toS() {
-        return "[" + x() + "," + y() + "]";
-    }
-}
-interface Colored {
-    int color();
-    Colored withColor(int val);
-}
-@Obj interface CPoint extends Point,Colored{}
-public class MarcoTest {
-    public static void main(String[] arg) {
-        CPoint p = CPoint.of(10, 20, 0);
-        Point p2 = p.withX(12);
-        p = p.withX(12);
-        System.out.println(p.toS());
-    }
-}
-
- interface A {void m();}
-interface B {void m();}
-interface C extends A, B {}

### 2015-12-09 UseMixinLombok/src/marcoTest/StaticTest.java

@@ -1,19 +1,0 @@
-package marcoTest;
-
-public interface StaticTest {
-  int foo();
-  class Foo{private static StaticTest of(int foo){
-    return new StaticTest(){public int foo(){return foo;}};
-  }}
-  static StaticTest of(int foo){
-    if(foo<0)foo=-foo;
-    return Foo.of(foo);
-  }
-  
-}
-
-interface StaticSee extends StaticTest{
-  static StaticTest boom(){
-    return StaticTest.Foo.of(-2);
-    }
-  }

### 2015-12-09 UseMixinLombok/src/overview/TestPoint.java

@@ -1,55 +1,0 @@
-package overview;
-
-import lombok.Obj;
-
-//BEGIN_POINT
-@Obj
-interface Point {
-    int x();
-    int y();
-}
-//END_POINT
-
-/****** generated
-//BEGIN_POINT_OF
-  // inside interface Point
-  static Point of(int X, int Y) {
-      return new Point() {
-          int _X = X;
-          public int X() {
-              return _X;
-          }
-          int _Y = Y;
-          public int Y() {
-              return _Y;
-          }
-      }; 
-  }  
-//END_POINT_OF
-*/
-
-/******* code without @Mixin
-//BEGIN_POINTIMPL
-class PointImpl implements Point {
-    private int _x;
-    private int _y;
-    public PointImpl(int x, int y) {
-        this._x = x;
-        this._y = y;
-    }
-    public int x() {
-        return _x;
-    }
-    public int y() {
-        return _y;
-    }
-}
-//END_POINTIMPL
- * */
-
-public class TestPoint {
-    public static void main(String[] args) {
-        Point p1 = Point.of(2, 3);
-        System.out.println("p1 = Point.of(2, 3);");
-    }
-}

### 2015-12-09 UseMixinLombok/src/pegasus/TestAnimal.java

@@ -1,4 +1,4 @@
-package test;
+package pegasus;
 
 import lombok.Obj;
 import static java.lang.System.out;

### 2015-12-09 UseMixinLombok/src/pegasus/simple/java8/Main.java

@@ -1,4 +1,4 @@
-package pegasusJava;
+package pegasus.simple.java8;
 
 import static java.lang.System.out;
 

### 2015-12-09 UseMixinLombok/src/pegasus/simple/lombok/Main.java

@@ -1,4 +1,4 @@
-package pegasusLombok;
+package pegasus.simple.lombok;
 
 import lombok.Obj;
 import static java.lang.System.out;

### 2015-12-09 UseMixinLombok/src/test/AutoTest.java

@@ -1,12 +1,9 @@
-package UnitTest;
+package test;
 
 import static org.junit.Assert.assertEquals;
-
 import org.junit.Test;
-
-import test.TestAnimal;
-import test.TestExpression;
-import test.TestLombok;
+import pegasus.TestAnimal;
+import casestudy.ep.TestExpression;
 
 public class AutoTest {
 	@Test
@@ -18,14 +15,7 @@ public class AutoTest {
 		s += "(3 + 4) = 7" + "\n";
 		assertEquals(s, TestExpression.runTest());
 	}
-	@Test
-	public void testLombok() {
-		String s = "p1 = Point.of(2, 3);" + "\n";
-		s += "print p1: 2 3" + "\n";
-		s += "p1.x(4); p1.y(5);" + "\n";
-		s += "print p1: 4 5" + "\n";
-		assertEquals(s, TestLombok.runTest());
-	}
+
 	@Test
 	public void testAnimal() {
 		String s = "horse initialized at 42, 0" + "\n";

### 2015-12-09 UseMixinLombok/src/test/TestLombok.java

@@ -1,97 +1,0 @@
-package test;
-
-import lombok.Obj;
-
-public class TestLombok {
-	public static String runTest() {
-		String res = "";
-		Point p1 = Point.of(2, 3);
-		res += "p1 = Point.of(2, 3);\n";
-		res += "print p1: " + p1.x() + " " + p1.y() + "\n";
-		p1.x(4); p1.y(5);
-		res += "p1.x(4); p1.y(5);\n";
-		res += "print p1: " + p1.x() + " " + p1.y() + "\n";
-		return res;
-	}
-}
-
-@Obj
-interface Point {
-    int x();
-    int y();
-    void x(int x);
-    void y(int y);
-    default int distance() {
-        return (int) Math.sqrt(x() * x() + y() * y());
-    }
-}
-
-/****** generated
-    // inside interface Point
-    // Yanlin: it generated withX, withY and clone methods even if users are not 
-    // requiring them. I think this may not be a proper design decision, because 
-    // it enforces users to have this methods, which are breaking encapsulation.
-    // I think refined withX withY clone methods should be generated inside 
-    // Point3D only Point is declaring these methods. 
-    default Point withX(int x) {
-        return of(x, y());
-    }
-    default Point withY(int y) {
-        return of(x(), y);
-    }
-    Point clone();
-    static Point of(int x, int y) {
-        return new Point() {
-            int _x = x;
-            public int x() {
-                return _x;
-            }
-            int _y = y;
-            public int y() {
-                return _y;
-            }
-            public void X(int x) {
-                _x = x;
-            }
-            public void Y(int y) {
-                _y = y;
-            }
-            @Override
-            public Point clone() {
-                return of(x(), y());
-            }
-        }; 
-    }
-*/
-
-class PointImpl implements Point {
-    private int _x;
-    private int _y;
-    public PointImpl(int x, int y) {
-        this._x = x;
-        this._y = y;
-    }
-    public int x() {
-        return _x;
-    }
-    public int y() {
-        return _y;
-    }
-    public Point withX(int x) {
-        x(x);
-        return this;
-    }
-    public void x(int x) {
-        _x = x;
-    }
-    public void y(int y) {
-        _y = y;
-    }
-    public Point withY(int y) {
-        y(y);
-        return this;
-    }
-    public Point clone() {
-        return new PointImpl(_x, _y);
-    }
-}

### 2016-01-26 OOPSLA16/paper.tex
OOPSLA template.

\documentclass[10pt,preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.
% nocopyrightspace ...

\usepackage{amsmath,amssymb}
\usepackage{balance,bbm,booktabs}
\usepackage{color}
\usepackage{etex}
\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathpartir,mathptmx,mdwlist,morefloats,multirow}
\usepackage{pstricks,pst-node,pst-tree,pgf}
\usepackage{rotating}
\usepackage{stmaryrd}
\usepackage{theorem}
\usepackage{url}
\usepackage{verbatim}
\usepackage{wrapfig}
\usepackage{xspace}

\usepackage[normalem]{ulem}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{dirtree}

\newcommand{\Or}{\ |\ }

\newcommand{\cL}{{\cal L}}
\input{predef/def.tex}
\input{predef/java.tex}
\hyphenation{}

\pagestyle{plain}

\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Title Text}
\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

\input{sections/Introduction.tex}

\input{sections/Formalization.tex}

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}

\end{document}

### 2016-01-26 OOPSLA16/predef/java.tex
OOPSLA template.

\lstset{ %
language=Java,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
deletekeywords={label}
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

### 2016-01-26 OOPSLA16/sections/Formalization.tex
OOPSLA template.

\section{Formal Semantics}\label{sec:formalization}

\begin{figure*}[t]
\[\begin{array}{llll}
T & ::= & I_1.\cdots.I_n                                                                                                         & \hspace{1in} \textrm{path} \\
e & ::= & x \Or e.m(\overline{e}) \Or T.m(\overline{e}) \Or T.\mathbf{super}.m(\overline{e}) \Or x = e;e' \Or obj         & \hspace{1in} \textrm{expressions} \\
obj & ::= & \mathbf{new}\ T()\ \{\overline{field}\ mh_1\{\mathbf{return}\ e_1;\}\cdots mh_n\{\mathbf{return}\ e_n;\}\}& \hspace{1in} \textrm{object creation} \\
field & ::= & T\ f=x;                                                                                              & \hspace{1in}  \textrm{field declaration} \\
\mathbb{I} & ::= & ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}& \hspace{1in} \textrm{interface declaration} \\
meth & ::= & \mathbf{static}\ mh\{\mathbf{return}\ e;\} \Or \mathbf{default}\ mh\{\mathbf{return}\ e;\} \Or mh; & \hspace{1in} \textrm{method declaration} \\
mh & ::= & T_0\ m(T_1\ x_1\cdots T_n\ x_n)  & \hspace{1in} \textrm{method header} \\
ann & ::= & \mathbf{@Family} \Or \emptyset & \hspace{1in} \textrm{annotations} \\
\Gamma & ::= & x_1:T_1\cdots x_n:T_n & \hspace{1in} \textrm{environment}
\end{array}\]
\caption{Grammar.}
\end{figure*}

\begin{figure*}[t]
\[\begin{array}{lll}
Exact(T_1.I,\ T_2) & = & Exact(I,\ Exact(T_1,\ T_2)) \\
Exact(I,\ I_1.\cdots.I_n) & = & Exact2(I,\ I_1.\cdots.I_i) \\
& & \ \ \ \ \textrm{for the smallest }i\in[0,n],\forall j\in(i,n],Exact2(I,\ I_1.\cdots.I_j) = \emptyset \\
Exact2(I,\ \emptyset) & = & CT(I) \\
Exact2(I,\ T_0.I_0) & = & \mathbb{I} \\
& & \ \ \ \ \textrm{if } \mathbb{I}\in\overline{\mathbb{I}} \textrm{ and } name(\mathbb{I})=I \\
& & Exact(I,\ Exact(T,\ T_0)) \\
& & \ \ \ \ \textrm{if } T\in\overline{T} \textrm{ and }\forall T'\in\overline{T}, Exact(I,\ Exact(T',\ T_0))\in\{\ Exact(I,\ Exact(T,\ T_0)),\ \emptyset\ \} \\
& & \textrm{where } CT(T_0.I_0) = ann\ \mathbf{interface}\ I_0\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}
\end{array}\]
\caption{Typing rules.}
\end{figure*}

### 2016-01-26 OOPSLA16/sections/Introduction.tex
OOPSLA template.

--- 2015-12-09 paper/sections/Introduction.tex
@@
 
-Java 8 introduced \emph{default methods}, allowing interfaces to have
-method implementations. The main motivation behind the introduction of
-default methods in Java 8 is \emph{interface evolution}. That is, to
-allow interfaces to be extended over time while preserving backwards
-compatibility. It soon became clear that default methods could also be
-used to emulate something similar to \emph{traits}~\cite{scharli03traits}. The
-original notion of traits by Scharli et al. prescribes, among other
-things, that: 1) a trait provides a set of methods that implement
-behaviour; and 2) a trait does not specify any state variables, so the
-methods provided by traits do not access state variables
-directly. Java 8 interfaces follow similar principles too. Indeed, a
-detailed description of how to emulate trait-oriented programming in
-Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
-team designing default methods, was also fully aware of that secondary
-use of interfaces, but it was not their objective to model traits:
-``The key goal of adding default methods to Java was "interface
-evolution", not "poor man's traits"''~\cite{goetz13default}. As a result
-they were happy to support the secondary use of interfaces with
-default methods as long as it did not make the implementation and
-language more complex.
-
-Still, the design is quite conservative and appears to be quite limited
- in its current form to model advanced forms of multiple inheritance.
-Indeed, our own personal experience of combining default methods 
-and multiple interface inheritance in Java to achieve multiple implementation 
-inheritance is that many workarounds and boilerplate code are needed. 
-In particular, we encountered difficulties because:
-
-\begin{itemize}
-
-\item {\em Interfaces have no constructors.} As a result classes are 
-still required to create objects, leading to substantial boilerplate 
-code for initialization.
-
-\item {\em Interfaces do not have state.} This creates a tension between 
- using multiple inheritance and having state. Using setter and
-  getter methods is a way out of this tension, but this workaround
-  requires tedious boilerplate classes that later implement those
-  methods.
-
-\item {\em Useful, general purpose methods require special care in
-  the presence of subtyping.} Methods such as
-  \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
-  internal state of an object, but they also require their return types to be
-  refined in subtypes.
-
-\end{itemize}
-
-\noindent Clearly, a way around those difficulties would be to
-change Java and just remove these limitations. Scala's own
-notion of traits, for example, allows state in traits. Of course adding
-state (and other features) to interfaces would complicate the language
-and require changes to the compiler, and this would go beyond the
-goals of Java 8 development team.
-
-This paper takes a different approach. Rather than trying to get
-around the difficulties by changing the language in fundamental ways,
-we show that, with a simple language feature, default methods and
-interface inheritance are in fact very expressive. Our proposed
-language feature enables powerful object-oriented idioms, using
-multiple inheritance. We call the language feature \emph{object
-  interfaces}, because such interfaces can be instantiated directly,
-without the need for an explicit class definition. Moreover, object
-interfaces support \emph{abstract state operations}, providing a way
-to use multiple inheritance with state in Java. The abstract state
-operations include various common utility methods (such as getters and
-setters, or clone-like methods). In the presence of subtyping, such
-operations often require special care, as their types need to be
-refined. Object interfaces provide support for type-refinement and can
-automatically produce code that deals with type-refinement
-adequately. %With object interfaces, many Java programs can be built
-%without using a single class!
-
-Object interfaces do not require changes to the Java runtime or compiler, 
-and they also do not introduce any new syntax. All three features of object interfaces are
-achieved by reinterpreting existing Java syntax, and are translated
-into regular Java code without loss of type-safety. Since no new
-syntax is introduced, it would be incorrect to call object interfaces
-a language extension or syntactic sugar. So we use the term
-\emph{language tuning} instead. Language tuning sits in between a
-lightweight language extension and a glorified library. Language
-tuning can offer many features usually implemented by a real language
-extension, but because it does not modify the language syntax
-pre-existing tools can work transparently on the tuned language.  To
-exploit the full benefits of language tuning, our prototype
-implementation of object interfaces uses Java annotations to do AST
-rewriting, allowing existing Java tools (such as IDEs) to work
-out-of-the-box with our implementation. As a result we could
-experiment object interfaces with several interesting Java programs,
-and conduct various case studies.
-
-To formalize object interfaces, we propose Classless Java (CJ): a
-FeatherweightJava-style~\cite{Igarashi01FJ} calculus, which captures the essence
-of interfaces with default methods. The semantics of object interfaces
-is given as a syntax-directed translation from CJ to itself. In the
-resulting CJ code all object interfaces are translated into regular CJ
-(and Java) interfaces with default methods. The translation is proved
-to be type-safe, ensuring that the translation does not
-introduce type-errors in client code. 
-%CJ's usefulness goes beyond serving as a
-%calculus to formalize object interfaces. During the development
-%process of CJ, we encountered a bug in the implementation of default
-%methods for the Eclipse Compiler for Java (ECJ). For the program revealing the 
-%bug, ECJ behaves differently from both our formalization and Oracle's 
-%Java compiler.
-
-To evaluate the usefulness of object interfaces, we illustrate 3
-applications. The first application is a simple 
-solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
-extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
-application shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
-can be easily defined using object interfaces. The last
-application is a larger case study for a simple Maze game implemented with 
-multiple inheritance. For the last application we show that there is a
-significant reduction in the numbers of lines of code when compared 
-to an existing implementation~\cite{bono14} using plain Java 8. 
-Noteworthy, all applications are implemented 
-without defining a single class!
-
-In summary, the contributions of this paper are:
-\begin{itemize}
-
-\item {\bf Object Interfaces:} A simple feature that allows various
-  powerful multiple-inheritance programming idioms to be expressed
-  conveniently in Java.
-
-\item {\bf Classless Java (CJ):} A simple formal calculus that models 
-the essential features of Java 8 interfaces with default methods, and 
-can be used to formally define the translation of object interfaces. 
-We prove several properties of the translation\footnote{Proofs and prototype implementation are available in
-  the supplementary materials.}.
-
-\item {\bf Implementation and Case Studies:} We have a prototype
-  implementation of object interfaces, using Java
-  annotations and AST rewriting. Moreover the usefulness of object interfaces is
-  illustrated through various examples and case studies.
-
-\item{\bf Language Tuning:} We identify the concept of language tuning 
-and describe object interfaces as an example. We also discuss 
-how other existing approaches, such as the annotations in project 
-Lombok~\cite{lombok},  can be viewed as language tuning.
-
-\end{itemize}
+The text of the paper begins here.

### 2016-01-26 OOPSLA16/experiments/Lombok_and_AnnoProcessor/Test.java

@@ -1,0 +1,13 @@
+import lombok.Obj;
+
+import com.zewei.annotation.processor.Algebra;
+
+@Algebra@Obj
+interface Test {
+	int x();
+}
+
+// @Obj: generate the of() method
+// @Algebra: print out all the method names
+// log1.txt: @Obj@Algebra
+// log2.txt: @Algebra@Obj

### 2016-02-23 OOPSLA16/paper.tex
OOPSLA format.

@@
 
-\usepackage{amsmath,amssymb}
-\usepackage{balance,bbm,booktabs}
+\usepackage{listings}
+\usepackage{multicol}
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+\usepackage[table,xcdraw]{xcolor}
 \usepackage{color}
-\usepackage{etex}
+\usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{stmaryrd}
+\usepackage{graphicx}
+\usepackage{amssymb}
 \usepackage{fancyvrb}
+\usepackage{url}
+\usepackage{pstricks,pst-node,pst-tree}
+\usepackage{bbm}
+\usepackage{pgf}
+\usepackage{multirow}
+
+\usepackage{listings}
+\usepackage{verbatim}
 \usepackage{graphicx}
-\usepackage{listings}
-\usepackage{mathpartir,mathptmx,mdwlist,morefloats,multirow}
-\usepackage{pstricks,pst-node,pst-tree,pgf}
-\usepackage{rotating}
-\usepackage{stmaryrd}
-\usepackage{theorem}
-\usepackage{url}
-\usepackage{verbatim}
 \usepackage{wrapfig}
+
+\usepackage[T1]{fontenc}
+\usepackage[scaled=0.85]{beramono}
+\usepackage{mathpartir}
 \usepackage{xspace}
 
-\usepackage[normalem]{ulem}
-\usepackage[T1]{fontenc}
-\usepackage[scaled=0.85]{beramono}
-\usepackage{dirtree}
+\input{predef/def.tex}
+\input{predef/java.tex}
 
@@
 \newcommand{\cL}{{\cal L}}
-\input{predef/def.tex}
-\input{predef/java.tex}
 \hyphenation{}
@@
 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
+\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
+\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
+\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
+
+\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
+\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
+\newcommand\tsem[1]{\llbracket #1 \rrbracket}
+\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
+\newcommand\nat[0]{\mathbb{N}}
+\newcommand\unit[0]{\mathbbm{1}}
+
+\newcommand\mixin{\mixinAnn\xspace}
+\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
+\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
+
+%\theoremstyle{plain}
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}{Lemma}
+\newtheorem{thm2}{Theorem}
+\newtheorem{lem2}{Lemma}
 
@@
 
-\title{Title Text}
+\title{Classless Java: Tuning Java Interfaces}
 \subtitle{Subtitle Text, if any}
@@
 \begin{abstract}
-This is the text of the abstract.
+Java 8 introduced \emph{default methods}, allowing interfaces to have method
+  implementations. When combined with (multiple) interface inheritance, default
+  methods provide a basic form of multiple inheritance. However, using this
+  combination to simulate more advanced forms of multiple inheritance quickly becomes cumbersome, and
+  appears to be quite restricted.
+
+  This paper shows that, with a simple language feature, default methods and
+  interface inheritance are in fact very expressive. Our proposed language
+  feature, called \emph{object interfaces}, enables powerful object-oriented
+  idioms, using multiple inheritance, to be expressed conveniently in
+  Java. Object interfaces refine conventional Java interfaces in three different
+  ways. Firstly, object interfaces have their own object instantiation
+  mechanism, providing an alternative to class constructors. Secondly, object
+  interfaces support \emph{abstract state operations}, providing a way to use
+  multiple inheritance with state in Java. Finally, object interfaces allow type
+  refinements that are often tricky to model in conventional class-based
+  approaches. Interestingly, object interfaces do not require changes to the
+  runtime, and they also do not introduce any new syntax: all three features are
+  achieved by reinterpreting existing Java syntax, and are translated into
+  regular Java code without loss of type-safety. Since no new syntax is
+  introduced, it would be incorrect to call object interfaces a language
+  extension or syntactic sugar. So we use the term \emph{language tuning} to
+  characterize this kind of language feature. An implementation of object
+  interfaces using Java annotations and a formalization of the static and
+  dynamic semantics are presented. Moreover, the usefulness of object interfaces
+  is illustrated through various examples.
 \end{abstract}
@@
 
+\input{sections/Overview.tex}
+
 \input{sections/Formalization.tex}
 
+\input{sections/Translation.tex}
+
+\input{sections/Implementation.tex}
+
+\input{sections/CaseStudy.tex}
+
+\input{sections/RelatedWork.tex}
+
+\input{sections/FutureWork.tex}
+
+\input{sections/Conclusion.tex}
+
 \appendix
-\section{Appendix Title}
-
-This is the text of the appendix, if you need one.
+\input{sections/Appendix.tex}
 
@@
 \bibliographystyle{abbrvnat}
+\bibliography{paper}
 
@@
 
-\begin{thebibliography}{}
-\softraggedright
+%\begin{thebibliography}{}
+%\softraggedright
 
-\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
-P. Q. Smith, and X. Y. Jones. ...reference text...
+%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
+%P. Q. Smith, and X. Y. Jones. ...reference text...
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2016-02-23 OOPSLA16/sections/CaseStudy.tex
OOPSLA format.

--- 2015-12-09 paper/sections/CaseStudy.tex
@@
 \begin{tabular}{l|l}
-\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 &
-\lstinputlisting[linerange=53-63]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=53-63]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
@@
 
-\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=87-89]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=87-89]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 
@@
 \texttt{select}, \texttt{from} and \texttt{where} clauses:
-\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
+\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
 % \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
@@
 
-\lstinputlisting[linerange=31-31]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
+\lstinputlisting[linerange=31-31]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
 
@@
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
-\lstinputlisting[linerange=14-16]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+\lstinputlisting[linerange=14-16]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
 In this way, when a query created using \texttt{ExtendedDatabase}, 
@@
 
-\lstinputlisting[linerange=37-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\lstinputlisting[linerange=37-38]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
 

### 2016-02-23 OOPSLA16/sections/Formalization.tex
OOPSLA format.

@@
-\section{Formal Semantics}\label{sec:formalization}
+\section{Interaction of Interface Methods with Interface Composition}
+Before formalizing Classless Java and object interfaces, it is helpful
+to informally discuss the behavior of methods in Java 8
+interfaces. In particular it is useful to understand how Java 8
+interfaces differ from conventional trait models.
 
-\begin{figure*}[t]
-\[\begin{array}{llll}
-T & ::= & I_1.\cdots.I_n                                                                                                         & \hspace{1in} \textrm{path} \\
-e & ::= & x \Or e.m(\overline{e}) \Or T.m(\overline{e}) \Or T.\mathbf{super}.m(\overline{e}) \Or x = e;e' \Or obj         & \hspace{1in} \textrm{expressions} \\
-obj & ::= & \mathbf{new}\ T()\ \{\overline{field}\ mh_1\{\mathbf{return}\ e_1;\}\cdots mh_n\{\mathbf{return}\ e_n;\}\}& \hspace{1in} \textrm{object creation} \\
-field & ::= & T\ f=x;                                                                                              & \hspace{1in}  \textrm{field declaration} \\
-\mathbb{I} & ::= & ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}& \hspace{1in} \textrm{interface declaration} \\
-meth & ::= & \mathbf{static}\ mh\{\mathbf{return}\ e;\} \Or \mathbf{default}\ mh\{\mathbf{return}\ e;\} \Or mh; & \hspace{1in} \textrm{method declaration} \\
-mh & ::= & T_0\ m(T_1\ x_1\cdots T_n\ x_n)  & \hspace{1in} \textrm{method header} \\
-ann & ::= & \mathbf{@Family} \Or \emptyset & \hspace{1in} \textrm{annotations} \\
-\Gamma & ::= & x_1:T_1\cdots x_n:T_n & \hspace{1in} \textrm{environment}
-\end{array}\]
-\caption{Grammar.}
-\end{figure*}
+%We show several interesting cases
+%and summarize compilation result of method composition into 3 categories:
+%conflict error, accepted (both abstract) and conservative error. Meanwhile we
+%also show different composition handling mechanism among traits, javac and ECJ.
 
-\begin{figure*}[t]
-\[\begin{array}{lll}
-Exact(T_1.I,\ T_2) & = & Exact(I,\ Exact(T_1,\ T_2)) \\
-Exact(I,\ I_1.\cdots.I_n) & = & Exact2(I,\ I_1.\cdots.I_i) \\
-& & \ \ \ \ \textrm{for the smallest }i\in[0,n],\forall j\in(i,n],Exact2(I,\ I_1.\cdots.I_j) = \emptyset \\
-Exact2(I,\ \emptyset) & = & CT(I) \\
-Exact2(I,\ T_0.I_0) & = & \mathbb{I} \\
-& & \ \ \ \ \textrm{if } \mathbb{I}\in\overline{\mathbb{I}} \textrm{ and } name(\mathbb{I})=I \\
-& & Exact(I,\ Exact(T,\ T_0)) \\
-& & \ \ \ \ \textrm{if } T\in\overline{T} \textrm{ and }\forall T'\in\overline{T}, Exact(I,\ Exact(T',\ T_0))\in\{\ Exact(I,\ Exact(T,\ T_0)),\ \emptyset\ \} \\
-& & \textrm{where } CT(T_0.I_0) = ann\ \mathbf{interface}\ I_0\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}
-\end{array}\]
-\caption{Typing rules.}
-\end{figure*}
+\subsection{Methods in Java 8 Interfaces}
+In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
+
+\paragraph{Static methods} are handled in a very clean way: they are visible only in
+  the interface in which they are explicitly defined. This means the following code
+  is ill-typed.
+\begin{lstlisting}
+interface A0 { static int m(){return 1;} }
+interface B0 extends A0 {}
+... B0.m()//ill typed
+\end{lstlisting}
+This is different from the way static methods are handled in classes. Here
+static methods have simply no interaction with interface
+composition (\Q@extends@ or \Q@implements@).
+
+\paragraph{Abstract method} composition is accepted when there exists a most specific one.
+  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+
+\paragraph{Default methods} conflict with any other default or abstract method. For
+  example the following code is rejected due to method conflicts.
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is different from what happens in most trait models, where \Q@D2@
+would be accepted, and the implementation in \Q@A2@ would be part of the
+behavior of \Q@D2@.
+
+\paragraph{Resolving conflicts:}
+A method in the current interface wins over any method in its
+super-interfaces, provided that the method
+is the most specific one. This method also overrides conflict due to
+inheritance. For example, the following code is accepted, but would be rejected
+(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
+\texttt{E3}.
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
+\end{lstlisting}
+
+\subsection{Classifying Outcomes of Interface Composition}
+%When interfaces are composed and methods with the same name (and
+%signature) exist, there are 3 possible outcomes.
+%
+We now try to classify possible outcomes for composition of methods with the same name (and
+signature).
+We will use the following (correct) declarations:
+\begin{lstlisting}
+interface A1 { T m(); }
+interface A2 extends A1 { default T m(){ ... } }
+interface A3 extends A2 { T m(); }
+
+interface B1 { default T m(){ ... } }
+interface B2 extends B1 { T m(); }
+interface B3 extends B2 { default T m(){ ... } }
+\end{lstlisting}
+
+\noindent
+What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
+\Q@B@${}_j$?\\*
+% Nine representative cases are shown next:\\*
+\noindent
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
+\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
+\end{tabular}
+\\*
+%We try to classify the results in the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
+\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
+  trait models.
+\item \textbf{conservative error} happens when only one method is implemented
+  (leaving another one abstract), which is different from what we would expect in
+  a trait model, but is coherent with the conservative idea that a method
+  defined in an interface should not silently satisfy a method in another one.
+\end{itemize}
+
+\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
+ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
+(leading Java 8 designer) we have confirmed that the expected behavior is
+rejection, hence this is a bug in ECJ. This bug was also reported by
+  others and is fixed in the ECJ developer branch, but not released as a stable
+  version yet.
+%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
+%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+%named as \texttt{m} which are either abstract or default methods. If we define a
+%new interface \texttt{M} that extends two of these interfaces, then the method
+%overriding result is shown in the table. For example, row2 col2 means \texttt{M
+%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+%following this way of interpretation.
+
+% The interesting case in row4 col2 where
+%Indeed, to be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
+%We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
+%We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
+%the experimental result is \texttt{B3.m}. Because if we think consistently, this
+
+%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+%In our approach, we choose to not model this strange behaviour (a bug?).
+%Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
+%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java changes its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
+
+\section{Formal Semantics}\label{sec:formal}
+
+This section presents a formalization of Classless Java: a minimal
+FeatherweightJava-like calculus which models the essence of Java
+interfaces with default methods. This formalization is used in
+Section~\ref{sec:translation} to define the semantics of object interfaces.
+
+\begin{figure}[t]
+\saveSpaceFig
+\begin{grammar}
+\production{
+\e
+}{
+  \x\mid\MCall\e\m\es\mid\MCall{\C}\m\es\mid\MCall{\C\QM{.super}}\m\es\mid\x\QM=\e\QM;\e'\mid\obj
+  }{expressions}\\
+\production{
+\obj
+}{
+\QM{new}\ \C\oR\cR\oC\fields\
+\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
+\ldots
+\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
+\cC
+  }{object creation}\\
+\production{\field}{\T\ \f \QM= \x\QM;}{field declaration}\\
+\production{
+\II
+}{
+ \ann\ \QM{interface}\ \C\ \QM{extends}\ \Cs\ \oC \methods\ \cC
+  }{interface declaration}\\
+\production{
+\method
+}{
+ \QM{static}\ \mh\ \oC\QM{return}\ \e\QM{;}\!\cC
+\mid
+\QM{default}\ \mh\ \oC\QM{return}\ \e\QM{;}\cC
+\mid
+\mh\QM{;}
+  }{method declaration}\\
+\production{
+\mh
+}{
+ \T_0\ \m\ \oR\T_1\ \x_1\ldots\T_n\ \x_n\cR
+  }{method header}\\
+\production{
+\ann
+}{
+  \mixinAnn|\emptyset
+  }{annotations}\\
+\production{\Gamma}{
+\x_1{:}\C_1\ldots\x_n{:}\C_n
+}{environment}
+\end{grammar}
+\caption{Grammar of Classless Java}
+\label{Grammar}
+\end{figure}
+
+\subsection{Syntax}
+
+Figure~\ref{Grammar} shows the syntax of Classless Java.
+The syntax formalizes a minimal
+subset of Java 8, focusing on interfaces, default methods and object
+creation literals.  There is no syntax for classes.
+To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all map to a single set of identifiers as in Java.  Expressions
+consist of conventional constructs such as variables ($\x$), method
+calls ($\e\QM.\m\QM(\es)$) and static method calls
+($\C\QM.\m\QM(\es)$). For simplicity the degenerate case of calling a
+static method over the $\this$ receiver is not considered.  A more
+interesting type of expressions is super calls
+($\C\QM{.super.}\m\QM(\es)$), whose semantics is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
+dispatching to the version of the method as visible in the interface
+$\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
+modeled. In the syntax of field updates $\x$ is expected to be a
+field name. After updating the field $\x$ using the value of $\e$, the
+expression $\e'$ is executed. To blend the statement based nature of
+Java and the expression based nature of our language, we consider a
+method body of the form \Q@return@ $\x\QM=\e\QM;\e'$ to represent
+$\x\QM=\e\QM;\QM{return}\ \e'$ in Java.  Finally, there is an object
+initialization expression from an interface $\C$, where (for
+simplicity) all the fields are initialized with a variable present in
+scope.
+To  be fully compatible with Java, the concrete syntax for an interface
+  declaration with empty supertype list  would also
+  omit the \Q@extends@ keyword.
+% The single
+%non-Java 8 piece of syntax is the \mixin annotation, which is the only
+%one interesting piece of syntax in this article.
+%??? how is @Mixin not java8? in the sense that is java5?
+  Following standard
+practice, we consider a global Interface Table (\metaVar{IT}) mapping
+from interface names $\C$ to interface declarations $\II$.
+
+The environment $\Gamma$ is a mapping from variables to types.  As
+usual, we allow a functional notation for $\Gamma$ to do variable
+lookup.  Moreover, to help us define auxiliary functions, a functional
+notation is also allowed for a set of methods $\methods$, using the
+method name $\m$ as a key.  That is, we define $\methods(\m)=\method$
+iff there is a unique $\method\in\methods$ whose name is $\m$.  For
+convenience, we define $\methods(\m)=\none$ otherwise; moreover
+$\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their names.
+
+\subsection{Typing}
+
+Typing statement $\Gamma \vdash \e\in\C$ reads ``in the environment
+$\Gamma$, expression $\e$ has type $\C$.''.
+Before discussing the typing rules we discuss some of the used notation.
+As a shortcut, we write
+$\Gamma \vdash \e\in\C<:\C'$ instead of $\Gamma \vdash \e\in\C$ and
+$\C<:\C'$.
+
+We omit the definition of
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.
+The auxiliary notation $\Gamma^\mh$ trivially
+extracts the environment from a method header, by collecting the all types
+and names of the method parameters.  The
+notation $\m^\mh$ and $\C^\mh$ denotes respectively, extracting the
+method name and the return type from a method header. $\mBody(\m,\C)$,
+defined in Section~\ref{sec:auxiliary},
+returns the full method declaration as seen by $\C$, that is the
+method $\m$ can be declared in $\C$ or inherited from another
+interface.
+$\textsf{mtype}(\m,\C)$ and $\textsf{mtypeS}(\m,\C)$ return the type
+signature from a method (using $\mBody(\m,\C)$ internally).
+$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
+$\textsf{mtypeS}(\m,\C)$ only for static ones. We use $\dom(\C)$ to
+denote the set of methods that are defined for type $\C$, that is:
+$\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
+
+\begin{figure}[t]
+\saveSpaceFig
+$
+\begin{array}{l}
+
+%% T-Invk
+\inferrule[(T-Invk)]{
+ \Gamma \vdash \e \in \C_0 \\\\
+\forall i\in 1..n\ \ \Gamma \vdash \e_i \in \_<:\C_i \\\\
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+%\textsf{mmodifier}(\m,\C) \neq \textbf{static}
+ }{
+ \Gamma \vdash \e\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C }
+\quad\quad
+
+%%T-StaticInvk
+\inferrule[(T-StaticInvk)]{
+\forall i\in 1..n\  \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
+\textsf{mtypeS}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\to\! \C
+%\textsf{mmodifier}(\m,\C) = \textbf{static}
+}{
+\Gamma \vdash \C_0\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
+\quad\quad
+
+%%T-SuperInvk
+\inferrule[(T-SuperInvk)]{
+\Gamma(\this) <: \C_0 \\\\
+\forall i\in 1..n\ \ \Gamma \vdash \e_i\in \_<:\C_i \\\\
+  \textsf{mtype}(\m,\C_0) \!=\! \C_1\ldots\C_n \!\!\to\! \C
+%\textsf{mmodifier}(\m,\C_0) \neq \textbf{static} \\\\
+}{\Gamma \vdash \C_0\QM.\QM{super}\QM.\m\QM(\e_1\ldots\e_n\QM) \in \C}
+
+%%T-Var
+\\[5ex]
+\inferrule[(T-Var)]{
+\Gamma(\x)=\C
+}{
+\Gamma \vdash \x \in\C}
+\quad\quad
+
+%%T-Obj
+\inferrule[(T-Obj)]{
+\forall i\in 1..k\ \ \Gamma(\x_i)\subtype\T_i\\\\
+\forall i\in 1..n\ \
+%\Gamma_i
+\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
+\vdash\e_i\in \_\subtype\C^{\mh_i}\\\\
+\sigvalid(\mh_1\ldots\mh_n,I)\quad\quad\quad\quad
+%\forall i\in 1..n\ \mh_i\subtype\mBody(\m^{\mh_i},\C)\\\\
+\alldefined(\mh_1\ldots\mh_n,I)
+%\forall\m\mbox{ such that }
+%\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
+%\forall i\in 1\ldots n\ \Gamma_i=\Gamma,\f_1{:}\T_1,\ldots,\f_k{:}\T_k,\,\QM{this}{:}\C,\Gamma^{\mh_i}
+}{
+\Gamma \vdash\QM{new}\ \C\oR\cR\oC\T_1\ \f_1\QM=\x_1\QM;\ldots\T_k\ \f_k\QM=\x_k\QM;\
+\mh_1\oC\QM{return}\ \e_1\QM{;}\!\cC
+\ldots
+\mh_n\oC\QM{return}\ \e_n\QM{;}\!\cC
+\cC
+\in\C
+}
+\\[5ex]
+
+%%T-update
+\quad
+\inferrule[(T-update)]{
+\Gamma \vdash \e\in\_<:\Gamma(\x)\\\\
+\Gamma \vdash \e'\in\C
+}{
+\Gamma \vdash \x\QM=\e\QM;\e'\in\C }
+\quad\quad\quad
+
+%%T-Intf
+ \inferrule[(T-Intf)]{
+IT(\C) = \ann\ \QM{interface}\ \C\ \QM{extends}\ \C_1\ldots\C_n \
+\oC\methods\ \cC\\\\
+ \forall \QM{default}\ \mh\ \oC\QM{return}\ \e\QM;\cC \in \methods,
+\ \ \Gamma^{\mh},\,\QM{this}{:}\C\vdash\e\in \_\subtype\C^{\mh} \\\\
+ \forall \QM{static}\ \mh\ \oC\QM{return}\ \e\QM;\cC \in \methods,
+\ \ \Gamma^{\mh}\vdash\e\in \_\subtype\C^{\mh} \\\\
+\dom(\C)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)
+ }{
+\C \text{ OK}
+}
+\end{array}$
+\caption{CJ Typing}
+\label{ET}
+\end{figure}
+
+In Figure~\ref{ET} we show the typing rules.  We discuss the
+most interesting rules, that is \rn{t-Obj} and \rn{t-Intf}. Rule
+\rn{t-Obj} is the most complex typing rule. Firstly, we need to
+ensure that all field initializations are type correct, by looking up the type of
+each variable assigned to a field in the typing environment and verifying that such type is a
+subtype of the field type. Finally, we check that all method bodies are
+well-typed. To do this the environment used to check the method body
+needs to be extended appropriately: we add all fields and their types;
+add $\this:I$; and add the arguments (and types) of the respective
+method.
+Now we need to check if the object is a valid extension for that specific interface.
+This can be logically divided into two steps.
+First  we check that all method headers are valid
+with respect to the corresponding method already present in $\C$:
+
+\noindent$\begin{array}{l}
+\InTextDef{10em}{\sigvalid(\mh_1\ldots\mh_n,I)
+}{
+\forall i\in 1..n\ \ \mh_i\QM;\subtype\mBody(\m^{\mh_i},\C)
+}
+\end{array}$
+
+\noindent Here we require that for all newly declared methods, there is a method
+with the same name defined in the interface $\C$, and that such method is a
+supertype of the newly introduced one. We define subtyping between methods in a
+general form that will also be useful later.
+
+\noindent$\begin{array}{l}
+\InTextDef{18em}{
+\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR\QM; \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR\QM;
+}{
+\T\subtype \T'
+}\\
+\InTextDef{18em}{\method \subtype
+\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}
+}{
+\method\subtype\mh\QM;
+}\\
+\InTextDef{18em}{\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method
+}{
+\mh\QM;\subtype\method
+}\\
+\end{array}$
+
+\noindent We allow return type specialization as introduced in Java 5. A method
+header with return type $I$ is a subtype of another method header with return
+type $I'$ if all parameter types are the same, and $I <: I'$. A default method
+$\method_1$ is a subtype of another default method $\method_2$ iff
+$\mh^{\method_1}$ is a subtype of $\mh^{\method_2}$. Secondly, we check that all
+abstract methods (which need to be explicitly overridden) in the interface have
+been implemented: %That is, we define a method with the same name.
+
+\noindent$\begin{array}{l}
+\InTextDef{11em}{\alldefined(\mh_1\ldots\mh_n,I)}{
+\forall\m\mbox{ such that } \mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m}
+\end{array}$
+
+The rule \rn{t-intf} checks that an interface $\C$ is correctly
+typed.  First we check that the body of all default and static
+methods are well typed.  Then we check that $\dom(\C)$ is the same as
+$\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$.  This is not a
+trivial check, since $\dom(\C)$ is defined using $\mBody$, which would be
+undefined in many cases: notably if a method $\method\in\methods$ is
+not compatible with some method in $\dom(\C_1)\ldots\dom(\C_n)$ or if
+there are methods in any $\dom(\C_i)$ and $\dom(\C_j)$ ($i,j\in 1..n$) conflict.
+
+\subsection{Auxiliary Definitions}\label{sec:auxiliary}
+
+\begin{comment}
+\subsubsection{Auxiliary function: \textsf{mtype}}
+- \textsf{mtype(m, C)} : the signature of method m in C.
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \emptyset } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
+\end{comment}
+
+Defining \mBody{} is not trivial, and requires quite a lot of attention to the
+specific model of Java interfaces, and to how it differs w.r.t. Java Class model.
+$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
+interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
+\noindent$\begin{array}{l}
+\InTextDef{15ex}{
+\mBody(\m,\C_0)
+}{
+\override(\methods(\m),
+\tops(\m,\Cs))
+}\\
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n\ \oC\methods\ \cC \mbox{ and }
+ \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n
+
+}
+\end{array}$
+
+\noindent The definition of $\mBody$ reconstructs the full set of supertypes $\Cs$ and then delegates the work to two other auxiliary functions:
+ $\tops(m,\Cs)$ and $\override(\method,\methods)$.
+
+\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
+the non-static ones that are not reachable by another, less specific superinterface.
+Since the second parameter of \tops{} is a set, we can choose an arbitrary element to be $\C_0$.
+In the definition we denote by $\Aux{originalMethod}(\m,\C)=\method$ the non-static method called $\m$ defined directly in $\C$.
+%\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
+%super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
+%Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
+Formally:
+
+\noindent$\begin{array}{l}
+\InTextDef{20ex}{
+\Aux{originalMethod}(\m,\C_0)}{
+\method
+}\\
+\InTextWith{\metaVar{IT}(\C_0) =
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
+ \\
+
+\InTextDef{44ex}{
+\Aux{originalMethod}(\m,\C_0)\in\tops(\m,\C_0\dots\C_n)}{
+ }{}\\
+\tab\tab
+\not\exists i\in1..n \mbox{ such that }
+\Aux{originalMethod}(\m,\C_i) \mbox{ is defined and }
+\C_i \subtype \C_0
+
+%\\
+%\InTextDef{13ex}{
+%\tops(\m,\C_0)}{
+%}{
+%\{\ \get(\m, \C)\ |\ \C\in\super(\C_0), \get(\m, \C) \ne\emptyset,\mbox{ and }
+%}\\
+%\hspace{2.1in}\forall\C'\in\super(\C_0)\mbox{ and }\get(\m,\C')\ne\emptyset: \C'<:\C\Rightarrow\C'=\C\ \}\\
+%\InTextDef{13ex}{
+%\super(\C_0)}{
+%}{
+%\{\ \C'\ |\ \C'\in\Cs,\mbox{ or }\exists\C\in\Cs,\C'\in\super(\C)\ \}
+%}\\
+%\InTextWith{\metaVar{IT}(\C_0) =
+%\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}\\
+%\InTextDef{13ex}{
+%\get(\m,\C_0)}{
+%}{
+%\method
+%}\\
+%\InTextWith{\metaVar{IT}(\C_0) =
+%\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
+\end{array}$
+
+%\paragraph{\shadow{}} chooses the most specific version of a method,
+%that is the unique version available, or a conflicted version from a
+%set of possibilities. In this case all the conflicting defintions are returned.
+%We do not model overloading, so it is an error if multiple versions are available with different parameter types. Formally:
+%
+%\noindent
+%$\begin{array}{l}
+%\InTextDef{15ex}{
+%\shadow()}{\none}\\
+%\InTextDef{15ex}{\shadow(\method)}{\method}\\
+%\InTextDef{15ex}{\shadow(\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
+%\InTextDef{15ex}{\shadow(\methods)}{\methods}\\
+%\InTextWith{\mbox{ the former case not applicable} }
+%%\methods\mbox{ not of the form }\overline{\mh\QM;}
+%%\mbox{ and }
+%%\Aux{mostSpecific}(\methods)\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\}}\\
+%\end{array}$
+%
+\paragraph{override} models how a method in an interface can
+override implementations in its superinterfaces, even in the case of
+conflicts. Note how the special value $\none$ is used, and how (the 5th
+case) overriding can solve a conflict.
+% The notation in the last case below
+%means $\method'$ is not static, and is supertype of $\method$.
+
+\noindent$\begin{array}{l}
+\InTextDef{20ex}{\override(\none,\emptyset)}{\none}\\
+\InTextDef{20ex}{\override(\method,\emptyset)}{\method}\\
+\InTextDef{20ex}{\override(\none,\method)}{\method}\\
+\InTextDef{20ex}{\override(\none,\overline{\mh\QM;})}{\Aux{mostSpecific}(\overline{\mh\QM;})}\\
+\InTextDef{20ex}{\override(\method,\methods)}{\method}\\
+\InTextWith{
+\forall \method' \in \methods :
+%\method'\in\{\mh\QM;,\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}, \conflicted\ \mh\QM; \},
+\method\subtype\method'
+}
+\end{array}$
+
+\noindent The definition $\Aux{mostSpecific}$ returns the most
+specific method whose type is the subtype of all the
+others.  Since method subtyping is a partial ordering,
+$\Aux{mostSpecific}$ may not be defined, this in turn forces us to rely on the last clause of $\override$; otherwise
+the whole \mBody{} would not be defined for that specific $\m$.
+Rule \rn{t-intf} relies on this behavior.
+
+\noindent$\begin{array}{l}
+\InTextDef{18ex}{\Aux{mostSpecific}(\methods)}{\method}\\
+\InTextWith{\method \in \methods\ \mand\ \forall \method' \in \methods :  \method \subtype \method'}
+\end{array}$
+
+To illustrate the mechanism of $\mBody$, we present an example. We compute $\mBody(\QM{m},\QM{D})$:
+\begin{lstlisting}
+interface A { Object m(); }
+interface B extends A { default Object m() {return this.m();} }
+interface C extends A {}
+interface D extends B, C { String m(); }
+\end{lstlisting}
+\vspace{-2ex}
+\begin{itemize}
+\item First $\{\QM{A,B,C}\}$, the full set of supertypes of $\QM{D}$ is obtained.
+\item Then we compute 
+$\tops(\QM{m},\{\QM{A,B,C}\})= $
+\Q@default Object m() {...}@,
+ that is $\QM{B.m}$.
+That is, we do not consider either
+$\QM{C.m}$ (since $\QM{m}$ is not declared directly in $\QM{C}$, hence $\Aux{originalMethod}(\QM{m},\QM{C})$ is undefined) or $\QM{B.m}$ 
+(that is a subtype of $\QM{A}$, thus $\QM{B.m}$ hides $\QM{A.m}$).
+\item The final step computes $\override(\QM{D.m},\QM{B.m})=\QM{D.m}$,
+by the last case of $\override$ we get that $\QM{D.m}$ hides $\QM{B.m}$ successfully ($\QM{String}$ is a subtype of $\QM{Object}$). Finally we get $\mBody(\QM{m},\QM{D}) = \QM{D.m}$.
+\end{itemize}
+
+%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+%exmples to motivate our design
+%interface A{static String m(){return "A";}}
+%interface C extends A{
+%	default String dm(){
+%		  this.m();//wrong in java
+%		  A.m();
+%		  C.m();//wrong in java
+%		}
+%}
+%
+%
+%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+
+\begin{comment}
+%Now it is correct, but may be we do not need it?
+We abbreviate typing statements on
+sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
+shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
+\end{comment}
+
+\begin{comment}
+\subsubsection{Method Typing}
+
+\[
+\inferrule
+{ }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-Meth)}
+\]
+
+\[
+\inferrule
+{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0}
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} \\\\ \Gamma \vdash \textbf{this}:I }
+\quad \textsc{(T-MethBody)}
+\]
+
+\[
+\inferrule
+{IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-MethExt)}
+\]
+
+\[
+\inferrule
+{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\\\
+IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\\\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} }
+\quad \textsc{(T-MethBodyExt)}
+\]
+
+\[ \inferrule
+{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
+ \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq \error \\\\
+ \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
+ \textsf{mbody}(m,I) \neq \error }
+{I \text{ OK}}
+\quad \textsc{(T-Intf)}
+ \]
+
+Interface $I$ type checks well, if:
+\begin{itemize}
+\item All its super-interfaces $\overline{J}$ OK.
+\item All methods inside interface $I$ are OK.
+\item All methods that $I$ is inheriting from super-interfaces are OK.
+\end{itemize}
+\subsubsection{Subtyping}
+
+\[ \inferrule{}{T <: T} \]
+
+\[ \inferrule{S <: T \\ T <: U}{S <: U}\]
+
+\[ \inferrule{\emph{ann} \spc \textbf{interface} \spc C_0 \spc \textbf{extends} \spc C_1,...,C_k \{...\}}
+{C_0 <: C_1 \\ ... \\ C_0 <: C_k} \]
+\end{comment}
+%\subsubsection{Interface Table}

### 2016-02-24 OOPSLA16/sections/Overview.tex
OOPSLA format.
Figures.

--- 2015-12-09 paper/sections/Overview.tex
@@
 
-\begin{figure}
+\begin{figure*}
+\centering
 \saveSpaceFig
 \begin{tabular}{c|c}
-\includegraphics[height=4cm]{PegasusDetail.pdf}\hspace{20pt} &
+\includegraphics[height=4cm]{pdfs/PegasusDetail.pdf}\hspace{20pt} &
 \begin{minipage}{7cm}
@@
 \caption{The animal system (left: complete structure, right: code for simplified animal system).}\label{fig:pegasus}
-\end{figure}
+\end{figure*}
 
@@
 
-\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+\lstinputlisting[linerange=8-10]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
 
@@
 
-\begin{figure}
+\begin{figure*}
+\centering
 \saveSpaceFig
@@
 
-\end{figure}
+\end{figure*}
 

### 2016-03-02 UseMixinLombok/src/futurework/TestStateInitialization.java
cleanup

@@ -1,5 +1,7 @@
 package futurework;
 
+import static java.lang.System.out;
+
 //BEGIN_STATE_INIT
 interface Box { 
     default int val() { return 0; } //provided
@@ -25,5 +27,20 @@ interface X {
 }
 
 public class TestStateInitialization {
+    public static void main(String[] args) {
+        out.println(new A4(){}.val());
+    }
 }
 
+interface A1 { int val(); }
+interface A2 extends A1 { default int val() { return 3; } }
+interface A3 extends A1 { default int val() { return 4; } }
+interface A4 extends A2, A3 {}
+
+interface B1 { 
+    //@Default(5)
+    int val();
+}
+interface B2 extends B1 {}
+interface B3 extends B1 {}
+interface B4 extends B2, B3 {}

### 2016-03-03 OOPSLA16/sections/Appendix.tex
Move formalization to appendix.

--- 2015-12-08 paper/sections/Appendix.tex
@@
 \item All field initializations are type correct.
-By the definition of $\ofMethod(\C^\II)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
+By the definition of $\ofMethod(\C^\II)$ in Appendix~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
 \item All method bodies are well-typed.
@@
     \item Typing of the \QM{with-} method of an arbitrary field $m_i$.
-By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C^\II\QM{.of(}\es_i\QM{);\}}\]
+By Appendix~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C^\II\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C^\II\QM{.of(}\es_i\QM{);\}}\]
         $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C^\II.\QM{of}$ method.

### 2016-03-03 OOPSLA16/sections/Appendix_Formalization.tex
Move formalization to appendix.

--- 2016-02-23 OOPSLA16/sections/Formalization.tex
@@
-\section{Interaction of Interface Methods with Interface Composition}
-Before formalizing Classless Java and object interfaces, it is helpful
-to informally discuss the behavior of methods in Java 8
-interfaces. In particular it is useful to understand how Java 8
-interfaces differ from conventional trait models.
+\clearpage
 
-%We show several interesting cases
-%and summarize compilation result of method composition into 3 categories:
-%conflict error, accepted (both abstract) and conservative error. Meanwhile we
-%also show different composition handling mechanism among traits, javac and ECJ.
-
-\subsection{Methods in Java 8 Interfaces}
-In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
-
-\paragraph{Static methods} are handled in a very clean way: they are visible only in
-  the interface in which they are explicitly defined. This means the following code
-  is ill-typed.
-\begin{lstlisting}
-interface A0 { static int m(){return 1;} }
-interface B0 extends A0 {}
-... B0.m()//ill typed
-\end{lstlisting}
-This is different from the way static methods are handled in classes. Here
-static methods have simply no interaction with interface
-composition (\Q@extends@ or \Q@implements@).
-
-\paragraph{Abstract method} composition is accepted when there exists a most specific one.
-  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
-\begin{lstlisting}
-interface A1 { Object m(); }
-interface B1 { Integer m();}
-interface C1 extends A1,B1 {} //accepted
-\end{lstlisting}
-
-\paragraph{Default methods} conflict with any other default or abstract method. For
-  example the following code is rejected due to method conflicts.
-\begin{lstlisting}
-interface A2 { default int m() {return 1;}}
-interface B2 { int m(); }
-interface C2 { default int m() {return 2;}}
-interface D2 extends A2,B2 {} //rejected due to conflicting methods
-interface E2 extends A2,C2 {} //rejected due to conflicting methods
-\end{lstlisting}
-Note how this is different from what happens in most trait models, where \Q@D2@
-would be accepted, and the implementation in \Q@A2@ would be part of the
-behavior of \Q@D2@.
-
-\paragraph{Resolving conflicts:}
-A method in the current interface wins over any method in its
-super-interfaces, provided that the method
-is the most specific one. This method also overrides conflict due to
-inheritance. For example, the following code is accepted, but would be rejected
-(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
-\texttt{E3}.
-\begin{lstlisting}
-interface D3 extends A2,B2 { int m(); } //accepted
-interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
-\end{lstlisting}
-
-\subsection{Classifying Outcomes of Interface Composition}
-%When interfaces are composed and methods with the same name (and
-%signature) exist, there are 3 possible outcomes.
-%
-We now try to classify possible outcomes for composition of methods with the same name (and
-signature).
-We will use the following (correct) declarations:
-\begin{lstlisting}
-interface A1 { T m(); }
-interface A2 extends A1 { default T m(){ ... } }
-interface A3 extends A2 { T m(); }
-
-interface B1 { default T m(){ ... } }
-interface B2 extends B1 { T m(); }
-interface B3 extends B2 { default T m(){ ... } }
-\end{lstlisting}
-
-\noindent
-What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
-\Q@B@${}_j$?\\*
-% Nine representative cases are shown next:\\*
-\noindent
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
-\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
-\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
-\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
-\end{tabular}
-\\*
-%We try to classify the results in the table:
-\begin{itemize}
-\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
-\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
-  trait models.
-\item \textbf{conservative error} happens when only one method is implemented
-  (leaving another one abstract), which is different from what we would expect in
-  a trait model, but is coherent with the conservative idea that a method
-  defined in an interface should not silently satisfy a method in another one.
-\end{itemize}
-
-\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
-the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
-ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
-(leading Java 8 designer) we have confirmed that the expected behavior is
-rejection, hence this is a bug in ECJ. This bug was also reported by
-  others and is fixed in the ECJ developer branch, but not released as a stable
-  version yet.
-%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
-%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
-%named as \texttt{m} which are either abstract or default methods. If we define a
-%new interface \texttt{M} that extends two of these interfaces, then the method
-%overriding result is shown in the table. For example, row2 col2 means \texttt{M
-%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
-%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
-%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
-%following this way of interpretation.
-
-% The interesting case in row4 col2 where
-%Indeed, to be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
-%We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
-%We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
-%the experimental result is \texttt{B3.m}. Because if we think consistently, this
-
-%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
-
-%In our approach, we choose to not model this strange behaviour (a bug?).
-%Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
-%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java changes its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
-
-\section{Formal Semantics}\label{sec:formal}
-
-This section presents a formalization of Classless Java: a minimal
-FeatherweightJava-like calculus which models the essence of Java
-interfaces with default methods. This formalization is used in
-Section~\ref{sec:translation} to define the semantics of object interfaces.
-
-\begin{figure}[t]
+\begin{figure*}[t]
+\centering
 \saveSpaceFig
@@
 \label{Grammar}
-\end{figure}
+\end{figure*}
 
-\subsection{Syntax}
-
-Figure~\ref{Grammar} shows the syntax of Classless Java.
-The syntax formalizes a minimal
-subset of Java 8, focusing on interfaces, default methods and object
-creation literals.  There is no syntax for classes.
-To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all map to a single set of identifiers as in Java.  Expressions
-consist of conventional constructs such as variables ($\x$), method
-calls ($\e\QM.\m\QM(\es)$) and static method calls
-($\C\QM.\m\QM(\es)$). For simplicity the degenerate case of calling a
-static method over the $\this$ receiver is not considered.  A more
-interesting type of expressions is super calls
-($\C\QM{.super.}\m\QM(\es)$), whose semantics is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
-dispatching to the version of the method as visible in the interface
-$\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
-modeled. In the syntax of field updates $\x$ is expected to be a
-field name. After updating the field $\x$ using the value of $\e$, the
-expression $\e'$ is executed. To blend the statement based nature of
-Java and the expression based nature of our language, we consider a
-method body of the form \Q@return@ $\x\QM=\e\QM;\e'$ to represent
-$\x\QM=\e\QM;\QM{return}\ \e'$ in Java.  Finally, there is an object
-initialization expression from an interface $\C$, where (for
-simplicity) all the fields are initialized with a variable present in
-scope.
-To  be fully compatible with Java, the concrete syntax for an interface
-  declaration with empty supertype list  would also
-  omit the \Q@extends@ keyword.
-% The single
-%non-Java 8 piece of syntax is the \mixin annotation, which is the only
-%one interesting piece of syntax in this article.
-%??? how is @Mixin not java8? in the sense that is java5?
-  Following standard
-practice, we consider a global Interface Table (\metaVar{IT}) mapping
-from interface names $\C$ to interface declarations $\II$.
-
-The environment $\Gamma$ is a mapping from variables to types.  As
-usual, we allow a functional notation for $\Gamma$ to do variable
-lookup.  Moreover, to help us define auxiliary functions, a functional
-notation is also allowed for a set of methods $\methods$, using the
-method name $\m$ as a key.  That is, we define $\methods(\m)=\method$
-iff there is a unique $\method\in\methods$ whose name is $\m$.  For
-convenience, we define $\methods(\m)=\none$ otherwise; moreover
-$\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
-For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their names.
-
-\subsection{Typing}
-
-Typing statement $\Gamma \vdash \e\in\C$ reads ``in the environment
-$\Gamma$, expression $\e$ has type $\C$.''.
-Before discussing the typing rules we discuss some of the used notation.
-As a shortcut, we write
-$\Gamma \vdash \e\in\C<:\C'$ instead of $\Gamma \vdash \e\in\C$ and
-$\C<:\C'$.
-
-We omit the definition of
-the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.
-The auxiliary notation $\Gamma^\mh$ trivially
-extracts the environment from a method header, by collecting the all types
-and names of the method parameters.  The
-notation $\m^\mh$ and $\C^\mh$ denotes respectively, extracting the
-method name and the return type from a method header. $\mBody(\m,\C)$,
-defined in Section~\ref{sec:auxiliary},
-returns the full method declaration as seen by $\C$, that is the
-method $\m$ can be declared in $\C$ or inherited from another
-interface.
-$\textsf{mtype}(\m,\C)$ and $\textsf{mtypeS}(\m,\C)$ return the type
-signature from a method (using $\mBody(\m,\C)$ internally).
-$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
-$\textsf{mtypeS}(\m,\C)$ only for static ones. We use $\dom(\C)$ to
-denote the set of methods that are defined for type $\C$, that is:
-$\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
-
-\begin{figure}[t]
+\begin{figure*}[t]
+\centering
 \saveSpaceFig
@@
 \label{ET}
-\end{figure}
+\end{figure*}
+
+\section{Formal Semantics}\label{sec:formal}
+
+This section presents a formalization of Classless Java, \delete{a minimal
+FeatherweightJava-like calculus} which models the essence of Java
+interfaces with default methods. This formalization is used \delete{in
+Section~\ref{sec:translation}} to define the semantics of object interfaces.
+
+\subsection{Syntax}
+
+Figure~\ref{Grammar} shows the syntax of Classless Java.
+The syntax formalizes a minimal
+subset of Java 8, focusing on interfaces, default methods and object
+creation literals.  There is no syntax for classes.
+To help readability we use many metavariables to represent identifiers: $C,x,f$ and $m$; however they all map to a single set of identifiers as in Java.  Expressions
+consist of conventional constructs such as variables ($\x$), method
+calls ($\e\QM.\m\QM(\es)$) and static method calls
+($\C\QM.\m\QM(\es)$). For simplicity the degenerate case of calling a
+static method over the $\this$ receiver is not considered.  A more
+interesting type of expressions is super calls
+($\C\QM{.super.}\m\QM(\es)$), whose semantics is to call the (non-static) method $\m$ over the $\this$ receiver, but statically
+dispatching to the version of the method as visible in the interface
+$\C$. A simple form of field updates ($\x\QM=\e\QM;\e'$) is also
+modeled. In the syntax of field updates $\x$ is expected to be a
+field name. After updating the field $\x$ using the value of $\e$, the
+expression $\e'$ is executed. To blend the statement based nature of
+Java and the expression based nature of our language, we consider a
+method body of the form \Q@return@ $\x\QM=\e\QM;\e'$ to represent
+$\x\QM=\e\QM;\QM{return}\ \e'$ in Java.  Finally, there is an object
+initialization expression from an interface $\C$, where (for
+simplicity) all the fields are initialized with a variable present in
+scope.
+To  be fully compatible with Java, the concrete syntax for an interface
+  declaration with empty supertype list  would also
+  omit the \Q@extends@ keyword.
+% The single
+%non-Java 8 piece of syntax is the \mixin annotation, which is the only
+%one interesting piece of syntax in this article.
+%??? how is @Mixin not java8? in the sense that is java5?
+  Following standard
+practice, we consider a global Interface Table (\metaVar{IT}) mapping
+from interface names $\C$ to interface declarations $\II$.
+
+The environment $\Gamma$ is a mapping from variables to types.  As
+usual, we allow a functional notation for $\Gamma$ to do variable
+lookup.  Moreover, to help us define auxiliary functions, a functional
+notation is also allowed for a set of methods $\methods$, using the
+method name $\m$ as a key.  That is, we define $\methods(\m)=\method$
+iff there is a unique $\method\in\methods$ whose name is $\m$.  For
+convenience, we define $\methods(\m)=\none$ otherwise; moreover
+$\m\in\dom(\methods)\ \miff\ \methods(\m)=\method$.
+For simplicity, we do not model overloading, thus for an interface to be well formed its methods must be uniquely identified by their names.
+
+\subsection{Typing}
+
+Typing statement $\Gamma \vdash \e\in\C$ reads ``in the environment
+$\Gamma$, expression $\e$ has type $\C$.''.
+Before discussing the typing rules we discuss some of the used notation.
+As a shortcut, we write
+$\Gamma \vdash \e\in\C<:\C'$ instead of $\Gamma \vdash \e\in\C$ and
+$\C<:\C'$.
+
+We omit the definition of
+the usual traditional subtyping relation between interfaces, that is the transitive and reflexive closure of the declared \Q@extends@ relation.
+The auxiliary notation $\Gamma^\mh$ trivially
+extracts the environment from a method header, by collecting the all types
+and names of the method parameters.  The
+notation $\m^\mh$ and $\C^\mh$ denotes respectively, extracting the
+method name and the return type from a method header. $\mBody(\m,\C)$,
+defined in Appendix~\ref{sec:auxiliary},
+returns the full method declaration as seen by $\C$, that is the
+method $\m$ can be declared in $\C$ or inherited from another
+interface.
+$\textsf{mtype}(\m,\C)$ and $\textsf{mtypeS}(\m,\C)$ return the type
+signature from a method (using $\mBody(\m,\C)$ internally).
+$\textsf{mtype}(\m,\C)$ is defined only for non static methods, while
+$\textsf{mtypeS}(\m,\C)$ only for static ones. We use $\dom(\C)$ to
+denote the set of methods that are defined for type $\C$, that is:
+$\m\in\dom(\C)\ \miff \ \mBody(\m,\C)=\method$.
 
@@
 \noindent$\begin{array}{l}
-\InTextDef{10em}{\sigvalid(\mh_1\ldots\mh_n,I)
-}{
-\forall i\in 1..n\ \ \mh_i\QM;\subtype\mBody(\m^{\mh_i},\C)
-}
+\InTextDef{9em}{\sigvalid(\mh_1\ldots\mh_n,I)
+}{}\\
+\hspace{.3in}\forall i\in 1..n\ \ \mh_i\QM;\subtype\mBody(\m^{\mh_i},\C)
 \end{array}$
@@
 \noindent$\begin{array}{l}
-\InTextDef{18em}{
+\InTextDef{17em}{
 \T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR\QM; \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR\QM;
@@
 }\\
-\InTextDef{18em}{\method \subtype
+\InTextDef{15em}{\method \subtype
 \QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}
@@
 }\\
-\InTextDef{18em}{\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method
+\InTextDef{15em}{\QM{default}\ \mh\,\mbox{\Q@\{return \_;\}@}\subtype\method
 }{
@@
 \InTextDef{11em}{\alldefined(\mh_1\ldots\mh_n,I)}{
-\forall\m\mbox{ such that } \mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m}
+\forall\m\mbox{ such that } }\\
+\hspace{.3in}\mBody(\m,\C)=\mh\QM; \exists i\in 1..n\ \m^{\mh_i}=\m
 \end{array}$
@@
 
-\begin{comment}
-\subsubsection{Auxiliary function: \textsf{mtype}}
-- \textsf{mtype(m, C)} : the signature of method m in C.
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
-{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
-  m \notin M}
-{ \textsf{mtype(m,T)} = \emptyset } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\
-  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
-{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
-
-\[ \inferrule{
-  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
-  \overline{M} \} \\
-  m \notin M}
-{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
-\end{comment}
-
 Defining \mBody{} is not trivial, and requires quite a lot of attention to the
@@
 \InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n\ \oC\methods\ \cC \mbox{ and }
- \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \C_1\ldots\C_n}\\
+\hspace{1.2in}\oC\methods\ \cC \mbox{ and } \C\in\Cs \mbox{ if } \C_i\subtype\C, i\in 1..n
 
-}
 \end{array}$
@@
 \InTextWith{\metaVar{IT}(\C_0) =
-\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
+\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC,}\\
+\hspace{1.2in}\method\in\methods\mbox{ not static},\m=\m^\method
  \\
 
-\InTextDef{44ex}{
+\InTextDef{41ex}{
 \Aux{originalMethod}(\m,\C_0)\in\tops(\m,\C_0\dots\C_n)}{
@@
 \not\exists i\in1..n \mbox{ such that }
-\Aux{originalMethod}(\m,\C_i) \mbox{ is defined and }
-\C_i \subtype \C_0
+\Aux{originalMethod}(\m,\C_i) \mbox{ is defined }\\
+\hspace{1.4in}\mbox{and } \C_i \subtype \C_0
 
@@
 \end{lstlisting}
-\vspace{-2ex}
+
 \begin{itemize}
@@
 \end{itemize}
-
-%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
-%exmples to motivate our design
-%interface A{static String m(){return "A";}}
-%interface C extends A{
-%	default String dm(){
-%		  this.m();//wrong in java
-%		  A.m();
-%		  C.m();//wrong in java
-%		}
-%}
-%
-%
-%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
-
-\begin{comment}
-%Now it is correct, but may be we do not need it?
-We abbreviate typing statements on
-sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
-shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
-\end{comment}
-
-\begin{comment}
-\subsubsection{Method Typing}
-
-\[
-\inferrule
-{ }
-{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
-\quad \textsc{(T-Meth)}
-\]
-
-\[
-\inferrule
-{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0}
-{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
-    I} \\\\ \Gamma \vdash \textbf{this}:I }
-\quad \textsc{(T-MethBody)}
-\]
-
-\[
-\inferrule
-{IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
-\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
-T_0 <: U_0 }
-{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
-\quad \textsc{(T-MethExt)}
-\]
-
-\[
-\inferrule
-{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\\\
-IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\\\
-\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
-T_0 <: U_0 }
-{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
-    I} }
-\quad \textsc{(T-MethBodyExt)}
-\]
-
-\[ \inferrule
-{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
- \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
- \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq \error \\\\
- \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
- \textsf{mbody}(m,I) \neq \error }
-{I \text{ OK}}
-\quad \textsc{(T-Intf)}
- \]
-
-Interface $I$ type checks well, if:
-\begin{itemize}
-\item All its super-interfaces $\overline{J}$ OK.
-\item All methods inside interface $I$ are OK.
-\item All methods that $I$ is inheriting from super-interfaces are OK.
-\end{itemize}
-\subsubsection{Subtyping}
-
-\[ \inferrule{}{T <: T} \]
-
-\[ \inferrule{S <: T \\ T <: U}{S <: U}\]
-
-\[ \inferrule{\emph{ann} \spc \textbf{interface} \spc C_0 \spc \textbf{extends} \spc C_1,...,C_k \{...\}}
-{C_0 <: C_1 \\ ... \\ C_0 <: C_k} \]
-\end{comment}
-%\subsubsection{Interface Table}

### 2016-03-03 OOPSLA16/sections/Formalization.tex
Figures.
Move formalization to appendix.

--- 2016-01-26 OOPSLA16/sections/Formalization.tex
@@
-\section{Formal Semantics}\label{sec:formalization}
+\section{Interaction of Interface Methods with Interface Composition}
+Before formalizing Classless Java and object interfaces, it is helpful
+to informally discuss the behavior of methods in Java 8
+interfaces. In particular it is useful to understand how Java 8
+interfaces differ from conventional trait models.
 
-\begin{figure*}[t]
-\[\begin{array}{llll}
-T & ::= & I_1.\cdots.I_n                                                                                                         & \hspace{1in} \textrm{path} \\
-e & ::= & x \Or e.m(\overline{e}) \Or T.m(\overline{e}) \Or T.\mathbf{super}.m(\overline{e}) \Or x = e;e' \Or obj         & \hspace{1in} \textrm{expressions} \\
-obj & ::= & \mathbf{new}\ T()\ \{\overline{field}\ mh_1\{\mathbf{return}\ e_1;\}\cdots mh_n\{\mathbf{return}\ e_n;\}\}& \hspace{1in} \textrm{object creation} \\
-field & ::= & T\ f=x;                                                                                              & \hspace{1in}  \textrm{field declaration} \\
-\mathbb{I} & ::= & ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}& \hspace{1in} \textrm{interface declaration} \\
-meth & ::= & \mathbf{static}\ mh\{\mathbf{return}\ e;\} \Or \mathbf{default}\ mh\{\mathbf{return}\ e;\} \Or mh; & \hspace{1in} \textrm{method declaration} \\
-mh & ::= & T_0\ m(T_1\ x_1\cdots T_n\ x_n)  & \hspace{1in} \textrm{method header} \\
-ann & ::= & \mathbf{@Family} \Or \emptyset & \hspace{1in} \textrm{annotations} \\
-\Gamma & ::= & x_1:T_1\cdots x_n:T_n & \hspace{1in} \textrm{environment}
-\end{array}\]
-\caption{Grammar.}
+%We show several interesting cases
+%and summarize compilation result of method composition into 3 categories:
+%conflict error, accepted (both abstract) and conservative error. Meanwhile we
+%also show different composition handling mechanism among traits, javac and ECJ.
+
+\subsection{Methods in Java 8 Interfaces}
+In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
+
+\paragraph{Static methods} are handled in a very clean way: they are visible only in
+  the interface in which they are explicitly defined. This means the following code
+  is ill-typed.
+\begin{lstlisting}
+interface A0 { static int m(){return 1;} }
+interface B0 extends A0 {}
+... B0.m()//ill typed
+\end{lstlisting}
+This is different from the way static methods are handled in classes. Here
+static methods have simply no interaction with interface
+composition (\Q@extends@ or \Q@implements@).
+
+\paragraph{Abstract method} composition is accepted when there exists a most specific one.
+  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+
+\paragraph{Default methods} conflict with any other default or abstract method. For
+  example the following code is rejected due to method conflicts.
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is different from what happens in most trait models, where \Q@D2@
+would be accepted, and the implementation in \Q@A2@ would be part of the
+behavior of \Q@D2@.
+
+\paragraph{Resolving conflicts:}
+A method in the current interface wins over any method in its
+super-interfaces, provided that the method
+is the most specific one. This method also overrides conflict due to
+inheritance. For example, the following code is accepted, but would be rejected
+(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
+\texttt{E3}.
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
+\end{lstlisting}
+
+\subsection{Classifying Outcomes of Interface Composition}
+%When interfaces are composed and methods with the same name (and
+%signature) exist, there are 3 possible outcomes.
+%
+We now try to classify possible outcomes for composition of methods with the same name (and
+signature).
+We will use the following (correct) declarations:
+\begin{lstlisting}
+interface A1 { T m(); }
+interface A2 extends A1 { default T m(){ ... } }
+interface A3 extends A2 { T m(); }
+
+interface B1 { default T m(){ ... } }
+interface B2 extends B1 { T m(); }
+interface B3 extends B2 { default T m(){ ... } }
+\end{lstlisting}
+
+\noindent
+What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
+\Q@B@${}_j$?
+% Nine representative cases are shown next:\\*
+\begin{figure*}[htbp]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
+\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
+\end{tabular}
 \end{figure*}
+%We try to classify the results in the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
+\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
+  trait models.
+\item \textbf{conservative error} happens when only one method is implemented
+  (leaving another one abstract), which is different from what we would expect in
+  a trait model, but is coherent with the conservative idea that a method
+  defined in an interface should not silently satisfy a method in another one.
+\end{itemize}
 
-\begin{figure*}[t]
-\[\begin{array}{lll}
-Exact(T_1.I,\ T_2) & = & Exact(I,\ Exact(T_1,\ T_2)) \\
-Exact(I,\ I_1.\cdots.I_n) & = & Exact2(I,\ I_1.\cdots.I_i) \\
-& & \ \ \ \ \textrm{for the smallest }i\in[0,n],\forall j\in(i,n],Exact2(I,\ I_1.\cdots.I_j) = \emptyset \\
-Exact2(I,\ \emptyset) & = & CT(I) \\
-Exact2(I,\ T_0.I_0) & = & \mathbb{I} \\
-& & \ \ \ \ \textrm{if } \mathbb{I}\in\overline{\mathbb{I}} \textrm{ and } name(\mathbb{I})=I \\
-& & Exact(I,\ Exact(T,\ T_0)) \\
-& & \ \ \ \ \textrm{if } T\in\overline{T} \textrm{ and }\forall T'\in\overline{T}, Exact(I,\ Exact(T',\ T_0))\in\{\ Exact(I,\ Exact(T,\ T_0)),\ \emptyset\ \} \\
-& & \textrm{where } CT(T_0.I_0) = ann\ \mathbf{interface}\ I_0\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}
-\end{array}\]
-\caption{Typing rules.}
-\end{figure*}
+\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
+ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
+(leading Java 8 designer) we have confirmed that the expected behavior is
+rejection, hence this is a bug in ECJ. This bug was also reported by
+  others and is fixed in the ECJ developer branch, but not released as a stable
+  version yet.
+%Table~\ref{table:javabug} shows the method overriding bug in Java. In the
+%example, there are 6 interfaces \texttt{A1,A2,A3,B1,B2,B3} with methods all
+%named as \texttt{m} which are either abstract or default methods. If we define a
+%new interface \texttt{M} that extends two of these interfaces, then the method
+%overriding result is shown in the table. For example, row2 col2 means \texttt{M
+%  extends A1,B1}. The result is \texttt{ERROR} because the abstract method
+%\texttt{T m();} in \texttt{A1} conflicts with the method \texttt{default T m()
+%  {...}} in \texttt{B1}. Readers may also figure out other extending cases by
+%following this way of interpretation.
+
+% The interesting case in row4 col2 where
+%Indeed, to be coherent with the idea of \textbf{conservative error}, the case should not be accepted.
+%We do not see how this should behave differently from \textbf{B1},\textbf{A1}, and \textbf{B3},\textbf{A3}.
+%We fear that the only retro-compatible fix for this strange behaviour is to accept all the cases of \textbf{conservative error} in a future version of Java.
+%the experimental result is \texttt{B3.m}. Because if we think consistently, this
+
+%\lstinputlisting[linerange=4-9]{../UseMixinLombok/src/methodshadowing/test5/Test5.java} % APPLY:linerange=JAVABUG
+
+%In our approach, we choose to not model this strange behaviour (a bug?).
+%Our auxiliary function $\mBody(\m,\C)$ enforce the \textbf{conservative error} strategy.
+%The rest of our formalization is parametric with the definition of $\mBody(\m,\C)$, thus if Java changes its resolution strategy to a more permissive one, only minor adaptations in $\mBody(\m,\C)$ would be needed.
+
+\begin{comment}
+\subsubsection{Auxiliary function: \textsf{mtype}}
+- \textsf{mtype(m, C)} : the signature of method m in C.
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \emptyset } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C \text{ extends } C_1,...,C_k \{ \overline{M} \} \\
+  E \spc m(\overline{D} \spc \overline{x}) \{ \text{return } e; \} \in M}
+{ \textsf{mtype(m,T)} = \overline{D} \to E } \]
+
+\[ \inferrule{
+  IT(T) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
+  \overline{M} \} \\
+  m \notin M}
+{ \textsf{mtype(m,T)} = \bigcup \textsf{mtype}(m,\overline{D}) } \]
+\end{comment}
+
+%The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
+%exmples to motivate our design
+%interface A{static String m(){return "A";}}
+%interface C extends A{
+%	default String dm(){
+%		  this.m();//wrong in java
+%		  A.m();
+%		  C.m();//wrong in java
+%		}
+%}
+%
+%
+%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
+
+\begin{comment}
+%Now it is correct, but may be we do not need it?
+We abbreviate typing statements on
+sequences in a simple way, writing $\Gamma \vdash \overline{t}:\overline{C}$ as
+shorthand for $\Gamma \vdash t_1:C_1,..., \Gamma \vdash t_n:C_n$.
+\end{comment}
+
+\begin{comment}
+\subsubsection{Method Typing}
+
+\[
+\inferrule
+{ }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-Meth)}
+\]
+
+\[
+\inferrule
+{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0}
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} \\\\ \Gamma \vdash \textbf{this}:I }
+\quad \textsc{(T-MethBody)}
+\]
+
+\[
+\inferrule
+{IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
+\quad \textsc{(T-MethExt)}
+\]
+
+\[
+\inferrule
+{\textbf{this}:I, \overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\\\
+IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\\\
+\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
+T_0 <: U_0 }
+{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
+    I} }
+\quad \textsc{(T-MethBodyExt)}
+\]
+
+\[ \inferrule
+{\textbf{interface } I \textbf{ extends } \overline{J} \{ \overline{M} \} \\\\
+ \forall J_i \in \overline{J}, J_i \text{ OK} \\\\
+ \forall m \in \overline{M}, \textsf{mbody}(m,I) \neq \error \\\\
+ \forall J_i \in \overline{J}, \forall m \text{ inside } J_i,
+ \textsf{mbody}(m,I) \neq \error }
+{I \text{ OK}}
+\quad \textsc{(T-Intf)}
+ \]
+
+Interface $I$ type checks well, if:
+\begin{itemize}
+\item All its super-interfaces $\overline{J}$ OK.
+\item All methods inside interface $I$ are OK.
+\item All methods that $I$ is inheriting from super-interfaces are OK.
+\end{itemize}
+\subsubsection{Subtyping}
+
+\[ \inferrule{}{T <: T} \]
+
+\[ \inferrule{S <: T \\ T <: U}{S <: U}\]
+
+\[ \inferrule{\emph{ann} \spc \textbf{interface} \spc C_0 \spc \textbf{extends} \spc C_1,...,C_k \{...\}}
+{C_0 <: C_1 \\ ... \\ C_0 <: C_k} \]
+\end{comment}
+%\subsubsection{Interface Table}

### 2016-03-03 OOPSLA16/sections/Translation.tex
Move formalization to appendix.

--- 2015-12-09 paper/sections/Translation.tex
@@
 
-This section shows what the \mixin annotation generates and presents a
-formal definition for most of the generated methods. Since the formalized part
-of Classless Java does not consider
-% setters, so we do not formally define the generation of the setters.
-%We do not include
-casts or \Q@instanceof@, the \Q@with@ method is not included in the
-formal translation. For the same reason \Q@void@ returning setters are
-not included, since they are just a minor variation over the more
-interesting fluent setters, and they would require special handling
-just for the conventional \Q@void@ type.
-
-\subsection{Translation}
-For the purposes of the formalization, the translation is divided into
-two parts for more convenient discussion on formal properties later. To this aim we introduce the annotation
-$\weakAnn$. Its role is only in the translation process, hence is
-not part of the Classless Java language.  $\weakAnn$ generates the
-constructor method $\QM{of}$, while \mixin automatically refines the
-return types and calls $\weakAnn$.
-
-\noindent$\begin{array}{l}
+\begin{figure*}[t]\label{figure:translation}
+\centering
+$\begin{array}{l}
 \InTextDef{16em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
@@
 \end{array}$
+\caption{The translation functions of $\mixinAnn$ and $\weakAnn$.}
+\end{figure*}
 
-\noindent Note that it is necessary to explicitly check if the interface is valid
-for annotation:
+This section gives an overview of what $\mixinAnn$ generates in Classless Java. We provide a
+formalization of Classless Java in Appendix~\ref{sec:formal}, which models the essence of
+Java interfaces with default methods, including the syntax as well as the typing rules.
+Since the formalized part of Classless Java does not consider casts or \Q@instanceof@,
+the \Q@with@ method is not included in the
+formal translation. For the same reason \Q@void@ returning setters are
+not included, since they are just a minor variation over the more
+interesting fluent setters, and they would require special handling
+just for the conventional \Q@void@ type.
 
-\noindent$\begin{array}{ll}
-\InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; =
-            \mBody(\m, \C_0),\mbox{ one of the following cases is satisfied:}}\\
-\tab\tab
-\isField(\method), \isWith(\method, \C_0) \mbox{ or }
-\isSetter(\method,\C_0)\\
-\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
-\mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
+\subsection{Translation}
 
-%\isClone:&\isClone(\C\ \QM{clone}\oR\cR\QM;, \C_0)\tab \mif\ \C_0 <: \C \\
-%\isImplemented:&\isImplemented(\method) \tab\mbox{iff }\method\mbox{ not of form }\mh\QM;
-%\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-%&\isImplemented(\QM{static}\ \mh\mbox{\Q@\{return \_;\}@}) = \QM{true} \\
-\end{array}$
+For the purposes of the formalization, the translation is divided into
+two parts for more convenient discussion on formal properties later. To this aim we introduce the annotation
+$\weakAnn$. Its role is only in the translation process, hence is
+not part of the Classless Java language.  $\weakAnn$ generates the
+constructor method $\QM{of}$, while \mixin automatically refines the
+return types and calls $\weakAnn$.
 
-\noindent That is, we can categorize all abstract methods in a pattern that we
-know how to implement: it is either a field getter, a with method or a setter.
+Figure~\ref{figure:translation} presents the translation functions. In the first function, $\mixinAnn$
+injects refined methods to interface $\C_0$, while in the second function, $\weakAnn$ invokes 
+$\ofMethod(\C_0)$ and generates the $\QM{of}$
+method for $\C_0$, if such a method does not exist in its domain, and all the abstract methods are
+valid for the annotation. The formal definition is available in Appendix~\ref{subsec:translation}.
 
-Moreover, we check that the method \Q@of@ is not already defined by the user.
-In the formalization an existing definition of the \Q@of@ method is an error. However,
-in the prototype (which also needs to account for overloading), the check is more complex
-as it just checks that an \Q@of@ method with the same signature of the one being
-generated is not already present.
-
-We write $\QM{with#}\m$ to append $\m$ to
-\QM{with}, following the camelCase rule. The first letter of $\m$
-must be lower-case and is changed to upper-case upon appending. For
-example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
-are \QM{with} and all identifiers of the form $\QM{with#}\m$.
-
-\paragraph{The $\otherMethod$ function:}
-$\otherMethod(\C_0,\methods)$ is defined as follows:
-
-\noindent$\begin{array}{ll}
-\InTextDef{16em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{
- \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
-\hspace{2.65in}{
-\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
-\InTextDef{16em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
-\hspace{2.65in}{
-\mbox{ and } \QM_\m\notin\dom(\methods)}\\
-\end{array}$
-
-\noindent The methods generated in the interface are \Q@with-@ and setters. %\Q@clone@.
-%A complete formalization would also generate the \Q@with@.
-The methods are generated when they are unimplemented in $\C_0$, because
-the return types need to be refined.
-To determine whether the methods need to be generated,
-we check if such \Q@with-@ or setter methods %\Q@clone@
-are required by $\C_0$, but not declared directly in $\C_0$.
-
-\paragraph{The $\ofMethod$ function:}\label{subsec:ofmethod}
-The function $\ofMethod$ generates the method \QM{of}, as an object factory. To
-avoid boring digressions into well-known ways to find unique
-names, %for the sake of this formalization
-we assume that all methods with no parameters do not start with an underscore,
-and we prefix method names with underscores to obtain valid parameter names for
-\QM{of}.
-
-\noindent$\begin{array}{l}
-\InTextDef{5em}{\ofMethod(\C_0)}{
- \QM{static}\ \C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
-\QM{\{}
-\QM{return new}\ \C_0 \oR\cR\ \QM{\{} }\\
-\tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
-\tab\tab
-\C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots
-\C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
-\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
-\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
-%\tab\tab\cloneMethod(\C_0,\es)\\
-%\tab\tab\withMethod(\C_0)\\
-\tab\QM{\};\}} \\
-\InTextWith{\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();} = \fieldsFunc(\C_0) 
-\mbox{ and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n}
-\end{array}$
-
-\noindent Note that, the function $\fieldsFunc(\C_0)$ denotes all the fields in the current interface:
-
-\noindent$\begin{array}{ll}
-\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
-\isField(\method)\ \mand\
-%}\\\InTextWith{
-\method=\mBody(m^\method,\C_0)
-}
-\end{array}$
-
-\noindent For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
-  \begin{itemize}
-   \item $\m_i$ is the field name, and has type $\C_i$.
-   \item $\m_i$\QM{()} is the getter and just returns the current field value.
-   \item if a method \Q@with#@$\m_i()$ is required, then it is implemented by calling the \Q@of@ method using
-    the current value for all the fields except for $\m_i$. Such new value is
-    provided as a parameter. This corresponds to the expressions $\es_i$.
-\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modeling overloading.
-%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
-   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
-   %\item \marco{should we just formalize setters?}
-   \end{itemize}
-
-The auxiliary functions are defined below. Note that we do not need to check if some header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
-
-\noindent$\begin{array}{l}
-\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{
-\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return}\ \C_0\QM{.of(}\es\QM{);\}}} \\
-\InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ having the form }\mh\QM;}\\
-\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{\emptyset\mbox{ otherwise}}\\
-\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{
-\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
- \m\QM{= _val;return this;\}}} \\
-\InTextWith{
-\mBody(\QM_\m,\C_0) \mbox{ having the form }\mh\QM;}\\
-\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{\emptyset\mbox{ otherwise}}\\
-%\cloneMethod:&\cloneMethod(\C_0,\es)=
-%\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
-%&\mbox{iff }
-%\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
-%&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
-\end{array}$
-
-\subsection{Other Features}
-
-We do not formally model non-fluent setters and the \Q@with@ method.
-An informal explanation of how those methods are generated is given next:
-\begin{itemize}
-\item For methods inside the interface with the form \Q@void @$\m$\QM($\C\ \x$\QM{);}:
-  \begin{itemize}
-    \item Check if method $\C\ \m$\Q@();@ exists. If not, generate error (that
-      is, $\valid(\C_0)$ is false).
-    \item Generate the implemented setter method inside \Q@of@:\\*
-      \Q@public void @$\m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val;}@\\*
-      There is no need to refine the return type for non-fluent setters, thus we
-      do not need to generate the method header in the interface body itself.
-    \end{itemize}
-\item For methods with the form $\C'\ $\QM{with(}$\C\ \x$\QM{);}:
-  \begin{itemize}
-   \item $\C$ must be an interface type (no classes or primitive types).
-    \item As before, check that $\C'$ is a supertype of the current interface type $\C_0$.
-    \item Generate implemented \Q@with@ method inside \Q@of@:\\*
-           \Q@public @$\C_0\ $\Q@with(@$\C$\Q@ _val) { @\\*
-           \Q@  if(_val instanceof @$\C_0$\Q@){return (@$\C_0$\Q@)_val;}@\\*
-${}_{}$\Q@  return @$\C_0$\Q@.of(@$\e_1\ldots\e_n$\Q@);}@\\*
- with $\e_i=$\Q@_val.@$\m_i$\Q@()@ if $\C$ has a $\m_i$\Q@()@ method where
- $\m_1\ldots\m_n$ are fields of $\C_0$; otherwise $\e_i=\m_i$.
-    \item If needed, as for \Q@with-@ and setters, generate the method headers with refined return types in the interface.
- \end{itemize}
-
-%\item For methods with the form $\C'\ \m$\QM($\C\ \x$\QM{);}:
- % \begin{itemize}
-  %  \item As for before, check if exist method $\C\ \m$\Q@();@. Also, check that $\C'$ is a supertype of the current interface type $\C_0$.
-   % \item Generate implemented setter method inside \Q@of@:\\*
-    %       \Q@public @$\C_0\ \m$\Q@(@$\C$\Q@ _val) { @$\m$\Q@=_val; return this;}@
-   % \item If needed, as for \Q@with-@ and clone, generate the method header with refined return type in the interface.
- % \end{itemize}
-\end{itemize}
+The function $\ofMethod$ generates the static method $\QM{of}$, as an object factory. It detects all the
+field methods of $\C_0$, take them as arguments, and in the return statement is an anonymous class which generates getters,
+fluent setters and with-methods in certain cases. Some other features of $\mixinAnn$, including non-fluent setters and the $\QM{with}$
+method, however, are not formalized in Classless Java. Appendix~\ref{subsec:otherfeatures} gives an informal explanation of those methods.
 
 \subsection{Guarantees}
+
 To understand to what extent our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
@@
 \subsection{Results}
+
 To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
@@
 \mixin tuning guarantees only self and client coherence. The formal
-theorem proofs are available in the supplementary materials.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
+theorem proofs are available in the supplementary materials.
+
+%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
 

### 2016-03-04 OOPSLA16/sections/Translation.tex
more functions in main body

@@
 
+\begin{figure*}[t]\label{figure:valid}
+\centering
+\hspace{-.3in}
+\begin{minipage}[b]{0.5\textwidth}
+$\begin{array}{l}
+\InTextDef{16em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{}\\
+\hspace{.25in}\isWith(\mBody(\QM{with#}\m, \C_0), \C_0),\
+\QM{with#}\m\notin\dom(\methods)\\
+\InTextDef{16em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
+\otherMethod(\C_0,\methods)}{}\\
+\hspace{.25in}\isSetter(\mBody(\QM_\m, \C_0), \C_0),\ \QM_\m\notin\dom(\methods)\\
+\InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; =
+            \mBody(\m, \C_0),}\\
+            \hspace{.24in}\mbox{ one of the following cases is satisfied:}\\
+\tab\tab
+\isField(\method), \isWith(\method, \C_0) \mbox{ or }
+\isSetter(\method,\C_0)\\
+\end{array}$
+\end{minipage}
+\vline
+\hspace{.1in}
+\begin{minipage}[b]{0.43\textwidth}
+$\begin{array}{l}
+\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
+\mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{}\\
+\hspace{.3in}\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)\\
+\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{}\\
+\hspace{.3in}\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)\\
+\end{array}$
+\end{minipage}
+\caption{The $\otherMethod$ and $\valid$ functions (left) and auxiliary functions (right).}
+\end{figure*}
+
 This section gives an overview of what $\mixinAnn$ generates in Classless Java. We provide a
@@
 just for the conventional \Q@void@ type.
+
+\begin{figure*}[t]\label{figure:ofmethod}
+\centering
+\hspace{-.2in}
+\begin{minipage}[b]{0.53\textwidth}
+$\begin{array}{l}
+\InTextDef{5em}{\ofMethod(\C_0)}{
+ \QM{static}\ \C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
+\QM{\{}}\\
+\tab\QM{return new}\ \C_0 \oR\cR\ \QM{\{}\\
+\tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
+\tab\tab
+\C_1\ \m_1\oR\cR\ \QM{\{return }\ \m_1\QM{;\}}\ \ldots
+\C_n\ \m_n\oR\cR\ \QM{\{return }\ \m_n\QM{;\}}\\
+\tab\tab\withMethod(\C_1,\m_1,\C_0,\es_1)\ldots\withMethod(\C_n,\m_n,\C_0,\es_n)\\
+\tab\tab\setterMethod(\C_1,\m_1,\C_0)\ldots\setterMethod(\C_n,\m_n,\C_0)\\
+%\tab\tab\cloneMethod(\C_0,\es)\\
+%\tab\tab\withMethod(\C_0)\\
+\tab\QM{\};\}} \\
+\InTextWith{\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();} = \fieldsFunc(\C_0)}\\
+\hspace{.5in}\mbox{ and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n
+\end{array}$
+\end{minipage}
+\vline
+\hspace{.2in}
+\begin{minipage}[b]{0.4\textwidth}
+$\begin{array}{l}
+\InTextDef{16ex}{\method\in\fieldsFunc(\C_0)}{}\\
+\hspace{.3in}\isField(\method)\ \mand\
+\method=\mBody(m^\method,\C_0)\\
+\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{}\\
+\hspace{.2in}\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+\QM{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+\InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ having the form }\mh\QM;}\\
+\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{\emptyset\mbox{ otherwise}}\\
+\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{}\\
+\hspace{.2in}\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+ \m\QM{= _val;return this;\}} \\
+\InTextWith{
+\mBody(\QM_\m,\C_0) \mbox{ having the form }\mh\QM;}\\
+\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{\emptyset\mbox{ otherwise}}\\
+%\cloneMethod:&\cloneMethod(\C_0,\es)=
+%\C_0\ \QM{clone()\{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
+%&\mbox{iff }
+%\mBody(\QM{clone},\C_0) \mbox{ is of form }\mh\QM;\\
+%&\cloneMethod(\C_0,\es)=\emptyset\mbox{ otherwise}\\
+\end{array}$
+\end{minipage}
+\caption{The generated $\QM{of}$ method (left) and auxiliary functions (right).}
+\end{figure*}
 
@@
 method for $\C_0$, if such a method does not exist in its domain, and all the abstract methods are
-valid for the annotation. The formal definition is available in Appendix~\ref{subsec:translation}.
+valid for the annotation. Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod}\haoyuan{Refs are wrong!?} present more details on the auxiliary
+functions. The formal definition is available in Appendix~\ref{subsec:translation}.
 
@@
 field methods of $\C_0$, take them as arguments, and in the return statement is an anonymous class which generates getters,
-fluent setters and with-methods in certain cases. Some other features of $\mixinAnn$, including non-fluent setters and the $\QM{with}$
-method, however, are not formalized in Classless Java. Appendix~\ref{subsec:otherfeatures} gives an informal explanation of those methods.
+fluent setters and with-methods in certain cases.
+Some other features of $\mixinAnn$, including non-fluent setters and the $\QM{with}$ method, however, are not formalized in Classless Java.
+Appendix~\ref{subsec:otherfeatures} gives an informal explanation of those methods.
 

### 2016-03-04 OOPSLA16/sections/Introduction.tex
First introduction fix

@@
 
-The text of the paper begins here.
+Object oriented languages always strive to offer great code reuse,
+they aim to couple flexibility and rigour, expressive power and modular reasoning.
+Two main idea have emerged to this end: prototype based  and class based language.
+In prototype based language an objects inherits from other objects,
+and thus objects own both behaviour and state (and objects is all you have).
+In class based languages an object is instance of a specific class, and classes inherits from other classes.
+Here objects own the state, while classes contains behaviour and the structure of the state.
+
+We present here a third alternative: the concept of
+ \textbf{Interface based object oriented languages}, where 
+  objects implements interfaces, and they own implementation for the behaviour,
+that is structurally defined in their interface.
+
+  State is just a special kind of behaviour, and is accessed by getter and setter methods.
+
+  The literature of traits provide good examples on how easy and modular can be combining
+  multiple sources of behaviour, and the literature over multiple inheritance show us how
+  hard it is to modularly combine multiple sources of \textbf{and} state.
+  However, our object are the only responsible to define 
+  the ultimate behaviour or a method, and if such method is just a getter/setter.
+
+We could define our approach as a novel language, with new syntax and semantic, but this would have a
+steep learning curve; so for the sake of providing a more accessible explanation, we will encode our ideas
+over Java, relying on the new features of Java 8; namely
+interface \emph{static methods}, and \emph{default methods}, allowing interfaces to have
+method implementations. While the motivation for default methods was
+to allow interfaces to be extended over time while preserving backwards
+compatibility (\emph{interface evolution}),
+ default methods can also emulate \emph{traits}~\cite{scharli03traits}.
+\footnote{The
+original notion of traits by Scharli et al. prescribes, among other
+things, that: 1) a trait provides a set of methods that implement
+behavior; and 2) a trait does not specify any state variables, so the
+methods provided by traits do not access state variables
+directly. Java 8 interfaces follow similar principles too. Indeed, a
+detailed description of how to emulate trait-oriented programming in
+Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
+team designing default methods, was also fully aware of that secondary
+use of interfaces, but it was not their objective to model traits:
+``The key goal of adding default methods to Java was "interface
+evolution", not "poor man's traits"''~\cite{goetz13default}. As a result, 
+they were happy to support the secondary use of interfaces with
+default methods as long as it did not make the implementation and
+language more complex.}
+
+%Still, the design is quite conservative and appears to be quite limited
+% in its current form to model advanced forms of multiple inheritance.
+%Indeed, our own personal experience of combining default methods 
+%and multiple interface inheritance in Java to achieve multiple implementation 
+%inheritance is that many workarounds and boilerplate code are needed. 
+%In particular, we encountered difficulties because:
+
+%\begin{itemize}
+%
+%\item {\em Interfaces have no constructors.} As a result, classes are 
+%still required to create objects, leading to substantial boilerplate 
+%code for initialization.
+%
+%\item {\em Interfaces do not have state.} This creates a tension between 
+% using multiple inheritance and having state. Using setter and
+%  getter methods is a way out of this tension, but this workaround
+%  requires tedious boilerplate classes that later implement those
+%  methods.
+%
+%\item {\em Useful, general purpose methods require special care in
+%  the presence of subtyping.} Methods such as
+%  \emph{fluent} setters~\cite{fowler2005fluentinterface}, not only require access to the
+%  internal state of an object, but they also require their return types to be
+%  refined in subtypes.
+%
+%\end{itemize}
+
+%\noindent Clearly, a way around those difficulties would be to change
+%Java and just remove these limitations. Scala's own notion of
+%traits~\cite{scala-overview}, for example, allows state in traits. Of
+%course adding state (and other features) to interfaces would
+%complicate the language and require changes to the compiler, and this
+%would go beyond the goals of Java 8 development team.
+
+%This paper takes a different approach. Rather than trying to get
+%around the difficulties by changing the language in fundamental ways,
+%we show that, with a simple language feature, default methods and
+%interface inheritance are in fact very expressive.
+
+We propose a programming style, where we never use classes\footnote{More precisely, we never use the \Q@class@ keyword.}.
+We call this restricted version of Java \emph{class-less java}.
+Our proposed style enables powerful object-oriented idioms, similar to 
+multiple inheritance.
+
+Since Java was not designed to be used in this way, our style can get pretty verbose, especially about
+object instantiation.
+We will annotate with \Q|@Obj| certain interfaces and rely on
+the Annotation processing to generate boring and repetitive code
+allowing interface instantiation.
+We call such annotated interface \emph{object interfaces};
+they support \emph{abstract state operations}, providing a way
+to use multiple inheritance with state in Java. The abstract state
+operations include various common utility methods (such as getters and
+setters, or clone-like methods). In the presence of subtyping, such
+operations often require special care, as their types need to be
+refined. Object interfaces provide support for type-refinement and can
+automatically produce code that deals with type-refinement
+adequately. %With object interfaces, many Java programs can be built
+%without using a single class!
+
+%
+%Object interfaces do not require changes to the Java runtime or compiler, 
+%and they also do not introduce any new syntax. All three features of object interfaces are
+%achieved by reinterpreting existing Java syntax, and are translated
+%into regular Java code without loss of type-safety. Since no new
+%syntax is introduced, it would be incorrect to call object interfaces
+%a language extension or syntactic sugar. So we use the term
+%\emph{language tuning} instead. Language tuning sits in between a
+%lightweight language extension and a glorified library. Language
+%tuning can offer many features usually implemented by a real language
+%extension, but because it does not modify the language syntax
+%pre-existing tools can work transparently on the tuned language.  To
+%exploit the full benefits of language tuning,
+
+We formally define the behaviour of our \Q|Obj| annotation and we provide a
+prototype
+implementation using Java annotations to do AST
+rewriting, allowing existing Java tools (such as IDEs) to work
+out-of-the-box with our implementation. As a result, we could
+experiment object interfaces with several interesting Java programs,
+and conduct various case studies.
+
+%To formalize object interfaces, we propose Classless Java (CJ): a
+%FeatherweightJava-style~\cite{Igarashi01FJ} calculus, which captures the essence
+%of interfaces with default methods. The semantics of object interfaces
+%is given as a syntax-directed translation from CJ to itself. In the
+%resulting CJ code, all object interfaces are translated into regular CJ
+%(and Java) interfaces with default methods. The translation is proved
+%to be type-safe, ensuring that the translation does not
+%introduce type-errors in client code. 
+%CJ's usefulness goes beyond serving as a
+%calculus to formalize object interfaces. During the development
+%process of CJ, we encountered a bug in the implementation of default
+%methods for the Eclipse Compiler for Java (ECJ). For the program revealing the 
+%bug, ECJ behaves differently from both our formalization and Oracle's 
+%Java compiler.
+
+To evaluate the usefulness of object interfaces, we illustrate 3
+applications. The first application is a simple 
+solution to the Expression Problem~\cite{wadler98expression}, supporting independent 
+extensibility~\cite{zenger05independentlyextensible}, and without boilerplate code. The second
+application shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
+can be easily defined using object interfaces. The last
+application is a larger case study for a simple Maze game implemented with 
+multiple inheritance. For the last application we show that there is a
+significant reduction in the numbers of lines of code when compared 
+to an existing implementation~\cite{bono14} using plain Java 8. 
+Noteworthy, all applications are implemented 
+without defining a single class!
+
+In summary, the contributions of this paper are:
+\begin{itemize}
+
+\item {\bf Object Interfaces:} A simple feature that allows various
+  powerful multiple-inheritance programming idioms to be expressed
+  conveniently in Java. We provide examples, informal and formal description.
+
+\item{\bf Type preservation guarantees:}
+We discuss our formalization of a subset of Java8 type system and how we use this
+to characterize safety properties about our annotations.
+%\item {\bf Classless Java (CJ):} A simple formal calculus that models 
+%the essential features of Java 8 interfaces with default methods, and 
+%can be used to formally define the translation of object interfaces. 
+%We prove several properties of the translation\footnote{Proofs and prototype implementation are available in
+%  the supplementary materials.}.
+
+\item {\bf Implementation and Case Studies:} We have a prototype
+  implementation of object interfaces, using Java
+  annotations and AST rewriting. Moreover, the usefulness of object interfaces is
+  illustrated through various examples and case studies.
+
+%\item{\bf Language Tuning:} We identify the concept of language tuning 
+%and describe object interfaces as an example. We also discuss 
+%how other existing approaches, such as the annotations in project 
+%Lombok~\cite{lombok},  can be viewed as language tuning.
+
+\end{itemize}

### 2016-03-08 OOPSLA16/predef/java.tex

@@
+\lstdefinelanguage{JavaScala}{
+  morekeywords={public,int,interface,implements,default,
+    abstract,case,void,catch,class,def,static,%
+    do,else,extends,false,final,finally,%
+    for,if,implicit,import,match,mixin,%
+    new,null,object,override,package,%
+    private,protected,requires,return,sealed,%
+    super,this,throw,trait,true,try,%
+    type,var,while,yield},
+  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
+  sensitive=true,
+  morecomment=[l]{//},
+  morecomment=[n]{/*}{*/},
+  morestring=[b]",
+  morestring=[b]',
+  morestring=[b]"""
+}
+
 \lstset{ %
-language=Java,                % choose the language of the code
+language=JavaScala,                % choose the language of the code
 columns=flexible,
@@
 showtabs=false,                 % show tabs within strings adding particular underscores
+morekeywords={var},
 %  frame=single,                   % adds a frame around the code
@@
 keywordstyle=\ttfamily\bfseries,
-deletekeywords={label}
+aboveskip=3pt,
+belowskip=-1pt
 % commentstyle=\color{Gray},

### 2016-03-08 OOPSLA16/sections/Compare.tex

--- 2015-11-15 paper/sections/Compare.tex
@@
  \textbf{exclusion}.
-The author of ~\cite{bono14} agree that our revised version is cleaner, typesafe and more direct.
+The author of ~\cite{bono14} agree that our revised version is cleaner, typesafe and more direct.\bruno{is this something to 
+say in the paper?}
 
-\newcommand\shortItem{\vspace{-1ex}}
+%%\newcommand\shortItem{\vspace{-1ex}}
 \begin{itemize}
@@
 
-There are other limitations of our current approach, but they may be addressed in future work. See \marco{reference to future work when there}
+There are other limitations of our current approach, but they may be addressed
+in future work (see Section~\ref{sec:futurework}).

### 2016-03-08 OOPSLA16/sections/Formalization_DoNotDeleteMe.tex

\section{Formal Semantics}\label{sec:formalization}

\begin{figure*}[t]
\[\begin{array}{llll}
T & ::= & I_1.\cdots.I_n                                                                                                         & \hspace{1in} \textrm{path} \\
e & ::= & x \Or e.m(\overline{e}) \Or T.m(\overline{e}) \Or T.\mathbf{super}.m(\overline{e}) \Or x = e;e' \Or obj         & \hspace{1in} \textrm{expressions} \\
obj & ::= & \mathbf{new}\ T()\ \{\overline{field}\ mh_1\{\mathbf{return}\ e_1;\}\cdots mh_n\{\mathbf{return}\ e_n;\}\}& \hspace{1in} \textrm{object creation} \\
field & ::= & T\ f=x;                                                                                              & \hspace{1in}  \textrm{field declaration} \\
\mathbb{I} & ::= & ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}& \hspace{1in} \textrm{interface declaration} \\
meth & ::= & \mathbf{static}\ mh\{\mathbf{return}\ e;\} \Or \mathbf{default}\ mh\{\mathbf{return}\ e;\} \Or mh; & \hspace{1in} \textrm{method declaration} \\
mh & ::= & T_0\ m(T_1\ x_1\cdots T_n\ x_n)  & \hspace{1in} \textrm{method header} \\
ann & ::= & \mathbf{@Family} \Or \emptyset & \hspace{1in} \textrm{annotations} \\
\Gamma & ::= & x_1:T_1\cdots x_n:T_n & \hspace{1in} \textrm{environment}
\end{array}\]
\caption{Grammar.}
\end{figure*}

\begin{figure*}[t]
\[\begin{array}{lll}
Exact(T_1.I,\ T_2) & = & Exact2(I,\ Exact(T_1,\ T_2)) \\
Exact(I,\ I_1.\cdots.I_n) & = & Exact2(I,\ I_1.\cdots.I_i) \\
& & \ \ \ \ \textrm{for the smallest }i\in[0,n],\forall j\in(i,n],Exact2(I,\ I_1.\cdots.I_j) = \emptyset \\
Exact2(I,\ \emptyset) & = & I \\
& & \ \ \ \ \textrm{if } CT(I) = ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}\\
Exact2(I,\ T_0.I_0) & = & T_0.I_0.I \\
& & \ \ \ \ \textrm{if } \mathbb{I}\in\overline{\mathbb{I}} \textrm{ and } name(\mathbb{I})=I \\
& & Exact2(I,\ Exact(T,\ T_0)) \\
& & \ \ \ \ \textrm{if } T\in\overline{T} \textrm{ and }\forall T'\in\overline{T}, Exact(I,\ Exact(T',\ T_0))\in\{\ Exact(I,\ Exact(T,\ T_0)),\ \emptyset\ \} \\
& & \textrm{where } CT(T_0.I_0) = ann\ \mathbf{interface}\ I_0\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}
\end{array}\]
\caption{Typing rules.}
\end{figure*}

### 2016-03-08 OOPSLA16/sections/Overview.tex

@@
 
-This section illustrates how object interfaces, expressed using the \mixinAnn{} annotation, enable powerful
-idioms using multiple inheritance in Java.
-To propose a standard example for multiple inheritance, we show \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
+This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
+idioms similar to multiple inheritance and field type refinement.
+We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
 Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
-Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s.
+Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
 Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
 mythology) as a kind of \Q@Animal@ with the skills
-of both \Q@Horse@s and \Q@Bird@s. A simple class diagram illustrating
+of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
 the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
@@
 
-Before modeling the complete animal system, we  start with a
+Before modelling the complete animal system, we  start with a
 simplified version without locations. This version serves the purpose of illustrating how
@@
 methods. Moreover, because interfaces support multiple interface
-inheritance, the interface for \texttt{Pegasus} can inherit behavior
+inheritance, the interface for \texttt{Pegasus} can inherit behaviour
 from both \texttt{Horse} and \texttt{Bird}. Although Java interfaces

### 2016-03-08 OOPSLA16/sections/Formalization.tex
reviewd up to 2.6.

@@
 Before formalizing Classless Java and object interfaces, it is helpful
-to informally discuss the behavior of methods in Java 8
-interfaces. In particular it is useful to understand how Java 8
-interfaces differ from conventional trait models.
+to informally discuss the behaviour of Java8
+interfaces, in contrast with conventional trait models.
 

### 2016-03-09 OOPSLA16/sections/Introduction.tex

@@
 they aim to couple flexibility and rigour, expressive power and modular reasoning.
-Two main idea have emerged to this end: prototype based  and class based language.
-In prototype based language an objects inherits from other objects,
+Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
+In prototype based languages an objects inherits from other objects,
 and thus objects own both behaviour and state (and objects is all you have).
@@
 We present here a third alternative: the concept of
- \textbf{Interface based object oriented languages}, where 
+ \textbf{Interface based object oriented languages} (IB), where 
   objects implements interfaces, and they own implementation for the behaviour,
@@
 
-  The literature of traits provide good examples on how easy and modular can be combining
-  multiple sources of behaviour, and the literature over multiple inheritance show us how
-  hard it is to modularly combine multiple sources of \textbf{and} state.
+  The literature provide good examples on how easy and modular can be combining
+  multiple sources of pure behaviour (traits[]), and how
+  hard it is to modularly combine multiple sources of behaviour \textbf{and} state (multiple inheritance[]).
   However, our object are the only responsible to define 
@@
 
-We could define our approach as a novel language, with new syntax and semantic, but this would have a
-steep learning curve; so for the sake of providing a more accessible explanation, we will encode our ideas
+Our idea could be explained by defining a novel language, with new syntax and semantic.
+However, this would have a
+steep learning curve; for the sake of providing a more accessible explanation, we will encode
+\footnote{
+In the same sense, you can encode prototype based programming over class based with
+the strategy pattern, and class based programming with prototypes by creating ``class'' objects.
+} our ideas
 over Java, relying on the new features of Java 8; namely
@@
 We propose a programming style, where we never use classes\footnote{More precisely, we never use the \Q@class@ keyword.}.
-We call this restricted version of Java \emph{class-less java}.
+We call this restricted version of Java \emph{Class-less Java}.
 Our proposed style enables powerful object-oriented idioms, similar to 
@@
 
-Since Java was not designed to be used in this way, our style can get pretty verbose, especially about
+Since Java was not designed to be used in this way, our style can be verbose, especially about
 object instantiation.

### 2016-03-09 OOPSLA16/sections/Overview.tex

@@
 \begin{lstlisting}
-interface Pegasus extends Horse, Bird { // generated code not visible to users
+interface Pegasus extends Horse, Bird { 
+  // generated code not visible to users
     static Pegasus of() { return new Pegasus() {}; }
@@
 The animal system modeled so far is a simplified version of the
-system presented in the left-side of Figure~\ref{fig:pegasus}, which still does not appear to
-justify the \mixinAnn{} annotation.  Now moving on to modeling
-the complete animal system where \Q@Animal@s will include a \Q@location@
-representing their positions in space, we shall
-see that modeling stateful components using plain Java 8 quickly becomes
-cumbersome. The \mixinAnn{} annotation comes to rescue here and
- avoids significant amounts of boilerplate code.
+system presented in the left-side of Figure~\ref{fig:pegasus}.
+Here you still can not appreciate the advantage of IB.
+Lets model now the complete animal system where \Q@Animal@s will include a \Q@location@
+representing their positions in space. We will use 2D points to keep track of locations using coordinates.
 
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-2D points should keep track of their coordinates.  The usual
+Here we will illustrate how points are modelled with
+interfaces. In IB we do not talk about state directly; state is accessed
+using abstract (getter) methods. Note how in   The usual
 approach to model points in Java would be to use a class with fields for the
-coordinates. However here we will illustrate how points are modeled with
-interfaces. Since Java disallows fields inside interfaces, state is modeled
-using abstract (getter) methods:
+coordinates.
 
@@
 
-\noindent Unfortunately, creating a new point object is cumbersome, even
+\noindent Our encoding over Java is now inconvenient: creating a new point object is cumbersome, even
 with anonymous classes:
@@
 \begin{lstlisting}
-Point2D p = new Point2D() { public int x(){return 4;} public int y(){return 2;} }
+Point2D p = new Point2D() {
+  public int x(){return 4;} public int y(){return 2;}  }
 \end{lstlisting}
@@
 interface Point2D { int x(); int y();
-    static Point2D of(int x, int y) { return new Point2D() {
+  static Point2D of(int x, int y) {
+    return new Point2D() {
       public int x(){return x;}
-        public int y(){return y;}};}
-}
+      public int y(){return y;}
+    };  }  }
 \end{lstlisting}
@@
     int x(); int y();
-    default Point2D withX(int val){ ... } default Point2D withY(int val){ ... } }
+    default Point2D withX(int val){ ... }
+    default Point2D withY(int val){ ... } }
 \end{lstlisting}
@@
     int x(); int y();
-    default Point2D withX(int val){ ... } default Point2D withY(int val){ ... }
-    static Point2D of(int _x, int _y){ return new Point2D(){
+    default Point2D withX(int val){ ... }
+    default Point2D withY(int val){ ... }
+    static Point2D of(int _x, int _y){
+      return new Point2D(){
         int x=_x;    int y=_y;
-        public int x(){return x;}    public int y(){return y;} }; } }
+        public int x(){return x;}
+        public int y(){return y;} }; } }
 \end{lstlisting}
@@
 \noindent Only code for methods needing implementation is generated. Thus,
-programmers can easily customize the behavior for their special needs.
+programmers can easily customize the behaviour for their special needs.
+Also, since \mixin interfaces offer the \Q@of@ factory method, only interfaces where all the abstract methods
+can be synthesized can be \mixin. A non \mixin interface is like an abstract class in Java.
 %\end{comment}
@@
 \caption{Abstract state operations, for a field \texttt{x}, allowed by the \mixin
-  annotation. }
+  annotation. \marco{I disagree with this table, for example we miss the with} }
 
@@
 interface extends the \Q@Animal@ interface, the notion of points needs to
-be refined. Such kind of refinement usually poses a challenge
+be refined. Such kind of refinement is challenging
 in typical class-based approaches. Fortunately, with object interfaces,
@@
 \noindent From the type perspective, the key is the covariant method
-overriding of \texttt{location()}. However, from the semantic
+overriding of \texttt{location()}. However, from the semantics
 perspective the key is the implementation for the setter with the old
@@
 \lstinputlisting[linerange=74-74]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+
+\subsection{Abstract state operations}
+Defining an convenient list of abstract state operation is not an easy task.
+Of course we need getters.
+Currently we are offering both setters and Fluent setters. May be fluent setters are enough, and we can drop the
+concept of setters?
+We offer with withers and the with.
+What can be a nice syntax?
+\marco{my coments from now on}
+A language natively supporting Object interfaces may offer the following syntax:
+\begin{lstlisting}
+obj interface Foo{
+  int bar;//just getter
+  var int bar;//also setter(fluent)
+  with int bar;//also wither
+  with var int bar;//getter setter wither
+  with/var int bar=2;//default value, not asked in of
+  invariant{/*stuff that should always hold*/}
+  updater Bar, Beer//multi setter
+  wither Foo, Beer//multi with
+  factory of(...){call generate of}// generated of is now private
+}
+\end{lstlisting}
 

### 2016-03-10 OOPSLA16/sections/Introduction.tex
Some edits and comments

@@
 
+\bruno{Generaly speaking I like the flow of the new introduction!
+I would say that we do need to defend ourselfves against Scala traits, 
+but the differences are probably in three aspects:
+
+\begin{itemize}
+
+\item The abstract state operations and the novel technique to deal with 
+covariant refinement of setters; Scala traits do support state (variables and 
+immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
+mutable state does not support type-refinement.
+
+\item Automatic type-refinements (although Scala does also support \emph{manual} type-refinement 
+of immutable fields);
+
+\item We support constructors; whereas Scala traits do not. This is one key reason why classes 
+are still necessary in a lnaguage like Scala.
+
+\end{itemize}
+
+One final point about the introduction is that language tuning is no longer 
+mentioned. I think this is fine, but we have to discuss the strategy.
+
+}
 Object oriented languages always strive to offer great code reuse,

### 2016-03-10 OOPSLA16/sections/Overview.tex
Some edits and comments

@@
 We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
-Some kinds of animals are \Q@Horse@s and \Q@Bird@s.
+Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
 Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
@@
 the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
+\bruno{Should we provide a one sentence summary in the abstract of how much code 
+is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}
 
@@
 interfaces. In IB we do not talk about state directly; state is accessed
-using abstract (getter) methods. Note how in   The usual
-approach to model points in Java would be to use a class with fields for the
+using abstract (getter) methods. \bruno{unfinished sentence: Note how in}   The usual
+approach to model points in Java is to use a class with fields for the
 coordinates.
@@
 
+\bruno{At some point we should talk about the point that was discussed in the reviews}
 \paragraph{\Q@Point3D@ and properties updaters}
@@
 
-\subsection{Advanced Multiple Inheritance with Object Interfaces}
+\subsection{Advanced Multiple Inheritance}
 Finally, defining \texttt{Pegasus} is as simple as we did in the simplified
@@
 
-\subsection{Abstract state operations}
+\subsection{Abstract state operations}\bruno{Is this something to go into a discussion later on?}
 Defining an convenient list of abstract state operation is not an easy task.

### 2016-03-10 OOPSLA16/sections/Formalization.tex
Some edits and comments

@@
 \section{Interaction of Interface Methods with Interface Composition}
+\bruno{I don't think this section makes alot of sense in the context 
+of the new paper!}
 Before formalizing Classless Java and object interfaces, it is helpful

### 2016-03-10 OOPSLA16/sections/Formalization.tex

@@
 \bruno{I don't think this section makes alot of sense in the context 
-of the new paper!}
+of the new paper! I think perhaps some of this could be mentioned 
+in related work.
+
+In my opinion, instead of this section, we could have one section 
+that demonstrates our implementation and its advantages.
+
+\begin{itemize}
+
+\item Show a diagram of how the approach works?
+
+\item Syntax and type system reuse; (good for error messages, for example)
+\item On-the-fly AST rewriting (means no new files are created, and functionality
+is immediatelly available for IDE's for example); this is better than conventional 
+annotation processors.
+\item Eclipse integration; code completion.
+\item No integration issues in typical pre-processors; (the process is transparent; 
+no need for an external tool that generates code in a first step).
+\item Separate compilation (which relates to the properties?)
+
+\end{itemize}
+
+}
 Before formalizing Classless Java and object interfaces, it is helpful

### 2016-03-10 OOPSLA16/sections/Implementation.tex
OOPSLA format.

--- 2015-12-08 paper/sections/Implementation.tex
@@
 all the methods inside the interface as well as the generated ones.  The custom
-transformation is easy and convenient to use.  For example this means that the
+transformation is easy and convenient to use.  For example, this means that the
 IDE functionality for content assist and autocomplete will work for the newly generated
@@
   Java annotation processors do not support.
+\end{itemize}
+
+\noindent \textcolor{red}{Another big concern of using Lombok in our approach is that, the modularity, code quality and maintainability
+are highly improved in contrast with preprocessors. For the last decades, we have seen a lot of convenience in
+programs using preprocessors; it is easy to learn and use, lightweight in code size, and applicable in various
+languages including C, Java, and so on. Nevertheless, preprocessors are heavily criticized in literature in
+terms of code quality and maintainability being affected and subtle errors being introduced. In [ref], the main
+aspects of criticism are discussed including:}
+
+\begin{itemize}
+\item \textbf{Separation of concerns.} \textcolor{red}{Preprocessors can affect the modularity of implementation code. Whereas in our approach,
+each feature can be realized by a single annotation with its handler, and meanwhile a declared Java class or interface
+can be applied with multiple annotations, which ensures the modularity.}
+\item \textbf{Sensitivity to subtle errors.} \textcolor{red}{Preprocessors like cpp, which introduces the syntax of ``\#ifdef'', are likely to
+bring very simple syntax errors or type errors. In our Lombok-based approach, annotation processing is applied between
+parsing and type-checking, where internal ASTs are modified. Lombok uses an internal AST for Java, with respect to
+Java syntax, hence syntax errors are avoided. On the other hand, possible type errors introduced by Lombok handlers
+can be observed by Java type system from compilation, which brings convenience in practice.}
+\item \textbf{Obfuscated source code.} \textcolor{red}{Implementations with preprocessors obfuscate the source code, making code less readable
+and hard to maintain with further changes. In our case, since modularity is supported, the handler code can be debugged
+and maintained for different features respectively. On the other hand, annotations do not introduce much confusion in
+the source code, while the information of annotated interface from compilation can still be captured in the "Outline" window
+of Eclipse or other IDEs.}
+\item \textbf{Lack of reuse.} \textcolor{red}{Reusability is yet another concern in using preprocessors. In Lombok, implementations of features are
+encapsulated in various annotation handlers, in which case some behaviours are allowed to reuse the code by invoking methods
+in other handlers, where tedious replicated code is avoided.}
 \end{itemize}

### 2016-03-13 OOPSLA16/sections/RelatedWork.tex
Content reorganization

--- 2015-12-09 paper/sections/RelatedWork.tex
@@
 \section{Related Work}\label{sec:related}
-In this section we discuss related work and comparison to Classless Java.
+In this section we discuss related work and how it compares to Classless Java.
 
@@
 interfaces and reabstraction of default methods are also modeled.
+
+\section{Interaction of Interface Methods with Interface Composition}
+\bruno{I don't think this section makes alot of sense in the context 
+of the new paper! I think perhaps some of this could be mentioned 
+in related work.
+
+In my opinion, instead of this section, we could have one section 
+that demonstrates our implementation and its advantages.
+
+\begin{itemize}
+
+\item Show a diagram of how the approach works?
+
+\item Syntax and type system reuse; (good for error messages, for example)
+\item On-the-fly AST rewriting (means no new files are created, and functionality
+is immediatelly available for IDE's for example); this is better than conventional 
+annotation processors.
+\item Eclipse integration; code completion.
+\item No integration issues in typical pre-processors; (the process is transparent; 
+no need for an external tool that generates code in a first step).
+\item Separate compilation (which relates to the properties?)
+
+\end{itemize}
+
+}
+Before formalizing Classless Java and object interfaces, it is helpful
+to informally discuss the behaviour of Java8
+interfaces, in contrast with conventional trait models.
+
+%We show several interesting cases
+%and summarize compilation result of method composition into 3 categories:
+%conflict error, accepted (both abstract) and conservative error. Meanwhile we
+%also show different composition handling mechanism among traits, javac and ECJ.
+
+\subsection{Methods in Java 8 Interfaces}
+In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
+
+\paragraph{Static methods} are handled in a very clean way: they are visible only in
+  the interface in which they are explicitly defined. This means the following code
+  is ill-typed.
+\begin{lstlisting}
+interface A0 { static int m(){return 1;} }
+interface B0 extends A0 {}
+... B0.m()//ill typed
+\end{lstlisting}
+This is different from the way static methods are handled in classes. Here
+static methods have simply no interaction with interface
+composition (\Q@extends@ or \Q@implements@).
+
+\paragraph{Abstract method} composition is accepted when there exists a most specific one.
+  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+
+\paragraph{Default methods} conflict with any other default or abstract method. For
+  example the following code is rejected due to method conflicts.
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is different from what happens in most trait models, where \Q@D2@
+would be accepted, and the implementation in \Q@A2@ would be part of the
+behavior of \Q@D2@.
+
+\paragraph{Resolving conflicts:}
+A method in the current interface wins over any method in its
+super-interfaces, provided that the method
+is the most specific one. This method also overrides conflict due to
+inheritance. For example, the following code is accepted, but would be rejected
+(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
+\texttt{E3}.
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
+\end{lstlisting}
+
+\subsection{Classifying Outcomes of Interface Composition}
+%When interfaces are composed and methods with the same name (and
+%signature) exist, there are 3 possible outcomes.
+%
+We now try to classify possible outcomes for composition of methods with the same name (and
+signature).
+We will use the following (correct) declarations:
+\begin{lstlisting}
+interface A1 { T m(); }
+interface A2 extends A1 { default T m(){ ... } }
+interface A3 extends A2 { T m(); }
+
+interface B1 { default T m(){ ... } }
+interface B2 extends B1 { T m(); }
+interface B3 extends B2 { default T m(){ ... } }
+\end{lstlisting}
+
+\noindent
+What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
+\Q@B@${}_j$?
+% Nine representative cases are shown next:\\*
+\begin{figure*}[htbp]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
+\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
+\end{tabular}
+\end{figure*}
+%We try to classify the results in the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
+\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
+  trait models.
+\item \textbf{conservative error} happens when only one method is implemented
+  (leaving another one abstract), which is different from what we would expect in
+  a trait model, but is coherent with the conservative idea that a method
+  defined in an interface should not silently satisfy a method in another one.
+\end{itemize}
+
+\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
+ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
+(leading Java 8 designer) we have confirmed that the expected behavior is
+rejection, hence this is a bug in ECJ. This bug was also reported by
+  others and is fixed in the ECJ developer branch, but not released as a stable
+  version yet.

### 2016-03-13 OOPSLA16/sections/FutureWork.tex
Reviewd red H text, plus some fix to discussions

--- 2015-12-09 paper/sections/FutureWork.tex
@@
 %\subsection{Qualifiers in Methods} %{Private state
-\paragraph{Qualifiers in Methods} %{Private state
-The biggest limitation of our approach is the inability to model qualifiers
-for class methods (private, protected, synchronized, etc.). For example, the absence
-of support for private/protected methods in Java 8 interfaces forces all
-members of interfaces to be public, including static methods. Since we use
-abstract methods to encode state, our state is always all public. Still, because 
-the state can only be accessed by methods, it 
+\paragraph{Encapsulation} %{Private state
+Since we are using Java8 to encode IB, we are unable to directly support qualifies
+like private or protected. In Java 9, where private methods will be allowed in
+interfaces; abstract state operations are, however, abstract methods,
+so they can not be private.
+Still, because the state can only be accessed by methods, it 
 is impossible for the user to know if a certain method maps directly to a field
-or if it has a default implementation.  If the user wants a constructor that
-does not directly maps to the fields, (as for secondary constructors in Scala)
-he can simply define its own \Q@of@ method and delegate on the generated one:
+or if it has a default implementation.
+
+As shown by Ferruccio in [...] interfaces are enough to support privateness by a
+ variation of the facade pattern:
 \begin{lstlisting}
-@Obj interface Point {  int x(); int y();
-    static Point of(int val){return Point.of(val,val);} }
+@Obj interface InternalData implements ExposedData{/*all the methods*/}//package visibility
+public interface ExposedData{
+  /*just the exposed interface*/
+  static ExposedData of(...){return InternalData.of(...);}
+  }
 \end{lstlisting}
-However, the generated \Q@of@ method would also be present and public.  If a
-future version of Java was to support \emph{static private methods in
-  interfaces} we could extend our code generation to handle also encapsulation.
-Currently, it is possible to use a public nested class with private static
-methods inside, but this is ugly and cumbersome. One possibility is that the
-annotation processor takes methods with a \Q!@Private! annotation, and turns it into
-static private methods of a nested class. In this extension, also the \Q@of@
-method could be made private following the same pattern.
 
-% \yanlin{future work on state initialization removed for now. Because I think the
-%   other two future works are fine. But this one is still not clear and
-%   controversial. The proposed workaround forces users to make decision: if he
-%   wants state with initialization then he also has to define the setter. Or
-%   else, it is ambiguous: could be interpreted as a state initializer or just
-%   a normal default method.}
-\begin{comment}
-\subsection{State initialization}
-As discussed before, the user can trivially define its own \Q@of@ method, and
-initialize a portion of the state with default values.  However, the
-initialization code would not be reusable, and subinterfaces would have
-to repeat such initialization code.  If a field has no setters, a simple
-alternative is to just define the ``field'' as a default method as in
+This is a safe and explicit way to list all the public interfaces and to hide all the implementation
+details from outside of the package.\footnote{
+Note how \Q@ExposedData.of@ do not need to map directly to the fields, and can also perform 
+invariant checking.}
+
+Using packages to as a boundary we can make all object interfaces ``with privateness''
+invisible types for the rest of the program.
+This preserve encapsulation, but limits inheritance:
+package protected interfaces can not be seen,
+thus can not be implemented from outside the package.
+
+Using Java9 private interface methods
+\footnote{Or some ugly pattern available today like public static classes with private methods}
+one can define (and inherit) private behaviour; but private state inheritance in IB poses a logical challenge:
+How can the heir produce initialization code for a field whose very existence is secret?
+
+CB get around this problem in two ways: the constructor have to call a super-constructor, and
+some fields can be pre-initialized.
+The first idea would not provide any encapsulation benefit in IB,
+since generated \Q@of@ methods expose all the field information anyway.
+
+Pre-initialized fields could be supported by IB if we support a cache expression,
+that is \Q@cache(e)@ would evaluate the expression a single time for each object, and cache the result
+somewhere.\footnote{
+For example, in a the compiled bytecode for objects
+could have an extra field for each cached expression in its methods.
+Such cache expression would be quite useful for other reasons,
+including replacing involved initialization patterns.}
+
+In this way
+
 \begin{lstlisting}
-@Obj interface Box{ default int val(){return 0;} }
+@Obj interface MyData{
+  private String f();//assuming Java9
+  private MyData f(String val);
+  }
 \end{lstlisting}
-If setters are required, a possible extension of our code expansion could
-recognize a field if the getter is provided and the setter is required, and
-could generate the following code:
-\lstinputlisting[linerange=4-11]{../UseMixinLombok/src/test/TestStateInitialization.java}% APPLY:linerange=STATE_INIT
+Would be expanded into 
+\begin{lstlisting}
+@Obj interface MyData{
+  private Box<String> _f(){return cache(Box<String>.of("hello"));}
+  private String f(){return _f().val();}
+  private MyData f(String val){return _f().val(val);}
+  }
+\end{lstlisting}
 
-We are unsure about the value of this workaround because of its trickiness. In
-order to define a state with initialization, users have to define a method trusting that it will be overridden later with a behavioral change, but
-such change is visible only after the first time the setter is called. Moreover, this code would cache
-the result instead of re-computing it every time. This can be very tricky in a
-non-functional setting.
-%\bruno{Yanlin please polish text/break long sentences.}
-\end{comment}
+Where we could have a predefined \Q@Box@ type as\\*
+\Q|@Obj interface Box<T>{ T val(); Box<T> val(T val);}|
 
@@
 \end{lstlisting}
+
+\paragraph{A real IB language}
+In this article we presented the concept of IB as a programming pattern over Java.
+We do not wish to define here a complete IB language, just to show the reader IB as
+a promising way to design a new oo language.
+The first task is to define the set of abstract state operations.
+Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
+clean function programming style that would be otherwise too cumbersome.
+
+Currently we are offering both setters and Fluent setters.
+We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
+
+Manually defining abstract state methods can be verbose, we could offer a convenient syntax to 
+declare get/set/with.
+For example
+\begin{lstlisting}
+obj interface Foo{
+  String bar;//== String bar(); Foo withBar(String val);
+  var String beer;//also fluent setter
+  }
+\end{lstlisting}
+
+We may also want to offer a compact syntax for property updaters and generalized with methods;
+\begin{lstlisting}
+obj interface Foo{
+  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
+  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
+  }
+\end{lstlisting}
+Finally, certain methods may have a specific meaning, for example
+if a private method called \Q@init@ is present, it will be called after construction,
+and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
+called in some order.

### 2016-03-13 OOPSLA16/sections/Implementation.tex
Reviewd red H text, plus some fix to discussions

--- 2015-12-08 paper/sections/Implementation.tex
@@
   Java annotation processors do not support.
+\end{itemize}
+
+\noindent 
+Lombok offers several advantages over conventional code preprocessors/macros, 
+and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok based
+ast rewriting.
+modularity, code quality and maintainability
+\begin{itemize}
+\item \textbf{Modularity}
+While general preprocessing can act across module boundaries, Lombock act modularly on each class, and it makes
+sense to apply the transformation to one class/interface at a time, and only to annotated classes/interface.
+This allows library code to be reused without the need of being
+reprocessed and recompiled.
+
+\item \textbf{Syntax and type errors.} 
+Some preprocessors (like the C one) can produce syntactically invalid code.
+Lombok ensures only syntactically valid code is produced; however, type errors can appear.
+In our annotation, we proven that we do not introduce new errors in clients (see [] for details).
+
+\item \textbf{Obfuscated source code.}
+Preprocessors bring great power, that can easily be misused producing 
+code particularly hard to understand.
+Lombok requires to start from Java Syntax, and can re interpret it.
+This is already better, since the syntax is preserved, but the reinterpretation could be surprising for 
+badly designed Lombok based rewritings.
+However, the``Outline'' window of Eclipse or other IDEs (and thus their auto-completion facilities)
+will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.
+
+In our annotation, we only enhance code:
+we satisfy the behaviour of abstract methods, we add methods and we refine return types.
+We consider this to be quite easy to follow, since it is similar to what happens in normal inheritance.
+
+%\item \textbf{Lack of reuse.}  %not sure here... I think most preprocessors support decent reuse, even the C one
+%Reusability is yet another concern in using preprocessors.
+% In Lombok, implementations of features are
+%encapsulated in various annotation handlers,
+% in which case some behaviours are allowed to reuse the code by invoking methods
+%in other handlers, where tedious replicated code is avoided.
 \end{itemize}

### 2016-03-13 OOPSLA16/sections/Overview.tex
Reviewd red H text, plus some fix to discussions

@@
 
-\subsection{Abstract state operations}\bruno{Is this something to go into a discussion later on?}
-Defining an convenient list of abstract state operation is not an easy task.
-Of course we need getters.
-Currently we are offering both setters and Fluent setters. May be fluent setters are enough, and we can drop the
-concept of setters?
-We offer with withers and the with.
-What can be a nice syntax?
-\marco{my coments from now on}
-A language natively supporting Object interfaces may offer the following syntax:
-\begin{lstlisting}
-obj interface Foo{
-  int bar;//just getter
-  var int bar;//also setter(fluent)
-  with int bar;//also wither
-  with var int bar;//getter setter wither
-  with/var int bar=2;//default value, not asked in of
-  invariant{/*stuff that should always hold*/}
-  updater Bar, Beer//multi setter
-  wither Foo, Beer//multi with
-  factory of(...){call generate of}// generated of is now private
-}
-\end{lstlisting}
-
 \begin{comment}

### 2016-03-15 OOPSLA16/sections/Implementation.tex
Some more text. Table and flow chart.

@@
 \section{Implementation}
+
+\begin{figure*}[t]\label{fig:lombok}
+\centering
+\includegraphics[width=5in]{pdfs/lombok.png}
+\caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]}
+\end{figure*}
 
@@
 
-The annotation we created is \mixin. In Eclipse, with an interface annotated by
+The annotation we created is \mixin. Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
+flow of \mixin annotation processing in Lombok. In Eclipse, with an interface annotated by
 \mixin, the automatic annotation processing is performed transparently and the information of
@@
 Lombok offers several advantages over conventional code preprocessors/macros,
-and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok based
-ast rewriting.
-modularity, code quality and maintainability
+and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok-based
+AST rewriting.
 \begin{itemize}
-\item \textbf{Modularity}
-While general preprocessing can act across module boundaries, Lombock act modularly on each class, and it makes
-sense to apply the transformation to one class/interface at a time, and only to annotated classes/interface.
+\item \textbf{Lack of modularity.}
+While general preprocessing can act across module boundaries, Lombok acts modularly on each class, and it makes
+sense to apply the transformation to one class/interface at a time, and only to annotated classes/interfaces.
 This allows library code to be reused without the need of being
@@
 Lombok ensures only syntactically valid code is produced; however, type errors can appear.
-In our annotation, we proven that we do not introduce new errors in clients (see [] for details).
+In our annotation, we prove that we do not introduce new errors in clients (see Section~\ref{subsec:guarantees} and Section~\ref{subsec:results} for details).
 
 \item \textbf{Obfuscated source code.}
-Preprocessors bring great power, that can easily be misused producing 
-code particularly hard to understand.
-Lombok requires to start from Java Syntax, and can re interpret it.
+Preprocessors bring great power, which can easily be misused producing
+code particularly hard to understand. Thus code quality and maintainability are reduced.
+Lombok requires to start from Java syntax, and can reinterpret it.
 This is already better, since the syntax is preserved, but the reinterpretation could be surprising for
-badly designed Lombok based rewritings.
+badly designed Lombok-based rewritings.
 However, the``Outline'' window of Eclipse or other IDEs (and thus their auto-completion facilities)
 will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.
+
+\item \textbf{No reuse of the type system.}
+As we mentioned above, badly designed rewritings can arise from the great power of Lombok. A simple piece of source code
+\begin{lstlisting}
+interface M { int m(); }
+\end{lstlisting}
+can be reinterpreted as
+\begin{lstlisting}
+interface M { void m(String s); }
+\end{lstlisting}
+in which case the type of method \Q@m@ is changed. Our \mixin annotation does not introduce this kind of rewritings,
+and hence the type system is reused. Moreover, Lombok can also modify unbounded types, which is easy to understand,
+for instance, the following code
+\begin{lstlisting}
+interface M { T m(); } // T is unbounded
+\end{lstlisting}
+is transformed into
+\begin{lstlisting}
+interface M { int m(); } // No error message
+\end{lstlisting}
+in which case the user will see the unbounded type in source code, but without error message from the compilation, since
+Lombok has modified the return type of \Q@m@. However, our \mixin annotation can still keep such errors and warnings.
 

### 2016-03-15 OOPSLA16/sections/Overview.tex
Some more text. Table and flow chart.

@@
 state are available to users. Object interfaces provide support for four
-different types of abstract state operations, which are summarized in
+different types of abstract state operations \textcolor{red}{together with two types of
+general operations}, which are summarized in
 Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
@@
 \saveSpaceFig
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
- & \textbf{Example}                  & \textbf{Description } \\ \hline
-\textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \hline
-{\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
-object; updates field \texttt{x} to \texttt{val}.             \\ \hline
-\textbf{setters}        & \Q@void x(int val);@ & Sets the field
-\texttt{x} to a  new value \texttt{val}.        \\ \hline
-\textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
+& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
+\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
+& {\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
+object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
+& \textbf{setters}        & \Q@void x(int val);@ & Sets the field
+\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
+& \textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
 \texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
+\multirow{2}{*}{Other operations} & \textbf{factory methods} & \begin{tabular}{l}\Q@static Point2D of(int _x,@\\ \hspace{.8in}\Q@int _y) {...}@\end{tabular} & Generate factory method for \texttt{Point2D}.
+\\ \cline{2-4}
+& \textbf{general updaters}        & \Q@Point3D with(Point2D val);@ & Update all fields
+that match the object \texttt{val}.        \\ \hline
 \end{tabular}
 
-\caption{Abstract state operations, for a field \texttt{x}, allowed by the \mixin
+\caption{\textcolor{red}{(edited)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
   annotation. \marco{I disagree with this table, for example we miss the with} }

### 2016-03-15 OOPSLA16/sections/Implementation.tex
Reviewed the new H text, improved a little the examples

@@
 \includegraphics[width=5in]{pdfs/lombok.png}
-\caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]}
+\caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]
+\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
+but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
+May be this should be "obj" processing?
+}}
 \end{figure*}
@@
 
-\item \textbf{No reuse of the type system.}
-As we mentioned above, badly designed rewritings can arise from the great power of Lombok. A simple piece of source code
-\begin{lstlisting}
-interface M { int m(); }
-\end{lstlisting}
-can be reinterpreted as
-\begin{lstlisting}
-interface M { void m(String s); }
-\end{lstlisting}
-in which case the type of method \Q@m@ is changed. Our \mixin annotation does not introduce this kind of rewritings,
-and hence the type system is reused. Moreover, Lombok can also modify unbounded types, which is easy to understand,
-for instance, the following code
-\begin{lstlisting}
-interface M { T m(); } // T is unbounded
-\end{lstlisting}
-is transformed into
-\begin{lstlisting}
-interface M { int m(); } // No error message
-\end{lstlisting}
-in which case the user will see the unbounded type in source code, but without error message from the compilation, since
-Lombok has modified the return type of \Q@m@. However, our \mixin annotation can still keep such errors and warnings.
-
-In our annotation, we only enhance code:
+Our annotation reinterpret the syntax for the sole goal of enhance/complete code:
 we satisfy the behaviour of abstract methods, we add methods and we refine return types.
 We consider this to be quite easy to follow, since it is similar to what happens in normal inheritance.
+
+\marco{The section No reuse of the type system
+is controversial.
+We do need to repeat the type checking, plus we aim to make untypable stuff well typed
+(for example anyone using the of method would not be well typed before).}
+%\item \textbf{No reuse of the type system.}
+%As we mentioned above, badly designed rewritings can arise from the great power of Lombok. A simple piece of source code
+%\begin{lstlisting}
+%interface M { int m(); }
+%\end{lstlisting}
+%can be reinterpreted as
+%\begin{lstlisting}
+%interface M { void m(String s); }
+%\end{lstlisting}
+%in which case the type of method \Q@m@ is changed. Our \mixin annotation does not introduce this kind of rewritings,
+%and hence the type system is reused. Moreover, Lombok can also modify unbounded types, which is easy to understand,
+%for instance, the following code
+%\begin{lstlisting}
+%interface M { T m(); } // T is unbounded
+%\end{lstlisting}
+%is transformed into
+%\begin{lstlisting}
+%interface M { int m(); } // No error message
+%\end{lstlisting}
+%in which case the user will see the unbounded type in source code, but without error message from the compilation, since
+%Lombok has modified the return type of \Q@m@. However, our \mixin annotation can still keep such errors and warnings.
 

### 2016-03-17 OOPSLA16/sections/Formalization.tex

--- 2016-03-08 OOPSLA16/sections/Formalization.tex
@@
-\section{Interaction of Interface Methods with Interface Composition}
-Before formalizing Classless Java and object interfaces, it is helpful
-to informally discuss the behaviour of Java8
-interfaces, in contrast with conventional trait models.
-
-%We show several interesting cases
-%and summarize compilation result of method composition into 3 categories:
-%conflict error, accepted (both abstract) and conservative error. Meanwhile we
-%also show different composition handling mechanism among traits, javac and ECJ.
-
-\subsection{Methods in Java 8 Interfaces}
-In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
-
-\paragraph{Static methods} are handled in a very clean way: they are visible only in
-  the interface in which they are explicitly defined. This means the following code
-  is ill-typed.
-\begin{lstlisting}
-interface A0 { static int m(){return 1;} }
-interface B0 extends A0 {}
-... B0.m()//ill typed
-\end{lstlisting}
-This is different from the way static methods are handled in classes. Here
-static methods have simply no interaction with interface
-composition (\Q@extends@ or \Q@implements@).
-
-\paragraph{Abstract method} composition is accepted when there exists a most specific one.
-  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
-\begin{lstlisting}
-interface A1 { Object m(); }
-interface B1 { Integer m();}
-interface C1 extends A1,B1 {} //accepted
-\end{lstlisting}
-
-\paragraph{Default methods} conflict with any other default or abstract method. For
-  example the following code is rejected due to method conflicts.
-\begin{lstlisting}
-interface A2 { default int m() {return 1;}}
-interface B2 { int m(); }
-interface C2 { default int m() {return 2;}}
-interface D2 extends A2,B2 {} //rejected due to conflicting methods
-interface E2 extends A2,C2 {} //rejected due to conflicting methods
-\end{lstlisting}
-Note how this is different from what happens in most trait models, where \Q@D2@
-would be accepted, and the implementation in \Q@A2@ would be part of the
-behavior of \Q@D2@.
-
-\paragraph{Resolving conflicts:}
-A method in the current interface wins over any method in its
-super-interfaces, provided that the method
-is the most specific one. This method also overrides conflict due to
-inheritance. For example, the following code is accepted, but would be rejected
-(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
-\texttt{E3}.
-\begin{lstlisting}
-interface D3 extends A2,B2 { int m(); } //accepted
-interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
-\end{lstlisting}
-
-\subsection{Classifying Outcomes of Interface Composition}
-%When interfaces are composed and methods with the same name (and
-%signature) exist, there are 3 possible outcomes.
-%
-We now try to classify possible outcomes for composition of methods with the same name (and
-signature).
-We will use the following (correct) declarations:
-\begin{lstlisting}
-interface A1 { T m(); }
-interface A2 extends A1 { default T m(){ ... } }
-interface A3 extends A2 { T m(); }
-
-interface B1 { default T m(){ ... } }
-interface B2 extends B1 { T m(); }
-interface B3 extends B2 { default T m(){ ... } }
-\end{lstlisting}
-
-\noindent
-What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
-\Q@B@${}_j$?
-% Nine representative cases are shown next:\\*
-\begin{figure*}[htbp]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
-\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
-\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
-\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
-\end{tabular}
-\end{figure*}
-%We try to classify the results in the table:
-\begin{itemize}
-\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
-\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
-  trait models.
-\item \textbf{conservative error} happens when only one method is implemented
-  (leaving another one abstract), which is different from what we would expect in
-  a trait model, but is coherent with the conservative idea that a method
-  defined in an interface should not silently satisfy a method in another one.
-\end{itemize}
-
-\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
-the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
-ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
-(leading Java 8 designer) we have confirmed that the expected behavior is
-rejection, hence this is a bug in ECJ. This bug was also reported by
-  others and is fixed in the ECJ developer branch, but not released as a stable
-  version yet.
 %Table~\ref{table:javabug} shows the method overriding bug in Java. In the

### 2016-03-17 OOPSLA16/sections/Introduction.tex

--- 2016-03-09 OOPSLA16/sections/Introduction.tex
@@
 \section{Introduction}\label{sec:intro}
+
+\bruno{need to defend ourselfves against Scala traits}
+\marco{I tried to improve a little, also in this sense,
+but I still think is better to defend ourselves mostly in the related work section, some of your text under may end up there}
+
+\bruno{The abstract state operations and the novel technique to deal with 
+covariant refinement of setters; Scala traits do support state (variables and 
+immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
+mutable state does not support type-refinement.
+}
 
@@
 and thus objects own both behaviour and state (and objects is all you have).
+
 In class based languages an object is instance of a specific class, and classes inherits from other classes.
@@
 
-  State is just a special kind of behaviour, and is accessed by getter and setter methods.
-
   The literature provide good examples on how easy and modular can be combining
@@
   hard it is to modularly combine multiple sources of behaviour \textbf{and} state (multiple inheritance[]).
-  However, our object are the only responsible to define 
-  the ultimate behaviour or a method, and if such method is just a getter/setter.
+
+  In IB, state is just a special kind of behaviour implementation.
+  Object are the only responsible to define 
+  the ultimate behaviour of a method and, for example, if such method is just a setter.
+  Anything related to state is completely contained in the instances and do not leak in the interfaces.
 
@@
 language more complex.}
+\marcoT{We could use Scala Traits to encode IB, but the simpler Java8 model is a better starting point.}
+\footnote{
+\marco{may go in related work}
+Scala traits can declare \Q@def, val, var@ and \Q@lazy val@ 
+However, while \Q@def@s can be overridden by \Q@val@/\Q@var@s, 
+\Q@val@/\Q@var@s can not be overridden.\marco{we need to check this}
+A\Q@val@/\Q@var@ member is predestined to be instance state, and this
+is in sharp contrast with our model where the state is only a private detail of instances, and
+do not contaminate interface/trait/classes.}
+\footnote{Scala traits have no constructor, but 
+\Q@var@s and (\Q@lazy@) \Q@val@s can be ``pre-initialized''.}
 
@@
 Our proposed style enables powerful object-oriented idioms, similar to 
-multiple inheritance.
+(stateful) multiple inheritance and field type refinement 
+(allowing a kind of covariant setters refinement).
 
-Since Java was not designed to be used in this way, our style can be verbose, especially about
-object instantiation.
-We will annotate with \Q|@Obj| certain interfaces and rely on
-the Annotation processing to generate boring and repetitive code
-allowing interface instantiation.
-We call such annotated interface \emph{object interfaces};
-they support \emph{abstract state operations}, providing a way
-to use multiple inheritance with state in Java. The abstract state
+As for CB, in IB programmers do not define objects directly,but delegate the task to
+\emph{object interfaces}, whose role is similar to non abstract classes in CB.
+
+The idea is that all abstract operations in an Object interface will be
+interpreted as \emph{abstract state operations};  The abstract state
 operations include various common utility methods (such as getters and
@@
 automatically produce code that deals with type-refinement
-adequately. %With object interfaces, many Java programs can be built
+adequately, while Java (and Scala) require explicit type-refinement.
+
+  %With object interfaces, many Java programs can be built
 %without using a single class!
+
+The Java annotation processors allows us to stick to pure Java8: by
+annotating with \Q|@Obj| certain interfaces
+we can generate boring and repetitive code
+allowing interface instantiation and type refinement; such code should not be needed in the first place
+in a real IB language.
+%Since Java was not designed to be used in this way, our style can be verbose, especially about object instantiation.
 
@@
 
+We can extend our approach to support encapsulation and privateness;
+various possibilities are discussed in the last part of the paper.
+
 In summary, the contributions of this paper are:
@@
 
-\item {\bf Object Interfaces:} A simple feature that allows various
+\item {\bf IB and Object Interfaces:} a novel take on object orientation, allowing
   powerful multiple-inheritance programming idioms to be expressed

### 2016-03-17 OOPSLA16/sections/Overview.tex

@@
 is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}
-
+\marco{Not in the abstract. But I think is good to have that number and to put in in the example section.
+I will set up the sentence and someone can compute the number.}
 \subsection{Simple Multiple Inheritance with Default
@@
     int x(); int y();
-    default Point2D withX(int val){ ... }
-    default Point2D withY(int val){ ... } }
+    default Point2D withX(int val){ /*myCode*/ }
+    default Point2D withY(int val){ /*myCode*/ } }
 \end{lstlisting}
@@
     int x(); int y();
-    default Point2D withX(int val){ ... }
-    default Point2D withY(int val){ ... }
+    default Point2D withX(int val){ /*myCode*/ }
+    default Point2D withY(int val){ /*myCode*/ }
     static Point2D of(int _x, int _y){
@@
 \paragraph{Summary} Dealing with state (mutable or not) in object interfaces
-relies on a notion of abstract state, where only \emph{methods} that interact with
-state are available to users. Object interfaces provide support for four
+relies on a notion of abstract state:
+\marcoT{Method bodies can only refer to other methods;
+Objects are (conceptually) closures and the state is 
+composed by captured local variables.}
+Methods interacting with those are called \emph{abstract state operations}.
+
+Object interfaces provide support for four
 different types of abstract state operations \textcolor{red}{together with two types of
@@
 & \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
-\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
-& {\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
+\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
+& {\bf withers}        &   \Q@Point2D withX(int val)@                & Clones
 object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
-& \textbf{setters}        & \Q@void x(int val);@ & Sets the field
+& \textbf{setters}        & \Q@void x(int val)@ & Sets the field
 \texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
-& \textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
+& \textbf{fluent setters}        & \Q@Point2D x(int val)@ &Sets the field
 \texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
-\multirow{2}{*}{Other operations} & \textbf{factory methods} & \begin{tabular}{l}\Q@static Point2D of(int _x,@\\ \hspace{.8in}\Q@int _y) {...}@\end{tabular} & Generate factory method for \texttt{Point2D}.
-\\ \cline{2-4}
-& \textbf{general updaters}        & \Q@Point3D with(Point2D val);@ & Update all fields
-that match the object \texttt{val}.        \\ \hline
+\multirow{3}{*}{Other operations} &
+\textbf{factory methods} & 
+%\begin{tabular}{l}
+\Q@static Point2D of(int _x,int _y)@
+%\end{tabular}
+ & Generate factory method for \texttt{Point2D}.        \\
+\cline{2-4}
+& \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@ & Update all fields
+that match fields in \texttt{val}.        \\
+\cline{2-4}
+& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ & Set all fields
+that match fields in \texttt{val}.        \\ \hline
 \end{tabular}
 
-\caption{\textcolor{red}{(edited)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
-  annotation. \marco{I disagree with this table, for example we miss the with} }
+\caption{\textcolor{red}{(edited2)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
+  annotation. }
 
@@
 signature (\Q@location(Point2D)@). The key to the setter
-implementation is a new type of \Q@with@ method, called a property updater.
+implementation is a new type of \Q@with@ method, called
+ a (functional) property updater.
 
 \bruno{At some point we should talk about the point that was discussed in the reviews}
+\marco{you mean the reviewer question on the list.size?}
 \paragraph{\Q@Point3D@ and properties updaters}
@@
 
+ \marcoT{Symmetrically,
+ we could offer an imperative property updater
+ that calls the setters instead of the withers.
+ \Q@Point3D set(Point2D val)@.
+ }
+
+ \marcoT{
+ Property updaters never break class invariants, since they
+ internally call operations that was already deemed
+ safe by the programmer. For example a list object
+ would not offer a setter for its size field, thus
+ a property updater would not attempt to set it.}
+
 \begin{figure}
+\marco{I can not find the declaration int x();
+also we can change the order of the code to write the non generated stuff first.}
 \saveSpaceFig
@@
 Just to give a feeling of how much mechanical code \mixin is generating, we show the
-generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}. Writing such code by hand is error-prone. For
+generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}.
+\marcoT{Overall, for the whole 
+animals-with-locations, an @@@ lines code example,
+the generate/completed code is composed by
+@@@ lines.}
+
+   Writing such code by hand is error-prone. For
 example a distracted programmer may swap the arguments of calls to

### 2016-03-17 OOPSLA16/sections/Introduction.tex
Polished introduction

@@
 
-Object oriented languages always strive to offer great code reuse,
-they aim to couple flexibility and rigour, expressive power and modular reasoning.
-Two main idea have emerged to this end: prototype based (PB) and class based (CB) languages.
-In prototype based languages an objects inherits from other objects,
-and thus objects own both behaviour and state (and objects is all you have).
+Object oriented languages have always strived to offer great code reuse.
+They aim to couple flexibility and rigour, expressive power and
+modular reasoning.  Two main ideas have emerged to this end: prototype
+based (PB)~\cite{} and class based (CB) languages~\cite{}.  In prototype based
+languages objects inherit from other objects, and thus objects own
+both behaviour and state (and objects is all you have).
+In class based languages an object is instance of a specific class,
+and classes inherit from other classes.  Here objects own the state,
+while classes contain behaviour and the structure of the state.
 
-In class based languages an object is instance of a specific class, and classes inherits from other classes.
-Here objects own the state, while classes contains behaviour and the structure of the state.
+We present here a third alternative: the concept of \textbf{Interface
+  based object oriented languages} (IB), where objects implements
+interfaces, and they own the implementation for the behaviour, which
+is structurally defined in their interface.
 
-We present here a third alternative: the concept of
- \textbf{Interface based object oriented languages} (IB), where 
-  objects implements interfaces, and they own implementation for the behaviour,
-that is structurally defined in their interface.
+The literature provide good examples on how easy and modular can be
+combining multiple sources of pure behaviour (traits~\cite{}), and how hard
+it is to modularly combine multiple sources of behaviour \textbf{and}
+state (multiple inheritance~\cite{}).\bruno{needs strenghtning}
 
-  The literature provide good examples on how easy and modular can be combining
-  multiple sources of pure behaviour (traits[]), and how
-  hard it is to modularly combine multiple sources of behaviour \textbf{and} state (multiple inheritance[]).
+To retain the easiness and modularity of combining multiple sources of
+pure behaviour, in IB state is just a special kind of behaviour
+implementation. Objects are the only responsibles to define the
+ultimate behaviour of a method and, for example, if such method is
+just a setter. Anything related to state is completely contained in
+the instances and do not leak in the interfaces.\bruno{Not sure if
+  this is totally clear. Implicit here is the fact that objects are
+  never defined explicitly, I think. And the behaviour is introduced 
+implicitly.}
 
-  In IB, state is just a special kind of behaviour implementation.
-  Object are the only responsible to define 
-  the ultimate behaviour of a method and, for example, if such method is just a setter.
-  Anything related to state is completely contained in the instances and do not leak in the interfaces.
-
-Our idea could be explained by defining a novel language, with new syntax and semantic.
-However, this would have a
-steep learning curve; for the sake of providing a more accessible explanation, we will encode
-\footnote{
-In the same sense, you can encode prototype based programming over class based with
-the strategy pattern, and class based programming with prototypes by creating ``class'' objects.
-} our ideas
-over Java, relying on the new features of Java 8; namely
-interface \emph{static methods}, and \emph{default methods}, allowing interfaces to have
-method implementations. While the motivation for default methods was
-to allow interfaces to be extended over time while preserving backwards
-compatibility (\emph{interface evolution}),
+Our idea could be explained by defining a novel language, with new
+syntax and semantic.  However, this would have a steep learning curve;
+for the sake of providing a more accessible explanation, we will
+encode \footnote{ In the same sense, you can encode prototype based
+  programming over class based with the strategy pattern, and class
+  based programming with prototypes by creating ``class'' objects.  }
+our ideas over Java, relying on the new features of Java 8; namely
+interface \emph{static methods}, and \emph{default methods}, allowing
+interfaces to have method implementations. While the motivation for
+default methods was to allow interfaces to be extended over time while
+preserving backwards compatibility (\emph{interface evolution}),
 default methods can also emulate \emph{traits}~\cite{scharli03traits}.
-\footnote{The
-original notion of traits by Scharli et al. prescribes, among other
-things, that: 1) a trait provides a set of methods that implement
-behavior; and 2) a trait does not specify any state variables, so the
-methods provided by traits do not access state variables
-directly. Java 8 interfaces follow similar principles too. Indeed, a
-detailed description of how to emulate trait-oriented programming in
-Java 8 can be found in the work by Bono et al.~\cite{bono14}. The Java 8
-team designing default methods, was also fully aware of that secondary
-use of interfaces, but it was not their objective to model traits:
-``The key goal of adding default methods to Java was "interface
-evolution", not "poor man's traits"''~\cite{goetz13default}. As a result, 
-they were happy to support the secondary use of interfaces with
-default methods as long as it did not make the implementation and
-language more complex.}
-\marcoT{We could use Scala Traits to encode IB, but the simpler Java8 model is a better starting point.}
-\footnote{
-\marco{may go in related work}
-Scala traits can declare \Q@def, val, var@ and \Q@lazy val@ 
-However, while \Q@def@s can be overridden by \Q@val@/\Q@var@s, 
-\Q@val@/\Q@var@s can not be overridden.\marco{we need to check this}
-A\Q@val@/\Q@var@ member is predestined to be instance state, and this
-is in sharp contrast with our model where the state is only a private detail of instances, and
-do not contaminate interface/trait/classes.}
-\footnote{Scala traits have no constructor, but 
-\Q@var@s and (\Q@lazy@) \Q@val@s can be ``pre-initialized''.}
+\footnote{The original notion of traits by Scharli et al. prescribes,
+  among other things, that: 1) a trait provides a set of methods that
+  implement behavior; and 2) a trait does not specify any state
+  variables, so the methods provided by traits do not access state
+  variables directly. Java 8 interfaces follow similar principles
+  too. Indeed, a detailed description of how to emulate trait-oriented
+  programming in Java 8 can be found in the work by Bono et
+  al.~\cite{bono14}. The Java 8 team designing default methods, was
+  also fully aware of that secondary use of interfaces, but it was not
+  their objective to model traits: ``The key goal of adding default
+  methods to Java was "interface evolution", not "poor man's
+  traits"''~\cite{goetz13default}. As a result, they were happy to
+  support the secondary use of interfaces with default methods as long
+  as it did not make the implementation and language more complex.}
+\marcoT{We could use Scala Traits to encode IB, but the simpler Java8
+  model is a better starting point.}  \footnote{ \marco{may go in
+    related work} Scala traits can declare \Q@def, val, var@ and
+  \Q@lazy val@ However, while \Q@def@s can be overridden by
+  \Q@val@/\Q@var@s, \Q@val@/\Q@var@s can not be overridden.\marco{we
+    need to check this} A\Q@val@/\Q@var@ member is predestined to be
+  instance state, and this is in sharp contrast with our model where
+  the state is only a private detail of instances, and do not
+  contaminate interface/trait/classes.}  \footnote{Scala traits have
+  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
+  ``pre-initialized''.}
 
@@
 
-We propose a programming style, where we never use classes\footnote{More precisely, we never use the \Q@class@ keyword.}.
-We call this restricted version of Java \emph{Class-less Java}.
-Our proposed style enables powerful object-oriented idioms, similar to 
-(stateful) multiple inheritance and field type refinement 
-(allowing a kind of covariant setters refinement).
+In the context of Java, what we propose is essentially a programming
+style, where we never use classes\footnote{More precisely, we never
+  use the \Q@class@ keyword.}.  We call this restricted version of
+Java \emph{Class-less Java}.  Our proposed style enables powerful
+object-oriented idioms, similar to (stateful) multiple inheritance and
+field type refinement (allowing a kind of covariant setters
+refinement).
 
-As for CB, in IB programmers do not define objects directly,but delegate the task to
-\emph{object interfaces}, whose role is similar to non abstract classes in CB.
+As for CB, in IB programmers do not define objects directly, but
+delegate the task to \emph{object interfaces}, whose role is similar
+to non abstract classes in CB.
 
-The idea is that all abstract operations in an Object interface will be
-interpreted as \emph{abstract state operations};  The abstract state
-operations include various common utility methods (such as getters and
-setters, or clone-like methods). In the presence of subtyping, such
-operations often require special care, as their types need to be
-refined. Object interfaces provide support for type-refinement and can
-automatically produce code that deals with type-refinement
-adequately, while Java (and Scala) require explicit type-refinement.
+To model state, the idea is that all abstract operations in an Object
+interface will be interpreted as \emph{abstract state operations}; The
+abstract state operations include various common utility methods (such
+as getters and setters, or clone-like methods). In the presence of
+subtyping, such operations often require special care, as their types
+need to be refined. Object interfaces provide support for
+type-refinement and can automatically produce code that deals with
+type-refinement adequately, while Java (and Scala) require explicit
+type-refinement.
 
@@
 
-The Java annotation processors allows us to stick to pure Java8: by
-annotating with \Q|@Obj| certain interfaces
-we can generate boring and repetitive code
-allowing interface instantiation and type refinement; such code should not be needed in the first place
-in a real IB language.
+Using Java annotation processors, we produce an implementation of
+Classless Java, which allow us to stick to pure Java8. By annotating
+with \Q|@Obj| the interfaces that represent object interfaces, we can
+generate boring and repetitive code for interface instantiation and
+type refinement.  Such code should not be needed in the first place in
+a real IB language.
+
 %Since Java was not designed to be used in this way, our style can be verbose, especially about object instantiation.
@@
 We formally define the behaviour of our \Q|Obj| annotation and we provide a
-prototype
-implementation using Java annotations to do AST
+prototype implementation using Java annotations to do AST
 rewriting, allowing existing Java tools (such as IDEs) to work
-out-of-the-box with our implementation. As a result, we could
+out-of-the-box with our implementation. As a result, we
 experiment object interfaces with several interesting Java programs,
@@
 
-To evaluate the usefulness of object interfaces, we illustrate 3
+To evaluate the usefulness of object interfaces, we illustrate
+3\bruno{needs updates}
 applications. The first application is a simple 
@@
 
-\item{\bf Type preservation guarantees:}
+\item{\bf Type preservation guarantees:}\bruno{needs rephrasing}
 We discuss our formalization of a subset of Java8 type system and how we use this

### 2016-03-18 OOPSLA16/sections/Implementation.tex
section 3 now much better

@@
-\section{Implementation}
+\section{A practical bridge between IB and CB}
 
-\begin{figure*}[t]\label{fig:lombok}
+\begin{figure}[t]\label{fig:lombok}
 \centering
-\includegraphics[width=5in]{pdfs/lombok.png}
+\includegraphics[width=3in]{pdfs/lombok.png}
 \caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]
@@
 }}
-\end{figure*}
+\end{figure}
 
-Our implementation is based on an extension of Lombok. The Lombok
-project~\cite{lombok} is a Java library that aims at removing (or
+Even if convinced of the advantages of IB,
+the practical convenience
+of using existing libraries points us into building IB support
+inside of an opportune CB language.\footnote{
+Rewriting libraries and application in this language/model
+would result in more reusable libraries, but it would be a taunting task.}
+
+We believe Java is a good language for such integration, since it supports compilation agents[citeSomehting].
+
+That is, Java libraries can interact with the Java compilation process,
+acting as a man in the middle between the 
+generation of the abstract syntax tree and the Java Bytecode.
+\marco{is this correct?}
+This process is facilitated by frameworks like Lombock~\cite{lombok}:
+a Java library that aims at removing (or
 reducing) Java boilerplate code via
-annotations. There are a number of annotations provided by the
-original Lombok, including \Q:@Getter:, \Q:@Setter:,
-\Q:@ToString: for generating getters, setters and \QM{toString}
-methods, respectively.  Furthermore, Lombok provides a number of
-interfaces for users to create custom transformations, as extensions
-to the original framework.
-A transformation is based on a handler, which acts on the AST for the
-annotated node and returns a modified AST for analysis and
-generation afterwards. Such a handler can either be a Javac handler or
-an Eclipse handler.
+annotations.
+Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
+flow of \mixin annotation processing in Lombok.
+%There are a number of annotations provided by the
+%original Lombok, including \Q:@Getter:, \Q:@Setter:,
+%\Q:@ToString: for generating getters, setters and \QM{toString}
+%methods, respectively.  Furthermore, Lombok provides a number of
+%interfaces for users to create custom transformations, as extensions
+%to the original framework.
+%A transformation is based on a handler, which acts on the AST for the
+%annotated node and returns a modified AST for analysis and
+%generation afterwards. Such a handler can either be a Javac handler or
+%an Eclipse handler.
 
-The annotation we created is \mixin. Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
-flow of \mixin annotation processing in Lombok. In Eclipse, with an interface annotated by
-\mixin, the automatic annotation processing is performed transparently and the information of
-the interface from compilation is captured in the ``Outline'' window. This includes
-all the methods inside the interface as well as the generated ones.  The custom
-transformation is easy and convenient to use.  For example, this means that the
-IDE functionality for content assist and autocomplete will work for the newly generated
-methods. The biggest reasons to use Lombok rather than using a conventional Java
- annotation processor are:
-\begin{itemize}
+Using Lombok we created our \mixin annotation, that supports object
+interfaces as described in the former examples.
+
+Compilation Agents have advantages with respect to
+alternatives like extending the language, using a macro pre processors 
+or using the Java Annotation processors.\marco{are we sure that annotation processors are not compilation agents too?}
+
+Our disciplined \mixin annotation offer some more advantages with respect to arbitrary (Compilation agent driven) AST rewriting.
+
 %this seems like a commercial spot
@@
 %  compile time to modify bytecode.
-\item Lombok modifies the generation process of the class files, by directly modifying the AST.
+
+\paragraph{Auxiliary files/informations do not pollutes the program:}
+Compilation Agents alters the generation process of the class files,
+by directly modifying the AST.
 Neither the source code is modified nor new Java files are generated.
-\item Moreover, and probably more importantly, Lombok is capable of generating
+Moreover, and probably more importantly, Lombok supports generation of
   code \emph{inside} a class/interface, which conventional
   Java annotation processors do not support.
-\end{itemize}
 
-\noindent
-Lombok offers several advantages over conventional code preprocessors/macros,
-and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok-based
-AST rewriting.
-\begin{itemize}
-\item \textbf{Lack of modularity.}
-While general preprocessing can act across module boundaries, Lombok acts modularly on each class, and it makes
+\paragraph{Modularity:}
+While general preprocessing/macros can act across module boundaries,
+Compilation agents act modularly on each class/compilation unit, and it makes
 sense to apply the transformation to one class/interface at a time, and only to annotated classes/interfaces.
 This allows library code to be reused without the need of being
-reprocessed and recompiled.
+reprocessed and recompiled, making our approach 100\% compatible with existing Java libraries, than can be used 
+and extended normally. Of course, Java libraries can also receive and use instances of object interfaces as normal objects.
 
-\item \textbf{Syntax and type errors.}
-Some preprocessors (like the C one) can produce syntactically invalid code.
-Lombok ensures only syntactically valid code is produced; however, type errors can appear.
-In our annotation, we prove that we do not introduce new errors in clients (see Section~\ref{subsec:guarantees} and Section~\ref{subsec:results} for details).
+\paragraph{Tool support:}
+Features integrate/supported directly in the language are often also supported by (most) tools.
+For example in Eclipse, the processing is performed transparently and the information of
+the interface from compilation is captured in the ``Outline'' window. This includes
+all the methods inside the interface as well as the generated ones.  
+This also means that the IDE functionality for content assist and autocomplete
+will work for the newly generated methods.
 
-\item \textbf{Obfuscated source code.}
+\paragraph{Clarity against Obfuscation:}
 Preprocessors bring great power, which can easily be misused producing
 code particularly hard to understand. Thus code quality and maintainability are reduced.
-Lombok requires to start from Java syntax, and can reinterpret it.
-This is already better, since the syntax is preserved, but the reinterpretation could be surprising for
-badly designed Lombok-based rewritings.
-However, the``Outline'' window of Eclipse or other IDEs (and thus their auto-completion facilities)
+Compilation agents start from Java syntax, but they can reinterpret it.
+Preserving the syntax avoid syntactic conflicts, and allows may tools to transparently work.
+
+The reinterpretation could still be surprising for badly designed rewritings.
+However, ``Outline'' windows/ auto-completion facilities 
 will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.
@@
 
-\marco{The section No reuse of the type system
-is controversial.
-We do need to repeat the type checking, plus we aim to make untypable stuff well typed
-(for example anyone using the of method would not be well typed before).}
+We believe that refactoring operations like rename and move
+should work transparently in conjunction with our annotaiton, since they rely on 
+the overall type structure of the class, that we do not arbitrarily modify but just complete.
+
+%\marco{The section No reuse of the type system
+%is controversial.
+%We do need to repeat the type checking, plus we aim to make untypable stuff well typed
+%(for example anyone using the of method would not be well typed before).}
 %\item \textbf{No reuse of the type system.}
@@
 %in other handlers, where tedious replicated code is avoided.
-\end{itemize}
 
-\paragraph{Limitations}
+\paragraph{Syntax and Type errors:}
+Some preprocessors (like the C one) can produce syntactically invalid code.
+Lombok ensures only syntactically valid code is produced; however, type errors can appear.
+
+Our annotation is guaranteed to not introduce errors in generate code and client code.
+See the following for details, where we identify tree form of types guarantees.
+
+\paragraph{Self coherence}: the generated code itself is well-typed; that is,
+  type errors are not present in code the user has not written.
+In our case it means that either \mixin{} produces (in a controlled way) an
+understandable error, or the interface can be successfully annotated and the generated code is well-typed.
+We guarantee \textit{Self coherence}.
+
+\paragraph{Client coherence}: all the client code (for example method calls)
+  that is well-typed before code generation is also well-typed after the generation.
+The annotation just adds more behaviour without removing any functionality.
+We guarantee \textit{Client coherence}.
+
+\paragraph{Heir coherence}: interfaces (and in general classes) inheriting the instrumented code are well-typed if they were well-typed without the instrumentation.
+This would forbid adding any (default or abstract) method to the annotated interfaces, including type refinement.
+\mixin  does not guarantee \textit{Heir coherence}.
+Indeed consider the following:
+
+\begin{lstlisting}
+interface A { int x(); A withX(int x); }
+@Obj interface B extends A {}
+interface C extends B { A withX(int x); }
+\end{lstlisting}
+
+\noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
+This would break \Q@C@. \\*
+Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
+would be correct before translation, but would be ill-typed after the translation.
+
+Our automatic type refinement
+is a useful and convenient feature, but not transparent to the heirs of the annotated interface.
+They need to be aware of the annotation semantics and provide the right type while refining methods.
+Please, note how Java libraries always break heir coherence during evolution
+ (is just enough to add any method to any non final class), and still claim backward compatibility
+ (false in theory but statistically true in practice).
+We think refining return types breaks heir coherence ``less" than normal library evolution.
+
+\paragraph{Type preservation:}
+We preferred to introduce self,client and hair coherence instead of referring to conventional 
+type preservation,
+to better model how our approach behave in a object oriented software ecosystem with inheritance,
+where only some units may be translated/expanded.
+Note inheritance crucial influence in hair coherence, and how client coherence allows 
+for only some units to be translated/expanded.
+
+This means that even a library designer can decide to start using IB and still 
+retain backward compatibility with its clients.
+
+\paragraph{Limitations}\marco{ to reorganize with updated limitations}
 Our prototype implementation using Lombok has certain limitations:
@@
 
-\paragraph{Comparison with other Lombok annotations}
+\paragraph{Comparison with other Lombok annotations}\marco{if we keep it, will go in related work}
 The Lombok project provides a set of predefined annotations, including constructor
@@
 
-\paragraph{Lombok does language tuning}
-We consider Lombok to be the most developed example of language
-tuning.  While the authors of Lombok do not introduce a specific term
-for what they are doing, their slogan \emph{``Spice up your java''}
-seems to be in line with the philosophy of language tuning. Some
-other examples of language tuning in Lombok include the \Q@val@ type,
-similar to \Q@auto@ in C\# or C++04.  Another library doing language
-tuning is CoFoJa~\cite{cofoja}, where annotations are used to insert
-pre-post conditions in generated bytecode.
+%\paragraph{Lombok does language tuning}
+%We consider Lombok to be the most developed example of language
+%tuning.  While the authors of Lombok do not introduce a specific term
+%for what they are doing, their slogan \emph{``Spice up your java''}
+%seems to be in line with the philosophy of language tuning. Some
+%other examples of language tuning in Lombok include the \Q@val@ type,
+%similar to \Q@auto@ in C\# or C++04.  Another library doing language
+%tuning is CoFoJa~\cite{cofoja}, where annotations are used to insert
+%pre-post conditions in generated bytecode.

### 2016-03-20 OOPSLA16/sections/Implementation.tex
local changes

@@
 a Java library that aims at removing (or
+>>>>>>> 769243b65e4aa45abf264bc79b016c61525e225f
 reducing) Java boilerplate code via

### 2016-03-20 OOPSLA16/sections/Introduction.tex
Polishing

@@
 We present here a third alternative: the concept of \textbf{Interface
-  based object oriented languages} (IB), where objects implements
+  based object oriented languages} (IB), where objects implement
 interfaces, and they own the implementation for the behaviour, which
-is structurally defined in their interface.
-
-The literature provide good examples on how easy and modular can be
-combining multiple sources of pure behaviour (traits~\cite{}), and how hard
-it is to modularly combine multiple sources of behaviour \textbf{and}
-state (multiple inheritance~\cite{}).\bruno{needs strenghtning}
+is structurally defined in their interface. In IB programmers do not define objects directly, but
+delegate the task to \emph{object interfaces}, whose role is similar
+to non abstract classes in CB languages. Key motivations for IB
+languages are improved \emph{modularity} and \emph{multiple inheritance}~\bruno{I'm using multiple
+  inheritance in a more abstract way than you Marco.}  The literature
+provides good examples on how easy and modular it is to combine
+multiple sources of pure behaviour, using mechanisms such as
+traits~\cite{scharli03traits}. In Java multiple \emph{interface} inheritance has been
+supported since inception, and in Java8 default methods~\cite{} bring some of
+the advantages of traits into Java. In contrast, the literature~\cite{} is
+also rich on how hard it is to modularly combine multiple sources of
+behaviour \textbf{and} state with multiple \emph{implementation}
+inheritance of classes.
 
@@
 
-Our idea could be explained by defining a novel language, with new
-syntax and semantic.  However, this would have a steep learning curve;
-for the sake of providing a more accessible explanation, we will
-encode \footnote{ In the same sense, you can encode prototype based
-  programming over class based with the strategy pattern, and class
-  based programming with prototypes by creating ``class'' objects.  }
-our ideas over Java, relying on the new features of Java 8; namely
-interface \emph{static methods}, and \emph{default methods}, allowing
-interfaces to have method implementations. While the motivation for
-default methods was to allow interfaces to be extended over time while
-preserving backwards compatibility (\emph{interface evolution}),
-default methods can also emulate \emph{traits}~\cite{scharli03traits}.
-\footnote{The original notion of traits by Scharli et al. prescribes,
-  among other things, that: 1) a trait provides a set of methods that
-  implement behavior; and 2) a trait does not specify any state
-  variables, so the methods provided by traits do not access state
-  variables directly. Java 8 interfaces follow similar principles
-  too. Indeed, a detailed description of how to emulate trait-oriented
-  programming in Java 8 can be found in the work by Bono et
-  al.~\cite{bono14}. The Java 8 team designing default methods, was
-  also fully aware of that secondary use of interfaces, but it was not
-  their objective to model traits: ``The key goal of adding default
-  methods to Java was "interface evolution", not "poor man's
-  traits"''~\cite{goetz13default}. As a result, they were happy to
-  support the secondary use of interfaces with default methods as long
-  as it did not make the implementation and language more complex.}
-\marcoT{We could use Scala Traits to encode IB, but the simpler Java8
-  model is a better starting point.}  \footnote{ \marco{may go in
-    related work} Scala traits can declare \Q@def, val, var@ and
-  \Q@lazy val@ However, while \Q@def@s can be overridden by
-  \Q@val@/\Q@var@s, \Q@val@/\Q@var@s can not be overridden.\marco{we
-    need to check this} A\Q@val@/\Q@var@ member is predestined to be
-  instance state, and this is in sharp contrast with our model where
-  the state is only a private detail of instances, and do not
-  contaminate interface/trait/classes.}  \footnote{Scala traits have
-  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
-  ``pre-initialized''.}
+%\footnote{ In the same sense,
+%  you can encode prototype based programming over class based with the
+%  strategy pattern, and class based programming with prototypes by
+%  creating ``class'' objects.}
+
+IB could be explained by defining a novel language, with new syntax
+and semantics. However, this would have a steep learning curve.  We
+take a different approach instead. For the sake of providing a more
+accessible explanation, we will embed our ideas directly into Java. 
+Our IB embedding relies on the
+new features of Java 8 for interfaces: interface \emph{static methods}; and
+\emph{default methods}, which allow interfaces to have method
+implementations.
 
@@
   use the \Q@class@ keyword.}.  We call this restricted version of
-Java \emph{Class-less Java}.  Our proposed style enables powerful
+Java \emph{Classless Java}.  
+Our proposed style enables powerful
 object-oriented idioms, similar to (stateful) multiple inheritance and
 field type refinement (allowing a kind of covariant setters
-refinement).
+refinement). \bruno{needs strenghtening}
 
-As for CB, in IB programmers do not define objects directly, but
-delegate the task to \emph{object interfaces}, whose role is similar
-to non abstract classes in CB.
-
-To model state, the idea is that all abstract operations in an Object
-interface will be interpreted as \emph{abstract state operations}; The
+A key challenge lies in how to model state, which is fundamental to
+have stateful objects. In IB and Classless Java, the idea is that all abstract operations in an Object
+interface will be interpreted as \emph{abstract state operations}. The
 abstract state operations include various common utility methods (such
@@
 type-refinement adequately, while Java (and Scala) require explicit
-type-refinement.
+type-refinement.\bruno{strenghtening needed!}
 
@@
 type refinement. Such code should not be needed in the first place in
-a real IB language.
+a real IB language, and the annotation processor allows us to
+transparently hide it from Java programmers. 
+The implementation works by performing \emph{on-the-fly} AST
+rewriting, allowing existing Java tools (such as IDEs) to work
+out-of-the-box with our implementation. Moreover, the 
+implementation blends Java's conventional CB style and IB smoothly. 
+As a result, we experiment object interfaces with several interesting Java programs,
+and conduct various case studies. 
+Finally, we also formally define the behaviour of our \Q|Obj| implementation.
 
@@
 %exploit the full benefits of language tuning,
-
-We formally define the behaviour of our \Q|Obj| annotation and we provide a
-prototype implementation using Java annotations to do AST
-rewriting, allowing existing Java tools (such as IDEs) to work
-out-of-the-box with our implementation. As a result, we
-experiment object interfaces with several interesting Java programs,
-and conduct various case studies.
 

### 2016-03-20 OOPSLA16/sections/RelatedWork.tex
Polishing

@@
 
+\bruno{next comes some text that may be handy for related work}
+
+While the motivation for default methods was to allow
+interfaces to be extended over time while preserving backwards
+compatibility (\emph{interface evolution}), default methods can also
+emulate \emph{traits}~\cite{scharli03traits}.  \footnote{The original
+  notion of traits by Scharli et al. prescribes, among other things,
+  that: 1) a trait provides a set of methods that implement behavior;
+  and 2) a trait does not specify any state variables, so the methods
+  provided by traits do not access state variables directly. Java 8
+  interfaces follow similar principles too. Indeed, a detailed
+  description of how to emulate trait-oriented programming in Java 8
+  can be found in the work by Bono et al.~\cite{bono14}. The Java 8
+  team designing default methods, was also fully aware of that
+  secondary use of interfaces, but it was not their objective to model
+  traits: ``The key goal of adding default methods to Java was
+  "interface evolution", not "poor man's
+  traits"''~\cite{goetz13default}. As a result, they were happy to
+  support the secondary use of interfaces with default methods as long
+  as it did not make the implementation and language more complex.}
+\marcoT{We could use Scala Traits to encode IB, but the simpler Java8
+  model is a better starting point.}  \footnote{ \marco{may go in
+    related work} Scala traits can declare \Q@def, val, var@ and
+  \Q@lazy val@ However, while \Q@def@s can be overridden by
+  \Q@val@/\Q@var@s, \Q@val@/\Q@var@s can not be overridden.\marco{we
+    need to check this} A\Q@val@/\Q@var@ member is predestined to be
+  instance state, and this is in sharp contrast with our model where
+  the state is only a private detail of instances, and do not
+  contaminate interface/trait/classes.}  \footnote{Scala traits have
+  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
+  ``pre-initialized''.}
+
 \section{Interaction of Interface Methods with Interface Composition}

### 2016-03-21 OOPSLA16/sections/Introduction.tex
More polishing for intro

@@
 \section{Introduction}\label{sec:intro}
-
-\bruno{need to defend ourselfves against Scala traits}
-\marco{I tried to improve a little, also in this sense,
-but I still think is better to defend ourselves mostly in the related work section, some of your text under may end up there}
-
-\bruno{The abstract state operations and the novel technique to deal with 
-covariant refinement of setters; Scala traits do support state (variables and 
-immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
-mutable state does not support type-refinement.
-}
 
@@
 We present here a third alternative: the concept of \textbf{Interface
-  based object oriented languages} (IB), where objects implement
-interfaces, and they own the implementation for the behaviour, which
-is structurally defined in their interface. In IB programmers do not define objects directly, but
+  based} (IB) object oriented languages, where objects implement
+interfaces directly. In IB interfaces own the implementation for the behaviour, which
+is structurally defined in their interface. Programmers do not define objects directly, but
 delegate the task to \emph{object interfaces}, whose role is similar
@@
 multiple sources of pure behaviour, using mechanisms such as
-traits~\cite{scharli03traits}. In Java multiple \emph{interface} inheritance has been
+traits~\cite{scharli03traits}. 
+In Java multiple \emph{interface} inheritance has been
 supported since inception, and in Java8 default methods~\cite{} bring some of
@@
 inheritance of classes. 
+%Traits or Java8 interfaces still assume a CB model: 
+%But traits still require classes, which
+%are responsible for object construction and adding state.
 
@@
 
+\bruno{To be mentioned: 3 features of IB: Interface-based programming with multiple
+  inheritance; automatic type-refinements; a new approach that allows 
+\emph{type-safe} covariant refinement of state.
+}
+
   %With object interfaces, many Java programs can be built
@@
 and conduct various case studies. 
-Finally, we also formally define the behaviour of our \Q|Obj| implementation.
+Finally, we also formally define the behaviour of our \Q|Obj| annotation.
 

### 2016-03-21 OOPSLA16/sections/Overview.tex
Section 2

@@
 
-This section illustrates how our programming style, supported by the \mixinAnn{} annotation, enable powerful
-idioms similar to multiple inheritance and field type refinement.
-We propose a standard example: \Q@Animal@s with a two-dimensional \Q@Point2D@ representing their \Q@location@.
-Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
-Birds can \Q@fly@, thus their locations need to be three-dimensional \Q@Point3D@s (field type refinement).
-Finally, we model \Q@Pegasus@ (one of the best-known creatures in Greek
-mythology) as a kind of \Q@Animal@ with the skills
-of both \Q@Horse@s and \Q@Bird@s (multiple inheritance). A simple class diagram illustrating
-the basic system is given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some research argues in favor of using subtyping for modeling taxonomies, other research argues against this practice, we do not wish to take sides in this argument, but to provide an engaging example.}
-\bruno{Should we provide a one sentence summary in the abstract of how much code
-is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}
-\marco{Not in the abstract. But I think is good to have that number and to put in in the example section.
-I will set up the sentence and someone can compute the number.}
+This section illustrates how our programming style, supported by the
+\mixinAnn{} annotation, enable powerful programming idioms based on multiple
+inheritance and type refinements.  We propose a standard example:
+\Q@Animal@s with a two-dimensional \Q@Point2D@ representing their
+\Q@location@.  Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
+Birds can \Q@fly@, thus their locations need to be three-dimensional
+\Q@Point3D@s (field type refinement).  Finally, we model \Q@Pegasus@
+(one of the best-known creatures in Greek mythology) as a kind of
+\Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s (multiple
+inheritance). A simple class diagram illustrating the basic system is
+given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some
+  research argues in favor of using subtyping for modeling taxonomies,
+  other research argues against this practice, we do not wish to take
+  sides in this argument, but to provide an engaging example.}
+
+%\bruno{Should we provide a one sentence summary in the abstract of how much code
+%is needed in Java (without Obj) vs the approach with CJ, for the Pegasus example?}
+%\marco{Not in the abstract. But I think is good to have that number and to put in in the example section.
+%I will set up the sentence and someone can compute the number.}
+
 \subsection{Simple Multiple Inheritance with Default
@@
 \lstinputlisting[linerange=8-10]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
-
+\bruno{I dislike the code formating here.}
 \noindent The effect of the annotations is that a static \emph{factory} method called
@@
 system presented in the left-side of Figure~\ref{fig:pegasus}.
-Here you still can not appreciate the advantage of IB.
-Lets model now the complete animal system where \Q@Animal@s will include a \Q@location@
-representing their positions in space. We will use 2D points to keep track of locations using coordinates.
+The example is still not sufficient to appreciate the advantages of IB
+programming.
+Lets now model the complete animal system where \Q@Animal@s includes a \Q@location@
+representing their positions in space. We use 2D points to keep track of locations using coordinates.
 
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-Here we will illustrate how points are modelled with
-interfaces. In IB we do not talk about state directly; state is accessed
-using abstract (getter) methods. \bruno{unfinished sentence: Note how in}   The usual
-approach to model points in Java is to use a class with fields for the
-coordinates.
+Here we will illustrate how points are modelled with interfaces. In IB
+we do not talk about state directly. Instead state is accessed and
+manipulated using abstract methods.  The usual approach to model
+points in Java is to use a class with fields for the coordinates. 
+In Classless Java interfaces are used instead:
 
@@
 
-\noindent Our encoding over Java is now inconvenient: creating a new point object is cumbersome, even
+\noindent The encoding over Java is now inconvenient: creating a new point object is cumbersome, even
 with anonymous classes:
@@
 Point2D p = new Point2D() {
-  public int x(){return 4;} public int y(){return 2;}  }
+  int x(){return 4;} public int y(){return 2;}  }
 \end{lstlisting}
@@
 \paragraph{\Q@Point2D@ with object interfaces}
-This obvious ``constructor'' code can be automatically generated by our \mixin
+This obvious ``constructor'' code can be automatically generated by the \mixin
 annotation.  By annotating the interface \Q@Point2D@, a variation of the shown
@@
 data-structures. As such, it is
-supported in our code generation as \Q@with-@ methods. For example:
+supported in by \mixin as \Q@with-@ methods. For example:
 \begin{lstlisting}
@@
     int x();    int y(); // getters
-    Point2D withX(int val);    Point2D withY(int val); //with methods
+     //with methods
+    Point2D withX(int val);    Point2D withY(int val);
 }
@@
 \begin{comment}
-is expanded by our annotation into\footnote{
+is expanded by \mixin into\footnote{
 Note how we actually generate a real field \Q@int x=_x;@.
@@
 
-\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
-relies on a notion of abstract state:
-\marcoT{Method bodies can only refer to other methods;
-Objects are (conceptually) closures and the state is 
-composed by captured local variables.}
-Methods interacting with those are called \emph{abstract state operations}.
-
-Object interfaces provide support for four
-different types of abstract state operations \textcolor{red}{together with two types of
-general operations}, which are summarized in
-Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
-naming conventions and the types of the methods. Fluent setters are a variant of
-conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.
-
-\begin{figure*}
-\centering
-\saveSpaceFig
-\begin{tabular}{|l|l|l|l|}
-\hline
-& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
-\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
-& {\bf withers}        &   \Q@Point2D withX(int val)@                & Clones
-object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
-& \textbf{setters}        & \Q@void x(int val)@ & Sets the field
-\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
-& \textbf{fluent setters}        & \Q@Point2D x(int val)@ &Sets the field
-\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
-\multirow{3}{*}{Other operations} &
-\textbf{factory methods} & 
-%\begin{tabular}{l}
-\Q@static Point2D of(int _x,int _y)@
-%\end{tabular}
- & Generate factory method for \texttt{Point2D}.        \\
-\cline{2-4}
-& \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@ & Update all fields
-that match fields in \texttt{val}.        \\
-\cline{2-4}
-& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ & Set all fields
-that match fields in \texttt{val}.        \\ \hline
-\end{tabular}
-
-\caption{\textcolor{red}{(edited2)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
-  annotation. }
-
-\label{fig:abstractstate}
-
-\end{figure*}
+In summary, dealing with state (mutable or not) in object interfaces
+relies on a notion of abstract state, and state is not directly
+available to programmers. Instead programmers use methods, called
+\emph{abstract state operations}, to interact with state.
 
@@
 interface extends the \Q@Animal@ interface, the notion of points needs to
-be refined. Such kind of refinement is challenging
+be \emph{refined}. Such kind of refinement is challenging
 in typical class-based approaches. Fortunately, with object interfaces,
@@
 \lstinputlisting[linerange=67-70]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
-
+\bruno{Don't like the code format here either.}
 %Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
@@
 
-\bruno{At some point we should talk about the point that was discussed in the reviews}
-\marco{you mean the reviewer question on the list.size?}
 \paragraph{\Q@Point3D@ and properties updaters}
@@
 
+Property updaters never break class invariants, since they
+internally call operations that were already deemed
+safe by the programmer. For example a list object
+would not offer a setter for its \texttt{size} field (which should be kept hidden), thus
+a property updater would not attempt to set it.
+
 % To implement the old setter in a convenient way, \mixin supports one
@@
  \Q@Point3D set(Point2D val)@.
- }
-
- \marcoT{
- Property updaters never break class invariants, since they
- internally call operations that was already deemed
- safe by the programmer. For example a list object
- would not offer a setter for its size field, thus
- a property updater would not attempt to set it.}
+ }\bruno{we could or we do? if we could then the table should not
+   contain this operation; if we do the table is ok}
 
@@
 \Q@this@.
+
+\paragraph{Summary of operations in Classless Java} 
+%Method bodies can only refer to other methods;
+%Objects are (conceptually) closures and the state is 
+%composed by captured local variables.}
+%Methods interacting with state are called \emph{abstract state operations}.
+Object interfaces provide support for
+different types of abstract state operations: four field-based state
+operations; and  two types of
+general updating operations. Furthermore object interfaces support
+direct object intantiation via \texttt{of} factory methods. 
+Figure~\ref{fig:abstractstate} summarizes the six operations supported 
+by \mixin. The field-based abstract state operations are determined by
+naming conventions and the types of the methods. Fluent setters are a variant of
+conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.
+
+\begin{figure*}
+\centering
+\saveSpaceFig
+\begin{tabular}{|l|l|l|l|}
+\hline
+& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
+\multirow{4}{*}{\parbox{2.5cm}{State operations (for a field \texttt{x})}} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.
+\\ \cline{2-4}
+& {\bf withers}        &   \Q@Point2D withX(int val)@                & Clones
+object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
+& \textbf{setters}        & \Q@void x(int val)@ & Sets the field
+\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
+& \textbf{fluent setters}        & \Q@Point2D x(int val)@ &Sets the field
+\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
+\multirow{3}{*}{Other operations} &
+\textbf{factory methods} & 
+%\begin{tabular}{l}
+\Q@static Point2D of(int _x,int _y)@
+%\end{tabular}
+ & Factory method (generated).        \\
+\cline{2-4}
+& \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@
+& Update all matching fields in \texttt{val}.        \\
+\cline{2-4}
+& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ &
+Set all matching fields in \texttt{val}.        \\ \hline
+\end{tabular}
+
+\caption{Abstract state operations for a field \texttt{x}, together with other operations, supported by the \mixin
+  annotation. }
+
+\label{fig:abstractstate}
+
+\end{figure*}
 
@@
 
-\subsection{Our Approach}
+\subsection{The Classless Java Approach}
 Instead of writing a whole another class to provide the implementation for
@@
 
-With our approach, we provide a Java annotation \mixin to provide default
+With \CJ, we provide a Java annotation \mixin to provide default
 implementations for various methods and a mechanism to instantiate

### 2016-03-21 OOPSLA16/sections/Overview.tex
Section 2 code style.

@@
 
-\lstinputlisting[linerange=8-10]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
+\lstinputlisting[linerange=8-12]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
 \bruno{I dislike the code formating here.}
@@
 Point2D p = new Point2D() {
-  int x(){return 4;} public int y(){return 2;}  }
+  int x() {return 4;}
+  public int y() {return 2;}
+}
 \end{lstlisting}
@@
     int x();    int y(); // getters
-     //with methods
-    Point2D withX(int val);    Point2D withY(int val);
+    // with methods
+    Point2D withX(int val);
+    Point2D withY(int val);
 }
@@
 % Again, we model this member field with getter and setter methods:
-\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-60]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
 
@@
 
-\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 
-\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=71-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 \bruno{Don't like the code format here either.}
@@
 also we can change the order of the code to write the non generated stuff first.}
-\saveSpaceFig
-\lstinputlisting[linerange=79-91]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
+%\saveSpaceFig
+\lstinputlisting[linerange=89-115]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
 \caption{Generated boilerplate code.}
@@
 
-\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=84-84]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=58-58]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
+\lstinputlisting[linerange=58-60]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=ANIMAL
 
-\lstinputlisting[linerange=62-63]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=67-70]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=71-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=74-74]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=84-84]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2016-03-21 OOPSLA16/sections/Implementation.tex
Fix figure 4.
Implementation

--- 2016-03-18 OOPSLA16/sections/Implementation.tex
@@
-\section{A practical bridge between IB and CB}
+\section{Bridging between IB and CB in Java}
 
@@
 May be this should be "obj" processing?
-}}
+}}\bruno{The figure is blurred and the font size is too
+small. Hoayuan, please improve the figure.}
 \end{figure}
 
-Even if convinced of the advantages of IB,
-the practical convenience
-of using existing libraries points us into building IB support
-inside of an opportune CB language.\footnote{
-Rewriting libraries and application in this language/model
-would result in more reusable libraries, but it would be a taunting task.}
+Even if convinced of the advantages of IB, the practical convenience
+of using existing libraries points us into building IB support inside
+of an opportune CB language like Java.
+%\footnote{ Rewriting libraries and
+%  application in this language/model would result in more reusable
+%  libraries, but it would be a taunting task.}
+While creating a new language or language extension would also
+illustrate the point of IB (and would be more elegant), it would also
+require significant amounts of engineering to achieve a similar level
+of integration and tool support as Java.
+This section illustrates how we integrate IB directly into Java instead. 
+Disciplined use
 
-We believe Java is a good language for such integration, since it supports compilation agents[citeSomehting].
+\subsection{Compilation Agents}
 
+Java is a good language for such integration, since it supports compilation agents\cite{compilationagents}.
 That is, Java libraries can interact with the Java compilation process,

### 2016-03-21 OOPSLA16/sections/Introduction.tex

@@
 
-Object oriented languages have always strived to offer great code reuse.
+\marcoT{Object oriented languages always strive to offer great code reuse.}
 They aim to couple flexibility and rigour, expressive power and
@@
 languages are improved \emph{modularity} and \emph{multiple inheritance}~\bruno{I'm using multiple
-  inheritance in a more abstract way than you Marco.}  The literature
+  inheritance in a more abstract way than you Marco.}\marco{Yes, but will the reviewer think in the abstract or in the
+  concrete way about MI?}  The literature
 provides good examples on how easy and modular it is to combine
@@
 pure behaviour, in IB state is just a special kind of behaviour
-implementation. Objects are the only responsibles to define the
+implementation.
+\marco{NO NO!
+the advantage is not just multiple composition 
+(by the way, that is the term you are searching for, not multiple inheritance)
+By not having state reified in the ideal/platonic level, it is not fixed
+what is state and what is not, thus you can (model an encoding of) REMOVE fields and you can refine
+fields.}
+ Objects are the only responsibles to define the
 ultimate behaviour of a method and, for example, if such method is
@@
 implicitly.}
-
+\marco{we need to discuss, not clear what you mean}
 %\footnote{ In the same sense,
@@
 transparently hide it from Java programmers. 
-The implementation works by performing \emph{on-the-fly} AST
+The implementation works by performing \emph{on-the-fly}\marco{False, we
+generate the instrumented bytecode. on the fly ast rewriting is a term that means 
+that you modify the bytecode while you load it} AST
 rewriting, allowing existing Java tools (such as IDEs) to work

### 2016-03-21 OOPSLA16/sections/Implementation.tex
Section 3

--- 2016-03-18 OOPSLA16/sections/Implementation.tex
@@
-\section{A practical bridge between IB and CB}
+\section{Bridging between IB and CB in Java}
 
-\begin{figure}[t]\label{fig:lombok}
-\centering
-\includegraphics[width=3in]{pdfs/lombok.png}
-\caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]
-\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
-but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
-May be this should be "obj" processing?
-}}
-\end{figure}
+%\footnote{ Rewriting libraries and
+%  application in this language/model would result in more reusable
+%  libraries, but it would be a taunting task.}
+Creating a new language or language extension would be an
+elegant way to illustrate the point of IB. However, 
+significant amounts of engineering would be needed to build a practical
+language, and to achieve a similar level of integration and tool support
+as Java. To have an approach that can both illustrate IB
+programming and be practical, we have instead implemented Classless Java. 
+This section presents the Classless Java design. 
+Disciplined use of Classless Java (that is avoiding class
+definitions as done in Section~\ref{}), illustrates what \emph{pure} IB is like. 
+However in Classless Java CB and IB programming can be mixed together, and all 
+the practical convenience of using existing Java libraries, the full
+Java language, and IDE's are available in Classless Java. 
+The key to our implementation is the use of compilation agents, which 
+are supported in Java, and allow us to rewrite the Java AST just
+before compilation.
 
-Even if convinced of the advantages of IB,
-the practical convenience
-of using existing libraries points us into building IB support
-inside of an opportune CB language.\footnote{
-Rewriting libraries and application in this language/model
-would result in more reusable libraries, but it would be a taunting task.}
-
-We believe Java is a good language for such integration, since it supports compilation agents[citeSomehting].
-
+\subsection{Compilation Agents}
+Java supports compilation agents\cite{compilationagents}.
 That is, Java libraries can interact with the Java compilation process,
@@
 generation of the abstract syntax tree and the Java Bytecode.
-\marco{is this correct?}
-This process is facilitated by frameworks like Lombock~\cite{lombok}:
+\marco{is this correct?}\bruno{my impression is that Lombok happens
+  \emph{before} the AST is compiled (not between AST and bytecode generation)}
+This process is facilitated by frameworks like Lombok~\cite{lombok}:
 a Java library that aims at removing (or
 reducing) Java boilerplate code via
-annotations.
+annotations. Classless Java was created using Lombok.
 Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
-flow of \mixin annotation processing in Lombok.
+flow of \mixin annotation processing in Lombok.\bruno{Haoyuan, please
+  explain the diagram.}
+
 %There are a number of annotations provided by the
@@
 
-Using Lombok we created our \mixin annotation, that supports object
-interfaces as described in the former examples.
+\begin{figure}[t]
+\centering
+\includegraphics[width=3in]{pdfs/lombok2.png}
+\caption{The flow chart of Lombok annotation processing. 
+\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
+but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
+May be this should be "obj" processing?
+}}\bruno{Haoyuan, please remove references to Eclipse in the Figure,
+that way it works both for Eclipse and Javac}
+\label{fig:lombok}
+\end{figure}
 
-Compilation Agents have advantages with respect to
-alternatives like extending the language, using a macro pre processors 
-or using the Java Annotation processors.\marco{are we sure that annotation processors are not compilation agents too?}
-
-Our disciplined \mixin annotation offer some more advantages with respect to arbitrary (Compilation agent driven) AST rewriting.
+\paragraph{Advantages of Lombok}
+The Lombok compilation agent has advantages with respect to alternatives like
+pre-processors, or other Java annotation processors. In some sense
+Lombok can be seen as a poor man's macro system~\cite{} for Java. 
+\bruno{Marco, do you agree with calling Lombok a poor man's macro system?}
 
@@
 
-\paragraph{Auxiliary files/informations do not pollutes the program:}
-Compilation Agents alters the generation process of the class files,
-by directly modifying the AST.
-Neither the source code is modified nor new Java files are generated.
-Moreover, and probably more importantly, Lombok supports generation of
-  code \emph{inside} a class/interface, which conventional
-  Java annotation processors do not support.
+\paragraph{Direct modification of the AST:}
+Lombok alters the generation process of the class files,
+by directly modifying the AST. Neither the source code is modified nor 
+new Java files are generated. Moreover, and probably more importantly, 
+Lombok supports generation of code \emph{inside} a class/interface, 
+which conventional Java annotation processors do not support this. For
+example, the standard \texttt{javax.annotation} processor, which is part of the
+Java platform, only allows generation of \emph{new code}, and the
+new code has to be written in \emph{new files}. Modification and/or
+reinterpretation of existing code is not supported.
 
 \paragraph{Modularity:}
-While general preprocessing/macros can act across module boundaries,
-Compilation agents act modularly on each class/compilation unit, and it makes
-sense to apply the transformation to one class/interface at a time, and only to annotated classes/interfaces.
-This allows library code to be reused without the need of being
-reprocessed and recompiled, making our approach 100\% compatible with existing Java libraries, than can be used 
-and extended normally. Of course, Java libraries can also receive and use instances of object interfaces as normal objects.
+While general preprocessing acts across module boundaries, compilation
+agents act modularly on each class/compilation unit. It makes sense to
+apply the transformations to one class/interface at a time, and only to
+annotated classes/interfaces. This allows library code to be reused
+without the need of being reprocessed and recompiled, making our
+approach 100\% compatible with existing Java libraries, which can be
+used and extended normally. Of course, Java libraries can also receive
+and use instances of object interfaces as normal objects.
 
 \paragraph{Tool support:}
-Features integrate/supported directly in the language are often also supported by (most) tools.
-For example in Eclipse, the processing is performed transparently and the information of
-the interface from compilation is captured in the ``Outline'' window. This includes
-all the methods inside the interface as well as the generated ones.  
-This also means that the IDE functionality for content assist and autocomplete
-will work for the newly generated methods.
+Features written in Lombok integrate and are supported directly in the
+language, and are often also supported by (most) tools.  For example in Eclipse, the processing is
+performed transparently and the information of the interface from
+compilation is captured in the ``Outline'' window.\bruno{Haoyuan: show
+and explain figure here.} This includes all
+the methods inside the interface as well as the generated ones.  This
+also means that the IDE functionality for content assist and
+autocomplete will work for the newly generated methods.
 
-\paragraph{Clarity against Obfuscation:}
+\paragraph{Clarity against obfuscation:}
 Preprocessors bring great power, which can easily be misused producing
@@
 Compilation agents start from Java syntax, but they can reinterpret it.
-Preserving the syntax avoid syntactic conflicts, and allows may tools to transparently work.
+Preserving the syntax avoid syntactic conflicts, and allowing many
+tools to work transparently.
 
-The reinterpretation could still be surprising for badly designed rewritings.
-However, ``Outline'' windows/ auto-completion facilities 
-will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.
+\subsection{AST Reinterpretation in Classless Java}
 
-Our annotation reinterpret the syntax for the sole goal of enhance/complete code:
-we satisfy the behaviour of abstract methods, we add methods and we refine return types.
-We consider this to be quite easy to follow, since it is similar to what happens in normal inheritance.
+Ofcourse, careless reinterpretation of the AST could still be
+surprising for badly designed rewritings.  Classless Java reinterprets
+the syntax with the sole goal of \emph{enhancing and completing code}:
+we satisfy the behaviour of abstract methods; we add method
+implementations; and we refine return types.  We consider this to be
+quite easy to follow and reason about, since it is similar to what
+happens in normal inheritance.  Refactoring operations like renaming
+and moving should work transparently in conjunction with our
+annotaiton, since they rely on the overall type structure of the
+class, that we do not arbitrarily modify but just complete.
 
-We believe that refactoring operations like rename and move
-should work transparently in conjunction with our annotaiton, since they rely on 
-the overall type structure of the class, that we do not arbitrarily modify but just complete.
+Thus, in addition to the advantages of Lombok, Classless Java offers
+some more advantages with respect to arbitrary (Compilation agent driven) AST rewriting.
 
@@
 
-\paragraph{Syntax and Type errors:}
+\paragraph{Syntax and Type Errors:}
 Some preprocessors (like the C one) can produce syntactically invalid code.
-Lombok ensures only syntactically valid code is produced; however, type errors can appear.
+Lombok ensures only syntactically valid code is produced. %; however, type errors can appear.
+Classless Java additionally guarantees that no type errors are introduced
+in generated code and client code. We discuss these two guarantees in
+more detail next:
 
-Our annotation is guaranteed to not introduce errors in generate code and client code.
-See the following for details, where we identify tree form of types guarantees.
+\begin{itemize}
 
-\paragraph{Self coherence}: the generated code itself is well-typed; that is,
+\item{\bf Self coherence}: the generated code itself is well-typed. That is,
   type errors are not present in code the user has not written.
@@
 understandable error, or the interface can be successfully annotated and the generated code is well-typed.
-We guarantee \textit{Self coherence}.
 
-\paragraph{Client coherence}: all the client code (for example method calls)
+\item{\bf Client coherence}: all the client code (for example method calls)
   that is well-typed before code generation is also well-typed after the generation.
 The annotation just adds more behaviour without removing any functionality.
-We guarantee \textit{Client coherence}.
 
-\paragraph{Heir coherence}: interfaces (and in general classes) inheriting the instrumented code are well-typed if they were well-typed without the instrumentation.
-This would forbid adding any (default or abstract) method to the annotated interfaces, including type refinement.
-\mixin  does not guarantee \textit{Heir coherence}.
-Indeed consider the following:
+\end{itemize}
+
+\paragraph{Heir coherence:} Another form of guarantee that could be
+useful in AST rewriting would be heir coherence. That is, interfaces
+(and in general classes) inheriting the instrumented code are
+well-typed if they were well-typed without the instrumentation.
+Classless Java \emph{does not} guarantee heir coherence.  The reason
+is that this would forbid adding any (default or abstract) method to
+the annotated interfaces, including type refinement. Indeed consider
+the following:
 
@@
 \noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@. \\*
+This would break \Q@C@. 
 Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
@@
 
-Our automatic type refinement
-is a useful and convenient feature, but not transparent to the heirs of the annotated interface.
-They need to be aware of the annotation semantics and provide the right type while refining methods.
-Please, note how Java libraries always break heir coherence during evolution
- (is just enough to add any method to any non final class), and still claim backward compatibility
- (false in theory but statistically true in practice).
-We think refining return types breaks heir coherence ``less" than normal library evolution.
+Our automatic type refinement is a useful and convenient feature, but
+not transparent to the heirs of the annotated interface.  They need to
+be aware of the annotation semantics and provide the right type while
+refining methods. To support heir coherence, Classless Java would need 
+to give up automatic type refinement, which is an essential part of IB programming.
+However, the reader should that note that Java libraries almost always break heir
+coherence during evolution and still claim backward compatibility (false in
+theory but statistically true in practice). In practice to break heir
+coherence it is just enough to add any method to any
+non final class of a Java library.  We think refining return
+types breaks heir coherence ``less" than normal library evolution.
 
-\paragraph{Type preservation:}
+In Section~\ref{} we make the claims the guarantees of Classless Java
+more formal.
+
+\paragraph{Type preservation:}\bruno{Move to next section?}
 We preferred to introduce self,client and hair coherence instead of referring to conventional
@@
 
-\paragraph{Limitations}\marco{ to reorganize with updated limitations}
+\subsection{Limitations}\marco{ to reorganize with updated limitations}
 Our prototype implementation using Lombok has certain limitations:

### 2016-03-21 OOPSLA16/sections/Implementation.tex
Section 3

@@
-\section{Bridging between IB and CB in Java}
+\section{Bridging between IB and CB in Java}\label{sec:imp}
 
@@
 are supported in Java, and allow us to rewrite the Java AST just
-before compilation.
+before compilation. We discuss advantages and limitations of our approach.
 
@@
 
-\paragraph{Type preservation:}\bruno{Move to next section?}
-We preferred to introduce self,client and hair coherence instead of referring to conventional
-type preservation,
-to better model how our approach behave in a object oriented software ecosystem with inheritance,
-where only some units may be translated/expanded.
-Note inheritance crucial influence in hair coherence, and how client coherence allows
-for only some units to be translated/expanded.
-
-This means that even a library designer can decide to start using IB and still
-retain backward compatibility with its clients.
-
 \subsection{Limitations}\marco{ to reorganize with updated limitations}

### 2016-03-21 OOPSLA16/sections/Translation.tex
Some more text. Table and flow chart.

@@
 
-\begin{figure*}[t]\label{figure:translation}
+\begin{figure*}[t]
 \centering
@@
 \caption{The translation functions of $\mixinAnn$ and $\weakAnn$.}
+\label{figure:translation}
 \end{figure*}
 
-\begin{figure*}[t]\label{figure:valid}
+\begin{figure*}[t]
 \centering
@@
 \end{minipage}
-\caption{The $\otherMethod$ and $\valid$ functions (left) and auxiliary functions (right).}
+\caption{The $\otherMethod$ and $\valid$ functions (left) and
+  auxiliary functions (right).}
+\label{figure:valid}
 \end{figure*}
 
-This section gives an overview of what $\mixinAnn$ generates in Classless Java. We provide a
-formalization of Classless Java in Appendix~\ref{sec:formal}, which models the essence of
-Java interfaces with default methods, including the syntax as well as the typing rules.
+This section gives an overview of what $\mixinAnn$ generates in
+Classless Java, and what properties are guaranteed in the translation.
+
+We provide a formalization of Classless Java in Appendix~\ref{sec:formal}, which models the essence of
+Java interfaces with default methods, including the syntax as well as
+the typing rules. Classless Java is just a proper subset of Java 8, so
+it is easy to understand the translation presented in this section
+without the syntax and typing rules of Classless Java.
 Since the formalized part of Classless Java does not consider casts or \Q@instanceof@,
@@
 
-\begin{figure*}[t]\label{figure:ofmethod}
+\begin{figure*}[t]
 \centering
@@
 \end{minipage}
-\caption{The generated $\QM{of}$ method (left) and auxiliary functions (right).}
+\caption{The generated $\QM{of}$ method (left) and auxiliary functions
+  (right).}
+\label{figure:ofmethod}
 \end{figure*}
@@
 method for $\C_0$, if such a method does not exist in its domain, and all the abstract methods are
-valid for the annotation. Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod}\haoyuan{Refs are wrong!?} present more details on the auxiliary
-functions. The formal definition is available in Appendix~\ref{subsec:translation}.
+valid for the annotation.
 
-The function $\ofMethod$ generates the static method $\QM{of}$, as an object factory. It detects all the
-field methods of $\C_0$, take them as arguments, and in the return statement is an anonymous class which generates getters,
-fluent setters and with-methods in certain cases.
-Some other features of $\mixinAnn$, including non-fluent setters and the $\QM{with}$ method, however, are not formalized in Classless Java.
-Appendix~\ref{subsec:otherfeatures} gives an informal explanation of those methods.
+Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod} present more details on the auxiliary
+functions. \bruno{if the figures are mentioned here, what should
+  follow is the explanation on what is one the figures. Please fix
+  this Haoyuan and makes sure that everything that is on the figures
+  is explained here. }
+The formal definition is available in
+Appendix~\ref{subsec:translation}.\bruno{what formal definition?}
 
-\subsection{Guarantees}
+The function $\ofMethod$ generates the static method $\QM{of}$, as an
+object factory. It detects all the field methods of $\C_0$, take them
+as arguments, and in the return statement is an anonymous class which
+generates getters, fluent setters and with-methods in certain
+cases. Some other features of $\mixinAnn$, including non-fluent
+setters and the $\QM{with}$ method, however, are not formalized in
+Classless Java. Appendix~\ref{subsec:otherfeatures} gives an informal
+explanation of those methods.
 
-To understand to what extent our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
-\begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; that is,
-  type errors are not present in code the user has not written.
-In our case it means that either \mixin{} produces (in a controlled way) an
-understandable error, or the interface can be successfully annotated and the generated code is well-typed.
-We guarantee \textit{Self coherence}.
+\bruno{I think there is explanation missing in this subsection;
+  probably text which is now in the appendix. Haoyuan, please take
+  care of this.}
 
-\item \textit{Client coherence}: all the client code (for example method calls)
-  that is well-typed before code generation is also well-typed after the generation.
-The annotation just adds more behavior without removing any functionality.
-We guarantee \textit{Client coherence}.
+\subsection{Results}\label{subsec:results}
 
-\item \textit{Heir coherence}: interfaces (and in general classes) inheriting the instrumented code are well-typed if they were well-typed without the instrumentation.
-This would forbid adding any (default or abstract) method to the annotated interfaces, including type refinement.
-\mixin  does not guarantee \textit{Heir coherence}.
-Indeed consider the following:
-
-\begin{lstlisting}
-interface A { int x(); A withX(int x); }
-@Obj interface B extends A {}
-interface C extends B { A withX(int x); }
-\end{lstlisting}
-
-\noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@. \\*
-Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
-would be correct before translation, but would be ill-typed after the translation.
-\end{itemize}
-
-\noindent This means that our automatic type refinement
-is a useful and convenient feature, but not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantics and provide the right type while refining methods.
-
-\subsection{Results}
+Classless Java provides some guarantees regarding the generated
+code. Essentially, Classless Java ensures the \emph{self coherence}
+and \emph{client coherence} properties informally introduced in
+Section~\ref{sec:imp}. Furthermore, we can show that 
+\emph{if there are no type-refinements}, then \emph{heir coherence} also
+holds. The result about heir coherence is possible to prove 
+because the translation is split into two parts. In essence heir
+coherence is a property of the translation of  $\weakAnn$, but not of 
+\mixin.
 
@@
 
-\begin{thm}[@ObjOf tuning]
+\bruno{Do we have good names for the theorems Marco? Cannot be called
+  tunning now!}
+\begin{thm}[@ObjOf]
 If a given interface table $\II$ IT is OK
@@
 
-\begin{thm}[@Obj tuning]
+\begin{thm}[@Obj]
 If a given interface table $\II$ IT is OK
@@
 no subtypes and no objects of that type are created, then type safety
+\bruno{of the generated code?}
 is guaranteed after the successful translation.
@@
 step of \mixin, namely what $\weakAnn$ does in the formalization, is
-guaranteed to be type-safe for the three kinds of coherence by the $\weakAnn$  tuning theorem.
-\mixin
-tuning is more interesting: since \mixin does not guarantee heir
+guaranteed to be type-safe for the three kinds of coherence by the $\weakAnn$ theorem.
+The \mixin theorem is more interesting: since \mixin does not guarantee heir
 coherence, we explicitly exclude the presence of heirs. In this way
-\mixin tuning guarantees only self and client coherence. The formal
-theorem proofs are available in the supplementary materials.
+the \mixin theorem guarantees only self and client coherence. The formal
+theorem proofs are available in the supplementary
+materials\bruno{point to appendix instead}.
+
+\paragraph{Type preservation:} 
+Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
+type preservation theorems~\cite{}. The reason is 
+to better model how our approach behaves in a object oriented software ecosystem with inheritance,
+where only some units may be translated/expanded.
+Note that inheritance crucial influence in heir coherence, and how client coherence allows
+for only some units to be translated/expanded.\bruno{I don't
+  understand the point of the last sentence? Rephrase?}
+This means that even a library designer can decide to start using IB
+programming and still
+retain backward compatibility with its clients.
 

### 2016-03-21 OOPSLA16/sections/CaseStudy.tex
added one paragraph for the new case study C0

@@
   print operation).}\label{fig:ep}
-\end{figure}% \bruno{Yanlin: code is overflowing the margins. Note that
+\end{figure*}% \bruno{Yanlin: code is overflowing the margins. Note that
   % there are 2 spare lines that can be used on the right!}
@@
 \vspace{5pt}
-\hspace{80pt}\begin{tabular}{ccc}
+\begin{tabular}{ccc}
 \hline
@@
 To ensure a fair comparison, we used the same coding style as Bono et al.'s.
+
+\subsection{Refactoring a Compiler}
+This case study applies our approach to a slightly larger code base of a simple
+one-pass compiler for \lstinline{C0} (a subset of \lstinline{C}). This compiler
+is a handwritten, monolithic compiler used for educational purposes at Aarhus
+University, Denmark\footnote{Source code available at
+  http://cs.au.dk/~mis/dOvs/Czero.java}. \lstinline{C0} is restricted to
+integers, several control structures and function declaration/definition and
+basic I/O statements. All the code (including parsing, code generation for
+various structures, error handling) is entangled in one big file. Following our
+approach with \mixin, we refactored the code to several smaller interfaces
+(including \lstinline{Constants}, \lstinline{MemberFields}, interfaces for
+parsing different structures, etc). After the refactoring, the new code does not
+reduce code as previous case studies. SLOC for original code without comments
+and blank lines is 828, while ours is 830 (for fair comparison, we put all
+interfaces into one file when calculating SLOC). However, the case study shows
+two things: Classless Java can be applied to real code base such as compilers;
+code becomes more modular without the sacrifice of code amount/simplicity.

### 2016-03-21 CZero/src/BaseOpers.java

@@ -1,4 +1,4 @@
-package c0.cj;
+
 
 public interface BaseOpers extends MemberFields, Constants {
     default int nextChar() {
@@ -57,7 +57,7 @@ public interface BaseOpers extends MemberFields, Constants {
                 tKind(tERROR());
                 break; 
             }
-            tKind(tCONST);
+            tKind(tCONST());
             tIntValue(c());
             nextChar();
             if (c()!='\'') {
@@ -69,7 +69,7 @@ public interface BaseOpers extends MemberFields, Constants {
             case '=': nextChar(); 
             if (c()=='=') { 
                 nextChar();
-                tKind(tEQ);
+                tKind(tEQ());
             } else {
                 tKind(tASSIGN());
             }
@@ -79,7 +79,7 @@ public interface BaseOpers extends MemberFields, Constants {
                 nextChar();
                 tKind(tLEQ());
             } else {
-                tKind(tLT);
+                tKind(tLT());
             }
             break;
             case '>': nextChar();
@@ -87,7 +87,7 @@ public interface BaseOpers extends MemberFields, Constants {
                 nextChar();
                 tKind(tGEQ());
             } else {
-                tKind(tGT);
+                tKind(tGT());
             }
             break;
             case '!': nextChar();
@@ -160,46 +160,43 @@ public interface BaseOpers extends MemberFields, Constants {
     }
 
     default void printToken() {
-        switch (tKind()) {
-
-           if (tKind() == tLPAR()) System.out.print("("); break;
-            else if (tKind() == tRPAR()) System.out.print(")"); break;
-            else if (tKind() == tASSIGN()) System.out.print("="); break;
-            else if (tKind() == tSEMI()) System.out.print(";"); break;
-            else if (tKind() == tCOMMA()) System.out.print(","); break;
-            else if (tKind() == tEQ()) System.out.print("=="); break;
-            else if (tKind() == tNE()) System.out.print("!="); break;
-            else if (tKind() == tID()) System.out.print(tIdValue()); break;
-            else if (tKind() == tCONST()) System.out.print(tIntValue()); break;
-            else if (tKind() == tCHAR()) System.out.print("char"); break;
-            else if (tKind() == tADD()) System.out.print("+"); break;
-            else if (tKind() == tSUB()) System.out.print("-"); break;
[…]

### 2016-03-21 CZero/src/Constants.java

@@ -1,4 +1,4 @@
-package c0.cj;
+
 
 import java.io.PrintWriter;
 

### 2016-03-21 CZero/src/MemberFields.java

@@ -1,4 +1,4 @@
-package c0.cj;
+
 
 import java.io.FileReader;
 import java.io.PrintWriter;

### 2016-03-21 CZero/src/ParseExpression.java

@@ -1,4 +1,4 @@
-package c0.cj;
+
 
 import java.util.Map;
 

### 2016-03-21 CZero/src/ParseFunctions.java

@@ -1,4 +1,4 @@
-package c0.cj;
+
 
 import java.util.HashMap;
 import java.util.Map;

### 2016-03-21 CZero/src/c0/cj/BaseOpers.java

@@ -1,202 +1,0 @@
-package c0.cj;
-
-public interface BaseOpers extends MemberFields, Constants {
-    default int nextChar() {
-        try {
-            c(in().read());
-        } catch (Exception e) {
-            c(-1);
-        }
-        // fix line and col counting
-        // (\n and \r on Windows ...)
-        if (c()=='\n') {
-            tLine(tLine()+1);
-            tCol(1);
-        } else tCol(tCol()+1);
-        return c();
-    }
-
-    default int nextToken() {
-        switch (c()) {
-            case '(': tKind(tLPAR());   nextChar(); break;
-            case ')': tKind(tRPAR());   nextChar(); break;
-            case ';': tKind(tSEMI());   nextChar(); break;
-            case ',': tKind(tCOMMA());  nextChar(); break;
-            case '+': tKind(tADD());    nextChar(); break;
-            case '-': tKind(tSUB());    nextChar(); break;
-            case '*': tKind(tMUL());    nextChar(); break;
-            case '/': tKind(tDIV());    nextChar(); break;
-            case '%': tKind(tMOD());    nextChar(); break;
-            case '{': tKind(tLBRACK()); nextChar(); break;
-            case '}': tKind(tRBRACK()); nextChar(); break;
-            case '|': tKind(tOR());     nextChar(); break;
-            case '&': tKind(tAND());    nextChar(); break;
-            case '#': tKind(tSHARP());  nextChar(); break;
-            case '.': tKind(tDOT());    nextChar(); break;
-            case '\n':
-            case '\r':
-            case ' ': tKind(tWHITE());  nextChar(); break;
-            case '\'': nextChar();
-            if (c()=='\\') {
-                nextChar();
-                if (c()!='n') {
-                    tKind(tERROR());
-                    break; 
-                }
-                nextChar();
-                if (c()!='\'') {
-                    tKind(tERROR());
-                    break; 
-                }
-                nextChar();
-                tKind(tCONST());
-                tIntValue(10);
-                break;
-            } 
-            if (c()==-1 || c()=='\'') { 
-                tKind(tERROR());
-                break; 
-            }
-            tKind(tCONST());
-            tIntValue(c());
-            nextChar();
-            if (c()!='\'') {
-                tKind(tERROR());
-                break;
-            }
-            nextChar();
-            break;
-            case '=': nextChar(); 
-            if (c()=='=') { 
-                nextChar();
[…]

### 2016-03-21 OOPSLA16/sections/Implementation.tex
Section 3: flow chart, screenshot and text.

@@
   explain the diagram.}
+\haoyuan{Source code in a Java program is first parsed into its syntactic structure,
+which we call abstract syntax tree (AST). The AST is passed to Lombok, and a corresponding
+AST defined in Lombok is generated for the annotated node. An annotation like \mixin is declared
+inside Lombok, together with an implementation of its (Eclipse or Javac) handler. The handler is
+free to modify the information of the annotated node, or even inject new nodes (like methods, inner classes,
+etc). Finally, Java compiler works on the modified AST to generate bytecode. During the compilation,
+no source code is changed, and no new files are created, bytecode takes the only difference.}
 
@@
 \centering
-\includegraphics[width=3in]{pdfs/lombok2.png}
-\caption{The flow chart of Lombok annotation processing. 
+\includegraphics[width=3in]{pdfs/lombok3.png}
+\caption{The flow chart of \mixin annotation processing.
 \marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
@@
 }}\bruno{Haoyuan, please remove references to Eclipse in the Figure,
-that way it works both for Eclipse and Javac}
+that way it works both for Eclipse and Javac}\haoyuan{done}
 \label{fig:lombok}
@@
 and explain figure here.} This includes all
-the methods inside the interface as well as the generated ones.  This
-also means that the IDE functionality for content assist and
+the methods inside the interface as well as the generated ones. \haoyuan{In Figure~\ref{fig:screenshot} (wrong),
+the screenshot presents the Outline window of Eclipse, where the generated methods including \lstinline{Point2D.of}, \lstinline{Point3D.of},
+\lstinline{Point3D.withX} and \lstinline{Point3D.withY} are visible to users, showing their types and modifiers.} 
+This also means that the IDE functionality for content assist and
 autocomplete will work for the newly generated methods.
+
+\begin{figure}[h]\label{fig:screenshot}
+\centering
+\includegraphics[width=3.3in]{pdfs/screenshot2.png}
+\caption{Generated methods shown in the Outline window of Eclipse.}
+\end{figure}
 

### 2016-03-22 OOPSLA16/sections/Introduction.tex
Intro improved

@@
 interfaces directly. In IB interfaces own the implementation for the behaviour, which
-is structurally defined in their interface. Programmers do not define objects directly, but
+is structurally defined in their interface.
+
+\marcoT{
+A key challenge lies in how to model state, which is fundamental to
+have stateful objects. 
+Programmers do not define objects directly, but
 delegate the task to \emph{object interfaces}, whose role is similar
-to non abstract classes in CB languages. Key motivations for IB
-languages are improved \emph{modularity} and \emph{multiple inheritance}~\bruno{I'm using multiple
-  inheritance in a more abstract way than you Marco.}\marco{Yes, but will the reviewer think in the abstract or in the
-  concrete way about MI?}  The literature
+to non abstract classes in CB languages.
+All abstract operations in an object interface will be interpreted as \emph{abstract state operations}. The
+abstract state operations include various common utility methods (such
+as getters and setters, or clone-like methods). In the presence of
+subtyping, such operations often require special care, as their types
+need to be refined. Object interfaces provide support for
+type-refinement and can automatically produce code that deals with
+type-refinement adequately, while CB (as in Java, Scala and C\#) require 
+verbose explicit type-refinement.
+We believe that such verbosity hindered and slowed down
+the discovery of useful programming pattens involving type-refinement\cite{ trivial solution to the EP}.
+}
+
+Finally, object interfaces also provide static factory methods instantiating objects.
+
+\marcoT{%
+Objects are the only responsible to define the
+ultimate behaviour of a method and, for example, if such method is
+just a setter. Anything related to state is completely contained in
+the instances and do not leak in the inheritance logic.
+In CB, the structure of the state is fixed and can be only
+grown by inheritance; on the opposite, in IB the state is never fixed,
+and methods that ``looks like'' abstract setters and getters
+can always receive an explicit implementation down in the inheritance chain,
+improving \textbf{modularity and flexibility}.
+
+That is, the concept of abstract state is more fluid.
+In this paper we show how to support type-safe
+\textbf{field removal},
+\textbf{field type refinement} allowing a kind of covariant setters refinement,
+and \textbf{multiple inheritance}.}
+
+Indeed the literature
 provides good examples on how easy and modular it is to combine
@@
 
-To retain the easiness and modularity of combining multiple sources of
-pure behaviour, in IB state is just a special kind of behaviour
-implementation.
-\marco{NO NO!
-the advantage is not just multiple composition 
-(by the way, that is the term you are searching for, not multiple inheritance)
-By not having state reified in the ideal/platonic level, it is not fixed
-what is state and what is not, thus you can (model an encoding of) REMOVE fields and you can refine
-fields.}
- Objects are the only responsibles to define the
-ultimate behaviour of a method and, for example, if such method is
-just a setter. Anything related to state is completely contained in
-the instances and do not leak in the interfaces.\bruno{Not sure if
+%To retain the easiness and modularity of combining multiple sources of
+%pure behaviour, in IB state is just a special kind of behaviour
+%implementation.
+%\marco{NO NO!
+%the advantage is not just multiple composition 
+%(by the way, that is the term you are searching for, not multiple inheritance)
+%By not having state reified in the ideal/platonic level, it is not fixed
+%what is state and what is not, thus you can (model an encoding of) REMOVE fields and you can refine
+%fields.}
+% Objects are the only responsibles to define the
+%ultimate behaviour of a method and, for example, if such method is
+%just a setter. Anything related to state is completely contained in
+%the instances and do not leak in the interfaces.
+
+\bruno{To be mentioned: 3 features of IB: Interface-based programming with multiple
+  inheritance; automatic type-refinements; a new approach that allows 
+\emph{type-safe} covariant refinement of state.
+}
+
+\bruno{Not sure if
   this is totally clear. Implicit here is the fact that objects are
@@
 Java \emph{Classless Java}.
-Our proposed style enables powerful
-object-oriented idioms, similar to (stateful) multiple inheritance and
-field type refinement (allowing a kind of covariant setters
-refinement). \bruno{needs strenghtening}
-
-A key challenge lies in how to model state, which is fundamental to
-have stateful objects. In IB and Classless Java, the idea is that all abstract operations in an Object
-interface will be interpreted as \emph{abstract state operations}. The
-abstract state operations include various common utility methods (such
-as getters and setters, or clone-like methods). In the presence of
-subtyping, such operations often require special care, as their types
-need to be refined. Object interfaces provide support for
-type-refinement and can automatically produce code that deals with
-type-refinement adequately, while Java (and Scala) require explicit
-type-refinement.\bruno{strenghtening needed!}
-
-\bruno{To be mentioned: 3 features of IB: Interface-based programming with multiple
-  inheritance; automatic type-refinements; a new approach that allows 
-\emph{type-safe} covariant refinement of state.
-}
 

### 2016-03-22 OOPSLA16/sections/Implementation.tex
Couple of fixes

@@
 \caption{The flow chart of \mixin annotation processing.
-\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
-but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
-May be this should be "obj" processing?
-}}\bruno{Haoyuan, please remove references to Eclipse in the Figure,
-that way it works both for Eclipse and Javac}\haoyuan{done}
+}
 \label{fig:lombok}
@@
 \bruno{Marco, do you agree with calling Lombok a poor man's macro system?}
-
+\marco{No, it is as powerful as lisp macros, and have a bigger (may be even  a better) design.}
 %this seems like a commercial spot
@@
 Compilation agents start from Java syntax, but they can reinterpret it.
-Preserving the syntax avoid syntactic conflicts, and allowing many
+Preserving the syntax avoid syntactic conflicts, and \marcoT{allows} many
 tools to work transparently.
@@
 to give up automatic type refinement, which is an essential part of IB programming.
-However, the reader should that note that Java libraries almost always break heir
+However, the reader should note that Java libraries almost always break heir
 coherence during evolution and still claim backward compatibility (false in
@@
 
-In Section~\ref{} we make the claims the guarantees of Classless Java
-more formal.
+In Section~\ref{} \marcoT{we provide a formal definition for our
+safety claims.}
 

### 2016-03-22 OOPSLA16/sections/Translation.tex
Couple of fixes

@@
 
-We provide a formalization of Classless Java in Appendix~\ref{sec:formal}, which models the essence of
-Java interfaces with default methods, including the syntax as well as
-the typing rules. Classless Java is just a proper subset of Java 8, so
+\marcoT{We formalize syntax and typing for 
+ Classless Java} in Appendix~\ref{sec:formal}, which models the essence of
+Java interfaces with default methods.
+%, including the syntax as well as
+%the typing rules.
+ Classless Java is just a proper subset of Java 8, so
 it is easy to understand the translation presented in this section
@@
 
+\marcoT{Since our properties are about preserving typing,
+we do not need to formalize Classless Java semantic to proof our statements.}
 \begin{figure*}[t]

### 2016-03-22 OOPSLA16/sections/FutureWork.tex
little fixes

@@
 Since we are using Java8 to encode IB, we are unable to directly support qualifies
-like private or protected. In Java 9, where private methods will be allowed in
-interfaces; abstract state operations are, however, abstract methods,
+like private or protected. In Java 9 private methods will be allowed in
+interfaces. Abstract state operations are, however, abstract methods,
 so they can not be private.
@@
 
-Using packages to as a boundary we can make all object interfaces ``with privateness''
+Using packages as a boundary we can make all object interfaces ``with privateness''
 invisible types for the rest of the program.
@@
 
-Using Java9 private interface methods
+Using Java9 private interface methods%
 \footnote{Or some ugly pattern available today like public static classes with private methods}
@@
 In this way
-
+\marco{I fixed the code down, was wrong, there was no mention of hello in the original source}
 \begin{lstlisting}
-@Obj interface MyData{
-  private String f();//assuming Java9
-  private MyData f(String val);
+@Obj interface Text{
+  private String msg();//assuming Java9
+  private Text msg(String val);{f("hello");}
   }
@@
 \begin{lstlisting}
-@Obj interface MyData{
-  private Box<String> _f(){return cache(Box<String>.of("hello"));}
-  private String f(){return _f().val();}
-  private MyData f(String val){return _f().val(val);}
+@Obj interface Text{
+  private Box<String> _msg(){
+    return cache(Box<String>.of("hello"));}
+  private String msg(){return _msg().val();}
+  private Text msg(String val){
+    return _msg().val(val);}
   }

### 2016-03-22 OOPSLA16/sections/Translation.tex
little fixes

@@
 \item An interface table
-IT is OK if under such interface table, all interfaces are OK.
+IT is OK if under such interface table, all interfaces are OK,
+\marcoT{that is, well typed}.
 \item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
@@
 no subtypes and no objects of that type are created, then type safety
-\bruno{of the generated code?}
+\marcoT{of the generated code}
 is guaranteed after the successful translation.
@@
 where only some units may be translated/expanded.
-Note that inheritance crucial influence in heir coherence, and how client coherence allows
-for only some units to be translated/expanded.\bruno{I don't
+Note inheritance crucial influence in heir coherence, and how client coherence allows
+for only some units to be translated/expanded.
+\bruno{I don't
   understand the point of the last sentence? Rephrase?}
-This means that even a library designer can decide to start using IB
-programming and still
-retain backward compatibility with its clients.
+  \marco{ok I tried to rephase the next sentence, how is now?}
+This means that developers and designers of Java libraries and frameworks
+can start using IB (and our \mixin annotation) in the evolution of their products
+and still retain backward compatibility with their clients.
 

### 2016-03-22 OOPSLA16/sections/Appendix_Formalization.tex
Some related work reorganization

@@
 \end{figure*}
+
+\section{Expanded related work}\label{sec:ERW}
+
+\paragraph{Meta-Programming Competes with Language Extensions}
+The most obvious solution to adding features to a language is via syntactic
+extensions. Syntactic extensions are often implemented as
+desugarings to the base language. For example,
+the Scala compiler was extended to directly support XML syntax.
+However, when syntactic extensions are independently created it is hard
+to combine multiple extensions into one.
+SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language that
+aims at making the creation and composition of syntactic sugar
+extensions easy, by allowing programmers to extend Java with custom
+features (typically for DSLs). However SugarJ's goals are different
+from language tuning: SugarJ aims at creating and composing new syntax;
+whereas language tuning merely reinterprets existing syntax. It is
+clear that reinterpreting existing syntax only can be limiting
+for some applications. However, when language tuning is possible
+it has the advantage that existing tools for the language work
+out-of-the-box (since the syntax is still the same); and composition
+of independently developed tunings is straightforward.
+
+Scala-Virtualized~\cite{Rompf2012scalavirtualized} is an extension to Scala,
+which allows blending shallow and deep embedding of DSLs. It redefines
+some of Scala's language constructs to method calls,
+which can be overridden by DSL implementer. Thus Scala-Virtualized
+can also reinterpret syntax, and be seen as a form of language tuning.
+However, although many Scala's language constructs are supported,
+not all language constructs can be virtualized.
+
+%These extensions are admittedly useful, however they are essentially
+%creating a fork of the language, and rarely the new fork gains enough
+%traction to become the main language release. % On this
+% topic we mention
+
+When the base language has a flexible enough syntax and a fast and
+powerful enough reflection mechanism, we may just need to play with
+operator overloading and other language tricks to discover that the
+language feature we need can be expressed as a simple library in our
+language. An example of this is SQLAlchemy~\cite{bayer2012sqlalchemy} in
+Python, which uses operator overloading to dynamically turn normal python expressions into database queries
+without requiring any syntactic extensions to Python.
+Java-like languages tend to sit in the middle of two extremes:
+libraries can not influence the type system, so many solutions valid
+in Python or other dynamic languages are not applicable, or have the
+cost of losing type-safety.
+
+In Java-like languages compile time code generation comes at the rescue: if, for a certain feature
+(\mixin in our case), it is possible to use the original language syntax to
+\emph{express/describe} any specific instantiation of such feature (annotating a
+class and providing getters), then we can insert in the compilation process a tool
+that examines and enriches the code before compilation. No need to modify the
+original source (for example we can work on temporary files).
+Java is a particular
+good candidate for this kind of manipulation since it already provide ways to
+define and integrate such tools in its own compilation process via annotation
+processing. In this way
+there is no need of temporary files, and there is a well-defined way of putting
+multiple extensions together.
+
+Other languages offer even stronger support for safe code manipulation:
+Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
+all allow to execute code at compile time. They generate code that is
+transparently integrated in the program that is being
+generated/processed/compiled. In particular, MetaFjig offers a property called
+\emph{meta-level-soundness}, ensuring by construction that library code (even if wrong or unreasonable) never generates ill-typed code. This is roughly equivalent to
+\textit{Self coherence}, that we have to manually prove.
+Since MetaFjig is not working on annotated classes, there is not a ``semantic with/without annotations''. Our \mixin tuning theorem does not make sense in such context.
+
+\paragraph{Formalization of Java 8}
+We provide a simple formalization for a subset of Java including
+default/static interface methods and object initialization literals
+(often called anonymous local inner classes).  A similar formalization
+was drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize
+defender (default) methods in Java. In their formalization, classes
+and interfaces can have only one method \Q@m()@ without arguments, so
+as to simplify method overloading and renaming. Classless Java is more
+general, as it supports multiple methods with arguments, it supports
+static methods, and features such as multiple inheritance of
+interfaces and reabstraction of default methods are also modeled.
+
+\section{Interaction of Interface Methods with Interface Composition}
+\bruno{I don't think this section makes alot of sense in the context 
+of the new paper! I think perhaps some of this could be mentioned 
+in related work.
+
+In my opinion, instead of this section, we could have one section 
+that demonstrates our implementation and its advantages.
+
+\begin{itemize}
+
+\item Show a diagram of how the approach works?
+
+\item Syntax and type system reuse; (good for error messages, for example)
+\item On-the-fly AST rewriting (means no new files are created, and functionality
+is immediatelly available for IDE's for example); this is better than conventional 
+annotation processors.
+\item Eclipse integration; code completion.
+\item No integration issues in typical pre-processors; (the process is transparent; 
+no need for an external tool that generates code in a first step).
+\item Separate compilation (which relates to the properties?)
+
+\end{itemize}
+
+}
+Before formalizing Classless Java and object interfaces, it is helpful
+to informally discuss the behaviour of Java8
+interfaces, in contrast with conventional trait models.
+
+%We show several interesting cases
+%and summarize compilation result of method composition into 3 categories:
+%conflict error, accepted (both abstract) and conservative error. Meanwhile we
+%also show different composition handling mechanism among traits, javac and ECJ.
+
+\subsection{Methods in Java 8 Interfaces}
+In Java 8 interfaces there are three types of methods: abstract, default, and static methods. Default and static methods were not allowed in interfaces in previous versions of Java.
+
+\paragraph{Static methods} are handled in a very clean way: they are visible only in
+  the interface in which they are explicitly defined. This means the following code
+  is ill-typed.
+\begin{lstlisting}
+interface A0 { static int m(){return 1;} }
+interface B0 extends A0 {}
+... B0.m()//ill typed
+\end{lstlisting}
+This is different from the way static methods are handled in classes. Here
+static methods have simply no interaction with interface
+composition (\Q@extends@ or \Q@implements@).
+
+\paragraph{Abstract method} composition is accepted when there exists a most specific one.
+  For example, here method \texttt{Integer m()} from \texttt{B1} is visible in \Q@C1@.
+\begin{lstlisting}
+interface A1 { Object m(); }
+interface B1 { Integer m();}
+interface C1 extends A1,B1 {} //accepted
+\end{lstlisting}
+
+\paragraph{Default methods} conflict with any other default or abstract method. For
+  example the following code is rejected due to method conflicts.
+\begin{lstlisting}
+interface A2 { default int m() {return 1;}}
+interface B2 { int m(); }
+interface C2 { default int m() {return 2;}}
+interface D2 extends A2,B2 {} //rejected due to conflicting methods
+interface E2 extends A2,C2 {} //rejected due to conflicting methods
+\end{lstlisting}
+Note how this is different from what happens in most trait models, where \Q@D2@
+would be accepted, and the implementation in \Q@A2@ would be part of the
+behavior of \Q@D2@.
+
+\paragraph{Resolving conflicts:}
+A method in the current interface wins over any method in its
+super-interfaces, provided that the method
+is the most specific one. This method also overrides conflict due to
+inheritance. For example, the following code is accepted, but would be rejected
+(see before) if the method \Q@m@ was not redefined in \texttt{D3} and
+\texttt{E3}.
+\begin{lstlisting}
+interface D3 extends A2,B2 { int m(); } //accepted
+interface E3 extends A2,C2 { default int m(){return 42;} } //accepted
+\end{lstlisting}
+
+\subsection{Classifying Outcomes of Interface Composition}
+%When interfaces are composed and methods with the same name (and
+%signature) exist, there are 3 possible outcomes.
+%
+We now try to classify possible outcomes for composition of methods with the same name (and
+signature).
+We will use the following (correct) declarations:
+\begin{lstlisting}
+interface A1 { T m(); }
+interface A2 extends A1 { default T m(){ ... } }
+interface A3 extends A2 { T m(); }
+
+interface B1 { default T m(){ ... } }
+interface B2 extends B1 { T m(); }
+interface B3 extends B2 { default T m(){ ... } }
+\end{lstlisting}
+
+\noindent
+What happens if a new interface \Q@M@ extends one \Q@A@${}_i$ and one
+\Q@B@${}_j$?
+% Nine representative cases are shown next:\\*
+\begin{figure*}[htbp]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\textbf{M extends} & \textbf{A1}                  & \textbf{A2} & \textbf{A3} \\ \hline
+\textbf{B1}        & conservative error                     & conflict error      & conservative error       \\ \hline
+\textbf{B2}        & both abstract (accepted)     & conservative error       & both abstract (accepted)       \\ \hline
+\textbf{B3}        & \textbf{conservative error} &conflict  error       & conservative error      \\ \hline
+\end{tabular}
+\end{figure*}
+%We try to classify the results in the table:
+\begin{itemize}
+\item \textbf{conflict error} happens when the methods from both interfaces are implemented, which is also an error in most trait models.
+\item \textbf{both abstract (accepted)} happens when the methods from both interfaces are abstract, which is also considered correct in all
+  trait models.
+\item \textbf{conservative error} happens when only one method is implemented
+  (leaving another one abstract), which is different from what we would expect in
+  a trait model, but is coherent with the conservative idea that a method
+  defined in an interface should not silently satisfy a method in another one.
+\end{itemize}
+
+\paragraph{A bug:} During our experimentation, we found a bug in ECJ (Eclipse compiler for Java):
+the case \textbf{M} extending \textbf{B3} and \textbf{A1} is accepted by
+ECJ4.5.1 and rejected by javac.  By email communication with Brian Goetz
+(leading Java 8 designer) we have confirmed that the expected behavior is
+rejection, hence this is a bug in ECJ. This bug was also reported by
+  others and is fixed in the ECJ developer branch, but not released as a stable
+  version yet.
 

### 2016-03-22 OOPSLA16/sections/RelatedWork.tex
Some related work reorganization

--- 2015-12-09 paper/sections/RelatedWork.tex
@@
 
-\emph{Traits vs Object Interfaces.}
+\paragraph{Traits vs Object Interfaces.}
 We consider object interfaces to be an alternative to traits or mixins.  In the
@@
 constructors adds expressivity, which is not available in approaches
-such as Scala's traits/mixins.
+such as Scala's traits/mixins.\footnote{Scala traits have
+  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
+  ``pre-initialized''.}
 
@@
 %in future work (see Section~\ref{sec:futurework}).
+
+% \footnote{The original
+%  notion of traits by Scharli et al. prescribes, among other things,
+%  that: 1) a trait provides a set of methods that implement behavior;
+%  and 2) a trait does not specify any state variables, so the methods
+%  provided by traits do not access state variables directly. Java 8
+%  interfaces follow similar principles too. Indeed, a detailed
+%  description of how to emulate trait-oriented programming in Java 8
+%  can be found in the work by Bono et al.~\cite{bono14}. The Java 8
+%  team designing default methods, was also fully aware of that
+%  secondary use of interfaces, but it was not their objective to model
+%  traits: ``The key goal of adding default methods to Java was
+%  "interface evolution", not "poor man's
+%  traits"''~\cite{goetz13default}. As a result, they were happy to
+%  support the secondary use of interfaces with default methods as long
+%  as it did not make the implementation and language more complex.}
+\marcoT{We could use Scala Traits to encode IB, but the simpler Java8
+  model is a better starting point: scala traits can declare \Q@def, val, var@ and
+  \Q@lazy val@ However, while \Q@def@s can be overridden by
+  \Q@val@/\Q@var@s, \Q@val@/\Q@var@s can not be overridden.
+   A\Q@val@/\Q@var@ member is predestined to be
+  instance state, and this is in sharp contrast with our model where
+  the state is only a private detail of instances, and do not
+  contaminate interface/trait/classes.
+  }
 
@@
 \end{comment}
-
-\paragraph{Meta-Programming Competes with Language Extensions}
-The most obvious solution to adding features to a language is via syntactic
-extensions. Syntactic extensions are often implemented as
-desugarings to the base language. For example,
-the Scala compiler was extended to directly support XML syntax.
-However, when syntactic extensions are independently created it is hard
-to combine multiple extensions into one.
-SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language that
-aims at making the creation and composition of syntactic sugar
-extensions easy, by allowing programmers to extend Java with custom
-features (typically for DSLs). However SugarJ's goals are different
-from language tuning: SugarJ aims at creating and composing new syntax;
-whereas language tuning merely reinterprets existing syntax. It is
-clear that reinterpreting existing syntax only can be limiting
-for some applications. However, when language tuning is possible
-it has the advantage that existing tools for the language work
-out-of-the-box (since the syntax is still the same); and composition
-of independently developed tunings is straightforward.
-
-Scala-Virtualized~\cite{Rompf2012scalavirtualized} is an extension to Scala,
-which allows blending shallow and deep embedding of DSLs. It redefines
-some of Scala's language constructs to method calls,
-which can be overridden by DSL implementer. Thus Scala-Virtualized
-can also reinterpret syntax, and be seen as a form of language tuning.
-However, although many Scala's language constructs are supported,
-not all language constructs can be virtualized.
-
-%These extensions are admittedly useful, however they are essentially
-%creating a fork of the language, and rarely the new fork gains enough
-%traction to become the main language release. % On this
-% topic we mention
-
-When the base language has a flexible enough syntax and a fast and
-powerful enough reflection mechanism, we may just need to play with
-operator overloading and other language tricks to discover that the
-language feature we need can be expressed as a simple library in our
-language. An example of this is SQLAlchemy~\cite{bayer2012sqlalchemy} in
-Python, which uses operator overloading to dynamically turn normal python expressions into database queries
-without requiring any syntactic extensions to Python.
-Java-like languages tend to sit in the middle of two extremes:
-libraries can not influence the type system, so many solutions valid
-in Python or other dynamic languages are not applicable, or have the
-cost of losing type-safety.
-
-In Java-like languages compile time code generation comes at the rescue: if, for a certain feature
-(\mixin in our case), it is possible to use the original language syntax to
-\emph{express/describe} any specific instantiation of such feature (annotating a
-class and providing getters), then we can insert in the compilation process a tool
-that examines and enriches the code before compilation. No need to modify the
-original source (for example we can work on temporary files).
-Java is a particular
-good candidate for this kind of manipulation since it already provide ways to
-define and integrate such tools in its own compilation process via annotation
-processing. In this way
-there is no need of temporary files, and there is a well-defined way of putting
-multiple extensions together.
-
-Other languages offer even stronger support for safe code manipulation:
-Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
-all allow to execute code at compile time. They generate code that is
-transparently integrated in the program that is being
-generated/processed/compiled. In particular, MetaFjig offers a property called
-\emph{meta-level-soundness}, ensuring by construction that library code (even if wrong or unreasonable) never generates ill-typed code. This is roughly equivalent to
-\textit{Self coherence}, that we have to manually prove.
-Since MetaFjig is not working on annotated classes, there is not a ``semantic with/without annotations''. Our \mixin tuning theorem does not make sense in such context.
-
-\paragraph{Formalization of Java 8}
-We provide a simple formalization for a subset of Java including
-default/static interface methods and object initialization literals
-(often called anonymous local inner classes).  A similar formalization
-was drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize
-defender (default) methods in Java. In their formalization, classes
-and interfaces can have only one method \Q@m()@ without arguments, so
-as to simplify method overloading and renaming. Classless Java is more
-general, as it supports multiple methods with arguments, it supports
-static methods, and features such as multiple inheritance of
-interfaces and reabstraction of default methods are also modeled.

### 2016-03-22 OOPSLA16/sections/FutureWork.tex
Some reorganization, and some sad sacrifice to save some space

@@
   /*just the exposed interface*/
-  static ExposedData of(...){return InternalData.of(...);}
-  }
+  static ExposedData of(...){return InternalData.of(...);}  }
 \end{lstlisting}
@@
 
-\paragraph{A real IB language}
-In this article we presented the concept of IB as a programming pattern over Java.
-We do not wish to define here a complete IB language, just to show the reader IB as
-a promising way to design a new oo language.
-The first task is to define the set of abstract state operations.
-Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
-clean function programming style that would be otherwise too cumbersome.
-
-Currently we are offering both setters and Fluent setters.
-We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
-
-Manually defining abstract state methods can be verbose, we could offer a convenient syntax to 
-declare get/set/with.
-For example
-\begin{lstlisting}
-obj interface Foo{
-  String bar;//== String bar(); Foo withBar(String val);
-  var String beer;//also fluent setter
-  }
-\end{lstlisting}
-
-We may also want to offer a compact syntax for property updaters and generalized with methods;
-\begin{lstlisting}
-obj interface Foo{
-  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
-  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
-  }
-\end{lstlisting}
-Finally, certain methods may have a specific meaning, for example
-if a private method called \Q@init@ is present, it will be called after construction,
-and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
-called in some order.
+%\paragraph{A real IB language}
+%In this article we presented the concept of IB as a programming pattern over Java.
+%We do not wish to define here a complete IB language, just to show the reader IB as
+%a promising way to design a new oo language.
+%The first task is to define the set of abstract state operations.
+%Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
+%clean function programming style that would be otherwise too cumbersome.
+%
+%Currently we are offering both setters and Fluent setters.
+%We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
+%
+%Manually defining abstract state methods can be verbose, we could offer a convenient syntax to 
+%declare get/set/with.
+%For example
+%\begin{lstlisting}
+%obj interface Foo{
+%  String bar;//== String bar(); Foo withBar(String val);
+%  var String beer;//also fluent setter
+%  }
+%\end{lstlisting}
+%
+%We may also want to offer a compact syntax for property updaters and generalized with methods;
+%\begin{lstlisting}
+%obj interface Foo{
+%  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
+%  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
+%  }
+%\end{lstlisting}
+%Finally, certain methods may have a specific meaning, for example
+%if a private method called \Q@init@ is present, it will be called after construction,
+%and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
+%called in some order.

### 2016-03-22 OOPSLA16/sections/Implementation.tex
Some reorganization, and some sad sacrifice to save some space

@@
 
-\paragraph{Comparison with other Lombok annotations}\marco{if we keep it, will go in related work}
-The Lombok project provides a set of predefined annotations, including constructor
-generators similar as ours (e.g., \Q:@NoArgsConstructor:,
-\Q:@RequiredArgsConstructor: and \Q:@AllArgsConstructor:). They
-generate various kinds of constructors for \emph{classes}, with or without
-constructor arguments. This set of annotations is of great use, especially when
-used together with other features provided in Lombok (e.g.,
-\Q:@Data:). Moreover, the implementation of these annotations in Lombok
-gives us hints on how to implement \mixin. However, none of these annotations
-can model what we are doing with \mixin - generating constructor-methods
-(\textbf{of}) for \emph{interfaces}. Apart from constructors, \mixin also
-provides other convenient features (including generating fluent setters, type
-refinement, etc), which the base Lombok project does not provide.
-Finally, while \mixin is formalized, none of Lombok's annotations have been
-studied in a formal way.
-
 %\paragraph{Lombok does language tuning}

### 2016-03-22 OOPSLA16/sections/Introduction.tex
Some reorganization, and some sad sacrifice to save some space

@@
 We believe that such verbosity hindered and slowed down
-the discovery of useful programming pattens involving type-refinement\cite{ trivial solution to the EP}.
+the discovery of useful programming pattens involving type-refinement\cite{trivialsolutiontotheEP}.
 }

### 2016-03-22 OOPSLA16/sections/RelatedWork.tex
Some reorganization, and some sad sacrifice to save some space

@@
 \begin{comment}
+
+\paragraph{Comparison with other Lombok annotations}\marco{if we keep it, will go in related work}
+The Lombok project provides a set of predefined annotations, including constructor
+generators similar as ours (e.g., \Q:@NoArgsConstructor:,
+\Q:@RequiredArgsConstructor: and \Q:@AllArgsConstructor:). They
+generate various kinds of constructors for \emph{classes}, with or without
+constructor arguments. This set of annotations is of great use, especially when
+used together with other features provided in Lombok (e.g.,
+\Q:@Data:). Moreover, the implementation of these annotations in Lombok
+gives us hints on how to implement \mixin. However, none of these annotations
+can model what we are doing with \mixin - generating constructor-methods
+(\textbf{of}) for \emph{interfaces}. Apart from constructors, \mixin also
+provides other convenient features (including generating fluent setters, type
+refinement, etc), which the base Lombok project does not provide.
+Finally, while \mixin is formalized, none of Lombok's annotations have been
+studied in a formal way.
+
 \subsection{Type-Directed Translations/Syntactic Sugar}

### 2016-03-22 OOPSLA16/sections/Implementation.tex
Limitations.

@@
 \subsection{Limitations}\marco{ to reorganize with updated limitations}
+\haoyuan{Should we only present the separate compilation and simple generics features in this (limitation) section?}
 Our prototype implementation using Lombok has certain limitations:
 \begin{itemize}
-\item The prototype does not support separate compilation yet. Currently all
+\item \haoyuan{The annotation \mixin does not support separate compilation yet. Using \mixin, all
   related interfaces have to appear in a single Java file. Therefore, changes to
-  a single interface would require re-compiling the whole file. This compilation
-  limitation is not caused by our algorithm. It is a Lombok implementation related
-  issue: in Lombok it is hard to capture a type declaration from its reference,
-  even harder when the type declaration is in other files (we have not found a
-  way to do this yet).
+  a single interface would require re-compiling the whole file. Nevertheless, to support separate compilation we
+  studied an existing yet experimental feature \lstinline{@Delegate}, reused the name and modified its behavior.
+  Now files in the same package can be compiled and \lstinline{@Delegate} has the same behavior as \mixin. A limitation is
+  that users need to put references to the super types of the annotated type together with the annotation, for instance,
+  \lstinline{@Delegate(types = Point2D.class)}. Another limitation is the methods generated by \lstinline{@Delegate} will
+  not be visible in the Outline window, but they can still be auto-completed in Eclipse.}
 \item At this stage our implementation only realizes the Eclipse handler and our
@@
   \texttt{javac} is missing.
-\item The current implementation does not take type-parameters into
-  consideration, thus it does not support generics yet.
+\item \haoyuan{The current implementation only supports simple generics, which means type parameters can be used, however, the generic
+method typing is neither formalized in this paper nor implemented, but delegated to the Java compiler.}
 \end{itemize}

### 2016-03-22 OOPSLA16/sections/CaseStudy.tex
little fixes

@@
 \section{Case Studies}
-In this section we conduct three case studies which reveal various
+In this section we conduct \numOfCaseStudies case studies which reveal various
 advantages of using the \mixin annotation. The first case study provides a
@@
 \subsection{A Maze Game}
-The last case study is a simplified variant of a Maze game, which is often
+This case study is a simplified variant of a Maze game, which is often
 used~\cite{gof,bono14}
@@
 \subsection{Refactoring a Compiler}
-This case study applies our approach to a slightly larger code base of a simple
+The last case study applies our approach to a slightly larger code base of a simple
 one-pass compiler for \lstinline{C0} (a subset of \lstinline{C}). This compiler
@@
 code becomes more modular without the sacrifice of code amount/simplicity.
+\marco{Why we do not cite the result of the former study? the one with object algebraes
+where the number of lines grown quite a while?}

### 2016-03-22 OOPSLA16/sections/Introduction.tex

@@
 To evaluate the usefulness of object interfaces, we illustrate
-3\bruno{needs updates}
+\numOfCaseStudies \bruno{needs updates}\yanlin{updated} 
 applications. The first application is a simple 
@@
 application shows how embedded DSLs using fluent interfaces~\cite{fowler2005fluentinterface} 
-can be easily defined using object interfaces. The last
-application is a larger case study for a simple Maze game implemented with 
-multiple inheritance. For the last application we show that there is a
+can be easily defined using object interfaces. The third
+application is a case study for a simple Maze game implemented with 
+multiple inheritance. For this application we show that there is a
 significant reduction in the numbers of lines of code when compared 
-to an existing implementation~\cite{bono14} using plain Java 8. 
-Noteworthy, all applications are implemented 
+to an existing implementation~\cite{bono14} using plain Java 8. The last
+application is a larger case study of refactoring of a simple one-pass compiler, showing that our
+approach can scale to real code base. Noteworthy, all applications are implemented 
 without defining a single class!

### 2016-03-22 OOPSLA16/sections/Implementation.tex
Label in figure.
Figure 5 fixed.

@@
 annotations. Classless Java was created using Lombok.
-Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
+Figure~\ref{fig:lombok} illustrates the
 flow of \mixin annotation processing in Lombok.\bruno{Haoyuan, please
@@
 and explain figure here.} This includes all
-the methods inside the interface as well as the generated ones. \haoyuan{In Figure~\ref{fig:screenshot} (wrong),
-the screenshot presents the Outline window of Eclipse, where the generated methods including \lstinline{Point2D.of}, \lstinline{Point3D.of},
-\lstinline{Point3D.withX} and \lstinline{Point3D.withY} are visible to users, showing their types and modifiers.}
-This also means that the IDE functionality for content assist and
-autocomplete will work for the newly generated methods.
+the methods inside the interface as well as the generated ones. \haoyuan{In Figure~\ref{fig:screenshot},
+\mixin generates \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@. These
+methods are also visible to users, showing their types and modifiers in Outline window. Moreover, as a useful IDE feature,
+the auto-completion also works for these newly generated methods.}
 
-\begin{figure}[h]\label{fig:screenshot}
+\begin{figure}[h]
 \centering
-\includegraphics[width=3.3in]{pdfs/screenshot2.png}
-\caption{Generated methods shown in the Outline window of Eclipse.}
+\includegraphics[width=3.3in]{pdfs/screenshot4.png}
+\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
+\label{fig:screenshot}
 \end{figure}

### 2016-03-22 OOPSLA16/sections/FutureWork.tex
Text in Section 4. Code style.

@@
 \begin{lstlisting}
-@Obj interface InternalData implements ExposedData{/*all the methods*/}//package visibility
-public interface ExposedData{
+@Obj interface InternalData implements ExposedData {
+    /*all the methods*/} //package visibility
+public interface ExposedData {
   /*just the exposed interface*/
-  static ExposedData of(...){return InternalData.of(...);}  }
+  static ExposedData of(...) {
+    return InternalData.of(...);}  }
 \end{lstlisting}
@@
         return new Point() {...
-            public Point clone() { return Point.super.clone();}};} }
+            public Point clone() {
+                return Point.super.clone();}};} }
 \end{lstlisting}

### 2016-03-22 UseMixinLombok/src/casestudy/ep/TestExpression.java

@@ -45,8 +45,10 @@ interface Exp { int eval(); }
 //END_EXPRESSION_INIT
 
 //BEGIN_EXPRESSION_SUB
-@Obj interface Sub extends Exp { Exp e1(); Exp e2();
-    default int eval() {return e1().eval() - e2().eval();} }
+@Obj interface Sub extends Exp {
+	Exp e1(); Exp e2();
+	default int eval() {
+		return e1().eval() - e2().eval();} }
 //END_EXPRESSION_SUB
 
 //BEGIN_EXPRESSION_PRINT

### 2016-03-23 OOPSLA16/sections/CaseStudy.tex

@@
 &
-\lstinputlisting[linerange=53-63]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=55-65]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
@@
 
-\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=48-51]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 \begin{lstlisting}[]
-interface ExpC extends Exp { List<Integer> collectLit(); }
+interface ExpC extends Exp {
+    List<Integer> collectLit(); }
 @Obj interface LitC extends Lit, ExpC {...}
-@Obj interface AddC extends Add, ExpC {ExpC e1(); ExpC e2(); ...}
+@Obj interface AddC extends Add, ExpC {
+    ExpC e1(); ExpC e2(); ...}
 \end{lstlisting}
@@
 
-\lstinputlisting[linerange=87-89]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=89-91]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 
@@
 
-\lstinputlisting[linerange=31-31]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
+\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1
 
@@
 \texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
-\lstinputlisting[linerange=14-16]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
+\lstinputlisting[linerange=14-18]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
 In this way, when a query created using \texttt{ExtendedDatabase},
@@
 
-\lstinputlisting[linerange=37-38]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
 

### 2016-03-23 OOPSLA16/sections/Introduction.tex

@@
 
-\marcoT{Object oriented languages always strive to offer great code reuse.}
+Object oriented languages strive to offer great code reuse.
 They aim to couple flexibility and rigour, expressive power and
@@
 
-We present here a third alternative: the concept of \textbf{Interface
-  based} (IB) object oriented languages, where objects implement
-interfaces directly. In IB interfaces own the implementation for the behaviour, which
-is structurally defined in their interface.
+This paper presents a third alternative: the concept of
+\textbf{Interface based} (IB) object oriented languages, where objects
+implement interfaces directly. In IB interfaces own the implementation
+for the behaviour, which is structurally defined in their
+interface. Programmers do not define objects directly, but delegate
+the task to \emph{object interfaces}, whose role is similar to non
+abstract classes in CB languages. Instantiation of objects is
+done using static factory methods in object interfaces.
 
-\marcoT{
-A key challenge lies in how to model state, which is fundamental to
-have stateful objects. 
-Programmers do not define objects directly, but
-delegate the task to \emph{object interfaces}, whose role is similar
-to non abstract classes in CB languages.
-All abstract operations in an object interface will be interpreted as \emph{abstract state operations}. The
-abstract state operations include various common utility methods (such
-as getters and setters, or clone-like methods). In the presence of
-subtyping, such operations often require special care, as their types
-need to be refined. Object interfaces provide support for
-type-refinement and can automatically produce code that deals with
-type-refinement adequately, while CB (as in Java, Scala and C\#) require 
-verbose explicit type-refinement.
-We believe that such verbosity hindered and slowed down
-the discovery of useful programming pattens involving type-refinement\cite{trivialsolutiontotheEP}.
-}
+A key challenge in IB programming lies in how to model state, which is
+fundamental to have stateful objects. All abstract operations in an
+object interface are interpreted as \emph{abstract state
+  operations}. The abstract state operations include various common
+utility methods (such as getters and setters, or clone-like
+methods). Objects are the only responsible to define the ultimate
+behaviour of a method and, for example, if such method is just a
+setter. Anything related to state is completely contained in the
+instances and does not leak into the inheritance logic.  In CB
+programming, the structure of the state is fixed and can be only grown
+by inheritance.  In contrast, in IB programming the state is never
+fixed, and methods that ``looks like'' abstract setters and getters
+can always receive an explicit implementation down in the inheritance
+chain, improving \textbf{modularity and flexibility}.  That is, the
+concept of abstract state is more fluid.
 
-Finally, object interfaces also provide static factory methods instantiating objects.
+In the presence of subtyping, abstract state operations often require
+special care, as their types need to be refined. Object interfaces
+provide support for type-refinement and can automatically produce code
+that deals with type-refinement adequately. In contrast, typical CB
+languages, such as Java or Scala, require verbose explicit
+type-refinement. We believe that such verbosity hindered and slowed
+down the discovery of useful programming pattens involving
+type-refinement. In particular, a recently discovered
+solution~\cite{trivialsolutiontotheEP} to the Expression
+Problem~\cite{} in Java-like languages, shows how easy it is to solve
+the problem using only type-refinement. However it took nearly 20
+years since the formulation of the problem for that
+solution to be presented in the literature. In IB programming, due to
+automatic type-refinement, that solution should have been obvious from
+day one.
 
+One important advantage of the use of abstract state operations and
+automatic type-refinement is that it offers a new form of
+\emph{type-safe covariant mutable state}. That is, in IB programming,
+it is possible to refine \emph{mutable} ``fields'' in subtypes. This is
+typically forbidden in CB programming languages (including Java and
+Scala), since it is widely known that \emph{naive} type-refinement of
+mutable fields is not type-safe. While covariant refinement of mutable
+fields is supported by some type systems~\cite{}, this requires
+significant complexity and restrictions to ensure that all uses of
+covariant state are indeed type-safe.
+
+\begin{comment}
 \marcoT{%
-Objects are the only responsible to define the
-ultimate behaviour of a method and, for example, if such method is
-just a setter. Anything related to state is completely contained in
-the instances and do not leak in the inheritance logic.
-In CB, the structure of the state is fixed and can be only
-grown by inheritance; on the opposite, in IB the state is never fixed,
-and methods that ``looks like'' abstract setters and getters
-can always receive an explicit implementation down in the inheritance chain,
-improving \textbf{modularity and flexibility}.
 
-That is, the concept of abstract state is more fluid.
 In this paper we show how to support type-safe
@@
 and \textbf{multiple inheritance}.}
+\end{comment}
 
-Indeed the literature
-provides good examples on how easy and modular it is to combine
-multiple sources of pure behaviour, using mechanisms such as
-traits~\cite{scharli03traits}. 
-In Java multiple \emph{interface} inheritance has been
-supported since inception, and in Java8 default methods~\cite{} bring some of
-the advantages of traits into Java. In contrast, the literature~\cite{} is
-also rich on how hard it is to modularly combine multiple sources of
-behaviour \textbf{and} state with multiple \emph{implementation}
-inheritance of classes. 
+Finally, another advantage of IB programming is its support for
+multiple inheritance. The literature provides good examples on how
+easy and modular it is to combine multiple sources of pure behaviour,
+using mechanisms such as traits~\cite{scharli03traits}. In Java
+multiple \emph{interface} inheritance has been supported since
+inception, and in Java8 default methods~\cite{} bring some of the
+advantages of traits into Java. In contrast, the literature~\cite{}
+is also rich on how hard it is to modularly combine multiple sources
+of behaviour \textbf{and} state with multiple \emph{implementation}
+inheritance of classes. In IB programming there is only multiple
+interface inheritance, yet programmers can still use state via the
+abstract state operations. IB programming enables powerful 
+idioms using multiple inheritance and state.
+
 %Traits or Java8 interfaces still assume a CB model: 
@@
 
-\bruno{To be mentioned: 3 features of IB: Interface-based programming with multiple
-  inheritance; automatic type-refinements; a new approach that allows 
-\emph{type-safe} covariant refinement of state.
-}
-
-\bruno{Not sure if
-  this is totally clear. Implicit here is the fact that objects are
-  never defined explicitly, I think. And the behaviour is introduced 
-implicitly.}
-\marco{we need to discuss, not clear what you mean}
-%\footnote{ In the same sense,
-%  you can encode prototype based programming over class based with the
-%  strategy pattern, and class based programming with prototypes by
-%  creating ``class'' objects.}
-
 IB could be explained by defining a novel language, with new syntax
@@
 \emph{default methods}, which allow interfaces to have method
-implementations.
+implementations. In the context of Java, what we propose is a programming
+style, where we never use classes\footnote{More precisely, we never
+  use the \Q@class@ keyword.}.  We call this restricted version of
+Java \emph{Classless Java}.
 
@@
 
-In the context of Java, what we propose is essentially a programming
-style, where we never use classes\footnote{More precisely, we never
-  use the \Q@class@ keyword.}.  We call this restricted version of
-Java \emph{Classless Java}.
-
   %With object interfaces, many Java programs can be built
@@
 a real IB language, and the annotation processor allows us to
-transparently hide it from Java programmers. 
-The implementation works by performing \emph{on-the-fly}\marco{False, we
-generate the instrumented bytecode. on the fly ast rewriting is a term that means 
-that you modify the bytecode while you load it} AST
-rewriting, allowing existing Java tools (such as IDEs) to work
-out-of-the-box with our implementation. Moreover, the 
+transparently hide it from Java programmers.  The implementation works
+by performing AST rewriting, allowing most existing Java tools (such as
+IDEs) to work out-of-the-box with our implementation. Moreover, the
 implementation blends Java's conventional CB style and IB smoothly.
-As a result, we experiment object interfaces with several interesting Java programs,
-and conduct various case studies. 
-Finally, we also formally define the behaviour of our \Q|Obj| annotation.
+As a result, we experiment object interfaces with several interesting
+Java programs, and conduct various case studies.  Finally, we also
+formally define the behaviour of our \Q|Obj| annotation and discuss
+its properties.
 
@@
 
-We can extend our approach to support encapsulation and privateness;
-various possibilities are discussed in the last part of the paper.
+While the Java embedding has obvious advantages from the practical
+point-of-view, it also imposes some limitations that a new IB language
+would not have. In particular, supporting proper encapsulation is
+difficult in Java due to limitations of Java interfaces. We discuss
+these limitations, possible workarounds, and native language support in Section~\ref{}.
 
@@
 \item {\bf IB and Object Interfaces:} a novel take on object orientation, allowing
-  powerful multiple-inheritance programming idioms to be expressed
-  conveniently in Java. We provide examples, informal and formal description.
+  powerful programming idioms using multiple-inheritance, 
+  type-refinement and abstract state operations.
 
-\item{\bf Type preservation guarantees:}\bruno{needs rephrasing}
-We discuss our formalization of a subset of Java8 type system and how we use this
-to characterize safety properties about our annotations.
+\item {\bf Classless Java:} A practical realization of IB in
+  Java. Classless Java is implemented using annotation processing, 
+  allowing most tools to work transparently with our approach. 
+  We provide examples, informal and formal descriptions of Classless
+  Java.
+
+\item {\bf Type-Safe Covariant Mutable State:} We show how the 
+ combination of abstract state operations and type-refinement enables 
+ a form of mutable fields that can be covariantly refined in a type-safe way.
+
+%\item{\bf Type preservation guarantees:}\bruno{needs rephrasing}
+%We discuss our formalization of a subset of Java8 type system and how we use this
+%to characterize safety properties about our annotations.
 %\item {\bf Classless Java (CJ):} A simple formal calculus that models 
@@
 
-\item {\bf Implementation and Case Studies:} We have a prototype
-  implementation of object interfaces, using Java
-  annotations and AST rewriting. Moreover, the usefulness of object interfaces is
-  illustrated through various examples and case studies.
+\item {\bf Case Studies:} We illustrate the usefulness of IB
+  through various examples and case studies.
 

### 2016-03-23 OOPSLA16/sections/Translation.tex

@@
   is explained here. }
-The formal definition is available in
-Appendix~\ref{subsec:translation}.\bruno{what formal definition?}
+\haoyuan{In Figure~\ref{figure:valid}, the $\otherMethod$ function is defined in the first two
+equations. This function generates unimplemented \Q@with-@ and fluent setters in the interface.
+They are generated as abstract methods because the return types need to be refined.
+To determine whether the methods need to be generated,
+we check if such \Q@with-@ or setter methods
+are required by $\C_0$, but not declared directly in $\C_0$. The third equation gives the definition
+of $\valid$. To generate the \Q@of@ method, it is necessary for $\weakAnn$ to explicitly check if all
+abstract methods are valid. That is, we can categorize them in a pattern that we
+know how to implement: it is either a field getter, a with method or a setter. Auxiliary checkers are
+defined on the right side. Note that we write $\QM{with#}\m$ to append $\m$ to
+\QM{with}, following the camelCase rule. The first letter of $\m$
+must be lower-case and is changed to upper-case upon appending. For
+example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
+are \QM{with} and all identifiers of the form $\QM{with#}\m$.
+}
 
-The function $\ofMethod$ generates the static method $\QM{of}$, as an
+In Figure~\ref{figure:ofmethod}, the function $\ofMethod$ generates the static method $\QM{of}$, as an
 object factory. It detects all the field methods of $\C_0$, take them
@@
   probably text which is now in the appendix. Haoyuan, please take
-  care of this.}
+  care of this.}\haoyuan{fixed}
 

### 2016-03-23 OOPSLA16/sections/Introduction.tex
references

@@
 modular reasoning.  Two main ideas have emerged to this end: prototype
-based (PB)~\cite{} and class based (CB) languages~\cite{}.  In prototype based
+based (PB)~\cite{Ungar87self} and class based (CB) languages such as
+Java or Scala.  In prototype based
 languages objects inherit from other objects, and thus objects own
@@
 type-refinement. In particular, a recently discovered
-solution~\cite{trivialsolutiontotheEP} to the Expression
-Problem~\cite{} in Java-like languages, shows how easy it is to solve
+solution~\cite{eptrivially} to the Expression
+Problem~\cite{wadler98expression} in Java-like languages, shows how easy it is to solve
 the problem using only type-refinement. However it took nearly 20
@@
 mutable fields is not type-safe. While covariant refinement of mutable
-fields is supported by some type systems~\cite{}, this requires
+fields is supported by some type systems~\cite{bruce98astatically,McDirmid01jiazzi,ernst01family,ernst06virtual}, this requires
 significant complexity and restrictions to ensure that all uses of
@@
 multiple \emph{interface} inheritance has been supported since
-inception, and in Java8 default methods~\cite{} bring some of the
-advantages of traits into Java. In contrast, the literature~\cite{}
+inception, and in Java8 default methods~\cite{goetz12fdefenders} bring some of the
+advantages of traits into Java. In contrast, the literature~\cite{Sak89dis,bracha90mixin,malayeri2009cz}
 is also rich on how hard it is to modularly combine multiple sources
@@
 difficult in Java due to limitations of Java interfaces. We discuss
-these limitations, possible workarounds, and native language support in Section~\ref{}.
+these limitations, possible workarounds, and native language support in Section~\ref{sec:futurework}.
 

### 2016-03-23 OOPSLA16/paper.tex
Move formalization to appendix.
Some edits and comments
Added missing appendix
Numeric citation style.
Some work in introduction
Content reorganization
Reviewd red H text, plus some fix to discussions
Changes
Polishing
Section 2
Some related work reorganization
Some reorganization, and some sad sacrifice to save some space
abstract

--- 2016-01-26 OOPSLA16/paper.tex
@@
-\documentclass[10pt,preprint]{sigplanconf}
+\documentclass[10pt,preprint,nocopyrightspace,numbers]{sigplanconf}
 
@@
 
-\usepackage{amsmath,amssymb}
-\usepackage{balance,bbm,booktabs}
+\usepackage{listings}
+\usepackage{multicol}
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+\usepackage[table,xcdraw]{xcolor}
 \usepackage{color}
-\usepackage{etex}
+\usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{stmaryrd}
+\usepackage{graphicx}
+\usepackage{amssymb}
 \usepackage{fancyvrb}
+\usepackage{url}
+\usepackage{pstricks,pst-node,pst-tree}
+\usepackage{bbm}
+\usepackage{pgf}
+\usepackage{multirow}
+
+\usepackage{listings}
+\usepackage{verbatim}
 \usepackage{graphicx}
-\usepackage{listings}
-\usepackage{mathpartir,mathptmx,mdwlist,morefloats,multirow}
-\usepackage{pstricks,pst-node,pst-tree,pgf}
-\usepackage{rotating}
-\usepackage{stmaryrd}
-\usepackage{theorem}
-\usepackage{url}
-\usepackage{verbatim}
 \usepackage{wrapfig}
+\usepackage[normalem]{ulem}
+
+\usepackage[T1]{fontenc}
+\usepackage[scaled=0.85]{beramono}
+\usepackage{mathpartir}
 \usepackage{xspace}
 
-\usepackage[normalem]{ulem}
-\usepackage[T1]{fontenc}
-\usepackage[scaled=0.85]{beramono}
-\usepackage{dirtree}
+\input{predef/def.tex}
+\input{predef/java.tex}
 
@@
 \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
+\newcommand{\authorText}[2]{{\color{#1}#2}}
 \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
+\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
+\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
+\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
+\newcommand\brunoT[1]{\authorText{red}{#1}}
+\newcommand\yanlinT[1]{\authorText{purple}{#1}}
+\newcommand\marcoT[1]{\authorText{blue}{#1}}
+\newcommand\haoyuanT[1]{\authorText{cyan}{#1}}
+
+\newcommand\delete[1]{\textcolor{red}{\sout{#1}}}
+
+\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
+\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
+\newcommand\tsem[1]{\llbracket #1 \rrbracket}
+\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
+\newcommand\nat[0]{\mathbb{N}}
+\newcommand\unit[0]{\mathbbm{1}}
+
+\newcommand\mixin{\mixinAnn\xspace}
+%\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
+\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
+
+%\theoremstyle{plain}
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}{Lemma}
+\newtheorem{thm2}{Theorem}
+\newtheorem{lem2}{Lemma}
 
@@
 
-\title{Title Text}
-\subtitle{Subtitle Text, if any}
+\title{Classless Java}
+\subtitle{Interface Based Programming for the Masses}
 
-\authorinfo{Name1}
-           {Affiliation1}
-           {Email1}
-\authorinfo{Name2\and Name3}
-           {Affiliation2/3}
-           {Email2/3}
+\authorinfo{}
+           {}
+           {}
+%\authorinfo{Name2\and Name3}
+%           {Affiliation2/3}
+%           {Email2/3}
 
@@
 \begin{abstract}
-This is the text of the abstract.
+
+  This paper presents Interface Based (IB) object-oriented (OO)
+  programming. Unlike class based OO programming, in IB programming there
+  is no need for classes. To support state, IB programming does not
+  use fields directly. Instead it uses \emph{abstract state
+    operations}, such as getters, setters or clone-like methods. This
+  new way to deal with state allows flexibility not typically
+  available in CB languages.  In IB state (including mutable state)
+  can be automatically type-refined in subtypes. The combination of a
+  purely IB style of programming and automatic type-refinement enables
+  powerful idioms using multiple inheritance and state.  To introduce
+  IB to programmers we created Classless Java: an embeding of IB
+  directly into Java. Classless Java uses annotation processing for
+  code generation and relies on the new features of Java 8 for
+  interfaces. We illustrate the usefulness of IB through various
+  examples and case studies.
+
 \end{abstract}
 
+\begin{comment}
 \category{CR-number}{subcategory}{third-level}
@@
 keyword1, keyword2
+\end{comment}
 
@@
 
+\input{sections/Overview.tex}
+
+\input{sections/Implementation.tex}
+
 \input{sections/Formalization.tex}
 
-\appendix
-\section{Appendix Title}
+\input{sections/Translation.tex}
 
-This is the text of the appendix, if you need one.
+\input{sections/CaseStudy.tex}
 
-\acks
+\input{sections/FutureWork.tex}
 
-Acknowledgments, if needed.
+\input{sections/RelatedWork.tex}
 
+\input{sections/Conclusion.tex}
+
+%\acks
+%
 % We recommend abbrvnat bibliography style.
@@
 
-\begin{thebibliography}{}
-\softraggedright
+%\begin{thebibliography}{}
+%\softraggedright
 
-\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
-P. Q. Smith, and X. Y. Jones. ...reference text...
+%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
+%P. Q. Smith, and X. Y. Jones. ...reference text...
 
-\end{thebibliography}
+%\end{thebibliography}
+
+\appendix
+\input{sections/Appendix_Formalization.tex}
+\input{sections/Appendix_Translation.tex}
+\input{sections/Appendix.tex}
 

### 2016-03-23 OOPSLA16/sections/Introduction.tex
Minor fixes and typos

@@
 type-refinement. We believe that such verbosity hindered and slowed
-down the discovery of useful programming pattens involving
+down the discovery of useful programming patterns involving
 type-refinement. In particular, a recently discovered
@@
 solution to be presented in the literature. In IB programming, due to
-automatic type-refinement, that solution should have been obvious from
-day one.
+its emphasis on type-refinement, that solution should have been much
+more obvious.
 

### 2016-03-23 OOPSLA16/sections/Overview.tex
Section 3: flow chart, screenshot and text.
Section 2

@@
 This section illustrates how our programming style, supported by the
-\mixinAnn{} annotation, enable powerful programming idioms based on multiple
+\mixinAnn{} annotation, enables powerful programming idioms based on multiple
 inheritance and type refinements.  We propose a standard example:
@@
 \lstinputlisting[linerange=8-12]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
-\bruno{I dislike the code formating here.}
 \noindent The effect of the annotations is that a static \emph{factory} method called
@@
 Also, since \mixin interfaces offer the \Q@of@ factory method, only interfaces where all the abstract methods
-can be synthesized can be \mixin. A non \mixin interface is like an abstract class in Java.
+can be synthesized can be object interfaces. A non \mixin interface is like an abstract class in Java.
 %\end{comment}
@@
 
-\noindent Here we declare abstract getter and setter for the mutable ``field''
+\noindent Here we declare an abstract getter and setter for the mutable ``field''
 \Q@location@.  Without the \mixin annotation, there is no convenient way to
@@
 \lstinputlisting[linerange=71-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
-\bruno{Don't like the code format here either.}
 %Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
@@
 
- \marcoT{Symmetrically,
- we could offer an imperative property updater
- that calls the setters instead of the withers.
- \Q@Point3D set(Point2D val)@.
- }\bruno{we could or we do? if we could then the table should not
-   contain this operation; if we do the table is ok}
+%Symmetrically, we could offer an imperative property updater that
+%calls the setters instead of the withers.  \Q@Point3D set(Point2D
+%val)@.
 
 \begin{figure}
-\marco{I can not find the declaration int x();
-also we can change the order of the code to write the non generated stuff first.}
 %\saveSpaceFig
@@
 generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}.
-\marcoT{Overall, for the whole
-animals-with-locations, an @@@ lines code example,
-the generate/completed code is composed by
-@@@ lines.}
-
+%\marcoT{Overall, for the whole
+%animals-with-locations, an @@@ lines code example,
+%the generate/completed code is composed by
+%@@@ lines.}
 Writing such code by hand is error-prone. For
@@
 \Q@this@.
-
-\paragraph{Summary of operations in Classless Java}
-%Method bodies can only refer to other methods;
-%Objects are (conceptually) closures and the state is
-%composed by captured local variables.}
-%Methods interacting with state are called \emph{abstract state operations}.
-Object interfaces provide support for
-different types of abstract state operations: four field-based state
-operations; and  two types of
-general updating operations. Furthermore object interfaces support
-direct object intantiation via \texttt{of} factory methods.
-Figure~\ref{fig:abstractstate} summarizes the six operations supported
-by \mixin. The field-based abstract state operations are determined by
-naming conventions and the types of the methods. Fluent setters are a variant of
-conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.
 
@@
 & \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@
-& Update all matching fields in \texttt{val}.        \\
-\cline{2-4}
-& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ &
-Set all matching fields in \texttt{val}.        \\ \hline
+& Update all matching fields in \texttt{val}.        \\ \hline
+%\cline{2-4}
+%& \textbf{imperative updaters}        & \Q@Point3D set(Point2D val)@ &
+%Set all matching fields in \texttt{val}.        \\ \hline
 \end{tabular}
@@
 \end{figure*}
+
+\paragraph{Summary of operations in Classless Java}
+%Method bodies can only refer to other methods;
+%Objects are (conceptually) closures and the state is
+%composed by captured local variables.}
+%Methods interacting with state are called \emph{abstract state operations}.
+In summary, object interfaces provide support for
+different types of abstract state operations: four field-based state
+operations; and functional updaters. Furthermore object interfaces support
+direct object intantiation via \texttt{of} factory methods.
+Figure~\ref{fig:abstractstate} summarizes the six operations supported
+by \mixin. The field-based abstract state operations are determined by
+naming conventions and the types of the methods. Fluent setters are a variant of
+conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.
 

### 2016-03-23 OOPSLA16/sections/Implementation.tex
Ref to the blog. Didn't find the author name.

@@
 as Java. To have an approach that can both illustrate IB
-programming and be practical, we have instead implemented Classless Java.
-This section presents the Classless Java design.
+programming and be practical, we have instead implemented 
+\mixin as an annotation over Java8, and a \emph{compilation agent}.
+That is, the Classless Java style of programming
+is supported by a Java8 library.
+
 Disciplined use of Classless Java (that is avoiding class
 definitions as done in Section~\ref{}), illustrates what \emph{pure} IB is like.
-However in Classless Java CB and IB programming can be mixed together, and all
-the practical convenience of using existing Java libraries, the full
-Java language, and IDE's are available in Classless Java.
+However, using \mixin, CB and IB programming can be mixed together, 
+harvesting the practical convenience of using existing Java libraries, the full
+Java language, and IDE's support.
 The key to our implementation is the use of compilation agents, which
@@
 generation of the abstract syntax tree and the Java Bytecode.
-\marco{is this correct?}\bruno{my impression is that Lombok happens
-  \emph{before} the AST is compiled (not between AST and bytecode generation)}
+
 This process is facilitated by frameworks like Lombok~\cite{lombok}:
@@
 reducing) Java boilerplate code via
-annotations. Classless Java was created using Lombok.
-Figure~\ref{fig:lombok} illustrates the
-flow of \mixin annotation processing in Lombok.\bruno{Haoyuan, please
-  explain the diagram.}
-\haoyuan{Source code in a Java program is first parsed into its syntactic structure,
-which we call abstract syntax tree (AST). The AST is passed to Lombok, and a corresponding
-AST defined in Lombok is generated for the annotated node. An annotation like \mixin is declared
-inside Lombok, together with an implementation of its (Eclipse or Javac) handler. The handler is
+annotations. \mixin was created using Lombok.
+As you can see explained in Figure~\ref{fig:lombok}~\cite{neildo2011blog},
+first java source code is parsed into an abstract syntax tree (AST).
+The AST is then captured by Lombok:
+each annotated node is passed to
+the corresponding (Eclipse or Javac) handler. The handler is
 free to modify the information of the annotated node, or even inject new nodes (like methods, inner classes,
-etc). Finally, Java compiler works on the modified AST to generate bytecode. During the compilation,
-no source code is changed, and no new files are created, bytecode takes the only difference.}
+etc). Finally, the Java compiler works on the modified AST to generate bytecode.
+%Note how during the compilation,
+%no source code is changed, and no new source/temporary files are created.
 
@@
 The Lombok compilation agent has advantages with respect to alternatives like
-pre-processors, or other Java annotation processors. In some sense
-Lombok can be seen as a poor man's macro system~\cite{} for Java.
-\bruno{Marco, do you agree with calling Lombok a poor man's macro system?}
-\marco{No, it is as powerful as lisp macros, and have a bigger (may be even  a better) design.}
-%this seems like a commercial spot
-%\item Lombok is byte-code based instead of source-code based, which makes client
-%  code concise and easy to maintain. Such code generation is performed at
-%  compile time to modify bytecode.
+pre-processors, or other Java annotation processors.
+Lombok offers in Java an expressive power similar to that of Scala/Lisp macros;
+except of course, for the syntactic convenience of quote/unquote templating.
 
@@
 performed transparently and the information of the interface from
-compilation is captured in the ``Outline'' window.\bruno{Haoyuan: show
-and explain figure here.} This includes all
-the methods inside the interface as well as the generated ones. \haoyuan{In Figure~\ref{fig:screenshot},
-\mixin generates \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@. These
-methods are also visible to users, showing their types and modifiers in Outline window. Moreover, as a useful IDE feature,
-the auto-completion also works for these newly generated methods.}
+compilation is captured in the ``Outline'' window.
+This includes all
+the methods inside the interface as well as the generated ones.
+In Figure~\ref{fig:screenshot},
+\mixin generates \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@.
+These methods are also visible to users, showing their types and modifiers in Outline window.
+Moreover, as a useful IDE feature, the auto-completion also works for these newly generated methods.
 
@@
 
-\subsection{AST Reinterpretation in Classless Java}
+\subsection{\mixin AST Reinterpretation}
 
-Ofcourse, careless reinterpretation of the AST could still be
-surprising for badly designed rewritings.  Classless Java reinterprets
+Of course, careless reinterpretation of the AST could still be
+surprising for badly designed rewritings.  \mixin reinterprets
 the syntax with the sole goal of \emph{enhancing and completing code}:
@@
 and moving should work transparently in conjunction with our
-annotaiton, since they rely on the overall type structure of the
+annotation, since they rely on the overall type structure of the
 class, that we do not arbitrarily modify but just complete.
@@
 well-typed if they were well-typed without the instrumentation.
-Classless Java \emph{does not} guarantee heir coherence.  The reason
+Our rewriting \emph{does not} guarantee heir coherence.  The reason
 is that this would forbid adding any (default or abstract) method to
@@
 be aware of the annotation semantics and provide the right type while
-refining methods. To support heir coherence, Classless Java would need
+refining methods. To support heir coherence, we would need
 to give up automatic type refinement, which is an essential part of IB programming.
@@
 
-\subsection{Limitations}\marco{ to reorganize with updated limitations}
-\haoyuan{Should we only present the separate compilation and simple generics features in this (limitation) section?}
+\subsection{Limitations}
 Our prototype implementation using Lombok has certain limitations:
 \begin{itemize}
-\item \haoyuan{The annotation \mixin does not support separate compilation yet. Using \mixin, all
-  related interfaces have to appear in a single Java file. Therefore, changes to
-  a single interface would require re-compiling the whole file. Nevertheless, to support separate compilation we
-  studied an existing yet experimental feature \lstinline{@Delegate}, reused the name and modified its behavior.
+\item Eclipse has its own compiler (ejc), alternative to javac.
+  Lombok allows to write handlers for either javac of ejc.
+  At this stage our implementation only realizes the Eclipse handler and our
+  experiments are all conducted in Eclipse. The implementation for
+  the \texttt{javac} handler is still missing.
+\item The current implementation supports generics in a simplified manner: 
+type parameters can be used but the generic
+method typing is neither formalized in this paper nor explicitly checked by our 
+implementation, but simply  delegated to the Java compiler.
+
+\item 
+Lombok offers only limited/experimental support for separate compilation, that is
+to access information of code defined in different files.
+In the same spirit, we have a mature \mixin annotation, that
+does not support separate compilation yet: it requires that all
+  related interfaces have to appear in a single Java file.
+We also offer a less polished annotation, 
+that trying to reuse the (yet experimental) Lombock annotation \lstinline{@Delegate}, reused the name and modified its behavior.
   Now files in the same package can be compiled and \lstinline{@Delegate} has the same behavior as \mixin. A limitation is
@@
   \lstinline{@Delegate(types = Point2D.class)}. Another limitation is the methods generated by \lstinline{@Delegate} will
-  not be visible in the Outline window, but they can still be auto-completed in Eclipse.}
-\item At this stage our implementation only realizes the Eclipse handler and our
-  experiments are all conducted in Eclipse. The implementation for
-  \texttt{javac} is missing.
-\item \haoyuan{The current implementation only supports simple generics, which means type parameters can be used, however, the generic
-method typing is neither formalized in this paper nor implemented, but delegated to the Java compiler.}
+  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
 \end{itemize}

### 2016-03-23 OOPSLA16/sections/Implementation.tex
Notation ok?

@@
   related interfaces have to appear in a single Java file.
-We also offer a less polished annotation, 
-that trying to reuse the (yet experimental) Lombock annotation \lstinline{@Delegate}, reused the name and modified its behavior.
-  Now files in the same package can be compiled and \lstinline{@Delegate} has the same behavior as \mixin. A limitation is
-  that users need to put references to the super types of the annotated type together with the annotation, for instance,
-  \lstinline{@Delegate(types = Point2D.class)}. Another limitation is the methods generated by \lstinline{@Delegate} will
-  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
+Reusing the logic inside the yet experimental Lombock annotation \lstinline{@Delegate},
+we also offer a less polished annotation supporting
+separate compilation for files in the same package.
+%  has the same behavior as \mixin. A limitation is
+%  that users need to put references to the super types
+% of the annotated type together with the annotation, for instance,
+%  \lstinline{@Delegate(types = Point2D.class)}.
+%   Another limitation is the methods generated by \lstinline{@Delegate} will
+%  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
 \end{itemize}

### 2016-03-23 OOPSLA16/sections/Translation.tex
Notation ok?

--- 2016-03-22 OOPSLA16/sections/Translation.tex
@@
 
-This section gives an overview of what $\mixinAnn$ generates in
-Classless Java, and what properties are guaranteed in the translation.
+This section gives an overview of what $\mixinAnn$ generates,
+ and what formal properties are guaranteed in the translation.
 
@@
 Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod} present more details on the auxiliary
-functions. \bruno{if the figures are mentioned here, what should
-  follow is the explanation on what is one the figures. Please fix
-  this Haoyuan and makes sure that everything that is on the figures
-  is explained here. }
-The formal definition is available in
-Appendix~\ref{subsec:translation}.\bruno{what formal definition?}
+functions. 
+The first two points of Figure~\ref{figure:valid}
+defines function $\otherMethod$;
+ This function generates unimplemented
+ \Q@with-@ and fluent setters in the interface, where the 
+return types has been refined.
 
-The function $\ofMethod$ generates the static method $\QM{of}$, as an
-object factory. It detects all the field methods of $\C_0$, take them
-as arguments, and in the return statement is an anonymous class which
-generates getters, fluent setters and with-methods in certain
-cases. Some other features of $\mixinAnn$, including non-fluent
-setters and the $\QM{with}$ method, however, are not formalized in
-Classless Java. Appendix~\ref{subsec:otherfeatures} gives an informal
-explanation of those methods.
+To determine whether a method need to be generated,
+we check if such \Q@with-@ or setter methods
+require an implementation in $\C_0$, but not declared directly in $\C_0$.
 
-\bruno{I think there is explanation missing in this subsection;
-  probably text which is now in the appendix. Haoyuan, please take
-  care of this.}
+The third point gives the definition
+of $\valid$:
+It is valid to annotate an interface if 
+%To generate the \Q@of@ method, it is necessary for $\weakAnn$ to explicitly check if
+all abstract methods (that is, all those
+requiring an implementation) are valid.
+That is, we can categorize them in a pattern that we
+know how to implement (right column):
+it is either a field getter (first point),
+a with method (second point) or a setter (third point).
+
+Note that we write $\QM{with#}\m$ to append $\m$ to
+\QM{with}, following the camelCase rule. The first letter of $\m$
+must be lower-case and is changed to upper-case upon appending. For
+example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
+are \QM{with} and all identifiers of the form $\QM{with#}\m$.
+
+Figure~\ref{figure:ofmethod},
+defines the $\ofMethod$  function, that generates the static method $\QM{of}$ as an
+object factory.
+It detects all the field methods of $\C_0$ and use them to synthesize
+its arguments.
+The return statement instantiate an anonymous class which
+generates the needed getters, fluent setters and with-methods.
+
+The right column first point collect the getter methods,
+the second and third point generate implementations for with methods
+if needed; similarly, the fourth and fifth point generate fluent setters if needed.
+
+Some other features of $\mixinAnn$, including non-fluent
+setters and the $\QM{with}$ method are not formalized here.
+Appendix~\ref{subsec:otherfeatures} gives an detailed but 
+informal explanation of generation for those methods.
 
@@
 \begin{itemize}
-\item We denote with $\C^{\II}$ the name of an interface.
+\item We denote with $\C^{\II}$ and $m^\method$ the name of an interface and of a method.
 \item An interface table
@@
 
-\bruno{Do we have good names for the theorems Marco? Cannot be called
-  tunning now!}
 \begin{thm}[@ObjOf]

### 2016-03-23 OOPSLA16/sections/Translation.tex
section 3,4 some work completed today

@@
 where only some units may be translated/expanded.
-Note inheritance crucial influence in heir coherence, and how client coherence allows
-for only some units to be translated/expanded.
-\bruno{I don't
-  understand the point of the last sentence? Rephrase?}
-  \marco{ok I tried to rephase the next sentence, how is now?}
+Note inheritance crucial influence in heir coherence, and 
+our formulation of client coherence allows to discuss about intermediate
+stages where only some code units are translated/expanded, while conventional 
+type preservation refers only to the completely translated program.
 This means that developers and designers of Java libraries and frameworks

### 2016-03-23 OOPSLA16/sections/CaseStudy.tex
added a paragraph for the last case study

@@
 
+\begin{comment}
 \subsection{Refactoring a Compiler}
@@
 where the number of lines grown quite a while?}
+\end{comment}
+
+\subsection{Refactoring an Interpreter}
+The last case study applies our approach to a slightly larger code base of an
+interpreter for a Lisp-like language
+\lstinline{Mumbler}\footnote{https://github.com/cesquivias/mumbler/tree/master/simple},
+which is created for tutorial purpose of Truffle. Keeping a balance between
+simplicity and useful features, \lstinline{Mumbler} contains numbers, booleans,
+lists (encoding function calls and special forms such as if-expression, lambdas,
+etc). In the original code base, only one operation \texttt{eval} is
+supported. We add one more operation \texttt{print} to the code base. Then we
+modularize the code using our techniques to convert as many classes into
+interfaces, separate out \texttt{eval} and \texttt{print} operation then combine
+them. Although the code size is slightly increased (by 2.2\% SLOC), the
+modularity is greatly increased.

### 2016-03-23 OOPSLA16/sections/CaseStudy.tex
revised text related to case studies

@@
 independent extensions in Java.  The second case study shows how to model an
-embedded DSL for SQL languages with fluent interfaces. Finally, the third
+embedded DSL for SQL languages with fluent interfaces. The third
 case study models a simple game, and compares our implementation with an
 existing one, showing that the amount of code is reduced significantly using \mixin.
+Finally, the fourth one refactors an interpreter, showing that our approach  
+benefits modularity and scales to larger and real code base.
 
@@
 \begin{tabular}{l|l}
-\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
+\lstinputlisting[linerange=33-43]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
 &
-\lstinputlisting[linerange=55-65]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
+\lstinputlisting[linerange=54-64]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
 \end{tabular}
@@
 
-\lstinputlisting[linerange=48-51]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
+\lstinputlisting[linerange=47-50]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB
 
@@
 
-\lstinputlisting[linerange=89-91]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
+\lstinputlisting[linerange=88-]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 

### 2016-03-23 OOPSLA16/sections/Introduction.tex
revised text related to case studies

@@
 to an existing implementation~\cite{bono14} using plain Java 8. The last
-application is a larger case study of refactoring of a simple one-pass compiler, showing that our
-approach can scale to real code base. Noteworthy, all applications are implemented 
-without defining a single class!
+application is a relatively larger case study of refactoring of a simple interpreter, showing that our
+approach can benefit modularity and scale to real code base. Noteworthy, the first three applications are implemented 
+without defining a single class! The last one contains several classes that cannot be 
+converted to interfaces due to the limitation of our implementation of \mixin.
 

### 2016-03-23 UseMixinLombok/src/casestudy/ep/mutability/TestExpression.java

@@ -41,7 +41,7 @@ public class TestExpression {
     }
 }
 
-//BEGIN_EXPRESSION_INIT-
+//BEGIN_EXPRESSION_INIT2
 interface Exp { int eval(); Exp with(Exp val);}
 @Obj interface Lit extends Exp {
     int x(); void x(int val);
@@ -53,16 +53,16 @@ interface Exp { int eval(); Exp with(Exp val);}
         return e1().eval() + e2().eval();
     }
 }
-//END_EXPRESSION_INIT-
+//END_EXPRESSION_INIT2
 
-//BEGIN_EXPRESSION_SUB-
+//BEGIN_EXPRESSION_SUB2
 @Obj interface Sub extends Exp {
 	Exp e1(); Exp e2();
 	default int eval() {
 		return e1().eval() - e2().eval();} }
-//END_EXPRESSION_SUB-
+//END_EXPRESSION_SUB2
 
-//BEGIN_EXPRESSION_PRINT-
+//BEGIN_EXPRESSION_PRINT2
 interface ExpP extends Exp {String print();}
 @Obj interface LitP extends Lit, ExpP {
     default String print() {return "" + x();}
@@ -77,9 +77,9 @@ interface ExpP extends Exp {String print();}
     default void e1(Exp val) { e1(e1().with(val)); }
     default void e2(Exp val) { e2(e2().with(val)); }
 }
-//END_EXPRESSION_PRINT-
+//END_EXPRESSION_PRINT2
 
-//BEGIN_EXPRESSION_COLLECTLIT-
+//BEGIN_EXPRESSION_COLLECTLIT2
 interface ExpC extends Exp { List<Integer> collectLit(); }
 @Obj interface LitC extends Lit, ExpC {
     default List<Integer> collectLit() {
@@ -100,9 +100,9 @@ interface ExpC extends Exp { List<Integer> collectLit(); }
     default void e1(Exp val) { e1(e1().with(val)); }
     default void e2(Exp val) { e2(e2().with(val)); }
 }
-//END_EXPRESSION_COLLECTLIT-
+//END_EXPRESSION_COLLECTLIT2
 
-//BEGIN_INDEPENDENT_EXTENSIBILITY-
+//BEGIN_INDEPENDENT_EXTENSIBILITY2
 interface ExpPC extends ExpP, ExpC {
     ExpPC with(Exp val);
 }
[…]

### 2016-03-23 OOPSLA16/sections/Introduction.tex
related work

--- 2016-03-23 OOPSLA16/sections/Introduction.tex
@@
 mutable fields is not type-safe. While covariant refinement of mutable
-fields is supported by some type systems~\cite{bruce98astatically,McDirmid01jiazzi,ernst01family,ernst06virtual}, this requires
+fields is supported by some type systems~\cite{bruce98astatically,bruce1994paradigmatic,ernst06virtual,Saito2013933}, this requires
 significant complexity and restrictions to ensure that all uses of

### 2016-03-23 OOPSLA16/sections/RelatedWork.tex
Text in Section 4. Code style.

@@
 \section{Related Work}\label{sec:related}
-In this section we discuss related work and how it compares to Classless Java.
+%In this section we discuss related work and how it compares to Classless Java.
 
@@
 
+\paragraph{Models of Object-Oriented Programming}
+There are traditionally two big styles of OOP: class-based and
+prototype-based. The class-based tradition started with Simula~\cite{Dahl66simula}, 
+and is the dominant style of OOP today. Major languages such 
+as Java, C\#, C++ or Scala favour a class based OOP style. Very often
+class-based languages are statically typed and they have interfaces or
+some other similar mechanisms such as traits. However, in contrast to IB
+programming, interfaces or traits in CB programming are not meant to
+replace classes, but rather, they complement classes. Normally
+interfaces or traits define the signatures of the methods and
+\emph{may} have some behaviour (such as in Java8 interfaces or Scala
+traits). However classes are still responsible for object
+instantiation and handling state. Prototype-based languages started
+with Self~\cite{Ungar87self}, and are still widely used today via languages like
+JavaScript. Frequentely prototype-based languages are dynamically 
+typed. IB programming is clearly closer to CB programming. However the
+key difference is that classes are not needed in IB
+programming. Object interfaces recover the responsability of object
+instantiation, and abstract state operations provide a way to deal
+with state. A primary motivation of IB programming is to improve 
+modularity and flexibility. Central to these goals are the 
+more expressive and automated forms of type-refinement, and support 
+for multiple interface inheritance.
+
 \paragraph{Multiple Inheritance in Object Oriented Languages}
-Many authors have argued in favor or against multiple inheritance.  Multiple
+Many authors have argued in favor or against multiple inheritance.
+Multiple \emph{implementation}
 inheritance is very expressive, but difficult to model and implement, and can
-cause difficulty (including the famous diamond (fork-join)
+cause difficulties (including the famous diamond (fork-join)
 problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
@@
 %\paragraph{Mixins.}
-Mixins are a more restricted model than the C++ approach. Mixins allow to name
+Mixins are more restricted than the C++ approach. Mixins allow naming
 components that can be applied to various classes as reusable functionality
@@
 problem by disallowing constructor parameters, causing no ambiguity in cases
-such as diamond problem. However this approach has limited expressiveness, and
+such as the diamond problem. However this approach has limited expressiveness, and
 suffers from all the problems of linearized mixin composition. % Other languages, such as
@@
 number of classes, but also the class hierarchy complexity
-increases. \mixin does not complicate the class structure, and state
+increases. IB programming does not complicate the hierarchical structure, and state
 can also coexist with multiple inheritance.
@@
 resolution. Former work~\cite{bono14} provides details on mimicking the trait
-algebra through Java 8 interfaces.  We briefly recall the main points of their
+algebra through Java 8 interfaces.
+
+\begin{comment}
+We briefly recall the main points of their
 encoding; however we propose a different representation of \textbf{exclusion}.
@@
     \begin{lstlisting}
-    interface A { int x(); }    interface B { int y(); }    interface C extends A, B {}
+    interface A { int x(); }
+    interface B { int y(); }
+    interface C extends A, B {}
     \end{lstlisting}
@@
     interface B { default int m() {return 2;} }
-    interface C extends A, B { default int m() {return B.super.m();} }
+    interface C extends A, B {
+        default int m() {return B.super.m();} }
     \end{lstlisting}
@@
     interface A { default int m() {return 1;} }
-    interface B extends A { default int k() {return A.super.m();} }
+    interface B extends A {
+        default int k() {return A.super.m();} }
     \end{lstlisting}
@@
 \end{itemize}
+\end{comment}
 
@@
 %  as it did not make the implementation and language more complex.}
+
+\begin{comment}
 \marcoT{We could use Scala Traits to encode IB, but the simpler Java8
@@
   }
+\end{comment}
 
@@
 This demolishes most advantages of interfaces.}
-In certain situations, object interfaces allow automatic refinement for \emph{return
-types}. This is part of a bigger topic in class-based languages: expressing and
+Object interfaces support automatic type-refinement.
+Type refinement is part of a bigger topic in class-based languages: expressing and
 preserving type recursion and (nominal/structural) subtyping at the same time.
@@
 object-oriented programming languages. Object interfaces use covariant-return types to
-simulate some uses of MyType. But our approach only works for refining
+simulate some uses of MyType. But this approach only works for refining
 return types, whereas MyType is more general, as it also works for
-parameter types. Nevertheless, as illustrated with our examples and
+parameter types. Our approach to covariantly refine state can recover
+some of the additional expressiveness of MyType. As illustrated with our examples and
 case studies, object interfaces are still very useful in many
-practical applications.
+practical applications, yet they do not require additional 
+complexity from the type system.
+
 % The addition of MyType to a language will allow easy definition of

### 2016-03-23 OOPSLA16/sections/Appendix_Formalization.tex
Appendices A and B commented.

@@
 
+%% Appendices A & B commented
+\begin{comment}
 \section{Expanded related work}\label{sec:ERW}
@@
 
+\end{comment}
+
 \section{Formal Semantics}\label{sec:formal}
 
-This section presents a formalization of Classless Java, \delete{a minimal
-FeatherweightJava-like calculus} which models the essence of Java
-interfaces with default methods. This formalization is used \delete{in
-Section~\ref{sec:translation}} to define the semantics of object interfaces.
+This section presents a formalization of Classless Java, which models the essence of Java
+interfaces with default methods. This formalization is used to define the semantics of object interfaces.
 

### 2016-03-23 OOPSLA16/sections/FutureWork.tex
All for this night. I have done some improvement in section Future work. It can be read by BHY to see if it makes any sense.

--- 2016-03-22 OOPSLA16/sections/FutureWork.tex
@@
 \section{Future work}\label{sec:futurework}
-In this section we discuss potential future work.
 \begin{comment}
@@
 \paragraph{Encapsulation} %{Private state
-Since we are using Java8 to encode IB, we are unable to directly support qualifies
-like private or protected. In Java 9 private methods will be allowed in
-interfaces. Abstract state operations are, however, abstract methods,
-so they can not be private.
-Still, because the state can only be accessed by methods, it 
-is impossible for the user to know if a certain method maps directly to a field
-or if it has a default implementation.
+%Encapsulation 
+%Since we are using Java8 to encode IB, we are unable to directly support qualifies
+%like private or protected. In Java 9 private methods will be allowed in
+%interfaces. Abstract state operations are, however, abstract methods,
+%so they can not be private.
+%Still, because the state can only be accessed by methods, it
+%is impossible for the user to know if a certain method maps directly to a field
+%or if it has a default implementation.
 
-As shown by Ferruccio in [...] interfaces are enough to support privateness by a
+As shown in~\cite{BettiniDSS13} interfaces are enough to support encapsulation/privateness by a
  variation of the facade pattern:
 \begin{lstlisting}
-@Obj interface InternalData implements ExposedData{/*all the methods*/}//package visibility
-public interface ExposedData{
+@Obj interface InternalData implements ExposedData {
+    /*all the methods*/} //package visibility
+public interface ExposedData {
   /*just the exposed interface*/
-  static ExposedData of(...){return InternalData.of(...);}
-  }
+  static ExposedData of(...) {
+    return InternalData.of(...);}  }
 \end{lstlisting}
@@
 
-Using packages as a boundary we can make all object interfaces ``with privateness''
-invisible types for the rest of the program.
+Using packages as a boundary we can make all object interfaces \emph{where some form of privateness is desirable}
+as invisible types for the rest of the program.
 This preserve encapsulation, but limits inheritance:
@@
 
-Using Java9 private interface methods%
-\footnote{Or some ugly pattern available today like public static classes with private methods}
-one can define (and inherit) private behaviour; but private state inheritance in IB poses a logical challenge:
-How can the heir produce initialization code for a field whose very existence is secret?
+This can raise questions: keeping fields private is a staple of CB object oriented programming;
+by forcing the user/heir to rely only on methods, we avoid (a part)
+of the fragile base class problem and we obtain representation independence.
 
-CB get around this problem in two ways: the constructor have to call a super-constructor, and
-some fields can be pre-initialized.
-The first idea would not provide any encapsulation benefit in IB,
-since generated \Q@of@ methods expose all the field information anyway.
+In IB fields do not exists; abstract state methods like getters and setters are abstract:
+it does not make sense to make them private; however, since they are methods,
+the base interface may be able to evolve freely. 
+To understand if a fragile base interface problem exists in IB, we need more research.
 
-Pre-initialized fields could be supported by IB if we support a cache expression,
-that is \Q@cache(e)@ would evaluate the expression a single time for each object, and cache the result
-somewhere.\footnote{
-For example, in a the compiled bytecode for objects
-could have an extra field for each cached expression in its methods.
-Such cache expression would be quite useful for other reasons,
-including replacing involved initialization patterns.}
-
-In this way
-\marco{I fixed the code down, was wrong, there was no mention of hello in the original source}
-\begin{lstlisting}
-@Obj interface Text{
-  private String msg();//assuming Java9
-  private Text msg(String val);{f("hello");}
-  }
-\end{lstlisting}
-Would be expanded into 
-\begin{lstlisting}
-@Obj interface Text{
-  private Box<String> _msg(){
-    return cache(Box<String>.of("hello"));}
-  private String msg(){return _msg().val();}
-  private Text msg(String val){
-    return _msg().val(val);}
-  }
-\end{lstlisting}
-
-Where we could have a predefined \Q@Box@ type as\\*
-\Q|@Obj interface Box<T>{ T val(); Box<T> val(T val);}|
+%
+%Using Java9 private interface methods%
+%\footnote{Or some ugly pattern available today like public static classes with private methods}
+%one can define (and inherit) private behaviour; but private state inheritance in IB poses a logical challenge:
+%How can the heir produce initialization code for a field whose very existence is secret?
+%
+%CB get around this problem in two ways: the constructor have to call a super-constructor, and
+%some fields can be pre-initialized.
+%The first idea would not provide any encapsulation benefit in IB,
+%since generated \Q@of@ methods expose all the field information anyway.
+%
+%Pre-initialized fields could be supported by IB if we support a cache expression,
+%that is \Q@cache(e)@ would evaluate the expression a single time for each object, and cache the result
+%somewhere.\footnote{
+%For example, in a the compiled bytecode for objects
+%could have an extra field for each cached expression in its methods.
+%Such cache expression would be quite useful for other reasons,
+%including replacing involved initialization patterns.}
+%
+%In this way
+%\marco{I fixed the code down, was wrong, there was no mention of hello in the original source}
+%\begin{lstlisting}
+%@Obj interface Text{
+%  private String msg();//assuming Java9
+%  private Text msg(String val);{f("hello");}
+%  }
+%\end{lstlisting}
+%Would be expanded into
+%\begin{lstlisting}
+%@Obj interface Text{
+%  private Box<String> _msg(){
+%    return cache(Box<String>.of("hello"));}
+%  private String msg(){return _msg().val();}
+%  private Text msg(String val){
+%    return _msg().val(val);}
+%  }
+%\end{lstlisting}
+%
+%Where we could have a predefined \Q@Box@ type as\\*
+%\Q|@Obj interface Box<T>{ T val(); Box<T> val(T val);}|
+%
 
@@
 In this article we presented the concept of IB as a programming pattern over Java.
+An alternative would be to have a language providing IB as its only paradigm.
 We do not wish to define here a complete IB language, just to show the reader IB as
 a promising way to design a new oo language.
-The first task is to define the set of abstract state operations.
+To define such language, the first task is to define the set of abstract state operations.
 Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
@@
 
-Currently we are offering both setters and Fluent setters.
+Currently we are offering both setters and fluent setters.
 We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
@@
 Manually defining abstract state methods can be verbose, we could offer a convenient syntax to
-declare get/set/with.
+declare in a compact way get/set/with.
 For example
@@
 obj interface Foo{
-  String bar;//== String bar(); Foo withBar(String val);
-  var String beer;//also fluent setter
+  String bar;// exapanded as getter and with:
+  //String bar(); Foo withBar(String val);
+  var String beer;//as before plus fluent setter
   }
 \end{lstlisting}
+
+It could be possible to offer pre initialized abstract state, for example
+\begin{lstlisting}
+obj interface Foo{
+  var String bar="hello";
+  }
+\end{lstlisting}
+
+could be a short hand for writing
+
+\begin{lstlisting}
+obj interface Foo{
+  String bar();
+  Foo withBar(String val);
+  Foo bar(String val);
+  String initBar(){return "hello";}
+  }
+\end{lstlisting}
+and the automatically generated \Q@of@ method will use \Q@initX@ methods to initialize fields
+instead of generating the corresponding parameter.
 
@@
 obj interface Foo{
-  updater Bar, Beer//== Foo set(Bar val); Foo set(Beer val);
-  with Foo, Beer//== Foo with(Bar val); Foo with(Beer val);
+  updater Bar, Beer//declares imperative updaters:
+  //Foo set(Bar val); Foo set(Beer val);
+  with Foo, Beer//declared functional updaters:
+  //Foo with(Bar val); Foo with(Beer val);
   }
@@
 Finally, certain methods may have a specific meaning, for example
-if a private method called \Q@init@ is present, it will be called after construction,
-and it may be handled specially during multiple implementations, so that all relevant \Q@init@ methods are
+if a method called \Q@postInit@ is present, it will be called after construction,
+and it may be handled specially during multiple implementations, so that all relevant \Q@postInit@ methods are
 called in some order.

### 2016-03-23 OOPSLA16/sections/Implementation.tex
Section 3

@@
 Disciplined use of Classless Java (that is avoiding class
-definitions as done in Section~\ref{}), illustrates what \emph{pure} IB is like.
+declarations as done in Section~\ref{sec:ep}), illustrates what \emph{pure} IB is like.
 However, using \mixin, CB and IB programming can be mixed together, 
@@
 \subsection{Compilation Agents}
-Java supports compilation agents\cite{compilationagents}.
+Java supports compilation agents.
 That is, Java libraries can interact with the Java compilation process,
@@
 annotations. \mixin was created using Lombok.
-As you can see explained in Figure~\ref{fig:lombok}~\cite{neildo2011blog},
-first java source code is parsed into an abstract syntax tree (AST).
+Figure~\ref{fig:lombok}~\cite{neildo2011blog} illustrates the flow of
+the \mixin annotation.
+First java source code is parsed into an abstract syntax tree (AST).
 The AST is then captured by Lombok:
@@
 Lombok supports generation of code \emph{inside} a class/interface,
-which conventional Java annotation processors do not support this. For
+which conventional Java annotation processors do not support. For
 example, the standard \texttt{javax.annotation} processor, which is part of the
@@
 
-\paragraph{Modularity:}
+\paragraph{Modularity}
 While general preprocessing acts across module boundaries, compilation
@@
 
-\paragraph{Tool support:}
+\begin{figure*}[t]
+\centering
+\includegraphics[width=6in]{pdfs/screenshot4.png}
+\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
+\label{fig:screenshot}
+\end{figure*}
+
+\paragraph{Tool support}
 Features written in Lombok integrate and are supported directly in the
@@
 In Figure~\ref{fig:screenshot},
-\mixin generates \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@.
+\mixin generates an \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@.
 These methods are also visible to users, showing their types and modifiers in Outline window.
 Moreover, as a useful IDE feature, the auto-completion also works for these newly generated methods.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=3.3in]{pdfs/screenshot4.png}
-\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
-\label{fig:screenshot}
-\end{figure}
 
@@
 Compilation agents start from Java syntax, but they can reinterpret it.
-Preserving the syntax avoid syntactic conflicts, and \marcoT{allows} many
+Preserving the syntax avoids syntactic conflicts, and allows many
 tools to work transparently.
@@
 \item{\bf Self coherence}: the generated code itself is well-typed. That is,
-  type errors are not present in code the user has not written.
+  type errors are not present in code the user has not written (for
+  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
 In our case it means that either \mixin{} produces (in a controlled way) an
@@
 \paragraph{Heir coherence:} Another form of guarantee that could be
-useful in AST rewriting would be heir coherence. That is, interfaces
+useful in AST rewriting is heir coherence. That is, interfaces
 (and in general classes) inheriting the instrumented code are
 well-typed if they were well-typed without the instrumentation.
-Our rewriting \emph{does not} guarantee heir coherence.  The reason
+In a strict sense, our rewriting \emph{does not} guarantee heir coherence.  The reason
 is that this would forbid adding any (default or abstract) method to
-the annotated interfaces, including type refinement. Indeed consider
+the annotated interfaces, or even doing type refinement. Indeed consider
 the following:
@@
 non final class of a Java library.  We think refining return
-types breaks heir coherence ``less" than normal library evolution.
+types breaks heir coherence ``less" than normal library evolution, and
+if no automatic type-refinements are needed, then \mixin can claim a
+form of heir coherence.
 
-In Section~\ref{} \marcoT{we provide a formal definition for our
-safety claims.}
+Section~\ref{sec:translation} provide a formal definition for our
+safety claims.
 
 \subsection{Limitations}
-Our prototype implementation using Lombok has certain limitations:
+Our prototype implementation has certain limitations:
 \begin{itemize}
 \item Eclipse has its own compiler (ejc), alternative to javac.
-  Lombok allows to write handlers for either javac of ejc.
+  Lombok allows writing handlers for either javac of ejc.
   At this stage our implementation only realizes the Eclipse handler and our
@@
 implementation, but simply  delegated to the Java compiler.
-
 \item 
@@
   related interfaces have to appear in a single Java file.
-Reusing the logic inside the yet experimental Lombock annotation \lstinline{@Delegate},
+Reusing the logic inside the yet experimental Lombok annotation \lstinline{@Delegate},
 we also offer a less polished annotation supporting

### 2016-03-23 OOPSLA16/paper.tex
Minor fixes and typos
Section 3
some local changes

@@
 \usepackage{multirow}
+\usepackage{enumitem}
 
@@
 \usepackage{xspace}
+
+\setlist[itemize]{noitemsep,nolistsep}
+\setlist[enumerate]{noitemsep,nolistsep}
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\vspace{-3pt}\oldparagraph{#1}}
 
@@
 
-  This paper presents Interface Based (IB) object-oriented (OO)
+  This paper introduces Interface Based (IB) object-oriented (OO)
   programming. Unlike class based OO programming, in IB programming there
@@
   directly into Java. Classless Java uses annotation processing for
-  code generation and relies on the new features of Java 8 for
+  code generation, and relies on new features of Java 8 for
   interfaces. We illustrate the usefulness of IB through various

### 2016-03-23 OOPSLA16/sections/Introduction.tex
revised S1: typos, grammar errors

@@
 languages objects inherit from other objects, and thus objects own
-both behaviour and state (and objects is all you have).
-In class based languages an object is instance of a specific class,
+both behaviour and state (and objects are all you have).
+In class based languages an object is an instance of a specific class,
 and classes inherit from other classes.  Here objects own the state,
@@
 interface. Programmers do not define objects directly, but delegate
-the task to \emph{object interfaces}, whose role is similar to non
-abstract classes in CB languages. Instantiation of objects is
+the task to \emph{object interfaces}, whose role is similar to non-abstract 
+classes in CB languages. Instantiation of objects is
 done using static factory methods in object interfaces.
@@
 utility methods (such as getters and setters, or clone-like
-methods). Objects are the only responsible to define the ultimate
-behaviour of a method and, for example, if such method is just a
+methods). Objects are only responsible to define the ultimate
+behaviour of a method and, for example, if such a method is just a
 setter. Anything related to state is completely contained in the
@@
 by inheritance.  In contrast, in IB programming the state is never
-fixed, and methods that ``looks like'' abstract setters and getters
+fixed, and methods that look like abstract setters and getters
 can always receive an explicit implementation down in the inheritance
@@
 multiple \emph{interface} inheritance has been supported since
-inception, and in Java8 default methods~\cite{goetz12fdefenders} bring some of the
+inception, and in Java 8 default methods~\cite{goetz12fdefenders} bring some of the
 advantages of traits into Java. In contrast, the literature~\cite{Sak89dis,bracha90mixin,malayeri2009cz}
@@
 Our IB embedding relies on the
-new features of Java 8 for interfaces: interface \emph{static methods}; and
+new features of Java 8 for interfaces: interface \emph{static methods} and
 \emph{default methods}, which allow interfaces to have method
@@
 Using Java annotation processors, we produce an implementation of
-Classless Java, which allow us to stick to pure Java8. By annotating
+Classless Java, which allow us to stick to pure Java 8. By annotating
 with \Q|@Obj| the interfaces that represent object interfaces, we can
@@
 Java programs, and conduct various case studies.  Finally, we also
-formally define the behaviour of our \Q|Obj| annotation and discuss
+formally define the behaviour of our \mixin annotation and discuss
 its properties.
@@
 multiple inheritance. For this application we show that there is a
-significant reduction in the numbers of lines of code when compared 
+significant reduction in the number of lines of code when compared 
 to an existing implementation~\cite{bono14} using plain Java 8. The last
 application is a relatively larger case study of refactoring of a simple interpreter, showing that our
-approach can benefit modularity and scale to real code base. Noteworthy, the first three applications are implemented 
+approach can benefit modularity and scale to real code base. Noteworthy is the fact that the first three applications are implemented 
 without defining a single class! The last one contains several classes that cannot be 

### 2016-03-23 OOPSLA16/sections/Introduction.tex
revert some S1 changes and revise S2

@@
 Our IB embedding relies on the
-new features of Java 8 for interfaces: interface \emph{static methods} and
+new features of Java 8 for interfaces: interface \emph{static methods}; and
 \emph{default methods}, which allow interfaces to have method
@@
 application is a relatively larger case study of refactoring of a simple interpreter, showing that our
-approach can benefit modularity and scale to real code base. Noteworthy is the fact that the first three applications are implemented 
+approach can benefit modularity and scale to real code base. Noteworthy, the first three applications are implemented 
 without defining a single class! The last one contains several classes that cannot be 

### 2016-03-23 OOPSLA16/sections/Overview.tex
revert some S1 changes and revise S2

@@
 To model the animal system with object interfaces all that a user
-needs to do is to add a \mixinAnn{} annotation to the \texttt{Horse},
+needs to do is to add an \mixinAnn{} annotation to the \texttt{Horse},
 \texttt{Bird}, and \texttt{Pegasus} interfaces:
@@
 programming.
-Lets now model the complete animal system where \Q@Animal@s includes a \Q@location@
-representing their positions in space. We use 2D points to keep track of locations using coordinates.
+Now we model the complete animal system where an \Q@Animal@ includes a \Q@location@
+representing its position in space. We use 2D points to keep track of locations using coordinates.
 
@@
 Point2D p = new Point2D() {
-  int x() {return 4;}
+  public int x() {return 4;}
   public int y() {return 2;}
@@
 simple-minded constructor. \mixin first looks at the abstract methods and detects
-what are the fields, then generates an \Q@of@ method with one parameter for each
+what the fields are, then generates an \Q@of@ method with one parameter for each
 of them. That is, we can just write
@@
 data-structures. As such, it is
-supported in by \mixin as \Q@with-@ methods. For example:
+supported by \mixin as \Q@with-@ methods. For example:
 \begin{lstlisting}
@@
     int x();    int y(); // getters
-    // with methods
+    // with- methods
     Point2D withX(int val);
@@
 
-\noindent If there is a large number of fields, \texttt{with} methods
+\noindent If there is a large number of fields, \texttt{with-} methods
 will save programmers from writing large amounts of tedious code that
@@
 and an implementation of \texttt{run()} is defined using a \Q@default@
-method. The implementation of \texttt{run()} further illustrates the convenience of \texttt{with} methods:
+method. The implementation of \texttt{run()} further illustrates the convenience of \texttt{with-} methods:
 
@@
 
-\paragraph{\Q@Point3D@ and properties updaters}
+\paragraph{\Q@Point3D@ and property updaters}
 The \Q@Point3D@ interface is defined as follows:
@@
 this idea, the method \Q@with(Point2D)@ is an example of a (functional)
-properties updater: it takes a certain type of object and returns a copy of the
+property updater: it takes a certain type of object and returns a copy of the
 current object where all the fields that match fields in the parameter
@@
 \Q@Point3D.of@.  Note how \Q@with-@ methods are automatically refined in their
-return type, so that code like:
+return types, so that code like:
 

### 2016-03-23 OOPSLA16/sections/Implementation.tex
revise S3

@@
 %  libraries, but it would be a taunting task.}
-Creating a new language or language extension would be an
+Creating a new language or a language extension would be an
 elegant way to illustrate the point of IB. However,
@@
 programming and be practical, we have instead implemented 
-\mixin as an annotation over Java8, and a \emph{compilation agent}.
+\mixin as an annotation over Java 8, and a \emph{compilation agent}.
 That is, the Classless Java style of programming
-is supported by a Java8 library.
+is supported by a Java 8 library.
 
@@
 The key to our implementation is the use of compilation agents, which
-are supported in Java, and allow us to rewrite the Java AST just
+ allows us to rewrite the Java AST just
 before compilation. We discuss advantages and limitations of our approach.
@@
 \subsection{Compilation Agents}
-Java supports compilation agents.
-That is, Java libraries can interact with the Java compilation process,
+Java supports compilation agents, where Java libraries can interact with the Java compilation process,
 acting as a man in the middle between the
-generation of the abstract syntax tree and the Java Bytecode.
+generation of the abstract syntax tree and the Java bytecode.
 
@@
 the syntax with the sole goal of \emph{enhancing and completing code}:
-we satisfy the behaviour of abstract methods; we add method
-implementations; and we refine return types.  We consider this to be
+we satisfy the behaviour of abstract methods; add method
+implementations; and refine return types.  We consider this to be
 quite easy to follow and reason about, since it is similar to what
@@
 annotation, since they rely on the overall type structure of the
-class, that we do not arbitrarily modify but just complete.
+class, which we do not arbitrarily modify but just complete.
 
@@
 
-\paragraph{Syntax and Type Errors:}
+\paragraph{Syntax and type errors}:
 Some preprocessors (like the C one) can produce syntactically invalid code.
@@
 
-\item{\bf Self coherence}: the generated code itself is well-typed. That is,
+\item{\bf Self coherence}: The generated code itself is well-typed. That is,
   type errors are not present in code the user has not written (for
@@
 
-\item{\bf Client coherence}: all the client code (for example method calls)
+\item{\bf Client coherence}: All the client code (for example method calls)
   that is well-typed before code generation is also well-typed after the generation.
@@
 
-\paragraph{Heir coherence:} Another form of guarantee that could be
+\paragraph{Heir coherence}: Another form of guarantee that could be
 useful in AST rewriting is heir coherence. That is, interfaces
@@
 Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
-would be correct before translation, but would be ill-typed after the translation.
+would be correct before translation, but ill-typed after the translation.
 
@@
 coherence during evolution and still claim backward compatibility (false in
-theory but statistically true in practice). In practice to break heir
-coherence it is just enough to add any method to any
-non final class of a Java library.  We think refining return
-types breaks heir coherence ``less" than normal library evolution, and
+theory but statistically true in practice). In practice, adding any method to any
+non final class of a Java library is enough to break heir
+coherence.  We think return type refinement breaks heir coherence ``less" than normal library evolution, and
 if no automatic type-refinements are needed, then \mixin can claim a
@@
 \item Eclipse has its own compiler (ejc), alternative to javac.
-  Lombok allows writing handlers for either javac of ejc.
+  Lombok allows writing handlers for either javac or ejc.
   At this stage our implementation only realizes the Eclipse handler and our

### 2016-03-23 OOPSLA16/sections/Translation.tex
revise S4

@@
 The first two points of Figure~\ref{figure:valid}
-defines function $\otherMethod$;
+define function $\otherMethod$;
  This function generates unimplemented
@@
 
-To determine whether a method need to be generated,
+To determine whether a method needs to be generated,
 we check if such \Q@with-@ or setter methods
@@
 of $\valid$:
-It is valid to annotate an interface if 
+it is valid to annotate an interface if 
 %To generate the \Q@of@ method, it is necessary for $\weakAnn$ to explicitly check if
@@
 
-Figure~\ref{figure:ofmethod},
+Figure~\ref{figure:ofmethod} 
 defines the $\ofMethod$  function, that generates the static method $\QM{of}$ as an
@@
 its arguments.
-The return statement instantiate an anonymous class which
+The return statement instantiates an anonymous class which
 generates the needed getters, fluent setters and with-methods.
 
-The right column first point collect the getter methods,
-the second and third point generate implementations for with methods
+The right column first point collects the getter methods,
+the second and third point generate implementations for \texttt{with-} methods
 if needed; similarly, the fourth and fifth point generate fluent setters if needed.
@@
 \item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
-\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
+\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiates it.
 \end{itemize}
@@
 Informally, the theorems mean that for a client program that
-typechecks before the translation is applied, if the annotated type has
+type-checks before the translation is applied, if the annotated type has
 no subtypes and no objects of that type are created, then type safety

### 2016-03-24 OOPSLA16/sections/Conclusion.tex
Conclusions first pass

--- 2015-12-09 paper/sections/Conclusion.tex
@@
 was probably overlooked: the concept of class
-(in Java) is now (almost) redundant and unneeded.  We define a subset of Java,
-called Classless Java, where programs and (reusable) libraries can be
-easily defined and used.  To avoid syntactic boilerplate
+(in Java) is now (almost) redundant and unneeded.
+We proposed a programming style, called ClassLess Java, where
+truly object oriented programs and (reusable) libraries
+can be defined and used without ever defining a single class.
+
+However, using this programming style directly in Java is very verbose.
+To avoid syntactic boilerplate
 caused by Java not being originally designed to work without classes,
@@
 for various methods (e.g. getters, setters, with-methods) and a
-mechanism to instantiate objects. The \mixin annotation helps programmers
+mechanism to instantiate objects. 
+We leverage on annotation processing and the Lombok library, in this way
+\mixin is just a normal Java library; thus our proposed style can be integrated
+in any Java project.
+
+The \mixin annotation helps programmers
 to write less cumbersome code while coding in Classless Java. Indeed, 
@@
 annotation can be less cumbersome than full Java.
+
+Class less Java is just a programming style, but is 
+showing the way for a new flavour of object orientation:
+We propose Interface Based Object Oriented languages (IB),
+as opposed to Class Based or Prototype Based.
+
+In IB state is not modelled at the platonic/ideal level
+but is handled excursively by instances.
+
+This unlock useful code reuse patterns, as shown in~\ref{sec:ep}.
+
 Interestingly, without classes there is also no subclassing. This scratches an old
@@
   Classless Java does not have classes purely in the Java sense.  
- Classless Java encourages coding in a more flexible way by either
- keeping a higher abstraction level (interfaces are a more abstract
- concept than classes), or relying on concrete object initialization
- (the \Q@new I(){...}@ construct).
+% Classless Java encourages coding in a more flexible way by either
+% keeping a higher abstraction level (interfaces are a more abstract
+% concept than classes), or relying on concrete object initialization
+% (the \Q@new I(){...}@ construct).
 
- More generally, we identify the concept of \emph{language tuning}.
- We identify libraries that are already performing language tuning (Lombok and Cofoja), and 
- we forecast many different kinds of language tuning will emerge on suitable platforms like Java or the C\# CLR.
- We identify various kinds of safety guarantees that can be offered by language tuning, but the door is open for more flavors of safety guarantees to emerge.
+% More generally, we identify the concept of \emph{language tuning}.
+% We identify libraries that are already performing language tuning (Lombok and Cofoja), and 
+% we forecast many different kinds of language tuning will emerge on suitable platforms like Java or the C\# CLR.
+% We identify various kinds of safety guarantees that can be offered by language tuning, but the door is open for more flavors of safety guarantees to emerge.

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
add references
Some spacing tweking.
Case study

@@
-\section{Case Studies}
-In this section we conduct \numOfCaseStudies case studies which reveal various
-advantages of using the \mixin annotation. The first case study provides a
-simple way to solve the Expression Problem while supporting multiple,
-independent extensions in Java.  The second case study shows how to model an
-embedded DSL for SQL languages with fluent interfaces. The third
-case study models a simple game, and compares our implementation with an
-existing one, showing that the amount of code is reduced significantly using \mixin.
-Finally, the fourth one refactors an interpreter, showing that our approach  
-benefits modularity and scales to larger and real code base.
+\section{Applications and Case Studies}
 
-\subsection{A Trivial Solution to the Expression Problem with Object Interfaces}
+This section illustrates applications and larger case studies for
+Classless Java. The first application shows how a useful pattern,
+using multiple inheritance and type-refinement, can be conviniently
+encoded in Classless Java. The second application. shows how to model an
+embedded DSL based on fluent APIs. The two larger case studies, take 
+existing projects and refactor them into Classless Java. The first
+case study shows a significant reduction in code size, while the
+second case study maintains the same amount of code, but increases 
+modularity.
+
+\subsection{The Expression Problem with Object Interfaces}
 
@@
 \lstinline{Mumbler}\footnote{https://github.com/cesquivias/mumbler/tree/master/simple},
-which is created for tutorial purpose of Truffle. Keeping a balance between
+which is created for tutorial purpose of Truffle~\cite{wurthinger2013one}. Keeping a balance between
 simplicity and useful features, \lstinline{Mumbler} contains numbers, booleans,

### 2016-03-24 OOPSLA16/sections/Translation.tex
Some spacing tweking.
Section 4

@@
 \begin{figure*}[t]
+\saveSpaceFig
 \centering
@@
 \begin{figure*}[t]
+\saveSpaceFig
 \centering
-\hspace{-.3in}
 \begin{minipage}[b]{0.5\textwidth}
@@
 \label{figure:valid}
+\saveSpaceFig
 \end{figure*}
@@
 
-\marcoT{We formalize syntax and typing for
- Classless Java} in Appendix~\ref{sec:formal}, which models the essence of
+We formalize syntax and typing for
+ Classless Java in Appendix~\ref{sec:formal}, which models the essence of
 Java interfaces with default methods.
@@
 just for the conventional \Q@void@ type.
-
-\marcoT{Since our properties are about preserving typing,
-we do not need to formalize Classless Java semantic to proof our statements.}
+Since our properties are about preserving typing,
+we do not need to formalize Classless Java semantics to prove our statements.
 \begin{figure*}[t]
@@
 \label{figure:ofmethod}
+\saveSpaceFig
 \end{figure*}
@@
 
-Figure~\ref{figure:translation} presents the translation functions. In the first function, $\mixinAnn$
-injects refined methods to interface $\C_0$, while in the second function, $\weakAnn$ invokes
+Figure~\ref{figure:translation} presents the translation. In
+the first function, $\mixinAnn$
+injects refined methods to interface $\C_0$. The second function, $\weakAnn$ invokes
 $\ofMethod(\C_0)$ and generates the $\QM{of}$
@@
 
-Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod} present more details on the auxiliary
+Figure~\ref{figure:valid} presents more details on the auxiliary
 functions. 
 The first two points of Figure~\ref{figure:valid}
-define function $\otherMethod$;
+define function $\otherMethod$.
  This function generates unimplemented
  \Q@with-@ and fluent setters in the interface, where the 
-return types has been refined.
-
+return types have been refined.
 To determine whether a method needs to be generated,
 we check if such \Q@with-@ or setter methods
-require an implementation in $\C_0$, but not declared directly in $\C_0$.
-
+require an implementation in $\C_0$, but are not declared directly in $\C_0$.
 The third point gives the definition
@@
 a with method (second point) or a setter (third point).
-
 Note that we write $\QM{with#}\m$ to append $\m$ to
@@
 generates the needed getters, fluent setters and with-methods.
-
 The right column first point collects the getter methods,
@@
 IT is OK if under such interface table, all interfaces are OK,
-\marcoT{that is, well typed}.
+that is, well typed.
 \item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
@@
 no subtypes and no objects of that type are created, then type safety
-\marcoT{of the generated code}
-is guaranteed after the successful translation.
+of the generated code is guaranteed after the successful translation.
 
@@
 the \mixin theorem guarantees only self and client coherence. The formal
-theorem proofs are available in the supplementary
-materials\bruno{point to appendix instead}.
+theorem proofs are available in Appendix~\ref{sec:appendix}.
 
@@
 Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
-type preservation theorems~\cite{}. The reason is
+type preservation theorems. The reason is
 to better model how our approach behaves in a object oriented software ecosystem with inheritance,
 where only some units may be translated/expanded.
-Note inheritance crucial influence in heir coherence, and 
-our formulation of client coherence allows to discuss about intermediate
-stages where only some code units are translated/expanded, while conventional 
-type preservation refers only to the completely translated program.
-This means that developers and designers of Java libraries and frameworks
+Note inheritance's crucial influence in heir coherence.
+Our formulation of client coherence allows us to discuss about intermediate
+stages where only some code units are translated/expanded. Conventional 
+type preservation refers only to completely translated program.
+Our coherence guarantees mean that developers and designers of Java libraries and frameworks
 can start using IB (and our \mixin annotation) in the evolution of their products

### 2016-03-24 OOPSLA16/sections/Introduction.tex
Introduction

@@
 They aim to couple flexibility and rigour, expressive power and
-modular reasoning.  Two main ideas have emerged to this end: prototype
+modular reasoning.  Two main ideas emerged to this end: prototype
 based (PB)~\cite{Ungar87self} and class based (CB) languages such as
 Java or Scala.  In prototype based
-languages objects inherit from other objects, and thus objects own
+languages objects inherit from other objects. Thus objects own
 both behaviour and state (and objects are all you have).
@@
 One important advantage of the use of abstract state operations and
-automatic type-refinement is that it offers a new form of
+type-refinement is that it allows a new approach to
 \emph{type-safe covariant mutable state}. That is, in IB programming,
-it is possible to refine \emph{mutable} ``fields'' in subtypes. This is
+it is possible to type-refine \emph{mutable} ``fields'' in subtypes. This is
 typically forbidden in CB programming languages (including Java and
@@
 with \Q|@Obj| the interfaces that represent object interfaces, we can
-generate boring and repetitive code for interface instantiation and
+automatically generate code for interface instantiation and
 type refinement. Such code should not be needed in the first place in
@@
 
+\begin{comment}
 To evaluate the usefulness of object interfaces, we illustrate
@@
 converted to interfaces due to the limitation of our implementation of \mixin.
+\end{comment}
 
@@
  combination of abstract state operations and type-refinement enables 
- a form of mutable fields that can be covariantly refined in a type-safe way.
+ a form of mutable state that can be covariantly refined in a type-safe way.
 
@@
 
-\item {\bf Case Studies:} We illustrate the usefulness of IB
+\item {\bf Applications and Case Studies:} We illustrate the usefulness of IB
   through various examples and case studies.

### 2016-03-24 OOPSLA16/sections/Implementation.tex
Some spacing tweking.
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

@@
 \begin{figure}[t]
+\saveSpaceFig
 \centering
@@
 \label{fig:lombok}
+\saveSpaceFig
 \end{figure}
@@
 
-\paragraph{Direct modification of the AST:}
+\paragraph{Direct modification of the AST}
 Lombok alters the generation process of the class files,
@@
 \begin{figure*}[t]
+\saveSpaceFig
+\saveSpaceFig
 \centering
@@
 \label{fig:screenshot}
+\saveSpaceFig
 \end{figure*}
@@
 
-\paragraph{Clarity against obfuscation:}
+\paragraph{Clarity against obfuscation}
 Preprocessors bring great power, which can easily be misused producing
@@
 
-\paragraph{Syntax and type errors}:
+\paragraph{Syntax and type errors}
 Some preprocessors (like the C one) can produce syntactically invalid code.
@@
 
-\item{\bf Self coherence}: The generated code itself is well-typed. That is,
+\item{\bf Self coherence}: the generated code itself is well-typed. That is,
   type errors are not present in code the user has not written (for
@@
 
-\item{\bf Client coherence}: All the client code (for example method calls)
+\item{\bf Client coherence}: all the client code (for example method calls)
   that is well-typed before code generation is also well-typed after the generation.
@@
 
-\paragraph{Heir coherence}: Another form of guarantee that could be
+\paragraph{Heir coherence} Another form of guarantee that could be
 useful in AST rewriting is heir coherence. That is, interfaces

### 2016-03-24 OOPSLA16/sections/Introduction.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

--- 2016-03-23 OOPSLA16/sections/Introduction.tex
@@
 
-\item {\bf Classless Java:} A practical realization of IB in
+\item {\bf Classless Java:} a practical realization of IB in
   Java. Classless Java is implemented using annotation processing, 
@@
 
-\item {\bf Type-Safe Covariant Mutable State:} We show how the 
+\item {\bf Type-safe covariant mutable state:} we show how the 
  combination of abstract state operations and type-refinement enables 
@@
 
-\item {\bf Case Studies:} We illustrate the usefulness of IB
+\item {\bf Case studies:} we illustrate the usefulness of IB
   through various examples and case studies.

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

@@
 
-\paragraph{Models of Object-Oriented Programming}
+\paragraph{Models of Object-Oriented programming}
 There are traditionally two big styles of OOP: class-based and
@@
 
-\paragraph{Multiple Inheritance in Object Oriented Languages}
+\paragraph{Multiple inheritance in Object-Oriented languages}
 Many authors have argued in favor or against multiple inheritance.

### 2016-03-24 OOPSLA16/sections/Translation.tex
unify title/paragraph/enumerate capitalization: The rule is: - Capitalization All Section/Subsection Titles. - No capitalization for paragraph/enumerate titles. - No capitalization for the first word after a colon ":" (There's difference in this point between American and British English, I think we should stick to one).

@@
 
-\paragraph{Type preservation:}
+\paragraph{Type preservation}
 Note that we preferred to introduce self, client and heir coherence instead of referring to conventional

### 2016-03-24 OOPSLA16/paper.tex

@@
   available in CB languages.  In IB state (including mutable state)
-  can be automatically type-refined in subtypes. The combination of a
-  purely IB style of programming and automatic type-refinement enables
+  can be type-refined in subtypes. The combination of a
+  purely IB style of programming and type-refinement enables
   powerful idioms using multiple inheritance and state.  To introduce

### 2016-03-24 OOPSLA16/sections/Introduction.tex

--- 2016-03-24 OOPSLA16/sections/Introduction.tex
@@
 
-\item {\bf Applications and Case Studies:} We illustrate the usefulness of IB
+\item {\bf Applications and case studies:} we illustrate the usefulness of IB
   through various examples and case studies.

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
Spellcheck.

@@
   print operation).}\label{fig:ep}
+\saveSpaceFig
 \end{figure*}% \bruno{Yanlin: code is overflowing the margins. Note that
@@
 \begin{enumerate}
-\item Instead of forcing programmers to hand write code using \textbf{return
+\item Instead of forcing programmers to hand-write code using \textbf{return
     this}, our approach with \mixin annotation removes this verbosity and
@@
 \lstinputlisting[linerange=14-18]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
-In this way, when a query created using \texttt{ExtendedDatabase},
+In this way, when a query is created using \texttt{ExtendedDatabase},
 all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling
@@
 modularize the code using our techniques to convert as many classes into
-interfaces, separate out \texttt{eval} and \texttt{print} operation then combine
+interfaces, separate out \texttt{eval} and \texttt{print} operations then combine
 them. Although the code size is slightly increased (by 2.2\% SLOC), the

### 2016-03-24 OOPSLA16/sections/FutureWork.tex
Spellcheck.

@@
 as invisible types for the rest of the program.
-This preserve encapsulation, but limits inheritance:
-package protected interfaces can not be seen,
-thus can not be implemented from outside the package.
+This preserves encapsulation, but limits inheritance:
+package protected interfaces cannot be seen,
+thus cannot be implemented from outside the package.
 
@@
 \paragraph{Clone, toString, equals and hashCode}
-Methods originally defined in Java class \Q@Object@, as \Q@clone@ and
+Methods originally defined in Java class \Q@Object@, such as \Q@clone@ and
 \Q@toString@, can be supported by our approach with special care. If an
@@
 We do not wish to define here a complete IB language, just to show the reader IB as
-a promising way to design a new oo language.
+a promising way to design a new OO language.
 To define such language, the first task is to define the set of abstract state operations.

### 2016-03-24 OOPSLA16/sections/Implementation.tex
Spellcheck.

@@
 to access information of code defined in different files.
-In the same spirit, we have a mature \mixin annotation, that
+In the same spirit, we have a mature \mixin annotation, which
 does not support separate compilation yet: it requires that all

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
Spellcheck.

@@
 interfaces or traits define the signatures of the methods and
-\emph{may} have some behaviour (such as in Java8 interfaces or Scala
+\emph{may} have some behaviour (such as in Java 8 interfaces or Scala
 traits). However classes are still responsible for object
@@
 with Self~\cite{Ungar87self}, and are still widely used today via languages like
-JavaScript. Frequentely prototype-based languages are dynamically 
+JavaScript. Frequently prototype-based languages are dynamically
 typed. IB programming is clearly closer to CB programming. However the
 key difference is that classes are not needed in IB
-programming. Object interfaces recover the responsability of object
+programming. Object interfaces recover the responsibility of object
 instantiation, and abstract state operations provide a way to deal
@@
 new features: exact statements and nonheritable methods. Both are
-related to our work: 1) any method generated inside of the \Q@of@ method is indeed
+related to our work: 1) any method generated inside the \Q@of@ method is indeed
 non-inheritable since there is no class name to extend from; 2) exact

### 2016-03-24 OOPSLA16/sections/Conclusion.tex
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex

@@
 We proposed a programming style, called ClassLess Java, where
-truly object oriented programs and (reusable) libraries
+truly \objectoriented programs and (reusable) libraries
 can be defined and used without ever defining a single class.
@@
 showing the way for a new flavour of object orientation:
-We propose Interface Based Object Oriented languages (IB),
-as opposed to Class Based or Prototype Based.
+We propose \interfacebased \objectoriented languages (IB),
+as opposed to \classbased or \prototypebased.
 

### 2016-03-24 OOPSLA16/sections/FutureWork.tex
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex

@@
 
-This can raise questions: keeping fields private is a staple of CB object oriented programming;
+This can raise questions: keeping fields private is a staple of CB \objectoriented programming;
 by forcing the user/heir to rely only on methods, we avoid (a part)

### 2016-03-24 OOPSLA16/sections/Introduction.tex
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex

@@
 
-Object oriented languages strive to offer great code reuse.
+\Objectoriented languages strive to offer great code reuse.
 They aim to couple flexibility and rigour, expressive power and
 modular reasoning.  Two main ideas emerged to this end: prototype
-based (PB)~\cite{Ungar87self} and class based (CB) languages such as
-Java or Scala.  In prototype based
+based (PB)~\cite{Ungar87self} and \classbased (CB) languages such as
+Java or Scala.  In \prototypebased
 languages objects inherit from other objects. Thus objects own
 both behaviour and state (and objects are all you have).
-In class based languages an object is an instance of a specific class,
+In \classbased languages an object is an instance of a specific class,
 and classes inherit from other classes.  Here objects own the state,
@@
 This paper presents a third alternative: the concept of
-\textbf{Interface based} (IB) object oriented languages, where objects
+\textbf{\interfacebased} (IB) \objectoriented languages, where objects 
 implement interfaces directly. In IB interfaces own the implementation

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex

@@
 
-\paragraph{Models of Object-Oriented programming}
-There are traditionally two big styles of OOP: class-based and
-prototype-based. The class-based tradition started with Simula~\cite{Dahl66simula},
+\paragraph{Models of \objectoriented programming}
+There are traditionally two big styles of OOP: \classbased and
+\prototypebased. The \classbased tradition started with Simula~\cite{Dahl66simula},
 and is the dominant style of OOP today. Major languages such
-as Java, C\#, C++ or Scala favour a class based OOP style. Very often
-class-based languages are statically typed and they have interfaces or
+as Java, C\#, C++ or Scala favour a \classbased OOP style. Very often
+\classbased languages are statically typed and they have interfaces or
 some other similar mechanisms such as traits. However, in contrast to IB
@@
 traits). However classes are still responsible for object
-instantiation and handling state. Prototype-based languages started
+instantiation and handling state. \Prototypebased languages started
 with Self~\cite{Ungar87self}, and are still widely used today via languages like
-JavaScript. Frequently prototype-based languages are dynamically
+JavaScript. Frequently \prototypebased languages are dynamically
 typed. IB programming is clearly closer to CB programming. However the
@@
 
-\paragraph{Multiple inheritance in Object-Oriented languages}
+\paragraph{Multiple inheritance in \objectoriented languages}
 Many authors have argued in favor or against multiple inheritance.
@@
 Object interfaces support automatic type-refinement.
-Type refinement is part of a bigger topic in class-based languages: expressing and
+Type refinement is part of a bigger topic in \classbased languages: expressing and
 preserving type recursion and (nominal/structural) subtyping at the same time.
@@
 invoking a method with MyType in a parameter position, the exact type of the
-receiver must be known.  This is a big limitation in class based object oriented
-programming, and is exasperated by the interface-based programming we propose: no
+receiver must be known.  This is a big limitation in \classbased \objectoriented
+programming, and is exasperated by the \interfacebased programming we propose: no
 type is ever going to be exact since classes are not explicitly used. A recent
@@
 Admittedly, MyType greatly enhances the expressiveness and extensibility of
-object-oriented programming languages. Object interfaces use covariant-return types to
+\objectoriented programming languages. Object interfaces use covariant-return types to
 simulate some uses of MyType. But this approach only works for refining

### 2016-03-24 OOPSLA16/sections/Translation.tex
Spellcheck.
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex

@@
 Figure~\ref{figure:ofmethod}
-defines the $\ofMethod$  function, that generates the static method $\QM{of}$ as an
+defines the $\ofMethod$  function, which generates the static method $\QM{of}$ as an
 object factory.
@@
 type preservation theorems. The reason is
-to better model how our approach behaves in a object oriented software ecosystem with inheritance,
+to better model how our approach behaves in a \objectoriented software ecosystem with inheritance,
 where only some units may be translated/expanded.

### 2016-03-24 OOPSLA16/paper.tex
please use these commands in further editions to unify hyphen usage: \objectoriented \interfacebased \classbased \prototypebased ... you may refer to file predef/def.tex
small fixes

@@
 \title{Classless Java}
-\subtitle{Interface Based Programming for the Masses}
+\subtitle{\InterfaceBased Programming for the Masses}
 
@@
 
-  This paper introduces Interface Based (IB) object-oriented (OO)
-  programming. Unlike class based OO programming, in IB programming there
+  This paper introduces \interfacebased (IB) \objectoriented (OO)
+  programming. Unlike \classbased OO programming, in IB programming there
   is no need for classes. To support state, IB programming does not

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
local changes

@@
 using multiple inheritance and type-refinement, can be conviniently
-encoded in Classless Java. The second application. shows how to model an
-embedded DSL based on fluent APIs. The two larger case studies, take
+encoded in Classless Java. The second application shows how to model
+embedded DSLs based on fluent APIs. The two larger case studies, take
 existing projects and refactor them into Classless Java. The first
 case study shows a significant reduction in code size, while the
-second case study maintains the same amount of code, but increases
+second case study maintains the same amount of code, but improves
 modularity.
@@
 
-The \emph{Expression Problem} (EP)~\cite{wadler98expression} is a well-known
+As a first application for Classless Java, we illustrate a useful
+programming pattern that improves modularity and extensibility of
+programs. This useful pattern is based on an existing solution to 
+the \emph{Expression Problem} (EP)~\cite{wadler98expression}, which is a well-known
 problem about modular extensibility issues in software evolution. Recently, a
@@
 to use. The \mixin annotation is enough to remove the boilerplate code, making
-the presented approach very appealing.
+the presented approach very appealing. Our last case study, presented
+in Section~\ref{subsec:int}, is essentially a (much larger) application of this
+pattern to an existing program. Here we illustrate the pattern in the
+much smaller Expression Problem.
 
@@
 
-\subsection{Refactoring an Interpreter}
+\subsection{Refactoring an Interpreter}\label{subsec:int}
 The last case study applies our approach to a slightly larger code base of an

### 2016-03-24 OOPSLA16/sections/Overview.tex
Conclusions first pass

@@
 \section{A Running Example: Animals}\label{sec:ep}
-
 \begin{figure*}
@@
 be \emph{refined}. Such kind of refinement is challenging
-in typical class-based approaches. Fortunately, with object interfaces,
+in typical \classbased approaches. Fortunately, with object interfaces,
 we are able to provide a simple and effective solution.
 
-\paragraph{Unsatisfactory class-based solutions to field type refinement}
+\paragraph{Unsatisfactory \classbased solutions to field type refinement}
 In Java if we want to define an animal class with a field we have a set of
@@
 \begin{figure*}
+\saveSpaceFig
 \centering
-\saveSpaceFig
 \begin{tabular}{|l|l|l|l|}

### 2016-03-24 OOPSLA16/paper.tex
abstract fix

@@
 
-  This paper introduces \interfacebased (IB) \objectoriented (OO)
-  programming. Unlike \classbased OO programming, in IB programming there
-  is no need for classes. To support state, IB programming does not
+  This paper introduces \interfacebased \objectoriented 
+  programming (in short IB). Unlike class based OO programming, in IB there
+  is no need for classes. To support state, IB does not
   use fields directly. Instead it uses \emph{abstract state

### 2016-03-24 OOPSLA16/sections/Introduction.tex
IB fix

@@
 They aim to couple flexibility and rigour, expressive power and
-modular reasoning.  Two main ideas emerged to this end: prototype
-based (PB)~\cite{Ungar87self} and \classbased (CB) languages such as
-Java or Scala.  In \prototypebased
+modular reasoning.  Two main ideas emerged to this end: 
+\prototypebased (PB)~\cite{Ungar87self} and \classbased (CB) languages such as
+Java, C\# or Scala.  In \prototypebased
 languages objects inherit from other objects. Thus objects own
@@
 This paper presents a third alternative: the concept of
-\textbf{\interfacebased} (IB) \objectoriented languages, where objects 
+\textbf{\interfacebased} \objectoriented programming languages (in short IB), where objects 
 implement interfaces directly. In IB interfaces own the implementation
@@
 
-A key challenge in IB programming lies in how to model state, which is
+A key challenge in IB lies in how to model state, which is
 fundamental to have stateful objects. All abstract operations in an
@@
 programming, the structure of the state is fixed and can be only grown
-by inheritance.  In contrast, in IB programming the state is never
+by inheritance.  In contrast, in IB the state is never
 fixed, and methods that look like abstract setters and getters
@@
 years since the formulation of the problem for that
-solution to be presented in the literature. In IB programming, due to
+solution to be presented in the literature. In IB, due to
 its emphasis on type-refinement, that solution should have been much
@@
 type-refinement is that it allows a new approach to
-\emph{type-safe covariant mutable state}. That is, in IB programming,
+\emph{type-safe covariant mutable state}. That is, in IB,
 it is possible to type-refine \emph{mutable} ``fields'' in subtypes. This is
@@
 
-Finally, another advantage of IB programming is its support for
+Finally, another advantage of IB is its support for
 multiple inheritance. The literature provides good examples on how
@@
 of behaviour \textbf{and} state with multiple \emph{implementation}
-inheritance of classes. In IB programming there is only multiple
+inheritance of classes. In IB there is only multiple
 interface inheritance, yet programmers can still use state via the
-abstract state operations. IB programming enables powerful 
+abstract state operations. IB enables powerful 
 idioms using multiple inheritance and state.

### 2016-03-24 OOPSLA16/sections/Introduction.tex
CB and Scala repetition fixed

@@
 the task to \emph{object interfaces}, whose role is similar to non-abstract 
-classes in CB languages. Instantiation of objects is
+classes in CB. Instantiation of objects is
 done using static factory methods in object interfaces.
@@
 setter. Anything related to state is completely contained in the
-instances and does not leak into the inheritance logic.  In CB
-programming, the structure of the state is fixed and can be only grown
+instances and does not leak into the inheritance logic.  In CB, the structure of the state is fixed and can be only grown
 by inheritance.  In contrast, in IB the state is never
@@
 provide support for type-refinement and can automatically produce code
-that deals with type-refinement adequately. In contrast, typical CB
-languages, such as Java or Scala, require verbose explicit
-type-refinement. We believe that such verbosity hindered and slowed
+that deals with type-refinement adequately. In contrast,
+ verbose explicit type-refinement
+ is typically required in CB.
+  We believe that such verbosity hindered and slowed
 down the discovery of useful programming patterns involving
@@
 it is possible to type-refine \emph{mutable} ``fields'' in subtypes. This is
-typically forbidden in CB programming languages (including Java and
-Scala), since it is widely known that \emph{naive} type-refinement of
+typically forbidden in CB: it is widely known that \emph{naive} type-refinement of
 mutable fields is not type-safe. While covariant refinement of mutable

### 2016-03-24 OOPSLA16/sections/Introduction.tex
Improved intro in classlessJava

@@
 utility methods (such as getters and setters, or clone-like
-methods). Objects are only responsible to define the ultimate
+methods). Objects are the only responsible to define the ultimate
 behaviour of a method and, for example, if such a method is just a
@@
   Java.
+  Existing Java projects can use our approach and still be
+  backward compatible with their client, in a way that is formally specified
+  by our safety properties.
 

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
Section 5

@@
 
-\subsection{The Expression Problem with Object Interfaces}
+\subsection{The Expression Problem with Object Interfaces}\label{subsec:ep}
 
@@
 \subsection{Refactoring a Compiler}
-The last case study applies our approach to a slightly larger code base of a simple
+The last case study applies our approach to a larger code base of a simple
 one-pass compiler for \lstinline{C0} (a subset of \lstinline{C}). This compiler
@@
 \subsection{Refactoring an Interpreter}\label{subsec:int}
-The last case study applies our approach to a slightly larger code base of an
-interpreter for a Lisp-like language
+The last case study applies refactors the code from an interpreter for
+a Lisp-like language
 \lstinline{Mumbler}\footnote{https://github.com/cesquivias/mumbler/tree/master/simple},
-which is created for tutorial purpose of Truffle~\cite{wurthinger2013one}. Keeping a balance between
-simplicity and useful features, \lstinline{Mumbler} contains numbers, booleans,
-lists (encoding function calls and special forms such as if-expression, lambdas,
-etc). In the original code base, only one operation \texttt{eval} is
-supported. We add one more operation \texttt{print} to the code base. Then we
-modularize the code using our techniques to convert as many classes into
-interfaces, separate out \texttt{eval} and \texttt{print} operations then combine
-them. Although the code size is slightly increased (by 2.2\% SLOC), the
-modularity is greatly increased.
+which is created as a tutorial for the Truffle
+Framework~\cite{wurthinger2013one}.  Keeping a balance between
+simplicity and useful features, \lstinline{Mumbler} contains numbers,
+booleans, lists (encoding function calls and special forms such as
+if-expression, lambdas, etc). In the original code base, which
+consists of 626 SLOC of Java, only one operation \texttt{eval} is
+supported. Extending Mumbler to support one more operation, such as a
+pretty printer, would normally require changing the existing code base
+directly.
+
+Our refactoring applies the pattern presented in
+Section~\ref{subsec:ep} to the existing Mumbler code base to improve
+its modularity and extensibility. Using the refactored code base it is 
+becomes possible to add new operations modularly, and to support
+independent extensibility. We add one more
+operation \texttt{print} to both the original and the refactored code
+base. In the original code base the pretty printer is added
+non-modularly by modifying the existing code. In the refactored code
+base the pretty printer is added modularly.
+Although the code in the refactored version is slightly increased (by 2.2\% SLOC), the
+modularity is greatly increased, allowing for improved reusability and maintainability.

### 2016-03-24 OOPSLA16/predef/java.tex
Section 6

@@
     super,this,throw,trait,true,try,%
-    type,var,while,yield},
+    type,var,while,yield,obj,updater},
   otherkeywords={=>,<-,<\%,<:,>:,\#,@},

### 2016-03-24 OOPSLA16/sections/FutureWork.tex
Section 6

@@
 \section{Future work}\label{sec:futurework}
+
+This section discusses some possible future work, including native
+language support for IB.
+
 \begin{comment}
+
 \subsection{Encapsulation}
+
 The biggest limitation of our approach is the inability to model visibility restrictions. For example, the absence
@@
 
-As shown in~\cite{BettiniDSS13} interfaces are enough to support encapsulation/privateness by a
- variation of the facade pattern:
+A limitation of our approach, inherited from Java, is the lack of some
+support for encapsulation: in Java interfaces all methods are public.
+However, it is possible to workaround the
+limitations of interfaces by a
+ variation of the facade pattern~\cite{BettiniDSS13}:
 \begin{lstlisting}
@@
 
-This is a safe and explicit way to list all the public interfaces and to hide all the implementation
+\noindent This is a safe and explicit way to list all the public interfaces and to hide all the implementation
 details from outside of the package.\footnote{
-Note how \Q@ExposedData.of@ do not need to map directly to the fields, and can also perform
+Note how \Q@ExposedData.of@ does not need to map directly to the fields, and can also perform
 invariant checking.}
-
 Using packages as a boundary we can make all object interfaces \emph{where some form of privateness is desirable}
@@
 thus cannot be implemented from outside the package.
-
-This can raise questions: keeping fields private is a staple of CB \objectoriented programming;
-by forcing the user/heir to rely only on methods, we avoid (a part)
+This can raise questions: keeping fields private is a staple of CB.
+By forcing the user/heir to rely only on methods, we avoid (a part)
 of the fragile base class problem and we obtain representation independence.
-
-In IB fields do not exists; abstract state methods like getters and setters are abstract:
-it does not make sense to make them private; however, since they are methods,
-the base interface may be able to evolve freely.
-To understand if a fragile base interface problem exists in IB, we need more research.
+In IB fields do not exists; abstract state methods like getters and
+setters are abstract: it does not make sense to make them private;
+however, since they are methods, the base interface may be able to
+evolve freely.  To understand if a fragile base interface problem
+exists in IB, we need more research.
 
@@
 %\subsection{Clone, toString, equals and hashCode}
+
+\begin{comment}
 \paragraph{Clone, toString, equals and hashCode}
@@
 \end{lstlisting}
+\end{comment}
 
 \paragraph{A real IB language}
-In this article we presented the concept of IB as a programming pattern over Java.
+In this paper we presented the concept of IB as a programming pattern over Java.
 An alternative would be to have a language providing IB as its only paradigm.
@@
 To define such language, the first task is to define the set of abstract state operations.
-Of course we need getters; \Q@withX@ methods seams to be pretty convenient, and they enable a
-clean function programming style that would be otherwise too cumbersome.
-
+Of course we need getters; \Q@withX@ methods are also convenient, and they enable a
+clean functional programming style that would be otherwise too cumbersome.
 Currently we are offering both setters and fluent setters.
-We believe be fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
-
+We believe fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
 Manually defining abstract state methods can be verbose, we could offer a convenient syntax to
 declare in a compact way get/set/with.
-For example
+For example:
 \begin{lstlisting}
 obj interface Foo{
-  String bar;// exapanded as getter and with:
+  String bar;// expanded as getter and with:
   //String bar(); Foo withBar(String val);
@@
 
-It could be possible to offer pre initialized abstract state, for example
+\noindent It should be possible to offer pre initialized abstract state, for example
 \begin{lstlisting}
-obj interface Foo{
-  var String bar="hello";
-  }
+obj interface Foo{ var String bar="hello"; }
 \end{lstlisting}
-
-could be a short hand for writing
+\noindent could be a short hand for writing
 
@@
 obj interface Foo{
-  updater Bar, Beer//declares imperative updaters:
+  updater Bar, Beer;//declares imperative updaters:
   //Foo set(Bar val); Foo set(Beer val);
-  with Foo, Beer//declared functional updaters:
+  with Foo, Beer;//declared functional updaters:
   //Foo with(Bar val); Foo with(Beer val);
@@
 \end{lstlisting}
-Finally, certain methods may have a specific meaning, for example
-if a method called \Q@postInit@ is present, it will be called after construction,
-and it may be handled specially during multiple implementations, so that all relevant \Q@postInit@ methods are
+Finally, certain methods may have a specific meaning. For example
+if a method called \Q@postInit@ is present, it will be called after construction.
+It may be handled specially during multiple implementations, so that all relevant \Q@postInit@ methods are
 called in some order.

### 2016-03-24 OOPSLA16/sections/FutureWork.tex
Some little changes to sect 6

@@
 
-This section discusses some possible future work, including native
-language support for IB.
+%This section discusses some possible future work, including native
+%language support for IB.
 
@@
 limitations of interfaces by a
- variation of the facade pattern~\cite{BettiniDSS13}:
+ variation of the facade pattern (as suggested by~\cite{BettiniDSS13}):
 \begin{lstlisting}
@@
 We may also want to offer a compact syntax for property updaters and generalized with methods;
-\begin{lstlisting}
+\begin{lstlisting}[morekeywords={with}]
 obj interface Foo{

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
fixes on RW

@@
 \classbased languages are statically typed and they have interfaces or
-some other similar mechanisms such as traits. However, in contrast to IB
-programming, interfaces or traits in CB programming are not meant to
+some other similar mechanisms such as traits. However, in contrast to
+IB, interfaces or traits in CB are not meant to
 replace classes, but rather, they complement classes. Normally
@@
 JavaScript. Frequently \prototypebased languages are dynamically
-typed. IB programming is clearly closer to CB programming. However the
-key difference is that classes are not needed in IB
-programming. Object interfaces recover the responsibility of object
+typed. IB is clearly closer to CB. However the
+key difference is that classes are not needed in IB. Object interfaces recover the responsibility of object
 instantiation, and abstract state operations provide a way to deal
-with state. A primary motivation of IB programming is to improve
+with state. A primary motivation of IB is to improve
 modularity and flexibility. Central to these goals are the
@@
 raises the complexity, especially when constructors and fields
-are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearized mixins.
+are considered~\cite{marco09FJigsaw}. Scala traits~\cite{scala-overview} are in fact more like linearized mixins.
 Scala avoids the object initialization
@@
 number of classes, but also the class hierarchy complexity
-increases. IB programming does not complicate the hierarchical structure, and state
+increases. IB does not complicate the hierarchical structure, and state
 can also coexist with multiple inheritance.

### 2016-03-24 OOPSLA16/sections/Conclusion.tex
13 pages

@@
 (in Java) is now (almost) redundant and unneeded.
-We proposed a programming style, called ClassLess Java, where
+We proposed a programming style, called Classless Java, where
 truly \objectoriented programs and (reusable) libraries
@@
 in any Java project.
-
 The \mixin annotation helps programmers
@@
 
-Class less Java is just a programming style, but is 
-showing the way for a new flavour of object orientation:
-We propose \interfacebased \objectoriented languages (IB),
-as opposed to \classbased or \prototypebased.
+Classless Java is just a programming style, but is showing the way
+for a new flavour of object orientation: We propose \interfacebased
+\objectoriented languages (IB), as opposed to \classbased or
+\prototypebased.  In IB state is not modelled at the platonic/ideal
+level but is handled excursively by instances.  This unlocks useful
+code reuse patterns, as shown in Section~\ref{sec:ep}.
+An interesting avenue for future work would be to design a new
+language based on IB. With a proper language design we would not need 
+to restrict ourselfves to the limitations of Java and its syntax.
 
-In IB state is not modelled at the platonic/ideal level
-but is handled excursively by instances.
+%Interestingly, without classes there is also no subclassing. This scratches an old
+% itching point in the long struggle of subtyping versus subclassing:
+%  according to some authors, from a software engineering perspective,
+%  interfaces are just a kind of classes. Others consider more
+%  opportune to consider interfaces as pure types. We do not know how to conciliate
+%  those two viewpoints and Classless Java design.
+%  Classless Java does not have classes purely in the Java sense.
 
-This unlock useful code reuse patterns, as shown in~\ref{sec:ep}.
-
-Interestingly, without classes there is also no subclassing. This scratches an old
-  itching point in the long struggle of subtyping versus subclassing:
-  according to some authors, from a software engineering perspective,
-  interfaces are just a kind of classes. Others consider more
-  opportune to consider interfaces as pure types. We do not know how to conciliate
-  those two viewpoints and Classless Java design.
-  Classless Java does not have classes purely in the Java sense.  
 % Classless Java encourages coding in a more flexible way by either

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
13 pages

@@
 %\paragraph{C++ Approach.}
-C++ tries to provide a general solution to multiple inheritance by
+C++ has a general solution to multiple inheritance by
 virtual inheritance, dealing with the diamond problem by keeping only
@@
 suffers from the object initialization problem~\cite{malayeri2009cz}.
-It bypasses all constructor calls to virtual superclasses, which can
-potentially cause serious semantic errors. In our approach, the \mixin
+It bypasses constructor calls to virtual superclasses, which can
+cause serious semantic errors. In our approach, the \mixin
 annotation has full control over object initialization,
@@
 Malayeri and Aldrich proposed a model CZ~\cite{malayeri2009cz} which
-aims to do multiple inheritance without the diamond problem. They
-divide inheritance into two separate concepts: inheritance dependency
+aims to do multiple inheritance without the diamond problem. 
+Inheritance is divided into two concepts: inheritance dependency
 and implementation inheritance. Using a combination of
 \texttt{requires} and \texttt{extends}, a program with diamond
-inheritance can be transformed to one without diamonds. Moreover,
+inheritance is transformed to one without diamonds. Moreover,
 fields and multiple inheritance can coexist. However untangling
@@
 increases. IB does not complicate the hierarchical structure, and state
-can also coexist with multiple inheritance.
+also coexists with multiple inheritance.
 
@@
 constructors adds expressivity, which is not available in approaches
-such as Scala's traits/mixins.\footnote{Scala traits have
-  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
-  ``pre-initialized''.}
+such as Scala's traits/mixins.
+%\footnote{Scala traits have
+%  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
+%  ``pre-initialized''.}
 
@@
 preserving type recursion and (nominal/structural) subtyping at the same time.
-
 One famous attempt in this direction is provided by

### 2016-03-24 OOPSLA16/paper.tex
fixes on RW
typo in abstract

@@
 \usepackage{listings}
+\usepackage{xspace}
 \usepackage{multicol}
@@
 \usepackage{mathpartir}
-\usepackage{xspace}
 
@@
   new way to deal with state allows flexibility not typically
-  available in CB languages.  In IB state (including mutable state)
+  available in CB.  In IB state (including mutable state)
   can be type-refined in subtypes. The combination of a
@@
   powerful idioms using multiple inheritance and state.  To introduce
-  IB to programmers we created Classless Java: an embeding of IB
+  IB to programmers we created Classless Java: an embedding of IB
   directly into Java. Classless Java uses annotation processing for

### 2016-03-24 OOPSLA16/sections/Introduction.tex
typo in abstract

@@
 \Objectoriented languages strive to offer great code reuse.
-They aim to couple flexibility and rigour, expressive power and
+They couple flexibility and rigour, expressive power and
 modular reasoning.  Two main ideas emerged to this end: 
@@
 In \classbased languages an object is an instance of a specific class,
-and classes inherit from other classes.  Here objects own the state,
+and classes inherit from other classes.  Objects own state,
 while classes contain behaviour and the structure of the state.
@@
 This paper presents a third alternative: the concept of
-\textbf{\interfacebased} \objectoriented programming languages (in short IB), where objects 
+\emph{\interfacebased} \objectoriented programming languages (in short IB), where objects 
 implement interfaces directly. In IB interfaces own the implementation
@@
 the task to \emph{object interfaces}, whose role is similar to non-abstract 
-classes in CB. Instantiation of objects is
+classes in \classbased \objectoriented programming languages (in short
+CB). Instantiation of objects is
 done using static factory methods in object interfaces.

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
mini fix
added a table for Mumbler

@@
 reductions were due to the replacement of instantiation classes with
-generated \texttt{of} methods. The number of source lines of code (SLOC)
+generated \texttt{of} methods. The number of source lines of code 
+(SLOC\footnote{Only Java code is counted, excluding comments and blank lines. })
 was reduced by 40\% due to both the removal of instantiation overhead and
@@
 \subsection{Refactoring an Interpreter}\label{subsec:int}
-The last case study applies refactors the code from an interpreter for
+The last case study refactors the code from an interpreter for
 a Lisp-like language
@@
 non-modularly by modifying the existing code. In the refactored code
-base the pretty printer is added modularly.
-Although the code in the refactored version is slightly increased (by 2.2\% SLOC), the
+base the pretty printer is added modularly. We list the result in the following
+table:
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+\rowcolor[HTML]{C0C0C0} 
+Code               & SLOC & Code               & SLOC \\ \hline
+old (eval)         & 626  & new (eval)         & 560  \\ \hline
+old (eval + print) & 661  & new (eval + print) & 677  \\ \hline
+\end{tabular}
+\noindent Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
 modularity is greatly increased, allowing for improved reusability and maintainability.

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
spell and grammar check

--- 2016-03-24 OOPSLA16/sections/CaseStudy.tex
@@
 Classless Java. The first application shows how a useful pattern,
-using multiple inheritance and type-refinement, can be conviniently
+using multiple inheritance and type-refinement, can be conveniently
 encoded in Classless Java. The second application shows how to model
@@
 
-As a first application for Classless Java, we illustrate a useful
+As the first application for Classless Java, we illustrate a useful
 programming pattern that improves modularity and extensibility of
@@
 Section~\ref{subsec:ep} to the existing Mumbler code base to improve
-its modularity and extensibility. Using the refactored code base it is 
+its modularity and extensibility. Using the refactored code base it
 becomes possible to add new operations modularly, and to support

### 2016-03-24 OOPSLA16/sections/Implementation.tex
spell and grammar check

@@
 Thus, in addition to the advantages of Lombok, Classless Java offers
-some more advantages with respect to arbitrary (Compilation agent driven) AST rewriting.
+some more advantages with respect to arbitrary (compilation agent driven) AST rewriting.
 
@@
 theory but statistically true in practice). In practice, adding any method to any
-non final class of a Java library is enough to break heir
+non-final class of a Java library is enough to break heir
 coherence.  We think return type refinement breaks heir coherence ``less" than normal library evolution, and

### 2016-03-24 OOPSLA16/sections/Overview.tex
spell and grammar check

@@
 Here we will illustrate how points are modelled with interfaces. In IB
-we do not talk about state directly. Instead state is accessed and
+we do not talk about state directly. Instead, state is accessed and
 manipulated using abstract methods.  The usual approach to model
@@
 \paragraph{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
-2D points are mathematical entities, thus we choose immutable data structure to
+2D points are mathematical entities, thus we choose an immutable data structure to
 model them. However animals are real world entities, and when an animal moves,
@@
 operations; and functional updaters. Furthermore object interfaces support
-direct object intantiation via \texttt{of} factory methods.
+direct object instantiation via \texttt{of} factory methods.
 Figure~\ref{fig:abstractstate} summarizes the six operations supported

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
space.

--- 2016-03-24 OOPSLA16/sections/CaseStudy.tex
@@
 \begin{tabular}{ccc}
-\hline
+\hline \rowcolor[HTML]{C0C0C0}
             & SLOC   & \# of classes/interfaces \\ \hline
@@
 \subsection{Refactoring an Interpreter}\label{subsec:int}
+\begin{comment}
+The last case study refactors the code from an interpreter for
+a Lisp-like language
+\lstinline{Mumbler}\footnote{https://github.com/cesquivias/mumbler/tree/master/simple},
+which is created as a tutorial for the Truffle
+Framework~\cite{wurthinger2013one}.  Keeping a balance between
+simplicity and useful features, \lstinline{Mumbler} contains numbers,
+booleans, lists (encoding function calls and special forms such as
+if-expression, lambdas, etc). In the original code base, which
+consists of 626 SLOC of Java, only one operation \texttt{eval} is
+supported. Extending Mumbler to support one more operation, such as a
+pretty printer (making it 661 SLOC), would normally require changing the existing code base
+directly.
+
+Our refactoring applies the pattern presented in
+Section~\ref{subsec:ep} to the existing Mumbler code base to improve
+its modularity and extensibility. Using the refactored code base (560 SLOC) it is 
+becomes possible to add new operations modularly, and to support
+independent extensibility (677 SLOC). We add one more
+operation \texttt{print} to both the original and the refactored code
+base. In the original code base the pretty printer is added
+non-modularly by modifying the existing code. In the refactored code
+base the pretty printer is added modularly. 
+Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
+modularity is greatly increased, allowing for improved reusability and maintainability.
+\end{comment}
+
 The last case study refactors the code from an interpreter for
@@
 non-modularly by modifying the existing code. In the refactored code
-base the pretty printer is added modularly. We list the result in the following
-table:
+base the pretty printer is added modularly. 
+Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
+modularity is greatly increased, allowing for improved reusability and maintainability.
+We list the results in the following table:
 
-\begin{tabular}{|l|l|l|l|}
+\begin{tabular}{llll}
 \hline
@@
 Code               & SLOC & Code               & SLOC \\ \hline
-old (eval)         & 626  & new (eval)         & 560  \\ \hline
+old (eval)         & 626  & new (eval)         & 560  \\ 
+\rowcolor[HTML]{C0C0C0}
 old (eval + print) & 661  & new (eval + print) & 677  \\ \hline
 \end{tabular}
-\noindent Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
-modularity is greatly increased, allowing for improved reusability and maintainability.

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
fix

@@
 supported. Extending Mumbler to support one more operation, such as a
-pretty printer, would normally require changing the existing code base
+pretty printer \texttt{print}, would normally require changing the existing code base
 directly.
@@
 
+\vspace{0.5em}
 \begin{tabular}{llll}
@@
 Code               & SLOC & Code               & SLOC \\ \hline
-old (eval)         & 626  & new (eval)         & 560  \\ 
+old (\texttt{eval})         & 626  & new (\texttt{eval})         & 560  \\ 
 \rowcolor[HTML]{C0C0C0}
-old (eval + print) & 661  & new (eval + print) & 677  \\ \hline
+old (\texttt{eval} + \texttt{print}) & 661  & new (\texttt{eval} + \texttt{print}) & 677  \\ \hline
 \end{tabular}

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
13pages

@@
 \vspace{0.5em}
-\begin{tabular}{llll}
+\begin{tabular}{cccc}
 \hline
@@
 Code               & SLOC & Code               & SLOC \\ \hline
-old (\texttt{eval})         & 626  & new (\texttt{eval})         & 560  \\ 
+old (\texttt{eval})         & 626  & old (\texttt{eval} + \texttt{print}) & 661  \\ 
 \rowcolor[HTML]{C0C0C0}
-old (\texttt{eval} + \texttt{print}) & 661  & new (\texttt{eval} + \texttt{print}) & 677  \\ \hline
+new (\texttt{eval})         & 560  & new (\texttt{eval} + \texttt{print}) & 677  \\ \hline
 \end{tabular}

### 2016-03-24 OOPSLA16/sections/Conclusion.tex
spell and grammar check
13pages

@@
 \section{Conclusion}\label{sec:conclusion}
-
 Before Java 8, concrete methods and static methods were not allowed
@@
 language based on IB. With a proper language design we would not need
-to restrict ourselfves to the limitations of Java and its syntax.
+to restrict ourselves to the limitations of Java and its syntax.
 

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
table small fix

--- 2016-03-24 OOPSLA16/sections/CaseStudy.tex
@@
 \vspace{0.5em}
+{\small
 \begin{tabular}{llll}
@@
 Code               & SLOC & Code               & SLOC \\ \hline
-old (\texttt{eval})         & 626  & new (\texttt{eval})         & 560  \\ 
+original (\texttt{eval})         & 626  & original (\texttt{eval}+\texttt{print}) & 661  \\ 
 \rowcolor[HTML]{C0C0C0}
-old (\texttt{eval} + \texttt{print}) & 661  & new (\texttt{eval} + \texttt{print}) & 677  \\ \hline
+refactored (\texttt{eval})         & 560  & refactored (\texttt{eval}+\texttt{print}) & 677  \\ \hline
 \end{tabular}
+}

### 2016-03-24 OOPSLA16/sections/CaseStudy.tex
spellcheck

@@
 Section~\ref{subsec:ep} to the existing Mumbler code base to improve
-its modularity and extensibility. Using the refactored code base (560 SLOC) it is 
+its modularity and extensibility. Using the refactored code base (560 SLOC) it  
 becomes possible to add new operations modularly, and to support
@@
 original (\texttt{eval})         & 626  & original (\texttt{eval}+\texttt{print}) & 661  \\ 
-\rowcolor[HTML]{C0C0C0}
 refactored (\texttt{eval})         & 560  & refactored (\texttt{eval}+\texttt{print}) & 677  \\ \hline

### 2016-03-24 OOPSLA16/sections/Implementation.tex
space.
spellcheck

@@
 the \mixin annotation.
-First java source code is parsed into an abstract syntax tree (AST).
+First Java source code is parsed into an abstract syntax tree (AST).
 The AST is then captured by Lombok:
@@
 new code has to be written in \emph{new files}. Modification and/or
-reinterpretation of existing code is not supported.
+reinterpretation of existing code are not supported.
 
@@
 \centering
-\includegraphics[width=6in]{pdfs/screenshot4.png}
+\includegraphics[width=5in]{pdfs/screenshot4.png}
 \caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
@@
   example \texttt{of} methods in Figure~\ref{fig:screenshot}).
-In our case it means that either \mixin{} produces (in a controlled way) an
-understandable error, or the interface can be successfully annotated and the generated code is well-typed.
+In our case, it means that either \mixin{} produces (in a controlled way) an
+understandable error or the interface can be successfully annotated and the generated code is well-typed.
 

### 2016-03-24 OOPSLA16/sections/RelatedWork.tex
spellcheck

@@
 instantiation and handling state. \Prototypebased languages started
-with Self~\cite{Ungar87self}, and are still widely used today via languages like
+with Self~\cite{Ungar87self} and are still widely used today via languages like
 JavaScript. Frequently \prototypebased languages are dynamically
@@
 for various flavors of multiple inheritance in Java. Traits offer an algebra
-of composition operations like sum, alias and exclusion, providing explicit conflict
+of composition operations like sum, alias, and exclusion, providing explicit conflict
 resolution. Former work~\cite{bono14} provides details on mimicking the trait

### 2016-06-09 GPCE2016/sections/Appendix_Translation.tex
preparing for GPCE

--- 2015-12-09 paper/sections/Translation.tex
@@
-\section{What  \mixin Generates}\label{sec:translation}
+\section{What  \mixin Generates}\label{sec:appendixtranslation}
 
-This section shows what the \mixin annotation generates and presents a
-formal definition for most of the generated methods. Since the formalized part
-of Classless Java does not consider
-% setters, so we do not formally define the generation of the setters.
-%We do not include
-casts or \Q@instanceof@, the \Q@with@ method is not included in the
-formal translation. For the same reason \Q@void@ returning setters are
-not included, since they are just a minor variation over the more
-interesting fluent setters, and they would require special handling
-just for the conventional \Q@void@ type.
+This section presents a formal definition for most of the generated methods by $\mixinAnn$.
 
-\subsection{Translation}
-For the purposes of the formalization, the translation is divided into
-two parts for more convenient discussion on formal properties later. To this aim we introduce the annotation
-$\weakAnn$. Its role is only in the translation process, hence is
-not part of the Classless Java language.  $\weakAnn$ generates the
-constructor method $\QM{of}$, while \mixin automatically refines the
-return types and calls $\weakAnn$.
+\subsection{Translation}\label{subsec:translation}
 
-\noindent$\begin{array}{l}
-\InTextDef{16em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
-}{
-[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
-\methods\ \methods' \cC
-]\!]}\\
-\InTextWith{\methods'=\otherMethod(\C_0,\methods)}\\
-
-\InTextDef{16em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
-}{
-\QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
-\methods\ \ofMethod(\C_0) \cC
-}\\
-\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\C_0) }
-\end{array}$
-
-\noindent Note that it is necessary to explicitly check if the interface is valid
+The translation functions of $\mixinAnn$ and $\weakAnn$
+are presented in Figure~\ref{figure:translation}. Note that
+it is necessary to explicitly check if the interface is valid
 for annotation:
@@
 \InTextDef{7ex}{\valid(\C_0)}{\forall \m\in\dom(\C_0),\mbox{ if }\mh\QM; =
-            \mBody(\m, \C_0),\mbox{ one of the following cases is satisfied:}}\\
+            \mBody(\m, \C_0),}\\
+            \hspace{.24in}\mbox{ one of the following cases is satisfied:}\\
 \tab\tab
@@
 \mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
-\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
-\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
-\ \mand\ \mnot\ \specialName(\m)}\\
+\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{}\\
+\hspace{.3in}\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)\\
+\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{}\\
+\hspace{.3in}\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
+\ \mand\ \mnot\ \specialName(\m)\\
 
@@
 \InTextDef{16em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{
- \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
-\hspace{2.65in}{
-\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
+\otherMethod(\C_0,\methods)}{}\\
+\hspace{.25in}\isWith(\mBody(\QM{with#}\m, \C_0), \C_0),\ 
+\QM{with#}\m\notin\dom(\methods)\\
 \InTextDef{16em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
-\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
-\hspace{2.65in}{
-\mbox{ and } \QM_\m\notin\dom(\methods)}\\
+\otherMethod(\C_0,\methods)}{}\\
+\hspace{.25in}\isSetter(\mBody(\QM_\m, \C_0), \C_0),\ \QM_\m\notin\dom(\methods)\\
 \end{array}$
@@
 \QM{of}.
-
 \noindent$\begin{array}{l}
@@
  \QM{static}\ \C_0\ \QM{of} \oR \C_1\ \QM_\m_1\QM,\ldots \C_n\ \QM_\m_n\cR\
-\QM{\{}
-\QM{return new}\ \C_0 \oR\cR\ \QM{\{} }\\
+\QM{\{}}\\
+\tab\QM{return new}\ \C_0 \oR\cR\ \QM{\{}\\
 \tab\tab \C_1\ \m_1 = \QM_\m_1\QM;\ldots \C_n\ \m_n = \QM_\m_n\QM; \\
@@
 \tab\QM{\};\}} \\
-\InTextWith{\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();} = \fieldsFunc(\C_0) 
-\mbox{ and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n}
+\InTextWith{\C_1\ \m_1\QM{();},\ldots \C_n\ \m_n\QM{();} = \fieldsFunc(\C_0)}\\
+\hspace{.5in}\mbox{ and }\es_i=\m_1\QM,\ldots\QM, \m_{i-1}\QM,\QM{_val,}\m_{i+1}\QM,\ldots\QM, \m_n
 \end{array}$
@@
 \noindent$\begin{array}{ll}
-\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
-\isField(\method)\ \mand\
-%}\\\InTextWith{
+\InTextDef{16ex}{\method\in\fieldsFunc(\C_0)}{}\\
+\hspace{.3in}\isField(\method)\ \mand\
 \method=\mBody(m^\method,\C_0)
-}
 \end{array}$
@@
 \noindent$\begin{array}{l}
-\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{
-\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
-\QM{return}\ \C_0\QM{.of(}\es\QM{);\}}} \\
+\InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{}\\
+\hspace{.2in}\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+\QM{return}\ \C_0\QM{.of(}\es\QM{);\}} \\
 \InTextWith{\mBody(\QM{with#}\m,\C_0) \mbox{ having the form }\mh\QM;}\\
 \InTextDef{10em}{\withMethod(\C,\m,\C_0,\es)}{\emptyset\mbox{ otherwise}}\\
-\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{
-\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
- \m\QM{= _val;return this;\}}} \\
+\InTextDef{10em}{\setterMethod(\C,\m,\C_0)}{}\\
+\hspace{.2in}\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\ \QM{\{}
+ \m\QM{= _val;return this;\}} \\
 \InTextWith{
@@
 
-\subsection{Other Features}
+\subsection{Other Features}\label{subsec:otherfeatures}
 
@@
 \end{itemize}
-
-\subsection{Guarantees}
-To understand to what extent our approach is correct (and more in general, what it means to say that a language tuning is correct) we identify three types of guarantees:
-\begin{itemize}
-\item \textit{Self coherence}: the generated code itself is well-typed; that is,
-  type errors are not present in code the user has not written.
-In our case it means that either \mixin{} produces (in a controlled way) an
-understandable error, or the interface can be successfully annotated and the generated code is well-typed.
-We guarantee \textit{Self coherence}.
-
-\item \textit{Client coherence}: all the client code (for example method calls)
-  that is well-typed before code generation is also well-typed after the generation.
-The annotation just adds more behavior without removing any functionality.
-We guarantee \textit{Client coherence}.
-
-\item \textit{Heir coherence}: interfaces (and in general classes) inheriting the instrumented code are well-typed if they were well-typed without the instrumentation.
-This would forbid adding any (default or abstract) method to the annotated interfaces, including type refinement.
-\mixin  does not guarantee \textit{Heir coherence}.
-Indeed consider the following:
-
-\begin{lstlisting}
-interface A { int x(); A withX(int x); }
-@Obj interface B extends A {}
-interface C extends B { A withX(int x); }
-\end{lstlisting}
-
-\noindent This code is correct before the translation, but \mixin would  generate in \Q@B@  a method ``\Q@B withX(int x);@''.
-This would break \Q@C@. \\*
-Similarly, an expression of the form ``\Q@new B(){.. A withX(int x){..}}@''
-would be correct before translation, but would be ill-typed after the translation.
-\end{itemize}
-
-\noindent This means that our automatic type refinement
-is a useful and convenient feature, but not transparent to the heirs of the annotated interface. They need to be aware of the annotation semantics and provide the right type while refining methods.
-
-\subsection{Results}
-To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
-\begin{itemize}
-\item We denote with $\C^{\II}$ the name of an interface.
-\item An interface table
-IT is OK if under such interface table, all interfaces are OK.
-\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
-\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
-\end{itemize}
-
-\begin{thm}[@ObjOf tuning]
-If a given interface table $\II$ IT is OK
- where $\II$ has $\weakAnn$,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,
-then the interface table $[\![\II]\!]$ IT is OK.
-\end{thm}
-
-\begin{thm}[@Obj tuning]
-If a given interface table $\II$ IT is OK
- where $\II$ has $\mixinAnn$,
-$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$, and there is no heir of $\C^{\II}$,
-then the interface table $[\![\II]\!]$ IT is OK.
-\end{thm}
-
-Informally, the theorems mean that for a client program that
-typechecks before the translation is applied, if the annotated type has
-no subtypes and no objects of that type are created, then type safety
-is guaranteed after the successful translation.
-
-The second
-step of \mixin, namely what $\weakAnn$ does in the formalization, is
-guaranteed to be type-safe for the three kinds of coherence by the $\weakAnn$  tuning theorem.
-\mixin
-tuning is more interesting: since \mixin does not guarantee heir
-coherence, we explicitly exclude the presence of heirs. In this way
-\mixin tuning guarantees only self and client coherence. The formal
-theorem proofs are available in the supplementary materials.%~\ref{subsec:lemma1},~\ref{subsec:lemma2} and~\ref{subsec:theorem}.\\
-
-\begin{comment}
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=5in]{screenshot.png}
-\caption{Screenshot.}\label{screenshot_png}
-\end{figure}
-
-\haoyuan{I tried to understand the current algorithm, and did more experiments in eclipse.
-Now I borrow some ideas from the current version, and give a new version of the algorithm in text. See below.
-
-(1) I guess the function \textsf{tops} is not necessary. The first step is still
-\[\textsf{mbody}(m,C_i)\in\overline{meth}\textrm{ (excluding \textbf{static} methods)}\]
-
-(2) Assume the context is ``interface $C_0$ extends $\overline{C}$ \{$meth'$;...\}''. First handle
-\[\textsf{override}(meth',\overline{meth}) \eqno{(*)}\]
-
-(3) If $meth'\ne\none$, $(*)$ returns $meth'$ if
-\[\forall meth\in\overline{meth},meth'\subtype meth\]
-even if there are conflicts in $\overline{meth}$.
-
-(4) If $meth'=\none$, we need to figure out
-\[\textsf{mostSpecific}(\overline{meth})\]
-and it should be the one that ``overrides'' all the others in $\overline{meth}$.
-It means we should not only deal with the return types of methods, but also look into the subtyping relation of interfaces.
-But for abstract methods, only return types are taken into consideration.
-}
-\end{comment}
-
-%\text{\yanlin{shouldn't mostSpecific be: $\forall \method' \in \methods : \method \subtype
-%  \method'$ ?}}
-
-%(2) If $body_1.\textsf{returnType}=body_2.\textsf{returnType}$, \textsf{shadow} tends to return a default method. If both $body_1$ and $body_2$ are default methods, \textsf{shadow} throws an error.
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{default}\\
-%\textsf{shadow}(body_1, body_2)=body_1 \hspace{.1in} & \textsf{if }body_1.\textsf{modifier}=\textbf{default} \\
-%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in} & \textsf{if }body_2.\textsf{modifier}=\textbf{default} \\
-%\textsf{shadow}(body_1, body_2)=body_1\textsf{ (or }body_2\textsf{)} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}
-%
-%(3) If $body_1.\textsf{returnType}<:body_2.\textsf{returnType}$, \textsf{shadow} tends to choose the one with the subtype (namely $body_1$), but only when both methods are abstract, otherwise it gives an error.
-The other direction $body_2.\textsf{returnType}<:body_1.\textsf{returnType}$ follows the same rule.
-It also gives an error if there is no subtyping relationship between two return types.
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-%& \textsf{and }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-%\textsf{shadow}(body_1, body_2)=body_2 & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\emptyset\\
-%& \textsf{and }body_2.\textsf{returnType}<:body_1.\textsf{returnType}\\
-%\textsf{shadow}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}
-
-%\subsubsection{Auxiliary function: \textsf{replace}}
-%
-%The \textsf{replace} function takes two same methods (with the same name and types of arguments), and gives the result of the first method overriding the second one.
-%
-%\begin{equation*}
-%\begin{array}{ll}
-%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\\
-%\textsf{replace}(body_1, body_2)=body_2 & \textsf{if }body_1=\emptyset\\
-%\textsf{replace}(body_1, body_2)=body_1 & \textsf{if }body_1.\textsf{returnType}<:body_2.\textsf{returnType}\\
-%\textsf{replace}(body_1, body_2)=\textsf{ERROR} \hspace{.1in} & \textsf{otherwise}
-%\end{array}
-%\end{equation*}

### 2016-06-09 GPCE2016/sections/FutureWork.tex
preparing for GPCE

--- 2016-03-24 OOPSLA16/sections/FutureWork.tex
@@
 of the fragile base class problem and we obtain representation independence.
-In IB fields do not exists; abstract state methods like getters and
+In IB fields do not exist; abstract state methods like getters and
 setters are abstract: it does not make sense to make them private;
@@
 Currently we are offering both setters and fluent setters.
-We believe fluent setters to be sufficient; we could drop the concept of (non fluent)setters.
+We believe fluent setters to be sufficient; we could drop the concept of (non-fluent) setters.
 Manually defining abstract state methods can be verbose, we could offer a convenient syntax to

### 2016-06-09 GPCE2016/sections/Introduction.tex
preparing for GPCE

--- 2016-03-24 OOPSLA16/sections/Introduction.tex
@@
 Using Java annotation processors, we produce an implementation of
-Classless Java, which allow us to stick to pure Java 8. By annotating
+Classless Java, which allows us to stick to pure Java 8. By annotating
 with \Q|@Obj| the interfaces that represent object interfaces, we can
@@
 As a result, we experiment object interfaces with several interesting
-Java programs, and conduct various case studies.  Finally, we also
+Java programs and conduct various case studies.  Finally, we also
 formally define the behaviour of our \mixin annotation and discuss

### 2016-06-09 GPCE2016/experiments/Separate_Compilation/Test1.java

@@ -1,0 +1,17 @@
+import java.lang.reflect.Method;
+
+
+public class Test1 {
+	
+	public int n() { return 1; }
+	
+	public static void main(String[] args) {
+		Method[] res = Test2.class.getMethods();
+		if (res != null) {
+			for (Method m : res) {
+				System.out.println(m.getName()); // m, n included
+			}
+		}
+	}
+	
+}

### 2016-06-09 GPCE2016/sections/Introduction.tex
Revised abstract and intro

@@
 modular reasoning.  Two main ideas emerged to this end:
-\prototypebased (PB)~\cite{Ungar87self} and \classbased (CB) languages such as
+\prototypebased (PB)~\cite{Ungar87self} and \classbased languages such as
 Java, C\# or Scala.  In \prototypebased
@@
 This paper presents a third alternative: the concept of
-\emph{\interfacebased} \objectoriented programming languages (in short IB), where objects
+\emph{\interfacebased} \objectoriented programming languages (IB), where objects
 implement interfaces directly. In IB interfaces own the implementation
@@
 the task to \emph{object interfaces}, whose role is similar to non-abstract
-classes in \classbased \objectoriented programming languages (in short
-CB). Instantiation of objects is
-done using static factory methods in object interfaces.
+classes in \classbased \objectoriented programming languages
+(CB). Objects 
+are instantiated by static factory methods in object interfaces.
 
@@
 down the discovery of useful programming patterns involving
-type-refinement. In particular, a recently discovered
+type-refinement. A recently discovered
 solution~\cite{eptrivially} to the Expression
@@
 
-One important advantage of the use of abstract state operations and
+One advantage of abstract state operations and
 type-refinement is that it allows a new approach to
@@
 
-Finally, another advantage of IB is its support for
-multiple inheritance. The literature provides good examples on how
-easy and modular it is to combine multiple sources of pure behaviour,
+Another advantage of IB is its support for
+multiple inheritance. The literature is rich on how
+easy it is to combine multiple sources of pure behaviour,
 using mechanisms such as traits~\cite{scharli03traits}. In Java
@@
 inheritance of classes. In IB there is only multiple
-interface inheritance, yet programmers can still use state via the
+interface inheritance, yet programmers can still use state via 
 abstract state operations. IB enables powerful
@@
 implementations. In the context of Java, what we propose is a programming
-style, where we never use classes\footnote{More precisely, we never
-  use the \Q@class@ keyword.}.  We call this restricted version of
+style, where we never use classes (more precisely, we never
+  use the \Q@class@ keyword).  We call this restricted version of
 Java \emph{Classless Java}.
@@
 Java programs and conduct various case studies.  Finally, we also
-formally define the behaviour of our \mixin annotation and discuss
+discuss the behaviour of our \mixin annotation and 
 its properties.
@@
 would not have. In particular, supporting proper encapsulation is
-difficult in Java due to limitations of Java interfaces. We discuss
-these limitations, possible workarounds, and native language support in Section~\ref{sec:futurework}.
+difficult in Java due to limitations of Java interfaces. In particular
+in Java interfaces the visibility of all methods is public. Thus
+modelling private state is difficult in current Java 8. However,
+ using existing design patterns~\cite{BettiniDSS13} we can emulate hiding
+methods from interfaces. Furthermore Java 9 will allow private methods
+on interfaces~\cite{srikanth16private}, greatly improving the support for
+encapsulation.
+
+%We discuss
+%these limitations, possible workarounds, and native language support in Section~\ref{sec:futurework}.
 
@@
 
-\item {\bf IB and Object Interfaces:} a novel take on object orientation, allowing
+\item {\bf IB and Object Interfaces:} which enable
   powerful programming idioms using multiple-inheritance,
@@
   allowing most tools to work transparently with our approach.
-  We provide examples, informal and formal descriptions of Classless
-  Java.
   Existing Java projects can use our approach and still be
-  backward compatible with their client, in a way that is formally specified
+  backward compatible with their clients, in a way that is specified
   by our safety properties.

### 2016-06-20 GPCE2016/sections/CaseStudy.tex
compressing down to 10 pages.

--- 2016-03-24 OOPSLA16/sections/CaseStudy.tex
@@
 \paragraph{Independent Extensibility}
-To show that our approach supports independent extensibility, we first define a
-new operation \texttt{collectLit}, which collects all
-literal components in an expression. For space reasons,
-we omit the definitions of the methods:
+To show that our approach supports independent extensibility, a
+new operation \texttt{collectLit} which collects all
+literal components in an expression is defined. For space reasons,
+we omit some code:
 
@@
 
-\noindent Now we combine the two extensions (\texttt{print} and
-\texttt{collectLit}) together:
+\noindent Now we combine the two extensions together:
+%(\texttt{print} and \texttt{collectLit})
 
@@
 
-Note that the code for instantiation is automatically generated by \mixin.
+Note that the code for instantiation is generated by \mixin.
 Creating a simple expression of type \texttt{ExpPC} is
@@
 Since the style of fluent interfaces was invented in Smalltalk as method
-cascading, more and more languages came to support fluent interfaces, including
-JavaScript, Java, C++, D, Ruby, Scala, etc. In most languages, to create fluent
+cascading, more and more languages (Java, C++, Scala, etc) came to support fluent interfaces. In most languages, to create fluent
 interfaces, programmers have to either hand-write everything or create a wrapper
@@
 fluent APIs. However most of them only provide a fixed set of predefined fluent
-interfaces. Fluflu enables the creation of a fluent API and implements control
-over method chaining by using Java annotations. However methods that returns
-\textbf{this} are still hand-written.
+interfaces. %Fluflu enables the creation of a fluent API and implements control
+% over method chaining by using Java annotations. However methods that returns
+% \textbf{this} are still hand-written.
 
@@
 \vspace{5pt}
+{\small
 \begin{tabular}{ccc}
@@
 \end{tabular}
+}
 \vspace{5pt}

### 2016-06-20 GPCE2016/sections/Implementation.tex
compressing down to 10 pages.

--- 2016-03-24 OOPSLA16/sections/Implementation.tex
@@
 %  libraries, but it would be a taunting task.}
-Creating a new language or a language extension would be an
+Creating a new language/extension would be an
 elegant way to illustrate the point of IB. However,
 significant amounts of engineering would be needed to build a practical
-language, and to achieve a similar level of integration and tool support
-as Java. To have an approach that can both illustrate IB
-programming and be practical, we have instead implemented
-\mixin as an annotation over Java 8, and a \emph{compilation agent}.
+language and achieve a similar level of integration and tool support
+as Java. To be practical, %To have an approach that can both illustrate IB programming and be practical, 
+we have instead implemented
+\mixin as an annotation in Java 8, and a \emph{compilation agent}.
 That is, the Classless Java style of programming
-is supported by a Java 8 library.
+is supported by library.
 
-Disciplined use of Classless Java (that is avoiding class
-declarations as done in Section~\ref{sec:ep}), illustrates what \emph{pure} IB is like.
+Disciplined use of Classless Java (avoiding class
+declarations as done in Section~\ref{sec:ep}) illustrates what \emph{pure} IB is like.
 However, using \mixin, CB and IB programming can be mixed together,
 harvesting the practical convenience of using existing Java libraries, the full
-Java language, and IDE's support.
-The key to our implementation is the use of compilation agents, which
+Java language, and IDE support.
+The key to our implementation is compilation agents, which
  allows us to rewrite the Java AST just
@@
 acting as a man in the middle between the
-generation of the abstract syntax tree and the Java bytecode.
+generation of AST and bytecode.
 
 This process is facilitated by frameworks like Lombok~\cite{lombok}:
-a Java library that aims at removing (or
-reducing) Java boilerplate code via
+a Java library that aims at reducing Java boilerplate code via 
 annotations. \mixin was created using Lombok.
@@
 \paragraph{Advantages of Lombok}
-The Lombok compilation agent has advantages with respect to alternatives like
-pre-processors, or other Java annotation processors.
-Lombok offers in Java an expressive power similar to that of Scala/Lisp macros;
-except of course, for the syntactic convenience of quote/unquote templating.
+The Lombok compilation agent has advantages over 
+pre-processors or other Java annotation processors.
+Lombok offers in Java an expressive power similar to that of Scala/Lisp macros,
+except, for the syntactic convenience of quote/unquote templating. Compilation
+agents act modularly on each compilation unit, applying transformations to one 
+annotated class/interface at a time. This allows library code to be reused
+without the need of being reprocessed and recompiled, making our
+approach 100\% compatible with existing Java libraries. 
+In Eclipse, the processing is 
+performed transparently and the information of the interface from
+compilation is captured in the ``Outline'' window as shown in Figure~\ref{fig:screenshot}..
+This includes all the methods inside the interface as well as the generated ones.
 
+\begin{figure*}[t]
+\saveSpaceFig
+\saveSpaceFig
+\centering
+\includegraphics[width=4.5in]{pdfs/screenshot4.png}
+\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
+\label{fig:screenshot}
+\saveSpaceFig
+\end{figure*}
+
+\begin{comment}
 \paragraph{Direct modification of the AST}
@@
 
-\begin{figure*}[t]
-\saveSpaceFig
-\saveSpaceFig
-\centering
-\includegraphics[width=5in]{pdfs/screenshot4.png}
-\caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
-\label{fig:screenshot}
-\saveSpaceFig
-\end{figure*}
-
 \paragraph{Tool support}
@@
 tools to work transparently.
+\end{comment}
 
@@
 form of heir coherence.
-
 Section~\ref{sec:translation} provide a formal definition for our
@@
 \begin{itemize}
-\item Eclipse has its own compiler (ejc), alternative to javac.
-  Lombok allows writing handlers for either javac or ejc.
-  At this stage our implementation only realizes the Eclipse handler and our
-  experiments are all conducted in Eclipse. The implementation for
-  the \texttt{javac} handler is still missing.
-\item The current implementation supports generics in a simplified manner:
-type parameters can be used but the generic
-method typing is neither formalized in this paper nor explicitly checked by our
-implementation, but simply  delegated to the Java compiler.
+\item Lombok allows writing handlers for either javac or ejc(Eclipse's own compiler).
+Our current implementation only realizes ejc version. The implementation for
+  the \texttt{javac} version is still missing.
+\item Simple generics is supported:
+type parameters can be used, but generic
+method typing %is neither formalized in this paper nor 
+not explicitly checked by \mixin, but simply delegated to the Java compiler.
 \item
@@
 to access information of code defined in different files.
-In the same spirit, we have a mature \mixin annotation, which
-does not support separate compilation yet: it requires that all
+Limited by Lombok, \mixin 
+%In the same spirit, we have a mature \mixin annotation, which
+%does not support separate compilation yet: it 
+requires that all
   related interfaces have to appear in a single Java file.
-Reusing the logic inside the yet experimental Lombok annotation \lstinline{@Delegate},
+Reusing the logic inside the experimental Lombok annotation \lstinline{@Delegate},
 we also offer a less polished annotation supporting

### 2016-06-20 GPCE2016/sections/Overview.tex
preparing for GPCE
compressing down to 10 pages.

--- 2016-03-24 OOPSLA16/sections/Overview.tex
@@
 \begin{tabular}{c|c}
-\includegraphics[height=4cm]{pdfs/PegasusDetail.pdf}\hspace{20pt} &
+\includegraphics[height=3.3cm]{pdfs/PegasusDetail.pdf}\hspace{0pt} &
 \begin{minipage}{7cm}
-\vspace{-90pt}
-\lstinputlisting[linerange=6-17]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
+\vspace{-95pt}
+\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
 %basicstyle=\ttfamily\scriptsize
@@
 \caption{The animal system (left: complete structure, right: code for simplified animal system).}\label{fig:pegasus}
+\saveSpaceFig
 \end{figure*}
 
-This section illustrates how our programming style, supported by the
-\mixinAnn{} annotation, enables powerful programming idioms based on multiple
+This section illustrates how our programming style, supported by 
+\mixinAnn{}, enables powerful programming idioms based on multiple
 inheritance and type refinements.  We propose a standard example:
-\Q@Animal@s with a two-dimensional \Q@Point2D@ representing their
+\Q@Animal@s with a 2-dimensional \Q@Point2D@ representing their
 \Q@location@.  Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
-Birds can \Q@fly@, thus their locations need to be three-dimensional
+Birds can \Q@fly@, thus their locations need to be 3-dimensional
 \Q@Point3D@s (field type refinement). Finally, we model \Q@Pegasus@
-(one of the best-known creatures in Greek mythology) as a kind of
+(a well-known creature in Greek mythology) as a kind of
 \Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s (multiple
@@
 Before modelling the complete animal system, we  start with a
-simplified version without locations. This version serves the purpose of illustrating how
+simple version without locations. This version serves the purpose of illustrating how
 Java 8 default methods can already model simple forms of multiple inheritance.
@@
 functionality from both \texttt{Horse} and \texttt{Bird}.
-A first attempt to model the animal system in Java 8 is given on the right side
+A first attempt to model the animal system is given on the right side
 of Figure~\ref{fig:pegasus}.
@@
 
-\lstinputlisting[linerange=21-23]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
+\lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
 
@@
 repetitive code can be encapsulated in a method. A generalization of the
-\texttt{of} static factory method is appropriate in this case:
+\texttt{of} static factory method is appropriate:% in this case:
 \begin{lstlisting}
@@
 \paragraph{\Q@Point2D@ with object interfaces}
-This obvious ``constructor'' code can be automatically generated by the \mixin
+This obvious ``constructor'' code can be generated by the \mixin
 annotation.  By annotating the interface \Q@Point2D@, a variation of the shown
@@
 \begin{figure}
-%\saveSpaceFig
+\saveSpaceFig
 \lstinputlisting[linerange=89-115]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
@@
 \label{fig:boilerplate}
+\saveSpaceFig
 \end{figure}
@@
 through interface \texttt{Horse} and \texttt{Bird}. Using \mixin annotation,
-actually these is no code at all programmers have to write. The idea of using
+actually there is no code that programmers have to write at all. The idea of using
 default methods inside interfaces was proposed in ~\cite{}. It enables us to do

### 2016-06-20 UseMixinLombok/src/pegasus/simple/java8/Main.java

@@ -5,14 +5,10 @@ import static java.lang.System.out;
 //BEGIN_PEGASUS_JAVA
 interface Animal {} // no points yet!
 interface Horse extends Animal {
-    default void run() {
-        out.println("running!");
-    }
+    default void run(){out.println("run!");}
 }
 interface Bird extends Animal {
-    default void fly() {
-        out.println("flying!");
-    }
+    default void fly(){out.println("fly!");}
 }
 interface Pegasus extends Horse, Bird {}
 //END_PEGASUS_JAVA

### 2016-06-21 GPCE2016/sections/Introduction.tex
revise. refer to appendix material in external link.

@@
 utility methods (such as getters and setters, or clone-like
-methods). Objects are the only responsible to define the ultimate
-behaviour of a method and, for example, if such a method is just a
-setter. Anything related to state is completely contained in the
+methods). Objects are only responsible to define the ultimate
+behaviour of a method. % and, for example, if such a method is just a setter. 
+Anything related to state is completely contained in the
 instances and does not leak into the inheritance logic.  In CB, the structure of the state is fixed and can be only grown
 by inheritance.  In contrast, in IB the state is never
-fixed, and methods that look like abstract setters and getters
+fixed, and methods such as abstract setters and getters
 can always receive an explicit implementation down in the inheritance
@@
 Our IB embedding relies on the
-new features of Java 8 for interfaces: interface \emph{static methods}; and
+new features of Java 8: interface \emph{static methods} and
 \emph{default methods}, which allow interfaces to have method
@@
 Classless Java, which allows us to stick to pure Java 8. By annotating
-with \Q|@Obj| the interfaces that represent object interfaces, we can
-automatically generate code for interface instantiation and
-type refinement. Such code should not be needed in the first place in
+ the interfaces that represent object interfaces with \mixin, code for interface instantiation and
+type refinement can be automatically generated. Such code should not be needed in the first place in
 a real IB language, and the annotation processor allows us to
@@
   allowing most tools to work transparently with our approach. 
+  We provide the formal semantics here~\footnote{See url?? Formal semantics for Classless Java is provided in Section A}.
   Existing Java projects can use our approach and still be

### 2016-06-24 GPCE2016/sections/CaseStudy.tex
Marco improvements.
I can not remember the macro to add comments, used @@@

@@
 
+{\bf @@@@@@@@@@@@FIX HERE, add AddPC and LitPC @@@@@@@}
+
 \noindent \texttt{ExpPC} is the new expression interface supporting
@@
 \lstinputlisting[linerange=39-39]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
+
+Smalltalk and Dart offer method cascading~[??];
+at the price of introducing additional syntax,
+this avoid the need for fluent setters;
+method cascading (as setters in general) 
+intrinsically rely  on an imperative setting.
+Our approach supports both fluent setters and (functional) 
+fluent withers.
 

### 2016-06-24 GPCE2016/sections/Implementation.tex
revise. refer to appendix material in external link.
Marco improvements.
I can not remember the macro to add comments, used @@@

@@
 form of heir coherence.
-Section~\ref{sec:translation} provide a formal definition for our
-safety claims.
+%Section~\ref{sec:translation} 
+We provide a formal definition for our safety claims here~\footnote{See url?? Section B}.
 
@@
 \item
+{\bf @@@English need improvement here@@@@@@@}
 Lombok offers only limited/experimental support for separate compilation, that is

### 2016-06-24 GPCE2016/sections/Introduction.tex
Marco improvements.
I can not remember the macro to add comments, used @@@

@@
 
-In the presence of subtyping, abstract state operations often require
-special care, as their types need to be refined. Object interfaces
-provide support for type-refinement and can automatically produce code
-that deals with type-refinement adequately. In contrast,
- verbose explicit type-refinement
- is typically required in CB.
+Object interfaces
+provide support for automatic type-refinement.
+In contrast, in CB 
+special care and verbose explicit type-refinement is be required to
+produce code that deals with subtyping adequately.
+
   We believe that such verbosity hindered and slowed

### 2016-06-24 GPCE2016/sections/Overview.tex
revise. refer to appendix material in external link.
Marco improvements.
I can not remember the macro to add comments, used @@@

@@
 \Q@Animal@s with a 2-dimensional \Q@Point2D@ representing their
-\Q@location@.  Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
+\Q@location@, subtypes \Q@Horses@, \Q@Bird@s, and \Q@Pegasus@.
 Birds can \Q@fly@, thus their locations need to be 3-dimensional
-\Q@Point3D@s (field type refinement). Finally, we model \Q@Pegasus@
+\Q@Point3D@s (field type refinement). We model \Q@Pegasus@
 (a well-known creature in Greek mythology) as a kind of
@@
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-Here we will illustrate how points are modelled with interfaces. In IB
-we do not talk about state directly. Instead, state is accessed and
+Here we will illustrate how points are modelled with interfaces.
+ In IB
+ state is accessed and
 manipulated using abstract methods.  The usual approach to model
@@
 
-\noindent The encoding over Java is now inconvenient: creating a new point object is cumbersome, even
+\noindent The encoding over pure Java is now inconvenient: creating a new point object is cumbersome, even
 with anonymous classes:
@@
 \noindent More precisely, a field or factory parameter is generated for every
-abstract method that takes no parameters (except for methods with special
-names). An example of using \Q@Point2D@ is:
+abstract method that takes no parameters.
+%(except for methods with special
+%names).
+ An example of using \Q@Point2D@ is:
 \begin{lstlisting}
@@
 \end{comment}
-Moreover, if the programmer wants a different implementation, he may
+Moreover, if the programmer wants a different implementation, they may
 provide an alternative implementation using \Q@default@ methods. For example:
@@
     default Point2D withX(int val){ /*myCode*/ }
-    default Point2D withY(int val){ /*myCode*/ } }
+    Point2D withY(int val); }
 \end{lstlisting}
@@
     default Point2D withX(int val){ /*myCode*/ }
-    default Point2D withY(int val){ /*myCode*/ }
+    Point2D withY(int val);
     static Point2D of(int _x, int _y){
@@
         public int x(){return x;}
-        public int y(){return y;} }; } }
+        public int y(){return y;} 
+        public Point2D withY(int val){
+          return of(x(),val);}  }; } }
 \end{lstlisting}
@@
 
+\paragraph{Property updater as Injections}
+
+By declaring fields in Newspeak and other languages,
+the programmers specifies field signatures and (critically) the intention of storing such information. Then the language generate getter and setter.
+Once state is abstracted away,
+it is well known that is possible to replace state
+access with computation, but the type of the field stays the same.
+
+Our fluid approach to state is specular:
+the intention of storing the information is not expressed by the programmer and set in stone, 
+but can vary by inheritance:
+We generate the idea of the field
+just starting from getters and setters signatures.
+
+In this case, we leverage on our fluid state to changing the underling type of the field,
+ and our \Q@with@ is providing
+the right injection from the old type to the new one.
+
 \subsection{Advanced Multiple Inheritance}

### 2016-06-28 GPCE2016/sections/CaseStudy.tex
fix table overflow
revise. fixed comments by Marco.

@@
 
-\lstinputlisting[linerange=88-]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
-
-{\bf @@@@@@@@@@@@FIX HERE, add AddPC and LitPC @@@@@@@}
+\lstinputlisting[linerange=88-90]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY
 
@@
 \vspace{5pt}
-{\small
+{\fontsize{8}{10}\selectfont
 \begin{tabular}{ccc}
@@
 \vspace{0.5em}
-{\small
+{\fontsize{8}{10}\selectfont
 \begin{tabular}{llll}

### 2016-06-28 UseMixinLombok/src/casestudy/ep/TestExpression.java

@@ -88,3 +88,4 @@ interface ExpC extends Exp { List<Integer> collectLit(); }
 interface ExpPC extends ExpP, ExpC {}
 @Obj interface LitPC extends ExpPC, LitP, LitC {}
 @Obj interface AddPC extends ExpPC, AddP, AddC { ExpPC e1(); ExpPC e2(); }
+//END_INDEPENDENT_EXTENSIBILITY

### 2016-06-29 GPCE2016/sections/Introduction.tex
revise. fixed comments by Marco.
merge

@@
 
-Object interfaces
-provide support for automatic type-refinement.
-In contrast, in CB 
-special care and verbose explicit type-refinement is be required to
-produce code that deals with subtyping adequately.
-
-  We believe that such verbosity hindered and slowed
-down the discovery of useful programming patterns involving
-type-refinement. A recently discovered
-solution~\cite{eptrivially} to the Expression
-Problem~\cite{wadler98expression} in Java-like languages, shows how easy it is to solve
-the problem using only type-refinement. However it took nearly 20
-years since the formulation of the problem for that
-solution to be presented in the literature. In IB, due to
-its emphasis on type-refinement, that solution should have been much
-more obvious.
+Object interfaces provide support for automatic type-refinement.  In
+contrast, in CB special care and verbose explicit type-refinement is
+required to produce code that deals with subtyping adequately. We
+believe that such verbosity hindered and slowed down the discovery of
+useful programming patterns involving type-refinement. A recently
+discovered solution~\cite{eptrivially} to the Expression
+Problem~\cite{wadler98expression} in Java-like languages, shows how
+easy it is to solve the problem using only type-refinement. However it
+took nearly 20 years since the formulation of the problem for that
+solution to be presented in the literature. In IB, due to its emphasis
+on type-refinement, that solution should have been more obvious.
 
@@
 
-Another advantage of IB is its support for
+Another advantage of IB is support for
 multiple inheritance. The literature is rich on how
@@
 multiple \emph{interface} inheritance has been supported since
-inception, and in Java 8 default methods~\cite{goetz12fdefenders} bring some of the
-advantages of traits into Java. In contrast, the literature~\cite{Sak89dis,bracha90mixin,malayeri2009cz}
-is also rich on how hard it is to modularly combine multiple sources
+inception, and Java 8 default methods~\cite{goetz12fdefenders} bring some of the
+advantages of traits to Java. The literature~\cite{Sak89dis,bracha90mixin,malayeri2009cz}
+is also rich on how hard it is to combine multiple sources
 of behaviour \textbf{and} state with multiple \emph{implementation}

### 2016-06-29 GPCE2016/sections/Implementation.tex

@@
 \item 
-{\bf @@@English need improvement here@@@@@@@}
-Lombok offers only limited/experimental support for separate compilation, that is
-to access information of code defined in different files.
-Limited by Lombok, \mixin 
+Due to limited support in Lombok for separate compilation, i.e., 
+accessing information of code defined in different files, \mixin 
 %In the same spirit, we have a mature \mixin annotation, which

### 2016-06-29 GPCE2016/sections/Introduction.tex
author-info & revise.

--- 2016-06-24 GPCE2016/sections/Introduction.tex
@@
 for the behaviour, which is structurally defined in their
-interface. Programmers do not define objects directly, but delegate
+interface. Programmers do not define objects directly but delegate
 the task to \emph{object interfaces}, whose role is similar to non-abstract
@@
 In contrast, in CB 
-special care and verbose explicit type-refinement is be required to
+special care and verbose explicit type-refinement are required to
 produce code that deals with subtyping adequately.
@@
 solution~\cite{eptrivially} to the Expression
-Problem~\cite{wadler98expression} in Java-like languages, shows how easy it is to solve
+Problem~\cite{wadler98expression} in Java-like languages shows how easy it is to solve
 the problem using only type-refinement. However it took nearly 20

### 2016-06-29 GPCE2016/sections/Introduction.tex
merge

--- 2016-06-29 GPCE2016/sections/Introduction.tex
@@
 point-of-view, it also imposes some limitations that a new IB language
-would not have. In particular, supporting proper encapsulation is
+would not have. Supporting proper encapsulation is
 difficult in Java due to limitations of Java interfaces. In particular
@@
   allowing most tools to work transparently with our approach. 
-  We provide the formal semantics here~\footnote{See url?? Formal semantics for Classless Java is provided in Section A}.
   Existing Java projects can use our approach and still be
   backward compatible with their clients, in a way that is specified
-  by our safety properties.
+  by our safety properties. A formal semantics is available in an extended report~\footnote{url}.
 

### 2016-06-29 GPCE2016/sections/Implementation.tex

@@
 harvesting the practical convenience of using existing Java libraries, the full
-Java language, and IDE support.
+Java language and IDE support.
 The key to our implementation is compilation agents, which

### 2016-06-29 GPCE2016/sections/Overview.tex

@@
 \end{comment}
-Moreover, if the programmer wants a different implementation, they may
+Moreover, if the programmer wants a different implementation, he may
 provide an alternative implementation using \Q@default@ methods. For example:
@@
 
-\paragraph{Property updater as Injections}
+% \paragraph{Property updater as Injections}
 
-By declaring fields in Newspeak and other languages,
-the programmers specifies field signatures and (critically) the intention of storing such information. Then the language generate getter and setter.
-Once state is abstracted away,
-it is well known that is possible to replace state
-access with computation, but the type of the field stays the same.
+% By declaring fields in Newspeak and other languages,
+% the programmers specifies field signatures and (critically) the intention of storing such information. Then the language generate getter and setter.
+% Once state is abstracted away,
+% it is well known that is possible to replace state
+% access with computation, but the type of the field stays the same.
 
-Our fluid approach to state is specular:
-the intention of storing the information is not expressed by the programmer and set in stone, 
-but can vary by inheritance:
-We generate the idea of the field
-just starting from getters and setters signatures.
+% Our fluid approach to state is specular:
+% the intention of storing the information is not expressed by the programmer and set in stone, 
+% but can vary by inheritance:
+% We generate the idea of the field
+% just starting from getters and setters signatures.
 
-In this case, we leverage on our fluid state to changing the underling type of the field,
- and our \Q@with@ is providing
-the right injection from the old type to the new one.
+% In this case, we leverage on our fluid state to changing the underling type of the field,
+%  and our \Q@with@ is providing
+% the right injection from the old type to the new one.
 

### 2016-06-29 GPCE2016/sections/Introduction.tex
Introduction and Abstract

@@
 Another advantage of IB is support for
-multiple inheritance. The literature is rich on how
+multiple inheritance. The literature shows how
 easy it is to combine multiple sources of pure behaviour,
-using mechanisms such as traits~\cite{scharli03traits}. In Java
+using traits~\cite{scharli03traits}. In Java
 multiple \emph{interface} inheritance has been supported since
@@
 methods from interfaces. Furthermore Java 9 will allow private methods
-on interfaces~\cite{srikanth16private}, greatly improving the support for
-encapsulation.
+on interfaces~\cite{srikanth16private}.
 
@@
   backward compatible with their clients, in a way that is specified
-  by our safety properties. A formal semantics is available in an extended report~\footnote{url}.
+  by our safety properties. A formal translation to Java is available in an extended report\footnote{url}.
 

### 2016-06-29 GPCE2016/sections/Overview.tex
Text saving

@@
 inheritance). A simple class diagram illustrating the basic system is
-given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some
-  research argues in favor of using subtyping for modeling taxonomies,
-  other research argues against this practice, we do not wish to take
-  sides in this argument, but to provide an engaging example.}
+given on the left side of Figure~\ref{fig:pegasus}.
+%\footnote{Some
+%  research argues in favor of using subtyping for modeling taxonomies,
+%  other research argues against this practice, we do not wish to take
+%  sides in this argument, but to provide an engaging example.}
 
@@
 Now we model the complete animal system where an \Q@Animal@ includes a \Q@location@
-representing its position in space. We use 2D points to keep track of locations using coordinates.
+representing its position in space. We use 2D points to keep track of locations.
 
 \paragraph{\Q@Point2D@: simple immutable data with fields}
-Here we will illustrate how points are modelled with interfaces.
+Points can be modelled with interfaces.
  In IB
@@
 
-\noindent The encoding over pure Java is now inconvenient: creating a new point object is cumbersome, even
+\noindent The encoding over Java is now inconvenient: creating a new point object is cumbersome, even
 with anonymous classes:
@@
 \paragraph{\Q@Point2D@ with object interfaces}
-This obvious ``constructor'' code can be generated by the \mixin
+This obvious ``constructor'' code is generated by the \mixin
 annotation.  By annotating the interface \Q@Point2D@, a variation of the shown
@@
 what the fields are, then generates an \Q@of@ method with one parameter for each
-of them. That is, we can just write
+of them. We can just write:
 
@@
 
-\noindent More precisely, a field or factory parameter is generated for every
+\noindent A field or factory parameter is generated for every
 abstract method that takes no parameters.
@@
 %names).
- An example of using \Q@Point2D@ is:
+ An example of using \Q@Point2D@, where we ``clone'' an existing point
+ but use
+ \Q@42@ as the x-coordinate, is:
 \begin{lstlisting}
@@
 \end{lstlisting}
-\noindent where we return a new point, using \Q@42@ as x-coordinate,
-and taking all the other information (only \Q@y@ in this case) from
-another point.
 
@@
 data-structures. As such, it is
-supported by \mixin as \Q@with-@ methods. For example:
+supported by \mixin as \Q@with-@ methods: 
 \begin{lstlisting}
@@
 2D points are mathematical entities, thus we choose an immutable data structure to
-model them. However animals are real world entities, and when an animal moves,
+model them. Animals are real world entities, and when an animal moves,
 it is the \emph{same} animal with a different location. We model this with
@@
 \noindent Note how the \texttt{of}, \texttt{withX} and
-\texttt{location} methods (all generated automatically) provide a
+\texttt{location} methods (generated automatically) give a
 basic interface for dealing with animals.
 
-In summary, dealing with state (mutable or not) in object interfaces
+In summary, state (mutable or not) in object interfaces
 relies on a notion of abstract state, and state is not directly
@@
 
-Thus, with object interfaces, the code for the \Q@Bird@ interface is:
+\noindent Thus the code for the \Q@Bird@ interface is:
 
@@
 \paragraph{Generated boilerplate}
-Just to give a feeling of how much mechanical code \mixin is generating, we show the
-generated code for the \texttt{Point3D} in Figure~\ref{fig:boilerplate}.
+To give an idea of how much code \mixin is generating, we show the
+generated code for \texttt{Point3D} in Figure~\ref{fig:boilerplate}.
 %\marcoT{Overall, for the whole
@@
 different types of abstract state operations: four field-based state
-operations; and functional updaters. Furthermore object interfaces support
-direct object instantiation via \texttt{of} factory methods.
+operations; and functional updaters. Object instantiation is directly
+supported by \texttt{of} factory methods.
 Figure~\ref{fig:abstractstate} summarizes the six operations supported

### 2016-06-29 GPCE2016/sections/Implementation.tex
Section 3 with recovered text

@@
 Java language and IDE support.
-The key to our implementation is compilation agents, which
+The key to our implementation are compilation agents, which
  allows us to rewrite the Java AST just
@@
 \paragraph{Advantages of Lombok}
-The Lombok compilation agent has advantages over 
-pre-processors or other Java annotation processors.
-Lombok offers in Java an expressive power similar to that of Scala/Lisp macros,
-except, for the syntactic convenience of quote/unquote templating. Compilation
-agents act modularly on each compilation unit, applying transformations to one 
-annotated class/interface at a time. This allows library code to be reused
+The Lombok compilation agent has advantages with respect to alternatives like
+pre-processors, or other Java annotation processors.
+Lombok offers in Java an expressive power similar to that of Scala/Lisp macros;
+except, for the syntactic convenience of quote/unquote templating.
+
+\paragraph{Direct modification of the AST}
+Lombok alters the generation process of the class files,
+by directly modifying the AST. Neither the source code is modified nor
+new Java files are generated. Moreover, and probably more importantly,
+Lombok supports generation of code \emph{inside} a class/interface,
+which conventional Java annotation processors, such as \texttt{javax.annotation}, do not support.
+
+\paragraph{Modularity}
+While general preprocessing acts across module boundaries, compilation
+agents act modularly on each class/compilation unit. It makes sense to
+apply the transformations to one class/interface at a time, and only to
+annotated classes/interfaces. This allows library code to be reused
 without the need of being reprocessed and recompiled, making our
-approach 100\% compatible with existing Java libraries. 
+approach 100\% compatible with existing Java libraries, which can be
+used and extended normally.
+
+\paragraph{Tool support}
+Features written in Lombok integrate and are supported directly in the
+language, and are often also supported by (most) tools.  
+In Figure~\ref{fig:screenshot},
+\mixin generates an \Q@of@ method in \Q@Point2D@, and \Q@of@, \Q@withX@, \Q@withY@ methods in \Q@Point3D@.
 In Eclipse, the processing is
 performed transparently and the information of the interface from
-compilation is captured in the ``Outline'' window as shown in Figure~\ref{fig:screenshot}..
-This includes all the methods inside the interface as well as the generated ones.
+compilation is captured in the ``Outline'' window.
+This includes all
+the methods inside the interface as well as the generated ones.
+Moreover, as a useful IDE feature, the auto-completion also works for these newly generated methods.
+
+\begin{comment}
+\paragraph{Clarity against obfuscation}
+Preprocessors bring great power, which can easily be misused producing
+code hard to understand. Thus code quality and maintainability are reduced.
+Compilation agents start from Java syntax, but they can reinterpret it.
+Preserving the syntax avoids syntactic conflicts, allowing many
+tools to work transparently.
+\end{comment}
 
@@
 However, the reader should note that Java libraries almost always break heir
-coherence during evolution and still claim backward compatibility (false in
-theory but statistically true in practice). In practice, adding any method to any
+coherence during evolution and still claim backward compatibility. In practice, adding any method to any
 non-final class of a Java library is enough to break heir
@@
 %Section~\ref{sec:translation} 
-We provide a formal definition for our safety claims here~\footnote{See url?? Section B}.
+Formal definition/proofs for our safety claims are found in the
+technical report.
 
@@
 we also offer a less polished annotation supporting
-separate compilation for files in the same package.
+separate compilation.
 %  has the same behavior as \mixin. A limitation is

### 2016-06-29 GPCE2016/sections/CaseStudy.tex
author-info & revise.
recovered and edited marco text in S3 and S4

@@
 
-Smalltalk and Dart offer method cascading~[??];
-at the price of introducing additional syntax,
-this avoid the need for fluent setters;
-method cascading (as setters in general) 
-intrinsically rely  on an imperative setting.
+Smalltalk and Dart offer method cascading and avoid the need for fluent setters.
+It is achieved at the price of introducing additional syntax and 
+intrinsically relies on an imperative setting. 
 Our approach supports both fluent setters and (functional) 

### 2016-06-29 GPCE2016/sections/Overview.tex
recovered and edited marco text in S3 and S4

@@
 
-% \paragraph{Property updater as Injections}
+\paragraph{Property updater as Injections}
 
-% By declaring fields in Newspeak and other languages,
-% the programmers specifies field signatures and (critically) the intention of storing such information. Then the language generate getter and setter.
-% Once state is abstracted away,
-% it is well known that is possible to replace state
-% access with computation, but the type of the field stays the same.
-
+By declaring fields in Newspeak and other languages,
+the programmers specifies field signatures and (critically) the intention of storing such information. 
+Then the language generates getters and setters.
+Once state is abstracted away, it is well known that state
+access can be replaced with computation, but the type of the field stays the same.
 % Our fluid approach to state is specular:
-% the intention of storing the information is not expressed by the programmer and set in stone, 
-% but can vary by inheritance:
-% We generate the idea of the field
-% just starting from getters and setters signatures.
-
-% In this case, we leverage on our fluid state to changing the underling type of the field,
-%  and our \Q@with@ is providing
-% the right injection from the old type to the new one.
+In our approach,
+the intention of storing the information is not expressed by the programmer and set in stone, 
+but can vary by inheritance: we generate the idea of the field
+just starting from signatures of getters and setters.
+In this case, we leverage on our fluid state to changing the underling type of the field,
+ and our \Q@with@ provides the right injection from the old type to the new.
 

### 2016-06-29 GPCE2016/sections/Overview.tex
Revised Section 4

@@
 
-\paragraph{Property updater as Injections}
-
-By declaring fields in Newspeak and other languages,
-the programmers specifies field signatures and (critically) the intention of storing such information. 
-Then the language generates getters and setters.
-Once state is abstracted away, it is well known that state
-access can be replaced with computation, but the type of the field stays the same.
-% Our fluid approach to state is specular:
-In our approach,
-the intention of storing the information is not expressed by the programmer and set in stone, 
-but can vary by inheritance: we generate the idea of the field
-just starting from signatures of getters and setters.
-In this case, we leverage on our fluid state to changing the underling type of the field,
- and our \Q@with@ provides the right injection from the old type to the new.
-
 \subsection{Advanced Multiple Inheritance}

### 2016-06-29 GPCE2016/sections/RelatedWork.tex
preparing for GPCE
Revised Section 4

--- 2016-03-24 OOPSLA16/sections/RelatedWork.tex
@@
 
+\begin{comment}
 \paragraph{Models of \objectoriented programming}
@@
 for multiple interface inheritance.
+\end{comment}
 
@@
 
+\paragraph{Automatically generated Getters and Setters}
+\bruno{Need to discuss automatic generation of setter and
+  getters, and then discuss this text about injections. See reviews
+  and reply.}
+By declaring fields in Newspeak and other languages,
+the programmers specifies field signatures and (critically) the intention of storing such information. 
+Then the language generates getters and setters.
+Once state is abstracted away, it is well known that state
+access can be replaced with computation, but the type of the field stays the same.
+% Our fluid approach to state is specular:
+In our approach,
+the intention of storing the information is not expressed by the programmer and set in stone, 
+but can vary by inheritance: we generate the idea of the field
+just starting from signatures of getters and setters.
+In this case, we leverage on our fluid state to changing the underling type of the field,
+ and our \Q@with@ provides the right injection from the old type to the new.
+
 \paragraph{ThisType and MyType}

### 2016-06-29 GPCE2016/predef/java.tex
revise & save space

--- 2016-03-24 OOPSLA16/predef/java.tex
@@
 lineskip=-1pt,
-basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
+basicstyle=\ttfamily\fontsize{8}{10}\selectfont, %\small,       % the size of the fonts that are used for the code
 numbers=none,                   % where to put the line-numbers

### 2016-06-29 GPCE2016/sections/RelatedWork.tex
revise & save space

--- 2016-03-24 OOPSLA16/sections/RelatedWork.tex
@@
 inheritance is very expressive, but difficult to model and implement, and can
-cause difficulties (including the famous diamond (fork-join)
+cause difficulties (including the famous diamond %(fork-join)
 problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
@@
 virtual inheritance, dealing with the diamond problem by keeping only
-one copy of the base class~\cite{ellis1990annotated}. However C++
-suffers from the object initialization problem~\cite{malayeri2009cz}.
+one copy of the base class~\cite{ellis1990annotated}. However it 
+suffers from object initialization problem~\cite{malayeri2009cz}.
 It bypasses constructor calls to virtual superclasses, which can
@@
 annotation has full control over object initialization,
-and the mechanism is transparent to users. Moreover, customized factory
-methods are also allowed: if users are not satisfied with the default
-generated \texttt{of} method, they can implement their own.
+and the mechanism is transparent to users. Moreover, if users are not satisfied with the default
+ \texttt{of} method, they can provide customized factory methods.
 

### 2016-06-29 GPCE2016/sections/CaseStudy.tex

@@
 \texttt{e1,e2} must be refined to \texttt{ExpPC}.
-
-Note that the code for instantiation is generated by \mixin.
 Creating a simple expression of type \texttt{ExpPC} is
@@
 \end{lstlisting}
-\noindent Without our approach, tedious instantiation code would need
+\noindent Without Classless Java, tedious instantiation code would need
 to be defined manually.
@@
 
-Smalltalk and Dart offer method cascading and avoid the need for fluent setters.
-It is achieved at the price of introducing additional syntax and 
+Languages like Smalltalk and Dart offer method cascading and avoid the need for fluent setters.
+This is achieved at the price of introducing additional syntax and 
 intrinsically relies on an imperative setting. 
@@
 
-\noindent The \mixin annotation allowed us to reduce the interfaces/classes used
-in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
-reductions were due to the replacement of instantiation classes with
+\noindent The \mixin annotation reduced the interfaces/classes used
+in Bono et al.'s implementation by 21.4\% (from 14 to 11), due to the replacement of instantiation classes with
 generated \texttt{of} methods. The number of source lines of code 
-(SLOC\footnote{Only Java code is counted, excluding comments and blank lines. })
+(SLOC)
 was reduced by 40\% due to both the removal of instantiation overhead and
 generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}.
-To ensure a fair comparison, we used the same coding style as Bono et al.'s.
+%To ensure a fair comparison, we used the same coding style as Bono et al.'s.
 
@@
 Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
-modularity is greatly increased, allowing for improved reusability and maintainability.
-We list the results in the following table:
+modularity is greatly increased, allowing for improved reusability and maintainability:
 

### 2016-06-29 GPCE2016/sections/RelatedWork.tex
merge

--- 2016-06-29 GPCE2016/sections/RelatedWork.tex
@@
 Multiple \emph{implementation}
-inheritance is very expressive, but difficult to model and implement, and can
+inheritance is very expressive, but difficult to model and implement. It can
 cause difficulties (including the famous diamond (fork-join)
 problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
-programs. To conciliate the need for expressive power and simplicity, many
+programs. To allow for expressive power and simplicity, many
 models have been proposed, including C++ virtual inheritance,

### 2016-06-29 GPCE2016/sections/Conclusion.tex
preparing for GPCE

--- 2016-03-24 OOPSLA16/sections/Conclusion.tex
@@
 caused by Java not being originally designed to work without classes,
-we introduce a new annotation, \mixin, which provides default implementations
+we introduce a new \mixin annotation. \mixin provides default implementations
 for various methods (e.g. getters, setters, with-methods) and a
@@
 We leverage on annotation processing and the Lombok library, in this way
-\mixin is just a normal Java library; thus our proposed style can be integrated
-in any Java project.
+\mixin is just a normal Java library.
 The \mixin annotation helps programmers
@@
 \prototypebased.  In IB state is not modelled at the platonic/ideal
-level but is handled excursively by instances.  This unlocks useful
+level but is handled exclusively by instances.  This unlocks useful
 code reuse patterns, as shown in Section~\ref{sec:ep}.

### 2016-06-29 GPCE2016/sections/RelatedWork.tex

@@
 inheritance is very expressive, but difficult to model and implement. It can
-cause difficulties (including the famous diamond (fork-join)
+cause difficulties (including the famous diamond
 problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.) in reasoning about
@@
 annotation has full control over object initialization,
-and the mechanism is transparent to users. Moreover, customized factory
-methods are also allowed: if users are not satisfied with the default
-generated \texttt{of} method, they can implement their own.
+and the mechanism is transparent to users. If users are not satisfied with the default
+ \texttt{of} method, customized factory methods can be provided.
 
@@
 algebra through Java 8 interfaces.
-
 \begin{comment}
@@
 \end{comment}
-
 There are also proposals for extending Java with traits. For
@@
 parameter types. Our approach to covariantly refine state can recover
-some of the additional expressiveness of MyType. As illustrated with our examples and
-case studies, object interfaces are still very useful in many
+some of the additional expressiveness of MyType. As illustrated with our examples, object interfaces are still very useful in many
 practical applications, yet they do not require additional

### 2016-06-29 GPCE2016/sections/RelatedWork.tex
paragraph on generated setter/getters in RW.

@@
 
-\paragraph{Automatically generated Getters and Setters}
-\bruno{Need to discuss automatic generation of setter and
-  getters, and then discuss this text about injections. See reviews
-  and reply.}
-By declaring fields in Newspeak and other languages,
-the programmers specifies field signatures and (critically) the intention of storing such information. 
-Then the language generates getters and setters.
+\paragraph{Automatic generation of getters and setters}
+% \bruno{Need to discuss automatic generation of setter and
+%   getters, and then discuss this text about injections. See reviews
+%   and reply.}
+is an old idea used in languages such as 
+Self~\cite{Ungar1987self}, Dart~\cite{Dart} and Newspeak~\cite{Bracha_thenewspeak}.
+% By declaring fields in Newspeak and other languages,
+The programmers specifies field signatures and (critically) the intention of storing such information,  
+then the language generates getters and setters.
 Once state is abstracted away, it is well known that state
 access can be replaced with computation, but the type of the field stays the same.
-% Our fluid approach to state is specular:
-In our approach,
-the intention of storing the information is not expressed by the programmer and set in stone, 
-but can vary by inheritance: we generate the idea of the field
-just starting from signatures of getters and setters.
-In this case, we leverage on our fluid state to changing the underling type of the field,
- and our \Q@with@ provides the right injection from the old type to the new.
+We do the opposite, the idea of the field is generated starting from signatures of 
+getters and setters. % Our fluid approach to state is specular:
+In our approach, the intention of storing the information is not expressed by the programmer and set in stone  
+but can vary by inheritance. 
+In this case, the underling type of the field can be changed by our fluid state, 
+ and \Q@with@ methods provide the right injection from the old type to the new.
 

### 2016-06-29 GPCE2016/sections/Implementation.tex
Revised Section 4
Minor edits

@@
 \centering
-\includegraphics[width=4.5in]{pdfs/screenshot4.png}
+\includegraphics[width=5in]{pdfs/screenshot4.png}
 \caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
@@
 be aware of the annotation semantics and provide the right type while
-refining methods. To support heir coherence, we would need
+refining methods. To support heir coherence, we need
 to give up automatic type refinement, which is an essential part of IB programming.
-However, the reader should note that Java libraries almost always break heir
+However, Java libraries almost always break heir
 coherence during evolution and still claim backward compatibility. In practice, adding any method to any
@@
 %Section~\ref{sec:translation} 
-Formal definition/proofs for our safety claims are found in the
+Formal definition/proofs for our safety claims are in the
 technical report.

### 2016-06-29 GPCE2016/sections/Conclusion.tex
revise RelatedWork and Conclusion. Now it is 10 pages!

@@
 \section{Conclusion}\label{sec:conclusion}
-Before Java 8, concrete methods and static methods were not allowed
+Before Java 8, concrete methods were not allowed
 to appear in interfaces.  Java 8 allows static interface methods and
-introduces \emph{default methods}, which enables implementations
-inside interfaces. This had an important positive consequence that
-was probably overlooked: the concept of class
+introduces \emph{default methods}, which enable implementations
+inside interfaces. An important positive consequence that
+was probably overlooked is that the concept of class
 (in Java) is now (almost) redundant and unneeded.
-We proposed a programming style, called Classless Java, where
+We propose a programming style, called Classless Java, where
 truly \objectoriented programs and (reusable) libraries
@@
 caused by Java not being originally designed to work without classes,
-we introduce a new \mixin annotation. \mixin provides default implementations
+we introduce the \mixin annotation that provides default implementations
 for various methods (e.g. getters, setters, with-methods) and a
@@
 
-Classless Java is just a programming style, but is showing the way
-for a new flavour of object orientation: We propose \interfacebased
+Classless Java is just a programming style, but showing the way
+to a new flavour of OOP: we propose \interfacebased
 \objectoriented languages (IB), as opposed to \classbased or
-\prototypebased.  In IB state is not modelled at the platonic/ideal
-level but is handled exclusively by instances.  This unlocks useful
+\prototypebased. In IB state is not modelled at the platonic
+level but handled exclusively by instances.  This unlocks useful
 code reuse patterns, as shown in Section~\ref{sec:ep}.
 An interesting avenue for future work would be to design a new
-language based on IB. With a proper language design we would not need
-to restrict ourselves to the limitations of Java and its syntax.
+language based on IB. With a proper language design, IB would not be restricted 
+by limitations of Java and its syntax.
 

### 2016-06-29 GPCE2016/sections/RelatedWork.tex
revise RelatedWork and Conclusion. Now it is 10 pages!

@@
 \paragraph{Traits vs Object Interfaces.}
-We consider object interfaces to be an alternative to traits or mixins.  In the
+We consider object interfaces an alternative to traits or mixins. In 
 trait model two concepts (traits and classes) coexist and cooperate. Some
 authors~\cite{BettiniDSS13} see this as good language design fostering good
-software development by helping programmers to think about the structure of
-their programs.  However, other authors see the need of two concepts and the
+software development by helping programmers think about the program structures. 
+However, others see the need of two concepts and the
 absence of state as drawbacks of this model~\cite{malayeri2009cz}.  Object
-interfaces are units of reuse, and at the same time they provide factory methods
+interfaces are units of reuse, and meanwhile provide factory methods
 for instantiation and support state. Our approach promotes the use of
-interfaces instead of classes, in order to rely on the modular composition
+interfaces in order to exploit the modular composition
 offered by interfaces. Since Java was designed for classes, a direct classless
 programming style is verbose and unnatural. However, annotation-driven code
-generation is enough to overcome this difficulty, and the resulting programming
-style encourages modularity, composability and reusability, by keeping a strong
-OO feel. In that sense, we promote object interfaces as being both units of
-reusable code and object factories. Our practical experience is that, in Java,
-separating the two notions leads to a lot of boilerplate code, and is quite
+generation is enough to overcome this difficulty and the resulting programming
+style encourages modularity, composability and reusability. %, by keeping a strong OO feel. 
+In that sense, we promote object interfaces as being both units of
+reuse and object factories. Our practical experience shows that 
+separating the two notions leads to lots of boilerplate code, and is quite
 limiting when multiple inheritance with state is required.  Abstract state
@@
 preserving type recursion and (nominal/structural) subtyping at the same time.
-One famous attempt in this direction is provided by
+One famous attempt in this direction is 
 \emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
 \textbf{this}, changing its meaning along with inheritance. However when
-invoking a method with MyType in a parameter position, the exact type of the
-receiver must be known.  This is a big limitation in \classbased \objectoriented
-programming, and is exasperated by the \interfacebased programming we propose: no
+invoking a method with MyType in parameter positions, the exact type of the
+receiver must be known. This is a big limitation in \classbased OO
+programming and is exasperated by the \interfacebased programming we propose: no
 type is ever going to be exact since classes are not explicitly used. A recent
-article~\cite{Saito2013933} lights up this topic, proposing two
+article~\cite{Saito2013933} proposes two
 new features: exact statements and nonheritable methods. Both are
@@
 environment.
-Admittedly, MyType greatly enhances the expressiveness and extensibility of
-\objectoriented programming languages. Object interfaces use covariant-return types to
+Admittedly, MyType greatly enhances the expressivity and extensibility of
+\objectoriented programming languages. Object interfaces %use covariant-return types to
 simulate some uses of MyType. But this approach only works for refining
@@
 parameter types. Our approach to covariantly refine state can recover
-some of the additional expressiveness of MyType. As illustrated with our examples, object interfaces are still very useful in many
+some of the additional expressivity of MyType. As illustrated with our examples, object interfaces are still very useful in many
 practical applications, yet they do not require additional

### 2016-06-30 GPCE2016/sections/CaseStudy.tex
revise all sections.

@@
 encoded in Classless Java. The second application shows how to model
-embedded DSLs based on fluent APIs. The two larger case studies, take
-existing projects and refactor them into Classless Java. The first
-case study shows a significant reduction in code size, while the
-second case study maintains the same amount of code, but improves
+embedded DSLs based on fluent APIs. Then two larger case studies refactor 
+existing projects into Classless Java. The first
+one shows a significant reduction in code size, while the
+second one maintains the same amount of code, but improves
 modularity.

### 2016-06-30 GPCE2016/sections/Implementation.tex
revise all sections.

@@
 The key to our implementation are compilation agents, which
- allows us to rewrite the Java AST just
+ allow us to rewrite the Java AST just
 before compilation. We discuss advantages and limitations of our approach.
@@
 annotated classes/interfaces. This allows library code to be reused
-without the need of being reprocessed and recompiled, making our
+without being reprocessed and recompiled, making our
 approach 100\% compatible with existing Java libraries, which can be
@@
 Features written in Lombok integrate and are supported directly in the
-language, and are often also supported by (most) tools.  
+language and are also supported by most tools.  
 In Figure~\ref{fig:screenshot},
@@
 This includes all
-the methods inside the interface as well as the generated ones.
+the methods inside the interface, including the generated ones.
 Moreover, as a useful IDE feature, the auto-completion also works for these newly generated methods.
@@
 Thus, in addition to the advantages of Lombok, Classless Java offers
-some more advantages with respect to arbitrary (compilation agent driven) AST rewriting.
+ more advantages with respect to arbitrary (compilation agent driven) AST rewriting.
 
@@
 Formal definition/proofs for our safety claims are in the
-technical report.
+technical report.%~\footnote{http://www.cs.hku.hk/research/techreps/document/TR-2016-02.pdf}.
 
@@
 type parameters can be used, but generic
-method typing %is neither formalized in this paper nor 
+method typing is %neither formalized in this paper nor 
 not explicitly checked by \mixin, but simply delegated to the Java compiler.

### 2016-06-30 GPCE2016/sections/Introduction.tex
Minor edits
revise all sections.

@@
 Anything related to state is completely contained in the
-instances and does not leak into the inheritance logic.  In CB, the structure of the state is fixed and can be only grown
+instances and does not leak into the inheritance logic.  In CB, the structure of the state is fixed and can only grow
 by inheritance.  In contrast, in IB the state is never
@@
 discovered solution~\cite{eptrivially} to the Expression
-Problem~\cite{wadler98expression} in Java-like languages, shows how
+Problem~\cite{wadler98expression} in Java-like languages shows how
 easy it is to solve the problem using only type-refinement. However it
@@
 typically forbidden in CB: it is widely known that \emph{naive} type-refinement of
-mutable fields is not type-safe. While covariant refinement of mutable
+mutable fields is not type-safe. Although covariant refinement of mutable
 fields is supported by some type systems~\cite{bruce98astatically,bruce1994paradigmatic,ernst06virtual,Saito2013933}, this requires
@@
 
-Another advantage of IB is support for
+Another advantage of IB is the support for
 multiple inheritance. The literature shows how
-easy it is to combine multiple sources of pure behaviour,
+easy it is to combine multiple sources of pure behaviour
 using traits~\cite{scharli03traits}. In Java
@@
 methods from interfaces. Furthermore Java 9 will allow private methods
-on interfaces~\cite{srikanth16private}.
+in interfaces~\cite{srikanth16private}.
 
@@
   backward compatible with their clients, in a way that is specified
-  by our safety properties. A formal translation to Java is available in an extended report\footnote{url}.
+  by our safety properties.
 
@@
 \item {\bf Applications and case studies:} we illustrate the usefulness of IB
-  through various examples and case studies.
+  through various examples and case studies. The implementation, case
+  studies, and an extended
+  technical report with a formal translation to Java are 
+  online:
+
+{\footnotesize \url{https://github.com/YanlinWang/classless-java}}
 

### 2016-06-30 GPCE2016/sections/Overview.tex
revise & save space
revise all sections.

@@
 
-\noindent Note that the generated code is transparent to a user, who
-only sees the original code with the \mixin annotation. Compared to the pure
+\noindent Note that the generated code is transparent to users, who
+only see the original code with the \mixin annotation. Compared to the pure
 Java solution in Section~\ref{sec:simple}, the solution using object interfaces
@@
 
-The animal system modeled so far is a simplified version of the
+The animal system modelled so far is a simplified version of the
 system presented in the left-side of Figure~\ref{fig:pegasus}.
@@
 
-\noindent Here we declare an abstract getter and setter for the mutable ``field''
+\noindent Here we declare an abstract getter and a setter for the mutable ``field''
 \Q@location@.  Without the \mixin annotation, there is no convenient way to
@@
 
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=64-66]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
@@
 
-\lstinputlisting[linerange=71-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=70-79]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 %Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
@@
 \saveSpaceFig
-\lstinputlisting[linerange=89-115]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
+\lstinputlisting[linerange=88-114]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
 \caption{Generated boilerplate code.}
@@
 \centering
+{%\fontsize{8}{10}\selectfont
 \begin{tabular}{|l|l|l|l|}
@@
 \end{tabular}
-
+}
 \caption{Abstract state operations for a field \texttt{x}, together with other operations, supported by the \mixin
@@
 
-\lstinputlisting[linerange=84-84]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=83-83]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
+\lstinputlisting[linerange=64-66]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=HORSE
 
-\lstinputlisting[linerange=71-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=70-79]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=84-84]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=83-83]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2016-06-30 GPCE2016/sections/CaseStudy.tex
grammar

@@
 new solution~\cite{eptrivially} using only covariant type refinement was
-proposed. When this solution is modeled with interfaces and default methods, it
+proposed. When this solution is modelled with interfaces and default methods, it
 can even provide independent extensibility~\cite{zenger05independentlyextensible}: the ability to assemble a system
@@
 In the formulation of the EP, there is an initial system that models
-arithmetic expressions with only literals and addition, and an initial
+arithmetic expressions with only literals, addition, and an initial
 operation \texttt{eval} for expression evaluation.
@@
 In this way, when a query is created using \texttt{ExtendedDatabase},
-all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling
+all the fluent setters return the correct type instead of the old \texttt{Database} type, which would prevent calling
 \texttt{orderBy}.
@@
 et. al~\cite{bono14}, and the code for that implementation is
-available online. We reimplemented the game using \mixin. Due to space
+available online. We refactored the game using \mixin. Due to space
 constraints, we omit the code here. The following table summarizes

### 2016-06-30 GPCE2016/sections/Implementation.tex
grammar

@@
 Java language and IDE support.
-The key to our implementation are compilation agents, which
+The key to our implementation is compilation agents, which
  allow us to rewrite the Java AST just
-before compilation. We discuss advantages and limitations of our approach.
+before compilation. We discuss the advantages and limitations of our approach.
 
@@
 annotated classes/interfaces. This allows library code to be reused
-without being reprocessed and recompiled, making our
+without being reprocessed or recompiled, making our
 approach 100\% compatible with existing Java libraries, which can be
@@
 
-\item{\bf Self coherence}: the generated code itself is well-typed. That is,
-  type errors are not present in code the user has not written (for
-  example \texttt{of} methods in Figure~\ref{fig:screenshot}).
+\item{\bf Self coherence}: the generated code itself is well-typed. 
 In our case, it means that either \mixin{} produces (in a controlled way) an
-understandable error or the interface can be successfully annotated and the generated code is well-typed.
+understandable error or the interface can be successfully annotated and the generated code
+ (e.g. the \texttt{of} methods in Figure~\ref{fig:screenshot}) is well-typed.
 
@@
 \item Simple generics is supported:
-type parameters can be used, but generic
-method typing is %neither formalized in this paper nor 
-not explicitly checked by \mixin, but simply delegated to the Java compiler.
+type parameters can be used, but generic method typing is 
+delegated to the Java compiler instead of  %neither formalized in this paper nor 
+being explicitly checked by \mixin.
 \item 

### 2016-06-30 GPCE2016/sections/Overview.tex
grammar

@@
 Also, since \mixin interfaces offer the \Q@of@ factory method, only interfaces where all the abstract methods
-can be synthesized can be object interfaces. A non \mixin interface is like an abstract class in Java.
+can be synthesized can be object interfaces. A non-\mixin interface is like an abstract class in Java.
 %\end{comment}
@@
 & \textbf{functional updaters}        & \Q@Point3D with(Point2D val)@
-& Update all matching fields in \texttt{val}.        \\ \hline
+& Updates all matching fields in \texttt{val}.        \\ \hline
 %\cline{2-4}

### 2016-06-30 GPCE2016/sections/RelatedWork.tex
grammar

@@
 mixin inheritance cannot provide a satisfactory resolution in some cases and
-restricts the flexibility of mixin composition. To fight this limitation, an
+restricts the flexibility of mixin composition. To fight against this limitation, an
 algebra of mixin operators is introduced~\cite{ancona2002calculus}, but this

### 2016-06-30 GPCE2016/predef/java.tex
adjust space, make figure bigger.

--- 2016-03-24 OOPSLA16/predef/java.tex
@@
 keywordstyle=\ttfamily\bfseries,
-aboveskip=3pt,
-belowskip=-1pt
+aboveskip=1pt,
+belowskip=1pt
 % commentstyle=\color{Gray},

### 2016-09-05 GPCE2016/sections/Introduction.tex
grammar
Text.

@@
 Object interfaces provide support for automatic type-refinement.  In
-contrast, in CB special care and verbose explicit type-refinement is
+contrast, in CB special care and verbose explicit type-refinement are
 required to produce code that deals with subtyping adequately. We
@@
   use the \Q@class@ keyword).  We call this restricted version of
-Java \emph{Classless Java}.
+Java \emph{Classless Java}. \textcolor{red}{Classless Java provides us
+with a way to address the famous diamond problem~\cite{bracha90mixin,Sak89dis}
+based on Java 8's interface model: interfaces can offer different implementations
+using the default methods, whereas conflicts are supposed to be explicitly resolved,
+otherwise the program will be rejected by the compiler. An interface that inherits
+a few default implementations of a method from its supertypes can define a new implemention
+to hide the conflicts among them, or use the \Q@super@ keyword to specify the one it wants. Examples
+and more discussion will be presented afterwards.}
 
@@
   technical report with a formal translation to Java are
-  online:
-
-{\footnotesize \url{https://github.com/YanlinWang/classless-java}}
+  online. \footnote{\url{https://github.com/YanlinWang/classless-java}}
 

### 2016-09-05 GPCE2016/sections/Overview.tex
Text.

@@
 
+\textcolor{red}{Note that we are breaking the naming conventions for getters and setters in Java;
+they have the same names as their fields in our case, which makes code more concise,
+but also presents a nice form of fluent setters with method chaining (see Section~\ref{sec:dsls}).
+Fluent setters are useful, and the difference from \Q@with-@ methods is that they do not create new objects.}
+
 \subsection{Advanced Multiple Inheritance}

### 2016-09-07 GPCE2016/sections/CaseStudy.tex
address reviews

@@
 non-modularly by modifying the existing code. In the refactored code
-base the pretty printer is added modularly. 
-Although the code in the refactored version is slightly increased (by 2.4\% SLOC), the
+base, the pretty printer is added modularly with more interfaces, thus 
+the code is slightly increased by 2.4\% SLOC. However, the
 modularity is greatly increased, allowing for improved reusability and maintainability:

### 2016-09-07 GPCE2016/sections/Introduction.tex
address reviews

@@
 believe that such verbosity hindered and slowed down the discovery of
-useful programming patterns involving type-refinement. A recently
-discovered solution~\cite{eptrivially} to the Expression
-Problem~\cite{wadler98expression} in Java-like languages shows how
+useful programming patterns involving type-refinement. Our previous work~\cite{eptrivially} 
+on the Expression Problem~\cite{wadler98expression} in Java-like languages shows how
 easy it is to solve the problem using only type-refinement. However it
@@
 implementation blends Java's conventional CB style and IB smoothly.
-As a result, we experiment object interfaces with several interesting
+As a result, we apply object interfaces to several interesting
 Java programs and conduct various case studies. Finally, we also

### 2016-09-09 GPCE2016/sections/Introduction.tex
cite technical report

@@
 \item {\bf Applications and case studies:} we illustrate the usefulness of IB
-  through various examples and case studies. The implementation, case
-  studies, and an extended
-  technical report with a formal translation to Java are
-  online. \footnote{\url{https://github.com/YanlinWang/classless-java}}
+  through various examples and case studies\footnote{\url{https://github.com/YanlinWang/classless-java}}. 
+  An extended version with a formal translation to Java can be found in the 
+  companion technical report~\cite{classless}.
 

### 2016-09-15 GPCE2016/sections/Introduction.tex
Introduction revised

@@
 They couple flexibility and rigour, expressive power and
-modular reasoning.  Two main ideas emerged to this end:
+modular reasoning.  Two main OO models emerged to this end:
 \prototypebased (PB)~\cite{Ungar87self} and \classbased languages such as
@@
 
-This paper presents a third alternative: the concept of
-\emph{\interfacebased} \objectoriented programming languages (IB), where objects
-implement interfaces directly. In IB interfaces own the implementation
-for the behaviour, which is structurally defined in their
-interface. Programmers do not define objects directly but delegate
-the task to \emph{object interfaces}, whose role is similar to non-abstract
-classes in \classbased \objectoriented programming languages
-(CB). Objects
-are instantiated by static factory methods in object interfaces.
+Regardeless of the OO model, inheritance is a key mechanism is OO 
+languages. Inheritance provides a modularization mechanism, which 
+is used to reuse implementations from inherited classes/objects. 
+Unfortunatelly, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
+multiple inheritance, especially when combined with state, has several
+tricky issues, including several variants of the famous 
+\emph{diamond problem}~\cite{}. Many of the problems related to 
+inheritance arize from the direct use of \emph{fields} to model state. 
+Inheriting two fields of the same name raises 
+the question of whether the two fields should be kept, or only one 
+field should exists (and if so, which value should be picked). Initialization 
+of the fields is also problematic, since initialization code may be inherited from 
+multiple parents. Finally, an additional problem with mutable fields is that their 
+type cannot be type-refined in extensions, which can cause modularity problems
+~\cite{wadler98expression,eptrivially}.
 
-A key challenge in IB lies in how to model state, which is
+To address those limitations, this paper presents a third alternative
+OO model called \emph{\interfacebased} \objectoriented programming
+languages (IB), where objects implement interfaces directly and fields
+are not directly supported. In IB interfaces own the implementation for the
+behaviour, which is structurally defined in their
+interface. Programmers do not define objects directly but delegate the
+task to \emph{object interfaces}, whose role is similar to
+non-abstract classes in \classbased \objectoriented programming
+languages (CB). Objects are instantiated by static factory methods in
+object interfaces.
+
+Due to the absence of fields, a key challenge in IB lies in how to model state, which is
 fundamental to have stateful objects. All abstract operations in an
@@
 \emph{type-safe covariant mutable state}. That is, in IB,
-it is possible to type-refine \emph{mutable} ``fields'' in subtypes. This is
+it is possible to type-refine \emph{mutable} state in subtypes. This is
 typically forbidden in CB: it is widely known that \emph{naive} type-refinement of
@@
 
+\begin{comment}
 Another advantage of IB is the support for
@@
 idioms using multiple inheritance and state.
+\end{comment}
 
@@
   use the \Q@class@ keyword).  We call this restricted version of
-Java \emph{Classless Java}. \textcolor{red}{Classless Java provides us
+Java \emph{Classless Java}. 
+\begin{comment}
+\textcolor{red}{Classless Java provides us
 with a way to address the famous diamond problem~\cite{bracha90mixin,Sak89dis}
@@
 and more discussion will be presented afterwards.}
+\end{comment}
 
@@
 Using Java annotation processors, we produce an implementation of
-Classless Java, which allows us to stick to pure Java 8. By annotating
- the interfaces that represent object interfaces with \mixin, code for interface instantiation and
-type refinement can be automatically generated. Such code should not be needed in the first place in
-a real IB language, and the annotation processor allows us to
-transparently hide it from Java programmers. The implementation works
+Classless Java, which allows us to stick to pure Java 8. 
+%By annotating
+% the interfaces that represent object interfaces with \mixin, code for interface instantiation and
+%type refinement can be automatically generated. Such code should not be needed in the first place in
+%a real IB language, and the annotation processor allows us to
+%transparently hide it from Java programmers. 
+The implementation works
 by performing AST rewriting, allowing most existing Java tools (such as
@@
 implementation blends Java's conventional CB style and IB smoothly.
-As a result, we apply object interfaces to several interesting
+We apply object interfaces to several interesting
 Java programs and conduct various case studies. Finally, we also
-discuss the behaviour of our \mixin annotation and
+discuss the behaviour of Classless Java and
 its properties.
@@
 
+\begin{comment}
 While the Java embedding has obvious advantages from the practical
@@
 in interfaces~\cite{srikanth16private}.
+\end{comment}
 

### 2016-09-15 GPCE2016/sections/Introduction.tex
revisions

@@
 tricky issues, including several variants of the famous 
-\emph{diamond problem}~\cite{}. Many of the problems related to 
+\emph{diamond problem}~\cite{bracha90mixin,Sak89dis}. Many of the problems related to 
 inheritance arize from the direct use of \emph{fields} to model state. 
@@
 the question of whether the two fields should be kept, or only one 
-field should exists (and if so, which value should be picked). Initialization 
+field should exist (and if so, which value should be picked). Initialization 
 of the fields is also problematic, since initialization code may be inherited from 

### 2016-09-15 GPCE2016/sections/Introduction.tex
Marco comments

@@
 \emph{diamond problem}~\cite{bracha90mixin,Sak89dis}. Many of the problems related to 
-inheritance arize from the direct use of \emph{fields} to model state. 
-Inheriting two fields of the same name raises 
+inheritance arise from the direct use of \emph{fields} to model state. 
+Inheriting two fields with the same name raises 
 the question of whether the two fields should be kept, or only one 
-field should exist (and if so, which value should be picked). Initialization 
+field should exist. Initialization 
 of the fields is also problematic, since initialization code may be inherited from 

### 2016-09-16 GPCE2016/sections/CaseStudy.tex
Acknowledgement.

@@
 
-\paragraph{Initial System}
+\para{Initial System}
 In the formulation of the EP, there is an initial system that models
@@
 
-\paragraph{Adding a New Type of Expressions}
+\para{Adding a New Type of Expressions}
 In the OO paradigm, it is easy to add new types of expressions. For example, the
@@
 %\marco{text under need to be improved}
-\paragraph{Adding a New Operation} The difficulty of the EP in OO
+\para{Adding a New Operation} The difficulty of the EP in OO
 languages arises from adding new operations. For example, adding a pretty printing
@@
 
-\paragraph{Independent Extensibility}
+\para{Independent Extensibility}
 To show that our approach supports independent extensibility, a
@@
 
-\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
+\para{Extending the Query Language} The previous query language can be extended with a new feature
 \texttt{orderBy} which orders the result records by a field that users

### 2016-09-16 GPCE2016/sections/Implementation.tex
adjust space, make figure bigger.
Acknowledgement.

@@
 
-\paragraph{Advantages of Lombok}
+\para{Advantages of Lombok}
 The Lombok compilation agent has advantages with respect to alternatives like
@@
 
-\paragraph{Direct modification of the AST}
+\para{Direct modification of the AST}
 Lombok alters the generation process of the class files,
@@
 
-\paragraph{Modularity}
+\para{Modularity}
 While general preprocessing acts across module boundaries, compilation
@@
 
-\paragraph{Tool support}
+\para{Tool support}
 Features written in Lombok integrate and are supported directly in the
@@
 \centering
-\includegraphics[width=5in]{pdfs/screenshot4.png}
+\includegraphics[width=5.2in]{pdfs/screenshot4.png}
 \caption{Generated methods shown in the Outline window of Eclipse and auto-completion.}
@@
 
-\paragraph{Syntax and type errors}
+\para{Syntax and type errors}
 Some preprocessors (like the C one) can produce syntactically invalid code.
@@
 
-\paragraph{Heir coherence} Another form of guarantee that could be
+\para{Heir coherence} Another form of guarantee that could be
 useful in AST rewriting is heir coherence. That is, interfaces

### 2016-09-16 GPCE2016/sections/Overview.tex
revisions
Acknowledgement.

@@
 
-\paragraph{Instantiation}
+\para{Instantiation}
 To use \texttt{Horse}, \texttt{Bird} and \texttt{Pegasus}, some
@@
 
-\paragraph{\Q@Point2D@: simple immutable data with fields}
+\para{\Q@Point2D@: simple immutable data with fields}
 Points can be modelled with interfaces.
@@
 \vspace{-5pt}
-\paragraph{\Q@Point2D@ with object interfaces}
+\para{\Q@Point2D@ with object interfaces}
 This obvious ``constructor'' code is generated by the \mixin
@@
 
-\paragraph{\texttt{with-} methods in object interfaces}
+\para{\texttt{with-} methods in object interfaces}
 The pattern of creating a new object by reusing most information from an old
@@
 
-\paragraph{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
+\para{\Q@Animal@ and \Q@Horse@: simple mutable data with fields}
 2D points are mathematical entities, thus we choose an immutable data structure to
@@
 
-\paragraph{Unsatisfactory \classbased solutions to field type refinement}
+\para{Unsatisfactory \classbased solutions to field type refinement}
 In Java if we want to define an animal class with a field we have a set of
@@
 
-\paragraph{Field type refinement with object interfaces}
+\para{Field type refinement with object interfaces}
 Object interfaces address the challenge of type-refinement as follows:
@@
 
-\paragraph{\Q@Point3D@ and property updaters}
+\para{\Q@Point3D@ and property updaters}
 The \Q@Point3D@ interface is defined as follows:
@@
 
-\paragraph{Generated boilerplate}
+\para{Generated boilerplate}
 To give an idea of how much code \mixin is generating, we show the
@@
 
-\paragraph{Summary of operations in Classless Java}
+\para{Summary of operations in Classless Java}
 %Method bodies can only refer to other methods;
@@
 
+\begin{comment}
 \textcolor{red}{Note that we are breaking the naming conventions for getters and setters in Java;
@@
 Fluent setters are useful, and the difference from \Q@with-@ methods is that they do not create new objects.}
+\end{comment}
 

### 2016-09-16 GPCE2016/sections/RelatedWork.tex
Acknowledgement.

@@
 
-\paragraph{Multiple inheritance in \objectoriented languages}
+\para{Multiple inheritance in \objectoriented languages}
 Many authors have argued in favor or against multiple inheritance.
@@
 
-\paragraph{Traits vs Object Interfaces.}
+\para{Traits vs Object Interfaces.}
 We consider object interfaces an alternative to traits or mixins. In
@@
 
-\paragraph{Automatic generation of getters and setters}
+\para{Automatic generation of getters and setters}
 % \bruno{Need to discuss automatic generation of setter and
@@
 
-\paragraph{ThisType and MyType}
+\para{ThisType and MyType}
 %\marco{ this section is unfair.

### 2016-09-16 GPCE2016/sections/CaseStudy.tex
consistency.

@@
 
-\para{Initial System}
+\para{Initial system}
 In the formulation of the EP, there is an initial system that models
@@
 
-\para{Adding a New Type of Expressions}
+\para{Adding a new type of expressions}
 In the OO paradigm, it is easy to add new types of expressions. For example, the
@@
 %\marco{text under need to be improved}
-\para{Adding a New Operation} The difficulty of the EP in OO
+\para{Adding a new operation} The difficulty of the EP in OO
 languages arises from adding new operations. For example, adding a pretty printing
@@
 
-\para{Independent Extensibility}
+\para{Independent extensibility}
 To show that our approach supports independent extensibility, a
@@
 
-\para{Extending the Query Language} The previous query language can be extended with a new feature
+\para{Extending the query language} The previous query language can be extended with a new feature
 \texttt{orderBy} which orders the result records by a field that users
@@
 a Lisp-like language
-\lstinline{Mumbler}\footnote{https://github.com/cesquivias/mumbler/tree/master/simple},
+\lstinline{Mumbler}\footnote{\url{https://github.com/cesquivias/mumbler/tree/master/simple}},
 which is created as a tutorial for the Truffle

### 2016-09-16 GPCE2016/sections/Overview.tex
Tables with captions.

@@
 \centering
-{%\fontsize{8}{10}\selectfont
+\hspace{.01in}{%\fontsize{8}{10}\selectfont
 \begin{tabular}{|l|l|l|l|}
@@
 & \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
-\multirow{4}{*}{\parbox{2.5cm}{State operations (for a field \texttt{x})}} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.
+\multirow{4}{*}{\parbox{2.3cm}{State operations (for a field \texttt{x})}} & \textbf{``fields''/getters}        &   \Q@int x()@                  & Retrieves value from field \texttt{x}.
 \\ \cline{2-4}
@@
 & \textbf{fluent setters}        & \Q@Point2D x(int val)@ &Sets the field
-\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
+\texttt{x} to \texttt{val} and returns \texttt{\this}.  \hspace{-.1in}         \\ \hline
 \multirow{3}{*}{Other operations} &

### 2016-09-16 GPCE2016/sections/Introduction.tex
spellcheck

@@
 
-Regardeless of the OO model, inheritance is a key mechanism is OO 
+Regardless of the OO model, inheritance is a key mechanism is OO 
 languages. Inheritance provides a modularization mechanism, which 
 is used to reuse implementations from inherited classes/objects. 
-Unfortunatelly, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
+Unfortunately, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
 multiple inheritance, especially when combined with state, has several

### 2016-09-16 GPCE2016/sections/RelatedWork.tex
consistency.
spellcheck

@@
 
-\para{Traits vs Object Interfaces.}
+\para{Traits vs object interfaces.}
 We consider object interfaces an alternative to traits or mixins. In
@@
 but can vary by inheritance.
-In this case, the underling type of the field can be changed by our fluid state,
+In this case, the underlying type of the field can be changed by our fluid state,
  and \Q@with@ methods provide the right injection from the old type to the new.

### 2016-09-16 GPCE2016/sections/CaseStudy.tex
Tables with captions.
spellcheck

@@
 new solution~\cite{eptrivially} using only covariant type refinement was
-proposed. When this solution is modelled with interfaces and default methods, it
+proposed. When this solution is modeled with interfaces and default methods, it
 can even provide independent extensibility~\cite{zenger05independentlyextensible}: the ability to assemble a system
@@
 
-\vspace{5pt}
+\begin{table}[t]
+\centering
 {\fontsize{8}{10}\selectfont
@@
 }
-\vspace{5pt}
+\caption{AAA.}
+\end{table}
 
@@
 
-\vspace{0.5em}
+\begin{table}[t]
+\centering
 {\fontsize{8}{10}\selectfont
@@
 }
+\caption{AAA.}
+\end{table}

### 2016-09-16 GPCE2016/sections/Conclusion.tex
spellcheck

@@
 Classless Java is just a programming style, but showing the way
-to a new flavour of OOP: we propose \interfacebased
+to a new flavor of OOP: we propose \interfacebased
 \objectoriented languages (IB), as opposed to \classbased or
-\prototypebased. In IB state is not modelled at the platonic
+\prototypebased. In IB state is not modeled at the platonic
 level but handled exclusively by instances.  This unlocks useful

### 2016-09-16 GPCE2016/sections/Implementation.tex
spellcheck

@@
 the syntax with the sole goal of \emph{enhancing and completing code}:
-we satisfy the behaviour of abstract methods; add method
+we satisfy the behavior of abstract methods; add method
 implementations; and refine return types.  We consider this to be
@@
 %encapsulated in various annotation handlers,
-% in which case some behaviours are allowed to reuse the code by invoking methods
+% in which case some behaviors are allowed to reuse the code by invoking methods
 %in other handlers, where tedious replicated code is avoided.
@@
   that is well-typed before code generation is also well-typed after the generation.
-The annotation just adds more behaviour without removing any functionality.
+The annotation just adds more behavior without removing any functionality.
 

### 2016-09-16 GPCE2016/sections/Introduction.tex
spellcheck

@@
 \Objectoriented languages strive to offer great code reuse.
-They couple flexibility and rigour, expressive power and
+They couple flexibility and rigor, expressive power and
 modular reasoning.  Two main OO models emerged to this end:
@@
 languages objects inherit from other objects. Thus objects own
-both behaviour and state (and objects are all you have).
+both behavior and state (and objects are all you have).
 In \classbased languages an object is an instance of a specific class,
 and classes inherit from other classes.  Objects own state,
-while classes contain behaviour and the structure of the state.
+while classes contain behavior and the structure of the state.
 
@@
 are not directly supported. In IB interfaces own the implementation for the
-behaviour, which is structurally defined in their
+behavior, which is structurally defined in their
 interface. Programmers do not define objects directly but delegate the
@@
 methods). Objects are only responsible to define the ultimate
-behaviour of a method. % and, for example, if such a method is just a setter.
+behavior of a method. % and, for example, if such a method is just a setter.
 Anything related to state is completely contained in the
@@
 multiple inheritance. The literature shows how
-easy it is to combine multiple sources of pure behaviour
+easy it is to combine multiple sources of pure behavior
 using traits~\cite{scharli03traits}. In Java
@@
 is also rich on how hard it is to combine multiple sources
-of behaviour \textbf{and} state with multiple \emph{implementation}
+of behavior \textbf{and} state with multiple \emph{implementation}
 inheritance of classes. In IB there is only multiple
@@
 %To retain the easiness and modularity of combining multiple sources of
-%pure behaviour, in IB state is just a special kind of behaviour
+%pure behavior, in IB state is just a special kind of behavior
 %implementation.
@@
 % Objects are the only responsibles to define the
-%ultimate behaviour of a method and, for example, if such method is
+%ultimate behavior of a method and, for example, if such method is
 %just a setter. Anything related to state is completely contained in
@@
 Java programs and conduct various case studies. Finally, we also
-discuss the behaviour of Classless Java and
+discuss the behavior of Classless Java and
 its properties.

### 2016-09-16 GPCE2016/sections/Overview.tex
spellcheck

@@
 respectively. Pegasus can not only \emph{run} but also \emph{fly}! This is the
-place where \emph{``multiple inheritance''} is needed, because
+place where \emph{``multiple inheritance''} is needed because
 \texttt{Pegasus} needs to obtain \texttt{fly} and \texttt{run}
@@
 
-The animal system modelled so far is a simplified version of the
+The animal system modeled so far is a simplified version of the
 system presented in the left-side of Figure~\ref{fig:pegasus}.
@@
 \para{\Q@Point2D@: simple immutable data with fields}
-Points can be modelled with interfaces.
+Points can be modeled with interfaces.
  In IB
@@
 \noindent Only code for methods needing implementation is generated. Thus,
-programmers can easily customize the behaviour for their special needs.
+programmers can easily customize the behavior for their special needs.
 Also, since \mixin interfaces offer the \Q@of@ factory method, only interfaces where all the abstract methods

### 2016-09-16 GPCE2016/predef/java.tex
space

@@
 \lstset{ %
-language=JavaScala,                % choose the language of the code
+language=Java,                % choose the language of the code
 columns=flexible,
@@
 keywordstyle=\ttfamily\bfseries,
-aboveskip=1pt,
-belowskip=1pt
+aboveskip=0.5pt,
+belowskip=-1.5pt
 % commentstyle=\color{Gray},

### 2016-09-16 GPCE2016/sections/CaseStudy.tex
space

@@
+\vspace{-5pt}
 \section{Applications and Case Studies}
@@
 
+\savespace
 \subsection{The Expression Problem with Object Interfaces}\label{subsec:ep}
@@
 
+\savespace
 \subsection{Embedded DSLs with Fluent Interfaces}\label{sec:dsls}
@@
 
+\savespace
 \subsection{A Maze Game}
@@
 available online. We refactored the game using \mixin. Due to space
-constraints, we omit the code here. The following table summarizes
-the number of lines of code and classes/interfaces in each implementation:
+constraints, we omit the code here. Table~\ref{tab:maze} summarizes
+the number of lines of code and classes/interfaces in each implementation.
+\noindent The \mixin annotation reduced the interfaces/classes used
+in Bono et al.'s implementation by 21.4\% (from 14 to 11), due to the replacement of instantiation classes with
+generated \texttt{of} methods. The number of source lines of code
+(SLOC) was reduced by 40\% due to both the removal of instantiation overhead and
+generation of getters/setters.
 
-\begin{table}[t]
+\begin{table}[h]
 \centering
@@
 }
-\caption{AAA.}
+\caption{Maze game code size comparison}
+\label{tab:maze}
+\vspace{-3pt}
 \end{table}
-
-\noindent The \mixin annotation reduced the interfaces/classes used
-in Bono et al.'s implementation by 21.4\% (from 14 to 11), due to the replacement of instantiation classes with
-generated \texttt{of} methods. The number of source lines of code
-(SLOC)
-was reduced by 40\% due to both the removal of instantiation overhead and
-generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}.
-%To ensure a fair comparison, we used the same coding style as Bono et al.'s.
 
@@
 
+\savespace
 \subsection{Refactoring an Interpreter}\label{subsec:int}
@@
 base. In the original code base the pretty printer is added
-non-modularly by modifying the existing code. In the refactored code
-base, the pretty printer is added modularly with more interfaces, thus
+non-modularly by modifying the existing code. As shown in table~\ref{tab:intp} 
+the pretty printer in the refactored code is added modularly with more interfaces, thus
 the code is slightly increased by 2.4\% SLOC. However, the
-modularity is greatly increased, allowing for improved reusability and maintainability:
+modularity is greatly increased, allowing for improved reusability and maintainability.
 
-\begin{table}[t]
+\begin{table}[h]
 \centering
@@
 }
-\caption{AAA.}
+\caption{Interpreter code size comparison}
+\label{tab:intp}
 \end{table}

### 2016-09-16 GPCE2016/sections/Overview.tex
space
Final version.

@@
 \section{A Running Example: Animals}\label{sec:ep}
-\begin{figure*}
-\centering
-\saveSpaceFig
-\begin{tabular}{c|c}
-\includegraphics[height=3.3cm]{pdfs/PegasusDetail.pdf}\hspace{0pt} &
-\begin{minipage}{7cm}
-\vspace{-95pt}
-\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
-%basicstyle=\ttfamily\scriptsize
-\end{minipage}
-\end{tabular}
-\caption{The animal system (left: complete structure, right: code for simplified animal system).}\label{fig:pegasus}
-\saveSpaceFig
-\end{figure*}
-
 This section illustrates how our programming style, supported by
@@
 %I will set up the sentence and someone can compute the number.}
+
+\begin{figure*}[t]
+\centering
+% \saveSpaceFig
+\vspace{-1ex}
+\begin{tabular}{c|c}
+\includegraphics[height=3.3cm]{pdfs/PegasusDetail.pdf}\hspace{0pt} &
+\begin{minipage}{7cm}
+\vspace{-95pt}
+\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
+%basicstyle=\ttfamily\scriptsize
+\end{minipage}
+\end{tabular}
+\caption{The animal system (left: complete structure, right: code for simplified animal system).}\label{fig:pegasus}
+\saveSpaceFig
+\end{figure*}
 
@@
   // generated code not visible to users
-  static Pegasus of() { return new Pegasus() {}; }
-}
+  static Pegasus of() { return new Pegasus() {}; } }
 \end{lstlisting}
@@
   public int x() {return 4;}
-  public int y() {return 2;}
-}
+  public int y() {return 2;} }
 \end{lstlisting}
@@
 
-\vspace{-5pt}
+\vspace{-3pt}
 \para{\Q@Point2D@ with object interfaces}
@@
     Point2D withX(int val);
-    Point2D withY(int val);
-}
+    Point2D withY(int val); }
 \end{lstlisting}
@@
 
-\lstinputlisting[linerange=70-79]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=70-76]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 %Interface \texttt{Point3D} extends \texttt{Point2D} with a new abstract method
@@
 \begin{figure}
-\saveSpaceFig
-\lstinputlisting[linerange=88-114]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
+\vspace{-1.5ex}
+\lstinputlisting[linerange=85-111]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=GENERATED_POINT3D
 \caption{Generated boilerplate code.}
@@
 
-\begin{figure*}
-\saveSpaceFig
+\begin{figure*}[t]
+% \saveSpaceFig
+\vspace{-1ex}
 \centering
@@
 \label{fig:abstractstate}
-
+\saveSpaceFig
 \end{figure*}
@@
 
-\lstinputlisting[linerange=83-83]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=80-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 
@@
 
-\lstinputlisting[linerange=70-79]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
+\lstinputlisting[linerange=70-76]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=BIRD
 
-\lstinputlisting[linerange=83-83]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
+\lstinputlisting[linerange=80-80]{../UseMixinLombok/src/pegasus/TestAnimal.java}% APPLY:linerange=PEGASUS
 

### 2016-10-16 poster/splash2016.tex
SPLASH poster

\documentclass[portrait,final,a0paper,fontscale=0.277]{baposter}

\usepackage{listings}

\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{bm}
\usepackage{url}

\usepackage{graphicx}
\usepackage{multicol}

%\usepackage{times}
%\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

\usepackage{xspace}

\usepackage{listings}

\lstset{ %
language=Java,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=0pt,
belowskip=0pt,
frame=single %border
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

\newcommand{\captionfont}{\footnotesize}

\graphicspath{{images/}{../images/}}
\usetikzlibrary{calc}

\newcommand{\SET}[1]  {\ensuremath{\mathcal{#1}}}
\newcommand{\MAT}[1]  {\ensuremath{\boldsymbol{#1}}}
\newcommand{\VEC}[1]  {\ensuremath{\boldsymbol{#1}}}
\newcommand{\Video}{\SET{V}}
\newcommand{\video}{\VEC{f}}
\newcommand{\track}{x}
\newcommand{\Track}{\SET T}
\newcommand{\LMs}{\SET L}
\newcommand{\lm}{l}
\newcommand{\PosE}{\SET P}
\newcommand{\posE}{\VEC p}
\newcommand{\negE}{\VEC n}
\newcommand{\NegE}{\SET N}
\newcommand{\Occluded}{\SET O}
\newcommand{\occluded}{o}
\newcommand\mixin{\textbf{\texttt{@Obj }}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{1.5em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

\input{symbols}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Define some colors

%\definecolor{lightblue}{cmyk}{0.83,0.24,0,0.12}
\definecolor{lightblue}{rgb}{0.145,0.6666,1}

% Draw a video
\newlength{\FSZ}
\newcommand{\drawvideo}[3]{% [0 0.25 0.5 0.75 1 1.25 1.5]
   \noindent\pgfmathsetlength{\FSZ}{\linewidth/#2}
   \begin{tikzpicture}[outer sep=0pt,inner sep=0pt,x=\FSZ,y=\FSZ]
   \draw[color=lightblue!50!black] (0,0) node[outer sep=0pt,inner sep=0pt,text width=\linewidth,minimum height=0] (video) {\noindent#3};
   \path [fill=lightblue!50!black,line width=0pt]
     (video.north west) rectangle ([yshift=\FSZ] video.north east)
    \foreach \x in {1,2,...,#2} {
      {[rounded corners=0.6] ($(video.north west)+(-0.7,0.8)+(\x,0)$) rectangle +(0.4,-0.6)}
    }
;
   \path [fill=lightblue!50!black,line width=0pt]
     ([yshift=-1\FSZ] video.south west) rectangle (video.south east)
    \foreach \x in {1,2,...,#2} {
      {[rounded corners=0.6] ($(video.south west)+(-0.7,-0.2)+(\x,0)$) rectangle +(0.4,-0.6)}
    }
;
   \foreach \x in {1,...,#1} {
     \draw[color=lightblue!50!black] ([xshift=\x\linewidth/#1] video.north west) -- ([xshift=\x\linewidth/#1] video.south west);
   }
   \foreach \x in {0,#1} {
     \draw[color=lightblue!50!black] ([xshift=\x\linewidth/#1,yshift=1\FSZ] video.north west) -- ([xshift=\x\linewidth/#1,yshift=-1\FSZ] video.south west);
   }
   \end{tikzpicture}
}

\hyphenation{resolution occlusions}
%%
\begin{poster}%
  % Poster Options
  {
  % Show grid to help with alignment
  grid=false,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=white,
  bgColorTwo=white,
  borderColor=lightblue,
  headerColorOne=black,
  headerColorTwo=lightblue,
  headerFontColor=white,
  boxColorOne=white,
  boxColorTwo=lightblue,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=true,
  headerborder=closed,
  headerheight=0.1\textheight,
%  textfont=\sc, An example of changing the text font
  headershape=roundedright,
  headershade=shadelr,
  headerfont=\Large\bf\textsc, %Sans Serif
  textfont={\setlength{\parindent}{1.5em}},
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  {\includegraphics[height=7em]{pdfs/vuw.png}}
  % Title
  {\bf\textsc{Classless Java: Interface-based Programming for the Masses}\vspace{0.1em}}
  % Authors
  {\text{Yanlin Wang {,} Haoyuan Zhang {,} Marco Servetto {,} Bruno C. d. S. Oliveira}}
  % \text{The University of Hong Kong, China}}
  % \text{The University of Hong Kong, China: \{ylwang,hyzhang,bruno\}@cs.hku.hk}\\
  % \text{Victoria University of Wellington, New Zealand: marco.servetto@ecs.vuw.ac.nz}}
  % University logo
  {% The makebox allows the title to flow into the logo, this is a hack because of the L shaped logo.
    \includegraphics[width=7em,height=7em]{pdfs/hku.png}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position
%%% towards an already declared box. So if you have a box attached to the
%%% bottom, one to the top and a third one which should be in between, you
%%% have to specify the top and bottom boxes before you specify the middle
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=lightblue,line width=0.03em] (0,0) circle(0.18em);}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Motivation}{name=motivation,column=0,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

   \begin{enumerate}\compresslist
      \item Generally speaking, multiple inheritance with fields is hard. Moreover, there 
      are problems with field type refinement, constructors, etc.
      \item Java support multiple inheritance with default methods in a limited way.
      \item Other models (traits, mixins, etc) have certain limitatations.
      \item Our approach is related with traits. But we use operations on state to simulate fields 
      which traits do not support.
   \end{enumerate}
   \vspace{0.3em}
 }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Contributions}{name=contribution,column=0,below=motivation}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \begin{enumerate}\compresslist
    \item IB and Object Interfaces: enable powerful programming idioms using multiple-inheritance, type-refinement and abstract state operations.
    \item Classless Java: a practical realization of IB in Java.
    \item Type-safe covariant mutable state
    \item Applications and case studies
   \end{enumerate}
   \vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Implementation}{name=implementation,column=0,below=contribution}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Java supports compilation agents, where Java libraries can interact with the Java compilation process,
acting as a man in the middle between the
generation of AST and bytecode.

This process is facilitated by frameworks like Lombok~\cite{lombok}:
a Java library that aims at reducing Java boilerplate code via
annotations. \mixin was created using Lombok.

\noindent{\centering\includegraphics[width=\linewidth]{pdfs/lombok3.png}\\}
\vspace{0.1em}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{UML Diagram}{name=uml,column=0,below=implementation}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent{\centering\includegraphics[width=0.8\linewidth]{pdfs/PegasusDetail.pdf}\\}
\vspace{0.1em}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Results (Partial)}{name=result,column=0,below=uml,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the result of the maze game case study, both SLOC and number of interfaces are greatly reduced:
\noindent{\includegraphics[width=0.95\linewidth]{pdfs/result.png}\\}
\vspace{0.1em}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Object Interfaces and Instantiation}{name=Model,column=1,span=2,row=0}{
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
\lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
\lstinputlisting[linerange=8-12]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
\lstinputlisting[linerange=1-5]{./src.txt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Object interfaces with state (immutable data)}{name=results,column=1,span=2,row=0,below=Model}{
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  \begin{multicols}{2}
%  \noindent{\centering\includegraphics[width=1\linewidth]{images/mymodel.pdf}\\}
% The process of generating an academic document is as follows: for each citation in that document, firstly sample a topic $\topic_k$ according to the distribution from paper\_topic distribution $\delta(\topic ; \paper)$ or author\_topic distribution $\zeta(\topic ; \authorvar)$ or venue\_topic distribution $\psi(\topic ; \venue)$  based on the document type. Then, draw a citation $\citationvar$ from the sampled topic distribution $\phi(:;\topic_k)$ in topic\_citation distribution $\phi(\citationvar;\topic)$.
% \end{multicols}

\lstinputlisting[linerange=7-11]{./src.txt}
\lstinputlisting[linerange=13-15]{./src.txt}
\vspace{0.1em}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{More in the paper}{name=more,column=1,span=2,row=0,below=results}{
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent{\centering\includegraphics[width=0.95\linewidth]{table.png}\\}

\begin{multicols}{2}
\noindent{\centering\includegraphics[width=1.15\linewidth]{pdfs/screenshot4.png}\\}
\begin{itemize}\compresslist
\item Techniques for field type refinement.
\item Formalization and proofs.
\item Case studies and applications.
    \small{
    \begin{itemize}\compresslist
    \item The Expression Problem
    \item Embedded DSLs with Fluent Interfaces
    \item A Maze Game
    \item Refactoring an Interpreter
    \end{itemize}
    }
\end{itemize}
\end{multicols}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Results(Partial)}{name=result2,column=1,below=more}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second case study, we found that embedding DSLs with our implementation of 
fluent interfaces is quite straightforward:
\lstinputlisting[linerange=17-19]{./src.txt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Limitations}{name=limitations,column=1,below=result2,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}\compresslist
\item Current implementation only realizes ejc version.
\item Generics is not fully supported.
\item Separate compilation is an experimental feature with current Lombok.
\end{itemize}
\vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Acknowledgments}{name=acknowledge,column=2,below=more}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This work is sponsored by the Hong Kong Research Grant Council Early Career 
Scheme project number 27200514. The first and second author are supported by 
the SIG PAC stduent travel grant.
   \vspace{0.3em}
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Contacts}{name=acknowledge,column=2,below=acknowledge,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent{\textbf{The University of Hong Kong}} \\
Yanlin Wang, Haoyuan Zhang, Bruno C. d. S. Oliveira: \emph{\{ylwang,hyzhang,bruno\}@cs.hku.hk} \\
\textbf{Victoria University of Wellington} \\
Marco Servetto: \emph{marco.servetto@ecs.vuw.ac.nz}
}

\end{poster}
\end{document}

### 2016-10-17 poster/splash2016.tex
fix

@@
   % Eye Catcher
-  {\includegraphics[height=7em]{pdfs/vuw.png}}
+  {\includegraphics[height=7em]{pdfs/hku.png}}
   % Title
@@
   {% The makebox allows the title to flow into the logo, this is a hack because of the L shaped logo.
-    \includegraphics[width=7em,height=7em]{pdfs/hku.png}
+    \includegraphics[width=7em,height=7em]{pdfs/vuw.png}
   }
@@
 
-This process is facilitated by frameworks like Lombok~\cite{lombok}:
+This process is facilitated by frameworks like Lombok:
 a Java library that aims at reducing Java boilerplate code via
@@
 This work is sponsored by the Hong Kong Research Grant Council Early Career 
-Scheme project number 27200514. The first and second author are supported by 
+Scheme project number 27200514. The first and second authors are supported by 
 the SIG PAC stduent travel grant.

### 2017-03-30 FACS/paper.tex
Latex

%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}

\newcommand{\mynotes}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\bruno[1]{\mynotes{Bruno}{red}{#1}}
\newcommand\marco[1]{\mynotes{Marco}{blue}{#1}}

\newcommand{\syndef}{$::=$}

\newcommand\name{{\bf DTrait}\xspace}

%
\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%

\title{Separating Use and Reuse to Improve Both}
%
\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)

\author{Marco Servetto\inst{1} \and Bruno C. d. S. Oliveira\inst{2}}
%
\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Marco Servetto and Bruno C. d. S. Oliveira}
%
\institute{Victoria University of Wellington, New Zealand,\\
\email{marco.servetto@ecs.vuw.ac.nz},\\ WWW home page:
\texttt{https://ecs.victoria.ac.nz/Main/MarcoServetto}
\and
The University of Hong Kong, Hong Kong\\}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Abstract here!
\dots
\keywords{Code Reuse, Object-Oriented Programming}
\end{abstract}

\section{Introduction}

Historically, we have seen a lot of focus on the importance of
separing subtype from subclassing~\cite{cook}.  This is claimed to be
good for code reuse, design and reasoning.  While there is no problem
in subtyping without subclassing, in most OO languages code reuse
(inheritance/extends) imply subtyping.  Consider the following (Java)
code:

\begin{lstlisting}[language=Java]
  class A{ int ma(){return Utils.m(this);} }
  class Utils{static int m(A a){..}}
\end{lstlisting}

This code seems correct, and there is no subtyping/subclassing.
Now, lets add a class B

\begin{lstlisting}[language=Java]
  class B extends A{ int mb(){return this.ma();} }
\end{lstlisting}

Class \lstinline{B} now have a method \lstinline{ma} inherited from \lstinline{A}.
This method passes \lstinline{this} as \lstinline{A}; we can see an invocation of \lstinline{ma}
inside \lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.
The execution will eventually call \lstinline{Util.m} with a \lstinline{B} instance.
This can be correct only if \lstinline{B} is a subtype of \lstinline{A}.
Thus, every OO language with the minimal features 
exposed in this example (using this, extends, method call)
is forced to accept that subclassing implies subtyping.
Note how this is true also in C++, where is possible to
"extends privately". Such is a limitation of the visibility of
subtyping but not over subtyping itself, and the former example
would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.
\marco{Cite some work of bruceTHisType and show how he also fails.}
\bruno{I was expecting some text/argument here on why subclassing
  implies subtyping is bad. Perhaps another example where subclassing 
must not imply subtyping to type-check.}

We will show here a simple design to completley uncouple subtyping
and subclassing in a nominally typed OO language,
by dividing code designed for USE
from code designed for REUSE.
First we show a minimal language, then we show how mutually recursive types
are supported, how state/constructors/fields are supported,
how we can extend the language with nested classes.
Finally we show 42, a full blown language build around our ideas of reuse.

Our design leverage on traits~\cite{}: a well know mechanisms for pure
code reuse

\section{Overview of \name}

Present basic examples and syntax?

\section{A Decoupled Trait Calculus}\bruno{Probably we need a name 
for the language/calculus}

Lets consider a simple language where trait names start lowercase (t)
and class names start uppercase (C).  
The syntax of the language is:\\

\begin{tabular}{lrcl}
      Code Literal & $L$ & \syndef & ${{\bf interface}?~{\bf implements}~\overline{C}~\overline{M}}$ \\
%      Types & $\tau$ & \syndef & $C$ \\
      Method Declaration & $M$ & \syndef & ${\bf class}?~{\bf method}~C~m(C~\overline{x})~e?$ \\

      Traits & $T$ & \syndef & $t \mid L$ \\
      Trait Declarations & $TD$ & \syndef & $t:L \mid t:Sum~\overline{T}$ \\
      Class Declarations & $CD$ & \syndef & $C:L \mid C:Sum~\overline{T}$ \\
      Expressions & $e$ & \syndef & $x \mid C \mid e.m(\overline{e})$\\
      Declarations & $D$ & \syndef & $CD \mid TD$\\
\end{tabular}

\bruno{Original is below for double-checking.}
\begin{verbatim}
L::={interface? implements Ts Ms} //code literal
T::=C //types are class names
M::= class? method T m(Txs) e? //method declaration
tL::=t | L  //trait name or literal
TD::= t:L | t:Sum tLs  //trait declaration
CD::= C:L | C:Sum tLs
e::= x | T | e.m(es)
D::=CD | TD
\end{verbatim}

To declare a trait or a class, we can use a code literal (or a trait
expression).  Traits come with various operators (restrict, hide,
alias) but for now we focus on the single operator Sum, taking a set
of trait names or literals and composing them in an \emph{associative}
and \emph{commutative} way.  Sum is the simplest and most elegant
composition operator.  $Sum \overline{T}$ composes the content of $\overline{T}$
by taking the union of the methods and the union of the implementations.
The Sum can not be applied if multiple versions of the same method are
present in different traits.  An exception is done for abstract methods:
methods where the implementation e is missing; In this case, if the
headers are compatible the implemented version is selected.  In a sum
of two abstract methods with compatible headers, the one with the more
specific type is selected. We will discuss later the formal details of
Sum.

A version of this language where there are no traits can be seen 
as a restriction/variation of FJ~\cite{}.

As you can see code literals can be marked as interfaces.  We will call class
interface, or simply interface a class declaration of for C:{interface
  ...}.

While classes are typed assuming \lstinline{this} is of the nominal type of the
class, trait declarations, do not introduce any nominal type.  \lstinline{this}
in a trait is typed with a special type \lstinline{This} that is visible only
inside such trait~\bruno{{\bf This} does not show up anywhere in
  the syntax!}.  Syntactically, a Literal can use the \lstinline{This} type,
and when the trait is reuse to create a class, This will become just
an alias for the class name.

In a program $\overline{D}$, a code literal is not well typed if it
implements some \lstinline{C} that is not an interface; that is,
implementing interface is the only way to induce subtyping in this
language literals can implement classes.

\subsection{Code Use and Code Reuse in \name}

As you can see there is no extends keyword, and \lstinline{Sum} can not contain class names.
In a sense, all classes are final (as in Java).
That is, using a trait is the only way to induce code reuse.

Lets now try to encode the example in Section~\ref{} e in our language:

\begin{lstlisting}
 A:{ method int ma(){return Utils.m(this);} }
 Utils:{ class method int m(A a){..}}
\end{lstlisting}

This code is fine, but there is no way to add a B reusing code of A, since
A is designed for code \emph{use} and not \emph{reuse}. Lets try
again, but this time aiming at code reuse:

\begin{lstlisting}
 ta:{ method int ma(){return Utils.m(this);} }
 A:SUM ta
 Utils:{ class method int m(A a){..}}
\end{lstlisting}

This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
ta have no knowlegde of \lstinline{A}. Lets try again

\begin{lstlisting}
 IA:{interface method int ma()}
 ta:{implements IA
   method int ma(){return Utils.m(this);} }
 A:SUM ta
 Utils:{ class method int m(IA a){..}}
\end{lstlisting}

This code works: Utils is using an interface IA and the trait ta
is implementing it. It is also possible to add a B as follow
Abeit other semantic models for traits~\cite{} have been proposed,
here we use the flattening model.
This means that 
 A:SUM ta
  reduces/is equivalent to/is flatted into
 A:{implements IA method int ma(){return Utils.m(this);} }
This code seems correct, and there is no subtyping/subclassing.
Now, lets add a class B

\begin{lstlisting}
  B:Sum ta, { int mb(){return this.ma();} }
\end{lstlisting}

This also works.  B reuses the code of ta, but has no knowledge of A.
Since B reuses ta, and ta implements IA, also B implements IA.  By
flattening, B is reduced to

\begin{lstlisting}
B:{implements IA
  method int ma(){return Utils.m(this);}
  int mb(){return this.ma();} } 
\end{lstlisting}

\subsection{Remarks on Typing}

Now some important notes on typing:

-traits are \emph{well-typed} before being reused. For example in

\begin{lstlisting}
t:{int m() 2 
   int n() this.m()+1}
\end{lstlisting}

\noindent t is well typed since m() is declared inside of t, while

\begin{lstlisting}
t1:{int n() this.m()+1} would be ill typed.
\end{lstlisting}

\noindent Note how in the former example ta is well typed just by knowing itself and IA.

-in class expressions C:Sum tLs
L in tLs are not well typed (on their own) before flattening, only
the result is supposed to be well typed.
Also Java have the same behaviour:
for example in

\begin{lstlisting}[language=Java]
  class B extends A{ int mb(){return this.ma();} }
\end{lstlisting}

\noindent we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for ma.
In our example,with the trait t of before

\begin{lstlisting}
C: Use t {int k() this.n()+this.m()}
\end{lstlisting}
\bruno{Do we mean Sum instead of Use?}

\noindent would be well typed, even if n,m are not defined inside
\begin{lstlisting}
{int k() this.n()+this.m()}
\end{lstlisting}
This is not the case in may similar works in literature~\cite{} where the
literals have to be self complete. In this case we would have been forced to
declare abstract methods n and m.

This idea that code literals wrote
in the declaration of a class
do not have to be well typed on their own also lets us
reason about mutually recursive types, as shown in Section~\ref{}.

-if a class is declared by using C:Sum ts, that is, without literals,
If the Sum is successfull, C is well typed, no need of further checking.
-on the other side, if a class is declared by C:Sum tLs, with
L1..Ln in tLs, and after successfull flattening C:L is not well typed,
the error was originally present in one of L1..Ln.

Recursive types:
OO language leverage on recursive types most of the times.
For example in a pure OO language, String may offer a Int size()
method, and Int may offer a String toString() method.

This means that is not possible to type in (full) isolation classes
String and Int.

The most expressive compilation process may divide the classes in groups of mutually 
dependent classes.
Each group may also depend from a number of other groups.
This would form a Direct Aciclyc Graph of groups.
To type a group, we first need to type all depended groups, then
we can extract the structure/signature/structural type of all
the classes of the group.
Now, with the information of the depended groups and the one extracted
from the current group, it is possible to typecheck the implementation
 of each class in the group.

In this model, it is reasonable to assume that flattening happens group by group, before
extracting the class signatures.

Here we go for a much simpler simple top down execution/interpretation for flattening, where flattening
happen one at the time, and classes are typechecked where their type is first needed.

For example

A:{int ma(B b) b.mb()+1}
tb:{int mb() 2}
tc:{int mc(A a,B b) a.ma(b)}
B: Sum tb
C: Sum tc, {method int hello() 1}

In this scenario, since we go top down, we first need to generate B.
To generate B, we need to use tb;
In order modularly ensure well typedness,
we require tb to be well typed at this stage. If tb was not well typed
a compilation error could be generated at this stage.
In this moment, A can not be compiled/checked alone,
we need informations about B, but A is not used in tb,
thus we do not need to type A and we can type tb with
 the available informations and proceed to generate B.
Now, we need to generate C, and we need to ensure well typedness of tc.
Now B is alreay well typed (since generated by Sum tb, with no Ls),
and A can be typed;  finally tc can be typed and used.
If sum could not be performed (for example it tc had a method hello too)
a compilation error could be generated at this stage.

On the opposite side, if B and C was swapped, as in
C: Sum tc, {method int hello() 1}  
B: Sum tb
now the first task would be to generate C, but 
to type tc we need to know the type of A and B.
But they are both unavailable: B is still not computed and 
A can not be compiled/checked alone, without information about B.
A compilation error would be generated, on the line of "flattening of C
requires tc, tc requires A,B, B is still in need of flattening".

In this example, a more expressive compilation process 
could compute a dependency graph and, if possible, reorganize the list,
but for simplicity lets consider to always provide the declarations
in the right order, if one exists.

Some may find the requirement of the existence of an order restrictive;
An example of a "morally correct" program where no right order exists is the following:

t:{ int mt(A a) a.ma()}
A:Sum t {int ma() 1}

We expect two (related) criticism to our compilation model:
1
In a system without inference for method types,
if the result of composition operators depends only on the
structural shape of their input (as for Sum)
is indeed possible to optimistically compute the resulting structural shape of the classes
and use this information to type involved examples like the former.
We believe such typing discipline could be fragile,
and could make human understanding the code reuse process much harder/involved.
2
In the world of strongly typed languages we could be tempted to
first check that all can go well, and then perform the flattening.
This would however be overcompicated for no observable difference:
Indeed, in the former example there is no difference
between
  First checking B and produce B code (that also contains B structural shape),
  then use B shape to check C and produce C code
or a more involved
  First check B and discover B structural shape as result of the checking,
  then use B shape to check C.
  Finally produce both B and C code.

\section{Managing State}

The Sum operator is very elegant, but our research community is struggling to
make it work with constructor/state/fields.
The goals fo this struggle are as follow:

\begin{itemize}

\item keep sum associative and commutative.
\item allowing a class to create instances of itself.
\item actually initialize objects, not leaving no null fields
\item managing fields in a way that borrows the elengance of summing methods
\item make easy to add new fields

\end{itemize}

In the related work we will show some alternative ways to handle state.
However the purest solution just requires methods:
  The idea is that
  the trait code just uses getter/setters/factories, while leaving
  to classes the role to finally define the fields/constructors.
  That is, the  the class has syntax richer that the trait one,
  allowing declaration for fields and constructors.
  This approach is very powefull[see class less java]

  Pro: This approach is associative and commutative, even self construction
  can be allowed if the trait requires a static/class method
  returning This; the class will then implement this method by forwarding
  a call to the constructor.
  Negative: writing the class code with the constructors and
   fields and getter/setters and factories can be quite tedious.
   There is no way for a trait to specify a default value for a field,
   the class need to handle all the state, even state that is conceptually
   "private" of such trait.

  for example

\begin{lstlisting}
  pointSum: { method int x(); method int y();
    static method This of(int x,int y);
    method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul: { method int x(); method int y();
    static method This of(int x,int y);
    method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
\end{lstlisting}

\subsection{A First Attempt at Composition}

\begin{lstlisting}
  Point:Use pointSum,pointMul
\end{lstlisting}

\noindent would fail since methods x,y and of are still abstract.
In this model the user is required to write something similar to

\begin{lstlisting}
  CPoint:Use point,colored, {
    int x; int y;
    method int x()x; method int y()y;
    class method This of(int x, int y)
      new Point(x,y);
    constructor Point(int x, int y){
      this.x=x;this.y=y;this.color=color;
      }
    }
\end{lstlisting}

\noindent after a while programming in this style, those "fixpoint" close the state in
the obvious way classes becomes quite repetitive, and one wonder
if it could be possible to automatically generate such code[]
In our model we go one step further:
In our model there is no need to generate code, or to explicitly
write down constructors and fields; there is not even syntax for those
constructor.
The idea is that any class that "could" be completed in the obvious way
is a complete "coherent" class.
Others call abstract a class that have abstract methods.
We call abstract a class that have a set of abstract methods that are not
coherent, that is, you can not interpret those as factory,getters and setters.

In the detail: a class with no abstract method is coherent, and like Java Math
will just be usefull for calling class/static methods.
a class with a single abstract class method returning This
is coherent if all the other abstract methods can be seen as "abstract state
operations" over one of its argument.
For example,
if there is a class method This of(int x, int y) as before,
then
a method int x() is an abstract state method, getter for x.
a method Void x(int that) is abstract state method, setter for x.
while getters and setters are fundamental operations, we can immagine
more operations to be supported:

\begin{lstlisting}
method This withX(int that) 
\end{lstlisting}
may be a "wither", doing a functional field update
  method Void update(int x,int) may do two field update at a time
  method This clone() may do a shallow clone of the object.
  We are not sure what is the best set of abstract state operations yet.

  lets play with the points of before, to see what good can we do with the current
  instruments:

\begin{lstlisting}
  pointSum: { method int x(); method int y();
    static method This of(int x,int y);
    method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul: { method int x(); method int y();
    static method This of(int x,int y);
    method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
  PointAlgebra:Sum pointSum,pointMul
\end{lstlisting}

  As you can see, we can declare the methods independently and compose the result
  as we wish. However we have to repeat the abstract methods x,y and of.
  In addition of Sum,Mul we may want many operations over points; can we improve our reuse
  and not repeat such abstract definitions? of course!

\begin{lstlisting}
  p: { method int x(); method int y();
    static method This of(int x,int y);
    }
  pointSum:Sum p, { method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul:Sum p, { method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
  pointDiv: ...
  PointAlgebra:Sum pointSum,pointMul,pointDiv,...
\end{lstlisting}

now our code is fully modularized, and each trait handle exactly one method.

What happens if we want to add fields instead of just operations?

\begin{lstlisting}
  colored:{ method Color color() }
  Point:Sum pointSum,colored
\end{lstlisting}

This first attempt does not work: the abstract color method
is not a getter for any of the parameters of

\begin{lstlisting}
  class method This of(int x,int y) 
\end{lstlisting}

A solution is to provide a richer factory:

\begin{lstlisting}
  CPoint:Sum pointSum,colored,{
    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/)
    class method This of(int x, int y,Color color)
    }
\end{lstlisting}

where we assume to support overloading on different parameter number.
This is a good solution, we think is better that any alternatives in literature,
however the method CPoint.sum reset the color to red.
What should be the behaviour in this case?
If instead of writing This.of() we used 
this.withX(newX).withY(newY) we could preserve the color from this.
Not sure if that would be better.
If the point designer could predict this kind of extension, then we could
use the following design:  
\begin{lstlisting}
  p: { method int x(); method int y();
    method This withX(int that);
    method This withY(int that);
    static method This of(int x,int y);
    method This merge(This other);
    }
  pointSum:Sum p, { method This sum(This other)
      this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
    }
  colored:{method Color color();
    method This withColor(Color that);
    method This merge(This other)
      this.withColor(this.color().mix(that.color());
    }
  CPoint:/*as before*/
\end{lstlisting}  
  Now we can merge colors, or any other kind of state we may want to add
  following this pattern.

\end{document}

### 2017-03-31 FACS/Examples.tex
division in files and some improvement

\section{A Decoupled Trait Calculus}
In order to illustrate how we plan to properly divide code reuse/inheritance/subclassing from subtyping,
lets consider a simple language where trait names start lowercase \mt
and class names start uppercase \mC.  
The syntax of the language is:

\begin{bnf}
\prodFull\mTD{\mt\terminalCode{:}\mL \mid \mt\terminalCode{:} \use\ \overline\mV}{Trait Decl}\\
\prodFull\mCD{\mC\terminalCode{:}\mL \mid \mC\terminalCode{:} \use\ \overline\mV}{Class Decl}\\
\prodFull\mV{\mt \mid \mL}{Code Value}\\
\prodFull\mL{\Opt{\terminalCode{interface}}\ \terminalCode{implements} \overline\mT\ \overline\mMD}{Code Literal}\\
\prodFull\mT{\mC}{types are class names}\\
\prodFull\mMD{\Opt{\terminalCode{class}}\ \terminalCode{method}\ \mT\ \mm\oR\overline{\mT\,\mx}\cR \Opt\me}{Method Decl}\\

\prodFull\me{\mx\mid\mT\mid\me\terminalCode{.}\mm\oR\overline\me\cR}{expressions}\\
\prodFull\mD{\mCD\mid\mTD}{Declaration}\\
\end{bnf}

To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
expression.  Traits come with various operators (restrict, hide,
alias) but for now we focus on the single operator 
$\use$, taking a set
of code values: that is trait names \mt\ or literals \mL\ and composing them in an \emph{associative}
and \emph{commutative} way.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
composition operator.  $\use\ \overline\mV$ composes the content of $\overline\mV$
by taking the union of the methods and the union of the implementations.

Use can not be applied if multiple versions of the same method are
present in different traits.  An exception is done for abstract methods:
methods where the implementation \me\ is missing; In this case (if the
headers are compatible) the implemented version is selected.  In a sum
of two abstract methods with compatible headers, the one with the more
specific type is selected. We will discuss later the formal details of
Sum.\marco{actually, this makes the sum not strongly associative...}

Code literals \mL\ can be marked as interfaces.  We will call class
interface, or simply interface a class declaration of for \Q@C:{interface  ...}@.
  Then we have a set of implemented interfaces and a set of member declarations; in this simple language, just methods.
If there are no implemented interfaces, in the concrete syntax we will omit the \Q@implements@ keyword.

Methods \mMD can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very usefull in the context of code composition.
To denote a method as abstract, instead of an optional keyword we just omit the implementation \me.

A version of this language where there are no traits can be seen 
as a restriction/variation of FJ~\cite{}.

Basic well formedness rules applies:
\begin{itemize}
\item all the traits and classes have unique names in a program $\overline\mD$,
\item all method parameters have unique names and the special parameter name \Q@this@ is not declared
 in the parameter list,
\item all methods in a code literal have unique names,
\item all used variables are in scope.
\end{itemize}
Those rules can be applied on any given \mL individually and in full isolation.

We expect the type system to enforce 
\begin{itemize}
\item subtyping between interfaces and classes,
\item method call typechecking,
\item no circular implementation of interfaces,
\item type signature of methods from interfaces can be refined following the well known variant-contravariant rules,
\item only interfaces can be implemented.
\item \marco{I'm sure I'm missing something}
\end{itemize}
We will see later the details on when exactly in the compilation process any code literal is typechecked.
While classes are typed assuming \lstinline{this} is of the nominal type of the
class, trait declarations, do not introduce any nominal type.  \lstinline{this}
in a trait is typed with a special type \lstinline{This} that is visible only
inside such trait. Syntactically, \Q@This@ is just a special, reserved, class name $\mC$.
A Literal can use the \lstinline{This} type,
and when the trait is reused to create a class, \Q@This@ will become just
an alias for the class name.

For the sake of simplicity, method bodies are just simple expressions \me, whose can be just variables and method calls.
We do not think that there is any problem to extend our language with other constructs; we use minimal expressions to keep our example simple and compact. In order to make our examples more engaging, we feel free to use the type int, numeric literals and their operations.

\subsection{Code Use and Code Reuse in \name}

The great absent in our language is the  \Q@extends@ keyword.
In a sense, all classes are final (as in Java).
Moreover  \use\ can not contain class names.
That is, using a trait is the only way to induce code reuse.

Lets now try to encode the example in Section~\ref{} e in our language:

\begin{lstlisting}
 A:{ method int ma() Utils.m(this) }//note, no {return _}
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

This code is fine, but there is no way to add a \Q@B@ reusing code of \Q@A@, since
\Q@A@ is designed for code \emph{use} and not \emph{reuse}. Lets try
again, but this time aiming at code reuse:

\begin{lstlisting}
 ta:{ method int ma() Utils.m(this) }//type error
 A:Use ta
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
\Q@ta@ have no knowlegde of \lstinline{A}. Lets try again

\begin{lstlisting}
 IA:{interface method int ma()}//interface with abstract method
 ta:{implements IA
   method int ma() Utils.m(this) }
 A:Use ta
 Utils:{ class method int m(IA a)/*method body here*/}
\end{lstlisting}

This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
is implementing it. It is also possible to add a \Q@B@ as follow
\begin{lstlisting}
  B:Use ta, { int mb(){return this.ma();} }
\end{lstlisting}
This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.

\paragraph*{Semantic of Use}
Abeit alternative semantic models for traits~\cite{} have been proposed,
here we use the flattening model.
This means that 
\begin{lstlisting}
A:Use ta
B:Use ta, { int mb(){return this.ma();} }
\end{lstlisting}

  reduces/is equivalent to/is flatted into

 \begin{lstlisting}
A:{implements IA method int ma() Utils.m(this) }
B:{implements IA
  method int ma() Utils.m(this)
  int mb() this.ma() }

 \end{lstlisting}

This code seems correct, and there is no mention of the trait \Q@ta@. In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.

\subsection{Remarks on Typing}
Our typing discipline is 
what distinguish our approach from a simple minded code composition macro(cite)
or a rigid module composition(cite may be some of my work).

The are two core ideas:
\paragraph*{traits are \emph{well-typed} before being reused.}${}_{}$\\*
 For example in

\begin{lstlisting}
t:{int m() 2 
   int n() this.m()+1}
\end{lstlisting}

\noindent \Q@t@ is well typed since \Q@m()@ is declared inside of \Q@t@, while

\begin{lstlisting}
t1:{int n() this.m()+1} 
\end{lstlisting}
\noindent would be ill typed.

\noindent Note how in the former example \Q@ta@ is well typed just by knowing itself and \Q@IA@.

\paragraph*{class expressions are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
In class expressions  $\use\ \overline\mV$
an \mL\ in $\overline\mV$ is not well typed before flattening, only the result is supposed to be well typed.
While this seams a very dangerous approach at first, consider that also Java have the same behaviour:
for example in
\begin{lstlisting}[language=Java]
  class A{ int m() {return 2;}  int n(){return this.m()+1;} }
  class B extends A{ int mb(){return this.ma();} }
\end{lstlisting}

\noindent in \Q@B@ we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for \Q@ma()@.
In our example,using the trait \Q@t@ of before

\begin{lstlisting}
C: Use t {int k() this.n()+this.m()}
\end{lstlisting}
\noindent would be correct: even if n,m are not defined inside
\Q@{int k() this.n()+this.m()}@,
the result of the flattening is well typed.

This is not the case in many similar works in literature~\cite{} where the
literals have to be self complete. In this case we would have been forced to
declare abstract methods \Q@n@ and \Q@m@.

Our typing strategy have two important properties:
\begin{itemize}
\item if a class is declared by using $\mC : \use\ \overline\mt$, that is, without literals,
and the flattening is successfull, \mC\ is well typed, no need of further checking.
\item on the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
$\mL_1\ldots\mL_n \in \overline\mV$, and after successfull flattening $\mC : \mL$ is not well typed,
the error was originally present in one of $\mL_1\ldots\mL_n$.
This also means that as an optimization strategy
 we may remember what method bodies come from traits and what method bodies come from code literals, in order to typecheck only the latter.
 \end{itemize}

We will see now how this idea that code literals wrote in the declaration of a class
do not have to be well typed on their own also lets us reason about mutually recursive types.

 \section{Recursive types}

OO language leverage on recursive types most of the times.
For example in a pure OO language, \Q@String@ may offers a \Q@Int size()@
method, and \Q@Int@ may offer a \Q@String toString()@ method.

This means that is not possible to type in (full) isolation classes
\Q@String@ and \Q@Int@.

The most expressive compilation process may divide the classes in groups of mutually 
dependent classes.
Each group may also depend from a number of other groups.
This would form a Direct Aciclyc Graph of groups.
To type a group, we first need to type all depended groups, then
we can extract the structure/signature/structural type of all
the classes of the group.
Now, with the information of the depended groups and the one extracted
from the current group, it is possible to typecheck the implementation
 of each class in the group.

In this model, it is reasonable to assume that flattening happens group by group, before
extracting the class signatures.

Here we go for a much simpler simple top down execution/interpretation for flattening, where flattening
happen one at the time, and classes are typechecked where their type is first needed.

For example
\begin{lstlisting}
A:{int ma(B b) b.mb()+1}
tb:{int mb() 2}
tc:{int mc(A a,B b) a.ma(b)}
B: Use tb
C: Use tc, {method int hello() 1}
\end{lstlisting}
In this scenario, since we go top down, we first need to generate \Q@B@.
To generate \Q@B@, we need to use \Q@tb@;
In order modularly ensure well typedness,
we require \Q@tb@ to be well typed at this stage. If \Q@tb@ was not well typed
a compilation error could be generated at this stage.
In this moment, \Q@A@ can not be compiled/checked alone,
we need informations about \Q@B@, but \Q@A@ is not used in \Q@tb@,
thus we do not need to type \Q@A@ and we can type \Q@tb@ with
 the available informations and proceed to generate \Q@B@.
Now, we need to generate \Q@C@, and we need to ensure well typedness of \Q@tc@.
Now \Q@B@ is alreay well typed (since generated by \use \Q@tb@, with no \mL),
and \Q@A@ can be typed;  finally \Q@tc@ can be typed and used.
If \use\ could not be performed (for example it \Q@tc@ had a method \Q@hello@ too)
a compilation error could be generated at this stage.

On the opposite side, if \Q@B@ and \Q@C@ was swapped, as in
\begin{lstlisting}
C: Use tc, {method int hello() 1}  
B: Use tb
\end{lstlisting}

now the first task would be to generate \Q@C@, but 
to type \Q@tc@ we need to know the type of \Q@A@ and \Q@B@.
But they are both unavailable: \Q@B@ is still not computed and 
\Q@A@ can not be compiled/checked alone, without information about \Q@B@.
A compilation error would be generated, on the line of ``flattening of \Q@C@
requires \Q@tc@, \Q@tc@ requires \Q@A@,\Q@B@, but \Q@B@ is still in need of flattening".

In this example, a more expressive compilation/precompilation process 
could compute a dependency graph and, if possible, reorganize the list,
but for simplicity lets consider to always provide the declarations
in the right order, if one exists.

\paragraph*{Criticism: existence of an order is restrictive.}${}_{}$\\*
Some may find the requirement of the existence of an order restrictive;
An example of a ``morally correct" program where no right order exists is the following:
\begin{lstlisting}
t:{ int mt(A a) a.ma()}
A:Use t {int ma() 1}
\end{lstlisting}

In a system without inference for method types,
if the result of composition operators depends only on the
structural shape of their input (as for \use)
is indeed possible to optimistically compute the resulting structural shape of the classes
and use this information to type involved examples like the former.
We stick to our simple approach, since we believe such typing discipline would be fragile,
and could make human understanding the code reuse process much harder/involved.
Indeed we just wrote a program where the correctness of trait \Q@t@ depends of 
\Q@A@, that is in turn generated using trait \Q@t@.

\paragraph*{Criticism: it would be better to typecheck before flatteningi.}${}_{}$\\*
In the world of strongly typed languages we could be tempted to
first check that all can go well, and then perform the flattening.
This would however be overcompicated for no observable difference:
Indeed, in the \Q@A,B,C@ example above there is no difference
between
\begin{itemize}
\item  (1)First check \Q@B@ and produce \Q@B@ code (that also contains \Q@B@ structural shape),
  (2) then use \Q@B@ shape to check \Q@C@ and produce \Q@C@ code
or a more involved
\item  (1)First check \Q@B@ and discover just \Q@B@ structural shape as result of the checking,
  (2)then use \Q@B@ shape to check \Q@C@.
  (3) Finally produce both \Q@B@ and \Q@C@ code.
\end{itemize}

Note how since our only point of relaxation is the class literal, there is no way an error can ``move around'' and be duplicated during the compilation process.
In particular, our approach allows for safe libraries of traits and classes to be fully typechecked, deployed and reused by multiple clients: no type error will emerge from library code.
On the other side, we do not enforce the programmer to write always self-contained code where all the abstract method definition are explicitly declared.

\section{Managing State}

The Sum operator is very elegant, but our research community is struggling to
make it work with constructor/state/fields.
The goals fo this struggle are as follow:

\begin{itemize}

\item keep sum associative and commutative.
\item allowing a class to create instances of itself.
\item actually initialize objects, not leaving no null fields
\item managing fields in a way that borrows the elengance of summing methods
\item make easy to add new fields

\end{itemize}

In the related work we will show some alternative ways to handle state.
However the purest solution just requires methods:
  The idea is that
  the trait code just uses getter/setters/factories, while leaving
  to classes the role to finally define the fields/constructors.
  That is, the  the class has syntax richer that the trait one,
  allowing declaration for fields and constructors.
  This approach is very powefull[see class less java]

  Pro: This approach is associative and commutative, even self construction
  can be allowed if the trait requires a static/class method
  returning This; the class will then implement this method by forwarding
  a call to the constructor.
  Negative: writing the class code with the constructors and
   fields and getter/setters and factories can be quite tedious.
   There is no way for a trait to specify a default value for a field,
   the class need to handle all the state, even state that is conceptually
   "private" of such trait.

  for example

\begin{lstlisting}
  pointSum: { method int x(); method int y();
    static method This of(int x,int y);
    method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul: { method int x(); method int y();
    static method This of(int x,int y);
    method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
\end{lstlisting}

\subsection{A First Attempt at Composition}

\begin{lstlisting}
  Point:Use pointSum,pointMul
\end{lstlisting}

\noindent would fail since methods x,y and of are still abstract.
In this model the user is required to write something similar to

\begin{lstlisting}
  CPoint:Use point,colored, {
    int x; int y;
    method int x()x; method int y()y;
    class method This of(int x, int y)
      new Point(x,y);
    constructor Point(int x, int y){
      this.x=x;this.y=y;this.color=color;
      }
    }
\end{lstlisting}

\noindent after a while programming in this style, those "fixpoint" close the state in
the obvious way classes becomes quite repetitive, and one wonder
if it could be possible to automatically generate such code[]
In our model we go one step further:
In our model there is no need to generate code, or to explicitly
write down constructors and fields; there is not even syntax for those
constructor.
The idea is that any class that "could" be completed in the obvious way
is a complete "coherent" class.
Others call abstract a class that have abstract methods.
We call abstract a class that have a set of abstract methods that are not
coherent, that is, you can not interpret those as factory,getters and setters.

In the detail: a class with no abstract method is coherent, and like Java Math
will just be usefull for calling class/static methods.
a class with a single abstract class method returning This
is coherent if all the other abstract methods can be seen as "abstract state
operations" over one of its argument.
For example,
if there is a class method This of(int x, int y) as before,
then
a method int x() is an abstract state method, getter for x.
a method Void x(int that) is abstract state method, setter for x.
while getters and setters are fundamental operations, we can immagine
more operations to be supported:

\begin{lstlisting}
method This withX(int that) 
\end{lstlisting}
may be a "wither", doing a functional field update
  method Void update(int x,int) may do two field update at a time
  method This clone() may do a shallow clone of the object.
  We are not sure what is the best set of abstract state operations yet.

  lets play with the points of before, to see what good can we do with the current
  instruments:

\begin{lstlisting}
  pointSum: { method int x(); method int y();
    static method This of(int x,int y);
    method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul: { method int x(); method int y();
    static method This of(int x,int y);
    method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
  PointAlgebra:Sum pointSum,pointMul
\end{lstlisting}

  As you can see, we can declare the methods independently and compose the result
  as we wish. However we have to repeat the abstract methods x,y and of.
  In addition of Sum,Mul we may want many operations over points; can we improve our reuse
  and not repeat such abstract definitions? of course!

\begin{lstlisting}
  p: { method int x(); method int y();
    static method This of(int x,int y);
    }
  pointSum:Sum p, { method This sum(This other)
      This.of(this.x+other.x,this.y+other.y);
    }
  pointMul:Sum p, { method This mul(This other)
      This.of(this.x*other.x,this.y*other.y);
    }
  pointDiv: ...
  PointAlgebra:Sum pointSum,pointMul,pointDiv,...
\end{lstlisting}

now our code is fully modularized, and each trait handle exactly one method.

What happens if we want to add fields instead of just operations?

\begin{lstlisting}
  colored:{ method Color color() }
  Point:Sum pointSum,colored
\end{lstlisting}

This first attempt does not work: the abstract color method
is not a getter for any of the parameters of

\begin{lstlisting}
  class method This of(int x,int y) 
\end{lstlisting}

A solution is to provide a richer factory:

\begin{lstlisting}
  CPoint:Sum pointSum,colored,{
    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/)
    class method This of(int x, int y,Color color)
    }
\end{lstlisting}

where we assume to support overloading on different parameter number.
This is a good solution, we think is better that any alternatives in literature,
however the method CPoint.sum reset the color to red.
What should be the behaviour in this case?
If instead of writing This.of() we used 
this.withX(newX).withY(newY) we could preserve the color from this.
Not sure if that would be better.
If the point designer could predict this kind of extension, then we could
use the following design:  
\begin{lstlisting}
  p: { method int x(); method int y();
    method This withX(int that);
    method This withY(int that);
    static method This of(int x,int y);
    method This merge(This other);
    }
  pointSum:Sum p, { method This sum(This other)
      this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
    }
  colored:{method Color color();
    method This withColor(Color that);
    method This merge(This other)
      this.withColor(this.color().mix(that.color());
    }
  CPoint:/*as before*/
\end{lstlisting}  
  Now we can merge colors, or any other kind of state we may want to add
  following this pattern.

### 2017-03-31 FACS/Intro.tex
division in files and some improvement

--- 2017-03-30 FACS/paper.tex
@@
-%
-\documentclass{llncs}
-%
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{listings}
-\usepackage{color}
-\usepackage{xspace}
-
-\newcommand{\mynotes}[3]{{\color{#2} {\sc #1}: #3}}
-\newcommand\bruno[1]{\mynotes{Bruno}{red}{#1}}
-\newcommand\marco[1]{\mynotes{Marco}{blue}{#1}}
-
-\newcommand{\syndef}{$::=$}
-
-\newcommand\name{{\bf DTrait}\xspace}
-
-%
-\begin{document}
-%
-\frontmatter          % for the preliminaries
-%
-\pagestyle{headings}  % switches on printing of running heads
-%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
-%
-
-\title{Separating Use and Reuse to Improve Both}
-%
-\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
-
-\author{Marco Servetto\inst{1} \and Bruno C. d. S. Oliveira\inst{2}}
-%
-\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Marco Servetto and Bruno C. d. S. Oliveira}
-%
-\institute{Victoria University of Wellington, New Zealand,\\
-\email{marco.servetto@ecs.vuw.ac.nz},\\ WWW home page:
-\texttt{https://ecs.victoria.ac.nz/Main/MarcoServetto}
-\and
-The University of Hong Kong, Hong Kong\\}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-Abstract here!
-\dots
-\keywords{Code Reuse, Object-Oriented Programming}
-\end{abstract}
-
 \section{Introduction}
@@
 This code seems correct, and there is no subtyping/subclassing.
-Now, lets add a class B
+Now, lets add a class \Q@B@
 
@@
 and subclassing in a nominally typed OO language,
-by dividing code designed for USE
-from code designed for REUSE.
+by dividing code designed for \textbf{USE}
+from code designed for \textbf{REUSE}.\footnote{We need to talk of unanticipated extensions?}
 First we show a minimal language, then we show how mutually recursive types
@@
 code reuse
-
-\section{Overview of \name}
-
-Present basic examples and syntax?
-
-\section{A Decoupled Trait Calculus}\bruno{Probably we need a name 
-for the language/calculus}
-
-Lets consider a simple language where trait names start lowercase (t)
-and class names start uppercase (C).  
-The syntax of the language is:\\
-
-\begin{tabular}{lrcl}
-      Code Literal & $L$ & \syndef & ${{\bf interface}?~{\bf implements}~\overline{C}~\overline{M}}$ \\
-%      Types & $\tau$ & \syndef & $C$ \\
-      Method Declaration & $M$ & \syndef & ${\bf class}?~{\bf method}~C~m(C~\overline{x})~e?$ \\
-
-      Traits & $T$ & \syndef & $t \mid L$ \\
-      Trait Declarations & $TD$ & \syndef & $t:L \mid t:Sum~\overline{T}$ \\
-      Class Declarations & $CD$ & \syndef & $C:L \mid C:Sum~\overline{T}$ \\
-      Expressions & $e$ & \syndef & $x \mid C \mid e.m(\overline{e})$\\
-      Declarations & $D$ & \syndef & $CD \mid TD$\\
-\end{tabular}
-
-\bruno{Original is below for double-checking.}
-\begin{verbatim}
-L::={interface? implements Ts Ms} //code literal
-T::=C //types are class names
-M::= class? method T m(Txs) e? //method declaration
-tL::=t | L  //trait name or literal
-TD::= t:L | t:Sum tLs  //trait declaration
-CD::= C:L | C:Sum tLs
-e::= x | T | e.m(es)
-D::=CD | TD
-\end{verbatim}
-
-To declare a trait or a class, we can use a code literal (or a trait
-expression).  Traits come with various operators (restrict, hide,
-alias) but for now we focus on the single operator Sum, taking a set
-of trait names or literals and composing them in an \emph{associative}
-and \emph{commutative} way.  Sum is the simplest and most elegant
-composition operator.  $Sum \overline{T}$ composes the content of $\overline{T}$
-by taking the union of the methods and the union of the implementations.
-The Sum can not be applied if multiple versions of the same method are
-present in different traits.  An exception is done for abstract methods:
-methods where the implementation e is missing; In this case, if the
-headers are compatible the implemented version is selected.  In a sum
-of two abstract methods with compatible headers, the one with the more
-specific type is selected. We will discuss later the formal details of
-Sum.
-
-A version of this language where there are no traits can be seen 
-as a restriction/variation of FJ~\cite{}.
-
-As you can see code literals can be marked as interfaces.  We will call class
-interface, or simply interface a class declaration of for C:{interface
-  ...}.
-
-While classes are typed assuming \lstinline{this} is of the nominal type of the
-class, trait declarations, do not introduce any nominal type.  \lstinline{this}
-in a trait is typed with a special type \lstinline{This} that is visible only
-inside such trait~\bruno{{\bf This} does not show up anywhere in
-  the syntax!}.  Syntactically, a Literal can use the \lstinline{This} type,
-and when the trait is reuse to create a class, This will become just
-an alias for the class name.
-
-In a program $\overline{D}$, a code literal is not well typed if it
-implements some \lstinline{C} that is not an interface; that is,
-implementing interface is the only way to induce subtyping in this
-language literals can implement classes.
-
-\subsection{Code Use and Code Reuse in \name}
-
-As you can see there is no extends keyword, and \lstinline{Sum} can not contain class names.
-In a sense, all classes are final (as in Java).
-That is, using a trait is the only way to induce code reuse.
-
-Lets now try to encode the example in Section~\ref{} e in our language:
-
-\begin{lstlisting}
- A:{ method int ma(){return Utils.m(this);} }
- Utils:{ class method int m(A a){..}}
-\end{lstlisting}
-
-This code is fine, but there is no way to add a B reusing code of A, since
-A is designed for code \emph{use} and not \emph{reuse}. Lets try
-again, but this time aiming at code reuse:
-
-\begin{lstlisting}
- ta:{ method int ma(){return Utils.m(this);} }
- A:SUM ta
- Utils:{ class method int m(A a){..}}
-\end{lstlisting}
-
-This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
-ta have no knowlegde of \lstinline{A}. Lets try again
-
-\begin{lstlisting}
- IA:{interface method int ma()}
- ta:{implements IA
-   method int ma(){return Utils.m(this);} }
- A:SUM ta
- Utils:{ class method int m(IA a){..}}
-\end{lstlisting}
-
-This code works: Utils is using an interface IA and the trait ta
-is implementing it. It is also possible to add a B as follow
-Abeit other semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
- A:SUM ta
-  reduces/is equivalent to/is flatted into
- A:{implements IA method int ma(){return Utils.m(this);} }
-This code seems correct, and there is no subtyping/subclassing.
-Now, lets add a class B
-
-\begin{lstlisting}
-  B:Sum ta, { int mb(){return this.ma();} }
-\end{lstlisting}
-
-This also works.  B reuses the code of ta, but has no knowledge of A.
-Since B reuses ta, and ta implements IA, also B implements IA.  By
-flattening, B is reduced to
-
-\begin{lstlisting}
-B:{implements IA
-  method int ma(){return Utils.m(this);}
-  int mb(){return this.ma();} } 
-\end{lstlisting}
-
-\subsection{Remarks on Typing}
-
-Now some important notes on typing:
-
--traits are \emph{well-typed} before being reused. For example in
-
-\begin{lstlisting}
-t:{int m() 2 
-   int n() this.m()+1}
-\end{lstlisting}
-
-\noindent t is well typed since m() is declared inside of t, while
-
-\begin{lstlisting}
-t1:{int n() this.m()+1} would be ill typed.
-\end{lstlisting}
-
-\noindent Note how in the former example ta is well typed just by knowing itself and IA.
-
--in class expressions C:Sum tLs
-L in tLs are not well typed (on their own) before flattening, only
-the result is supposed to be well typed.
-Also Java have the same behaviour:
-for example in
-
-\begin{lstlisting}[language=Java]
-  class B extends A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
-\noindent we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for ma.
-In our example,with the trait t of before
-
-\begin{lstlisting}
-C: Use t {int k() this.n()+this.m()}
-\end{lstlisting}
-\bruno{Do we mean Sum instead of Use?}
-
-\noindent would be well typed, even if n,m are not defined inside
-\begin{lstlisting}
-{int k() this.n()+this.m()}
-\end{lstlisting}
-This is not the case in may similar works in literature~\cite{} where the
-literals have to be self complete. In this case we would have been forced to
-declare abstract methods n and m.
-
-This idea that code literals wrote
-in the declaration of a class
-do not have to be well typed on their own also lets us
-reason about mutually recursive types, as shown in Section~\ref{}.
-
--if a class is declared by using C:Sum ts, that is, without literals,
-If the Sum is successfull, C is well typed, no need of further checking.
--on the other side, if a class is declared by C:Sum tLs, with
-L1..Ln in tLs, and after successfull flattening C:L is not well typed,
-the error was originally present in one of L1..Ln.
-
-Recursive types:
-OO language leverage on recursive types most of the times.
-For example in a pure OO language, String may offer a Int size()
-method, and Int may offer a String toString() method.
-
-This means that is not possible to type in (full) isolation classes
-String and Int.
-
-The most expressive compilation process may divide the classes in groups of mutually 
-dependent classes.
-Each group may also depend from a number of other groups.
-This would form a Direct Aciclyc Graph of groups.
-To type a group, we first need to type all depended groups, then
-we can extract the structure/signature/structural type of all
-the classes of the group.
-Now, with the information of the depended groups and the one extracted
-from the current group, it is possible to typecheck the implementation
- of each class in the group.
-
-In this model, it is reasonable to assume that flattening happens group by group, before
-extracting the class signatures.
-
-Here we go for a much simpler simple top down execution/interpretation for flattening, where flattening
-happen one at the time, and classes are typechecked where their type is first needed.
-
-For example
-
-A:{int ma(B b) b.mb()+1}
-tb:{int mb() 2}
-tc:{int mc(A a,B b) a.ma(b)}
-B: Sum tb
-C: Sum tc, {method int hello() 1}
-
-In this scenario, since we go top down, we first need to generate B.
-To generate B, we need to use tb;
-In order modularly ensure well typedness,
-we require tb to be well typed at this stage. If tb was not well typed
-a compilation error could be generated at this stage.
-In this moment, A can not be compiled/checked alone,
-we need informations about B, but A is not used in tb,
-thus we do not need to type A and we can type tb with
- the available informations and proceed to generate B.
-Now, we need to generate C, and we need to ensure well typedness of tc.
-Now B is alreay well typed (since generated by Sum tb, with no Ls),
-and A can be typed;  finally tc can be typed and used.
-If sum could not be performed (for example it tc had a method hello too)
-a compilation error could be generated at this stage.
-
-On the opposite side, if B and C was swapped, as in
-C: Sum tc, {method int hello() 1}  
-B: Sum tb
-now the first task would be to generate C, but 
-to type tc we need to know the type of A and B.
-But they are both unavailable: B is still not computed and 
-A can not be compiled/checked alone, without information about B.
-A compilation error would be generated, on the line of "flattening of C
-requires tc, tc requires A,B, B is still in need of flattening".
-
-In this example, a more expressive compilation process 
-could compute a dependency graph and, if possible, reorganize the list,
-but for simplicity lets consider to always provide the declarations
-in the right order, if one exists.
-
-Some may find the requirement of the existence of an order restrictive;
-An example of a "morally correct" program where no right order exists is the following:
-
-t:{ int mt(A a) a.ma()}
-A:Sum t {int ma() 1}
-
-We expect two (related) criticism to our compilation model:
-1
-In a system without inference for method types,
-if the result of composition operators depends only on the
-structural shape of their input (as for Sum)
-is indeed possible to optimistically compute the resulting structural shape of the classes
-and use this information to type involved examples like the former.
-We believe such typing discipline could be fragile,
-and could make human understanding the code reuse process much harder/involved.
-2
-In the world of strongly typed languages we could be tempted to
-first check that all can go well, and then perform the flattening.
-This would however be overcompicated for no observable difference:
-Indeed, in the former example there is no difference
-between
-  First checking B and produce B code (that also contains B structural shape),
-  then use B shape to check C and produce C code
-or a more involved
-  First check B and discover B structural shape as result of the checking,
-  then use B shape to check C.
-  Finally produce both B and C code.
-
-\section{Managing State}
-
-The Sum operator is very elegant, but our research community is struggling to
-make it work with constructor/state/fields.
-The goals fo this struggle are as follow:
-
-\begin{itemize}
-
-\item keep sum associative and commutative.
-\item allowing a class to create instances of itself.
-\item actually initialize objects, not leaving no null fields
-\item managing fields in a way that borrows the elengance of summing methods
-\item make easy to add new fields
-
-\end{itemize}
-
-In the related work we will show some alternative ways to handle state.
-However the purest solution just requires methods:
-  The idea is that
-  the trait code just uses getter/setters/factories, while leaving
-  to classes the role to finally define the fields/constructors.
-  That is, the  the class has syntax richer that the trait one,
-  allowing declaration for fields and constructors.
-  This approach is very powefull[see class less java]
-
-  Pro: This approach is associative and commutative, even self construction
-  can be allowed if the trait requires a static/class method
-  returning This; the class will then implement this method by forwarding
-  a call to the constructor.
-  Negative: writing the class code with the constructors and
-   fields and getter/setters and factories can be quite tedious.
-   There is no way for a trait to specify a default value for a field,
-   the class need to handle all the state, even state that is conceptually
-   "private" of such trait.
-
-  for example
-
-\begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-\end{lstlisting}
-
-\subsection{A First Attempt at Composition}
-
-\begin{lstlisting}
-  Point:Use pointSum,pointMul
-\end{lstlisting}
-
-\noindent would fail since methods x,y and of are still abstract.
-In this model the user is required to write something similar to
-
-\begin{lstlisting}
-  CPoint:Use point,colored, {
-    int x; int y;
-    method int x()x; method int y()y;
-    class method This of(int x, int y)
-      new Point(x,y);
-    constructor Point(int x, int y){
-      this.x=x;this.y=y;this.color=color;
-      }
-    }
-\end{lstlisting}
-
-\noindent after a while programming in this style, those "fixpoint" close the state in
-the obvious way classes becomes quite repetitive, and one wonder
-if it could be possible to automatically generate such code[]
-In our model we go one step further:
-In our model there is no need to generate code, or to explicitly
-write down constructors and fields; there is not even syntax for those
-constructor.
-The idea is that any class that "could" be completed in the obvious way
-is a complete "coherent" class.
-Others call abstract a class that have abstract methods.
-We call abstract a class that have a set of abstract methods that are not
-coherent, that is, you can not interpret those as factory,getters and setters.
-
-In the detail: a class with no abstract method is coherent, and like Java Math
-will just be usefull for calling class/static methods.
-a class with a single abstract class method returning This
-is coherent if all the other abstract methods can be seen as "abstract state
-operations" over one of its argument.
-For example,
-if there is a class method This of(int x, int y) as before,
-then
-a method int x() is an abstract state method, getter for x.
-a method Void x(int that) is abstract state method, setter for x.
-while getters and setters are fundamental operations, we can immagine
-more operations to be supported:
-
-\begin{lstlisting}
-method This withX(int that) 
-\end{lstlisting}
-may be a "wither", doing a functional field update
-  method Void update(int x,int) may do two field update at a time
-  method This clone() may do a shallow clone of the object.
-  We are not sure what is the best set of abstract state operations yet.
-
-  lets play with the points of before, to see what good can we do with the current
-  instruments:
-
-\begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-  PointAlgebra:Sum pointSum,pointMul
-\end{lstlisting}
-
-  As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods x,y and of.
-  In addition of Sum,Mul we may want many operations over points; can we improve our reuse
-  and not repeat such abstract definitions? of course!
-
-\begin{lstlisting}
-  p: { method int x(); method int y();
-    static method This of(int x,int y);
-    }
-  pointSum:Sum p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul:Sum p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-  pointDiv: ...
-  PointAlgebra:Sum pointSum,pointMul,pointDiv,...
-\end{lstlisting}
-
-now our code is fully modularized, and each trait handle exactly one method.
-
-What happens if we want to add fields instead of just operations?
-
-\begin{lstlisting}
-  colored:{ method Color color() }
-  Point:Sum pointSum,colored
-\end{lstlisting}
-
-This first attempt does not work: the abstract color method
-is not a getter for any of the parameters of
-
-\begin{lstlisting}
-  class method This of(int x,int y) 
-\end{lstlisting}
-
-A solution is to provide a richer factory:
-
-\begin{lstlisting}
-  CPoint:Sum pointSum,colored,{
-    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/)
-    class method This of(int x, int y,Color color)
-    }
-\end{lstlisting}
-
-where we assume to support overloading on different parameter number.
-This is a good solution, we think is better that any alternatives in literature,
-however the method CPoint.sum reset the color to red.
-What should be the behaviour in this case?
-If instead of writing This.of() we used 
-this.withX(newX).withY(newY) we could preserve the color from this.
-Not sure if that would be better.
-If the point designer could predict this kind of extension, then we could
-use the following design:  
-\begin{lstlisting}
-  p: { method int x(); method int y();
-    method This withX(int that);
-    method This withY(int that);
-    static method This of(int x,int y);
-    method This merge(This other);
-    }
-  pointSum:Sum p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
-    }
-  colored:{method Color color();
-    method This withColor(Color that);
-    method This merge(This other)
-      this.withColor(this.color().mix(that.color());
-    }
-  CPoint:/*as before*/
-\end{lstlisting}  
-  Now we can merge colors, or any other kind of state we may want to add
-  following this pattern.
-
-\end{document}

### 2017-03-31 FACS/paper.tex
division in files and some improvement

@@
 \usepackage{xspace}
+\usepackage{times}
+\lstset{language=Java,
+  basicstyle=\ttfamily\small,%\scriptsize,
+  keywordstyle=\bfseries,%\color{darkRed},
+  showstringspaces=false,
+  mathescape=true,
+  xleftmargin=0pt,
+  xrightmargin=0pt,
+  breaklines=false,
+  breakatwhitespace=false,
+  breakautoindent=false,
+  linewidth=4\textwidth,% should be enough
+%  identifierstyle=\idstyle
+ morekeywords={method,Use,This}
+}
 
+%marco
+\newcommand{\Q}{\lstinline}
+\newcommand\Opt[1]{#1 ?}
+%bnf
+\newenvironment{bnf}{$\begin{array}{lcll}}{\end{array}$}
+\newcommand{\prodFull}[3]{#1&::=&#2&\mbox{#3}}
+\newcommand{\prodInline}[2]{#1::=#2}
+\newcommand{\prodNextLine}[2]{&&#1&\mbox{#2}}
+\newcommand{\terminal}[1]{%
+\ensuremath{$\texttt{#1}$}%
+}
+\newcommand{\terminalCode}[1]{\mbox{\lstinline{#1}}}
+\newcommand{\metavariable}[1]{%
+\ensuremath{\mathit{#1}}%
+}
+%----------------------
+\newcommand\mL{\metavariable{L}}
+\newcommand\mC{\metavariable{C}}
+\newcommand\mT{\metavariable{T}}
+\newcommand\mV{\metavariable{V}} %code value t or L
+\newcommand\mMD{\metavariable{MD}}
+\newcommand\mTD{\metavariable{TD}}
+\newcommand\mCD{\metavariable{CD}}
+\newcommand\mD{\metavariable{D}}
+\newcommand\me{\metavariable{e}}
+\newcommand\mx{\metavariable{x}}
+\newcommand\mm{\metavariable{m}}
+\newcommand\mt{\metavariable{t}}
+\newcommand\use{\terminalCode{Use}}
+\newcommand\oC{\mbox{\lstinline@{@}}
+\newcommand\cC{\mbox{\lstinline@}@}}
+\newcommand\oR{\mbox{\lstinline@(@}}
+\newcommand\cR{\mbox{\lstinline@)@}}
+%--------------------------
 \newcommand{\mynotes}[3]{{\color{#2} {\sc #1}: #3}}
@@
 %
-\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
+\titlerunning{Separating Use and Reuse to Improve Both}  % abbreviated title (for running head)
 
@@
 %
-\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
+\authorrunning{M. Servetto and B. Oliveira} % abbreviated author list (for running head)
 %
@@
 
-\section{Introduction}
-
-Historically, we have seen a lot of focus on the importance of
-separing subtype from subclassing~\cite{cook}.  This is claimed to be
-good for code reuse, design and reasoning.  While there is no problem
-in subtyping without subclassing, in most OO languages code reuse
-(inheritance/extends) imply subtyping.  Consider the following (Java)
-code:
-
-\begin{lstlisting}[language=Java]
-  class A{ int ma(){return Utils.m(this);} }
-  class Utils{static int m(A a){..}}
-\end{lstlisting}
-
-This code seems correct, and there is no subtyping/subclassing.
-Now, lets add a class B
-
-\begin{lstlisting}[language=Java]
-  class B extends A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
-Class \lstinline{B} now have a method \lstinline{ma} inherited from \lstinline{A}.
-This method passes \lstinline{this} as \lstinline{A}; we can see an invocation of \lstinline{ma}
-inside \lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.
-The execution will eventually call \lstinline{Util.m} with a \lstinline{B} instance.
-This can be correct only if \lstinline{B} is a subtype of \lstinline{A}.
-Thus, every OO language with the minimal features 
-exposed in this example (using this, extends, method call)
-is forced to accept that subclassing implies subtyping.
-Note how this is true also in C++, where is possible to
-"extends privately". Such is a limitation of the visibility of
-subtyping but not over subtyping itself, and the former example
-would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.
-\marco{Cite some work of bruceTHisType and show how he also fails.}
-\bruno{I was expecting some text/argument here on why subclassing
-  implies subtyping is bad. Perhaps another example where subclassing 
-must not imply subtyping to type-check.}
-
-We will show here a simple design to completley uncouple subtyping
-and subclassing in a nominally typed OO language,
-by dividing code designed for USE
-from code designed for REUSE.
-First we show a minimal language, then we show how mutually recursive types
-are supported, how state/constructors/fields are supported,
-how we can extend the language with nested classes.
-Finally we show 42, a full blown language build around our ideas of reuse.
-
-Our design leverage on traits~\cite{}: a well know mechanisms for pure
-code reuse
-
-\section{Overview of \name}
-
-Present basic examples and syntax?
-
-\section{A Decoupled Trait Calculus}\bruno{Probably we need a name 
-for the language/calculus}
-
-Lets consider a simple language where trait names start lowercase (t)
-and class names start uppercase (C).  
-The syntax of the language is:\\
-
-\begin{tabular}{lrcl}
-      Code Literal & $L$ & \syndef & ${{\bf interface}?~{\bf implements}~\overline{C}~\overline{M}}$ \\
-%      Types & $\tau$ & \syndef & $C$ \\
-      Method Declaration & $M$ & \syndef & ${\bf class}?~{\bf method}~C~m(C~\overline{x})~e?$ \\
-
-      Traits & $T$ & \syndef & $t \mid L$ \\
-      Trait Declarations & $TD$ & \syndef & $t:L \mid t:Sum~\overline{T}$ \\
-      Class Declarations & $CD$ & \syndef & $C:L \mid C:Sum~\overline{T}$ \\
-      Expressions & $e$ & \syndef & $x \mid C \mid e.m(\overline{e})$\\
-      Declarations & $D$ & \syndef & $CD \mid TD$\\
-\end{tabular}
-
-\bruno{Original is below for double-checking.}
-\begin{verbatim}
-L::={interface? implements Ts Ms} //code literal
-T::=C //types are class names
-M::= class? method T m(Txs) e? //method declaration
-tL::=t | L  //trait name or literal
-TD::= t:L | t:Sum tLs  //trait declaration
-CD::= C:L | C:Sum tLs
-e::= x | T | e.m(es)
-D::=CD | TD
-\end{verbatim}
-
-To declare a trait or a class, we can use a code literal (or a trait
-expression).  Traits come with various operators (restrict, hide,
-alias) but for now we focus on the single operator Sum, taking a set
-of trait names or literals and composing them in an \emph{associative}
-and \emph{commutative} way.  Sum is the simplest and most elegant
-composition operator.  $Sum \overline{T}$ composes the content of $\overline{T}$
-by taking the union of the methods and the union of the implementations.
-The Sum can not be applied if multiple versions of the same method are
-present in different traits.  An exception is done for abstract methods:
-methods where the implementation e is missing; In this case, if the
-headers are compatible the implemented version is selected.  In a sum
-of two abstract methods with compatible headers, the one with the more
-specific type is selected. We will discuss later the formal details of
-Sum.
-
-A version of this language where there are no traits can be seen 
-as a restriction/variation of FJ~\cite{}.
-
-As you can see code literals can be marked as interfaces.  We will call class
-interface, or simply interface a class declaration of for C:{interface
-  ...}.
-
-While classes are typed assuming \lstinline{this} is of the nominal type of the
-class, trait declarations, do not introduce any nominal type.  \lstinline{this}
-in a trait is typed with a special type \lstinline{This} that is visible only
-inside such trait~\bruno{{\bf This} does not show up anywhere in
-  the syntax!}.  Syntactically, a Literal can use the \lstinline{This} type,
-and when the trait is reuse to create a class, This will become just
-an alias for the class name.
-
-In a program $\overline{D}$, a code literal is not well typed if it
-implements some \lstinline{C} that is not an interface; that is,
-implementing interface is the only way to induce subtyping in this
-language literals can implement classes.
-
-\subsection{Code Use and Code Reuse in \name}
-
-As you can see there is no extends keyword, and \lstinline{Sum} can not contain class names.
-In a sense, all classes are final (as in Java).
-That is, using a trait is the only way to induce code reuse.
-
-Lets now try to encode the example in Section~\ref{} e in our language:
-
-\begin{lstlisting}
- A:{ method int ma(){return Utils.m(this);} }
- Utils:{ class method int m(A a){..}}
-\end{lstlisting}
-
-This code is fine, but there is no way to add a B reusing code of A, since
-A is designed for code \emph{use} and not \emph{reuse}. Lets try
-again, but this time aiming at code reuse:
-
-\begin{lstlisting}
- ta:{ method int ma(){return Utils.m(this);} }
- A:SUM ta
- Utils:{ class method int m(A a){..}}
-\end{lstlisting}
-
-This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
-ta have no knowlegde of \lstinline{A}. Lets try again
-
-\begin{lstlisting}
- IA:{interface method int ma()}
- ta:{implements IA
-   method int ma(){return Utils.m(this);} }
- A:SUM ta
- Utils:{ class method int m(IA a){..}}
-\end{lstlisting}
-
-This code works: Utils is using an interface IA and the trait ta
-is implementing it. It is also possible to add a B as follow
-Abeit other semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
- A:SUM ta
-  reduces/is equivalent to/is flatted into
- A:{implements IA method int ma(){return Utils.m(this);} }
-This code seems correct, and there is no subtyping/subclassing.
-Now, lets add a class B
-
-\begin{lstlisting}
-  B:Sum ta, { int mb(){return this.ma();} }
-\end{lstlisting}
-
-This also works.  B reuses the code of ta, but has no knowledge of A.
-Since B reuses ta, and ta implements IA, also B implements IA.  By
-flattening, B is reduced to
-
-\begin{lstlisting}
-B:{implements IA
-  method int ma(){return Utils.m(this);}
-  int mb(){return this.ma();} } 
-\end{lstlisting}
-
-\subsection{Remarks on Typing}
-
-Now some important notes on typing:
-
--traits are \emph{well-typed} before being reused. For example in
-
-\begin{lstlisting}
-t:{int m() 2 
-   int n() this.m()+1}
-\end{lstlisting}
-
-\noindent t is well typed since m() is declared inside of t, while
-
-\begin{lstlisting}
-t1:{int n() this.m()+1} would be ill typed.
-\end{lstlisting}
-
-\noindent Note how in the former example ta is well typed just by knowing itself and IA.
-
--in class expressions C:Sum tLs
-L in tLs are not well typed (on their own) before flattening, only
-the result is supposed to be well typed.
-Also Java have the same behaviour:
-for example in
-
-\begin{lstlisting}[language=Java]
-  class B extends A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
-\noindent we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for ma.
-In our example,with the trait t of before
-
-\begin{lstlisting}
-C: Use t {int k() this.n()+this.m()}
-\end{lstlisting}
-\bruno{Do we mean Sum instead of Use?}
-
-\noindent would be well typed, even if n,m are not defined inside
-\begin{lstlisting}
-{int k() this.n()+this.m()}
-\end{lstlisting}
-This is not the case in may similar works in literature~\cite{} where the
-literals have to be self complete. In this case we would have been forced to
-declare abstract methods n and m.
-
-This idea that code literals wrote
-in the declaration of a class
-do not have to be well typed on their own also lets us
-reason about mutually recursive types, as shown in Section~\ref{}.
-
--if a class is declared by using C:Sum ts, that is, without literals,
-If the Sum is successfull, C is well typed, no need of further checking.
--on the other side, if a class is declared by C:Sum tLs, with
-L1..Ln in tLs, and after successfull flattening C:L is not well typed,
-the error was originally present in one of L1..Ln.
-
-Recursive types:
-OO language leverage on recursive types most of the times.
-For example in a pure OO language, String may offer a Int size()
-method, and Int may offer a String toString() method.
-
-This means that is not possible to type in (full) isolation classes
-String and Int.
-
-The most expressive compilation process may divide the classes in groups of mutually 
-dependent classes.
-Each group may also depend from a number of other groups.
-This would form a Direct Aciclyc Graph of groups.
-To type a group, we first need to type all depended groups, then
-we can extract the structure/signature/structural type of all
-the classes of the group.
-Now, with the information of the depended groups and the one extracted
-from the current group, it is possible to typecheck the implementation
- of each class in the group.
-
-In this model, it is reasonable to assume that flattening happens group by group, before
-extracting the class signatures.
-
-Here we go for a much simpler simple top down execution/interpretation for flattening, where flattening
-happen one at the time, and classes are typechecked where their type is first needed.
-
-For example
-
-A:{int ma(B b) b.mb()+1}
-tb:{int mb() 2}
-tc:{int mc(A a,B b) a.ma(b)}
-B: Sum tb
-C: Sum tc, {method int hello() 1}
-
-In this scenario, since we go top down, we first need to generate B.
-To generate B, we need to use tb;
-In order modularly ensure well typedness,
-we require tb to be well typed at this stage. If tb was not well typed
-a compilation error could be generated at this stage.
-In this moment, A can not be compiled/checked alone,
-we need informations about B, but A is not used in tb,
-thus we do not need to type A and we can type tb with
- the available informations and proceed to generate B.
-Now, we need to generate C, and we need to ensure well typedness of tc.
-Now B is alreay well typed (since generated by Sum tb, with no Ls),
-and A can be typed;  finally tc can be typed and used.
-If sum could not be performed (for example it tc had a method hello too)
-a compilation error could be generated at this stage.
-
-On the opposite side, if B and C was swapped, as in
-C: Sum tc, {method int hello() 1}  
-B: Sum tb
-now the first task would be to generate C, but 
-to type tc we need to know the type of A and B.
-But they are both unavailable: B is still not computed and 
-A can not be compiled/checked alone, without information about B.
-A compilation error would be generated, on the line of "flattening of C
-requires tc, tc requires A,B, B is still in need of flattening".
-
-In this example, a more expressive compilation process 
-could compute a dependency graph and, if possible, reorganize the list,
-but for simplicity lets consider to always provide the declarations
-in the right order, if one exists.
-
-Some may find the requirement of the existence of an order restrictive;
-An example of a "morally correct" program where no right order exists is the following:
-
-t:{ int mt(A a) a.ma()}
-A:Sum t {int ma() 1}
-
-We expect two (related) criticism to our compilation model:
-1
-In a system without inference for method types,
-if the result of composition operators depends only on the
-structural shape of their input (as for Sum)
-is indeed possible to optimistically compute the resulting structural shape of the classes
-and use this information to type involved examples like the former.
-We believe such typing discipline could be fragile,
-and could make human understanding the code reuse process much harder/involved.
-2
-In the world of strongly typed languages we could be tempted to
-first check that all can go well, and then perform the flattening.
-This would however be overcompicated for no observable difference:
-Indeed, in the former example there is no difference
-between
-  First checking B and produce B code (that also contains B structural shape),
-  then use B shape to check C and produce C code
-or a more involved
-  First check B and discover B structural shape as result of the checking,
-  then use B shape to check C.
-  Finally produce both B and C code.
-
-\section{Managing State}
-
-The Sum operator is very elegant, but our research community is struggling to
-make it work with constructor/state/fields.
-The goals fo this struggle are as follow:
-
-\begin{itemize}
-
-\item keep sum associative and commutative.
-\item allowing a class to create instances of itself.
-\item actually initialize objects, not leaving no null fields
-\item managing fields in a way that borrows the elengance of summing methods
-\item make easy to add new fields
-
-\end{itemize}
-
-In the related work we will show some alternative ways to handle state.
-However the purest solution just requires methods:
-  The idea is that
-  the trait code just uses getter/setters/factories, while leaving
-  to classes the role to finally define the fields/constructors.
-  That is, the  the class has syntax richer that the trait one,
-  allowing declaration for fields and constructors.
-  This approach is very powefull[see class less java]
-
-  Pro: This approach is associative and commutative, even self construction
-  can be allowed if the trait requires a static/class method
-  returning This; the class will then implement this method by forwarding
-  a call to the constructor.
-  Negative: writing the class code with the constructors and
-   fields and getter/setters and factories can be quite tedious.
-   There is no way for a trait to specify a default value for a field,
-   the class need to handle all the state, even state that is conceptually
-   "private" of such trait.
-
-  for example
-
-\begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-\end{lstlisting}
-
-\subsection{A First Attempt at Composition}
-
-\begin{lstlisting}
-  Point:Use pointSum,pointMul
-\end{lstlisting}
-
-\noindent would fail since methods x,y and of are still abstract.
-In this model the user is required to write something similar to
-
-\begin{lstlisting}
-  CPoint:Use point,colored, {
-    int x; int y;
-    method int x()x; method int y()y;
-    class method This of(int x, int y)
-      new Point(x,y);
-    constructor Point(int x, int y){
-      this.x=x;this.y=y;this.color=color;
-      }
-    }
-\end{lstlisting}
-
-\noindent after a while programming in this style, those "fixpoint" close the state in
-the obvious way classes becomes quite repetitive, and one wonder
-if it could be possible to automatically generate such code[]
-In our model we go one step further:
-In our model there is no need to generate code, or to explicitly
-write down constructors and fields; there is not even syntax for those
-constructor.
-The idea is that any class that "could" be completed in the obvious way
-is a complete "coherent" class.
-Others call abstract a class that have abstract methods.
-We call abstract a class that have a set of abstract methods that are not
-coherent, that is, you can not interpret those as factory,getters and setters.
-
-In the detail: a class with no abstract method is coherent, and like Java Math
-will just be usefull for calling class/static methods.
-a class with a single abstract class method returning This
-is coherent if all the other abstract methods can be seen as "abstract state
-operations" over one of its argument.
-For example,
-if there is a class method This of(int x, int y) as before,
-then
-a method int x() is an abstract state method, getter for x.
-a method Void x(int that) is abstract state method, setter for x.
-while getters and setters are fundamental operations, we can immagine
-more operations to be supported:
-
-\begin{lstlisting}
-method This withX(int that) 
-\end{lstlisting}
-may be a "wither", doing a functional field update
-  method Void update(int x,int) may do two field update at a time
-  method This clone() may do a shallow clone of the object.
-  We are not sure what is the best set of abstract state operations yet.
-
-  lets play with the points of before, to see what good can we do with the current
-  instruments:
-
-\begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-  PointAlgebra:Sum pointSum,pointMul
-\end{lstlisting}
-
-  As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods x,y and of.
-  In addition of Sum,Mul we may want many operations over points; can we improve our reuse
-  and not repeat such abstract definitions? of course!
-
-\begin{lstlisting}
-  p: { method int x(); method int y();
-    static method This of(int x,int y);
-    }
-  pointSum:Sum p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
-    }
-  pointMul:Sum p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
-    }
-  pointDiv: ...
-  PointAlgebra:Sum pointSum,pointMul,pointDiv,...
-\end{lstlisting}
-
-now our code is fully modularized, and each trait handle exactly one method.
-
-What happens if we want to add fields instead of just operations?
-
-\begin{lstlisting}
-  colored:{ method Color color() }
-  Point:Sum pointSum,colored
-\end{lstlisting}
-
-This first attempt does not work: the abstract color method
-is not a getter for any of the parameters of
-
-\begin{lstlisting}
-  class method This of(int x,int y) 
-\end{lstlisting}
-
-A solution is to provide a richer factory:
-
-\begin{lstlisting}
-  CPoint:Sum pointSum,colored,{
-    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/)
-    class method This of(int x, int y,Color color)
-    }
-\end{lstlisting}
-
-where we assume to support overloading on different parameter number.
-This is a good solution, we think is better that any alternatives in literature,
-however the method CPoint.sum reset the color to red.
-What should be the behaviour in this case?
-If instead of writing This.of() we used 
-this.withX(newX).withY(newY) we could preserve the color from this.
-Not sure if that would be better.
-If the point designer could predict this kind of extension, then we could
-use the following design:  
-\begin{lstlisting}
-  p: { method int x(); method int y();
-    method This withX(int that);
-    method This withY(int that);
-    static method This of(int x,int y);
-    method This merge(This other);
-    }
-  pointSum:Sum p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
-    }
-  colored:{method Color color();
-    method This withColor(Color that);
-    method This merge(This other)
-      this.withColor(this.color().mix(that.color());
-    }
-  CPoint:/*as before*/
-\end{lstlisting}  
-  Now we can merge colors, or any other kind of state we may want to add
-  following this pattern.
+\input{Intro}
+\input{Examples}
+\input{RelatedWork}
+\input{Conclusion}
 

### 2017-03-31 FACS/Examples.tex
.

@@
 by taking the union of the methods and the union of the implementations.
+In our simple mode, we consider an error trying to merge a class and an interface.
 
@@
 specific type is selected. We will discuss later the formal details of
-Sum.\marco{actually, this makes the sum not strongly associative...}
+\use.\marco{actually, this makes the sum not strongly associative...}
 
@@
 
-Note how since our only point of relaxation is the class literal, there is no way an error can ``move around'' and be duplicated during the compilation process.
+Note that we can reuse code only by naming traits; but our only point of relaxation is the class literal: there is no way an error can ``move around'' and be duplicated during the compilation process.
 In particular, our approach allows for safe libraries of traits and classes to be fully typechecked, deployed and reused by multiple clients: no type error will emerge from library code.
@@
 \section{Managing State}
-
-The Sum operator is very elegant, but our research community is struggling to
-make it work with constructor/state/fields.
-The goals fo this struggle are as follow:
+This idea of summing pieces of code is very elegant,
+and has proven very successfull in module composition languages~\cite{} but our research community is struggling to
+make it work with object state (constructor and fields) while acheving the following goals:
 
 \begin{itemize}
-
-\item keep sum associative and commutative.
-\item allowing a class to create instances of itself.
-\item actually initialize objects, not leaving no null fields
-\item managing fields in a way that borrows the elengance of summing methods
-\item make easy to add new fields
-
+\item keep sum associative and commutative,
+\item allowing a class to create instances of itself,
+\item actually initialize objects, leaving no null fields,
+\item managing fields in a way that borrows the elengance of summing methods,
+\item make easy to add new fields.
 \end{itemize}
@@
 
-  Pro: This approach is associative and commutative, even self construction
+\paragraph*{Advantages:} 
+this approach is associative and commutative, even self construction
   can be allowed if the trait requires a static/class method
@@
   a call to the constructor.
-  Negative: writing the class code with the constructors and
+
+\paragraph*{Negatives:} writing the class code with the constructors and
    fields and getter/setters and factories can be quite tedious.
-   There is no way for a trait to specify a default value for a field,
+   Moreover, there is no way for a trait to specify a default value for a field,
    the class need to handle all the state, even state that is conceptually
@@
 
-  for example
+Here in the following an exaple of such approach:
 
 \begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
+  pointSum: { method int x()  method int y()
+    class method This of(int x,int y)
     method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
+      This.of(this.x+other.x,this.y+other.y)
     }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
+  pointMul: { method int x() method int y()
+    class method This of(int x,int y)
     method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
+      This.of(this.x*other.x,this.y*other.y)
     }
@@
 
+As you see, all the state operations are represented as abstract methods.
+
 \subsection{A First Attempt at Composition}
-
+According to the general ideas expressed before,
 \begin{lstlisting}
@@
 
-\noindent would fail since methods x,y and of are still abstract.
-In this model the user is required to write something similar to
+\noindent would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract.
+In this mindset, the user would be required to write something similar to
 
 \begin{lstlisting}
-  CPoint:Use point,colored, {
+  CPoint:Use point,colored, {//not our suggested solution
     int x; int y;
@@
 
-\noindent after a while programming in this style, those "fixpoint" close the state in
-the obvious way classes becomes quite repetitive, and one wonder
-if it could be possible to automatically generate such code[]
-In our model we go one step further:
-In our model there is no need to generate code, or to explicitly
+\noindent after a while programming in this style, 
+writing those obvious ``close the state'' classes become a
+repetivie boring job, and one wonder
+if it could be possible to automatically generate such code[].
+Indeed those classes are just a form of ``fixpoint''.
+
+In our model we go one step further: there is no need to generate code, or to explicitly
 write down constructors and fields; there is not even syntax for those
 constructor.
-The idea is that any class that "could" be completed in the obvious way
-is a complete "coherent" class.
-Others call abstract a class that have abstract methods.
-We call abstract a class that have a set of abstract methods that are not
-coherent, that is, you can not interpret those as factory,getters and setters.
+The idea is that any class that ``could" be completed in the obvious way
+\emph{is a complete ``coherent" class}.
+In most other languages, a class is abstract if have abstract methods.
+Instead, we call abstract a class whose set of abstract methods is not
+coherent, that is, can not be automatically recognized as factory,getters and setters.
 
-In the detail: a class with no abstract method is coherent, and like Java Math
-will just be usefull for calling class/static methods.
-a class with a single abstract class method returning This
-is coherent if all the other abstract methods can be seen as "abstract state
-operations" over one of its argument.
+\paragraph*{Detaild definition of coherent:}
+\begin{itemize}
+\item a class with no abstract method is coherent, and like Java \Q@Math@.
+Will just be usefull for calling class/static methods.
+\item a class with a single abstract \Q@class@ method returning \Q@This@
+is coherent if all the other abstract methods can be seen as \emph{abstract state
+operations} over one of its argument.
 For example,
-if there is a class method This of(int x, int y) as before,
+if there is a \Q@class method This of(int x, int y)@ as before,
 then
-a method int x() is an abstract state method, getter for x.
-a method Void x(int that) is abstract state method, setter for x.
-while getters and setters are fundamental operations, we can immagine
-more operations to be supported:
+\item a method \Q@int x()@ is intepreted as an abstract state method: a getter for \Q@x@.
+\item a method \Q@Void x(int that)@ is a setter for x.
+\end{itemize}
 
-\begin{lstlisting}
-method This withX(int that) 
-\end{lstlisting}
-may be a "wither", doing a functional field update
-  method Void update(int x,int) may do two field update at a time
-  method This clone() may do a shallow clone of the object.
-  We are not sure what is the best set of abstract state operations yet.
+While getters and setters are fundamental operations, we can immagine
+more operations to be supported; for example
+\begin{itemize}
+\item \Q@method This withX(int that)@
+may be a ``wither", doing a functional field update.
+\item \Q@method Void update(int x,int)@
+may do two field update at a time.
+\item\Q@method This clone()@ may do a shallow clone of the object.
+\end{itemize}
+
+We are not sure what is the best set of abstract state operations yet, but we think this could become a very interesting area of research.
 
@@
 \begin{lstlisting}
-  pointSum: { method int x(); method int y();
-    static method This of(int x,int y);
+ //same code as before for pointSum and pointMul
+  pointSum: { method int x() method int y()
+    static method This of(int x,int y)
     method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
+      This.of(this.x+other.x,this.y+other.y)
     }
-  pointMul: { method int x(); method int y();
-    static method This of(int x,int y);
+  pointMul: { method int x() method int y()//look we are repeating
+    static method This of(int x,int y)//the abstract method declarations.
     method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
+      This.of(this.x*other.x,this.y*other.y)
     }
@@
   As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods x,y and of.
-  In addition of Sum,Mul we may want many operations over points; can we improve our reuse
+  as we wish. However we have to repeat the abstract methods \Q@x@,\Q@y@ and \Q@of@.
+  In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
   and not repeat such abstract definitions? of course!
@@
 \begin{lstlisting}
-  p: { method int x(); method int y();
-    static method This of(int x,int y);
+  p: { method int x() method int y()
+    static method This of(int x,int y)
     }
-  pointSum:Sum p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y);
+  pointSum:Use p, { method This sum(This other)
+      This.of(this.x+other.x,this.y+other.y)
     }
-  pointMul:Sum p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y);
+  pointMul:Use p, { method This mul(This other)
+      This.of(this.x*other.x,this.y*other.y)
     }
   pointDiv: ...
-  PointAlgebra:Sum pointSum,pointMul,pointDiv,...
+  PointAlgebra:Use pointSum,pointMul,pointDiv,...
 \end{lstlisting}
 
-now our code is fully modularized, and each trait handle exactly one method.
+Now our code is fully modularized, and each trait handle exactly one method.
 
@@
   colored:{ method Color color() }
-  Point:Sum pointSum,colored
+  Point:Sum pointSum,colored //fails
 \end{lstlisting}
@@
 is not a getter for any of the parameters of 
-
-\begin{lstlisting}
-  class method This of(int x,int y) 
-\end{lstlisting}
-
+\Q@ class method This of(int x,int y)@
 A solution is to provide a richer factory:
@@
 What should be the behaviour in this case?
-If instead of writing This.of() we used 
-this.withX(newX).withY(newY) we could preserve the color from this.
-Not sure if that would be better.
-If the point designer could predict this kind of extension, then we could
-use the following design:  
+If we support withers, instead of writing This.of() we can use
+this.withX(newX).withY(newY) in order to preserve the color from \Q@this@.
+Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
+
+If the point designer could predict this kind of extension, then we could use the following design:  
 \begin{lstlisting}
-  p: { method int x(); method int y();
-    method This withX(int that);
-    method This withY(int that);
-    static method This of(int x,int y);
-    method This merge(This other);
+  p: { method int x() method int y()
+    method This withX(int that)
+    method This withY(int that)
+    static method This of(int x,int y)
+    method This merge(This other)
     }
-  pointSum:Sum p, { method This sum(This other)
+  pointSum:Use p, { method This sum(This other)
       this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
     }
-  colored:{method Color color();
-    method This withColor(Color that);
+  colored:{method Color color()
+    method This withColor(Color that)
     method This merge(This other)
@@
   following this pattern.
+  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
+  need to compose the merge operation inside of both of them.
+  The simple model we are presenting could accomodate this with an
+  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
+This is similar to the \emph{override} operation present in the original trait model[].

### 2017-03-31 FACS/Extensions.tex
.

\section{Extensions to our model}
  One of the main feature of our simple reuse/use model is that it can be
  easly extended. One simple but amazingly expressive extension is nested classes

\subsection{Nested classes}

A nested class will be another kind of member in the Literal, so  
the grammar can be updated as following:

\begin{bnf}
\prodFull\mMD{
\Opt{\terminalCode{class}}\ \terminalCode{method}\ \mT\ \mm\oR\overline{\mT\,\mx}\cR \Opt\me
\mid \mCD
}{Member Decl}\\
\prodFull\mT{\mC\mid\mC\terminalCode{.}\mT}{types are now paths}
\end{bnf}\\

The general idea is that by composing code with \use, nested classes with the same name are recursivelly composed.
Note that while we have nested classes, we do not have nested traits: all traits are still
at top level.
Untypable/unresolved Traits are also the only``dependency"
the type system keeps track of, this means that when a nested class at an arbirary
nested level is flattend, as in
\Q@C:{ D:{ E:Use t1,t2,L}}@
t1 and t2 must be defined before C at top level; and they may require classes (and their
nested) defined before C. This means that the type system can still consider
the class table as a simple map from Types T to their definition.

This extension lets us challenge the expression problem[]:
in the expression problem we have data variants and operations and....

Let see how to easily encode and solve the expression problem:

\begin{lstlisting}
exp:{Exp:{interface}}//Exp declared once, reused everywhere
lit:Use exp,{ Lit:{implements Exp //Exp not explicitly declared
    class method Lit of(int inner) //Lit abstract state
    method int inner()}
  }
sum:Use  exp,{ Sum:{implements Exp 
    class method This of(Exp left, Exp right)
    method Exp left() //Sum abstract state
    method Exp right()}
  }  
uminus:Use exp,{ UMinus:{implements Exp 
    class method This of(exp inner)//and so on for
    method Exp inner()}//all the needed datavariants
  }

expToS:{Exp:{interface method String toString()}}
//concept of toString declared once

sumToS:Use sum,expToS,{ Sum:{implements Exp//with Exp.toString
    method String toString()//just the implementation of the
      left.toString()+"+"+right.toString()//specific method
  }
uminusToS:...//implement toString for all the datavariants

expEval:{Exp:{interface method int eval()}}
//declare the next operation and implement it for all the datavariant
\end{lstlisting}

Now that you have nicely modularized the code, just compose all the traits you need.
\Q@MySolution:Use sumToS,litToS@ //sum,lit and exp traits are alread included

The expression problem presented up to now is the traditional challenge proposed by [];
this has been criticized to not really address the fundamental issues since it does not handle ....
Now we show how we can go behond the traditional expression problem by encoding transformer methods:
For example, lets add 1 to all literals
\begin{lstlisting}
expAdd1:{Exp:{interface method Exp add1()}}
sumAdd1:Use sum,expAdd1,{Sum:{implements Exp
    method Exp add1()
      Sum.of(left.add1,right.add1())
  }
litAdd1:Use lit,expAdd1,{Lit:{implements Exp
    method Exp add1()
      Lit.of(inner()+1);
    }

MySolutionAdd1:Use sumToS,litToS,sumAdd1,litAdd1
\end{lstlisting}

This nicely solve our problem. 
However, notice how if we wished to add many similar operations we would 
have to repeat the propagation code (as in \Q@sumAdd1@) many times
just changing the name of the operation.
In the next section we will show how to improve on this point.

\section{More composition operators}
\use is amazing, elegant and simple, but the system can be easily enriched with more 
operators: while most approaches in literator present a fixed set of operators, in our
system we do not need such restriction.
We just need to be sure that every newly added operator respects the following criteria:
if starts form well typed code, it produces well typed code.\marco{not enough? rephrease to respect the two quality at the start?}

---------
Restrict make abstract method

--------
Rename, like refactoring
Rename can invoke sum inside
---------
Redirect, like generics
Redirect can be multiple.

With redirect+rename we can have the general operator propagator

operation:{//for sum and lit, easy to extends as before
  T:{}
  Exp:{interface method Exp op(T x)}}
  Sum:Sum sum,{ extends Exp sum,expAdd1,{
    method Exp op(T x)
      Sum.of(left.op(x),right.op(x));
    }
  Lit:Sum lit,{
    method Exp op(T x)
      this;
  }
Now, to have my addN I can
opAddn: Sum
  operation[redirect T to Int]
    [rename Exp.op(x) to addN(x)][restrict Lit.op(x)], {
  Lit:{method Exp addN(Int x) Lit.of(inner())+x}
  }

Full power of redirect

serviceCombinator:{ S:{implements Service
  method R report()
  }

  R:{method R combine(R that)
    class method R empty()}

  ListS:list[redirect Elem to S]

  class R method(ListS ss){//here we use extended java like syntax
    R r=R.empty()
    for(S s in ss){
      s.performService();
      r=r.combine(s.report())
      }
    return r;
  }
}
PaintingService:serviceCombinator[redirect S to Painting.Service]

To encode the former generic code in java you need to write
the following headeche inducing interfaces for RService and Report (and still we miss the empty method)
and require that the services you want to serve implement those.
  interface Service{ void performService();}
  interface Report<R extends Report<R>>{R combine(R that);}
  interface RService<R extends Report<R>> extends Service{ R report();}

### 2017-04-02 FACS/Extensions.tex
.

@@
 \section{More composition operators}
-\use is amazing, elegant and simple, but the system can be easily enriched with more 
-operators: while most approaches in literator present a fixed set of operators, in our
-system we do not need such restriction.
+\use is amazing, elegant and simple, but our system can be easily enriched with more 
+operators: while most approaches in literature presents a fixed set of operators, we do not need such restriction.
 We just need to be sure that every newly added operator respects the following criteria:
-if starts form well typed code, it produces well typed code.\marco{not enough? rephrease to respect the two quality at the start?}
 
----------
-Restrict make abstract method
+\begin{itemize}
+\item As for \use, the operato does not need to be total, but if it fails it needs to provide an error that will be reported to the programmer.
+\item When the operator takes in input only traits (they are going to be well typed), if a result is produced,
+ such result is well typed.
+\item When the operator takes in input also code literal, if a non well typed result is produced,
+the type error must be tracked back to code in one of those non typed yet code literals.
+ \end{itemize}
 
---------
-Rename, like refactoring
-Rename can invoke sum inside
----------
-Redirect, like generics
-Redirect can be multiple.
+Here we see some example of other composition operators:
+\paragraph*{Restrict}${}_{}$\\*
+Restrict makes a method abstract
+\begin{lstlisting}
+t:{method bool geq(This x) x.leq(this)   method bool leq(This x) x.qeq(this) }
+C:Use t[restrict geq],{method bool geq(This x)  /*actual geq impl*/}
+\end{lstlisting}
+
+A variant of this operator allows to move the implementation to another name.
+This is very usefull to implement \Q@super@.
+This also is enough to suppor the example of before, when we try to compose \Q@colored@
+with \Q@flavoured@.
+
+\begin{lstlisting}
+t:{method int count() 0 }
+C:Use t[restrict count as _count],{method int count() this._count()+1}
+\end{lstlisting}
+Note how we are leveraging on the fact that the code literal does not need to be complete, thus we can just call \Q@_count()@.
+
+\paragraph*{Rename}${}_{}$\\*
+Rename allows to make some form of ``compile time'' refactoring
+There are a lots of different forms of rename in literature,
+sometime allowing only to rename specific methods, sometime allowing to rename
+nested classes into other nested classes either at the same or at a different nesting level.
+Renaming in the context of nested classes also means that when renaming a method of an interface, all the 
+nested classes implementing such interface inside of that code literal need to be adjusted.
+Renaming need to rename not only the method headers, but all the method calls inside of method bodies.
+At first glimps, this seams to be not always possible since we are considering to be able to apply those
+operators also to non well typed code.
+However, if the expression language is simple enough, it is possible to pre process the code to
+annotate the expected receiver type on all method calls by doing a purelly sintactic analysis
+on a single code literal in isolation. 
+All the expression whose type is guessed to be out of the border of the literal can stay unannotated; they are not going to be renamed anyway.
+
+\begin{lstlisting}
+t:{ I:{interface method int mI() }
+     A:{implements I  method int mI() 42}
+     B:{ method int mB(I i, A a, C c) i.mI()+a.mI()+c.mI()}
+     //mB would be annotated i[I].mI()+a[A].mI()+c.mI()
+     }
+D:t[rename A.mI kI]
+\end{lstlisting}
+ Notice how we are sure that c does not implements I since it is invisible from the outside: traits does not introduce nominal types!
+
+ We expect the flattened version for \Q@D@ to be
+\begin{lstlisting}
+D:{ I:{interface method int kI() }
+     A:{implements I  method int kI() 42}
+     B:{ method int mB(I i, A a, C c) i.kI()+a.kI()+c.mI()}
+     }
+\end{lstlisting}
+
+\paragraph*{Redirect}${}_{}$\\*
+Redirect allows to emulate generics; the main idea is that a (fully abstract) class can be redirect to another one external to the trait/code literal.
+For example a linked list can be implement as
+\begin{lstlisting}
+list:{ Elem:{}
+     Cell:{class method Cell of(Elem e,Cell c) 
+       method Elem e()  method Cell c()
+       }
+   method Elem get(int x) ...
+   ...more methods..
+   }
+ListString:list[redirect Elem to String]
+\end{lstlisting}
+
+%An expressive form of Redirect can be multiple, that is, can redirect may interdependent classes at the same time.
+%We show a graph example, where also we can show how to propagate generics:
+%For example
+%\begin{lstlisting}
+%t:{ method boolean reachable(Node start, Node end)/*implements reachability*/
+%     Node:{method ListEdge nodes()}
+%     Edge:{method Node in()  method Node out()}
+%     ListEdge:list[redirect Elem to Edge]
+%     }
+%\end{lstlisting}
+
+%Redirect can be multiple.
+
+\paragraph*{Application on the expression problem}${}_{}$\\*
 
@@
   Exp:{interface method Exp op(T x)}}
-  Sum:Sum sum,{ extends Exp sum,expAdd1,{
+  Sum:Use sum,{ extends Exp sum,expAdd1,{
     method Exp op(T x)
-      Sum.of(left.op(x),right.op(x));
-    }
-  Lit:Sum lit,{
-    method Exp op(T x)
-      this;
+      Sum.of(left.op(x),right.op(x))  }
+  Lit:Use lit,{
+    method Exp op(T x)  this
   }
 Now, to have my addN I can
-opAddn: Sum
+opAddn: Use
   operation[redirect T to Int]

### 2017-04-03 FACS/Conclusion.tex
added example in intro

\section{Conclusions and practical applications}

In this paper we explained a simple model to 
radically decouple inheritance/code reuse and subtyping.

We show how our model can be easly extended with more operators, if they behave well with respect to a simple predicate.

In order to focus on the explanation of the high level concepts, this paper do not present formalism for our approach.
Formalism for similar approaches can be found in[deefjig, myThesis].

We are working on 42: a full blown language 
that leverage on the ideas presented in this paper to obtain reliable and understandble metaprogramming.
Formalization (in progress) for such language can be found at URLURLURL.
42 extands our model allowing
flattening to execute arbitrary computations.
In such model we do not need an explicit notion of traits: they are encoded as method returning a code literal.

  In addition to this, 42 have feature less related to code composition, like
  a strong type system supporting aliasing mutability[] and circularity control[],
   checked exceptions[], errors (unchecked exceptions) with strong exception safety for errors[].

42 do not have a finite set of composion operators; they can be added using the built in support for native method calls. They can be dynamically checked to verify that they are well behaved according to our predicate, or they can be trusted to acheive efficiency.

%\item more operators can be accomodated without changing the general model

### 2017-04-03 FACS/Extensions.tex
added example in intro

@@
 \section{More composition operators}
-\use is amazing, elegant and simple, but our system can be easily enriched with more 
+\use\ is amazing, elegant and simple, but our system can be easily enriched with more 
 operators: while most approaches in literature presents a fixed set of operators, we do not need such restriction.
@@
 \paragraph*{Application on the expression problem}${}_{}$\\*
-
-With redirect+rename we can have the general operator propagator
-
+With redirect, rename and restrict we can have the general operator propagator
+\begin{lstlisting}
 operation:{//for sum and lit, easy to extends as before
@@
   }
+\end{lstlisting}
+
 Now, to have my addN I can
+
+\begin{lstlisting}
 opAddn: Use
@@
   }
+\end{lstlisting}
 
-Full power of redirect
+\paragraph*{Full power of redirect}${}_{}$\\*
+An expressive form of Redirect can be multiple, that is, can redirect may interdependent classes at the same time.
+We show an example where a specific kind of \Q@Service@ can produce a \Q@Report@, and 
+\Q@Report@s can be combined together.
+The goal is to execute a list of such services and produce a collated report.
+This example also show how to propagate generics:
 
-serviceCombinator:{ S:{implements Service
-  method R report()
-  }
+\begin{lstlisting}
+Service:{interface method Void performService()}
+serviceCombinator:{
+  S:{implements Service method R report()  }
 
-  R:{method R combine(R that)
-    class method R empty()}
+  R:{method R combine(R that)   class method R empty() }
 
@@
 
-  class R method(ListS ss){//here we use extended java like syntax
+  class method R doAll(ListS ss){//here we use extended java like syntax
     R r=R.empty()
@@
 }
-PaintingService:serviceCombinator[redirect S to Painting.Service]
+PaintingService:serviceCombinator[redirect S to PaintingService]
+PaintingService:{... method PaintingReport report()..}
+PaintingReport:{..}
+\end{lstlisting}
+
+The flattened version of \Q@PaintingService@ would look like:
+\begin{lstlisting}
+PaintingService:{
+  ListS:/*the expansion of list[redirect Elem to PaintingService]*/
+
+  class method PaintingReport doAll(ListS ss){
+    PaintingReport r=PaintingReport.empty()
+    for(PaintingService s in ss){
+      s.performService();
+      r=r.combine(s.report())
+      }
+    return r;
+  }
+}
+\end{lstlisting}
+Where you can note how redirect figured out \Q@R=PaintingReport@ by comparing the structural shape of
+classes \Q@PaintingService@ and \Q@S@.
 
 To encode the former generic code in java you need to write
-the following headeche inducing interfaces for RService and Report (and still we miss the empty method)
+the following headeche inducing interfaces for RService and Report.
 and require that the services you want to serve implement those.
+\begin{lstlisting}
 interface Service{ void performService();}
@@
 interface RService<R extends Report<R>> extends Service{ R report();}
+\end{lstlisting}
+Note how we still can not encode the method \Q@empty@.

### 2017-04-03 FACS/Intro.tex
added example in intro

@@
 must not imply subtyping to type-check.}
+An iconic example on why connecting inheritance/code reuse and subtpying is problematic is
+provided by the historic[]:
+A reasonable implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@ keeping tracks of how many times an element occurs.
+We would just add some state and override a couple of methods.
+
+However, our subclassing would break LSP[]: not all bags are sets!
+Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
+and making both \Q@Bag@ and \Q@Set@ inherit from it.
+This looks innatural, since \Q@Set@ would extend it without adding anything,
+and we would be suprtised to find a use of the type \Q@AbstractSetOrBag@.
+Worst, if we was to constantly apply this mentalty, we would introduce a very high number
+of abstract classes that are not supposed to be used as types, and that will clutter the 
+public interface of our classes and our code project as a whole.
 
@@
 code reuse
+
+\subsection{List of contribution}
+\marco{text to improve, but is nice to keep a list}
+\begin{itemize}
+\item simple approach to fully separate inheritance and subtyping
+\item easy to grasp model of compilation were typing is late but not too late (lazy?)
+\item clean and elegant handling of state in trat/module composition language
+\item extension with nested class is natural and powefull
+\item more operators can be accomodated without changing the general model
+\end{itemize}

### 2017-04-03 FACS/RelatedWork.tex
added example in intro

\section{Related Work}
The literature on code reuse is too vast to let us do justice of it in a fiew pages,
Our work is clearly inspired on Traits[sharly]; they are in turn insipired on module composition languages[]
and multiple inheritance[].

Our idea of fully separate inheritance and subtyping has some precedence
In all the line of work of Ferruccio [cite many] neither traits nor classes introduce nominal types,
that are provided only by interfaces.
In DeepFjig[] classes introduce nominal types but in an innatural way, the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class.

Other work cite[something on family polimorphism, scala?]
do no completly break the relation between inheritance and subtyping, but only prevent subtyping where 
it would be unsound.

Our compilation model is a clear step forward with respect to 
\cite[metaFjig, deepFjig,myThesis], where all code was required to declare all the abstract dependencies all of the time.

%\item clean and elegant handling of state in trat/module composition language
The idea of abstract state operation has emerged from CJ[], and offer a clean solution to handling state
in a trait composition setting.
Note how abstract state operation are different from just hiding fields under getter and setters: 
in our model the programmer simply never declare what is the state of the class, not even what information is stored in fields.
The state is computed by the system as an overall result of the whole code composition process.

In literature there has been many attempts to add state in traits/module composition languages.
\begin{itemize}  
\item No constructor: all the fields start at null/a default specified value.
  In this model fields are like another kind of member, and two fields
  with identical type can be merged by sum; \Q@new C()@ can be used for all classes, and \Q@init@ method may be called later, as in
  \Q@Point p=new Point(); p.init(10,30)@.

  To its credit, this simple approach is commutative and associative and do not disrupt elegance of summing methods.
  However, objects are created "broken" and the user is trusted with fixing them.
  While is easy to add fields, the load of inizializing them is on the user; moreover
    all the objects are intrinsically mutable, so this model is unfriendly
    to a functional programming style.
\item Constructor compose fields:
In this approach the fields are declared but not initialized, and
a canonical constrctor constructor taking a value for each field and just initializing such field
is automatically generated in the resulting class.

In this approach it is easy to add fields, however this approach is associative but not commutative, since the order of composition will change the order of the fields; moreover there is no support for self construction
since the signature of the constuctor change during composition.

\item Constructor can be composed if they offers the same exact parameters:
In this approach traits declare field and constructors.
The constructor initialize the fields but can do any other computation.
Traits whose constructors have the same signature can be composed.
The composed constructor will execute both constructor bodies in order.

This approach is designed to allows Self construction.
It is also associative and mostly commutative: the order of composition is relevant only
in the order of execution of side effects during construction.
However the requirement that the constructors need the same parameters 
hamper reuse, and if a field is added, its initial value need to be
somehow syntetized from the old constructor parameters.

\end{itemize}

%\item extension with nested class is natural and powefull
We claim that our solution to the expression problem is the most natural in literature to date:
While a similar sintax can be achieved with the scandinavian style[], their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system[Lightwait family polimporphism?].

Our close contendant is DeepFJig[]: all our gain over their model is based on our relaxation over abstract signatures.
Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}

### 2017-04-03 FACS/Examples.tex
.

@@
 \begin{lstlisting}
-  B:Use ta, { int mb(){return this.ma();} }
+  B:Use ta, { method int mb(){return this.ma();} }
 \end{lstlisting}
@@
 A:Use ta
-B:Use ta, { int mb(){return this.ma();} }
+B:Use ta, { method int mb(){return this.ma();} }
 \end{lstlisting}
@@
   method int ma() Utils.m(this)
-  int mb() this.ma() }
+  method int mb() this.ma() }
 
@@
 \begin{lstlisting}
-t:{int m() 2 
-   int n() this.m()+1}
+t:{method int m() 2 
+   method int n() this.m()+1}
 \end{lstlisting}
@@
 \begin{lstlisting}
-t1:{int n() this.m()+1} 
+t1:{method int n() this.m()+1} 
 \end{lstlisting}
@@
 \begin{lstlisting}
-C: Use t {int k() this.n()+this.m()}
+C: Use t {method int k() this.n()+this.m()}
 \end{lstlisting}
 \noindent would be correct: even if n,m are not defined inside
-\Q@{int k() this.n()+this.m()}@,
+\Q@{method int k() this.n()+this.m()}@,
 the result of the flattening is well typed.
@@
 \begin{lstlisting}
-A:{int ma(B b) b.mb()+1}
-tb:{int mb() 2}
-tc:{int mc(A a,B b) a.ma(b)}
+A:{method int ma(B b) b.mb()+1}
+tb:{method int mb() 2}
+tc:{method int mc(A a,B b) a.ma(b)}
 B: Use tb
@@
 \item  (1)First check \Q@B@ and produce \Q@B@ code (that also contains \Q@B@ structural shape),
-  (2) then use \Q@B@ shape to check \Q@C@ and produce \Q@C@ code
+  (2) then use \Q@B@ shape to check \Q@C@ and produce \Q@C@ code;\ 
 or a more involved
@@
 \begin{lstlisting}
-  CPoint:Use point,colored, {//not our suggested solution
-    int x; int y;
-    method int x()x; method int y()y;
+  CPoint:Use pointSum,pointMul, {//not our suggested solution
+    int x   int y
+    method int x()x   method int y()y
     class method This of(int x, int y)
-      new Point(x,y);
-    constructor Point(int x, int y){
-      this.x=x;this.y=y;this.color=color;
-      }
+      new Point(x,y)
+    constructor Point(int x, int y){ this.x=x   this.y=y }
     }
@@
 may be a ``wither", doing a functional field update.
-\item \Q@method Void update(int x,int)@
+\item \Q@method Void update(int x,int y)@
 may do two field update at a time.
@@
   pointSum: { method int x() method int y()
-    static method This of(int x,int y)
+    class method This of(int x,int y)
     method This sum(This other)
@@
   pointMul: { method int x() method int y()//look we are repeating
-    static method This of(int x,int y)//the abstract method declarations.
+    class method This of(int x,int y)//the abstract method declarations.
     method This mul(This other)
@@
   p: { method int x() method int y()
-    static method This of(int x,int y)
+    class method This of(int x,int y)
     }
@@
 \begin{lstlisting}
-  CPoint:Sum pointSum,colored,{
-    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/)
+  CPoint:Use pointSum,colored,{
+    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/))
     class method This of(int x, int y,Color color)
@@
 This is a good solution, we think is better that any alternatives in literature,
-however the method CPoint.sum reset the color to red.
+however the method \Q@CPoint.sum@ resets the color to red.
 What should be the behaviour in this case?
-If we support withers, instead of writing This.of() we can use
-this.withX(newX).withY(newY) in order to preserve the color from \Q@this@.
+If we support withers, instead of writing \Q@This.of()@ we can use
+\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
 Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
@@
   pointSum:Use p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y);
+      this.merge(other).withX(this.x+other.x).withY(this.y+other.y)
     }
@@
     method This merge(This other)
-      this.withColor(this.color().mix(that.color());
+      this.withColor(this.color().mix(that.color())
     }

### 2017-04-03 FACS/Extensions.tex
.

@@
 Now that you have nicely modularized the code, just compose all the traits you need.
-\Q@MySolution:Use sumToS,litToS@ //sum,lit and exp traits are alread included
+\begin{lstlisting}
+MySolution:Use sumToS,litToS
+//sum,lit and exp traits are alread included
+\end{lstlisting}
 
@@
 
-\section{More composition operators}
+\subsection{More composition operators}
 \use\ is amazing, elegant and simple, but our system can be easily enriched with more 
@@
 \begin{lstlisting}
-t:{method int count() 0 }
-C:Use t[restrict count as _count],{method int count() this._count()+1}
+  colored:{/*as before*/}
+  flavoured:{
+  method Flavour flavour()
+    method This withFlavour(Flavour that)
+    method This merge(This other)
+      this.withFlavour(this.flavour().mix(that.flavour())
+    }
+  CPoint:Use colored[restrict merge as _1merge],flavoured[restrict merge as _2merge],
+  pointSum,{
+    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/))
+    class method This of(int x, int y,Color color)
+    method This merge(This other)
+      this._1merge(other)._2merge(other)
+    }
 \end{lstlisting}
-Note how we are leveraging on the fact that the code literal does not need to be complete, thus we can just call \Q@_count()@.
+
+Note how we are leveraging on the fact that the code literal does not need to be complete, thus we can just call \Q@_1merge()@ and \Q@_2merge()@
 
@@
 \end{lstlisting}
+
+Hide can be seen as a variation of rename, where the method/class is renamed to a fresh unguessable name.
 

### 2017-04-03 FACS/Intro.tex
.

@@
 exposed in this example (using this, extends, method call)
-is forced to accept that subclassing implies subtyping.
+is forced to accept that subclassing implies subtyping.\footnote{
 Note how this is true also in C++, where is possible to
@@
 subtyping but not over subtyping itself, and the former example
-would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.
+would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.}
 \marco{Cite some work of bruceTHisType and show how he also fails.}
@@
 This looks innatural, since \Q@Set@ would extend it without adding anything,
-and we would be suprtised to find a use of the type \Q@AbstractSetOrBag@.
+and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
 Worst, if we was to constantly apply this mentalty, we would introduce a very high number

### 2017-04-04 FACS/Intro.tex
bib

@@
 An iconic example on why connecting inheritance/code reuse and subtpying is problematic is
-provided by the historic[]:
+provided by the historic\cite{LaLonde:1991:SSS:110673.110679}:
 A reasonable implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@ keeping tracks of how many times an element occurs.

### 2017-04-04 FACS/Examples.tex
.

@@
 A version of this language where there are no traits can be seen 
-as a restriction/variation of FJ~\cite{}.
+as a restriction/variation of FJ~\cite{igarashi2001featherweight}.
 
@@
 Our typing discipline is 
-what distinguish our approach from a simple minded code composition macro(cite)
-or a rigid module composition(cite may be some of my work).
+what distinguish our approach from a simple minded code composition macro~\cite{bawden1999quasiquotation}
+or a rigid module composition~\cite{ancona2002calculus}
 
@@
 In class expressions  $\use\ \overline\mV$
-an \mL\ in $\overline\mV$ is not well typed before flattening, only the result is supposed to be well typed.
+an \mL\ in $\overline\mV$ is not typechecked before flattening, and only the result is expected to be well-typed.
 While this seams a very dangerous approach at first, consider that also Java have the same behaviour:
@@
 \item on the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
-$\mL_1\ldots\mL_n \in \overline\mV$, and after successfull flattening $\mC : \mL$ is not well typed,
-the error was originally present in one of $\mL_1\ldots\mL_n$.
+$\mL_1\ldots\mL_n \in \overline\mV$, and after successfull flattening $\mC : \mL$ can not be typechecked,
+then the issue was originally present in one of $\mL_1\ldots\mL_n$.
+It may be that the result is intrinsically ill-typed, if of the methods in $\mL_1\ldots\mL_n$ is not well typed,
+but it may also happens that a type referred from one of those methods is declared \emph{after} the current class. As we will see later, this is how our relaxation allows to support recursive types.
+
 This also means that as an optimization strategy

### 2017-04-04 FACS/Intro.tex
.

@@
 
-However, our subclassing would break LSP[]: not all bags are sets!
+However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
 Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
@@
 
-Our design leverage on traits~\cite{}: a well know mechanisms for pure
+Our design leverage on traits~\cite{ducasse2006traits}: a well know mechanisms for pure
 code reuse

### 2017-04-04 FACS/Examples.tex
.

@@
 This idea of summing pieces of code is very elegant,
-and has proven very successfull in module composition languages~\cite{} but our research community is struggling to
+and has proven very successfull in module composition languages~\cite{ancona2002calculus} but our research community is struggling to
 make it work with object state (constructor and fields) while acheving the following goals:
@@
   allowing declaration for fields and constructors.
-  This approach is very powefull[see class less java]
+  This approach is very powefull~\cite{wang2016classless}
 
@@
 repetivie boring job, and one wonder
-if it could be possible to automatically generate such code[].
+if it could be possible to automatically generate such code~\cite{wang2016classless}.
 Indeed those classes are just a form of ``fixpoint''.
@@
   extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
-This is similar to the \emph{override} operation present in the original trait model[].
+This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.

### 2017-04-04 FACS/RelatedWork.tex
.

@@
 
-Other work cite[something on family polimorphism, scala?]
+Other work cite~\cite{odersky2008programming,nystrom2006j}
 do no completly break the relation between inheritance and subtyping, but only prevent subtyping where 
@@
 %\item clean and elegant handling of state in trat/module composition language
-The idea of abstract state operation has emerged from CJ[], and offer a clean solution to handling state
+The idea of abstract state operation has emerged from CJ~\cite{wang2016classless}, and offer a clean solution to handling state
 in a trait composition setting.
@@
 We claim that our solution to the expression problem is the most natural in literature to date:
-While a similar sintax can be achieved with the scandinavian style[], their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system[Lightwait family polimporphism?].
+While a similar sintax can be achieved with the scandinavian style~\cite{ernst2004expression}, their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system~\cite{igarashi2005lightweight}.
 

### 2017-04-04 FACS/RelatedWork.tex
.

@@
 The literature on code reuse is too vast to let us do justice of it in a fiew pages,
-Our work is clearly inspired on Traits[sharly]; they are in turn insipired on module composition languages[]
-and multiple inheritance[].
+Our work is clearly inspired on Traits~\cite{ducasse2006traits}; they are in turn insipired on module composition languages~\cite{ancona2002calculus}
+and multiple inheritance~\cite{}.
 
@@
 
-Our close contendant is DeepFJig[]: all our gain over their model is based on our relaxation over abstract signatures.
+Our close contendant is DeepFJig~\cite{corradi2011deepfjig}: all our gain over their model is based on our relaxation over abstract signatures.
 Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}

### 2017-04-27 FACS/Extensions.tex

@@
 
-This extension lets us challenge the expression problem[]:
+This extension lets us challenge the expression problem~\cite{wadler1998expression}:
 in the expression problem we have data variants and operations and....
@@
 
-The expression problem presented up to now is the traditional challenge proposed by [];
+The expression problem presented up to now is the traditional challenge proposed by~\cite{wadler1998expression};
 this has been criticized to not really address the fundamental issues since it does not handle ....
@@
 
-Here we see some example of other composition operators:
+Here we see some example of other composition operators, inspired from~\cite{servetto2014meta}:
 \paragraph*{Restrict}${}_{}$\\*

### 2017-04-27 GPCE2016/sections/Conclusion.tex

@@
+\vspace{-5pt}
 \section{Conclusion}\label{sec:conclusion}
-Before Java 8, concrete methods were not allowed
-to appear in interfaces.  Java 8 allows static interface methods and
-introduces \emph{default methods}, which enable implementations
-inside interfaces. An important positive consequence that
+From Java 8, static and default methods are allowed in interfaces, 
+which enable implementations inside interfaces. An important positive consequence that
 was probably overlooked is that the concept of class
@@
 
+\begin{comment}
 Classless Java is just a programming style, but showing the way
@@
 by limitations of Java and its syntax.
+\end{comment}
 

### 2017-04-27 poster/splash2016.tex

@@
 \newcommand{\compresslist}{%
-\setlength{\itemsep}{1pt}%
+\setlength{\itemsep}{0pt}%
 \setlength{\parskip}{0pt}%
@@
   % Column spacing
-  colspacing=1em,
+  colspacing=0.5em,
   % Color style
@@
 
-   \begin{enumerate}\compresslist
-      \item Generally speaking, multiple inheritance with fields is hard. Moreover, there 
-      are problems with field type refinement, constructors, etc.
-      \item Java support multiple inheritance with default methods in a limited way.
-      \item Other models (traits, mixins, etc) have certain limitatations.
-      \item Our approach is related with traits. But we use operations on state to simulate fields 
-      which traits do not support.
+  \noindent \begin{enumerate}\compresslist
+      \item Multiple inheritance with state is hard. 
+      \noindent \begin{itemize}\compresslist
+        \item Trait is a nice  model but without state.
+        \item Multiple inheritance in C++ is complex. Scala traits have fields but no constructors.
+        \item Java supports multiple inheritance with default methods in a limited way.
+      \end{itemize}
+      \item Fields usually cannot be type-refined, which prevents easy solutions to the Expression Problem.
+      % 
+      % \item Other models (traits, mixins, etc) have certain limitations.
+      % \item Our approach is related with traits. But we use operations on state to simulate fields 
+      % which traits do not support.
    \end{enumerate}
-   \vspace{0.3em}
  }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-   \begin{enumerate}\compresslist
-    \item IB and Object Interfaces: enable powerful programming idioms using multiple-inheritance, type-refinement and abstract state operations.
+  \noindent \begin{enumerate}\compresslist
+  \item Abstract state operations: 
+      \noindent \begin{itemize} \compresslist
+          \item a new way to think about state, via operations instead of fields directly
+          \item easy to combine with \emph{multiple (trait) inheritance}
+          \item supports constructors, co-variant type-refinement of state
+      \end{itemize}    
+  % \item Classless Java: a Java extension that allows an interface-based programming style
+    % \item IB and Object Interfaces: enable powerful programming idioms using multiple-inheritance, type-refinement and abstract state operations.
     \item Classless Java: a practical realization of IB in Java.
-    \item Type-safe covariant mutable state
-    \item Applications and case studies
+    % \item Type-safe covariant mutable state
+    % \item Applications and case studies
    \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the result of the maze game case study, both SLOC and number of interfaces are greatly reduced:
-\noindent{\includegraphics[width=0.95\linewidth]{pdfs/result.png}\\}
+In the maze game case study, both SLOC and \# of interfaces are greatly reduced:
+\noindent{\centering\includegraphics[width=0.81\linewidth]{pdfs/result.png}\\}
 \vspace{0.1em}
@@
 
+\vspace{-0.5em}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
-\lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
+% \lstinputlisting[linerange=6-13]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_JAVA
+% \lstinputlisting[linerange=17-19]{../UseMixinLombok/src/pegasus/simple/java8/Main.java}% APPLY:linerange=PEGASUS_INST
 \lstinputlisting[linerange=8-12]{../UseMixinLombok/src/pegasus/simple/lombok/Main.java}% APPLY:linerange=PEGASUS_LOMBOK
-\lstinputlisting[linerange=1-5]{./src.txt}
+\lstinputlisting[linerange=1-5]{./import_src/src1.txt}
 }
 
+\vspace{-0.5em}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{Object interfaces with state (immutable data)}{name=results,column=1,span=2,row=0,below=Model}{
+\headerbox{Object Interfaces with State (immutable data)}{name=results,column=1,span=2,row=0,below=Model}{
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\lstinputlisting[linerange=7-11]{./src.txt}
-\lstinputlisting[linerange=13-15]{./src.txt}
-\vspace{0.1em}
+\lstinputlisting[linerange=7-11]{./import_src/src1.txt}
+\lstinputlisting[linerange=13-15]{./import_src/src1.txt}
 }
 
+\vspace{-0.5em}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\headerbox{More in the paper}{name=more,column=1,span=2,row=0,below=results}{
+\headerbox{with- Methods}{name=with,column=1,span=2,row=0,below=results}{
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\lstinputlisting[linerange=1-5]{./import_src/src2.txt}
+}
+
+\vspace{-0.5em}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\headerbox{Mutable Data \& Field Type Refinement}{name=refine,column=1,span=2,row=0,below=with}{
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\lstinputlisting[linerange=7-10]{./import_src/src2.txt}
+}
+\vspace{-0.5em}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\headerbox{More in the paper}{name=more,column=1,span=2,row=0,below=refine}{
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \item Case studies and applications.
-    \small{
-    \begin{itemize}\compresslist
-    \item The Expression Problem
-    \item Embedded DSLs with Fluent Interfaces
-    \item A Maze Game
-    \item Refactoring an Interpreter
-    \end{itemize}
+    \small{\newline
+    % \begin{itemize}\compresslist
+    % \item The Expression Problem
+    % \item Embedded DSLs with Fluent Interfaces
+    % \item A Maze Game
+    % \item Refactoring an Interpreter
+    % \end{itemize}
+    [The Expression Problem]\newline
+    [Embedded DSLs with Fluent Interfaces]\newline
+    [A Maze Game]\newline
+    [Refactoring an Interpreter] 
     }
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Results(Partial)}{name=result2,column=1,below=more}{
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   \headerbox{Results(Partial)}{name=result2,column=1, span=2, below=more}{
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 2nd case study, embedding DSLs with our implementation of 
+% fluent interfaces is quite straightforward:
+% \lstinputlisting[linerange=17-18]{./import_src/src1.txt}
+% }
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   \headerbox{Limitations}{name=limitations,column=1,below=result2,above=bottom}{
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{itemize}\compresslist
+% \item Current implementation only realizes ejc version.
+% \item Generics is not fully supported.
+% \item Separate compilation is an experimental feature with current Lombok.
+% \end{itemize}
+% \vspace{0.3em}
+%   }
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the second case study, we found that embedding DSLs with our implementation of 
-fluent interfaces is quite straightforward:
-\lstinputlisting[linerange=17-19]{./src.txt}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Limitations}{name=limitations,column=1,below=result2,above=bottom}{
+  \headerbox{Acknowledgments}{name=acknowledge,column=1, span=2, below=more}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}\compresslist
-\item Current implementation only realizes ejc version.
-\item Generics is not fully supported.
-\item Separate compilation is an experimental feature with current Lombok.
-\end{itemize}
-\vspace{0.3em}
+This work is sponsored by the Hong Kong Research Grant Council Early Career 
+Scheme project number 27200514. The first and second authors are supported by 
+the SIG PAC student travel grant.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Acknowledgments}{name=acknowledge,column=2,below=more}{
+  \headerbox{Contacts}{name=acknowledge,column=1,span=2,below=acknowledge,above=bottom}{
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-This work is sponsored by the Hong Kong Research Grant Council Early Career 
-Scheme project number 27200514. The first and second authors are supported by 
-the SIG PAC stduent travel grant.
-   \vspace{0.3em}
-  }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Contacts}{name=acknowledge,column=2,below=acknowledge,above=bottom}{
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\noindent{\textbf{The University of Hong Kong}} \\
+\noindent \textbf{The University of Hong Kong} \\
 Yanlin Wang, Haoyuan Zhang, Bruno C. d. S. Oliveira: \emph{\{ylwang,hyzhang,bruno\}@cs.hku.hk} \\
+\newline
 \textbf{Victoria University of Wellington} \\

### 2017-04-28 FACS/Examples.tex
Comments/changes

@@
 In order to illustrate how we plan to properly divide code reuse/inheritance/subclassing from subtyping,
-lets consider a simple language where trait names start lowercase \mt
-and class names start uppercase \mC.  
+lets consider a simple language where trait names start with a lowercase \mt
+and class names start with an uppercase \mC.  
 The syntax of the language is:
 
+\begin{comment}
 \begin{bnf}
@@
 \end{bnf}
+\end{comment}
 
@@
 by taking the union of the methods and the union of the implementations.
-In our simple mode, we consider an error trying to merge a class and an interface.
+In our simple mode\bruno{you mean model?}, we consider an error trying to merge a class and an interface.
 
@@
 present in different traits.  An exception is done for abstract methods:
-methods where the implementation \me\ is missing; In this case (if the
+methods where the implementation \me\ is missing. In this case (if the
 headers are compatible) the implemented version is selected.  In a sum
@@
 \use.\marco{actually, this makes the sum not strongly associative...}
+\bruno{ok: would be nice to see an example?}
 
@@
 interface, or simply interface a class declaration of for \Q@C:{interface  ...}@.
-  Then we have a set of implemented interfaces and a set of member declarations; in this simple language, just methods.
+  Then we have a set of implemented interfaces and a set of member
+  declarations. In this simple language, the only members are methods.
 If there are no implemented interfaces, in the concrete syntax we will omit the \Q@implements@ keyword.
 
-Methods \mMD can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very usefull in the context of code composition.
+Methods \mMD~can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very usefull in the context of code composition.
 To denote a method as abstract, instead of an optional keyword we just omit the implementation \me.
@@
 
-Basic well formedness rules applies:
+\paragraph{Well-formedness}
+Basic well formedness rules apply:
 \begin{itemize}
@@
 \end{itemize}
-Those rules can be applied on any given \mL individually and in full isolation.
+Those rules can be applied on any given \mL~individually and in full isolation.
 
-We expect the type system to enforce 
+We expect the type system to enforce: 
 \begin{itemize}
@@
 \item only interfaces can be implemented.
+\item \bruno{method conflict detection in traits?}
 \item \marco{I'm sure I'm missing something}
@@
 
-For the sake of simplicity, method bodies are just simple expressions \me, whose can be just variables and method calls.
-We do not think that there is any problem to extend our language with other constructs; we use minimal expressions to keep our example simple and compact.
-In order to make our examples more engaging, we feel free to use the type int, numeric literals and their operations.
+For the sake of simplicity, method bodies are just simple expressions
+\me: they can be just variables and method calls.
+We do not think that there is any problem to extend our language with
+other constructs~\bruno{don't you actually have evidence that this is
+  the case from 42?}. We use minimal expressions to keep our example
+simple and compact. In order to make our examples more engaging, we
+feel free to use the type \lstinline{int}, numeric literals and their operations.
 
@@
 In a sense, all classes are final (as in Java).
-Moreover  \use\ can not contain class names.
+Moreover  \use\ cannot contain class names.
 That is, using a trait is the only way to induce code reuse.
@@
 This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. It is also possible to add a \Q@B@ as follow
+is implementing it. It is also possible to add a \Q@B@ as follows
 \begin{lstlisting}
@@
 Our typing discipline is 
-what distinguish our approach from a simple minded code composition macro~\cite{bawden1999quasiquotation}
+what distinguishes our approach from a simple minded code composition macro~\cite{bawden1999quasiquotation}
 or a rigid module composition~\cite{ancona2002calculus}
@@
 
-\noindent Note how in the former example \Q@ta@ is well typed just by knowing itself and \Q@IA@.
+\noindent Note how in the former example \Q@ta@ is well typed just by
+knowing itself and \Q@IA@.\bruno{why IA? it is not used in this
+  example, right?}
 
@@
 an \mL\ in $\overline\mV$ is not typechecked before flattening, and only the result is expected to be well-typed.
-While this seams a very dangerous approach at first, consider that also Java have the same behaviour:
+While this seems a very dangerous approach at first, consider that also Java have the same behaviour:
 for example in
@@
 literals have to be self complete. In this case we would have been forced to
-declare abstract methods \Q@n@ and \Q@m@.
+declare abstract methods \Q@n@ and \Q@m@.\bruno{right. this is a good point}
 
-Our typing strategy have two important properties:
+Our typing strategy has two important properties:
 \begin{itemize}
@@
 It may be that the result is intrinsically ill-typed, if of the methods in $\mL_1\ldots\mL_n$ is not well typed,
-but it may also happens that a type referred from one of those methods is declared \emph{after} the current class. As we will see later, this is how our relaxation allows to support recursive types.
+but it may also happen that a type referred from one of those methods is declared \emph{after} the current class. As we will see later, this is how our relaxation allows to support recursive types.
 

### 2017-04-28 FACS/Intro.tex
Comments/changes

@@
 
-Historically, we have seen a lot of focus on the importance of
+Historically, there has been a lot of focus on the importance of
 separing subtype from subclassing~\cite{cook}.  This is claimed to be
@@
 in subtyping without subclassing, in most OO languages code reuse
-(inheritance/extends) imply subtyping.  Consider the following (Java)
-code:
+(inheritance/extends) implies subtyping in a fundamental way. 
+Consider the following (Java) code:
 
@@
 
+In class \lstinline{A}, the method \lstinline{m} passes \lstinline{this} as \lstinline{A}.
 This code seems correct, and there is no subtyping/subclassing.
@@
 
-Class \lstinline{B} now have a method \lstinline{ma} inherited from \lstinline{A}.
-This method passes \lstinline{this} as \lstinline{A}; we can see an invocation of \lstinline{ma}
-inside \lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.
-The execution will eventually call \lstinline{Util.m} with a \lstinline{B} instance.
-This can be correct only if \lstinline{B} is a subtype of \lstinline{A}.
-Thus, every OO language with the minimal features 
-exposed in this example (using this, extends, method call)
-is forced to accept that subclassing implies subtyping.\footnote{
-Note how this is true also in C++, where is possible to
-"extends privately". Such is a limitation of the visibility of
-subtyping but not over subtyping itself, and the former example
-would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.}
+Class \lstinline{B} inherits the method \lstinline{ma} from
+\lstinline{A}.  We can see an invocation of \lstinline{ma} inside
+\lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.  The
+execution will eventually call \lstinline{Util.m} with a \lstinline{B}
+instance. This can be correct only if \lstinline{B} is a subtype of
+\lstinline{A}. Thus, every OO language with the minimal features
+exposed in this example (using \lstinline{this}, extends and method calls) is forced
+to accept that subclassing implies subtyping.\footnote{ Note how this
+  is true also in C++, where is possible to "extends privately". Such
+  is a limitation of the visibility of subtyping but not over
+  subtyping itself, and the former example would be accepted by C++
+  even if \lstinline{B} was to "privately extends" \lstinline{A}.}
 \marco{Cite some work of bruceTHisType and show how he also fails.}
@@
   must not imply subtyping to type-check.}
-An iconic example on why connecting inheritance/code reuse and subtpying is problematic is
-provided by the historic\cite{LaLonde:1991:SSS:110673.110679}:
-A reasonable implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@ keeping tracks of how many times an element occurs.
-We would just add some state and override a couple of methods.
+
+An iconic example on why connecting inheritance/code reuse and
+subtpying is problematic is provided by the
+historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
+implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
+keeping tracks of how many times an element occurs.  We would just add
+some state and override a couple of methods.\bruno{Are we going to
+  present this example in the paper solved in 42, for example? 
+I think I would expect to see it.}
 
@@
 and making both \Q@Bag@ and \Q@Set@ inherit from it.
-This looks innatural, since \Q@Set@ would extend it without adding anything,
+This looks unnatural, since \Q@Set@ would extend it without adding anything,
 and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
@@
 
-We will show here a simple design to completley uncouple subtyping
-and subclassing in a nominally typed OO language,
-by dividing code designed for \textbf{USE}
-from code designed for \textbf{REUSE}.\footnote{We need to talk of unanticipated extensions?}
-First we show a minimal language, then we show how mutually recursive types
+This paper aims at showing a simple language design to completley uncouple subtyping
+and subclassing in a nominally typed OO language. The key idea is to
+divide between code designed for \textbf{USE}
+and code designed for \textbf{REUSE}.\marco{We need to talk of unanticipated extensions?}
+First we show a minimal language that illustrates the basic ideas. 
+Then we show how mutually recursive types
 are supported, how state/constructors/fields are supported,
 how we can extend the language with nested classes.
-Finally we show 42, a full blown language build around our ideas of reuse.
+Finally we show 42, a full blown language build around our ideas of
+reuse.
 
 Our design leverage on traits~\cite{ducasse2006traits}: a well know mechanisms for pure
-code reuse
+code reuse.
 
@@
 \item simple approach to fully separate inheritance and subtyping
-\item easy to grasp model of compilation were typing is late but not too late (lazy?)
+\item easy to grasp model of compilation were typing is late but not
+  too late (lazy?)\bruno{this idea is important: it's a key design
+    decision. We need to talk about it in the introduction and offer
+    some motivation/justification. }
 \item clean and elegant handling of state in trat/module composition language
-\item extension with nested class is natural and powefull
+\item extension with nested classes is natural and
+  powefull\bruno{wondering if this is necessary. I think there are
+    several interesting/novel ideas already. Covering too much will
+    result on not saying enough about each individual ideas.}
 \item more operators can be accomodated without changing the general model

### 2017-05-04 FACS/Intro.tex
Intro

@@
 
+In mainstream OO languages like Java, C++ or C\# subclassing 
+implies subtyping. For example, in Java a subclass definition, such as:
+
+\begin{lstlisting}
+class A extends B {}
+\end{lstlisting}
+
+\noindent does two things at the same time:
+1) {\bf inheriting} code from class \lstinline{B}; and {\bf creating 
+a subtype} of \emph{B}. Therefore in a language like Java 
+a subclass is \emph{always} a subtype of the extended class.
+
 Historically, there has been a lot of focus on the importance of
-separing subtype from subclassing~\cite{cook}.  This is claimed to be
-good for code reuse, design and reasoning.  While there is no problem
+separing subtyping from subclassing~\cite{cook}.  This is claimed to be
+good for code reuse, design and reasoning. 
+While there is no problem
 in subtyping without subclassing, in most OO languages code reuse
-(inheritance/extends) implies subtyping in a fundamental way. 
-Consider the following (Java) code:
+(inheritance/extends) implies subtyping in a fundamental way because 
+of what we call the \emph{self-reference problem}. The self-reference
+problem is illustrated by the following (Java) code:
 
@@
 
-Class \lstinline{B} inherits the method \lstinline{ma} from
-\lstinline{A}.  We can see an invocation of \lstinline{ma} inside
-\lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.  The
-execution will eventually call \lstinline{Util.m} with a \lstinline{B}
-instance. This can be correct only if \lstinline{B} is a subtype of
-\lstinline{A}. Thus, every OO language with the minimal features
-exposed in this example (using \lstinline{this}, extends and method calls) is forced
-to accept that subclassing implies subtyping.\footnote{ Note how this
-  is true also in C++, where is possible to "extends privately". Such
-  is a limitation of the visibility of subtyping but not over
-  subtyping itself, and the former example would be accepted by C++
-  even if \lstinline{B} was to "privately extends" \lstinline{A}.}
+%%Class \lstinline{B} does two things at the same time: 
+%%1) it {\bf inherits} the method \lstinline{ma} from
+%%\lstinline{A}; and 2) it creates a {\bf subtype} of \lstinline{A}.
+We can see an invocation of \lstinline{ma} inside
+\lstinline{mb}, where \lstinline{this} is of type \lstinline{B}. 
+The execution will eventually call \lstinline{Util.m} with an
+instance of \lstinline{B}. This can be correct only if \lstinline{B} is a subtype of
+\lstinline{A}.
+
+If Java were to support a mechanism to allow reuse/inheritance 
+without introducing subtyping, such as:
+
+\begin{lstlisting}[language=Java]
+  class B inherits A{ int mb(){return this.ma();} }
+\end{lstlisting}
+
+\noindent Then an invocation of 
+\lstinline{mb} would be type-unsafe (i.e. it would 
+result in a run-time type error). 
+Note that here the intention of using the imaginary keyword {\bf
+  inherits} is to allow the code from \lstinline{A} to be inherited 
+without \lstinline{B} becoming a subtype of \lstinline{A}. 
+However this breaks type-safety. The problem is that the
+self-reference \lstinline{this} in class \lstinline{B} has 
+type \lstinline{B}. Thus, when \lstinline{this} is passed as an argument to 
+the method \lstinline{Utils.m} as a result of the invocation of
+\lstinline{mb}, it will have a type that is incompatible with the
+expected argument of type \lstinline{A}.
+
+What the self-reference problem illustrates is that adopting a more flexible
+OO model where subclassing does not imply subtyping is not trivial. 
+If we are to separate subclassing from subtyping then a more substantial
+change in the language design is necessary.  Every OO language with
+the minimal features exposed in the example (using \lstinline{this},
+extends and method calls) is forced to accept that subclassing implies
+subtyping.\footnote{ Note how this is true also in C++, where is
+  possible to "extends privately". Such is a limitation of the
+  visibility of subtyping but not over subtyping itself, and the
+  former example would be accepted by C++ even if \lstinline{B} was to
+  "privately extends" \lstinline{A}.}
+
 \marco{Cite some work of bruceTHisType and show how he also fails.}
-\bruno{I was expecting some text/argument here on why subclassing
-  implies subtyping is bad. Perhaps another example where subclassing
-  must not imply subtyping to type-check.}
 
+\begin{comment}
 An iconic example on why connecting inheritance/code reuse and
@@
 public interface of our classes and our code project as a whole.
+\end{comment}
 

### 2017-05-04 FACS/Intro.tex
Intro

@@
   "privately extends" \lstinline{A}.}
+In essence we believe that a problem with the design of classes in languages like Java
+is that classes do too many things at once. In particular they act
+both as units of \emph{use} and \emph{reuse}.
+That is, in languages like Java, classes can be
+instantiated (using {\bf new}) to create objects that \emph{use} some 
+functionality. Classes can also be subclassed to provide \emph{reuse}
+of code.
 
@@
 
-This paper aims at showing a simple language design to completley uncouple subtyping
-and subclassing in a nominally typed OO language. The key idea is to
-divide between code designed for \textbf{USE}
-and code designed for \textbf{REUSE}.\marco{We need to talk of unanticipated extensions?}
+This paper aims at showing a simple language design, called \name, to
+completley uncouples subtyping and subclassing in a nominally typed OO
+language. The key idea is to divide between code designed for
+\textbf{USE} and code designed for \textbf{REUSE}. 
+In \name there are two separate concepts: classes
+and traits~\cite{}. Classes are meant for code use, and cannot be used
+for reuse. In some sense classes in \name are like final classes in
+Java. Traits are used for code reuse and multiple traits can be
+composed to form a class which can then be instantiated. 
+\marco{We need to talk of unanticipated extensions?}
+\bruno{Talk more about typing aspects here. Summarize the important 
+aspects of the design of \name.}
+
 First we show a minimal language that illustrates the basic ideas. 
@@
 \end{itemize}
+
+\section{Overview}
+
+This section presents the overview of \name. In name

### 2017-05-04 FACS/Intro.tex
Section 2

@@
 (inheritance/extends) implies subtyping in a fundamental way because 
-of what we call the \emph{self-reference problem}. The self-reference
+of what we call the \emph{self-reference subclassing problem}. This 
 problem is illustrated by the following (Java) code:
@@
 We can see an invocation of \lstinline{ma} inside
-\lstinline{mb}, where \lstinline{this} is of type \lstinline{B}. 
+\lstinline{mb}, where the self-reference \lstinline{this} is of type \lstinline{B}. 
 The execution will eventually call \lstinline{Util.m} with an
-instance of \lstinline{B}. This can be correct only if \lstinline{B} is a subtype of
+instance of \lstinline{B}. However, this can be correct only if \lstinline{B} is a subtype of
 \lstinline{A}.
 
-If Java were to support a mechanism to allow reuse/inheritance 
+If Java was to support a mechanism to allow reuse/inheritance 
 without introducing subtyping, such as:
@@
 
-What the self-reference problem illustrates is that adopting a more flexible
+What the self-reference subclassing problem illustrates is that
+adopting a more flexible
+nominally typed
 OO model where subclassing does not imply subtyping is not trivial. 
@@
 This paper aims at showing a simple language design, called \name, to
-completley uncouples subtyping and subclassing in a nominally typed OO
+completely decouple subtyping and subclassing in a nominally typed OO
 language. The key idea is to divide between code designed for
@@
 for reuse. In some sense classes in \name are like final classes in
-Java. Traits are used for code reuse and multiple traits can be
-composed to form a class which can then be instantiated. 
+Java. Traits are meant for code reuse and multiple traits can be
+composed to form a class which can then be instantiated. Traits 
+cannot be instantiated (or used) directly. Such design allows the
+subtyping and code reuse to be treated separately, which in terms 
+brings several benefits in terms of flexibility and code reuse.
 \marco{We need to talk of unanticipated extensions?}
@@
 
-This section presents the overview of \name. In name
+This section presents the overview of \name, and it illustrates the
+key ideas of its design. In particular we illustrate how to separate code use and 
+code reuse in \name, and how this allows improving both.
+
+\subsection{The Design of \name: Decoupling Use from Reuse}
+
+\paragraph{Classes in \name: A mechanism for code Use}
+\name has a substantially different design from Java-like languages.
+The concept of a class in \name provides a mechanism for code use
+only. This means that in \name there is actually no subclassing, and
+and classes are roughly equivalent to final classes in Java.  Thus,
+compared to Java-like languages, the most noticeable difference is the
+absence of the \Q@extends@ keyword in \name.
+
+To illustrate classes in \name consider the example in Section~\ref{}:
+
+\begin{lstlisting}
+ A: { method int ma() Utils.m(this) }//note, no {return _}
+ Utils:{ class method int m(A a)/*method body here*/ }
+\end{lstlisting}
+
+\noindent Classes in \name use a different declaration style compared
+to Java. Class declarations have a name (which must always start with
+an uppercase letter) and a code literal, which is used to specify the
+definitions of the class. For instance, in the class declaration
+for \lstinline{A}, the name of the class is \lstinline{A} and the code 
+literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
+class. There are important differences to Java-like languages in the way 
+classes and code-literals are type-checked, as we shall see next. 
+However for this example, things still work in a similar way to Java.
+
+The \name code above is fine, but there is no way to add a \Q@B@ reusing code of \Q@A@, since
+\Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
+the Java code, there is no way to introduce a subclass
+\lstinline{B}. At first, this may seem like a severe restriction, but
+\name has a different mechanism for code \emph{reuse} that 
+is more appropriate when code reuse is intended.
+
+\paragraph{Traits in \name: A mechanim of code Reuse}
+Lets try again, but this time aiming at code reuse. Name 
+has a form of traits for code reuse. Trait declarations 
+look very much like class declarations, but trait names 
+start with a lowercase letter. An obvious first attempt 
+to model the example in Section~\ref{} (for a Java programmer) 
+with traits is:
+
+\begin{lstlisting}
+ ta:{ method int ma() Utils.m(this) }//type error
+ A:Use ta
+ Utils:{ class method int m(A a)/*method body here*/ }
+\end{lstlisting}
+
+\noindent Here \lstinline{ta} is a trait intended to replace the
+original class \lstinline{A} so that the code of the method
+\lstinline{ma} can be reused. Then the class \lstinline{A} 
+is created by inheriting the code from the trait using the keyword 
+{\bf Use}. Note that \use\ cannot contain class names: only trait
+names are allowed.
+That is, using a trait is the only way to induce code reuse.
+Unfortunatelly, this code does not work, 
+because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
+\Q@ta@ has no knowlegde of \lstinline{A}.
+%: in \name all declarations 
+%are sequentially type-checked and if a definition occurs afterwards
+%another, then it is not visible on all previous definitions.
+
+\paragraph{Type of the self-reference} The code above does not work
+because the way type-checking works in \name is that type-checking of 
+code literals is independent of the class/trait names associated to
+it. A Java programmer may expect that the type of \lstinline{this} 
+in the previous definition of \lstinline{ta} is \lstinline{ta}
+itself. However, this intuition brought from Java is wrong in \name
+for two reasons:
+
+\begin{itemize}
+
+\item The first reason is that {\bf traits are not types}. Traits in
+  \name are simply units of reuse and cannot be used as types. Types
+  are only introduced by class/interface declarations.
+
+\item The second reason is that {\bf the type of self-references is the structural type of the
+    code literal.} Unlike Java-like languages, where the body of a
+  class declaration is intrinsically coupled with the class itself, in
+  \name code literals are first-class\bruno{Is it appropriate to say
+    that they are first class?} and are type-checked independently
+  using a form of structural typing.\bruno{Say how we recover nominality?} 
+\end{itemize}
+
+With this in mind, we can try to model the example in Section~\ref{} again:
+
+\begin{lstlisting}
+ IA:{interface method int ma()}//interface with abstract method
+ ta:{implements IA
+   method int ma() Utils.m(this) }
+ A:Use ta
+ Utils:{ class method int m(IA a)/*method body here*/}
+\end{lstlisting}
+
+This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
+is implementing it. It is also possible to add a \Q@B@ as follows
+\begin{lstlisting}
+  B:Use ta, { method int mb(){return this.ma();} }
+\end{lstlisting}
+This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
+Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.
+
+\paragraph*{Semantic of Use}
+Abeit alternative semantic models for traits~\cite{} have been proposed,
+here we use the flattening model.
+This means that 
+\begin{lstlisting}
+A:Use ta
+B:Use ta, { method int mb(){return this.ma();} }
+\end{lstlisting}
+
+\noindent reduces/is equivalent to/is flatted into
+
+ \begin{lstlisting}
+A:{implements IA method int ma() Utils.m(this) }
+B:{implements IA
+  method int ma() Utils.m(this)
+  method int mb() this.ma() }
+
+ \end{lstlisting}
+
+This code seems correct, and there is no mention of the trait \Q@ta@.
+In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.
+
+\subsection{Recursive Types}

### 2017-05-04 FACS/Intro.tex
Added note

@@
 (inheritance/extends) implies subtyping in a fundamental way because 
-of what we call the \emph{self-reference subclassing problem}. This 
+of what we call the self-reference \emph{leaking problem}. This 
 problem is illustrated by the following (Java) code:
@@
 expected argument of type \lstinline{A}.  
+\bruno{
+We can also mention what happens in structural types:
+"System with structural types would *be required to* guarantee that
+the structural type of A is a supertype of the structural type of B”.
+}
 
-What the self-reference subclassing problem illustrates is that
+What the leaking problem illustrates is that
 adopting a more flexible

### 2017-05-08 FACS/Intro.tex
Intro

@@
 separing subtyping from subclassing~\cite{cook}.  This is claimed to be
-good for code reuse, design and reasoning. 
-While there is no problem
-in subtyping without subclassing, in most OO languages code reuse
-(inheritance/extends) implies subtyping in a fundamental way because 
-of what we call the self-reference \emph{leaking problem}. This 
-problem is illustrated by the following (Java) code:
+good for code reuse, design and reasoning. There are at
+least two distinct situations where the separation of subtyping and 
+subclassing is helpful.
+
+\begin{itemize}
+
+\item {\bf Allowing reuse even when subtyping is impossible:} 
+In some situations a subclass contains methods whose signatures 
+are incompatible with the superclass, yet inheritance is still
+possible. A typical example, which was illustrated by Cook et al., are 
+classes with \emph{binary methods}~\cite{}.
+
+\item {\bf Preventing unintended subtyping:} For certain classes we
+  would like to inherit code without creating a subtype even if, from
+  the typing point of view, subtyping is still possible. A typical
+  example~\cite{} of this are methods for collection classes such as Sets and
+  Bags. Bag implementations can often inherit 
+  from Set implementations, and the interfaces of the two collection types are
+  similar and type compatible. 
+  However, from the logical point-of-view a Bag is \emph{not a
+    subtype} of a Set.
+
+\end{itemize}
+
+Type systems based on structural typing can deal with the first
+situation well, but not the second. Since structural subtyping
+accounts for the types of the methods only, a Bag would be a subtype
+of a Set if the two interfaces are type compatible. For dealing with
+the second situation nominal subtyping is preferable. With nominal
+subtyping an explicit subtyping relation must be signalled by the
+programmer. Thus if subtyping is not desired, the idea is that 
+programmer can simply not declare a subtyping relationship.
+
+While there is no problem in subtyping without subclassing, in the design
+of most nominal OO languages subtyping implies subtyping in a
+fundamental way. This is because of what we call the self-reference
+\emph{leaking problem}. The leaking problem is illustrated by the following
+(Java) code:
 
@@
 \paragraph{Classes in \name: A mechanism for code Use}
-\name has a substantially different design from Java-like languages.
+%\name has a substantially different design from Java-like languages.
 The concept of a class in \name provides a mechanism for code use
@@
 
-\noindent Classes in \name use a different declaration style compared
-to Java. Class declarations have a name (which must always start with
+\noindent Classes in \name use a slightly different declaration style compared
+to Java. In \name there is no \lstinline{class} keyword. 
+Class declarations have a name (which must always start with
 an uppercase letter) and a code literal, which is used to specify the
@@
 literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
-class. There are important differences to Java-like languages in the way 
+class. There are, however, some 
+important differences to Java-like languages in the way 
 classes and code-literals are type-checked, as we shall see next. 
-However for this example, things still work in a similar way to Java.
+Nevertheless, for this example, things still work in a similar way to Java.
 
@@
 \Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
-the Java code, there is no way to introduce a subclass
-\lstinline{B}. At first, this may seem like a severe restriction, but
+the Java code, introducing a subclass
+\lstinline{B} is not possible. At first, this may seem like a severe restriction, but
 \name has a different mechanism for code \emph{reuse} that 
@@
 \Q@ta@ has no knowlegde of \lstinline{A}.
+\bruno{Can we summarize in a table what are the differences between
+  traits and classes? For example: 
+traits: do not introduce types; units of reuse; cannot be instantiated
+classes: introduce types; can be instantiated. 
+Also it would be good to motivate those differences more from the 
+technical point of view. For example: the fact that traits do not
+introduce types is it mostly a philosophical difference, or does it 
+actually buy you something from the technical aspect.
+}
 %: in \name all declarations 
@@
 
-\item The second reason is that {\bf the type of self-references is the structural type of the
+\item The second reason is that {\bf the type of self-references is the self type of the
     code literal.} Unlike Java-like languages, where the body of a
@@
   \name code literals are first-class\bruno{Is it appropriate to say
-    that they are first class?} and are type-checked independently
-  using a form of structural typing.\bruno{Say how we recover nominality?} 
+    that they are first class?} and are type-checked independently.
+  \name has a notion of self-type, which is closely related to
+  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
+  in \name the self-reference does not have the type of the class
+  being defined, but rather it has the self-type.
 \end{itemize}
@@
 This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. It is also possible to add a \Q@B@ as follows
+is implementing it. Thus the self-type of the trait \Q@ta@ will be a
+subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
 \begin{lstlisting}
@@
 \paragraph*{Semantic of Use}
-Abeit alternative semantic models for traits~\cite{} have been proposed,
+Albeit alternative semantic models for traits~\cite{} have been proposed,
 here we use the flattening model.
@@
 
-This code seems correct, and there is no mention of the trait \Q@ta@.
-In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.
+ This code seems correct, and there is no mention of the trait
+ \Q@ta@. In some sense, all the information about code
+ reuse/subclassing is just a private implementation detail of \Q@A@
+ and \Q@B@; while subtyping is part of the class interface.
 
-\subsection{Recursive Types}
+\subsection{Self Types, Binary Methods and the Need to Separate
+  Inheritance from Subtyping}
+
+So, far it appears that we have not gained much from decoupling
+reuse from subtyping.
+
+As discussed in Section~\ref{} the type of the self-reference is the
+self-type. Like languages with ThisType, \name allows explicit
+references to the self type. This is useful in several situations: 
+For example it can be helpful in the presence of binary
+methods~\cite{}, or methods that return an instance of the 
+class being defined.

### 2017-05-08 FACS/Intro.tex
Overview

@@
 
-\item {\bf Allowing reuse even when subtyping is impossible:} 
+\item {\bf Allowing inheritance/reuse even when subtyping is impossible:} 
 In some situations a subclass contains methods whose signatures 
@@
 
+\name solution for the leaking problem shows
+
+\subsection{Preventing Unintended Subtyping}
+As stated in Section~\ref{} a benefit of a nominally typed OO language
+that separates inheritance from subtyping is that it is possible to 
+have reuse, while preventing unintended subtyping.
+
+An iconic example on why connecting inheritance/code reuse and
+subtpying is problematic is provided by the
+historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
+implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
+keeping tracks of how many times an element occurs.  We would just add
+some state and override a couple of methods.\bruno{Are we going to
+  present this example in the paper solved in 42, for example? 
+I think I would expect to see it.}
+
+However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
+Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
+and making both \Q@Bag@ and \Q@Set@ inherit from it.
+This looks unnatural, since \Q@Set@ would extend it without adding anything,
+and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
+Worst, if we was to constantly apply this mentalty, we would introduce a very high number
+of abstract classes that are not supposed to be used as types, and that will clutter the 
+public interface of our classes and our code project as a whole.
+
+\begin{lstlisting}
+class Set {.../* Elem put() boolean isIn(Elem)*/ ...}
+class Bag extends Set{.../*override something to keep track of duplications*/..}
+\end{lstlisting}
+
+but now,
+
+\begin{lstlisting}
+Set mySet=new Bag(); //OK for the type system but not for LSP
+\end{lstlisting}
+
+as a pleasaruble accident, avoid such code gift us simple support for
+This type and (in the extensions with nested classes seen later)
+family polimporphism.
+
 \subsection{Self Types, Binary Methods and the Need to Separate

### 2017-05-11 FACS/Intro.tex
Some more work on Sections 2 3 and 4

@@
 
-\section{Overview}
+\section{The Design of \name: Separating Use and Reuse}
 
@@
 key ideas of its design. In particular we illustrate how to separate code use and 
-code reuse in \name, and how this allows improving both.
+code reuse in \name. %, and how this allows improving both.
 
-\subsection{The Design of \name: Decoupling Use from Reuse}
+%\subsection{The Design of \name: Decoupling Use from Reuse}
 
-\paragraph{Classes in \name: A mechanism for code Use}
+\subsection{Classes in \name: A mechanism for code Use}
 %\name has a substantially different design from Java-like languages.
@@
 only. This means that in \name there is actually no subclassing, and
-and classes are roughly equivalent to final classes in Java.  Thus,
+classes are roughly equivalent to final classes in Java.  Thus,
 compared to Java-like languages, the most noticeable difference is the
@@
 
-The \name code above is fine, but there is no way to add a \Q@B@ reusing code of \Q@A@, since
+The \name code above is fine, but there is no way to add a class 
+\Q@B@ reusing the code of \Q@A@, since
 \Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
@@
 
-\paragraph{Traits in \name: A mechanim of code Reuse}
-Lets try again, but this time aiming at code reuse. Name 
-has a form of traits for code reuse. Trait declarations 
+\subsection{Traits in \name: A mechanim of code Reuse}
+
+Unlike classes traits in \name cannot be instantiated and do not introduce new
+types. However they provide code reuse.
+%So, lets try again encoding the code for the leaking problem, but this
+%time aiming at code reuse. 
+Trait declarations 
 look very much like class declarations, but trait names 
@@
 to model the example in Section~\ref{} (for a Java programmer) 
-with traits is:
+with traits and code reuse is:
 
@@
 Unfortunatelly, this code does not work, 
-because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
-\Q@ta@ has no knowlegde of \lstinline{A}.
-\bruno{Can we summarize in a table what are the differences between
-  traits and classes? For example: 
-traits: do not introduce types; units of reuse; cannot be instantiated
-classes: introduce types; can be instantiated. 
-Also it would be good to motivate those differences more from the 
-technical point of view. For example: the fact that traits do not
-introduce types is it mostly a philosophical difference, or does it 
-actually buy you something from the technical aspect.
-}
+because \lstinline{Utils.m} requires an \lstinline{A} and the type of \lstinline{this} in
+\Q@ta@ has no relationship to the type \lstinline{A}.
+Since the trait name is not a type, no code external to that trait can
+refer to it. This is one of the key design decisions in \name. 
+The point is that the trait can refer to the program via \lstinline{this}, but the
+program is agnostic to what the trait type is going to be, so it can
+be later assigned to any (or many) classes. This improves the
+flexibility of reuse as illustrated in Section~\ref{}. However, 
+to solve the more immediate typing issue above we need one more
+round of refactoring, as we shall see next.
+
 %: in \name all declarations 
@@
  IA:{interface method int ma()}//interface with abstract method
+ Utils:{ class method int m(IA a)/*method body here*/}
  ta:{implements IA
@@
  A:Use ta
- Utils:{ class method int m(IA a)/*method body here*/}
 \end{lstlisting}
@@
 This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. Thus the self-type of the trait \Q@ta@ will be a
+is implementing it. Thus the self-type of the trait \Q@ta@ (and the
+type of \lstinline{this}) will be a
 subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
@@
 
-\name solution for the leaking problem shows
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{c|c|c|c}
+              & Instantiable (Use) & Unit of Reuse & Introduces New Type \\
+\hline
+Class  & Yes              & No          & Yes           \\
+Trait   & No               & Yes         & No            \\
+\end{tabular}
+\end{center}
+\caption{A comparision between traits and classes in \name.}
+\label{fig:compare}
+\end{figure}
 
-\subsection{Preventing Unintended Subtyping}
-As stated in Section~\ref{} a benefit of a nominally typed OO language
+To finish this section, Figure \ref{fig:compare} provides a summary of
+the differences between classes and traits. The comparison focus on
+the roles of traits and classes with respect to instantiation,
+reusability and whether the declarations also introduce new types or
+not.
+
+\section{Improving Use by Preventing Unintended Subtyping}
+%%Preventing Unintended Subtyping}
+As stated in Section~\ref{} a possible benefit of a nominally typed OO language
 that separates inheritance from subtyping is that it is possible to 
-have reuse, while preventing unintended subtyping.
+have reuse, while preventing unintended subtyping. The design of \name 
+allows this, thus it \emph{improves the use} of classes, while
+retaining the benefits of reuse when compared to Java-like languages. 
+Here we illustrate this benefit by modelling a simplified version of
+Set and Bag collections.
+
+\bruno{Plan: Show simplified Java code. Show how the Java code forces
+  us to make Bags a subtype of Set, if we want reuse. 
+Then show the 42/language in this paper/ code and solution.  
+}
 
@@
 
-\subsection{Self Types, Binary Methods and the Need to Separate
+\section{Improving Reuse when Subtyping is Impossible}
+
+\name retains the ability to allow reuse even when subtyping is
+impossible. That is, in some situations a subclass contains methods
+whose signatures are incompatible with the superclass, yet inheritance
+is still possible. This type of situations is the primary motivator
+for previous work aiming at separating inheritance from subtyping, and
+it happens, for example, in classes with \emph{binary
+  methods}~\cite{}.  Despite the adoption of a nominal approach and
+unlike other nominally typed languages, the leaking problem does not
+prevent \name from enjoying such kind of improved reuse.
+
+To illustrate how \name allows for improved reuse, we consider a
+pattern to combine \emph{multiple trait inheritance} and \emph{state}:
+a challenging problem that has limited the flexibility of traits and
+reuse in the past.
+
+\subsection{Managing State}
+
+This idea of summing pieces of code is very elegant, and has proven
+very successfull in module composition
+languages~\cite{ancona2002calculus}\bruno{and several trait models?}. 
+However our research community is
+struggling to make it work with object state (constructor and fields)
+while acheving the following goals:
+
+\begin{itemize}
+\item keep sum associative and commutative,
+\item allowing a class to create instances of itself,
+\item actually initialize objects, leaving no null fields,
+\item managing fields in a way that borrows the elegance of summing methods,
+\item make it easy to add new fields.
+\end{itemize}
+
+In the related work we will show some alternative ways to handle state.
+However the purest solution just requires methods:
+  The idea is that
+  the trait code just uses getter/setters/factories, while leaving
+  to classes the role to finally define the fields/constructors.
+  That is, the  the class has syntax richer that the trait one,
+  allowing declaration for fields and constructors.
+  This approach is very powefull~\cite{wang2016classless}
+
+\paragraph*{Advantages:} 
+this approach is associative and commutative, even self construction
+  can be allowed if the trait requires a static/class method
+  returning This; the class will then implement this method by forwarding
+  a call to the constructor.
+
+\paragraph*{Negatives:} writing the class code with the constructors and
+   fields and getter/setters and factories can be quite tedious.
+   Moreover, there is no way for a trait to specify a default value for a field,
+   the class need to handle all the state, even state that is conceptually
+   "private" of such trait.
+
+Here in the following an example of such approach:
+
+\begin{lstlisting}
+  pointSum: { method int x()  method int y()
+    class method This of(int x,int y)
+    method This sum(This other)
+      This.of(this.x+other.x,this.y+other.y)
+    }
+  pointMul: { method int x() method int y()
+    class method This of(int x,int y)
+    method This mul(This other)
+      This.of(this.x*other.x,this.y*other.y)
+    }
+\end{lstlisting}
+
+As you see, all the state operations are represented as abstract methods.
+
+\subsection{A First Attempt at Composition}
+According to the general ideas expressed before,
+\begin{lstlisting}
+  Point:Use pointSum,pointMul
+\end{lstlisting}
+
+\noindent would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract.
+In this mindset, the user would be required to write something similar to
+
+\begin{lstlisting}
+  CPoint:Use pointSum,pointMul, {//not our suggested solution
+    int x   int y
+    method int x()x   method int y()y
+    class method This of(int x, int y)
+      new Point(x,y)
+    constructor Point(int x, int y){ this.x=x   this.y=y }
+    }
+\end{lstlisting}
+
+\noindent after a while programming in this style, 
+writing those obvious ``close the state'' classes become a
+repetivie boring job, and one wonder
+if it could be possible to automatically generate such code~\cite{wang2016classless}.
+Indeed those classes are just a form of ``fixpoint''.
+
+In our model we go one step further: there is no need to generate code, or to explicitly
+write down constructors and fields; there is not even syntax for those
+constructor.
+The idea is that any class that ``could" be completed in the obvious way
+\emph{is a complete ``coherent" class}.
+In most other languages, a class is abstract if have abstract methods.
+Instead, we call abstract a class whose set of abstract methods is not
+coherent, that is, can not be automatically recognized as factory,getters and setters.
+
+\paragraph*{Detaild definition of coherent:}
+\begin{itemize}
+\item a class with no abstract method is coherent, and like Java \Q@Math@.
+Will just be usefull for calling class/static methods.
+\item a class with a single abstract \Q@class@ method returning \Q@This@
+is coherent if all the other abstract methods can be seen as \emph{abstract state
+operations} over one of its argument.
+For example,
+if there is a \Q@class method This of(int x, int y)@ as before,
+then
+\item a method \Q@int x()@ is intepreted as an abstract state method: a getter for \Q@x@.
+\item a method \Q@Void x(int that)@ is a setter for x.
+\end{itemize}
+
+While getters and setters are fundamental operations, we can immagine
+more operations to be supported; for example
+\begin{itemize}
+\item \Q@method This withX(int that)@
+may be a ``wither", doing a functional field update.
+\item \Q@method Void update(int x,int y)@
+may do two field update at a time.
+\item\Q@method This clone()@ may do a shallow clone of the object.
+\end{itemize}
+
+We are not sure what is the best set of abstract state operations yet, but we think this could become a very interesting area of research.
+
+  lets play with the points of before, to see what good can we do with the current
+  instruments:
+
+\begin{lstlisting}
+ //same code as before for pointSum and pointMul
+  pointSum: { method int x() method int y()
+    class method This of(int x,int y)
+    method This sum(This other)
+      This.of(this.x+other.x,this.y+other.y)
+    }
+  pointMul: { method int x() method int y()//look we are repeating
+    class method This of(int x,int y)//the abstract method declarations.
+    method This mul(This other)
+      This.of(this.x*other.x,this.y*other.y)
+    }
+  PointAlgebra:Sum pointSum,pointMul
+\end{lstlisting}
+
+  As you can see, we can declare the methods independently and compose the result
+  as we wish. However we have to repeat the abstract methods \Q@x@,\Q@y@ and \Q@of@.
+  In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
+  and not repeat such abstract definitions? of course!
+
+\begin{lstlisting}
+  p: { method int x() method int y()
+    class method This of(int x,int y)
+    }
+  pointSum:Use p, { method This sum(This other)
+      This.of(this.x+other.x,this.y+other.y)
+    }
+  pointMul:Use p, { method This mul(This other)
+      This.of(this.x*other.x,this.y*other.y)
+    }
+  pointDiv: ...
+  PointAlgebra:Use pointSum,pointMul,pointDiv,...
+\end{lstlisting}
+
+Now our code is fully modularized, and each trait handle exactly one method.
+
+What happens if we want to add fields instead of just operations?
+
+\begin{lstlisting}
+  colored:{ method Color color() }
+  Point:Sum pointSum,colored //fails
+\end{lstlisting}
+
+This first attempt does not work: the abstract color method
+is not a getter for any of the parameters of 
+\Q@ class method This of(int x,int y)@
+A solution is to provide a richer factory:
+
+\begin{lstlisting}
+  CPoint:Use pointSum,colored,{
+    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/))
+    class method This of(int x, int y,Color color)
+    }
+\end{lstlisting}
+
+where we assume to support overloading on different parameter number.
+This is a good solution, we think is better that any alternatives in literature,
+however the method \Q@CPoint.sum@ resets the color to red.
+What should be the behaviour in this case?
+If we support withers, instead of writing \Q@This.of()@ we can use
+\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
+Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
+
+If the point designer could predict this kind of extension, then we could use the following design:  
+\begin{lstlisting}
+  p: { method int x() method int y()
+    method This withX(int that)
+    method This withY(int that)
+    static method This of(int x,int y)
+    method This merge(This other)
+    }
+  pointSum:Use p, { method This sum(This other)
+      this.merge(other).withX(this.x+other.x).withY(this.y+other.y)
+    }
+  colored:{method Color color()
+    method This withColor(Color that)
+    method This merge(This other)
+      this.withColor(this.color().mix(that.color())
+    }
+  CPoint:/*as before*/
+\end{lstlisting}  
+  Now we can merge colors, or any other kind of state we may want to add
+  following this pattern.
+  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
+  need to compose the merge operation inside of both of them.
+  The simple model we are presenting could accomodate this with an
+  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
+This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
+
+\section{Old: Self Types, Binary Methods and the Need to Separate
   Inheritance from Subtyping}

### 2017-05-18 FACS/paper.tex
.
bib
Comments/changes
Intro
Abstract

@@
 \usepackage{times}
+\usepackage{comment}
 \lstset{language=Java,
@@
 \newcommand{\prodFull}[3]{#1&::=&#2&\mbox{#3}}
+%\newcommand{\prodFull}[3]{#1&::=&\mbox{#2}&\mbox{#3}}
 \newcommand{\prodInline}[2]{#1::=#2}
@@
 \begin{abstract}
-Abstract here!
-\dots
+In most object oriented languages subclassing/code-reuse implies
+subtyping. This is considered by many a conceptual design error, but it
+seems required by technical reasons.
+
+This paper synthesizes an OO language design for a nominally typed
+language that allows code-reuse and subtyping to be decoupled using a
+form of \emph{traits}. This design improves on previous work by
+Damiani and Servetto by proving a simpler and more natural model for
+OO programmers. We provide a novel and elegant solution to uniformly
+handle behavior and state within trait composition.  By making our
+type system distinguish between code use and code reuse we can avoid
+many redundant abstract declarations that were required in previous
+work by Servetto and Damiani, while supporting self construction,
+binary methods and recursive types.
+
+\begin{comment}
+Two distinct groups of researchers independently [.marco..] and
+[..ferruccio.] used traits to provide
+a sound language design where code-reuse and subtyping are decoupled.
+In this work we distill their work in a simpler model that is far more
+natural for a OO programmer.
+For example, in [ marco ] class A {A m(){return this;}} is not well typed code,
+and in [.ferruccio..] all the logic about object construction have to
+be provided after all trait composition.
+
+We provide an novel and elegant solution to uniformly handle behavior
+and state within trait composition,
+Making our type system distinguish between used code and reused code
+we  can avoid many redundant
+abstract declarations that was required by [ marco ] and [ ferruccio ],
+while supporting self construction, binary methods and recursive types;
+
+We then show that our model can easily support nested classes and a
+rich composition algebra as in [ marco ]
+\end{comment}
+
 \keywords{Code Reuse, Object-Oriented Programming}
@@
 \input{Examples}
+\input{Extensions}
 \input{RelatedWork}
@@
 
+\section*{Bibliography}
+\bibliographystyle{elsart-num-sort}
+\bibliography{main}
+
 \end{document}

### 2017-05-18 FACS/paper.tex
Abstract

@@
 subtyping. This is considered by many a conceptual design error, but it
-seems required by technical reasons.
+seems required for technical reasons.
 
@@
 Damiani and Servetto by proving a simpler and more natural model for
-OO programmers. We provide a novel and elegant solution to uniformly
-handle behavior and state within trait composition.  By making our
+OO programmers. By making our
 type system distinguish between code use and code reuse we can avoid
@@
 work by Servetto and Damiani, while supporting self construction,
-binary methods and recursive types.
+binary methods and recursive types. Moreover, we provide a novel and elegant solution to uniformly
+handle behavior and state within trait composition.
 

### 2017-05-18 FACS/paper.tex
Abstract

@@
 subtyping. This is considered by many a conceptual design error, but it
-seems required for technical reasons.
+seems required for technical reasons, due to what we call the
+\emph{leaking problem}. The leaking problem shows that the separating
+code-reuse from subtyping is non-trivial and requires a significant
+departure from the OO models in existing mainstream OO languages.
 
@@
 form of \emph{traits}. This design improves on previous work by
-Damiani and Servetto by proving a simpler and more natural model for
+Damiani and Servetto by providing a simpler and more natural model for
 OO programmers. By making our

### 2017-05-18 FACS/paper.tex
new abstract

@@
 \begin{abstract}
-In most object oriented languages subclassing/code-reuse implies
+In most object oriented languages subclassing/inheritance implies
 subtyping. This is considered by many a conceptual design error, but it
 seems required for technical reasons, due to what we call the
-\emph{leaking problem}. The leaking problem shows that the separating
+\emph{this-leaking problem}, showing that separating
 code-reuse from subtyping is non-trivial and requires a significant
@@
 
-This paper synthesizes an OO language design for a nominally typed
-language that allows code-reuse and subtyping to be decoupled using a
-form of \emph{traits}. This design improves on previous work by
-Damiani and Servetto by providing a simpler and more natural model for
-OO programmers. By making our
-type system distinguish between code use and code reuse we can avoid
-many redundant abstract declarations that were required in previous
-work by Servetto and Damiani, while supporting self construction,
-binary methods and recursive types. Moreover, we provide a novel and elegant solution to uniformly
+We are aware of at least 3 independently designed research languages 
+that address this limitation:Delta-Trait, Package Template, DeepFjig.
+
+Leveraging on \emph{traits}, in this work we aim to synthesizes
+the main ideas of those very different language designs.
+Our result is a nominally typed minimalistic language, natural to OO programmers.
+By making our type system distinguish between code-use and code-reuse
+we can separe inheritance and subtyping, while avoiding 
+redundant abstract declarations (required in Delta-Trait and DeepFjig), while supporting self construction,
+binary methods and recursive types.
+Moreover, we provide a novel and elegant solution to uniformly
 handle behavior and state within trait composition.

### 2017-05-18 FACS/paper.tex
local
.

@@
 the main ideas of those very different language designs.
-Our result is a nominally typed minimalistic language, natural to OO programmers.
+Our result is a nominally typed minimalist language, natural to OO programmers.
 By making our type system distinguish between code-use and code-reuse
-we can separe inheritance and subtyping, while avoiding 
+we can separate inheritance and subtyping, while avoiding 
 redundant abstract declarations (required in Delta-Trait and DeepFjig), while supporting self construction,

### 2017-05-18 FACS/paper.tex

@@
 seems required for technical reasons, due to what we call the
-\emph{this-leaking problem}, showing that separating
+\emph{this-leaking problem}. The leaking problem shows that separating
 code-reuse from subtyping is non-trivial and requires a significant
 departure from the OO models in existing mainstream OO languages.
-
 We are aware of at least 3 independently designed research languages 
 that address this limitation:Delta-Trait, Package Template, DeepFjig.
-
 Leveraging on \emph{traits}, in this work we aim to synthesizes

### 2017-05-19 FACS/Intro.tex
.

@@
 
-\section{Improving Use by Preventing Unintended Subtyping}
-%%Preventing Unintended Subtyping}
-As stated in Section~\ref{} a possible benefit of a nominally typed OO language
-that separates inheritance from subtyping is that it is possible to 
-have reuse, while preventing unintended subtyping. The design of \name 
-allows this, thus it \emph{improves the use} of classes, while
-retaining the benefits of reuse when compared to Java-like languages. 
-Here we illustrate this benefit by modelling a simplified version of
-Set and Bag collections.
-
-\bruno{Plan: Show simplified Java code. Show how the Java code forces
-  us to make Bags a subtype of Set, if we want reuse. 
-Then show the 42/language in this paper/ code and solution.  
-}
-
-An iconic example on why connecting inheritance/code reuse and
-subtpying is problematic is provided by the
-historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
-implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
-keeping tracks of how many times an element occurs.  We would just add
-some state and override a couple of methods.\bruno{Are we going to
-  present this example in the paper solved in 42, for example? 
-I think I would expect to see it.}
-
-However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
-Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
-and making both \Q@Bag@ and \Q@Set@ inherit from it.
-This looks unnatural, since \Q@Set@ would extend it without adding anything,
-and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
-Worst, if we was to constantly apply this mentalty, we would introduce a very high number
-of abstract classes that are not supposed to be used as types, and that will clutter the 
-public interface of our classes and our code project as a whole.
-
-\begin{lstlisting}
-class Set {.../* Elem put() boolean isIn(Elem)*/ ...}
-class Bag extends Set{.../*override something to keep track of duplications*/..}
-\end{lstlisting}
-
-but now,
-
-\begin{lstlisting}
-Set mySet=new Bag(); //OK for the type system but not for LSP
-\end{lstlisting}
-
-as a pleasaruble accident, avoid such code gift us simple support for
-This type and (in the extensions with nested classes seen later)
-family polimporphism.
+\input{section3}
 

### 2017-05-19 FACS/section3.tex
.

\section{Improving Use by Preventing Unintended Subtyping}
%%Preventing Unintended Subtyping}
As stated in Section~\ref{} a possible benefit of a nominally typed OO language
that separates inheritance from subtyping is that it is possible to 
have reuse, while preventing unintended subtyping. The design of \name 
allows this, thus it \emph{improves the use} of classes, while
retaining the benefits of reuse when compared to Java-like languages. 
Here we illustrate this benefit by modelling a simplified version of
Set and Bag collections.

\bruno{Plan: Show simplified Java code. Show how the Java code forces
  us to make Bags a subtype of Set, if we want reuse. 
Then show the 42/language in this paper/ code and solution.  
}

An iconic example on why connecting inheritance/code reuse and
subtpying is problematic is provided by the
historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
keeping tracks of how many times an element occurs.  We would just add
some state and override a couple of methods.\bruno{Are we going to
  present this example in the paper solved in 42, for example? 
I think I would expect to see it.}

However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
and making both \Q@Bag@ and \Q@Set@ inherit from it.
This looks unnatural, since \Q@Set@ would extend it without adding anything,
and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
Worst, if we was to constantly apply this mentalty, we would introduce a very high number
of abstract classes that are not supposed to be used as types, and that will clutter the 
public interface of our classes and our code project as a whole.

\begin{lstlisting}
class Set {.../* Elem put() boolean isIn(Elem)*/ ...}
class Bag extends Set{.../*override something to keep track of duplications*/..}
\end{lstlisting}

but now,

\begin{lstlisting}
Set mySet=new Bag(); //OK for the type system but not for LSP
\end{lstlisting}

as a pleasaruble accident, avoid such code gift us simple support for
This type and (in the extensions with nested classes seen later)
family polimporphism.

### 2017-05-19 FACS/section3.tex
.

@@
 \section{Improving Use by Preventing Unintended Subtyping}
-%%Preventing Unintended Subtyping}
-As stated in Section~\ref{} a possible benefit of a nominally typed OO language
-that separates inheritance from subtyping is that it is possible to 
-have reuse, while preventing unintended subtyping. The design of \name 
-allows this, thus it \emph{improves the use} of classes, while
-retaining the benefits of reuse when compared to Java-like languages. 
+
+%As stated in Section~\ref{} 
+To have code-reuse, while preventing unintended subtyping
+in a nominally typed OO language we need to
+ separates inheritance from subtyping.
+
+Thus, \name  \emph{improves the use} of classes, while
+retaining the benefits of code-reuse when compared to Java-like languages. 
 Here we illustrate this benefit by modelling a simplified version of
@@
 
-\bruno{Plan: Show simplified Java code. Show how the Java code forces
-  us to make Bags a subtype of Set, if we want reuse. 
-Then show the 42/language in this paper/ code and solution.  
-}
-
+\paragraph{Set/Bag broken solution}
 An iconic example on why connecting inheritance/code reuse and
@@
 keeping tracks of how many times an element occurs.  We would just add
-some state and override a couple of methods.\bruno{Are we going to
-  present this example in the paper solved in 42, for example? 
-I think I would expect to see it.}
+some state and override a couple of methods.
+For example in Java one could have
 
-However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
+\begin{lstlisting}
+class Set {...//usual hashmap implementation
+  private Elem[] hashMap;
+  Elem put() /*body*/
+  boolean isIn(Elem)/*body*/
+  ...}
+class Bag extends Set{...//for each element in the hash map,
+  private int[] countMap;// keep track of how many times is there
+  @Override Elem put()/*body*/
+  int howManyTimes(Elem)/*body*/
+  ...}\end{lstlisting}
+
+Coding \Q@Bag@ in this way can avoid a lot of code duplication, but we induced unintented subtyping: indeed
+
+\begin{lstlisting}
+Set mySet=new Bag(); //OK for the type system but not for LSP
+\end{lstlisting}
+Since Subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
+
+\paragraph{Set/Bag fixed solution still broken}
 Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
-and making both \Q@Bag@ and \Q@Set@ inherit from it.
-This looks unnatural, since \Q@Set@ would extend it without adding anything,
+and making both \Q@Bag@ and \Q@Set@ inherit from it:
+\begin{lstlisting}
+abstract class AbstractSetOrBag{/*old set code goes here*/}
+class Set extends AbstractSetOrBag{} //empty body? just repeat constructors?
+class Bag extends AbstractSetOrBag{/*old bag code goes here*/}
+..
+AbstractSetOrBag unexpected=new Bag(); //AbstractSetOrBag type not designed to be used.
+\end{lstlisting}
+
+This looks unnatural, since \Q@Set@ would extend \Q@AbstractSetOrBag@ without adding anything,
 and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
@@
 public interface of our classes and our code project as a whole.
+In our example now the information \Q@Set<:AbstractSetOrBag@ is present in the public interface
+of the class \Q@Set@, but is not needed to use the class properly.
 
+Moreover, we have not really solved the original problem but only moved it further away. For example: one day  we may need bags that can only store up to 5 copies of the same element.
+We are now at the starting point again:
+\begin{itemize}
+\item either we inser \Q@class Bag5 extends Bag@ and we break LSP,
+\item we duplicate the code of the \Q@Bag@ implementation with minimal adjustments in \Q@class Bag5 extends AbstractSetOrBag@
+\item or we introduce a
+\Q@abstract class BagN extends AbstractSetOrBag@ and \Q@class Bag5 extends BagN@
+and we modify \Q@Bag@ so that  \Q@class Bag extends BagN@.
+Note that this last solution is changing the public interface of the formerly released \Q@Bag@ class, and
+this may even break retro-compatibility (if a client program was using reflection in a poor way)
+\end{itemize}
+
+\paragraph{Set/Bag in \name}
+Instead, in our proposed approach, if we was to originally declare
 \begin{lstlisting}
-class Set {.../* Elem put() boolean isIn(Elem)*/ ...}
-class Bag extends Set{.../*override something to keep track of duplications*/..}
+Set:{/*set implementation*/} 
 \end{lstlisting}
+Then our code would be impossible to reuse in the first place for any user of our library.
+We consider this an advantage, since unintended code reuse runs into underdocumented behaviour near all of the times\footnote{See
+``Design and document for inheritance or else prohibit it''\cite{https://www.infoq.com/articles/API-Design-Joshua-Bloch}:
+self use of public methods is rarelly documented, thus is hard to understand the effects of overriding a library metod\cite{https://www.infoq.com/articles/API-Design-Joshua-Bloch}.
+}
 
-but now,
+If the designer of the set class wishes to make it reusable, it can do it explicitly by providing a set trait:
+\begin{lstlisting}
+set:{/*set implementation*/} 
+Set:Use set
+\end{lstlisting}
+Notice how since \Q@set@ can never be used as a type, there is no reason to give it a fancy-future-aware name like
+\Q@AbstractSetOrBag@.
+When bag will be added, the code will look either like
+\begin{lstlisting}
+set:{/*set implementation*/} 
+Set:Use set
+Bag: Use set, {/*bag implementation*/}
+\end{lstlisting}
+or 
+\begin{lstlisting}
+set:{/*set implementation*/} 
+Set:Use set
+bag: Use set, {/*bag implementation*/}
+Bag: Use bag
+\end{lstlisting}
+Notice how, thanks to flattening, the resulting code for \Q@Bag@ is identical in both versions, thus if we are the developers of bags, we can temporarly go for the first version, and move when needed to the second without adding new undesired complexity for our old clients.
 
-\begin{lstlisting}
-Set mySet=new Bag(); //OK for the type system but not for LSP
-\end{lstlisting}
-
-as a pleasaruble accident, avoid such code gift us simple support for
-This type and (in the extensions with nested classes seen later)
-family polimporphism.
+%As a pleasaruble accident, avoid such code gift us simple support for
+%This type and (in the extensions with nested classes seen later)
+%family polimporphism.

### 2017-05-19 FACS/Intro.tex
.

@@
 \end{center}
+More complete table here, we can put it where you prefer
+\begin{center}
+\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c}
+&\Rotated{direct instantation}
+&\Rotated{self instantiation}
+&\Rotated{unit of use}
+&\Rotated{unit of reuse}
+&\Rotated{introduce type}
+&\Rotated{induced type is this type}
+&\Rotated{binary methods}
+&\Rotated{reuse induce subtype}
+&\Rotated{well-typed before imported}
+&\Rotated{well-typed before composed} 
+\\
+\hline
+java/scala class&Y&X&Y&Y&Y&Y&X&Y&Y&X\\
+java8 interface &X&X&X&Y&Y&Y&X&Y&Y&X\\
+scala trait        &X&X&X&Y&Y&Y&X&Y&X&X\\
+original trait     &X&X&X&Y&-&-&X&X&-&-\\
+Ferruccio trait  &X&X&X&Y&X&-&X&X&Y&Y\\
+42 trait            &X&Y&X&Y&X&-&Y&X&Y&X\\
+42 class           &Y&Y&Y&X&Y&Y&Y&-&Y&X\\
+module composition
+                      &-&-&Y&Y&-&-&-&-&Y&Y\\
+deepFJig class &Y&Y&Y&Y&Y&X&Y&X&Y&Y\\
+package template
+                      &X&Y&X&Y&X&-&X&X&Y&X\\
+
+\end{tabular}
+\end{center}
+
 \caption{A comparision between traits and classes in \name.}

### 2017-05-19 FACS/Intro.tex
.

--- 2017-05-19 FACS/Intro.tex
@@
 \end{itemize}
-
-\section{The Design of \name: Separating Use and Reuse}
-
-This section presents the overview of \name, and it illustrates the
-key ideas of its design. In particular we illustrate how to separate code use and 
-code reuse in \name. %, and how this allows improving both.
-
-%\subsection{The Design of \name: Decoupling Use from Reuse}
-
-\subsection{Classes in \name: A mechanism for code Use}
-%\name has a substantially different design from Java-like languages.
-The concept of a class in \name provides a mechanism for code use
-only. This means that in \name there is actually no subclassing, and
-classes are roughly equivalent to final classes in Java.  Thus,
-compared to Java-like languages, the most noticeable difference is the
-absence of the \Q@extends@ keyword in \name.
-
-To illustrate classes in \name consider the example in Section~\ref{}:
-
-\begin{lstlisting}
- A: { method int ma() Utils.m(this) }//note, no {return _}
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-\noindent Classes in \name use a slightly different declaration style compared
-to Java. In \name there is no \lstinline{class} keyword. 
-Class declarations have a name (which must always start with
-an uppercase letter) and a code literal, which is used to specify the
-definitions of the class. For instance, in the class declaration
-for \lstinline{A}, the name of the class is \lstinline{A} and the code 
-literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
-class. There are, however, some 
-important differences to Java-like languages in the way 
-classes and code-literals are type-checked, as we shall see next. 
-Nevertheless, for this example, things still work in a similar way to Java.
-
-The \name code above is fine, but there is no way to add a class 
-\Q@B@ reusing the code of \Q@A@, since
-\Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
-the Java code, introducing a subclass
-\lstinline{B} is not possible. At first, this may seem like a severe restriction, but
-\name has a different mechanism for code \emph{reuse} that 
-is more appropriate when code reuse is intended.
-
-\subsection{Traits in \name: A mechanim of code Reuse}
-
-Unlike classes traits in \name cannot be instantiated and do not introduce new
-types. However they provide code reuse.
-%So, lets try again encoding the code for the leaking problem, but this
-%time aiming at code reuse. 
-Trait declarations 
-look very much like class declarations, but trait names 
-start with a lowercase letter. An obvious first attempt 
-to model the example in Section~\ref{} (for a Java programmer) 
-with traits and code reuse is:
-
-\begin{lstlisting}
- ta:{ method int ma() Utils.m(this) }//type error
- A:Use ta
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-\noindent Here \lstinline{ta} is a trait intended to replace the
-original class \lstinline{A} so that the code of the method
-\lstinline{ma} can be reused. Then the class \lstinline{A} 
-is created by inheriting the code from the trait using the keyword 
-{\bf Use}. Note that \use\ cannot contain class names: only trait
-names are allowed.
-That is, using a trait is the only way to induce code reuse.
-Unfortunatelly, this code does not work, 
-because \lstinline{Utils.m} requires an \lstinline{A} and the type of \lstinline{this} in
-\Q@ta@ has no relationship to the type \lstinline{A}.
-Since the trait name is not a type, no code external to that trait can
-refer to it. This is one of the key design decisions in \name. 
-The point is that the trait can refer to the program via \lstinline{this}, but the
-program is agnostic to what the trait type is going to be, so it can
-be later assigned to any (or many) classes. This improves the
-flexibility of reuse as illustrated in Section~\ref{}. However, 
-to solve the more immediate typing issue above we need one more
-round of refactoring, as we shall see next.
-
-%: in \name all declarations 
-%are sequentially type-checked and if a definition occurs afterwards
-%another, then it is not visible on all previous definitions.
-
-\paragraph{Type of the self-reference} The code above does not work
-because the way type-checking works in \name is that type-checking of 
-code literals is independent of the class/trait names associated to
-it. A Java programmer may expect that the type of \lstinline{this} 
-in the previous definition of \lstinline{ta} is \lstinline{ta}
-itself. However, this intuition brought from Java is wrong in \name
-for two reasons:
-
-\begin{itemize}
-
-\item The first reason is that {\bf traits are not types}. Traits in
-  \name are simply units of reuse and cannot be used as types. Types
-  are only introduced by class/interface declarations.
-
-\item The second reason is that {\bf the type of self-references is the self type of the
-    code literal.} Unlike Java-like languages, where the body of a
-  class declaration is intrinsically coupled with the class itself, in
-  \name code literals are first-class\bruno{Is it appropriate to say
-    that they are first class?} and are type-checked independently.
-  \name has a notion of self-type, which is closely related to
-  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
-  in \name the self-reference does not have the type of the class
-  being defined, but rather it has the self-type.
-\end{itemize}
-
-With this in mind, we can try to model the example in Section~\ref{} again:
-
-\begin{lstlisting}
- IA:{interface method int ma()}//interface with abstract method
- Utils:{ class method int m(IA a)/*method body here*/}
- ta:{implements IA
-   method int ma() Utils.m(this) }
- A:Use ta
-\end{lstlisting}
-
-This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. Thus the self-type of the trait \Q@ta@ (and the
-type of \lstinline{this}) will be a
-subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
-\begin{lstlisting}
-  B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
-Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.
-
-\paragraph*{Semantic of Use}
-Albeit alternative semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
-\begin{lstlisting}
-A:Use ta
-B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-
-\noindent reduces/is equivalent to/is flatted into
-
- \begin{lstlisting}
-A:{implements IA method int ma() Utils.m(this) }
-B:{implements IA
-  method int ma() Utils.m(this)
-  method int mb() this.ma() }
-
- \end{lstlisting}
-
- This code seems correct, and there is no mention of the trait
- \Q@ta@. In some sense, all the information about code
- reuse/subclassing is just a private implementation detail of \Q@A@
- and \Q@B@; while subtyping is part of the class interface.
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{c|c|c|c}
-              & Instantiable (Use) & Unit of Reuse & Introduces New Type \\
-\hline
-Class  & Yes              & No          & Yes           \\
-Trait   & No               & Yes         & No            \\
-\end{tabular}
-\end{center}
-\caption{A comparision between traits and classes in \name.}
-\label{fig:compare}
-\end{figure}
-
-To finish this section, Figure \ref{fig:compare} provides a summary of
-the differences between classes and traits. The comparison focus on
-the roles of traits and classes with respect to instantiation,
-reusability and whether the declarations also introduce new types or
-not.
-
+\input{section2}
 \input{section3}

### 2017-05-19 FACS/section2.tex
.

\section{The Design of \name: Separating Use and Reuse}

This section presents the overview of \name, and it illustrates the
key ideas of its design. In particular we illustrate how to separate code use and 
code reuse in \name. %, and how this allows improving both.

%\subsection{The Design of \name: Decoupling Use from Reuse}

\subsection{Classes in \name: A mechanism for code Use}
%\name has a substantially different design from Java-like languages.
The concept of a class in \name provides a mechanism for code use
only. This means that in \name there is actually no subclassing, and
classes are roughly equivalent to final classes in Java.  Thus,
compared to Java-like languages, the most noticeable difference is the
absence of the \Q@extends@ keyword in \name.

To illustrate classes in \name consider the example in Section~\ref{}:

\begin{lstlisting}
 A: { method int ma() Utils.m(this) }//note, no {return _}
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

\noindent Classes in \name use a slightly different declaration style compared
to Java. In \name there is no \lstinline{class} keyword. 
Class declarations have a name (which must always start with
an uppercase letter) and a code literal, which is used to specify the
definitions of the class. For instance, in the class declaration
for \lstinline{A}, the name of the class is \lstinline{A} and the code 
literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
class. There are, however, some 
important differences to Java-like languages in the way 
classes and code-literals are type-checked, as we shall see next. 
Nevertheless, for this example, things still work in a similar way to Java.

The \name code above is fine, but there is no way to add a class 
\Q@B@ reusing the code of \Q@A@, since
\Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
the Java code, introducing a subclass
\lstinline{B} is not possible. At first, this may seem like a severe restriction, but
\name has a different mechanism for code \emph{reuse} that 
is more appropriate when code reuse is intended.

\subsection{Traits in \name: A mechanim of code Reuse}

Unlike classes traits in \name cannot be instantiated and do not introduce new
types. However they provide code reuse.
%So, lets try again encoding the code for the leaking problem, but this
%time aiming at code reuse. 
Trait declarations 
look very much like class declarations, but trait names 
start with a lowercase letter. An obvious first attempt 
to model the example in Section~\ref{} (for a Java programmer) 
with traits and code reuse is:

\begin{lstlisting}
 ta:{ method int ma() Utils.m(this) }//type error
 A:Use ta
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

\noindent Here \lstinline{ta} is a trait intended to replace the
original class \lstinline{A} so that the code of the method
\lstinline{ma} can be reused. Then the class \lstinline{A} 
is created by inheriting the code from the trait using the keyword 
{\bf Use}. Note that \use\ cannot contain class names: only trait
names are allowed.
That is, using a trait is the only way to induce code reuse.
Unfortunatelly, this code does not work, 
because \lstinline{Utils.m} requires an \lstinline{A} and the type of \lstinline{this} in
\Q@ta@ has no relationship to the type \lstinline{A}.
Since the trait name is not a type, no code external to that trait can
refer to it. This is one of the key design decisions in \name. 
The point is that the trait can refer to the program via \lstinline{this}, but the
program is agnostic to what the trait type is going to be, so it can
be later assigned to any (or many) classes. This improves the
flexibility of reuse as illustrated in Section~\ref{}. However, 
to solve the more immediate typing issue above we need one more
round of refactoring, as we shall see next.

%: in \name all declarations 
%are sequentially type-checked and if a definition occurs afterwards
%another, then it is not visible on all previous definitions.

\paragraph{Type of the self-reference} The code above does not work
because the way type-checking works in \name is that type-checking of 
code literals is independent of the class/trait names associated to
it. A Java programmer may expect that the type of \lstinline{this} 
in the previous definition of \lstinline{ta} is \lstinline{ta}
itself. However, this intuition brought from Java is wrong in \name
for two reasons:

\begin{itemize}

\item The first reason is that {\bf traits are not types}. Traits in
  \name are simply units of reuse and cannot be used as types. Types
  are only introduced by class/interface declarations.

\item The second reason is that {\bf the type of self-references is the self type of the
    code literal.} Unlike Java-like languages, where the body of a
  class declaration is intrinsically coupled with the class itself, in
  \name code literals are first-class\bruno{Is it appropriate to say
    that they are first class?} and are type-checked independently.
  \name has a notion of self-type, which is closely related to
  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
  in \name the self-reference does not have the type of the class
  being defined, but rather it has the self-type.
\end{itemize}

With this in mind, we can try to model the example in Section~\ref{} again:

\begin{lstlisting}
 IA:{interface method int ma()}//interface with abstract method
 Utils:{ class method int m(IA a)/*method body here*/}
 ta:{implements IA
   method int ma() Utils.m(this) }
 A:Use ta
\end{lstlisting}

This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
is implementing it. Thus the self-type of the trait \Q@ta@ (and the
type of \lstinline{this}) will be a
subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
\begin{lstlisting}
  B:Use ta, { method int mb(){return this.ma();} }
\end{lstlisting}
This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.

\paragraph*{Semantic of Use}
Albeit alternative semantic models for traits~\cite{} have been proposed,
here we use the flattening model.
This means that 
\begin{lstlisting}
A:Use ta
B:Use ta, { method int mb(){return this.ma();} }
\end{lstlisting}

\noindent reduces/is equivalent to/is flatted into

 \begin{lstlisting}
A:{implements IA method int ma() Utils.m(this) }
B:{implements IA
  method int ma() Utils.m(this)
  method int mb() this.ma() }

 \end{lstlisting}

 This code seems correct, and there is no mention of the trait
 \Q@ta@. In some sense, all the information about code
 reuse/subclassing is just a private implementation detail of \Q@A@
 and \Q@B@; while subtyping is part of the class interface.

\begin{figure}[t]
\begin{center}
\begin{tabular}{c|c|c|c}
              & Instantiable (Use) & Unit of Reuse & Introduces New Type \\
\hline
Class  & Yes              & No          & Yes           \\
Trait   & No               & Yes         & No            \\
\end{tabular}
\end{center}
More complete table here, we can put it where you prefer
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c}
&\Rotated{direct instantation}
&\Rotated{self instantiation}
&\Rotated{unit of use}
&\Rotated{unit of reuse}
&\Rotated{introduce type}
&\Rotated{induced type is this type}
&\Rotated{binary methods}
&\Rotated{reuse induce subtype}
&\Rotated{well-typed before imported}
&\Rotated{well-typed before composed} 
\\
\hline
java/scala class&3Y&X&Y&Y&Y&Y&X&Y&Y&X\\
java8 interface &X&X&X&Y&Y&Y&X&Y&Y&X\\
scala trait        &X&X&X&Y&Y&Y&X&Y&X&X\\
original trait     &X&X&X&Y&-&-&X&X&-&-\\
Ferruccio trait  &X&X&X&Y&X&-&X&X&Y&Y\\
42 trait            &X&Y&X&Y&X&-&Y&X&Y&X\\
42 class           &Y&Y&Y&X&Y&Y&Y&-&Y&X\\
module composition
                      &-&-&Y&Y&-&-&-&-&Y&Y\\
deepFJig class &Y&Y&Y&Y&Y&X&Y&X&Y&Y\\
package template
                      &X&Y&X&Y&X&-&X&X&Y&X\\

\end{tabular}
\end{center}

\caption{A comparision between traits and classes in \name.}
\label{fig:compare}
\end{figure}

To finish this section, Figure \ref{fig:compare} provides a summary of
the differences between classes and traits. The comparison focus on
the roles of traits and classes with respect to instantiation,
reusability and whether the declarations also introduce new types or
not.

### 2017-05-19 FACS/section2.tex
.

@@
 key ideas of its design. In particular we illustrate how to separate code use and 
-code reuse in \name. %, and how this allows improving both.
+code reuse.
+% in \name. %, and how this allows improving both.
 
@@
 %\name has a substantially different design from Java-like languages.
-The concept of a class in \name provides a mechanism for code use
-only. This means that in \name there is actually no subclassing, and
+The concept of a class in \name provides a mechanism for code-use
+only. This means that there is actually no subclassing:
 classes are roughly equivalent to final classes in Java.  Thus,
@@
 \noindent Classes in \name use a slightly different declaration style compared
-to Java. In \name there is no \lstinline{class} keyword. 
-Class declarations have a name (which must always start with
-an uppercase letter) and a code literal, which is used to specify the
-definitions of the class. For instance, in the class declaration
+to Java: there is no \lstinline{class} keyword, and a colon separe the class name (which must always start with
+an uppercase letter) and the class implementation, which is used to specify the
+definitions of the class. In our example, in the class declaration
 for \lstinline{A}, the name of the class is \lstinline{A} and the code 
 literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
-class. There are, however, some 
-important differences to Java-like languages in the way 
-classes and code-literals are type-checked, as we shall see next. 
-Nevertheless, for this example, things still work in a similar way to Java.
+class.
+
+%We will see next some 
+%important differences to Java-like languages in the way 
+%classes and code-literals are type-checked, as we shall see next. 
+%Nevertheless, for this example, things still work in a similar way to Java.
 
@@
 
-Unlike classes traits in \name cannot be instantiated and do not introduce new
+Unlike classes, traits in \name cannot be instantiated and do not introduce new
 types. However they provide code reuse.
@@
 \Q@ta@ has no relationship to the type \lstinline{A}.
-Since the trait name is not a type, no code external to that trait can
+
+A Java programmer may then try to write
+\begin{lstlisting}
+ ta:{ method int ma() Utils.m(this) }//type error
+ A:Use ta
+ Utils:{ class method int m(ta a)//syntax error
+   /*method body here*/ }
+\end{lstlisting}
+But this does not work either, indeed \Q@ta@ is not a type in the first place, since it is a (lowercase) trait name.
+Indeed
+since the trait name is not a type, no code external to that trait can
 refer to it. This is one of the key design decisions in \name.
-The point is that the trait can refer to the program via \lstinline{this}, but the
-program is agnostic to what the trait type is going to be, so it can
-be later assigned to any (or many) classes. This improves the
-flexibility of reuse as illustrated in Section~\ref{}. However, 
-to solve the more immediate typing issue above we need one more
-round of refactoring, as we shall see next.
+
+% this text may go after, while explaining the This type?
+%Code inside traits/classes can refer to (uppercase) class names, but not to (lowercase) trait names.
+%%The point is that the trait can refer to the program via \lstinline{this},
+%??????????????
+% This means that the
+%program is agnostic to what the trait type is going to be, so it can
+%be later assigned to any (or many) classes. This improves the
+%flexibility of reuse as illustrated in Section~\ref{}.
+%However, 
+%to solve the more immediate typing issue above we need one more round of refactoring, as we shall see next.
 
@@
 
-\paragraph{Type of the self-reference} The code above does not work
-because the way type-checking works in \name is that type-checking of 
-code literals is independent of the class/trait names associated to
-it. A Java programmer may expect that the type of \lstinline{this} 
-in the previous definition of \lstinline{ta} is \lstinline{ta}
-itself. However, this intuition brought from Java is wrong in \name
-for two reasons:
+%\paragraph{Type of the self-reference} 
+%The code above does not work.
+%because the way type-checking works in \name is that type-checking of 
+%code literals is independent of the class/trait names associated to it.
+%A Java programmer may expect that the type of \lstinline{this} 
+%in the previous definition of \lstinline{ta} is \lstinline{ta}
+%itself. However, this intuition brought from Java is wrong in \name
+%for two reasons:
 
-\begin{itemize}
+%\begin{itemize}
 
-\item The first reason is that {\bf traits are not types}. Traits in
-  \name are simply units of reuse and cannot be used as types. Types
-  are only introduced by class/interface declarations.
+%\item {\bf Traits are not types}. Traits in
+%  \name are simply units of reuse and cannot be used as types. Types
+%  are only introduced by class/interface declarations.
 
-\item The second reason is that {\bf the type of self-references is the self type of the
-    code literal.} Unlike Java-like languages, where the body of a
-  class declaration is intrinsically coupled with the class itself, in
-  \name code literals are first-class\bruno{Is it appropriate to say
-    that they are first class?} and are type-checked independently.
-  \name has a notion of self-type, which is closely related to
-  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
-  in \name the self-reference does not have the type of the class
-  being defined, but rather it has the self-type.
-\end{itemize}
+%\item The second reason is that {\bf the type of self-references is the self type of the
+%    code literal.} Unlike Java-like languages, where the body of a
+%  class declaration is intrinsically coupled with the class itself, in
+%  \name code literals are first-class\bruno{Is it appropriate to say
+%    that they are first class?} and are type-checked independently.
+%  \name has a notion of self-type, which is closely related to
+%  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
+%  in \name the self-reference does not have the type of the class
+%  being defined, but rather it has the self-type.
+%\end{itemize}
 
@@
 
-This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. Thus the self-type of the trait \Q@ta@ (and the
-type of \lstinline{this}) will be a
-subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
+This code works: \Q@Utils@ rely over interface \Q@IA@ and the trait \Q@ta@
+implements \Q@IA@.
+\Q@ta@ is well typed: independently of what class name(s) will be associated to its code, we knew such class(es) will implement \Q@IA@, thus while typechecking \Q@Utils.m(this)@ we can assume
+\Q@this<:IA@.
+ It is also possible to add a \Q@B@ as follows
 \begin{lstlisting}
@@
 Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.
+
+Later, in section \ref{} we will provide a more formal detail over type system. For now notice that in the former example the code is correct even if no method called \Q@ma@ is explicitly declared.
+The idea is that such method is imported from trait \Q@ta@, exactly as in the java equivalent
+\begin{lstlisting}[language=Java]
+  class B extends A{ int mb(){return this.ma();} }  
+\end{lstlisting}
+method \Q@ma@ is imported from \Q@A@.
+This concept is natural for a Java programmer, but was not supported in previous work \cite{deep, ferruccio}
 
@@
  This code seems correct, and there is no mention of the trait
- \Q@ta@. In some sense, all the information about code
- reuse/subclassing is just a private implementation detail of \Q@A@
+ \Q@ta@. In some sense, all the information about code-reuse/inheritance
+  is just a private implementation detail of \Q@A@
  and \Q@B@; while subtyping is part of the class interface.
+
+%To finish this section, Figure \ref{fig:compare} provides a summary of
+%the differences between classes and traits. The comparison focus on
+%the roles of traits and classes with respect to instantiation,
+%reusability and whether the declarations also introduce new types or
+%not.
 
@@
 \end{figure}
-
-To finish this section, Figure \ref{fig:compare} provides a summary of
-the differences between classes and traits. The comparison focus on
-the roles of traits and classes with respect to instantiation,
-reusability and whether the declarations also introduce new types or
-not.

### 2017-05-20 FACS/Intro.tex
.

@@
 \input{section3}
-
-\section{Improving Reuse when Subtyping is Impossible}
-
-\name retains the ability to allow reuse even when subtyping is
-impossible. That is, in some situations a subclass contains methods
-whose signatures are incompatible with the superclass, yet inheritance
-is still possible. This type of situations is the primary motivator
-for previous work aiming at separating inheritance from subtyping, and
-it happens, for example, in classes with \emph{binary
-  methods}~\cite{}.  Despite the adoption of a nominal approach and
-unlike other nominally typed languages, the leaking problem does not
-prevent \name from enjoying such kind of improved reuse.
-
-To illustrate how \name allows for improved reuse, we consider a
-pattern to combine \emph{multiple trait inheritance} and \emph{state}:
-a challenging problem that has limited the flexibility of traits and
-reuse in the past.
-
-\subsection{Managing State}
-
-This idea of summing pieces of code is very elegant, and has proven
-very successfull in module composition
-languages~\cite{ancona2002calculus}\bruno{and several trait models?}. 
-However our research community is
-struggling to make it work with object state (constructor and fields)
-while acheving the following goals:
-
-\begin{itemize}
-\item keep sum associative and commutative,
-\item allowing a class to create instances of itself,
-\item actually initialize objects, leaving no null fields,
-\item managing fields in a way that borrows the elegance of summing methods,
-\item make it easy to add new fields.
-\end{itemize}
-
-In the related work we will show some alternative ways to handle state.
-However the purest solution just requires methods:
-  The idea is that
-  the trait code just uses getter/setters/factories, while leaving
-  to classes the role to finally define the fields/constructors.
-  That is, the  the class has syntax richer that the trait one,
-  allowing declaration for fields and constructors.
-  This approach is very powefull~\cite{wang2016classless}
-
-\paragraph*{Advantages:} 
-this approach is associative and commutative, even self construction
-  can be allowed if the trait requires a static/class method
-  returning This; the class will then implement this method by forwarding
-  a call to the constructor.
-
-\paragraph*{Negatives:} writing the class code with the constructors and
-   fields and getter/setters and factories can be quite tedious.
-   Moreover, there is no way for a trait to specify a default value for a field,
-   the class need to handle all the state, even state that is conceptually
-   "private" of such trait.
-
-Here in the following an example of such approach:
-
-\begin{lstlisting}
-  pointSum: { method int x()  method int y()
-    class method This of(int x,int y)
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul: { method int x() method int y()
-    class method This of(int x,int y)
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-\end{lstlisting}
-
-As you see, all the state operations are represented as abstract methods.
-
-\subsection{A First Attempt at Composition}
-According to the general ideas expressed before,
-\begin{lstlisting}
-  Point:Use pointSum,pointMul
-\end{lstlisting}
-
-\noindent would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract.
-In this mindset, the user would be required to write something similar to
-
-\begin{lstlisting}
-  CPoint:Use pointSum,pointMul, {//not our suggested solution
-    int x   int y
-    method int x()x   method int y()y
-    class method This of(int x, int y)
-      new Point(x,y)
-    constructor Point(int x, int y){ this.x=x   this.y=y }
-    }
-\end{lstlisting}
-
-\noindent after a while programming in this style, 
-writing those obvious ``close the state'' classes become a
-repetivie boring job, and one wonder
-if it could be possible to automatically generate such code~\cite{wang2016classless}.
-Indeed those classes are just a form of ``fixpoint''.
-
-In our model we go one step further: there is no need to generate code, or to explicitly
-write down constructors and fields; there is not even syntax for those
-constructor.
-The idea is that any class that ``could" be completed in the obvious way
-\emph{is a complete ``coherent" class}.
-In most other languages, a class is abstract if have abstract methods.
-Instead, we call abstract a class whose set of abstract methods is not
-coherent, that is, can not be automatically recognized as factory,getters and setters.
-
-\paragraph*{Detaild definition of coherent:}
-\begin{itemize}
-\item a class with no abstract method is coherent, and like Java \Q@Math@.
-Will just be usefull for calling class/static methods.
-\item a class with a single abstract \Q@class@ method returning \Q@This@
-is coherent if all the other abstract methods can be seen as \emph{abstract state
-operations} over one of its argument.
-For example,
-if there is a \Q@class method This of(int x, int y)@ as before,
-then
-\item a method \Q@int x()@ is intepreted as an abstract state method: a getter for \Q@x@.
-\item a method \Q@Void x(int that)@ is a setter for x.
-\end{itemize}
-
-While getters and setters are fundamental operations, we can immagine
-more operations to be supported; for example
-\begin{itemize}
-\item \Q@method This withX(int that)@
-may be a ``wither", doing a functional field update.
-\item \Q@method Void update(int x,int y)@
-may do two field update at a time.
-\item\Q@method This clone()@ may do a shallow clone of the object.
-\end{itemize}
-
-We are not sure what is the best set of abstract state operations yet, but we think this could become a very interesting area of research.
-
-  lets play with the points of before, to see what good can we do with the current
-  instruments:
-
-\begin{lstlisting}
- //same code as before for pointSum and pointMul
-  pointSum: { method int x() method int y()
-    class method This of(int x,int y)
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul: { method int x() method int y()//look we are repeating
-    class method This of(int x,int y)//the abstract method declarations.
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-  PointAlgebra:Sum pointSum,pointMul
-\end{lstlisting}
-
-  As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods \Q@x@,\Q@y@ and \Q@of@.
-  In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
-  and not repeat such abstract definitions? of course!
-
-\begin{lstlisting}
-  p: { method int x() method int y()
-    class method This of(int x,int y)
-    }
-  pointSum:Use p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul:Use p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-  pointDiv: ...
-  PointAlgebra:Use pointSum,pointMul,pointDiv,...
-\end{lstlisting}
-
-Now our code is fully modularized, and each trait handle exactly one method.
-
-What happens if we want to add fields instead of just operations?
-
-\begin{lstlisting}
-  colored:{ method Color color() }
-  Point:Sum pointSum,colored //fails
-\end{lstlisting}
-
-This first attempt does not work: the abstract color method
-is not a getter for any of the parameters of 
-\Q@ class method This of(int x,int y)@
-A solution is to provide a richer factory:
-
-\begin{lstlisting}
-  CPoint:Use pointSum,colored,{
-    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/))
-    class method This of(int x, int y,Color color)
-    }
-\end{lstlisting}
-
-where we assume to support overloading on different parameter number.
-This is a good solution, we think is better that any alternatives in literature,
-however the method \Q@CPoint.sum@ resets the color to red.
-What should be the behaviour in this case?
-If we support withers, instead of writing \Q@This.of()@ we can use
-\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
-Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
-
-If the point designer could predict this kind of extension, then we could use the following design:  
-\begin{lstlisting}
-  p: { method int x() method int y()
-    method This withX(int that)
-    method This withY(int that)
-    static method This of(int x,int y)
-    method This merge(This other)
-    }
-  pointSum:Use p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y)
-    }
-  colored:{method Color color()
-    method This withColor(Color that)
-    method This merge(This other)
-      this.withColor(this.color().mix(that.color())
-    }
-  CPoint:/*as before*/
-\end{lstlisting}  
-  Now we can merge colors, or any other kind of state we may want to add
-  following this pattern.
-  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
-  need to compose the merge operation inside of both of them.
-  The simple model we are presenting could accomodate this with an
-  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
-This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
+\input{section4}
 

### 2017-05-20 FACS/section4.tex
.

--- 2017-04-04 FACS/Examples.tex
@@
-\section{A Decoupled Trait Calculus}
-In order to illustrate how we plan to properly divide code reuse/inheritance/subclassing from subtyping,
-lets consider a simple language where trait names start lowercase \mt
-and class names start uppercase \mC.  
-The syntax of the language is:
+\section{Improving Reuse when Subtyping is Impossible}
 
-\begin{bnf}
-\prodFull\mTD{\mt\terminalCode{:}\mL \mid \mt\terminalCode{:} \use\ \overline\mV}{Trait Decl}\\
-\prodFull\mCD{\mC\terminalCode{:}\mL \mid \mC\terminalCode{:} \use\ \overline\mV}{Class Decl}\\
-\prodFull\mV{\mt \mid \mL}{Code Value}\\
-\prodFull\mL{\Opt{\terminalCode{interface}}\ \terminalCode{implements} \overline\mT\ \overline\mMD}{Code Literal}\\
-\prodFull\mT{\mC}{types are class names}\\
-\prodFull\mMD{\Opt{\terminalCode{class}}\ \terminalCode{method}\ \mT\ \mm\oR\overline{\mT\,\mx}\cR \Opt\me}{Method Decl}\\
+\name retains the ability to allow reuse even when subtyping is
+impossible. That is, in some situations a subclass contains methods
+whose signatures are incompatible with the superclass, yet inheritance
+is still possible. This type of situations is the primary motivator
+for previous work aiming at separating inheritance from subtyping, and
+it happens, for example, in classes with \emph{binary
+  methods}~\cite{}.  Despite the adoption of a nominal approach and
+unlike other nominally typed languages, the leaking problem does not
+prevent \name from enjoying such kind of improved reuse.
 
-\prodFull\me{\mx\mid\mT\mid\me\terminalCode{.}\mm\oR\overline\me\cR}{expressions}\\
-\prodFull\mD{\mCD\mid\mTD}{Declaration}\\
-\end{bnf}
+To illustrate how \name allows for improved reuse, we consider a
+pattern to combine \emph{multiple trait inheritance} and \emph{state}:
+a challenging problem that has limited the flexibility of traits and
+reuse in the past.
 
-To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
-expression.  Traits come with various operators (restrict, hide,
-alias) but for now we focus on the single operator 
-$\use$, taking a set
-of code values: that is trait names \mt\ or literals \mL\ and composing them in an \emph{associative}
-and \emph{commutative} way.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
-composition operator.  $\use\ \overline\mV$ composes the content of $\overline\mV$
-by taking the union of the methods and the union of the implementations.
-In our simple mode, we consider an error trying to merge a class and an interface.
+\subsection{Managing State}
 
-Use can not be applied if multiple versions of the same method are
-present in different traits.  An exception is done for abstract methods:
-methods where the implementation \me\ is missing; In this case (if the
-headers are compatible) the implemented version is selected.  In a sum
-of two abstract methods with compatible headers, the one with the more
-specific type is selected. We will discuss later the formal details of
-\use.\marco{actually, this makes the sum not strongly associative...}
-
-Code literals \mL\ can be marked as interfaces.  We will call class
-interface, or simply interface a class declaration of for \Q@C:{interface  ...}@.
-  Then we have a set of implemented interfaces and a set of member declarations; in this simple language, just methods.
-If there are no implemented interfaces, in the concrete syntax we will omit the \Q@implements@ keyword.
-
-Methods \mMD can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very usefull in the context of code composition.
-To denote a method as abstract, instead of an optional keyword we just omit the implementation \me.
-
-A version of this language where there are no traits can be seen 
-as a restriction/variation of FJ~\cite{igarashi2001featherweight}.
-
-Basic well formedness rules applies:
-\begin{itemize}
-\item all the traits and classes have unique names in a program $\overline\mD$,
-\item all method parameters have unique names and the special parameter name \Q@this@ is not declared
- in the parameter list,
-\item all methods in a code literal have unique names,
-\item all used variables are in scope.
-\end{itemize}
-Those rules can be applied on any given \mL individually and in full isolation.
-
-We expect the type system to enforce 
-\begin{itemize}
-\item subtyping between interfaces and classes,
-\item method call typechecking,
-\item no circular implementation of interfaces,
-\item type signature of methods from interfaces can be refined following the well known variant-contravariant rules,
-\item only interfaces can be implemented.
-\item \marco{I'm sure I'm missing something}
-\end{itemize}
-We will see later the details on when exactly in the compilation process any code literal is typechecked.
-While classes are typed assuming \lstinline{this} is of the nominal type of the
-class, trait declarations, do not introduce any nominal type.  \lstinline{this}
-in a trait is typed with a special type \lstinline{This} that is visible only
-inside such trait. Syntactically, \Q@This@ is just a special, reserved, class name $\mC$.
-A Literal can use the \lstinline{This} type,
-and when the trait is reused to create a class, \Q@This@ will become just
-an alias for the class name.
-
-For the sake of simplicity, method bodies are just simple expressions \me, whose can be just variables and method calls.
-We do not think that there is any problem to extend our language with other constructs; we use minimal expressions to keep our example simple and compact.
-In order to make our examples more engaging, we feel free to use the type int, numeric literals and their operations.
-
-\subsection{Code Use and Code Reuse in \name}
-
-The great absent in our language is the  \Q@extends@ keyword.
-In a sense, all classes are final (as in Java).
-Moreover  \use\ can not contain class names.
-That is, using a trait is the only way to induce code reuse.
-
-Lets now try to encode the example in Section~\ref{} e in our language:
-
-\begin{lstlisting}
- A:{ method int ma() Utils.m(this) }//note, no {return _}
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-This code is fine, but there is no way to add a \Q@B@ reusing code of \Q@A@, since
-\Q@A@ is designed for code \emph{use} and not \emph{reuse}. Lets try
-again, but this time aiming at code reuse:
-
-\begin{lstlisting}
- ta:{ method int ma() Utils.m(this) }//type error
- A:Use ta
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
-\Q@ta@ have no knowlegde of \lstinline{A}. Lets try again
-
-\begin{lstlisting}
- IA:{interface method int ma()}//interface with abstract method
- ta:{implements IA
-   method int ma() Utils.m(this) }
- A:Use ta
- Utils:{ class method int m(IA a)/*method body here*/}
-\end{lstlisting}
-
-This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. It is also possible to add a \Q@B@ as follow
-\begin{lstlisting}
-  B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
-Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.
-
-\paragraph*{Semantic of Use}
-Abeit alternative semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
-\begin{lstlisting}
-A:Use ta
-B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-
-  reduces/is equivalent to/is flatted into
-
- \begin{lstlisting}
-A:{implements IA method int ma() Utils.m(this) }
-B:{implements IA
-  method int ma() Utils.m(this)
-  method int mb() this.ma() }
-
- \end{lstlisting}
-
-This code seems correct, and there is no mention of the trait \Q@ta@.
-In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.
-
-\subsection{Remarks on Typing}
-Our typing discipline is 
-what distinguish our approach from a simple minded code composition macro~\cite{bawden1999quasiquotation}
-or a rigid module composition~\cite{ancona2002calculus}
-
-The are two core ideas:
-\paragraph*{traits are \emph{well-typed} before being reused.}${}_{}$\\*
- For example in
-
-\begin{lstlisting}
-t:{method int m() 2 
-   method int n() this.m()+1}
-\end{lstlisting}
-
-\noindent \Q@t@ is well typed since \Q@m()@ is declared inside of \Q@t@, while
-
-\begin{lstlisting}
-t1:{method int n() this.m()+1} 
-\end{lstlisting}
-\noindent would be ill typed.
-
-\noindent Note how in the former example \Q@ta@ is well typed just by knowing itself and \Q@IA@.
-
-\paragraph*{class expressions are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
-In class expressions  $\use\ \overline\mV$
-an \mL\ in $\overline\mV$ is not typechecked before flattening, and only the result is expected to be well-typed.
-While this seams a very dangerous approach at first, consider that also Java have the same behaviour:
-for example in
-\begin{lstlisting}[language=Java]
-  class A{ int m() {return 2;}  int n(){return this.m()+1;} }
-  class B extends A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
-\noindent in \Q@B@ we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for \Q@ma()@.
-In our example,using the trait \Q@t@ of before
-
-\begin{lstlisting}
-C: Use t {method int k() this.n()+this.m()}
-\end{lstlisting}
-\noindent would be correct: even if n,m are not defined inside
-\Q@{method int k() this.n()+this.m()}@,
-the result of the flattening is well typed.
-
-This is not the case in many similar works in literature~\cite{} where the
-literals have to be self complete. In this case we would have been forced to
-declare abstract methods \Q@n@ and \Q@m@.
-
-Our typing strategy have two important properties:
-\begin{itemize}
-\item if a class is declared by using $\mC : \use\ \overline\mt$, that is, without literals,
-and the flattening is successfull, \mC\ is well typed, no need of further checking.
-\item on the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
-$\mL_1\ldots\mL_n \in \overline\mV$, and after successfull flattening $\mC : \mL$ can not be typechecked,
-then the issue was originally present in one of $\mL_1\ldots\mL_n$.
-It may be that the result is intrinsically ill-typed, if of the methods in $\mL_1\ldots\mL_n$ is not well typed,
-but it may also happens that a type referred from one of those methods is declared \emph{after} the current class. As we will see later, this is how our relaxation allows to support recursive types.
-
-This also means that as an optimization strategy
- we may remember what method bodies come from traits and what method bodies come from code literals, in order to typecheck only the latter.
- \end{itemize}
-
-We will see now how this idea that code literals wrote in the declaration of a class
-do not have to be well typed on their own also lets us reason about mutually recursive types.
-
- \section{Recursive types}
-
-OO language leverage on recursive types most of the times.
-For example in a pure OO language, \Q@String@ may offers a \Q@Int size()@
-method, and \Q@Int@ may offer a \Q@String toString()@ method.
-
-This means that is not possible to type in (full) isolation classes
-\Q@String@ and \Q@Int@.
-
-The most expressive compilation process may divide the classes in groups of mutually 
-dependent classes.
-Each group may also depend from a number of other groups.
-This would form a Direct Aciclyc Graph of groups.
-To type a group, we first need to type all depended groups, then
-we can extract the structure/signature/structural type of all
-the classes of the group.
-Now, with the information of the depended groups and the one extracted
-from the current group, it is possible to typecheck the implementation
- of each class in the group.
-
-In this model, it is reasonable to assume that flattening happens group by group, before
-extracting the class signatures.
-
-Here we go for a much simpler simple top down execution/interpretation for flattening, where flattening
-happen one at the time, and classes are typechecked where their type is first needed.
-
-For example
-\begin{lstlisting}
-A:{method int ma(B b) b.mb()+1}
-tb:{method int mb() 2}
-tc:{method int mc(A a,B b) a.ma(b)}
-B: Use tb
-C: Use tc, {method int hello() 1}
-\end{lstlisting}
-In this scenario, since we go top down, we first need to generate \Q@B@.
-To generate \Q@B@, we need to use \Q@tb@;
-In order modularly ensure well typedness,
-we require \Q@tb@ to be well typed at this stage. If \Q@tb@ was not well typed
-a compilation error could be generated at this stage.
-In this moment, \Q@A@ can not be compiled/checked alone,
-we need informations about \Q@B@, but \Q@A@ is not used in \Q@tb@,
-thus we do not need to type \Q@A@ and we can type \Q@tb@ with
- the available informations and proceed to generate \Q@B@.
-Now, we need to generate \Q@C@, and we need to ensure well typedness of \Q@tc@.
-Now \Q@B@ is alreay well typed (since generated by \use \Q@tb@, with no \mL),
-and \Q@A@ can be typed;  finally \Q@tc@ can be typed and used.
-If \use\ could not be performed (for example it \Q@tc@ had a method \Q@hello@ too)
-a compilation error could be generated at this stage.
-
-On the opposite side, if \Q@B@ and \Q@C@ was swapped, as in
-\begin{lstlisting}
-C: Use tc, {method int hello() 1}  
-B: Use tb
-\end{lstlisting}
-
-now the first task would be to generate \Q@C@, but 
-to type \Q@tc@ we need to know the type of \Q@A@ and \Q@B@.
-But they are both unavailable: \Q@B@ is still not computed and 
-\Q@A@ can not be compiled/checked alone, without information about \Q@B@.
-A compilation error would be generated, on the line of ``flattening of \Q@C@
-requires \Q@tc@, \Q@tc@ requires \Q@A@,\Q@B@, but \Q@B@ is still in need of flattening".
-
-In this example, a more expressive compilation/precompilation process 
-could compute a dependency graph and, if possible, reorganize the list,
-but for simplicity lets consider to always provide the declarations
-in the right order, if one exists.
-
-\paragraph*{Criticism: existence of an order is restrictive.}${}_{}$\\*
-Some may find the requirement of the existence of an order restrictive;
-An example of a ``morally correct" program where no right order exists is the following:
-\begin{lstlisting}
-t:{ int mt(A a) a.ma()}
-A:Use t {int ma() 1}
-\end{lstlisting}
-
-In a system without inference for method types,
-if the result of composition operators depends only on the
-structural shape of their input (as for \use)
-is indeed possible to optimistically compute the resulting structural shape of the classes
-and use this information to type involved examples like the former.
-We stick to our simple approach, since we believe such typing discipline would be fragile,
-and could make human understanding the code reuse process much harder/involved.
-Indeed we just wrote a program where the correctness of trait \Q@t@ depends of 
-\Q@A@, that is in turn generated using trait \Q@t@.
-
-\paragraph*{Criticism: it would be better to typecheck before flatteningi.}${}_{}$\\*
-In the world of strongly typed languages we could be tempted to
-first check that all can go well, and then perform the flattening.
-This would however be overcompicated for no observable difference:
-Indeed, in the \Q@A,B,C@ example above there is no difference
-between
-\begin{itemize}
-\item  (1)First check \Q@B@ and produce \Q@B@ code (that also contains \Q@B@ structural shape),
-  (2) then use \Q@B@ shape to check \Q@C@ and produce \Q@C@ code;\ 
-or a more involved
-\item  (1)First check \Q@B@ and discover just \Q@B@ structural shape as result of the checking,
-  (2)then use \Q@B@ shape to check \Q@C@.
-  (3) Finally produce both \Q@B@ and \Q@C@ code.
-\end{itemize}
-
-Note that we can reuse code only by naming traits; but our only point of relaxation is the class literal: there is no way an error can ``move around'' and be duplicated during the compilation process.
-In particular, our approach allows for safe libraries of traits and classes to be fully typechecked, deployed and reused by multiple clients: no type error will emerge from library code.
-On the other side, we do not enforce the programmer to write always self-contained code where all the abstract method definition are explicitly declared.
-
-\section{Managing State}
-This idea of summing pieces of code is very elegant,
-and has proven very successfull in module composition languages~\cite{ancona2002calculus} but our research community is struggling to
-make it work with object state (constructor and fields) while acheving the following goals:
+This idea of summing pieces of code is very elegant, and has proven
+very successfull in module composition
+languages~\cite{ancona2002calculus}\bruno{and several trait models?}. 
+However our research community is
+struggling to make it work with object state (constructor and fields)
+while acheving the following goals:
 
@@
 \item actually initialize objects, leaving no null fields,
-\item managing fields in a way that borrows the elengance of summing methods,
-\item make easy to add new fields.
+\item managing fields in a way that borrows the elegance of summing methods,
+\item make it easy to add new fields.
 \end{itemize}
@@
 
-Here in the following an exaple of such approach:
+Here in the following an example of such approach:
 

### 2017-05-20 FACS/section4.tex
.

@@
-\section{Improving Reuse when Subtyping is Impossible}
+\section{Allowing Reuse against Subtyping}
+Before we discussed how traits does not induce a new (externally visible) type.
+However, locally in the trait the programmer can use the special type \Q@This@ in order to denote the 
+type of \Q@this@.
+That is: the program is agnostic to what the \Q@This@ type is going to be, so it can
+be later assigned to any (or many) classes. 
+The idea is that during flattening, \Q@This@ will be replaced with the actual class name.
+In this way, \name allows reuse even when subtyping is
+impossible, that is, when a method parameter has type \Q@This@ (also called \emph{binary
+  methods}~\cite{}). 
+This type of situations is the primary motivator
+for previous work aiming at separating inheritance from subtyping.
+Leveraging on \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smootly integrate state and traits:
 
-\name retains the ability to allow reuse even when subtyping is
-impossible. That is, in some situations a subclass contains methods
-whose signatures are incompatible with the superclass, yet inheritance
-is still possible. This type of situations is the primary motivator
-for previous work aiming at separating inheritance from subtyping, and
-it happens, for example, in classes with \emph{binary
-  methods}~\cite{}.  Despite the adoption of a nominal approach and
-unlike other nominally typed languages, the leaking problem does not
-prevent \name from enjoying such kind of improved reuse.
+  %????Despite the adoption of a nominal approach and unlike other nominally typed languages, the leaking problem does not prevent \name from enjoying such kind of improved reuse.
 
-To illustrate how \name allows for improved reuse, we consider a
-pattern to combine \emph{multiple trait inheritance} and \emph{state}:
+%To illustrate how \name allows for improved reuse, we consider a
+%pattern to combine \emph{multiple trait inheritance} and \emph{state}:
 a challenging problem that has limited the flexibility of traits and
 reuse in the past.
+%First we show how self-instantiation and state are already avaible in the simplest trait model,
+%then we show how to radically improve the
 
@@
 
-\paragraph*{Advantages:} 
-this approach is associative and commutative, even self construction
-  can be allowed if the trait requires a static/class method
-  returning This; the class will then implement this method by forwarding
-  a call to the constructor.
-
-\paragraph*{Negatives:} writing the class code with the constructors and
-   fields and getter/setters and factories can be quite tedious.
-   Moreover, there is no way for a trait to specify a default value for a field,
-   the class need to handle all the state, even state that is conceptually
-   "private" of such trait.
-
-Here in the following an example of such approach:
-
 \begin{lstlisting}
-  pointSum: { method int x()  method int y()
-    class method This of(int x,int y)
+  pointSum: { method int x()  method int y()//getters
+    class method This of(int x,int y)//factory method
     method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
+      This.of(this.x+other.x,this.y+other.y)//self instantiation
     }
@@
     int x   int y
-    method int x()x   method int y()y
-    class method This of(int x, int y)
-      new Point(x,y)
+    method int x() x       
+    method int y() y
+    class method This of(int x, int y) new Point(x,y)
     constructor Point(int x, int y){ this.x=x   this.y=y }
@@
 
-\noindent after a while programming in this style, 
-writing those obvious ``close the state'' classes become a
-repetivie boring job, and one wonder
-if it could be possible to automatically generate such code~\cite{wang2016classless}.
-Indeed those classes are just a form of ``fixpoint''.
+\paragraph*{Advantages:} 
+this approach is associative and commutative, even self construction
+  can be allowed if the trait requires a static/class method
+  returning \Q@This@; the class will then implement this method by forwarding
+  a call to the constructor.
+
+\paragraph*{Negatives:} writing those obvious ``close the state/fixpoint'' class 
+   with the constructors and  fields and getter/setters and factories can be quite tedious.
+   %Moreover, there is no way for a trait to specify a default value for a field,
+   %the class need to handle all the state, even state that is conceptually
+   %"private" of such trait. 
+   Former work consider to automatically generate such code~\cite{wang2016classless}.
 
@@
 write down constructors and fields; there is not even syntax for those
-constructor.
+constructs.
 The idea is that any class that ``could" be completed in the obvious way
@@
 \item \Q@method This withX(int that)@
-may be a ``wither", doing a functional field update.
+may be a ``wither", doing a functional field update, that is create a new instance that is like \Q@this@ but where field \Q@x@ has now \Q@that@ value.
 \item \Q@method Void update(int x,int y)@
@@
 
-  lets play with the points of before, to see what good can we do with the current
-  instruments:
+With our approach to handle the state, the following code can be directly composed:
 

### 2017-05-20 FACS/Intro.tex
.

@@
 \input{section2}
+\input{section4}
 \input{section3}
-\input{section4}
-
-\section{Old: Self Types, Binary Methods and the Need to Separate
-  Inheritance from Subtyping}
-
-So, far it appears that we have not gained much from decoupling
-reuse from subtyping.
-
-As discussed in Section~\ref{} the type of the self-reference is the
-self-type. Like languages with ThisType, \name allows explicit
-references to the self type. This is useful in several situations: 
-For example it can be helpful in the presence of binary
-methods~\cite{}, or methods that return an instance of the 
-class being defined.

### 2017-05-20 FACS/section4.tex
.

@@
 \begin{itemize}
-\item keep sum associative and commutative,
-\item allowing a class to create instances of itself,
+%complicated discussions on this point \item keep sum associative and commutative,
+\item managing fields in a way that borrows the elegance of summing methods,
 \item actually initialize objects, leaving no null fields,
-\item managing fields in a way that borrows the elegance of summing methods,
-\item make it easy to add new fields.
+\item make it easy to add new fields,
+\item allowing a class to create instances of itself.
 \end{itemize}
@@
     class method This of(int x,int y)//factory method
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)//self instantiation
+    method This sum(This that)
+      This.of(this.x()+that.x(),this.y()+that.y())//self instantiation
     }
@@
     class method This of(int x,int y)
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
+    method This mul(This that)
+      This.of(this.x()*that.x(),this.y()*that.y())
     }
@@
 
+\paragraph{The solution in \name:}${}_{}$\\
 With our approach to handle the state, the following code can be directly composed:
@@
     class method This of(int x,int y)
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
+    method This sum(This that)
+      This.of(this.x()+that.x(),this.y()+that.y())
     }
@@
     class method This of(int x,int y)//the abstract method declarations.
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
+    method This mul(This that)
+      This.of(this.x()*that.x(),this.y()*that.y())
     }
@@
   As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods \Q@x@,\Q@y@ and \Q@of@.
+  as we wish. In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
   In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
@@
     }
-  pointSum:Use p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
+  pointSum:Use p, { method This sum(This that)
+      This.of(this.x()+that.x(),this.y()+that.y())
     }
-  pointMul:Use p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
+  pointMul:Use p, { method This mul(This that)
+      This.of(this.x()*that.x(),this.y()*that.y())
     }
@@
 
+\paragraph{Adding state:}${}_{}$\\
 What happens if we want to add fields instead of just operations?
@@
   colored:{ method Color color() }
-  Point:Sum pointSum,colored //fails
+  Point:Use pointSum,colored //Failure: resulting class not coherent
 \end{lstlisting}
@@
 where we assume to support overloading on different parameter number.
-This is a good solution, we think is better that any alternatives in literature,
-however the method \Q@CPoint.sum@ resets the color to red.
+This is a good solution, indeed we think is better that any alternatives in literature,
+however the method \Q@CPoint.sum@ resets the color to red: we call the \Q@of(int,int)@ method, that now delegate to \Q@of(int,int,Color)@ by passing red as the default field value.
 What should be the behaviour in this case?
-If we support withers, instead of writing \Q@This.of()@ we can use
+If our abstract state supports withers, instead of writing \Q@This.of(...)@ we can use
 \Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
-Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
+This solution is still not sadisfactory: this design ignore the color from \Q@that@.
 
-If the point designer could predict this kind of extension, then we could use the following design:  
+If the point designer could predict this kind of extension, then we could use the following better design:  
 \begin{lstlisting}
-  p: { method int x() method int y()
-    method This withX(int that)
-    method This withY(int that)
-    static method This of(int x,int y)
-    method This merge(This other)
+  p: { method int x() method int y() //getters
+    method This withX(int that) method This withY(int that)//withers
+    class method This of(int x,int y)
+    method This merge(This that) //new method merge!
     }
-  pointSum:Use p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y)
+  pointSum:Use p, { method This sum(This that)
+      this.merge(that).withX(this.x()+that.x()).withY(this.y()+that.y())
     }
@@
     method This withColor(Color that)
-    method This merge(This other)
+    method This merge(This that) //how to merge color handled here
       this.withColor(this.color().mix(that.color())
@@
   following this pattern.
+
   In order to compose, let say \Q@colored@ with \Q@flavored@ we would
@@
 This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
+
+In the following you can see how to mix colors and flavours. The syntax \Q@_2merge@ and   \Q@_3merge@
+call the version of merge as defined in the second/third element of \use.
+\begin{lstlisting}
+  p: {/*as before*/ }
+  pointSum:Use p, {/*as before*/ }
+  colored:{/*as before*/}
+  flavored:{method Flavor flavor() //very similar to colored
+    method This withFlavor(Flavor that)
+    method This merge(This that) //how to merge flavors handled here
+      this.withFlavor(that.flavor())}//here we just inherit ``that'' flavor
+  FCPoint:Use pointSum,colored,flavored{
+    class method This of(int x,int y)
+      This.of(x,y,Color.of(/*red*/),Flavor.none())
+    class method This of(int x, int y,Color color,Flavor flavor)
+    //we can resolve the conflict about two implementations for merge
+    //by proving our own implementation here
+    method This merge(This that) this._2merge(that)._3merge(that)
+    }
+\end{lstlisting}

### 2017-05-20 FACS/Intro.tex
.

@@
 programmer. Thus if subtyping is not desired, the idea is that 
-programmer can simply not declare a subtyping relationship.
+programmer can simply {\bf not} declare a subtyping relationship.
 
@@
 of most nominal OO languages subtyping implies subtyping in a
-fundamental way. This is because of what we call the self-reference
-\emph{leaking problem}. The leaking problem is illustrated by the following
+fundamental technical way. This is because of what we call the
+\emph{this leaking problem}, illustrated by the following
 (Java) code:
@@
 
-In class \lstinline{A}, the method \lstinline{m} passes \lstinline{this} as \lstinline{A}.
+Method \lstinline{A.ma} passes \lstinline{this} as \lstinline{A} to \Q@Util.m@.
 This code seems correct, and there is no subtyping/subclassing.
@@
 %%\lstinline{A}; and 2) it creates a {\bf subtype} of \lstinline{A}.
-We can see an invocation of \lstinline{ma} inside
-\lstinline{mb}, where the self-reference \lstinline{this} is of type \lstinline{B}. 
+We can see an invocation of \lstinline{A.ma} inside
+\lstinline{B.mb}, where the self-reference \lstinline{this} is of type \lstinline{B}. 
 The execution will eventually call \lstinline{Util.m} with an
@@
 
-If Java was to support a mechanism to allow reuse/inheritance 
-without introducing subtyping, such as:
+%If Java was to support a mechanism to allow reuse/inheritance 
+%without introducing subtyping, such as:
+%
+%\begin{lstlisting}[language=Java]
+%  class B inherits A{ int mb(){return this.ma();} }
+%\end{lstlisting}
+%
+%\noindent Then an invocation of 
+%\lstinline{mb} would be type-unsafe (i.e. it would 
+%result in a run-time type error). 
+%Note that here the intention of using the imaginary keyword {\bf
+%  inherits} is to allow the code from \lstinline{A} to be inherited 
+%without \lstinline{B} becoming a subtype of \lstinline{A}. 
+%However this breaks type-safety. The problem is that the
+%self-reference \lstinline{this} in class \lstinline{B} has 
+%type \lstinline{B}. Thus, when \lstinline{this} is passed as an argument to 
+%the method \lstinline{Utils.m} as a result of the invocation of
+%\lstinline{mb}, it will have a type that is incompatible with the
+%expected argument of type \lstinline{A}.
 
-\begin{lstlisting}[language=Java]
-  class B inherits A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
-\noindent Then an invocation of 
-\lstinline{mb} would be type-unsafe (i.e. it would 
-result in a run-time type error). 
-Note that here the intention of using the imaginary keyword {\bf
-  inherits} is to allow the code from \lstinline{A} to be inherited 
-without \lstinline{B} becoming a subtype of \lstinline{A}. 
-However this breaks type-safety. The problem is that the
+If in Java code reuse was not introducing subtyping an invocation of 
+\lstinline{B.mb} would result in a run-time type error.
+The problem is that the
 self-reference \lstinline{this} in class \lstinline{B} has 
@@
 expected argument of type \lstinline{A}.
-\bruno{
-We can also mention what happens in structural types:
-"System with structural types would *be required to* guarantee that
-the structural type of A is a supertype of the structural type of B”.
-}
 
-What the leaking problem illustrates is that
-adopting a more flexible
-nominally typed
-OO model where subclassing does not imply subtyping is not trivial. 
-If we are to separate subclassing from subtyping then a more substantial
-change in the language design is necessary.  Every OO language with
-the minimal features exposed in the example (using \lstinline{this},
+Every OO language with the minimal features exposed in the example (using \lstinline{this},
 extends and method calls) is forced to accept that subclassing implies
@@
   "privately extends" \lstinline{A}.}
+
 In essence we believe that a problem with the design of classes in languages like Java
@@
 both as units of \emph{use} and \emph{reuse}.
-That is, in languages like Java, classes can be
-instantiated (using {\bf new}) to create objects that \emph{use} some 
-functionality. Classes can also be subclassed to provide \emph{reuse}
+That is, in languages like Java, class names can be
+\emph{use}d as type and instantiated.
+Classes can also be subclassed to provide \emph{reuse}
 of code.
 
-\marco{Cite some work of bruceTHisType and show how he also fails.}
-
-\begin{comment}
-An iconic example on why connecting inheritance/code reuse and
-subtpying is problematic is provided by the
-historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
-implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
-keeping tracks of how many times an element occurs.  We would just add
-some state and override a couple of methods.\bruno{Are we going to
-  present this example in the paper solved in 42, for example? 
-I think I would expect to see it.}
-
-However, our subclassing would break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
-Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
-and making both \Q@Bag@ and \Q@Set@ inherit from it.
-This looks unnatural, since \Q@Set@ would extend it without adding anything,
-and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
-Worst, if we was to constantly apply this mentalty, we would introduce a very high number
-of abstract classes that are not supposed to be used as types, and that will clutter the 
-public interface of our classes and our code project as a whole.
-\end{comment}
+Adopting a more flexible nominally typed OO model where subclassing does not imply subtyping is not trivial,
+and a more substantial change in the language design is necessary.
 
@@
 cannot be instantiated (or used) directly. Such design allows the
-subtyping and code reuse to be treated separately, which in terms 
+subtyping and code reuse to be treated separately, which in turn
 brings several benefits in terms of flexibility and code reuse.
-\marco{We need to talk of unanticipated extensions?}
-\bruno{Talk more about typing aspects here. Summarize the important 
-aspects of the design of \name.}
+%\marco{We need to talk of unanticipated extensions?}
+%\bruno{Talk more about typing aspects here. Summarize the important 
+%aspects of the design of \name.}
 

### 2017-05-20 FACS/RelatedWork.tex
new abstract
.

@@
 Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}
+
+Add that: original trait model has no self construction and types introduction not relevant since dynamicly typed language
+
+\bruno{
+We can also mention what happens in structural types:
+"System with structural types would *be required to* guarantee that
+the structural type of A is a supertype of the structural type of B”.
+}
+
+\marco{Cite some work of bruceTHisType and show how he also fails to separe reuse and use}

### 2017-05-20 FACS/Examples.tex
Intro
Section 2
Some more work on Sections 2 3 and 4
.

@@
-\section{A Decoupled Trait Calculus}
-In order to illustrate how we plan to properly divide code reuse/inheritance/subclassing from subtyping,
-lets consider a simple language where trait names start with a lowercase \mt
-and class names start with an uppercase \mC.  
-The syntax of the language is:
+\section{Intuitions on formalization}
+In this article we dedicate more space to examples and informal presentation and motivations;
+so we do not have space to provide a full formalizations.
+We will provide here some hints on how the formalization works.
 
-\begin{comment}
+In the following, we present a very simplified grammar:
+%\begin{bnf}
+%\prodFull{aa}{bb}{Declaration}\\
+%\end{bnf}
+
+%\begin{comment}
 \begin{bnf}
@@
 \end{bnf}
-\end{comment}
+%\end{comment}
 
@@
 expression.  Traits come with various operators (restrict, hide,
-alias) but for now we focus on the single operator 
+alias) but in this work we focus on the single operator 
 $\use$, taking a set
-of code values: that is trait names \mt\ or literals \mL\ and composing them in an \emph{associative}
-and \emph{commutative} way.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
-composition operator.  $\use\ \overline\mV$ composes the content of $\overline\mV$
+of code values: that is trait names \mt\ or literals \mL\ and composing them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
+trait composition operator.  $\use\ \overline\mV$ composes the content of $\overline\mV$
 by taking the union of the methods and the union of the implementations.
-In our simple mode\bruno{you mean model?}, we consider an error trying to merge a class and an interface.
 
-Use can not be applied if multiple versions of the same method are
+\use can not be applied if multiple versions of the same method are
 present in different traits.  An exception is done for abstract methods:
@@
 of two abstract methods with compatible headers, the one with the more
-specific type is selected. We will discuss later the formal details of
-\use.\marco{actually, this makes the sum not strongly associative...}
-\bruno{ok: would be nice to see an example?}
+specific type is selected.
 
-Code literals \mL\ can be marked as interfaces.  We will call class
-interface, or simply interface a class declaration of for \Q@C:{interface  ...}@.
+Code literals \mL\ can be marked as interfaces. 
+That is, the interface keyword is inside the curly brakets, so an upper case name associated with an interface literal is a class-interface, while a lowercase one is a trait-interface.
+In our simple model, we consider an error trying to merge an interface with a non-interface.
  Then we have a set of implemented interfaces and a set of member
@@
 \item all methods in a code literal have unique names,
-\item all used variables are in scope.
+\item all used variables are in scope,
+\item all methods in an interface are abstract, and there are no interface class methods.
 \end{itemize}
@@
 \item only interfaces can be implemented.
-\item \bruno{method conflict detection in traits?}
 \item \marco{I'm sure I'm missing something}
 \end{itemize}
-We will see later the details on when exactly in the compilation process any code literal is typechecked.
 While classes are typed assuming \lstinline{this} is of the nominal type of the
@@
 A Literal can use the \lstinline{This} type,
-and when the trait is reused to create a class, \Q@This@ will become just
-an alias for the class name.
+and when flattening completes creating a class definition, \Q@This@ will be replaced with such class name.
 
@@
 \me: they can be just variables and method calls.
-We do not think that there is any problem to extend our language with
-other constructs~\bruno{don't you actually have evidence that this is
-  the case from 42?}. We use minimal expressions to keep our example
-simple and compact. In order to make our examples more engaging, we
-feel free to use the type \lstinline{int}, numeric literals and their operations.
 
+\begin{comment}
 \subsection{Code Use and Code Reuse in \name}
@@
 
-This code is fine, but there is no way to add a \Q@B@ reusing code of \Q@A@, since
+This code is fine, but there is no way to add a class \Q@B@ reusing
+the code of \Q@A@, since
 \Q@A@ is designed for code \emph{use} and not \emph{reuse}. Lets try
@@
 In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.
+\end{comment}
 
@@
 what distinguishes our approach from a simple minded code composition macro~\cite{bawden1999quasiquotation}
-or a rigid module composition~\cite{ancona2002calculus}
+or a rigid module composition~\cite{ancona2002calculus}.
 
 The are two core ideas:
-\paragraph*{traits are \emph{well-typed} before being reused.}${}_{}$\\*
+\paragraph*{1: traits are \emph{well-typed} before being reused.}${}_{}$\\*
  For example in
@@
 
-\noindent Note how in the former example \Q@ta@ is well typed just by
-knowing itself and \Q@IA@.\bruno{why IA? it is not used in this
-  example, right?}
-
-\paragraph*{class expressions are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
+\paragraph*{2: code literals are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
 In class expressions  $\use\ \overline\mV$
@@
 \noindent in \Q@B@ we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for \Q@ma()@.
-In our example,using the trait \Q@t@ of before
+In our example, using the trait \Q@t@ of before
 
@@
 literals have to be self complete. In this case we would have been forced to
-declare abstract methods \Q@n@ and \Q@m@.\bruno{right. this is a good point}
+declare abstract methods \Q@n@ and \Q@m@.
 
@@
 then the issue was originally present in one of $\mL_1\ldots\mL_n$.
-It may be that the result is intrinsically ill-typed, if of the methods in $\mL_1\ldots\mL_n$ is not well typed,
-but it may also happen that a type referred from one of those methods is declared \emph{after} the current class. As we will see later, this is how our relaxation allows to support recursive types.
+It may be that the result is intrinsically ill-typed, if one of the methods in $\mL_1\ldots\mL_n$ is not well typed,
+but it may also happen that a type referred from one of those methods is declared \emph{after} the current class.
+As we will see later, this is how our relaxation allows to support (indirectly) recursive types.
 
@@
 
-We will see now how this idea that code literals wrote in the declaration of a class
-do not have to be well typed on their own also lets us reason about mutually recursive types.
-
- \section{Recursive types}
+ \subsection{Recursive types}
 
@@
 happen one at the time, and classes are typechecked where their type is first needed.
+That is, In our approach typing and flattening interleaves. We assume our compilation process to stop as soon as 
+an error arise. There are two main kinds of errors: Type errors (like method not found) or Composition errors (like summing two conflicting implementation for the same method).
 
@@
 To generate \Q@B@, we need to use \Q@tb@;
-In order modularly ensure well typedness,
+In order to modularly ensure well typedness,
 we require \Q@tb@ to be well typed at this stage. If \Q@tb@ was not well typed
-a compilation error could be generated at this stage.
+a type error could be generated at this stage.
 In this moment, \Q@A@ can not be compiled/checked alone,
@@
 If \use\ could not be performed (for example it \Q@tc@ had a method \Q@hello@ too)
-a compilation error could be generated at this stage.
+a composition error could be generated at this stage.
 
@@
 \Q@A@ can not be compiled/checked alone, without information about \Q@B@.
-A compilation error would be generated, on the line of ``flattening of \Q@C@
+A type error would be generated, on the line of ``flattening of \Q@C@
 requires \Q@tc@, \Q@tc@ requires \Q@A@,\Q@B@, but \Q@B@ is still in need of flattening".
@@
 We stick to our simple approach, since we believe such typing discipline would be fragile,
-and could make human understanding the code reuse process much harder/involved.
-Indeed we just wrote a program where the correctness of trait \Q@t@ depends of 
+and could make human understanding the code-reuse process much harder/involved.
+Indeed we just wrote an involved program where the correctness of trait \Q@t@ depends of 
 \Q@A@, that is in turn generated using trait \Q@t@.
 
-\paragraph*{Criticism: it would be better to typecheck before flatteningi.}${}_{}$\\*
+\paragraph*{Criticism: it would be better to typecheck before flattening.}${}_{}$\\*
 In the world of strongly typed languages we could be tempted to
@@
 
+\begin{comment}
 \section{Managing State}
@@
 This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
+\end{comment}

### 2017-05-20 FACS/RelatedWork.tex
.

@@
+\section{Extensions and practical experience}
+The model presented here is very easy to extend:
+For example nested classes can be easly added,giving us enough power to easly solve the expression problem as in \cite{deepfjig}. More composition operators can be added in addition to use. Indeed
+we just need to be sure that every newly added operator respects the following criteria:
+
+\begin{itemize}
+\item As for \use, the operato does not need to be total, but if it fails it needs to provide an error that will be reported to the programmer.
+\item When the operator takes in input only traits (they are going to be well typed), if a result is produced,
+ such result is well typed.
+\item When the operator takes in input also code literal, if a non well typed result is produced,
+the type error must be tracked back to code in one of those non typed yet code literals.
+ \end{itemize}
+ That is, our model can easly support all the operators of \cite{deep}.
+
+ Our simplified model represent the conceptual core of  a novel programming language (http://L42.is)
+ we are currently working at.
+ L42 contains the extensions described before and much more, and provide a concrete example of the usability of our model.
+
+ The main difference is that code literal are first class values in L42, thus we do not need traits as a language concept, but just as a method returning such code literal.
+
 \section{Related Work}
@@
 that are provided only by interfaces.
-In DeepFjig[] classes introduce nominal types but in an innatural way, the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class.
+In DeepFjig[] classes introduce nominal types but in an innatural way, the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class: that is in DeepFjig 
+\Q@A:{ method A m()this}@ is not well typed.
 
@@
 Our compilation model is a clear step forward with respect to 
-\cite[metaFjig, deepFjig,myThesis], where all code was required to declare all the abstract dependencies all of the time.
+\cite[metaFjig, deepFjig,myThesis,allferruccioWorkBefore2017], where all code was required to declare all the abstract dependencies all of the time.
 

### 2017-05-21 FACS/RelatedWork.tex
.

@@
 and multiple inheritance~\cite{}.
+
+We are aware of at least 3 independently designed research languages 
+that address this limitation:Delta-Trait(DT)\cite{3ferruccio works}, Package Template(PT)\cite{}, DeepFjig\cite{}.
+Leveraging on \emph{traits}, in this work we aim to synthesizes
+the main ideas of those very different language designs.
+
+\begin{itemize}
+\item deep+ {\bf A simple uniform syntax for code literals}
+DeepFjig is best in this sense, since DT have separate syntax for class literals, trait literals and record literals.
+PT on the other side build on top of full Java, thus has a very involved syntax.
+Thanks to our novel represantation of state, our approach offers a much simpler and uniform syntax then
+all others, where evertyhing is basically just a method.
+\item deep-
+{\bf Reusable code can not be ``used'', that is instantiated or used as a type.}
+This happens in DT and in PT, but not in DeepFjig. To allow reusable code to be directly 
+usable, in DeepFjig
+classes introduce nominal types in an innatural way: the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class: that is in DeepFjig 
+\Q@A:{ method A m()this}@ is not well typed. This is because
+\Q@B: Use A@ flattens to \Q@B:{ method A m()this}@, that is clearly not well typed.
+Looking to this example is clear why we need reusable code to be agnostic of its name.
+Then, either reusable code have no name (DT, PT and this work)
+or all code is reusable and usable, and all code need to be auckwardly agnostic of its name.
+
+\item PT+
+{\bf Requiring abstract signatures is a left over of module composition mindset.}
+DT and Deepfjig comes form a tradition of functional module composition, where 
+modules are typed in isolation under an environments, and then the composition is performed.
+As we show in this work, this is ends up requiring verbose repetition of abstract signatures, 
+that for hightly modularized code may end up composing most of the program.
+Simple Java (and thus PT, since it is an add on over Java) show us a better way:
+the meaning of names can be understood from the reuse context.
+The typing strategy of PT offers the same advantages of our typing model, 
+but is more involved and indirect. This may be caused by the
+heavy task of integrating with full Java.
+\item PT-{\bf Composition algebra.}
+The idea of using composition operators over atomic values as in an arithmetic expression is very powerfull,
+and makes easy to extend languages with more operators. Deepfjig and DT embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
+The typing strategy of PT also seams to be connected with this decision, so it would be hard to move their approach in a composition algebra setting.
+\item DT- {\bf Naming your type, even if you have none.}
+Both Deepfjig and PT allows a class to refer to its name, abait is more unobvious in PT and is unclear 
+...
+\item DT+ real separation use-reuse
+
+\end{itemize}
 

### 2017-05-21 FACS/RelatedWork.tex
first hit to the related work

@@
 and multiple inheritance~\cite{}.
+
+\subsection{Inheritance and subtyping}
 
@@
 The typing strategy of PT also seams to be connected with this decision, so it would be hard to move their approach in a composition algebra setting.
-\item DT- {\bf Naming your type, even if you have none.}
-Both Deepfjig and PT allows a class to refer to its name, abait is more unobvious in PT and is unclear 
-...
-\item DT+ real separation use-reuse
+\item DT- {\bf Naming your type, even if you have none yet.}
+Both Deepfjig and PT allows a class to refer to its name, abait is more unobvious in PT since you have to introduce both a package and a classes to express it.
+This allows to encode binary methods, express patterns like withers or fluent setters and to instantiate instances of the (future) class(es)  using the reused code.
+\item DT+{\bf Complete ontological separation between use and reuse}
+While all 3 works allows to separate inheritance and subtyping only DT properly enforce 
+separation between use (classes and interfaces) and reuse (traits).
+This is because in DeepFjig all classes are both units of use and reuse (however, subtype is not induced)
+and PT imports all the complexity of Java, so abait is possible to separate use and reuse, the model have powerfull but unobvous implications where (conventional Java) \Q@extends@ and PT are used together
 
 \end{itemize}
-
-Our idea of fully separate inheritance and subtyping has some precedence
-In all the line of work of Ferruccio [cite many] neither traits nor classes introduce nominal types,
-that are provided only by interfaces.
-In DeepFjig[] classes introduce nominal types but in an innatural way, the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class: that is in DeepFjig 
-\Q@A:{ method A m()this}@ is not well typed.
-
-Other work cite~\cite{odersky2008programming,nystrom2006j}
-do no completly break the relation between inheritance and subtyping, but only prevent subtyping where 
+With the exception of those 3 lines of work, to the best of our understading
+other famous work in literature, like~\cite{odersky2008programming,nystrom2006j}
+does no completly break the relation between inheritance and subtyping, but only prevent subtyping only where 
 it would be unsound.
 
-Our compilation model is a clear step forward with respect to 
-\cite[metaFjig, deepFjig,myThesis,allferruccioWorkBefore2017], where all code was required to declare all the abstract dependencies all of the time.
+\subsection{State and traits}
 
-%\item clean and elegant handling of state in trat/module composition language
 The idea of abstract state operation has emerged from CJ~\cite{wang2016classless}, and offer a clean solution to handling state
@@
 \item No constructor: all the fields start at null/a default specified value.
-  In this model fields are like another kind of member, and two fields
-  with identical type can be merged by sum; \Q@new C()@ can be used for all classes, and \Q@init@ method may be called later, as in
+  Fields are just like another kind of (abstract) member, and two fields
+  with identical type can be merged by sum/use; \Q@new C()@ can be used for all classes, and \Q@init@ methods may be called later, as in
   \Q@Point p=new Point(); p.init(10,30)@.
@@
 \item Constructor compose fields:
-In this approach the fields are declared but not initialized, and
-a canonical constrctor constructor taking a value for each field and just initializing such field
+In this approach (used by \cite{FJig}) the fields are declared but not initialized, and
+a canonical constrctor constructor (as in FJ) taking a value for each field and just initializing such field
 is automatically generated in the resulting class.
 
-In this approach it is easy to add fields, however this approach is associative but not commutative, since the order of composition will change the order of the fields; moreover there is no support for self construction
+It is easy to add fields, however this model is associative but not commutative: composition order influence field order, and thus the construtor signature.
+Self construction is not possible 
 since the signature of the constuctor change during composition.
@@
 \item Constructor can be composed if they offers the same exact parameters:
-In this approach traits declare field and constructors.
+In this approach (used by DeepFJig) traits declare field and constructors.
 The constructor initialize the fields but can do any other computation.
@@
 This approach is designed to allows Self construction.
-It is also associative and mostly commutative: the order of composition is relevant only
-in the order of execution of side effects during construction.
-However the requirement that the constructors need the same parameters 
+It is also associative and mostly commutative: composition order only influence execution order of side effects during construction.
+However constructors composition requires identical constructor signature: this
 hamper reuse, and if a field is added, its initial value need to be
-somehow syntetized from the old constructor parameters.
+somehow syntetized from the constructor parameters.
 
@@
 %\item extension with nested class is natural and powefull
-We claim that our solution to the expression problem is the most natural in literature to date:
-While a similar sintax can be achieved with the scandinavian style~\cite{ernst2004expression}, their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system~\cite{igarashi2005lightweight}.
+%We claim that our solution to the expression problem is the most natural in literature to date:
+%While a similar sintax can be achieved with the scandinavian style~\cite{ernst2004expression}, their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system~\cite{igarashi2005lightweight}.
 
-Our close contendant is DeepFJig~\cite{corradi2011deepfjig}: all our gain over their model is based on our relaxation over abstract signatures.
-Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}
+%Our close contendant is DeepFJig~\cite{corradi2011deepfjig}: all our gain over their model is based on our relaxation over abstract signatures.
+%Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}
 
-Add that: original trait model has no self construction and types introduction not relevant since dynamicly typed language
+%Add that: original trait model has no self construction and types introduction not relevant since dynamicly typed language
+
+\subsection{Tablular comparision of many approaches}
+\begin{minipage}[t]{0.4\textwidth}
+In this table we show if some approach or part-there-of support certain critical features:
+Direct instantation (as in \Q@new C()@),
+Self instantiation (as in \Q@new This()@),
+Unit of use, Unit of reuse,
+Introduce type and if the Induced type is the type of \Q@this@,
+support for binary methods,
+does inheritance induce subtype?,
+is code required to be well-typed before being inherited /imported in a new context?
+is code required to be well-typed before composed with other code?
+
+We use Y and X to mean yes and no, and we use ``-'' were the question is not really applicable to the current approach.
+For example the original trait model was untyped, so typing questions makes little sense.
+\end{minipage}
+%second column
+\begin{minipage}[t]{0.6\textwidth}
+${}_{}$
+\begin{center}
+\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c}
+&\Rotated{direct instantation}
+&\Rotated{self instantiation}
+&\Rotated{unit of use}
+&\Rotated{unit of reuse}
+&\Rotated{introduce type}
+&\Rotated{induced type is this type}
+&\Rotated{binary methods}
+&\Rotated{inheritance induce subtype}
+&\Rotated{well-typed before imported}
+&\Rotated{well-typed before composed} 
+\\
+\hline
+java/scala class&Y&X&Y&Y&Y&Y&X&Y&Y&X\\
+java8 interface &X&X&X&Y&Y&Y&X&Y&Y&X\\
+scala trait        &X&X&X&Y&Y&Y&X&Y&X&X\\
+original trait     &X&X&X&Y&-&-&X&X&-&-\\
+Ferruccio trait  &X&X&X&Y&X&-&X&X&Y&Y\\
+\name trait            &X&Y&X&Y&X&-&Y&X&Y&X\\
+\name class           &Y&Y&Y&X&Y&Y&Y&-&Y&X\\
+module composition
+                      &-&-&Y&Y&-&-&-&-&Y&Y\\
+deepFJig class &Y&Y&Y&Y&Y&X&Y&X&Y&Y\\
+package template
+                      &X&Y&X&Y&X&-&X&X&Y&X\\
+
+\end{tabular}
+\end{center}
+\end{minipage}
 

### 2017-05-21 FACS/section2.tex
first hit to the related work

@@
 %not.
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{c|c|c|c}
-              & Instantiable (Use) & Unit of Reuse & Introduces New Type \\
-\hline
-Class  & Yes              & No          & Yes           \\
-Trait   & No               & Yes         & No            \\
-\end{tabular}
-\end{center}
-More complete table here, we can put it where you prefer
-\begin{center}
-\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c}
-&\Rotated{direct instantation}
-&\Rotated{self instantiation}
-&\Rotated{unit of use}
-&\Rotated{unit of reuse}
-&\Rotated{introduce type}
-&\Rotated{induced type is this type}
-&\Rotated{binary methods}
-&\Rotated{reuse induce subtype}
-&\Rotated{well-typed before imported}
-&\Rotated{well-typed before composed} 
-\\
-\hline
-java/scala class&3Y&X&Y&Y&Y&Y&X&Y&Y&X\\
-java8 interface &X&X&X&Y&Y&Y&X&Y&Y&X\\
-scala trait        &X&X&X&Y&Y&Y&X&Y&X&X\\
-original trait     &X&X&X&Y&-&-&X&X&-&-\\
-Ferruccio trait  &X&X&X&Y&X&-&X&X&Y&Y\\
-42 trait            &X&Y&X&Y&X&-&Y&X&Y&X\\
-42 class           &Y&Y&Y&X&Y&Y&Y&-&Y&X\\
-module composition
-                      &-&-&Y&Y&-&-&-&-&Y&Y\\
-deepFJig class &Y&Y&Y&Y&Y&X&Y&X&Y&Y\\
-package template
-                      &X&Y&X&Y&X&-&X&X&Y&X\\
-
-\end{tabular}
-\end{center}
-
-\caption{A comparision between traits and classes in \name.}
-\label{fig:compare}
-\end{figure}

### 2017-05-22 FACS/Examples.tex
Revising Initial sections

@@
-\section{Intuitions on formalization}
+\section{Intuitions on formalization}\label{sec:formal}
 In this article we dedicate more space to examples and informal presentation and motivations;

### 2017-05-22 FACS/Intro.tex
.
Revising Initial sections

@@
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 
@@
 are incompatible with the superclass, yet inheritance is still
-possible. A typical example, which was illustrated by Cook et al., are 
-classes with \emph{binary methods}~\cite{}.
+possible. A typical example, which was illustrated by Cook et al.~\cite{cook}, are 
+classes with \emph{binary methods}~\cite{bruce96binary}.
 
@@
   the typing point of view, subtyping is still possible. A typical
-  example~\cite{} of this are methods for collection classes such as Sets and
+  example~\cite{LaLonde:1991:SSS:110673.110679} of this are methods for collection classes such as Sets and
   Bags. Bag implementations can often inherit 
@@
 of most nominal OO languages subtyping implies subtyping in a
-fundamental technical way. This is because of what we call the
-\emph{this leaking problem}, illustrated by the following
+fundamental way. This is because of what we call the
+\emph{this-leaking problem}, illustrated by the following
 (Java) code:
@@
 
-If in Java code reuse was not introducing subtyping an invocation of 
+\bruno{You removed my lengthier explanation. I think this makes things
+less clear.}
+If in Java code reuse was not introducing subtyping, then an invocation of 
 \lstinline{B.mb} would result in a run-time type error.
@@
 extends and method calls) is forced to accept that subclassing implies
-subtyping.\footnote{ Note how this is true also in C++, where is
+subtyping. Note how this is true also in C++, where is
   possible to "extends privately". Such is a limitation of the
@@
   former example would be accepted by C++ even if \lstinline{B} was to
-  "privately extends" \lstinline{A}.}
+  "privately extends" \lstinline{A}.
 
-In essence we believe that a problem with the design of classes in languages like Java
-is that classes do too many things at once. In particular they act
-both as units of \emph{use} and \emph{reuse}.
-That is, in languages like Java, class names can be
-\emph{use}d as type and instantiated.
-Classes can also be subclassed to provide \emph{reuse}
-of code.
+In essence we believe that a problem with the design of classes in
+languages like Java is that classes do too many things at once. In
+particular they act both as units of \emph{use} and \emph{reuse}.
+That is, in languages like Java, class names can be \emph{use}d as
+type and instantiated.  Classes can also be subclassed to provide
+\emph{reuse} of code.
 
-Adopting a more flexible nominally typed OO model where subclassing does not imply subtyping is not trivial,
-and a more substantial change in the language design is necessary.
+What the this-leaking problem shows is that 
+adopting a more flexible nominally typed OO model where subclassing
+does not imply subtyping is not trivial, and a more substantial change
+in the language design is necessary. 
+We are aware of at least 3 independently designed research languages 
+that address this limitation: Delta-Trait~\cite{Damiani2014}, Package Templates~\cite{KrogdahlMS09}, DeepFjig~\cite{deep}.
+\bruno{Marco, please expand this a bit. Check if references are correct!}
 
@@
 In \name there are two separate concepts: classes
-and traits~\cite{}. Classes are meant for code use, and cannot be used
+and traits~\cite{Traits:ECOOP2003}. Classes are meant for code use, and cannot be used
 for reuse. In some sense classes in \name are like final classes in
@@
 
+\bruno{Need to say something about state?}
+
+The language design of \name is adopted by the 42 programming 
+language, All the examples shown in this paper can be run in 42.
+The compiler for 42 is available at:
+
+\url{http://l42.is}
+
+\begin{comment}
 First we show a minimal language that illustrates the basic ideas. 
@@
 code reuse.
+\end{comment}
 
-\subsection{List of contribution}
-\marco{text to improve, but is nice to keep a list}
+\subsection{Contributions:}
 \begin{itemize}
+\item We identify the {\bf this-leaking} problem and argue why it
+  makes the separation of subclassing and subtyping difficult.
+\item Inspired by previous language designs that solve the
+  this-leaking problem, we synthesize the key ideas in {\bf a novel and
+  simple language design}. 
+\item We illustrate how the new design {\bf improves both code use and code
+  reuse}.
+\item We propose {\bf a clean and elegant approach to handle of state} in a
+  languages that support traits.
+\item The new language design is {\bf adopted by the 42 language}. All the
+  examples in the paper are valid 42 programs.
+
+\begin{comment}
 \item simple approach to fully separate inheritance and subtyping
@@
     result on not saying enough about each individual ideas.}
-\item more operators can be accomodated without changing the general model
+\item more operators can be accomodated without changing the general
+  model
+\end{comment}
 \end{itemize}
-\input{section2}
-\input{section4}
-\input{section3}

### 2017-05-22 FACS/paper.tex
.
Revising Initial sections

@@
 \usepackage{comment}
+\usepackage{rotating} 
 \lstset{language=Java,
@@
 }
-
+%
+\newcommand\Rotated[1]{\begin{turn}{90}\begin{minipage}{12em}#1\end{minipage}\end{turn}}
+%
 %marco
@@
 
-\newcommand\name{{\bf DTrait}\xspace}
+\newcommand\name{{\bf $42_{\mu}$}\xspace}
 
@@
 seems required for technical reasons, due to what we call the
-\emph{this-leaking problem}. The leaking problem shows that separating
+\emph{this-leaking problem}. The this-leaking problem shows that separating
 code-reuse from subtyping is non-trivial and requires a significant
@@
 that address this limitation:Delta-Trait, Package Template, DeepFjig.
-Leveraging on \emph{traits}, in this work we aim to synthesizes
+Leveraging on \emph{traits}, in this work we aim to synthesize
 the main ideas of those very different language designs.
@@
 we can separate inheritance and subtyping, while avoiding 
-redundant abstract declarations (required in Delta-Trait and DeepFjig), while supporting self construction,
-binary methods and recursive types.
+redundant abstract declarations (required in Delta-Trait and
+DeepFjig). At the same time self construction,
+binary methods and recursive types are also supported.
 Moreover, we provide a novel and elegant solution to uniformly
 handle behavior and state within trait composition.
+These ideas have been implemented in the 42 language, which we use 
+throughout the paper for code examples.\bruno{Is it true that all the code
+examples are valid 42?}
 
@@
 \input{Intro}
+\input{section2}
+\input{section3}
+\input{section4}
 \input{Examples}
-\input{Extensions}
+%\input{Extensions}
 \input{RelatedWork}
@@
 \section*{Bibliography}
-\bibliographystyle{elsart-num-sort}
+\bibliographystyle{plain}
 \bibliography{main}

### 2017-05-22 FACS/section2.tex
Revising Initial sections

@@
 key ideas of its design. In particular we illustrate how to separate code use and 
-code reuse.
+code reuse, and how \name solves the this-leaking problem.
 % in \name. %, and how this allows improving both.
@@
 
-To illustrate classes in \name consider the example in Section~\ref{}:
+To illustrate classes in \name consider the example in Section~\ref{sec:intro}:
 
@@
 \noindent Classes in \name use a slightly different declaration style compared
-to Java: there is no \lstinline{class} keyword, and a colon separe the class name (which must always start with
+to Java: there is no \lstinline{class} keyword, and a colon separates the class name (which must always start with
 an uppercase letter) and the class implementation, which is used to specify the
@@
 start with a lowercase letter. An obvious first attempt 
-to model the example in Section~\ref{} (for a Java programmer) 
+to model the example in Section~\ref{sec:intro} 
 with traits and code reuse is:
@@
 \end{lstlisting}
-But this does not work either, indeed \Q@ta@ is not a type in the first place, since it is a (lowercase) trait name.
+But this does not work either: \Q@ta@ is not a type in the first place, since it is a (lowercase) trait name.
 Indeed
@@
 
-With this in mind, we can try to model the example in Section~\ref{} again:
+With this in mind, we can try to model the example again:
 
@@
  IA:{interface method int ma()}//interface with abstract method
- Utils:{ class method int m(IA a)/*method body here*/}
+ Utils:{class method int m(IA a)/*method body here*/}
  ta:{implements IA
@@
 
-This code works: \Q@Utils@ rely over interface \Q@IA@ and the trait \Q@ta@
+This code works: \Q@Utils@ relies on interface \Q@IA@ and the trait \Q@ta@
 implements \Q@IA@.
-\Q@ta@ is well typed: independently of what class name(s) will be associated to its code, we knew such class(es) will implement \Q@IA@, thus while typechecking \Q@Utils.m(this)@ we can assume
+\Q@ta@ is well typed: independently of what class name(s) will be
+associated to its code, we know that such class(es) will implement
+\Q@IA@. 
+Therefore, while typechecking \Q@Utils.m(this)@ we can assume
 \Q@this<:IA@.
@@
 
-Later, in section \ref{} we will provide a more formal detail over type system. For now notice that in the former example the code is correct even if no method called \Q@ma@ is explicitly declared.
-The idea is that such method is imported from trait \Q@ta@, exactly as in the java equivalent
+Later, in Section \ref{sec:formal} we will provide more details on the type
+system. 
+For now notice that in the former example the code is correct even if
+no method called \Q@ma@ is explicitly declared\bruno{but it would not
+  work if the \lstinline{ta} was not in the use list, right?}.
+The idea is that such method is imported from trait \Q@ta@, exactly as in the Java equivalent
 \begin{lstlisting}[language=Java]
@@
 \end{lstlisting}
-method \Q@ma@ is imported from \Q@A@.
-This concept is natural for a Java programmer, but was not supported in previous work \cite{deep, ferruccio}
+where method \Q@ma@ is imported from \Q@A@.
+This concept is natural for a Java programmer, but was not supported
+in previous work \cite{Damiani2014,deep}. Those works require all
+dependencies in code literals to be explicitly declared, so that the
+code literal is completely self-contained. However, this results in
+many spurious abstract method declarations.\bruno{Improved text. 
+Please double-check! You may actually show how the code with the extra 
+explicit declarations would look like! 
+}
 
 \paragraph*{Semantic of Use}
-Albeit alternative semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
+Albeit alternative semantic models for traits~\cite{Traits:ECOOP2003} have been proposed,
+here we use the flattening model. This means that 
 \begin{lstlisting}
@@
 
- This code seems correct, and there is no mention of the trait
+ This code is correct, and there is no mention of the trait
  \Q@ta@. In some sense, all the information about code-reuse/inheritance
@@
  and \Q@B@; while subtyping is part of the class interface.
+\bruno{I don't think we are making  a strong point here, and this
+  discussion seems inconsequent at this point!}
 

### 2017-05-22 FACS/section3.tex
Revising Initial sections

@@
-\section{Improving Use by Preventing Unintended Subtyping}
+\section{Improving Use}
 
+This section illustrates how 
 %As stated in Section~\ref{} 
-To have code-reuse, while preventing unintended subtyping
-in a nominally typed OO language we need to
- separates inheritance from subtyping.
-
-Thus, \name  \emph{improves the use} of classes, while
+%To have code-reuse, while preventing unintended subtyping
+%in a nominally typed OO language we need to
+% separate inheritance from subtyping.
+\name \emph{improves the use} of classes, while
 retaining the benefits of code-reuse when compared to Java-like languages. 
 Here we illustrate this benefit by modelling a simplified version of
-Set and Bag collections.
+Set and Bag collections first in Java, and then in \name.
+The main benefit of the \name solution is that we can get reuse 
+without introducing subtyping between Bags and Sets. This improves the 
+use of Bags and eliminates logical errors arizing from incorrect
+subtyping relations that are allowed in the Java solution.
 
-\paragraph{Set/Bag broken solution}
+\subsection{Sets and Bags in Java}
 An iconic example on why connecting inheritance/code reuse and
-subtpying is problematic is provided by the
-historic\cite{LaLonde:1991:SSS:110673.110679}: A reasonable
-implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@
-keeping tracks of how many times an element occurs.  We would just add
-some state and override a couple of methods.
-For example in Java one could have
+subtpying is problematic is provided by
+LaLonde~\cite{LaLonde:1991:SSS:110673.110679}.  A reasonable
+implementation for a \Q@Set@ is easy to extend into a \Q@Bag@ by
+keeping track of how many times an element occurs.  We just need to
+add some state and override a few methods.
+
+For example in Java one could have:
 
@@
   ...}\end{lstlisting}
+\bruno{did you mean get instead of put?}
 
-Coding \Q@Bag@ in this way can avoid a lot of code duplication, but we induced unintented subtyping: indeed
+\noindent Coding \Q@Bag@ in this way can avoid a lot of code
+duplication, but \emph{we induced unintented subtyping}! 
+Indeed, the following is allowed:
 
@@
 \end{lstlisting}
-Since Subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
+Since subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
 
-\paragraph{Set/Bag fixed solution still broken}
-Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
+\paragraph{A (broken) attempt to fix the Problem in Java.}
+Of course, one could \emph{retroactivelly} fix this problem by introducing \Q@AbstractSetOrBag@
 and making both \Q@Bag@ and \Q@Set@ inherit from it:
@@
 abstract class AbstractSetOrBag{/*old set code goes here*/}
-class Set extends AbstractSetOrBag{} //empty body? just repeat constructors?
+class Set extends AbstractSetOrBag{} //empty body
 class Bag extends AbstractSetOrBag{/*old bag code goes here*/}
-..
-AbstractSetOrBag unexpected=new Bag(); //AbstractSetOrBag type not designed to be used.
+...
+//AbstractSetOrBag type not designed to be used.
+AbstractSetOrBag unexpected=new Bag(); 
 \end{lstlisting}
@@
 and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
-Worst, if we was to constantly apply this mentalty, we would introduce a very high number
-of abstract classes that are not supposed to be used as types, and that will clutter the 
+Worst, if we are to constantly apply this mentalty, we would introduce a very high number
+of abstract classes that are not supposed to be used as types. Those classes would clutter the 
 public interface of our classes and our code project as a whole.
-In our example now the information \Q@Set<:AbstractSetOrBag@ is present in the public interface
-of the class \Q@Set@, but is not needed to use the class properly.
+In our example the information \Q@Set<:AbstractSetOrBag@ would be present in the public interface
+of the class \Q@Set@, but it is not needed to use the class properly!
 
-Moreover, we have not really solved the original problem but only moved it further away. For example: one day  we may need bags that can only store up to 5 copies of the same element.
+Moreover, the original problem is not really solved, but only moved 
+further away. For example, one day  we may need bags that can only store up to 5 copies of the same element.
 We are now at the starting point again:
 \begin{itemize}
-\item either we inser \Q@class Bag5 extends Bag@ and we break LSP,
-\item we duplicate the code of the \Q@Bag@ implementation with minimal adjustments in \Q@class Bag5 extends AbstractSetOrBag@
+\item either we insert \Q@class Bag5 extends Bag@ and we break LSP; 
+\item or we duplicate the code of the \Q@Bag@ implementation with minimal
+  adjustments in \Q@class Bag5 extends AbstractSetOrBag@;
 \item or we introduce a
@@
 Note that this last solution is changing the public interface of the formerly released \Q@Bag@ class, and
-this may even break retro-compatibility (if a client program was using reflection in a poor way)
+this may even break retro-compatibility (if a client program was using
+reflection, for example).
 \end{itemize}
 
-\paragraph{Set/Bag in \name}
-Instead, in our proposed approach, if we was to originally declare
+\subsection{Sets and Bags in \name}
+Instead, in \name, if we was to originally declare
 \begin{lstlisting}
@@
 Then our code would be impossible to reuse in the first place for any user of our library.
-We consider this an advantage, since unintended code reuse runs into underdocumented behaviour near all of the times\footnote{See
-``Design and document for inheritance or else prohibit it''\cite{https://www.infoq.com/articles/API-Design-Joshua-Bloch}:
-self use of public methods is rarelly documented, thus is hard to understand the effects of overriding a library metod\cite{https://www.infoq.com/articles/API-Design-Joshua-Bloch}.
-}
+We consider this an advantage, since unintended code reuse runs into underdocumented behaviour nearly all the time\footnote{See
+``Design and document for inheritance or else prohibit
+it''\cite{Bloch08}: the
+self use of public methods is rarelly documented, thus is hard to understand the effects of overriding a library metod.
+}!
 

### 2017-05-22 FACS/section4.tex
Section 4 (in progress)

@@
-\section{Allowing Reuse against Subtyping}
-Before we discussed how traits does not induce a new (externally visible) type.
-However, locally in the trait the programmer can use the special type \Q@This@ in order to denote the 
+\section{Improving Reuse}
+\name allows reuse even when subtyping is impossible.
+In \name traits do not induce a new (externally visible) type.
+However, locally in the trait the programmer can use the special self-type \Q@This@ in order to denote the 
 type of \Q@this@.
@@
 impossible, that is, when a method parameter has type \Q@This@ (also called \emph{binary
-  methods}~\cite{}). 
+  methods}~\cite{bruce96binary}). 
 This type of situations is the primary motivator
 for previous work aiming at separating inheritance from subtyping.
-Leveraging on \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smootly integrate state and traits:
-
-  %????Despite the adoption of a nominal approach and unlike other nominally typed languages, the leaking problem does not prevent \name from enjoying such kind of improved reuse.
-
-%To illustrate how \name allows for improved reuse, we consider a
-%pattern to combine \emph{multiple trait inheritance} and \emph{state}:
-a challenging problem that has limited the flexibility of traits and
+Leveraging on the \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smootly integrate state and traits: a challenging problem that has limited the flexibility of traits and
 reuse in the past.
@@
 
-This idea of summing pieces of code is very elegant, and has proven
-very successfull in module composition
-languages~\cite{ancona2002calculus}\bruno{and several trait models?}. 
-However our research community is
-struggling to make it work with object state (constructor and fields)
-while acheving the following goals:
+To illustrate how \name improves reuse, we will show a novel approach
+to deal with \emph{state} in traits.  The idea of summing pieces of
+code is very elegant, and has proven very successfull in module
+composition languages~\cite{ancona2002calculus} and several trait
+models~\cite{Traits:ECOOP2003}\bruno{others}.  However the research
+community is struggling to make it work with object state (constructor
+and fields) while achieving the following goals:
 
@@
 %complicated discussions on this point \item keep sum associative and commutative,
-\item managing fields in a way that borrows the elegance of summing methods,
-\item actually initialize objects, leaving no null fields,
-\item make it easy to add new fields,
+\item managing fields in a way that borrows the elegance of summing methods;
+\item actually initialize objects, leaving no null fields;
+\item making it easy to add new fields;
 \item allowing a class to create instances of itself.
@@
 
-In the related work we will show some alternative ways to handle state.
-However the purest solution just requires methods:
-  The idea is that
-  the trait code just uses getter/setters/factories, while leaving
-  to classes the role to finally define the fields/constructors.
-  That is, the  the class has syntax richer that the trait one,
-  allowing declaration for fields and constructors.
-  This approach is very powefull~\cite{wang2016classless}
+In the related work we will show some alternative ways to handle
+state.  However the purest solution just requires methods. The idea is
+that the trait code just uses getter/setters/factories, while leaving
+to classes the role to finally define the fields/constructors. That
+is, the class has syntax richer that the trait one, allowing
+declaration for fields and constructors.  This approach is very
+powerful as illustrated by Wang et al.~\cite{wang2016classless}.
+
+\paragraph{Immutable Points} Consider, for example, two simple 
+traits that deal with \emph{immutable point} objects. That is, points
+in the cartesian plane (with coordinates \lstinline{x} and
+\lstinline{y}). The first trait provides a \emph{binary method} that 
+sums the point object with another point to return a new point. 
+The second trait provides a similar operation that does multiplication 
+instead.
 
@@
 
-As you see, all the state operations are represented as abstract methods.
+\noindent As we can see, all the state operations (the getters for the 
+\lstinline{x} and \lstinline{y} coordinates) are represented as abstract methods.
+Moreover there is an \lstinline{of} method which acts as a constructor
+for points. \bruno{I think we need to explain what a class method is
+  here, and how it differs from Java's static methods! mention
+  flatenning; importantly they can be abstract!}
 
 \subsection{A First Attempt at Composition}
-According to the general ideas expressed before,
+According to the general ideas about trait composition presented in
+Section~\ref{sec:separate}, a first attempt at composing the two traits providing
+two different operations on points is:
 \begin{lstlisting}
@@
 
-\noindent would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract.
-In this mindset, the user would be required to write something similar to
+\noindent However this fails since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract!
+Instead a user could write something similar to:
 
@@
 
-\paragraph*{Advantages:} 
-this approach is associative and commutative, even self construction
+\noindent This approach works, and it as some advantages, but also
+some disadvantages:
+
+\begin{itemize}
+
+\item {\bf Advantages:} This approach is associative and commutative, even self construction
   can be allowed if the trait requires a static/class method
-  returning \Q@This@; the class will then implement this method by forwarding
-  a call to the constructor.
+  returning \Q@This@. The class will then implement the methods returning \Q@This@
+  by forwarding a call to the constructor.
 
-\paragraph*{Negatives:} writing those obvious ``close the state/fixpoint'' class 
+\item {\bf Disadvantages:} Writing those obvious definitions to close
+  the state/fixpoint in the class 
    with the constructors and fields and getter/setters and factories can be quite tedious.
@@
    %"private" of such trait. 
-   Former work consider to automatically generate such code~\cite{wang2016classless}.
+   Previous work shows that such code can be automatically generated~\cite{wang2016classless}.
 
-In our model we go one step further: there is no need to generate code, or to explicitly
-write down constructors and fields; there is not even syntax for those
-constructs.
-The idea is that any class that ``could" be completed in the obvious way
-\emph{is a complete ``coherent" class}.
-In most other languages, a class is abstract if have abstract methods.
-Instead, we call abstract a class whose set of abstract methods is not
-coherent, that is, can not be automatically recognized as factory,getters and setters.
+\end{itemize}
 
-\paragraph*{Detaild definition of coherent:}
+\subsection{The \name approach to State}
+
+In our model we go one step further: there is no need to generate
+code, or to explicitly write down constructors and fields. In fact in
+\name there is not even syntax for those constructs!  The idea is that
+any class that ``can" be completed in the obvious way to form \emph{a
+  complete ``coherent" class}.  In most other languages, a class is
+abstract if it has abstract methods.  Instead, we call a class
+abstract only when the set of abstract methods is not coherent. That
+is, the unimplemented methods cannot not be automatically recognized
+as factory, getters and setters. Getters and setters are called
+\emph{abstract state operations}.
+
+\paragraph*{Coherent classes} A more detailed definition of coherent
+classes is given next:
 \begin{itemize}
-\item a class with no abstract method is coherent, and like Java \Q@Math@.
-Will just be usefull for calling class/static methods.
+\item a class with no abstract methods is coherent (just like Java
+  \Q@Math@, for example). Such classes are useful for calling class/static methods.
 \item a class with a single abstract \Q@class@ method returning \Q@This@
@@
 then
-\item a method \Q@int x()@ is intepreted as an abstract state method: a getter for \Q@x@.
-\item a method \Q@Void x(int that)@ is a setter for x.
+\begin{itemize}
+\item a method \Q@int x()@ is interpreted as an abstract state method: a \emph{getter} for \Q@x@.
+\item a method \Q@Void x(int that)@ is a \emph{setter} for x.
+\end{itemize}
 \end{itemize}
@@
 While getters and setters are fundamental operations, we can immagine
-more operations to be supported; for example
+more operations to be supported; for example:
 \begin{itemize}
@@
 
-We are not sure what is the best set of abstract state operations yet, but we think this could become a very interesting area of research.
+We are not sure what is the best set of abstract state operations yet,
+but we think this could become a very interesting area of research.
+The work by Wang et al.~\cite{wang2016classless} explores a particular
+set of such abstract state operations.
 
-\paragraph{The solution in \name:}${}_{}$\\
-With our approach to handle the state, the following code can be directly composed:
+\paragraph{Points in \name:}
+In \name and with our approach to handle the state, 
+\lstinline{pointSum} and \lstinline{pointMul} can be directly composed:
 
 \begin{lstlisting}
- //same code as before for pointSum and pointMul
-  pointSum: { method int x() method int y()
-    class method This of(int x,int y)
-    method This sum(This that)
-      This.of(this.x()+that.x(),this.y()+that.y())
-    }
-  pointMul: { method int x() method int y()//look we are repeating
-    class method This of(int x,int y)//the abstract method declarations.
-    method This mul(This that)
-      This.of(this.x()*that.x(),this.y()*that.y())
-    }
   PointAlgebra:Sum pointSum,pointMul
@@
 
-  As you can see, we can declare the methods independently and compose the result
+  As we can see, we can declare the methods independently and compose the result
   as we wish. In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
-  In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
+  In addition to \Q@sum@ and \Q@mul@ we may want many operations over points; can we improve our reuse
   and not repeat such abstract definitions? of course!

### 2017-05-23 FACS/Intro.tex
.

@@
 
-\bruno{You removed my lengthier explanation. I think this makes things
-less clear.}
-If in Java code reuse was not introducing subtyping, then an invocation of 
+As a though experiment, imagine that Java code-reuse (extends keyword) was not introducing subtyping: then an invocation of 
 \lstinline{B.mb} would result in a run-time type error.
@@
 We are aware of at least 3 independently designed research languages 
-that address this limitation: Delta-Trait~\cite{Damiani2014}, Package Templates~\cite{KrogdahlMS09}, DeepFjig~\cite{deep}.
-\bruno{Marco, please expand this a bit. Check if references are correct!}
+that address this limitation
+\begin{itemize}
+\item Delta-Trait (DT)~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}:
+Each language constructs  has one single responsibilty: classes instantiate objects,
+interfaces induce norminal types, records express state and traits are unit of reuse.
+\item Package Templates (PT)~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}:
+An extension of (full) Java where new packages can be ``syntetized'' by mixing
+and integrating code templates. As an extension of java, PT allows but not requires
+separation of inheritance and subtyping.
+\item DeepFjig~\cite{deep,servetto2014meta,DBLP:journals/iandc/LagorioSZ12}:
+A module composition language supporting composing nested classes, where the main idea is that
+nested classes with the same name are recursivelly composed.
+\end{itemize}
 
@@
   this-leaking problem, we synthesize the key ideas in {\bf a novel and
-  simple language design}. 
+  simple language design}. This language is the logic core of the language 42, and 
+  all the examples in the paper can be directly encoded as valid 42 programs.
+
 \item We illustrate how the new design {\bf improves both code use and code
@@
   languages that support traits.
-\item The new language design is {\bf adopted by the 42 language}. All the
-  examples in the paper are valid 42 programs.
+%\item The new language design is {\bf adopted by the 42 language}. All the
+ % examples in the paper are valid 42 programs.
 

### 2017-05-23 FACS/RelatedWork.tex
.

@@
 We are aware of at least 3 independently designed research languages 
-that address this limitation:Delta-Trait(DT)\cite{3ferruccio works}, Package Template(PT)\cite{}, DeepFjig\cite{}.
+that address this limitation:Delta-Trait(DT)\cite{ferruccioworks}, Package Template(PT)\cite{}, DeepFjig\cite{}.
 Leveraging on \emph{traits}, in this work we aim to synthesizes

### 2017-05-23 FACS/section2.tex
Section 4 (in progress)
.

@@
-\section{The Design of \name: Separating Use and Reuse}
+\section{The Design of \name: Separating Use and Reuse}\label{sec:separate}
 
@@
 For now notice that in the former example the code is correct even if
-no method called \Q@ma@ is explicitly declared\bruno{but it would not
-  work if the \lstinline{ta} was not in the use list, right?}.
+no method called \Q@ma@ is explicitly declared.
+DeepFJig\cite{} and DT\cite{} would require instead to explicitly declare an abstract \Q@ma@ as in
+\begin{lstlisting}
+  B:Use ta, { method int ma()//not required by us
+      method int mb(){return this.ma();} }
+\end{lstlisting}
+
 The idea is that such method is imported from trait \Q@ta@, exactly as in the Java equivalent
@@
 code literal is completely self-contained. However, this results in
-many spurious abstract method declarations.\bruno{Improved text. 
-Please double-check! You may actually show how the code with the extra 
-explicit declarations would look like! 
-}
+many spurious abstract method declarations.
 
@@
 
- This code is correct, and there is no mention of the trait
- \Q@ta@. In some sense, all the information about code-reuse/inheritance
+ This code is correct, and {\bf in the resulting code there is no mention of the trait
+ \Q@ta@}. In some sense, all the information about code-reuse/inheritance
   is just a private implementation detail of \Q@A@
  and \Q@B@; while subtyping is part of the class interface.
-\bruno{I don't think we are making  a strong point here, and this
-  discussion seems inconsequent at this point!}
 

### 2017-05-23 FACS/section3.tex
.

@@
   private Elem[] hashMap;
-  Elem put() /*body*/
-  boolean isIn(Elem)/*body*/
+  void put(Elem e) /*body*/
+  boolean isIn(Elem e)/*body*/
   ...}
@@
   private int[] countMap;// keep track of how many times is there
-  @Override Elem put()/*body*/
-  int howManyTimes(Elem)/*body*/
+  @Override void put(Elem e)/*body*/
+  int howManyTimes(Elem e)/*body*/
   ...}\end{lstlisting}
-\bruno{did you mean get instead of put?}
 
@@
 \end{lstlisting}
-Notice how, thanks to flattening, the resulting code for \Q@Bag@ is identical in both versions, thus if we are the developers of bags, we can temporarly go for the first version, and move when needed to the second without adding new undesired complexity for our old clients.
+Notice how, thanks to flattening, the resulting code for \Q@Bag@ is identical in both versions, 
+and as shown in Section 2, there is no trace of trait \Q@bag@.
+Thus if we are the developers of bags, we can temporarly go for the first version, and move when needed to the second without adding new undesired complexity for our old clients.
 

### 2017-05-23 FACS/paper.tex
.

@@
 seems required for technical reasons, due to what we call the
-\emph{this-leaking problem}. The this-leaking problem shows that separating
+\emph{this-leaking problem}. It shows that separating
 code-reuse from subtyping is non-trivial and requires a significant
@@
 handle behavior and state within trait composition.
-These ideas have been implemented in the 42 language, which we use 
-throughout the paper for code examples.\bruno{Is it true that all the code
-examples are valid 42?}
-
-\begin{comment}
-Two distinct groups of researchers independently [.marco..] and
-[..ferruccio.] used traits to provide
-a sound language design where code-reuse and subtyping are decoupled.
-In this work we distill their work in a simpler model that is far more
-natural for a OO programmer.
-For example, in [ marco ] class A {A m(){return this;}} is not well typed code,
-and in [.ferruccio..] all the logic about object construction have to
-be provided after all trait composition.
-
-We provide an novel and elegant solution to uniformly handle behavior
-and state within trait composition,
-Making our type system distinguish between used code and reused code
-we  can avoid many redundant
-abstract declarations that was required by [ marco ] and [ ferruccio ],
-while supporting self construction, binary methods and recursive types;
-
-We then show that our model can easily support nested classes and a
-rich composition algebra as in [ marco ]
-\end{comment}
+These ideas have been implemented in the 42 language, 
+that supports all the examples we show in the paper.
 

### 2017-05-23 FACS/section4.tex
Section 4

@@
 \end{lstlisting}
+\bruno{We talk about withers later on. So I think we should consider
+  having withers in this code, so that readers can understand what 
+withers are!}
 
@@
 \begin{lstlisting}
-  PointAlgebra:Sum pointSum,pointMul
+  PointAlgebra:Use pointSum,pointMul
 \end{lstlisting}
@@
   As we can see, we can declare the methods independently and compose the result
-  as we wish. In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
-  In addition to \Q@sum@ and \Q@mul@ we may want many operations over points; can we improve our reuse
-  and not repeat such abstract definitions? of course!
+  as we wish.
+
+  \paragraph{Improved solution} In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
+  Moreover, in addition to \Q@sum@ and \Q@mul@ we may want many
+  operations over points. It is possible to improve reuse
+  and not repeat such abstract definitions by abstracting the common
+  abstract definitions into a trait \Q@p@:
 
@@
 
-Now our code is fully modularized, and each trait handle exactly one method.
+Now the code is fully modularized, and each trait handles exactly one method.
 
-\paragraph{Adding state:}${}_{}$\\
-What happens if we want to add fields instead of just operations?
+\subsection{State Extensibility}
+Programmers may want to extend points with more state. For example 
+they may want to add colors to the points. A first attempt at doing
+this would be:
 
@@
 is not a getter for any of the parameters of 
-\Q@ class method This of(int x,int y)@
+\Q@ class method This of(int x,int y)@. 
 A solution is to provide a richer factory:
@@
 
-where we assume to support overloading on different parameter number.
-This is a good solution, indeed we think is better that any alternatives in literature,
-however the method \Q@CPoint.sum@ resets the color to red: we call the \Q@of(int,int)@ method, that now delegate to \Q@of(int,int,Color)@ by passing red as the default field value.
-What should be the behaviour in this case?
-If our abstract state supports withers, instead of writing \Q@This.of(...)@ we can use
-\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
-This solution is still not sadisfactory: this design ignore the color from \Q@that@.
+\noindent 
+where we assume support for overloading on different number of parameters.
+This is a reasonable solution, however the method \Q@CPoint.sum@ resets
+the color to red: we call the \Q@of(int,int)@ method, that now
+delegates to \Q@of(int,int,Color)@ by passing red as the default field
+value.  What should be the behaviour in this case?  If our abstract
+state supports withers, instead of writing \Q@This.of(...)@ we can use
+\Q@this.withX(newX).withY(newY)@ in order to preserve the color from
+\Q@this@.  This solution is still not satisfactory: this design ignores
+the color from \Q@that@.
 
-If the point designer could predict this kind of extension, then we could use the following better design:  
+\paragraph{A better design}
+If the point designer is designing for reuse and extensibility, then 
+a better design would be the following:  
 \begin{lstlisting}
@@
 \end{lstlisting}  
-  Now we can merge colors, or any other kind of state we may want to add
-  following this pattern.
+  This design allows merging colors, or any other kind of state we may want to add
+  following this pattern.\bruno{worried that withers are not explained
+  enough.}
 
+\paragraph{Independent Extensibility}
+  Ofcourse, quite frequentely there can be multiple independent
+  extensions~\cite{Zenger-Odersky2005} that need to be composed. Lets suppose that 
+  we could have a notion of flavoured points as well.   
   In order to compose, let say \Q@colored@ with \Q@flavored@ we would

### 2017-05-23 FACS/Intro.tex
.

@@
 In mainstream OO languages like Java, C++ or C\# subclassing 
-implies subtyping. For example, in Java a subclass definition, such as:
-
-\begin{lstlisting}
-class A extends B {}
-\end{lstlisting}
-
+implies subtyping. For example, in Java a subclass definition, such as 
+\Q@class A extends B {}@
 \noindent does two things at the same time:
-1) {\bf inheriting} code from class \lstinline{B}; and {\bf creating 
+1) {\bf inherits} \lstinline{B} code; and {\bf create
 a subtype} of \emph{B}. Therefore in a language like Java 
@@
 separing subtyping from subclassing~\cite{cook}.  This is claimed to be
-good for code reuse, design and reasoning. There are at
+good for code-reuse, design and reasoning. There are at
 least two distinct situations where the separation of subtyping and 
@@
 extends and method calls) is forced to accept that subclassing implies
-subtyping. Note how this is true also in C++, where is
-  possible to "extends privately". Such is a limitation of the
-  visibility of subtyping but not over subtyping itself, and the
-  former example would be accepted by C++ even if \lstinline{B} was to
-  "privately extends" \lstinline{A}.
+subtyping.\footnote{C++ allows to "extends privately", but it is a limitation over
+  subtyping visibility, not over subtyping itself.  The
+  former example would be accepted even if \lstinline{B} was to
+  "privately extends" \lstinline{A}}.
 
 In essence we believe that a problem with the design of classes in
-languages like Java is that classes do too many things at once. In
-particular they act both as units of \emph{use} and \emph{reuse}.
+languages like Java is that classes
+% do too many things at once. In particular they 
+act both as units of \emph{use} and \emph{reuse}.
 That is, in languages like Java, class names can be \emph{use}d as
@@
 \item Delta-Trait (DT)~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}:
-Each language constructs  has one single responsibilty: classes instantiate objects,
-interfaces induce norminal types, records express state and traits are unit of reuse.
+Each  constructs  has one single responsibilty: classes instantiate objects,
+interfaces induce types, records express state and traits are unit of reuse.
 \item Package Templates (PT)~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}:

### 2017-05-23 FACS/paper.tex
.

@@
 \usepackage{rotating} 
+\addtolength{\textwidth}{0.5ex}
+\usepackage{enumitem}
+\setlist{nosep}
+\setlist{leftmargin=0ex}
+
 \lstset{language=Java,
@@
 \begin{abstract}
-In most object oriented languages subclassing/inheritance implies
+In most OO languages subclassing/inheritance implies
 subtyping. This is considered by many a conceptual design error, but it
@@
 \emph{this-leaking problem}. It shows that separating
-code-reuse from subtyping is non-trivial and requires a significant
+inheritance from subtyping is non-trivial and requires a significant
 departure from the OO models in existing mainstream OO languages.
@@
 that address this limitation:Delta-Trait, Package Template, DeepFjig.
-Leveraging on \emph{traits}, in this work we aim to synthesize
-the main ideas of those very different language designs.
-Our result is a nominally typed minimalist language, natural to OO programmers.
+In this work we synthesize
+the main ideas of those very different language designs into a nominally typed minimalist language,
+natural to OO programmers.
 By making our type system distinguish between code-use and code-reuse
 we can separate inheritance and subtyping, while avoiding 
-redundant abstract declarations (required in Delta-Trait and
-DeepFjig). At the same time self construction,
+redundant abstract declarations required in Delta-Trait and
+DeepFjig. At the same time self construction,
 binary methods and recursive types are also supported.
@@
 
-\section*{Bibliography}
 \bibliographystyle{plain}

### 2017-05-23 FACS/Intro.tex
.

@@
 languages like Java is that classes
-% do too many things at once. In particular they 
-act both as units of \emph{use} and \emph{reuse}.
-That is, in languages like Java, class names can be \emph{use}d as
-type and instantiated.  Classes can also be subclassed to provide
+do too many things at once. In particular they 
+act both as units of \emph{use} and \emph{reuse}:
+ classes can be \emph{use}d as
+type and instantiated; classes can also be subclassed to provide
 \emph{reuse} of code.

### 2017-05-23 FACS/section4.tex
.

@@
 composition languages~\cite{ancona2002calculus} and several trait
-models~\cite{Traits:ECOOP2003}\bruno{others}.  However the research
+models~\cite{Traits:ECOOP2003,Bergel2007,BETTINI2013521,DBLP:journals/iandc/LagorioSZ12}.  However the research
 community is struggling to make it work with object state (constructor
@@
 
-\begin{itemize}
+%\begin{itemize}
 %complicated discussions on this point \item keep sum associative and commutative,
-\item managing fields in a way that borrows the elegance of summing methods;
-\item actually initialize objects, leaving no null fields;
-\item making it easy to add new fields;
-\item allowing a class to create instances of itself.
-\end{itemize}
+%\item
+(1) managing fields in a way that borrows the elegance of summing methods;
+%\item
+(2) actually initialize objects, leaving no null fields;
+%\item
+(3) making it easy to add new fields;
+%\item
+(4) allowing a class to create instances of itself.
+%\end{itemize}
 
@@
     }
-  pointMul: { method int x() method int y()
-    class method This of(int x,int y)
+  pointMul: { method int x() method int y()//repeating getters
+    class method This of(int x,int y)//repeating factory
     method This mul(This that)
@@
 \noindent As we can see, all the state operations (the getters for the 
-\lstinline{x} and \lstinline{y} coordinates) are represented as abstract methods.
-Moreover there is an \lstinline{of} method which acts as a constructor
-for points. \bruno{I think we need to explain what a class method is
-  here, and how it differs from Java's static methods! mention
-  flatenning; importantly they can be abstract!}
+\lstinline{x} and \lstinline{y} coordinates) are represented as {\bf abstract} methods.
+Notice the abstract \Q@class method This of(..)@ which acts as a constructor
+for points:
+A class method is similar to a \Q@static@ method in Java but can be abstract. 
+As for instance methods, they are late bound:  flattening can provide an implementation for them.
+abstract class method are very similar to the original concept of member functions in the module composition setting.
 
@@
 withers are!}
+\marco{with withers it will look more complicated}
 
@@
 \name there is not even syntax for those constructs!  The idea is that
-any class that ``can" be completed in the obvious way to form \emph{a
+any class that ``can" be completed in the obvious way  is \emph{a
   complete ``coherent" class}.  In most other languages, a class is
@@
 is, the unimplemented methods cannot not be automatically recognized
-as factory, getters and setters. Getters and setters are called
+as factory, getters and setters. Methods recognized as factory, getters and setters are called
 \emph{abstract state operations}.
@@
 \item \Q@method This withX(int that)@
-may be a ``wither", doing a functional field update, that is create a new instance that is like \Q@this@ but where field \Q@x@ has now \Q@that@ value.
+may be a ``wither", doing a functional field update: it creates a new instance that is like \Q@this@ but where field \Q@x@ has now \Q@that@ value.
 \item \Q@method Void update(int x,int y)@
@@
 \paragraph{Independent Extensibility}
-  Ofcourse, quite frequentely there can be multiple independent
+  Of course, quite frequentely there can be multiple independent
   extensions~\cite{Zenger-Odersky2005} that need to be composed. Lets suppose that 

### 2017-05-24 FACS/Examples.tex
.

@@
 \prodFull\mV{\mt \mid \mL}{Code Value}\\
-\prodFull\mL{\Opt{\terminalCode{interface}}\ \terminalCode{implements} \overline\mT\ \overline\mMD}{Code Literal}\\
+\prodFull\mL{
+\oC
+\Opt{\terminalCode{interface}}\ \terminalCode{implements} \overline\mT\ \overline\mMD
+\cC
+}{Code Literal}\\
 \prodFull\mT{\mC}{types are class names}\\
@@
 
-\use can not be applied if multiple versions of the same method are
+\use\ can not be applied if multiple versions of the same method are
 present in different traits.  An exception is done for abstract methods:
@@
 \begin{itemize}
-\item all the traits and classes have unique names in a program $\overline\mD$,
 \item all method parameters have unique names and the special parameter name \Q@this@ is not declared
@@
 \begin{itemize}
+\item all the traits and classes have unique names in a program $\overline\mD$, and the special class name
+\Q@This@ is reserved,
+\item all used types correspond to class declarations in the program, or are \Q@This@, 
 \item subtyping between interfaces and classes,
@@
 For the sake of simplicity, method bodies are just simple expressions
-\me: they can be just variables and method calls.
+\me: they can be just variables, types and method calls. We need types as part of expressions in order to use them as receivers for class methods.
 
@@
 
-This is not the case in many similar works in literature~\cite{} where the
+This is not the case in many similar works in literature~\cite{deep,Bettini2015282,Bergel2007} where the
 literals have to be self complete. In this case we would have been forced to
@@
 \begin{itemize}
-\item if a class is declared by using $\mC : \use\ \overline\mt$, that is, without literals,
+\item If a class is declared by using $\mC : \use\ \overline\mt$, that is, without literals,
 and the flattening is successfull, \mC\ is well typed, no need of further checking.
-\item on the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
+\item On the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
 $\mL_1\ldots\mL_n \in \overline\mV$, and after successfull flattening $\mC : \mL$ can not be typechecked,
@@
 
-OO language leverage on recursive types most of the times.
+OO languages leverage on recursive types most of the times.
 For example in a pure OO language, \Q@String@ may offers a \Q@Int size()@
@@
 Now, we need to generate \Q@C@, and we need to ensure well typedness of \Q@tc@.
-Now \Q@B@ is alreay well typed (since generated by \use \Q@tb@, with no \mL),
+Now \Q@B@ is alreay well typed (since generated by \use\ \Q@tb@, with no \mL),
 and \Q@A@ can be typed;  finally \Q@tc@ can be typed and used.
@@
 
-Note that we can reuse code only by naming traits; but our only point of relaxation is the class literal: there is no way an error can ``move around'' and be duplicated during the compilation process.
+Note that we can reuse code only by naming traits; but our only point of relaxation is {\bf only} the code literal: there is no way an error can ``move around'' and be duplicated during the compilation process.
 In particular, our approach allows for safe libraries of traits and classes to be fully typechecked, deployed and reused by multiple clients: no type error will emerge from library code.

### 2017-05-25 FACS/Conclusion.tex
.

@@
-\section{Conclusions and practical applications}
+\section{Conclusions, extensions and practical applications}
 
@@
 
-We show how our model can be easly extended with more operators, if they behave well with respect to a simple predicate.
+The model presented here is very easy to extend:
+For example nested classes can be easly added,giving us enough power to easly solve the expression problem as in \cite{deep}. More composition operators can be added in addition to \use.
+In particular all the sophisticated operators of \cite{deep} could be added.
+ Indeed we can add any operator respecting following criteria:
 
-In order to focus on the explanation of the high level concepts, this paper do not present formalism for our approach.
-Formalism for similar approaches can be found in[deefjig, myThesis].
+\begin{itemize}
+\item As for \use, the operator does not need to be total, but if it fails it needs to provide an error that will be reported to the programmer.
+\item When the operator takes in input only traits (they are going to be well typed), if a result is produced,
+ such result is well typed.
+\item When the operator takes in input also code literals, if a non well typed result is produced,
+the type error must be tracked back to code in one of those non typed yet code literals.
+ \end{itemize}
 
-We are working on 42: a full blown language 
+ Our simplified model represent the conceptual core of  a novel full blown programming language 
+ (\verb@http://L42.is@),
 that leverage on the ideas presented in this paper to obtain reliable and understandble metaprogramming.
-Formalization (in progress) for such language can be found at URLURLURL.
-42 extands our model allowing
+Formalization (in progress) for such language can be found at 
+\verb@github.com/ElvisResearchGroup/L42/tree/master/Main/formal@.
+L42 extands our model allowing
 flattening to execute arbitrary computations.
 In such model we do not need an explicit notion of traits: they are encoded as method returning a code literal.
+L42 also have feature less related to code composition, like
+  a strong type system supporting aliasing mutability and circularity control,
+   checked exceptions, errors (unchecked exceptions) with strong-exception-safety for errors.
 
-  In addition to this, 42 have feature less related to code composition, like
-  a strong type system supporting aliasing mutability[] and circularity control[],
-   checked exceptions[], errors (unchecked exceptions) with strong exception safety for errors[].
-
-42 do not have a finite set of composion operators; they can be added using the built in support for native method calls.
+L42 do not have a finite set of composion operators; they can be added using the built in support for native method calls.
 They can be dynamically checked to verify that they are well behaved according to our predicate, or they can be trusted to acheive efficiency.
-
-%\item more operators can be accomodated without changing the general model

### 2017-05-25 FACS/RelatedWork.tex
.

@@
-\section{Extensions and practical experience}
-The model presented here is very easy to extend:
-For example nested classes can be easly added,giving us enough power to easly solve the expression problem as in \cite{deepfjig}. More composition operators can be added in addition to use. Indeed
-we just need to be sure that every newly added operator respects the following criteria:
-
-\begin{itemize}
-\item As for \use, the operato does not need to be total, but if it fails it needs to provide an error that will be reported to the programmer.
-\item When the operator takes in input only traits (they are going to be well typed), if a result is produced,
- such result is well typed.
-\item When the operator takes in input also code literal, if a non well typed result is produced,
-the type error must be tracked back to code in one of those non typed yet code literals.
- \end{itemize}
- That is, our model can easly support all the operators of \cite{deep}.
-
- Our simplified model represent the conceptual core of  a novel programming language (http://L42.is)
- we are currently working at.
- L42 contains the extensions described before and much more, and provide a concrete example of the usability of our model.
-
- The main difference is that code literal are first class values in L42, thus we do not need traits as a language concept, but just as a method returning such code literal.
-
 \section{Related Work}

### 2017-05-25 FACS/Intro.tex
.

@@
 \section{Introduction}\label{sec:intro}
-
+\saveSpace\saveSpace
 In mainstream OO languages like Java, C++ or C\# subclassing 
@@
 
-Historically, there has been a lot of focus on the importance of
+Historically, there has been a lot of focus on
 separing subtyping from subclassing~\cite{cook}.  This is claimed to be
@@
 Method \lstinline{A.ma} passes \lstinline{this} as \lstinline{A} to \Q@Util.m@.
-This code seems correct, and there is no subtyping/subclassing.
-Now, lets add a class \Q@B@
+This code seems correct, and there is no subtyping/subclassing.  Now, lets add a class \Q@B@
 
@@
 
-In essence we believe that a problem with the design of classes in
-languages like Java is that classes
+In essence we believe that in
+languages like Java classes
 do too many things at once. In particular they 
@@
 \item DeepFjig~\cite{deep,servetto2014meta,DBLP:journals/iandc/LagorioSZ12}:
-A module composition language supporting composing nested classes, where the main idea is that
+A module composition language where the main idea is that
 nested classes with the same name are recursivelly composed.
@@
 brings several benefits in terms of flexibility and code reuse.
+In summary, our contributions are:
+
 %\marco{We need to talk of unanticipated extensions?}
@@
 
-\bruno{Need to say something about state?}
+%\bruno{Need to say something about state?}
 
-The language design of \name is adopted by the 42 programming 
-language, All the examples shown in this paper can be run in 42.
-The compiler for 42 is available at:
+%The language design of \name is adopted by the 42 programming 
+%language, All the examples shown in this paper can be run in 42.
+%The compiler for 42 is available at:
 
-\url{http://l42.is}
+%\url{http://l42.is}
 
-\begin{comment}
-First we show a minimal language that illustrates the basic ideas. 
-Then we show how mutually recursive types
-are supported, how state/constructors/fields are supported,
-how we can extend the language with nested classes.
-Finally we show 42, a full blown language build around our ideas of
-reuse.
-
-Our design leverage on traits~\cite{ducasse2006traits}: a well know mechanisms for pure
-code reuse.
-\end{comment}
-
-\subsection{Contributions:}
 \begin{itemize}
@@
   makes the separation of subclassing and subtyping difficult.
-\item Inspired by previous language designs that solve the
-  this-leaking problem, we synthesize the key ideas in {\bf a novel and
+\item We synthesize the key ideas of previous designs solving the
+  this-leaking problemin into {\bf a novel and
   simple language design}. This language is the logic core of the language 42, and 
@@
   reuse}.
-\item We propose {\bf a clean and elegant approach to handle of state} in a
-  languages that support traits.
+\item We propose {\bf a clean and elegant approach to handle of state} in a trait based language.
+
 %\item The new language design is {\bf adopted by the 42 language}. All the
  % examples in the paper are valid 42 programs. 
+\end{itemize}
 
-\begin{comment}
-\item simple approach to fully separate inheritance and subtyping
-\item easy to grasp model of compilation were typing is late but not
-  too late (lazy?)\bruno{this idea is important: it's a key design
-    decision. We need to talk about it in the introduction and offer
-    some motivation/justification. }
-\item clean and elegant handling of state in trat/module composition language
-\item extension with nested classes is natural and
-  powefull\bruno{wondering if this is necessary. I think there are
-    several interesting/novel ideas already. Covering too much will
-    result on not saying enough about each individual ideas.}
-\item more operators can be accomodated without changing the general
-  model
-\end{comment}
-\end{itemize}
+\saveSpace
+\saveSpace

### 2017-05-25 FACS/paper.tex
.

@@
 \setlist{leftmargin=0ex}
+\setlist[itemize]{noitemsep,nolistsep}
+\setlist[enumerate]{noitemsep,nolistsep}
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\vspace{-5pt}\oldparagraph{#1}\vspace{-5pt}}
+\let\oldsubsection\subsection
+\renewcommand{\subsection}[1]{\vspace{-5pt}\oldsubsection{#1}\vspace{-5pt}}
 
+\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
+%
 \lstset{language=Java,
-  basicstyle=\ttfamily\small,%\scriptsize,
+  basicstyle=\ttfamily\footnotesize,%\small,%\scriptsize,
   keywordstyle=\bfseries,%\color{darkRed},
@@
 %
+\newcommand\saveSpace{\vspace{-3pt}}
 \newcommand\Rotated[1]{\begin{turn}{90}\begin{minipage}{12em}#1\end{minipage}\end{turn}}
@@
 \newcommand\use{\terminalCode{Use}}
-\newcommand\oC{\mbox{\lstinline@{@}}
-\newcommand\cC{\mbox{\lstinline@}@}}
+\newcommand\oC{\mbox{\lstinline@\{@}}
+\newcommand\cC{\mbox{\lstinline@\}@}}
 \newcommand\oR{\mbox{\lstinline@(@}}
@@
 \institute{Victoria University of Wellington, New Zealand,\\
-\email{marco.servetto@ecs.vuw.ac.nz},\\ WWW home page:
-\texttt{https://ecs.victoria.ac.nz/Main/MarcoServetto}
 \and
@@
 \begin{abstract}
+\saveSpace\saveSpace\saveSpace
 In most OO languages subclassing/inheritance implies
@@
 We are aware of at least 3 independently designed research languages 
-that address this limitation:Delta-Trait, Package Template, DeepFjig.
-In this work we synthesize
-the main ideas of those very different language designs into a nominally typed minimalist language,
+addressing this limitation:Delta-Trait, Package Template, DeepFjig.
+Here we synthesize
+the main ideas of those very different designs into a nominally typed minimalist language,
 natural to OO programmers.
@@
 that supports all the examples we show in the paper.
+\saveSpace\saveSpace\saveSpace
 
 \keywords{Code Reuse, Object-Oriented Programming}
+\saveSpace\saveSpace\saveSpace
 \end{abstract}

### 2017-05-25 FACS/section3.tex
.

@@
 
-This section illustrates how 
-%As stated in Section~\ref{} 
-%To have code-reuse, while preventing unintended subtyping
-%in a nominally typed OO language we need to
-% separate inheritance from subtyping.
-\name \emph{improves the use} of classes, while
-retaining the benefits of code-reuse when compared to Java-like languages. 
-Here we illustrate this benefit by modelling a simplified version of
+To illustrates how 
+\name \emph{improves the use} of classes we model a simplified version of
 Set and Bag collections first in Java, and then in \name.
@@
 add some state and override a few methods.
-
 For example in Java one could have:
@@
   void put(Elem e) /*body*/
-  boolean isIn(Elem e)/*body*/
-  ...}
+  boolean isIn(Elem e)/*body*/}
 class Bag extends Set{...//for each element in the hash map,
@@
   @Override void put(Elem e)/*body*/
-  int howManyTimes(Elem e)/*body*/
-  ...}\end{lstlisting}
+  int howManyTimes(Elem e)/*body*/}
+\end{lstlisting}
 
-\noindent Coding \Q@Bag@ in this way can avoid a lot of code
+\noindent Coding \Q@Bag@ in this way avoids a lot of code
 duplication, but \emph{we induced unintented subtyping}! 
+Since subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
 Indeed, the following is allowed:
@@
 \end{lstlisting}
-Since subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
 
 \paragraph{A (broken) attempt to fix the Problem in Java.}
-Of course, one could \emph{retroactivelly} fix this problem by introducing \Q@AbstractSetOrBag@
+One could \emph{retroactivelly} fix this problem by introducing \Q@AbstractSetOrBag@
 and making both \Q@Bag@ and \Q@Set@ inherit from it:

### 2017-05-25 FACS/Intro.tex
paper

@@
 \noindent does two things at the same time:
-1) {\bf inherits} \lstinline{B} code; and {\bf create
+1) it {\bf inherits} code from \lstinline{B}; and it {\bf creates
 a subtype} of \emph{B}. Therefore in a language like Java 
@@
 Historically, there has been a lot of focus on
-separing subtyping from subclassing~\cite{cook}.  This is claimed to be
+separating subtyping from subclassing~\cite{cook}.  This is claimed to be
 good for code-reuse, design and reasoning. There are at
@@
 
-As a though experiment, imagine that Java code-reuse (extends keyword) was not introducing subtyping: then an invocation of 
+As a though experiment, imagine that Java code-reuse (the {\bf extends} keyword) was not introducing subtyping: then an invocation of 
 \lstinline{B.mb} would result in a run-time type error.
@@
 expected argument of type \lstinline{A}.  
-
-Every OO language with the minimal features exposed in the example (using \lstinline{this},
+Therefore, every OO language with the minimal features exposed in the example (using \lstinline{this},
 extends and method calls) is forced to accept that subclassing implies
-subtyping.\footnote{C++ allows to "extends privately", but it is a limitation over
+subtyping\footnote{C++ allows to "extends privately", but it is a limitation over
   subtyping visibility, not over subtyping itself.  The
@@
 
-In essence we believe that in
-languages like Java classes
-do too many things at once. In particular they 
-act both as units of \emph{use} and \emph{reuse}:
- classes can be \emph{use}d as
-type and instantiated; classes can also be subclassed to provide
-\emph{reuse} of code.
+In essence we believe that in languages like Java classes do too many
+things at once. In particular they act both as units of \emph{use} and
+\emph{reuse}: classes can be \emph{use}d as types and can be instantiated;
+classes can also be subclassed to provide \emph{reuse} of code.
 
-What the this-leaking problem shows is that 
-adopting a more flexible nominally typed OO model where subclassing
-does not imply subtyping is not trivial, and a more substantial change
-in the language design is necessary. 
-We are aware of at least 3 independently designed research languages 
-that address this limitation
+What the this-leaking problem shows is that adopting a more flexible
+nominally typed OO model where subclassing does not imply subtyping is
+not trivial, and a more substantial change in the language design is
+necessary.  We are aware of at least 3 independently designed research
+languages that address this limitation:
 \begin{itemize}
-\item Delta-Trait (DT)~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}:
-Each  constructs  has one single responsibilty: classes instantiate objects,
-interfaces induce types, records express state and traits are unit of reuse.
-\item Package Templates (PT)~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}:
-An extension of (full) Java where new packages can be ``syntetized'' by mixing
+\item In {\bf Delta-Trait (DT)}~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}
+each  construct has a single responsability: classes instantiate objects,
+interfaces induce types, records express state and traits are reuse units.
+\item {\bf Package Templates (PT)}~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}:
+are an extension of (full) Java where new packages can be ``syntetized'' by mixing
 and integrating code templates. As an extension of java, PT allows but not requires
 separation of inheritance and subtyping.
-\item DeepFjig~\cite{deep,servetto2014meta,DBLP:journals/iandc/LagorioSZ12}:
-A module composition language where the main idea is that
+\item {\bf
+    DeepFjig}~\cite{deep,servetto2014meta,DBLP:journals/iandc/LagorioSZ12} is
+a module composition language where the main idea is that
 nested classes with the same name are recursivelly composed.
 \end{itemize}
-
+\bruno{We need here a better motivation. From the reviewer perspactive
+he will think: ``So, there are already 3 solutions, why should I care
+about a new one?''}
 This paper aims at showing a simple language design, called \name, to
@@
   makes the separation of subclassing and subtyping difficult.
-\item We synthesize the key ideas of previous designs solving the
-  this-leaking problemin into {\bf a novel and
+\item We synthesize the key ideas of previous designs that solve the
+  this-leaking problem into {\bf a novel and
   simple language design}. This language is the logic core of the language 42, and 

### 2017-05-25 FACS/paper.tex
paper

@@
 seems required for technical reasons, due to what we call the
-\emph{this-leaking problem}. It shows that separating
+\emph{this-leaking problem}. This problem shows that separating
 inheritance from subtyping is non-trivial and requires a significant

### 2017-05-25 FACS/Examples.tex

@@
-\section{Intuitions on formalization}\label{sec:formal}
+\saveSpace\saveSpace\section{Intuitions on formalization}\label{sec:formal}
+\saveSpace\saveSpace
 In this article we dedicate more space to examples and informal presentation and motivations;
@@
 
-\begin{comment}
-\subsection{Code Use and Code Reuse in \name}
-
-The great absent in our language is the  \Q@extends@ keyword.
-In a sense, all classes are final (as in Java).
-Moreover  \use\ cannot contain class names.
-That is, using a trait is the only way to induce code reuse.
-
-Lets now try to encode the example in Section~\ref{} e in our language:
-
-\begin{lstlisting}
- A:{ method int ma() Utils.m(this) }//note, no {return _}
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-This code is fine, but there is no way to add a class \Q@B@ reusing
-the code of \Q@A@, since
-\Q@A@ is designed for code \emph{use} and not \emph{reuse}. Lets try
-again, but this time aiming at code reuse:
-
-\begin{lstlisting}
- ta:{ method int ma() Utils.m(this) }//type error
- A:Use ta
- Utils:{ class method int m(A a)/*method body here*/ }
-\end{lstlisting}
-
-This does not work, because \lstinline{Utils.m} requires an \lstinline{A} and \lstinline{this} in
-\Q@ta@ have no knowlegde of \lstinline{A}. Lets try again
-
-\begin{lstlisting}
- IA:{interface method int ma()}//interface with abstract method
- ta:{implements IA
-   method int ma() Utils.m(this) }
- A:Use ta
- Utils:{ class method int m(IA a)/*method body here*/}
-\end{lstlisting}
-
-This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
-is implementing it. It is also possible to add a \Q@B@ as follows
-\begin{lstlisting}
-  B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
-Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.
-
-\paragraph*{Semantic of Use}
-Abeit alternative semantic models for traits~\cite{} have been proposed,
-here we use the flattening model.
-This means that 
-\begin{lstlisting}
-A:Use ta
-B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
-
-  reduces/is equivalent to/is flatted into
-
- \begin{lstlisting}
-A:{implements IA method int ma() Utils.m(this) }
-B:{implements IA
-  method int ma() Utils.m(this)
-  method int mb() this.ma() }
-
- \end{lstlisting}
-
-This code seems correct, and there is no mention of the trait \Q@ta@.
-In some sense, all the information about code reuse/subclassing is just a private implementation detail of \Q@A@ and \Q@B@; while subtyping is part of the class interface.
-\end{comment}
-
 \subsection{Remarks on Typing}
@@
 The are two core ideas:
-\paragraph*{1: traits are \emph{well-typed} before being reused.}${}_{}$\\*
+\paragraph{1: traits are \emph{well-typed} before being reused.}${}_{}$\\*
  For example in
 
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 t:{method int m() 2 
    method int n() this.m()+1}
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 
@@
 
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 t1:{method int n() this.m()+1} 
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 \noindent would be ill typed.
 
-\paragraph*{2: code literals are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
+\paragraph{2: code literals are not required to be \emph{well-typed} before flattening.}${}_{}$\\*
 In class expressions  $\use\ \overline\mV$
@@
 for example in
-\begin{lstlisting}[language=Java]
+\saveSpace\begin{lstlisting}[language=Java]
   class A{ int m() {return 2;}  int n(){return this.m()+1;} }
   class B extends A{ int mb(){return this.ma();} }
-\end{lstlisting}
-
+\end{lstlisting}\saveSpace
 \noindent in \Q@B@ we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for \Q@ma()@.
@@
 
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 C: Use t {method int k() this.n()+this.m()}
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 \noindent would be correct: even if n,m are not defined inside
@@
 
-The most expressive compilation process may divide the classes in groups of mutually 
+The most expressive compilation process may divide the program in groups of mutually 
 dependent classes.
@@
  of each class in the group.
-
 In this model, it is reasonable to assume that flattening happens group by group, before
@@
 an error arise. There are two main kinds of errors: Type errors (like method not found) or Composition errors (like summing two conflicting implementation for the same method).
-
 For example
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 A:{method int ma(B b) b.mb()+1}
@@
 C: Use tc, {method int hello() 1}
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 In this scenario, since we go top down, we first need to generate \Q@B@.
@@
 a composition error could be generated at this stage.
-
 On the opposite side, if \Q@B@ and \Q@C@ was swapped, as in
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 C: Use tc, {method int hello() 1}  
 B: Use tb
-\end{lstlisting}
-
+\end{lstlisting}\saveSpace
+\noindent
 now the first task would be to generate \Q@C@, but 
@@
 But they are both unavailable: \Q@B@ is still not computed and 
-\Q@A@ can not be compiled/checked alone, without information about \Q@B@.
+\Q@A@ can not be compiled/checked without information about \Q@B@.
 A type error would be generated, on the line of ``flattening of \Q@C@
@@
 
-\paragraph*{Criticism: existence of an order is restrictive.}${}_{}$\\*
+\paragraph{Criticism: existence of an order is restrictive.}${}_{}$\\*
 Some may find the requirement of the existence of an order restrictive;
 An example of a ``morally correct" program where no right order exists is the following:
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 t:{ int mt(A a) a.ma()}
 A:Use t {int ma() 1}
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 
@@
 is indeed possible to optimistically compute the resulting structural shape of the classes
-and use this information to type involved examples like the former.
+and use it to type involved examples like the former.
 We stick to our simple approach, since we believe such typing discipline would be fragile,
@@
 
-\paragraph*{Criticism: it would be better to typecheck before flattening.}${}_{}$\\*
-In the world of strongly typed languages we could be tempted to
+\paragraph{Criticism: it would be better to typecheck before flattening.}${}_{}$\\*
+In the world of strongly typed languages we are tempted to
 first check that all can go well, and then perform the flattening.
@@
 On the other side, we do not enforce the programmer to write always self-contained code where all the abstract method definition are explicitly declared.
-
-\begin{comment}
-\section{Managing State}
-This idea of summing pieces of code is very elegant,
-and has proven very successfull in module composition languages~\cite{ancona2002calculus} but our research community is struggling to
-make it work with object state (constructor and fields) while acheving the following goals:
-
-\begin{itemize}
-\item keep sum associative and commutative,
-\item allowing a class to create instances of itself,
-\item actually initialize objects, leaving no null fields,
-\item managing fields in a way that borrows the elengance of summing methods,
-\item make easy to add new fields.
-\end{itemize}
-
-In the related work we will show some alternative ways to handle state.
-However the purest solution just requires methods:
-  The idea is that
-  the trait code just uses getter/setters/factories, while leaving
-  to classes the role to finally define the fields/constructors.
-  That is, the  the class has syntax richer that the trait one,
-  allowing declaration for fields and constructors.
-  This approach is very powefull~\cite{wang2016classless}
-
-\paragraph*{Advantages:} 
-this approach is associative and commutative, even self construction
-  can be allowed if the trait requires a static/class method
-  returning This; the class will then implement this method by forwarding
-  a call to the constructor.
-
-\paragraph*{Negatives:} writing the class code with the constructors and
-   fields and getter/setters and factories can be quite tedious.
-   Moreover, there is no way for a trait to specify a default value for a field,
-   the class need to handle all the state, even state that is conceptually
-   "private" of such trait.
-
-Here in the following an exaple of such approach:
-
-\begin{lstlisting}
-  pointSum: { method int x()  method int y()
-    class method This of(int x,int y)
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul: { method int x() method int y()
-    class method This of(int x,int y)
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-\end{lstlisting}
-
-As you see, all the state operations are represented as abstract methods.
-
-\subsection{A First Attempt at Composition}
-According to the general ideas expressed before,
-\begin{lstlisting}
-  Point:Use pointSum,pointMul
-\end{lstlisting}
-
-\noindent would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract.
-In this mindset, the user would be required to write something similar to
-
-\begin{lstlisting}
-  CPoint:Use pointSum,pointMul, {//not our suggested solution
-    int x   int y
-    method int x()x   method int y()y
-    class method This of(int x, int y)
-      new Point(x,y)
-    constructor Point(int x, int y){ this.x=x   this.y=y }
-    }
-\end{lstlisting}
-
-\noindent after a while programming in this style, 
-writing those obvious ``close the state'' classes become a
-repetivie boring job, and one wonder
-if it could be possible to automatically generate such code~\cite{wang2016classless}.
-Indeed those classes are just a form of ``fixpoint''.
-
-In our model we go one step further: there is no need to generate code, or to explicitly
-write down constructors and fields; there is not even syntax for those
-constructor.
-The idea is that any class that ``could" be completed in the obvious way
-\emph{is a complete ``coherent" class}.
-In most other languages, a class is abstract if have abstract methods.
-Instead, we call abstract a class whose set of abstract methods is not
-coherent, that is, can not be automatically recognized as factory,getters and setters.
-
-\paragraph*{Detaild definition of coherent:}
-\begin{itemize}
-\item a class with no abstract method is coherent, and like Java \Q@Math@.
-Will just be usefull for calling class/static methods.
-\item a class with a single abstract \Q@class@ method returning \Q@This@
-is coherent if all the other abstract methods can be seen as \emph{abstract state
-operations} over one of its argument.
-For example,
-if there is a \Q@class method This of(int x, int y)@ as before,
-then
-\item a method \Q@int x()@ is intepreted as an abstract state method: a getter for \Q@x@.
-\item a method \Q@Void x(int that)@ is a setter for x.
-\end{itemize}
-
-While getters and setters are fundamental operations, we can immagine
-more operations to be supported; for example
-\begin{itemize}
-\item \Q@method This withX(int that)@
-may be a ``wither", doing a functional field update.
-\item \Q@method Void update(int x,int y)@
-may do two field update at a time.
-\item\Q@method This clone()@ may do a shallow clone of the object.
-\end{itemize}
-
-We are not sure what is the best set of abstract state operations yet, but we think this could become a very interesting area of research.
-
-  lets play with the points of before, to see what good can we do with the current
-  instruments:
-
-\begin{lstlisting}
- //same code as before for pointSum and pointMul
-  pointSum: { method int x() method int y()
-    class method This of(int x,int y)
-    method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul: { method int x() method int y()//look we are repeating
-    class method This of(int x,int y)//the abstract method declarations.
-    method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-  PointAlgebra:Sum pointSum,pointMul
-\end{lstlisting}
-
-  As you can see, we can declare the methods independently and compose the result
-  as we wish. However we have to repeat the abstract methods \Q@x@,\Q@y@ and \Q@of@.
-  In addition of \Q@Sum@ and \Q@Mul@ we may want many operations over points; can we improve our reuse
-  and not repeat such abstract definitions? of course!
-
-\begin{lstlisting}
-  p: { method int x() method int y()
-    class method This of(int x,int y)
-    }
-  pointSum:Use p, { method This sum(This other)
-      This.of(this.x+other.x,this.y+other.y)
-    }
-  pointMul:Use p, { method This mul(This other)
-      This.of(this.x*other.x,this.y*other.y)
-    }
-  pointDiv: ...
-  PointAlgebra:Use pointSum,pointMul,pointDiv,...
-\end{lstlisting}
-
-Now our code is fully modularized, and each trait handle exactly one method.
-
-What happens if we want to add fields instead of just operations?
-
-\begin{lstlisting}
-  colored:{ method Color color() }
-  Point:Sum pointSum,colored //fails
-\end{lstlisting}
-
-This first attempt does not work: the abstract color method
-is not a getter for any of the parameters of 
-\Q@ class method This of(int x,int y)@
-A solution is to provide a richer factory:
-
-\begin{lstlisting}
-  CPoint:Use pointSum,colored,{
-    class method This of(int x,int y) This.of(x,y,Color.of(/*red*/))
-    class method This of(int x, int y,Color color)
-    }
-\end{lstlisting}
-
-where we assume to support overloading on different parameter number.
-This is a good solution, we think is better that any alternatives in literature,
-however the method \Q@CPoint.sum@ resets the color to red.
-What should be the behaviour in this case?
-If we support withers, instead of writing \Q@This.of()@ we can use
-\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
-Sadly, if we use this design inside of \Q@sum(This other)@ we would loose the color from \Q@other@.
-
-If the point designer could predict this kind of extension, then we could use the following design:  
-\begin{lstlisting}
-  p: { method int x() method int y()
-    method This withX(int that)
-    method This withY(int that)
-    static method This of(int x,int y)
-    method This merge(This other)
-    }
-  pointSum:Use p, { method This sum(This other)
-      this.merge(other).withX(this.x+other.x).withY(this.y+other.y)
-    }
-  colored:{method Color color()
-    method This withColor(Color that)
-    method This merge(This other)
-      this.withColor(this.color().mix(that.color())
-    }
-  CPoint:/*as before*/
-\end{lstlisting}  
-  Now we can merge colors, or any other kind of state we may want to add
-  following this pattern.
-  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
-  need to compose the merge operation inside of both of them.
-  The simple model we are presenting could accomodate this with an
-  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
-This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
-\end{comment}

### 2017-05-25 FACS/RelatedWork.tex

@@
+\saveSpace\saveSpace
 \section{Related Work}
-The literature on code reuse is too vast to let us do justice of it in a fiew pages,
-Our work is clearly inspired on Traits~\cite{ducasse2006traits}; they are in turn insipired on module composition languages~\cite{ancona2002calculus}
-and multiple inheritance~\cite{}.
-
+\saveSpace\saveSpace
+Literature on code reuse is too vast to let us do justice of it in fiew pages,
+Our work is inspired on Traits~\cite{ducasse2006traits}; they are in turn insipired on module composition languages~\cite{ancona2002calculus}.
+\saveSpace
 \subsection{Inheritance and subtyping}
-
+\saveSpace
 We are aware of at least 3 independently designed research languages 
@@
 Leveraging on \emph{traits}, in this work we aim to synthesizes
-the main ideas of those very different language designs.
+the main ideas of those very different designs.
 
 \begin{itemize}
-\item deep+ {\bf A simple uniform syntax for code literals}
+\item DeepFjig+ {\bf A simple uniform syntax for code literals}
 DeepFjig is best in this sense, since DT have separate syntax for class literals, trait literals and record literals.
@@
 all others, where evertyhing is basically just a method.
-\item deep-
+\item DeepFjig-
 {\bf Reusable code can not be ``used'', that is instantiated or used as a type.}

### 2017-05-25 FACS/section2.tex

@@
+\saveSpace\saveSpace\saveSpace
 \section{The Design of \name: Separating Use and Reuse}\label{sec:separate}
-
+\saveSpace\saveSpace\saveSpace
 This section presents the overview of \name, and it illustrates the
@@
 %\subsection{The Design of \name: Decoupling Use from Reuse}
-
+\saveSpace\saveSpace
 \subsection{Classes in \name: A mechanism for code Use}
+\saveSpace
 %\name has a substantially different design from Java-like languages.
@@
 absence of the \Q@extends@ keyword in \name. 
-
-To illustrate classes in \name consider the example in Section~\ref{sec:intro}:
-
+Consider the example in Section~\ref{sec:intro}:
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting} 
-
+\saveSpace
 \noindent Classes in \name use a slightly different declaration style compared
@@
 the Java code, introducing a subclass
-\lstinline{B} is not possible. At first, this may seem like a severe restriction, but
-\name has a different mechanism for code \emph{reuse} that 
-is more appropriate when code reuse is intended.
+\lstinline{B} is not possible. This may seem like a severe restriction, but
+\name has a different mechanism for \emph{code-reuse} that 
+is more appropriate when \emph{code-reuse} is intended.
 
@@
 with traits and code reuse is:
-
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\saveSpace
 \noindent Here \lstinline{ta} is a trait intended to replace the
@@
 \Q@ta@ has no relationship to the type \lstinline{A}.
-
-A Java programmer may then try to write
+A Java programmer may then try to write:
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
+\saveSpace
 But this does not work either: \Q@ta@ is not a type in the first place, since it is a (lowercase) trait name.
@@
 With this in mind, we can try to model the example again:
-
+\saveSpace
 \begin{lstlisting}
@@
  Utils:{class method int m(IA a)/*method body here*/}
- ta:{implements IA
+ ta:{implements IA //This line is the core of the solution
      method int ma() Utils.m(this) }
@@
 \end{lstlisting}
-
+\saveSpace
 This code works: \Q@Utils@ relies on interface \Q@IA@ and the trait \Q@ta@
@@
 \Q@this<:IA@.
- It is also possible to add a \Q@B@ as follows
+ It is also possible to add a \Q@B@ as follows:
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
+\saveSpace
 This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
@@
 no method called \Q@ma@ is explicitly declared.
-DeepFJig\cite{} and DT\cite{} would require instead to explicitly declare an abstract \Q@ma@ as in
+DeepFJig and DT would require instead to explicitly declare an abstract \Q@ma@:
+\saveSpace
 \begin{lstlisting}
@@
       method int mb(){return this.ma();} }
-\end{lstlisting}
-
+\end{lstlisting}\saveSpace
+\noindent
 The idea is that such method is imported from trait \Q@ta@, exactly as in the Java equivalent
-\begin{lstlisting}[language=Java]
+\saveSpace\begin{lstlisting}[language=Java]
   class B extends A{ int mb(){return this.ma();} }  
 \end{lstlisting}
+\saveSpace
 where method \Q@ma@ is imported from \Q@A@.
@@
 
-\paragraph*{Semantic of Use}
+\paragraph{Semantic of Use}
 Albeit alternative semantic models for traits~\cite{Traits:ECOOP2003} have been proposed,
 here we use the flattening model. This means that 
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 A:Use ta
 B:Use ta, { method int mb(){return this.ma();} }
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 
@@
 
- \begin{lstlisting}
+\saveSpace \begin{lstlisting}
 A:{implements IA method int ma() Utils.m(this) }
@@
   method int mb() this.ma() } 
-
  \end{lstlisting}
-
+\saveSpace 
  This code is correct, and {\bf in the resulting code there is no mention of the trait

### 2017-05-25 FACS/section3.tex

@@
 \item or we duplicate the code of the \Q@Bag@ implementation with minimal
-  adjustments in \Q@class Bag5 extends AbstractSetOrBag@;
+  adjustments in \\* \Q@class Bag5 extends AbstractSetOrBag@;
 \item or we introduce a
-\Q@abstract class BagN extends AbstractSetOrBag@ and \Q@class Bag5 extends BagN@
+\Q@abstract class BagN extends AbstractSetOrBag@ and \\*\Q@class Bag5 extends BagN@
 and we modify \Q@Bag@ so that  \Q@class Bag extends BagN@.
@@
 Instead, in \name, if we was to originally declare
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 Set:{/*set implementation*/} 
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 Then our code would be impossible to reuse in the first place for any user of our library.
@@
 }!
-
 If the designer of the set class wishes to make it reusable, it can do it explicitly by providing a set trait:
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 set:{/*set implementation*/} 
 Set:Use set
-\end{lstlisting}
-Notice how since \Q@set@ can never be used as a type, there is no reason to give it a fancy-future-aware name like
+\end{lstlisting}\saveSpace
+Since \Q@set@ can never be used as a type, there is no reason to give it a {\bf fancy-future-aware} name like
 \Q@AbstractSetOrBag@.
 When bag will be added, the code will look either like
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
 set:{/*set implementation*/} 
@@
 Bag: Use set, {/*bag implementation*/}
-\end{lstlisting}
+\end{lstlisting}\saveSpace\saveSpace
 or 
-\begin{lstlisting}
+\saveSpace\saveSpace\begin{lstlisting}
 set:{/*set implementation*/} 
@@
 Bag: Use bag
-\end{lstlisting}
+\end{lstlisting}\saveSpace
 Notice how, thanks to flattening, the resulting code for \Q@Bag@ is identical in both versions, 

### 2017-05-25 FACS/section4.tex

@@
-\section{Improving Reuse}
+\saveSpace\saveSpace\section{Improving Reuse}\saveSpace
 \name allows reuse even when subtyping is impossible.
@@
 instead.
-
+\saveSpace 
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\saveSpace
 \noindent As we can see, all the state operations (the getters for the 
@@
 two different operations on points is:
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
   Point:Use pointSum,pointMul
-\end{lstlisting}
+\end{lstlisting}  \saveSpace
 
@@
 \end{lstlisting}
-\bruno{We talk about withers later on. So I think we should consider
-  having withers in this code, so that readers can understand what 
-withers are!}
-\marco{with withers it will look more complicated}
+%\bruno{We talk about withers later on. So I think we should consider
+ % having withers in this code, so that readers can understand what 
+%withers are!}
+%\marco{with withers it will look more complicated}
 
@@
 
-\paragraph*{Coherent classes} A more detailed definition of coherent
+\paragraph{Coherent classes} A more detailed definition of coherent
 classes is given next:
@@
 \end{itemize}
-
+\noindent
 While getters and setters are fundamental operations, we can immagine
@@
 \lstinline{pointSum} and \lstinline{pointMul} can be directly composed:
-
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}  
-
-  As we can see, we can declare the methods independently and compose the result
+\saveSpace
+\noindent
+  Note how we can declare the methods independently and compose the result
   as we wish.
@@
   abstract definitions into a trait \Q@p@: 
-
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\saveSpace      
 Now the code is fully modularized, and each trait handles exactly one method.
@@
 this would be:
-
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\saveSpace
 This first attempt does not work: the abstract color method
@@
 A solution is to provide a richer factory:
-
+\saveSpace
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\saveSpace
 \noindent 
@@
 a better design would be the following:  
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
   p: { method int x() method int y() //getters
@@
   CPoint:/*as before*/
-\end{lstlisting}  
-  This design allows merging colors, or any other kind of state we may want to add
-  following this pattern.\bruno{worried that withers are not explained
-  enough.}
+\end{lstlisting}  \saveSpace
+  \noindent This design allows merging colors, or any other kind of state we may want to add
+  following this pattern.%\bruno{worried that withers are not explained enough.}
 
@@
 call the version of merge as defined in the second/third element of \use.
-\begin{lstlisting}
+\saveSpace\begin{lstlisting}
   p: {/*as before*/ }
@@
     }
-\end{lstlisting}
+\end{lstlisting}  \saveSpace\saveSpace

### 2017-05-25 FACS/Extensions.tex
.

@@
 Here we see some example of other composition operators, inspired from~\cite{servetto2014meta}:
-\paragraph*{Restrict}${}_{}$\\*
+\paragraph{Restrict}${}_{}$\\*
 Restrict makes a method abstract
@@
 
-\paragraph*{Rename}${}_{}$\\*
+\paragraph{Rename}${}_{}$\\*
 Rename allows to make some form of ``compile time'' refactoring
@@
 
-\paragraph*{Redirect}${}_{}$\\*
+\paragraph{Redirect}${}_{}$\\*
 Redirect allows to emulate generics; the main idea is that a (fully abstract) class can be redirect to another one external to the trait/code literal.
@@
 
-\paragraph*{Application on the expression problem}${}_{}$\\*
+\paragraph{Application on the expression problem}${}_{}$\\*
 With redirect, rename and restrict we can have the general operator propagator
@@
 
-\paragraph*{Full power of redirect}${}_{}$\\*
+\paragraph{Full power of redirect}${}_{}$\\*
 An expressive form of Redirect can be multiple, that is, can redirect may interdependent classes at the same time.

### 2017-05-25 FACS/Intro.tex
Revised RW

@@
 \item {\bf
-    DeepFjig}~\cite{deep,servetto2014meta,DBLP:journals/iandc/LagorioSZ12} is
+    DeepFjig}~\cite{deep,servetto2014meta,fjig} is
 a module composition language where the main idea is that

### 2017-05-25 FACS/RelatedWork.tex
Revised RW

@@
 \saveSpace\saveSpace
-Literature on code reuse is too vast to let us do justice of it in fiew pages,
-Our work is inspired on Traits~\cite{ducasse2006traits}; they are in turn insipired on module composition languages~\cite{ancona2002calculus}.
+Literature on code reuse is too vast to let us do justice of it in a few pages.
+Our work is inspired by traits~\cite{ducasse2006traits}, which in turn
+are inspired by module composition languages~\cite{ancona2002calculus}.
 \saveSpace
@@
 We are aware of at least 3 independently designed research languages 
-that address this limitation:Delta-Trait(DT)\cite{ferruccioworks}, Package Template(PT)\cite{}, DeepFjig\cite{}.
-Leveraging on \emph{traits}, in this work we aim to synthesizes
-the main ideas of those very different designs.
+that address the this-leaking problem: Delta-Trait~(DT)\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}, Package Templates~(PT)\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}, DeepFjig\cite{deep,servetto2014meta,fjig}.
+Leveraging on \emph{traits}, in this work we aim to synthesize
+the best ideas of those very different designs, while at the same time 
+coming up with a simpler and improved design for separating
+subclassing from subtyping. The following compares 
+various aspects of the language designs:
+\bruno{can we say that by borrowing ideas from various designs we come
+up with a simpler design, which also addresses various limitations of
+particular designs?}
 
@@
 \item DeepFjig+ {\bf A simple uniform syntax for code literals}
-DeepFjig is best in this sense, since DT have separate syntax for class literals, trait literals and record literals.
-PT on the other side build on top of full Java, thus has a very involved syntax.
-Thanks to our novel represantation of state, our approach offers a much simpler and uniform syntax then
-all others, where evertyhing is basically just a method.
+DeepFjig is best in this sense, since DT has separate syntax for class literals, trait literals and record literals.
+PT on the other hand is built on top of full Java, thus has a very
+involved syntax.
+\name leverages on DeepFjig's approach but,
+\emph{thanks to our novel representation of state}, \name also offers a much simpler and uniform syntax than
+all other approaches: everyhing is just a method.
 \item DeepFjig-
-{\bf Reusable code can not be ``used'', that is instantiated or used as a type.}
+{\bf Reusable code cannot be ``used'', that is instantiated or used as a type.}
 This happens in DT and in PT, but not in DeepFjig. To allow reusable code to be directly 
 usable, in DeepFjig
-classes introduce nominal types in an innatural way: the type of \Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the nominal type of its class: that is in DeepFjig 
+classes introduce nominal types in an unnatural way: the type of
+\Q@this@ is only \Q@This@ (sometimes called \Q@<>@) and not the
+nominal type of its class. 
+That is in DeepFjig 
 \Q@A:{ method A m()this}@ is not well typed. This is because
-\Q@B: Use A@ flattens to \Q@B:{ method A m()this}@, that is clearly not well typed.
+\Q@B: Use A@ flattens to \Q@B:{ method A m()this}@, which is clearly not well typed.
 Looking to this example is clear why we need reusable code to be agnostic of its name.
-Then, either reusable code have no name (DT, PT and this work)
-or all code is reusable and usable, and all code need to be auckwardly agnostic of its name.
+Then, either reusable code has no name (as in DT, PT and \name)
+or all code is reusable and usable, and all code needs to be awkwardly agnostic of its name.
 
@@
 {\bf Requiring abstract signatures is a left over of module composition mindset.}
-DT and Deepfjig comes form a tradition of functional module composition, where 
-modules are typed in isolation under an environments, and then the composition is performed.
-As we show in this work, this is ends up requiring verbose repetition of abstract signatures, 
-that for hightly modularized code may end up composing most of the program.
-Simple Java (and thus PT, since it is an add on over Java) show us a better way:
+DT and Deepfjig comes from a tradition of functional module composition, where 
+modules are typed in isolation under an environment, and then the composition is performed.
+As we show in this work, this ends up requiring verbose repetition of abstract signatures,
+which (for highly modularized code) may end up constituting most of the program.
+Simple Java (and thus PT, since it is a Java extension) shows us a better way:
 the meaning of names can be understood from the reuse context.
@@
 \item PT-{\bf Composition algebra.}
-The idea of using composition operators over atomic values as in an arithmetic expression is very powerfull,
-and makes easy to extend languages with more operators. Deepfjig and DT embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
-The typing strategy of PT also seams to be connected with this decision, so it would be hard to move their approach in a composition algebra setting.
-\item DT- {\bf Naming your type, even if you have none yet.}
-Both Deepfjig and PT allows a class to refer to its name, abait is more unobvious in PT since you have to introduce both a package and a classes to express it.
-This allows to encode binary methods, express patterns like withers or fluent setters and to instantiate instances of the (future) class(es)  using the reused code.
+The idea of using composition operators over atomic values as in an arithmetic expression is very powerful,
+and makes it easy to extend languages with more operators. Deepfjig and DT embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
+The typing strategy of PT also seems to be connected with this decision, so it would be hard to move their approach in a composition algebra setting.
+\item DT- {\bf Naming your type, even if there is none yet.}
+Both Deepfjig and PT allow a class to refer to its name, albeit this is
+less obvious in PT since both a package and a class have to be introduced to express it.
+This allows encoding binary methods, express patterns like withers or fluent setters and to instantiate instances of the (future) class(es)  using the reused code.
 \item DT+{\bf Complete ontological separation between use and reuse}
-While all 3 works allows to separate inheritance and subtyping only DT properly enforce 
+While all 3 works allow separating inheritance and subtyping only DT properly enforces 
 separation between use (classes and interfaces) and reuse (traits).
-This is because in DeepFjig all classes are both units of use and reuse (however, subtype is not induced)
-and PT imports all the complexity of Java, so abait is possible to separate use and reuse, the model have powerfull but unobvous implications where (conventional Java) \Q@extends@ and PT are used together
+This is because in DeepFjig all classes are both units of use and reuse (however, subtyping is not induced).
+PT imports all the complexity of Java, so although is possible to separate use and reuse, the model have powerful but non-obvious implications where (conventional Java) \Q@extends@ and PT are used together.
 
@@
 other famous work in literature, like~\cite{odersky2008programming,nystrom2006j}
-does no completly break the relation between inheritance and subtyping, but only prevent subtyping only where 
-it would be unsound.
+does not completely break the relation between inheritance and subtyping, but only prevents subtyping where 
+it would be unsound.\bruno{structural typing? Cook?}
 
@@
 
-The idea of abstract state operation has emerged from CJ~\cite{wang2016classless}, and offer a clean solution to handling state
+The idea of abstract state operations emerged from Classless
+Java~\cite{wang2016classless}. This approach offers a clean solution to handle state
 in a trait composition setting.
-Note how abstract state operation are different from just hiding fields under getter and setters: 
-in our model the programmer simply never declare what is the state of the class, not even what information is stored in fields.
+Note how abstract state operations are different from just hiding fields under getter and setters: 
+in our model the programmer simply never has to declare what is the state of the class, not even what information is stored in fields.
 The state is computed by the system as an overall result of the whole code composition process.
 
-In literature there has been many attempts to add state in traits/module composition languages.
+In the literature there have been many attempts to add state in traits/module composition languages:
 \begin{itemize}  
-\item No constructor: all the fields start at null/a default specified value.
+\item {\bf No constructor}: all the fields start with {\bf null} or a default specified value.
   Fields are just like another kind of (abstract) member, and two fields
-  with identical type can be merged by sum/use; \Q@new C()@ can be used for all classes, and \Q@init@ methods may be called later, as in
+  with identical types can be merged by sum/use; \Q@new C()@ can be used for all classes, and \Q@init@ methods may be called later, as in
   \Q@Point p=new Point(); p.init(10,30)@.
 
-  To its credit, this simple approach is commutative and associative and do not disrupt elegance of summing methods.
+  To its credit, this simple approach is commutative and associative and does not disrupt elegance of summing methods.
   However, objects are created "broken" and the user is trusted with fixing them.
@@
     to a functional programming style.
-\item Constructor compose fields:
-In this approach (used by \cite{FJig}) the fields are declared but not initialized, and
-a canonical constrctor constructor (as in FJ) taking a value for each field and just initializing such field
+\item {\bf Constructors compose fields}:
+In this approach (used by \cite{fjig}) the fields are declared but not initialized, and
+a canonical constructor constructor (as in FJ) taking a value for each field and just initializing such field
 is automatically generated in the resulting class.
+It is easy to add fields, however this model is associative but not commutative: composition order influences field order, and thus the construtor signature.
+Self construction is not possible 
+since the signature of the constuctors changes during composition.
 
-It is easy to add fields, however this model is associative but not commutative: composition order influence field order, and thus the construtor signature.
-Self construction is not possible 
-since the signature of the constuctor change during composition.
-
-\item Constructor can be composed if they offers the same exact parameters:
-In this approach (used by DeepFJig) traits declare field and constructors.
-The constructor initialize the fields but can do any other computation.
+\item {\bf Constructors can be composed if they offer the same exact parameters}:
+In this approach (used by DeepFJig) traits declare fields and constructors.
+The constructor initializes the fields but can do any other computation.
 Traits whose constructors have the same signature can be composed.
 The composed constructor will execute both constructor bodies in order.
-
 This approach is designed to allows Self construction.
-It is also associative and mostly commutative: composition order only influence execution order of side effects during construction.
-However constructors composition requires identical constructor signature: this
-hamper reuse, and if a field is added, its initial value need to be
+It is also associative and mostly commutative: composition order only influences execution order of side effects during construction.
+However constructor composition requires identical constructor signatures: this
+hampers reuse, and if a field is added, its initial value needs to be
 somehow syntetized from the constructor parameters.

### 2017-05-25 FACS/RelatedWork.tex
APLAS changes
changes discussed earlier

@@
 \saveSpace
-\subsection{Inheritance and subtyping}
+\subsection{Separating Inheritance and Subtyping}
 \saveSpace
@@
 \begin{itemize}
-\item DeepFjig+ {\bf A simple uniform syntax for code literals}
+\item {\bf A simple uniform syntax for code literals}
 DeepFjig is best in this sense, since DT has separate syntax for class literals, trait literals and record literals.
@@
 all other approaches: everyhing is just a method.
-\item DeepFjig-
+\item 
 {\bf Reusable code cannot be ``used'', that is instantiated or used as a type.}
@@
 
-\item PT+
+\item 
 {\bf Requiring abstract signatures is a left over of module composition mindset.}
@@
 heavy task of integrating with full Java.
-\item PT-{\bf Composition algebra.}
+\item {\bf Composition algebra.}
 The idea of using composition operators over atomic values as in an arithmetic expression is very powerful,
 and makes it easy to extend languages with more operators. Deepfjig and DT embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
-The typing strategy of PT also seems to be connected with this decision, so it would be hard to move their approach in a composition algebra setting.
-\item DT- {\bf Naming your type, even if there is none yet.}
-Both Deepfjig and PT allow a class to refer to its name, albeit this is
-less obvious in PT since both a package and a class have to be introduced to express it.
-This allows encoding binary methods, express patterns like withers or fluent setters and to instantiate instances of the (future) class(es)  using the reused code.
-\item DT+{\bf Complete ontological separation between use and reuse}
+The typing strategy of PT also seems to be connected with this
+decision, so it would be hard to move their approach in a composition
+algebra setting.
+\item {\bf Complete ontological separation between use and reuse}
 While all 3 works allow separating inheritance and subtyping only DT properly enforces 
@@
 PT imports all the complexity of Java, so although is possible to separate use and reuse, the model have powerful but non-obvious implications where (conventional Java) \Q@extends@ and PT are used together.
+\item {\bf Naming your type, even if there is none yet.}
+Both Deepfjig and PT allow a class to refer to its name, albeit this is
+less obvious in PT since both a package and a class have to be introduced to express it.
+This allows encoding binary methods, express patterns like withers or fluent setters and to instantiate instances of the (future) class(es)  using the reused code.
 
@@
 \begin{itemize}  
-\item {\bf No constructor}: all the fields start with {\bf null} or a default specified value.
+\item An early approach is to have {\bf no constructors}: all the fields start with {\bf null} or a default specified value.
   Fields are just like another kind of (abstract) member, and two fields
@@
 
+\subsection{ThisType with Subclassing implying Subtyping}
+
+Various Java/FJ extensions with ThisType here.
+
+\marco{Cite some work of bruceTHisType and show how he also fails to separe reuse and use}
+
+\begin{comment}
 \subsection{Tablular comparision of many approaches}
@@
 }
-
-\marco{Cite some work of bruceTHisType and show how he also fails to separe reuse and use}
+\end{comment}

### 2017-05-25 FACS/paper.tex
.
APLAS changes
changes discussed earlier

@@
 \setlist{nosep}
-\setlist{leftmargin=0ex}
+\setlist{leftmargin=1ex}
 \setlist[itemize]{noitemsep,nolistsep}
 \setlist[enumerate]{noitemsep,nolistsep}
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\vspace{-5pt}\oldparagraph{#1}\vspace{-5pt}}
+%\let\oldparagraph\paragraph
+%\renewcommand{\paragraph}[1]{\vspace{-5pt}\oldparagraph{#1}}
+\renewcommand{\paragraph}[1]{\noindent\emph{#1}}
+
 \let\oldsubsection\subsection
@@
 
+\begin{comment}
 \author{Marco Servetto\inst{1} \and Bruno C. d. S. Oliveira\inst{2}}
@@
 The University of Hong Kong, Hong Kong\\}
+\end{comment}
+\author{Authors omitted for double-bind review.}
 
@@
 addressing this limitation:Delta-Trait, Package Template, DeepFjig.
-Here we synthesize
+The goal of this paper is to synthesize and improve on
 the main ideas of those very different designs into a nominally typed minimalist language,
@@
 
-\keywords{Code Reuse, Object-Oriented Programming}
+%\keywords{Code Reuse, Object-Oriented Programming}
 \saveSpace\saveSpace\saveSpace

### 2017-05-25 FACS/Intro.tex
changes discussed earlier
Intro

@@
 \noindent does two things at the same time:
-1) it {\bf inherits} code from \lstinline{B}; and it {\bf creates
+it {\bf inherits} code from \lstinline{B}; and it {\bf creates
 a subtype} of \emph{B}. Therefore in a language like Java 
@@
 
-Type systems based on structural typing can deal with the first
+Type systems based on structural typing~\cite{cook} can deal with the first
 situation well, but not the second. Since structural subtyping
@@
 Method \lstinline{A.ma} passes \lstinline{this} as \lstinline{A} to \Q@Util.m@.
-This code seems correct, and there is no subtyping/subclassing.  Now, lets add a class \Q@B@
+This code is correct, and there is no subtyping/subclassing.  Now, lets add a class \Q@B@
 
@@
 %%\lstinline{A}; and 2) it creates a {\bf subtype} of \lstinline{A}.
-We can see an invocation of \lstinline{A.ma} inside
+\noindent We can see an invocation of \lstinline{A.ma} inside
 \lstinline{B.mb}, where the self-reference \lstinline{this} is of type \lstinline{B}. 
 The execution will eventually call \lstinline{Util.m} with an
-instance of \lstinline{B}. However, this can be correct only if \lstinline{B} is a subtype of
-\lstinline{A}.
+instance of \lstinline{B}. However, \emph{this can be correct only if \lstinline{B} is a subtype of
+\lstinline{A}}.
 
@@
 Therefore, every OO language with the minimal features exposed in the example (using \lstinline{this},
-extends and method calls) is forced to accept that subclassing implies
+{\bf extends} and method calls) is forced to accept that subclassing implies
 subtyping\footnote{C++ allows to "extends privately", but it is a limitation over
@@
 
-In essence we believe that in languages like Java classes do too many
+What the this-leaking problem shows is that adopting a more flexible
+nominally typed OO model where subclassing does not imply subtyping is
+not trivial, and a more substantial change in the language design is
+necessary.  In essence we believe that in languages like Java classes do too many
 things at once. In particular they act both as units of \emph{use} and
@@
 classes can also be subclassed to provide \emph{reuse} of code.
-
-What the this-leaking problem shows is that adopting a more flexible
-nominally typed OO model where subclassing does not imply subtyping is
-not trivial, and a more substantial change in the language design is
-necessary.  We are aware of at least 3 independently designed research
+We are aware of at least 3 independently designed research
 languages that address this limitation:
@@
 interfaces induce types, records express state and traits are reuse units.
-\item {\bf Package Templates (PT)}~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}:
-are an extension of (full) Java where new packages can be ``syntetized'' by mixing
-and integrating code templates. As an extension of java, PT allows but not requires
+\item {\bf Package Templates (PT)}~\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}
+are an extension of (full) Java where new packages can be ``synthesized'' by mixing
+and integrating code templates. As an extension of java, PT allows but not require
 separation of inheritance and subtyping.
@@
 \end{itemize}
-\bruno{We need here a better motivation. From the reviewer perspactive
-he will think: ``So, there are already 3 solutions, why should I care
-about a new one?''}
-This paper aims at showing a simple language design, called \name, to
-completely decouple subtyping and subclassing in a nominally typed OO
-language. The key idea is to divide between code designed for
-\textbf{USE} and code designed for \textbf{REUSE}. 
+This paper aims at showing a simple language design, called \name.
+Leveraging on traits, in this work we aim to synthesize the best ideas
+of those 3 very different designs, while at the same time coming up with a simpler and
+improved design for separating subclassing from subtyping.
+The keys ideas in \name are to divide between code designed for
+\textbf{USE} and code designed for \textbf{REUSE}, and a novel
+approach to state in traits that avoids the complexities introduced by
+fields and their initialization.
 In \name there are two separate concepts: classes
@@
 brings several benefits in terms of flexibility and code reuse.
+
+We focus on an example-driven presentation to illustrate how to
+improve use and reuse. Most of the hard technical aspects of the
+semantics have been studied in previous 
+work~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282,KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12,deep,servetto2014meta,fjig},
+and the design of \name is a simplification of the previously studied
+concepts. The language design ideas have been implemented in the 42 language, which supports all
+the examples we show in the paper, and is available at:
+
+\url{http://l42.is}
+
 In summary, our contributions are:

### 2017-05-25 FACS/paper.tex
Intro

@@
 %
+\usepackage{url}
 \usepackage{makeidx}  % allows for indexgeneration
@@
 We are aware of at least 3 independently designed research languages 
-addressing this limitation:Delta-Trait, Package Template, DeepFjig.
+addressing this limitation: \emph{Delta-Trait}, \emph{Package Template}, \emph{DeepFjig}.
 The goal of this paper is to synthesize and improve on
-the main ideas of those very different designs into a nominally typed minimalist language,
+the main ideas of those very different designs into a nominally typed
+minimalist language, called \name, which is quite
 natural to OO programmers.

### 2017-05-25 FACS/Intro.tex
Section 3

@@
 nominally typed OO model where subclassing does not imply subtyping is
-not trivial, and a more substantial change in the language design is
-necessary.  In essence we believe that in languages like Java classes do too many
+not trivial: a more substantial change in the language design is
+necessary.  In essence we believe that, in languages like Java, classes do too many
 things at once. In particular they act both as units of \emph{use} and
@@
 We are aware of at least 3 independently designed research
-languages that address this limitation:
+languages that address the this-leaking problem:
 \begin{itemize}
@@
 are an extension of (full) Java where new packages can be ``synthesized'' by mixing
-and integrating code templates. As an extension of java, PT allows but not require
+and integrating code templates. As an extension of Java, PT allows but not require
 separation of inheritance and subtyping.
@@
 work~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282,KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12,deep,servetto2014meta,fjig},
-and the design of \name is a simplification of the previously studied
-concepts. The language design ideas have been implemented in the 42 language, which supports all
+and the design of \name synthesises some of the previously studied
+concepts. We briefly sketch the key ideas of the formalization of
+\name in Section~\ref{sec:formal}.
+The language design ideas have been implemented in the 42 language, which supports all
 the examples we show in the paper, and is available at:

### 2017-05-25 FACS/section2.tex
Section 3

@@
 This concept is natural for a Java programmer, but was not supported
-in previous work \cite{Damiani2014,deep}. Those works require all
+in previous work \cite{BETTINI2013521,deep}. Those works require all
 dependencies in code literals to be explicitly declared, so that the
 code literal is completely self-contained. However, this results in
-many spurious abstract method declarations.
+many redundant abstract method declarations.
 
@@
 
-\noindent reduces/is equivalent to/is flatted into
+\noindent is equivalent to:
 

### 2017-05-25 FACS/section3.tex
Section 3

@@
 duplication, but \emph{we induced unintented subtyping}! 
-Since subclassing implies subtyping, our code break Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
+Since subclassing implies subtyping, our code breaks the Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
 Indeed, the following is allowed:
@@
 and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
-Worst, if we are to constantly apply this mentalty, we would introduce a very high number
+Worst, if we are to constantly apply this mentaly, we would introduce a very high number
 of abstract classes that are not supposed to be used as types. Those classes would clutter the 
@@
 In our example the information \Q@Set<:AbstractSetOrBag@ would be present in the public interface
-of the class \Q@Set@, but it is not needed to use the class properly!
+of the class \Q@Set@, but such information is not needed to use the class properly!
 
@@
 \begin{itemize}
-\item either we insert \Q@class Bag5 extends Bag@ and we break LSP; 
+\item either we insert \Q@class Bag5 extends Bag@ and we break the LSP; 
 \item or we duplicate the code of the \Q@Bag@ implementation with minimal
@@
 ``Design and document for inheritance or else prohibit
-it''\cite{Bloch08}: the
+it''~\cite{Bloch08}: the
 self use of public methods is rarelly documented, thus is hard to understand the effects of overriding a library metod.

### 2017-05-29 FACS/Examples.tex
Local changes

@@
 \saveSpace\saveSpace
-In this article we dedicate more space to examples and informal presentation and motivations;
-so we do not have space to provide a full formalizations.
-We will provide here some hints on how the formalization works.
+
+This section sketches some of the key ideas in \name.
+
+%In this article we dedicate more space to examples and informal presentation and motivations;
+%so we do not have space to provide a full formalizations.
+%We will provide here some hints on how the formalization works.
+
+\subsection{Syntax}
 
@@
 
-\paragraph{Well-formedness}
+\subsection{Well-formedness}
 Basic well formedness rules apply:

### 2017-05-29 FACS/section4.tex
Local changes

@@
 In this way, \name allows reuse even when subtyping is
-impossible, that is, when a method parameter has type \Q@This@ (also called \emph{binary
-  methods}~\cite{bruce96binary}). 
+impossible, that is, when a method parameter has type \Q@This@ (also
+called a \emph{binary
+  method}~\cite{bruce96binary}). 
 This type of situations is the primary motivator
 for previous work aiming at separating inheritance from subtyping.
-Leveraging on the \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smootly integrate state and traits: a challenging problem that has limited the flexibility of traits and
+Leveraging on the \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smoothly integrate state and traits: a challenging problem that has limited the flexibility of traits and
 reuse in the past.
@@
 composition languages~\cite{ancona2002calculus} and several trait
-models~\cite{Traits:ECOOP2003,Bergel2007,BETTINI2013521,DBLP:journals/iandc/LagorioSZ12}.  However the research
-community is struggling to make it work with object state (constructor
+models~\cite{Traits:ECOOP2003,Bergel2007,BETTINI2013521,fjig}.  However the research
+community is struggling to make it work with object state (constructors
 and fields) while achieving the following goals:
 
-%\begin{itemize}
+\begin{itemize}
 %complicated discussions on this point \item keep sum associative and commutative,
-%\item
-(1) managing fields in a way that borrows the elegance of summing methods;
-%\item
-(2) actually initialize objects, leaving no null fields;
-%\item
-(3) making it easy to add new fields;
-%\item
-(4) allowing a class to create instances of itself.
-%\end{itemize}
+\item managing fields in a way that borrows the elegance of summing methods;
+\item actually initialize objects, leaving no null fields;
+\item making it easy to add new fields;
+\item allowing a class to create instances of itself.
+\end{itemize}
 
 In the related work we will show some alternative ways to handle
-state.  However the purest solution just requires methods. The idea is
+state.  However the purest solution requires methods only. The idea is
 that the trait code just uses getter/setters/factories, while leaving
 to classes the role to finally define the fields/constructors. That
-is, the class has syntax richer that the trait one, allowing
+is, classes have syntax richer than traits, allowing
 declaration for fields and constructors.  This approach is very
@@
 
-\paragraph{Immutable Points} Consider, for example, two simple 
-traits that deal with \emph{immutable point} objects. That is, points
+\paragraph{Modeling Points} Consider, for example, two simple 
+traits that deal with \emph{point} objects. That is, points
 in the cartesian plane (with coordinates \lstinline{x} and
@@
 for points:
-A class method is similar to a \Q@static@ method in Java but can be abstract. 
+a class method is similar to a \Q@static@ method in Java but can be abstract. 
 As for instance methods, they are late bound:  flattening can provide an implementation for them.
-abstract class method are very similar to the original concept of member functions in the module composition setting.
+Abstract class methods are very similar to the original concept of member functions in the module composition setting.
 
-\subsection{A First Attempt at Composition}
+\subsection{A First Attempt at Composition with a Traditional Model of
+Traits}
 According to the general ideas about trait composition presented in
@@
 
-\noindent However this fails since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract!
+\noindent However \emph{in a traditional trait model} this would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract!
 Instead a user could write something similar to:
@@
 
-\subsection{The \name approach to State}
+\subsection{The \name Approach to State: Coherent Classes}
 
@@
 \noindent
-While getters and setters are fundamental operations, we can immagine
-more operations to be supported; for example:
+While getters and setters are fundamental operations, it is possible to
+support more operations. For example:
 \begin{itemize}
@@
 \item \Q@method Void update(int x,int y)@
-may do two field update at a time.
+may do two field updates at a time.
 \item\Q@method This clone()@ may do a shallow clone of the object.
@@
 In \name and with our approach to handle the state, 
-\lstinline{pointSum} and \lstinline{pointMul} can be directly composed:
+\lstinline{pointSum} and \lstinline{pointMul} can indeed be directly composed:
 \saveSpace
 \begin{lstlisting}
-  PointAlgebra:Use pointSum,pointMul
+  PointAlgebra:Use pointSum,pointMul //coherent class completion!
 \end{lstlisting}  
@@
 
-  \paragraph{Improved solution} In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
+  \paragraph{Improved solution} So far the current solution still
+  repeats the abstract methods \Q@x@, \Q@y@ and \Q@of@.
   Moreover, in addition to \Q@sum@ and \Q@mul@ we may want many
@@
   colored:{ method Color color() }
-  Point:Use pointSum,colored //Failure: resulting class not coherent
+  Point:Use pointSum,colored //Fails: class not coherent
 \end{lstlisting}
@@
 value.  What should be the behaviour in this case?  If our abstract
-state supports withers, instead of writing \Q@This.of(...)@ we can use
-\Q@this.withX(newX).withY(newY)@ in order to preserve the color from
+state supports withers, we can use
+\Q@this.withX(newX).withY(newY)@, instead of writing \Q@This.of(...)@, in order to preserve the color from
 \Q@this@.  This solution is still not satisfactory: this design ignores

### 2017-05-29 FACS/Examples.tex
Section 5

@@
 
-This section sketches some of the key ideas in \name.
+This section sketches some of the key ideas in \name more
+formally. \bruno{Could we point out here a few references of key 
+ideas that were presented in other work? This would justify the 
+formalization sketch better, I think.}
 
@@
 
-In the following, we present a very simplified grammar:
+In the following, we present a simplified grammar of \name:
 %\begin{bnf}
@@
 To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
-expression.  Traits come with various operators (restrict, hide,
-alias) but in this work we focus on the single operator 
+expression.  Traits in full 42 come with various operators (restrict, hide,
+alias) but in \name we focus on the single operator 
 $\use$, taking a set
 of code values: that is trait names \mt\ or literals \mL\ and composing them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
-trait composition operator.  $\use\ \overline\mV$ composes the content of $\overline\mV$
-by taking the union of the methods and the union of the implementations.
+trait composition operator~\bruno{Maybe some references here would
+  help}.
 
-\use\ can not be applied if multiple versions of the same method are
+$\use\ \overline\mV$ composes the content of $\overline\mV$
+by taking the union of the methods and the union of the
+implementations.\bruno{Also, for Use, wasn't this similar to your
+  previous work in FJig/DeepFJig?}
+\use\ cannot be applied if multiple versions of the same method are
 present in different traits.  An exception is done for abstract methods:
@@
 
-Methods \mMD~can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very usefull in the context of code composition.
+Methods \mMD~can be instance methods or \Q@class@ methods. A class method is similar to a \Q@static@ method in Java but can be abstract. This is very useful in the context of code composition.
 To denote a method as abstract, instead of an optional keyword we just omit the implementation \me.
@@
 \subsection{Well-formedness}
-Basic well formedness rules apply:
+In \name the basic well formedness rules apply:
 \begin{itemize}
@@
 \item only interfaces can be implemented.
-\item \marco{I'm sure I'm missing something}
 \end{itemize}
@@
 The are two core ideas:
-\paragraph{1: traits are \emph{well-typed} before being reused.}${}_{}$\\*
+
+\paragraph{1: traits are \emph{well-typed} before being reused.}
  For example in
@@
 an \mL\ in $\overline\mV$ is not typechecked before flattening, and only the result is expected to be well-typed.
-While this seems a very dangerous approach at first, consider that also Java have the same behaviour:
+While this seems a dangerous approach at first, consider that also Java have the same behaviour:
 for example in
@@
 \end{lstlisting}\saveSpace
-\noindent would be correct: even if n,m are not defined inside
+\noindent would be correct. In the code literal
 \Q@{method int k() this.n()+this.m()}@, 
-the result of the flattening is well typed.
-
+ even if \Q@n@, \Q@m@ are not locally defined, in 
+\name the result of the flattening is well typed.
 This is not the case in many similar works in literature~\cite{deep,Bettini2015282,Bergel2007} where the
-literals have to be self complete. In this case we would have been forced to
-declare abstract methods \Q@n@ and \Q@m@.
+literals have to be \emph{self complete}. In this case we would have been forced to
+declare abstract methods \Q@n@ and \Q@m@, even if \Q@t@ already 
+provides such methods.
 
@@
 \begin{itemize}
-\item If a class is declared by using $\mC : \use\ \overline\mt$, that is, without literals,
-and the flattening is successfull, \mC\ is well typed, no need of further checking.
+\item If a class is declared by using $\mC : \use\ \overline\mt$, that is without literals,
+and the flattening is successfull then \mC\ is well typed: there is no need of further checking.
 \item On the other side, if a class is declared by $\mC : \use\ \overline\mV$, with
@@
 It may be that the result is intrinsically ill-typed, if one of the methods in $\mL_1\ldots\mL_n$ is not well typed,
-but it may also happen that a type referred from one of those methods is declared \emph{after} the current class.
-As we will see later, this is how our relaxation allows to support (indirectly) recursive types.
+but it may also happen that a type referred from one of those methods
+is declared \emph{after} the current class. As we will see later, this
+is how our relaxation allows us to support (indirectly) recursive types.
 
@@
 OO languages leverage on recursive types most of the times.
-For example in a pure OO language, \Q@String@ may offers a \Q@Int size()@
+For example in a pure OO language, \Q@String@ may offer a \Q@Int size()@
 method, and \Q@Int@ may offer a \Q@String toString()@ method.
-
 This means that is not possible to type in (full) isolation classes
@@
 Each group may also depend from a number of other groups.
-This would form a Direct Aciclyc Graph of groups.
+This would form a Direct Acyclic Graph of groups.
 To type a group, we first need to type all depended groups, then
@@
 happen one at the time, and classes are typechecked where their type is first needed.
-That is, In our approach typing and flattening interleaves. We assume our compilation process to stop as soon as 
-an error arise. There are two main kinds of errors: Type errors (like method not found) or Composition errors (like summing two conflicting implementation for the same method).
+That is, in \name typing and flattening are interleaved. We assume our compilation process to stop as soon as 
+an error arises. There are two main kinds of errors: \emph{type errors} (like method not found) or \emph{composition errors} (like summing two conflicting implementation for the same method).
 For example
@@
 In this scenario, since we go top down, we first need to generate \Q@B@.
-To generate \Q@B@, we need to use \Q@tb@;
+To generate \Q@B@, we need to use \Q@tb@.
 In order to modularly ensure well typedness,
@@
 a type error could be generated at this stage.
-In this moment, \Q@A@ can not be compiled/checked alone,
-we need informations about \Q@B@, but \Q@A@ is not used in \Q@tb@,
+At this moment, \Q@A@ cannot be compiled/checked alone:
+information about \Q@B@ is needed, but \Q@A@ is not used in \Q@tb@,
 thus we do not need to type \Q@A@ and we can type \Q@tb@ with
@@
 Now, we need to generate \Q@C@, and we need to ensure well typedness of \Q@tc@.
-Now \Q@B@ is alreay well typed (since generated by \use\ \Q@tb@, with no \mL),
-and \Q@A@ can be typed;  finally \Q@tc@ can be typed and used.
+Now \Q@B@ is already well typed (since generated by \use\ \Q@tb@, with no \mL),
+and \Q@A@ can be typed. Finally \Q@tc@ can be typed and used.
 If \use\ could not be performed (for example it \Q@tc@ had a method \Q@hello@ too)
@@
 to type \Q@tc@ we need to know the type of \Q@A@ and \Q@B@.
-But they are both unavailable: \Q@B@ is still not computed and 
+However they are both unavailable: \Q@B@ is still not computed and 
 \Q@A@ can not be compiled/checked without information about \Q@B@.
@@
 first check that all can go well, and then perform the flattening.
-This would however be overcompicated for no observable difference:
+This would however be overcomplicated without any observable difference:
 Indeed, in the \Q@A,B,C@ example above there is no difference
@@
 In particular, our approach allows for safe libraries of traits and classes to be fully typechecked, deployed and reused by multiple clients: no type error will emerge from library code.
-On the other side, we do not enforce the programmer to write always self-contained code where all the abstract method definition are explicitly declared.
+On the other side, we do not enforce the programmer to always write self-contained code where all the abstract method definition are explicitly declared.

### 2017-05-29 FACS/Examples.tex
Section 5

@@
 an \mL\ in $\overline\mV$ is not typechecked before flattening, and only the result is expected to be well-typed.
-While this seems a dangerous approach at first, consider that also Java have the same behaviour:
+While this seems a dangerous approach at first, consider that also Java has the same behaviour:
 for example in
@@
 This is not the case in many similar works in literature~\cite{deep,Bettini2015282,Bergel2007} where the
-literals have to be \emph{self complete}. In this case we would have been forced to
+literals have to be \emph{self contained}. In this case we would have been forced to
 declare abstract methods \Q@n@ and \Q@m@, even if \Q@t@ already 

### 2017-05-29 FACS/section2.tex
typo

@@
 for \lstinline{A}, the name of the class is \lstinline{A} and the code 
-literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
+literal associated with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
 class.

### 2017-05-29 FACS/Examples.tex
Section 5
.

@@
 This section sketches some of the key ideas in \name more
-formally. \bruno{Could we point out here a few references of key 
-ideas that were presented in other work? This would justify the 
-formalization sketch better, I think.}
+formally. 
+A similarly interleaving approach can be seen in\cite{servetto2014meta} fully formalized.
 
@@
 
-To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
-expression.  Traits in full 42 come with various operators (restrict, hide,
+\noindent To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
+expression.  In full 42 we support with various operators (restrict, hide,
 alias) but in \name we focus on the single operator 
 $\use$, taking a set
-of code values: that is trait names \mt\ or literals \mL\ and composing them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
-trait composition operator~\bruno{Maybe some references here would
-  help}.
+of code values: that is trait names \mt\ or code literals \mL\ and composing them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
+trait composition operator~\cite{ducasse2006traits}.
 
@@
 by taking the union of the methods and the union of the
-implementations.\bruno{Also, for Use, wasn't this similar to your
-  previous work in FJig/DeepFJig?}
+implementations.
 \use\ cannot be applied if multiple versions of the same method are

### 2017-05-29 FACS/Conclusion.tex
.

@@
 Formalization (in progress) for such language can be found at 
-\verb@github.com/ElvisResearchGroup/L42/tree/master/Main/formal@.
-L42 extands our model allowing
+\verb@urlOmittedForDoubleBlindReview@.
+%\verb@github.com/ElvisResearchGroup/L42/tree/master/Main/formal@.
+L42 extends our model allowing
 flattening to execute arbitrary computations.
 In such model we do not need an explicit notion of traits: they are encoded as method returning a code literal.
-L42 also have feature less related to code composition, like
+L42 also have features less related to code composition, like
   a strong type system supporting aliasing mutability and circularity control,

### 2017-05-29 FACS/RelatedWork.tex
.

@@
 coming up with a simpler and improved design for separating
-subclassing from subtyping. The following compares 
+subclassing from subtyping, which also addresses various limitations of those
+3 particular language designs.
+The following compares 
 various aspects of the language designs:
-\bruno{can we say that by borrowing ideas from various designs we come
-up with a simpler design, which also addresses various limitations of
-particular designs?}
 
@@
 heavy task of integrating with full Java.
+Recent versions of Delta-Traits are trying to address this issue too.
 \item {\bf Composition algebra.}
@@
 \end{itemize}
-With the exception of those 3 lines of work, to the best of our understading
-other famous work in literature, like~\cite{odersky2008programming,nystrom2006j}
-does not completely break the relation between inheritance and subtyping, but only prevents subtyping where 
-it would be unsound.\bruno{structural typing? Cook?}
 
@@
 In this approach (used by \cite{fjig}) the fields are declared but not initialized, and
-a canonical constructor constructor (as in FJ) taking a value for each field and just initializing such field
+a canonical constructor (as in FJ) taking a value for each field and just initializing such field
 is automatically generated in the resulting class.
@@
 \subsection{ThisType with Subclassing implying Subtyping}
-
-Various Java/FJ extensions with ThisType here.
-
-\marco{Cite some work of bruceTHisType and show how he also fails to separe reuse and use}
+With the exception of those mentioned 3 lines of work, to the best of our understading
+other famous work in literature, like~\cite{odersky2008programming,nystrom2006j}
+does not completely break the relation between inheritance and subtyping, but only prevents subtyping where 
+it would be unsound.
+Recently work on ThisType \cite{ToplasAndIgarashi}
+also continue on this line. In those works, ``subtyping by subclassing'' is considered a valuable feature.
 

### 2017-06-06 FACS/Intro.tex
Revision

@@
 In mainstream OO languages like Java, C++ or C\# subclassing 
-implies subtyping. For example, in Java a subclass definition, such as 
+implies subtyping. For example, a Java subclass definition, such as 
 \Q@class A extends B {}@
@@
 \end{itemize}
-This paper aims at showing a simple language design, called \name.
+This paper aims at showing a simple language design, called \name,
+which addresses the this-leaking problem and decouples subtyping from inheritance.
 Leveraging on traits, in this work we aim to synthesize the best ideas
@@
 The keys ideas in \name are to divide between code designed for
-\textbf{USE} and code designed for \textbf{REUSE}, and a novel
+\textbf{use} and code designed for \textbf{reuse}, and a novel
 approach to state in traits that avoids the complexities introduced by
-fields and their initialization.
+fields and their initialization in previous approaches.
 In \name there are two separate concepts: classes
@@
 \name in Section~\ref{sec:formal}.
-The language design ideas have been implemented in the 42 language, which supports all
+The language design ideas have been implemented in the {\bf 42 language}, which supports all
 the examples we show in the paper, and is available at:
@@
   simple language design}. This language is the logic core of the language 42, and 
-  all the examples in the paper can be directly encoded as valid 42 programs.
+  all the examples in the paper can be encoded as valid 42 programs.
 

### 2017-06-06 FACS/paper.tex
Section 3
Revision

@@
 redundant abstract declarations required in Delta-Trait and
-DeepFjig. At the same time self construction,
-binary methods and recursive types are also supported.
+DeepFjig. At the same time \emph{self construction},
+\emph{binary methods} and \emph{recursive types} are also supported.
 Moreover, we provide a novel and elegant solution to uniformly
@@
 These ideas have been implemented in the 42 language, 
-that supports all the examples we show in the paper.
+which supports all the examples we show in the paper.
 \saveSpace\saveSpace\saveSpace

### 2017-06-06 FACS/section3.tex
paper

@@
 Set and Bag collections first in Java, and then in \name.
-The main benefit of the \name solution is that we can get reuse 
-without introducing subtyping between Bags and Sets. This improves the 
+The main benefit of the \name solution is that \emph{we can get reuse 
+without introducing subtyping between Bags and Sets}. This improves the 
 use of Bags and eliminates logical errors arizing from incorrect
@@
 \subsection{Sets and Bags in \name}
-Instead, in \name, if we was to originally declare
+Instead, in \name, if we were to originally declare
 \saveSpace\begin{lstlisting}
@@
 }!
-If the designer of the set class wishes to make it reusable, it can do it explicitly by providing a set trait:
+If the designer of the set class wishes to make it reusable, he/she can do it explicitly by providing a set trait:
 \saveSpace\begin{lstlisting}

### 2017-06-06 FACS/section4.tex
Revision
paper

@@
 \name allows reuse even when subtyping is impossible.
-In \name traits do not induce a new (externally visible) type.
-However, locally in the trait the programmer can use the special self-type \Q@This@ in order to denote the 
+\name traits do not induce a new (externally visible) type.
+However, locally in a trait, programmers can use the special self-type \Q@This@~\cite{bruce_1994,Saito:2009,ryu16ThisType} in order to denote the 
 type of \Q@this@.
-That is: the program is agnostic to what the \Q@This@ type is going to be, so it can
+That is, a program is agnostic to what the \Q@This@ type is, so that it can
 be later assigned to any (or many) classes. 
@@
 In this way, \name allows reuse even when subtyping is
-impossible, that is, when a method parameter has type \Q@This@ (also
-called a \emph{binary
-  method}~\cite{bruce96binary}). 
+impossible. For example for \emph{binary
+  methods}~\cite{bruce96binary} where the method parameter has type \Q@This@. 
 This type of situations is the primary motivator
-for previous work aiming at separating inheritance from subtyping.
+for previous work aiming at separating inheritance from subtyping~\cite{cook}.
 Leveraging on the \Q@This@ type, we can also provide self-instantiation (trait methods can create instances of the class using them) and smoothly integrate state and traits: a challenging problem that has limited the flexibility of traits and
@@
 As for instance methods, they are late bound:  flattening can provide an implementation for them.
-Abstract class methods are very similar to the original concept of member functions in the module composition setting.
+Abstract class methods are very similar to the original concept of member functions in the module composition setting~\cite{ancona2002calculus}.
 
@@
 
-\noindent However \emph{in a traditional trait model} this would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract!
+\noindent However \emph{in a traditional trait model}~\cite{Traits:ECOOP2003} this would fail since methods \Q@x@,\Q@y@ and \Q@of@ are still abstract!
 Instead a user could write something similar to:
@@
 
-\item {\bf Disadvantages:} Writing those obvious definitions to close
+\item {\bf Disadvantages:} Writing such obvious definitions to close
   the state/fixpoint in the class 
-   with the constructors and fields and getter/setters and factories can be quite tedious.
+   with the constructors and fields and getter/setters and factories is tedious.
    %Moreover, there is no way for a trait to specify a default value for a field,
@@
    %"private" of such trait. 
-   Previous work shows that such code can be automatically generated~\cite{wang2016classless}.
+   Previous work shows that such code can be automatically
+   generated~\cite{wang2016classless}.
+   Also a model with fields and constructors is necessarily
+   more complex than a model with methods only.
 
@@
 \begin{lstlisting}
-  PointAlgebra:Use pointSum,pointMul //coherent class completion!
+  //Same code as before! Works because resulting class is coherent.
+  PointAlgebra:Use pointSum,pointMul 
 \end{lstlisting}  
@@
 
-In the following you can see how to mix colors and flavours. The syntax \Q@_2merge@ and   \Q@_3merge@
+The following code shows how to mix colors and flavours. The syntax \Q@_2merge@ and   \Q@_3merge@
 call the version of merge as defined in the second/third element of \use.
@@
     method This withFlavor(Flavor that)
-    method This merge(This that) //how to merge flavors handled here
-      this.withFlavor(that.flavor())}//here we just inherit ``that'' flavor
+    method This merge(This that) //merging flavors handled here
+      this.withFlavor(that.flavor())}//inherits ``that'' flavor
   FCPoint:Use pointSum,colored,flavored{
@@
     class method This of(int x, int y,Color color,Flavor flavor)
-    //we can resolve the conflict about two implementations for merge
+    //resolves the conflict about two implementations for merge
     //by proving our own implementation here

### 2017-06-06 FACS/Examples.tex
Section 5

@@
 This section sketches some of the key ideas in \name more
-formally. 
-A similarly interleaving approach can be seen in\cite{servetto2014meta} fully formalized.
+formally. A similarly interleaving approach can be seen in~\cite{servetto2014meta} fully formalized.
 
@@
 \noindent To declare a trait \mTD\ or a class \mCD, we can use either a code literal \mL\ or a trait
-expression.  In full 42 we support with various operators (restrict, hide,
-alias) but in \name we focus on the single operator 
-$\use$, taking a set
-of code values: that is trait names \mt\ or code literals \mL\ and composing them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
+expression.  In full 42 we support various operators (restrict, hide,
+alias), but in \name we focus on the single operator 
+$\use$. $\use$ takes a set
+of code values: that is trait names \mt\ or code literals \mL\ and composes them.  This operation, sometimes called \emph{sum}, is the simplest and most elegant
 trait composition operator~\cite{ducasse2006traits}.
@@
 Those rules can be applied on any given \mL~individually and in full isolation.
-
 We expect the type system to enforce: 
@@
 or a rigid module composition~\cite{ancona2002calculus}.
-
-The are two core ideas:
+There are two core ideas:
 
@@
 \noindent in \Q@B@ we can call \lstinline{this.ma()} even if in the curly braces there is no declaration for \Q@ma()@.
-In our example, using the trait \Q@t@ of before
+In our example, using the trait \Q@t@
 
@@
 OO languages leverage on recursive types most of the times.
-For example in a pure OO language, \Q@String@ may offer a \Q@Int size()@
+In a pure OO language, \Q@String@ may offer a \Q@Int size()@
 method, and \Q@Int@ may offer a \Q@String toString()@ method.
-This means that is not possible to type in (full) isolation classes
-\Q@String@ and \Q@Int@.
+This means that typing classes 
+\Q@String@ and \Q@Int@ in isolation is not possible.
 
@@
 Each group may also depend from a number of other groups.
-This would form a Direct Acyclic Graph of groups.
+This would form a \emph{direct acyclic graph} of groups.
 To type a group, we first need to type all depended groups, then
@@
 a composition error could be generated at this stage.
-On the opposite side, if \Q@B@ and \Q@C@ was swapped, as in
+On the opposite side, if \Q@B@ and \Q@C@ were swapped, as in
 \saveSpace\begin{lstlisting}
@@
 However they are both unavailable: \Q@B@ is still not computed and 
-\Q@A@ can not be compiled/checked without information about \Q@B@.
+\Q@A@ cannot be compiled/checked without information about \Q@B@.
 A type error would be generated, on the line of ``flattening of \Q@C@

### 2017-06-06 FACS/Conclusion.tex
Formatting

@@
 radically decouple inheritance/code reuse and subtyping.
-
-The model presented here is very easy to extend:
-For example nested classes can be easly added,giving us enough power to easly solve the expression problem as in \cite{deep}. More composition operators can be added in addition to \use.
-In particular all the sophisticated operators of \cite{deep} could be added.
+The model presented here is easy to extend.
+For example nested classes can be easily added, giving us enough power
+to easly solve the Expression Problem~\cite{EP}, as demonstrated previously in
+DeepFJig~\cite{deep}. More composition operators can be added in addition to \use.
+In particular all the sophisticated operators of DeepFJig can be added.
  Indeed we can add any operator respecting following criteria:
@@
  such result is well typed.
-\item When the operator takes in input also code literals, if a non well typed result is produced,
-the type error must be tracked back to code in one of those non typed yet code literals.
+\item When the operator takes in input also code literals, if a non-well typed result is produced,
+the type error must be traced back to code in one of those not-yet typed code literals.
  \end{itemize}
 
- Our simplified model represent the conceptual core of  a novel full blown programming language 
- (\verb@http://L42.is@),
-that leverage on the ideas presented in this paper to obtain reliable and understandble metaprogramming.
-Formalization (in progress) for such language can be found at 
-\verb@urlOmittedForDoubleBlindReview@.
+ Our simplified model represents the conceptual core of  42: a novel full blown programming language,
+which leverages on the ideas presented in this paper to obtain reliable and understandable metaprogramming.
+Formalization (in progress) for full 42 can be found at
+\url{http://}\footnote{Ommitted for anonymous review}. 
+%\verb@urlOmittedForDoubleBlindReview@.
 %\verb@github.com/ElvisResearchGroup/L42/tree/master/Main/formal@.
-L42 extends our model allowing
+42 extends our model allowing
 flattening to execute arbitrary computations.
 In such model we do not need an explicit notion of traits: they are encoded as method returning a code literal.
-L42 also have features less related to code composition, like
+42 also have features less related to code composition, like
   a strong type system supporting aliasing mutability and circularity control,
@@
 
-L42 do not have a finite set of composion operators; they can be added using the built in support for native method calls.
-They can be dynamically checked to verify that they are well behaved according to our predicate, or they can be trusted to acheive efficiency.
+42 do not have a finite set of composion operators; they can be
+added using the built in support for native method calls. They can
+be dynamically checked to verify that they are well behaved
+according to our predicate, or they can be trusted to acheive
+efficiency.

### 2017-06-06 FACS/Intro.tex
Formatting

@@
 \section{Introduction}\label{sec:intro}
-\saveSpace\saveSpace
 In mainstream OO languages like Java, C++ or C\# subclassing 
@@
 \begin{itemize}
-\item In {\bf Delta-Trait (DT)}~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}
+\item In {\bf TraitRecordJ (TR)}~\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}
 each  construct has a single responsability: classes instantiate objects,
@@
 \item {\bf
-    DeepFjig}~\cite{deep,servetto2014meta,fjig} is
+    DeepFJig}~\cite{deep,servetto2014meta,fjig} is
 a module composition language where the main idea is that

### 2017-06-06 FACS/RelatedWork.tex
Revision
paper
Formatting

@@
 \subsection{Separating Inheritance and Subtyping}
-\saveSpace
 We are aware of at least 3 independently designed research languages 
-that address the this-leaking problem: Delta-Trait~(DT)\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}, Package Templates~(PT)\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}, DeepFjig\cite{deep,servetto2014meta,fjig}.
+that address the this-leaking problem: TraitRecordJ~(TR)\cite{Bettini:2010:ISP:1774088.1774530,BETTINI2013521,Bettini2015282}, Package Templates~(PT)\cite{KrogdahlMS09,DBLP:journals/taosd/AxelsenSKM12,DBLP:conf/gpce/AxelsenK12}, DeepFJig~\cite{deep,servetto2014meta,fjig}.
 Leveraging on \emph{traits}, in this work we aim to synthesize
@@
 \item {\bf A simple uniform syntax for code literals}
-DeepFjig is best in this sense, since DT has separate syntax for class literals, trait literals and record literals.
+DeepFJig is best in this sense, since TR has separate syntax for class literals, trait literals and record literals.
 PT on the other hand is built on top of full Java, thus has a very
 involved syntax.
-\name leverages on DeepFjig's approach but,
+\name leverages on DeepFJig's approach but,
 \emph{thanks to our novel representation of state}, \name also offers a much simpler and uniform syntax than
@@
 {\bf Reusable code cannot be ``used'', that is instantiated or used as a type.}
-This happens in DT and in PT, but not in DeepFjig. To allow reusable code to be directly 
-usable, in DeepFjig
+This happens in TR and in PT, but not in DeepFJig. To allow reusable code to be directly 
+usable, in DeepFJig
 classes introduce nominal types in an unnatural way: the type of
@@
 nominal type of its class. 
-That is in DeepFjig 
+That is in DeepFJig 
 \Q@A:{ method A m()this}@ is not well typed. This is because
@@
 Looking to this example is clear why we need reusable code to be agnostic of its name.
-Then, either reusable code has no name (as in DT, PT and \name)
+Then, either reusable code has no name (as in TR, PT and \name)
 or all code is reusable and usable, and all code needs to be awkwardly agnostic of its name.
@@
 {\bf Requiring abstract signatures is a left over of module composition mindset.}
-DT and Deepfjig comes from a tradition of functional module composition, where 
+TR and DeepFJig comes from a tradition of functional module composition, where 
 modules are typed in isolation under an environment, and then the composition is performed.
@@
 heavy task of integrating with full Java.
-Recent versions of Delta-Traits are trying to address this issue too.
+%Recent work based on TR is trying to address this issue too.
 \item {\bf Composition algebra.}
 The idea of using composition operators over atomic values as in an arithmetic expression is very powerful,
-and makes it easy to extend languages with more operators. Deepfjig and DT embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
+and makes it easy to extend languages with more operators. DeepFJig and TR embrace this idea, while PT takes the traditional Java/C++ approach of using enhanced class/package declaration syntax.
 The typing strategy of PT also seems to be connected with this
@@
 \item {\bf Complete ontological separation between use and reuse}
-While all 3 works allow separating inheritance and subtyping only DT properly enforces 
+While all 3 works allow separating inheritance and subtyping only TR properly enforces 
 separation between use (classes and interfaces) and reuse (traits).
-This is because in DeepFjig all classes are both units of use and reuse (however, subtyping is not induced).
+This is because in DeepFJig all classes are both units of use and reuse (however, subtyping is not induced).
 PT imports all the complexity of Java, so although is possible to separate use and reuse, the model have powerful but non-obvious implications where (conventional Java) \Q@extends@ and PT are used together.
-\item {\bf Naming your type, even if there is none yet.}
-Both Deepfjig and PT allow a class to refer to its name, albeit this is
+\item {\bf Naming the self type, even if there is none yet.}
+Both DeepFJig and PT allow a class to refer to its name, albeit this is
 less obvious in PT since both a package and a class have to be introduced to express it.
@@
 The composed constructor will execute both constructor bodies in order.
-This approach is designed to allows Self construction.
+This approach is designed to allow self construction.
 It is also associative and mostly commutative: composition order only influences execution order of side effects during construction.
@@
 it would be unsound.
-Recently work on ThisType \cite{ToplasAndIgarashi}
-also continue on this line. In those works, ``subtyping by subclassing'' is considered a valuable feature.
+Recent work on {\bf ThisType} \cite{Saito:2009,ryu16ThisType}
+also continues on this line.
+In those works, ``subtyping by subclassing'' is considered a valuable
+feature.
 

### 2017-06-06 FACS/paper.tex
Formatting

@@
 We are aware of at least 3 independently designed research languages 
-addressing this limitation: \emph{Delta-Trait}, \emph{Package Template}, \emph{DeepFjig}.
+addressing this limitation: \emph{TraitRecordJ}, \emph{Package Templates}, \emph{DeepFJig}.
 The goal of this paper is to synthesize and improve on
@@
 we can separate inheritance and subtyping, while avoiding 
-redundant abstract declarations required in Delta-Trait and
-DeepFjig. At the same time \emph{self construction},
+redundant abstract declarations required in TraitRecordJ and
+DeepFJig. At the same time \emph{self construction},
 \emph{binary methods} and \emph{recursive types} are also supported.

### 2017-06-06 FACS/section2.tex
paper
Formatting

@@
 no method called \Q@ma@ is explicitly declared.
-DeepFJig and DT would require instead to explicitly declare an abstract \Q@ma@:
+DeepFJig and TR would require instead to explicitly declare an abstract \Q@ma@:
 \saveSpace
@@
 \noindent
-The idea is that such method is imported from trait \Q@ta@, exactly as in the Java equivalent
+The idea in \name is that such method is imported from trait \Q@ta@, exactly as in the Java equivalent
 \saveSpace\begin{lstlisting}[language=Java]

### 2017-06-06 FACS/Intro.tex
Submission

@@
 a module composition language where the main idea is that
-nested classes with the same name are recursivelly composed.
+nested classes with the same name are recursively composed.
 \end{itemize}
@@
 approach to state in traits that avoids the complexities introduced by
-fields and their initialization in previous approaches.
+fields and their initialisation in previous approaches.
 In \name there are two separate concepts: classes
@@
 %\marco{We need to talk of unanticipated extensions?}
-%\bruno{Talk more about typing aspects here. Summarize the important 
+%\bruno{Talk more about typing aspects here. Summarise the important 
 %aspects of the design of \name.}

### 2017-06-06 FACS/RelatedWork.tex
Submission

@@
 \emph{thanks to our novel representation of state}, \name also offers a much simpler and uniform syntax than
-all other approaches: everyhing is just a method.
+all other approaches: everything is just a method.
 \item 
@@
   However, objects are created "broken" and the user is trusted with fixing them.
-  While is easy to add fields, the load of inizializing them is on the user; moreover
+  While is easy to add fields, the load of initializing them is on the user; moreover
     all the objects are intrinsically mutable, so this model is unfriendly
@@
 Self construction is not possible 
-since the signature of the constuctors changes during composition.
+since the signature of the constructors changes during composition.
 
@@
 hampers reuse, and if a field is added, its initial value needs to be
-somehow syntetized from the constructor parameters.
+somehow syntesized from the constructor parameters.
 
@@
 
-%\item extension with nested class is natural and powefull
+%\item extension with nested class is natural and powerful
 %We claim that our solution to the expression problem is the most natural in literature to date:
@@
 
-%Add that: original trait model has no self construction and types introduction not relevant since dynamicly typed language
+%Add that: original trait model has no self construction and types introduction not relevant since dynamically typed language
 
@@
 also continues on this line.
-In those works, ``subtyping by subclassing'' is considered a valuable
-feature.
+In those works, ``subtyping by subclassing'' is preserved, which means
+that those designs are more suitable to retain the programming model
+of mainstream OOP languages and backwards compatibility. The design 
+of \name (and 42) is a more radical departure of mainstream OOP, with
+the hope to improve both the mechanisms for use and reuse in OOP.
 

### 2017-06-06 FACS/section2.tex
Submission

@@
 
-\subsection{Traits in \name: A mechanim of code Reuse}
+\subsection{Traits in \name: A mechanism of code Reuse}
 
@@
 That is, using a trait is the only way to induce code reuse.
-Unfortunatelly, this code does not work, 
+Unfortunately, this code does not work, 
 because \lstinline{Utils.m} requires an \lstinline{A} and the type of \lstinline{this} in

### 2017-06-06 FACS/section3.tex
Submission

@@
 without introducing subtyping between Bags and Sets}. This improves the 
-use of Bags and eliminates logical errors arizing from incorrect
+use of Bags and eliminates logical errors arising from incorrect
 subtyping relations that are allowed in the Java solution.
@@
 \noindent Coding \Q@Bag@ in this way avoids a lot of code
-duplication, but \emph{we induced unintented subtyping}! 
+duplication, but \emph{we induced unintended subtyping}! 
 Since subclassing implies subtyping, our code breaks the Liskov substitution principle (LSP)~\cite{martin2000design}: not all bags are sets!
@@
 \paragraph{A (broken) attempt to fix the Problem in Java.}
-One could \emph{retroactivelly} fix this problem by introducing \Q@AbstractSetOrBag@
+One could \emph{retroactively} fix this problem by introducing \Q@AbstractSetOrBag@
 and making both \Q@Bag@ and \Q@Set@ inherit from it:
@@
 and we would be surprised to find a use of the type \Q@AbstractSetOrBag@.
-Worst, if we are to constantly apply this mentaly, we would introduce a very high number
+Worst, if we are to constantly apply this mentally, we would introduce a very high number
 of abstract classes that are not supposed to be used as types. Those classes would clutter the 
@@
 it''~\cite{Bloch08}: the
-self use of public methods is rarelly documented, thus is hard to understand the effects of overriding a library metod.
+self use of public methods is rarely documented, thus is hard to understand the effects of overriding a library method.
 }!
@@
 and as shown in Section 2, there is no trace of trait \Q@bag@.
-Thus if we are the developers of bags, we can temporarly go for the first version, and move when needed to the second without adding new undesired complexity for our old clients.
+Thus if we are the developers of bags, we can temporarily go for the first version, and move when needed to the second without adding new undesired complexity for our old clients.
 

### 2017-06-06 FACS/section4.tex
Submission

@@
 to deal with \emph{state} in traits.  The idea of summing pieces of
-code is very elegant, and has proven very successfull in module
+code is very elegant, and has proven very successful in module
 composition languages~\cite{ancona2002calculus} and several trait
@@
 
-\paragraph{Modeling Points} Consider, for example, two simple 
+\paragraph{Modelling Points} Consider, for example, two simple 
 traits that deal with \emph{point} objects. That is, points
@@
 abstract only when the set of abstract methods is not coherent. That
-is, the unimplemented methods cannot not be automatically recognized
-as factory, getters and setters. Methods recognized as factory, getters and setters are called
+is, the unimplemented methods cannot not be automatically recognised
+as factory, getters and setters. Methods recognised as factory, getters and setters are called
 \emph{abstract state operations}.
@@
 \end{lstlisting}  \saveSpace
-  \noindent This design allows merging colors, or any other kind of state we may want to add
+  \noindent This design allows merging colours, or any other kind of state we may want to add
   following this pattern.%\bruno{worried that withers are not explained enough.}
@@
 \paragraph{Independent Extensibility}
-  Of course, quite frequentely there can be multiple independent
+  Of course, quite frequently there can be multiple independent
   extensions~\cite{Zenger-Odersky2005} that need to be composed. Lets suppose that 
@@
   need to compose the merge operation inside of both of them.
-  The simple model we are presenting could accomodate this with an
+  The simple model we are presenting could accommodate this with an
   extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations.
@@
 
-The following code shows how to mix colors and flavours. The syntax \Q@_2merge@ and   \Q@_3merge@
+The following code shows how to mix colours and flavours. The syntax \Q@_2merge@ and   \Q@_3merge@
 call the version of merge as defined in the second/third element of \use.

### 2017-11-29 FACS/RelatedWork.tex
Ecoop style+re enabled extensions

@@
 
-%\item extension with nested class is natural and powerful
-%We claim that our solution to the expression problem is the most natural in literature to date:
-%While a similar sintax can be achieved with the scandinavian style~\cite{ernst2004expression}, their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system~\cite{igarashi2005lightweight}.
+Extension with nested class is natural and powerful
+We claim that our solution to the expression problem is the most natural in literature to date:
+While a similar sintax can be achieved with the scandinavian style~\cite{ernst2004expression}, their dependent type system makes reasoning quite complex, and indeed more recent solutions have accepted a more involved syntax in order to have an understandable type system~\cite{igarashi2005lightweight}.
 
-%Our close contendant is DeepFJig~\cite{corradi2011deepfjig}: all our gain over their model is based on our relaxation over abstract signatures.
-%Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}
+Our close contendant is DeepFJig~\cite{corradi2011deepfjig}: all our gain over their model is based on our relaxation over abstract signatures.
+Note how we solve the expression problem in the most radical way possible \marco{bruno, here you can insert some of the criticism over expression problem trivially and explain how is not applicable to us here.}
 
-%Add that: original trait model has no self construction and types introduction not relevant since dynamically typed language
+Add that: original trait model has no self construction and types introduction not relevant since dynamically typed language
 
@@
 
-\begin{comment}
 \subsection{Tablular comparision of many approaches}
@@
 }
-\end{comment}

### 2017-11-29 FACS/paper.tex
Submission
Ecoop style+re enabled extensions

@@
-%
-\documentclass{llncs}
-%
+\documentclass[a4paper,UKenglish]{lipics-v2016}
+%This is a template for producing LIPIcs articles. 
+%See lipics-manual.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+% for section-numbered lemmas etc., use "numberwithinsect"
+
+\usepackage{microtype}%if unwanted, comment out or use option "draft"
+
+%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+
+\bibliographystyle{plainurl}% the recommended bibstyle
+
 \usepackage{url}
@@
 %
-\newcommand\saveSpace{\vspace{-3pt}}
+\newcommand\saveSpace{}
+%\newcommand\saveSpace{\vspace{-3pt}}
+
 \newcommand\Rotated[1]{\begin{turn}{90}\begin{minipage}{12em}#1\end{minipage}\end{turn}}
@@
 
-%
-\begin{document}
-%
-\frontmatter          % for the preliminaries
-%
-\pagestyle{headings}  % switches on printing of running heads
-%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
-%
+% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Separating Use and Reuse to Improve Both}
+\titlerunning{Separating Use and Reuse to Improve Both}
 
-\title{Separating Use and Reuse to Improve Both}
-%
-\titlerunning{Separating Use and Reuse to Improve Both}  % abbreviated title (for running head)
-
+\author{Authors omitted for double-bind review.}
+%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
+%\author[1]{John Q. Open}
+%\author[2]{Joan R. Access}
+%\affil[1]{Dummy University Computing Laboratory, Address/City, Country\\
+%  \texttt{open@dummyuniversity.org}}
+%\affil[2]{Department of Informatics, Dummy College, Address/City, Country\\
+%  \texttt{access@dummycollege.org}}
+%\authorrunning{J.\,Q. Open and J.\,R. Access} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
 \begin{comment}
@@
 \end{comment}
-\author{Authors omitted for double-bind review.}
 
-\maketitle              % typeset the title of the contribution
+\Copyright{John Q. Open and Joan R. Access}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+
+\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html .
+E.g., cite as "F.1.1 Models of Computation".
+\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
+% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Acces}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+\maketitle
 
@@
 Moreover, we provide a novel and elegant solution to uniformly
-handle behavior and state within trait composition.
+handle behaviour and state within trait composition.
 These ideas have been implemented in the 42 language, 
@@
 \input{Examples}
-%\input{Extensions}
+\input{Extensions}
 \input{RelatedWork}
@@
 
-\bibliographystyle{plain}
+%\subparagraph*{Acknowledgements.}
+
+%I want to thank \dots
+
+%\appendix
+%\section{Morbi eros magna}
+
+%%
+%% Bibliography
+%%
+
+%% Please use bibtex,
+
 \bibliography{main}

### 2017-11-29 FACS/section2.tex
Ecoop style+re enabled extensions

@@
 \begin{lstlisting}
- A: { method int ma() Utils.m(this) }//note, no {return _}
- Utils:{ class method int m(A a)/*method body here*/ }
+ A: { method int ma() Utils.m(this) }//note, no need for {return _}
+ Utils:{ class method int m(A a)/*some method body here*/ }
 \end{lstlisting} 
@@
 % this text may go after, while explaining the This type?
-%Code inside traits/classes can refer to (uppercase) class names, but not to (lowercase) trait names.
-%%The point is that the trait can refer to the program via \lstinline{this},
+Code inside traits/classes can refer to (uppercase) class names, but not to (lowercase) trait names.
+The point is that the trait can refer to the program via \lstinline{this},
 %??????????????
-% This means that the
-%program is agnostic to what the trait type is going to be, so it can
-%be later assigned to any (or many) classes. This improves the
-%flexibility of reuse as illustrated in Section~\ref{}.
-%However, 
-%to solve the more immediate typing issue above we need one more round of refactoring, as we shall see next.
+ This means that the
+program is agnostic to what the trait type is going to be, so it can
+be later assigned to any (or many) classes. This improves the
+flexibility of reuse as illustrated in Section~\ref{}.
+However, 
+to solve the more immediate typing issue above we need one more round of refactoring, as we shall see next.
 
-%: in \name all declarations 
-%are sequentially type-checked and if a definition occurs afterwards
-%another, then it is not visible on all previous definitions.
+in \name all declarations 
+are sequentially type-checked and if a definition occurs afterwards
+another, then it is not visible on all previous definitions.
 
-%\paragraph{Type of the self-reference} 
-%The code above does not work.
-%because the way type-checking works in \name is that type-checking of 
-%code literals is independent of the class/trait names associated to it.
-%A Java programmer may expect that the type of \lstinline{this} 
-%in the previous definition of \lstinline{ta} is \lstinline{ta}
-%itself. However, this intuition brought from Java is wrong in \name
-%for two reasons:
+\paragraph{Type of the self-reference} 
+The code above does not work.
+because the way type-checking works in \name is that type-checking of 
+code literals is independent of the class/trait names associated to it.
+A Java programmer may expect that the type of \lstinline{this} 
+in the previous definition of \lstinline{ta} is \lstinline{ta}
+itself. However, this intuition brought from Java is wrong in \name
+for two reasons:
 
-%\begin{itemize}
+\begin{itemize}
 
-%\item {\bf Traits are not types}. Traits in
-%  \name are simply units of reuse and cannot be used as types. Types
-%  are only introduced by class/interface declarations.
+\item {\bf Traits are not types}. Traits in
+  \name are simply units of reuse and cannot be used as types. Types
+  are only introduced by class/interface declarations.
 
-%\item The second reason is that {\bf the type of self-references is the self type of the
-%    code literal.} Unlike Java-like languages, where the body of a
-%  class declaration is intrinsically coupled with the class itself, in
-%  \name code literals are first-class\bruno{Is it appropriate to say
-%    that they are first class?} and are type-checked independently.
-%  \name has a notion of self-type, which is closely related to
-%  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
-%  in \name the self-reference does not have the type of the class
-%  being defined, but rather it has the self-type.
-%\end{itemize}
+\item The second reason is that {\bf the type of self-references is the self type of the
+    code literal.} Unlike Java-like languages, where the body of a
+  class declaration is intrinsically coupled with the class itself, in
+  \name code literals are first-class\bruno{Is it appropriate to say
+    that they are first class?} and are type-checked independently.
+  \name has a notion of self-type, which is closely related to
+  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
+  in \name the self-reference does not have the type of the class
+  being defined, but rather it has the self-type.
+\end{itemize}
 
@@
 
-%To finish this section, Figure \ref{fig:compare} provides a summary of
-%the differences between classes and traits. The comparison focus on
-%the roles of traits and classes with respect to instantiation,
-%reusability and whether the declarations also introduce new types or
-%not.
+To finish this section, Figure \ref{fig:compare} provides a summary of
+the differences between classes and traits. The comparison focus on
+the roles of traits and classes with respect to instantiation,
+reusability and whether the declarations also introduce new types or
+not.
